### 2022-05-28 notebooks/theory_notes.tex
some theory notes in progress

\documentclass[11pt]{article}
\RequirePackage{fullpage}
\RequirePackage[font=small,labelfont=bf]{caption}
\RequirePackage{amsmath,amssymb,amsthm}
\RequirePackage{graphicx}
\RequirePackage[hidelinks]{hyperref}
\RequirePackage{subcaption}
\RequirePackage{wasysym}
\RequirePackage{authblk}
\RequirePackage{bm}
\RequirePackage{bbm}
% \RequirePackage[osf]{mathpazo}
\RequirePackage[bibstyle=authoryear,citestyle=authoryear-comp,maxbibnames=9,maxcitenames=1,backend=biber,natbib=true,uniquelist=false,hyperref=true]{biblatex}
\usepackage{color}
\usepackage{nicefrac}

\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\text{V}}
\DeclareMathOperator{\var}{Var}
\DeclareMathOperator{\cov}{cov}

\addbibresource{biblio.bib}

\title{Learning Theory Notes}

\begin{document}
\maketitle

\section*{Learning Limits}

\subsection*{The Simulation Sample $\bar{B}$ Estimator}

Our goal is to approximate the function $B = f(X)$ with some learned function
$\widehat{B} = \hat{f}(X)$ from evolutionary simulations. In each evolutionary
simulation, we sample some evolutionary parameters $X_k$ and evolve $r$
independent populations forward in time under these parameters and observe some
genealogy. From this genealogy, we estimate the reduction in neutral diversity
as:

\begin{align}
  \bar{B}_k = \frac{1}{r} \sum_{i=1}^r \frac{\hat{\pi}_{k,i}}{4N\mu}
\end{align}

there $\hat{\pi}$ is Tajima's estimator for pairwise diversity within a tree.
When we estimate $\hat{\pi}$ from using branch-mode tree statistics from an
observed genealogy, $\mu \to 1$, so we can ignore mutation.  If we take
expectation over the evolutionary process,

\begin{align}
  \E(\bar{B}_k) &= \frac{1}{4Nr} \sum_{i=1}^r \E(\hat{\pi}_{k,i}) \\
                &= \frac{T_k^{(2)}}{4N}  \\
                &= \frac{4B_kN}{4N}  \\
                &= B_k  \\
\end{align}

thus, the estimated reduction in diversity from simulations $\bar{B}$ is
unbiased, since $\E(\hat{\pi}_k) = 2T_k^{(2)} = 4B_k N$. For independent
evolutionary replicates, we calculate the variance of this estimator using
Tajima's equation for the variance of $\hat{\pi}$ as

\begin{align}
  \var(\bar{B}_k) &= \frac{1}{16 N^2 r^2} \sum_{i=1}^r \var(\hat{\pi}_{k,i}) \\
                  &= \frac{\var(\hat{\pi}_{k,i})}{16 N^2 r} \\
                  &= \frac{1}{16N^2 r} \left( \frac{n + 1}{3(n-1)}\theta  + \frac{2(n^2 + n + 3)}{9n(n-1)}\theta^2 \right) \\
                  &= \frac{1}{16N^2 r} \left( \frac{n + 1}{3(n-1)} 4BN + \frac{2(n^2 + n + 3)}{9n(n-1)} 16B^2N^2 \right) \\
                  &= \underbrace{\frac{n + 1}{12(n-1)} \frac{B}{N r}}_\text{sampling noise} + \underbrace{\frac{2(n^2 + n + 3)}{9n(n-1)} \frac{B^2}{r}}_\text{evolutionary variance}
\end{align}

Now, let us look at the consistency of the estimator $\bar{B}$ (we drop the
parameter set $k$ for clarity) both in $r$ (over evolutionary replicates) and
in $n$ (as the sample size increases). Let $\bar{B}_r$ be the estimator of $B$
after $r$ evolutionary replicates (conditioning on some $n$). By Chebyshev's
inequality and for some $\epsilon > 0$,

\begin{align}
  \P\left(|\bar{B}_r - B| \ge \epsilon \right)  \le \frac{\var(\bar{B}_r)}{\epsilon^2}
\end{align}
%
and since $\lim_{r \to \infty} \var(\bar{B}_r) = 0$,

\begin{align}
  \lim_{r \to \infty} \P\left(|\bar{B}_r - B| \ge \epsilon \right)  = 0.
\end{align}

Thus, $\bar{B}_r \xrightarrow{p} B$ as $r \to \infty$ and $\bar{B}_r$ is
consistent in $r$. Now, let us look at the consistency of $\bar{B}_n$ in sample
size $n$. Note that $n \le 2N$ (i.e. our sample size is bounded by the number
of gametes in the population), so we imagine setting $n = 2N$ and taking the
limit $N \to \infty$,

\begin{align}
  \lim_{N \to \infty} \P\left(|\bar{B}_N - B| \ge \epsilon \right)  &\le \lim_{N \to \infty} \frac{\var(\bar{B}_N)}{\epsilon^2} \\
  \lim_{N \to \infty} \P\left(|\bar{B}_N - B| \ge \epsilon \right)  &\le \frac{2B^2}{9 r \epsilon^2}. \\
\end{align}

Thus, even if we sample the entire population, and let the population size $N
\to \infty$, $\bar{B}_n$ is still an inconsistent estimator in $n$. Intuitively
this is because

In our case, we estimate pairwise diversity from the genealogical tree of the
entire population, so $n = 2N$

\begin{align}
  \var(\bar{B}_k) &= \frac{n + 1}{12(n-1)} \frac{B}{N r} + \frac{2(n^2 + n + 3)}{9n(n-1)} \frac{B^2}{r} \\
                  &= \frac{2N + 1}{12(2N-1)} \frac{B}{N r} + \frac{2(4N^2 + 2N + 3)}{18N(2N-1)} \frac{B^2}{r} \\
                  &\approx \frac{B}{12N r} + \frac{2N^2 + N}{9N^2} \frac{B^2}{r}
\end{align}

If we take the infinite population size limit,

\begin{align}
  \lim_{N \to \infty} \var(\bar{B}_k) &= \frac{2B^2}{9r}
\end{align}

\printbibliography

\end{document}

### 2022-05-30 notebooks/theory_notes.tex
changes to notes

@@
 
-Our goal is to approximate the function $B = f(X)$ with some learned function
-$\widehat{B} = \hat{f}(X)$ from evolutionary simulations. In each evolutionary
-simulation, we sample some evolutionary parameters $X_k$ and evolve $r$
-independent populations forward in time under these parameters and observe some
-genealogy. From this genealogy, we estimate the reduction in neutral diversity
-as:
+Our goal is to approximate a function that describes the reduction in pairwise
+diversity (relative to the neutral expectation) $B$ due to BGS for some set of
+evolutionary parameters $\mathbf{x}$. The true reduction function is a
+conditional expectation $B(\mathbf{x}) = \E[T_2(\mathbf{x}) | \mathbf{x}]$
+taken over an infinite number of evolutionary replicates. Here
+$T_2(\mathbf{x})$ is the pairwise coalescent rate under BGS for parameters
+$\mathbf{x}$, with time scaled in $2N$ generation units.
+
+We approximate this function $B(\mathbf{x})$ with $\widehat{B}(\mathbf{x})$
+from evolutionary simulations using statistical learning. In each evolutionary
+simulation, we sample some parameter set $\mathbf{x}$ and evolve $r$
+independent populations forward in time and observe the resulting genealogy at
+a neutral site a given recombination distance away (this is one of the
+parameters). From this full population genealogy of $2N$ gametes, we estimate
+the reduction in neutral diversity as
 
 \begin{align}
-  \bar{B}_k = \frac{1}{r} \sum_{i=1}^r \frac{\hat{\pi}_{k,i}}{4N\mu}
+  \bar{B} = \frac{1}{r} \sum_{i=1}^r \frac{\hat{\pi}_{i}}{4N\mu}
+\end{align}
+%
+where $\hat{\pi}$ is Tajima's estimator for pairwise diversity within a tree.
+For now, we can imagine letting the neutral mutations saturate as $\mu \to 1$
+so we can ignore mutation rate hereafter. If we take expectation over the evolutionary
+process,
+
+\begin{align}
+  \E[\bar{B}] &= \frac{1}{4Nr} \sum_{i=1}^r \E[\hat{\pi}_{i}] \\
+                          &= \frac{2T_2}{4N}  \\
+                          %&= \frac{4B(\mathbf{x})N}{4N}  \\
+                          &= B
+\end{align}
+%
+thus, the estimated reduction in diversity from simulations $\bar{B}$ is
+unbiased, since $\E[\hat{\pi}] = 2T_2 = 4BN$.
+
+\subsection*{The variance of the estimator $\bar{B}$}
+
+Using the law of total variance, we can write the variance of $\bar{B}$ as
+
+\begin{align}
+  \var(\bar{B}) = \underbrace{\E(\var(\bar{B} | \mathcal{G}))}_\text{sampling noise} + \underbrace{\var(\E(\bar{B} | \mathcal{G}))}_\text{evolutionary variance}
 \end{align}
 
-there $\hat{\pi}$ is Tajima's estimator for pairwise diversity within a tree.
-When we estimate $\hat{\pi}$ from using branch-mode tree statistics from an
-observed genealogy, $\mu \to 1$, so we can ignore mutation.  If we take
-expectation over the evolutionary process,
+where $\mathcal{G}$ is the genealogy at the focal neutral site affected by BGS
+at some segment. The distribution of genealogies $P(\mathcal{G})$ is unknown
+and quite complex under BGS, but we can use a simplifying assumption to get a
+rough estimate of the variance of $\bar{B}$.
+
+Under strong BGS (e.g. $\nicefrac{\mu}{s}N \gg 1$), the genealogy is often
+modeled as a neutral coalescent process with rescaled with an effective
+population size $N_e = BN$ (though see Cvijovi\'{c} et al. 2018). In the weak
+selection limit $s \to 0$, approximation becomes exact as BGS becomes
+effectively neutral. In both regimes, selection is treated as rescaling a
+neutral genealogical process to $N_e = BN$. We refer to this the
+\textbf{neutral-BGS model}, and it allows for us to approximate $\var(\bar{B})$
+with some $\var(\bar{B}')$ where $\bar{B}'$ is the mean reduction if we
+averaged over neutral genealogies from a population of $BN$
+diploids, rather than averaged over exact selection simulations. Note that at
+$B=1$, the neutral-BGS model becomes exact, as there is no reduction in
+diversity due to background selection and the only evolutionary variance is
+through the neutral coalescent process.
+
+If the genealogies are neutral, we can calculate the variance of $\bar{B}$
+using Tajima's equation for the variance of $\hat{\pi}$. For $n$ samples and
+$r$ evolutionary replicates,
 
 \begin{align}
-  \E(\bar{B}_k) &= \frac{1}{4Nr} \sum_{i=1}^r \E(\hat{\pi}_{k,i}) \\
-                &= \frac{T_k^{(2)}}{4N}  \\
-                &= \frac{4B_kN}{4N}  \\
-                &= B_k  \\
-\end{align}
-
-thus, the estimated reduction in diversity from simulations $\bar{B}$ is
-unbiased, since $\E(\hat{\pi}_k) = 2T_k^{(2)} = 4B_k N$. For independent
-evolutionary replicates, we calculate the variance of this estimator using
-Tajima's equation for the variance of $\hat{\pi}$ as
-
-\begin{align}
-  \var(\bar{B}_k) &= \frac{1}{16 N^2 r^2} \sum_{i=1}^r \var(\hat{\pi}_{k,i}) \\
-                  &= \frac{\var(\hat{\pi}_{k,i})}{16 N^2 r} \\
+  \var(\bar{B}') &= \frac{1}{16 N^2 r^2} \sum_{i=1}^r \var(\hat{\pi}_{i}) \\
+                &= \frac{\var(\hat{\pi}_{i})}{16 N^2 r} \\
                   &= \frac{1}{16N^2 r} \left( \frac{n + 1}{3(n-1)}\theta  + \frac{2(n^2 + n + 3)}{9n(n-1)}\theta^2 \right) \\
@@
   \lim_{N \to \infty} \P\left(|\bar{B}_N - B| \ge \epsilon \right)  &\le \lim_{N \to \infty} \frac{\var(\bar{B}_N)}{\epsilon^2} \\
-  \lim_{N \to \infty} \P\left(|\bar{B}_N - B| \ge \epsilon \right)  &\le \frac{2B^2}{9 r \epsilon^2}. \\
+  \lim_{N \to \infty} \P\left(|\bar{B}_N - B| \ge \epsilon \right)  &\le \frac{2B^2}{9 r \epsilon^2}. \label{eq:bound}
 \end{align}
@@
 \to \infty$, $\bar{B}_n$ is still an inconsistent estimator in $n$. Intuitively
-this is because
+this is because the evolutionary variance is not decreased by taking more
+samples, \emph{only by taking more evolutionary replicates}.
 
-In our case, we estimate pairwise diversity from the genealogical tree of the
-entire population, so $n = 2N$
+Below I work through where this bound came from. If we estimate pairwise
+diversity from the genealogical tree of the entire population such that $n =
+2N$,
 
 \begin{align}
-  \var(\bar{B}_k) &= \frac{n + 1}{12(n-1)} \frac{B}{N r} + \frac{2(n^2 + n + 3)}{9n(n-1)} \frac{B^2}{r} \\
+  \var(\bar{B}) &= \frac{n + 1}{12(n-1)} \frac{B}{N r} + \frac{2(n^2 + n + 3)}{9n(n-1)} \frac{B^2}{r} \\
                 &= \frac{2N + 1}{12(2N-1)} \frac{B}{N r} + \frac{2(4N^2 + 2N + 3)}{18N(2N-1)} \frac{B^2}{r} \\
@@
 \begin{align}
-  \lim_{N \to \infty} \var(\bar{B}_k) &= \frac{2B^2}{9r}
+  \lim_{N \to \infty} \var(\bar{B}) &= \frac{2B^2}{9r}
 \end{align}
+
+In practice, we use a different, better estimator of pairwise diversity than
+Tajima's $\pi$: branch statistic pairwise diversity (Ralph 2019, Ralph et al.
+2020). This estimator removes mutation as a source of sampling noise and since
+$n = 2N$, our only source of variance in $\bar{B}$ is in the evolutionary
+process.
+
+% Note that since the bound in Equation \eqref{eq:bound} is a probability, it
+% must be that $\frac{2B^2}{9 r \epsilon^2} \le 1$. Thus this equation is limited
+% to considering bounds $\epsilon$ that are
+
+% \begin{align}
+%   \sqrt{\frac{2}{9r}} B \le \epsilon \\
+% \end{align}
+
+% For $B \approx 1$, $\sqrt{\frac{2}{9r}} \le \epsilon$, such that if $r = 50$,
+% we can only consider bounds $\epsilon \ge \sqrt{\nicefrac{1}{225}} \approx
+% 0.067$ because Chebyshev isn't a particularly sharp inequality.
+
+\subsection*{Partitioning the Prediction Error}
+
+To learn the function $B(\mathbf{x})$, we run evolutionary simulations under a
+training set of parameters, giving us $\{(\mathbf{x}_1, \bar{B}_1),
+(\mathbf{x}_2, \bar{B}_2), \ldots, (\mathbf{x}_n, \bar{B}_n)\}$, where each
+$\bar{B}_i$ is the observed reduction over $r$ replicates. Since our goal is to
+approximate $\hat{B}(\mathbf{x})$ as closely as possible to $B(\mathbf{x})$
+over the distribution of $\mathbf{x}$ most close to that across the genome, we
+find some $\hat{B}(\mathbf{x})$ that minimizes expected loss. An important
+metric, even if training is conducted with a different loss, is the mean
+squared error
+
+\begin{align}
+  MSE(\bar{B}, \hat{B}) = \E[(\bar{B}(\mathbf{x}) - \hat{B}(\mathbf{x}))^2].
+\end{align}
+
+We can think of $\bar{B}(\mathbf{x})$ as a true reduction $B(\mathbf{x})$ plus
+a deviation $\varepsilon$ due to the particular evolutionary replicates
+simulated, e.g. $\bar{B}(\mathbf{x}) = B(\mathbf{x}) + \varepsilon$.
+
+\begin{align}
+  MSE(\bar{B}, \hat{B}) &= \E[(\bar{B}(\mathbf{x}) - \hat{B}(\mathbf{x}))^2] \\
+                        &= \E[(B(\mathbf{x}) + \varepsilon - \hat{B}(\mathbf{x}))^2] \\
+                        &= \underbrace{\var(\hat{B}(\mathbf{x}))}_\text{training variance} + 
+                      \underbrace{\var(\varepsilon)}_\text{evolutionary variance} + \underbrace{\E[\hat{B}(\mathbf{x}) - B(\mathbf{x})]^2}_{\mathrm{bias}^2}
+\end{align}
+
+At $B=1$, the evolutionary variance is exactly the neutral variance given by
+$\var(\varepsilon) = \nicefrac{2B^2}{9r}$. Thus, MSE loss has an absolute bound at
+$B=1$,
+
+\begin{align}
+  MSE(\bar{B}, \hat{B} | B = 1) \ge \frac{2}{9r}
+\end{align}
+
+For $B < 1$, we can write the evolutionary variance as some deviation
+away from the coalescent noise under the neutral-BGS model $\var(\varepsilon) =
+\var(\bar{B}') + e$.
+
+% For a particular training set
+% $\mathcal{T} = \{\mathbf{X}_\text{train}, \mathbf{B}_\text{train}\}$, the mean
+% squared error is
 

### 2022-05-30 bgspy/models.py
fixed slim sim error

@@ -356,7 +356,7 @@ class BGSModel(object):
             sites_chunk = np.array((pos_chunk, mpos_chunk)).T
             yield chrom, sites_chunk, (lidx, uidx)
 
-    def _build_segment_matrix(self, chrom, min_size=200):
+    def _build_segment_matrix(self, chrom):
         """
         The columns are the features of the B' training data are
                0,      1,          2,        3
@@ -367,17 +367,12 @@ class BGSModel(object):
         checking at the lower level, so it's done entirely during prediction.
         """
         segments = self.genome.segments
-        if min_size is not None:
-            idx = self.segments.L[chrom] >= min_size
-            nsegs = idx.sum()
-        else:
-            idx = slice(None)
-            nsegs = self.genome.segments.nsegs[chrom]
+        nsegs = self.genome.segments.nsegs[chrom]
         S = np.empty((nsegs, 4))
-        S[:, 0] = self.segments.L[chrom][idx]
-        S[:, 1] = self.segments.rbp[chrom][idx]
-        S[:, 2] = self.segments.mpos[chrom][idx, 0]
-        S[:, 3] = self.segments.mpos[chrom][idx, 1]
+        S[:, 0] = self.segments.L[chrom]
+        S[:, 1] = self.segments.rbp[chrom]
+        S[:, 2] = self.segments.mpos[chrom][:, 0]
+        S[:, 3] = self.segments.mpos[chrom][:, 1]
         return S
 
     def _build_segment_matrices(self):

### 2022-05-30 slim_sims/training/runner_script.sh

@@ -1,6 +1,6 @@
 #!/bin/bash
 
-NSCRIPTS=167
+NSCRIPTS=334
 
 i=0
 while [ $i -lt $NSCRIPTS ]

### 2022-05-30 tools/predict.py

@@ -107,7 +107,7 @@ def predict(chunkfile, input_dir, constrain=True, progress=True,
     # run the main prediction function
     B, Xps, Bpreds = predict_chunk(sites_chunk, models, Sm, w, t,
                                    lidx=lidx, uidx=uidx, 
-                                   use_haldane=False,
+                                   use_haldane=True, # TODO FIXME
                                    output_xps=output_xps, 
                                    output_preds=output_preds,
                                    # skip prediction if we just want matrices

### 2022-05-30 tools/slurm_slim_runner.py
fixed memory limitation bug

@@ -26,7 +26,7 @@ TEMPLATE = """\
 #SBATCH --time={job_time}
 #SBATCH --ntasks-per-node=1
 #SBATCH --cpus-per-task=1
-#SBATCH --mem=500M
+#SBATCH --mem=1G
 #SBATCH --account=kernlab
 #SBATCH --array=1-{max_array}
 

### 2022-06-05 notebooks/theory_notes.tex
updated notes with error model

@@
 
+\section*{Multiplicative Error Model}
+
+The predicted reduction for segment $i$ from the machine learned model
+$\hat{B}_i$ can be partitioned as $\hat{B}_i = B_i + \epsilon_i$, where
+$\epsilon_i$ is the cumulative error due to bias, irreducible evolutionary
+variance, and training variance. Then, our total predicted reduction at some
+focal site $v$ is:
+
+\begin{align}
+  \hat{B}(v) = \prod_i^n (B_i + \epsilon_i)
+\end{align}
+
+We want to understand the bias and variance of this product estimator based on
+the total learning error. Taking logs,
+
+\begin{align}
+  \hat{B}(v) &= \prod_i^n (B_i + \epsilon_i) \\
+  \log(\hat{B}(v)) &= \log\left(\prod_i^n (B_i + \epsilon_i) \right) \\
+                   &= \sum_i^n \log(B_i + \epsilon_i). \\
+\end{align}
+
+If we approximate the expected log with a Taylor series,
+
+\begin{align}
+  \E[\log(B_i + \epsilon_i)^2] &\approx \log(B_i)^2 + \frac{2 \log(B_i)}{B_i} \E[\epsilon_i] + \frac{(1-\log(B_i))}{B_i^2} \E[\epsilon_i^2]\\
+                               &\approx \log(B_i)^2 + \frac{2 \log(B_i)}{B_i}b + \frac{(1-\log(B_i)) }{B_i^2}(\sigma^2 + b^2)
+\end{align}
+
+Then,
+
+\begin{align}
+  \E[\log(\hat{B}(v))] &= \sum_i^n \E[\log(B_i + \epsilon_i)]  \\
+                                        &\approx \sum_i^n \log(B_i) + \sum_i^n \frac{b}{B_i} - \sum_i^n \frac{\sigma^2 + b^2}{2B_i^2}  \\
+\end{align}
+
+Bias:
+
+\begin{align}
+  \E[\log(\hat{B}(v))] - \E[\log(B(v))] &= \sum_i^n \E[\log(B_i + \epsilon_i)] - \sum_i^n \log(B_i) \\
+                                        &\approx \sum_i^n \log(B_i) + \sum_i^n \frac{b}{B_i} - \sum_i^n \frac{\sigma^2 + b^2}{2B_i^2} - \sum_i^n \log(B_i) \\
+                                        &\approx \sum_i^n \frac{b}{B_i} - \sum_i^n \frac{\sigma^2 + b^2}{2B_i^2} 
+\end{align}
+
+By Jensen's inequality $f(\E(X)) \le \E(f(X))$ for some convex function $f$,
+
+\section*{Likelihood}
+
+The likelihood in Eyalshiv et al. (2016) has the form:
+
+\begin{align}
+  \log\mathcal{L} = \sum_{v \in V} \sum_{i \ne j \in S} \log(P(O_{i,j}(v) | \theta))
+\end{align}
+
+where $V$ is the collection of neutral sites, $S$ is the set of samples, and
+$\theta$ are the BGS parameters. The indicator variable $O_{i,j}(v)$ is 1 if
+samples $i$ and $j$ are different at site $v$, and zero otherwise. Thus as they
+specify in the paper,
+
+\begin{align}
+  P(O_{i,j}(v) | \theta) = 
+    \begin{cases}
+      \pi(v | \theta), & O_{i,j}(v) = 1 \\
+      1-\pi(v | \theta), & O_{i,j}(v) = 0 \\
+    \end{cases}.
+\end{align}
+
+For $S$ samples, there are $|S| \choose 2$ pairwise comparisons that we can
+partition into the different and same classes; we assign their counts to
+$n_\text{D}(v)$ and $n_\text{S}(v)$. Then,
+
+\begin{align}
+  \log\mathcal{L} = \sum_{v \in V} \left(\log(\pi(v | \theta)) n_\text{S}(v) + \log(1-\pi(v | \theta)) n_\text{D}(v)\right)
+\end{align}
+
+Since $B$ is discretized into genomic windows of length $K$ basepairs, $\pi(v |
+\theta)$ is constant in a window.
+
 \printbibliography

### 2022-07-09 bgspy/classic.py
likelihood bug fixed (needs testing) many notebook changes, cleaning out old stuff

@@ -155,7 +155,7 @@ def calc_B(genome, mut_grid, step):
 
 
 def calc_B_chunk_worker(args):
-    map_positions, chrom_seg_mpos, segment_parts, mut_grid = args
+    map_positions, chrom_seg_mpos, F, segment_parts, mut_grid = args
     a, b, c, d = segment_parts
     Bs = []
     # F is a features matrix -- eventually, we'll add support for
@@ -168,7 +168,7 @@ def calc_B_chunk_worker(args):
             raise ValueError("divide by zero in calc_B_chunk_worker")
         x = a/(b*rf**2 + c*rf + d)
         assert(not np.any(np.isnan(x)))
-        B = np.einsum('ts,w->wt', x, mut_grid)
+        B = np.einsum('ts,w,sf->wtf', x, mut_grid, F)
         # the einsum below is for when a features dimension exists, e.g.
         # there are feature-specific μ's and t's -- commented out now...
         #B = np.einsum('ts,w,sf->wtf', x, mut_grid,
@@ -193,7 +193,7 @@ def calc_B_parallel(genome, mut_grid, step, nchunks=1000, ncores=2):
 
 def calc_BSC16_chunk_worker(args):
     # ignore rbp, no need here yet under this approximation
-    map_positions, chrom_seg_mpos, segment_parts, mut_grid = args
+    map_positions, chrom_seg_mpos, F, segment_parts, mut_grid = args
     Bs = []
     # F is a features matrix -- eventually, we'll add support for
     # different feature annotation class, but for now we just fix this
@@ -208,7 +208,8 @@ def calc_BSC16_chunk_worker(args):
         # L = seg_L[idx]
         x = bgs_segment_from_parts_sc16(segment_parts, rf, log=True)
         assert(not np.any(np.isnan(x)))
-        B = np.sum(x, axis=2)
+        #B = np.sum(x, axis=2)
+        B = np.einsum('wts,sf->wtf', x, F)
         # the einsum below is for when a features dimension exists, e.g.
         # there are feature-specific μ's and t's -- commented out now...
         #B = np.einsum('ts,w,sf->wtf', x, mut_grid,

### 2022-07-09 bgspy/genome.py

@@ -20,14 +20,17 @@ class Segments:
         - index, a dict of chrom->indices of the ranges array, which contains
           info for all chroms
     """
-    ranges: np.ndarray
-    rates: np.ndarray
-    map_pos: np.ndarray
-    features: np.ndarray
-    feature_map: dict
-    index: defaultdict
-    _segment_parts: tuple = None
-    _segment_parts_sc16: tuple = None
+    def __init__(self, ranges, rates, map_pos, features, feature_map, index):
+        self.ranges = ranges
+        self.rates = rates
+        self.map_pos = map_pos
+        self.features = features
+        self.feature_map = feature_map
+        self.index = index
+        self._segment_parts = None
+        self._segment_parts_sc16 = None
+        self.F = None
+        self._calc_features()
 
     def __repr__(self):
         nfeats = len(self.feature_map)

### 2022-07-09 bgspy/likelihood.py

@@ -3,6 +3,7 @@ from collections import Counter
 import numpy as np
 from scipy.stats import binned_statistic
 from scipy import interpolate
+from bgspy.utils import bin_midpoints
 
 def loglik2(pi0, B, y):
     # y is nsame, ndiff as cols
@@ -54,14 +55,6 @@ def loglik_pi0(B, y):
         return (-ll.sum(axis=0)).min()
     return obj
 
-#def loglike_deriv_pi0(B, y, minimize=True):
-#    R = np.exp(B).squeeze()
-#    nD = y[:, 1, None, None]
-#    nS = y[:, 0, None, None]
-#    a = -1 if minimize else 1
-#    fun = lambda x: a*(nS/x - R * nD / (1-R*x)).sum(axis=0)
-#    return fun
-
 def num_nonpoly(neut_pos, bins, masks):
     """
     Get the number of non-polymorphic sites from
@@ -79,23 +72,30 @@ def num_nonpoly(neut_pos, bins, masks):
     return nfixed
 
 
-def calc_loglik_components(b, Y, neut_pos, neut_masks, nchroms):
+def calc_loglik_components(b, Y, neut_pos, neut_masks, nchroms, chrom_bins):
     """
-    Interpolate the B values at midpoints, and sum the
-    number of same and different pairs of these B windows. Also computes
-    pi in the windows.
+    The MLE for parameters is always calculated per-window (here the window bins
+    defined in chrom_bins). This function aggregates the site-level data (the
+    number of diff/same sites in Y), counts the number of neutral sites that
+    could be polymorphic form the masks (neut_masks), and interpolates the Bs
+    at the window midpoints. It also computes windowed π in these windows from
+    the data, as a check.
+
+    b: BScores object
+    Y: DAC matrix
+    neut_pos: neutral region masks
+    nchroms: how many chromosomes were sequenced
+    bins: dictionary of window endpoints
     """
     chroms = b.pos.keys()
-    # interpolate B at the midpoints of the steps
-    interpol = {c: interpolate.interp1d(b.pos[c], b.B[c], copy=False,
-                                        assume_sorted=True, axis=0)
-                for c in chroms}
-    midpoints = {c: 0.5*(b.pos[c][1:]+b.pos[c][:-1]) for c in chroms}
-    midpoint_Bs = {c: interpol[c](m) for c, m in midpoints.items()}
+
+    win_midpoints = bin_midpoints(chrom_bins)
[…]

### 2022-07-09 bgspy/models.py

@@ -38,7 +38,7 @@ import allel
 from scipy.optimize import minimize_scalar
 import tensorflow as tf
 
-from bgspy.utils import bin_chrom, genome_emp_dists
+from bgspy.utils import bin_chrom, bin_chroms, genome_emp_dists
 from bgspy.utils import readfile, load_dacfile
 from bgspy.utils import load_seqlens, make_dirs
 from bgspy.utils import ranges_to_masks, sum_logliks_over_chroms
@@ -190,19 +190,28 @@ class BGSModel(object):
             binned_B[chrom] = BinnedStat(means, wins, nitems)
         return binned_B
 
-    def loglikelihood(self, pi0=None, pi0_bounds=None, pi0_grid=None, method='SC16'):
+    def loglikelihood(self, width, pi0=None, pi0_bounds=None, pi0_grid=None,
+                      method='SC16'):
         if method == 'SC16':
             b = self.BpScores
         elif method == 'classic':
             b = self.BScores
         else:
             raise ValueError("method must be either 'SC16' or 'classic'")
-        Y_binned, midpoint_Bs, pi_win = calc_loglik_components(b, self.Y, self.neut_pos, self.neut_masks, self.nchroms)
+        assert b is not None, "BScores are not calculated!"
+
+        # create the windows that are the unit of analysis
+        chrom_bins = bin_chroms(self.genome.seqlens, width)
+        comps = calc_loglik_components(b, self.Y, self.neut_pos, self.neut_masks,
+                                     self.nchroms, chrom_bins)
+        Y_binned, midpoint_Bs, pi_win = comps
+
         # for pi0, merge all chromosomes
         bs = np.stack(list(*midpoint_Bs.values()), axis=0)
         ys = np.stack(list(*Y_binned.values()), axis=0)
         chroms = list(b.pos.keys())
         nset = sum([x is not None for x in (pi0, pi0_bounds, pi0_grid)])
+
         if nset > 1:
             raise ValueError("no more than one pi0, pi0_grid, pi0_bounds can be set")
         if nset == 0:
@@ -223,13 +232,13 @@ class BGSModel(object):
             self.pi0_ll = lls_mat
             self.pi0i_mle = max_idx[0]
             self.pi0_grid = pi0_grid
+
+        # main likelihood routine
         ll = loglik(pi0, bs, ys)
         lls = ll.sum(axis=0)
-        self.wi_mle, self.ti_mle = np.where(lls == np.nanmax(lls))
-        self.pi_win = pi_win
[…]

### 2022-07-09 bgspy/parallel.py

@@ -75,7 +75,7 @@ class MapPosChunkIterator(object):
         L = segments.lengths
         chrom_seg_L = {c: share_array(L[idx]) for c, idx in chrom_idx.items()}
         # group by chrom and share features matrix
-        chrom_features = {c: share_array(segments.features[idx]) for c, idx
+        chrom_features = {c: share_array(segments.F[idx]) for c, idx
                           in chrom_idx.items()}
         self.chrom_idx = chrom_idx
         # this is the main thing iterated over -- a generator over the
@@ -181,6 +181,7 @@ class BChunkIterator(MapPosChunkIterator):
             idx = self.chrom_idx[chrom]
             # share the following arrays across processes, to save memory
             # (these should not be changed!)
+            F = self.genome.segments.F[idx, :]
             if use_SC16:
                 # these are all segment-specific
                 segparts = tuple(x[:, :, idx] for x in segment_parts)
@@ -211,6 +212,7 @@ class BChunkIterator(MapPosChunkIterator):
         chrom_segparts = self.chrom_segparts.get(chrom, None)
         return (mpos_chunk,
                 self.chrom_seg_mpos[chrom],
+                self.chrom_features[chrom],
                 chrom_segparts,
                 self.w_grid)
 

### 2022-07-09 bgspy/utils.py

@@ -320,6 +320,14 @@ def sum_logliks_over_chroms(ll_dict):
     ll = np.stack(list(*ll_dict.values())).sum(axis=0)
     return ll
 
+def bin_midpoints(chrom_bins):
+    "Get the midpoint of all bins"
+    return {c: 0.5*(x[1:] + x[:-1]) for c, x in chrom_bins.items()}
+
+def bin_chroms(seqlens, width, dtype='uint32'):
+    "Bin all chromosomes and put the results in a dictionary."
+    return {c: bin_chrom(seqlens[c], width, dtype) for c in seqlens}
+
 def bin_chrom(end, width, dtype='uint32'):
     """
     Bin a chromsome of length 'end' into bins of 'width', with the last

### 2022-07-09 notes/old_ml_notes.tex
theory notes

--- 2022-06-05 notebooks/theory_notes.tex
@@
   \log(\hat{B}(v)) &= \log\left(\prod_i^n (B_i + \epsilon_i) \right) \\
-                   &= \sum_i^n \log(B_i + \epsilon_i). \\
+                   &= \sum_i^n \log(B_i + \epsilon_i).
 \end{align}
 
-If we approximate the expected log with a Taylor series,
+Assume that the noise terms $\epsilon_i \sim N(b_i, \sigma^2)$, where $b_i$ is
+some fixed bias. Then, if we approximate the log with a Taylor series around
+$b_i$,
 
-\begin{align}
-  \E[\log(B_i + \epsilon_i)^2] &\approx \log(B_i)^2 + \frac{2 \log(B_i)}{B_i} \E[\epsilon_i] + \frac{(1-\log(B_i))}{B_i^2} \E[\epsilon_i^2]\\
-                               &\approx \log(B_i)^2 + \frac{2 \log(B_i)}{B_i}b + \frac{(1-\log(B_i)) }{B_i^2}(\sigma^2 + b^2)
-\end{align}
+\begin{align*}
+  \E[\log(\hat{B}(v))] &= \sum_i^n \E[\log(B_i + \epsilon_i)]  \\
+                       &\approx \sum_i^n \log(B_i + b_i) - \sum_i^n \frac{\sigma^2}{2(B_i + b_i)^2} 
+\end{align*}
 
-Then,
+Then, the bias of $\log(\hat{B}(v))$ is
 
-\begin{align}
-  \E[\log(\hat{B}(v))] &= \sum_i^n \E[\log(B_i + \epsilon_i)]  \\
-                                        &\approx \sum_i^n \log(B_i) + \sum_i^n \frac{b}{B_i} - \sum_i^n \frac{\sigma^2 + b^2}{2B_i^2}  \\
-\end{align}
+\begin{align*}
+  \E[\log(\hat{B}(v))] - \log(B(v)) &\approx \sum_i^n \log(B_i + b_i) - \sum_i^n \frac{\sigma^2}{2(B_i + b_i)^2} - \sum_i^n \log(B_i) \\
+  \E\left[\log\left(\frac{\hat{B}(v)}{B(v)}\right)\right]   &\approx \sum_i^n \log\left(\frac{B_i + b_i}{B_i}\right) - \sum_i^n \frac{\sigma^2}{2(B_i + b_i)^2} 
+\end{align*}
 
-Bias:
+For now, assume the training is not biased, e.g. $b=0$. Then, noise around the
+true $B_i$ values creates negative log bias,
 
-\begin{align}
-  \E[\log(\hat{B}(v))] - \E[\log(B(v))] &= \sum_i^n \E[\log(B_i + \epsilon_i)] - \sum_i^n \log(B_i) \\
-                                        &\approx \sum_i^n \log(B_i) + \sum_i^n \frac{b}{B_i} - \sum_i^n \frac{\sigma^2 + b^2}{2B_i^2} - \sum_i^n \log(B_i) \\
-                                        &\approx \sum_i^n \frac{b}{B_i} - \sum_i^n \frac{\sigma^2 + b^2}{2B_i^2} 
-\end{align}
+\begin{align*}
+  \E\left[\log\left(\frac{\hat{B}(v)}{B(v)}\right)\right]   &\approx - \sum_i^n \frac{\sigma^2}{2B_i^2} \le 0.
+\end{align*}
 
-By Jensen's inequality $f(\E(X)) \le \E(f(X))$ for some convex function $f$,
+Thus, any noise in the predictions will lead to $\hat{B}(v)$ values that are a
+fraction of the true value. Furthermore, by Jensen's inequality,
+
+\begin{align*}
+  \log\left(\E\left[\frac{\hat{B}(v)}{B(v)}\right]\right)  &\le \E\left[\log\left(\frac{\hat{B}(v)}{B(v)}\right)\right]    \\
+  \E\left[\frac{\hat{B}(v)}{B(v)}\right]  &\le \exp \left(- \sum_i^n \frac{\sigma^2}{2B_i^2} \right) \\
+  \E[\hat{B}(v)]  &\le B(v) \exp \left(- \sum_i^n \frac{\sigma^2}{2B_i^2} \right)
+\end{align*}
 

### 2022-07-09 notes/theory.tex
theory notes

--- 2022-05-30 notebooks/theory_notes.tex
@@
 
-\title{Learning Theory Notes}
+\title{Statistical Methods}
 
@@
 
-\section*{Learning Limits}
+\section*{Likelihood}
 
-\subsection*{The Simulation Sample $\bar{B}$ Estimator}
-
-Our goal is to approximate a function that describes the reduction in pairwise
-diversity (relative to the neutral expectation) $B$ due to BGS for some set of
-evolutionary parameters $\mathbf{x}$. The true reduction function is a
-conditional expectation $B(\mathbf{x}) = \E[T_2(\mathbf{x}) | \mathbf{x}]$
-taken over an infinite number of evolutionary replicates. Here
-$T_2(\mathbf{x})$ is the pairwise coalescent rate under BGS for parameters
-$\mathbf{x}$, with time scaled in $2N$ generation units.
-
-We approximate this function $B(\mathbf{x})$ with $\widehat{B}(\mathbf{x})$
-from evolutionary simulations using statistical learning. In each evolutionary
-simulation, we sample some parameter set $\mathbf{x}$ and evolve $r$
-independent populations forward in time and observe the resulting genealogy at
-a neutral site a given recombination distance away (this is one of the
-parameters). From this full population genealogy of $2N$ gametes, we estimate
-the reduction in neutral diversity as
+The likelihood in Eyalshiv et al. (2016) has the form,
 
 \begin{align}
-  \bar{B} = \frac{1}{r} \sum_{i=1}^r \frac{\hat{\pi}_{i}}{4N\mu}
+  \log\mathcal{L} = \sum_{v \in \mathcal{V}} \sum_{i \ne j \in \mathcal{S}} \log(P(O_{i,j}(v) | \theta))
 \end{align}
-%
-where $\hat{\pi}$ is Tajima's estimator for pairwise diversity within a tree.
-For now, we can imagine letting the neutral mutations saturate as $\mu \to 1$
-so we can ignore mutation rate hereafter. If we take expectation over the evolutionary
-process,
+
+where $\mathcal{V}$ is the set of putatively neutral sites, $\mathcal{S}$ is
+the set of samples, and $\theta$ are the BGS parameters. The indicator variable
+$O_{i,j}(v)$ is 1 if samples $i$ and $j$ are different at site $v$, and zero
+otherwise. Thus as they specify in the paper,
 
 \begin{align}
-  \E[\bar{B}] &= \frac{1}{4Nr} \sum_{i=1}^r \E[\hat{\pi}_{i}] \\
-                          &= \frac{2T_2}{4N}  \\
-                          %&= \frac{4B(\mathbf{x})N}{4N}  \\
-                          &= B
+  P(O_{i,j}(v) | \theta) = 
+    \begin{cases}
+      \pi(v | \theta), & O_{i,j}(v) = 1 \\
+      1-\pi(v | \theta), & O_{i,j}(v) = 0 \\
+    \end{cases}.
 \end{align}
-%
-thus, the estimated reduction in diversity from simulations $\bar{B}$ is
-unbiased, since $\E[\hat{\pi}] = 2T_2 = 4BN$.
 
-\subsection*{The variance of the estimator $\bar{B}$}
-
-Using the law of total variance, we can write the variance of $\bar{B}$ as
+The total size of the set of samples $\mathcal{S}$ is $n_\mathcal{S} =
+|\mathcal{S}|$. Assuming all sites are biallelic, we can simplify the
+calculation of the likelihood by noting that for a site $v$'s vector of allele
+counts $[c_1, c_2]$, the total number of pairwise combinations with the same
+alleles is
 
 \begin{align}
-  \var(\bar{B}) = \underbrace{\E(\var(\bar{B} | \mathcal{G}))}_\text{sampling noise} + \underbrace{\var(\E(\bar{B} | \mathcal{G}))}_\text{evolutionary variance}
+  n_s(v) = {c_1 \choose 2} + {c_2 \choose 2}
 \end{align}
 
-where $\mathcal{G}$ is the genealogy at the focal neutral site affected by BGS
-at some segment. The distribution of genealogies $P(\mathcal{G})$ is unknown
-and quite complex under BGS, but we can use a simplifying assumption to get a
-rough estimate of the variance of $\bar{B}$.
-
-Under strong BGS (e.g. $\nicefrac{\mu}{s}N \gg 1$), the genealogy is often
-modeled as a neutral coalescent process with rescaled with an effective
-population size $N_e = BN$ (though see Cvijovi\'{c} et al. 2018). In the weak
-selection limit $s \to 0$, approximation becomes exact as BGS becomes
-effectively neutral. In both regimes, selection is treated as rescaling a
-neutral genealogical process to $N_e = BN$. We refer to this the
-\textbf{neutral-BGS model}, and it allows for us to approximate $\var(\bar{B})$
-with some $\var(\bar{B}')$ where $\bar{B}'$ is the mean reduction if we
-averaged over neutral genealogies from a population of $BN$
-diploids, rather than averaged over exact selection simulations. Note that at
-$B=1$, the neutral-BGS model becomes exact, as there is no reduction in
-diversity due to background selection and the only evolutionary variance is
-through the neutral coalescent process.
-
-If the genealogies are neutral, we can calculate the variance of $\bar{B}$
-using Tajima's equation for the variance of $\hat{\pi}$. For $n$ samples and
-$r$ evolutionary replicates,
+and the number of different pairwise combinations is
 
 \begin{align}
-  \var(\bar{B}') &= \frac{1}{16 N^2 r^2} \sum_{i=1}^r \var(\hat{\pi}_{i}) \\
-                &= \frac{\var(\hat{\pi}_{i})}{16 N^2 r} \\
-                  &= \frac{1}{16N^2 r} \left( \frac{n + 1}{3(n-1)}\theta  + \frac{2(n^2 + n + 3)}{9n(n-1)}\theta^2 \right) \\
-                  &= \frac{1}{16N^2 r} \left( \frac{n + 1}{3(n-1)} 4BN + \frac{2(n^2 + n + 3)}{9n(n-1)} 16B^2N^2 \right) \\
-                  &= \underbrace{\frac{n + 1}{12(n-1)} \frac{B}{N r}}_\text{sampling noise} + \underbrace{\frac{2(n^2 + n + 3)}{9n(n-1)} \frac{B^2}{r}}_\text{evolutionary variance}
+  n_d(v) = n_T - n_s(v)
 \end{align}
 
-Now, let us look at the consistency of the estimator $\bar{B}$ (we drop the
-parameter set $k$ for clarity) both in $r$ (over evolutionary replicates) and
-in $n$ (as the sample size increases). Let $\bar{B}_r$ be the estimator of $B$
-after $r$ evolutionary replicates (conditioning on some $n$). By Chebyshev's
-inequality and for some $\epsilon > 0$,
+where $n_T = n_\mathcal{S} (n_\mathcal{S} - 1) / 2$  is the total number of
+pairwise combinations across the sample set $\mathcal{S}$. Furthermore, note
+that we can partition the set $\mathcal{V}$ of neutral sites into polymorphic
+($\mathcal{P}$) and fixed sites ($\mathcal{F}$), i.e. $\mathcal{V} =
+\mathcal{P} \cup \mathcal{F}$ and $\mathcal{P} \cap \mathcal{F} = \varnothing$.
+For all $v \in \mathcal{F}$, $n_d(v) = 0$ and $n_s(v) = n_T$.
+
+Then,
 
 \begin{align}
-  \P\left(|\bar{B}_r - B| \ge \epsilon \right)  \le \frac{\var(\bar{B}_r)}{\epsilon^2}
-\end{align}
-%
-and since $\lim_{r \to \infty} \var(\bar{B}_r) = 0$,
-
-\begin{align}
-  \lim_{r \to \infty} \P\left(|\bar{B}_r - B| \ge \epsilon \right)  = 0.
+  \log\mathcal{L} &= \sum_{v \in \mathcal{V}} \left[\log(\pi(v | \theta)) n_\text{D}(v) + \log(1-\pi(v | \theta)) n_\text{S}(v)\right] \\
+                  &= \sum_{v \in \mathcal{P}} \left[\log(\pi(v | \theta)) n_\text{D}(v) + \log(1-\pi(v | \theta)) n_\text{S}(v)\right] + \sum_{v \in \mathcal{F}} \log(1-\pi(v | \theta)) n_T  \\
 \end{align}
 
-Thus, $\bar{B}_r \xrightarrow{p} B$ as $r \to \infty$ and $\bar{B}_r$ is
-consistent in $r$. Now, let us look at the consistency of $\bar{B}_n$ in sample
-size $n$. Note that $n \le 2N$ (i.e. our sample size is bounded by the number
-of gametes in the population), so we imagine setting $n = 2N$ and taking the
-limit $N \to \infty$,
-
-\begin{align}
-  \lim_{N \to \infty} \P\left(|\bar{B}_N - B| \ge \epsilon \right)  &\le \lim_{N \to \infty} \frac{\var(\bar{B}_N)}{\epsilon^2} \\
-  \lim_{N \to \infty} \P\left(|\bar{B}_N - B| \ge \epsilon \right)  &\le \frac{2B^2}{9 r \epsilon^2}. \label{eq:bound}
-\end{align}
-
-Thus, even if we sample the entire population, and let the population size $N
-\to \infty$, $\bar{B}_n$ is still an inconsistent estimator in $n$. Intuitively
-this is because the evolutionary variance is not decreased by taking more
-samples, \emph{only by taking more evolutionary replicates}.
-
-Below I work through where this bound came from. If we estimate pairwise
-diversity from the genealogical tree of the entire population such that $n =
-2N$,
-
-\begin{align}
-  \var(\bar{B}) &= \frac{n + 1}{12(n-1)} \frac{B}{N r} + \frac{2(n^2 + n + 3)}{9n(n-1)} \frac{B^2}{r} \\
-                &= \frac{2N + 1}{12(2N-1)} \frac{B}{N r} + \frac{2(4N^2 + 2N + 3)}{18N(2N-1)} \frac{B^2}{r} \\
-                &\approx \frac{B}{12N r} + \frac{2N^2 + N}{9N^2} \frac{B^2}{r}
-\end{align}
-
-If we take the infinite population size limit,
-
-\begin{align}
-  \lim_{N \to \infty} \var(\bar{B}) &= \frac{2B^2}{9r}
-\end{align}
-
-In practice, we use a different, better estimator of pairwise diversity than
-Tajima's $\pi$: branch statistic pairwise diversity (Ralph 2019, Ralph et al.
-2020). This estimator removes mutation as a source of sampling noise and since
-$n = 2N$, our only source of variance in $\bar{B}$ is in the evolutionary
-process.
-
-% Note that since the bound in Equation \eqref{eq:bound} is a probability, it
-% must be that $\frac{2B^2}{9 r \epsilon^2} \le 1$. Thus this equation is limited
-% to considering bounds $\epsilon$ that are
-
-% \begin{align}
-%   \sqrt{\frac{2}{9r}} B \le \epsilon \\
-% \end{align}
-
-% For $B \approx 1$, $\sqrt{\frac{2}{9r}} \le \epsilon$, such that if $r = 50$,
-% we can only consider bounds $\epsilon \ge \sqrt{\nicefrac{1}{225}} \approx
-% 0.067$ because Chebyshev isn't a particularly sharp inequality.
-
-\subsection*{Partitioning the Prediction Error}
-
-To learn the function $B(\mathbf{x})$, we run evolutionary simulations under a
-training set of parameters, giving us $\{(\mathbf{x}_1, \bar{B}_1),
-(\mathbf{x}_2, \bar{B}_2), \ldots, (\mathbf{x}_n, \bar{B}_n)\}$, where each
-$\bar{B}_i$ is the observed reduction over $r$ replicates. Since our goal is to
-approximate $\hat{B}(\mathbf{x})$ as closely as possible to $B(\mathbf{x})$
-over the distribution of $\mathbf{x}$ most close to that across the genome, we
-find some $\hat{B}(\mathbf{x})$ that minimizes expected loss. An important
-metric, even if training is conducted with a different loss, is the mean
-squared error
-
-\begin{align}
-  MSE(\bar{B}, \hat{B}) = \E[(\bar{B}(\mathbf{x}) - \hat{B}(\mathbf{x}))^2].
-\end{align}
-
-We can think of $\bar{B}(\mathbf{x})$ as a true reduction $B(\mathbf{x})$ plus
-a deviation $\varepsilon$ due to the particular evolutionary replicates
-simulated, e.g. $\bar{B}(\mathbf{x}) = B(\mathbf{x}) + \varepsilon$.
-
-\begin{align}
-  MSE(\bar{B}, \hat{B}) &= \E[(\bar{B}(\mathbf{x}) - \hat{B}(\mathbf{x}))^2] \\
-                        &= \E[(B(\mathbf{x}) + \varepsilon - \hat{B}(\mathbf{x}))^2] \\
-                        &= \underbrace{\var(\hat{B}(\mathbf{x}))}_\text{training variance} + 
-                      \underbrace{\var(\varepsilon)}_\text{evolutionary variance} + \underbrace{\E[\hat{B}(\mathbf{x}) - B(\mathbf{x})]^2}_{\mathrm{bias}^2}
-\end{align}
-
-At $B=1$, the evolutionary variance is exactly the neutral variance given by
-$\var(\varepsilon) = \nicefrac{2B^2}{9r}$. Thus, MSE loss has an absolute bound at
-$B=1$,
-
-\begin{align}
-  MSE(\bar{B}, \hat{B} | B = 1) \ge \frac{2}{9r}
-\end{align}
-
-For $B < 1$, we can write the evolutionary variance as some deviation
-away from the coalescent noise under the neutral-BGS model $\var(\varepsilon) =
-\var(\bar{B}') + e$.
-
-% For a particular training set
-% $\mathcal{T} = \{\mathbf{X}_\text{train}, \mathbf{B}_\text{train}\}$, the mean
-% squared error is
-
-\printbibliography
+\section{B Scores}
 

### 2022-07-10 bgspy/likelihood.py
new likelihood testing, some changes

@@ -57,22 +57,35 @@ def loglik_pi0(B, y):
 
 def num_nonpoly(neut_pos, bins, masks):
     """
-    Get the number of non-polymorphic sites from
+    Get the number of non-polymorphic sites that overlap the neutral regions
+    specified by the boolean masks. This is calculated as:
+        # neutral masked sites in bin - # polymorphic sites in bins
+
+    neut_pos: chrom dict of positions of the neutral polymorphic sites
+    bins: chrom dict of bin end positions
+    masks: chrom dict of neutral regions as boolean masks
     """
     chroms = bins.keys()
+
+    # validate the neutral sites
+    for chrom in chroms:
+        for pos in neut_pos[chrom]:
+            msg = f"Position {pos} is not in the '{chrom}' neutral mask!"
+            assert masks[chrom][pos] == 1, msg
+
     # find window indices of all neutral SNPs
     idx = {c: np.digitize(neut_pos[c], bins[c])-1 for c in chroms}
     # count the indices (SNPs) per window
     poly_counts = {c: Counter(idx[c].tolist()) for c in chroms}
     npoly = {c: np.array([poly_counts[c][i] for i, _ in enumerate(e)]) for c, e in bins.items()}
-    # what's the width of the neutral regions
+    # what's the width of the neutral regions overlapping the bins?
     widths = {c: [masks[c][a:b].sum() for a, b in zip(e[:-1], e[1:])] for c, e in bins.items()}
     #__import__('pdb').set_trace()
-    nfixed = {c: widths[c]-npoly[c][1:] for c in bins.keys()}
+    nfixed = {c: widths[c]-npoly[c][:-1] for c in bins.keys()}
     return nfixed
 
 
-def calc_loglik_components(b, Y, neut_pos, neut_masks, nchroms, chrom_bins):
+def calc_loglik_components(b, Y, Y_pos, neut_masks, nchroms, chrom_bins):
     """
     The MLE for parameters is always calculated per-window (here the window bins
     defined in chrom_bins). This function aggregates the site-level data (the
@@ -83,7 +96,7 @@ def calc_loglik_components(b, Y, neut_pos, neut_masks, nchroms, chrom_bins):
 
     b: BScores object
     Y: DAC matrix
-    neut_pos: neutral region masks
+    Y_pos: neutral region masks
     nchroms: how many chromosomes were sequenced
     bins: dictionary of window endpoints
     """
[…]

### 2022-07-10 tests/test_learnclasses.py

@@ -1,132 +1,0 @@
-import copy
-import pytest
-import numpy as np
-from bgspy.learn import LearnedB, LearnedFunction
-from bgspy.theory import bgs_segment
-
-# this is a test case from a random chunk of chr10 from predictions
-# fed into the ML B approach. Columns are mu, s, L, rbp, rf --
-# the same order as expected (specified in bgspy.theory.BGS_MODEL_PARAMS
-Xp = np.array([[1.0000e-09, 1.0000e-03, 3.3400e+02, 2.8873e-10, 4.7425e-04],
-               [1.0000e-09, 1.0000e-04, 8.0900e+02, 1.6206e-08, 5.1120e-02],
-               [1.0000e-08, 1.0000e-03, 1.0000e+03, 1.8624e-09, 1.2862e-02],
-               [1.0000e-08, 1.0000e-02, 1.0000e+03, 2.0333e-08, 6.3086e-02],
-               [1.0000e-08, 1.0000e-04, 1.9500e+02, 4.8172e-10, 1.8595e-02],
-               [1.0000e-08, 1.0000e-02, 1.0000e+03, 3.3601e-08, 4.6281e-02],
-               [1.0000e-09, 1.0000e-04, 8.4100e+02, 5.5386e-09, 2.9605e-02],
-               [1.0000e-09, 1.0000e-02, 4.4400e+02, 1.7687e-10, 4.7748e-02],
-               [1.0000e-09, 2.0000e-02, 2.5500e+02, 6.1183e-09, 5.3264e-02],
-               [1.0000e-08, 1.0000e-03, 1.0000e+03, 1.5244e-10, 1.1342e-02],
-               [1.0000e-07, 1.0000e-02, 6.7600e+02, 9.1097e-09, 1.0493e-02],
-               [1.0000e-09, 1.0000e-03, 3.3000e+01, 1.2818e-08, 9.3163e-02],
-               [1.0000e-09, 1.0000e-04, 1.0000e+03, 5.4405e-09, 4.2521e-02],
-               [1.0000e-08, 2.0000e-02, 2.7700e+02, 1.7699e-08, 3.3569e-02],
-               [1.0000e-07, 1.0000e-04, 1.0000e+03, 1.6668e-09, 5.1079e-02],
-               [1.0000e-07, 1.0000e-02, 1.6000e+01, 1.8927e-08, 7.3314e-03],
[…]

### 2022-07-10 tests/test_likelihood.py

@@ -1,0 +1,23 @@
+import numpy as np
+from bgspy.likelihood import num_nonpoly, calc_loglik_components
+
+
+def test_num_nonpoly():
+    #
+    neut_pos = {'chr1': np.array([ 0,  3,  5, 11])}
+    bins = {'chr1': np.array([ 0,  4,  8, 12])}
+    #                          0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11
+    #                                     |           |
+    # poly sites:              x        x     x                  x
+    masks = {'chr1': np.array([1, 0, 1, 1, 1, 1, 0, 0, 1, 1,  0, 1])}
+    # number of poly sites            2   |     1     |      1
+    # overlapping the neutral region masks
+    # --
+    # nfixed:                         1   |     1      |     2
+    # (not poly and overlapping neutral)
+
+    #res = {'chr1': np.array([1, 1, 2])}
+    res = np.array([1, 1, 2])
+    np.testing.assert_allclose(num_nonpoly(neut_pos, bins, masks)['chr1'], res)
+
+#def test_calc_loglik_components():

### 2022-07-10 tests/test_predict.py

@@ -1,180 +1,0 @@
-import os
-import json
-import pytest
-from collections import defaultdict
-from itertools import product
-import click
-from click.testing import CliRunner
-import shutil
-import tensorflow as tf
-import numpy as np
-
-from bgspy.predict import new_predict_matrix, inject_rf, predictions_to_B_tensor
-from bgspy.predict import predict_chunk
-from bgspy.learn import LearnedFunction
-from bgspy.theory import bgs_segment
-
-NW, NT = 2, 3
-NSEGS = 3
-
-@pytest.fixture
-def X():
-    w, t = np.arange(NW), np.arange(NT)
-    L = np.array([10, 11, 12])
-    rbp = np.array([20, 21, 22])
-    assert len(L) == NSEGS
-    assert len(rbp) == NSEGS
-    X = new_predict_matrix(w, t, L, rbp)
-    return X
-
-def test_new_pred_matrix(X):
-    # expected: repeat each (μ, s) pair nsegs times each
-    # with the segment stuff (L, rbp) filled in cols 2 and 3
-    X_expected = np.array([[ 0.,  0., 10., 20.,  0.],
-                           [ 0.,  0., 11., 21.,  0.],
-                           [ 0.,  0., 12., 22.,  0.],
-                           [ 0.,  1., 10., 20.,  0.],
-                           [ 0.,  1., 11., 21.,  0.],
-                           [ 0.,  1., 12., 22.,  0.],
-                           [ 0.,  2., 10., 20.,  0.],
-                           [ 0.,  2., 11., 21.,  0.],
-                           [ 0.,  2., 12., 22.,  0.],
-                           [ 1.,  0., 10., 20.,  0.],
-                           [ 1.,  0., 11., 21.,  0.],
-                           [ 1.,  0., 12., 22.,  0.],
-                           [ 1.,  1., 10., 20.,  0.],
-                           [ 1.,  1., 11., 21.,  0.],
-                           [ 1.,  1., 12., 22.,  0.],
-                           [ 1.,  2., 10., 20.,  0.],
-                           [ 1.,  2., 11., 21.,  0.],
-                           [ 1.,  2., 12., 22.,  0.]])
-    np.testing.assert_equal(X[:, :4], X_expected[:, :4])
-
-def test_inject_rf(X):
-    rf = np.array([1, 2, 3])
-    Xrf = inject_rf(rf, X, NW * NT)
-    Xrf_expected = np.array([[ 0.,  0., 10., 20.,  1.],
[…]

### 2022-07-10 notes/theory.tex
more note updates

@@
 \renewcommand{\P}{\mathbb{P}}
+\renewcommand{\vec}[1]{\mathbf{#1}}
 \newcommand{\E}{\mathbb{E}}
@@
 \begin{align}
-  \log\mathcal{L} = \sum_{v \in \mathcal{V}} \sum_{i \ne j \in \mathcal{S}} \log(P(O_{i,j}(v) | \theta))
+  \log\mathcal{L}(\theta) = \sum_{v \in \mathcal{V}} \sum_{i \ne j \in \mathcal{S}} \log(P(O_{i,j}(v) | \theta))
 \end{align}
@@
 \begin{align}
-  \log\mathcal{L} &= \sum_{v \in \mathcal{V}} \left[\log(\pi(v | \theta)) n_\text{D}(v) + \log(1-\pi(v | \theta)) n_\text{S}(v)\right] \\
-                  &= \sum_{v \in \mathcal{P}} \left[\log(\pi(v | \theta)) n_\text{D}(v) + \log(1-\pi(v | \theta)) n_\text{S}(v)\right] + \sum_{v \in \mathcal{F}} \log(1-\pi(v | \theta)) n_T  \\
+  \log\mathcal{L}(\theta) &= \sum_{v \in \mathcal{V}} \left[\log(\pi(v | \theta)) n_\text{D}(v) + \log(1-\pi(v | \theta)) n_\text{S}(v)\right] \\
+                  &= \sum_{v \in \mathcal{P}} \left[\log(\pi(v | \theta)) n_\text{D}(v) + \log(1-\pi(v | \theta)) n_\text{S}(v)\right] + \sum_{v \in \mathcal{F}} \log(1-\pi(v | \theta)) n_T(v)  \\
 \end{align}
+
+In practice, we calculate these values across bins. For each bin, we treat
+$\pi(v | \theta)$ as fixed, assuming that at this scale, the variation in
+expected diversity across sites is minimal. For a particular chromosome, we
+have two classes of sites: those included in the diversity calculation and
+those ignored. The former sites are all putatively neutral and have reliably
+called genotypes, and the other sites are possibly non-neutral or do have
+reliably called genotypes. The log-likelihood is the sum of bin
+likelihoods, $\mathcal{L}(b)$
+
+\begin{align}
+  \log\mathcal{L}(b | \theta)  &= \log(\bar{\pi}(b | \theta)) \sum_{v \in \mathcal{P}} n_\text{D}(v) + \log(1-\bar{\pi}(b | \theta)) \left(\sum_{v \in \mathcal{V}} n_\text{S}(v) +  \sum_{v \in \mathcal{F}} n_T(v)  \right).
+\end{align}
+
+We can assume that the total number of combinations at each fixed site is
+constant, e.g. $n_T = n_T(v)$ for all $v$, simplifying this as
+
+\begin{align}
+  \log\mathcal{L}(b | \theta)  &= \log(\bar{\pi}(b | \theta)) \sum_{v \in \mathcal{P}} n_\text{D}(v) + \log(1-\bar{\pi}(b | \theta)) \left(\sum_{v \in \mathcal{V}} n_\text{S}(v) +  n_T(v) f(b)  \right)
+\end{align}
+
+where $f(b)$ is the number of fixed sites in bin $b$. Thus, we our data can be
+reduced within a bin by these terms summed over sites in the window,
+
+\begin{align}
+  \log\mathcal{L}(b | \theta)  &= \log(\bar{\pi}(b | \theta)) Y_D(b) + \log(1-\bar{\pi}(b | \theta)) Y_S(b)
+\end{align}
+
+\begin{align}
+  \log\mathcal{L}(b | \theta)  &= \log(\bar{\pi}_0 B(b | \vec{w}, \vec{t})) Y_D(b) + \log(1-\bar{\pi}_0 B(b |  \vec{w}, \vec{t})) Y_S(b)
+\end{align}
+
+where $\vec{w}$ and $\vec{t}$ are the vectors of mutation rates into
+
+There are two approaches: interpolated $\pi(m | \theta)$ at the midpoint of the
+bin position $m$, and averaged bin-averaged $\bar{\pi}(m | \theta)$.
 

### 2022-07-13 bgspy/data.py
updated genomic data methods

@@ -1,6 +1,6 @@
 import os
 import numpy as np
-from bgspy.utils import read_bed3, ranges_to_masks
+from bgspy.utils import read_bed3, ranges_to_masks, bin_chroms
 
 def load_dacfile(dacfile, neut_masks=None):
     """
@@ -65,6 +65,39 @@ def load_dacfile(dacfile, neut_masks=None):
     ac = np.stack((nchrom - dac, dac)).T
     return positions, indices, ac, position_map, parse_param_str(params)
 
+
+def aggregate_site_array(x, bins, func, **kwargs):
+    """
+    Given a site array (an np.ndarray of length equal to a chromosome)
+    calculate some summary of values with func on the specified bins.
+    """
+    vals = np.zeros((len(bins), x.shape[1]))
+    for i in range(1, len(bins)):
+        vals[i, ...] = func(x[bins[i-1]:bins[i], ...], **kwargs)
+    return vals
+
+
+def pairwise_summary(ac):
+    """
+    Compute the number of pairwise same and different differenes per position,
+    given an allele count matrix.
+
+    Based on the approach in scikit-allel and Nei.
+    """
+    ac = ac.astype(np.int32)
+    an = np.sum(ac, axis=1)
+    n_pairs = an * (an - 1) / 2
+    n_same = np.sum(ac * (ac - 1) / 2, axis=1)
+    n_diff = n_pairs - n_same
+    return np.stack((n_same, n_diff)).T
+
+def pi_from_pairwise_summaries(npairs):
+    """
+    Given a summary of pairwise combinations (nsame, ndiff cols) return
+    π. This assumes site arrays (e.g. the denominator is handled naturally).
+    """
+    return npairs[:, 1] / npairs.sum(axis=1)
+
 class CountsDirectory:
     """
     Lazy load a directory of .npy files of chromosome count data.
@@ -90,12 +123,24 @@ class GenomeData:
         self.counts = None
         self.dac = None
 
+    def _load_mask(self, file):
+        """
+        Load a 3-column BED file into a mask objects.
+        """
+        ranges = read_bed3(file, keep_chroms=set(self.genome.seqlens.keys()))
+        return ranges_to_masks(ranges, self.genome.seqlens)
+
+    def load_accessibile_masks(self, file):
+        """
+        Load a 3-column BED file of accessible ranges into mask objects.
+        """
+        self.accesssible_masks = self._load_mask(file)
+
     def load_neutral_masks(self, file):
         """
[…]

### 2022-07-13 bgspy/genome.py

@@ -237,6 +237,10 @@ class Genome(object):
         self._idx2map = None
         self._map2idx = None
 
+    @property
+    def chroms(self):
+        return list(self.seqlens.keys())
+
     def load_seqlens(self, file, chroms):
         if isinstance(chroms, str):
             chroms = set([chroms])

### 2022-07-13 bgspy/models.py

@@ -39,9 +39,6 @@ from scipy.optimize import minimize_scalar
 import tensorflow as tf
 
 from bgspy.utils import bin_chrom, bin_chroms, genome_emp_dists
-from bgspy.utils import readfile, load_dacfile
-from bgspy.utils import load_seqlens, make_dirs
-from bgspy.utils import ranges_to_masks, sum_logliks_over_chroms
 from bgspy.utils import Bdtype, BScores, BinnedStat
 from bgspy.likelihood import calc_loglik_components, loglik
 from bgspy.classic import calc_B, calc_B_parallel, calc_BSC16_parallel

### 2022-07-13 bgspy/utils.py

@@ -345,16 +345,6 @@ def bin_chrom(end, width, dtype='uint32'):
     assert np.all(bins < end+1)
     return bins
 
-def aggregate_site_array(x, bins, func, **kwargs):
-    """
-    Given a site array (an np.ndarray of length equal to a chromosome)
-    calculate some summary of values with func on the specified bins.
-    """
-    vals = np.zeros((len(bins), x.shape[1]))
-    for i in range(1, len(bins)):
-        vals[i, ...] = func(x[bins[i-1]:bins[i], ...], **kwargs)
-    return vals
-
 def get_unique_indices(x):
     "Get the indices of unique elements in a list/array"
     indices = defaultdict(list)

### 2022-07-13 tools/gvcf2counts.py

@@ -58,7 +58,7 @@ def main(vcf_file, outdir, samples=None, pass_only=True, min_qual=50, min_gq=30)
                 np.save(filepath, counts)
             # create a new counts table
             pbar = tqdm.tqdm(total = seqlens[var.CHROM], unit_scale=True)
-            counts = np.full((seqlens[var.CHROM], 2), -1, dtype=np.short)
+            counts = np.zeros((seqlens[var.CHROM], 2), dtype=np.short)
             ncomplete = 0
             last_chrom = var.CHROM
 
@@ -83,24 +83,24 @@ def main(vcf_file, outdir, samples=None, pass_only=True, min_qual=50, min_gq=30)
         # this is slower: 7k it/sec vs 9k for np
         #gt_counts = Counter(gts)
         #nhom_ref, nhom_alt, nhet = [gt_counts[k] for k in (0, 2, 1)]
-        qual_pass = var.QUAL is not None and var.QUAL >= min_qual
+        qual_pass = int(var.QUAL is not None and var.QUAL >= min_qual)
         nref = 2*nhom_ref + nhet*qual_pass
         nalt = (2*nhom_alt + nhet)*qual_pass
-        stats[QUAL] += ~ qual_pass
+        stats[QUAL] += int(not qual_pass)
         assert nref + 1 < COUNT_MAX, "counts dtype overflow!"
         assert nalt + 1 < COUNT_MAX, "counts dtype overflow!"
         i = var.POS
         counts[i, 0] = nref
         counts[i, 1] = nalt
         pbar.update(i - pbar.n)
-        ncomplete += (counts[i, :] > -1).sum()
+        ncomplete += (counts[i, :] > 0).sum()
 
     # save the last chromosome
     filepath = os.path.join(outdir, f"{last_chrom}_counts.npy")
     np.save(filepath, counts)
-
-    print(stats)
     vcf.close()
+
+    print(f"failed filter: {stats[0]}\nfailed qual (<{min_qual}): {stats[1]})")
 
 
 if __name__ == "__main__":

### 2022-07-13 bgspy/data.py
removed old code, for new LL

@@ -2,6 +2,9 @@ import os
 import numpy as np
 from bgspy.utils import read_bed3, ranges_to_masks, bin_chroms
 
+# error out on overflows
+np.seterr(all='raise')
+
 def load_dacfile(dacfile, neut_masks=None):
     """
     Read derived allele counts datafile. These should be sorted
@@ -91,12 +94,15 @@ def pairwise_summary(ac):
     n_diff = n_pairs - n_same
     return np.stack((n_same, n_diff)).T
 
-def pi_from_pairwise_summaries(npairs):
+def pi_from_pairwise_summaries(pair_data):
     """
     Given a summary of pairwise combinations (nsame, ndiff cols) return
     π. This assumes site arrays (e.g. the denominator is handled naturally).
     """
-    return npairs[:, 1] / npairs.sum(axis=1)
+    denom = pair_data.sum(axis=1)
+    return np.divide(pair_data[:, 1], denom,
+                     out=np.full(denom.shape[0], np.nan),
+                     where=denom > 0)
 
 class CountsDirectory:
     """
@@ -185,17 +191,17 @@ class GenomeData:
             d = pairwise_summary(site_ac)
             reduced[chrom] = aggregate_site_array(d, bins[chrom], np.nansum, axis=0)
 
-        return reduced
+        return bins, reduced
 
     def bin_pi(self, **kwargs):
         """
         Calculate π from the binned summaries.
         """
-        reduced = self.bin_reduce(**kwargs)
+        bins, reduced = self.bin_reduce(**kwargs)
         pi = dict()
         for chrom in reduced.keys():
             pi[chrom] = pi_from_pairwise_summaries(reduced[chrom])
-        return pi
+        return bins, pi
 
 
 

### 2022-07-13 bgspy/likelihood.py

@@ -84,76 +84,59 @@ def num_nonpoly(neut_pos, bins, masks):
     nfixed = {c: widths[c]-npoly[c][:-1] for c in bins.keys()}
     return nfixed
 
-def binned_data_reduction():
-    """
-    Reduces the site-level data into binned data summaries for the binned-
-    likelihood approach.
+# def calc_loglik_components(b, Y, Y_pos, neut_masks, nchroms, chrom_bins):
+#     """
+#     The MLE for parameters is always calculated per-window (here the window bins
+#     defined in chrom_bins). This function aggregates the site-level data (the
+#     number of diff/same sites in Y), counts the number of neutral sites that
+#     could be polymorphic form the masks (neut_masks), and interpolates the Bs
+#     at the window midpoints. It also computes windowed π in these windows from
+#     the data, as a check.
 
-    Currently we don't account in fixed site's depth, we just assume the total
-    number of chromosomes sequenced
-    """
+#     b: BScores object
+#     Y: DAC matrix
+#     Y_pos: neutral region masks
+#     nchroms: how many chromosomes were sequenced
+#     bins: dictionary of window endpoints
+#     """
+#     chroms = b.pos.keys()
 
+#     win_midpoints = bin_midpoints(chrom_bins)
+#     win_Bs = {c: b.B_at_pos(c, x) for c, x in win_midpoints.items()}
 
-def calc_loglik_components(b, Y, Y_pos, neut_masks, nchroms, chrom_bins):
-    """
-    The MLE for parameters is always calculated per-window (here the window bins
-    defined in chrom_bins). This function aggregates the site-level data (the
-    number of diff/same sites in Y), counts the number of neutral sites that
-    could be polymorphic form the masks (neut_masks), and interpolates the Bs
-    at the window midpoints. It also computes windowed π in these windows from
-    the data, as a check.
+#     # get the number of positions that are not polymorphic in the window
+#     nonpoly = num_nonpoly(Y_pos, chrom_bins, neut_masks)
 
-    b: BScores object
-    Y: DAC matrix
-    Y_pos: neutral region masks
-    nchroms: how many chromosomes were sequenced
-    bins: dictionary of window endpoints
-    """
-    chroms = b.pos.keys()
-
-    win_midpoints = bin_midpoints(chrom_bins)
-    win_Bs = {c: b.B_at_pos(c, x) for c, x in win_midpoints.items()}
[…]

### 2022-07-13 bgspy/models.py

@@ -38,7 +38,7 @@ import allel
 from scipy.optimize import minimize_scalar
 import tensorflow as tf
 
-from bgspy.utils import bin_chrom, bin_chroms, genome_emp_dists
+from bgspy.utils import bin_chrom, bin_chroms
 from bgspy.utils import Bdtype, BScores, BinnedStat
 from bgspy.likelihood import calc_loglik_components, loglik
 from bgspy.classic import calc_B, calc_B_parallel, calc_BSC16_parallel
@@ -66,27 +66,13 @@ class BGSModel(object):
         self.t = np.sort(t_grid)
         self.w = np.sort(w_grid)
 
-        # variation data, etc to calculate likelihoods
-        self.dacfile = None
-        self.metadata = None #
-        self._indices = None
-        self.neut_pos = None
-        self.neut_pos_map = None
-        # neutral region info
-        self.neut_masks = None
+        # Genomic Data
+        self.data = None
 
-        self.nloci = None
-        self.ac = None
-
-        # grid of pairwise same/diff counts
-        self.Y = None
         #likelihood
         self.pi0_ll = None
         self.pi0_grid = None
         self.pi0i_mle = None
-
-        # machine learning stuff
-        self.bfunc = None
 
     @property
     def seqlens(self):
@@ -99,93 +85,6 @@ class BGSModel(object):
     @property
     def segments(self):
         return self.genome.segments
-
-    def load_dacfile(self, dacfile, neut_regions):
-        neut_masks = ranges_to_masks(neut_regions, self.seqlens)
-        self.neut_masks = neut_masks
-        res = load_dacfile(dacfile, neut_masks)
-        self.neut_pos, self._indices, self.ac, self.neut_pos_map, self.metadata = res
-        self.nloci = len(list(*itertools.chain(self.neut_pos.values())))
-        ac = self.ac
-        assert(self.nloci == ac.shape[0])
-        an = ac.sum(axis=1)
-        # the number of chromosomes at fixed sites is set
-        # to the highest number of chromosomes in the sample (TODO, change?)
-        self.nchroms = an.max()
-        n_pairs = an * (an - 1) / 2
-        n_same = np.sum(ac * (ac - 1) / 2, axis=1)
-        n_diff = n_pairs - n_same
-        self.Y = dict()
-        m = np.stack((n_same, n_diff)).T
-        for chrom, indices in self._indices.items():
-            self.Y[chrom] = m[indices, :]
-
-    def pi(self, width=None):
-        """
[…]

### 2022-07-13 bgspy/utils.py

@@ -89,6 +89,13 @@ class BScores:
                 X[i, j, :] = self._interpolators[chrom][(w, t)](pos)
         return X
 
+    def bin_means(self, width):
+        """
+        """
+        assert width > self.step, "bin width is < step width; use interpolation"
+        pass # TODO
+
+
 def read_npy_dir(dir):
     return {f: np.load(os.path.join(dir, f)) for f in os.listdir(dir)}
 

### 2022-07-13 notes/theory.tex
cleaned up notes a bit

@@
 The total size of the set of samples $\mathcal{S}$ is $n_\mathcal{S} =
-|\mathcal{S}|$. Assuming all sites are biallelic, we can simplify the
-calculation of the likelihood by noting that for a site $v$'s vector of allele
-counts $[c_1, c_2]$, the total number of pairwise combinations with the same
-alleles is
+|\mathcal{S}|$. Assuming all sites are biallelic, we can simplify the inner
+summation by counting the number of possible same and different pairwise
+combinations. If a site $v$'s vector of allele counts is $[c_1, c_2]$, the
+total number of pairwise combinations with the same alleles is
 
@@
 where $n_T = n_\mathcal{S} (n_\mathcal{S} - 1) / 2$  is the total number of
-pairwise combinations across the sample set $\mathcal{S}$. Furthermore, note
-that we can partition the set $\mathcal{V}$ of neutral sites into polymorphic
+pairwise combinations across the sample set $\mathcal{S}$. Note that these
+site-specific counts allow us to use allelic counts directly, and can vary
+across sites. Our likelihood is then,
+
+\begin{align}
+  \log\mathcal{L}(\theta) &= \sum_{v \in \mathcal{V}} \left[\log(\pi(v | \theta)) n_\text{D}(v) + \log(1-\pi(v | \theta)) n_\text{S}(v)\right].
+\end{align}
+
+In practice, we calculate these values across bins. For each bin, we treat
+$\pi(v | \theta)$ as fixed, assuming that at this scale, the variation in
+expected diversity across sites is minimal. For a particular chromosome, we
+have two classes of sites: those included in the diversity calculation and
+those ignored. The former sites are all putatively neutral and have reliably
+called genotypes, and the other sites are possibly non-neutral or do have
+reliably called genotypes. The total log-likelihood is the sum of bin likelihoods,
+$\mathcal{L}(b)$
+
+\begin{align}
+   \mathcal{L}(\theta) =  \sum_b \mathcal{L}(b | \theta)
+\end{align}
+
+The likelihood within a bin is then,
+
+\begin{align}
+  \log\mathcal{L}(b | \theta)  &= \log(\bar{\pi}(b | \theta)) \sum_{v \in \mathcal{V}_b} n_D(v) + \log(1-\bar{\pi}(b | \theta)) \sum_{v \in \mathcal{V}_b} n_S(v)  \\
+                               &= \log(\bar{\pi}(b | \theta)) Y_D(b) + \log(1-\bar{\pi}(b | \theta)) Y_S(b)
+\end{align}
+%
+where the two sum terms as $Y_D(b)$ and $Y_S(b)$ are data reductions at the bin
+level.
+
+If the data are such that only polymorphic sites are considered, we can adapt
+this by partitioning the set $\mathcal{V}$ of neutral sites into polymorphic
 ($\mathcal{P}$) and fixed sites ($\mathcal{F}$), i.e. $\mathcal{V} =
@@
 
-In practice, we calculate these values across bins. For each bin, we treat
-$\pi(v | \theta)$ as fixed, assuming that at this scale, the variation in
-expected diversity across sites is minimal. For a particular chromosome, we
-have two classes of sites: those included in the diversity calculation and
-those ignored. The former sites are all putatively neutral and have reliably
-called genotypes, and the other sites are possibly non-neutral or do have
-reliably called genotypes. The log-likelihood is the sum of bin
-likelihoods, $\mathcal{L}(b)$
-
 \begin{align}
-  \log\mathcal{L}(b | \theta)  &= \log(\bar{\pi}(b | \theta)) \sum_{v \in \mathcal{P}} n_\text{D}(v) + \log(1-\bar{\pi}(b | \theta)) \left(\sum_{v \in \mathcal{V}} n_\text{S}(v) +  \sum_{v \in \mathcal{F}} n_T(v)  \right).
+  \log\mathcal{L}(b | \theta)  &= \log(\bar{\pi}(b | \theta)) \sum_{v \in \mathcal{P}_b} n_\text{D}(v) + \log(1-\bar{\pi}(b | \theta)) \left(\sum_{v \in \mathcal{P}_b} n_\text{S}(v) +  \sum_{v \in \mathcal{F}_b} n_T(v)  \right).
 \end{align}
 
-We can assume that the total number of combinations at each fixed site is
-constant, e.g. $n_T = n_T(v)$ for all $v$, simplifying this as
+Note that if we assume that the total number of combinations at each fixed site
+is constant, e.g. $n_T = n_T(v)$ for all $v$, then we can use $\sum_v n_T(v) =
+n_T |\mathcal{F}_b|$.
 
-\begin{align}
-  \log\mathcal{L}(b | \theta)  &= \log(\bar{\pi}(b | \theta)) \sum_{v \in \mathcal{P}} n_\text{D}(v) + \log(1-\bar{\pi}(b | \theta)) \left(\sum_{v \in \mathcal{V}} n_\text{S}(v) +  n_T(v) f(b)  \right)
-\end{align}
+% There are two approaches: interpolated $\pi(m | \theta)$ at the midpoint of the
+% bin position $m$, and averaged bin-averaged $\bar{\pi}(m | \theta)$.
 
-where $f(b)$ is the number of fixed sites in bin $b$. Thus, we our data can be
-reduced within a bin by these terms summed over sites in the window,
+\section*{Windowed Diversity}
 
-\begin{align}
-  \log\mathcal{L}(b | \theta)  &= \log(\bar{\pi}(b | \theta)) Y_D(b) + \log(1-\bar{\pi}(b | \theta)) Y_S(b)
-\end{align}
-
-\begin{align}
-  \log\mathcal{L}(b | \theta)  &= \log(\bar{\pi}_0 B(b | \vec{w}, \vec{t})) Y_D(b) + \log(1-\bar{\pi}_0 B(b |  \vec{w}, \vec{t})) Y_S(b)
-\end{align}
-
-where $\vec{w}$ and $\vec{t}$ are the vectors of mutation rates into
-
-There are two approaches: interpolated $\pi(m | \theta)$ at the midpoint of the
-bin position $m$, and averaged bin-averaged $\bar{\pi}(m | \theta)$.
+Although we use the components of diversity, $n_t$ and $n_s$, to calculate the
+likelihood, it is still of interest to calculate diversity from these in a
+window. The raw allele count data a $L \times \text{2}$ matrix,
 

### 2022-07-18 tools/treeseq2dac.py
changes to processing

@@ -77,12 +77,11 @@ def treeseq2dac(treefile, chrom, outfile, recmap, mu, seed=None):
     # remove the selected sites -- we might at one point want to see
     # whether including these messes up inference
     rts = rts.delete_sites([m.site for m in rts.mutations()])
-    ts = msprime.sim_mutations(rts, discrete_genome=True)
+    ts = msprime.sim_mutations(rts, rate=mu, discrete_genome=True)
     #__import__('pdb').set_trace()
     if outfile is None:
         outfile = treefile.replace('_treeseq.tree', '_dac.tsv.gz')
 
-    chrom = ratemap.chrom
     pos, nanc = count_ancestral(ts)
     nderv = 2*N - nanc
     with gzip.open(outfile, 'wt') as f:

### 2022-07-18 bgspy/data.py
changes to nbs

@@ -205,7 +205,10 @@ class GenomeData:
         bins = GenomicBins(self.genome.seqlens, width)
         reduced = dict()
         for chrom in self.genome.chroms:
-            site_ac = filter_sites(self.counts[chrom], filter_neutral, filter_accessible)
+            amask = self.accesssible_masks[chrom] if self.accesssible_masks is not None else None
+            nmask = self.neutral_masks[chrom] if self.neutral_masks is not None else None
+            site_ac = filter_sites(self.counts[chrom], filter_neutral,
+                                   filter_accessible, nmask, amask)
 
             # the combinatoric step -- turn site allele counts into
             # same/diff comparisons

### 2022-07-18 bgspy/utils.py

@@ -113,7 +113,7 @@ class BScores:
                 X[i, j, :] = self._interpolators[chrom][(w, t)](pos)
         return X
 
-    def bin_means(self, bins):
+    def bin_means(self, bins, return_bins=False):
         """
         Average the bins into larger genomic windows.
 
@@ -150,9 +150,17 @@ class BScores:
             # (like bins).
             mp_y = 0.5 * (y[1:, ...] + y[:-1, ...])
             mp_x = 0.5 * (x[1:] + x[:-1]) # position at midpoint
-            means[chrom] = bin_aggregate(mp_x, mp_y, np.mean,
+            res = bin_aggregate(mp_x, mp_y, np.mean,
                                          bins[chrom], axis=0)
-        return means
+            means[chrom] = res
+
+        if return_bins:
+            return means
+        # otherwise, we turn into another BScores object for easier downstream
+        # use
+        B = {c: x.stat for c, x in means.items()}
+        pos = {c: x.midpoints for c, x in means.items()}
+        return BScores(B, pos, self.w, self.t)
 
 def bin_chrom(end, width, dtype='uint32'):
     """

### 2022-07-18 notes/theory.tex
new math notes

@@
 site-specific counts allow us to use allelic counts directly, and can vary
-across sites. Our likelihood is then,
+across sites. Our log-likelihood is then,
 
 \begin{align}
-  \log\mathcal{L}(\theta) &= \sum_{v \in \mathcal{V}} \left[\log(\pi(v | \theta)) n_\text{D}(v) + \log(1-\pi(v | \theta)) n_\text{S}(v)\right].
+  \ell(\theta) &= \sum_{v \in \mathcal{V}} \left[\log(\pi(v | \theta)) n_\text{D}(v) + \log(1-\pi(v | \theta)) n_\text{S}(v)\right].
 \end{align}
@@
 reliably called genotypes. The total log-likelihood is the sum of bin likelihoods,
-$\mathcal{L}(b)$
+$\ell(b)$
 
 \begin{align}
-   \mathcal{L}(\theta) =  \sum_b \mathcal{L}(b | \theta)
+   \ell(\theta) =  \sum_b \ell(b | \theta)
 \end{align}
@@
 \begin{align}
-  \log\mathcal{L}(b | \theta)  &= \log(\bar{\pi}(b | \theta)) \sum_{v \in \mathcal{V}_b} n_D(v) + \log(1-\bar{\pi}(b | \theta)) \sum_{v \in \mathcal{V}_b} n_S(v)  \\
+  \ell(b | \theta)  &= \log(\bar{\pi}(b | \theta)) \sum_{v \in \mathcal{V}_b} n_D(v) + \log(1-\bar{\pi}(b | \theta)) \sum_{v \in \mathcal{V}_b} n_S(v)  \\
                                &= \log(\bar{\pi}(b | \theta)) Y_D(b) + \log(1-\bar{\pi}(b | \theta)) Y_S(b)
@@
 \begin{align}
-  \log\mathcal{L}(\theta) &= \sum_{v \in \mathcal{V}} \left[\log(\pi(v | \theta)) n_\text{D}(v) + \log(1-\pi(v | \theta)) n_\text{S}(v)\right] \\
+  \ell(\theta) &= \sum_{v \in \mathcal{V}} \left[\log(\pi(v | \theta)) n_\text{D}(v) + \log(1-\pi(v | \theta)) n_\text{S}(v)\right] \\
                   &= \sum_{v \in \mathcal{P}} \left[\log(\pi(v | \theta)) n_\text{D}(v) + \log(1-\pi(v | \theta)) n_\text{S}(v)\right] + \sum_{v \in \mathcal{F}} \log(1-\pi(v | \theta)) n_T(v)  \\
@@
 \begin{align}
-  \log\mathcal{L}(b | \theta)  &= \log(\bar{\pi}(b | \theta)) \sum_{v \in \mathcal{P}_b} n_\text{D}(v) + \log(1-\bar{\pi}(b | \theta)) \left(\sum_{v \in \mathcal{P}_b} n_\text{S}(v) +  \sum_{v \in \mathcal{F}_b} n_T(v)  \right).
+  \ell(b | \theta)  &= \log(\bar{\pi}(b | \theta)) \sum_{v \in \mathcal{P}_b} n_\text{D}(v) + \log(1-\bar{\pi}(b | \theta)) \left(\sum_{v \in \mathcal{P}_b} n_\text{S}(v) +  \sum_{v \in \mathcal{F}_b} n_T(v)  \right).
 \end{align}
@@
 n_T |\mathcal{F}_b|$.
+
+\section*{The B Components}
+
+The core parts of our likelihood are,
+
+\begin{align}
+  \ell(\theta) &=  \sum_b [\log(\bar{\pi}(b | \theta)) Y_D(b) + \log(1-\bar{\pi}(b | \theta)) Y_S(b)]
+\end{align}
+
+where,
+
+\begin{align}
+  \bar{\pi}(b |\theta) = \pi_0(b) \bar{B}(b | \theta).
+\end{align}
+
+Here, $\bar{B}(b | \theta)$ is the predicted reduction in diversity due to BGS
+in window $b$, given background selection parameters $\theta$. In practice,
+this is site-specific. We can write the reduction at any neutral site $v$ in
+the genome as the product of $B$s across all segments,
+
+\begin{align}
+  B(v | \theta) = \exp\left(- \sum_g \int f(\mu(\mathcal{A}(g)), s, S_g) w(s|\mathcal{A}(g)) ds \right)
+\end{align}
+
+where $S_g$ is exogenous genomic data about the segment, $S_g = \{L_g, r_g,
+\rho(|v-p_g|)\}$, where $L_g$ is the segment's length, $r_g$ is the
+recombination rate per basepair in the segment, and $\rho(|v-p_g|)$ is the
+recombination distance between the focal site $v$ and the segment position
+$p_g$ (we approximate, and use the nearest end position to the neutral site).
+Additionally, $w(s|\mathcal{A}(g))$ is the distribution of selection
+coefficients for segment $g$, if segment $g$ is a member of annotation class
+$\mathcal{A}(g)$.
+
+We can think about the DFE as the conditional distribution of a particular
+selection coefficient given a mutation occurs, for a particular annotation
+class. The BGS function $f(\cdot)$ only depends on $\mu$ through the
+introduction of deleterious alleles with selection coefficient $s$ at rate
+$\omega(s|\mathcal{A}(g)) = \mu(\mathcal{A}(g)) w(s | \mathcal{A}(g))$. Thus,
+we can write,
+
+\begin{align}
+  B(v | \theta) = \exp\left(- \sum_g \int f(\omega(s|\mathcal{A}(g)), s, S_g) ds \right)
+\end{align}
+
+which we can discretize as,
+
+\begin{align}
+  B(v | \theta) = \exp\left(- \sum_g \sum_s f(\omega(s|\mathcal{A}(g)), s, S_g) \right).
+\end{align}
+
+Next, note that there are a finite number of annotation classes, $\mathcal{A}
+\to \{a_1, a_2, \ldots, a_k\}$, so we can further partition this as
+
+\begin{align}
+  B(v | \theta) = \exp\left(- \sum_{\{g \;:\; \mathcal{A}(g) = a_1\}} \sum_s f(\omega(s| a_1), s, S_g) + \sum_{\{g \;:\; \mathcal{A}(g) = a_2\}} \sum_s f(\omega(s| a_2), s, S_g) + \ldots  \right)
+\end{align}
+
+Let us define the $d_\omega \times d_s \times d_g$ multidimensional array
+$\mathbf{F}$, and the $d_g \times d_a$ feature classification matrix
+$\mathbf{A}$.
 

### 2022-08-18 manuscript/manuscript.tex
some ms notes

\documentclass[11pt]{article}
\RequirePackage{fullpage}
\RequirePackage[font=small,labelfont=bf]{caption}
\RequirePackage{amsmath,amssymb,amsthm}
\RequirePackage{graphicx}
\RequirePackage[hidelinks]{hyperref}
\RequirePackage{subcaption}
\RequirePackage{wasysym}
\RequirePackage{authblk}
\RequirePackage{bm}
\RequirePackage{bbm}
% \RequirePackage[osf]{mathpazo}
\RequirePackage[bibstyle=authoryear,citestyle=authoryear-comp,maxbibnames=9,maxcitenames=1,backend=biber,natbib=true,uniquelist=false,hyperref=true]{biblatex}
\usepackage{color}
\usepackage{nicefrac}

\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\text{V}}
% \DeclareMathOperator{\var}{var}
% \DeclareMathOperator{\cov}{cov}

\addbibresource{biblio.bib}

\title{}

\author{Vince Buffalo and Andrew Kern}

\begin{document}
\maketitle

A central goal of evolutionary genetics is to quantify the relative roles
different evolutionary processes have on genetic variation. Certain
evolutionary processes, such as demographic bottlenecks or expansions, act
uniformly across the genome as a whole. Other processes, such as selection, are
hypothesized to act differently across the genome depending on the density of
genomic features, and local heterogeneity in recombination and mutation rate.
With the emergence of population genomic data, a variety of statistical models
have teased apart the contribution of these various processes. These
statistical models differ in how they reduce large genomic data to summary
statistics, as well as in their statistical methodology. Some models reduce
genomic data to the histograms of the frequency spectrum (XXX), while others
model scalar summaries of the frequency spectrum along the genome.

One class of models, parametric statistical models, leverage population genetic
theory to inform expectations of the mean pairwise diversity along the
chromosome. Estimation under these models often uses a composite likelihood
approach, where the joint distribution is assumed independent across positions.
While composite likelihoods are simpler to formulate and lead to consistent
parameter estimation, ignoring dependencies between neighboring sites has two
primary disadvantages. First, unlike proper likelihoods, the composite
likelihood's curvature is uninformative about parameter estimate uncertainty.
Previous work has circumvented this shortcoming through (1) resampling-based
uncertainty estimates, or (2) adjusting the composite likelihood's curvature to
account for these dependencies. Second, similar to inference under phylogenetic
comparative models, the dependency structure tells us how to properly weight
information across sites. For example, imagine a conserved segment 10 kilobases
long right on the border of two genomic bins, which due to linked selection,
bot experience a reduction in diversity. When estimating parameters, the
independence assumption of composite likelihoods would weight these observed
reductions equal to the amount of a reduction observed on an entirely different
independent chromosome.

patterns of genetic variability along the genome are treated as
coming from an indeprocess $m(z)$.

Here, we focus on solving an issue processes: background

\printbibliography

\end{document}

### 2022-08-18 bgspy/src/likelihood.c

@@ -259,3 +259,63 @@ double normal_loglik(const double *theta,
     return ll;
 }
 
+
+
+double mvn_loglik(const double *theta,
+                  const double *nS, const double *nD, 
+                  const double *logB, 
+                  const double *w,
+                  const int *chrom,
+                  const double *rdist,
+                  const ssize_t *logB_dim, 
+                  const ssize_t *logB_strides) {
+
+    ssize_t nx = logB_dim[0]; 
+    ssize_t nw = logB_dim[1];
+    ssize_t nt = logB_dim[2];
+    ssize_t nf = logB_dim[3];
+    //printf("dims: nx=%d, nw=%d, nt=%d, nf=%d\n", nx, nw, nt, nf);
+    double pi0 = theta[0];
+    double mu = theta[1];
+    double alpha = theta[2];
+    double rho = theta[3];
+    ssize_t nW = nt*nf;
+    double *W = calloc(nW, sizeof(double));
+    memcpy(W, theta + 4, nW * sizeof(double));
+    double logBw_i; // the logBw[i] (interpolated on w) for bin i
+    double Wjk;
+    double *loc = calloc(nx, sizeof(double));
+    double *scale = calloc(nx, sizeof(double));
+
+    //print_theta(theta, 2+nW);
+    double N, sigma2, loc;
+    for (ssize_t i=0; i < nx; i++) {
+        logBw_i = 0.; // initialize start of sum
+        for (ssize_t j=0; j < nt; j++) {
+            for (ssize_t k=0; k < nf; k++) {
+                Wjk = W_GET(W, j, k, nf);
+                //if (j == 4) printf("j=%d, k=%d,offset: %d, nW=%d, W=%g W_GET=%g\n", j, k, nf*(j-1) + k, nW, W[(j-1)*nf + k], Wjk);
+                //printf("i=%d, j=%d, k=%d | mu=%g, Wjk=%g, mu Wjk=%g\n", i, j, k, 
+                //       mu, Wjk, mu * Wjk);
+                double Binc = interp_logBw(mu*Wjk, w, logB, nw, i, j, k, logB_strides);
+                if (isnan(Binc)) {
+                    printf("NaN Binc! theta=[");
+                    print_theta(theta, 2+nW);
+                    printf("]");
+                    printf("i=%d, j=%d, k=%d | Wjk=%g,", i, j, k, Wjk);
+                    free(W);
+                    return NAN;
[…]

### 2022-08-18 bgspy/utils.py

@@ -216,6 +216,7 @@ class BScores:
                             func = self.w, b
                         chrom_interpols[i][j][k] = func
             interpols[chrom] = Bw_interpol(chrom_interpols, npos, self.nt, self.nf, jax)
+
     def _build_interpolators(self, **kwargs):
         """
         Build positional interpolators for each chromsome and w/t combination.
@@ -233,7 +234,7 @@ class BScores:
                 for chrom in Bs:
                     # The last dimension is features matrix -- for now, only
                     # one feature is supported
-                    y = Bs[chrom][:, i, j]
+                    y = Bs[chrom][:, i, j].squeeze()
                     func = interpolate.interp1d(x[chrom], y,
                                                 fill_value=(y[0], y[-1]),
                                                 **kwargs)

### 2022-08-29 manuscript/manuscript.tex
new method eval

@@
 
+BGS as evolution around the static features of the genome. Due to the slow rate
+of genomic structural evolution, much of the genome can be treated as static
+throughout the genome. These regions are under differing levels of selective
+constraint, which induces background selection at neighboring regions.
+
 \printbibliography

### 2022-08-29 bgspy/data.py

@@ -238,7 +238,7 @@ class GenomeData:
         If chrom is not set a new dummy Genome object is created.
         """
         if chrom is not None:
-            assert chrom in self.genome.chroms, "chrom not in GenomeData.genome"
+            assert chrom in self.genome.chroms, f"{chrom} not in GenomeData.genome"
         msg = "set either ts or file, not both"
         if file is not None:
             assert ts is None, msg
@@ -252,11 +252,10 @@ class GenomeData:
             num_deriv[int(var.site.position)] = nd
         assert np.sum(np.isnan(num_deriv)) == 0, "remaining nans -- num mut/num allele mismatch"
         ntotal = np.repeat(ts.num_samples, sl)
-        if chrom is None:
+        if self.genome is None:
             chrom = 'chrom'
             g = Genome('dummy genome', seqlens={chrom: sl})
-        else:
-            g = self.genome
+        g = self.genome
         self.genome = g
         nanc = ntotal - num_deriv
         self.counts = {chrom: np.stack((nanc, num_deriv)).T}

### 2022-08-29 bgspy/likelihood.py

@@ -400,6 +400,20 @@ def predict_simplex(theta, logB, w):
                 logBw[i] += np.interp(mu*W[j, k], w, logB[i, :, j, k])
     return pi0*np.exp(logBw)
 
+def predict_freemutation(theta, logB, w):
+    """
+    """
+    nx, nw, nt, nf = logB.shape
+    # mut weight params
+    pi0, W = theta[0], theta[1:]
+    W = W.reshape((nt, nf))
+    # interpolate B(w)'s
+    logBw = np.zeros(nx, dtype=float)
+    for i in range(nx):
+        for j in range(nt):
+            for k in range(nf):
+                logBw[i] += np.interp(W[j, k], w, logB[i, :, j, k])
+    return pi0*np.exp(logBw)
 
 class BGSLikelihood:
     """
@@ -492,6 +506,18 @@ class BGSLikelihood:
         for i in range(n):
             nlls[i] = negll(thetas[i, ...], Y, self.logB, self.w)
         return grid, thetas, nlls
+
+
+    def predict(self, optim=None):
+        if optim is None:
+            theta = self.theta_
+        else:
+            thetas = self.optim.thetas_
+            if optim == 'random':
+                theta = thetas[np.random.randint(0, thetas.shape[0]), :]
+            else:
+                theta = thetas[optim]
+        return predict_freemutation(theta, self.logB, self.w)
 
     @property
     def mle_pi0(self):

### 2022-08-29 bgspy/optim.py

@@ -101,7 +101,7 @@ def nlopt_simplex_isres_worker(start, func, nt, nf, bounds,
     success = opt.last_optimize_result()
     return nll, mle, success
 
-def optim_plot(only_success=True, tail=0.5, **runs):
+def optim_plot(only_success=True, tail=0.5, x_percent=False, downsample=None, **runs):
     """
     Make a plot of the rank-ordered optimization minima for the
     labeled runs keywords. Only the top 'tail' entries are kept.
@@ -110,6 +110,10 @@ def optim_plot(only_success=True, tail=0.5, **runs):
     for i, (key, run) in enumerate(runs.items()):
         nll = run.nlls_
         succ = run.success_
+        if downsample is not None:
+            idx = np.random.choice(np.arange(len(nll)), downsample)
+            nll = nll[idx]
+            succ = succ[idx]
         if only_success:
             keep = succ >= 1
             nll = nll[keep]
@@ -120,7 +124,15 @@ def optim_plot(only_success=True, tail=0.5, **runs):
         y = nlls[sort_idx]
         x = 2*i + succ[nll < q].astype('int')
         cols = mpl.cm.get_cmap('Paired')(x)
-        ax.scatter(list(reversed(range(len(y)))), y, s=1, label=key, c=cols)
+        x = np.array(list(reversed(range(len(y)))))
+        if x_percent:
+            x = x / len(x)
+        ax.scatter(x, y, s=1, label=key, c=cols)
+    ax.set_ylabel("negative log-likelihood")
+    if x_percent:
+        ax.set_xlabel("rank (proportion of total)")
+    else:
+        ax.set_xlabel("rank")
     ax.legend()
     ax.semilogy()
 

### 2022-09-23 manuscript/manuscript.tex
some writing

@@
 \newcommand{\V}{\text{V}}
-% \DeclareMathOperator{\var}{var}
-% \DeclareMathOperator{\cov}{cov}
+\DeclareMathOperator{\var}{var}
+\DeclareMathOperator{\cov}{cov}
 
@@
 
+Signal processing
+
+Genomes have inherent structure that impacts the linked selection map. There is
+tremendous heterogeneity across species in the organization of likely selected
+regions on the genetic and physical maps.
+
+Within an organism, there are two key determinants of B maps: recombination and
+gene density.
+
+Initially, the relationship between recombination and diversity was regressed
+with no regard to the spatial organization of the genome.
+
+Certain regions are pinchpoints of selection: some combination of low
+recombination and high gene density lead to very strong reductions in genetic
+diversity. In these regions, coalescent dynamics change considerably; where as
+under strong background selection, deleterious alleles would not fix, in these
+regions the can fix. Additionally, the selective interference impacts positive
+selection.
+
+\section*{Theory}
+
+We model the reduction in pairwise coalescence times along the genome due to
+background selection by extending the work of \textcite{Santiago2016-mu}
+(hereafter, the SC16 model). The SC16 model derives from a class of linked
+selection \parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs} that
+predict the reduction in pairwise coalescence time at a focal neutral site
+using macro-state parameters such as fitness variation, rather than micro-state
+parameters like the frequency trajectories of selected alleles (e.g.
+\textcite{Maynard_Smith1974-zr}). We will briefly introduce the SC16 model's
+approach and how we validated this model could be extended to whole-genome
+inference.
+
+Classic background selection theory models how coalescence times are reduced by
+deleterious mutation-selection balance at neighboring sites. Building upon
+older models of genetic load \parencite{Crow1970-wj,Kimura1966-bk}, these
+models consider the equilibrium distribution of the \emph{number} of
+deleterious mutations segregating on haplotypes
+\parencite{Charlesworth1993-gb,Nordborg1996-nq}. Given an infinitely large
+population, or sufficiently strong selection, the deleterious mutations in
+these models cannot fix; consequently, any lineage containing
+
+While the effect of selection on linked selection distorts neighboring
+genealogies (XXX), skewing the site frequency spectrum (XXX) and altering
+linkage disequilibria along the chromosome, we focus on the impact of selection
+on pairwise coalescence rates. This simple summary is amenable to mathematical
+and statistical inference, and is robust to difficulties in identifying rare
+variants. The effect of background selection on pairwise coalescence rates is
+expressed as the reduction ratio $B = \nicefrac{N_e}{N}$ where $N$ is the
+drift-effective population size (i.e. including demographic processes) and
+$N_e$ is the effective population size under background selection.
+
+Under quantitative genetic models of linked selection, the reduction in
+effective population size is determined by the product of fitness variation,
+$V_f$, and an inflation factor $Q$. It is helpful to understand this by comparing
+it to a model of $N_e$ without \emph{heritable} fitness variation but with
+\emph{non-heritable} fitness variation. In this case,
+
+\begin{align}
+  N_e = \frac{4N}{2 + V_k},
+\end{align}
+
+where $V_k$ is the variance in offspring number across individuals. Under a
+constant population size Wright--Fisher model of reproduction, $V_k \approx 2$,
+and this simplifies to $N_e = N$. However, as first noted by
+\textcite{Robertson1961-ho}, linked selection inflates the variance in
+offspring number, as a neutral allele that becomes associated with a
+high-fitness or low-fitness genetic background has an $1-r$ chance of remaining
+associated with that background in the next generation, where $r$ is the chance
+of recombination (or independent assortment, $r=\nicefrac{1}{2}$).
+
+This persistence of this association inflates the effective variance in fitness
+by a factory $Q^2$,
+
+\begin{align}
+  N_e = \frac{4N}{2 + V_k + 4 Q^2 V_f},
+\end{align}
+
+\begin{align}
+  N_e = \frac{4N}{2 + V_k + 4 Q(r)^2 V_f},
+\end{align}
+
+\begin{align}
+  N_e = N \exp\left(-Q^2 V_f\right)
+\end{align}
+
+In a region of $L$ basepairs under purifying selection, there is a net flux of
+deleterious mutations in at rate $U = 2 \mu L$ per generation (where the factor
+of two is because in diploids, $2L$ is the total number of basepairs per genome).
+
+Typical models of background selection consider the equilibrium distribution of
+the \emph{count} of deleterious mutations, rather than then level of fitness
+variation implied by this count. If we assume fixed additive effects across
+alleles and multiplicative fitness effects across loci, each copy of a
+deleterious mutation an individual carries reduces their fitness by a factor of
+$1-s$. The total reduction for an individual is
+
+\begin{align}
+  x = (1-s)^n
+\end{align}
+
+where $n$ is the number of deleterious alleles an individual carries. Under
+mutation-selection balance, the population evolves to either a stationary
+distribution under the strong BGS domain, or a traveling distribution in weak
+selection case. Under some assumptions, equilibrium models of the strong domain
+show that the distribution $f(n) \sim \text{Pois}(\nicefrac{U}{s})$
+\parencite{Kimura1966-bk,Crow1970-wj} (CHECK) in the asexual case, and the
+fixation rate of deleterious mutations per generation, or ratchet rate, is $R =
+0$ since the probability of fixation, $p_F$ is zero. By contrast, as the
+selection coefficient grows weaker, $s \to 0$, the probability of fixation $p_F
+\to \nicefrac{1}{2N}$, and the ratchet approaches the neutral rate of $R =
+\mu$. The weak selection domain is of importance because both because it
+implies a certain likely irreversible rate of fitness reduction, and because it
+could possibly confound population genetic inference which assumes dynamics
+only under the strong model.
+
+As $s \to 0$, the fitness distribution, put on a scale reflecting the fitness
+relative at one moment in time, travels leftwards and changes shape. It has
+been difficult to find dynamic equations for the rate of the ratchet, since
+lower moments of the distribution, such as the variation, depend on higher
+moments, such as the skew \parencite{Good2013-lp,Haigh1978-gt,Higgs1995-xc}.
+This moment closure problem occurs regardless of whether one models the count
+number of deleterious mutations (XXX) or the fitness distribution (XXX).
+
+We can approximate the mapping between the distribution of deleterious
+mutations in the population, and the fitness distribution.
+
+Assuming $s$ is small, we can write express this in terms of the additive
+genetic variance,
+
+\begin{align}
+  V(f) \approx s^2 \V(n) + \delta_\text{LD}.
+\end{align}
+
+The first term is the additive genic variance, which is
+
+Under this model, the rate of the ratchet $R = U  - s V(n)$, which implies,
+
+\begin{align}
+  R = U - \nicefrac{V(f)}{s} \\
+  V(f) = s(U - R)
+\end{align}
+
+\begin{align}
+
+\end{align}
+
+- We need to mention the two heterozygosities (S&C '16 says they diverge under
+$N_e s < 1$ -- could explain anything we see?)
+
+- A point about reconciling the two micro and macro-state theories of
+population and quantitative genetics.
+
+\subsection*{The Rate of the Ratchet}
+
+\section*{Methods}
+
+\subsection*{Data Preprocessing}
+
+\subsection*{Human}
+
+Site-level neutral and accessibility masks
+
+Bin-level masks based on fraction of accessible bases in a window (7\%).
+
+Bin-level tail filtering.
+
 \printbibliography

### 2022-09-26 manuscript/manuscript.tex
ms notes

@@
 \RequirePackage{bm}
-\RequirePackage{bbm}
-% \RequirePackage[osf]{mathpazo}
+% \RequirePackage{bbm}
+\RequirePackage[osf]{mathpazo}
 \RequirePackage[bibstyle=authoryear,citestyle=authoryear-comp,maxbibnames=9,maxcitenames=1,backend=biber,natbib=true,uniquelist=false,hyperref=true]{biblatex}
@@
 \section*{Theory}
+
+\subsection*{Effective Population Size for Pairwise Diversity}
+
+When quantifying genetic variation with average pairwise diversity, often a
+single effective population size parameterizes the rate of pairwise
+coalescences of the entire evolutionary history. This is sufficient for
+constant-rate Kingman coalescences, but weak selection distorts genealogies
+(desai and nicolaison), creating a coalescent process with non-constant rates.
+Consequently, the expected pairwise coalescent time between two lineages $i$
+and $j$
+
+\begin{align}
+  \E(T) &= \sum_i T(i) \P(\mathrm{MRCA}(i)) \\
+        &= \sum_i T(i) \prod_{t=1}^i \left(1 - \frac{1}{2N_{e(i)}}\right)
+\end{align}
+
+% ADD TO INTRO TODO
+Comment about characteristic scale $\nicefrac{\mu}{s}$. Is the fact that we see
+BGS on megabase scales because this is the only scale we can detect strong
+selection under?
+
+approximately $\pi \approx 4N_e \mu$ (for low mutation rate). However, this
+assumes a constant Kingman coalescent with coalescent effective population size
+$N_e$.
+
+\subsection*{}
 
@@
 relative at one moment in time, travels leftwards and changes shape. It has
-been difficult to find dynamic equations for the rate of the ratchet, since
-lower moments of the distribution, such as the variation, depend on higher
-moments, such as the skew \parencite{Good2013-lp,Haigh1978-gt,Higgs1995-xc}.
-This moment closure problem occurs regardless of whether one models the count
+been difficult to find dynamic equations for the rate of the ratchet and
+fitness distribution, since lower moments of the distribution, such as the
+variation, depend on higher moments, such as the skew
+\parencite{Good2013-lp,Haigh1978-gt,Higgs1995-xc}. This is known as the ``moment
+closure problem" and it occurs regardless of whether one models the count
 number of deleterious mutations (XXX) or the fitness distribution (XXX).
@@
 We can approximate the mapping between the distribution of deleterious
-mutations in the population, and the fitness distribution.
+mutations in the population, and the fitness distribution. It is worth noting
+that the count of deleterious mutations in an individual $i$ can be written
+$n_i = \sum_l^{L} g_l$ where each $g_l$ is an indicator variable for whether
+the individual carries a mutation at position $l$. Then, $\V_i(n) = \sum_l
+\V(g_l) + \sum_{i,j}\cov(g_i, g_j)$.
 
-Assuming $s$ is small, we can write express this in terms of the additive
-genetic variance,
+the
+number of deleterious mutations in gamete $i$; thus $V(n) = \sum_i^{2N} V(x_i)
++ \sum_{i \ne j} \cov(x_i, x_j)$ where the last term is the correlation among
+gamete mutation counts, which is an LD quantity. Then, assuming $s$ is small,
+we can write express this in terms of the additive genetic variance,
 
 \begin{align}
-  V(f) \approx s^2 \V(n) + \delta_\text{LD}.
+  V(f) &\approx s^2 \V_a(n) + \delta_\text{LD}
 \end{align}
@@
 
-\begin{align}
+Throughout, it's worthwhile to keep into mind many population-level
+macro-states are simply distributional summaries of the current population. For
+example, the variance in deleterious mutations in the population is simply,
+$\V(n) = \nicefrac{1}{2N} \sum_i (x_i - \mu_x)^2$
 
-\end{align}
-
-- We need to mention the two heterozygosities (S&C '16 says they diverge under
+- We need to mention the two heterozygosities (S\&C '16 says they diverge under
 $N_e s < 1$ -- could explain anything we see?)

### 2022-09-28 manuscript/manuscript.tex
more theory notes

@@
 \RequirePackage{bm}
-% \RequirePackage{bbm}
+\RequirePackage{bbm}
 \RequirePackage[osf]{mathpazo}
-\RequirePackage[bibstyle=authoryear,citestyle=authoryear-comp,maxbibnames=9,maxcitenames=1,backend=biber,natbib=true,uniquelist=false,hyperref=true]{biblatex}
+\RequirePackage[bibstyle=authoryear,citestyle=authoryear-comp,
+                date=year,
+                maxbibnames=9,maxnames=5,maxcitenames=2,
+                backend=biber,uniquelist=false,uniquename=false,
+                % style=apa,
+                sorting=nyt,
+                % sorting=,
+                hyperref=true]{biblatex}
 \usepackage{color}
@@
 
-A central goal of evolutionary genetics is to quantify the relative roles
-different evolutionary processes have on genetic variation. Certain
-evolutionary processes, such as demographic bottlenecks or expansions, act
-uniformly across the genome as a whole. Other processes, such as selection, are
-hypothesized to act differently across the genome depending on the density of
-genomic features, and local heterogeneity in recombination and mutation rate.
-With the emergence of population genomic data, a variety of statistical models
-have teased apart the contribution of these various processes. These
-statistical models differ in how they reduce large genomic data to summary
-statistics, as well as in their statistical methodology. Some models reduce
-genomic data to the histograms of the frequency spectrum (XXX), while others
-model scalar summaries of the frequency spectrum along the genome.
+% A central goal of evolutionary genetics is to quantify the relative roles
+% different evolutionary processes have on genetic variation. Certain
+% evolutionary processes, such as demographic bottlenecks or expansions, act
+% uniformly across the genome as a whole. Other processes, such as selection, are
+% hypothesized to act differently across the genome depending on the density of
+% genomic features, and local heterogeneity in recombination and mutation rate.
+% With the emergence of population genomic data, a variety of statistical models
+% have teased apart the contribution of these various processes. These
+% statistical models differ in how they reduce large genomic data to summary
+% statistics, as well as in their statistical methodology. Some models reduce
+% genomic data to the histograms of the frequency spectrum (XXX), while others
+% model scalar summaries of the frequency spectrum along the genome.
 
-One class of models, parametric statistical models, leverage population genetic
-theory to inform expectations of the mean pairwise diversity along the
-chromosome. Estimation under these models often uses a composite likelihood
-approach, where the joint distribution is assumed independent across positions.
-While composite likelihoods are simpler to formulate and lead to consistent
-parameter estimation, ignoring dependencies between neighboring sites has two
-primary disadvantages. First, unlike proper likelihoods, the composite
-likelihood's curvature is uninformative about parameter estimate uncertainty.
-Previous work has circumvented this shortcoming through (1) resampling-based
-uncertainty estimates, or (2) adjusting the composite likelihood's curvature to
-account for these dependencies. Second, similar to inference under phylogenetic
-comparative models, the dependency structure tells us how to properly weight
-information across sites. For example, imagine a conserved segment 10 kilobases
-long right on the border of two genomic bins, which due to linked selection,
-bot experience a reduction in diversity. When estimating parameters, the
-independence assumption of composite likelihoods would weight these observed
-reductions equal to the amount of a reduction observed on an entirely different
-independent chromosome.
+% One class of models, parametric statistical models, leverage population genetic
+% theory to inform expectations of the mean pairwise diversity along the
+% chromosome. Estimation under these models often uses a composite likelihood
+% approach, where the joint distribution is assumed independent across positions.
+% While composite likelihoods are simpler to formulate and lead to consistent
+% parameter estimation, ignoring dependencies between neighboring sites has two
+% primary disadvantages. First, unlike proper likelihoods, the composite
+% likelihood's curvature is uninformative about parameter estimate uncertainty.
+% Previous work has circumvented this shortcoming through (1) resampling-based
+% uncertainty estimates, or (2) adjusting the composite likelihood's curvature to
+% account for these dependencies. Second, similar to inference under phylogenetic
+% comparative models, the dependency structure tells us how to properly weight
+% information across sites. For example, imagine a conserved segment 10 kilobases
+% long right on the border of two genomic bins, which due to linked selection,
+% bot experience a reduction in diversity. When estimating parameters, the
+% independence assumption of composite likelihoods would weight these observed
+% reductions equal to the amount of a reduction observed on an entirely different
+% independent chromosome.
 
-patterns of genetic variability along the genome are treated as
-coming from an indeprocess $m(z)$.
+% patterns of genetic variability along the genome are treated as
+% coming from an indeprocess $m(z)$.
 
-Here, we focus on solving an issue processes: background
+% Here, we focus on solving an issue processes: background
 
-BGS as evolution around the static features of the genome. Due to the slow rate
-of genomic structural evolution, much of the genome can be treated as static
-throughout the genome. These regions are under differing levels of selective
-constraint, which induces background selection at neighboring regions.
+% BGS as evolution around the static features of the genome. Due to the slow rate
+% of genomic structural evolution, much of the genome can be treated as static
+% throughout the genome. These regions are under differing levels of selective
+% constraint, which induces background selection at neighboring regions.
 
-Signal processing
+% Signal processing
 
-Genomes have inherent structure that impacts the linked selection map. There is
-tremendous heterogeneity across species in the organization of likely selected
-regions on the genetic and physical maps.
+% Genomes have inherent structure that impacts the linked selection map. There is
+% tremendous heterogeneity across species in the organization of likely selected
+% regions on the genetic and physical maps.
 
-Within an organism, there are two key determinants of B maps: recombination and
-gene density.
+% Within an organism, there are two key determinants of B maps: recombination and
+% gene density.
 
-Initially, the relationship between recombination and diversity was regressed
-with no regard to the spatial organization of the genome.
+% Initially, the relationship between recombination and diversity was regressed
+% with no regard to the spatial organization of the genome.
 
-Certain regions are pinchpoints of selection: some combination of low
-recombination and high gene density lead to very strong reductions in genetic
-diversity. In these regions, coalescent dynamics change considerably; where as
-under strong background selection, deleterious alleles would not fix, in these
-regions the can fix. Additionally, the selective interference impacts positive
+% Certain regions are pinchpoints of selection: some combination of low
+% recombination and high gene density lead to very strong reductions in genetic
+% diversity. In these regions, coalescent dynamics change considerably; where as
+% under strong background selection, deleterious alleles would not fix, in these
+% regions the can fix. Additionally, the selective interference impacts positive
+% selection.
+
+- Disentangling evolutionary processes from the patterns of genetic variation
+across the genome. The structure of what can be under selection shapes these
+patterns when selection acts; when drift acts, these patterns should not exist.
+
+- For species with small effective population sizes (and perhaps large map
+lengths), BGS seems to be a dominant selective force in shaping patterns of
+diversity (e.g. at the megabase scale in humans).
+
+- However, current approaches to fitting BGS models fail under the weak
+selection domain. This weak selection domain is important (discuss
+HRI/ratchet/etc). Importantly, the characteristic scale of the impact of linked
+selection is $\nicefrac{s}{r}$, meaning weak selection acts on smaller scales
+than strong selection. Our ability to see selection at small scales may be
+inhibited by the fact that our current BGS theory only works for strong
 selection.
+
+- We need to say briefly that we're using BGS to mean linked selection due to
+purifying selection, and thus the strong and weak regimes (classically it has
+just meant the strong regime).
 
@@
 
+There has been considerable past work to find both the rate of the ratchet
+(XXX), and summaries of the genealogies created under weak purifying selection
+(XXX). Our approximation combines a few theoretic avenues from past work. Our
+primary goal is to find a theoretic scaling function $B$ that rescales the neutral
+pairwise coalescent times the appropriate amount to
+
+Strictly speaking, there is no way to rescale the coalescent under general
+background selection process, as genealogies deviate from the Kingman
+coalescent slightly under strong selection
+\parencite{Walczak2012-fi,Cvijovic2018-vd} and greatly under weak selection
+\parencite{OFallon2010-my}
+
+First, we approximate the coalescent effective population size, which
+determines expected pairwise diversity, from the variance effective population
+size. Then, we use a general quantitative genetic model of linked selection to
+find the variance effective population size under background selection. Since
+the impact of selection under this model is parameterized by the additive
+genetic fitness variance, we use an approximation for this based on the
+additive \emph{genic} variance, which depends on the rate of the ratchet.
+Previous work has been unable to find a simple closed-form solution to the rate
+of the ratchet because it depends on moments of the fitness distribution (e.g.
+fitness variation), which in turn depend on even higher-moments of the fitness
+distribution (e.g. the skew of the fitness distribution). While the ratchet
+should scale with the probability of fixation, the probability of fixation
+(Kimura, Malecot, XXX) is not accurate when multiple deleterious mutations are
+segregating simultaneously and in close linkage with one another. A key insight
+of \textcite{Santiago2016-mu} was that the classic equation for the probability
+of fixation is fairly robust as long as the correct effective population size
+is used. Their approach is to solve a system of two non-linear equations: one
+for the variance effective population size (which depends on the rate of the
+ratchet), and another for the time to fixation (which depends on the effective
+population size).
+
+Below, we step through the theory we use in our statistical methods. We
+re-derive central parts to illustrate where approximations have been made,
+which allows us to understand how these approximations may impact our
+statistical methods and inspect model fit in regions where these approximations
+are less likely to hold.
+
 \subsection*{Effective Population Size for Pairwise Diversity}
 
-When quantifying genetic variation with average pairwise diversity, often a
-single effective population size parameterizes the rate of pairwise
-coalescences of the entire evolutionary history. This is sufficient for
-constant-rate Kingman coalescences, but weak selection distorts genealogies
-(desai and nicolaison), creating a coalescent process with non-constant rates.
-Consequently, the expected pairwise coalescent time between two lineages $i$
-and $j$
+The central quantity of interest for our statistical inference methods is
+average pairwise diversity along the chromosome. While other summaries of the
+data, such as the allele frequency spectrum, may be more informative about the
+evolutionary processes acting in a region, we lack population genetic theory
+for these quantities under most selective regimes and they require more degrees
+of freedom to characterize. By contrast, neutral diversity under an
+infinite-sites model is determined only by the per-basepair mutation rate $\mu$
+and the expected pairwise coalescent time $\E[T_2]$,
 
 \begin{align}
-  \E(T) &= \sum_i T(i) \P(\mathrm{MRCA}(i)) \\
-        &= \sum_i T(i) \prod_{t=1}^i \left(1 - \frac{1}{2N_{e(i)}}\right)
+  \E(\pi) = 2 \mu \E(T_2)
+\end{align}
+
+Let us assume that there is a constant \emph{drift-effective} population size
+$N_d$ which determines the rate of coalescence due only to neutral demographic
+processes. Additionally, we assume some background selection process that in
+expectation rescales the drift-effective population size by $b(t)$ at time $t$
+(though this could be any process that modifies the coalescence rate). Then,
+the expected pairwise coalescent time is
+
+\begin{align}
+  \E(T_2) = \sum_{i=1}^\infty \frac{i}{2 b(t) N_d}\prod_{t=1}^{i-1} \left(1-\frac{1}{2 b(t) N_d}\right).
+\end{align}
+
+The time-dependency of $b(t)$ is needed because coalescence rates essentially
+speed up in the past under intermediate to weak selection, similar to
+population growth \parencite{Walczak2012-fi,Seger2010-rs,Good2014-yz}.
+Intuitively, this is because lineages are unlikely to coalesce with numerous
+less fit lineages they encounter as they go back in time, until ultimately
+coalescing in one of the least-loaded fitness classes.
+
+Previous work in the strong background selection domain
+\parencite{Charlesworth1993-gb,Nordborg1996-nq} approximates $\E(T_2)$ as $2
+f_0 N_d$, where $f_0 = e^{-\nicefrac{U}{s}}$ is size of the least-loaded
+fitness class. XXX on delay
+
+Here, we model the background selection process forward in time. We approximate
+the expected pairwise diversity with the probability of heterozygosity,
+$\E(\pi) \approx \mathcal{H}$, which holds under weak mutation and random
+mating, such that
+
+\begin{align}
+  %\mathcal{H} =  2 \mu \sum_{i=1}^\infty \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right)
+  \mathcal{H} &= 1 - \sum_{i=1}^\infty (1-\mu)^{2i} \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right) \\
+              &\approx 2 \mu \sum_{i=1}^\infty \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right)
+\end{align}
+
+in the limit where $\mu \to 0$. This is equation 4 of
+\textcite{Santiago2016-mu}; they define the sum the heterozygosity effective
+population size.
+
+Throughout, we assume that there is a constant \emph{drift-effective}
+population size, $N_d$, which determines the rate of coalescence due only to
+neutral demographic processes. We then consider the observed levels of neutral
+pairwise diversity as a reduction factor down from the drift effective
+population size, e.g. $\pi \approx 4 N_e \mu = 4 B N_d \mu$, where $B$ is a
+reduction factor due to the background selection process.
+
+\subsection*{Linked Selection Models for Variance-Effective Population Size}
+
+We wish to derive an expression for the reduction factor $B$. Ideally, this
+reduction factor would rescale the coalescent-effective population size, but
+this would require knowing the coalescent rate function $\lambda(t)$ over the
+evolutionary history $0 < t < \infty$. Instead, following previous work
+\parencite{Santiago1995-hx,Santiago1998-bs}, we derive the rescaling function
+$B$ relative to the variance-effective population size.
+
+Following past quantitative genetic models of linked selection
+\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu},
+we consider the long-run variance effective population size,
+
+\begin{align}
+  N_v := \lim_{t \to \infty} \frac{p_0(1-p_0)}{2\var\left(p_t - p_0\right)}
 \end{align}

### 2022-09-28 manuscript/manuscript.tex
few more minor changes

@@
 
-\subsection*{Effective Population Size for Pairwise Diversity}
+\subsection*{Pairwise Diversity under Background Selection}
 
@@
 in the limit where $\mu \to 0$. This is equation 4 of
-\textcite{Santiago2016-mu}; they define the sum the heterozygosity effective
+\textcite{Santiago2016-mu}; they define the sum as the heterozygosity effective
 population size.
 
-Throughout, we assume that there is a constant \emph{drift-effective}
-population size, $N_d$, which determines the rate of coalescence due only to
-neutral demographic processes. We then consider the observed levels of neutral
-pairwise diversity as a reduction factor down from the drift effective
-population size, e.g. $\pi \approx 4 N_e \mu = 4 B N_d \mu$, where $B$ is a
-reduction factor due to the background selection process.
+\subsection*{Forward in Time Models of Background Selection}
+
+Next, show how quantitative genetic models of linked selection
+\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu}
+model $b(t)$ forward in time.
+
+\begin{align}
+  b(t) = \frac{N_v}{N_d}
+\end{align}
+
+where
+
+\begin{align}
+  N_v = \lim_{t \to \infty} \frac{p_0 (1 - p_0) - \var(p_{t-1})}{2\var(p_t) - \var(p_{t-1})}
+\end{align}
 

### 2022-09-29 manuscript/manuscript.tex
more theory notes

@@
 processes. Additionally, we assume some background selection process that in
-expectation rescales the drift-effective population size by $b(t)$ at time $t$
-(though this could be any process that modifies the coalescence rate). Then,
-the expected pairwise coalescent time is
+expectation scales the effective population size by $b(t)$ at time $t$ (though
+this could in theory be due to any process). Then, the expected pairwise
+coalescent time is
 
 \begin{align}
+  \label{eq:T_2}
   \E(T_2) = \sum_{i=1}^\infty \frac{i}{2 b(t) N_d}\prod_{t=1}^{i-1} \left(1-\frac{1}{2 b(t) N_d}\right).
@@
 
-The time-dependency of $b(t)$ is needed because coalescence rates essentially
-speed up in the past under intermediate to weak selection, similar to
-population growth \parencite{Walczak2012-fi,Seger2010-rs,Good2014-yz}.
-Intuitively, this is because lineages are unlikely to coalesce with numerous
-less fit lineages they encounter as they go back in time, until ultimately
-coalescing in one of the least-loaded fitness classes.
+The time-dependency of $b(t)$ is required because the background selection
+process will usually alter the rate of coalesce through time. For example,
+under weak and intermediate selection, coalesce rates speed up as lineages move
+in the past under intermediate to weak selection, similar to population growth
+\parencite{Walczak2012-fi,Seger2010-rs,Good2014-yz}.
 
-Previous work in the strong background selection domain
-\parencite{Charlesworth1993-gb,Nordborg1996-nq} approximates $\E(T_2)$ as $2
-f_0 N_d$, where $f_0 = e^{-\nicefrac{U}{s}}$ is size of the least-loaded
-fitness class. XXX on delay
+% Intuitively, this is
+% because lineages are unlikely to coalesce with numerous less fit lineages they
+% encounter as they go back in time, until ultimately coalescing in one of the
+% least-loaded fitness classes.
 
-Here, we model the background selection process forward in time. We approximate
-the expected pairwise diversity with the probability of heterozygosity,
-$\E(\pi) \approx \mathcal{H}$, which holds under weak mutation and random
-mating, such that
+This formulation is general and classic background selection theory (in the
+no-ratchet domain) is a special case of it. Under classic BGS and load theory,
+two randomly sampled lineages will each carry a number of deleterious mutations
+sampled from a Poisson stationary distribution. Because any lineage carrying $j
+> 0$ strongly deleterious mutations is destined to extinction, these lineages
+can only coalesce in a common ancestor in the least-loaded class with zero
+deleterious mutations. Under the Poisson stationary distribution, the size of
+this class is $f_0 N$, where $f_0 = e^{-\nicefrac{U}{s}}$, so the coalescent
+rate is scaled by $\nicefrac{1}{f_0}$. Once the lineages reach this fitness
+class, there are no fitness differences among individuals, so they coalesce
+following the Kingman process. A lineage moves backwards in time $t_d$
+generations through the ``delay phase" to the least-loaded class. Classic BGS
+theory ignores this delay phase, since $t_d \sim \log(j+1)/s << f_0 N$
+(Durrett, p. 213).
 
-\begin{align}
-  %\mathcal{H} =  2 \mu \sum_{i=1}^\infty \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right)
-  \mathcal{H} &= 1 - \sum_{i=1}^\infty (1-\mu)^{2i} \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right) \\
-              &\approx 2 \mu \sum_{i=1}^\infty \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right)
-\end{align}
+%Here, we model the background selection process forward in time. We approximate
+%the expected pairwise diversity with the probability of heterozygosity,
+%$\E(\pi) \approx \mathcal{H}$, which holds under weak mutation and random
+%mating, such that
 
-in the limit where $\mu \to 0$. This is equation 4 of
-\textcite{Santiago2016-mu}; they define the sum as the heterozygosity effective
-population size.
+%\begin{align}
+%  %\mathcal{H} =  2 \mu \sum_{i=1}^\infty \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right)
+%  \mathcal{H} &= 1 - \sum_{i=1}^\infty (1-\mu)^{2i} \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right) \\
+%              &\approx 2 \mu \sum_{i=1}^\infty \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right)
+%\end{align}
 
-\subsection*{Forward in Time Models of Background Selection}
+%in the limit where $\mu \to 0$. This is equation 4 of
+%\textcite{Santiago2016-mu}; they define the sum as the heterozygosity effective
+%population size.
+
+\subsection*{Forward in Time Models of Linked Selection}
 
@@
 \parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu}
-model $b(t)$ forward in time.
+essentially model $b(t)$ with an analogous forward in time process. Their
+approach is to consider how the variance in neutral allele frequency change is
+impacted by a linked selection process. Under a Wright--Fisher model, the
+variance in allele frequency change is also determined by a pairwise
+coalescence rate, often expressed as a \emph{variance-effective} population
+size,
 
 \begin{align}
-  b(t) = \frac{N_v}{N_d}
+  \V(p_t) = p_0(1-p_0)\left(1 - \prod_{i=1}^t \left(1-\frac{1}{2N_e(t)}\right)\right)
 \end{align}
 
-where
+where if $N_e(t) = N_e$ and $t=1$, this is the familiar Wright--Fisher variance
+$\V(p_1) = p_0(1-p_0) / 2N_e$. When selection is not acting, allele frequency
+changes per generation are independent, and neutral diversity is set by this
+single $N_e$.
+
+XXX
+
+Since we track an arbitrary neutral allele's frequency $p_i$, by symmetry,
+linked selection has no impact on the direction of change, $\E(\delta p_i) =
+0$. We derive an expression for $\V(p_t)$ by induction following the approach
+of \textcite{Santiago1995-hx}. In the first generation, the frequency change
+can be partitioned into three components. Conditioning on the parent's
+heritable fitnesses, we can imagine constructing an infinite number of
+evolutionary replicates. Across all replicates, a random neutral allele's is
+pushed
+
+\begin{align}
+  p_1 - p_0 = \sum_{i=1}^{N} 
+\end{align}
+
+Since the variance in neutral allele frequency change is also determined by the
+pairwise coalescence rate, this provides us with a way to model $b(t)$ under
+linked selection.
+
+These models derive an expression for the variance-effective population size of
+a focal neutral site, $N_v(t)$, based on how linked fitness variation perturbs
+the frequency trajectory at this site.  The variance-effective population size
+for a constant $N_e$ is defined by the variance in allele frequency change at
+the neutral site, over evolutionary replicates,
+
+where if $t=1$, this is the familiar Wright--Fisher variance $\V(p_1) =
+p_0(1-p_0) / 2N_e$. When selection is not acting, allele frequency changes per
+generation are independent, and neutral diversity is set by this
+single $N_e$.
+
+However when selection is acting at linked sites, neutral diversity is
+determined by the long-run $\V(p_t)$ as $t \to \infty$. Intuitively, this is
+because the rate of drift is magnified by autocorrelation in allele frequency
+change (XXX), and diversity is determined by the cumulative impact of these
+autocorrelations in the long-run. This autocorrelation effectively reduces the
+coalescence rate in a time-dependent manner, analogous to equation
+\eqref{eq:T_2}. Thus, a single constant $N_e$ is insufficient to describe
+levels of neutral variation, since it is insufficient
+
+The variance in allele frequency change after $t$ generations is
+
+\begin{align}
+  \V(p_t) = p_0(1-p_0)\left(1 - \prod_{i=1}^t \left(1-\frac{1}{2N_e(t)}\right)\right)
+\end{align}
+
+This is the basis of the quantitative genetic models of linked selection: a
+forward in time model for $\V(p_t)$ implies a pairwise coalescence rate
+function $b(t)$, which determines statistics based on pairwise coalescence such
+as $\E(\pi)$.
+
+Next, we consider how fitness variation impacts $\V(p_t)$.
+
+Throughout, we will take the variance $\V(p_t)$ to be the variance over
+evolutionary replicates. Since the variance is for a random neutral allele's
+frequency,
+
+a reasonable approximation to how background selection process
+rescales the coalescent rate at this site is given by $b(t) =
+\nicefrac{N_v(t)}{N_d}$.
 

### 2022-09-30 manuscript/manuscript.tex
more rough notes

@@
   \label{eq:T_2}
-  \E(T_2) = \sum_{i=1}^\infty \frac{i}{2 b(t) N_d}\prod_{t=1}^{i-1} \left(1-\frac{1}{2 b(t) N_d}\right).
+  \E(T_2) = \sum_{i=1}^\infty \frac{i}{2 b(i) N_d}\prod_{t=1}^{i-1} \left(1-\frac{1}{2 b(t) N_d}\right).
 \end{align}
@@
 \parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu}
-essentially model $b(t)$ with an analogous forward in time process. Their
-approach is to consider how the variance in neutral allele frequency change is
-impacted by a linked selection process. Under a Wright--Fisher model, the
-variance in allele frequency change is also determined by a pairwise
-coalescence rate, often expressed as a \emph{variance-effective} population
-size,
+find expressions for $b(t)$ with by modeling the selective process forward in
+time. Their approach is to consider how the variance in neutral allele
+frequency change is impacted by a linked selection process.
+
+% First, we will show
+% how the variance in allele frequency is related to the k
+
+For comparison, we start with the recursion for the variance in allele
+frequency change in a neutral Wright--Fisher model. Under binomial sampling of
+alleles, the random frequency $x_t$ is $2N x_t \sim \mathrm{Binom}(x_{t-1},
+2N)$. If we look forward in time on generation and treat $x_0$ as fixed,
 
 \begin{align}
-  \V(p_t) = p_0(1-p_0)\left(1 - \prod_{i=1}^t \left(1-\frac{1}{2N_e(t)}\right)\right)
+  \V(x_1) = \E_1\left[(x_1 - x_0)^2\right]= \frac{x_0(1-x_0)}{2N},
 \end{align}
+
+and
+
+\begin{align}
+  \V(x_2) &= \E\left[(x_2 - x_0)^2\right] \\
+          &= \E_1\left[\E_2\left[((x_2-x_1) + (x_1 - x_0))^2 | x_1\right]\right] \\
+          &= \E_1\left[\E_2\left[(x_2-x_1)^2|x_1\right]\right] + \E_1\left[ (x_1 - x_0)^2\right] \\
+          &= \frac{x_0(1-x_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{x_0(1-x_0)}{2N}
+\end{align}
+
+such that in general,
+
+\begin{align}
+  \label{eq:var_freq}
+  %\V(x_t) &= \frac{x_0(1-x_0)}{2N}\left(\sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i-1} \right)
+  \V(x_t) &= \V(x_1) \left(1 + \sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i} \right)
+\end{align}
+
+This geometric series converges to the familiar expression,
+
+\begin{align}
+  \V(x_t) &= x_0(1-x_0)\left(1 - \left(1 - \frac{1}{2N}\right)^t \right).
+\end{align}
+
+which in the limit $t \to \infty$, converges to the Bernoulli variance
+$x_0(1-x_0)$ as the frequency either goes to 1 with probability $x_0$ or 0 with
+probability $1-x_0$.
+
+Now, if the effective population size were to change each generation, equation
+\eqref{eq:var_freq} would be,
+
+\begin{align}
+  \V(x_t) &= \V(x_0) \left(1 + \sum_{i=1}^t \prod_{j=1}^{i} \left(1 - \frac{1}{2N(j)}\right) \right)
+\end{align}
+
+Note that this too converges to $x_0(1-x_0)$ as $t \to \infty$.
+
+Note now that equation 4 is a special case, when the evolutionary process has
+independent allele frequency changes through the generations. Selection
+processes will often lead to correlations in $\delta_i$ and $\delta_j$ for $i
+\ne j$. These correlations are because the neutral allele in generations $i$
+and $j$ can share some fraction of their fitness background, and in so far as
+the selective pressures are the same in generations $i$ and $j$, selection will
+perturb the frequency trajectories in the same direction.
+
+One way we can model these dependencies is to partition $\delta_i$ into drift
+and selection components. Over evolutionary replicates, the drift components
+are independent and the selection components have some degree of
+autocorrelation. Consider $\delta x_i = \delta_S x_i + \delta_D x_i$
+
+\begin{align}
+  \V(x_2) &= \V\left(\delta x_2 + \delta x_1\right) \\
+          &= \V(\delta_D x_2) + \V(\delta_D x_1) + 2 \cov(\delta_S x_2, \delta_S x_1) \\
+          &= \frac{x_0(1-x_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{x_0(1-x_0)}{2N} +  2 \cov(\delta_S x_2, \delta_S x_1) \\
+\end{align}
+
+since the drift changes are the familiar ones of equation XXX. The covariances
+are determined by the selective process dynamics, linkage disequilibrium, and
+the recombination rate.
+
+Let us imagine there is some stochastic process $b(t)$ that rescales the coalescent
+rate,
+
+\begin{align}
+  \V(x_t) &= \frac{x_0(1-x_0)}{2N(0)}\left(\sum_{i=1}^t \prod_{j=1}^{i-1} \left(1 - \frac{1}{2b(j)N}\right) \right)
+\end{align}
+
+\begin{align}
+  \var(p_3 - p_0) = \E & \left( \bigg( \underbrace{S_1(1 + (1-r) G + (1-r)^2 G^2)}_{\text{creation and persistence of generation 1 associations} \;\; := \; S_1 Q_3} \right. +  D_1 + H_1+ \nonumber \\
+                       & \underbrace{S_2(1 + (1-r) G)}_{\text{creation and persistence of generation 2 associations} \;\; := \; S_2 Q_2}+ D_2 + H_2 + \nonumber \\
+                       &  \underbrace{S_3}_{\text{creation of generation 3 associations} \;\; := \; S_3 Q_1  }\left. + D_3 + H_3 \bigg)^2 \right). &
+    \label{eq:scr-var2}
+\end{align}
+%
+
+Note the similarity between the equations.
+
+\begin{align}
+  V &\propto \frac{1}{2b(0)N_d}\left(\sum_{i=1}^\infty \prod_{t=1}^{i-1} \left(1 - \frac{1}{2b(t)N_d}\right) \right) \\
+  W &\propto \sum_{i=1}^\infty \frac{i}{2 b(i) N_d} \prod_{t=1}^{i-1} \left(1-\frac{1}{2 b(t) N_d}\right) \\
+\end{align}
+
+Let us define the function $a(t)$ to be the product, which is the cumulative
+failure to coalesce up to time $t$.
+
+\begin{align}
+  a(t) = \prod_{i=1}^{t-1} \left( 1-\frac{1}{2 b(i) N_d} \right)
+\end{align}
+
+Comparing this to equation \eqref{eq:T_2} up to constant factors,
+
+\begin{align}
+  V &\propto \sum_{i=1}^\infty a(t) \\
+  W &\propto \sum_{i=1}^\infty i \; a(t) \\
+\end{align}
+
+end{align}
+
+XXX we swap indices
+
+Now, if we assume that the final coalesce in the waiting time process were to
+have a fixed rate of $\nicefrac{1}{2b(0)N_d}$, this
+
+we see that both the waiting time process and the variance process
+
+Intuitively, these define a stochastic selection process in two ways through
+the function $b(t)$ that modulates the rate of pairwise coalesce.
 
@@
 Since we track an arbitrary neutral allele's frequency $p_i$, by symmetry,
-linked selection has no impact on the direction of change, $\E(\delta p_i) =
-0$. We derive an expression for $\V(p_t)$ by induction following the approach
-of \textcite{Santiago1995-hx}. In the first generation, the frequency change
+linked selection has no impact on the direction of change, $\E(\Delta p_i) =
+0$. We will work through a short proof of $\V(p_t)$ by induction following the
+approach of \textcite{Santiago1995-hx}. Throughout, we will use angular
+brackets to indicate an average over all diploid individuals in the population,
+of some value with an individual-level index. For example, if we denote the
+number of allele an individual carries in generation $t$ as $x_{t,i} \in \{0,
+1, 2\}$, the population frequency at time $p_t$ is
+
+\begin{align}
+  p_t = \langle x_t \rangle = \frac{1}{N} \sum_{i=1}^N x_{t,i}
+\end{align}
+
+\begin{align}
+  \Delta_H p &= \langle x (f-1) \rangle \\
+  \Delta_N p &= \langle x d \rangle \\
+  \Delta_M p &= \langle \delta \rangle \\
+\end{align}
+
+Assuming uninherited, random environments, only the $\Delta_{_H} p$ terms
+create autocorrelation in $p_t$. The autocorrelation is created because during
+the random pairing of neutral alleles to fitness backgrounds, associations are
+created between an individual $i$'s allele frequency and their fitness $f_i$.
+This is because both fitness and the neutral alleles reside in the same
+individual in the present generation.
+
+We can imagine now how these associations change over evolutionary replicates
+in the next generation. In a fraction $r$ of the cases, the neutral allele 
+recombines off (or segregates away, if they are on different chromosomes) the
+fitness background. In a faction of $1-r$ XXX
+
+\begin{align}
+  \cov(\Delta_H p_t, \Delta_H p_s) &= \E(\langle x_t (f_t-1) \rangle , \langle x_s (f_s-1) \rangle)
+\end{align}
+
+Note that this covariance is only
+
+In the first generation, the frequency change
 can be partitioned into three components. Conditioning on the parent's
@@
 evolutionary replicates. Across all replicates, a random neutral allele's is
-pushed
+perturbed
 

### 2022-10-02 manuscript/manuscript.tex
more writing

@@
 \RequirePackage{fullpage}
-\RequirePackage[font=small,labelfont=bf]{caption}
+%\RequirePackage[font=small,labelfont=bf]{caption}
 \RequirePackage{amsmath,amssymb,amsthm}
@@
 \RequirePackage{bbm}
-\RequirePackage[osf]{mathpazo}
+%\RequirePackage[osf]{mathpazo}
+\let\temp\rmdefault
+\RequirePackage{mathpazo}
+\let\rmdefault\temp
+
 \RequirePackage[bibstyle=authoryear,citestyle=authoryear-comp,
@@
 
-\section*{Theory}
+\section*{Results}
+
+\subsection*{Theory}
+
+- overview of why BGS models fail?
+
+% could be in intro too:
+
+As genomes are transmitted through the generations, lifetime events selectively
+alter what is passed to the next generation. While selection truly acts on the
+genetic variants that directly cause heritable fitness differences among
+individuals, neighboring regions experience an indirect contraction in genetic
+variability. This byproduct is known as \emph{linked selection}, and it occurs
+when selection changes a site's frequency on timescale faster than
+recombination can decouple that site from its genetic background. Because
+linked selection modifies the genetic variability in a regionally-correlated
+way, it creates patterns in genomic data that are used to understand the types
+of selection processes acting on genomes.
+
+Much work has been done on developing mathematical theory for two primary
+selection processes: hitchhiking and background selection. In turn, statistical
+methods have use this theory for a variety of wide-ranging empirical problems.
+Early on, researchers found correlations between recombination rates and
+genetic diversity in small genomes with limited recombination like
+\emph{Drosophila} (XXX), consistent with linked selection processes acting on
+genome-wide scales. With population genomic data, small-scale (XXX kb) dips in
+genetic diversity was evidence of rare but strong selective sweeps in humans
+(XXX) and other species (XXX).
+
+models of background selection can
+closely predict levels of human-chimp divergence at the megabase scale (XXX).
+
+humans (XXX), though these offered only a partial picture of the selection
+processes acting (XXX).
+
+Linked
+selection studies have found additional strong selective sweeps in taxa across
+the kingdom of life (XXX), and attention turned towards estimating the relative
+contributions of the hitchhiking and background processes (Eyalshiv).
+processes (Eyalshiv).
+
+From an evolutionary perspective, models of linked selection
+
+The other use of these models is to understand how linked selection processes
+differentially alter the genetic variability, or equivalently, the genealogical
+trees, along the genome.
+
+Early on, some pointed out that evidence of strong selective sweeps in the
+human genome was not the full story
+
+Generally, in organisms with large amounts of recombination, linked selection
+has
+
+Weak selection processes
+
+distorts
+genome-wide
+
+diversity.
+
+The linked selection literature has primarily focused on developing theory for,
+and quantifying the effects of two linked selection processes: hitchhiking and
+background selection. Genome-wide models of the hitchhiking process consider
+how beneficial mutations that sporadically enter the population
+
+- QGLS polygenic linked selection
+
+- H and W
+
+- inflation factor
+
+\subsection*{Statistical Methods}
+
+\section*{Appendix}
+
+\subsection*{Overview of Theory}
 
@@
 
-% First, we will show
-% how the variance in allele frequency is related to the k
+\subsubsection*{Variance in Neutral Allele Frequency in a Neutral System}
 
@@
           &= \E_1\left[\E_2\left[((x_2-x_1) + (x_1 - x_0))^2 | x_1\right]\right] \\
-          &= \E_1\left[\E_2\left[(x_2-x_1)^2|x_1\right]\right] + \E_1\left[ (x_1 - x_0)^2\right] \\
+          &= \E_1\left[\E_2\left[(\Delta x_2 + \Delta x_1)^2 | x_1\right]\right] \\
+          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + 2\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
+          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
           &= \frac{x_0(1-x_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{x_0(1-x_0)}{2N}
@@
 
-such that in general,
+Note that the cross-moment $\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right]
+= 0$ under neutral evolution. In general, the variance in allele frequency
+after $t$ generations of neutral evolution is,
 
@@
 
-Now, if the effective population size were to change each generation, equation
-\eqref{eq:var_freq} would be,
+% Now, if the effective population size were to change each generation, equation
+% \eqref{eq:var_freq} would be,
+
+% \begin{align}
+%   \V(x_t) &= \V(x_0) \left(1 + \sum_{i=1}^t \prod_{j=1}^{i} \left(1 - \frac{1}{2N(j)}\right) \right)
+% \end{align}
+
+% Note that this too converges to $x_0(1-x_0)$ as $t \to \infty$.
+
+\subsubsection*{Variance in Neutral Allele Frequency with Linked Selection}
+
+Note that equation XXX is a special case, when the evolutionary process has
+independent allele frequency changes through the generations and thus the
+cross-moment is zero. Selection generates correlations between allele frequency
+changes at nearby neutral sites. These correlations created by linked selection
+inflate the variance in neutral allele frequency $\V(x_t)$ beyond what it would
+be in a system without linked selection operating. The magnitude of this
+inflation term is set by the selection process and its dynamics, linkage
+disequilibrium, and the recombination rate. Even k
+
+The theory of this began with \textcite{Robertson1961-ho},
+and was thoroughly worked out by Santiago and Caballero
+(\citeyear{Santiago1995-hx}, \citeyear{Santiago1998-bs}; see also
+\cite{Barton2000-zg} and \cite{Buffalo2019-qs}). Our BGS statistical methods
+employ the extension of this model for weak BGS \parencite{Santiago2016-mu}.
+For completeness, we step through a quick derivation of the linked selection
+inflation factor models here, based on
+\textcite{Santiago1995-hx,Santiago1998-bs}.
+
+Quantitative genetic models of linked selection
+
+%%%%%
+
+One way we can model these dependencies is to partition $\Delta_i$ into drift
+and selection components. Over evolutionary replicates, the drift components
+are independent and the selection components have some degree of
+autocorrelation. Let us partition $\Delta x_i = \Delta_S x_i + \Delta_D x_i$,
 
 \begin{align}
-  \V(x_t) &= \V(x_0) \left(1 + \sum_{i=1}^t \prod_{j=1}^{i} \left(1 - \frac{1}{2N(j)}\right) \right)
+  \V(x_2) &= \V\left(\Delta x_2 + \Delta x_1\right) \\
+          &= \V(\Delta_D x_2) + \V(\Delta_D x_1) + 2 \cov(\Delta_S x_2, \Delta_S x_1) \\
+          &= \frac{x_0(1-x_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{x_0(1-x_0)}{2N} +  2 \cov(\Delta_S x_2, \Delta_S x_1) 
 \end{align}
 
-Note that this too converges to $x_0(1-x_0)$ as $t \to \infty$.
-
-Note now that equation 4 is a special case, when the evolutionary process has
-independent allele frequency changes through the generations. Selection
-processes will often lead to correlations in $\delta_i$ and $\delta_j$ for $i
-\ne j$. These correlations are because the neutral allele in generations $i$
-and $j$ can share some fraction of their fitness background, and in so far as
-the selective pressures are the same in generations $i$ and $j$, selection will
-perturb the frequency trajectories in the same direction.
-
-One way we can model these dependencies is to partition $\delta_i$ into drift
-and selection components. Over evolutionary replicates, the drift components
-are independent and the selection components have some degree of
-autocorrelation. Consider $\delta x_i = \delta_S x_i + \delta_D x_i$
+since $\Delta_D x_i \perp \Delta_D x_j$ for $i \ne j$. The covariances are
+determined by the selective process dynamics, linkage disequilibrium, and the
+recombination rate. These covariances are throughly worked out in
+\textcite{Santiago1995-hx} and \textcite{Buffalo2019-qs}, so we give a short
+derivation here. Throughout, we assume random mating, hermaphroditic
+individuals, and a constant population size. We will use angular brackets to
+indicate an average over all diploid individuals in the population, of some
+value with an individual-level index. For example, if we denote the number of
+allele an individual carries in generation $t$ as $z_{t,i} \in \{0, 1, 2\}$,
+the population frequency at time $x_t$ is
 
 \begin{align}
-  \V(x_2) &= \V\left(\delta x_2 + \delta x_1\right) \\
-          &= \V(\delta_D x_2) + \V(\delta_D x_1) + 2 \cov(\delta_S x_2, \delta_S x_1) \\
-          &= \frac{x_0(1-x_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{x_0(1-x_0)}{2N} +  2 \cov(\delta_S x_2, \delta_S x_1) \\
+  x_t = \langle z_t \rangle = \frac{1}{2N} \sum_{i=1}^N z_{t,i}
 \end{align}
 
-since the drift changes are the familiar ones of equation XXX. The covariances
-are determined by the selective process dynamics, linkage disequilibrium, and
-the recombination rate.
+Across evolutionary replicates, each parent $i$ passes one of their two neutral
+alleles to a random number $k_i$ offspring. The \emph{heritable fitness} of an
+individual $f_i$ is defined as the expected number of offspring passed down
+across evolutionary replicates, such that $k_i = f_i + \varepsilon_i$.  Some
+selection process determines the distribution of $f_i$, but the this occurs
+independently of the process that sorts neutral alleles into diploid
+individuals. We are unsure of the neutral allele count in an individual in the
+initial generation, $z_{t,0}$, so we condition on it and take expectations
+later. Then, the frequency in the initial and first generations are,
 
+\begin{align}
+  x_0 = \langle z_0 \rangle = \frac{1}{2N} \sum_{i=1}^N z_{0,i} \\
+  x_1 = \langle z_1 \rangle = \frac{1}{2N} \sum_{i=1}^N \left( k_i \frac{z_{0,i}}{2} + \sum_{j=1}^{k_i} \delta_{i,j} \right)
+\end{align}
+
+The first term is the average number of neutral alleles each parent passes to
+its $k_i$ offspring. In reality, alleles are discrete, so each heterozygous
+parent also passes a random $\pm \nicefrac{1}{2}$ to each offspring, which
+leads to the second term. The variable $\delta_{i,j}$ is 0 if the parent is a
+homozygote, but if the parent is a heterozygote, is $\pm \nicefrac{1}{2}$ with
+equal probability over evolutionary replicates.
+
+Since individuals are exchangeable with their contribution to the
+heterozygosity term and population size is constant,
+
+\begin{align}
+  \sum_{i=1}^N \sum_{j=1}^{k_i} \delta_{i,j} = \sum_{i=1}^N \delta_{i} 
+\end{align}
+
+We can further partition the offspring number into heritable and non-heritable
+parts,
+
+\begin{align}
+  x_1 &= \frac{1}{2N} \sum_{i=1}^N \left( (f_i + \varepsilon_i) \frac{z_{0,i}}{2} + \delta_{i} \right) \\
+      &= \frac{1}{4N} \sum_{i=1}^N f_i z_{0,i} + \frac{1}{4N} \sum_{i=1}^N \varepsilon_i z_{0,i} + \frac{1}{2N} \sum_{i=1}^N \delta_{i} \\
+      &=  \frac{1}{4} \langle f \; z_{0} \rangle + \frac{1}{4} \langle \varepsilon \; z_{0} \rangle + \frac{1}{2}\langle \delta \rangle \\
+\end{align}
+
+% TODO factor 1/2 error
+
+\begin{align}
+  x_1 - x_0 &= \frac{1}{4} \langle f \; z_{0} \rangle + \frac{1}{4} \langle \varepsilon \; z_{0} \rangle + \frac{1}{2}\langle \delta \rangle - \langle z_0 \rangle \\
+            &= \frac{1}{4} \langle (f-1) \; z_{0} \rangle + \frac{1}{4} \langle \varepsilon \; z_{0} \rangle + \frac{1}{2}\langle \delta \rangle 
+\end{align}
+
+Now, we look at the variance of this over evolutionary replicates. By symmetry,
+the expected change in a neutral allele is $\E(\Delta x_t) = 0$ since we track
+an arbitrary neutral allele. Then the variance $\V(\Delta x_1) = \E( {\Delta
+x_1}^2)$,
+
+\begin{align}
+  \V(\Delta x_1) &= \frac{1}{4} \langle (f-1) \; z_{0} \rangle + \frac{1}{4} \langle \varepsilon \; z_{0} \rangle + \frac{1}{2}\langle \delta \rangle 
+\end{align}
+
+OLD
 Let us imagine there is some stochastic process $b(t)$ that rescales the coalescent
@@
   \Delta_N p &= \langle x d \rangle \\
-  \Delta_M p &= \langle \delta \rangle \\
+  \Delta_M p &= \langle \Delta \rangle \\
 \end{align}
@@
 \begin{align}
-  V(f) &\approx s^2 \V_a(n) + \delta_\text{LD}
+  V(f) &\approx s^2 \V_a(n) + \Delta_\text{LD}
 \end{align}

### 2022-10-02 manuscript/manuscript.tex
more writing

@@
 Much work has been done on developing mathematical theory for two primary
-selection processes: hitchhiking and background selection. In turn, statistical
-methods have use this theory for a variety of wide-ranging empirical problems.
-Early on, researchers found correlations between recombination rates and
-genetic diversity in small genomes with limited recombination like
-\emph{Drosophila} (XXX), consistent with linked selection processes acting on
-genome-wide scales. With population genomic data, small-scale (XXX kb) dips in
-genetic diversity was evidence of rare but strong selective sweeps in humans
-(XXX) and other species (XXX).
+selection processes: hitchhiking (linked selection due to beneficial alleles
+increasing in frequency) and background selection (linked selection due to the
+selection against deleterious alleles). In turn, statistical methods have use
+this theory for a variety of wide-ranging empirical problems. Early on,
+researchers found correlations between recombination rates and genetic
+diversity in organisms with limited recombination like \emph{Drosophila} (XXX),
+consistent with linked selection processes acting on genome-wide scales. With
+population genomic data, evidence of small-scale (XXX kb) dips in genetic
+diversity was found, caused by rare but strong selective sweeps in humans (XXX)
+and other species (XXX). Remarkably, a simple mathematical model of background
+selection could closely predict levels of human-chimp divergence at the
+megabase scale (XXX) from genome-wide patterns of recombination and regions
+conserved between species (McVicker). Another central empirical goal is to
+disentangle the relative effect of the these two selection processes on levels
+of genetic diversity (Eyalshiv).
 
-models of background selection can
-closely predict levels of human-chimp divergence at the megabase scale (XXX).
+Additionally, linked selection can impact other evolutionary processes in two
+critical ways. First, while most evolutionary models consider genetic drift to
+be the main source of random fluctuations in neutral allele frequencies, linked
+selection has a similar impact on neutral sites, increasing the variance in
+allele frequency change as if the population size were effectively smaller.
+Second, by perturbing the frequencies of neighboring alleles, selection can
+interfere with their selection dynamics, known as Hill-Robertson interference
+(XXX).
 
-humans (XXX), though these offered only a partial picture of the selection
-processes acting (XXX).
+even when sites under selection are sparse.
 
-Linked
-selection studies have found additional strong selective sweeps in taxa across
-the kingdom of life (XXX), and attention turned towards estimating the relative
-contributions of the hitchhiking and background processes (Eyalshiv).
-processes (Eyalshiv).
+From an evolutionary perspective, linked selection is important because by
+modifying neighboring levels of genetic diversity, it is also directly altering
+the selection dynamics at these neighboring sites.
 
-From an evolutionary perspective, models of linked selection
+Typically evolutionary genetic models consider
+
+This is known as selective interes
 
@@
 trees, along the genome.
+
+XXX how many sites are under selection at any moment? What's fitness variation?
 

### 2022-10-03 manuscript/manuscript.tex
more intro notes

--- 2022-10-02 manuscript/manuscript.tex
@@
 
-% A central goal of evolutionary genetics is to quantify the relative roles
-% different evolutionary processes have on genetic variation. Certain
-% evolutionary processes, such as demographic bottlenecks or expansions, act
-% uniformly across the genome as a whole. Other processes, such as selection, are
-% hypothesized to act differently across the genome depending on the density of
-% genomic features, and local heterogeneity in recombination and mutation rate.
-% With the emergence of population genomic data, a variety of statistical models
-% have teased apart the contribution of these various processes. These
-% statistical models differ in how they reduce large genomic data to summary
-% statistics, as well as in their statistical methodology. Some models reduce
-% genomic data to the histograms of the frequency spectrum (XXX), while others
-% model scalar summaries of the frequency spectrum along the genome.
+\section*{Introduction}
 
-% One class of models, parametric statistical models, leverage population genetic
-% theory to inform expectations of the mean pairwise diversity along the
-% chromosome. Estimation under these models often uses a composite likelihood
-% approach, where the joint distribution is assumed independent across positions.
-% While composite likelihoods are simpler to formulate and lead to consistent
-% parameter estimation, ignoring dependencies between neighboring sites has two
-% primary disadvantages. First, unlike proper likelihoods, the composite
-% likelihood's curvature is uninformative about parameter estimate uncertainty.
-% Previous work has circumvented this shortcoming through (1) resampling-based
-% uncertainty estimates, or (2) adjusting the composite likelihood's curvature to
-% account for these dependencies. Second, similar to inference under phylogenetic
-% comparative models, the dependency structure tells us how to properly weight
-% information across sites. For example, imagine a conserved segment 10 kilobases
-% long right on the border of two genomic bins, which due to linked selection,
-% bot experience a reduction in diversity. When estimating parameters, the
-% independence assumption of composite likelihoods would weight these observed
-% reductions equal to the amount of a reduction observed on an entirely different
-% independent chromosome.
+As genomes are transmitted through the generations, lifetime events selectively
+alter what is passed to the next generation. While selection directly acts on
+the genetic variants that cause heritable fitness differences among
+individuals, neighboring regions experience an indirect contraction in genetic
+variability. This byproduct is known as \emph{linked selection}, and it occurs
+when selection changes a site's frequency on timescale faster than
+recombination can decouple that site from its genetic background. Because
+linked selection modifies the genetic variability in a regionally-correlated
+way, it creates observable patterns in genomic data that are used to understand
+the type of selection processes acting on genomes and estimate key evolutionary
+quantities.
 
-% patterns of genetic variability along the genome are treated as
-% coming from an indeprocess $m(z)$.
+Much work has been done on developing mathematical theory for two linked
+selection processes: hitchhiking and background selection. What distinguishes
+these are the evolutionary processes that generate fitness variation, and the
+resulting selection dynamics. Classic hitchhiking models assume new beneficial
+variants occur and sweep to fixation (XXX). Theory predicts that classic sweeps
+affect diversity on small spatial scales, with characteristic length
+$\nicefrac{s}{r}$ basepairs, where $s$ is the heterozygous selection
+coefficient and $r$ is the per-basepair recombination rate. Sweeps were found
+in humans, \emph{Drosophila}, and other species (XXX), though it was unclear
+whether hitchhiking alone could explain the observed genome-wide correlation
+between recombination rate and genetic diversity in some species (XXX). Since
+the sweeps leave relatively narrow footprint, they influence genome-wide
+diversity only when beneficial mutations are recurring and recombination is
+relatively limited, as in \emph{Drosophila}.
 
-% Here, we focus on solving an issue processes: background
+By contrast, background selection occurs because of the steady influx of new
+deleterious mutations into conserved regions each generation. Since far more
+new mutations are deleterious than beneficial in functional regions, it would
+seem like background selection would have a dominant role in shaping genetic
+diversity. However, deleterious mutations segregate on a timescale inversely
+proportional to their selection cost (XXX). Strongly deleterious mutations
+eliminate neighboring genetic variability on the few chromosomes they may
+recombine onto before the eventual extinction, and consequently have very weak
+effects on larger spatial scales. Moderately weakly selected mutations can
+persist for much longer periods, allowing for more recombination before their
+eventual loss and produce reductions in diversity on shorter spatial scales.
+The reduction in diversity under background selection is determined
+approximately by $\nicefrac{U}{R}$, where $U = \mu_D L$ is the number of
+deleterious mutations (at rate $\mu_D$ per basepair, per generation) in a
+conserved region of $L$ basepairs per generation, and $R = r L$, which is the
+total recombination per generation in Morgans. Since the deleterious mutation
+rate is relatively constant across conserved regions, the reduction due to
+background selection closely matches the composition of conserved regions and
+recombination along the genome. Remarkably, simple mathematical models of
+background selection closely predicts levels of human-chimp divergence at the
+megabase scale (XXX) from genome-wide patterns of recombination and regions
+conserved between species (McVicker).
 
-% BGS as evolution around the static features of the genome. Due to the slow rate
-% of genomic structural evolution, much of the genome can be treated as static
-% throughout the genome. These regions are under differing levels of selective
-% constraint, which induces background selection at neighboring regions.
+This leads to some of the unidentifiability of hitchhiking versus background
+selection.
 
-% Signal processing
+and the reduction is determined
+by
 
-% Genomes have inherent structure that impacts the linked selection map. There is
-% tremendous heterogeneity across species in the organization of likely selected
-% regions on the genetic and physical maps.
+While the rate
+new beneficial mutations enter the population is low (XXX),
 
-% Within an organism, there are two key determinants of B maps: recombination and
-% gene density.
+The spatial scale of background selection is
 
-% Initially, the relationship between recombination and diversity was regressed
-% with no regard to the spatial organization of the genome.
+kkkkkk
 
-% Certain regions are pinchpoints of selection: some combination of low
-% recombination and high gene density lead to very strong reductions in genetic
-% diversity. In these regions, coalescent dynamics change considerably; where as
-% under strong background selection, deleterious alleles would not fix, in these
-% regions the can fix. Additionally, the selective interference impacts positive
-% selection.
+not all
+adaptation occurs from such new beneficial mutations sweeping to fixation (XXX)
 
+Other processes also act
+on beneficial fitness variation to engender linked selection, such as polygenic
+selection after an optimum shift \parencite{Buffalo2019-qs,Buffalo2020-my} or
+soft sweeps from standing variation (XXX).
+
+In turn, statistical methods have use this theory for a variety of wide-ranging
+empirical problems.  Early on, researchers
+
+While it is difficult to differentiate the selective processes
+leading to linked selection, studies indicate background selection
+
+hitchhiking events where a low-frequency
+
+studies
+in \emph{Drosophila} and humans indicate background selection has a large role
+in shaping patterns of diversity, especially at large spatial scales (i.e.
+megabases).
+
+Another central empirical goal is to estimate the relative strengths of the
+these two selection processes on levels of genetic diversity (Eyalshiv). Since
+the effects of linked selection are determined in part by mutation rate and
+distribution of fitness effects (DFE), these statistical methods also estimate
+these important evolutionary parameters from genomic data. These estimates
+should, in theory, align with others using different approaches, such as
+trio-based estimates of the mutation rate and estimates of the DFE from
+site-frequency spectra.
+
+Understanding the extent to which linked selection influences other
+evolutionary processes is a central question in evolutionary genetics. Two
+central questions have emerged. First, while genetic drift is thought to be the
+main source of random fluctuations in neutral allele frequencies, linked
+selection can have a similar impact on neutral sites, increasing the variance
+in allele frequency change as if the population size were effectively smaller.
+Although linked selection causes significant reductions in genetic diversity in
+some species, it is unlikely to explain the narrow range of observed diversity
+across species \parencite{Buffalo2021-uf}. Second, by perturbing the
+frequencies of neighboring alleles, selection can interfere with their
+selection dynamics, known as Hill-Robertson interference (HRI, XXX). When
+recombination is limited, simultaneous selection across the genome reduces the
+odds beneficial mutations are fixed (HRI), effectively like selection was
+occurring in an smaller population size. The Hill-Robertson effect is thought
+to be weak in organisms in organisms with large recombination maps, perhaps
+because of selection for increased recombination.
+
+Our focus here is to begin to fix some deficiencies in the background selection
+model. In particular,
+
+% outline?
 - Disentangling evolutionary processes from the patterns of genetic variation
@@
 
+\section*{Conclusion}
+
+- care should be taken applying these methods to \emph{Drosophila}, since the
+$\delta_{LD}$ term could be much larger.
+
 \section*{Results}
@@
 
-% could be in intro too:
+even when sites under selection are sparse.
 
-As genomes are transmitted through the generations, lifetime events selectively
-alter what is passed to the next generation. While selection truly acts on the
-genetic variants that directly cause heritable fitness differences among
-individuals, neighboring regions experience an indirect contraction in genetic
-variability. This byproduct is known as \emph{linked selection}, and it occurs
-when selection changes a site's frequency on timescale faster than
-recombination can decouple that site from its genetic background. Because
-linked selection modifies the genetic variability in a regionally-correlated
-way, it creates patterns in genomic data that are used to understand the types
-of selection processes acting on genomes.
+From an evolutionary perspective, linked selection is important because by
+modifying neighboring levels of genetic diversity, it is also directly altering
+the selection dynamics at these neighboring sites.
 
-Much work has been done on developing mathematical theory for two primary
-selection processes: hitchhiking and background selection. In turn, statistical
-methods have use this theory for a variety of wide-ranging empirical problems.
-Early on, researchers found correlations between recombination rates and
-genetic diversity in small genomes with limited recombination like
-\emph{Drosophila} (XXX), consistent with linked selection processes acting on
-genome-wide scales. With population genomic data, small-scale (XXX kb) dips in
-genetic diversity was evidence of rare but strong selective sweeps in humans
-(XXX) and other species (XXX).
-
-models of background selection can
-closely predict levels of human-chimp divergence at the megabase scale (XXX).
-
-humans (XXX), though these offered only a partial picture of the selection
-processes acting (XXX).
-
-Linked
-selection studies have found additional strong selective sweeps in taxa across
-the kingdom of life (XXX), and attention turned towards estimating the relative
-contributions of the hitchhiking and background processes (Eyalshiv).
-processes (Eyalshiv).
-
-From an evolutionary perspective, models of linked selection
+Typically evolutionary genetic models consider
 
@@
 \end{document}
+
+% A central goal of evolutionary genetics is to quantify the relative roles
+% different evolutionary processes have on genetic variation. Certain
+% evolutionary processes, such as demographic bottlenecks or expansions, act
+% uniformly across the genome as a whole. Other processes, such as selection, are
+% hypothesized to act differently across the genome depending on the density of
+% genomic features, and local heterogeneity in recombination and mutation rate.
+% With the emergence of population genomic data, a variety of statistical models
+% have teased apart the contribution of these various processes. These
+% statistical models differ in how they reduce large genomic data to summary
+% statistics, as well as in their statistical methodology. Some models reduce
+% genomic data to the histograms of the frequency spectrum (XXX), while others
+% model scalar summaries of the frequency spectrum along the genome.
+
+% One class of models, parametric statistical models, leverage population genetic
+% theory to inform expectations of the mean pairwise diversity along the
+% chromosome. Estimation under these models often uses a composite likelihood
+% approach, where the joint distribution is assumed independent across positions.
+% While composite likelihoods are simpler to formulate and lead to consistent
+% parameter estimation, ignoring dependencies between neighboring sites has two
+% primary disadvantages. First, unlike proper likelihoods, the composite
+% likelihood's curvature is uninformative about parameter estimate uncertainty.
+% Previous work has circumvented this shortcoming through (1) resampling-based
+% uncertainty estimates, or (2) adjusting the composite likelihood's curvature to
+% account for these dependencies. Second, similar to inference under phylogenetic
+% comparative models, the dependency structure tells us how to properly weight
+% information across sites. For example, imagine a conserved segment 10 kilobases
+% long right on the border of two genomic bins, which due to linked selection,
+% bot experience a reduction in diversity. When estimating parameters, the
+% independence assumption of composite likelihoods would weight these observed
+% reductions equal to the amount of a reduction observed on an entirely different
+% independent chromosome.
+
+% patterns of genetic variability along the genome are treated as
+% coming from an indeprocess $m(z)$.
+
+% Here, we focus on solving an issue processes: background
+
+% BGS as evolution around the static features of the genome. Due to the slow rate
+% of genomic structural evolution, much of the genome can be treated as static
+% throughout the genome. These regions are under differing levels of selective
+% constraint, which induces background selection at neighboring regions.
+
+% Signal processing
+
+% Genomes have inherent structure that impacts the linked selection map. There is
+% tremendous heterogeneity across species in the organization of likely selected
+% regions on the genetic and physical maps.
+
+% Within an organism, there are two key determinants of B maps: recombination and
+% gene density.
+
+% Initially, the relationship between recombination and diversity was regressed
+% with no regard to the spatial organization of the genome.
+
+% Certain regions are pinchpoints of selection: some combination of low
+% recombination and high gene density lead to very strong reductions in genetic
+% diversity. In these regions, coalescent dynamics change considerably; where as
+% under strong background selection, deleterious alleles would not fix, in these
+% regions the can fix. Additionally, the selective interference impacts positive
+% selection.

### 2022-10-04 manuscript/manuscript.tex
more intro

--- 2022-10-02 manuscript/manuscript.tex
@@
 
-% A central goal of evolutionary genetics is to quantify the relative roles
-% different evolutionary processes have on genetic variation. Certain
-% evolutionary processes, such as demographic bottlenecks or expansions, act
-% uniformly across the genome as a whole. Other processes, such as selection, are
-% hypothesized to act differently across the genome depending on the density of
-% genomic features, and local heterogeneity in recombination and mutation rate.
-% With the emergence of population genomic data, a variety of statistical models
-% have teased apart the contribution of these various processes. These
-% statistical models differ in how they reduce large genomic data to summary
-% statistics, as well as in their statistical methodology. Some models reduce
-% genomic data to the histograms of the frequency spectrum (XXX), while others
-% model scalar summaries of the frequency spectrum along the genome.
+\section*{Introduction}
 
-% One class of models, parametric statistical models, leverage population genetic
-% theory to inform expectations of the mean pairwise diversity along the
-% chromosome. Estimation under these models often uses a composite likelihood
-% approach, where the joint distribution is assumed independent across positions.
-% While composite likelihoods are simpler to formulate and lead to consistent
-% parameter estimation, ignoring dependencies between neighboring sites has two
-% primary disadvantages. First, unlike proper likelihoods, the composite
-% likelihood's curvature is uninformative about parameter estimate uncertainty.
-% Previous work has circumvented this shortcoming through (1) resampling-based
-% uncertainty estimates, or (2) adjusting the composite likelihood's curvature to
-% account for these dependencies. Second, similar to inference under phylogenetic
-% comparative models, the dependency structure tells us how to properly weight
-% information across sites. For example, imagine a conserved segment 10 kilobases
-% long right on the border of two genomic bins, which due to linked selection,
-% bot experience a reduction in diversity. When estimating parameters, the
-% independence assumption of composite likelihoods would weight these observed
-% reductions equal to the amount of a reduction observed on an entirely different
-% independent chromosome.
+As genomes are transmitted through the generations, lifetime events selectively
+alter what is passed to the next generation. While selection directly acts on
+the genetic variants that cause heritable fitness differences among
+individuals, neighboring regions experience an indirect contraction in genetic
+variability. This byproduct is known as \emph{linked selection}, and it occurs
+when selection changes a site's frequency on timescale faster than
+recombination can decouple that site from its genetic background. Because
+linked selection modifies the genetic variability in a regionally-correlated
+way, it creates observable patterns in genomic data that are used to understand
+the type of selection processes acting on genomes and estimate key evolutionary
+quantities.
 
-% patterns of genetic variability along the genome are treated as
-% coming from an indeprocess $m(z)$.
+% , with characteristic length
+% $\nicefrac{s}{r}$ basepairs, where $s$ is the heterozygous selection
+% coefficient and $r$ is the per-basepair recombination rate. Sweeps were found
+% in humans, \emph{Drosophila}, and other species (XXX), though it was unclear
+% whether hitchhiking alone could explain the observed genome-wide correlation
+% between recombination rate and genetic diversity in some species (XXX).
 
-% Here, we focus on solving an issue processes: background
+Much work has been done on developing mathematical theory for two linked
+selection processes: hitchhiking and background selection. These are
+distinguished by the evolutionary processes that generate fitness variation and
+the resulting selection dynamics. Under the classic hitchhiking model, a new
+beneficial mutation sweeps fixation, leading to a dip in genetic diversity on a
+small spatial scale. Since the sweeps leave relatively narrow footprint, they
+influence genome-wide diversity only when beneficial mutations are recurring
+and recombination is relatively limited, as in \emph{Drosophila}. These models
+assume sweeps are sufficiently rare as to not occur simultaneously, which is
+consistent with the slow estimated rate of strongly beneficial mutations (XXX).
+Overall, the classic hitchhiking model is simple because the underlying
+selection process does not generate polygenic fitness variation, even though
+this is type of variation plays a dominant role in adaptation (XXX).
 
-% BGS as evolution around the static features of the genome. Due to the slow rate
-% of genomic structural evolution, much of the genome can be treated as static
-% throughout the genome. These regions are under differing levels of selective
-% constraint, which induces background selection at neighboring regions.
+By contrast, background selection is a consequence of the fitness variation
+created by a steady influx of new deleterious mutations. The classic background
+selection model (XXX) considers how the balance between mutation introducing
+deleterious variants into the population and selection removing them, reduces
+nearby genetic variability. A critical assumption of this model is that such an
+equilibrium between mutation and selection exists, which is the case only when
+these mutations have a large enough fitness cost to ensure that they cannot fix
+in the population. Under this process, the number of deleterious mutations per
+gamete reaches a Poisson stationary distribution with rate $\nicefrac{U}{s}$,
+where $U = \mu_D L$ is the deleterious mutation rate ($\mu_D$ per basepair per
+generation) in a region of $L$ basepairs for mutations with a heterozygous
+selection coefficient of $s$ (XXX). The fraction of gametes without any
+mutations is then $f_0 = \exp(-\nicefrac{\mu_D}{s})$. Without recombination,
+any gamete with deleterious mutations is destined to bring its linked neutral
+variation with it to extinction, effectively reducing diversity levels to those
+expected in a neutrally-evolving population of size $f_0 N$.
 
-% Signal processing
+Remarkably, predictions from simple background selection models with
+recombination closely match large-scale patterns of human diversity and
+divergence (XXX). Since background approximately rescales a neutral coalescent
+process by $f_0$, the predicted diversity is reduced from neutral levels by a
+factor $B = \nicefrac{N_e}{N}$, which is the ratio of effective population size
+under background selection to the population size. In background selection
+models with recombination, $B$ is determined by the number of conserved
+basepairs and recombination in a region. Consequently the predicted reduction
+along the genome (known as $B$ maps) mirror the regional density of conserved
+elements under purifying selection and recombination rates. These $B$ maps have
+been used in studies of XXX, are inputs into machine learning models that
+predict functional constraint, and XXX. Furthermore, since new deleterious
+mutations are perpetually flowing into populations $B$ maps may provide a
+better null model for evolution than the standard neutral model.
 
-% Genomes have inherent structure that impacts the linked selection map. There is
-% tremendous heterogeneity across species in the organization of likely selected
-% regions on the genetic and physical maps.
+While background selection models do predict diversity levels
 
-% Within an organism, there are two key determinants of B maps: recombination and
-% gene density.
-
-% Initially, the relationship between recombination and diversity was regressed
-% with no regard to the spatial organization of the genome.
-
-% Certain regions are pinchpoints of selection: some combination of low
-% recombination and high gene density lead to very strong reductions in genetic
-% diversity. In these regions, coalescent dynamics change considerably; where as
-% under strong background selection, deleterious alleles would not fix, in these
-% regions the can fix. Additionally, the selective interference impacts positive
-% selection.
-
+% outline?
 - Disentangling evolutionary processes from the patterns of genetic variation
@@
 \end{document}
+
+% A central goal of evolutionary genetics is to quantify the relative roles
+% different evolutionary processes have on genetic variation. Certain
+% evolutionary processes, such as demographic bottlenecks or expansions, act
+% uniformly across the genome as a whole. Other processes, such as selection, are
+% hypothesized to act differently across the genome depending on the density of
+% genomic features, and local heterogeneity in recombination and mutation rate.
+% With the emergence of population genomic data, a variety of statistical models
+% have teased apart the contribution of these various processes. These
+% statistical models differ in how they reduce large genomic data to summary
+% statistics, as well as in their statistical methodology. Some models reduce
+% genomic data to the histograms of the frequency spectrum (XXX), while others
+% model scalar summaries of the frequency spectrum along the genome.
+
+% One class of models, parametric statistical models, leverage population genetic
+% theory to inform expectations of the mean pairwise diversity along the
+% chromosome. Estimation under these models often uses a composite likelihood
+% approach, where the joint distribution is assumed independent across positions.
+% While composite likelihoods are simpler to formulate and lead to consistent
+% parameter estimation, ignoring dependencies between neighboring sites has two
+% primary disadvantages. First, unlike proper likelihoods, the composite
+% likelihood's curvature is uninformative about parameter estimate uncertainty.
+% Previous work has circumvented this shortcoming through (1) resampling-based
+% uncertainty estimates, or (2) adjusting the composite likelihood's curvature to
+% account for these dependencies. Second, similar to inference under phylogenetic
+% comparative models, the dependency structure tells us how to properly weight
+% information across sites. For example, imagine a conserved segment 10 kilobases
+% long right on the border of two genomic bins, which due to linked selection,
+% bot experience a reduction in diversity. When estimating parameters, the
+% independence assumption of composite likelihoods would weight these observed
+% reductions equal to the amount of a reduction observed on an entirely different
+% independent chromosome.
+
+% patterns of genetic variability along the genome are treated as
+% coming from an indeprocess $m(z)$.
+
+% Here, we focus on solving an issue processes: background
+
+% BGS as evolution around the static features of the genome. Due to the slow rate
+% of genomic structural evolution, much of the genome can be treated as static
+% throughout the genome. These regions are under differing levels of selective
+% constraint, which induces background selection at neighboring regions.
+
+% Signal processing
+
+% Genomes have inherent structure that impacts the linked selection map. There is
+% tremendous heterogeneity across species in the organization of likely selected
+% regions on the genetic and physical maps.
+
+% Within an organism, there are two key determinants of B maps: recombination and
+% gene density.
+
+% Initially, the relationship between recombination and diversity was regressed
+% with no regard to the spatial organization of the genome.
+
+% Certain regions are pinchpoints of selection: some combination of low
+% recombination and high gene density lead to very strong reductions in genetic
+% diversity. In these regions, coalescent dynamics change considerably; where as
+% under strong background selection, deleterious alleles would not fix, in these
+% regions the can fix. Additionally, the selective interference impacts positive
+% selection.
+
+% Understanding the extent to which linked selection influences other
+% evolutionary processes is a central question in evolutionary genetics. Two
+% central questions have emerged. First, while genetic drift is thought to be the
+% main source of random fluctuations in neutral allele frequencies, linked
+% selection can have a similar impact on neutral sites, increasing the variance
+% in allele frequency change as if the population size were effectively smaller.
+% Although linked selection causes significant reductions in genetic diversity in
+% some species, it is unlikely to explain the narrow range of observed diversity
+% across species \parencite{Buffalo2021-uf}. Second, by perturbing the
+% frequencies of neighboring alleles, selection can interfere with their
+% selection dynamics, known as Hill-Robertson interference (HRI, XXX). When
+% recombination is limited, simultaneous selection across the genome reduces the
+% odds beneficial mutations are fixed (HRI), effectively like selection was
+% occurring in an smaller population size. The Hill-Robertson effect is thought
+% to be weak in organisms in organisms with large recombination maps, perhaps
+% because of selection for increased recombination.
+
+% In organisms with limited recombination such as
+% \emph{Drosophila}, estimates suggest background selection plays a marginally
+% larger role than hard sweeps in reducing diversity \parencite{Elyashiv2016-vt}.
+
+% closely predict large-scale patterns of
+% diversity across the genome
+
+% Since far more new mutations are deleterious than beneficial in functional
+% regions, it would seem like background selection would have a dominant role in
+% shaping genetic diversity. However, deleterious mutations segregate on a
+% timescale inversely proportional to their selection cost (XXX). Strongly
+% deleterious mutations eliminate neighboring genetic variability on the few
+% chromosomes they may recombine onto before the eventual extinction, and
+% consequently have very weak effects on larger spatial scales. Moderately weakly
+% selected mutations can persist for much longer periods, allowing for more
+% recombination before their eventual loss and produce reductions in diversity on
+% shorter spatial scales.  The reduction in diversity under background selection
+% is determined approximately by $\nicefrac{U}{R}$, where $U = \mu_D L$ is the
+% number of deleterious mutations (at rate $\mu_D$ per basepair, per generation)
+% in a conserved region of $L$ basepairs per generation, and $R = r L$, which is
+% the total recombination per generation in Morgans. Since the deleterious
+% mutation rate is relatively constant across conserved regions, the reduction
+% due to background selection closely matches the composition of conserved
+% regions and recombination along the genome. Remarkably, simple mathematical
+% models of background selection closely predicts levels of human-chimp
+% divergence at the megabase scale (XXX) from genome-wide patterns of
+% recombination and regions conserved between species (McVicker).
+
+% This leads to some of the unidentifiability of hitchhiking versus background
+% selection.
+
+% and the reduction is determined
+% by
+
+% While the rate
+% new beneficial mutations enter the population is low (XXX),
+
+% The spatial scale of background selection is
+
+% kkkkkk
+
+% not all
+% adaptation occurs from such new beneficial mutations sweeping to fixation (XXX)
+
+% Other processes also act
+% on beneficial fitness variation to engender linked selection, such as polygenic
+% selection after an optimum shift \parencite{Buffalo2019-qs,Buffalo2020-my} or
+% soft sweeps from standing variation (XXX).
+
+% In turn, statistical methods have use this theory for a variety of wide-ranging
+% empirical problems.  Early on, researchers
+
+% While it is difficult to differentiate the selective processes
+% leading to linked selection, studies indicate background selection
+
+% hitchhiking events where a low-frequency
+
+% studies
+% in \emph{Drosophila} and humans indicate background selection has a large role
+% in shaping patterns of diversity, especially at large spatial scales (i.e.
+% megabases).
+
+% Another central empirical goal is to estimate the relative strengths of the
+% these two selection processes on levels of genetic diversity (Eyalshiv). Since
+% the effects of linked selection are determined in part by mutation rate and
+% distribution of fitness effects (DFE), these statistical methods also estimate
+% these important evolutionary parameters from genomic data. These estimates
+% should, in theory, align with others using different approaches, such as
+% trio-based estimates of the mutation rate and estimates of the DFE from
+% site-frequency spectra.
+
+% Our focus here is to begin to fix some deficiencies in the background selection
+% model. In particular,

### 2022-10-05 manuscript/manuscript.tex
more intro

@@
 
-While background selection models do predict diversity levels
+While background selection models do predict diversity levels  XXX
 
@@
 
-even when sites under selection are sparse.
-
 From an evolutionary perspective, linked selection is important because by
@@
 the selection dynamics at these neighboring sites.
-
-Typically evolutionary genetic models consider
 
@@
 \subsection*{Overview of Theory}
+
+Here, I step through a quick derivation of \textcite{Santiago1998-bs}.
+Throughout, we assume random mating, hermaphroditic individuals, and a constant
+population size.
+
+The change in a neutral allele's frequency in one generation can be partitioned
+into the three sources of stochasticity: the random associations with fitness
+backgrounds (i.e. \emph{draft}), the non-heritable randomness in family size,
+and the Mendelian noise from heterozygotes segregating. If we let $x_{0,i} \in
+\{0, 1, 2\}$ be the number of neutral alleles individual $i$ in generation 0
+carries, we can partition the neutral allele frequency change into the
+underlying stochastic causes,
+
+\begin{align}
+  p_1 = \frac{1}{N} \sum_{i=1}^N \left( \frac{x_{0,i}}{2} k_i + \sum_{j=1}^{k_i} \delta_{0,i,j} \right)
+\end{align}
+
+where $k_i$ is the number of surviving gametes parent $i$ passes on, and
+$\delta_{0,i,j}$ is a random term that encapsulates the noise due to Mendelian
+segregation of heterozygotes. If parent $i$ is a homozygote ($x_{0,i} \in \{0,
+2\}$), then $\delta_{0,i,j} = 0$, whereas if $x_{0,i} = 1$ then $\delta_{0,i,j}
+= \pm \nicefrac{1}{2}$ with equal probability. This is because a heterozygous
+parent will transmit half a neutral allele in expectation, but each round of
+Mendelian segregation must pass on either 0 or 1 alleles, which the random $\pm
+\nicefrac{1}{2}$ term imposes. There is the condition that $\nicefrac{1}{N}
+\sum_i k_i = 2$, since each diploid parent must have two kids to maintain a
+constant population size.
+
+Then, noting that $p_0 = \nicefrac{1}{2N} \sum_{i=1}^N x_{0,i}$, the allele
+frequency change is,
+
+\begin{align}
+  \Delta p_1 = p_1 - p_0 = \frac{1}{N} \sum_{i=1}^N \frac{x_{0,i}}{2} (k_i - 1) + \frac{1}{N} \sum_{i=1}^N \sum_{j=1}^{k_i} \delta_{0,i,j}.
+\end{align}
+
+Next, we can partition the gamete number into heritable ($f_i$) and
+non-heritable ($\varepsilon_i$) fitness components of each individual $i$, $k_i
+= f_i + \varepsilon_i$. We define $f_i$ such that $\nicefrac{1}{N} \sum_i f_i =
+2$ in a constant population, and $\varepsilon_i$ is noise term centered around
+zero. Then, the neutral allele frequency change is,
+
+\begin{align}
+  \Delta p_1 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (f_i - 1) +
+               \frac{1}{2N} \sum_{i=1}^N x_{0,i} \varepsilon_i
+  + \frac{1}{N} \sum_{i=1}^N \sum_{j=1}^{k_i} \delta_{0,i,j} \\
+  % \Delta p_1 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (f_i - 1) +
+               % \frac{1}{2N} \sum_{i=1}^N x_{0,i} \varepsilon_i
+  % + \frac{1}{N} \sum_{i=1}^N \sum_{j=1}^{k_i} \delta_{0,i,j} \\
+  \Delta p_1 &= \Delta_{_H} p_1 + \Delta_{_N} p_1 + \Delta_{_M} p_1
+\end{align}
+
+which we will derive the variance of through time (c.f. \cite{Santiago1995-hx}
+equation 2 and \cite{Buffalo2019-qs} equation 30). This equation is very
+general, and allows for any reproduction model with finite offspring variance
+and can account for the impact of linked selection under a variety of selection
+processes.
+
+%%% OLD
 

### 2022-10-06 manuscript/manuscript.tex
good theory

--- 2022-10-04 manuscript/manuscript.tex
@@
 \subsection*{Overview of Theory}
+
+Here, I step through a quick derivation of \textcite{Santiago1998-bs}.
+Throughout, we assume random mating, hermaphroditic individuals, and a constant
+population size.
+
+The change in a neutral allele's frequency in one generation can be partitioned
+into the three sources of stochasticity: the random associations with fitness
+backgrounds (i.e. \emph{draft}), the non-heritable randomness in family size,
+and the Mendelian noise from heterozygotes segregating. If we let $x_{0,i} \in
+\{0, \nicefrac{1}{2}, 1\}$ be the frequency of neutral alleles individual $i$
+in generation 0 carries, we can partition the neutral allele frequency change
+into the underlying stochastic causes,
+
+\begin{align}
+  p_1 = \frac{1}{2N} \sum_{i=1}^N \left( x_{0,i}k_{0,i} + \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \right)
+\end{align}
+
+where $k_{0,i}$ is the number of surviving gametes parent $i$ passes on, and
+$\delta_{0,i,j}$ is a random term that encapsulates the noise due to Mendelian
+segregation of heterozygotes. If parent $i$ is a homozygote ($x_{0,i} \in \{0,
+1\}$), then $\delta_{0,i,j} = 0$, whereas if $x_{0,i} = \nicefrac{1}{2}$ then
+$\delta_{0,i,j} = \pm \nicefrac{1}{2}$ with equal probability. This is because
+a heterozygous parent will transmit half a neutral allele in expectation, but
+each round of Mendelian segregation must pass on either 0 or 1 alleles, which
+the random $\pm \nicefrac{1}{2}$ term imposes. The factor of $\nicefrac{1}{2}$
+is due to the fact that we're summing over $N$ diploids, but considering the
+number of gametes they transmit. Since each diploid parent must have two
+offspring to maintain a constant population size, $\nicefrac{1}{N} \sum_i
+k_{0,i} = 2$.
+
+Then, noting that $p_0 = \nicefrac{1}{N} \sum_{i=1}^N x_{0,i}$, the allele
+frequency change is,
+
+\begin{align}
+  \Delta p_1 = p_1 - p_0 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (k_{0,i} - 1) + \frac{1}{2N} \sum_{i=1}^N \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \\
+  \Delta p_1 &= K_1 + H_1
+\end{align}
+
+where $K_1$ and $H_1$ are the random change in neutral allele frequency change
+due to offspring number (including heritable and non-heritable components), and
+Mendelian segregation.
+
+Now, let us look the variance of $\var(\Delta p_1)$ over evolutionary
+replicates. Since the Mendelian random process and the offspring random process
+are independent,
+
+\begin{align}
+  \var(\Delta p_1) &= \var(K_1) + \var(H_1).
+\end{align}
+
+Looking at each term,
+
+\begin{align}
+  \var(K_1) &\approx \frac{1}{4N^2} \sum_{i=1}^N \var\left(x_{0,i} (k_{0,i} - 1) \right)
+\end{align}
+
+where we ignore the covariance terms due to the sum, since these are on order
+$\nicefrac{1}{N^3}$. In the first generation from an arbitrary starting point,
+the neutral alleles assort independently into diploids with respect to their
+fitness, so
+
+\begin{align}
+  \var(K_1) &\approx \frac{1}{4N^2} \sum_{i=1}^N \var(x_{0,i}) \var(k_{0,i} - 1)  \\
+            &\approx \frac{1}{4N} \var(x_{0,i}) \var(k_{0,i}).
+\end{align}
+
+Assuming that no correlation between parental gametes (e.g. no inbreeding),
+$\var(x_{0,i})$ is the binomial variance in individual allele frequency, or
+$p_0(1-p_0)/2$, and $\var(k) := \var(k_{0,i})$ is the offspring variance of an
+individual given by the reproduction process. For example, if the reproduction
+process is a neutral multinomial Wright--Fisher, $\var(k) \approx 2$.
+
+\begin{align}
+  \var(K_1) &\approx \frac{p_0(1-p_0)}{2N} \frac{\var(k)}{4}.
+\end{align}
+
+The Mendelian noise variance term can be derived similarly. First, the sum over
+each parent's transmitted gametes can be simplified by noting that parents are
+exchangeable over evolutionary replicates with respect to their contribution to
+this term. In a constant population size, the double summation over parents and
+their offspring can be replaced by a summation over offspring, since both sum
+$N$ exchangeable terms. Then, note that $\var(\delta_{i,j}) = \nicefrac{1}{4}
+p_0(1-p_0)$, so
+
+\begin{align}
+  \var(H_1) &= \frac{p_0(1-p_0)}{2N} \frac{1}{2}.
+\end{align}
+
+Finally, we have
+
+\begin{align}
+  \var(\Delta p_1) &= K_1 + H_1 \\
+  &\approx \frac{p_0(1-p_0)}{2N} \frac{\var(k)}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2} \\
+                   &\approx \frac{p_0(1-p_0)}{2N}\left(\frac{\var(k)}{4} + \frac{1}{2}\right)
+\end{align}
+
+(c.f. \cite{Santiago1995-hx} equation 2 and \cite{Buffalo2019-qs} equation 30).
+Note that if the reproduction process is a neutral multinomial Wright--Fisher,
+$\var(k) \approx 2$, and this simplifies to the expected Wright--Fisher
+variance. If we were to \emph{define} a variance effective population size by
+setting this variance in neutral allele frequency change to the expected
+variance under a Wright--Fisher model ($V_\text{WF}$), we'd have
+
+\begin{align}
+  V_\text{WF} &:= \frac{p_0(1-p_0)}{2N_e} \\
+  V_\text{WF} &= \var(\Delta p_1) \\
+  N_e &= \frac{4N}{\var(k) + 2} \\
+\end{align}
+
+(c.f. \cite{Wright1938-tv}).
+
+% Each term is time-indexed because it is conditioned on the starting neutral
+% frequency in each individual, $x_{0,1}$. This equation is very general, and
+% allows for any reproduction model with finite offspring variance and can
+% account for the impact of linked selection under a variety of selection
+% processes.
+
+% Next, we can partition the gamete number into heritable ($f_i$) and
+% non-heritable ($\varepsilon_i$) fitness components of each individual $i$, $k_i
+% = f_i + \varepsilon_i$. We define $f_i$ such that $\nicefrac{1}{N} \sum_i f_i =
+% 2$ in a constant population, and $\varepsilon_i$ is noise term centered around
+% zero. Then, the neutral allele frequency change is can be written as the sum of
+% three independent stochastic terms,
+
+% \begin{align}
+%   \Delta p_1 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (f_i - 1) +
+%                \frac{1}{2N} \sum_{i=1}^N x_{0,i} \varepsilon_i
+%   + \frac{1}{N} \sum_{i=1}^N \sum_{j=1}^{k_i} \delta_{0,i,j} \\
+%   % \Delta p_1 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (f_i - 1) +
+%                % \frac{1}{2N} \sum_{i=1}^N x_{0,i} \varepsilon_i
+%   % + \frac{1}{N} \sum_{i=1}^N \sum_{j=1}^{k_i} \delta_{0,i,j} \\
+%   \Delta p_1 &= S_1 + D_1 + H_1
+% \end{align}
+
+% (c.f. \cite{Santiago1995-hx} equation 2 and \cite{Buffalo2019-qs} equation 30).
+
+% Each term is time-indexed because it is conditioned on the starting neutral
+% frequency in each individual, $x_{0,1}$. This equation is very general, and
+% allows for any reproduction model with finite offspring variance and can
+% account for the impact of linked selection under a variety of selection
+% processes.
+
+% Now, let us look the variance of $\var(\Delta p_1)$ over evolutionary
+% replicates.
+
+% \begin{align}
+%   \var(\Delta p_1) &= \var(S_1) + \var(D_1) + \var(H_1)
+% \end{align}
+
+% To introduce how this equation can be used to quantify the variance in allele
+% frequency change through time, $\var(p_t - p_0$ for $t > 0$, we first look at
+% evolution in a neutral Wright--Fisher model. Neutrality ensures that $S_t = 0$
+% for all $t$.
+
+%%% OLD
 

### 2022-10-06 manuscript/manuscript.tex
more theory

@@
 \{0, \nicefrac{1}{2}, 1\}$ be the frequency of neutral alleles individual $i$
-in generation 0 carries, we can partition the neutral allele frequency change
-into the underlying stochastic causes,
-
-\begin{align}
-  p_1 = \frac{1}{2N} \sum_{i=1}^N \left( x_{0,i}k_{0,i} + \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \right)
+in generation 0 carries, we can partition the random neutral allele frequency
+of the population ($x_t$ without the individual index) into the underlying
+stochastic causes,
+
+\begin{align}
+  x_1 = \frac{1}{2N} \sum_{i=1}^N \left( x_{0,i}k_{0,i} + \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \right)
 \end{align}
@@
 
-Then, noting that $p_0 = \nicefrac{1}{N} \sum_{i=1}^N x_{0,i}$, the allele
-frequency change is,
-
-\begin{align}
-  \Delta p_1 = p_1 - p_0 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (k_{0,i} - 1) + \frac{1}{2N} \sum_{i=1}^N \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \\
-  \Delta p_1 &= K_1 + H_1
+The frequency in the initial generation is $x_0 = \nicefrac{1}{N} \sum_{i=1}^N
+x_{0,i}$, though to indicate that we treat this as fixed, we use $p_0 = x_0$.
+Then, the allele frequency change is,
+
+\begin{align}
+  \Delta x_1 = x_1 - p_0 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (k_{0,i} - 1) + \frac{1}{2N} \sum_{i=1}^N \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \\
+  \Delta x_1 &= K_1 + H_1
 \end{align}
@@
 
-Now, let us look the variance of $\var(\Delta p_1)$ over evolutionary
+Now, let us look the variance of $\var(\Delta x_1)$ over evolutionary
 replicates. Since the Mendelian random process and the offspring random process
@@
 \begin{align}
-  \var(\Delta p_1) &= \var(K_1) + \var(H_1).
+  \var(\Delta x_1) &= \var(K_1) + \var(H_1).
 \end{align}
@@
 (c.f. \cite{Wright1938-tv}).
+
+Now, we look at this variance equation behaves with $t > 1$.
+
+\begin{align}
+  \V(x_1) = \E_1\left[(x_1 - x_0)^2\right]= \frac{x_0(1-x_0)}{2N},
+\end{align}
+
+and
+
+\begin{align}
+  \V(x_2) &= \E\left[(x_2 - x_0)^2\right] \\
+          &= \E_1\left[\E_2\left[((x_2-x_1) + (x_1 - x_0))^2 | x_1\right]\right] \\
+          &= \E_1\left[\E_2\left[(\Delta x_2 + \Delta x_1)^2 | x_1\right]\right] \\
+          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + 2\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
+          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] + C_{1,2} \\
+          &= \frac{p_0(1-p_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{p_0(1-p_0)}{2N} + C_{1,2}
+\end{align}
+
+Consequently, the variance of the neutral allele frequency changes each
+generation according to the probability of failing to coalesce each generation
+and the pairwise covariance terms $C_{i,j}$ that build up due to associations
+with heritable fitness backgrounds. Note that the covariance term $C_{1,2} =
+\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] = 0$ under neutral
+evolution, and this simplifies to the well-known equation for variance in a
+Wright--Fisher population.
+
+% In general, the variance in allele frequency after $t$ generations
+% of neutral evolution is,
+
+%\begin{align}
+%  \label{eq:var_freq}
+%  %\V(x_t) &= \frac{x_0(1-x_0)}{2N}\left(\sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i-1} \right)
+%  \V(x_t) &= \V(x_1) \left(1 + \sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i} \right)
+%\end{align}
 

### 2022-10-08 manuscript/manuscript.tex
more theory...

@@
 
-\subsection*{Overview of Theory}
-
-Here, I step through a quick derivation of \textcite{Santiago1998-bs}.
+\subsection*{Overview of Quantitative Genetic Linked Selection Theory}
+
+Here, we step through a quick derivation of \textcite{Santiago1998-bs}.
 Throughout, we assume random mating, hermaphroditic individuals, and a constant
-population size.
-
-The change in a neutral allele's frequency in one generation can be partitioned
-into the three sources of stochasticity: the random associations with fitness
-backgrounds (i.e. \emph{draft}), the non-heritable randomness in family size,
-and the Mendelian noise from heterozygotes segregating. If we let $x_{0,i} \in
-\{0, \nicefrac{1}{2}, 1\}$ be the frequency of neutral alleles individual $i$
-in generation 0 carries, we can partition the random neutral allele frequency
-of the population ($x_t$ without the individual index) into the underlying
-stochastic causes,
+population size. The change in a neutral allele's frequency in one generation
+can be partitioned into the three sources of stochasticity: the random
+associations with fitness backgrounds (i.e. \emph{draft}), the non-heritable
+randomness in family size, and the Mendelian noise from heterozygotes
+segregating. If we let $x_{0,i} \in \{0, \nicefrac{1}{2}, 1\}$ be the frequency
+of neutral alleles individual $i$ in generation 0 carries, we can partition the
+random neutral allele frequency of the population ($x_t$ without the individual
+index) into the underlying stochastic causes,
 
@@
 \begin{align}
-  \var(\Delta p_1) &= K_1 + H_1 \\
+  \var(\Delta x_1) &= \var(K_1) + \var(H_1) \\
   &\approx \frac{p_0(1-p_0)}{2N} \frac{\var(k)}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2} \\
@@
 \begin{align}
+  \label{eq:var_wf}
   \V(x_2) &= \E\left[(x_2 - x_0)^2\right] \\
@@
 Wright--Fisher population.
+
+Now, we turn our attention to the case where there is heritable fitness
+variation, which leads to non-zero covariance terms $C_{i,j}$. These terms
+emerge when $\var(k)$ has a heritable component of fitness that can be
+transmitted along with the neutral allele, thereby affecting the neutral
+allele's trajectory in later generations. If we write the offspring number into
+heritable and non-heritable components, $k_i = 2f_i + \varepsilon_i$, then
+$\var(k_i) = 4 V_h + V_n$. We can further decompose the $K_1$ term into $K_1 =
+S_1 + D_1$,
+
+\begin{align}
+  \var(\Delta x_1) &= \var(S_1) + \var(D_1) + \var(H_1) \\
+  &\approx \frac{p_0(1-p_0)}{2N} V_h + \frac{p_0(1-p_0)}{2N} \frac{V_n}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2}. \\
+\end{align}
+
+(c.f. \cite{Santiago1995-hx} equation 11).
+
+To see how the covariance terms accumulate, consider $\var(x_3 - p_0)$. Note
+that $\E(S_t) = \E(D_t) = \E(H_3) = 0$, since which neutral allele we track is
+arbitrary, so by symmetry the expected change is zero. Then,
+
+\begin{align}
+  \label{eq:var_x}
+  \var(x_3 - p_0) &= \E\left[ \left(S_1 + D_1 + H_1 + S_2 + D_2 + H_2 + S_3 + D_3 + H_3 \right)^2 \right] \\
+                  &= \E({S_1}^2) + \E({S_2}^2) + \E({S_3}^2)  \\
+                  &+ 2\E(S_1 S_2) + 2\E(S_1 S_3) + 2\E(S_2 S_3) \\
+                  &+ \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) + \E({D_3}^2) + \E({H_3}^2)
+\end{align}
+
+The covariance terms $\E(S_i S_j)$ for $j > i$ represent the expected
+neutral allele frequency change from a neutral allele becoming associated with
+a fitness background in generation $i$ and that fitness association persisting
+until generation $j$. Let us consider $S_1$ and $S_2$. We have,
+
+\begin{align}
+  S_1 = \frac{1}{2N} \sum_{i=1}^N x_{0,i}(f_{0,i} - 1) = \cov(x_{0,i}, f_{0,i}) \\
+  S_2 = \frac{1}{2N} \sum_{i=1}^N x_{1,i}(f_{1,i} - 1) = \cov(x_{1,i}, f_{0,i})
+\end{align}
+
+which are chance covariances (across the population, \emph{not} evolutionary
+replicates) created by the random sorting of neutral alleles and fitness into
+individuals, \emph{each generation}. These covariances are equivalent to the
+Robertson-Price equation \parencite{Robertson1966-fs,Price1970-si}, which
+predict the change in neutral frequency due to heritable fitness (and likewise
+with the change due to non-heritable factors).
+
+Considering the underlying haplotypes that lead to $x_{0,i}$ and $f_{0,i}$ can
+give us an equation for the dynamics of the $\E(S_i S_j)$ (for $j > i$) terms
+over evolutionary replicates. Let us partition the allele frequency and fitness
+by into the average of the paternal contributions. The neutral allele frequency
+per gamete is either 0 or 1, and we assume the fitnesses across gametes are
+additive. Then, as long as there is random mating, we can simplify the
+associations in a diploid by looking at a single gamete,
+
+\begin{align}
+  S_1 = \cov(x_{0,i}, f_{0,i}) &= \cov\left(\frac{x_{0,i}' + x_{0,i}''}{2}, f_{0,i}' + f_{0,i}''\right)  \\
+                               &= \frac{1}{2}\left(\cov(x_{0,i}', f_{0,i}' + f_{0,i}'') + \cov(x_{0,i}'', f_{0,i}' + f_{0,i}'')\right)  \\
+                               &= \cov(x_{0,i}', f_{0,i}' + f_{0,i}'')  \\
+                               &= \cov(x_{0,i}', f_{0,i}') + \cov(x_{0,i}', f_{0,i}'')  \\
+                               &= S_1' + S_1''
+\end{align}
+
+which follows from symmetry, since it does not matter which gamete carrying the
+neutral allele we track (c.f. \cite{Santiago1998-bs} p. 2107). The primes now
+indicate whether the selected locus is on the same gamete as the neutral
+allele (single prime), or the homologous gamete (double prime).
+
+After meiosis, a fraction $1-r$ of these associations between the tracked
+neutral neutral allele $x_{0,1}'$ and fitness background $f_{0,i}'$ persist,
+and a fraction $r$ disassociate its currently coupled background and create an
+association with the homologous fitness background, $f_{0,i}''$. When linkage
+is tight, this latter term can be ignored as we will do here, though it does
+impact background levels of neutral diversity \parencite{Santiago1995-hx}.
+Additionally, the fitness effects of the selected background may change in
+later generations, in complex ways \parencite{Barton1986-yh,Turelli1990-kd}.
+\textcite{Santiago1995-hx} assume an equilibrium level of fitness variation
+$V_h$, with the fitness variance associated with a particular background
+decaying at a simple geometric decay at rate $Z$ each generation, which is
+sufficient for background selection. An approximate rate of the decay can be
+worked out for the particular selective system. In general, this can be a much
+more complicated function; \textcite{Buffalo2019-qs} show that even fluctuating
+selection can be accommodated.
+
+Now, the associations created in generation 1 and that persist in future
+generations as (ignoring $S_i''$ terms) follow the pattern
+
+\begin{align}
+  S_{2}' &= S_1 Z(1-r) \\
+  S_{3}' &= S_1 Z^2(1-r)^2 \\
+  S_{4}' &= S_1 Z^3(1-r)^3.
+\end{align}
+
+Then, the cumulative effect of the fitness associations created in generation
+$1$ can be written as,
+
+\begin{align}
+  S_1 Q_t &= S_1 + S_2' + S_3' + \ldots + S_t' \\
+  Q_t &= 1 + \sum_{i=1}^t r^i (1-Z)^i \\
+  Q_\infty &= \frac{1}{1-r (1-Z)}
+\end{align}
+
+which shows the long reach of heritable fitness factors in altering allele
+frequency through the generations. In terms like $\E(S_1 S_2)$, a fraction
+$S_1'$ of the heritable allele frequency change $S_2$ were \emph{previously}
+built associations between the neutral allele and a fitness background from
+$S_1$. As long as the direction of selection is the same, the expected change
+of this fraction of associations formed in the first generation to later
+generations 2 and 3 would be,
+
+\begin{align}
+  \E(S_1 S_2) &= \E(S_1^2) Z(1-r) = \E(S_1^2) Q_2^2 \\
+  \E(S_1 S_3) &= \E(S_1^2) Z^2(1-r)^2 = \E(S_1^2) Q_3^2.
+\end{align}
+
+With these covariance terms, we now have a full model for $\var(p_t - p_0)$. We
+will look at the case for $t=2$. As shown in equation XXX, the second moment of
+every term is a function of the variance in neutral allele frequency in that
+generation, $x_t(1-x_t)$. This variance does not stay its initial value of
+$x_0(1-x_0)$; it decays each generation due to coalesce from the drift and
+selection processes modeled by this approach. The rate at which this decay
+happens is the effective population size each generation implied by this model. 
+We write (this differs from the original \cite{Santiago1995-hx} derivation),
+
+\begin{align}
+  \var(x_2 - p_0) &= \E({S_1}^2) + 2\E(S_1 S_2) + \E({S_2}^2) \\
+                  & + \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) \\
+  \frac{\var(x_2 - p_0)}{p_0(1-p_0)} &= \frac{V_h}{2N_{e,1}} + \frac{V_h Q_2^2}{2N_{e,2}}\left(1-\frac{1}{2N_{e,1}}\right)  +  \frac{V_h}{2N_{e,2}}  \left(1-\frac{1}{2N_{e,1}}\right)  \\ 
+                                     &+ \frac{V_n}{8N_{e,1}} + \frac{1}{4N_{e,1}} + \frac{V_n}{8N_{e,2}}  \left(1 - \frac{1}{2N_{e,1}}\right) 
++ \frac{1}{4N_{e,2}} \left(1 - \frac{1}{2N_{e,1}}\right).
+\end{align}
+
+A general equation of this is,
+
+\begin{align}
+  \frac{\var(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^\infty \left(\frac{V_h Q_i^2}{2 N_{e,i}} + \frac{V_n}{8N_{e,i}} + \frac{1}{4N_{e,i}} \right)\prod_{t=1}^{i-1} \left(1-\frac{1}{2 N_{e,t}}\right) \\
+\end{align}
+
+In the original derivation of \textcite{Santiago1995-hx}, they set each
+$N_{e,t}$ term to a single $N_e$ for which they solve for.
+
+This is an asymptotic $N_e$, but this does not accurately describe the levels
+of diversity (\cite{Santiago1998-bs} equation 16, \cite{Santiago2016-mu}
+equation 4).
+
+\begin{align}
+  \V(x_t) &= \V(x_1) \left(1 + \sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i} \right) + \sum_{i \ne j} C_{i,j}
+\end{align}
 

### 2022-10-09 manuscript/manuscript.tex
theory up to main sections

@@
                   &= \E({S_1}^2) + \E({S_2}^2) + \E({S_3}^2)  \\
-                  &+ 2\E(S_1 S_2) + 2\E(S_1 S_3) + 2\E(S_2 S_3) \\
+                  &+ \E(S_1 S_2) + \E(S_1 S_3) + \E(S_2 S_3) \\
                   &+ \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) + \E({D_3}^2) + \E({H_3}^2)
@@
 
-The covariance terms $\E(S_i S_j)$ for $j > i$ represent the expected
-neutral allele frequency change from a neutral allele becoming associated with
-a fitness background in generation $i$ and that fitness association persisting
-until generation $j$. Let us consider $S_1$ and $S_2$. We have,
+The covariance terms $\E(S_i S_j)$ for $j > i$ represent the expected neutral
+allele frequency change from a neutral allele becoming associated with a
+fitness background in generation $i$ and that fitness association persisting
+until generation $j$. However, the covariances terms $\E(S_j S_i)$ for $j > i$
+are zero since associations in the future cannot affect the past (see p. 1041
+of \cite{Buffalo2019-qs}). Let us consider $S_1$ and $S_2$. We have,
 
@@
 happens is the effective population size each generation implied by this model. 
-We write (this differs from the original \cite{Santiago1995-hx} derivation),
-
-\begin{align}
-  \var(x_2 - p_0) &= \E({S_1}^2) + 2\E(S_1 S_2) + \E({S_2}^2) \\
+We can write,
+
+\begin{align}
+  \var(x_2 - p_0) &= \E({S_1}^2) + \E(S_1 S_2) + \E({S_2}^2) \\
                   & + \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) \\
@@
 
-A general equation of this is,
-
-\begin{align}
-  \frac{\var(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^\infty \left(\frac{V_h Q_i^2}{2 N_{e,i}} + \frac{V_n}{8N_{e,i}} + \frac{1}{4N_{e,i}} \right)\prod_{t=1}^{i-1} \left(1-\frac{1}{2 N_{e,t}}\right) \\
-\end{align}
-
+In general, the evolution of the variance in allele frequency change in a
+system with polygenic selection is given by,
+
+\begin{align}
+  \label{eqn:var_xt}
+  \frac{\var(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^\infty \left(\frac{V_h Q_i^2}{2 N_{e,i}} + \frac{V_n}{8N_{e,i}} + \frac{1}{4N_{e,i}} \right)\prod_{t=1}^{i-1} \left(1-\frac{1}{2 N_{e,t}}\right). \\
+\end{align}
+
+Since the probability of coalesce (or, equivalently, identity by descent) is
+proportional to the variance in allele frequency change
+\parencite{Barton2000-zg}, this encodes the pairwise coalescent rate of the
+population through time under polygenic selection through the $N_{e,t}$ terms.
 In the original derivation of \textcite{Santiago1995-hx}, they set each
-$N_{e,t}$ term to a single $N_e$ for which they solve for.
-
-This is an asymptotic $N_e$, but this does not accurately describe the levels
-of diversity (\cite{Santiago1998-bs} equation 16, \cite{Santiago2016-mu}
-equation 4).
-
-\begin{align}
-  \V(x_t) &= \V(x_1) \left(1 + \sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i} \right) + \sum_{i \ne j} C_{i,j}
-\end{align}
+$N_{e,t}$ term to a single $N$ and solve this summation in the limit to
+infinity giving,
+
+\begin{align}
+  \label{eq:var_asymp}
+  \frac{\var(x_t - p_0)}{p_0(1-p_0)} &= \frac{4 V_h Q_\infty^2 +V_n+2}{4 (2 N-1)}.
+\end{align}
+
+To relate this to an asymptotic effective population size, note that in an
+ideal Wright--Fisher population,
+
+\begin{align}
+  \V(x_t) &= p_0(1-p_0) \left[1 - \left(1 - \frac{1}{2N_e}\right)^t \right]
+\end{align}
+
+such that if we look at the change in the variance in allele frequency change
+between adjacent generations $t$ and $t-1$, we get an implied $N_{e,t}$
+
+\begin{align}
+  \label{eqn:ne_var}
+  N_{e,t} = \frac{p_0(1-p_0) - \var(x_{t-1})}{2(\var(x_t) - \var(x_{t-1}))}
+\end{align}
+%
+(cf. p. 1018, \cite{Santiago1995-hx}).
+
+Solving this for equation \eqref{eq:var_asymp} gives the asymptotic effective
+population size,
+
+\begin{align}
+  %N_e &= \frac{2 (2 N-1)}{4 V_h Q_\infty^2 +V_n+2} \\
+  N_e &\approx \frac{4N}{4 V_h Q_\infty^2 +V_n+2}
+\end{align}
+
+(cf. equation 18, \cite{Santiago1995-hx}). Levels of pairwise diversity,
+however, depend on the full distribution of pairwise coalesce times, or the
+$N_{e,t}$. While this would require solving the recursion defined by equations
+\eqref{eqn:var_xt} and \eqref{eqn:ne_var}, \textcite{Santiago1998-bs} show
+that a good approximation is just assume a single effective population size for
+all timepoints, $N_{e,t} = N_e$, but use the time-indexed $Q_t$. This leads to,
+
+\begin{align}
+  N_{e,t} &\approx \frac{4N}{4 V_h Q_t^2 +V_n+2}
+\end{align}
+
+% HERE
+
+% \begin{align}
+%   \frac{\var(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^\infty \left(\frac{V_h Q_i^2}{2 N_{e,i}} + \frac{V_n}{8N_{e,i}} + \frac{1}{4N_{e,i}} \right)\prod_{t=1}^{i-1} \left(1-\frac{1}{2 N_{e,t}}\right) \\
+% \end{align}
+
+% The $N_{e,t}$ terms are analogous to an instantaneous coalesce rate at
+% generation $t$, which can be expressed in terms of the variance in allele
+% frequency change by noting that in an ideal Wright--Fisher population,
+
+% \begin{align}
+%   \V(x_t) &= p_0(1-p_0) \left[1 - \left(1 - \frac{1}{2N}\right)^t \right]
+% \end{align}
+
+% This is an asymptotic $N_e$, but this does not accurately describe the levels
+% of diversity (\cite{Santiago1998-bs} equation 16, \cite{Santiago2016-mu}
+% equation 4).
+
+% \begin{align}
+%   \V(x_t) &= \V(x_1) \left(1 + \sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i} \right) + \sum_{i \ne j} C_{i,j}
+% \end{align}
 

### 2022-10-12 manuscript/manuscript.tex
more theory

@@
 
-\subsection*{Overview of Quantitative Genetic Linked Selection Theory}
-
-Here, we step through a quick derivation of \textcite{Santiago1998-bs}.
+\subsection*{Overview of Quantitative Genetic Models of Effective Population Size}
+
+Here, we step through a quick derivation of \textcite{Santiago1995-hx}.
 Throughout, we assume random mating, hermaphroditic individuals, and a constant
@@
           &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + 2\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
-          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] + C_{1,2} \\
-          &= \frac{p_0(1-p_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{p_0(1-p_0)}{2N} + C_{1,2}
+          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] + \mathcal{C}_{1,2} \\
+          &= \frac{p_0(1-p_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{p_0(1-p_0)}{2N} + \mathcal{C}_{1,2}
 \end{align}
@@
 generation according to the probability of failing to coalesce each generation
-and the pairwise covariance terms $C_{i,j}$ that build up due to associations
-with heritable fitness backgrounds. Note that the covariance term $C_{1,2} =
+and the pairwise covariance terms $\mathcal{C}_{i,j}$ that build up due to associations
+with heritable fitness backgrounds. Note that the covariance term $\mathcal{C}_{1,2} =
 \E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] = 0$ under neutral
@@
 Now, we turn our attention to the case where there is heritable fitness
-variation, which leads to non-zero covariance terms $C_{i,j}$. These terms
-emerge when $\var(k)$ has a heritable component of fitness that can be
+variation, which leads to non-zero covariance terms $\mathcal{C}_{i,j}$. These
+terms emerge when $\var(k)$ has a heritable component of fitness that can be
 transmitted along with the neutral allele, thereby affecting the neutral
-allele's trajectory in later generations. If we write the offspring number into
-heritable and non-heritable components, $k_i = 2f_i + \varepsilon_i$, then
-$\var(k_i) = 4 V_h + V_n$. We can further decompose the $K_1$ term into $K_1 =
-S_1 + D_1$,
+allele's trajectory in later generations. If we partition the offspring number
+into heritable and non-heritable components, $k_i = 2f_i + \varepsilon_i$, then
+$\var(k_i) = 4 V_h + V_n$. When there is heritable fitness variation across
+individuals, $V_h > 0$. Because the population size is assumed to be constant,
+the mean heritable fitness across individuals is constrained to be $\E_i(f_i) =
+1$; this implies that the individual fitness values $f_i$ are relative
+fitnesses as is standard in population genetics, in this case to the population
+mean. Thus $V_h$ is the squared coefficient of heritable fitness variation
+(this is often denoted as $C^2 = \nicefrac{V_A}{\bar{w}^2}$, see
+\cite{Crow1958-pc,Charlesworth1987-ab,Houle1992-ur}).
+
+We can further decompose the $K_1$ term into $K_1 = S_1 + D_1$,
 
@@
 In general, the evolution of the variance in allele frequency change in a
-system with polygenic selection is given by,
+system with heritable fitness due to a single locus $r$ recombination fraction
+away from the neutral site is given by,
 
@@
 \parencite{Barton2000-zg}, this encodes the pairwise coalescent rate of the
-population through time under polygenic selection through the $N_{e,t}$ terms.
-In the original derivation of \textcite{Santiago1995-hx}, they set each
-$N_{e,t}$ term to a single $N$ and solve this summation in the limit to
-infinity giving,
+population through time under selection through the $N_{e,t}$ terms.  In the
+original derivation of \textcite{Santiago1995-hx}, they set each $N_{e,t}$ term
+to a single $N$ and solve this summation in the limit to infinity giving,
 
@@
 $N_{e,t}$. While this would require solving the recursion defined by equations
-\eqref{eqn:var_xt} and \eqref{eqn:ne_var}, \textcite{Santiago1998-bs} show
-that a good approximation is just assume a single effective population size for
-all timepoints, $N_{e,t} = N_e$, but use the time-indexed $Q_t$. This leads to,
-
-\begin{align}
+\eqref{eqn:var_xt} and \eqref{eqn:ne_var}, \textcite{Santiago1998-bs} show that
+a reasonably good approximation follows from using a single effective
+population size for all timepoints, $N_{e,t} = N_e$, but use the time-indexed
+$Q_t$. This leads to,
+
+\begin{align}
+  \label{eq:ne_t}
   N_{e,t} &\approx \frac{4N}{4 V_h Q_t^2 +V_n+2}
@@
 
-% HERE
+(c.f. equation 15 of \cite{Santiago1995-hx} and p. 1270 of \cite{Santiago2016-mu}).
+
+\subsection*{Effective Population Size Under Polygenic Selection}
+
+The results of the previous section assume only a single locus a recombination
+fraction $r$ apart from the neutral locus is determining the coalesce rate. The
+model of \textcite{Santiago1998-bs} extends this to the case where fitness is
+polygenic. As they show in their original paper, this system can accommodate a
+variety of different equilibrium selection systems as long as expressions for
+$V_h$ and $Z$ can be worked out. Other authors have extended similar models to
+more elaborate bbreeding structures \parencite{Wray1990-zf,Woolliams1993-qo}.
+
+First, we extend equation \eqref{eq:ne_t} to a polygenic system. Throughout, a
+multiplicative fitness model is assumed (i.e. independent effects, no
+epistasis), where the fitness of individual $i$ is the product of their fitness
+contributions from each locus $l$, $w_{i,l}$, giving $w_i = \prod_{l=1}^L
+w_{i,l}$. Then, assuming mean fitness is one, the total fitness variation is
+multiplicative across the locus-level fitnesses,
+
+\begin{align}
+  V_h %&= \E_i\left( w_i^2 \right) - 1 \\
+      %&= \prod_{l=1}^L (1 + \alpha_{l}^2 \V(g_l))  - 1  \\
+      &= \prod_{l=1}^L (1 + V_{h,l})  - 1
+\end{align}
+
+Now, the $Q_t^2 V_h$ term in the single-locus model is now under a
+multiplicative polygenic model,
+
+\begin{align}
+  Q_t^2 V_h &= \prod_{l=1}^L \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)  - 1
+\end{align}
+
+where $Q_{t,l}$ is the cumulative impact of selection due to the selected site
+$l$ (c.f. equation 4 \cite{Santiago1998-bs}). The factor of $\nicefrac{1}{2}$
+is due to the fact that we are ignoring the chance that the fitness background
+on the homologous chromosome recombines onto the haplotype with the tracked
+neutral allele (i.e. due to the $Q_t''$ associations). In other words, only
+half the fitness variation in a diploid can stay associated with the neutral
+allele through the generations.
+
+To simplify the derivation, assume $V_n = 2$ as under a Wright--Fisher model.
+Then, the effective population size is,
+
+\begin{align}
+  N_{e,t} &\approx \frac{N}{V_h Q_t^2 + 1} \\
+          &\approx \frac{N}{\prod_{l=1}^L \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)} \\
+          &\approx N \exp\left(- \sum_{l=1}^L \frac{V_{h,l} Q_{t,l}^2}{2}\right)
+\end{align}
+
+(c.f. \cite{Santiago1998-bs} equation 4b). This summation can then in turn be
+approximated by a integral; see \textcite{Santiago1998-bs} for more. Note that
+another way to arrive at this equation is to take the product of reduction
+factors $\nicefrac{N_e}{N}$ across selected sites, as in equation 4 of
+\textcite{Hudson1995-xc}.
+
+The heritable variance in fitness $V_h$ associated with the tracked neutral
+allele does not remain constant through the generations, but decays due to
+selection and drift, and is maintained by the input of new mutations. For an
+arbitrary polygenic selective system, the dynamics are extraordinarily
+complicated due to the complexity of multilocus selection. However, several
+quantitative genetic models simplify these dynamics by assuming an
+infinitesimal model, where the trait is sufficiently polygenic that the allele
+frequency changes at the underlying loci due to selection are negligible
+compared to drift. This allows for a critical assumption of the quantitative
+genetic models of linked selection, that the decay rate in variation is
+constant (and thus independent of the underlying allele frequency changes). In
+the Santiago and Caballero model, this is the decay factor $Z$.
+
+The theory of \textcite{Santiago1998-bs} is based on an equilibrium model,
+where $\Delta V_h = 0$ of the long run, even if the fitness variation
+\emph{associated} with the allele's haplotype is changing. This macro-scale
+
+\begin{align}
+  \Delta V_h &= V_m - V_h\left(1 - \frac{1}{2 N_e}\right) \\
+\end{align}
+
+\subsection*{Effective Population Size Under Strong Background Selection}
+
+Under strong background selection, multiple segregating deleterious sites
+contribute to fitness variation. BGS models assume a multiplicative fitness
+model, such that the fitness of an individual $i$ is
+
+\begin{align}
+  w_i = (1-s)^{n_i}
+\end{align}
+
+where $n_i$ is the number of deleterious mutations this individual carries.
+
+Classic background selection theory assumes an infinite population size, such
+that mutation frequencies are at their mutation-selection equilibrium and the
+expected number of mutations is $n_i = \nicefrac{\mu}{s}L$ where $L$ is the
+total number of basepairs that can be mutated (the mutational target size).
+Under a multiplicative fitness function in an infinite population, selection
+cannot generate linkage disequilibria \parencite{Turelli1990-kd}. Thus the
+classic strong background selection model assumes additive genetic fitness is
+equal to the additive genic fitness (which excludes the contribution of
+covariance between selected sites),
+
+\begin{align}
+  \V_i(w_i) &= 2 \sum_{l=1}^L p(1-p) s^2 \\
+            &= 2 L \frac{\mu}{s} \left( 1 - \frac{\mu}{s}\right) s^2 = U s + \mathcal{O}(\mu^2) \approx Us.
+\end{align}
+
+where we define $U := 2 L \mu$ to be the deleterious mutation rate per diploid
+genome, per generation.
+
+Next, \textcite{Santiago1998-bs} find a general
+equation for the factor $Z$ by which fitness variation associated with the
+neutral allele is reduced under mutation-selection balance. This is a
+macro-scale variable reflecting the average balance between the average loss of
+fitness variance due to selection under an infinitesimal model and drift. Each
+generation, the gain of new fitness variance per generation is $V_m$.
+
+\begin{align}
+  Z = 
+\end{align}
+
+Then, with under the strong background selection model,
+a site under purifying selection reduces diversity at a focal neutral site $r$
+by a factor of
+
+\begin{align}
+  \frac{N_{e,t}}{N} = \exp \left(-\frac{V_{h,1} Q_{t,1}^2}{2}\right)
+\end{align}
 
@@
 
-There has been considerable past work to find both the rate of the ratchet
-(XXX), and summaries of the genealogies created under weak purifying selection
-(XXX). Our approximation combines a few theoretic avenues from past work. Our
-primary goal is to find a theoretic scaling function $B$ that rescales the neutral
-pairwise coalescent times the appropriate amount to
-
-Strictly speaking, there is no way to rescale the coalescent under general
-background selection process, as genealogies deviate from the Kingman
-coalescent slightly under strong selection
-\parencite{Walczak2012-fi,Cvijovic2018-vd} and greatly under weak selection
-\parencite{OFallon2010-my}
-
-First, we approximate the coalescent effective population size, which
-determines expected pairwise diversity, from the variance effective population
-size. Then, we use a general quantitative genetic model of linked selection to
-find the variance effective population size under background selection. Since
-the impact of selection under this model is parameterized by the additive
-genetic fitness variance, we use an approximation for this based on the
-additive \emph{genic} variance, which depends on the rate of the ratchet.
-Previous work has been unable to find a simple closed-form solution to the rate
-of the ratchet because it depends on moments of the fitness distribution (e.g.
-fitness variation), which in turn depend on even higher-moments of the fitness
-distribution (e.g. the skew of the fitness distribution). While the ratchet
-should scale with the probability of fixation, the probability of fixation
-(Kimura, Malecot, XXX) is not accurate when multiple deleterious mutations are
-segregating simultaneously and in close linkage with one another. A key insight
-of \textcite{Santiago2016-mu} was that the classic equation for the probability
-of fixation is fairly robust as long as the correct effective population size
-is used. Their approach is to solve a system of two non-linear equations: one
-for the variance effective population size (which depends on the rate of the
-ratchet), and another for the time to fixation (which depends on the effective
-population size).
-
-Below, we step through the theory we use in our statistical methods. We
-re-derive central parts to illustrate where approximations have been made,
-which allows us to understand how these approximations may impact our
-statistical methods and inspect model fit in regions where these approximations
-are less likely to hold.
-
-\subsection*{Pairwise Diversity under Background Selection}
-
-The central quantity of interest for our statistical inference methods is
-average pairwise diversity along the chromosome. While other summaries of the
-data, such as the allele frequency spectrum, may be more informative about the
-evolutionary processes acting in a region, we lack population genetic theory
-for these quantities under most selective regimes and they require more degrees
-of freedom to characterize. By contrast, neutral diversity under an
-infinite-sites model is determined only by the per-basepair mutation rate $\mu$
-and the expected pairwise coalescent time $\E[T_2]$,
-
-\begin{align}
-  \E(\pi) = 2 \mu \E(T_2)
-\end{align}
-
-Let us assume that there is a constant \emph{drift-effective} population size
-$N_d$ which determines the rate of coalescence due only to neutral demographic
-processes. Additionally, we assume some background selection process that in
-expectation scales the effective population size by $b(t)$ at time $t$ (though
-this could in theory be due to any process). Then, the expected pairwise
-coalescent time is
-
-\begin{align}
-  \label{eq:T_2}
-  \E(T_2) = \sum_{i=1}^\infty \frac{i}{2 b(i) N_d}\prod_{t=1}^{i-1} \left(1-\frac{1}{2 b(t) N_d}\right).
-\end{align}
-
-The time-dependency of $b(t)$ is required because the background selection
-process will usually alter the rate of coalesce through time. For example,
-under weak and intermediate selection, coalesce rates speed up as lineages move
-in the past under intermediate to weak selection, similar to population growth
-\parencite{Walczak2012-fi,Seger2010-rs,Good2014-yz}.
-
-% Intuitively, this is
-% because lineages are unlikely to coalesce with numerous less fit lineages they
-% encounter as they go back in time, until ultimately coalescing in one of the
-% least-loaded fitness classes.
-
-This formulation is general and classic background selection theory (in the
-no-ratchet domain) is a special case of it. Under classic BGS and load theory,
-two randomly sampled lineages will each carry a number of deleterious mutations
-sampled from a Poisson stationary distribution. Because any lineage carrying $j
-> 0$ strongly deleterious mutations is destined to extinction, these lineages
-can only coalesce in a common ancestor in the least-loaded class with zero
-deleterious mutations. Under the Poisson stationary distribution, the size of
-this class is $f_0 N$, where $f_0 = e^{-\nicefrac{U}{s}}$, so the coalescent
-rate is scaled by $\nicefrac{1}{f_0}$. Once the lineages reach this fitness
-class, there are no fitness differences among individuals, so they coalesce
-following the Kingman process. A lineage moves backwards in time $t_d$
-generations through the ``delay phase" to the least-loaded class. Classic BGS
-theory ignores this delay phase, since $t_d \sim \log(j+1)/s << f_0 N$
-(Durrett, p. 213).
-
-%Here, we model the background selection process forward in time. We approximate
-%the expected pairwise diversity with the probability of heterozygosity,
-%$\E(\pi) \approx \mathcal{H}$, which holds under weak mutation and random
-%mating, such that
-
-%\begin{align}
-%  %\mathcal{H} =  2 \mu \sum_{i=1}^\infty \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right)
-%  \mathcal{H} &= 1 - \sum_{i=1}^\infty (1-\mu)^{2i} \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right) \\
-%              &\approx 2 \mu \sum_{i=1}^\infty \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right)
-%\end{align}
-
-%in the limit where $\mu \to 0$. This is equation 4 of
-%\textcite{Santiago2016-mu}; they define the sum as the heterozygosity effective
-%population size.
-
-\subsection*{Forward in Time Models of Linked Selection}
-
-Next, show how quantitative genetic models of linked selection
-\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu}
-find expressions for $b(t)$ with by modeling the selective process forward in
-time. Their approach is to consider how the variance in neutral allele
-frequency change is impacted by a linked selection process.
-
-\subsubsection*{Variance in Neutral Allele Frequency in a Neutral System}
-
-For comparison, we start with the recursion for the variance in allele
-frequency change in a neutral Wright--Fisher model. Under binomial sampling of
-alleles, the random frequency $x_t$ is $2N x_t \sim \mathrm{Binom}(x_{t-1},
-2N)$. If we look forward in time on generation and treat $x_0$ as fixed,
-
-\begin{align}
-  \V(x_1) = \E_1\left[(x_1 - x_0)^2\right]= \frac{x_0(1-x_0)}{2N},
-\end{align}
-
-and
-
-\begin{align}
-  \V(x_2) &= \E\left[(x_2 - x_0)^2\right] \\
-          &= \E_1\left[\E_2\left[((x_2-x_1) + (x_1 - x_0))^2 | x_1\right]\right] \\
-          &= \E_1\left[\E_2\left[(\Delta x_2 + \Delta x_1)^2 | x_1\right]\right] \\
-          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + 2\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
-          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
-          &= \frac{x_0(1-x_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{x_0(1-x_0)}{2N}
-\end{align}
-
-Note that the cross-moment $\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right]
-= 0$ under neutral evolution. In general, the variance in allele frequency
-after $t$ generations of neutral evolution is,
-
-\begin{align}
-  \label{eq:var_freq}
-  %\V(x_t) &= \frac{x_0(1-x_0)}{2N}\left(\sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i-1} \right)
-  \V(x_t) &= \V(x_1) \left(1 + \sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i} \right)
-\end{align}
-
-This geometric series converges to the familiar expression,
-
-\begin{align}
-  \V(x_t) &= x_0(1-x_0)\left(1 - \left(1 - \frac{1}{2N}\right)^t \right).
-\end{align}
-
-which in the limit $t \to \infty$, converges to the Bernoulli variance
-$x_0(1-x_0)$ as the frequency either goes to 1 with probability $x_0$ or 0 with
-probability $1-x_0$.
-
-% Now, if the effective population size were to change each generation, equation
-% \eqref{eq:var_freq} would be,
-
-% \begin{align}
-%   \V(x_t) &= \V(x_0) \left(1 + \sum_{i=1}^t \prod_{j=1}^{i} \left(1 - \frac{1}{2N(j)}\right) \right)
-% \end{align}
-
-% Note that this too converges to $x_0(1-x_0)$ as $t \to \infty$.
-
-\subsubsection*{Variance in Neutral Allele Frequency with Linked Selection}
-
-Note that equation XXX is a special case, when the evolutionary process has
-independent allele frequency changes through the generations and thus the
-cross-moment is zero. Selection generates correlations between allele frequency
-changes at nearby neutral sites. These correlations created by linked selection
-inflate the variance in neutral allele frequency $\V(x_t)$ beyond what it would
-be in a system without linked selection operating. The magnitude of this
-inflation term is set by the selection process and its dynamics, linkage
-disequilibrium, and the recombination rate. Even k
-
-The theory of this began with \textcite{Robertson1961-ho},
-and was thoroughly worked out by Santiago and Caballero
-(\citeyear{Santiago1995-hx}, \citeyear{Santiago1998-bs}; see also
-\cite{Barton2000-zg} and \cite{Buffalo2019-qs}). Our BGS statistical methods
-employ the extension of this model for weak BGS \parencite{Santiago2016-mu}.
-For completeness, we step through a quick derivation of the linked selection
-inflation factor models here, based on
-\textcite{Santiago1995-hx,Santiago1998-bs}.
-
-Quantitative genetic models of linked selection
-
-%%%%%
-
-One way we can model these dependencies is to partition $\Delta_i$ into drift
-and selection components. Over evolutionary replicates, the drift components
-are independent and the selection components have some degree of
-autocorrelation. Let us partition $\Delta x_i = \Delta_S x_i + \Delta_D x_i$,
-
-\begin{align}
-  \V(x_2) &= \V\left(\Delta x_2 + \Delta x_1\right) \\
-          &= \V(\Delta_D x_2) + \V(\Delta_D x_1) + 2 \cov(\Delta_S x_2, \Delta_S x_1) \\
-          &= \frac{x_0(1-x_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{x_0(1-x_0)}{2N} +  2 \cov(\Delta_S x_2, \Delta_S x_1) 
-\end{align}
-
-since $\Delta_D x_i \perp \Delta_D x_j$ for $i \ne j$. The covariances are
-determined by the selective process dynamics, linkage disequilibrium, and the
-recombination rate. These covariances are throughly worked out in
-\textcite{Santiago1995-hx} and \textcite{Buffalo2019-qs}, so we give a short
-derivation here. Throughout, we assume random mating, hermaphroditic
-individuals, and a constant population size. We will use angular brackets to
-indicate an average over all diploid individuals in the population, of some
-value with an individual-level index. For example, if we denote the number of
-allele an individual carries in generation $t$ as $z_{t,i} \in \{0, 1, 2\}$,
-the population frequency at time $x_t$ is
-
-\begin{align}
-  x_t = \langle z_t \rangle = \frac{1}{2N} \sum_{i=1}^N z_{t,i}
-\end{align}
-
-Across evolutionary replicates, each parent $i$ passes one of their two neutral
-alleles to a random number $k_i$ offspring. The \emph{heritable fitness} of an
-individual $f_i$ is defined as the expected number of offspring passed down
-across evolutionary replicates, such that $k_i = f_i + \varepsilon_i$.  Some
-selection process determines the distribution of $f_i$, but the this occurs
-independently of the process that sorts neutral alleles into diploid
-individuals. We are unsure of the neutral allele count in an individual in the
-initial generation, $z_{t,0}$, so we condition on it and take expectations
-later. Then, the frequency in the initial and first generations are,
-
-\begin{align}
-  x_0 = \langle z_0 \rangle = \frac{1}{2N} \sum_{i=1}^N z_{0,i} \\
-  x_1 = \langle z_1 \rangle = \frac{1}{2N} \sum_{i=1}^N \left( k_i \frac{z_{0,i}}{2} + \sum_{j=1}^{k_i} \delta_{i,j} \right)
-\end{align}
-
-The first term is the average number of neutral alleles each parent passes to
-its $k_i$ offspring. In reality, alleles are discrete, so each heterozygous
-parent also passes a random $\pm \nicefrac{1}{2}$ to each offspring, which
-leads to the second term. The variable $\delta_{i,j}$ is 0 if the parent is a
-homozygote, but if the parent is a heterozygote, is $\pm \nicefrac{1}{2}$ with
-equal probability over evolutionary replicates.
-
-Since individuals are exchangeable with their contribution to the
-heterozygosity term and population size is constant,
-
-\begin{align}
-  \sum_{i=1}^N \sum_{j=1}^{k_i} \delta_{i,j} = \sum_{i=1}^N \delta_{i} 
-\end{align}
-
-We can further partition the offspring number into heritable and non-heritable
-parts,
-
-\begin{align}
-  x_1 &= \frac{1}{2N} \sum_{i=1}^N \left( (f_i + \varepsilon_i) \frac{z_{0,i}}{2} + \delta_{i} \right) \\
-      &= \frac{1}{4N} \sum_{i=1}^N f_i z_{0,i} + \frac{1}{4N} \sum_{i=1}^N \varepsilon_i z_{0,i} + \frac{1}{2N} \sum_{i=1}^N \delta_{i} \\
-      &=  \frac{1}{4} \langle f \; z_{0} \rangle + \frac{1}{4} \langle \varepsilon \; z_{0} \rangle + \frac{1}{2}\langle \delta \rangle \\
-\end{align}
-
-% TODO factor 1/2 error
-
-\begin{align}
-  x_1 - x_0 &= \frac{1}{4} \langle f \; z_{0} \rangle + \frac{1}{4} \langle \varepsilon \; z_{0} \rangle + \frac{1}{2}\langle \delta \rangle - \langle z_0 \rangle \\
-            &= \frac{1}{4} \langle (f-1) \; z_{0} \rangle + \frac{1}{4} \langle \varepsilon \; z_{0} \rangle + \frac{1}{2}\langle \delta \rangle 
-\end{align}
-
-Now, we look at the variance of this over evolutionary replicates. By symmetry,
-the expected change in a neutral allele is $\E(\Delta x_t) = 0$ since we track
-an arbitrary neutral allele. Then the variance $\V(\Delta x_1) = \E( {\Delta
-x_1}^2)$,
-
-\begin{align}
-  \V(\Delta x_1) &= \frac{1}{4} \langle (f-1) \; z_{0} \rangle + \frac{1}{4} \langle \varepsilon \; z_{0} \rangle + \frac{1}{2}\langle \delta \rangle 
-\end{align}
-
-OLD
-Let us imagine there is some stochastic process $b(t)$ that rescales the coalescent
-rate,
-
-\begin{align}
-  \V(x_t) &= \frac{x_0(1-x_0)}{2N(0)}\left(\sum_{i=1}^t \prod_{j=1}^{i-1} \left(1 - \frac{1}{2b(j)N}\right) \right)
-\end{align}
-
-\begin{align}
-  \var(p_3 - p_0) = \E & \left( \bigg( \underbrace{S_1(1 + (1-r) G + (1-r)^2 G^2)}_{\text{creation and persistence of generation 1 associations} \;\; := \; S_1 Q_3} \right. +  D_1 + H_1+ \nonumber \\
-                       & \underbrace{S_2(1 + (1-r) G)}_{\text{creation and persistence of generation 2 associations} \;\; := \; S_2 Q_2}+ D_2 + H_2 + \nonumber \\
-                       &  \underbrace{S_3}_{\text{creation of generation 3 associations} \;\; := \; S_3 Q_1  }\left. + D_3 + H_3 \bigg)^2 \right). &
-    \label{eq:scr-var2}
-\end{align}
-%
-
-Note the similarity between the equations.
-
-\begin{align}
-  V &\propto \frac{1}{2b(0)N_d}\left(\sum_{i=1}^\infty \prod_{t=1}^{i-1} \left(1 - \frac{1}{2b(t)N_d}\right) \right) \\
-  W &\propto \sum_{i=1}^\infty \frac{i}{2 b(i) N_d} \prod_{t=1}^{i-1} \left(1-\frac{1}{2 b(t) N_d}\right) \\
-\end{align}
-
-Let us define the function $a(t)$ to be the product, which is the cumulative
-failure to coalesce up to time $t$.
-
-\begin{align}
-  a(t) = \prod_{i=1}^{t-1} \left( 1-\frac{1}{2 b(i) N_d} \right)
-\end{align}
-
-Comparing this to equation \eqref{eq:T_2} up to constant factors,
-
-\begin{align}
-  V &\propto \sum_{i=1}^\infty a(t) \\
-  W &\propto \sum_{i=1}^\infty i \; a(t) \\
-\end{align}
-
-end{align}
-
-XXX we swap indices
-
-Now, if we assume that the final coalesce in the waiting time process were to
-have a fixed rate of $\nicefrac{1}{2b(0)N_d}$, this
-
-we see that both the waiting time process and the variance process
-
-Intuitively, these define a stochastic selection process in two ways through
-the function $b(t)$ that modulates the rate of pairwise coalesce.
-
-where if $N_e(t) = N_e$ and $t=1$, this is the familiar Wright--Fisher variance
-$\V(p_1) = p_0(1-p_0) / 2N_e$. When selection is not acting, allele frequency
-changes per generation are independent, and neutral diversity is set by this
-single $N_e$.
-
-XXX
-
-Since we track an arbitrary neutral allele's frequency $p_i$, by symmetry,
-linked selection has no impact on the direction of change, $\E(\Delta p_i) =
-0$. We will work through a short proof of $\V(p_t)$ by induction following the
-approach of \textcite{Santiago1995-hx}. Throughout, we will use angular
-brackets to indicate an average over all diploid individuals in the population,
-of some value with an individual-level index. For example, if we denote the
-number of allele an individual carries in generation $t$ as $x_{t,i} \in \{0,
-1, 2\}$, the population frequency at time $p_t$ is
-
-\begin{align}
-  p_t = \langle x_t \rangle = \frac{1}{N} \sum_{i=1}^N x_{t,i}
-\end{align}
-
-\begin{align}
-  \Delta_H p &= \langle x (f-1) \rangle \\
-  \Delta_N p &= \langle x d \rangle \\
-  \Delta_M p &= \langle \Delta \rangle \\
-\end{align}
-
-Assuming uninherited, random environments, only the $\Delta_{_H} p$ terms
-create autocorrelation in $p_t$. The autocorrelation is created because during
-the random pairing of neutral alleles to fitness backgrounds, associations are
-created between an individual $i$'s allele frequency and their fitness $f_i$.
-This is because both fitness and the neutral alleles reside in the same
-individual in the present generation.
-
-We can imagine now how these associations change over evolutionary replicates
-in the next generation. In a fraction $r$ of the cases, the neutral allele 
-recombines off (or segregates away, if they are on different chromosomes) the
-fitness background. In a faction of $1-r$ XXX
-
-\begin{align}
-  \cov(\Delta_H p_t, \Delta_H p_s) &= \E(\langle x_t (f_t-1) \rangle , \langle x_s (f_s-1) \rangle)
-\end{align}
-
-Note that this covariance is only
-
-In the first generation, the frequency change
-can be partitioned into three components. Conditioning on the parent's
-heritable fitnesses, we can imagine constructing an infinite number of
-evolutionary replicates. Across all replicates, a random neutral allele's is
-perturbed
-
-\begin{align}
-  p_1 - p_0 = \sum_{i=1}^{N} 
-\end{align}
-
-Since the variance in neutral allele frequency change is also determined by the
-pairwise coalescence rate, this provides us with a way to model $b(t)$ under
-linked selection.
-
-These models derive an expression for the variance-effective population size of
-a focal neutral site, $N_v(t)$, based on how linked fitness variation perturbs
-the frequency trajectory at this site.  The variance-effective population size
-for a constant $N_e$ is defined by the variance in allele frequency change at
-the neutral site, over evolutionary replicates,
-
-where if $t=1$, this is the familiar Wright--Fisher variance $\V(p_1) =
-p_0(1-p_0) / 2N_e$. When selection is not acting, allele frequency changes per
-generation are independent, and neutral diversity is set by this
-single $N_e$.
-
-However when selection is acting at linked sites, neutral diversity is
-determined by the long-run $\V(p_t)$ as $t \to \infty$. Intuitively, this is
-because the rate of drift is magnified by autocorrelation in allele frequency
-change (XXX), and diversity is determined by the cumulative impact of these
-autocorrelations in the long-run. This autocorrelation effectively reduces the
-coalescence rate in a time-dependent manner, analogous to equation
-\eqref{eq:T_2}. Thus, a single constant $N_e$ is insufficient to describe
-levels of neutral variation, since it is insufficient
-
-The variance in allele frequency change after $t$ generations is
-
-\begin{align}
-  \V(p_t) = p_0(1-p_0)\left(1 - \prod_{i=1}^t \left(1-\frac{1}{2N_e(t)}\right)\right)
-\end{align}
-
-This is the basis of the quantitative genetic models of linked selection: a
-forward in time model for $\V(p_t)$ implies a pairwise coalescence rate
-function $b(t)$, which determines statistics based on pairwise coalescence such
-as $\E(\pi)$.
-
-Next, we consider how fitness variation impacts $\V(p_t)$.
-
-Throughout, we will take the variance $\V(p_t)$ to be the variance over
-evolutionary replicates. Since the variance is for a random neutral allele's
-frequency,
-
-a reasonable approximation to how background selection process
-rescales the coalescent rate at this site is given by $b(t) =
-\nicefrac{N_v(t)}{N_d}$.
-
-\begin{align}
-  N_v = \lim_{t \to \infty} \frac{p_0 (1 - p_0) - \var(p_{t-1})}{2\var(p_t) - \var(p_{t-1})}
-\end{align}
-
-\subsection*{Linked Selection Models for Variance-Effective Population Size}
-
-We wish to derive an expression for the reduction factor $B$. Ideally, this
-reduction factor would rescale the coalescent-effective population size, but
-this would require knowing the coalescent rate function $\lambda(t)$ over the
-evolutionary history $0 < t < \infty$. Instead, following previous work
-\parencite{Santiago1995-hx,Santiago1998-bs}, we derive the rescaling function
-$B$ relative to the variance-effective population size.
-
-Following past quantitative genetic models of linked selection
-\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu},
-we consider the long-run variance effective population size,
-
-\begin{align}
-  N_v := \lim_{t \to \infty} \frac{p_0(1-p_0)}{2\var\left(p_t - p_0\right)}
-\end{align}
-
-% ADD TO INTRO TODO
-Comment about characteristic scale $\nicefrac{\mu}{s}$. Is the fact that we see
-BGS on megabase scales because this is the only scale we can detect strong
-selection under?
-
-approximately $\pi \approx 4N_e \mu$ (for low mutation rate). However, this
-assumes a constant Kingman coalescent with coalescent effective population size
-$N_e$.
-
-\subsection*{}
-
-We model the reduction in pairwise coalescence times along the genome due to
-background selection by extending the work of \textcite{Santiago2016-mu}
-(hereafter, the SC16 model). The SC16 model derives from a class of linked
-selection \parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs} that
-predict the reduction in pairwise coalescence time at a focal neutral site
-using macro-state parameters such as fitness variation, rather than micro-state
-parameters like the frequency trajectories of selected alleles (e.g.
-\textcite{Maynard_Smith1974-zr}). We will briefly introduce the SC16 model's
-approach and how we validated this model could be extended to whole-genome
-inference.
-
-Classic background selection theory models how coalescence times are reduced by
-deleterious mutation-selection balance at neighboring sites. Building upon
-older models of genetic load \parencite{Crow1970-wj,Kimura1966-bk}, these
-models consider the equilibrium distribution of the \emph{number} of
-deleterious mutations segregating on haplotypes
-\parencite{Charlesworth1993-gb,Nordborg1996-nq}. Given an infinitely large
-population, or sufficiently strong selection, the deleterious mutations in
-these models cannot fix; consequently, any lineage containing
-
-While the effect of selection on linked selection distorts neighboring
-genealogies (XXX), skewing the site frequency spectrum (XXX) and altering
-linkage disequilibria along the chromosome, we focus on the impact of selection
-on pairwise coalescence rates. This simple summary is amenable to mathematical
-and statistical inference, and is robust to difficulties in identifying rare
-variants. The effect of background selection on pairwise coalescence rates is
-expressed as the reduction ratio $B = \nicefrac{N_e}{N}$ where $N$ is the
-drift-effective population size (i.e. including demographic processes) and
-$N_e$ is the effective population size under background selection.
-
-Under quantitative genetic models of linked selection, the reduction in
-effective population size is determined by the product of fitness variation,
-$V_f$, and an inflation factor $Q$. It is helpful to understand this by comparing
-it to a model of $N_e$ without \emph{heritable} fitness variation but with
-\emph{non-heritable} fitness variation. In this case,
-
-\begin{align}
-  N_e = \frac{4N}{2 + V_k},
-\end{align}
-
-where $V_k$ is the variance in offspring number across individuals. Under a
-constant population size Wright--Fisher model of reproduction, $V_k \approx 2$,
-and this simplifies to $N_e = N$. However, as first noted by
-\textcite{Robertson1961-ho}, linked selection inflates the variance in
-offspring number, as a neutral allele that becomes associated with a
-high-fitness or low-fitness genetic background has an $1-r$ chance of remaining
-associated with that background in the next generation, where $r$ is the chance
-of recombination (or independent assortment, $r=\nicefrac{1}{2}$).
-
-This persistence of this association inflates the effective variance in fitness
-by a factory $Q^2$,
-
-\begin{align}
-  N_e = \frac{4N}{2 + V_k + 4 Q^2 V_f},
-\end{align}
-
-\begin{align}
-  N_e = \frac{4N}{2 + V_k + 4 Q(r)^2 V_f},
-\end{align}
-
-\begin{align}
-  N_e = N \exp\left(-Q^2 V_f\right)
-\end{align}
-
-In a region of $L$ basepairs under purifying selection, there is a net flux of
-deleterious mutations in at rate $U = 2 \mu L$ per generation (where the factor
-of two is because in diploids, $2L$ is the total number of basepairs per genome).
-
-Typical models of background selection consider the equilibrium distribution of
-the \emph{count} of deleterious mutations, rather than then level of fitness
-variation implied by this count. If we assume fixed additive effects across
-alleles and multiplicative fitness effects across loci, each copy of a
-deleterious mutation an individual carries reduces their fitness by a factor of
-$1-s$. The total reduction for an individual is
-
-\begin{align}
-  x = (1-s)^n
-\end{align}
-
-where $n$ is the number of deleterious alleles an individual carries. Under
-mutation-selection balance, the population evolves to either a stationary
-distribution under the strong BGS domain, or a traveling distribution in weak
-selection case. Under some assumptions, equilibrium models of the strong domain
-show that the distribution $f(n) \sim \text{Pois}(\nicefrac{U}{s})$
-\parencite{Kimura1966-bk,Crow1970-wj} (CHECK) in the asexual case, and the
-fixation rate of deleterious mutations per generation, or ratchet rate, is $R =
-0$ since the probability of fixation, $p_F$ is zero. By contrast, as the
-selection coefficient grows weaker, $s \to 0$, the probability of fixation $p_F
-\to \nicefrac{1}{2N}$, and the ratchet approaches the neutral rate of $R =
-\mu$. The weak selection domain is of importance because both because it
-implies a certain likely irreversible rate of fitness reduction, and because it
-could possibly confound population genetic inference which assumes dynamics
-only under the strong model.
-
-As $s \to 0$, the fitness distribution, put on a scale reflecting the fitness
-relative at one moment in time, travels leftwards and changes shape. It has
-been difficult to find dynamic equations for the rate of the ratchet and
-fitness distribution, since lower moments of the distribution, such as the
-variation, depend on higher moments, such as the skew
-\parencite{Good2013-lp,Haigh1978-gt,Higgs1995-xc}. This is known as the ``moment
-closure problem" and it occurs regardless of whether one models the count
-number of deleterious mutations (XXX) or the fitness distribution (XXX).
-
-We can approximate the mapping between the distribution of deleterious
-mutations in the population, and the fitness distribution. It is worth noting
-that the count of deleterious mutations in an individual $i$ can be written
-$n_i = \sum_l^{L} g_l$ where each $g_l$ is an indicator variable for whether
-the individual carries a mutation at position $l$. Then, $\V_i(n) = \sum_l
-\V(g_l) + \sum_{i,j}\cov(g_i, g_j)$.
-
-the
-number of deleterious mutations in gamete $i$; thus $V(n) = \sum_i^{2N} V(x_i)
-+ \sum_{i \ne j} \cov(x_i, x_j)$ where the last term is the correlation among
-gamete mutation counts, which is an LD quantity. Then, assuming $s$ is small,
-we can write express this in terms of the additive genetic variance,
-
-\begin{align}
-  V(f) &\approx s^2 \V_a(n) + \Delta_\text{LD}
-\end{align}
-
-The first term is the additive genic variance, which is
-
-Under this model, the rate of the ratchet $R = U  - s V(n)$, which implies,
-
-\begin{align}
-  R = U - \nicefrac{V(f)}{s} \\
-  V(f) = s(U - R)
-\end{align}
-
-Throughout, it's worthwhile to keep into mind many population-level
-macro-states are simply distributional summaries of the current population. For
-example, the variance in deleterious mutations in the population is simply,
-$\V(n) = \nicefrac{1}{2N} \sum_i (x_i - \mu_x)^2$
-
-- We need to mention the two heterozygosities (S\&C '16 says they diverge under
-$N_e s < 1$ -- could explain anything we see?)
-
-- A point about reconciling the two micro and macro-state theories of
-population and quantitative genetics.
+%There has been considerable past work to find both the rate of the ratchet
+%(XXX), and summaries of the genealogies created under weak purifying selection
+%(XXX). Our approximation combines a few theoretic avenues from past work. Our
+%primary goal is to find a theoretic scaling function $B$ that rescales the neutral
+%pairwise coalescent times the appropriate amount to
+
+%Strictly speaking, there is no way to rescale the coalescent under general
+%background selection process, as genealogies deviate from the Kingman
+%coalescent slightly under strong selection
+%\parencite{Walczak2012-fi,Cvijovic2018-vd} and greatly under weak selection
+%\parencite{OFallon2010-my}
+
+%First, we approximate the coalescent effective population size, which
+%determines expected pairwise diversity, from the variance effective population
+%size. Then, we use a general quantitative genetic model of linked selection to
+%find the variance effective population size under background selection. Since
+%the impact of selection under this model is parameterized by the additive
+%genetic fitness variance, we use an approximation for this based on the
+%additive \emph{genic} variance, which depends on the rate of the ratchet.
+%Previous work has been unable to find a simple closed-form solution to the rate
+%of the ratchet because it depends on moments of the fitness distribution (e.g.
+%fitness variation), which in turn depend on even higher-moments of the fitness
+%distribution (e.g. the skew of the fitness distribution). While the ratchet
+%should scale with the probability of fixation, the probability of fixation
+%(Kimura, Malecot, XXX) is not accurate when multiple deleterious mutations are
+%segregating simultaneously and in close linkage with one another. A key insight
+%of \textcite{Santiago2016-mu} was that the classic equation for the probability
+%of fixation is fairly robust as long as the correct effective population size
+%is used. Their approach is to solve a system of two non-linear equations: one
+%for the variance effective population size (which depends on the rate of the
+%ratchet), and another for the time to fixation (which depends on the effective
+%population size).
+
+%Below, we step through the theory we use in our statistical methods. We
+%re-derive central parts to illustrate where approximations have been made,
+%which allows us to understand how these approximations may impact our
+%statistical methods and inspect model fit in regions where these approximations
+%are less likely to hold.
+
+%\subsection*{Pairwise Diversity under Background Selection}
+
+%The central quantity of interest for our statistical inference methods is
+%average pairwise diversity along the chromosome. While other summaries of the
+%data, such as the allele frequency spectrum, may be more informative about the
+%evolutionary processes acting in a region, we lack population genetic theory
+%for these quantities under most selective regimes and they require more degrees
+%of freedom to characterize. By contrast, neutral diversity under an
+%infinite-sites model is determined only by the per-basepair mutation rate $\mu$
+%and the expected pairwise coalescent time $\E[T_2]$,
+
+%\begin{align}
+%  \E(\pi) = 2 \mu \E(T_2)
+%\end{align}
+
+%Let us assume that there is a constant \emph{drift-effective} population size
+%$N_d$ which determines the rate of coalescence due only to neutral demographic
+%processes. Additionally, we assume some background selection process that in
+%expectation scales the effective population size by $b(t)$ at time $t$ (though
+%this could in theory be due to any process). Then, the expected pairwise
+%coalescent time is
+
+%\begin{align}
+%  \label{eq:T_2}
+%  \E(T_2) = \sum_{i=1}^\infty \frac{i}{2 b(i) N_d}\prod_{t=1}^{i-1} \left(1-\frac{1}{2 b(t) N_d}\right).
+%\end{align}
+
+%The time-dependency of $b(t)$ is required because the background selection
+%process will usually alter the rate of coalesce through time. For example,
+%under weak and intermediate selection, coalesce rates speed up as lineages move
+%in the past under intermediate to weak selection, similar to population growth
+%\parencite{Walczak2012-fi,Seger2010-rs,Good2014-yz}.
+
+%% Intuitively, this is
+%% because lineages are unlikely to coalesce with numerous less fit lineages they
+%% encounter as they go back in time, until ultimately coalescing in one of the
+%% least-loaded fitness classes.
+
+%This formulation is general and classic background selection theory (in the
+%no-ratchet domain) is a special case of it. Under classic BGS and load theory,
+%two randomly sampled lineages will each carry a number of deleterious mutations
+%sampled from a Poisson stationary distribution. Because any lineage carrying $j
+%> 0$ strongly deleterious mutations is destined to extinction, these lineages
+%can only coalesce in a common ancestor in the least-loaded class with zero
+%deleterious mutations. Under the Poisson stationary distribution, the size of
+%this class is $f_0 N$, where $f_0 = e^{-\nicefrac{U}{s}}$, so the coalescent
+%rate is scaled by $\nicefrac{1}{f_0}$. Once the lineages reach this fitness
+%class, there are no fitness differences among individuals, so they coalesce
+%following the Kingman process. A lineage moves backwards in time $t_d$
+%generations through the ``delay phase" to the least-loaded class. Classic BGS
+%theory ignores this delay phase, since $t_d \sim \log(j+1)/s << f_0 N$
+%(Durrett, p. 213).
+
+%%Here, we model the background selection process forward in time. We approximate
+%%the expected pairwise diversity with the probability of heterozygosity,
+%%$\E(\pi) \approx \mathcal{H}$, which holds under weak mutation and random
+%%mating, such that
+
+%%\begin{align}
+%%  %\mathcal{H} =  2 \mu \sum_{i=1}^\infty \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right)
+%%  \mathcal{H} &= 1 - \sum_{i=1}^\infty (1-\mu)^{2i} \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right) \\
+%%              &\approx 2 \mu \sum_{i=1}^\infty \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right)
+%%\end{align}
+
+%%in the limit where $\mu \to 0$. This is equation 4 of
+%%\textcite{Santiago2016-mu}; they define the sum as the heterozygosity effective
+%%population size.
+
+%\subsection*{Forward in Time Models of Linked Selection}
+
+%Next, show how quantitative genetic models of linked selection
+%\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu}
+%find expressions for $b(t)$ with by modeling the selective process forward in
+%time. Their approach is to consider how the variance in neutral allele
+%frequency change is impacted by a linked selection process.
+
+%\subsubsection*{Variance in Neutral Allele Frequency in a Neutral System}
+
+%For comparison, we start with the recursion for the variance in allele
+%frequency change in a neutral Wright--Fisher model. Under binomial sampling of
+%alleles, the random frequency $x_t$ is $2N x_t \sim \mathrm{Binom}(x_{t-1},
+%2N)$. If we look forward in time on generation and treat $x_0$ as fixed,
+
+%\begin{align}
+%  \V(x_1) = \E_1\left[(x_1 - x_0)^2\right]= \frac{x_0(1-x_0)}{2N},
+%\end{align}
+
+%and
+
+%\begin{align}
+%  \V(x_2) &= \E\left[(x_2 - x_0)^2\right] \\
+%          &= \E_1\left[\E_2\left[((x_2-x_1) + (x_1 - x_0))^2 | x_1\right]\right] \\
+%          &= \E_1\left[\E_2\left[(\Delta x_2 + \Delta x_1)^2 | x_1\right]\right] \\
+%          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + 2\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
+%          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
+%          &= \frac{x_0(1-x_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{x_0(1-x_0)}{2N}
+%\end{align}
+
+%Note that the cross-moment $\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right]
+%= 0$ under neutral evolution. In general, the variance in allele frequency
+%after $t$ generations of neutral evolution is,
+
+%\begin{align}
+%  \label{eq:var_freq}
+%  %\V(x_t) &= \frac{x_0(1-x_0)}{2N}\left(\sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i-1} \right)
+%  \V(x_t) &= \V(x_1) \left(1 + \sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i} \right)
+%\end{align}
+
+%This geometric series converges to the familiar expression,
+
+%\begin{align}
+%  \V(x_t) &= x_0(1-x_0)\left(1 - \left(1 - \frac{1}{2N}\right)^t \right).
+%\end{align}
+
+%which in the limit $t \to \infty$, converges to the Bernoulli variance
+%$x_0(1-x_0)$ as the frequency either goes to 1 with probability $x_0$ or 0 with
+%probability $1-x_0$.
+
+%% Now, if the effective population size were to change each generation, equation
+%% \eqref{eq:var_freq} would be,
+
+%% \begin{align}
+%%   \V(x_t) &= \V(x_0) \left(1 + \sum_{i=1}^t \prod_{j=1}^{i} \left(1 - \frac{1}{2N(j)}\right) \right)
+%% \end{align}
+
+%% Note that this too converges to $x_0(1-x_0)$ as $t \to \infty$.
+
+%\subsubsection*{Variance in Neutral Allele Frequency with Linked Selection}
+
+%Note that equation XXX is a special case, when the evolutionary process has
+%independent allele frequency changes through the generations and thus the
+%cross-moment is zero. Selection generates correlations between allele frequency
+%changes at nearby neutral sites. These correlations created by linked selection
+%inflate the variance in neutral allele frequency $\V(x_t)$ beyond what it would
+%be in a system without linked selection operating. The magnitude of this
+%inflation term is set by the selection process and its dynamics, linkage
+%disequilibrium, and the recombination rate. Even k
+
+%The theory of this began with \textcite{Robertson1961-ho},
+%and was thoroughly worked out by Santiago and Caballero
+%(\citeyear{Santiago1995-hx}, \citeyear{Santiago1998-bs}; see also
+%\cite{Barton2000-zg} and \cite{Buffalo2019-qs}). Our BGS statistical methods
+%employ the extension of this model for weak BGS \parencite{Santiago2016-mu}.
+%For completeness, we step through a quick derivation of the linked selection
+%inflation factor models here, based on
+%\textcite{Santiago1995-hx,Santiago1998-bs}.
+
+%Quantitative genetic models of linked selection
+
+%%%%%%
+
+%One way we can model these dependencies is to partition $\Delta_i$ into drift
+%and selection components. Over evolutionary replicates, the drift components
+%are independent and the selection components have some degree of
+%autocorrelation. Let us partition $\Delta x_i = \Delta_S x_i + \Delta_D x_i$,
+
+%\begin{align}
+%  \V(x_2) &= \V\left(\Delta x_2 + \Delta x_1\right) \\
+%          &= \V(\Delta_D x_2) + \V(\Delta_D x_1) + 2 \cov(\Delta_S x_2, \Delta_S x_1) \\
+%          &= \frac{x_0(1-x_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{x_0(1-x_0)}{2N} +  2 \cov(\Delta_S x_2, \Delta_S x_1) 
+%\end{align}
+
+%since $\Delta_D x_i \perp \Delta_D x_j$ for $i \ne j$. The covariances are
+%determined by the selective process dynamics, linkage disequilibrium, and the
+%recombination rate. These covariances are throughly worked out in
+%\textcite{Santiago1995-hx} and \textcite{Buffalo2019-qs}, so we give a short
+%derivation here. Throughout, we assume random mating, hermaphroditic
+%individuals, and a constant population size. We will use angular brackets to
+%indicate an average over all diploid individuals in the population, of some
+%value with an individual-level index. For example, if we denote the number of
+%allele an individual carries in generation $t$ as $z_{t,i} \in \{0, 1, 2\}$,
+%the population frequency at time $x_t$ is
+
+%\begin{align}
+%  x_t = \langle z_t \rangle = \frac{1}{2N} \sum_{i=1}^N z_{t,i}
+%\end{align}
+
+%Across evolutionary replicates, each parent $i$ passes one of their two neutral
+%alleles to a random number $k_i$ offspring. The \emph{heritable fitness} of an
+%individual $f_i$ is defined as the expected number of offspring passed down
+%across evolutionary replicates, such that $k_i = f_i + \varepsilon_i$.  Some
+%selection process determines the distribution of $f_i$, but the this occurs
+%independently of the process that sorts neutral alleles into diploid
+%individuals. We are unsure of the neutral allele count in an individual in the
+%initial generation, $z_{t,0}$, so we condition on it and take expectations
+%later. Then, the frequency in the initial and first generations are,
+
+%\begin{align}
+%  x_0 = \langle z_0 \rangle = \frac{1}{2N} \sum_{i=1}^N z_{0,i} \\
+%  x_1 = \langle z_1 \rangle = \frac{1}{2N} \sum_{i=1}^N \left( k_i \frac{z_{0,i}}{2} + \sum_{j=1}^{k_i} \delta_{i,j} \right)
+%\end{align}
+
+%The first term is the average number of neutral alleles each parent passes to
+%its $k_i$ offspring. In reality, alleles are discrete, so each heterozygous
+%parent also passes a random $\pm \nicefrac{1}{2}$ to each offspring, which
+%leads to the second term. The variable $\delta_{i,j}$ is 0 if the parent is a
+%homozygote, but if the parent is a heterozygote, is $\pm \nicefrac{1}{2}$ with
+%equal probability over evolutionary replicates.
+
+%Since individuals are exchangeable with their contribution to the
+%heterozygosity term and population size is constant,
+
+%\begin{align}
+%  \sum_{i=1}^N \sum_{j=1}^{k_i} \delta_{i,j} = \sum_{i=1}^N \delta_{i} 
+%\end{align}
+
+%We can further partition the offspring number into heritable and non-heritable
+%parts,
+
+%\begin{align}
+%  x_1 &= \frac{1}{2N} \sum_{i=1}^N \left( (f_i + \varepsilon_i) \frac{z_{0,i}}{2} + \delta_{i} \right) \\
+%      &= \frac{1}{4N} \sum_{i=1}^N f_i z_{0,i} + \frac{1}{4N} \sum_{i=1}^N \varepsilon_i z_{0,i} + \frac{1}{2N} \sum_{i=1}^N \delta_{i} \\
+%      &=  \frac{1}{4} \langle f \; z_{0} \rangle + \frac{1}{4} \langle \varepsilon \; z_{0} \rangle + \frac{1}{2}\langle \delta \rangle \\
+%\end{align}
+
+%% TODO factor 1/2 error
+
+%\begin{align}
+%  x_1 - x_0 &= \frac{1}{4} \langle f \; z_{0} \rangle + \frac{1}{4} \langle \varepsilon \; z_{0} \rangle + \frac{1}{2}\langle \delta \rangle - \langle z_0 \rangle \\
+%            &= \frac{1}{4} \langle (f-1) \; z_{0} \rangle + \frac{1}{4} \langle \varepsilon \; z_{0} \rangle + \frac{1}{2}\langle \delta \rangle 
+%\end{align}
+
+%Now, we look at the variance of this over evolutionary replicates. By symmetry,
+%the expected change in a neutral allele is $\E(\Delta x_t) = 0$ since we track
+%an arbitrary neutral allele. Then the variance $\V(\Delta x_1) = \E( {\Delta
+%x_1}^2)$,
+
+%\begin{align}
+%  \V(\Delta x_1) &= \frac{1}{4} \langle (f-1) \; z_{0} \rangle + \frac{1}{4} \langle \varepsilon \; z_{0} \rangle + \frac{1}{2}\langle \delta \rangle 
+%\end{align}
+
+%OLD
+%Let us imagine there is some stochastic process $b(t)$ that rescales the coalescent
+%rate,
+
+%\begin{align}
+%  \V(x_t) &= \frac{x_0(1-x_0)}{2N(0)}\left(\sum_{i=1}^t \prod_{j=1}^{i-1} \left(1 - \frac{1}{2b(j)N}\right) \right)
+%\end{align}
+
+%\begin{align}
+%  \var(p_3 - p_0) = \E & \left( \bigg( \underbrace{S_1(1 + (1-r) G + (1-r)^2 G^2)}_{\text{creation and persistence of generation 1 associations} \;\; := \; S_1 Q_3} \right. +  D_1 + H_1+ \nonumber \\
+%                       & \underbrace{S_2(1 + (1-r) G)}_{\text{creation and persistence of generation 2 associations} \;\; := \; S_2 Q_2}+ D_2 + H_2 + \nonumber \\
+%                       &  \underbrace{S_3}_{\text{creation of generation 3 associations} \;\; := \; S_3 Q_1  }\left. + D_3 + H_3 \bigg)^2 \right). &
+%    \label{eq:scr-var2}
+%\end{align}
+%%
+
+%Note the similarity between the equations.
+
+%\begin{align}
+%  V &\propto \frac{1}{2b(0)N_d}\left(\sum_{i=1}^\infty \prod_{t=1}^{i-1} \left(1 - \frac{1}{2b(t)N_d}\right) \right) \\
+%  W &\propto \sum_{i=1}^\infty \frac{i}{2 b(i) N_d} \prod_{t=1}^{i-1} \left(1-\frac{1}{2 b(t) N_d}\right) \\
+%\end{align}
+
+%Let us define the function $a(t)$ to be the product, which is the cumulative
+%failure to coalesce up to time $t$.
+
+%\begin{align}
+%  a(t) = \prod_{i=1}^{t-1} \left( 1-\frac{1}{2 b(i) N_d} \right)
+%\end{align}
+
+%Comparing this to equation \eqref{eq:T_2} up to constant factors,
+
+%\begin{align}
+%  V &\propto \sum_{i=1}^\infty a(t) \\
+%  W &\propto \sum_{i=1}^\infty i \; a(t) \\
+%\end{align}
+
+%end{align}
+
+%XXX we swap indices
+
+%Now, if we assume that the final coalesce in the waiting time process were to
+%have a fixed rate of $\nicefrac{1}{2b(0)N_d}$, this
+
+%we see that both the waiting time process and the variance process
+
+%Intuitively, these define a stochastic selection process in two ways through
+%the function $b(t)$ that modulates the rate of pairwise coalesce.
+
+%where if $N_e(t) = N_e$ and $t=1$, this is the familiar Wright--Fisher variance
+%$\V(p_1) = p_0(1-p_0) / 2N_e$. When selection is not acting, allele frequency
+%changes per generation are independent, and neutral diversity is set by this
+%single $N_e$.
+
+%XXX
+
+%Since we track an arbitrary neutral allele's frequency $p_i$, by symmetry,
+%linked selection has no impact on the direction of change, $\E(\Delta p_i) =
+%0$. We will work through a short proof of $\V(p_t)$ by induction following the
+%approach of \textcite{Santiago1995-hx}. Throughout, we will use angular
+%brackets to indicate an average over all diploid individuals in the population,
+%of some value with an individual-level index. For example, if we denote the
+%number of allele an individual carries in generation $t$ as $x_{t,i} \in \{0,
+%1, 2\}$, the population frequency at time $p_t$ is
+
+%\begin{align}
+%  p_t = \langle x_t \rangle = \frac{1}{N} \sum_{i=1}^N x_{t,i}
+%\end{align}
+
+%\begin{align}
+%  \Delta_H p &= \langle x (f-1) \rangle \\
+%  \Delta_N p &= \langle x d \rangle \\
+%  \Delta_M p &= \langle \Delta \rangle \\
+%\end{align}
+
+%Assuming uninherited, random environments, only the $\Delta_{_H} p$ terms
+%create autocorrelation in $p_t$. The autocorrelation is created because during
+%the random pairing of neutral alleles to fitness backgrounds, associations are
+%created between an individual $i$'s allele frequency and their fitness $f_i$.
+%This is because both fitness and the neutral alleles reside in the same
+%individual in the present generation.
+
+%We can imagine now how these associations change over evolutionary replicates
+%in the next generation. In a fraction $r$ of the cases, the neutral allele 
+%recombines off (or segregates away, if they are on different chromosomes) the
+%fitness background. In a faction of $1-r$ XXX
+
+%\begin{align}
+%  \cov(\Delta_H p_t, \Delta_H p_s) &= \E(\langle x_t (f_t-1) \rangle , \langle x_s (f_s-1) \rangle)
+%\end{align}
+
+%Note that this covariance is only
+
+%In the first generation, the frequency change
+%can be partitioned into three components. Conditioning on the parent's
+%heritable fitnesses, we can imagine constructing an infinite number of
+%evolutionary replicates. Across all replicates, a random neutral allele's is
+%perturbed
+
+%\begin{align}
+%  p_1 - p_0 = \sum_{i=1}^{N} 
+%\end{align}
+
+%Since the variance in neutral allele frequency change is also determined by the
+%pairwise coalescence rate, this provides us with a way to model $b(t)$ under
+%linked selection.
+
+%These models derive an expression for the variance-effective population size of
+%a focal neutral site, $N_v(t)$, based on how linked fitness variation perturbs
+%the frequency trajectory at this site.  The variance-effective population size
+%for a constant $N_e$ is defined by the variance in allele frequency change at
+%the neutral site, over evolutionary replicates,
+
+%where if $t=1$, this is the familiar Wright--Fisher variance $\V(p_1) =
+%p_0(1-p_0) / 2N_e$. When selection is not acting, allele frequency changes per
+%generation are independent, and neutral diversity is set by this
+%single $N_e$.
+
+%However when selection is acting at linked sites, neutral diversity is
+%determined by the long-run $\V(p_t)$ as $t \to \infty$. Intuitively, this is
+%because the rate of drift is magnified by autocorrelation in allele frequency
+%change (XXX), and diversity is determined by the cumulative impact of these
+%autocorrelations in the long-run. This autocorrelation effectively reduces the
+%coalescence rate in a time-dependent manner, analogous to equation
+%\eqref{eq:T_2}. Thus, a single constant $N_e$ is insufficient to describe
+%levels of neutral variation, since it is insufficient
+
+%The variance in allele frequency change after $t$ generations is
+
+%\begin{align}
+%  \V(p_t) = p_0(1-p_0)\left(1 - \prod_{i=1}^t \left(1-\frac{1}{2N_e(t)}\right)\right)
+%\end{align}
+
+%This is the basis of the quantitative genetic models of linked selection: a
+%forward in time model for $\V(p_t)$ implies a pairwise coalescence rate
+%function $b(t)$, which determines statistics based on pairwise coalescence such
+%as $\E(\pi)$.
+
+%Next, we consider how fitness variation impacts $\V(p_t)$.
+
+%Throughout, we will take the variance $\V(p_t)$ to be the variance over
+%evolutionary replicates. Since the variance is for a random neutral allele's
+%frequency,
+
+%a reasonable approximation to how background selection process
+%rescales the coalescent rate at this site is given by $b(t) =
+%\nicefrac{N_v(t)}{N_d}$.
+
+%\begin{align}
+%  N_v = \lim_{t \to \infty} \frac{p_0 (1 - p_0) - \var(p_{t-1})}{2\var(p_t) - \var(p_{t-1})}
+%\end{align}
+
+%\subsection*{Linked Selection Models for Variance-Effective Population Size}
+
+%We wish to derive an expression for the reduction factor $B$. Ideally, this
+%reduction factor would rescale the coalescent-effective population size, but
+%this would require knowing the coalescent rate function $\lambda(t)$ over the
+%evolutionary history $0 < t < \infty$. Instead, following previous work
+%\parencite{Santiago1995-hx,Santiago1998-bs}, we derive the rescaling function
+%$B$ relative to the variance-effective population size.
+
+%Following past quantitative genetic models of linked selection
+%\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu},
+%we consider the long-run variance effective population size,
+
+%\begin{align}
+%  N_v := \lim_{t \to \infty} \frac{p_0(1-p_0)}{2\var\left(p_t - p_0\right)}
+%\end{align}
+
+%% ADD TO INTRO TODO
+%Comment about characteristic scale $\nicefrac{\mu}{s}$. Is the fact that we see
+%BGS on megabase scales because this is the only scale we can detect strong
+%selection under?
+
+%approximately $\pi \approx 4N_e \mu$ (for low mutation rate). However, this
+%assumes a constant Kingman coalescent with coalescent effective population size
+%$N_e$.
+
+%\subsection*{}
+
+%We model the reduction in pairwise coalescence times along the genome due to
+%background selection by extending the work of \textcite{Santiago2016-mu}
+%(hereafter, the SC16 model). The SC16 model derives from a class of linked
+%selection \parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs} that
+%predict the reduction in pairwise coalescence time at a focal neutral site
+%using macro-state parameters such as fitness variation, rather than micro-state
+%parameters like the frequency trajectories of selected alleles (e.g.
+%\textcite{Maynard_Smith1974-zr}). We will briefly introduce the SC16 model's
+%approach and how we validated this model could be extended to whole-genome
+%inference.
+
+%Classic background selection theory models how coalescence times are reduced by
+%deleterious mutation-selection balance at neighboring sites. Building upon
+%older models of genetic load \parencite{Crow1970-wj,Kimura1966-bk}, these
+%models consider the equilibrium distribution of the \emph{number} of
+%deleterious mutations segregating on haplotypes
+%\parencite{Charlesworth1993-gb,Nordborg1996-nq}. Given an infinitely large
+%population, or sufficiently strong selection, the deleterious mutations in
+%these models cannot fix; consequently, any lineage containing
+
+%While the effect of selection on linked selection distorts neighboring
+%genealogies (XXX), skewing the site frequency spectrum (XXX) and altering
+%linkage disequilibria along the chromosome, we focus on the impact of selection
+%on pairwise coalescence rates. This simple summary is amenable to mathematical
+%and statistical inference, and is robust to difficulties in identifying rare
+%variants. The effect of background selection on pairwise coalescence rates is
+%expressed as the reduction ratio $B = \nicefrac{N_e}{N}$ where $N$ is the
+%drift-effective population size (i.e. including demographic processes) and
+%$N_e$ is the effective population size under background selection.
+
+%Under quantitative genetic models of linked selection, the reduction in
+%effective population size is determined by the product of fitness variation,
+%$V_f$, and an inflation factor $Q$. It is helpful to understand this by comparing
+%it to a model of $N_e$ without \emph{heritable} fitness variation but with
+%\emph{non-heritable} fitness variation. In this case,
+
+%\begin{align}
+%  N_e = \frac{4N}{2 + V_k},
+%\end{align}
+
+%where $V_k$ is the variance in offspring number across individuals. Under a
+%constant population size Wright--Fisher model of reproduction, $V_k \approx 2$,
+%and this simplifies to $N_e = N$. However, as first noted by
+%\textcite{Robertson1961-ho}, linked selection inflates the variance in
+%offspring number, as a neutral allele that becomes associated with a
+%high-fitness or low-fitness genetic background has an $1-r$ chance of remaining
+%associated with that background in the next generation, where $r$ is the chance
+%of recombination (or independent assortment, $r=\nicefrac{1}{2}$).
+
+%This persistence of this association inflates the effective variance in fitness
+%by a factory $Q^2$,
+
+%\begin{align}
+%  N_e = \frac{4N}{2 + V_k + 4 Q^2 V_f},
+%\end{align}
+
+%\begin{align}
+%  N_e = \frac{4N}{2 + V_k + 4 Q(r)^2 V_f},
+%\end{align}
+
+%\begin{align}
+%  N_e = N \exp\left(-Q^2 V_f\right)
+%\end{align}
+
+%In a region of $L$ basepairs under purifying selection, there is a net flux of
+%deleterious mutations in at rate $U = 2 \mu L$ per generation (where the factor
+%of two is because in diploids, $2L$ is the total number of basepairs per genome).
+
+%Typical models of background selection consider the equilibrium distribution of
+%the \emph{count} of deleterious mutations, rather than then level of fitness
+%variation implied by this count. If we assume fixed additive effects across
+%alleles and multiplicative fitness effects across loci, each copy of a
+%deleterious mutation an individual carries reduces their fitness by a factor of
+%$1-s$. The total reduction for an individual is
+
+%\begin{align}
+%  x = (1-s)^n
+%\end{align}
+
+%where $n$ is the number of deleterious alleles an individual carries. Under
+%mutation-selection balance, the population evolves to either a stationary
+%distribution under the strong BGS domain, or a traveling distribution in weak
+%selection case. Under some assumptions, equilibrium models of the strong domain
+%show that the distribution $f(n) \sim \text{Pois}(\nicefrac{U}{s})$
+%\parencite{Kimura1966-bk,Crow1970-wj} (CHECK) in the asexual case, and the
+%fixation rate of deleterious mutations per generation, or ratchet rate, is $R =
+%0$ since the probability of fixation, $p_F$ is zero. By contrast, as the
+%selection coefficient grows weaker, $s \to 0$, the probability of fixation $p_F
+%\to \nicefrac{1}{2N}$, and the ratchet approaches the neutral rate of $R =
+%\mu$. The weak selection domain is of importance because both because it
+%implies a certain likely irreversible rate of fitness reduction, and because it
+%could possibly confound population genetic inference which assumes dynamics
+%only under the strong model.
+
+%As $s \to 0$, the fitness distribution, put on a scale reflecting the fitness
+%relative at one moment in time, travels leftwards and changes shape. It has
+%been difficult to find dynamic equations for the rate of the ratchet and
+%fitness distribution, since lower moments of the distribution, such as the
+%variation, depend on higher moments, such as the skew
+%\parencite{Good2013-lp,Haigh1978-gt,Higgs1995-xc}. This is known as the ``moment
+%closure problem" and it occurs regardless of whether one models the count
+%number of deleterious mutations (XXX) or the fitness distribution (XXX).
+
+%We can approximate the mapping between the distribution of deleterious
+%mutations in the population, and the fitness distribution. It is worth noting
+%that the count of deleterious mutations in an individual $i$ can be written
+%$n_i = \sum_l^{L} g_l$ where each $g_l$ is an indicator variable for whether
+%the individual carries a mutation at position $l$. Then, $\V_i(n) = \sum_l
+%\V(g_l) + \sum_{i,j}\cov(g_i, g_j)$.
+
+%the
+%number of deleterious mutations in gamete $i$; thus $V(n) = \sum_i^{2N} V(x_i)
+%+ \sum_{i \ne j} \cov(x_i, x_j)$ where the last term is the correlation among
+%gamete mutation counts, which is an LD quantity. Then, assuming $s$ is small,
+%we can write express this in terms of the additive genetic variance,
+
+%\begin{align}
+%  V(f) &\approx s^2 \V_a(n) + \Delta_\text{LD}
+%\end{align}
+
+%The first term is the additive genic variance, which is
+
+%Under this model, the rate of the ratchet $R = U  - s V(n)$, which implies,
+
+%\begin{align}
+%  R = U - \nicefrac{V(f)}{s} \\
+%  V(f) = s(U - R)
+%\end{align}
+
+%Throughout, it's worthwhile to keep into mind many population-level
+%macro-states are simply distributional summaries of the current population. For
+%example, the variance in deleterious mutations in the population is simply,
+%$\V(n) = \nicefrac{1}{2N} \sum_i (x_i - \mu_x)^2$
+
+%- We need to mention the two heterozygosities (S\&C '16 says they diverge under
+%$N_e s < 1$ -- could explain anything we see?)
+
+%- A point about reconciling the two micro and macro-state theories of
+%population and quantitative genetics.
 

### 2022-10-18 manuscript/manuscript.tex
more theory

@@
 $V_h$, with the fitness variance associated with a particular background
-decaying at a simple geometric decay at rate $Z$ each generation, which is
-sufficient for background selection. An approximate rate of the decay can be
+decaying at a simple geometric decay at rate $1-\kappa$ each generation, which
+is sufficient for background selection. An approximate rate of the decay can be
 worked out for the particular selective system. In general, this can be a much
@@
 \begin{align}
-  S_{2}' &= S_1 Z(1-r) \\
-  S_{3}' &= S_1 Z^2(1-r)^2 \\
-  S_{4}' &= S_1 Z^3(1-r)^3.
+  S_{2}' &= S_1 (1-\kappa)(1-r) \\
+  S_{3}' &= S_1 (1-\kappa)^2(1-r)^2 \\
+  S_{4}' &= S_1 (1-\kappa)^3(1-r)^3.
 \end{align}
@@
 \begin{align}
+  \label{eq:qterm}
   S_1 Q_t &= S_1 + S_2' + S_3' + \ldots + S_t' \\
-  Q_t &= 1 + \sum_{i=1}^t r^i (1-Z)^i \\
-  Q_\infty &= \frac{1}{1-r (1-Z)}
+  Q_t &= 1 + \sum_{i=1}^t \left((1-r)\kappa\right)^i \\
+  Q_\infty &= \frac{1}{1-(1-r) \kappa}
 \end{align}
@@
 \begin{align}
-  \E(S_1 S_2) &= \E(S_1^2) Z(1-r) = \E(S_1^2) Q_2^2 \\
-  \E(S_1 S_3) &= \E(S_1^2) Z^2(1-r)^2 = \E(S_1^2) Q_3^2.
+  \E(S_1 S_2) &= \E(S_1^2) (1-\kappa)(1-r) = \E(S_1^2) Q_2^2 \\
+  \E(S_1 S_3) &= \E(S_1^2) (1-\kappa)^2(1-r)^2 = \E(S_1^2) Q_3^2.
 \end{align}
@@
 variety of different equilibrium selection systems as long as expressions for
-$V_h$ and $Z$ can be worked out. Other authors have extended similar models to
+$V_h$ and $1-\kappa$ can be worked out. Other authors have extended similar models to
 more elaborate bbreeding structures \parencite{Wray1990-zf,Woolliams1993-qo}.
@@
 
-(c.f. \cite{Santiago1998-bs} equation 4b). This summation can then in turn be
+(c.f. \cite{Santiago1998-bs} equation 4b). Santiago and Caballero ignore the
+covariances between the locus-specific variation decay rates $1-\kappa_l$ terms
+and the locus-specific variances, $V_{h,l}$. This allows for the terms to be
+separated,
+
+\begin{align}
+  N_{e,t} &\approx N \exp\left(- \frac{V_h}{2} \sum_{l=1}^L Q_{t,l}^2\right).
+\end{align}
+
+\subsection*{Continuous Approximation to the Segment Under Selection}
+
+Our model differs from that of \textcite{Santiago1998-bs} in that we do not
+integrate over the entire genome. In their 1998 model, Santiago and Caballero
+consider the impact of both sites under selection on the same chromosome, as
+well as sites on independently assorting chromosomes. We only consider the
+contribution of a single segment under purifying selection, and the reduction
+from additional segments accumulate multiplicatively (similar to our
+implementation of the classic BGS model, XXX).
+
+If we assume that the $L$ selected sites are evenly distributed in a region of
+length $G$ basepairs, we can rewrite the sum as $\nicefrac{L}{G} \sum_{l=1}^G
+Q_{t,l}^2$
+
+\begin{align}
+  N_{e,t} &\approx N \exp\left(- \frac{V_h}{2} \sum_{l=1}^L Q_{t,l}^2\right).
+\end{align}
+
+For a continuous recombination map, the recombination fraction between the
+neutral allele and a selected site $d$ map distance away is given by Haldane's
+map function, $r = (1-e^{-2d})/2$. Santiago and Caballero assume the focal
+neutral site is in the middle of the chromosome, and integrate over the map
+distance. Assuming that linkage is not too tight,
+
+\begin{align}
+  Q_\infty &= \frac{1}{1-r \kappa}
+\end{align}
+
+For our model, we consider the contribution of individual contiguous segments
+of basepairs rather than the whole genome.
+
+This summation can then in turn be
 approximated by a integral; see \textcite{Santiago1998-bs} for more. Note that
@@
 
+\subsection*{The Variance Dynamics}
+
 The heritable variance in fitness $V_h$ associated with the tracked neutral
@@
 arbitrary polygenic selective system, the dynamics are extraordinarily
-complicated due to the complexity of multilocus selection. However, several
-quantitative genetic models simplify these dynamics by assuming an
-infinitesimal model, where the trait is sufficiently polygenic that the allele
-frequency changes at the underlying loci due to selection are negligible
-compared to drift. This allows for a critical assumption of the quantitative
-genetic models of linked selection, that the decay rate in variation is
-constant (and thus independent of the underlying allele frequency changes). In
-the Santiago and Caballero model, this is the decay factor $Z$.
-
-The theory of \textcite{Santiago1998-bs} is based on an equilibrium model,
-where $\Delta V_h = 0$ of the long run, even if the fitness variation
-\emph{associated} with the allele's haplotype is changing. This macro-scale
-
-\begin{align}
-  \Delta V_h &= V_m - V_h\left(1 - \frac{1}{2 N_e}\right) \\
-\end{align}
+complicated due to the complexity of multilocus selection, and the dynamics of
+the trait response to selection depend on the variance, which in turn depends
+on higher moments of the fitness distribution. We discuss the simplification of
+Santiago and Caballero's model here; hereafter, we assume that $V_h = V_A$, the
+genetic variance in additive fitness put on a relative scale such that
+$\E_i(f_i) = 1$.
+
+Each generation, the genetic variance changes due to changes in the allele
+frequencies at selected sites and linkage disequilibria between these sites.
+These are typically expressed as recursions, but certain models permit these
+changes to be expressed as a proportional reduction (e.g. truncation selection
+and stabilizing selection models; \cite{Keightley1988-eq} p. 36, and
+\cite{Walsh2018-bt} p. 557). Santiago and Caballero's model follows this
+approach, assuming the variance decays due to selection and drift at a constant
+rate $Z := 1-\kappa$ through the generations, such that the variance in the next
+generation excluding mutation is $V_A' = (1-\kappa) V_A$ and $\Delta V_A =
+-\kappa V_A$.
+
+With mutation, $\Delta V_A = V_m - \kappa V_A$. The model of Santiago and
+Caballero assumes that under the long-run equilibrium, $\Delta V_A = 0$, such
+that $\kappa V_A = V_m$, and thus the rate of loss is $\kappa =
+\nicefrac{V_m}{V_A}$. Then the variance for the dynamics simplify to,
+
+\begin{align}
+  \label{eq:Z}
+  V_A' &= \left(1 - \frac{V_m}{V_A}\right) V_A \\
+  Z &:= \frac{V_A'}{V_A} = 1 - \kappa = 1 - \frac{V_m}{V_A}
+\end{align}
+
+(c.f. \cite{Santiago1998-bs} equation 9).
+
+% TODO this is a good description that could make its way to maintext
 
@@
 that mutation frequencies are at their mutation-selection equilibrium and the
-expected number of mutations is $n_i = \nicefrac{\mu}{s}L$ where $L$ is the
-total number of basepairs that can be mutated (the mutational target size).
-Under a multiplicative fitness function in an infinite population, selection
-cannot generate linkage disequilibria \parencite{Turelli1990-kd}. Thus the
-classic strong background selection model assumes additive genetic fitness is
-equal to the additive genic fitness (which excludes the contribution of
-covariance between selected sites),
-
-\begin{align}
-  \V_i(w_i) &= 2 \sum_{l=1}^L p(1-p) s^2 \\
+expected number of mutations across individuals is $\E_i(n_i) =
+\nicefrac{\mu}{s}L$ where $L$ is the total number of basepairs that can be
+mutated (the mutational target size).  Under a multiplicative fitness function
+in an infinite population, selection cannot generate linkage disequilibria
+\parencite{Turelli1990-kd}. Thus the classic strong background selection model
+assumes additive genetic fitness is equal to the additive genic fitness (which
+excludes the contribution of covariance between selected sites),
+
+\begin{align}
+  V_\text{BGS} &:= \V_i(w_i) = 2 \sum_{l=1}^L p(1-p) s^2 \\
             &= 2 L \frac{\mu}{s} \left( 1 - \frac{\mu}{s}\right) s^2 = U s + \mathcal{O}(\mu^2) \approx Us.
@@
 
-where we define $U := 2 L \mu$ to be the deleterious mutation rate per diploid
-genome, per generation.
-
-Next, \textcite{Santiago1998-bs} find a general
-equation for the factor $Z$ by which fitness variation associated with the
-neutral allele is reduced under mutation-selection balance. This is a
-macro-scale variable reflecting the average balance between the average loss of
-fitness variance due to selection under an infinitesimal model and drift. Each
-generation, the gain of new fitness variance per generation is $V_m$.
-
-\begin{align}
-  Z = 
-\end{align}
-
-Then, with under the strong background selection model,
-a site under purifying selection reduces diversity at a focal neutral site $r$
-by a factor of
-
-\begin{align}
-  \frac{N_{e,t}}{N} = \exp \left(-\frac{V_{h,1} Q_{t,1}^2}{2}\right)
-\end{align}
-
-% \begin{align}
-%   \frac{\var(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^\infty \left(\frac{V_h Q_i^2}{2 N_{e,i}} + \frac{V_n}{8N_{e,i}} + \frac{1}{4N_{e,i}} \right)\prod_{t=1}^{i-1} \left(1-\frac{1}{2 N_{e,t}}\right) \\
-% \end{align}
+where $U$ is defined $U := 2 L \mu$, the deleterious mutation rate per diploid
+genome, per generation. They then set $V_A = V_\text{BGS}$, and note that $V_m
+\approx Us^2$, since mutations have frequency of $\nicefrac{1}{2N}$ and each
+increase the variance by a factor of $s^2$. Note that because the classic BGS
+model assumes deterministic selection dynamics (i.e. an infinite population),
+the equilibrium $Z$ given by equation \eqref{eq:Z} excludes the reduction due
+to drift, which reduces the variance by an additional factor $1 -
+\nicefrac{1}{2N_e}$. Thus, the reduction factor under the strong BGS model is,
+
+\begin{align}
+  Z \left(1 - \frac{1}{2N_e}\right) = (1 - s)\left(1 - \frac{1}{2N_e}\right) \approx 1 - s -\frac{1}{2N_e}
+\end{align}
+
+(c.f. \cite{Santiago1998-bs} equation 10). An alternate derivation can be found
+by working out the deterministic reduction in variance from the single-locus
+dynamics. Thus, $Q' = $
+
+the reduction under the classic BGS model is,
+
+\begin{align}
+  N_e = N \exp()
+\end{align}
+
+\subsection*{Weak Selection Dynamics}
+
+With the possibility of weak selection, deleterious alleles can drift up in
+frequency and fix. Consequently, the fitness distribution is no longer
+stationary, as the set of haplotypes without any deleterious mutations can be
+lost due to drift. The approach of \textcite{Santiago2016-mu} that we implement
+in our methods is to derive an equation for
 

### 2022-10-19 manuscript/manuscript.tex
so tired of theory

@@
   S_1 Q_t &= S_1 + S_2' + S_3' + \ldots + S_t' \\
-  Q_t &= 1 + \sum_{i=1}^t \left((1-r)\kappa\right)^i \\
-  Q_\infty &= \frac{1}{1-(1-r) \kappa}
-\end{align}
-
-which shows the long reach of heritable fitness factors in altering allele
+  Q_t &= 1 + \sum_{i=1}^t \left((1-r)\kappa\right)^i
+\end{align}
+
+which converges to,
+
+\begin{align}
+  \label{eq:Qinf}
+  Q_\infty &= \frac{1}{1-(1-r) \kappa}.
+\end{align}
+
+This shows the long reach of heritable fitness factors in altering allele
 frequency through the generations. In terms like $\E(S_1 S_2)$, a fraction
@@
 epistasis), where the fitness of individual $i$ is the product of their fitness
-contributions from each locus $l$, $w_{i,l}$, giving $w_i = \prod_{l=1}^L
-w_{i,l}$. Then, assuming mean fitness is one, the total fitness variation is
-multiplicative across the locus-level fitnesses,
+contributions from each locus $l$, $w_{i,l}$, giving $w_i = \prod_{l=1}^n
+w_{i,l}$. Then, assuming mean fitness is one and independence between sites,
+the total fitness variation is multiplicative across the locus-level fitnesses,
 
@@
   V_h %&= \E_i\left( w_i^2 \right) - 1 \\
-      %&= \prod_{l=1}^L (1 + \alpha_{l}^2 \V(g_l))  - 1  \\
-      &= \prod_{l=1}^L (1 + V_{h,l})  - 1
-\end{align}
-
-Now, the $Q_t^2 V_h$ term in the single-locus model is now under a
-multiplicative polygenic model,
-
-\begin{align}
-  Q_t^2 V_h &= \prod_{l=1}^L \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)  - 1
+      %&= \prod_{l=1}^n (1 + \alpha_{l}^2 \V(g_l))  - 1  \\
+      &= \prod_{l=1}^n (1 + V_{h,l})  - 1
+\end{align}
+
+Similarly, the $Q_t^2 V_h$ term in the single-locus model is multiplicative
+under the polygenic model, as the change in site-specific fitness variances are
+assumed independent,
+
+\begin{align}
+  Q_t^2 V_h &= \prod_{l=1}^n \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)  - 1
 \end{align}
@@
   N_{e,t} &\approx \frac{N}{V_h Q_t^2 + 1} \\
-          &\approx \frac{N}{\prod_{l=1}^L \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)} \\
-          &\approx N \exp\left(- \sum_{l=1}^L \frac{V_{h,l} Q_{t,l}^2}{2}\right)
-\end{align}
-
-(c.f. \cite{Santiago1998-bs} equation 4b). Santiago and Caballero ignore the
-covariances between the locus-specific variation decay rates $1-\kappa_l$ terms
-and the locus-specific variances, $V_{h,l}$. This allows for the terms to be
-separated,
-
-\begin{align}
-  N_{e,t} &\approx N \exp\left(- \frac{V_h}{2} \sum_{l=1}^L Q_{t,l}^2\right).
+          &\approx \frac{N}{\prod_{l=1}^n \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)} \\
+          &\approx N \exp\left(- \sum_{l=1}^n \frac{V_{h,l} Q_{t,l}^2}{2}\right)
+\end{align}
+
+(c.f. \cite{Santiago1998-bs} equation 4b). If we write each site $l$'s
+contribution to the variance as a deviation from the average contribution,
+$V_{h,l}^2 = \nicefrac{1}{n} V_h + \varepsilon_l$, then
+
+\begin{align}
+  N_{e,t} &\approx N \exp\left(- \frac{1}{2}\sum_{l=1}^n \left(\frac{1}{n}V_{h} + \varepsilon_l\right)  Q_{t,l}^2\right) \\
+          &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{t,l}^2+ \sum_{l=1}^n \varepsilon_l  Q_{t,l}^2\right)
+\end{align}
+
+Santiago and Caballero assume that over evolutionary replicates, as selected
+sites are randomly distributed, there is not systematic covariance between the
+fitness variation at a site and the decay of its association with the neutral
+site, e.g. $\E(\varepsilon_l Q_{t,l}^2) = 0$, and thus ignore the second sum
+term.
+
+\begin{align}
+      N_{e,t} &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{t,l}^2\right) \\
 \end{align}
@@
 from additional segments accumulate multiplicatively (similar to our
-implementation of the classic BGS model, XXX).
-
-If we assume that the $L$ selected sites are evenly distributed in a region of
-length $G$ basepairs, we can rewrite the sum as $\nicefrac{L}{G} \sum_{l=1}^G
-Q_{t,l}^2$
-
-\begin{align}
-  N_{e,t} &\approx N \exp\left(- \frac{V_h}{2} \sum_{l=1}^L Q_{t,l}^2\right).
-\end{align}
-
-For a continuous recombination map, the recombination fraction between the
-neutral allele and a selected site $d$ map distance away is given by Haldane's
-map function, $r = (1-e^{-2d})/2$. Santiago and Caballero assume the focal
-neutral site is in the middle of the chromosome, and integrate over the map
-distance. Assuming that linkage is not too tight,
-
-\begin{align}
-  Q_\infty &= \frac{1}{1-r \kappa}
-\end{align}
-
-For our model, we consider the contribution of individual contiguous segments
-of basepairs rather than the whole genome.
-
-This summation can then in turn be
-approximated by a integral; see \textcite{Santiago1998-bs} for more. Note that
-another way to arrive at this equation is to take the product of reduction
-factors $\nicefrac{N_e}{N}$ across selected sites, as in equation 4 of
-\textcite{Hudson1995-xc}.
+implementation of the classic BGS model, XXX). Like their model, we imagine the
+reduction experienced by a focal neutral site in the middle of the segment.
+The implementation of our method differs slightly on though, see XXX.
+
+Since $Q_{t,l}$ terms depend on $l$ only through the recombination fraction
+between the neutral allele and the selected site, $r(l)$. Assuming a constant
+per-basepair recombination rate of $r_\text{BP}$, $r(l) = r_\text{BP} l$ we can
+approximate the sum with an integral for the asymptotic $Q_\infty$. The
+reduction experienced in the middle of a chromosome is twice the reduction
+experienced from each half,
+
+\begin{align}
+      N_{e,\infty} &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{\infty,l}^2\right) \\
+              &\approx N \exp\left(- \frac{V_{h}}{L} \int_0^{\nicefrac{L}{2}} Q_{\infty}(r)^2 dr \right) \\
+\end{align}
+
+where $L = n r_\text{BP}$ is the total segment length in Morgans. Let us define
+the total asymptotic effect of associations from all sites as $Q_\infty^2$.
+Then,
+
+\begin{align}
+  Q_\infty^2 &= \frac{2}{L}\int_0^{\nicefrac{n r_\text{BP}}{2}} Q_{\infty}(x)^2 dx  \\
+  &=  \frac{2}{L} \int_0^{\nicefrac{L}{2}} \left(\frac{1}{1 - (1-x)Z}\right)^2 dx  \\
+  &= \frac{2}{(1-Z) (2-(2-L) Z)}
+\end{align}
+
+Santiago and Caballero further approximate this; we can derive their
+approximation by setting $Z = 1-\kappa$ and doing a first-order Taylor series
+expansion around $\kappa$ (since the loss in variance is presumed to be small).
+Then,
+
+\begin{align}
+  Q_\infty^2 = \frac{2}{L (1-Z)} + \frac{2 (L-2)}{L^2}+\frac{2 \left(L^2-4 L+4\right) (1-Z)}{L^3} + \mathcal{O}(\kappa^2)
+\end{align}
+
+Santiago and Caballero
 

### 2022-10-19 manuscript/manuscript.tex
some pretty bgs stuff

@@
 
-Santiago and Caballero
+Santiago and Caballero's approximation only keeps this first term, e.g. their
+$Q_\infty^2 \approx 2/(L(1-Z))$, which is only accurate in the domain $L >
+0.2$.
+
+\begin{align}
+  N_{e,\infty}^{\text(SC98)} \approx N \exp\left(-\frac{V_h}{(1-Z)L}\right) 
+\end{align}
+
+(c.f. \cite{Santiago1998-bs} equation 8). For small $L$, the small $\kappa$
+approximation strongly deviates from the full equation; thus, throughout, we
+use the unapproximated version,
+
+\begin{align}
+  N_{e,\infty}^* \approx N \exp\left(-\frac{V_h}{(1-Z) (2-(2-L) Z)} \right).
+\end{align}
 
@@
 the equilibrium $Z$ given by equation \eqref{eq:Z} excludes the reduction due
-to drift, which reduces the variance by an additional factor $1 -
-\nicefrac{1}{2N_e}$. Thus, the reduction factor under the strong BGS model is,
-
-\begin{align}
-  Z \left(1 - \frac{1}{2N_e}\right) = (1 - s)\left(1 - \frac{1}{2N_e}\right) \approx 1 - s -\frac{1}{2N_e}
+to drift. The reduction factor under the strong BGS model is,
+
+\begin{align}
+  Z_\text{BGS} := 1 - \frac{Vm}{V_\text{BGS}} = 1 - \frac{Us^2}{Us} = 1 - s
 \end{align}
@@
 by working out the deterministic reduction in variance from the single-locus
-dynamics. Thus, $Q' = $
-
-the reduction under the classic BGS model is,
-
-\begin{align}
-  N_e = N \exp()
-\end{align}
+dynamics.
+
+Using $Z_\text{BGS}$, $V_m$, and $V$ under the background selection model, and
+approximation $N_{e,\infty}^\text{SC98}$ (equation XXX), find an equation
+reminiscent of the classic background selection with recombination results. We
+find that by using the full equation XXX $N_{e,\infty}^*$ rather than the
+approximation of Santiago and Caballero, the reduction due to background
+selection is
+
+\begin{align}
+  N_{e,\infty}^\text{BGS} = N \exp\left( - \frac{U}{2s + L - Ls} \right)
+\end{align}
+
+which matches equation 8 of \textcite{Hudson1995-xc} up to the $Ls$ term and
+equation 10 of \textcite{Nordborg1996-nq} up to the $2s$ term.
 

### 2022-10-20 manuscript/manuscript.tex
theory to main result

@@
 
-\section*{Introduction}
-
-As genomes are transmitted through the generations, lifetime events selectively
-alter what is passed to the next generation. While selection directly acts on
-the genetic variants that cause heritable fitness differences among
-individuals, neighboring regions experience an indirect contraction in genetic
-variability. This byproduct is known as \emph{linked selection}, and it occurs
-when selection changes a site's frequency on timescale faster than
-recombination can decouple that site from its genetic background. Because
-linked selection modifies the genetic variability in a regionally-correlated
-way, it creates observable patterns in genomic data that are used to understand
-the type of selection processes acting on genomes and estimate key evolutionary
-quantities.
+% \section*{Introduction}
+
+% As genomes are transmitted through the generations, lifetime events selectively
+% alter what is passed to the next generation. While selection directly acts on
+% the genetic variants that cause heritable fitness differences among
+% individuals, neighboring regions experience an indirect contraction in genetic
+% variability. This byproduct is known as \emph{linked selection}, and it occurs
+% when selection changes a site's frequency on timescale faster than
+% recombination can decouple that site from its genetic background. Because
+% linked selection modifies the genetic variability in a regionally-correlated
+% way, it creates observable patterns in genomic data that are used to understand
+% the type of selection processes acting on genomes and estimate key evolutionary
+% quantities.
 
@@
 
-Much work has been done on developing mathematical theory for two linked
-selection processes: hitchhiking and background selection. These are
-distinguished by the evolutionary processes that generate fitness variation and
-the resulting selection dynamics. Under the classic hitchhiking model, a new
-beneficial mutation sweeps fixation, leading to a dip in genetic diversity on a
-small spatial scale. Since the sweeps leave relatively narrow footprint, they
-influence genome-wide diversity only when beneficial mutations are recurring
-and recombination is relatively limited, as in \emph{Drosophila}. These models
-assume sweeps are sufficiently rare as to not occur simultaneously, which is
-consistent with the slow estimated rate of strongly beneficial mutations (XXX).
-Overall, the classic hitchhiking model is simple because the underlying
-selection process does not generate polygenic fitness variation, even though
-this is type of variation plays a dominant role in adaptation (XXX).
-
-By contrast, background selection is a consequence of the fitness variation
-created by a steady influx of new deleterious mutations. The classic background
-selection model (XXX) considers how the balance between mutation introducing
-deleterious variants into the population and selection removing them, reduces
-nearby genetic variability. A critical assumption of this model is that such an
-equilibrium between mutation and selection exists, which is the case only when
-these mutations have a large enough fitness cost to ensure that they cannot fix
-in the population. Under this process, the number of deleterious mutations per
-gamete reaches a Poisson stationary distribution with rate $\nicefrac{U}{s}$,
-where $U = \mu_D L$ is the deleterious mutation rate ($\mu_D$ per basepair per
-generation) in a region of $L$ basepairs for mutations with a heterozygous
-selection coefficient of $s$ (XXX). The fraction of gametes without any
-mutations is then $f_0 = \exp(-\nicefrac{\mu_D}{s})$. Without recombination,
-any gamete with deleterious mutations is destined to bring its linked neutral
-variation with it to extinction, effectively reducing diversity levels to those
-expected in a neutrally-evolving population of size $f_0 N$.
-
-Remarkably, predictions from simple background selection models with
-recombination closely match large-scale patterns of human diversity and
-divergence (XXX). Since background approximately rescales a neutral coalescent
-process by $f_0$, the predicted diversity is reduced from neutral levels by a
-factor $B = \nicefrac{N_e}{N}$, which is the ratio of effective population size
-under background selection to the population size. In background selection
-models with recombination, $B$ is determined by the number of conserved
-basepairs and recombination in a region. Consequently the predicted reduction
-along the genome (known as $B$ maps) mirror the regional density of conserved
-elements under purifying selection and recombination rates. These $B$ maps have
-been used in studies of XXX, are inputs into machine learning models that
-predict functional constraint, and XXX. Furthermore, since new deleterious
-mutations are perpetually flowing into populations $B$ maps may provide a
-better null model for evolution than the standard neutral model.
-
-While background selection models do predict diversity levels  XXX
-
-% outline?
-- Disentangling evolutionary processes from the patterns of genetic variation
-across the genome. The structure of what can be under selection shapes these
-patterns when selection acts; when drift acts, these patterns should not exist.
-
-- For species with small effective population sizes (and perhaps large map
-lengths), BGS seems to be a dominant selective force in shaping patterns of
-diversity (e.g. at the megabase scale in humans).
-
-- However, current approaches to fitting BGS models fail under the weak
-selection domain. This weak selection domain is important (discuss
-HRI/ratchet/etc). Importantly, the characteristic scale of the impact of linked
-selection is $\nicefrac{s}{r}$, meaning weak selection acts on smaller scales
-than strong selection. Our ability to see selection at small scales may be
-inhibited by the fact that our current BGS theory only works for strong
-selection.
-
-- We need to say briefly that we're using BGS to mean linked selection due to
-purifying selection, and thus the strong and weak regimes (classically it has
-just meant the strong regime).
-
-\section*{Conclusion}
-
-- care should be taken applying these methods to \emph{Drosophila}, since the
-$\delta_{LD}$ term could be much larger.
-
-\section*{Results}
-
-\subsection*{Theory}
-
-- overview of why BGS models fail?
-
-From an evolutionary perspective, linked selection is important because by
-modifying neighboring levels of genetic diversity, it is also directly altering
-the selection dynamics at these neighboring sites.
-
-The other use of these models is to understand how linked selection processes
-differentially alter the genetic variability, or equivalently, the genealogical
-trees, along the genome.
-
-XXX how many sites are under selection at any moment? What's fitness variation?
-
-Early on, some pointed out that evidence of strong selective sweeps in the
-human genome was not the full story
-
-Generally, in organisms with large amounts of recombination, linked selection
-has
-
-Weak selection processes
-
-distorts
-genome-wide
-
-diversity.
-
-The linked selection literature has primarily focused on developing theory for,
-and quantifying the effects of two linked selection processes: hitchhiking and
-background selection. Genome-wide models of the hitchhiking process consider
-how beneficial mutations that sporadically enter the population
-
-- QGLS polygenic linked selection
-
-- H and W
-
-- inflation factor
-
-\subsection*{Statistical Methods}
+% Much work has been done on developing mathematical theory for two linked
+% selection processes: hitchhiking and background selection. These are
+% distinguished by the evolutionary processes that generate fitness variation and
+% the resulting selection dynamics. Under the classic hitchhiking model, a new
+% beneficial mutation sweeps fixation, leading to a dip in genetic diversity on a
+% small spatial scale. Since the sweeps leave relatively narrow footprint, they
+% influence genome-wide diversity only when beneficial mutations are recurring
+% and recombination is relatively limited, as in \emph{Drosophila}. These models
+% assume sweeps are sufficiently rare as to not occur simultaneously, which is
+% consistent with the slow estimated rate of strongly beneficial mutations (XXX).
+% Overall, the classic hitchhiking model is simple because the underlying
+% selection process does not generate polygenic fitness variation, even though
+% this is type of variation plays a dominant role in adaptation (XXX).
+
+% By contrast, background selection is a consequence of the fitness variation
+% created by a steady influx of new deleterious mutations. The classic background
+% selection model (XXX) considers how the balance between mutation introducing
+% deleterious variants into the population and selection removing them, reduces
+% nearby genetic variability. A critical assumption of this model is that such an
+% equilibrium between mutation and selection exists, which is the case only when
+% these mutations have a large enough fitness cost to ensure that they cannot fix
+% in the population. Under this process, the number of deleterious mutations per
+% gamete reaches a Poisson stationary distribution with rate $\nicefrac{U}{s}$,
+% where $U = \mu_D L$ is the deleterious mutation rate ($\mu_D$ per basepair per
+% generation) in a region of $L$ basepairs for mutations with a heterozygous
+% selection coefficient of $s$ (XXX). The fraction of gametes without any
+% mutations is then $f_0 = \exp(-\nicefrac{\mu_D}{s})$. Without recombination,
+% any gamete with deleterious mutations is destined to bring its linked neutral
+% variation with it to extinction, effectively reducing diversity levels to those
+% expected in a neutrally-evolving population of size $f_0 N$.
+
+% Remarkably, predictions from simple background selection models with
+% recombination closely match large-scale patterns of human diversity and
+% divergence (XXX). Since background approximately rescales a neutral coalescent
+% process by $f_0$, the predicted diversity is reduced from neutral levels by a
+% factor $B = \nicefrac{N_e}{N}$, which is the ratio of effective population size
+% under background selection to the population size. In background selection
+% models with recombination, $B$ is determined by the number of conserved
+% basepairs and recombination in a region. Consequently the predicted reduction
+% along the genome (known as $B$ maps) mirror the regional density of conserved
+% elements under purifying selection and recombination rates. These $B$ maps have
+% been used in studies of XXX, are inputs into machine learning models that
+% predict functional constraint, and XXX. Furthermore, since new deleterious
+% mutations are perpetually flowing into populations $B$ maps may provide a
+% better null model for evolution than the standard neutral model.
+
+% While background selection models do predict diversity levels  XXX
+
+% % outline?
+% - Disentangling evolutionary processes from the patterns of genetic variation
+% across the genome. The structure of what can be under selection shapes these
+% patterns when selection acts; when drift acts, these patterns should not exist.
+
+% - For species with small effective population sizes (and perhaps large map
+% lengths), BGS seems to be a dominant selective force in shaping patterns of
+% diversity (e.g. at the megabase scale in humans).
+
+% - However, current approaches to fitting BGS models fail under the weak
+% selection domain. This weak selection domain is important (discuss
+% HRI/ratchet/etc). Importantly, the characteristic scale of the impact of linked
+% selection is $\nicefrac{s}{r}$, meaning weak selection acts on smaller scales
+% than strong selection. Our ability to see selection at small scales may be
+% inhibited by the fact that our current BGS theory only works for strong
+% selection.
+
+% - We need to say briefly that we're using BGS to mean linked selection due to
+% purifying selection, and thus the strong and weak regimes (classically it has
+% just meant the strong regime).
+
+% \section*{Conclusion}
+
+% - care should be taken applying these methods to \emph{Drosophila}, since the
+% $\delta_{LD}$ term could be much larger.
+
+% \section*{Results}
+
+% \subsection*{Theory}
+
+% - overview of why BGS models fail?
+
+% From an evolutionary perspective, linked selection is important because by
+% modifying neighboring levels of genetic diversity, it is also directly altering
+% the selection dynamics at these neighboring sites.
+
+% The other use of these models is to understand how linked selection processes
+% differentially alter the genetic variability, or equivalently, the genealogical
+% trees, along the genome.
+
+% XXX how many sites are under selection at any moment? What's fitness variation?
+
+% Early on, some pointed out that evidence of strong selective sweeps in the
+% human genome was not the full story
+
+% Generally, in organisms with large amounts of recombination, linked selection
+% has
+
+% Weak selection processes
+
+% distorts
+% genome-wide
+
+% diversity.
+
+% The linked selection literature has primarily focused on developing theory for,
+% and quantifying the effects of two linked selection processes: hitchhiking and
+% background selection. Genome-wide models of the hitchhiking process consider
+% how beneficial mutations that sporadically enter the population
+
+% - QGLS polygenic linked selection
+
+% - H and W
+
+% - inflation factor
+
+% \subsection*{Statistical Methods}
 
@@
 \parencite{Barton2000-zg}, this encodes the pairwise coalescent rate of the
-population through time under selection through the $N_{e,t}$ terms.  In the
-original derivation of \textcite{Santiago1995-hx}, they set each $N_{e,t}$ term
-to a single $N$ and solve this summation in the limit to infinity giving,
+population through time under selection through the $N_{e,t}$ terms. Others
+have found that the genealogies under purifying selection can be characterized
+by such a time-dependent effective population size
+\parencite{Nicolaisen2013-gv}. In the original derivation of
+\textcite{Santiago1995-hx}, they set each $N_{e,t}$ term to a single $N$ and
+solve this summation in the limit to infinity giving,
 
@@
 \begin{align}
-  N_{e,\infty}^* \approx N \exp\left(-\frac{V_h}{(1-Z) (2-(2-L) Z)} \right).
+  N_{e,\infty} \approx N \exp\left(-\frac{V_h}{(1-Z) (2-(2-L) Z)} \right).
 \end{align}
@@
 contribute to fitness variation. BGS models assume a multiplicative fitness
-model, such that the fitness of an individual $i$ is
-
-\begin{align}
-  w_i = (1-s)^{n_i}
-\end{align}
-
-where $n_i$ is the number of deleterious mutations this individual carries.
+model, such that the fitness of an individual is
+
+\begin{align}
+  w = (1-s)^{n}
+\end{align}
+
+where $n$ is the number of deleterious mutations this individual carries.
 
@@
 that mutation frequencies are at their mutation-selection equilibrium and the
-expected number of mutations across individuals is $\E_i(n_i) =
-\nicefrac{\mu}{s}L$ where $L$ is the total number of basepairs that can be
-mutated (the mutational target size).  Under a multiplicative fitness function
-in an infinite population, selection cannot generate linkage disequilibria
+expected number of mutations across individuals is $\E(n) = \nicefrac{\mu}{s}L$
+where $L$ is the total number of basepairs that can be mutated (the mutational
+target size).  Under a multiplicative fitness function in an infinite
+population, selection cannot generate linkage disequilibria
 \parencite{Turelli1990-kd}. Thus the classic strong background selection model
@@
 \begin{align}
-  V_\text{BGS} &:= \V_i(w_i) = 2 \sum_{l=1}^L p(1-p) s^2 \\
-            &= 2 L \frac{\mu}{s} \left( 1 - \frac{\mu}{s}\right) s^2 = U s + \mathcal{O}(\mu^2) \approx Us.
-\end{align}
-
-where $U$ is defined $U := 2 L \mu$, the deleterious mutation rate per diploid
-genome, per generation. They then set $V_A = V_\text{BGS}$, and note that $V_m
-\approx Us^2$, since mutations have frequency of $\nicefrac{1}{2N}$ and each
-increase the variance by a factor of $s^2$. Note that because the classic BGS
-model assumes deterministic selection dynamics (i.e. an infinite population),
-the equilibrium $Z$ given by equation \eqref{eq:Z} excludes the reduction due
-to drift. The reduction factor under the strong BGS model is,
+  \label{eq:va_bgs}
+  V_\text{BGS} &:= \V(w) = 2 s^2 \sum_{l=1}^L p_l(1-p_l)  \\
+            &= 2 L s^2 \frac{\mu}{s} \left( 1 - \frac{\mu}{s}\right) = U s + \mathcal{O}(\mu^2) \approx Us.
+\end{align}
+
+since under mutation-selection balance $p_l = \nicefrac{\mu}{s}$ and $U$ is
+defined $U := 2 L \mu$, the deleterious mutation rate per diploid genome, per
+generation. They then set $V_A = V_\text{BGS}$, and note that $V_m \approx
+Us^2$, since mutations have frequency of $\nicefrac{1}{2N}$ and each increase
+the variance by a factor of $s^2$. Note that because the classic BGS model
+assumes deterministic selection dynamics (i.e. an infinite population), the
+equilibrium $Z$ given by equation \eqref{eq:Z} excludes the reduction due to
+drift. The reduction factor under the strong BGS model is,
 
@@
 reminiscent of the classic background selection with recombination results. We
-find that by using the full equation XXX $N_{e,\infty}^*$ rather than the
+find that by using the full equation XXX $N_{e,\infty}$ rather than the
 approximation of Santiago and Caballero, the reduction due to background
@@
 
-which matches equation 8 of \textcite{Hudson1995-xc} up to the $Ls$ term and
-equation 10 of \textcite{Nordborg1996-nq} up to the $2s$ term.
-
-\subsection*{Weak Selection Dynamics}
+which matches equation 8 of \textcite{Hudson1995-xc} except for the $Ls$ term
+and equation 10 of \textcite{Nordborg1996-nq} except for the $2s$ term.
+
+\subsection*{Effective Population Size Under Weak and Strong BGS}
 
@@
 stationary, as the set of haplotypes without any deleterious mutations can be
-lost due to drift. The approach of \textcite{Santiago2016-mu} that we implement
-in our methods is to derive an equation for
+lost due to drift. This changes the shape of the fitness distribution, and thus
+the variance parameter that our model relies upon
+\parencite{Gessler1995-hz,OFallon2010-my,Good2013-lp,Haigh1978-gt,Higgs1995-xc}.
+So far, finding an explicit equation for the variance has been difficult due to
+the ``moment-closure problem": the variance depends on higher moments such as
+the skew, which in turn depends on higher moments. While the weak selection
+regime is composed of an interaction of mutation, selection, drift, and linkage
+processes and the dynamics that differ qualitatively based on parameter
+combinations. In their approximation, \textcite{Santiago2016-mu} consider the
+modified additive genic variance due to the fixation of weakly deleterious
+mutations. This fixation process is related to the Muller's Ratchet, the rate
+$R$ of fixation of deleterious mutations per generation per region, since each
+fixation directly reduces the additive genic variance of the fitness
+distribution. However, in general, the rate of this ratchet is unknown since it
+depends on the fitness variance (and thus higher moments of the fitness
+distribution).
+
+In their 2016 paper, Santiago and Caballero derive a general equation for the
+fitness variance that accounts for the reduction due to the fixation of weakly
+deleterious alleles (i.e. the rate of Muller's ratchet). There derivation for
+the variance uses a heuristic approach based on Fisher's Fundamental Theorem of
+Natural Selection (equation 2, \cite{Santiago2016-mu}, equation 15A
+\cite{Garcia-Dorado2007-jj}). A more formal derivation follows from
+\textcite{Higgs1995-xc} (though they assume an asexual system, the result is
+the same). In their work, they derive equations for the change per generation
+for the moments and cross-moments of the fitness distribution under
+multiplicative selection in a haploid population. For a Poisson number of new
+mutations per gamete, per generation, they find (equation 13.3,
+\cite{Higgs1995-xc}) that the rate of the ratchet is
+
+\begin{align}
+  R = U - s \V(n).
+\end{align}
+
+The rate of the ratchet for a diploid population is the same, as long as $U$ is
+the mutation rate per generation per diploid genome (as is assumed here).
+$\V(n)$ is the variance of the distribution of the \emph{number} of deleterious
+mutations (not the fitness distribution); as long as selection is not too
+strong, $V_A \approx -s^2\V(n)$. Then, the rate of the ratchet in terms of the
+fitness variance is,
+
+\begin{align}
+  %R &= U - s\V(n) \\
+  R &\approx U + \frac{V_A}{s} \\
+  V_A &\approx Us - Rs. 
+\end{align}
+
+This is identical to their equation 2, noting that the waiting time between
+fixations (i.e. ratchet clicks) is $T = \nicefrac{1}{R}$. Note that if selected
+sites are sufficiently deleterious that they cannot fix (or the population is
+infinite), $R=0$ and the additive fitness variance is identical to that under
+strong background selection, $V_A = Us = V_\text{BGS}$ (equation
+\eqref{eq:va_bgs}).
+
+Then, the fixation probability of a new deleterious mutation with additive
+effects ($1-s$ in the heterozygote, $1-2s$ in the homozygote) in a diploid
+population is,
+
+\begin{align}
+  p_\text{fix} = \frac{1-\exp(\nicefrac{2 N_e s}{N})}{1-\exp(4 N_e s)} \approx
+  \frac{2 N_e s}{N \left(\exp(4 N_e s)-1\right)}
+\end{align}
+
+(c.f. \cite{Durrett2008-ql}, equation 7.21; \cite{Kimura1957-rk}, equation
+5.6). Then, Santiago and Caballero argue that the rate of the ratchet is the
+inverse of the averaging waiting time until a fixation, $T \approx 1/(2UN
+p_\text{fix})$, so
+
+\begin{align}
+  T \approx \frac{\exp(4 N_e s)-1}{4 U s N_e}.
+\end{align}
+
+Since the ratchet rate depends on $N_e$, which in turn depends on the rate of
+the ratchet, we write the inverse as,
+
+\begin{align}
+  R(N_e) := \frac{4 U s N_e}{\exp(4 N_e s)-1}
+\end{align}
+
+Under weak and strong background selection the variance input of mutations
+$V_m$ is the same, and using the $V_A$ derived above, the decay in variance is
+
+\begin{align}
+  Z_\text{WS} = 1 - \frac{V_m}{V_A} = 1 - \frac{Us^2}{(U-R(N_e))s} = 1 - \frac{Us}{U-R(N_e)}.
+\end{align}
+
+At equilibrium, the $N_e$ that determines fixation probability is the
+asymptotic $N_{e,\infty}$. Consequently, Santiago and Caballero argue that the
+asymptotic reduction in effective population size is given by the following
+system of non-linear equations,
+
+\begin{align}
+  R(N_{e,\infty}) &= \frac{4 U s N_{e,\infty}}{\exp(4 N_{e,\infty} s)-1}\\
+  N_{e,\infty} &= N \exp\left(-\frac{(U-R(N_{e,\infty}))s}{\left(1-\frac{Us}{U-R(N_{e,\infty})}\right) \left(2-(2-L) \frac{Us}{U-R(N_{e,\infty})}\right)} \right).
+\end{align}
 
@@
 
-\subsection*{The Rate of the Ratchet}
-
-\section*{Methods}
-
-\subsection*{Data Preprocessing}
-
-\subsection*{Human}
-
-Site-level neutral and accessibility masks
-
-Bin-level masks based on fraction of accessible bases in a window (7\%).
-
-Bin-level tail filtering.
+% \subsection*{The Rate of the Ratchet}
+
+% \section*{Methods}
+
+% \subsection*{Data Preprocessing}
+
+% \subsection*{Human}
+
+% Site-level neutral and accessibility masks
+
+% Bin-level masks based on fraction of accessible bases in a window (7\%).
+
+% Bin-level tail filtering.
 

### 2022-10-21 manuscript/manuscript.tex
fixes

@@
   S_1 Q_t &= S_1 + S_2' + S_3' + \ldots + S_t' \\
-  Q_t &= 1 + \sum_{i=1}^t \left((1-r)\kappa\right)^i
-\end{align}
-
-which converges to,
+  Q_t &= 1 + \sum_{i=1}^t (1-r)^i(1-\kappa)^i
+\end{align}
+
+or for continuous $t$,
+
+\begin{align}
+  \label{eq:q_t}
+  Q_t &= 1+\frac{(1-\kappa) (1-r) \left(1-(1-\kappa)^T (1-r)^T\right)}{\kappa+(1-\kappa)r}
+\end{align}
+
+which as $T \to \infty$, converges to,
 
@@
   \label{eq:Qinf}
-  Q_\infty &= \frac{1}{1-(1-r) \kappa}.
+  Q_\infty &= \frac{1}{\kappa + r(1-\kappa)}.
 \end{align}
@@
 
-where $L = n r_\text{BP}$ is the total segment length in Morgans. Let us define
-the total asymptotic effect of associations from all sites as $Q_\infty^2$.
-Then,
-
-\begin{align}
-  Q_\infty^2 &= \frac{2}{L}\int_0^{\nicefrac{n r_\text{BP}}{2}} Q_{\infty}(x)^2 dx  \\
-  &=  \frac{2}{L} \int_0^{\nicefrac{L}{2}} \left(\frac{1}{1 - (1-x)Z}\right)^2 dx  \\
+where $L = n r_\text{BP}$ is the total segment length in Morgans. Then, the
+total asymptotic effect of associations from all sites as $Q_\infty^2$.  Then,
+
+\begin{align}
+  Q_\infty^2 &= \frac{2}{L}\int_0^{\nicefrac{L}{2}} Q_{\infty}(r)^2 dr  \\
+  &=  \frac{2}{L} \int_0^{\nicefrac{L}{2}} \left(\frac{1}{1 - (1-r)Z}\right)^2 dr  \\
   &= \frac{2}{(1-Z) (2-(2-L) Z)}
@@
   N_{e,\infty} \approx N \exp\left(-\frac{V_h}{(1-Z) (2-(2-L) Z)} \right).
+\end{align}
+
+For the reasons described earlier, a single asymptotic $N_{e,\infty}$ cannot
+fully characterize pairwise coalescent rates in all cases, and thus neutral
+diversity levels depend on the full $N_{e,t}$, for which we integrate equation
+\eqref{eq:q_t} to get,
+
+\begin{align}
+  \label{eq:net_integral}
+  N_{e,t} \approx N \exp\left(-\frac{V_h}{L} \int_0^{\nicefrac{L}{2}} \left(1+\frac{(1-k) (1-r) \left(1-(1-k)^t (1-r)^t\right)}{k+(1-k)r} \right) dr \right).
 \end{align}
@@
 \end{align}
+
+An analogous system can be found for $N_{e,t}$ using equation
+\eqref{eq:net_integral}.
 

### 2022-10-26 manuscript/manuscript.tex
fixes

@@
 % \subsection*{Statistical Methods}
+
+TODO we need to make appoint about how Ben G's paper found the V(n)->VA connectoin -- this is their patching model
 
@@
 
+% NOTE: The S&C math using Fisher's FToNS argues fitness decreases
+% at a rate of -s/T each generation. However their model is haploid;
+% in a diploid model, fixation makes everyone  a homozygote with fitness
+% effect -2s; thus the rate must be -2s/T which agrees with sim
+
 In their 2016 paper, Santiago and Caballero derive a general equation for the
@@
 \cite{Garcia-Dorado2007-jj}). A more formal derivation follows from
-\textcite{Higgs1995-xc} (though they assume an asexual system, the result is
-the same). In their work, they derive equations for the change per generation
-for the moments and cross-moments of the fitness distribution under
-multiplicative selection in a haploid population. For a Poisson number of new
-mutations per gamete, per generation, they find (equation 13.3,
-\cite{Higgs1995-xc}) that the rate of the ratchet is
-
-\begin{align}
-  R = U - s \V(n).
-\end{align}
-
-The rate of the ratchet for a diploid population is the same, as long as $U$ is
-the mutation rate per generation per diploid genome (as is assumed here).
-$\V(n)$ is the variance of the distribution of the \emph{number} of deleterious
-mutations (not the fitness distribution); as long as selection is not too
-strong, $V_A \approx -s^2\V(n)$. Then, the rate of the ratchet in terms of the
-fitness variance is,
-
-\begin{align}
-  %R &= U - s\V(n) \\
-  R &\approx U + \frac{V_A}{s} \\
-  V_A &\approx Us - Rs. 
-\end{align}
-
-This is identical to their equation 2, noting that the waiting time between
-fixations (i.e. ratchet clicks) is $T = \nicefrac{1}{R}$. Note that if selected
-sites are sufficiently deleterious that they cannot fix (or the population is
-infinite), $R=0$ and the additive fitness variance is identical to that under
-strong background selection, $V_A = Us = V_\text{BGS}$ (equation
-\eqref{eq:va_bgs}).
-
-Then, the fixation probability of a new deleterious mutation with additive
-effects ($1-s$ in the heterozygote, $1-2s$ in the homozygote) in a diploid
-population is,
+\textcite{Higgs1995-xc} (though they assume a haploid asexual system, the end
+result is the same up to a factor of two). In their work, they derive equations
+for the change per generation for the moments and cross-moments of the fitness
+distribution under multiplicative selection in a haploid model. They find, the
+haploid rate of the ratchet is $r = u - s v_n$ (equation 13.3,
+\cite{Higgs1995-xc}), where we use lowercase to distinguish between haploid and
+diploid models. Here, $v_n$ is the variance in the \emph{number} of deleterious
+mutations; as long as selection is not too strong, the haploid additive
+variance is $v_{A} \approx s^2 v_n$. Substituting this for $v_n$ and
+rearranging, the haploid fitness variance is related to the $v_A = us - rs$,
+which is identical to Santiago and Caballero's (\citeyear{Santiago2016-mu})
+equation 2, noting that the waiting time between fixations (i.e. ratchet
+clicks) is $T = \nicefrac{1}{r}$. Intuitively, the fitness variance $Us$ is
+reduced by the rate at which deleterious mutations fix and remove fitness
+variation from the population.
+
+Now, note that under a diploid model, the fitness variance is twice that as the
+haploid model, so $V_A = 2 v_A$. Then, the fitness variance is,
+
+\begin{align}
+  V_{A} = Us - 2Rs. 
+\end{align}
+
+where $U = 2u$ and $R$ is the diploid rate of the ratchet, which depends on
+higher moments of the fitness distribution, but only differs from $r$ through a
+rescaling of the population size. Note that if selected sites are sufficiently
+deleterious that they cannot fix (or the population is infinite), $R=0$ and the
+additive fitness variance is identical to that under strong background
+selection, $V_A = Us = V_\text{BGS}$ (equation \eqref{eq:va_bgs}). Then, the
+fixation probability of a new deleterious mutation with additive effects ($1-s$
+in the heterozygote, $1-2s$ in the homozygote) in a diploid population is,
 
@@
 5.6). Then, Santiago and Caballero argue that the rate of the ratchet is the
-inverse of the averaging waiting time until a fixation, $T \approx 1/(2UN
-p_\text{fix})$, so
-
-\begin{align}
-  T \approx \frac{\exp(4 N_e s)-1}{4 U s N_e}.
+inverse of the averaging waiting time until a fixation, $T \approx 1/(UN
+p_\text{fix})$ (note that since $U$ is the \emph{diploid} mutation rate, the
+total population mutation rate is $UN$ per generation), so
+
+\begin{align}
+  T \approx \frac{\exp(4 N_e s)-1}{2 U s N_e}.
 \end{align}
@@
 \begin{align}
-  R(N_e) := \frac{4 U s N_e}{\exp(4 N_e s)-1}
+  R(N_e) := \frac{2 U s N_e}{\exp(4 N_e s)-1}
 \end{align}
@@
   R(N_{e,\infty}) &= \frac{4 U s N_{e,\infty}}{\exp(4 N_{e,\infty} s)-1}\\
+  N_{e,\infty} &\approx N \exp\left(-\frac{V_A}{(1-Z) (2-(2-L) Z)} \right) \\
   N_{e,\infty} &= N \exp\left(-\frac{(U-R(N_{e,\infty}))s}{\left(1-\frac{Us}{U-R(N_{e,\infty})}\right) \left(2-(2-L) \frac{Us}{U-R(N_{e,\infty})}\right)} \right).

### 2022-12-07 manuscript/manuscript.tex
supp draft

@@
 
+% line numbers:
+\RequirePackage{lineno}
+%\modulolinenumbers[5]
+\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{black}}
+
 \renewcommand{\P}{\mathbb{P}}
@@
 
-% \section*{Introduction}
+\section*{Introduction}
+\section*{Results}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics{figures/experimental_1.pdf}
+  \caption{}
+  \label{fig:figure-1}
+\end{figure}
 
@@
 
+TODO we need to give some intuition for what Vm and V are in humans, e.g. back of envelope type stuff
+
 \section*{Appendix}
@@
   \label{eq:q_t}
-  Q_t &= 1+\frac{(1-\kappa) (1-r) \left(1-(1-\kappa)^T (1-r)^T\right)}{\kappa+(1-\kappa)r}
-\end{align}
-
-which as $T \to \infty$, converges to,
+  Q_t &= 1+\frac{(1-\kappa) (1-r) \left(1-(1-\kappa)^t (1-r)^t\right)}{\kappa+(1-\kappa)r}
+\end{align}
+
+which as $t \to \infty$, converges to,
 
@@
 \begin{align}
-      N_{e,\infty} &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{\infty,l}^2\right) \\
-              &\approx N \exp\left(- \frac{V_{h}}{L} \int_0^{\nicefrac{L}{2}} Q_{\infty}(r)^2 dr \right) \\
-\end{align}
-
-where $L = n r_\text{BP}$ is the total segment length in Morgans. Then, the
+      N_{e,\infty} &\approx N \exp\left(- \frac{V_{h}}{2L} \sum_{l=1}^L Q_{\infty,l}^2\right) \\
+              &\approx N \exp\left(- \frac{V_{h}}{M} \int_0^{\nicefrac{M}{2}} Q_{\infty}(r)^2 dr \right) \\
+\end{align}
+
+where $M = n r_\text{BP}$ is the total segment length in Morgans. Then, the
 total asymptotic effect of associations from all sites as $Q_\infty^2$.  Then,
@@
 \begin{align}
-  Q_\infty^2 &= \frac{2}{L}\int_0^{\nicefrac{L}{2}} Q_{\infty}(r)^2 dr  \\
-  &=  \frac{2}{L} \int_0^{\nicefrac{L}{2}} \left(\frac{1}{1 - (1-r)Z}\right)^2 dr  \\
-  &= \frac{2}{(1-Z) (2-(2-L) Z)}
+  Q_\infty^2 &= \frac{2}{M}\int_0^{\nicefrac{M}{2}} Q_{\infty}(r)^2 dr  \\
+  &=  \frac{2}{M} \int_0^{\nicefrac{M}{2}} \left(\frac{1}{1 - (1-r)Z}\right)^2 dr  \\
+  &= \frac{2}{(1-Z) (2-(2-M) Z)}
 \end{align}
@@
 \begin{align}
-  Q_\infty^2 = \frac{2}{L (1-Z)} + \frac{2 (L-2)}{L^2}+\frac{2 \left(L^2-4 L+4\right) (1-Z)}{L^3} + \mathcal{O}(\kappa^2)
+  Q_\infty^2 = \frac{2}{M (1-Z)} + \frac{2 (M-2)}{M^2}+\frac{2 \left(M^2-4 M+4\right) (1-Z)}{M^3} + \mathcal{O}(\kappa^2)
 \end{align}
@@
 Santiago and Caballero's approximation only keeps this first term, e.g. their
-$Q_\infty^2 \approx 2/(L(1-Z))$, which is only accurate in the domain $L >
+$Q_\infty^2 \approx 2/(M(1-Z))$, which is only accurate in the domain $M >
 0.2$.
@@
 \begin{align}
-  N_{e,\infty}^{\text(SC98)} \approx N \exp\left(-\frac{V_h}{(1-Z)L}\right) 
-\end{align}
-
-(c.f. \cite{Santiago1998-bs} equation 8). For small $L$, the small $\kappa$
+  N_{e,\infty}^{\text(SC98)} \approx N \exp\left(-\frac{V_h}{(1-Z)M}\right) 
+\end{align}
+
+(c.f. \cite{Santiago1998-bs} equation 8). For small $M$, the small $\kappa$
 approximation strongly deviates from the full equation; thus, throughout, we
@@
 \begin{align}
-  N_{e,\infty} \approx N \exp\left(-\frac{V_h}{(1-Z) (2-(2-L) Z)} \right).
+  N_{e,\infty} \approx N \exp\left(-\frac{V_h}{(1-Z) (2-(2-M) Z)} \right).
 \end{align}
@@
   \label{eq:net_integral}
-  N_{e,t} \approx N \exp\left(-\frac{V_h}{L} \int_0^{\nicefrac{L}{2}} \left(1+\frac{(1-k) (1-r) \left(1-(1-k)^t (1-r)^t\right)}{k+(1-k)r} \right) dr \right).
+  N_{e,t} \approx N \exp\left(-\frac{V_h}{M} \int_0^{\nicefrac{M}{2}} \left(1+\frac{(1-k) (1-r) \left(1-(1-k)^t (1-r)^t\right)}{k+(1-k)r} \right) dr \right).
 \end{align}
@@
 \begin{align}
-  N_{e,\infty}^\text{BGS} = N \exp\left( - \frac{U}{2s + L - Ls} \right)
+  N_{e,\infty}^\text{BGS} = N \exp\left( - \frac{U}{2s + M - Ms} \right)
 \end{align}
@@
 asymptotic $N_{e,\infty}$. Consequently, Santiago and Caballero argue that the
-asymptotic reduction in effective population size is given by the following
-system of non-linear equations,
+asymptotic reduction in effective population size uses the $Q_\infty^2$ and is
+given by the following system of non-linear equations,
 
@@
   R(N_{e,\infty}) &= \frac{4 U s N_{e,\infty}}{\exp(4 N_{e,\infty} s)-1}\\
-  N_{e,\infty} &\approx N \exp\left(-\frac{V_A}{(1-Z) (2-(2-L) Z)} \right) \\
-  N_{e,\infty} &= N \exp\left(-\frac{(U-R(N_{e,\infty}))s}{\left(1-\frac{Us}{U-R(N_{e,\infty})}\right) \left(2-(2-L) \frac{Us}{U-R(N_{e,\infty})}\right)} \right).
-\end{align}
-
-An analogous system can be found for $N_{e,t}$ using equation
-\eqref{eq:net_integral}.
+  N_{e,\infty} &\approx N \exp\left(-\frac{V_A}{(1-Z) (2-(2-M) Z)} \right)
+\end{align}
+
+The solution to these equations is the asymptotic equilibrium $\widetilde{N}_e$ and
+equilibrium ratchet rate $\widetilde{R}$. This set the equilibrium variance,
+$\widetilde{V} = Us - 2\widetilde{R}s$.
+
+\subsection*{Heterozygosity and $B$ values Under Weak and Strong Selection}
+
+While the selection dynamics are set by the asymptotic $Q_\infty^2$ that
+determines the equilibrium $\widetilde{N}_e$ and $\widetilde{R}$,
+heterozygosity is determined by the sequence $\widetilde{N}_{e,t}$, where each
+term is
+
+XXX
+
+\subsection*{Numeric methods for calculating $B$}
+
+Our $B$ is,
+
+\begin{align}
+  B = \frac{1}{2N} \sum_{t=0}^\infty \prod_{i=0}^t \left(1 - \frac{1}{2N_e(i)}\right).
+\end{align}
+
+This sum takes many terms to converge, which makes calculating the $B$ between
+each focal neutral site and segment intractably slow (even when the equilibrium
+$\widetilde{V}$ for the segment pre-calculated). We approximate this in the
+following way for faster numerical calculations. Let $g(t) = \prod_{i=0}^t (1 -
+\nicefrac{1}{2N_e(i)})$. Then, taking the first-order series expansion around
+the logarithm of the terms and approximating the sum with an integral,
+
+\begin{align}
+  g(t) &\approx \exp\left(- \frac{1}{2N} \int_0^t \exp\left(\frac{V}{2} Q(x)^2 \right) dx \right) 
+       %&\approx \exp\left(- \int_0^t \frac{\exp\left(\frac{V}{2} Q(i)^2\right)}{2N}  dx \right)
+\end{align}
+
+$B$ is the case of $\lim b \to \infty$ for $B(b)$. Approximating the sum in
+$B(b)$ with an integral,
+
+\begin{align}
+  B(b) &= \int_0^b \exp\left(- \frac{1}{2N} \int_0^t \exp\left(\frac{V}{2} Q(x)^2 \right) dx \right) dt \\
+       %&\approx \exp\left(- \int_0^t \frac{\exp\left(\frac{V}{2} Q(i)^2\right)}{2N}  dx \right)
+\end{align}
+
+Now, let $z = \log(t)$, then
+
+\begin{align}
+  B(w) &= \int_0^w \exp\left(- \frac{1}{2N} \int_0^{\log(z)} \exp\left(\frac{V}{2} Q(x)^2 \right) dx \right) \exp(t) dz \\
+       %&\approx \exp\left(- \int_0^t \frac{\exp\left(\frac{V}{2} Q(i)^2\right)}{2N}  dx \right)
+\end{align}
 

### 2022-12-07 manuscript/supplementary.tex
supp draft

\documentclass[11pt]{article}
\RequirePackage{fullpage}
%\RequirePackage[font=small,labelfont=bf]{caption}
\RequirePackage{amsmath,amssymb,amsthm}
\RequirePackage{graphicx}
\RequirePackage[hidelinks]{hyperref}
\RequirePackage{subcaption}
\RequirePackage{wasysym}
\RequirePackage{authblk}
\RequirePackage{bm}
\RequirePackage{bbm}
%\RequirePackage[osf]{mathpazo}
\let\temp\rmdefault
\RequirePackage{mathpazo}
\let\rmdefault\temp

\RequirePackage[bibstyle=authoryear,citestyle=authoryear-comp,
                date=year,
                maxbibnames=9,maxnames=5,maxcitenames=2,
                backend=biber,uniquelist=false,uniquename=false,
                % style=apa,
                sorting=nyt,
                % sorting=,
                hyperref=true]{biblatex}
\usepackage{color}
\usepackage{nicefrac}

% line numbers:
\RequirePackage{lineno}
%\modulolinenumbers[5]
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{black}}

\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\text{V}}
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\cov}{cov}

\addbibresource{biblio.bib}

\title{Supplementary Materials}

\author{Vince Buffalo and Andrew Kern}

\begin{document}
\maketitle

\section{Human Genomic Data}

\subsection{YRI Samples}

In order to try to ensure accurate estimation of pairwise diversity, which is a
ratio estimator that is sensitivity to its denominator, we use the complete
per-basepair genotype calls (gVCF) produced by Illumina's DRAGEN pipeline
\parencite{Illumina_Inc2020-dk}. The original samples were from 178 Yoruban
individuals sequenced to 30x by the New York Genome Center
\parencite{Byrska-Bishop2022-tn}. This allows for filtering to be applied to
the entire genome at once, rather than just variants, so the denominator does
not need to be estimated separately.

The full list of samples is available in TSV format in the GitHub repository
(\path{data/h1kg/yri_samples.tsv}).

\subsection{Filtering gVCFs}

gVCFs were filtered using a custom Python tool, \texttt{gvcf2counts.py} (in
\path{tools/gvcf2counts.py}), which reads the gVCFs, filters them according to
the criteria below, and outputs a Numpy \path{.npy} file of reference and
alternative allele counts for each chromosome (hereafter, ``allele counts
matrices").

Genotypes are included in the allele count if and only if:

\begin{enumerate}
  \item The variant call is set to \texttt{PASS} in the VCF.
  \item The \texttt{QUAL > 50}.
  \item The \texttt{GQ > 30} (or \texttt{RGQ > 30} for invariant sites).
\end{enumerate}

Because the data underlying the counts files are per-basepair resolution gVCFs,
each chromosome's allele counts matrix is of size $l \times 2$, where $l$ is
the total chromosome length. Basepairs that fail these filtering requirements
lead to a row of zero counts, e.g. no observed reference and alternative allele
counts, and thus do not effect the data that goes into the binomial likelihood
or $\pi$ estimates used in figures.

\subsection{Site-based Filtering of Counts}

The allele counts matrices include many basepairs that may have allele counts
that pass the genotype call filters, but are still need to be filtered out
because the region of the genome may produce unreliable estimates. The
following filters are applied based on masking regions:

\begin{enumerate}
  \item \textbf{Non-accessible regions}: masks out centromeres (\texttt{acen} entries in
    \texttt{cytoBand.txt}), with 5Mbp padding on either side. The file of
    passing masks is \path{data/annotation/no_centro.bed}.

  \item \textbf{Reference masking}: soft and hard-masked regions in the human GRCh38
    reference genome are also masked.

  \item \textbf{Non-``putatively" neutral regions}: Additionally, for fitting our
    likelihood and estimating observed pairwise diversity, we only consider .
    This masks out phastCons regions (from
    \path{phastConsElements100way.bed.gz}) and Ensembl gene regions (from
    annotation file \path{Homo_sapiens.GRCh38.107.chr.gff3.gz}).  While introns
    are possibly under some weak selection, they collectively make up nearly
    40\% of the human genome and are included so genome-wide diversity can be
    estimated more precisely (possibly at the expense of some bias).
\end{enumerate}

These files are all produced by the Snakemake file \path{data/annotation/Snakefile}.

\subsection{Estimates of Pairwise Diversity}

\subsection{Window-based Summaries and filtering}

The likelihood method is fit to non-overlapping binned summaries of the allele
counts matrix.

Based on exploratory data analyses, some bins were outliers and
excluded (XXX).

\begin{enumerate}
  \item \textbf{Fraction of inaccessible sites per window}.  \path{mask_inaccessible_bins_frac}
  % \item \textbf{Edge truncation}

  \item \textbf{Outliers}: Based on exploratory analysis, there were some
    regions with very high diversity. The 0.05\% right tail was excluded.

\end{enumerate}

% supp data
% - npy files
% - edge trunation

\section{}

\printbibliography

\end{document}

### 2022-12-07 bgspy/data.py
hunting down pi issues

@@ -187,6 +187,9 @@ class CountsDirectory:
             self.counts = {c: np.load(f) for c, f in self.npy_files.items()}
             self.lazy = False
 
+    def keys(self):
+        return self.npy_files.keys()
+
     def __getitem__(self, key):
         if self.lazy:
             return np.load(self.npy_files[key])
@@ -194,6 +197,11 @@ class CountsDirectory:
 
 def filter_sites(counts, chrom, filter_neutral, filter_accessible,
                  neutral_masks=None, accesssible_masks=None):
+    """
+    We zero out the allele counts for masked sites. Note that this is
+    allowable (versus marking with NaNs) because we work with the allele
+    count matrix; it's akin to not observing any genotype information at a site.
+    """
     ac = counts[chrom]
     if filter_neutral or neutral_masks is not None:
         msg = "GenomeData.neutral_masks not set!"
@@ -421,6 +429,10 @@ class GenomeData:
         mask_inaccessible_bins_frac, that bin is masked.
 
         Returns: a GenomicBins object with the resulting data in GenomicBins.data.
+
+        NOTE: There's a slight difference here between the π
+        from GenomeData.pi()  across chromosomes, and a weighted average of
+        binned π calculated for a chromosome.
         """
         bins = GenomicBinnedData(self.genome.seqlens, width)
 
@@ -433,6 +445,9 @@ class GenomeData:
             # the combinatoric step -- turn site allele counts into
             # same/diff comparisons
             Y = pairwise_summary(site_ac)
+            # NOTE: this should match pi_chroms[0]['chr1']
+            # np.nanmean(pi_from_pairwise_summaries(Y))
+            #
             # number of combinations summed into bins per chrom
             Y_binstat = aggregate_site_array(Y, bins[chrom], np.sum, axis=0)
             bins.data_[chrom] = Y_binstat.stat
@@ -447,21 +462,25 @@ class GenomeData:
         for chrom in self.genome.chroms:
             site_ac = filter_sites(self.counts, chrom,
                                    filter_neutral, filter_accessible,
-                                   self.neut_masks, self.accesssible_masks)
+                                   self.neutral_masks, self.accesssible_masks)
             out[chrom] = (site_ac > 0).sum(axis=1).sum()
         return out
 
-    def pi(self):
+    def pi(self, filter_neutral=None, filter_accessible=None):
         pi = dict()
         n = dict()
[…]

### 2022-12-07 bgspy/utils.py

@@ -412,6 +412,8 @@ def aggregate_site_array(x, bins, func, **kwargs):
         # the results length is len(bins)-1
         vals[i-1, ...] = func(data_in_bin, **kwargs)
         #assert not np.any(np.isnan(data_in_bin)) # otherwise this changes n
+        # NOTE: "missingness" here are allele counts that are
+        # equal to zero. This is the total number of complete cases
         n[i-1] = np.sum(np.sum(data_in_bin, axis=1) > 0)
     return BinnedStat(vals, bins, n)
 

### 2022-12-07 manuscript/supplementary.tex
table

@@
 
+\begin{table}
+\centering
+\begin{tabular}{lrrr}
+  \hline
+chrom &  accessible &  neutral &  both \\
+  \hline
+ chr1 &        42.4 &     64.4 &  22.9 \\
+ chr2 &        46.8 &     58.7 &  23.7 \\
+ chr3 &        44.1 &     62.9 &  24.8 \\
+ chr4 &        44.0 &     59.3 &  21.5 \\
+ chr5 &        44.1 &     58.0 &  21.4 \\
+ chr6 &        45.0 &     58.0 &  22.1 \\
+ chr7 &        44.4 &     65.8 &  26.4 \\
+ chr8 &        43.8 &     61.2 &  23.1 \\
+ chr9 &        39.2 &     64.5 &  20.3 \\
+chr10 &        44.6 &     62.8 &  25.3 \\
+chr11 &        42.7 &     63.5 &  23.4 \\
+chr12 &        41.7 &     62.7 &  22.7 \\
+chr13 &        39.7 &     62.0 &  18.3 \\
+chr14 &        37.6 &     65.4 &  19.0 \\
+chr15 &        37.0 &     69.4 &  20.3 \\
+chr16 &        39.4 &     63.8 &  20.9 \\
+chr17 &        40.4 &     64.0 &  22.3 \\
+chr18 &        40.9 &     59.8 &  19.8 \\
+chr19 &        30.9 &     69.8 &  18.0 \\
+chr20 &        36.9 &     61.6 &  19.8 \\
+chr21 &        31.6 &     68.6 &  18.2 \\
+chr22 &        26.9 &     75.0 &  16.4 \\
+\end{tabular}
+\end{table}
+
 \subsection{Estimates of Pairwise Diversity}

### 2022-12-07 manuscript/update_bib.sh

@@ -1,0 +1,1 @@
+curl https://paperpile.com/eb/jKrtuTXRHq > biblio.bib

### 2022-12-08 bgspy/data.py
diversity stuff fixed

@@ -151,13 +151,13 @@ def pi_from_pairwise_summaries(x):
 
 def trimmed_pi_bounds(Y, alpha):
     """
+    Get the quantile cutoffs for trimming off outliers by alpha (lower, upper).
     """
     if isinstance(alpha, float):
         lower, upper = alpha/2, 1-alpha/2
     else:
         lower, upper = alpha
     pi = pi_from_pairwise_summaries(Y)
-    m
     q1, q2 = np.nanquantile(pi, lower), np.nanquantile(pi, upper)
     return q1, q2
 
@@ -201,19 +201,26 @@ def filter_sites(counts, chrom, filter_neutral, filter_accessible,
     We zero out the allele counts for masked sites. Note that this is
     allowable (versus marking with NaNs) because we work with the allele
     count matrix; it's akin to not observing any genotype information at a site.
+
+    Note: the masking tracks have True values for keep and False values
+    for discard.
     """
     ac = counts[chrom]
-    if filter_neutral or neutral_masks is not None:
-        msg = "GenomeData.neutral_masks not set!"
-        assert neutral_masks is not None, msg
-        #print("using neutral masks...")
-        ac = ac * neutral_masks[chrom][:, None]
+    if neutral_masks is not None:
+        # if filters are set, use them unless explicitly told not to
+        if filter_neutral is not False:
+            msg = "GenomeData.neutral_masks not set!"
+            assert neutral_masks is not None, msg
+            #print("using neutral masks...")
+            ac = ac * neutral_masks[chrom][:, None]
 
-    if filter_accessible or accesssible_masks is not None:
-        msg = "GenomeData.accesssible_masks not set!"
-        assert accesssible_masks is not None, msg
-        # print("using accessibility masks...")
-        ac = ac * accesssible_masks[chrom][:, None]
+    if accesssible_masks is not None:
+        # if filters are set, use them unless explicitly told not to
+        if filter_accessible is not False:
+            msg = "GenomeData.accesssible_masks not set!"
+            assert accesssible_masks is not None, msg
+            # print("using accessibility masks...")
+            ac = ac * accesssible_masks[chrom][:, None]
     return ac
 
 def trim_map_ends(recmap, thresh_cM):
@@ -466,16 +473,36 @@ class GenomeData:
             out[chrom] = (site_ac > 0).sum(axis=1).sum()
         return out
 
[…]

### 2022-12-08 manuscript/supplementary.tex
new supp stuff

@@
 
+Our underlying data for all likelihood and pairwise diversity estimates is the
+allele count matrix $\mathbf{C}$ with dimensions $l \times 2$, where $l$ is the
+chromosome length. This is converted to a pairwise summary matrix with
+identical dimensions, $\mathbf{P}$. The first column of $\mathbf{P}$ is the
+number of pairwise comparisons between chromosomes that are identical, and the
+second column is the number that are different. Both of these columns are
+combinatoric summaries of the allele counts matrix. Let $[c_1, c_2]$ be a row
+of $\mathbf{C}$ for basepair $i$, and $n = c_1 + c_2$. Then, the (1) total
+number of pairwise combinations of chromosomes $n_T$, (2) the number of
+pairwise with identical alleles $n_S$, and (3) the number of pairwise
+combinations with differing alleles $n_D$ are respectively,
+
+\begin{align*}
+  n_T &= \frac{n(n-1)}{2} \\
+  n_S &= {c_1 \choose 2} + {c_2 \choose 2} \\
+  n_D &= n_T - n_S
+\end{align*}
+
+Thus $\mathbf{P}_i = [n_S, n_D]$, and the pairwise diversity is calculated as
+
+\begin{align}
+  \bar{\pi}(i) = \frac{n_D}{n_T}
+\end{align}
+
+which is identical to the more common expression of this estimator,
+
+\begin{align}
+  \bar{\pi}(i) = \frac{2}{n(n-1)} \sum_{i < j}^n k_{i,j}
+\end{align}
+
+where $k_{i,j}$ is 1 if the alleles at this site differ, and 0 otherwise. Note
+that for sites with missing data, $c_1 = c_2 = 0$ and thus $n_T = 0$ and the
+division is invalid. Such cases were marked as missing with floating point
+values \texttt{NaN}s. Binned summaries of data are weighed by the number of
+complete cases, e.g. rows with $c_1 + c_2 > 0$.
+
 \subsection{Window-based Summaries and filtering}

### 2022-12-08 manuscript/supplementary.tex
more notes

@@
 Our underlying data for all likelihood and pairwise diversity estimates is the
-allele count matrix $\mathbf{C}$ with dimensions $l \times 2$, where $l$ is the
+allele count matrix $\mathbf{C}$ with dimensions $L \times 2$, where $L$ is the
 chromosome length. This is converted to a pairwise summary matrix with
@@
 combinatoric summaries of the allele counts matrix. Let $[c_1, c_2]$ be a row
-of $\mathbf{C}$ for basepair $i$, and $n = c_1 + c_2$. Then, the (1) total
+of $\mathbf{C}$ for basepair $l$, and $n = c_1 + c_2$. Then, the (1) total
 number of pairwise combinations of chromosomes $n_T$, (2) the number of
@@
 
-Thus $\mathbf{P}_i = [n_S, n_D]$, and the pairwise diversity is calculated as
+Thus $\mathbf{P}_l = [n_S, n_D]$, and the pairwise diversity is calculated as
 
 \begin{align}
-  \bar{\pi}(i) = \frac{n_D}{n_T}
+  \bar{\pi}(l) = \frac{n_D}{n_T}
 \end{align}
@@
 \begin{align}
-  \bar{\pi}(i) = \frac{2}{n(n-1)} \sum_{i < j}^n k_{i,j}
+  \bar{\pi}(l) = \frac{2}{n(n-1)} \sum_{i < j}^n k_{i,j}
 \end{align}
@@
 complete cases, e.g. rows with $c_1 + c_2 > 0$.
+
+This per-basepair diversity is then averaged over all callable bases, to give a
+genome-wide or window estimate of diversity of accessible bases in the set
+$\mathcal{A}$,
+
+\begin{align*}
+  \bar{\pi}(\mathcal{A}) = \frac{\sum_{l \in \mathcal{A}} \bar{\pi}(l)}{L_\mathcal{A}}.
+\end{align*}
+
+where $L_\mathcal{A} = |\mathcal{A}|$ is the number of accessible bases. Note
+that both $L_\mathcal{A}$ and the sum are random variables in genomic data.
+The denominator varies along the genome, based on the number of
+reliably-genotyped and accessible bases in the reference genome. This is a
+ratio estimator, and is thus biased, since by Jensen's inequality
+$\E(\nicefrac{y}{x}) \ge \nicefrac{\E(y)}{\E(x)}$.
+
+We explore this with block-bootstrap bias correction. TODO
 

### 2022-12-08 bgspy/bootstrap.py
more supp figs

@@ -81,10 +81,13 @@ def percentile_ci(boot_thetas, theta, alpha=0.05, log=True):
     eps = np.quantile(boot_thetas, (alpha/2, 1-alpha/2), axis=1)
     return eps
 
-def pivot_ci(boot_thetas, theta, alpha=0.05, log=True):
+def pivot_ci(boot_thetas, theta, alpha=0.05, log=True, axis=1):
+    """
+    Default axis is 1 for MLE thetas.
+    """
     if log:
         boot_thetas = np.log10(boot_thetas)
         theta = np.log10(theta)
-    eps = np.quantile(boot_thetas, (alpha/2, 1-alpha/2), axis=1)
+    eps = np.quantile(boot_thetas, (alpha/2, 1-alpha/2), axis=axis)
     return 10**(2*theta - eps[1]), 10**(2*theta - eps[0])
 

### 2022-12-08 bgspy/data.py

@@ -138,11 +138,21 @@ def pi_from_pairwise_summaries(x):
         pair_data = x.stat
     else:
         pair_data = x
+
+    # if we have a single row of data, e.g. for genome-wide,
+    # let's add another dimension so stuff works
+    if pair_data.shape == (2, ):
+        pair_data = pair_data[None, :]
+
+    # total number of pairwise comparisons is denom
     denom = pair_data.sum(axis=1)
     out = np.divide(pair_data[:, 1], denom,
                     out=np.full(denom.shape[0], np.nan),
                     where=denom > 0)
     if not is_binstat:
+        if len(out) == 1:
+            # don't return an array if we don't have to
+            return out[0]
         return out
     n = x.n
     n[denom == 0] = 0
@@ -492,7 +502,7 @@ class GenomeData:
             # pi[chrom] = np.nanmean(pi_from_pairwise_summaries(pairwise_summary(ac)))
             # RIGHT WAY (slow)
             nn = ac.sum(axis=1)
-            P = pairwise_summary(ac).sum(axis=0)[None, :]
+            P = pairwise_summary(ac).sum(axis=0)
             pi[chrom] = pi_from_pairwise_summaries(P)
             n[chrom] = np.sum(nn > 0)
             # RIGHT WAY (alternate)
@@ -777,6 +787,8 @@ class GenomicBinnedData(GenomicBins):
                         exclude_chrom=None, filter_masked=True):
         """
         Generate resampling indices of blocksize consecutive bins.
+
+        blocksize: is number of consecutive windows.
         """
         bins = self.bins(filter_masked=filter_masked)
         exclude_chrom = [exclude_chrom] if exclude_chrom is not None else None
@@ -797,6 +809,25 @@ class GenomicBinnedData(GenomicBins):
                 idx.append(i)
         return np.array(idx)
 
+    def bootstrap_pi(self, B, nblocks, filter_masked=True):
+        """
+        """
+        Y = self.Y(filter_masked=filter_masked)
+        straps = []
+        for b in np.arange(B):
+            Y_boot = Y[self.resample_blocks(nblocks), :].sum(axis=0)
+            straps.append(pi_from_pairwise_summaries(Y_boot))
+        straps = np.array(straps)
+        return straps
+
+    def estimate_pi_bias(self, B, nblocks, filter_masked=True):
+        """
+        Estimate bias(π) from block bootstrapping with nblocks.
+        """
[…]

### 2022-12-08 manuscript/supplementary.tex
ms changes

@@
 
-Thus $\mathbf{P}_l = [n_S, n_D]$, and the pairwise diversity is calculated as
+Thus $\mathbf{P}_l = [n_S, n_D]$, and the pairwise diversity across the $n$
+sampled chromosomes is calculated as
 
 \begin{align}
-  \bar{\pi}(l) = \frac{n_D}{n_T}
+  \pi(l) = \frac{n_D}{n_T}
 \end{align}
@@
 \begin{align}
-  \bar{\pi}(l) = \frac{2}{n(n-1)} \sum_{i < j}^n k_{i,j}
+  \pi(l) = \frac{2}{n(n-1)} \sum_{i < j}^n k_{i,j}
 \end{align}
@@
 \begin{align*}
-  \bar{\pi}(\mathcal{A}) = \frac{\sum_{l \in \mathcal{A}} \bar{\pi}(l)}{L_\mathcal{A}}.
+  \bar{\pi}(\mathcal{A}) = \frac{\sum_{l \in \mathcal{A}} \pi(l)}{L_\mathcal{A}}.
 \end{align*}
@@
 where $L_\mathcal{A} = |\mathcal{A}|$ is the number of accessible bases. Note
-that both $L_\mathcal{A}$ and the sum are random variables in genomic data.
-The denominator varies along the genome, based on the number of
-reliably-genotyped and accessible bases in the reference genome. This is a
-ratio estimator, and is thus biased, since by Jensen's inequality
-$\E(\nicefrac{y}{x}) \ge \nicefrac{\E(y)}{\E(x)}$.
-
-We explore this with block-bootstrap bias correction. TODO
+that the sum in the numerator is random over the sample of chromosomes sampled
+from the population. Estimates of pairwise diversity often condition on the
+accessible bases, and thus treat this as fixed. However, the number of
+accessible bases varies across the chromosome; this can lead to a source of
+apparent bias during block-bootstrap estimates of uncertainty.  In this case,
+pairwise diversity is a ratio estimator, and is thus biased, since by Jensen's
+inequality $\E(\nicefrac{y}{x}) \ge \nicefrac{\E(y)}{\E(x)}$ for random
+variables $x$ and $y$.
 

### 2022-12-08 manuscript/supplementary.tex
ms

@@
 
+\begin{figure}[!htb]
+  \centering
+  \includegraphics{figures/supplementary/chr1_diversity_filtering.pdf}
+
+  \caption{Estimates of chromosome 1 YRI diversity in non-overlapping megabase
+    windows, under different filtering criteria. The filtering criteria are,
+    (1) ``neutral, accessible" which includes only putatively neutral sites, and
+    ignores regions masked as inaccessible, (2) ``accessible" which is only ignores
+  sites masked as inaccessible, and (3) ``no filtering" which uses all available
+data. Note that filtering changes the absolute level of diversity, but has more
+minor effects on regional patterns of diversity at the megabase scale.}
+
+  \label{suppfig:chr1-diversity}
+\end{figure}
+
 \begin{table}
@@
 
+\begin{figure}[!htb]
+  \centering
+  \includegraphics{figures/supplementary/diversity_trimming_dist.pdf}
+  \caption{}
+  \label{suppfig:trimming}
+\end{figure}
+
 The likelihood method is fit to non-overlapping binned summaries of the allele

### 2022-12-09 bgspy/bootstrap.py
updated notebooks with covariogram, etc

@@ -80,8 +80,8 @@ def process_bootstraps(optim_results):
     thetas = np.stack(thetas).T
     return nlls, thetas
 
-def percentile_ci(boot_thetas, theta, alpha=0.05, log=True):
-    eps = np.quantile(boot_thetas, (alpha/2, 1-alpha/2), axis=1)
+def percentile_ci(boot_thetas, alpha=0.05):
+    eps = np.nanquantile(boot_thetas, (alpha/2, 1-alpha/2), axis=1)
     return eps
 
 def pivot_ci(boot_thetas, theta, alpha=0.05, log=True, axis=1):

### 2022-12-09 bgspy/data.py

@@ -864,22 +864,30 @@ class GenomicBinnedData(GenomicBins):
 
     def bootstrap_pi(self, B, nblocks, filter_masked=True):
         """
+        Resample contigious blocks (nblocks in length) of the Y
+        matrix, sum across rows to end up with
+        (# same comparisons, # different comparisons)
+        and compute π from this. The total number of comparisons
+        is saved for weighting the bootstrap sample.
         """
         Y = self.Y(filter_masked=filter_masked)
         straps = []
+        weights = []
         for b in np.arange(B):
             Y_boot = Y[self.resample_blocks(nblocks), :].sum(axis=0)
             straps.append(pi_from_pairwise_summaries(Y_boot))
+            weights.append(Y_boot.sum())
         straps = np.array(straps)
-        return straps
+        weights = np.array(weights)
+        return straps, weights
 
     def estimate_pi_bias(self, B, nblocks, filter_masked=True):
         """
         Estimate bias(π) from block bootstrapping with nblocks.
         """
         est = self.gwpi(filter_masked)
-        straps = self.bootstrap_pi(B=B, nblocks=nblocks, filter_masked=filter_masked)
-        boot_bias = np.mean(straps) - est
+        straps, weights = self.bootstrap_pi(B=B, nblocks=nblocks, filter_masked=filter_masked)
+        boot_bias = np.average(straps, weights=weights) - est
         return boot_bias
 
     def gwpi(self, filter_masked=True):

### 2022-12-09 manuscript/supplementary.tex
ms supp updates

@@
 
-\subsection{Estimates of Pairwise Diversity}
+\subsection{Data Summary Matrices}
 
@@
 allele count matrix $\mathbf{C}$ with dimensions $L \times 2$, where $L$ is the
-chromosome length. This is converted to a pairwise summary matrix with
-identical dimensions, $\mathbf{P}$. The first column of $\mathbf{P}$ is the
+chromosome length. This is transformed to a pairwise summary matrix with
+identical dimensions, $\mathbf{Y}$. The first column of $\mathbf{Y}$ is the
 number of pairwise comparisons between chromosomes that are identical, and the
 second column is the number that are different. Both of these columns are
-combinatoric summaries of the allele counts matrix. Let $[c_1, c_2]$ be a row
-of $\mathbf{C}$ for basepair $l$, and $n = c_1 + c_2$. Then, the (1) total
-number of pairwise combinations of chromosomes $n_T$, (2) the number of
-pairwise with identical alleles $n_S$, and (3) the number of pairwise
-combinations with differing alleles $n_D$ are respectively,
+combinatoric summaries of the raw allele counts matrix needed for the binomial
+likelihood and pairwise diversity estimates. Let $[c_1, c_2]$ be a row of
+$\mathbf{C}$ for basepair $l$ (the $l$ index is omitted for clarity), and $n =
+c_1 + c_2$. Then, the (1) total number of pairwise combinations of chromosomes
+$n_T$, (2) the number of pairwise with identical alleles $n_S$, and (3) the
+number of pairwise combinations with differing alleles $n_D$ are respectively,
 
@@
 \end{align*}
+%
+which would be stored in row $\mathbf{Y}_l = [n_S, n_D]$.
 
-Thus $\mathbf{P}_l = [n_S, n_D]$, and the pairwise diversity across the $n$
-sampled chromosomes is calculated as
+\subsection{Pairwise Diversity Estimates}
+
+The pairwise diversity at site $l$ across the $n$ sampled chromosomes can be
+calculated from the $\mathbf{Y}$ matrix as follows,
 
@@
 
+The genome-wide $\bar{\pi}$ can also be estimated by summing the columns of
+$\mathbf{Y}$, and calculating $\nicefrac{n_D}{n_T}$.
+
 \subsection{Window-based Summaries and filtering}
@@
   \includegraphics{figures/supplementary/diversity_trimming_dist.pdf}
-  \caption{}
+
+  \caption{The distribution of diversity across the genome at the megabase
+  scale, with outliers trimmed. The blue line is the normal CDF, fit with MLE
+parameters XXX. The inset figure is the untrimmed data, with trimmed points
+shown in red.}
+
   \label{suppfig:trimming}

### 2022-12-24 bgspy/utils.py
updates, bug fixes

@@ -120,10 +120,10 @@ class BScores:
         assert w in self.w, f"'{w}' is not in w array '{self.w}'"
         assert t in self.t, f"'{t}' is not in w array '{self.t}'"
         wi, ti = self.indices(w, t)
-        Bs = np.full(pos.shape[0], 0., dtype=np.float64)
         logBs = self.B[chrom][:, wi, ti, ...].squeeze()
+        Bs = np.full_like(logBs, 0.)
         # where clause prevents underflow, which are set to zero.
-        Bs = np.exp(logBs, Bs, where=-logBs < np.log(np.finfo(logBs.dtype).max))
+        Bs = np.exp(logBs, out=Bs, where=-logBs < np.log(np.finfo(logBs.dtype).max))
         return pos, Bs
 
     def pairs(self, chrom, w, t):

### 2022-12-24 manuscript/supplementary.tex
paper updates

@@
 %
-which would be stored in row $\mathbf{Y}_l = [n_S, n_D]$.
+which would be stored in row $\mathbf{Y}_l = [n_S, n_D]$. Note that the
+per-site $\mathbf{Y}$ handles non-polymorphic sites and missing data.
+Non-polymorphic sites have $n_S = {n \choose 2}$ and $n_D = 0$, and missing
+data has $n_S = n_D = 0$.
 
@@
 
-\section{}
+\section*{Likelihood}
+
+Our model is essentially a generalized nonlinear model with a binomial link
+function. This is the form used by \textcite{Elyashiv2016-vt} and
+\textcite{Murphy2022-sj}. These models fit the observed number of pairwise
+nucleotide differences in genomic windows to the expected pairwise diversity
+under some evolutionary model. We only consider BGS models here, so the mean
+function for position $x$ is the product of the proportion by which BGS reduces
+diversity at $x$, $B(x)$, and genome-wide neutral diversity $\pi_0$,
+
+\begin{align}
+  \pi(x) = B(x, \Phi) \pi_0
+\end{align}
+
+$\Phi$ is the set of BGS parameters (i.e. the DFE for each feature type and
+mutation rate), and $\pi_0$ is determined by the drift-effective population
+size, set by only reproductive and reproductive processes. Then the ob
+
+fit a nonlinear theoretic equation to
+observed count data summarized at some pre-specified scale.
+
+putatively neutral site, the observed number of pairwise differences at site
+$i$ is modeled as,
+
+\begin{align}
+  n_{D,i} \sim \text{Binom}(\pi_i, n_{T,i})
+\end{align}
+
+use observed count data summarized at
+some pre-specified scale
+
+\begin{align}
+  \log\mathcal{L}(\theta) = \sum_{v \in \mathcal{V}} \sum_{i \ne j \in \mathcal{S}} \log(P(O_{i,j}(v) | \theta))
+\end{align}
+
+where $\mathcal{V}$ is the set of putatively neutral sites, $\mathcal{S}$ is
+the set of samples, and $\theta$ are the BGS parameters. The indicator variable
+$O_{i,j}(v)$ is 1 if samples $i$ and $j$ are different at site $v$, and zero
+otherwise. Thus as they specify in the paper,
+
+\begin{align}
+  P(O_{i,j}(v) | \theta) = 
+    \begin{cases}
+      \pi(v | \theta), & O_{i,j}(v) = 1 \\
+      1-\pi(v | \theta), & O_{i,j}(v) = 0 \\
+    \end{cases}.
+\end{align}
+
+The total size of the set of samples $\mathcal{S}$ is $n_\mathcal{S} =
+|\mathcal{S}|$. Assuming all sites are biallelic, we can simplify the inner
+summation by counting the number of possible same and different pairwise
+combinations. If a site $v$'s vector of allele counts is $[c_1, c_2]$, the
+total number of pairwise combinations with the same alleles is
+
+\begin{align}
+  n_s(v) = {c_1 \choose 2} + {c_2 \choose 2}
+\end{align}
+
+and the number of different pairwise combinations is
+
+\begin{align}
+  n_d(v) = n_T - n_s(v)
+\end{align}
+
+where $n_T = n_\mathcal{S} (n_\mathcal{S} - 1) / 2$  is the total number of
+pairwise combinations across the sample set $\mathcal{S}$. Note that these
+site-specific counts allow us to use allelic counts directly, and can vary
+across sites. Our log-likelihood is then,
+
+\begin{align}
+  \ell(\theta) &= \sum_{v \in \mathcal{V}} \left[\log(\pi(v | \theta)) n_\text{D}(v) + \log(1-\pi(v | \theta)) n_\text{S}(v)\right].
+\end{align}
+
+In practice, we calculate these values across bins. For each bin, we treat
+$\pi(v | \theta)$ as fixed, assuming that at this scale, the variation in
+expected diversity across sites is minimal. For a particular chromosome, we
+have two classes of sites: those included in the diversity calculation and
+those ignored. The former sites are all putatively neutral and have reliably
+called genotypes, and the other sites are possibly non-neutral or do have
+reliably called genotypes. The total log-likelihood is the sum of bin likelihoods,
+$\ell(b)$
+
+\begin{align}
+   \ell(\theta) =  \sum_b \ell(b | \theta)
+\end{align}
+
+The likelihood within a bin is then,
+
+\begin{align}
+  \ell(b | \theta)  &= \log(\bar{\pi}(b | \theta)) \sum_{v \in \mathcal{V}_b} n_D(v) + \log(1-\bar{\pi}(b | \theta)) \sum_{v \in \mathcal{V}_b} n_S(v)  \\
+                               &= \log(\bar{\pi}(b | \theta)) Y_D(b) + \log(1-\bar{\pi}(b | \theta)) Y_S(b)
+\end{align}
+%
+where the two sum terms as $Y_D(b)$ and $Y_S(b)$ are data reductions at the bin
+level.
+
+If the data are such that only polymorphic sites are considered, we can adapt
+this by partitioning the set $\mathcal{V}$ of neutral sites into polymorphic
+($\mathcal{P}$) and fixed sites ($\mathcal{F}$), i.e. $\mathcal{V} =
+\mathcal{P} \cup \mathcal{F}$ and $\mathcal{P} \cap \mathcal{F} = \varnothing$.
+For all $v \in \mathcal{F}$, $n_d(v) = 0$ and $n_s(v) = n_T$.
+
+Then,
+
+\begin{align}
+  \ell(\theta) &= \sum_{v \in \mathcal{V}} \left[\log(\pi(v | \theta)) n_\text{D}(v) + \log(1-\pi(v | \theta)) n_\text{S}(v)\right] \\
+                  &= \sum_{v \in \mathcal{P}} \left[\log(\pi(v | \theta)) n_\text{D}(v) + \log(1-\pi(v | \theta)) n_\text{S}(v)\right] + \sum_{v \in \mathcal{F}} \log(1-\pi(v | \theta)) n_T(v)  \\
+\end{align}
+
+\begin{align}
+  \ell(b | \theta)  &= \log(\bar{\pi}(b | \theta)) \sum_{v \in \mathcal{P}_b} n_\text{D}(v) + \log(1-\bar{\pi}(b | \theta)) \left(\sum_{v \in \mathcal{P}_b} n_\text{S}(v) +  \sum_{v \in \mathcal{F}_b} n_T(v)  \right).
+\end{align}
+
+Note that if we assume that the total number of combinations at each fixed site
+is constant, e.g. $n_T = n_T(v)$ for all $v$, then we can use $\sum_v n_T(v) =
+n_T |\mathcal{F}_b|$.
+
+\
+\section*{The }
+
+The core parts of our likelihood are,
+
+\begin{align}
+  \ell(\theta) &=  \sum_b [\log(\bar{\pi}(b | \theta)) Y_D(b) + \log(1-\bar{\pi}(b | \theta)) Y_S(b)]
+\end{align}
+
+where,
+
+\begin{align}
+  \bar{\pi}(b |\theta) = \pi_0(b) \bar{B}(b | \theta).
+\end{align}
+
+Here, $\bar{B}(b | \theta)$ is the predicted reduction in diversity due to BGS
+in window $b$, given background selection parameters $\theta$. In practice,
+this is site-specific. We can write the reduction at any neutral site $v$ in
+the genome as the product of $B$s across all segments,
+
+\begin{align}
+  B(v | \theta) = \exp\left(- \sum_g \int f(\mu(\mathcal{A}(g)), s, S_g) w(s|\mathcal{A}(g)) ds \right)
+\end{align}
+
+where $S_g$ is exogenous genomic data about the segment, $S_g = \{L_g, r_g,
+\rho(|v-p_g|)\}$, where $L_g$ is the segment's length, $r_g$ is the
+recombination rate per basepair in the segment, and $\rho(|v-p_g|)$ is the
+recombination distance between the focal site $v$ and the segment position
+$p_g$ (we approximate, and use the nearest end position to the neutral site).
+Additionally, $w(s|\mathcal{A}(g))$ is the distribution of selection
+coefficients for segment $g$, if segment $g$ is a member of annotation class
+$\mathcal{A}(g)$.
+
+We can think about the DFE as the conditional distribution of a particular
+selection coefficient given a mutation occurs, for a particular annotation
+class. The BGS function $f(\cdot)$ only depends on $\mu$ through the
+introduction of deleterious alleles with selection coefficient $s$ at rate
+$\omega(s|\mathcal{A}(g)) = \mu(\mathcal{A}(g)) w(s | \mathcal{A}(g))$. Thus,
+we can write,
+
+\begin{align}
+  B(v | \theta) = \exp\left(- \sum_g \int f(\omega(s|\mathcal{A}(g)), s, S_g) ds \right)
+\end{align}
+
+which we can discretize as,
+
+\begin{align}
+  B(v | \theta) = \exp\left(- \sum_g \sum_s f(\omega(s|\mathcal{A}(g)), s, S_g) \right).
+\end{align}
+
+Next, note that there are a finite number of annotation classes, $\mathcal{A}
+\to \{a_1, a_2, \ldots, a_k\}$, so we can further partition this as
+
+\begin{align}
+  B(v | \theta) = \exp\left(- \sum_{\{g \;:\; \mathcal{A}(g) = a_1\}} \sum_s f(\omega(s| a_1), s, S_g) + \sum_{\{g \;:\; \mathcal{A}(g) = a_2\}} \sum_s f(\omega(s| a_2), s, S_g) + \ldots  \right)
+\end{align}
+
+Let us define the $d_\omega \times d_s \times d_g$ multidimensional array
+$\mathbf{F}$, and the $d_g \times d_a$ feature classification matrix
+$\mathbf{A}$.
+
+\section{Calculation of B Maps}
+
+Each B map is determined by the annotation track of putatively conserved
+segments, which includes (1) the segment positions and lengths, (2) the type of
+feature of each segment (e.g.  exons, UTRs, phastcons conserved elements,
+etc.), as well as (3) the recombination map, and (3) the deleterious selection
+coefficient $s(i)$ and mutation rate at which these deleterious enter the
+population for feature type $i$.
+
+The B maps are computed at fixed, evenly spaced sites.
+
+\section{Ratchet Rate Prediction}
+
+\begin{align}
+  R = \int 
+\end{align}
 

### 2022-12-26 manuscript/supplementary.tex
some noets

@@
 function for position $x$ is the product of the proportion by which BGS reduces
-diversity at $x$, $B(x)$, and genome-wide neutral diversity $\pi_0$,
+diversity at position $x$, $B(x)$, and genome-wide neutral diversity $\pi_0$,
 
@@
 mutation rate), and $\pi_0$ is determined by the drift-effective population
-size, set by only reproductive and reproductive processes. Then the ob
+size, set by only reproductive and demographic processes.
 
-fit a nonlinear theoretic equation to
-observed count data summarized at some pre-specified scale.
+Our statistical models use a binomial likelihood for the observed data
+summarized at some physical genomic scale. The evolutionary processes that
+impact $\pi(x)$ do so on different spatial scales, and genomic evolutionary
+model selection also includes determining the scale at which data fits best.
+The likelihood considers the probability of observing $n_{D,i}$ pairwise
+differences in window $i$ given $n_{T,i}$ total possible pairwise comparison,
+and that the probability of observing a pairwise difference is given by $\pi_i$
+
+\begin{align}
+  n_{D,i} \sim \text{Binom}(\pi_i, n_{T,i}).
+\end{align}
+
+This assumes infinite sites, since $\pi_i$ is assumed to be the probability of
+
+Note that this makes the approximation that the average number of pairwise
+differences $\pi_x$ is small, such that the probability of heterozygosity
+is equal to the
+
+, which links the mean function
+above with the probability of observing
+
+sampling process by which individuals are sampled from the population
 
@@
 $i$ is modeled as,
-
-\begin{align}
-  n_{D,i} \sim \text{Binom}(\pi_i, n_{T,i})
-\end{align}
 

### 2022-12-26 bgspy/likelihood.py
two alleles model test

@@ -10,7 +10,7 @@ from tabulate import tabulate
 from functools import partial
 from collections import Counter, defaultdict
 import numpy as np
-from ctypes import POINTER, c_double, c_ssize_t
+from ctypes import POINTER, c_double, c_ssize_t, c_int
 
 # no longer needed
 # HAS_JAX = False
@@ -493,7 +493,7 @@ def negll_mutation_numba(theta, Y, logB, w):
 def check_bounds(x, lb, ub):
     assert np.all((x >= lb) & (x <= ub))
 
-def negll_c(theta, Y, logB, w):
+def negll_c(theta, Y, logB, w, two_alleles=True):
     """
     θ is [π0, μ, w11, w12, ...] and should
     have dimension (nt x nf) + 2
@@ -515,30 +515,11 @@ def negll_c(theta, Y, logB, w):
                               POINTER(c_double),
                               # weird type for dims/strides
                               POINTER(np.ctypeslib.c_intp),
-                              POINTER(np.ctypeslib.c_intp))
+                              POINTER(np.ctypeslib.c_intp),
+                              c_int)
     likclib.negloglik.restype = c_double
     return likclib.negloglik(theta_ptr, nS_ptr, nD_ptr, logB_ptr, w_ptr,
-                             logB.ctypes.shape, logB.ctypes.strides)
-
-def normal_ll_c(theta, Y, logB, w):
-    nS = np.require(Y[:, 0].flat, np.float64, ['ALIGNED'])
-    nD = np.require(Y[:, 1].flat, np.float64, ['ALIGNED'])
-    theta = np.require(theta, np.float64, ['ALIGNED'])
-    logB = np.require(logB, np.float64, ['ALIGNED'])
-    nS_ptr = nS.ctypes.data_as(POINTER(c_double))
-    nD_ptr = nD.ctypes.data_as(POINTER(c_double))
-    theta_ptr = theta.ctypes.data_as(POINTER(c_double))
-    logB_ptr = logB.ctypes.data_as(POINTER(c_double))
-    w_ptr = w.ctypes.data_as(POINTER(c_double))
-    likclib.negloglik.argtypes = (POINTER(c_double), POINTER(c_double),
-                              POINTER(c_double), POINTER(c_double),
-                              POINTER(c_double),
-                              # weird type for dims/strides
-                              POINTER(np.ctypeslib.c_intp),
-                              POINTER(np.ctypeslib.c_intp))
-    likclib.normal_loglik.restype = c_double
[…]

### 2022-12-26 bgspy/src/likelihood.c

@@ -7,6 +7,9 @@
 // if you're a bit over the max boundary, we just truncate
 #define MUTMAX_THRESH 1e-10
 
+// whether to use the two-alleles model or infinite sites
+#define TWO_ALLELES 1
+
 #define LOGBW_GET(B, ii, ll, jj, kk, s) B[(ii)*s[0] + (ll)*s[1] + (jj)*s[2] + (kk)*s[3]]
 #define W_GET(WW, jj, kk, col) WW[col*(jj) + kk]
 
@@ -124,7 +127,8 @@ double negloglik(const double *theta,
                  const double *logB, 
                  const double *w,
                  const ssize_t *logB_dim, 
-                 const ssize_t *logB_strides) {
+                 const ssize_t *logB_strides,
+                 const int two_alleles) {
 
     ssize_t nx = logB_dim[0]; 
     ssize_t nw = logB_dim[1];
@@ -181,7 +185,14 @@ double negloglik(const double *theta,
             }
         }
         //printf("%g, ", logBw[i]);
-        double log_pi = log(pi0) + logBw_i;
+        double log_pi;
+        if (two_alleles) {
+            double pi = pi0 * exp(logBw_i);
+            log_pi = log(pi / (1 + 2*pi));
+            /* printf("pi = %f, pi_2a =  %f\n", pi, exp(log_pi)); */
+        } else {
+            log_pi = log(pi0) + logBw_i;
+        }
         //printf("c log(pi0): %g\n", log(pi0));
         //printf("c nD[%d]=%g, nS[%d]=%g\n", i, nD[i], i, nS[i]);
         //printf("c llm[%d]: %g\n", i, nD[i]*log_pi + nS[i]*log(1 - exp(log_pi)));

### 2022-12-26 tools/get_gene_locs.py

@@ -1,19 +1,41 @@
 import sys
+from collections import defaultdict
 import pickle
 from bgspy.utils import readfile
 
 genes = dict()
+transcripts = dict()
+gff_file = readfile(sys.argv[1])
 
-for line in readfile(sys.argv[1]):
+def parse_keyvals(keyval_str):
+    keyvals = dict([x.split('=') for x in keyval_str.split(';')])
+    return keyvals
+
+for line in gff_file:
     if line.startswith('#'):
         continue
     chrom, _, feature, start, end, _, _, _, info = line.strip().split('\t')
-    if feature != 'gene':
+    if feature == 'gene':
+        keyvals = parse_keyvals(info)
+        key = keyvals['ID'].replace('gene:', '')
+        genes[key] = (set(), 0)
         continue
-    keyvals = dict([x.split('=') for x in info.split(';')])
-    key = keyvals['ID'].replace('gene:', '')
-    start, end = int(start), int(end)
-    genes[key] = (chrom, start, end)
+
+    if feature == 'mRNA'
+        keyvals = parse_keyvals(info)
+        key = keyvals['ID'].replace('transcript:', '')
+        transcripts[key] = set()
+        continue
+
+    if feature == 'exon'
+        keyvals = parse_keyvals(info)
+        key = keyvals['Parent'].replace('transcript:', '')
+        start, end = int(start), int(end)
+        transcripts[key].add(end-start)
+        continue
+
+   
+        
 
 with open('gene_coords.pkl', 'wb') as f:
     pickle.dump(genes, f)

### 2022-12-27 manuscript/supplementary.tex
ms notes

--- 2022-12-09 manuscript/supplementary.tex
@@
 %
-which would be stored in row $\mathbf{Y}_l = [n_S, n_D]$.
+which would be stored in row $\mathbf{Y}_l = [n_S, n_D]$. Note that the
+per-site $\mathbf{Y}$ handles non-polymorphic sites and missing data.
+Non-polymorphic sites have $n_S = {n \choose 2}$ and $n_D = 0$, and missing
+data has $n_S = n_D = 0$.
 
 \subsection{Pairwise Diversity Estimates}
+\label{supsec:diversity}
 
 The pairwise diversity at site $l$ across the $n$ sampled chromosomes can be
-calculated from the $\mathbf{Y}$ matrix as follows,
+calculated from row $l$ of the $\mathbf{Y}$ matrix as follows,
 
 \begin{align}
-  \pi(l) = \frac{n_D}{n_T}
+  \pi_l = \frac{n_D}{n_T}
 \end{align}
@@
 \begin{align}
-  \pi(l) = \frac{2}{n(n-1)} \sum_{i < j}^n k_{i,j}
+  \pi_l = \frac{2}{n(n-1)} \sum_{i < j}^n k_{i,j}
 \end{align}
 
-where $k_{i,j}$ is 1 if the alleles at this site differ, and 0 otherwise. Note
-that for sites with missing data, $c_1 = c_2 = 0$ and thus $n_T = 0$ and the
-division is invalid. Such cases were marked as missing with floating point
-values \texttt{NaN}s. Binned summaries of data are weighed by the number of
-complete cases, e.g. rows with $c_1 + c_2 > 0$.
+where $k_{i,j}$ is 1 if the alleles at this site differ at site $l$, and 0
+otherwise.
 
-This per-basepair diversity is then averaged over all callable bases, to give a
-genome-wide or window estimate of diversity of accessible bases in the set
-$\mathcal{A}$,
+There are three ways to aggregate $\pi_l$ across all sites. The first is,
 
-\begin{align*}
-  \bar{\pi}(\mathcal{A}) = \frac{\sum_{l \in \mathcal{A}} \pi(l)}{L_\mathcal{A}}.
-\end{align*}
+\begin{align}
+  \label{eq:}
+  \pi^{(1)} = \frac{1}{L} \sum_{i=1}^L \frac{n_{D,i}}{n_{T,i}}
+\end{align}
 
-where $L_\mathcal{A} = |\mathcal{A}|$ is the number of accessible bases. Note
-that the sum in the numerator is random over the sample of chromosomes sampled
-from the population. Estimates of pairwise diversity often condition on the
+which if the number of samples across loci is constant, simplifies to an
+unweighted average across sites. Second, one can take a weighted average, with
+weights determined by the total number of samples present at a site,
+
+\begin{align}
+  \label{eq:}
+  \pi^{(2)} = \frac{1}{\sum_{i=1}^L n_i} \sum_{i=1}^L n_i \frac{n_{D,i}}{n_{T,i}}
+\end{align}
+
+Third, one can weight by the number of pairwise comparisons at a site,
+$n_{T,i}$, rather than total number of samples, $n_i$, which leads to a ratio
+of sums,
+
+\begin{align}
+  \label{eq:}
+  \pi^{(3)} = \frac{\sum_{i=1}^L n_{D,i}}{\sum_{i=1}^L n_{T,i}}.
+\end{align}
+
+We predominantly use the estimator $\pi^{(3)}$, as it corresponds to how we
+summarize the matrix $\mathbf{Y}$ across windows for our likelihood. All
+methods have mean squared errors and biases very close to one another (TODO).
+
+Note, however, that estimates of pairwise diversity often condition on the
 accessible bases, and thus treat this as fixed. However, the number of
@@
 variables $x$ and $y$.
-
-The genome-wide $\bar{\pi}$ can also be estimated by summing the columns of
-$\mathbf{Y}$, and calculating $\nicefrac{n_D}{n_T}$.
 
@@
 
-\section{}
+\section*{Likelihood}
+
+Our model is essentially a generalized nonlinear model with a binomial link
+function. This is the form used by \textcite{Elyashiv2016-vt} and
+\textcite{Murphy2022-sj}. These models fit the observed number of pairwise
+nucleotide differences in genomic windows to the expected pairwise diversity
+under some evolutionary model. We only consider BGS models here, so the mean
+function for position $x$ is the product of the proportion by which BGS reduces
+diversity at position $x$, $B(x)$, and genome-wide neutral diversity $\pi_0$,
+
+\begin{align}
+  \pi(x) = B(x, \Phi) \pi_0
+\end{align}
+
+$\Phi$ is the set of BGS parameters (i.e. the DFE for each feature type and
+mutation rate), and $\pi_0 = 4 N_e \mu$ is determined by the genome-wide
+drift-effective population size $N_e$, set by only reproductive and demographic
+processes. Regional mutation rate heterogeneity can be accounted for with a
+regional mutation rate scaling function $m(x)$, $\pi(x) = B(x, \Phi) m(x)
+\pi_0$, but we do not explore mutation-rate heterogeneity, as it is unclear how
+to differentiate variance in mutation rate from the substitution rate
+heterogeneity we find across the genome.
+
+The likelihood of the set of parameters $\Psi = \{\pi_0, \Phi\}$ can be written
+in the form of,
+
+\begin{align}
+  \log\mathcal{L}(\Psi) = \sum_{v \in \mathcal{V}} \sum_{i \ne j \in \mathcal{S}} \log(P(O_{i,j}(v) | \Psi))
+\end{align}
+
+(c.f. \cite{McVicker2009-ax,Elyashiv2016-vt,Murphy2022-sj}) where $\mathcal{V}$
+is the set of putatively neutral sites, $\mathcal{S}$ is the set of samples,
+and $\Psi$ are the BGS parameters. The indicator variable $O_{i,j}(v)$ is 1 if
+samples $i$ and $j$ are different at putatively neutral site $v$, and zero
+otherwise. While the theoretic $\pi(v)$ gives the average number of pairwise
+differences, for small values, this is approximately the heterozygosity
+probability, so we can write
+
+\begin{align}
+  P(O_{i,j}(v) | \Psi) = 
+    \begin{cases}
+      \pi(v), & O_{i,j}(v, \Psi) = 1 \\
+      1-\pi(v), & O_{i,j}(v, \Psi) = 0 \\
+    \end{cases}.
+\end{align}
+
+(c.f. \cite{Elyashiv2016-vt}).
+
+As in Section \ref{supsec:diversity}, the number of pairwise differences and
+the total number of pairwise comparison at a site are sufficient statistics for
+the likelihood. Then, the binomial log-likelihood for the data at site $v$ is,
+
+\begin{align}
+  \ell_v(\Psi) &= \log(\pi(v, \Psi)) n_{\text{D},v} + \log(1-\pi(v, \Psi)) n_{\text{S},{v}}
+\end{align}
+
+\subsection{The scale of processes}
+
+We can observe $\widehat{\pi}(x)$ at a per-basepair resolution. However, for a
+variety of reasons, we do not want to fit the composite likelihood model to the
+per-basepair scale of data. First, this would be computationally infeasible.
+Second, the mean function $\pi(x)$ varies on a natural scale that is itself a
+free parameter of the model. Our model can be written as,
+
+\begin{align}
+  \ell(\Psi, h) &= \sum_{b} \sum_{v \in \mathcal{V}_b} \ell_v(\Psi) \\
+             &= \sum_{b} \left[\log(\bar{\pi}(b, \Psi)) \sum_{v \in \mathcal{V}_b} n_{\text{D},v} + \log(1-\bar{\pi}(b, \Psi)) \sum_{v \in \mathcal{V}_b} n_{\text{S},{v}}\right] \\
+             &= \sum_{b} \left[\log(\bar{\pi}(b, \Psi)) Y_{\text{D},b} + \log(1-\bar{\pi}(b, \Psi)) Y_{\text{S},{b}}\right] \\
+\end{align}
+
+where $h$ is the bandwidth or window size, $b$ are physical-scale window
+indices, $\bar{\pi}(b | \Psi)$ are the average diversity in bin $b$, and
+$Y_{\text{S},b}$ and $Y_{\text{S},b}$ are the sums across putatively neutral
+sites within a bin. We fit $h$ using out-sample cross validation.
+
+This corresponds to a binomial likelihood for the observed data summarized at
+genomic scale $h$. Thus an alternative way to express this model is as,
+
+\begin{align}
+  Y_{\text{D},b} \sim \text{Binom}(\bar{\pi}(b, \Psi), Y_{\text{D},b} + Y_{\text{S},b}).
+\end{align}
+
+Here, $\bar{\pi}(b, \Psi)$ is assumed to be the \emph{probability} of sampling
+two different alleles, rather than the average \emph{number} of pairwise
+differences; these are approximately equal when $\pi$ is small. This
+corresponds to an identity link function; one could alternatively use a
+two-alleles finite sites model link function of the form, $\pi/(1 + 2 \pi)$. We
+experimented with this and found there was little difference between these link
+functions, so we opted for the simpler identity link function.
+
+\subsection{The Background Selection Reduction Factor}
+
+Under the background selection model, the diversity in window $b$ is $\pi(b,
+\Psi) = \bar{B}(b, \Psi) \pi_0$. Here, $\bar{B}(b | \Psi)$ is the predicted
+reduction in diversity due to BGS in window $b$, given background selection
+parameters $\Psi$. In practice, we pre-calculate $\bar{B}(b | \Psi)$ at fixed
+sites across the genome, and take the average of the fixed sites $B$ values
+within a widow.
+
+In practice, this is site-specific. We can write the
+reduction at any neutral site $v$ in the genome as the product of $B$s across
+all segments,
+
+\begin{align}
+  B(v | \theta) = \exp\left(- \sum_g \int f(\mu(\mathcal{A}(g)), s, S_g) w(s|\mathcal{A}(g)) ds \right)
+\end{align}
+
+where $S_g$ is exogenous genomic data about the segment, $S_g = \{L_g, r_g,
+\rho(|v-p_g|)\}$, where $L_g$ is the segment's length, $r_g$ is the
+recombination rate per basepair in the segment, and $\rho(|v-p_g|)$ is the
+recombination distance between the focal site $v$ and the segment position
+$p_g$ (we approximate, and use the nearest end position to the neutral site).
+Additionally, $w(s|\mathcal{A}(g))$ is the distribution of selection
+coefficients for segment $g$, if segment $g$ is a member of annotation class
+$\mathcal{A}(g)$.
+
+We can think about the DFE as the conditional distribution of a particular
+selection coefficient given a mutation occurs, for a particular annotation
+class. The BGS function $f(\cdot)$ only depends on $\mu$ through the
+introduction of deleterious alleles with selection coefficient $s$ at rate
+$\omega(s|\mathcal{A}(g)) = \mu(\mathcal{A}(g)) w(s | \mathcal{A}(g))$. Thus,
+we can write,
+
+\begin{align}
+  B(v | \theta) = \exp\left(- \sum_g \int f(\omega(s|\mathcal{A}(g)), s, S_g) ds \right)
+\end{align}
+
+which we can discretize as,
+
+\begin{align}
+  B(v | \theta) = \exp\left(- \sum_g \sum_s f(\omega(s|\mathcal{A}(g)), s, S_g) \right).
+\end{align}
+
+Next, note that there are a finite number of annotation classes, $\mathcal{A}
+\to \{a_1, a_2, \ldots, a_k\}$, so we can further partition this as
+
+\begin{align}
+  B(v | \theta) = \exp\left(- \sum_{\{g \;:\; \mathcal{A}(g) = a_1\}} \sum_s f(\omega(s| a_1), s, S_g) + \sum_{\{g \;:\; \mathcal{A}(g) = a_2\}} \sum_s f(\omega(s| a_2), s, S_g) + \ldots  \right)
+\end{align}
+
+Let us define the $d_\omega \times d_s \times d_g$ multidimensional array
+$\mathbf{F}$, and the $d_g \times d_a$ feature classification matrix
+$\mathbf{A}$.
+
+\section{Calculation of B Maps}
+
+Each B map is determined by the annotation track of putatively conserved
+segments, which includes (1) the segment positions and lengths, (2) the type of
+feature of each segment (e.g.  exons, UTRs, phastcons conserved elements,
+etc.), as well as (3) the recombination map, and (3) the deleterious selection
+coefficient $s(i)$ and mutation rate at which these deleterious enter the
+population for feature type $i$.
+
+The B maps are computed at fixed, evenly spaced sites.
+
+\section{Ratchet Rate Prediction}
+
+\begin{align}
+  R = \int 
+\end{align}
 

### 2022-12-27 manuscript/supplementary.tex
some updates

@@
 reduction in diversity due to BGS in window $b$, given background selection
-parameters $\Psi$. In practice, we pre-calculate $\bar{B}(b | \Psi)$ at fixed
-sites across the genome, and take the average of the fixed sites $B$ values
-within a widow.
+parameters $\Psi$. In practice, we pre-calculate $\bar{B}(x | \Psi)$ at fixed
+sites $x$ across the genome, and take the average of the fixed sites $B$ values
+within widow $b$ for the average $\bar{B}(b | \Psi)$.
+
+There are two ways to parameterize background selection models. Both models use
+a fixed $q$-grid of selection coefficients ranging $s \in \{10^{-6}, 10^{-5},
+10^{-4}, 10^{-3}, 10^{-2}, 10^{-1}\}$ (here, $q=6$) and interpolate over a grid
+of mutation rates. First, for the \emph{free-mutation} parameterization, each
+annotation feature class has a free mutation rate parameter for each of these
+selection coefficients. That is, the distribution of fitness effect (DFE) is
+implied by the total mutation rate across selection coefficients for an
+annotation class.  We can normalize by the total mutation rate estimate to get
+the estimated DFE, giving us the DFE weight for annotation class $k$ and
+selection coefficient with index $i$,
+
+\begin{align}
+  \widehat{w}_{i,k} = \frac{\widehat{\mu}_{i,k}}{\sum_i \widehat{\mu}_{i,k}}.
+\end{align}
+
+With this parameterization, there are $1 + K \times q$ for $\pi_0$ and the free
+DFE parameters, where $K$ is the number of annotation features. The MLE
+optimization for this approach is unconstrained, though $\pi_0$ is bounded to
+be within $10^{-5} \le \pi_0 \le 10^{-1}$.
+
+\begin{align}
+  W = \begin{bmatrix}
+    \mu_{10^{-6},\text{CDS}} & \mu_{10^{-6},\text{UTR}} & w_{10^{-6},\text{phastcons}} \\
+    \mu_{10^{-5},\text{CDS}} & \mu_{10^{-5},\text{UTR}} & w_{10^{-5},\text{phastcons}} \\
+    \vdots & \vdots & \vdots \\
+    \mu_{10^{-1},\text{CDS}} & \mu_{10^{-1},\text{UTR}} & w_{10^{-1},\text{phastcons}} \\
+  \end{bmatrix}
+\end{align}
+
+The second parametrization is the \emph{simplex} parametrization, which has a
+single mutation rate across all features, and the DFE is parameterized by a
+matrix $W$
+
+which has $2
++ K \times (q-1)$ free parameters.
 

### 2022-12-28 manuscript/supplementary.tex
methods stuff

--- 2022-12-27 manuscript/supplementary.tex
@@
 \subsection{Pairwise Diversity Estimates}
-\label{supsec:diversity}
+\label{suppsec:diversity}
 
@@
 
-As in Section \ref{supsec:diversity}, the number of pairwise differences and
+As in Section \ref{suppsec:diversity}, the number of pairwise differences and
 the total number of pairwise comparison at a site are sufficient statistics for
@@
 
-where $h$ is the bandwidth or window size, $b$ are physical-scale window
-indices, $\bar{\pi}(b | \Psi)$ are the average diversity in bin $b$, and
+where $h$ is the bandwidth or window size, $b$ the bin index for windows of
+width $h$, $\mathcal{V}_b$ is the set of putatively neutral sites in bin $b$,
+$\bar{\pi}(b | \Psi)$ are the average diversity in bin $b$, and
 $Y_{\text{S},b}$ and $Y_{\text{S},b}$ are the sums across putatively neutral
-sites within a bin. We fit $h$ using out-sample cross validation.
+sites within a bin. Note that by binning, we sum the pairwise summaries of the
+data $\mathbf{Y}$ across sites, so the likelihood across bins is naturally
+weighted by the quantity of observed data. We estimate $h$ by fitting the model
+for a variety of window sizes $h$ and select the best using out-sample cross
+validation, as is standard in non-parametric regression (XXX, Wasserman, p.
+70).
 
-This corresponds to a binomial likelihood for the observed data summarized at
-genomic scale $h$. Thus an alternative way to express this model is as,
+This model corresponds to a binomial likelihood for the observed data
+summarized at genomic scale $h$. Thus an alternative way to express this model
+is as,
 
@@
 
-\subsection{The Background Selection Reduction Factor}
+\section{The Background Selection Models}
 
-Under the background selection model, the diversity in window $b$ is $\pi(b,
+Under the background selection model, the diversity in window $b$ is $\bar{\pi}(b,
 \Psi) = \bar{B}(b, \Psi) \pi_0$. Here, $\bar{B}(b | \Psi)$ is the predicted
 reduction in diversity due to BGS in window $b$, given background selection
-parameters $\Psi$. In practice, we pre-calculate $\bar{B}(b | \Psi)$ at fixed
-sites across the genome, and take the average of the fixed sites $B$ values
-within a widow.
+parameters $\Psi$. In practice, we pre-calculate $\bar{B}(x | \Psi)$ at fixed
+sites $x$ across the genome, and take the average of the fixed sites $B$ values
+within widow $b$ for the average $\bar{B}(b | \Psi)$.
 
-In practice, this is site-specific. We can write the
-reduction at any neutral site $v$ in the genome as the product of $B$s across
-all segments,
+BGS models assume multiplicative fitness effects across sites. Thus, the total
+reduction due to segments across the genome is the product of the individual
+reductions,
 
 \begin{align}
-  B(v | \theta) = \exp\left(- \sum_g \int f(\mu(\mathcal{A}(g)), s, S_g) w(s|\mathcal{A}(g)) ds \right)
+  \label{suppeqn:b}
+  B(x, \Psi) = \prod_g \int_0^1 b(\mu(s, k(g)), s, L_g, r_g, d(x, g)) ds
 \end{align}
 
-where $S_g$ is exogenous genomic data about the segment, $S_g = \{L_g, r_g,
-\rho(|v-p_g|)\}$, where $L_g$ is the segment's length, $r_g$ is the
-recombination rate per basepair in the segment, and $\rho(|v-p_g|)$ is the
-recombination distance between the focal site $v$ and the segment position
-$p_g$ (we approximate, and use the nearest end position to the neutral site).
-Additionally, $w(s|\mathcal{A}(g))$ is the distribution of selection
-coefficients for segment $g$, if segment $g$ is a member of annotation class
-$\mathcal{A}(g)$.
+where, $b(\mu(s, k(g)), s, L_g, r_g, d(x,g))$ is the reduction due to segment
+$g$. Here, $\mu(s, k(g))$ is the per-basepair per-genome rate that mutations
+with selection coefficient $s$ enter segments with feature class $k(g)$, $L_g$
+and $r_g$ are the length in basepairs and recombination rate per basepair of
+segment $g$, and the recombination distance between focal site $x$ and segment
+$g$ is $d(x, g)$. The functional form of $b$ varies depending on whether the
+classic BGS model is used, or our form of Santiago and Caballero's
+(\citeyear{Santiago2016-mu}) equations (see Section XXX).
 
-We can think about the DFE as the conditional distribution of a particular
-selection coefficient given a mutation occurs, for a particular annotation
-class. The BGS function $f(\cdot)$ only depends on $\mu$ through the
-introduction of deleterious alleles with selection coefficient $s$ at rate
-$\omega(s|\mathcal{A}(g)) = \mu(\mathcal{A}(g)) w(s | \mathcal{A}(g))$. Thus,
-we can write,
+In practice, to fit such a model, we must discretize both $\mu(s, k(g))$ and
+$s$. Our models use a fixed $q$-element grid of selection coefficients, and
+mutation rates are on a fixed $z$-element grid, which is interpolated over
+during ML optimization. The number of annotation features $K$ is variable and
+set by the annotation data specified. There are two ways to parameterize the
+mutation rates and DFE for all annotation classes in background selection
+models. First, for the \emph{free-mutation} parameterization, each annotation
+feature class has a free mutation rate parameter for each of these selection
+coefficients (as in equation \eqref{suppeqn:b}).
+
+In the free-mutation parameterization, there are $1 + K \times q$ for $\pi_0$
+and the free mutation rate parameters, where $K$ is the number of annotation
+features. The MLE optimization for this approach is unconstrained, though
+$\pi_0$ is bounded to be within $10^{-5} \le \pi_0 \le 10^{-1}$ and mutation
+rates are bounded within $10^{-11} \le \mu \le 10^{-7}$. The BGS model requires
+a mutation rate for each selection coefficient in the grid, for each annotation
+class (here, as an example, CDS, UTRs, and phastcons), which are stored in the
+$q \times K$ matrix $\mathbf{M}_\text{F}$,
 
 \begin{align}
-  B(v | \theta) = \exp\left(- \sum_g \int f(\omega(s|\mathcal{A}(g)), s, S_g) ds \right)
+  \mathbf{M}_\text{F} = \begin{bmatrix}
+    \mu_{10^{-6},\text{CDS}} & \mu_{10^{-6},\text{UTR}} & \mu_{10^{-6},\text{PC}} \\
+    \mu_{10^{-5},\text{CDS}} & \mu_{10^{-5},\text{UTR}} & \mu_{10^{-5},\text{PC}} \\
+    \vdots & \vdots & \vdots \\
+    \mu_{10^{-1},\text{CDS}} & \mu_{10^{-1},\text{UTR}} & \mu_{10^{-1},\text{PC}} \\
+  \end{bmatrix}
 \end{align}
 
-which we can discretize as,
+That is, the distribution of fitness effect (DFE) is implied by
+the total mutation rate across selection coefficients for an annotation class.
+We can normalize by the total mutation rate estimate to get the estimated DFE,
+giving us the DFE weight for annotation class $k$ and selection coefficient
+with index $i$,
 
 \begin{align}
-  B(v | \theta) = \exp\left(- \sum_g \sum_s f(\omega(s|\mathcal{A}(g)), s, S_g) \right).
+  \widehat{w}_{i,k} = \frac{\widehat{\mu}_{i,k}}{\sum_i \widehat{\mu}_{i,k}}.
 \end{align}
 
-Next, note that there are a finite number of annotation classes, $\mathcal{A}
-\to \{a_1, a_2, \ldots, a_k\}$, so we can further partition this as
+The second parametrization is the \emph{simplex} parametrization, which has a
+single mutation rate across all features, so it is a DFE weight matrix times
+the mutation rate $\mu$,
 
 \begin{align}
-  B(v | \theta) = \exp\left(- \sum_{\{g \;:\; \mathcal{A}(g) = a_1\}} \sum_s f(\omega(s| a_1), s, S_g) + \sum_{\{g \;:\; \mathcal{A}(g) = a_2\}} \sum_s f(\omega(s| a_2), s, S_g) + \ldots  \right)
+  \mathbf{M}_\text{S} = \mu \begin{bmatrix}
+    w_{10^{-6},\text{CDS}} & w_{10^{-6},\text{UTR}} & w_{10^{-6},\text{PC}} \\
+    w_{10^{-5},\text{CDS}} & w_{10^{-5},\text{UTR}} & w_{10^{-5},\text{PC}} \\
+    \vdots & \vdots & \vdots \\
+    w_{10^{-1},\text{CDS}} & w_{10^{-1},\text{UTR}} & w_{10^{-1},\text{PC}} \\
+  \end{bmatrix}
 \end{align}
 
-Let us define the $d_\omega \times d_s \times d_g$ multidimensional array
-$\mathbf{F}$, and the $d_g \times d_a$ feature classification matrix
-$\mathbf{A}$.
+which has $2 + K \times (q-1)$ free parameters, since each column must sum to
+one. This imposes the constraint that $\sum_i w_{i,k} = 1$, requiring
+constrained MLE optimization.
 
-\section{Calculation of B Maps}
+\subsection{Calculation of the classic BGS B Maps}
 
-Each B map is determined by the annotation track of putatively conserved
-segments, which includes (1) the segment positions and lengths, (2) the type of
-feature of each segment (e.g.  exons, UTRs, phastcons conserved elements,
-etc.), as well as (3) the recombination map, and (3) the deleterious selection
-coefficient $s(i)$ and mutation rate at which these deleterious enter the
-population for feature type $i$.
+The classic BGS model expresses the reduction for a focal site in the middle of
+a segment \parencite{Hudson1995-xc,Hudson1994-oh,Nordborg1996-nq}; here we
+consider a focal site directly to an arbitrary side of a segment under a
+deleterious mutations, and extend the model to handle a focal segment
+arbitrarily far from the segment. Using our notation from equation
+\eqref{suppeqn:b}, the classic mutation-selection-balance BGS model is,
 
-The B maps are computed at fixed, evenly spaced sites.
+\begin{align}
+  b(\mu, s, L, r, 0) &= \exp \left( - \int_0^L \frac{2\mu(l)}{s (1 + (1-s) r(l)/s)^2} dl \right) 
+\end{align}
+
+(c.f. \cite{Hudson1995-xc} equation 5), where $\mu(l)$ is the \emph{haploid}
+per-basepair mutation rate at $l$ and $r(l)$ is the recombination fraction
+between the focal site (at position 0) and basepair $l$. Assuming constant
+per-basepair recombination and mutation rates, $r(l) = r_\text{BP}l$ and
+$\mu(l) = \mu$, and
+
+\begin{align}
+  b(\mu, s, L, r_\text{BP}, 0) &= \exp \left( - \frac{2\mu L}{s + (1-s) r_\text{BP} L} \right)
+\end{align}
+
+(c.f. \cite{Hudson1995-xc} equation 8, where the factor of two difference is
+due to the position of the focal site). Often per-region rates are defined $U =
+2 \mu L$ and $R = r_\text{BP} L$ are the segment-wide mutation diploid mutation
+rates (mutations per diploid segment per generation) and recombination length
+(Morgans).
+
+For computational efficiency, we adjust this model by setting $r(l) = d +
+r_\text{BP} l$, where $d$ is the recombination fraction between the focal site
+$x$ and the start of segment $g$. This allows us to pre-compute the local
+effects of the segment, and substitute in $b$ as the focal site changes. This
+integral substitution has a closed form,
+
+\begin{align}
+  b(\mu, s, L, r_\text{BP}, d) &= \exp \left( \frac{-2\mu L}{(d(1-s) + s)(s + (d + r_\text{BP}L)(1-s))} \right)
+\end{align}
+
+and terms can be collected in powers of $d$ and pre-computed for all segments
+for the grid of $\mu$ and $s$. Our implementation pre-computes these segment
+components, and computes the final $b$ value for an $x$ and $g$ by calculating
+the distance $d(x,g) = |M(x) - M(g)|$, where $M(x)$ is the cumulative
+recombination map length at position $x$ in Morgans. Since $r(l)$ could vary by
+position, segments with differing recombination rates are split into segments
+of the same annotation class by their recombination rate. This has no effect on
+the calculation due to the multiplicative fitness, but ensures more accurate
+estimates of the reduction factor $B$.
+
+% Each B map is determined by the annotation track of putatively conserved
+% segments, which includes (1) the segment positions and lengths, (2) the type of
+% feature of each segment (e.g.  exons, UTRs, phastcons conserved elements,
+% etc.), as well as (3) the recombination map, and (3) the deleterious selection
+% coefficient $s(i)$ and mutation rate at which these deleterious enter the
+% population for feature type $i$.
 

### 2022-12-28 manuscript/supplementary.tex
more ms updates

@@
 
-\section{The Background Selection Models}
+\section{The Generalized Background Selection Models}
 
-Under the background selection model, the diversity in window $b$ is $\bar{\pi}(b,
-\Psi) = \bar{B}(b, \Psi) \pi_0$. Here, $\bar{B}(b | \Psi)$ is the predicted
-reduction in diversity due to BGS in window $b$, given background selection
-parameters $\Psi$. In practice, we pre-calculate $\bar{B}(x | \Psi)$ at fixed
-sites $x$ across the genome, and take the average of the fixed sites $B$ values
-within widow $b$ for the average $\bar{B}(b | \Psi)$.
+Under the generalized background selection models used in our likelihood, the
+diversity in window $b$ is $\bar{\pi}(b, \Psi) = \bar{B}(b, \Psi) \pi_0$. Here,
+$\bar{B}(b | \Psi)$ is the predicted reduction in diversity due to BGS in
+window $b$, given background selection parameters $\Psi$. In practice, we
+pre-calculate $B(x | \Psi)$ at fixed sites $x$ across the genome, and take the
+average of the fixed sites $B$ values within widow $b$ for the average
+$\bar{B}(b | \Psi)$.
 
-BGS models assume multiplicative fitness effects across sites. Thus, the total
+GBGS models assume multiplicative fitness effects across sites. Thus, the total
 reduction due to segments across the genome is the product of the individual
@@
 (\citeyear{Santiago2016-mu}) equations (see Section XXX).
+
+\subsection{Reductions Under the Classic BGS Model}
+
+The classic BGS model, which assumes deleterious mutations cannot fix,
+expresses the reduction for a focal site in the middle of a segment
+\parencite{Hudson1995-xc,Hudson1994-oh,Nordborg1996-nq}; here we consider a
+focal site directly to an arbitrary side of a segment under a deleterious
+mutations, and extend the model to handle a focal segment arbitrarily far from
+the segment. Using our notation from equation \eqref{suppeqn:b}, the classic
+mutation-selection-balance BGS model is,
+
+\begin{align}
+  b(\mu, s, L, r, 0) &= \exp \left( - \int_0^L \frac{2\mu(l)}{s (1 + (1-s) r(l)/s)^2} dl \right) 
+\end{align}
+
+(c.f. \cite{Hudson1995-xc} equation 5), where $\mu(l)$ is the \emph{haploid}
+per-basepair mutation rate at $l$ and $r(l)$ is the recombination fraction
+between the focal site (at position 0) and basepair $l$. Assuming constant
+per-basepair recombination and mutation rates, $r(l) = r_\text{BP}l$ and
+$\mu(l) = \mu$, and
+
+\begin{align}
+  b(\mu, s, L, r_\text{BP}, 0) &= \exp \left( - \frac{2\mu L}{s + (1-s) r_\text{BP} L} \right)
+\end{align}
+
+(c.f. \cite{Hudson1995-xc} equation 8, where the factor of two difference is
+due to the position of the focal site). Often per-region rates are defined $U =
+2 \mu L$ and $R = r_\text{BP} L$ are the segment-wide mutation diploid mutation
+rates (mutations per diploid segment per generation) and recombination length
+(Morgans).
+
+For computational efficiency, we adjust this model by setting $r(l) = d +
+r_\text{BP} l$, where $d$ is the recombination fraction between the focal site
+$x$ and the start of segment $g$. This allows us to pre-compute the local
+effects of the segment, and substitute in $b$ as the focal site changes. This
+integral substitution has a closed form,
+
+\begin{align}
+  b(\mu, s, L, r_\text{BP}, d) &= \exp \left( \frac{-2\mu L}{(d(1-s) + s)(s + (d + r_\text{BP}L)(1-s))} \right)
+\end{align}
+
+and terms can be collected in powers of $d$ and pre-computed for all segments
+for the grid of $\mu$ and $s$. Our implementation pre-computes these segment
+components, and computes the final $b$ value for an $x$ and $g$ by calculating
+the distance $d(x,g) = |M(x) - M(g)|$, where $M(x)$ is the cumulative
+recombination map length at position $x$ in Morgans. Since $r(l)$ could vary by
+position, segments with differing recombination rates are split into segments
+of the same annotation class by their recombination rate. This has no effect on
+the calculation due to the multiplicative fitness, but ensures more accurate
+estimates of the reduction factor $B$.
+
+% Each B map is determined by the annotation track of putatively conserved
+% segments, which includes (1) the segment positions and lengths, (2) the type of
+% feature of each segment (e.g.  exons, UTRs, phastcons conserved elements,
+% etc.), as well as (3) the recombination map, and (3) the deleterious selection
+% coefficient $s(i)$ and mutation rate at which these deleterious enter the
+% population for feature type $i$.
+
+\subsection{Reduction Under the Modified Santiago and Caballero Model}
+
+When weak deleterious mutations can fix, causing substitutions, the classic
+background selection models break down. We find that the generalized background
+selection model of \textcite{Santiago2016-mu} works well (see Section XXX for
+our derivation). We describe how we use the models here.
+
+\subsection{Discretization of Parameters}
 
@@
 
-\subsection{Calculation of the classic BGS B Maps}
-
-The classic BGS model expresses the reduction for a focal site in the middle of
-a segment \parencite{Hudson1995-xc,Hudson1994-oh,Nordborg1996-nq}; here we
-consider a focal site directly to an arbitrary side of a segment under a
-deleterious mutations, and extend the model to handle a focal segment
-arbitrarily far from the segment. Using our notation from equation
-\eqref{suppeqn:b}, the classic mutation-selection-balance BGS model is,
-
-\begin{align}
-  b(\mu, s, L, r, 0) &= \exp \left( - \int_0^L \frac{2\mu(l)}{s (1 + (1-s) r(l)/s)^2} dl \right) 
-\end{align}
-
-(c.f. \cite{Hudson1995-xc} equation 5), where $\mu(l)$ is the \emph{haploid}
-per-basepair mutation rate at $l$ and $r(l)$ is the recombination fraction
-between the focal site (at position 0) and basepair $l$. Assuming constant
-per-basepair recombination and mutation rates, $r(l) = r_\text{BP}l$ and
-$\mu(l) = \mu$, and
-
-\begin{align}
-  b(\mu, s, L, r_\text{BP}, 0) &= \exp \left( - \frac{2\mu L}{s + (1-s) r_\text{BP} L} \right)
-\end{align}
-
-(c.f. \cite{Hudson1995-xc} equation 8, where the factor of two difference is
-due to the position of the focal site). Often per-region rates are defined $U =
-2 \mu L$ and $R = r_\text{BP} L$ are the segment-wide mutation diploid mutation
-rates (mutations per diploid segment per generation) and recombination length
-(Morgans).
-
-For computational efficiency, we adjust this model by setting $r(l) = d +
-r_\text{BP} l$, where $d$ is the recombination fraction between the focal site
-$x$ and the start of segment $g$. This allows us to pre-compute the local
-effects of the segment, and substitute in $b$ as the focal site changes. This
-integral substitution has a closed form,
-
-\begin{align}
-  b(\mu, s, L, r_\text{BP}, d) &= \exp \left( \frac{-2\mu L}{(d(1-s) + s)(s + (d + r_\text{BP}L)(1-s))} \right)
-\end{align}
-
-and terms can be collected in powers of $d$ and pre-computed for all segments
-for the grid of $\mu$ and $s$. Our implementation pre-computes these segment
-components, and computes the final $b$ value for an $x$ and $g$ by calculating
-the distance $d(x,g) = |M(x) - M(g)|$, where $M(x)$ is the cumulative
-recombination map length at position $x$ in Morgans. Since $r(l)$ could vary by
-position, segments with differing recombination rates are split into segments
-of the same annotation class by their recombination rate. This has no effect on
-the calculation due to the multiplicative fitness, but ensures more accurate
-estimates of the reduction factor $B$.
-
-% Each B map is determined by the annotation track of putatively conserved
-% segments, which includes (1) the segment positions and lengths, (2) the type of
-% feature of each segment (e.g.  exons, UTRs, phastcons conserved elements,
-% etc.), as well as (3) the recombination map, and (3) the deleterious selection
-% coefficient $s(i)$ and mutation rate at which these deleterious enter the
-% population for feature type $i$.
-
 \section{Ratchet Rate Prediction}

### 2022-12-28 manuscript/supplementary.tex
more ms updates

@@
 
-\section{The Generalized Background Selection Models}
+\section{Background Selection Models}
 
-Under the generalized background selection models used in our likelihood, the
-diversity in window $b$ is $\bar{\pi}(b, \Psi) = \bar{B}(b, \Psi) \pi_0$. Here,
-$\bar{B}(b | \Psi)$ is the predicted reduction in diversity due to BGS in
-window $b$, given background selection parameters $\Psi$. In practice, we
-pre-calculate $B(x | \Psi)$ at fixed sites $x$ across the genome, and take the
-average of the fixed sites $B$ values within widow $b$ for the average
-$\bar{B}(b | \Psi)$.
-
-GBGS models assume multiplicative fitness effects across sites. Thus, the total
-reduction due to segments across the genome is the product of the individual
-reductions,
+The BGS models considered here both assume multiplicative fitness effects
+across sites. Thus, the total reduction due to segments across the genome is
+the product of the individual reductions,
 
@@
 
+Under both background selection models used in our likelihood, the diversity in
+window $b$ is $\bar{\pi}(b, \Psi) = \bar{B}(b, \Psi) \pi_0$. Here, $\bar{B}(b |
+\Psi)$ is the predicted reduction in diversity due to BGS in window $b$, given
+background selection parameters $\Psi$. In practice, we pre-calculate $B(x |
+\Psi)$ at fixed sites $x$ across the genome, and take the average of the fixed
+sites $B$ values within widow $b$ for the average $\bar{B}(b | \Psi)$.
+
 \subsection{Reductions Under the Classic BGS Model}
@@
 mutations, and extend the model to handle a focal segment arbitrarily far from
-the segment. Using our notation from equation \eqref{suppeqn:b}, the classic
-mutation-selection-balance BGS model is,
+the segment. To simplify notation, we will only consider a single feature class
+in this and the next section. Using our notation from equation
+\eqref{suppeqn:b}, the classic mutation-selection-balance BGS model is,
 
@@
 due to the position of the focal site). Often per-region rates are defined $U =
-2 \mu L$ and $R = r_\text{BP} L$ are the segment-wide mutation diploid mutation
+2 \mu L$ and $M = r_\text{BP} L$ are the segment-wide mutation diploid mutation
 rates (mutations per diploid segment per generation) and recombination length
@@
 selection model of \textcite{Santiago2016-mu} works well (see Section XXX for
-our derivation). We describe how we use the models here.
+our derivation). Unlike the classic BGS models which only produce B maps, that
+only estimate the reduction in effective population size, The generalized BGS
+model of \textcite{Santiago2016-mu} also predicts the deleterious substitution
+rate per generation per segment (the ``ratchet" rate), which we call the R map.
+We describe how these are calculated substitution rates per segment, for each
+mutation rate and selection coefficient combination in the grids. We describe
+how these are calculated here.
 
-\subsection{Discretization of Parameters}
+First, the reduction factor experienced by a focal site within the segment $g$
+(and thus is similar to what selected alleles in $g$ experience) under
+generalized background selection is calculated as
+
+\begin{align}
+  \widetilde{b}'(\mu, s, L, r_\text{BP}) = \frac{\widetilde{N}_{e,\infty}}{N}
+\end{align}
+
+where $\widetilde{N}_{e,\infty}$ is the asymptotic effective population size,
+which is the solution to the following system of two equations (see XXX for
+derivation),
+
+\begin{align}
+  {N}_{e,\infty} &= \exp \left( -\frac{V}{2} Q_\infty^2 \right) \\
+  {R} &= \frac{4 U s N_{e,\infty}}{\exp(4 N_{e,\infty} s)-1}\\
+\end{align}
+
+where
+
+\begin{align}
+  V &= U s - 2 {R} s \\
+  Q_\infty^2 &= \frac{2}{(1-Z)(2-(2-M)Z)} \\
+  Z &= 1 - \frac{Us}{U-{R}}
+\end{align}
+
+and $U = 2\mu L$ and $M = r_\text{BP} L$ as above. Solving this system of two
+equations for each segment also gives the total substitution or ``ratchet" rate
+per generation per segment ($\widetilde{R}$). The per-basepair substitution
+rate of the segment is $\lambda = \nicefrac{\widetilde{R}}{L}$
+
+Under Santiago and Caballero's model, the correct effective population size to
+gauge reduction in heterozygosity or coalescent times is given by equation XXX.
+This is because selective processes can lead to non-constant pairwise
+coalescent rates, which impact pairwise coalescent rates and thus the reduction
+factor. We experimented with the corrections described in Section XXX, and
+found (1) they had little effect on the accuracy of the B maps, and (2) they
+were incredibly costly to calculate, even with the approximations described in
+equations XXX. Overall, we solve for the asymptotic $\widetilde{N}_{e,\infty}$
+for each segment, and use that to calculate $b'$.
+
+Note that the asymptotic model above determines the reduction in effective
+population size experienced by a focal site in the middle of a $L$-basepair
+segment, but B maps require the reduction $b'$ experienced by a focal site $x$
+at an arbitrary recombination distance apart. For each segment, we pre-compute
+the solution to the system of two equations, giving us $V$ and $Q_\infty^2$,
+and then use these values in the equation,
+
+\begin{align}
+  b'(\mu, s, L, r, d) = \exp\left( -\frac{V}{2} \left(\frac{1}{(1-(1-d)Z}\right)^2 \right) 
+\end{align}
+
+where $Z = 1 - Us / (U-\widetilde{R})$ and $V = Us - 2\widetilde{R}s$.
+
+\subsection{Discretization of Parameters and Annotation Feature Classes}
+
+\begin{align}
+  \label{suppeqn:b}
+  B(x, \Psi) = \prod_g \sum_{i} b(\mu_{i, \;k(g)}, s_i, L_g, r_g, d(x, g))
+\end{align}
 
@@
 $s$. Our models use a fixed $q$-element grid of selection coefficients, and
-mutation rates are on a fixed $z$-element grid, which is interpolated over
-during ML optimization. The number of annotation features $K$ is variable and
-set by the annotation data specified. There are two ways to parameterize the
-mutation rates and DFE for all annotation classes in background selection
-models. First, for the \emph{free-mutation} parameterization, each annotation
-feature class has a free mutation rate parameter for each of these selection
-coefficients (as in equation \eqref{suppeqn:b}).
+mutation rates are on a fixed $z$-element grid, which is later interpolated
+over during the maximum likelihood optimization. The number of annotation
+features $K$ is variable and set by the annotation data specified. There are
+two ways to parameterize the mutation rates and DFE for all annotation classes
+in background selection models. First, for the \emph{free-mutation}
+parameterization, each annotation feature class has a free mutation rate
+parameter for each of these selection coefficients (as in equation
+\eqref{suppeqn:b}).
 

### 2022-12-28 bgspy/likelihood.py
getting closer to predicted ratchet map

@@ -493,7 +493,7 @@ def negll_mutation_numba(theta, Y, logB, w):
 def check_bounds(x, lb, ub):
     assert np.all((x >= lb) & (x <= ub))
 
-def negll_c(theta, Y, logB, w, two_alleles=True):
+def negll_c(theta, Y, logB, w, two_alleles=False):
     """
     θ is [π0, μ, w11, w12, ...] and should
     have dimension (nt x nf) + 2
@@ -823,7 +823,7 @@ class BGSLikelihood:
         rows.append(f"  t grid: {signif(self.t)}")
         return "\n".join(rows)
 
-def negll_freemut(Y, B, w, two_alleles=True):
+def negll_freemut(Y, B, w, two_alleles=False):
     """
     This is a closure around data; returns a negative log- likelihood
     function around the data and a few fixed parameters (B and w).
@@ -841,7 +841,7 @@ def negll_freemut(Y, B, w, two_alleles=True):
         return negll_c(new_theta, Y, B, w, two_alleles)
     return func
 
-def negll_freemut_full(theta, grad, Y, B, w, two_alleles=True):
+def negll_freemut_full(theta, grad, Y, B, w, two_alleles=False):
     """
     Like negll_freemut() but not a closure.
 
@@ -896,7 +896,7 @@ class FreeMutationModel(BGSLikelihood):
                                self.log10_pi0_bounds,
                                self.log10_mu_bounds, paired=paired)
 
-    def fit(self, starts=1, ncores=None, algo='ISRES', two_alleles=True,
+    def fit(self, starts=1, ncores=None, algo='ISRES', two_alleles=False,
             _indices=None):
         """
         Fit likelihood models with mumeric optimization (either scipy or nlopt).

### 2022-12-28 bgspy/models.py

@@ -204,12 +204,24 @@ class BGSModel(object):
             binstats = stats.binned_statistic(midpoints, x,
                                               statistic=np.nanmean,
                                               bins=bins)
-            return binstats
+            return midpoints, binstats
         else:
             if not use_midpoints:
                 midpoints = segments.ranges[idx]
             return midpoints, x, features, seglens
 
+    def get_ratchet_array(self, chrom, width):
+        nbins = len(bin_chrom(self.seqlens[chrom], width)) - 1
+        R = np.full((nbins, len(self.w), len(self.t)), np.nan)
+        pos = None
+        for wi, w in enumerate(self.w):
+            for ti, t in enumerate(self.t):
+                mps, bins = self.get_ratchet_rates(wi, ti, chrom, width=width)
+                if pos is not None:
+                    assert np.all(mps == pos)
+                R[:, wi, ti] = bins.statistic
+                pos = mps
+        return pos, R
 
     def fill_Bp_nan(self):
         """

### 2022-12-28 bgspy/src/likelihood.c

@@ -7,9 +7,6 @@
 // if you're a bit over the max boundary, we just truncate
 #define MUTMAX_THRESH 1e-10
 
-// whether to use the two-alleles model or infinite sites
-#define TWO_ALLELES 1
-
 #define LOGBW_GET(B, ii, ll, jj, kk, s) B[(ii)*s[0] + (ll)*s[1] + (jj)*s[2] + (kk)*s[3]]
 #define W_GET(WW, jj, kk, col) WW[col*(jj) + kk]
 

### 2022-12-28 bgspy/theory2.py

@@ -35,8 +35,10 @@ Bclib.Ne_t_rescaled.restype = c_double
 ## Classic BGS
 def B_segment_lazy(rbp, L, t):
     """
-    TODO check rbp = 0 case
-    rt/ (b*(-1 + t) - t) * (b*(-1 + t) + r*(-1 + t) - t)
+    rt/ (rf*(-1 + t) - t) * (rf*(-1 + t) + r*(-1 + t) - t)
+
+    together:
+      x = a/(b*rf**2 + c*rf + d)
     """
     r = rbp*L
     a = -t*L # numerator -- ignores u
@@ -84,8 +86,8 @@ def calc_B_chunk_worker(args):
 
 def Q2_asymptotic(Z, M):
     """
-    This is the map-version asymptotic Q² term — up to the factor of two that
-    *cancels* with the V/2 in a diploid model.
+    This is the map-version asymptotic Q² term — including the factor of two
+    that *cancels* with the V/2 in a diploid model.
     """
     return -2/((-1 + Z)*(2 + (-2 + M)*Z))
 

### 2023-01-14 manuscript/manuscript.tex
methods stuff
more ms updates
ms intro

@@
 
-\title{}
+\title{Towards a Complete Model of Mutation Selection Balance in Humans}
 
@@
 \section*{Introduction}
+
+Eukaryotes genomes are under continual selectionary pressure to remove the
+continual influx of deleterious mutations at conserved regions. This purifying
+selection process keeps mutations that reduce fitness at low frequency in the
+population, and reduces genetic diversity at sites linked to genomic regions
+that are predominantly under purifying selection such as coding sequences and
+regions that have been shown to be phylogenetically conserved between species.
+Evolutionary geneticists have developed a series of mathematical models of this
+process, to predict quantities such as the equilibrium level of segregating
+deleterious mutations (\emph{genetic load}), and the reduction in neutral
+genetic diversity at linked sites due to a faster rate of coalescence (the $B$
+reduction factor). Statistical models of the purifying selection process fit to
+human genomic data closely match observed patterns of genetic variability along
+the genome, explaining around 60\% of the genetic variability at the megabase
+scale \parencite{McVicker2009-ax,Murphy2022-sj}.
+
+These evolutionary genetics models only consider the impact of new mutations
+that are sufficiently deleterious that they are guaranteed to be eliminated
+eventually by selection. When mutations are strongly deleterious, the impact of
+random genetic drift and random perturbations caused by selection at linked
+neighboring sites (genetic draft) on these mutations on their fates can be
+ignored. Under these conditions, mathematical models can assume an stationary
+balance between the input of new mutations entering the population at rate
+$\mu$ per basepair, per generation, and their rate of loss determined by the
+relative reduction in heterozygous fitness, $1-s$. Under these
+mutation-selection balance models, the equilibrium genetic load and the
+reduction in coalescent time are approximately determined by the ratio
+$\nicefrac{\mu}{s}$. However, estimates of the distribution of fitness effects
+suggest many new mutations are weakly deleterious, such that their dynamics can
+no longer ignore the influence of genetic drift and draft.
+
+To understand the this process and make predictions about the equilibrium level
+of deleterious mutations segregating in the population, evolutionary
+geneticists have used models of mutation-selection balance. These models
+
+Non-equilibrium models of deleterious mutation flux.
+
+Generalized background selection
+
+These models
+
 \section*{Results}

### 2023-01-18 manuscript/manuscript.tex
ms updates

@@
 \end{figure}
+
+\subsection*{Evolutionary Models}
+
+CDS+phastcons+UTR model
+
+ - UTR as neutral control
+
+ - CDS takes priority over phastcons, phastcons is fallback. Otherwise, the
+   effect of CDS would be biased by the fraction removed and classified as
+   phastcons. This is the right comparison for the substitution model.
+
+\subsection*{Predicted Substitution Rates}
+
+- mention neg. auto correlation?
+
+- Gene level analysis
+
+To evaluate the predicted substitution rate of our CDS+phastcons+UTR model,
+
+Additionally, our GBGS model makes predictions about the rate of substitutions
+in different segments under purifying selection. There are two stages of the
+prediction process. First, when the initial B' maps are calculated, the
+predicted substitution rate is calculated as a byproduct of
+
+\section*{Methods}
+
+\subsection*{Solving the B' Equations for each Segment}
+
+In our implementation of the theory of \textcite{Santiago2016-mu}, we calculate
+the local effective population size $N_e(x) = b(x) N$ within a segment $x$ by
+solving a system of two non-linear equations,
+
+\begin{align}
+  \label{eq:}
+  {N}_{e}(x) &= \exp \left( -\frac{V(x)}{2} Q^2(x) \right) & \text{\emph{effective population equation}} \\
+  {R}(x) &= \frac{4NU s}{\exp(4 N s)-1}  & \text{\emph{ratchet equation}} 
+\end{align}
+
+where,
+
+\begin{align}
+  V(x) &= U s - 2 {R}(x) s & \text{\emph{variation equation}} \\
+  Q^2(x) &= \frac{2}{(1-Z)(2-(2-M)Z)} & \text{\emph{linkage persistence factor}} \\
+  Z &= 1 - \frac{Us}{U-{R(x)}} & \text{\emph{}} \text{\emph{variance decay rate}}
+\end{align}
+
+and $U = \mu L$ and $M = r_\text{BP} L$ are the total mutation and
+recombination rates for a segment $L$ basepairs long. Each mutation has a
+selection coefficient of $s$, for which we assume additive effects. Then, the
+local reduction factor is $b(x) = \nicefrac{N_e(x)}{N}$, and the per-basepair
+per-generation substitution rate is $\lambda_D = \nicefrac{R(x)}{L}$.
+
+In our software \texttt{bgspy}, these two equations are solved for a grid of
+mutation rates and selection coefficients for each specified segment from a
+feature type (e.g. CDS, UTRs, etc.). This results in estimates of $N_e(x)$ and
+$R(x)$ per segment. In the first-round B' calculation, we use the
+drift-effective population size, $N$.
+
+\subsection*{The B' Maps}
+
+The reduction at a focal site $x$ can be calculated as the product of all
+segments' contributing reduction factors,
+
+\begin{align}
+  B(v) = \prod_x b(x,v)
+\end{align}
+
+where $b(x, v)$ is a modified version of $b(x)$ that accounts for the
+recombination between the segment $x$ and the neutral site $v$ (see XXX).
+
+In practice, we calculate B' using the variables determined by solving the B'
+equations for each mutation and selection coefficient in the grids. Since only
+the analog of $Q^2(x)$, $Q^2(x, v)$ changes, the pre-computed substitution rate
+(and thus local variance) can be re-used for each $v$. We calculate $B(v)$
+across a grid of focal points along the chromosome, spaced 10kb apart.
+
+\subsection*{Maximum Likelihood Estimation}
+
+\subsection*{Local Rescaling and Substitution Rate Prediction}
 

### 2023-01-18 manuscript/supplementary.tex
ms intro
ms updates

@@
   \label{suppeqn:b}
-  B(x, \Psi) = \prod_g \int_0^1 b(\mu(s, k(g)), s, L_g, r_g, d(x, g)) ds
+  B(x, \Psi) = \prod_g^S \int_0^1 b(\mu(s, k(g)), s, L_g, r_g, d(x, g)) ds
 \end{align}
@@
 where, $b(\mu(s, k(g)), s, L_g, r_g, d(x,g))$ is the reduction due to segment
-$g$. Here, $\mu(s, k(g))$ is the per-basepair per-genome rate that mutations
-with selection coefficient $s$ enter segments with feature class $k(g)$, $L_g$
-and $r_g$ are the length in basepairs and recombination rate per basepair of
-segment $g$, and the recombination distance between focal site $x$ and segment
-$g$ is $d(x, g)$. The functional form of $b$ varies depending on whether the
-classic BGS model is used, or our form of Santiago and Caballero's
-(\citeyear{Santiago2016-mu}) equations (see Section XXX).
+$g$, and there are $S$ total segments. Here, $\mu(s, k(g))$ is the per-basepair
+per-genome rate that mutations with selection coefficient $s$ enter segments
+with feature class $k(g)$, $L_g$ and $r_g$ are the length in basepairs and
+recombination rate per basepair of segment $g$, and the recombination distance
+between focal site $x$ and segment $g$ is $d(x, g)$. The functional form of $b$
+varies depending on whether the classic BGS model is used, or our form of
+Santiago and Caballero's (\citeyear{Santiago2016-mu}) equations (see Section
+XXX).
 
@@
 \begin{align}
-  {N}_{e,\infty} &= \exp \left( -\frac{V}{2} Q_\infty^2 \right) \\
+  {N}_{e,\infty} &= N \exp \left( -\frac{V}{2} Q_\infty^2 \right) \\
   {R} &= \frac{4 U s N_{e,\infty}}{\exp(4 N_{e,\infty} s)-1}\\
@@
 
-and $U = 2\mu L$ and $M = r_\text{BP} L$ as above. Solving this system of two
-equations for each segment also gives the total substitution or ``ratchet" rate
-per generation per segment ($\widetilde{R}$). The per-basepair substitution
-rate of the segment is $\lambda = \nicefrac{\widetilde{R}}{L}$
+and $U = 2\mu L$ and $M = r_\text{BP} L$ as above. In this section, we assume
+$\mu$ is the mutation rate to variants with heterozygous selection coefficient
+$s$ to simplify notation; later, $\mu$ refers to the genome-wide average
+mutation rate and the mutation rate is scaled by the DFE. Solving this system
+of two equations for each segment also gives the total substitution or
+``ratchet" rate per generation per segment ($\widetilde{R}$). The per-basepair
+substitution rate of a segment with mutation rate $\mu$ and selection
+coefficient $s$ is $\lambda_d(\mu, s) = \nicefrac{\widetilde{R}}{L}$
 
@@
 In practice, to fit such a model, we must discretize both $\mu(s, k(g))$ and
-$s$. Our models use a fixed $q$-element grid of selection coefficients, and
-mutation rates are on a fixed $z$-element grid, which is later interpolated
+$s$. Our models use a fixed $m_s$-element grid of selection coefficients, and
+mutation rates are on a fixed $m_\mu$-element grid, which is later interpolated
 over during the maximum likelihood optimization. The number of annotation
@@
 
-In the free-mutation parameterization, there are $1 + K \times q$ for $\pi_0$
+In the free-mutation parameterization, there are $1 + K \times m_s$ for $\pi_0$
 and the free mutation rate parameters, where $K$ is the number of annotation
@@
 class (here, as an example, CDS, UTRs, and phastcons), which are stored in the
-$q \times K$ matrix $\mathbf{M}_\text{F}$,
+$m_s \times K$ matrix $\mathbf{M}_\text{F}$,
 
@@
 
-which has $2 + K \times (q-1)$ free parameters, since each column must sum to
+which has $2 + K \times (m_s-1)$ free parameters, since each column must sum to
 one. This imposes the constraint that $\sum_i w_{i,k} = 1$, requiring
@@
 
+After rescaling each segment $N_e$ by the local predicted reduction
+$\hat{B}(x)$, the ratchet rates and $B'$ values are re-calculated.
+
+The rescaled $B'$ calculation outputs a $m_\mu \times m_s \times S$
+multidimensional array $\mathbf{R}$ of ratchet rates per segment. These are
+rescaled by the segment lengths $\mathbf{l}$, giving the per-basepair mutation
+rate array. The per-basepair ratchet rate given maximum likelihood estimates of
+$\hat{\mu}$ and $\widehat{\mathbf{W}}$ is used to predicted the ratchet rate
+for each segment $g$. The DFE estimate for feature $k$ is a column vector of
+$\widehat{\mathbf{W}}$, i.e. $\widehat{\mathbf{w}}_k$
+
 \begin{align}
-  R = \int 
+  \lambda_D(g) = 
 \end{align}
+
+The predicted ratchet rate for segment $g$ $\lambda_d(g)$ is based on these
+values.
 

### 2023-01-18 bgspy/likelihood.py
non-parametric method updates, outlier filtering for binning... etc

@@ -604,6 +604,26 @@ class BGSLikelihood:
     def nf(self):
         return self.dim()[2]
 
+    def predict_B_at_pos(self, chrom, pos, **kwargs):
+        """
+        Predict B from this fit, via interpolation.
+
+        TODO (low priority): sometimes outputs B > 1 due to
+        interpolation issues.
+        """
+        defaults = {'kind': 'quadratic',
+                    'assume_sorted': True,
+                    'bounds_error': False,
+                    'copy': False}
+        kwargs = {**defaults, **kwargs}
+        mids = self.bins.midpoints()
+        idx = self.bins.chrom_indices(chrom)
+        y = self.predict(B=True)
+        func = interpolate.interp1d(mids[chrom], y[idx],
+                                    fill_value=(y[0], y[-1]),
+                                    **kwargs)
+        return func(pos)
+
     def _load_optim(self, optim_res):
         """
         Taken an OptimResult() object.
@@ -1013,38 +1033,6 @@ class FreeMutationModel(BGSLikelihood):
             # rescale so B is returned, π0 = 1
             theta[0] = 1.
         return predict_freemutation(theta, self.logB, self.w)
-
-    def predict_B_at_pos(self, chrom, pos, **kwargs):
-        defaults = {'kind': 'quadratic',
-                    'assume_sorted': True,
-                    'bounds_error': False,
-                    'copy': False}
-        kwargs = {**defaults, **kwargs}
-        mids = self.bins.midpoints()
-        idx = self.bins.chrom_indices(chrom)
-        y = self.predict(B=True)
-        func = interpolate.interp1d(mids[chrom], y[idx],
-                                    fill_value=(y[0], y[-1]),
-                                    **kwargs)
-        return func(pos)
-
-
-    def predict_R(self, R, optim=None, theta=None):
-        """
-        """
-        if theta is not None:
-            return predict_freemutation(theta, self.logB, self.w)
-        if optim is None:
-            theta = self.theta_
-        else:
-            thetas = self.optim.thetas_
-            if optim == 'random':
-                theta = thetas[np.random.randint(0, thetas.shape[0]), :]
-            else:
-                theta = thetas[optim]
-        return predict_freemutation(theta, self.logB, self.w)
-
-
 
     def __repr__(self):
         base_rows = super().__repr__()

### 2023-01-18 bgspy/nonparametric.py

@@ -4,6 +4,7 @@ import warnings
 from sklearn.model_selection import KFold, LeaveOneOut
 from collections import defaultdict
 from bgspy.utils import binned_statistic, cutbins
+from bgspy.plots import get_figax
 
 def gaussian_kernel(x):
     # for underflow
@@ -83,4 +84,13 @@ def bin_kfolds(x, y, bins=np.arange(5, 100, 2), n_splits=100, **bin_args):
 
     msedat = {nb: np.mean(v) for nb, v in mses.items()}
     return list(msedat.keys()), list(msedat.values())
+
+def kfolds_results(kf_res, figax=None):
+    fig, ax = get_figax(figax)
+    ax.plot(*kf_res)
+    b, mse = kf_res
+    best_idx = np.argmin(mse)
+    best = b[best_idx]
+    ax.scatter(b[best_idx], mse[best_idx])
+    return b[best_idx]
 

### 2023-01-18 bgspy/utils.py

@@ -1171,9 +1171,22 @@ def binned_summaries(x, y, nbins, method='interval',
                      funs={'mean': np.nanmean,
                            'median': np.nanmedian,
                            'sd': np.nanstd,
-                           'n': lambda x: np.sum(np.isfinite(x))}):
+                           'n': lambda x: np.sum(np.isfinite(x))},
+                     cut_tail=None):
     """
+
+    cut_tail: if None, no modifications. If this is a tuple of probabilities,
+               this will use censor() to remove x and y values that fall outside
+               the tails.
+
+
+    Notes: I find cut_tail is often very important for interval-based binning,
+    as the bin-width is based on the total range, and thus very sensitive
+    to outliers. I think cut_tail should be used in most of these cases.
     """
+    if cut_tail is not None:
+        idx = censor(x, cut_tail, return_idx=True)
+        x, y = x[idx], y[idx]
     bins = cutbins(x, nbins, method)
     cols = defaultdict(list)
     cols['start'] = bins[1:]
@@ -1195,11 +1208,14 @@ def corr(x, y):
     keep = np.isfinite(x) & np.isfinite(y)
     return pearsonr(x[keep], y[keep]), spearmanr(x[keep], y[keep])
 
-def censor(x, probs):
+def censor(x, probs, return_idx=False):
     """
     Trim of tails
     """
     assert len(probs) == 2
     l, u = np.quantile(x, probs)
-    return x[(x >= l) & (x <= u)]
+    idx = (x >= l) & (x <= u)
+    if return_idx:
+        return idx
+    return x[idx]
 

### 2023-01-18 bgspy/utils.py
latenight bug fixes

@@ -1172,20 +1172,20 @@ def binned_summaries(x, y, nbins, method='interval',
                            'median': np.nanmedian,
                            'sd': np.nanstd,
                            'n': lambda x: np.sum(np.isfinite(x))},
-                     cut_tail=None):
+                     cut_tails=None):
     """
 
-    cut_tail: if None, no modifications. If this is a tuple of probabilities,
+    cut_tails: if None, no modifications. If this is a tuple of probabilities,
                this will use censor() to remove x and y values that fall outside
                the tails.
 
 
-    Notes: I find cut_tail is often very important for interval-based binning,
+    Notes: I find cut_tails is often very important for interval-based binning,
     as the bin-width is based on the total range, and thus very sensitive
-    to outliers. I think cut_tail should be used in most of these cases.
+    to outliers. I think cut_tails should be used in most of these cases.
     """
-    if cut_tail is not None:
-        idx = censor(x, cut_tail, return_idx=True)
+    if cut_tails is not None:
+        idx = censor(x, cut_tails, return_idx=True)
         x, y = x[idx], y[idx]
     bins = cutbins(x, nbins, method)
     cols = defaultdict(list)

### 2023-01-28 manuscript/manuscript.tex
ms updates

@@
 
+\begin{abstract}
+
+Previous work in population genomics has established that negative selection
+against against new deleterious variants introduced by mutation profoundly
+shapes large-scale patterns of variation across the genome. However, the
+mathematical models that describe this process only consider variants so
+deleterious that they cannot drift up in frequency and ultimately fix. In
+reality, there number of mildly deleterious mutations
+
+\end{abstract}
+
 \section*{Introduction}
@@
   \label{eq:}
-  {N}_{e}(x) &= \exp \left( -\frac{V(x)}{2} Q^2(x) \right) & \text{\emph{effective population equation}} \\
-  {R}(x) &= \frac{4NU s}{\exp(4 N s)-1}  & \text{\emph{ratchet equation}} 
+  {N}_{e}(x) &= N \exp \left( -\frac{V(x)}{2} Q^2(x) \right) & \text{\emph{effective population equation}} \\
+  {R}(x) &= \frac{4N_e(x)U s}{\exp(4 N_e(x) s)-1}  & \text{\emph{ratchet equation}} 
 \end{align}
@@
 \begin{align}
-  V(x) &= U s - 2 {R}(x) s & \text{\emph{variation equation}} \\
-  Q^2(x) &= \frac{2}{(1-Z)(2-(2-M)Z)} & \text{\emph{linkage persistence factor}} \\
+  V(x) &= U s - 2 {R}(x) s & \text{\emph{fitness variance equation}} \\
+  Q^2(x) &= \frac{2}{(1-Z)(2-(2-M)Z)} & \text{\emph{linkage inflation factor}} \\
   Z &= 1 - \frac{Us}{U-{R(x)}} & \text{\emph{}} \text{\emph{variance decay rate}}

### 2023-01-28 bgspy/models.py
figure changes, etc

@@ -25,37 +25,33 @@ group. The set a_B(i_B) is the collection of sites in with membership in
 annotation set i_B.
 """
 
-from collections import defaultdict, namedtuple, Counter
-import multiprocessing
-import os
 import pickle
 import warnings
-import itertools
-import tqdm
-import time
 import numpy as np
-from scipy.optimize import minimize_scalar
 import scipy.stats as stats
-import pandas as pd
 
-from bgspy.utils import signif
-from bgspy.utils import Bdtype, BScores, BinnedStat, bin_chrom
+from bgspy.utils import Bdtype, BScores, bin_chrom
 from bgspy.parallel import calc_B_parallel, calc_BSC16_parallel
 from bgspy.substitution import ratchet_df
+
 
 class BGSModel(object):
     """
     BGSModel contains the the segments under purifying selection
     to compute the B and B' values.
     """
+
     def __init__(self, genome, t_grid=None, w_grid=None, split_length=None):
         # main genome data needed to calculate B
         self.genome = genome
         assert self.genome.is_complete(), "genome is missing data!"
-        diff_split_lengths = genome.split_length is not None and split_length != genome.split_length
+        diff_split_lengths = (genome.split_length is not None and
+                              split_length != genome.split_length)
         if self.genome.segments is None or diff_split_lengths:
             if diff_split_lengths:
-                warnings.warn("supplied Genome object has segment split lengths that differ from that specified -- resegmenting")
+                warnings.warn("supplied Genome object has segment split "
+                              "lengths that differ from that specified "
+                              "-- resegmenting")
             self.genome.create_segments(split_length=split_length)
         # stuff for B
         self.Bs = None
@@ -70,7 +66,8 @@ class BGSModel(object):
 
     @property
     def features(self):
-        features = [self.segments.inverse_feature_map[i] for i in range(self.nf)]
+        features = [self.segments.inverse_feature_map[i] for
+                    i in range(self.nf)]
         return features
 
     @property
@@ -102,7 +99,8 @@ class BGSModel(object):
 
     @property
     def BScores(self):
-        return BScores(self.Bs, self.B_pos, self.w, self.t, self.features, self.step)
+        return BScores(self.Bs, self.B_pos, self.w,
+                       self.t, self.features, self.step)
 
     @property
     def BpScores(self):
@@ -117,8 +115,9 @@ class BGSModel(object):
         if self.Bs is None or self.B_pos is None:
[…]

### 2023-01-28 bgspy/plots.py

@@ -1,29 +1,30 @@
 import numpy as np
-from math import ceil
-import itertools
-import pandas as pd
 import matplotlib.pyplot as plt
-from matplotlib import cm
 import statsmodels.api as sm
-import scipy.stats as stats
-from bgspy.theory import B_var_limit
-from bgspy.utils import signif, mean_ratio
+from bgspy.utils import mean_ratio
+import matplotlib as mpl
+lowess = sm.nonparametric.lowess
 
-# from https://writing.stackexchange.com/questions/21658/what-is-the-image-size-in-scientific-paper-if-indicated-as-a-single-1-5-or-2-c 
-# in mm
-mm_to_inches = lambda x: 0.0393701 * x
+#  ---- some figure size preset for the paper
 
-img_size = dict(one = mm_to_inches(90), onehalf = mm_to_inches(140), two = mm_to_inches(190))
 
-asp_ratio = dict(golden = (1 + np.sqrt(5))/2, one=1, two=2)
+def mm_to_inches(x):
+    # from https://writing.stackexchange.com/questions/21658/what-is-the-image-size-in-scientific-paper-if-indicated-as-a-single-1-5-or-2-c 
+    return 0.0393701 * x
 
-sizes = {(k, ar): np.round((v, v/asp_ratio[ar]), 4) for k, v in img_size.items() for ar in asp_ratio}
+
+img_size = dict(one=mm_to_inches(90),
+                onehalf=mm_to_inches(140),
+                two=mm_to_inches(190))
+
+asp_ratio = dict(golden=(1 + np.sqrt(5))/2, one=1, two=2)
+sizes = {(k, ar): np.round((v, v/asp_ratio[ar]), 4) for
+         k, v in img_size.items() for ar in asp_ratio}
+
 
 def center_scale(x):
     return (x-x.mean())/x.std()
 
-
-lowess = sm.nonparametric.lowess
 
 def get_figax(figax, **kwargs):
     if figax is None:
@@ -31,6 +32,33 @@ def get_figax(figax, **kwargs):
     else:
         fig, ax = figax
     return fig, ax
+
+
+def smooth(x, y, frac=None):
+    """
+    Smooth a line for visual clarity.
+    """
+    if frac is None:
+        return x, y
+    sx, sy = lowess(y, x, frac=frac).T
+    return sx, sy
+
+
+def chromosome_ticks_formatter(scale):
+    """
[…]

### 2023-01-28 bgspy/utils.py

@@ -1133,12 +1133,15 @@ def bin2midpoints(x):
     return 0.5*(x.bin_edges[1:] + x.bin_edges[:-1])
 
 
-def bin2pairs(x):
+def bin2pairs(x, use_mean_ratio=False):
     """
     Convenience function to take a BinnedStatistics object and return the
     binned midpoints and statistic together as a pair.
     """
-    return bin2midpoints(x), x.statistic
+    y = x.statistic
+    if use_mean_ratio:
+        y = mean_ratio(y)
+    return bin2midpoints(x), y
 
 def logbins(x, nbins, density=True, remove_nan=True):
     """

### 2023-01-31 manuscript/manuscript.tex
minor stuff
ms changes

@@
 deleterious that they cannot drift up in frequency and ultimately fix. In
-reality, there number of mildly deleterious mutations
+reality, some deleterious variants can fix if the randomness of the
+evolutionary process overpowers selection. There are two primary sources of
+such randomness in the evolutionary process, genetic drift and genetic draft,
+but both are ignored by current mutation-selection models. Here, we introduce a
+new whole-genome general mutation-selection-drift model that extends and
+implements previous theory by \textcite{Santiago2016-mu}. This model estimates
+the strength of selection at specified sites by modeling large-scale
+distortions in chromosomal patterns of genetic variability across the genome.
+Unlike previous models, our models can estimate the distribution of 
+fitness effects at conserved sites. Additionally, because our 
+model incorporates drift and draft, our model makes explicit predictions 
+about the substitution rate of weakly deleterious mutations and allows us to 
+test XXX.
+
+A full evolutionary model of deleterious mutation must not only describe the
+mutation rate and distribution of fitness effects, but also the substitution
+rate for mildly deleterious mutations. Our model is fit to genome wide
+patterns of genetic diversity across the genome, estimating the 
+mutation rate, neutral diversity level, and the distribution of fitness effects. 
+Additionally, because our model allows for substitution, our model makes 
+predictions about the substitution rate in conserved features.
+
+Previous work has established that selection against strongly deleterious
+mutations across the genome profoundly shapes large-scale patterns of genetic
+variability in the human genome. However,
+
+selection against new deleterious mtuationss
+
+In reality, natural selection in some 
+parts of the genome effectively speeds up evolution, which
+
+In
+reality, the number of mildly deleterious mutations
+
+increases the pace of ev
+
+Here extend the theory of \textcite{Santiago2016-mu} to work across 
+whole-genomes. This model estimates the reduction in genetic diversity 
+from mutation-selection-drift processes
 
@@
 \section*{Introduction}
+
+emphasis on whole genome models
 
@@
 \subsection*{Evolutionary Models}
+
+We model the diversity at punitively neutral sites. These are sites in regions
+we have \emph{a prior} reason to believe have a DFE nearly entirely
+concentrated in the neutral range. While this may not be the case in reality,
+it is necessary given the computational demands of modeling whole-genome
+selection.
 

### 2023-01-31 bgspy/likelihood.py
dfe plot, etc

@@ -26,7 +26,7 @@ from bgspy.utils import signif, load_seqlens
 from bgspy.data import pi_from_pairwise_summaries, GenomicBinnedData
 from bgspy.optim import run_optims, nlopt_mutation_worker, nlopt_simplex_worker
 from bgspy.plots import model_diagnostic_plots, predict_chrom_plot
-from bgspy.plots import resid_fitted_plot
+from bgspy.plots import resid_fitted_plot, get_figax
 from bgspy.bootstrap import process_bootstraps, pivot_ci
 from bgspy.models import BGSModel
 from bgspy.sim_utils import mutate_simulated_tree
@@ -832,7 +832,25 @@ class BGSLikelihood:
                 fpath = os.path.join(loo_fits_dir, f"mle_loo_{chrom}.pkl")
                 new_fit.save(fpath)
         return np.array(r2s)
-           
+
+    def dfe_plot(self, figax=None):
+        """
+        TODO
+        """
+        fig, ax = get_figax(figax)
+        xt = np.log10(self.t)
+
+        nf = self.nf
+        pad = 0.03
+        w = 1/nf - pad  # width of bars, for each feature with allowance
+        hw = w/2
+        for i in range(nf):
+            feat = self.features[i]
+            ax.bar(xt - 1/(nf/2) + i/nf, self.mle_W[:, i], align='edge', width=w, label=feat)
+        ax.set_xticks(np.log10(self.t), [f"$10^{{{int(x)}}}$" for x in xt])
+        ax.set_ylabel('probability')
+        ax.legend()
+
     def R2(self, _idx=None, **kwargs):
         """
         The R² value of the predictions against actual results.

### 2023-02-08 bgspy/likelihood.py
new super good mle

@@ -26,6 +26,7 @@ from bgspy.data import GenomeData
 from bgspy.utils import signif, load_seqlens, load_pickle
 from bgspy.data import pi_from_pairwise_summaries, GenomicBinnedData
 from bgspy.optim import run_optims, nlopt_mutation_worker, nlopt_simplex_worker, nlopt_softmax_worker
+from bgspy.optim import scipy_softmax_worker
 from bgspy.plots import model_diagnostic_plots, predict_chrom_plot
 from bgspy.plots import resid_fitted_plot, get_figax
 from bgspy.plots import chrom_resid_plot
@@ -674,7 +675,7 @@ def check_bounds(x, lb, ub):
     assert np.all((x >= lb) & (x <= ub))
 
 def negll_c(theta, Y, logB, w, two_alleles=False,
-            version2=True):
+            version=2):
     """
     θ is [π0, μ, w11, w12, ...] and should
     have dimension (nt x nf) + 2
@@ -706,14 +707,21 @@ def negll_c(theta, Y, logB, w, two_alleles=False,
     likclib.negloglik2.argtypes = negloglik_argtypes
     likclib.negloglik2.restype = negloglik_restype
 
+    likclib.negloglik3.argtypes = negloglik_argtypes
+    likclib.negloglik3.restype = negloglik_restype
+
     args = (theta_ptr, nS_ptr, nD_ptr, logB_ptr, w_ptr,
             logB.ctypes.shape, logB.ctypes.strides,
             int(two_alleles))
 
-    if not version2:
+    if version == 1:
         return likclib.negloglik(*args)
-    else:
+    elif version == 2:
         return likclib.negloglik2(*args)
+    elif version == 3:
+        return likclib.negloglik3(*args)
+    else:
+        raise ValueError("unsupported version")
 
 
 def predict_simplex(theta, logB, w, mu=None):
@@ -785,7 +793,7 @@ class BGSLikelihood:
 
         try:
             assert logB.ndim == 4
-            assert logB.shape[1] == w.size
+            #assert logB.shape[1] == w.size # TODO FIX
             assert logB.shape[2] == t.size
         except AssertionError:
             msg = "B dimensions ({logB.shape}) do not match (supplied w and t dimensions)"
@@ -1126,7 +1134,7 @@ def negll_freemut_full(theta, grad, Y, B, w, two_alleles=False):
     return negll_c(new_theta, Y, B, w, two_alleles)
 
 
[…]

### 2023-02-08 bgspy/optim.py

@@ -11,11 +11,12 @@ from tabulate import tabulate
 import nlopt
 from scipy.optimize import minimize
 
-# nlopt has different error messages; we use this for scipy to
-# but just force to 'success' (1) or 'failure' (-1)
+# nlopt has different error messages; we use this for scipy too
+# which uses booleans we convert to integers
 NL_OPT_CODES = {1:'success', 2:'stopval reached', 3:'ftol reached',
                 4:'xtol reached', 5:'max eval', 6:'maxtime reached', -1:'failure',
-                -2:'invalid args', -3:'out of memory', -4:'forced stop'}
+               -2:'invalid args', -3:'out of memory', -4:'forced stop',
+                0: 'scipy failure'}
 
 
 def extract_opt_info(x):
@@ -35,7 +36,8 @@ def extract_opt_info(x):
 def array_all(x):
     return tuple([np.array(a) for a in x])
 
-def run_optims(workerfunc, starts, progress=True, ncores=50):
+def run_optims(workerfunc, starts, progress=True, ncores=50,
+               return_raw=False):
     """
     Parallel optimization.
 
@@ -60,6 +62,8 @@ def run_optims(workerfunc, starts, progress=True, ncores=50):
             res = list(map(workerfunc, starts))
 
     nlls, thetas, success = array_all(zip(*map(extract_opt_info, res)))
+    if return_raw:
+        return nlls, thetas, success
     return OptimResult(nlls, thetas, success, np.array(starts))
 
 
@@ -230,9 +234,12 @@ def nlopt_softmax_worker(start, func, nt, nf,
     success = opt.last_optimize_result()
     return nll, mle, success
 
+def scipy_softmax_worker(start, func, nt, nf):
+    res = minimize(func, start, method='L-BFGS-B')
+    return res.fun, res.x, res.success
 
 
-def nlopt_simplex_worker(start, func, nt, nf, bounds, 
+def nlopt_simplex_worker(start, func, nt, nf, bounds,
                          log10_W_bounds, mu=None,
                          constraint_tol=1e-3, xtol_rel=1e-3,
                          maxeval=1000000, algo='ISRES'):
@@ -354,25 +361,28 @@ def optim_diagnotics_plot(fit, top_n=100):
 
     
     for i in range(nf):
-        ax[i].imshow(dfes[:, :, i-1].T, cmap='inferno')
[…]

### 2023-02-08 bgspy/src/likelihood.c

@@ -172,7 +172,7 @@ double negloglik(const double *theta,
                 logBw_i += Binc;
             }
         }
-
+        //printf("%g, ", logBw[i]);
         double log_pi;
         if (two_alleles) {
             double pi = pi0 * exp(logBw_i);
@@ -184,13 +184,8 @@ double negloglik(const double *theta,
         //printf("c log(pi0): %g\n", log(pi0));
         //printf("c nD[%d]=%g, nS[%d]=%g\n", i, nD[i], i, nS[i]);
         //printf("c llm[%d]: %g\n", i, nD[i]*log_pi + nS[i]*log(1 - exp(log_pi)));
-        
-        // compute the final likelihood for this window
-        double obs_pi = nD[i] / (nS[i] + nD[i]);
-        double pi = exp(log_pi);
-        //ll += pow(obs_pi - pi, 2.);
-        ll += pow(obs_pi - pi, 2.);
-    }
+        ll += nD[i]*log_pi + nS[i]*log1p(-exp(log_pi));
+    }                
     free(W);
     fflush(stdout);
     return -ll;
@@ -302,6 +297,63 @@ double negloglik2(const double *theta,
     }
 
     free(w_idx);
+    free(W);
+    free(strides);
+    fflush(stdout);
+    return -ll;
+}
+
+
+double negloglik3(const double *theta,
+                  const double *nS, const double *nD, 
+                  const double *logB, 
+                  const double *w,
+                  const ssize_t *logB_dim, 
+                  const ssize_t *logB_strides,
+                  const int two_alleles) {
+
+    ssize_t nx = logB_dim[0]; 
+    ssize_t nw = logB_dim[1];
+    assert(nw == 1);
+    ssize_t nt = logB_dim[2];
+    ssize_t nf = logB_dim[3];
+
+    double pi0 = theta[0];
+    double mu = theta[1];
+    ssize_t nW = nt*nf;
+    double *W = calloc(nW, sizeof(double));
+    memcpy(W, theta + 2, nW * sizeof(double));
+    double logBw_i;
+    double Wjk;
+    double ll = 0;
+    double log_pi;
+
+    // variables we need for interpolation in this loop
+    ssize_t *strides = malloc(4 * sizeof(ssize_t));
+    for (int i=0; i<4; i++) strides[i] = logB_strides[i] / sizeof(double);
+ 
+    // Iterate through all windows, summing B values across features and
+    // selection coefficients per feature
+    //
[…]

### 2023-02-08 tests/test_likelihood.py

@@ -138,8 +138,8 @@ def test_new_likelihood_versus_old():
     for _ in range(20):
         theta_jitter = np.random.normal(0, 1e-6) + theta
         theta_jitter[:1] = theta[:1]
-        old_results = negll_c(theta_jitter, Y, B, w, version2=False)
-        new_results = negll_c(theta_jitter, Y, B, w, version2=True)
+        old_results = negll_c(theta_jitter, Y, B, w, version=1)
+        new_results = negll_c(theta_jitter, Y, B, w, version=2)
         if not np.isnan(old_results):
             # we ignore different NAN behavior
             assert(old_results == new_results), theta_jitter

### 2023-02-08 manuscript/manuscript.tex
ms updates

--- 2023-01-28 manuscript/manuscript.tex
@@
 shapes large-scale patterns of variation across the genome. However, the
-mathematical models that describe this process only consider variants so
-deleterious that they cannot drift up in frequency and ultimately fix. In
-reality, there number of mildly deleterious mutations
+mathematical models that describe this process only consider variation so
+deleterious that it cannot drift up in frequency and ultimately fix in the
+population. In reality, many new mutations are only mildly deleterious, and can
+fix in the population if random perturbations from genetic drift or selection
+at neighboring sites (genetic draft) overpowers the selection against them.
+Despite the large role negative selection plays across the human genome, no
+existing evolutionary models consider the joint effects of mutation, selection,
+and genetic drift and drift across the genome. Consequently, population genetic
+models cannot describe the ultimate fate of every new variant ---eventual
+extinction or fixation in the population--- and thus cannot connect to observed
+rates of molecular evolution.
+
+whole-genome variance models.
+
+This is because, in the very long run, all mutations must be accounted for.
+
+Here, we introduce a new whole-genome general mutation-selection-draft-drift
+(MSD$^2$) model that extends and implements previous theory by
+\textcite{Santiago2016-mu}.
+
+Unlike classic mutation-selection models which cannot accurately capture the
+dynamics of weakly deleterious mutation, MSD$^2$ models can, and consequently
+can provide estimates of the distribution of selection coefficients of new
+mutations. Additionally, because our model incorporates drift and draft, our
+model makes explicit predictions about the substitution rate of weakly
+deleterious mutations and allows us to test XXX.
 
@@
 \section*{Introduction}
+
+If the human evolutionary process were strongly impacted by Hill--Robertson
+Interference, we should expect that substitution rates are highest in regions
+experiencing the most selection, which are regions when $B(x)$ is the lowest.
+However, our model
+
+emphasis on whole genome models
 
@@
 
+At every spot in the genome the processes of mutation-selection-draft-drift
+against alleles deleterious in most environments reduce pairwise diversity. We
+can imagine some unobserved reduction map, $B(x)$ across the genome. Here we
+derive theory for this map under these processes and show how statistical
+models can be fit using this theory.
+
 To understand the this process and make predictions about the equilibrium level
@@
 
-\begin{figure}[!htb]
-  \centering
-  \includegraphics{figures/experimental_1.pdf}
-  \caption{}
-  \label{fig:figure-1}
-\end{figure}
+%\begin{figure}[!htb]
+%  \centering
+%  \includegraphics{figures/experimental_1.pdf}
+%  \caption{}
+%  \label{fig:figure-1}
+%\end{figure}
 
 \subsection*{Evolutionary Models}
+
+We model the diversity at punitively neutral sites. These are sites in regions
+we have \emph{a prior} reason to believe have a DFE nearly entirely
+concentrated in the neutral range. While this may not be the case in reality,
+it is necessary given the computational demands of modeling whole-genome
+selection.
 
@@
 
+\section*{Discussion}
+
+It is worth considering why this model performs well for models of negative
+selection. The fitness variance equation, $V(x) = (U-2R)s$ tells us that under
+MSD$^$2 models, the long-run variance is set by the balance of mutation and
+fixation. Whether a particular basepair's behavior follows this equation
+depends upon whether the long-run average is close to ``typical" for that
+basepair at a given time. Good things fix so quickly, that we're left with a
+constant churning of new variation
+
 \section*{Methods}
@@
   {N}_{e}(x) &= N \exp \left( -\frac{V(x)}{2} Q^2(x) \right) & \text{\emph{effective population equation}} \\
-  {R}(x) &= \frac{4N_e(x)U s}{\exp(4 N_e(x) s)-1}  & \text{\emph{ratchet equation}} 
+  {R}(x) &= \frac{4N_e(x)U s}{\exp(4 N_e(x) s)-1}  & \text{\emph{substitution equation}} 
 \end{align}
@@
 \begin{align}
-  V(x) &= U s - 2 {R}(x) s & \text{\emph{fitness variance equation}} \\
+  %V(x) &= U s - 2 {R}(x) s & \text{\emph{fitness variance equation}} \\
+  V(x) &= (U  - 2 {R}(x)) s & \text{\emph{fitness variance equation}} \\
   Q^2(x) &= \frac{2}{(1-Z)(2-(2-M)Z)} & \text{\emph{linkage inflation factor}} \\
@@
 local reduction factor is $b(x) = \nicefrac{N_e(x)}{N}$, and the per-basepair
-per-generation substitution rate is $\lambda_D = \nicefrac{R(x)}{L}$.
+per-generation substitution rate is $\lambda_D = \nicefrac{R(x)}{L}$. Note that
+in the variance equation, the long-run variance is the balance between mutation
+and twice the loss due to substitution, $(U-2R)s$, since for selective balance
+$V(x) = 0$ for $s>0$, the per-basepair substitution rate must be half the mutation rate,
+since each substitution removes twice the variance introduced by a mutation in
+a diploid species.
 
@@
 feature type (e.g. CDS, UTRs, etc.). This results in estimates of $N_e(x)$ and
-$R(x)$ per segment. In the first-round B' calculation, we use the
-drift-effective population size, $N$.
+$R(x)$ per segment. In the first-round B' calculation, we use a single
+pre-specified drift-effective population size, $N$. The B' maps created here at
+this step reflect selection only in the segment, ignoring the impact of
+selection at other segments. By contrast, the locally-rescaled B' maps solve
+these equations, assuming the drift-effective population size is $B(x)N$. These
+reflect the segment's dynamics if they were effected by selection at other
+segments under selection in the genome. Consequently, we can compare and
+contrast models with local-recalling to test whether we achieve a better fit if
+we account selective interference from other segments.
 
@@
 
+\subsection*{Assessing the B' Maps}
+
+We first assessed the B' maps against the observed average reduction across
+$XXX$ forward-simulations, $\bar{B}_\text{sim}(x)$. The mean-squared error has
+two sources, (1) bias from a breakdown in our theory $B(x)$ for the average
+reduction, and (2) variance from the simulations. We show that for most of the
+parameter space, the measured mean-squared error between simulation averages
+and our B' maps is at the theoretic minimum. This minimum variance is entirely
+from the coalescent noise when estimating $\bar{B}\text{sim}(x), which
+
 \subsection*{Maximum Likelihood Estimation}
@@
 \subsection*{Local Rescaling and Substitution Rate Prediction}
+
+\subsection*{Substitution Rate Predictions}
+
+Our MLE fit $\hat{\Phi}$ also implies a set of corresponding substitution rate
+predictions for each segment under mutation-selection-draft process. There are
+two sources of variation in these predictions in this model. First, there is
+the variation across the pre-specified functional categories, e.g. CDS, UTRs,
+and PhastCons. Second, there is region-specific variation
 
@@
 \begin{align}
-  T \approx \frac{\exp(4 N_e s)-1}{2 U s N_e}.
+  T \approx \frac{\exp(4 N_e s)-1}{2 U s N_e}
 \end{align}
@@
 Since the ratchet rate depends on $N_e$, which in turn depends on the rate of
-the ratchet, we write the inverse as,
-
-\begin{align}
-  R(N_e) := \frac{2 U s N_e}{\exp(4 N_e s)-1}
+the ratchet, we write the ratchet, or substitution, rate per basepair per
+generation as,
+
+\begin{align}
+  R(N_e) := \frac{2 U s N_e}{\exp(4 N_e s)-1}.
 \end{align}
@@
 
-The solution to these equations is the asymptotic equilibrium $\widetilde{N}_e$ and
-equilibrium ratchet rate $\widetilde{R}$. This set the equilibrium variance,
-$\widetilde{V} = Us - 2\widetilde{R}s$.
+The solution to these equations is the asymptotic equilibrium $\widetilde{N}_e$
+and equilibrium ratchet rate $\widetilde{R}$. This sets the equilibrium
+variance, $\widetilde{V} = Us - 2\widetilde{R}s$, or $\widetilde{V} = (U -
+2\widetilde{R})s$. Note that because $p_\text{fix} \le \nicefrac{1}{2N}$ for
+deleterious to neutral variation, $2R \le U$, or the diploid mutation rate must
+be greater than twice the per-basepair per-generation substitution rate for
+equilibrium variance $\widetilde{V} > 0$ for $s>0$.
 

### 2023-02-08 manuscript/supplementary.tex
ms updates

@@
 
-  \item \textbf{Reference masking}: soft and hard-masked regions in the human GRCh38
-    reference genome are also masked.
+  \item \textbf{Reference masking}: soft and hard-masked regions in the human
+      GRCh38 reference genome are also masked. Soft-masked regions were
+      determined by Ensembl (XXX), which uses Repeat Masker (XXX) and Dust
+      (XXX).
 
@@
 These files are all produced by the Snakemake file \path{data/annotation/Snakefile}.
+Note that the levels of accessible putatively neutral bases
 
@@
 data $\mathbf{Y}$ across sites, so the likelihood across bins is naturally
-weighted by the quantity of observed data. We estimate $h$ by fitting the model
-for a variety of window sizes $h$ and select the best using out-sample cross
-validation, as is standard in non-parametric regression (XXX, Wasserman, p.
-70).
+weighted by the quantity of observed data.
+
+% TODO
+%We estimate $h$ by fitting the model
+%for a variety of window sizes $h$ and select the best using out-sample cross
+%validation, as is standard in non-parametric regression (XXX, Wasserman, p.
+%70).
 
@@
   \label{suppeqn:b}
-  B(x, \Psi) = \prod_g^S \int_0^1 b(\mu(s, k(g)), s, L_g, r_g, d(x, g)) ds
+  B(x, \Psi) &= \prod_g^S \exp \left( \int_0^1 b(\mu(s, k(g)), s, L_g, r_g, d(x, g)) ds \right) \\
+             &= \exp \left( \sum_g^S \int_0^1 b(\mu(s, k(g)), s, L_g, r_g, d(x, g)) ds \right)
 \end{align}
 
-where, $b(\mu(s, k(g)), s, L_g, r_g, d(x,g))$ is the reduction due to segment
-$g$, and there are $S$ total segments. Here, $\mu(s, k(g))$ is the per-basepair
-per-genome rate that mutations with selection coefficient $s$ enter segments
-with feature class $k(g)$, $L_g$ and $r_g$ are the length in basepairs and
-recombination rate per basepair of segment $g$, and the recombination distance
-between focal site $x$ and segment $g$ is $d(x, g)$. The functional form of $b$
-varies depending on whether the classic BGS model is used, or our form of
-Santiago and Caballero's (\citeyear{Santiago2016-mu}) equations (see Section
-XXX).
+where, $b(\mu(s, k(g)), s, L_g, r_g, d(x,g))$ is the log reduction due to
+segment $g$, and there are $S$ total segments. Here, $\mu(s, k(g))$ is the
+per-basepair per-genome rate that mutations with selection coefficient $s$
+enter segments with feature class $k(g)$, $L_g$ and $r_g$ are the length in
+basepairs and recombination rate per basepair of segment $g$, and the
+recombination distance between focal site $x$ and segment $g$ is $d(x, g)$. The
+functional form of $b$ varies depending on whether the classic BGS model is
+used, or our form of Santiago and Caballero's (\citeyear{Santiago2016-mu})
+equations (see Section XXX).
 
@@
 \begin{align}
-  b(\mu, s, L, r, 0) &= \exp \left( - \int_0^L \frac{2\mu(l)}{s (1 + (1-s) r(l)/s)^2} dl \right) 
+  b(\mu, s, L, r, 0) &=  - \int_0^L \frac{2\mu(l)}{s (1 + (1-s) r(l)/s)^2} dl 
 \end{align}
@@
 \begin{align}
-  b(\mu, s, L, r_\text{BP}, 0) &= \exp \left( - \frac{2\mu L}{s + (1-s) r_\text{BP} L} \right)
+  b(\mu, s, L, r_\text{BP}, 0) &= - \frac{2\mu L}{s + (1-s) r_\text{BP} L} 
 \end{align}
@@
 \begin{align}
-  b(\mu, s, L, r_\text{BP}, d) &= \exp \left( \frac{-2\mu L}{(d(1-s) + s)(s + (d + r_\text{BP}L)(1-s))} \right)
+  b(\mu, s, L, r_\text{BP}, d) &= \frac{-2\mu L}{(d(1-s) + s)(s + (d + r_\text{BP}L)(1-s))}
 \end{align}
@@
 \begin{align}
-  \widetilde{b}'(\mu, s, L, r_\text{BP}) = \frac{\widetilde{N}_{e,\infty}}{N}
+  \widetilde{b}'(\mu, s, L, r_\text{BP}) = \log\left(\frac{\widetilde{N}_{e,\infty}}{N} \right)
 \end{align}
@@
 \begin{align}
-  {N}_{e,\infty} &= N \exp \left( -\frac{V}{2} Q_\infty^2 \right) \\
-  {R} &= \frac{4 U s N_{e,\infty}}{\exp(4 N_{e,\infty} s)-1}\\
+    {N}_{e,\infty} &= N \exp \left( -\frac{V}{2} Q_\infty^2 \right) \label{eqn:Ne_VQ2} \\
+  {R} &= \frac{4 U s N_{e,\infty}}{\exp(4 N_{e,\infty} s)-1} \nonumber \\
 \end{align}
@@
 the solution to the system of two equations, giving us $V$ and $Q_\infty^2$,
-and then use these values in the equation,
+and then use these values in the equation \eqref{eqn:Ne_VQ2},
 
 \begin{align} 
-  b'(\mu, s, L, r, d) = \exp\left( -\frac{V}{2} \left(\frac{1}{(1-(1-d)Z}\right)^2 \right) 
+    b'(\mu, s, L, r, d) = -\frac{V}{2} \left(\frac{1}{(1-(1-d)Z}\right)^2 
 \end{align}
 
-where $Z = 1 - Us / (U-\widetilde{R})$ and $V = Us - 2\widetilde{R}s$.
+where $Z = 1 - Us / (U-\widetilde{R})$ and $V = Us - 2\widetilde{R}s$. TODO
+check d.
 
@@
   \label{suppeqn:b}
-  B(x, \Psi) = \prod_g \sum_{i} b(\mu_{i, \;k(g)}, s_i, L_g, r_g, d(x, g))
+  B(x, \Psi) = \exp \left( \sum_g^S \sum_{i} b\left( \mu_{i, \;k(g)}, s_i, L_g, r_g, d(x, g) \right)\right)
 \end{align}
+
+%TODO double check
 

### 2023-02-18 bgspy/genome.py
few nan related bugs fixed, new asserts added

@@ -112,6 +112,7 @@ class Segments:
         """
         L = self.lengths
         rbp = self.rates
+        assert np.all(~np.isnan(rbp))
         if t.ndim == 1:
             t = t[:, None]
         print(f"calculating classic B components...\t", end='', flush=True)
@@ -351,7 +352,9 @@ def process_annotation(features, recmap, split_length=None):
     map_pos = np.concatenate(map_pos, axis=0)
     assert(map_pos.shape[0] == ranges.shape[0])
     print(f"done.")
+    assert np.sum(np.isnan(map_pos)) == 0, "some interpolatd map positions are nan!"
     rates = np.array(split_rates, dtype='float32')
+    assert np.sum(np.isnan(split_rates)) == 0, "some segment recombination rates are nan!"
     features = np.array([feature_map[x] for x in split_features])
     return Segments(ranges, rates, map_pos, features, feature_map, index)
 

### 2023-02-18 bgspy/recmap.py

@@ -117,7 +117,7 @@ class RecMap(object):
             ends.insert(0, 0)  # add in the left-most position
             # the hapmap format format is each line is rate between it and *next*
             # so popping in a zero means we don't know the rate -- fill with nan
-            rec_rates.insert(0, np.nan)
+            rec_rates.insert(0, 0)
             # if we don't go to end, we add that in -- note that the last
             # rate we have is to end.
             if ends[-1] < self.seqlens[chrom]:
@@ -126,14 +126,14 @@ class RecMap(object):
             rec_rates = self.conversion_factor*np.array(rec_rates)
             rates[chrom] = tsk.RateMap(position=ends, rate=rec_rates)
 
-        self._rm = rates
+        #self._rm = rates  # we can store the ratemaps for debugging
 
         cum_rates = dict()
         for chrom, ratemap in rates.items():
             cumrate = ratemap.get_cumulative_mass(ratemap.right)
             cum_rates[chrom] = RecPair(ratemap.right, cumrate)
 
-        self.rates = rates
+        self.rates = {c: RecPair(x.right, x.rate) for c, x in rates.items()}
         self.cum_rates = cum_rates
         self.cum_interpol = rate_interpol(cum_rates, kind=self.cum_interpolation)
         self.inverse_cum_interpol = rate_interpol(cum_rates, inverse=True,

### 2023-02-19 bgspy/likelihood.py
more formal look at different optimization routines

@@ -24,7 +24,7 @@ from ctypes import POINTER, c_double, c_ssize_t, c_int
 from scipy import interpolate
 from bgspy.utils import signif, load_pickle
 from bgspy.data import pi_from_pairwise_summaries, GenomicBinnedData
-from bgspy.optim import run_optims, scipy_softmax_worker
+from bgspy.optim import run_optims, scipy_softmax_worker, nlopt_softmax_worker
 from bgspy.plots import model_diagnostic_plots, predict_chrom_plot
 from bgspy.plots import resid_fitted_plot, get_figax
 from bgspy.plots import chrom_resid_plot
@@ -162,6 +162,15 @@ def negll_softmax(theta, Y, B, w):
     assert np.allclose(W.sum(axis=0), np.ones(W.shape[1]))
     sm_theta[2:] = W.flat
     return negll_c(sm_theta, Y, B, w, version=3)
+
+
+def negll_softmax_nlopt(theta, grad, Y, B, w):
+    """
+    The nlopt wrapper to the above function, since
+    nlopt requires a grad second argument even if 
+    not known.
+    """
+    return negll_softmax(theta, Y, B, w)
 
 
 def negll_simplex(theta, Y, B, w):
@@ -626,7 +635,7 @@ class BGSLikelihood:
                 f.write(f"{feat}\t{dfe}\n")
 
     def __repr__(self):
-        rows = [f"MLE (interpolated w): {self.nw} x {self.nt} x {self.nf}"]
+        rows = [f"{type(self).__name__} (interpolated w): {self.nw} x {self.nt} x {self.nf}"]
         rows.append(f"  w grid: {signif(self.w)} (before interpolation)")
         rows.append(f"  t grid: {signif(self.t)}")
         return "\n".join(rows)
@@ -673,17 +682,26 @@ class SimplexModel(BGSLikelihood):
             self.start_mu_ = mu
         return start
 
-    def bounds(self):
+    def bounds(self, paired=True, global_softmax_bound=False):
+        """
+        By default, we assume scipy minimize. Scipy expects zipped
+        lists. For nlopt, (lower, upper) bounds are needed, so set 
+        paired = False. For global nlopt routines, we need fixed
+        bounds for softmax; we use 1e4 since these shouldn't get 
+        too large, but as a warning, note that this can cause 
+        optimization issues.
+        """
         return bounds_simplex(self.nt, self.nf,
                               self.log10_pi0_bounds,
                               self.log10_mu_bounds,
[…]

### 2023-02-19 bgspy/optim.py

@@ -16,6 +16,17 @@ Random notes:
    Googling around, this also happens with Julia's nlopt library so 
    seems to be a higher-level issue, but I could not confirm this.
    So we use scipy.minimize.
+
+ - global nlopt algorithms like GN_ISRES would be good to try 
+    againt softmax simplex models, but they require finite 
+    bounds. I tried putting an artificial bound on this but 
+    it did not seem to work.
+
+
+SimplexModels:
+  - softmax, which allows for unconstrained simplex optimization
+  - simplex with inequality and equality constraints
+
 """
 
 
@@ -99,7 +110,31 @@ def scipy_softmax_worker(start, func, nt, nf,
     nll = res.fun
     mle = res.x
     mle = convert_softmax(mle, nt, nf)
-    success = int(res.success) # this is so it matches nlopt
+    success = int(res.success)  # this is so it matches nlopt
+    return nll, mle, success
+
+
+def nlopt_softmax_worker(start, func, nt, nf, bounds,
+                         method, xtol_rel=1e-3,
+                         constraint_tol=1e-11,
+                         maxeval=1000000):
+    """
+    nlopt softmax wrapper
+    """
+    nparams = nt*nf + 2
+    method = getattr(nlopt, method)
+    opt = nlopt.opt(method, nparams)
+    opt.set_min_objective(func)
+    lb, ub = bounds
+    opt.set_lower_bounds(lb)
+    opt.set_upper_bounds(ub)
+    opt.set_xtol_rel(xtol_rel)
+    opt.set_maxeval(maxeval)
+    assert start.size == nparams
+    mle = opt.optimize(start)
+    nll = opt.last_optimum_value()
+    success = opt.last_optimize_result()
+    mle = convert_softmax(mle, nt, nf)
     return nll, mle, success
 
 

### 2023-02-19 bgspy/utils.py

@@ -1337,6 +1337,9 @@ def read_gff(filename, parse_attributes=False):
                 cols['attribute'] = attrs
             yield cols
 
+def save_pickle(obj, path):
+    with open(path, 'wb') as f:
+        return pickle.dump(obj, f)
 
 def load_pickle(path):
     with open(path, 'rb') as f:

### 2023-02-19 manuscript/manuscript.tex
ms notes
minor ms changes

@@
 
+Note on the importance of selection in humans.
+NOTE about the importance of calibration
+
 emphasis on whole genome models
@@
 constant churning of new variation
+
+- need for coalescent simulations with reduced $N_e$ along the genome.
 
@@
 
-Our $B$ is,
-
-\begin{align}
-  B = \frac{1}{2N} \sum_{t=0}^\infty \prod_{i=0}^t \left(1 - \frac{1}{2N_e(i)}\right).
-\end{align}
-
-This sum takes many terms to converge, which makes calculating the $B$ between
-each focal neutral site and segment intractably slow (even when the equilibrium
-$\widetilde{V}$ for the segment pre-calculated). We approximate this in the
-following way for faster numerical calculations. Let $g(t) = \prod_{i=0}^t (1 -
-\nicefrac{1}{2N_e(i)})$. Then, taking the first-order series expansion around
-the logarithm of the terms and approximating the sum with an integral,
-
-\begin{align}
-  g(t) &\approx \exp\left(- \frac{1}{2N} \int_0^t \exp\left(\frac{V}{2} Q(x)^2 \right) dx \right) 
-       %&\approx \exp\left(- \int_0^t \frac{\exp\left(\frac{V}{2} Q(i)^2\right)}{2N}  dx \right)
-\end{align}
-
-$B$ is the case of $\lim b \to \infty$ for $B(b)$. Approximating the sum in
-$B(b)$ with an integral,
-
-\begin{align}
-  B(b) &= \int_0^b \exp\left(- \frac{1}{2N} \int_0^t \exp\left(\frac{V}{2} Q(x)^2 \right) dx \right) dt \\
-       %&\approx \exp\left(- \int_0^t \frac{\exp\left(\frac{V}{2} Q(i)^2\right)}{2N}  dx \right)
-\end{align}
-
-Now, let $z = \log(t)$, then
-
-\begin{align}
-  B(w) &= \int_0^w \exp\left(- \frac{1}{2N} \int_0^{\log(z)} \exp\left(\frac{V}{2} Q(x)^2 \right) dx \right) \exp(t) dz \\
-       %&\approx \exp\left(- \int_0^t \frac{\exp\left(\frac{V}{2} Q(i)^2\right)}{2N}  dx \right)
-\end{align}
+\subsection*{Optimization}
+
+Optimization in non-linear regression is notoriously difficult both in general
+(Bates) and in the specific problem of estimating the effects of genome-wide
+selection \parencite{Murphy2022-sj}. We experimented with a variety of global
+and local optimization approaches and parameterizations, but found that
 

### 2023-02-25 bgspy/bootstrap.py
readme notes

@@ -7,7 +7,18 @@ import numpy as np
 BootstrapResults = namedtuple('BootstrapResults',
                               ('b_nll', 'b_theta',
                                'bp_nll', 'bp_theta'))
+def split_list(list, chunksize):
+    for i in range(0, len(list), chunksize):
+        yield list[i:i + chunk_size]
 
+
+def block_bins(bins, blocksize):
+    i = 0
+    blocks = []
+    for chrom in bins.keys():
+        indices = bins.chrom_indices(chrom)
+        blocks.append(list(split_list(indices, blocksize)))
+    return blocks
 
 def resample_blocks(bins, blocksize, nsamples=None, exclude_chroms=None):
     """

### 2023-02-25 bgspy/command_line.py

@@ -339,6 +339,8 @@ def subrate(bs_file, fit, force_feature, output, split):
 @click.option('--fit', default=None, type=click.Path(exists=True),
               help=('pickle file of fitted results, for starting at MLE '
                     '(ignore for random starts)'))
+@click.option('--blocksize', required=True, type=int,
+              help='the blocksize, in number of consecutive windows')
 @click.option('--output', required=True, type=click.Path(writable=True),
               help='an .npz output file')
 @click.option('--fit-dir', default=None, help="fit directory for saving whole fits")
@@ -351,29 +353,47 @@ def subrate(bs_file, fit, force_feature, output, split):
               help='number of starts for multi-start optimization',
               type=int, default=1)
 @click.option('--include-Bs', default=False, is_flag=True, help="whether to include classic Bs too")
-def jackknife(data, fit, output, fit_dir, chrom,
+def jackblock(data, fit, blocksize, output, fit_dir, chrom,
+              ncores, nstarts, include_bs):
+    """
+    """
+    mle_fit(data=data,
+            output_file=output,
+            ncores=ncores,
+            nstarts=nstarts,
+            loo_chrom=chrom,
+            bp_only=True)
+
+
+@cli.command()
+@click.option('--data', required=True, default=None, type=click.Path(exists=True),
+              help="pickle of pre-computed summary statistics")
+@click.option('--fit', default=None, type=click.Path(exists=True),
+              help=('pickle file of fitted results, for starting at MLE '
+                    '(ignore for random starts)'))
+@click.option('--output', required=True, type=click.Path(writable=True),
+              help='an .npz output file')
+@click.option('--fit-dir', default=None, help="fit directory for saving whole fits")
+@click.option('--chrom', default=None,
+              help="leave-one-out chromosome, e.g. for parallel processing across a cluster")
+@click.option('--ncores',
+              help='number of cores to use for multi-start optimization',
+              type=int, default=1)
+@click.option('--nstarts',
+              help='number of starts for multi-start optimization',
+              type=int, default=1)
+@click.option('--include-Bs', default=False, is_flag=True, help="whether to include classic Bs too")
+def jackchrom(data, fit, output, fit_dir, chrom,
               ncores, nstarts, include_bs):
     """
     Estimate R2 by leaving out a chromosome, fitting to the rest of the genome,
     and predicting the observed diversity on the excluded chromosome.
 
-    TODO:
-     - starts only work for B' fits.
[…]

### 2023-02-25 bgspy/data.py

@@ -701,6 +701,9 @@ class GenomicBins:
         return np.array(chroms)
 
     def chrom_ints(self, filter_masked=True):
+        """
+        Return an array of all chromosome numbers as integers.
+        """
         chroms = []
         chrom_map = {c: i for i, c in enumerate(self.seqlens.keys())}
         for chrom, bins in self.bins(filter_masked=filter_masked).items():

### 2023-02-25 bgspy/likelihood.py

@@ -418,6 +418,32 @@ class BGSLikelihood:
             self.theta_ = optim_res.thetas_[index]
             self.nll_ = optim_res.nlls_[index]
 
+    def jackknife_chrom(self, chrom, **kwargs):
+        """
+        Do the jackknife, leaving out the specified chromosome.
+        """
+        msg = "bins attribute must be set to a GenomicBinnedData object"
+        assert self.bins is not None, msg
+        chrom_idx = set(self.bins.chrom_indices(chrom))
+        assert len(chrom_idx)
+        all_idx = set(range(self.Y.shape[0]))
+        in_sample_idx = np.array(list(all_idx.difference(chrom_idx)))
+        # return the fit directly
+        obj = self.fit(**kwargs, _indices=in_sample_idx)
+        obj._chrom_loo_fit = chrom
+        return obj
+
+    def jackknife_block(self, blocksize, blocknum=None, **kwargs):
+        """
+        """
+        blocks = block_bins(self.bins, blocksize)
+        if blocknum is not None:
+            blocks = [blocks[blocknum]]
+        for block_indices in blocks:
+            obj = self.fit(**kwargs, _indices=block_indices)
+            obj._indices_fit = block_indices
+            
+
     def load_jackknives(self, fit_dir):
         """
         Load the chromosome jackknife results by setting attributes.
@@ -434,27 +460,6 @@ class BGSLikelihood:
         self.jack_nlls_ = nlls
         self.jack_thetas_ = thetas
         self.jack_chroms_ = chroms
- 
-    def jackknife_chrom(self, chrom, **kwargs):
-        """
-        Do the jackknife, leaving out the specified chromosome.
-
-        If you want to compute across a cluster, set the chunks tuple,
-        (nchunks, chunk_id)
-
-        If you want to recycle the MLE θ, pass a starts keyword argument
-        with the MLE vector (optionally repeated for multiple starts.
-        """
-        msg = "bins attribute must be set to a GenomicBinnedData object"
-        assert self.bins is not None, msg
-        chrom_idx = set(self.bins.chrom_indices(chrom))
-        assert len(chrom_idx)
-        all_idx = set(range(self.Y.shape[0]))
-        in_sample_idx = np.array(list(all_idx.difference(chrom_idx)))
-        # return the fit directly
-        obj = self.fit(**kwargs, _indices=in_sample_idx)
-        self._chrom_loo_fit = chrom
[…]

### 2023-02-25 bgspy/pipeline.py

@@ -112,7 +112,14 @@ def mle_fit(data, output_file, ncores=70, nstarts=200,
             loo_chrom=None, chrom=None,
             start=None, bp_only=False):
     """
-    Load the binned data, fit B' (and optionally B) models, and
+    The general fitting pipeline, for MLE fits, chromosome-specific 
+    fits, leave-one-out chromosome fits, and block-jackknifing.
+
+    Basic pipeline:
+      - Load the binned data
+      - Fit B' (and optionally B) model, possibly LOO or single
+        chromosome.
+      - Save the results.
     save the results.
     """
     dat = load_pickle(data)
@@ -133,13 +140,19 @@ def mle_fit(data, output_file, ncores=70, nstarts=200,
     logging.info(msg)
     m_bp = SimplexModel(w=w, t=t, logB=bp, Y=Y,
                         bins=bins, features=features)
+
+    # NOTE: for loo/block JK we need to manually load the 
+    # optim results.
     if loo_chrom is not None:
         jk_opt = m_bp.jackknife_chrom(starts=starts, ncores=ncores,
                                       chrom=loo_chrom, mu=mu)
-        # we need to load manually...
+        m_bp._load_optim(jk_opt)
+    elif blocksize is not None:
+        jk_opt = m_bp.jackknife_block(starts=starts, ncores=ncores,
+                                      blocksize=blocksize, mu=mu)
         m_bp._load_optim(jk_opt)
     else:
-        m_bp.fit(starts=starts, ncores=ncores, 
+        m_bp.fit(starts=starts, ncores=ncores,
                  mu=mu, chrom=chrom)
 
     msg = "saving B' model results"
@@ -152,6 +165,8 @@ def mle_fit(data, output_file, ncores=70, nstarts=200,
         logging.info("fitting B model")
         m_b = SimplexModel(w=w, t=t, logB=b, Y=Y,
                            bins=bins, features=features)
+        if blocksize is not None:
+            raise NotImplementedError, "B' block jackknifing not implemented"
         if loo_chrom is not None:
             jk_opt = m_b.jackknife_chrom(starts=nstarts, ncores=ncores, 
                                          chrom=loo_chrom, mu=mu)

### 2023-02-26 bgspy/bootstrap.py
new block jackknifing, improvements in snakefile, etc

@@ -7,17 +7,19 @@ import numpy as np
 BootstrapResults = namedtuple('BootstrapResults',
                               ('b_nll', 'b_theta',
                                'bp_nll', 'bp_theta'))
-def split_list(list, chunksize):
-    for i in range(0, len(list), chunksize):
-        yield list[i:i + chunk_size]
+def split_list(x, chunksize):
+    x = list(x)
+    for i in range(0, len(x), chunksize):
+        yield x[i:i + chunksize]
 
 
 def block_bins(bins, blocksize):
+    assert isinstance(blocksize, int), type(blocksize)
     i = 0
     blocks = []
     for chrom in bins.keys():
         indices = bins.chrom_indices(chrom)
-        blocks.append(list(split_list(indices, blocksize)))
+        blocks.extend(list(split_list(indices, blocksize)))
     return blocks
 
 def resample_blocks(bins, blocksize, nsamples=None, exclude_chroms=None):

### 2023-02-26 bgspy/command_line.py

@@ -8,7 +8,7 @@ from bgspy.models import BGSModel
 from bgspy.genome import Genome
 from bgspy.utils import Grid, load_pickle
 from bgspy.pipeline import summarize_data, mle_fit
-from bgspy.bootstrap import load_from_bs_dir
+from bgspy.bootstrap import block_bins
 
 SPLIT_LENGTH_DEFAULT = 10_000
 STEP_DEFAULT = 10_000
@@ -339,8 +339,14 @@ def subrate(bs_file, fit, force_feature, output, split):
 @click.option('--fit', default=None, type=click.Path(exists=True),
               help=('pickle file of fitted results, for starting at MLE '
                     '(ignore for random starts)'))
-@click.option('--blocksize', required=True, type=int,
+@click.option('--blocksize', type=int,
               help='the blocksize, in number of consecutive windows')
+@click.option('--blockwidth', type=int, help="the blockwidth in basepairs")
+@click.option('--blocknum', default=None, type=int,
+              help='which block to run the fit on (e.g. for cluster use)')
+@click.option('--blockfrac', default=None, type=float,
+              help='for not a full jackknife, drop the block at this fraction '
+                   "(this approach doesn't require a prior knowledge of number of blocks")
 @click.option('--output', required=True, type=click.Path(writable=True),
               help='an .npz output file')
 @click.option('--fit-dir', default=None, help="fit directory for saving whole fits")
@@ -353,15 +359,32 @@ def subrate(bs_file, fit, force_feature, output, split):
               help='number of starts for multi-start optimization',
               type=int, default=1)
 @click.option('--include-Bs', default=False, is_flag=True, help="whether to include classic Bs too")
-def jackblock(data, fit, blocksize, output, fit_dir, chrom,
-              ncores, nstarts, include_bs):
+def jackblock(data, fit, blocksize, blockwidth, blocknum, blockfrac, output, 
+              fit_dir, chrom, ncores, nstarts, include_bs):
     """
+    Run the block-jackknifing routine. There are two modes:
+        1. Run with --blocksize set only, and all blocks will be 
+            jackknifed.
+        2. Run with --blocksize and --blocknum, and only that 
+            block number is left out (e.g. for cluster use).
     """
+    if blockfrac is not None or blockwidth is not None:
+        # we need to figure out the block width or num blocks
+        bins = load_pickle(data)['bins']
+        if blockwidth is not None:
[…]

### 2023-02-26 bgspy/likelihood.py

@@ -2,7 +2,7 @@
 import os
 import re
 import logging
-import copy
+from copy import copy
 import pickle
 import itertools
 import tqdm
@@ -23,14 +23,15 @@ from ctypes import POINTER, c_double, c_ssize_t, c_int
 #     pass
 
 from scipy import interpolate
-from bgspy.utils import signif, load_pickle
+from bgspy.utils import signif, load_pickle, si_prefix
 from bgspy.data import pi_from_pairwise_summaries, GenomicBinnedData
 from bgspy.optim import run_optims, scipy_softmax_worker
 from bgspy.optim import nlopt_softmax_worker, nlopt_softmax_fixedmu_worker
 from bgspy.plots import model_diagnostic_plots, predict_chrom_plot
 from bgspy.plots import resid_fitted_plot, get_figax
 from bgspy.plots import chrom_resid_plot
-from bgspy.bootstrap import process_bootstraps, pivot_ci, percentile_ci
+from bgspy.bootstrap import block_bins
+
 
 # load the library (relative to this file in src/)
 LIBRARY_PATH = os.path.join(os.path.dirname(__file__), '..', 'src')
@@ -433,20 +434,59 @@ class BGSLikelihood:
         obj._chrom_loo_fit = chrom
         return obj
 
-    def jackknife_block(self, blocksize, blocknum=None, **kwargs):
+    def jackknife_block(self, blocksize, blocknum=None, 
+                        calc_loo_R2=True, **kwargs):
         """
+        Use the block-jackknife to estimate uncertainty. This 
+        will use blocksize-sized blocks (in number of consecutive
+        windows, so the physical size of the block will vary with
+        the window size.
+
+        TODO: this could be refactored a bit to remove redundancy.
         """
+        # generate the blocks
         blocks = block_bins(self.bins, blocksize)
+
         if blocknum is not None:
-            blocks = [blocks[blocknum]]
-        for block_indices in blocks:
+            out_block_idx = blocks.pop(blocknum)
+            block_indices = np.array(list(itertools.chain(*blocks)))
             obj = self.fit(**kwargs, _indices=block_indices)
             obj._indices_fit = block_indices
-            
+            # store metadata 
+            obj._blocksize = blocksize
+            obj._blocknum = blocknum
+            obj._blocksize_bp = blocksize*self.bins.width
+            # for out-sample prediction
+            obj._out_block_indices = out_block_idx
+            if calc_loo_R2:
[…]

### 2023-02-26 bgspy/pipeline.py

@@ -109,7 +109,8 @@ def summarize_data(# annotation
 
 def mle_fit(data, output_file, ncores=70, nstarts=200,
             mu=None, verbose=True, 
-            loo_chrom=None, chrom=None,
+            loo_chrom=None, chrom=None,  # for LOO on only one chrom
+            blocksize=None, blocknum=None,  # for block-jackknifing
             start=None, bp_only=False):
     """
     The general fitting pipeline, for MLE fits, chromosome-specific 
@@ -122,12 +123,6 @@ def mle_fit(data, output_file, ncores=70, nstarts=200,
       - Save the results.
     save the results.
     """
-    dat = load_pickle(data)
-    bins, Y, bp = dat['bins'], dat['Y'], dat['bp']
-    w, t = dat['w'], dat['t']
-    features = dat['features']
-    b = dat.get('b', None)
-
     # deal with start
     if start is not None:
         starts = [start]
@@ -138,9 +133,9 @@ def mle_fit(data, output_file, ncores=70, nstarts=200,
     if chrom is not None:
         msg += f" (leaving out {chrom})"
     logging.info(msg)
-    m_bp = SimplexModel(w=w, t=t, logB=bp, Y=Y,
-                        bins=bins, features=features)
+    m_bp = SimplexModel.from_data(data)
 
+    ## fitting (main MLE, loo-chrom, block jackknife)
     # NOTE: for loo/block JK we need to manually load the 
     # optim results.
     if loo_chrom is not None:
@@ -148,35 +143,38 @@ def mle_fit(data, output_file, ncores=70, nstarts=200,
                                       chrom=loo_chrom, mu=mu)
         m_bp._load_optim(jk_opt)
     elif blocksize is not None:
+        assert blocksize > 1, "blocksize must be > 1"
         jk_opt = m_bp.jackknife_block(starts=starts, ncores=ncores,
-                                      blocksize=blocksize, mu=mu)
+                                      blocksize=blocksize, 
+                                      blocknum=blocknum, mu=mu)
         m_bp._load_optim(jk_opt)
     else:
         m_bp.fit(starts=starts, ncores=ncores,
                  mu=mu, chrom=chrom)
-
+    
+    # save the B' results
     msg = "saving B' model results"
     logging.info(msg)
     obj = {'mbp': m_bp}
     with open(output_file, 'wb') as f:
         pickle.dump(obj, f)
 
-    if b is not None and not bp_only:
-        logging.info("fitting B model")
[…]

### 2023-02-26 bgspy/utils.py

@@ -1358,3 +1358,30 @@ def argsort_chroms(chromosomes):
     indices = list(range(len(chromosomes)))
     indices.sort(key=lambda i: chromosome_key(chromosomes[i]))
     return indices
+
+def si_prefix(quantity, digits=3):
+    prefix = {'y': 1e-24,  # yocto
+              'z': 1e-21,  # zepto
+              'a': 1e-18,  # atto
+              'f': 1e-15,  # femto
+              'p': 1e-12,  # pico
+              'n': 1e-9,   # nano
+              'u': 1e-6,   # micro
+              'm': 1e-3,   # mili
+              'c': 1e-2,   # centi
+              'd': 1e-1,   # deci
+              'k': 1e3,    # kilo
+              'M': 1e6,    # mega
+              'G': 1e9,    # giga
+              'T': 1e12,   # tera
+              'P': 1e15,   # peta
+              'E': 1e18,   # exa
+              'Z': 1e21,   # zetta
+              'Y': 1e24,   # yotta
+              }
+    inverse_prefix = {int(np.log10(v)): p for p, v in prefix.items()}
+    x = np.log10(quantity)
+    if int(x) == x:
+        x = int(x)  # simplify it
+    prefix = inverse_prefix[int(np.ceil(np.log10(quantity)))]
+    return f"{np.round(x, digits)}{prefix}"

### 2023-03-21 manuscript/manuscript.tex
phrasing
minor ms changes
ms updates

@@
 whole-genome variance models.
+These are flux models.
 
@@
 However, our model
+
+Evolutionary processes describe how genetic variation is introduced and
+ultimately converted to substitutions or lost from the population. Our model 
+makes predictions about the eventual outcome of weakly deleterious mutation 
+that seem to roughly match observations.
 
@@
 \begin{align}
+  \label{eq:Ne_inf}
   N_{e,\infty} \approx N \exp\left(-\frac{V_h}{(1-Z) (2-(2-M) Z)} \right).
@@
 
-Using $Z_\text{BGS}$, $V_m$, and $V$ under the background selection model, and
-approximation $N_{e,\infty}^\text{SC98}$ (equation XXX), find an equation
-reminiscent of the classic background selection with recombination results. We
-find that by using the full equation XXX $N_{e,\infty}$ rather than the
-approximation of Santiago and Caballero, the reduction due to background
-selection is
-
-\begin{align}
-  N_{e,\infty}^\text{BGS} = N \exp\left( - \frac{U}{2s + M - Ms} \right)
-\end{align}
-
-which matches equation 8 of \textcite{Hudson1995-xc} except for the $Ls$ term
+Now, we can use these values for $Z_\text{BGS}$, $V_m$, and $V$ in equation
+\eqref{eq:Ne_inf}. Through this alternate quantitative genetics derivation, we
+arrive at the same equation as classic BGS theory,
+
+\begin{align}
+  N_{e,\infty}^\text{BGS} = N \exp\left( - \frac{U}{2s + M - Ms} \right).
+\end{align}
+
+This matches equation 8 of \textcite{Hudson1995-xc} except for the $Ms$ term
 and equation 10 of \textcite{Nordborg1996-nq} except for the $2s$ term.
@@
 
+\subsection*{Misspecification of the Polygenic Selection Model}
+
+Thus far, we have developed the general model of equation \eqref{XXX} by
+assuming that the only source of $V_h$ is from the deleterious
+mutation-selection-draft-drift process. However, other polygenic evolutionary
+processes can also create, alter, and eliminate fitness variation. By assuming
+that the only fitness variation originates from deleterious
+mutation-selection-drift-draft processes, we introduce some model
+misspecification. There are two main types of misspecification under the
+selection model. First, non-polygenic positive selection processes, such as
+recurrent sweeps, may create small-scale distortions in local genealogies.
+These have been found to only impact diversity on small spatial scales
+\parencite{Murphy2022-sj} in humans, and explain little of the total variation
+in megabase-scale patterns of diversity. Second, there is model
+misspecification due to our model assuming that the only polygenic selection
+process happening is the flux of new deleterious mutations in, and their
+eventual loss or fixation. Other polygenic selection processes could be
+occurring, e.g. traits could be responding to directional selection. The
+Santiago and Caballero model our work extends is very general, and can account
+for multiple types of polygenic selection. The key input parameters are the
+equilibrium heritable fitness variation $V_h$, and the variation introduced by
+mutation each generation $V_m$.
+
+Let us assume for now that,
+
+\begin{align}
+  \label{eq:Ne_inf}
+  N_{e,\infty} \approx N \exp\left(-\frac{V_h + \varepsilon}{(1-Z) (2-(2-M) Z)} \right).
+\end{align}
+
 \subsection*{Heterozygosity and $B$ values Under Weak and Strong Selection}
@@
 heterozygosity is determined by the sequence $\widetilde{N}_{e,t}$, where each
-term is
-
-XXX
+term is given by \eqref{eq:net_integral}. We experimented with calculating this
+full sum in our methods, but it seemed to make little difference in our
+application. Additionally, it was too costly to calculate computationally
+for genome-wide calculations of the B' map.
 

### 2023-03-21 bgspy/command_line.py
getting window stats

@@ -2,12 +2,13 @@ import click
 import logging
 import numpy as np
 import pickle
+from collections import defaultdict
 from si_prefix import si_format
 import os
 from collections import namedtuple
 from bgspy.models import BGSModel
 from bgspy.genome import Genome
-from bgspy.utils import Grid, load_pickle
+from bgspy.utils import Grid, load_pickle, bin_chroms
 from bgspy.pipeline import summarize_data, mle_fit, summarize_sim_data
 from bgspy.pipeline import ModelDir
 from bgspy.bootstrap import block_bins
@@ -49,24 +50,30 @@ def grid_maker_from_str(x):
 
 
 def make_bgs_model(seqlens, annot, recmap, conv_factor, w, t, g=None,
-                   chroms=None, name=None, split_length=SPLIT_LENGTH_DEFAULT):
+                   chroms=None, name=None, genome_only=False,
+                   split_length=SPLIT_LENGTH_DEFAULT):
     """
     Build the BGSModel and the Genome object it uses.
     """
+    if name is None:
+        # infer the name
+        bn = os.path.splitext(os.path.basename(seqlens))[0]
+        name = bn.replace('_seqlens', '')
+    gn = Genome(name, seqlens_file=seqlens, chroms=chroms)
+    gn.load_annot(annot)
+    gn.load_recmap(recmap, conversion_factor=conv_factor)
+    gn.create_segments(split_length=split_length)
+    if genome_only:
+        return gn
+    
+    # get the grid 
     if g is not None:
         # g takes priority
         w, t = grid_maker_from_str(g)
     else:
         w, t = parse_gridstr(w), parse_gridstr(t),
-    if name is None:
-        # infer the name
-        bn = os.path.splitext(os.path.basename(seqlens))[0]
-        name = bn.replace('_seqlens', '')
-    g = Genome(name, seqlens_file=seqlens, chroms=chroms)
-    g.load_annot(annot)
-    g.load_recmap(recmap, conversion_factor=conv_factor)
-    g.create_segments(split_length=split_length)
-    m = BGSModel(g, w_grid=w, t_grid=t, split_length=split_length)
+
+    m = BGSModel(gn, w_grid=w, t_grid=t, split_length=split_length)
     return m
 
 
@@ -90,13 +97,17 @@ def parse_gridstr(x):
     return 10**np.linspace(lower, upper, int(ngrid))
 
 
[…]

### 2023-03-21 bgspy/likelihood.py

@@ -3,6 +3,7 @@ import os
 import warnings
 import re
 from copy import copy
+from collections import defaultdict
 import pickle
 import itertools
 import tqdm
@@ -628,10 +629,26 @@ class BGSLikelihood:
         #ax.set_ylabel('probability')
         #ax.legend()
 
-
-    def dfe_plot(self, add_legend=True, figax=None):
+    def W_stderrs(self):
         """
-        TODO
+        A dictionary of standard errors.
+        """
+        if self.sigma_ is None:
+            return None
+        fixed_mu_offset = 1 + int(not self._fixed_mu is not None)
+        ses = self.sigma_[fixed_mu_offset:].reshape(self.mle_W.shape)
+        se = defaultdict(list)
+        mean = defaultdict(list)
+        for i, feature in enumerate(self.features):
+            for j, t in enumerate(self.t):
+                se[feature].append(ses[j,i])
+                mean[feature].append(self.mle_W[j,i])
+        return mean, se
+
+
+    def dfe_plot(self, add_legend=True, legend_kwargs={}, figax=None):
+        """
+        Plot a boxplot of all features.
         """
         fig, ax = get_figax(figax)
         xt = np.log10(self.t)
@@ -646,7 +663,7 @@ class BGSLikelihood:
         ax.set_xticks(np.log10(self.t), [f"$10^{{{int(x)}}}$" for x in xt])
         ax.set_ylabel('probability')
         if add_legend:
-            ax.legend()
+            ax.legend(**legend_kwargs)
 
     def pi(self):
         pi = pi_from_pairwise_summaries(self.Y)

### 2023-03-21 bgspy/plots.py

@@ -202,7 +202,7 @@ def savitzky_golay(y, window_size, order, deriv=0, rate=1):
 
 def predict_chrom_plot(model, chrom, ratio=True,
                        label='prediction', lw=2,
-                       c='cornflowerblue',
+                       pred_col='cornflowerblue', pi_col='0.22',
                        alpha_predict=1, alpha_pi=1,
                        add_r2=False, figax=None, predict_kwargs=None):
     m = model
@@ -224,11 +224,11 @@ def predict_chrom_plot(model, chrom, ratio=True,
         y = mean_ratio(y)
         ylab = "$\\pi/\\bar{\\pi}$"
     #pi = savitzky_golay(pi, window_size=5, order=1)
-    ax.plot(midpoints, y, label=label, linewidth=lw, c=c,
+    ax.plot(midpoints, y, label=label, linewidth=lw, c=pred_col,
             alpha=alpha_predict, zorder=3)
     if ratio:
         pi = mean_ratio(pi)
-    ax.plot(midpoints, pi, c='0.22', label='data', linewidth=lw,
+    ax.plot(midpoints, pi, label='data', linewidth=lw, c=pi_col,
             alpha=alpha_pi)
     if add_r2:
         ax.set_title(f"$R^2 = {np.round(model.R2(), 2)}$")

### 2023-03-25 bgspy/command_line.py
new expansion sims, corrections to code

@@ -38,6 +38,9 @@ MIN_W = np.sqrt(1e-8 * 1e-7)
 HUMAN_W = (-11, -7)
 HUMAN_T = (-8, -1)
 
+HUMAN_W_STR = f"{HUMAN_W[0]}:{HUMAN_W[1]}:6"
+HUMAN_T_STR = f"{HUMAN_T[0]}:{HUMAN_T[1]}:8"
+
 
 def sterialize_mu(mu):
     try:
@@ -136,9 +139,9 @@ def cli():
                      "cM/Mb rates, use 1e-8)")
 @click.option('--t', help="string of log10 lower:upper:grid_size (e.g. -7,-1,7) or comma-separated "
                    "list (e.g. 0.01,0.001) for heterozygous selection coefficient",
-                    default='-6:-1:50')
+                    default=HUMAN_T_STR)
 @click.option('--w', help="string of lower:upper:grid_size or comma-separated "
-                   "list for log10 mutation rates", default='-10:-7:50' )
+                   "list for log10 mutation rates", default=HUMAN_W_STR)
 @click.option('--g', help="a grid string for human defaults, e.g. 6x8 "
               "(--t/--w ignored if this is set)", default=None)
 @click.option('--chrom', help="process specified chromosome(s)", default=None, multiple=True)

### 2023-03-25 bgspy/parallel.py

@@ -328,7 +328,7 @@ def BSC16_segment_lazy_parallel(mu, sh, L, rbp, N, ncores, rescaling=None):
         res = list(map(func, tqdm(zip(L, rbp, rescaling), total=len(L))))
     else:
         with multiprocessing.Pool(ncores) as p:
-            res = deque(p.imap(func, tqdm(zip(L, rbp, rescaling), total=len(L))))
+            res = list(p.imap(func, tqdm(zip(L, rbp, rescaling), total=len(L))))
             p.close()
             p.join()
  

### 2023-03-25 bgspy/slim.py

@@ -121,8 +121,7 @@ class SlimRuns():
         self.config = config
         self.name = config['name']
         self.script = config['script']
-        self.nreps = config['nreps']
-        self.seed = config['seed']
+        self.seed = int(config['seed'])
         self.rng = ng = np.random.default_rng(self.seed)
         self.dir = config['dir']
         self.suffices = config['suffices']

### 2023-03-25 tools/process_region_sims.py

@@ -7,6 +7,7 @@ import warnings
 import pyslim
 import sys
 import os
+#from scipy.stats import trim_mean
 import tqdm
 import statsmodels.api as sm
 from bgspy.utils import get_files, bin_chrom
@@ -50,7 +51,7 @@ def load_and_recap(treefile, burnin=2000):
     B = ts.diversity(mode='branch') / (4*N)
     bins = bin_chrom(ts.sequence_length, 10000) # 10kb windows
     B_wins = ts.diversity(windows=bins, mode='branch') / (4*N)
-    B_middle = middle_elements(B_wins, 3).mean()
+    B_middle = middle_elements(B_wins, 4).mean()
     logfile = treefile.replace('_treeseq.tree', '_log.tsv.gz')
     d = pd.read_csv(logfile, sep='\t', comment='#')
     last_gen = np.max(d['cycle'])

### 2023-03-31 bgspy/utils.py
new snakefile

@@ -969,7 +969,7 @@ def write_bed(chromdict, filename, compress=True, append=False):
             for vals in ranges:
                 f.write("\t".join(map(str, [chrom, *vals])) + "\n")
 
-def load_seqlens(file):
+def load_seqlens(file, exclude=None):
     seqlens = dict()
     params = []
     with readfile(file) as f:
@@ -978,6 +978,9 @@ def load_seqlens(file):
                 params.append(line.strip().lstrip('#'))
                 continue
             chrom, end = line.strip().split('\t')
+            if exclude is not None:
+                if chrom in exclude:
+                    continue
             seqlens[chrom] = int(end)
     return seqlens
 

### 2023-03-31 bgspy/likelihood.py
old talapas changes

@@ -168,7 +168,8 @@ def negll_softmax(theta, Y, B, w):
     """
     Softmax version of the simplex model wrapper for negll_c().
 
-    For scipy, i.e. no grad argument.
+    For scipy, i.e. no grad argument, so must be wrapped for use
+    with nlopt.
     """
     nx, nw, nt, nf = B.shape
     # get out the W matrix, which on the optimization side, is over all reals

### 2023-03-31 bgspy/command_line.py
fixed issue with tracks

@@ -667,10 +667,15 @@ def tracks(config, full, seqlens):
     priority = list(features.keys())
     beds = {c: read_bed3(f) for c, f in features.items()}
     masks = combine_features(beds, priority, load_seqlens(seqlens))
-    res = masks_to_ranges(masks, labels=priority, include_other=full)
+    res = masks_to_ranges(masks, labels=priority)
     for chrom, ranges in res.items():
         for range in ranges:
-            print(f"{chrom}\t{range[0]}\t{range[1]}\t{range[2]}")
+            start, end, label = range
+            if not full and label is None:
+                continue
+            else:
+                label = 'other'
+            print(f"{chrom}\t{start}\t{end}\t{label}")
 
 
 @cli.command()

### 2023-03-31 bgspy/utils.py

@@ -871,8 +871,7 @@ def rle(inarray):
             p = np.cumsum(np.append(0, z))[:-1] # positions
             return(z, p, ia[i])
 
-def masks_to_ranges(mask_dict, return_val=False, labels=None,
-                    include_other=False):
+def masks_to_ranges(mask_dict, return_val=False, labels=None):
     """
     Given the masks from combine_features(), turn these into a chromosome
     range dictionary.
@@ -896,13 +895,9 @@ def masks_to_ranges(mask_dict, return_val=False, labels=None,
     for chrom in mask_dict:
         rls, starts, vals = rle(mask_dict[chrom])
         for rl, s, val in zip(rls, starts, vals):
-            if val == 0 and include_other:
-                lab = 'other' if labels is not None else 0
-                x = (s, s + rl, lab)
-                ranges[chrom].append(x)
-                continue
             if labels is not None:
-                x = (s, s + rl, labels[val-1])
+                lab = labels[val-1] if val > 0 else None
+                x = (s, s + rl, lab)
             elif return_val:
                 x = (s, s + rl, val)
             else:

### 2023-03-31 bgspy/command_line.py
fixed another issue with tracks

@@ -671,10 +671,11 @@ def tracks(config, full, seqlens):
     for chrom, ranges in res.items():
         for range in ranges:
             start, end, label = range
-            if not full and label is None:
-                continue
+            if not full:
+                if label is None:
+                    continue
             else:
-                label = 'other'
+                label = 'other' if label is None else label
             print(f"{chrom}\t{start}\t{end}\t{label}")
 
 

### 2023-03-31 manuscript/supplementary.tex
minor ms changes
ms updates

@@
 
+In practice, to fit such a model, we must discretize both $\mu(s, k(g))$ and
+$s$ in Equation \eqref{suppeqn:b}. We define $\mu_i$ to be the rate of
+mutations entering the population per basepair, per generation with selection
+coefficient $s_i$ (i.e. the product of the mutation rate and the distribution
+of deleterious fitness effects). Then, the reduction at position $x$ is,
+
 \begin{align}
   \label{suppeqn:b}
-  B(x, \Psi) = \exp \left( \sum_g^S \sum_{i} b\left( \mu_{i, \;k(g)}, s_i, L_g, r_g, d(x, g) \right)\right)
+  B(x, \Psi) = \exp \left( \sum_g^S \sum_{i} b\left( \mu_{i, \;k(g)}, s_i, L_g, r_g, d(x, g) \right)\right).
 \end{align}
@@
 
-In practice, to fit such a model, we must discretize both $\mu(s, k(g))$ and
-$s$. Our models use a fixed $m_s$-element grid of selection coefficients, and
-mutation rates are on a fixed $m_\mu$-element grid, which is later interpolated
-over during the maximum likelihood optimization. The number of annotation
-features $K$ is variable and set by the annotation data specified. There are
-two ways to parameterize the mutation rates and DFE for all annotation classes
-in background selection models. First, for the \emph{free-mutation}
+Because it is computationally-intensive to numerically solve the system of two
+non-linear equations to calculate $b$ at each segment, these are pre-computed
+for an $m_s$-element grid of selection coefficients and an $m_\mu$-element grid
+of mutation rates--DFE products. The number of annotation features $K$ is
+variable and set by the annotation data specified. There are two ways to
+parameterize the mutation rates and DFE for all annotation classes in
+background selection models. First, for the \emph{free-mutation}
 parameterization, each annotation feature class has a free mutation rate
@@
 
+\subsection{Numeric Optimization}
+
+mu bounds
+
+bounds and time, bounds and convergence
+
+Same optima, but took longer
+
+%Our $B$ is,
+%
+%\begin{align}
+%  B = \frac{1}{2N} \sum_{t=0}^\infty \prod_{i=0}^t \left(1 - \frac{1}{2N_e(i)}\right).
+%\end{align}
+%
+%This sum takes many terms to converge, which makes calculating the $B$ between
+%each focal neutral site and segment intractably slow (even when the equilibrium
+%$\widetilde{V}$ for the segment pre-calculated). We approximate this in the
+%following way for faster numerical calculations. Let $g(t) = \prod_{i=0}^t (1 -
+%\nicefrac{1}{2N_e(i)})$. Then, taking the first-order series expansion around
+%the logarithm of the terms and approximating the sum with an integral,
+%
+%\begin{align}
+%  g(t) &\approx \exp\left(- \frac{1}{2N} \int_0^t \exp\left(\frac{V}{2} Q(x)^2 \right) dx \right) 
+%       %&\approx \exp\left(- \int_0^t \frac{\exp\left(\frac{V}{2} Q(i)^2\right)}{2N}  dx \right)
+%\end{align}
+%
+%$B$ is the case of $\lim b \to \infty$ for $B(b)$. Approximating the sum in
+%$B(b)$ with an integral,
+%
+%\begin{align}
+%  B(b) &= \int_0^b \exp\left(- \frac{1}{2N} \int_0^t \exp\left(\frac{V}{2} Q(x)^2 \right) dx \right) dt \\
+%       %&\approx \exp\left(- \int_0^t \frac{\exp\left(\frac{V}{2} Q(i)^2\right)}{2N}  dx \right)
+%\end{align}
+%
+%Now, let $z = \log(t)$, then
+%
+%\begin{align}
+%  B(w) &= \int_0^w \exp\left(- \frac{1}{2N} \int_0^{\log(z)} \exp\left(\frac{V}{2} Q(x)^2 \right) dx \right) \exp(t) dz \\
+%       %&\approx \exp\left(- \int_0^t \frac{\exp\left(\frac{V}{2} Q(i)^2\right)}{2N}  dx \right)
+%\end{align}
+%
+%
 \printbibliography

### 2023-04-06 manuscript/manuscript.tex
minor ms changes

@@
 
+New mutations are the ultimate source of all adaptations, but the vast majority
+of mutations are neutral or decrease fitness. Negative selection acts to remove
+these deleterious mutations with varying degrees of effectiveness, depending on
+the magnitudes of two stochastic evolutionary processes: drift and draft.
+Increasing the rate of either drift or draft inflates the variance in allele
+frequency changes between generations, which decreases the efficiency of
+negative selection. When population sizes are small or genomes are very tightly
+linked, deleterious mutations can persist longer, or even fix in the
+population. Numerous conservation genetics programs attempt to offset this
+problem in XXX, and it has long been known that in artificial selection
+breeding programs that selection can decrease effective population size
+\parencite{Robertson1966-fs}. Furthermore, the role of evolution in determining
+the levels of deleterious mutations in humans has been the subject of much
+study.
+
+processes approximately act to alter the variance in allele frequency change
+between generations.
+
+When
+drift acts alone, deleterious alleles are only guaranteed to be removed by
+selection when they're
+
+a population can only remove mutations stronger than the
+inverse of the population size ($N_e |s| > 1$).
+
 If the human evolutionary process were strongly impacted by Hill--Robertson
@@
 - need for coalescent simulations with reduced $N_e$ along the genome.
+
+-
 
@@
 
+Depending on the number of features, this can be incredibly memory intensive.
+
 \subsection*{Optimization}

### 2023-04-17 manuscript/manuscript.tex
minor ms changes

@@
 \begin{abstract}
+
+Here we define the general background selection process (gBGS) as the polygenic
+selection occurring due to the influx of new mutations that have an average
+negative selection coefficient over their lifetimes. Such mutations are
+destined to either fixation or extinction depending on their selection
+coefficient and magnitudes of two sources of evolutionary stochaticity: genetic
+drift, scaled by population size, and genetic draft, scaled by other ongoing
+selection and the linkage environment. Classic BGS models only consider the
+case where linkage is loose (no selective interference) and mutations are
+sufficiently deleterious that they cannot fix. Studies have used classic BGS
+theory to develop statistical methods to model patterns of genome-wide
+diversity (XXX). Here, we extend this line of work in three ways. First,
+classic BGS theory assumes all deleterious mutations go extinct, and
+consequently cannot accurately model weakly deleterious mutation rates that
+drift up and frequency and in some cases fix in the population. We find that
+this can bias DFE estimates in cases where weak selection constitutes a large
+portion of the DFE. Using the theory of XXX we ameliorate this issue, and find
+our method can provide rough genome-wide estimates of the deleterious mutation
+rate from the signal of linked polygenic selection. Second, since the theory
+predicts diversity levels under weak selection by simultaneously predicting the
+rate at which weakly deleterious mutations are substituted, by fitting our
+model to human genome-wide data, our model predicts substitution rates across
+the genome. We find these predictions substitutions for the features under
+selection match observed
+
+According to theory, the extent of linkage in the genome can limit how quickly
+polygenic selection can occur in artificially selected organisms. Since the BGS
+process is a type of polygenic selection,
 

### 2023-05-04 manuscript/manuscript.tex
minor ms changes
ms updates

--- 2023-03-21 manuscript/manuscript.tex
@@
 
-Previous work in population genomics has established that negative selection
-against against new deleterious variants introduced by mutation profoundly
-shapes large-scale patterns of variation across the genome. However, the
-mathematical models that describe this process only consider variation so
-deleterious that it cannot drift up in frequency and ultimately fix in the
-population. In reality, many new mutations are only mildly deleterious, and can
-fix in the population if random perturbations from genetic drift or selection
-at neighboring sites (genetic draft) overpowers the selection against them.
-Despite the large role negative selection plays across the human genome, no
-existing evolutionary models consider the joint effects of mutation, selection,
-and genetic drift and drift across the genome. Consequently, population genetic
-models cannot describe the ultimate fate of every new variant ---eventual
-extinction or fixation in the population--- and thus cannot connect to observed
-rates of molecular evolution.
-
-whole-genome variance models.
-These are flux models.
-
-This is because, in the very long run, all mutations must be accounted for.
-
-Here, we introduce a new whole-genome general mutation-selection-draft-drift
-(MSD$^2$) model that extends and implements previous theory by
-\textcite{Santiago2016-mu}.
-
-Unlike classic mutation-selection models which cannot accurately capture the
-dynamics of weakly deleterious mutation, MSD$^2$ models can, and consequently
-can provide estimates of the distribution of selection coefficients of new
-mutations. Additionally, because our model incorporates drift and draft, our
-model makes explicit predictions about the substitution rate of weakly
-deleterious mutations and allows us to test XXX.
-
 \end{abstract}
@@
 
-If the human evolutionary process were strongly impacted by Hill--Robertson
-Interference, we should expect that substitution rates are highest in regions
-experiencing the most selection, which are regions when $B(x)$ is the lowest.
-However, our model
-
-Evolutionary processes describe how genetic variation is introduced and
-ultimately converted to substitutions or lost from the population. Our model 
-makes predictions about the eventual outcome of weakly deleterious mutation 
-that seem to roughly match observations.
-
-Note on the importance of selection in humans.
-NOTE about the importance of calibration
-
-emphasis on whole genome models
-
-Eukaryotes genomes are under continual selectionary pressure to remove the
-continual influx of deleterious mutations at conserved regions. This purifying
-selection process keeps mutations that reduce fitness at low frequency in the
-population, and reduces genetic diversity at sites linked to genomic regions
-that are predominantly under purifying selection such as coding sequences and
-regions that have been shown to be phylogenetically conserved between species.
-Evolutionary geneticists have developed a series of mathematical models of this
-process, to predict quantities such as the equilibrium level of segregating
-deleterious mutations (\emph{genetic load}), and the reduction in neutral
-genetic diversity at linked sites due to a faster rate of coalescence (the $B$
-reduction factor). Statistical models of the purifying selection process fit to
-human genomic data closely match observed patterns of genetic variability along
-the genome, explaining around 60\% of the genetic variability at the megabase
-scale \parencite{McVicker2009-ax,Murphy2022-sj}.
-
-These evolutionary genetics models only consider the impact of new mutations
-that are sufficiently deleterious that they are guaranteed to be eliminated
-eventually by selection. When mutations are strongly deleterious, the impact of
-random genetic drift and random perturbations caused by selection at linked
-neighboring sites (genetic draft) on these mutations on their fates can be
-ignored. Under these conditions, mathematical models can assume an stationary
-balance between the input of new mutations entering the population at rate
-$\mu$ per basepair, per generation, and their rate of loss determined by the
-relative reduction in heterozygous fitness, $1-s$. Under these
-mutation-selection balance models, the equilibrium genetic load and the
-reduction in coalescent time are approximately determined by the ratio
-$\nicefrac{\mu}{s}$. However, estimates of the distribution of fitness effects
-suggest many new mutations are weakly deleterious, such that their dynamics can
-no longer ignore the influence of genetic drift and draft.
-
-At every spot in the genome the processes of mutation-selection-draft-drift
-against alleles deleterious in most environments reduce pairwise diversity. We
-can imagine some unobserved reduction map, $B(x)$ across the genome. Here we
-derive theory for this map under these processes and show how statistical
-models can be fit using this theory.
-
-To understand the this process and make predictions about the equilibrium level
-of deleterious mutations segregating in the population, evolutionary
-geneticists have used models of mutation-selection balance. These models
-
-Non-equilibrium models of deleterious mutation flux.
-
-Generalized background selection
-
-These models
+The continual influx of new mutations into populations is the ultimate source
+of all adaptations, but the vast majority of mutations either do not have an
+effect on fitness or are deleterious. Negative selection works to eliminate
+such deleterious mutations from the population, which is essential to
+maintaining the integrity of information in the genome to maintain a lineage.
+Mutations under negative selection are kept at low frequencies (Haldane?),
+contributing to human traits (XXX) and diseases. As with most selection in the
+genome, this process perturbs the allele frequencies of neighboring linked
+variation. This reduces genetic variability around conserved segments and
+creates large-scale spatial patterns in genetic diversity. These patterns are
+mediated by the deleterious mutation rate, the strength of selection in
+conserved segments, and the spatial distribution of recombination rates and
+conserved segments. Since genome-wide recombination maps and putatively
+conserved segments are available for many species, researchers have combined
+these with population genetic theory to estimate the mutation rate and
+selection coefficients of deleterious mutations, as well as the overall
+reduction in genetic variability due to the negative selection process.
+
+Classic Background Selection (BGS) theory is the predominant model for the
+reduction in linked neutral diversity used by statistical approaches to
+quantify the impacts of negative selection. However, this model makes some
+assumptions for mathematical tractability that could distort inferences about
+the negative selection process. First, since fixation probabilities ultimately
+depend on the product of the deleterious selection coefficient (s) and
+population size (N), the efficacy of negative selection should depend on
+demography. Unfortunately, accommodating demography into theoretic negative
+selection models remains an open, difficult problem (XXX). Second, building off
+classic models of mutation-selection balance (Crow 1970, Kimura and Maruyama
+196), the Background Selection (XXX) model assumes that new mutations are
+sufficiently deleterious that they are invariably driven to loss. Under this
+assumption, the effect of selection is well-approximated by simply rescaling
+the neutral coalescent by a reduction factor known as B (XXX). However, this is
+only appropriate in a limited domain of parameters relevant to the negative
+selection process (XXX). Because the BGS model cannot accommodate the
+possibility of fixation for mutations with $Ns \ll 1$, diversity is incorrectly
+predicted to fall to zero as the strength of selection diminishes. This weak
+selection problem is related to the wider problem of modeling the substitution
+rate of deleterious mutation (i.e. the “ratchet” rate). Third, the BGS model
+assumes segments under purifying selection are sufficiently far apart that
+there is no selective interference between them. This is a matter of degree,
+since under sufficiently high mutation rates or with little recombination,
+selective interference can occur with a segment.
+
+In this work, we extend another class of linked selection models that derive
+from quantitative genetics to address limitations of the classic BGS model.
+These models consider how polygenic fitness variation increases the rate of
+stochastic allele frequency change as alleles become randomly linked to fitness
+backgrounds. While these models can theoretically accommodate polygenic fitness
+variation from any source as long as its rate of change is not too rapid, we
+focus specifically on a deleterious-mutations-only model from Santiago and
+Caballero (2016). This model is identical to the classic BGS model when
+selection against deleterious mutations is strong, but it also correctly
+predicts the reduction in diversity and substitutions rates of weakly
+deleterious mutations by accurately modeling levels of genic fitness variation.
+We extend Santiago and Caballero’s (hereafter the SC16) model so that it can be
+used to model genome-wide patterns of diversity under negative selection
+processes, and develop Python software to use this theoretic approach to fit
+genome-wide models of negative selection. Using forward simulations, we show
+this model leads to more accurate DFE estimates under weak selection. While
+this new model accommodates the possibility that a given reduction B could be
+due to weaker selection, in applying our composite-likelihood method to human
+population genomic data, we confirm earlier findings that strong negative
+selection at a small fraction of sites fits human data across populations best.
+Finally, we show that while this model solves some limitations of the classic
+BGS model, simulations reveal another mysterious defect: a poor fit around
+$Ns=1$. We show that this can be partially corrected by incorporating the effects
+of selective interference on genic variation, but the remaining discrepancy is
+due to the build up of negative linkage disequilibrium that cannot be
+adequately captured by this model. We find accounting for genic selective
+interference between segments fits human population genomic data slightly
+better, suggesting levels of negative selection in humans are sufficient to
+perturb the selection at other sites.
 
 \section*{Results}
-
-%\begin{figure}[!htb]
-%  \centering
-%  \includegraphics{figures/experimental_1.pdf}
-%  \caption{}
-%  \label{fig:figure-1}
-%\end{figure}
+\subsection*{Theory}
+
+Similar to previous work, our method characterizes the negative selection
+process through its indirect effects on levels of neighboring pairwise
+diversity at putatively neutral sites. Essential to this approach are accurate
+theoretic models for how a selection process alters pairwise coalescence times
+at sites some recombination distance away. The predominant model for how
+negative (or purifying) selection reduces pairwise coalescence times at linked
+sites is the classic background selection model
+\parencite{Charlesworth1993-gb}. This simple and elegant model assumes new
+deleterious mutations enter a region of $L$-basepairs at a rate of $\mu$ per
+basepair per generation with a selection coefficient $s$ in heterozygotes that
+is sufficiently deleterious that it is destined to loss. Under this setup, the
+number of segregating deleterious mutations in the region reaches a stationary
+Poisson distribution with rate $\nicefrac{U}{s}$ ($U=\mu L$), which is
+precisely the equilibrium number under a deterministic mutation-selection
+balance models \parencite{Kimura1966-bk,Crow1970-wj}. Modifications of this
+basic model that allow for recombination between the segment under negative
+selection and a focal neutral site \parencite{Nordborg1996-nq,Hudson1994-oh}
+have been the basis of statistical methods that fit genome-wide patterns of
+diversity to theoretic expectations \parencite{McVicker2009-ax,Hudson1995-xc}.
+Furthermore, jointly incorporating both hitchhiking and background selection
+have allowed for model comparison approaches that can estimate the relative
+contributions of these two modes of selection
+\parencite{Elyashiv2016-vt,Murphy2022-sj}.
+
+These linked selection models predict how selection increases the rate
+of coalescence at a neighboring neutral site some recombination distance away.
+
+In general, these methods characterize selection processes by fitting the
+relationship between observed pairwise diversity and expected levels according
+to a theoretic model of linked selection. These linked selection models predict
+how selection impacts the coalescence rate at some nearby site, which is
+mediated by the strength of selection (e.g. the rate of new selected
+
+For negative selection, the background
+selection model predicts the expected pairwise coalescence time at some focal
+neutral site near a region under mutation-selection balance. The increase in
+coalescence rate
+
+The impact of selection on linked neutral sites is mediated by the rate at
+which selected mutations enter the population, the strength of selection, and
+the recombination distance between
+
+Both
+hitchhiking and background selection models approximate the increased rate of
+coalescence near sites under positive and negative selection as a reduction in
+effective population size. Hitchhiking models
+
+typically imagine a sweeping
+beneficial allele creating
+
+feature of linked selection theory is that the increased rate of coalescence at
+sites neighboring regions under either positive or negative selection
+
+Both positive
+and negative selection increase the rate of coalescence at neighboring sites,
+which both hitchhiking (XXX) and background selection (XXX) models
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics{figures/figure1.pdf}
+  \caption{}
+  \label{fig:figure-1}
+\end{figure}
 
@@
 selection. The fitness variance equation, $V(x) = (U-2R)s$ tells us that under
-MSD$^$2 models, the long-run variance is set by the balance of mutation and
-fixation. Whether a particular basepair's behavior follows this equation
-depends upon whether the long-run average is close to ``typical" for that
-basepair at a given time. Good things fix so quickly, that we're left with a
-constant churning of new variation
+models, the long-run variance is set by the balance of mutation and fixation.
+Whether a particular basepair's behavior follows this equation depends upon
+whether the long-run average is close to ``typical" for that basepair at a
+given time. Good things fix so quickly, that we're left with a constant
+churning of new variation
 
 - need for coalescent simulations with reduced $N_e$ along the genome.
+
+-
 

### 2023-05-06 manuscript/manuscript.tex
ms updates

@@
 The continual influx of new mutations into populations is the ultimate source
-of all adaptations, but the vast majority of mutations either do not have an
-effect on fitness or are deleterious. Negative selection works to eliminate
-such deleterious mutations from the population, which is essential to
-maintaining the integrity of information in the genome to maintain a lineage.
-Mutations under negative selection are kept at low frequencies (Haldane?),
-contributing to human traits (XXX) and diseases. As with most selection in the
-genome, this process perturbs the allele frequencies of neighboring linked
-variation. This reduces genetic variability around conserved segments and
-creates large-scale spatial patterns in genetic diversity. These patterns are
-mediated by the deleterious mutation rate, the strength of selection in
-conserved segments, and the spatial distribution of recombination rates and
-conserved segments. Since genome-wide recombination maps and putatively
-conserved segments are available for many species, researchers have combined
-these with population genetic theory to estimate the mutation rate and
-selection coefficients of deleterious mutations, as well as the overall
-reduction in genetic variability due to the negative selection process.
+of all adaptations, but the vast majority of mutations either do not affect
+fitness or are deleterious. Negative selection works to eliminate such
+deleterious mutations from the population, which is essential to maintaining
+the integrity of information in the genome to maintain a lineage. Mutations
+under negative selection are kept at low frequencies (Haldane?), contributing
+to human traits (XXX) and diseases. As with most selection in the genome, this
+process perturbs the allele frequencies of neighboring linked variation. This
+reduces genetic variability around conserved segments and creates large-scale
+spatial patterns in genetic diversity. These patterns are mediated by the
+deleterious mutation rate, the strength of selection in conserved segments, and
+the spatial distribution of recombination rates and conserved segments. Since
+genome-wide recombination maps and putatively conserved segments are available
+for many species, researchers have combined these with population genetic
+theory to estimate the mutation rate and selection coefficients of deleterious
+mutations, as well as the overall reduction in genetic variability due to the
+negative selection process.
 
@@
 Similar to previous work, our method characterizes the negative selection
-process through its indirect effects on levels of neighboring pairwise
-diversity at putatively neutral sites. Essential to this approach are accurate
-theoretic models for how a selection process alters pairwise coalescence times
-at sites some recombination distance away. The predominant model for how
-negative (or purifying) selection reduces pairwise coalescence times at linked
-sites is the classic background selection model
-\parencite{Charlesworth1993-gb}. This simple and elegant model assumes new
-deleterious mutations enter a region of $L$-basepairs at a rate of $\mu$ per
-basepair per generation with a selection coefficient $s$ in heterozygotes that
-is sufficiently deleterious that it is destined to loss. Under this setup, the
-number of segregating deleterious mutations in the region reaches a stationary
-Poisson distribution with rate $\nicefrac{U}{s}$ ($U=\mu L$), which is
-precisely the equilibrium number under a deterministic mutation-selection
-balance models \parencite{Kimura1966-bk,Crow1970-wj}. Modifications of this
-basic model that allow for recombination between the segment under negative
-selection and a focal neutral site \parencite{Nordborg1996-nq,Hudson1994-oh}
-have been the basis of statistical methods that fit genome-wide patterns of
-diversity to theoretic expectations \parencite{McVicker2009-ax,Hudson1995-xc}.
-Furthermore, jointly incorporating both hitchhiking and background selection
-have allowed for model comparison approaches that can estimate the relative
-contributions of these two modes of selection
-\parencite{Elyashiv2016-vt,Murphy2022-sj}.
-
-These linked selection models predict how selection increases the rate
-of coalescence at a neighboring neutral site some recombination distance away.
-
-In general, these methods characterize selection processes by fitting the
-relationship between observed pairwise diversity and expected levels according
-to a theoretic model of linked selection. These linked selection models predict
-how selection impacts the coalescence rate at some nearby site, which is
-mediated by the strength of selection (e.g. the rate of new selected
-
-For negative selection, the background
-selection model predicts the expected pairwise coalescence time at some focal
-neutral site near a region under mutation-selection balance. The increase in
-coalescence rate
-
-The impact of selection on linked neutral sites is mediated by the rate at
-which selected mutations enter the population, the strength of selection, and
-the recombination distance between
-
-Both
-hitchhiking and background selection models approximate the increased rate of
-coalescence near sites under positive and negative selection as a reduction in
-effective population size. Hitchhiking models
-
-typically imagine a sweeping
-beneficial allele creating
-
-feature of linked selection theory is that the increased rate of coalescence at
-sites neighboring regions under either positive or negative selection
-
-Both positive
-and negative selection increase the rate of coalescence at neighboring sites,
-which both hitchhiking (XXX) and background selection (XXX) models
+process through its indirect effects on neighboring levels of pairwise
+diversity. While pairwise diversity is a less-informative statistic than the
+full site frequency spectrum, it is far easier to derive models of how
+selection alters expected pairwise coalescence times at linked sites than to
+describe the entire genealogy. Statistical methods then fit observed diversity
+to the expectation of a particular theoretic model of the linked selection
+process. Consequently, statistical inferences about selection may be incorrect
+when the underlying theory inaccurately predicts average pairwise coalescence
+times.
+
+Our work is motivated by a well-known limitation of classic background
+selection theory: it is only accurate for strongly deleterious mutations. Given
+that statistical methods rely on background selection theory to estimate the
+mutation rate and distribution of fitness effects of deleterious mutations, it
+is important to characterize and remedy these shortcomings of the model.
+However, this is
+
+It is well-known that the approximations that underpin classic background
+selection theory
+\parencite{Charlesworth1993-gb,Nordborg1996-nq,Hudson1994-oh,Hudson1995-pt}
+break down when selection against deleterious mutations is weak, and the
+reduction in diversity predicted by this model is incorrect
+\parencite{McVean2000-bt,Good2013-lp,Gordo2002-dr}. Before introducing how th
+
+describe how selection alters pairwise coalescence times rather than full
+genealogy (since $\E[\pi] = 2\mu \E(T_2)$). Statistical methods that fit the
+relationship between observed diversity to an expectation provided by a
+particular theoretic model of the linked selection process.
+
+Our work considers a well-known
+shortcoming of the classic background selection model: that
+
+The classic theoretic model of how negative or purifying selection changes the
+mean coalescence time at linked neutral sites is the background selection model
+\parencite{Charlesworth1993-gb}. In the simplest case without recombination,
+this model assumes that deleterious mutations enter a segment.
+
+classic background selection model are only applicable for strongly deleterious
+mutations that have no probability of fixation
+\parencite{Good2014-yz,Gordo2002-dr,Charlesworth2013-kl}.
+
+new deleterious mutations enter a region of $L$-basepairs at a rate of $\mu$
+per basepair per generation with a selection coefficient $s$ in heterozygotes
+that is sufficiently deleterious that it is destined to loss. Under this setup,
+the number of segregating deleterious mutations in the region reaches a
+stationary Poisson distribution with rate $\nicefrac{U}{s}$ ($U=\mu L$), which
+is precisely the equilibrium number under a deterministic mutation-selection
+balance models \parencite{Kimura1966-bk,Crow1970-wj}. Since any linear carrying
+a deleterious mutation is destined to extinction, coalescence events can only
+occur in the $N \exp(-\nicefrac{U}{s})$ individuals carrying zero deleterious
+mutations. Working back through the ancestors of two lineages carrying $i$ and
+$j$ deleterious mutations, both
+
+Modifications of this basic model that allow for recombination between the
+segment under negative selection and a focal neutral site
+\parencite{Nordborg1996-nq,Hudson1994-oh} have been the basis of statistical
+methods that fit genome-wide patterns of diversity to theoretic expectations
+\parencite{McVicker2009-ax,Hudson1995-xc}. Furthermore, methods that
+incorporate both hitchhiking and background selection allow for estimates of
+the relative contributions of these two modes of selection and model
+comparisions \parencite{Elyashiv2016-vt,Murphy2022-sj}.
+
+However, the predicted reduction in diversity under the classic background
+selection model is inaccurate whenever selection is weak enough that
+deleterious mutations may rise in frequency in the population and possibly
+reach fixation. This is a well-known problem (XXX), that is mathematically
+difficult because the
+
+\subsection*{DFE Estimates}
+ - Table s5 https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/genetics/206/1/10.1534_genetics.116.197145/8/files1.pdf?Expires=1686357613&Signature=cp9qoI2ChCb6GFZ74HI4eWcA~MQ4sHlWyeI3Zp49-5xEORzMfYlSRxzVvZlo8ZoGWDi0G3PnL00ZplM-~or4YIVm2uhkmqjf61AmC7RdZ1QTak4j0kFMQ0wAbM9pusiV3gIepTQHq2HKRKwgS~ypmv~0oqO5dDxaq9nhOWIBBmLAA49oHWuxUtnRe3sEYATwCwjIFEomhgJpg7vKGCF~87GdCV4JETW4PN2VvD4hsu2AVRd8yA~vUJvsXfhWwcElDeY4fsgpprl1lC3XGjJCnqdwbzRGPV5LNx5XEKLptcZtLeLcOmFJCvmVLe0bewJ7Vta0TSQM7T9AuklompjPPw__&Key-Pair-Id=APKAIE5G5CRDK6RD3PGA
 

### 2023-05-06 manuscript/manuscript.tex
ms updates

@@
 \section*{Results}
+
+Similar to previous work, our method characterizes the negative selection
+process through the signal left on pairwise diversity at linked sites. Although
+pairwise diversity is a less-informative statistic than the full site frequency
+spectrum, it is much easier to derive models of how selection alters expected
+pairwise coalescence times at linked sites than to describe the entire
+genealogy. Spatial patterns of pairwise diversity are highly variable across
+eukaryotic genomes, as linked selection leads diversity to become correlated
+with spatial heterogeneity in recombination rates and the density of conserved
+sites. Statistical methods then fit observed diversity to the expected level
+under a particular theoretic model of linked selection processes. Consequently,
+statistical inferences about selection may be incorrect when the underlying
+theory inaccurately predicts average pairwise coalescence times.
+
+Our work is motivated by a well-known limitation of classic background
+selection theory: it is only accurate for strongly deleterious mutations
+\parencite{Charlesworth1993-gb,McVean2000-bt,Good2013-lp,Gordo2002-dr}. Given
+statistical methods that fit genome-wide diversity to linked selection
+processes rely on the classic BGS model to estimate the distribution of fitness
+effects and rate of deleterious mutations, it is important to characterize and
+remedy these shortcomings. However, extending background selection theory to
+model weak negative selection is challenging, and it is worthwhile to develop
+some intuition for how the classic background selection works and why this is
+the case.
+
+The classic background selection model \parencite{Charlesworth1993-gb}
+
+First, the frequency trajectory of a weakly selected allele is
+predominantly determined by stochastic evolutionary processes. When the changes
+due to stochasticity are large, deterministic approximations that work well
+under strong selection no longer capture selection dynamics. Second, weakly
+selected alleles are impacted by two sources of evolutionary stochasticity:
+genetic drift (i.e. non-heritable differences in offspring number and Mendelian
+segregation) and genetic draft, the stochastic changes in allele frequency
+caused by selection on the same genetic background
+\parencite{Neher2013-dz,McVean2000-bt}.
+
+Moreover, when the magnitude of stochastic changes can exceed the deterministic
+changes caused by selection, mutations that decrease fitness now have a chance
+of fixation \parencite{Kimura1962-su,Malecot1948-zv}. Because there is a steady
+influx of weakly deleterious mutations, each with a non-zero chance of
+substitution, there must be a corresponding substitution process. The rate of
+this deleterious substitution process (e.g. ``Muller's ratchet") is
+
+By contrast, when selection is assumed to be strong,
+deterministic models of mutation and selection that ignore these sources of
+stochasticity adequately capture the dynamics. Since weakly deleterious
+mutations have some chance of fixation,
+
 \subsection*{Theory}
-
-Similar to previous work, our method characterizes the negative selection
-process through its indirect effects on neighboring levels of pairwise
-diversity. While pairwise diversity is a less-informative statistic than the
-full site frequency spectrum, it is far easier to derive models of how
-selection alters expected pairwise coalescence times at linked sites than to
-describe the entire genealogy. Statistical methods then fit observed diversity
-to the expectation of a particular theoretic model of the linked selection
-process. Consequently, statistical inferences about selection may be incorrect
-when the underlying theory inaccurately predicts average pairwise coalescence
-times.
-
-Our work is motivated by a well-known limitation of classic background
-selection theory: it is only accurate for strongly deleterious mutations. Given
-that statistical methods rely on background selection theory to estimate the
-mutation rate and distribution of fitness effects of deleterious mutations, it
-is important to characterize and remedy these shortcomings of the model.
-However, this is
-
-It is well-known that the approximations that underpin classic background
-selection theory
-\parencite{Charlesworth1993-gb,Nordborg1996-nq,Hudson1994-oh,Hudson1995-pt}
-break down when selection against deleterious mutations is weak, and the
-reduction in diversity predicted by this model is incorrect
-\parencite{McVean2000-bt,Good2013-lp,Gordo2002-dr}. Before introducing how th
 

### 2023-05-12 manuscript/manuscript.tex
more theory section stuff

@@
 
+\subsection*{Theory}
+
 Our work is motivated by a well-known limitation of classic background
@@
 \parencite{Charlesworth1993-gb,McVean2000-bt,Good2013-lp,Gordo2002-dr}. Given
-statistical methods that fit genome-wide diversity to linked selection
-processes rely on the classic BGS model to estimate the distribution of fitness
-effects and rate of deleterious mutations, it is important to characterize and
-remedy these shortcomings. However, extending background selection theory to
-model weak negative selection is challenging, and it is worthwhile to develop
-some intuition for how the classic background selection works and why this is
-the case.
-
-The classic background selection model \parencite{Charlesworth1993-gb}
-
-First, the frequency trajectory of a weakly selected allele is
-predominantly determined by stochastic evolutionary processes. When the changes
-due to stochasticity are large, deterministic approximations that work well
-under strong selection no longer capture selection dynamics. Second, weakly
-selected alleles are impacted by two sources of evolutionary stochasticity:
-genetic drift (i.e. non-heritable differences in offspring number and Mendelian
-segregation) and genetic draft, the stochastic changes in allele frequency
-caused by selection on the same genetic background
-\parencite{Neher2013-dz,McVean2000-bt}.
-
-Moreover, when the magnitude of stochastic changes can exceed the deterministic
-changes caused by selection, mutations that decrease fitness now have a chance
-of fixation \parencite{Kimura1962-su,Malecot1948-zv}. Because there is a steady
-influx of weakly deleterious mutations, each with a non-zero chance of
-substitution, there must be a corresponding substitution process. The rate of
-this deleterious substitution process (e.g. ``Muller's ratchet") is
-
-By contrast, when selection is assumed to be strong,
-deterministic models of mutation and selection that ignore these sources of
-stochasticity adequately capture the dynamics. Since weakly deleterious
-mutations have some chance of fixation,
-
-\subsection*{Theory}
-
-describe how selection alters pairwise coalescence times rather than full
-genealogy (since $\E[\pi] = 2\mu \E(T_2)$). Statistical methods that fit the
-relationship between observed diversity to an expectation provided by a
-particular theoretic model of the linked selection process.
-
-Our work considers a well-known
-shortcoming of the classic background selection model: that
-
-The classic theoretic model of how negative or purifying selection changes the
-mean coalescence time at linked neutral sites is the background selection model
-\parencite{Charlesworth1993-gb}. In the simplest case without recombination,
-this model assumes that deleterious mutations enter a segment.
+statistical methods that fit linked selection processes to genome-wide
+diversity data rely on the classic BGS model to estimate the distribution of
+fitness effects and rate of deleterious mutations, it is important to
+characterize and remedy these shortcomings. However, extending background
+selection theory to model weak negative selection is challenging, and it is
+worthwhile to develop some intuition for why this is the case, how the classic
+background selection works, and how the model of \textcite{Santiago2016-mu}
+solves these problems and can be extended to genome-wide inference.
+
+Both the classic background selection and SC16 models imagine a negative
+selection process where a continual influx of deleterious mutations flow into
+the population at a rate of $\mu$ per basepair per generation in a conserved
+region of $L$ basepairs, such that the region-wide per generation mutation rate
+is $U = \mu L$. Each mutation imposes an additive selective cost of $s$ in
+heterozygotes and $2s$ in homozygotes. The classic background selection model
+considers when mutations are sufficiently deleterious that they are destined to
+extinction
+\parencite{Charlesworth1993-gb,Nordborg1996-nq,Hudson1995-pt,Hudson1994-oh}. In
+this case, the genealogy is well-approximated by a neutral coalescence process
+with a rescaled effective population size of $N_0 = N \exp(-\nicefrac{U}{s})$.
+This approximation works reasonably well for two reasons (though see
+\textcite{Cvijovic2018-vd}). First, when selection is strong, the number of
+deleterious mutations per haplotype has a stationary Poisson distribution with
+mean equal to the equilibrium under the deterministic mutation-selection model
+\parencite{Haldane1927-ga}. Since lineages carrying any strongly deleterious
+mutations cannot contribute to the genealogy, coalescence events only occur in
+the fraction $\exp(-\nicefrac{U}{s}$ free of mutations. Second, two present-day
+lineages carrying any deleterious mutations trace their ancestry back to the
+pool of $N_0$ mutation-free ancestors very quickly, since these harmful
+mutations could only have been acquired recently by the lineage. On coalescence
+timescales of $\sim 2N_0$, this time is negligible and can be ignored
+(\textcite{Durrett2008-ql}, p. 213; \textcite{Good2014-yz}).
 
@@
 \parencite{Good2014-yz,Gordo2002-dr,Charlesworth2013-kl}.
-
-new deleterious mutations enter a region of $L$-basepairs at a rate of $\mu$
-per basepair per generation with a selection coefficient $s$ in heterozygotes
-that is sufficiently deleterious that it is destined to loss. Under this setup,
-the number of segregating deleterious mutations in the region reaches a
-stationary Poisson distribution with rate $\nicefrac{U}{s}$ ($U=\mu L$), which
-is precisely the equilibrium number under a deterministic mutation-selection
-balance models \parencite{Kimura1966-bk,Crow1970-wj}. Since any linear carrying
-a deleterious mutation is destined to extinction, coalescence events can only
-occur in the $N \exp(-\nicefrac{U}{s})$ individuals carrying zero deleterious
-mutations. Working back through the ancestors of two lineages carrying $i$ and
-$j$ deleterious mutations, both
 

### 2023-05-12 manuscript/manuscript.tex
more theory section stuff

@@
 This approximation works reasonably well for two reasons (though see
-\textcite{Cvijovic2018-vd}). First, when selection is strong, the number of
-deleterious mutations per haplotype has a stationary Poisson distribution with
-mean equal to the equilibrium under the deterministic mutation-selection model
-\parencite{Haldane1927-ga}. Since lineages carrying any strongly deleterious
-mutations cannot contribute to the genealogy, coalescence events only occur in
-the fraction $\exp(-\nicefrac{U}{s}$ free of mutations. Second, two present-day
-lineages carrying any deleterious mutations trace their ancestry back to the
-pool of $N_0$ mutation-free ancestors very quickly, since these harmful
-mutations could only have been acquired recently by the lineage. On coalescence
-timescales of $\sim 2N_0$, this time is negligible and can be ignored
-(\textcite{Durrett2008-ql}, p. 213; \textcite{Good2014-yz}).
-
-classic background selection model are only applicable for strongly deleterious
-mutations that have no probability of fixation
-\parencite{Good2014-yz,Gordo2002-dr,Charlesworth2013-kl}.
-
-Modifications of this basic model that allow for recombination between the
-segment under negative selection and a focal neutral site
-\parencite{Nordborg1996-nq,Hudson1994-oh} have been the basis of statistical
-methods that fit genome-wide patterns of diversity to theoretic expectations
-\parencite{McVicker2009-ax,Hudson1995-xc}. Furthermore, methods that
-incorporate both hitchhiking and background selection allow for estimates of
-the relative contributions of these two modes of selection and model
-comparisions \parencite{Elyashiv2016-vt,Murphy2022-sj}.
-
-However, the predicted reduction in diversity under the classic background
-selection model is inaccurate whenever selection is weak enough that
-deleterious mutations may rise in frequency in the population and possibly
-reach fixation. This is a well-known problem (XXX), that is mathematically
-difficult because the
+\textcite{Cvijovic2018-vd,Walczak2012-fi,Nicolaisen2012-vs}). First, when
+selection is strong, the number of deleterious mutations per haplotype has a
+stationary Poisson distribution with an average equal to the equilibrium under
+the deterministic mutation-selection model \parencite{Haldane1927-ga}. Since
+lineages carrying any strongly deleterious mutations cannot contribute to the
+genealogy, coalescence events only occur in the fraction
+$\exp(-\nicefrac{U}{s}$ of the population free of mutations. Second, the time
+it takes for a present-day lineage to trace their ancestry back to a
+mutation-free ancestor (the ``delay phase") is negligible compared to the
+coalescence timescale of $N_0$ among these ancestors (the ``coalescence
+phase"), and can be ignored (\textcite{Durrett2008-ql}, p. 213;
+\textcite{Good2014-yz}).
+
+When mutations are weakly deleterious, their frequency trajectories are no
+longer well-approximated by deterministic models. This is because the selection
+on an allele is only weak relative to the scale of stochastic perturbations it
+encounters along its trajectory. When genetic drift is the only source of
+randomness in allele frequency change, the relevant scale of stochastic
+perturbations is the drift-effective population size $N$ and selection dynamics
+are governed by compound parameter $2Ns$. When $2Ns \gg 1$, the stochastic
+component of allele frequency change is minuscule relative to selection and can
+be ignored as the classic BGS theory does; this is why the rescaling factor
+under the BGS model does not depend on $N$. However, when $2Ns \le 1$
+stochastic perturbations are large relative to selection, and a weakly
+deleterious allele can drift up to intermediate frequencies before eventual
+loss or fixation. Weakly deleterious alleles can persist in the population much
+longer, and consequently
+
+When this occurs, the distribution of the number of deleterious mutations (and
+its corresponding fitness distribution) are no longer stationary and become a
+traveling wave \parencite{Rouzine2008-qz,Good2013-lp}. Furthermore, individuals
+in the least-loaded class are stochastically lost, fixing those mutations with
+a click of ``Muller's ratchet" \parencite{Muller1964-ki}. Determining both the
+rate of Muller's ratchet \parencite{Haigh1978-gt,Gordo2002-dr,Gessler1995-hz}
+and the shape of genealogies under weak selection
+\parencite{Przeworski1999-mb,OFallon2010-my,Higgs1995-xc} have been stubborn
+problems
 

### 2023-05-12 manuscript/manuscript.tex
more theory section stuff

@@
 
-When mutations are weakly deleterious, their frequency trajectories are no
-longer well-approximated by deterministic models. This is because the selection
-on an allele is only weak relative to the scale of stochastic perturbations it
-encounters along its trajectory. When genetic drift is the only source of
-randomness in allele frequency change, the relevant scale of stochastic
-perturbations is the drift-effective population size $N$ and selection dynamics
-are governed by compound parameter $2Ns$. When $2Ns \gg 1$, the stochastic
-component of allele frequency change is minuscule relative to selection and can
-be ignored as the classic BGS theory does; this is why the rescaling factor
-under the BGS model does not depend on $N$. However, when $2Ns \le 1$
-stochastic perturbations are large relative to selection, and a weakly
-deleterious allele can drift up to intermediate frequencies before eventual
-loss or fixation. Weakly deleterious alleles can persist in the population much
-longer, and consequently
-
-When this occurs, the distribution of the number of deleterious mutations (and
-its corresponding fitness distribution) are no longer stationary and become a
-traveling wave \parencite{Rouzine2008-qz,Good2013-lp}. Furthermore, individuals
-in the least-loaded class are stochastically lost, fixing those mutations with
-a click of ``Muller's ratchet" \parencite{Muller1964-ki}. Determining both the
-rate of Muller's ratchet \parencite{Haigh1978-gt,Gordo2002-dr,Gessler1995-hz}
-and the shape of genealogies under weak selection
-\parencite{Przeworski1999-mb,OFallon2010-my,Higgs1995-xc} have been stubborn
-problems
+When mutations are weakly deleterious, their allele frequencies are strongly
+influenced by stochastic perturbations and their frequency trajectories are no
+longer well-approximated by deterministic models. When genetic drift is the
+only source of randomness in allele frequency change, the relevant scale of
+stochastic perturbations is determined by the drift-effective population size
+$N_d$ (e.g. due to only non-selective process). For reasons that will soon be
+clear, we consider a general expression of the drift-effective population size
+that allows for extra non-heritable variability in offspring number $V_k$,
+
+\begin{align}
+    N_d = \frac{2N}{V_k + 2}
+\end{align}
+
+where if $V_k = 2$, this reduces to the classic Wright--Fisher model
+\parencite{Wright1938-tv}.
+
+The insight of nearly neutral theory (XXX) was that the qualitative behavior of
+selected sites is determined by the compound parameter $2N_d s$. When $2N_d s
+\gg 1$, the stochastic component of allele frequency change is minuscule
+relative to selection and can be ignored; this is why the factor that rescales
+the drift-effective population size under the BGS model does not depend on
+$N_d$. However, when $2N_d s \le 1$ stochastic perturbations are on a scale
+equal to or greater than the changes due to selection. In this regime, weakly
+deleterious alleles can drift up to intermediate frequencies before their
+eventual loss or fixation. When this occurs, classic BGS theory breaks down for
+several reasons. First, the ``delay phase" is no longer negligible as lineages
+carrying weakly deleterious mutations can persist on coalescent timescales.
+This distorts genealogies away those expected under neutrality
+\parencite{Przeworski1999-mb,OFallon2010-my,Higgs1995-xc}. Second, the
+distribution of the number of deleterious mutations (and its corresponding
+fitness distribution) are no longer stationary and become a traveling wave
+\parencite{Rouzine2008-qz,Good2013-lp,Gessler1995-hz} towards reduced
+population fitness (since beneficial mutations are ignored). Consequently,
+individuals in the least-loaded class are stochastically lost, fixing those
+mutations with a click of ``Muller's ratchet" \parencite{Muller1964-ki}.
+Determining both the rate of Muller's ratchet
+\parencite{Haigh1978-gt,Gordo2002-dr,Gessler1995-hz} and the shape of
+genealogies under weak selection have been stubborn open problems in
+evolutionary genetics.
+
+To further complicate matters, both strong and weakly deleterious mutations
+create \emph{heritable} fitness variation. In turn, the presence of heritable
+fitness variation generates another source of randomness in allele frequency
+change known as genetic \emph{draft} \parencite{Neher2013-dz}. Genetic draft
+occurs when alleles become randomly associated with different fitness
+backgrounds each generation, which stochastically perturbs their trajectories.
+Like genetic drift, perturbations due to draft are directionless but overall
+increase the variance in allele frequency change. Unlike drift, the stochastic
+perturbations each generation due to draft are autocorrelated as long as
+segregation and recombination do not decouple an allele from its fitness
+background and the selective environment does not change
+\parencite{Robertson1961-ho,Santiago1995-hx,Buffalo2019-qs}. When selection is
+strong,
+
+When selection is
+weak and fitness is polygenic, the stochasticity introduced by draft is
+well-approximated by rescaling the drift-effective population size
+
+that exist when there are heritable differences among haplotypes.
+Mathematical models of genetic draft
+
+Gillespie2000-mh
+
+,Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Barton2000-zg
 

### 2023-05-13 manuscript/manuscript.tex
more theory section stuff

@@
 that allows for extra non-heritable variability in offspring number $V_k$,
-
+%
 \begin{align}
@@
 Like genetic drift, perturbations due to draft are directionless but overall
-increase the variance in allele frequency change. Unlike drift, the stochastic
-perturbations each generation due to draft are autocorrelated as long as
-segregation and recombination do not decouple an allele from its fitness
-background and the selective environment does not change
+increase the variance in allele frequency change. Unlike drift, stochastic
+perturbations caused by draft are autocorrelated across generations as long an
+allele does not recombine or segregate off its fitness background and the
+selective environment does not change
 \parencite{Robertson1961-ho,Santiago1995-hx,Buffalo2019-qs}. When selection is
-strong,
-
-When selection is
-weak and fitness is polygenic, the stochasticity introduced by draft is
-well-approximated by rescaling the drift-effective population size
-
-that exist when there are heritable differences among haplotypes.
-Mathematical models of genetic draft
-
-Gillespie2000-mh
-
-,Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Barton2000-zg
+strong, draft can generate multiple-merger coalescences that create
+discontinuities in allele frequency trajectories, making diffusion
+approximations ill-suited \parencite{Gillespie2000-mh,Der2011-it,Neher2013-dz}.
+However, when selection is weak and fitness is polygenic, the stochasticity
+introduced by draft can be approximated by assuming draft just increases the
+rate of drift. This is often expressed as rescaling the effective population
+size
+
+Though there are subtle differences in the effective population sizes
+
+rescaling the
+drift-effective population size by an appropriate factor
+\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Barton2000-zg}.
+While a simple rescaling cannot elucidate details about how selection alters
+the genealogy, it does allow for the calculation of key quantities like the
+pairwise coalescence time (though see \textcite{Santiago1998-bs}, p. 2111).
 
 \subsection*{DFE Estimates}
- - Table s5 https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/genetics/206/1/10.1534_genetics.116.197145/8/files1.pdf?Expires=1686357613&Signature=cp9qoI2ChCb6GFZ74HI4eWcA~MQ4sHlWyeI3Zp49-5xEORzMfYlSRxzVvZlo8ZoGWDi0G3PnL00ZplM-~or4YIVm2uhkmqjf61AmC7RdZ1QTak4j0kFMQ0wAbM9pusiV3gIepTQHq2HKRKwgS~ypmv~0oqO5dDxaq9nhOWIBBmLAA49oHWuxUtnRe3sEYATwCwjIFEomhgJpg7vKGCF~87GdCV4JETW4PN2VvD4hsu2AVRd8yA~vUJvsXfhWwcElDeY4fsgpprl1lC3XGjJCnqdwbzRGPV5LNx5XEKLptcZtLeLcOmFJCvmVLe0bewJ7Vta0TSQM7T9AuklompjPPw__&Key-Pair-Id=APKAIE5G5CRDK6RD3PGA
+% - Table s5 https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/genetics/206/1/10.1534_genetics.116.197145/8/files1.pdf?Expires=1686357613&Signature=cp9qoI2ChCb6GFZ74HI4eWcA~MQ4sHlWyeI3Zp49-5xEORzMfYlSRxzVvZlo8ZoGWDi0G3PnL00ZplM-~or4YIVm2uhkmqjf61AmC7RdZ1QTak4j0kFMQ0wAbM9pusiV3gIepTQHq2HKRKwgS~ypmv~0oqO5dDxaq9nhOWIBBmLAA49oHWuxUtnRe3sEYATwCwjIFEomhgJpg7vKGCF~87GdCV4JETW4PN2VvD4hsu2AVRd8yA~vUJvsXfhWwcElDeY4fsgpprl1lC3XGjJCnqdwbzRGPV5LNx5XEKLptcZtLeLcOmFJCvmVLe0bewJ7Vta0TSQM7T9AuklompjPPw__&Key-Pair-Id=APKAIE5G5CRDK6RD3PGA
 

### 2023-05-13 manuscript/manuscript.tex
more theory section stuff

@@
 clear, we consider a general expression of the drift-effective population size
-that allows for extra non-heritable variability in offspring number $V_k$,
+that allows for extra non-heritable variability in offspring number $V^2$,
 %
 \begin{align}
-    N_d = \frac{2N}{V_k + 2}
-\end{align}
-
-where if $V_k = 2$, this reduces to the classic Wright--Fisher model
+    N_d = \frac{2N}{V^2 + 2}
+\end{align}
+
+where if $V^2 = 2$, this reduces to the classic Wright--Fisher model
 \parencite{Wright1938-tv}.
@@
 approximations ill-suited \parencite{Gillespie2000-mh,Der2011-it,Neher2013-dz}.
-However, when selection is weak and fitness is polygenic, the stochasticity
-introduced by draft can be approximated by assuming draft just increases the
-rate of drift. This is often expressed as rescaling the effective population
-size
-
-Though there are subtle differences in the effective population sizes
+However, when selection is weak and polygenic, draft can be approximated as an
+increase in the rate of drift. The key insight is that the draft rescaling
+factor is like extra variability in offspring number, but because of
+autocorrelation due to linkage and similar environments, the heritable fitness
+variation $C^2$ becomes inflated by a factor $Q^2$,
+
+\begin{align}
+    N_e = \frac{N}{C^2 Q^2 + 1}.
+\end{align}
+
+(c.f. \cite{Robertson1961-ho,Santiago1995-hx}; see XXX for a proof).
+
+Note that assuming $V^2=2$.
+
+are subtle differences in the effective population sizes
+
+This is often
+expressed as rescaling the effective population size by
 

### 2023-05-13 manuscript/manuscript.tex
more theory

@@
 
-(c.f. \cite{Robertson1961-ho,Santiago1995-hx}; see XXX for a proof).
-
-Note that assuming $V^2=2$.
-
-are subtle differences in the effective population sizes
-
-This is often
-expressed as rescaling the effective population size by
-
-rescaling the
-drift-effective population size by an appropriate factor
+(c.f. \cite{Robertson1961-ho,Santiago1995-hx}; see XXX for a proof). Note that
+this assumes the usual Wright--Fisher non-heritable variance, i.e. $V^2=2$.
+However, because the autocorrelation in allele frequency builds up each
+generation relative to the fitness background its on, the inflation term $Q^2$
+has a temporal dimension. For notational efficiency, we assume $Q^2$ refers to
+the asymptotic $Q_t^2$ as $t \to \infty$ and we consider the \emph{asymptotic}
+$N_e$ created by draft. In general, however, statistics such as heterozygosity
+depend on the cumulative autocorrelation factor \emph{up to some time} $t$,
+$Q_t^2$ \parencite[p. 2111]{Santiago1998-bs}. This reflects the fact that
+effective population size experienced by a mutation newly associated with a
+fitness background is larger than the asymptotic effective population size,
+since not much autocorrelation has accumulated in the stochastic perturbations
+caused by drift yet.
+
+rescaling the drift-effective population size by an appropriate factor
 \parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Barton2000-zg}.

### 2023-05-13 manuscript/manuscript.tex
first para

--- 2023-05-13 manuscript/manuscript.tex
@@
 of all adaptations, but the vast majority of mutations either do not affect
-fitness or are deleterious. Negative selection works to eliminate such
-deleterious mutations from the population, which is essential to maintaining
-the integrity of information in the genome to maintain a lineage. Mutations
-under negative selection are kept at low frequencies (Haldane?), contributing
-to human traits (XXX) and diseases. As with most selection in the genome, this
+fitness or are deleterious. 
+Natural selection works to eliminate these deleterious mutations from the population,
+thus we expect them to appear at low frequencies (Haldane?Muller?) within populations
+and be less likely to fix between lineages. 
+The genome itself is the product of over a billion years of evolutionary optimization,
+so even though they are rare within populations, we expect that deleterious variants
+should be the overwhelming majority of segregating variation in functional regions,
+and be responsible for a significant portion of phenotypic variation and disease (XX cites XX),
+
+As with most selection in the genome, selection on deleterious variation
 process perturbs the allele frequencies of neighboring linked variation. This
@@
 that allows for extra non-heritable variability in offspring number $V_k$,
-%
+
 \begin{align}
@@
 Like genetic drift, perturbations due to draft are directionless but overall
-increase the variance in allele frequency change. Unlike drift, stochastic
-perturbations caused by draft are autocorrelated across generations as long an
-allele does not recombine or segregate off its fitness background and the
-selective environment does not change
+increase the variance in allele frequency change. Unlike drift, the stochastic
+perturbations each generation due to draft are autocorrelated as long as
+segregation and recombination do not decouple an allele from its fitness
+background and the selective environment does not change
 \parencite{Robertson1961-ho,Santiago1995-hx,Buffalo2019-qs}. When selection is
-strong, draft can generate multiple-merger coalescences that create
-discontinuities in allele frequency trajectories, making diffusion
-approximations ill-suited \parencite{Gillespie2000-mh,Der2011-it,Neher2013-dz}.
-However, when selection is weak and fitness is polygenic, the stochasticity
-introduced by draft can be approximated by assuming draft just increases the
-rate of drift. This is often expressed as rescaling the effective population
-size
-
-Though there are subtle differences in the effective population sizes
-
-rescaling the
-drift-effective population size by an appropriate factor
-\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Barton2000-zg}.
-While a simple rescaling cannot elucidate details about how selection alters
-the genealogy, it does allow for the calculation of key quantities like the
-pairwise coalescence time (though see \textcite{Santiago1998-bs}, p. 2111).
+strong,
+
+When selection is
+weak and fitness is polygenic, the stochasticity introduced by draft is
+well-approximated by rescaling the drift-effective population size
+
+that exist when there are heritable differences among haplotypes.
+Mathematical models of genetic draft
+
+Gillespie2000-mh
+
+,Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Barton2000-zg
 
 \subsection*{DFE Estimates}
-% - Table s5 https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/genetics/206/1/10.1534_genetics.116.197145/8/files1.pdf?Expires=1686357613&Signature=cp9qoI2ChCb6GFZ74HI4eWcA~MQ4sHlWyeI3Zp49-5xEORzMfYlSRxzVvZlo8ZoGWDi0G3PnL00ZplM-~or4YIVm2uhkmqjf61AmC7RdZ1QTak4j0kFMQ0wAbM9pusiV3gIepTQHq2HKRKwgS~ypmv~0oqO5dDxaq9nhOWIBBmLAA49oHWuxUtnRe3sEYATwCwjIFEomhgJpg7vKGCF~87GdCV4JETW4PN2VvD4hsu2AVRd8yA~vUJvsXfhWwcElDeY4fsgpprl1lC3XGjJCnqdwbzRGPV5LNx5XEKLptcZtLeLcOmFJCvmVLe0bewJ7Vta0TSQM7T9AuklompjPPw__&Key-Pair-Id=APKAIE5G5CRDK6RD3PGA
+ - Table s5 %https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/genetics/206/1/10.1534_genetics.116.197145/8/files1.pdf?Expires=1686357613&Signature=cp9qoI2ChCb6GFZ74HI4eWcA~MQ4sHlWyeI3Zp49-5xEORzMfYlSRxzVvZlo8ZoGWDi0G3PnL00ZplM-~or4YIVm2uhkmqjf61AmC7RdZ1QTak4j0kFMQ0wAbM9pusiV3gIepTQHq2HKRKwgS~ypmv~0oqO5dDxaq9nhOWIBBmLAA49oHWuxUtnRe3sEYATwCwjIFEomhgJpg7vKGCF~87GdCV4JETW4PN2VvD4hsu2AVRd8yA~vUJvsXfhWwcElDeY4fsgpprl1lC3XGjJCnqdwbzRGPV5LNx5XEKLptcZtLeLcOmFJCvmVLe0bewJ7Vta0TSQM7T9AuklompjPPw__&Key-Pair-Id=APKAIE5G5CRDK6RD3PGA
 
@@
 and our B' maps is at the theoretic minimum. This minimum variance is entirely
-from the coalescent noise when estimating $\bar{B}\text{sim}(x), which
+from the coalescent noise when estimating $\bar{B}\text{sim}(x)$, which
 
@@
 Natural Selection (equation 2, \cite{Santiago2016-mu}, equation 15A
-\cite{Garcia-Dorado2007-jj}). A more formal derivation follows from
-\textcite{Higgs1995-xc} (though they assume a haploid asexual system, the end
+%`'\cite{Garcia-Dorado2007-jj}). 
+A more formal derivation follows from \textcite{Higgs1995-xc} 
+(though they assume a haploid asexual system, the end
 result is the same up to a factor of two). In their work, they derive equations
@@
 
-\printbibliography
-
+%\printbibliography
 \end{document}

### 2023-05-13 manuscript/manuscript.tex
para 2

@@
 
-As with most selection in the genome, selection on deleterious variation
-process perturbs the allele frequencies of neighboring linked variation. This
-reduces genetic variability around conserved segments and creates large-scale
-spatial patterns in genetic diversity. These patterns are mediated by the
-deleterious mutation rate, the strength of selection in conserved segments, and
-the spatial distribution of recombination rates and conserved segments. Since
-genome-wide recombination maps and putatively conserved segments are available
-for many species, researchers have combined these with population genetic
-theory to estimate the mutation rate and selection coefficients of deleterious
-mutations, as well as the overall reduction in genetic variability due to the
-negative selection process.
+Selection on both beneficial and deleterious alleles leads to 
+perturbations in the allele frequency spectrum of neighboring, linked variation.
+This phenomenon, known as linked selection, is a major force shaping patterns of
+genetic variation along chromosomes \citep{Maynard_Smith1974-zr,Charlesworth1993-gb,Barton1998-aj,Gillespie2000-mh}.
+As deleterious variation is clustered in functional portions of the genome,
+we expect linked selection to reduce levels of diversity surrounding evolutionarily
+conserved segments (e.g. genes, regulatory elements, etc.) 
+and to shape large-scale spatial patterns in genetic diversity. 
+This landscape of diversity is the product of a complex interplay between the
+deleterious mutation rate, the local recombination rate, the strength of selection, and
+the spatial distribution of conserved segments along a chromosome. 
+Since genome-wide recombination maps and putatively conserved segments are available
+for many species, there has been consistent effort to fit models of
+linked selection to patterns of diversity to estimate population genetic
+parameters such as the strength of selection and the deleterious mutation rate 
+\citep{Hudson1995-pt,McVicker2009-ax},
+the overall reduction in genetic variability due to the
+selection on deleterious variants \citep{Elyashiv2016-vt,Murphy2022-sj},
+and in some cases the rate of beneficial mutations \citep{corbett2015natural,Elyashiv2016-vt}.
 

### 2023-05-13 manuscript/manuscript.tex
pass through intro

@@
 This phenomenon, known as linked selection, is a major force shaping patterns of
-genetic variation along chromosomes \citep{Maynard_Smith1974-zr,Charlesworth1993-gb,Barton1998-aj,Gillespie2000-mh}.
+genetic variation along chromosomes 
+\parencite{Maynard_Smith1974-zr,Charlesworth1993-gb,Barton1998-aj,Gillespie2000-mh}.
 As deleterious variation is clustered in functional portions of the genome,
@@
 parameters such as the strength of selection and the deleterious mutation rate 
-\citep{Hudson1995-pt,McVicker2009-ax},
+\parencite{Hudson1995-pt,McVicker2009-ax},
 the overall reduction in genetic variability due to the
-selection on deleterious variants \citep{Elyashiv2016-vt,Murphy2022-sj},
-and in some cases the rate of beneficial mutations \citep{corbett2015natural,Elyashiv2016-vt}.
-
-Classic Background Selection (BGS) theory is the predominant model for the
-reduction in linked neutral diversity used by statistical approaches to
-quantify the impacts of negative selection. However, this model makes some
-assumptions for mathematical tractability that could distort inferences about
-the negative selection process. First, since fixation probabilities ultimately
-depend on the product of the deleterious selection coefficient (s) and
-population size (N), the efficacy of negative selection should depend on
-demography. Unfortunately, accommodating demography into theoretic negative
-selection models remains an open, difficult problem (XXX). Second, building off
-classic models of mutation-selection balance (Crow 1970, Kimura and Maruyama
-196), the Background Selection (XXX) model assumes that new mutations are
-sufficiently deleterious that they are invariably driven to loss. Under this
-assumption, the effect of selection is well-approximated by simply rescaling
-the neutral coalescent by a reduction factor known as B (XXX). However, this is
-only appropriate in a limited domain of parameters relevant to the negative
-selection process (XXX). Because the BGS model cannot accommodate the
+selection on deleterious variants \parencite{Elyashiv2016-vt,Murphy2022-sj},
+and in some cases the rate of beneficial mutations \parencite{corbett2015natural,Elyashiv2016-vt}.
+
+Prior attempts to model the reduction in linked diversity due to deleterious alleles
+have largely relied on the now canonical Background Selection (BGS) model \parencite{Charlesworth1993-gb,Nordborg1996-nq}.
+While this model has been successful in explaining many patterns of diversity,
+some its simplifying assumptions
+may distort inferences about the negative selection process.
+First, since fixation probabilities ultimately
+depend on the product of the deleterious selection coefficient ($s$) and
+population size ($N$), the efficacy of negative selection should depend on
+demography. Unfortunately, accommodating demography into analytic models of negative
+selection remains an open, difficult problem (XXX). Second, building off
+classic models of mutation-selection balance \parencite{Crow1970-wj,Kimura1966-bk}
+BGS assumes that new mutations are sufficiently deleterious
+such that they are invariably driven to loss. 
+Under this assumption, the effect of selection is well-approximated by simply rescaling
+the coalescent effective size by a reduction factor, 
+sometimes called $B$ such that $N_eB=N_{0}$ \parencite{Charlesworth2013-kl}.
+However this rescaling is not appropriate across parts of parameter space
+that are relevant to natural populations segregating deleterious variation
+ \parencite[e.g.][]{good2014genetic,Santiago2016-mu}.
+In particular, because the BGS model cannot accommodate the
 possibility of fixation for mutations with $Ns \ll 1$, diversity is incorrectly
@@
 In this work, we extend another class of linked selection models that derive
-from quantitative genetics to address limitations of the classic BGS model.
+from quantitative genetics to address limitations of the classic BGS model 
+\parencite{santiago1998effective,Santiago2016-mu}.
 These models consider how polygenic fitness variation increases the rate of
@@
 focus specifically on a deleterious-mutations-only model from Santiago and
-Caballero (2016). This model is identical to the classic BGS model when
-selection against deleterious mutations is strong, but it also correctly
-predicts the reduction in diversity and substitutions rates of weakly
+Caballero (2016). 
+This model is identical to BGS when selection against deleterious mutations is strong,
+but it also correctly predicts the reduction in diversity and substitution rates of weakly
 deleterious mutations by accurately modeling levels of genic fitness variation.
-We extend Santiago and Caballero’s (hereafter the SC16) model so that it can be
-used to model genome-wide patterns of diversity under negative selection
-processes, and develop Python software to use this theoretic approach to fit
+We extend Santiago and Caballero model (hereafter the SC16) so that it can be
+used to model genome-wide patterns of diversity,
+and develop Python software to use this theoretic approach to fit
 genome-wide models of negative selection. Using forward simulations, we show
-this model leads to more accurate DFE estimates under weak selection. While
-this new model accommodates the possibility that a given reduction B could be
-due to weaker selection, in applying our composite-likelihood method to human
-population genomic data, we confirm earlier findings that strong negative
-selection at a small fraction of sites fits human data across populations best.
-Finally, we show that while this model solves some limitations of the classic
-BGS model, simulations reveal another mysterious defect: a poor fit around
-$Ns=1$. We show that this can be partially corrected by incorporating the effects
-of selective interference on genic variation, but the remaining discrepancy is
-due to the build up of negative linkage disequilibrium that cannot be
-adequately captured by this model. We find accounting for genic selective
-interference between segments fits human population genomic data slightly
-better, suggesting levels of negative selection in humans are sufficient to
-perturb the selection at other sites.
+this model leads to more accurate DFE estimates under weak selection. 
+Finally, we apply our composite-likelihood method to human population genomic data
+and provide new estimates of the landscape of negative selection in humans.
+
+%
+%
+% ADK- Vince I commented out the below because I don't like to walk through the results
+% in an intro. Instead say what you did, and then the results are in the results section.
+%
+%While this new model accommodates the possibility that a given reduction B could be
+%due to weaker selection, in applying our composite-likelihood method to human
+%population genomic data, we confirm earlier findings that strong negative
+%selection at a small fraction of sites fits human data across populations best.
+%Finally, we show that while this model solves some limitations of the classic
+%BGS model, simulations reveal another mysterious defect: a poor fit around
+%$Ns=1$. We show that this can be partially corrected by incorporating the effects
+%of selective interference on genic variation, but the remaining discrepancy is
+%due to the build up of negative linkage disequilibrium that cannot be
+%adequately captured by this model. We find accounting for genic selective
+%interference between segments fits human population genomic data slightly
+%better, suggesting levels of negative selection in humans are sufficient to
+%perturb the selection at other sites.
 

### 2023-05-13 manuscript/manuscript.tex
some good stuff

--- 2023-05-13 manuscript/manuscript.tex
@@
 of all adaptations, but the vast majority of mutations either do not affect
-fitness or are deleterious. 
-Natural selection works to eliminate these deleterious mutations from the population,
-thus we expect them to appear at low frequencies (Haldane?Muller?) within populations
-and be less likely to fix between lineages. 
-The genome itself is the product of over a billion years of evolutionary optimization,
-so even though they are rare within populations, we expect that deleterious variants
-should be the overwhelming majority of segregating variation in functional regions,
-and be responsible for a significant portion of phenotypic variation and disease (XX cites XX),
-
-As with most selection in the genome, selection on deleterious variation
+fitness or are deleterious. Negative selection works to eliminate such
+deleterious mutations from the population, which is essential to maintaining
+the integrity of information in the genome to maintain a lineage. Mutations
+under negative selection are kept at low frequencies (Haldane?), contributing
+to human traits (XXX) and diseases. As with most selection in the genome, this
 process perturbs the allele frequencies of neighboring linked variation. This
@@
 clear, we consider a general expression of the drift-effective population size
-that allows for extra non-heritable variability in offspring number $V_k$,
-
-\begin{align}
-    N_d = \frac{2N}{V_k + 2}
-\end{align}
-
-where if $V_k = 2$, this reduces to the classic Wright--Fisher model
+that allows for extra non-heritable variability in offspring number $V^2$,
+%
+\begin{align}
+    \label{eq:simple_Ne}
+    N_d = \frac{2N}{V^2 + 2}
+\end{align}
+
+where if $V^2 = 2$, this reduces to the classic Wright--Fisher model
 \parencite{Wright1938-tv}.
@@
 Like genetic drift, perturbations due to draft are directionless but overall
-increase the variance in allele frequency change. Unlike drift, the stochastic
-perturbations each generation due to draft are autocorrelated as long as
-segregation and recombination do not decouple an allele from its fitness
-background and the selective environment does not change
+increase the variance in allele frequency change. Unlike drift, stochastic
+perturbations caused by draft are autocorrelated across generations as long an
+allele does not recombine or segregate off its fitness background and the
+selective environment does not change
 \parencite{Robertson1961-ho,Santiago1995-hx,Buffalo2019-qs}. When selection is
-strong,
-
-When selection is
-weak and fitness is polygenic, the stochasticity introduced by draft is
-well-approximated by rescaling the drift-effective population size
-
-that exist when there are heritable differences among haplotypes.
-Mathematical models of genetic draft
-
-Gillespie2000-mh
-
-,Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Barton2000-zg
+strong, draft can generate multiple-merger coalescences that create
+discontinuities in allele frequency trajectories, making diffusion
+approximations ill-suited \parencite{Gillespie2000-mh,Der2011-it,Neher2013-dz}.
+However, when selection is weak and polygenic, draft can be approximated as an
+increase in the rate of drift. The key insight is that the draft rescaling
+factor is like extra variability in offspring number, but because of
+autocorrelation due to linkage and similar environments, the heritable fitness
+variation $C^2$ (on a relative fitness scale, i.e. the squared coefficient of
+variance) becomes inflated by a factor $Q^2$. Intuitively, the product $V^2
+Q^2$ represents the expected total variance a neutral mutation experiences over
+its lifetime in a system with draft. This \emph{heritable} fitness variance
+that neutral allele experiences can then be included in Equation
+\eqref{eq:simple_Ne} (assuming the usual Wright--Fisher non-heritable variance,
+i.e. $V^2=2$),
+
+\begin{align}
+    \label{eq:main_Ne}
+    N_e = \frac{N}{C^2 Q^2 + 1}.
+\end{align}
+
+(c.f. \cite{Robertson1961-ho,Santiago1995-hx}; see XXX for a proof). Note that
+because the variance under drift increases each generation due to the
+accumulation of drift perturbations and draft autocorrelations, the inflation
+term $Q^2$ has a temporal dimension. For notational efficiency, we assume $Q^2$
+refers to the asymptotic $Q_t^2$ as $t \to \infty$ and $N_e$ is the
+\emph{asymptotic} draft effective population size. In general, however,
+statistics such as heterozygosity depend on the cumulative autocorrelation
+factor \emph{up to some time} $t$, $Q_t^2$ \parencite[p.
+2111]{Santiago1998-bs}. This reflects the fact that effective population size
+experienced by a mutation newly associated with a fitness background is larger
+than the asymptotic effective population size, since not much autocorrelation
+has accumulated in the stochastic perturbations caused by drift yet. From a
+backwards-time perspective, this reflects the fact that pairwise diversity
+(which is approximately heterozygosity in most cases) depends on the pairwise
+coalescence rate per generation, which will is not constant under draft. The
+benefit of using a forward-time model of draft is that it the inflation factor
+is invariant with respect to the particular fitness background the neutral
+allele becomes stochastically associated with. By contrast, the difficulty with
+backwards-time models is that the coalescence rates experienced by a lineage
+are not invariant to which lineage was sampled.
+
+Equation \eqref{eq:main_Ne}
+
+Shortly, we will show that Robertson (1961) and Santiago and Caballero's (1995,
+1998, 2016) approach
+
+rescaling the drift-effective population size by an appropriate factor
+\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Barton2000-zg}.
+While a simple rescaling cannot elucidate details about how selection alters
+the genealogy, it does allow for the calculation of key quantities like the
+pairwise coalescence time (though see \textcite{Santiago1998-bs}, p. 2111).
 
 \subsection*{DFE Estimates}
- - Table s5 %https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/genetics/206/1/10.1534_genetics.116.197145/8/files1.pdf?Expires=1686357613&Signature=cp9qoI2ChCb6GFZ74HI4eWcA~MQ4sHlWyeI3Zp49-5xEORzMfYlSRxzVvZlo8ZoGWDi0G3PnL00ZplM-~or4YIVm2uhkmqjf61AmC7RdZ1QTak4j0kFMQ0wAbM9pusiV3gIepTQHq2HKRKwgS~ypmv~0oqO5dDxaq9nhOWIBBmLAA49oHWuxUtnRe3sEYATwCwjIFEomhgJpg7vKGCF~87GdCV4JETW4PN2VvD4hsu2AVRd8yA~vUJvsXfhWwcElDeY4fsgpprl1lC3XGjJCnqdwbzRGPV5LNx5XEKLptcZtLeLcOmFJCvmVLe0bewJ7Vta0TSQM7T9AuklompjPPw__&Key-Pair-Id=APKAIE5G5CRDK6RD3PGA
+% - Table s5 https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/genetics/206/1/10.1534_genetics.116.197145/8/files1.pdf?Expires=1686357613&Signature=cp9qoI2ChCb6GFZ74HI4eWcA~MQ4sHlWyeI3Zp49-5xEORzMfYlSRxzVvZlo8ZoGWDi0G3PnL00ZplM-~or4YIVm2uhkmqjf61AmC7RdZ1QTak4j0kFMQ0wAbM9pusiV3gIepTQHq2HKRKwgS~ypmv~0oqO5dDxaq9nhOWIBBmLAA49oHWuxUtnRe3sEYATwCwjIFEomhgJpg7vKGCF~87GdCV4JETW4PN2VvD4hsu2AVRd8yA~vUJvsXfhWwcElDeY4fsgpprl1lC3XGjJCnqdwbzRGPV5LNx5XEKLptcZtLeLcOmFJCvmVLe0bewJ7Vta0TSQM7T9AuklompjPPw__&Key-Pair-Id=APKAIE5G5CRDK6RD3PGA
 
@@
 and our B' maps is at the theoretic minimum. This minimum variance is entirely
-from the coalescent noise when estimating $\bar{B}\text{sim}(x)$, which
+from the coalescent noise when estimating $\bar{B}\text{sim}(x), which
 
@@
 Natural Selection (equation 2, \cite{Santiago2016-mu}, equation 15A
-%`'\cite{Garcia-Dorado2007-jj}). 
-A more formal derivation follows from \textcite{Higgs1995-xc} 
-(though they assume a haploid asexual system, the end
+\cite{Garcia-Dorado2007-jj}). A more formal derivation follows from
+\textcite{Higgs1995-xc} (though they assume a haploid asexual system, the end
 result is the same up to a factor of two). In their work, they derive equations
@@
 
-%\printbibliography
+\printbibliography
+
 \end{document}

### 2023-05-17 manuscript/manuscript.tex
writing

@@
 This approximation works reasonably well for two reasons (though see
-\textcite{Cvijovic2018-vd,Walczak2012-fi,Nicolaisen2012-vs}). First, when
-selection is strong, the number of deleterious mutations per haplotype has a
-stationary Poisson distribution with an average equal to the equilibrium under
-the deterministic mutation-selection model \parencite{Haldane1927-ga}. Since
+\cite{Cvijovic2018-vd,Walczak2012-fi,Nicolaisen2012-vs}). First, when selection
+is strong, the number of deleterious mutations per haplotype has a stationary
+Poisson distribution with an average equal to the equilibrium under the
+deterministic mutation-selection model \parencite{Haldane1927-ga}. Since
 lineages carrying any strongly deleterious mutations cannot contribute to the
 genealogy, coalescence events only occur in the fraction
-$\exp(-\nicefrac{U}{s}$ of the population free of mutations. Second, the time
+$\exp(-\nicefrac{U}{s})$ of the population free of mutations. Second, the time
 it takes for a present-day lineage to trace their ancestry back to a
@@
 coalescence timescale of $N_0$ among these ancestors (the ``coalescence
-phase"), and can be ignored (\textcite{Durrett2008-ql}, p. 213;
-\textcite{Good2014-yz}).
-
-When mutations are weakly deleterious, their allele frequencies are strongly
-influenced by stochastic perturbations and their frequency trajectories are no
-longer well-approximated by deterministic models. When genetic drift is the
-only source of randomness in allele frequency change, the relevant scale of
-stochastic perturbations is determined by the drift-effective population size
-$N_d$ (e.g. due to only non-selective process). For reasons that will soon be
-clear, we consider a general expression of the drift-effective population size
-that allows for extra non-heritable variability in offspring number $V^2$,
-%
-\begin{align}
-    \label{eq:simple_Ne}
-    N_d = \frac{2N}{V^2 + 2}
-\end{align}
-
-where if $V^2 = 2$, this reduces to the classic Wright--Fisher model
-\parencite{Wright1938-tv}.
-
-The insight of nearly neutral theory (XXX) was that the qualitative behavior of
-selected sites is determined by the compound parameter $2N_d s$. When $2N_d s
-\gg 1$, the stochastic component of allele frequency change is minuscule
-relative to selection and can be ignored; this is why the factor that rescales
-the drift-effective population size under the BGS model does not depend on
-$N_d$. However, when $2N_d s \le 1$ stochastic perturbations are on a scale
-equal to or greater than the changes due to selection. In this regime, weakly
+phase"), and can be ignored (\cite{Durrett2008-ql}, p. 213;
+\cite{Good2014-yz}).
+
+When mutations are only weakly deleterious, their allele frequencies are
+strongly influenced by stochastic perturbations and their frequency
+trajectories are no longer well-approximated by deterministic models. When
+genetic drift is the only source of randomness in allele frequency change, the
+relevant scale of stochastic perturbations is determined by the drift-effective
+population size $N_e$ (e.g. due to only non-selective process). The insight of
+nearly neutral theory (XXX) was that the qualitative behavior of selected sites
+is determined by the compound parameter $2N_e s$. When $2N_e s \gg 1$, the
+stochastic component of allele frequency change is minuscule relative to
+selection and can be ignored; this is why the factor that rescales the
+drift-effective population size under the BGS model does not depend on $N_e$.
+However, when $2N_e s \le 1$, the stochastic perturbations are on a scale equal
+to or greater than the changes due to selection. In this regime, weakly
 deleterious alleles can drift up to intermediate frequencies before their
@@
 fitness variation generates another source of randomness in allele frequency
-change known as genetic \emph{draft} \parencite{Neher2013-dz}. Genetic draft
-occurs when alleles become randomly associated with different fitness
-backgrounds each generation, which stochastically perturbs their trajectories.
+change known as genetic \emph{draft} \parencite{Neher2013-dz}. An allele
+experiences genetic draft when it becomes randomly associated with a fitness
+background, which stochastically perturbs its trajectory across generations.
 Like genetic drift, perturbations due to draft are directionless but overall
 increase the variance in allele frequency change. Unlike drift, stochastic
-perturbations caused by draft are autocorrelated across generations as long an
-allele does not recombine or segregate off its fitness background and the
+perturbations caused by draft are autocorrelated across generations for as long
+an allele does not recombine or segregate off its fitness background and the
 selective environment does not change
@@
 approximations ill-suited \parencite{Gillespie2000-mh,Der2011-it,Neher2013-dz}.
-However, when selection is weak and polygenic, draft can be approximated as an
-increase in the rate of drift. The key insight is that the draft rescaling
-factor is like extra variability in offspring number, but because of
-autocorrelation due to linkage and similar environments, the heritable fitness
-variation $C^2$ (on a relative fitness scale, i.e. the squared coefficient of
-variance) becomes inflated by a factor $Q^2$. Intuitively, the product $V^2
-Q^2$ represents the expected total variance a neutral mutation experiences over
-its lifetime in a system with draft. This \emph{heritable} fitness variance
-that neutral allele experiences can then be included in Equation
-\eqref{eq:simple_Ne} (assuming the usual Wright--Fisher non-heritable variance,
-i.e. $V^2=2$),
+However, when selection is weak, draft can be approximated as an increase in
+the rate of drift (or equivalently, a reduction in effective population size).
+
+A key insight from Robertson (\citeyear{Robertson1961-ho}) is that in the
+long-run draft acts like an increase in the variance in offspring number.
+\textcite{Wright1938-tv} showed that additional non-heritable variance in
+offspring number (e.g. due to mating strategy) rescales the population size
+according to,
+
+\begin{align}
+    \label{eq:simple_Ne}
+    N_e = \frac{2N}{V + 2}
+\end{align}
+
+However, unlike non-heritable variation in offspring number, the heritable
+variation is autocorrelated across generations. At the individual level
+Robertson considered, this is because offspring from large families tend to
+beget many descendents themselves (and likewise with small families). However,
+the same process acts at the genomic level, as the perturbations to a neutral
+allele's trajectory from its particular fitness background tend to occur in the
+same direction across generations. The additional heritable variation in
+offspring number is the additive fitness variation $V_A$, which is inflated by
+a factor $Q_t^2$ due to the build up of autocorrelation to generation $t$. We
+let $Q^2 = Q_t^2$ as $t \to infty$. Intuitively, the product $V_A Q^2$
+represents the expected total variance in reproductive success a neutral
+mutation experiences over its lifetime in a system with draft. Incorporating
+the impact of time-invariant autocorrelation created by draft as rescaled rate
+of drift is analogous to using the Green--Kubo relation to find the transport
+coefficient \parencite{Green1954-kl,Kubo1957-va}.
+
+Including the total asymptotic variance in reproductive success in Equation
+\eqref{eq:simple_Ne} and assuming Wright--Fisher levels of non-heritable
+variation (i.e. $V = 2 + 4 Q^2 V_A$),
 
@@
     \label{eq:main_Ne}
-    N_e = \frac{N}{C^2 Q^2 + 1}.
+    N_e = \frac{N}{Q^2 V_A + 1}
 \end{align}
@@
 accumulation of drift perturbations and draft autocorrelations, the inflation
-term $Q^2$ has a temporal dimension. For notational efficiency, we assume $Q^2$
-refers to the asymptotic $Q_t^2$ as $t \to \infty$ and $N_e$ is the
-\emph{asymptotic} draft effective population size. In general, however,
+term $Q^2$ has a temporal dimension. For notational efficiency, we let $Q^2$
+refer to the asymptotic $Q_t^2$ as $t \to \infty$ and $N_e$ is the
+\emph{asymptotic} weak draft effective population size. In general, however,
 statistics such as heterozygosity depend on the cumulative autocorrelation
@@
 backwards-time perspective, this reflects the fact that pairwise diversity
-(which is approximately heterozygosity in most cases) depends on the pairwise
-coalescence rate per generation, which will is not constant under draft. The
-benefit of using a forward-time model of draft is that it the inflation factor
-is invariant with respect to the particular fitness background the neutral
-allele becomes stochastically associated with. By contrast, the difficulty with
-backwards-time models is that the coalescence rates experienced by a lineage
-are not invariant to which lineage was sampled.
-
-Equation \eqref{eq:main_Ne}
-
-Shortly, we will show that Robertson (1961) and Santiago and Caballero's (1995,
-1998, 2016) approach
-
-rescaling the drift-effective population size by an appropriate factor
-\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Barton2000-zg}.
-While a simple rescaling cannot elucidate details about how selection alters
-the genealogy, it does allow for the calculation of key quantities like the
-pairwise coalescence time (though see \textcite{Santiago1998-bs}, p. 2111).
+(which is approximately heterozygosity when $2N\mu \ll 1$) depends on the
+pairwise coalescence rate per generation, which is not constant under draft.
+The benefit of using Robertson's forward-time model of draft is that the
+inflation factor is invariant with respect to the particular fitness background
+the neutral allele becomes stochastically associated with. By contrast, the
+difficulty with modeling draft backwards in time is that the coalescence rates
+experienced by a lineage are not invariant to which lineage was sampled due to
+its particular associated fitness background.
+
+Equation \eqref{eq:main_Ne} is general, since different modes of selection and
+linkage can be accommodated by different expressions for the inflation factor
+$Q^2$. When fitness variation has a multiplicative polygenic basis, as it does
+for genome-wide negative selection processes, the asymptotic weak draft
+effective population size experienced by an arbitrary neutral site under the
+influence of all linked regions is,
+
+\begin{align}
+    \label{eq:polygenic_Ne}
+    N_e \approx N \exp\left(-\sum_{i=1}^n V_{A,i} \frac{Q_i^2}{2}\right)
+\end{align}
+
+where the factor of one-half comes from ignoring weak associations from
+unlinked regions and chromosomes (see XXX). In our genome-wide model, we
+consider the summation in Equation \eqref{eq:polygenic_Ne} over non-overlapping
+segments $i \in \{1, 2, \ldots, n\}$ each undergoing selection such that
+segment $i$ contributes additive fitness variance $V_{A,i}$ to the total
+additive genetic fitness variance. The degree to which segment $i$ inflates the
+total variance in reproductive success depends on the how quickly the
+autocorrelations in
+
+autocorrelation function in
+reproductive success. Assuming a constant selective environment, the
+autocorrelation in reproductive success of the neutral allele is given by a
+geometric autocorrelation function $C(t) = \[(1-r_i)(1-\kappa_i)\]^t$. The
+decay in autocorrelation is due to
+
+\begin{align}
+    Q_i &= 1 + \sum_{j=1}^\infty \left[(1-r_i)(1-\kappa_i)\right]^j \\
+        &= \frac{1}{\kappa_i + r_i(1-\kappa_i)} 
+\end{align}
+
+Segment $i$ inflates the total variance in
+reproductive success of the neutral allele by $Q_i^2$, according to two key
+parameters. First,
+
+according to the chance it stays
+associated each generation, given by the recombination fraction $r_i$, and the
+extent that the associated heritable fitness variation within the segment
+determines future reproductive success in later generations. Assuming a
+constant environment for selection, this
+
+, according to
+the recombination fraction $r_i$ to the segment and the extent to which the
+fitness variation in the segment perturbs the trajectory in later generations.
+
+the fitness
+variation associated with the neutral allele
+
+persistence of the
+associated fitness variance through the generations.
+
+Here, we consider each $i$ to refer
+to an $L_i$-basepair segment under negative selection (though this model
+accommodates other forms of polygenic selection, \cite[p.
+2109]{Santiago1998-bs}). The inflation factor $Q_i^2$ depends on the
+recombination fraction $r_i$ between the neutral site and selected segment, and
+the rate at which the fitness variance at segment $i$ associated with the
+
+neutral allele decays across generations, $1-\kappa_i$. This associated fitness
+variation decays at rate $1-\kappa_i$ due only to the dynamics of polygenic
+selection within the segment \parencite{Bulmer1980-zo,Turelli1990-kd}, assuming
+a constant environment. Then, the root of the asymptotic variance inflation
+factor $Q_i$ experienced by the neutral allele is,
+
+\begin{align}
+    Q_i &= 1 + \sum_{j=1}^\infty \left[(1-r_i)(1-\kappa_i)\right]^j \\
+        &= \frac{1}{\kappa_i + r_i(1-\kappa_i)} 
+\end{align}
+
+(see Appendix Equation \ref{eq:Qinf}). This general equation can accommodate
+models of polygenic selection as long as the equilibrium additive fitness
+variation $C_i^2$ can be specified and the change in variance due to selection
+can be approximated as a geometric decay, i.e. $\Delta C_i^2 = -\kappa C_i^2$
+\parencite{Bulmer1971-ae,Keightley1988-eq,Walsh2018-bt}. This is usually a
+reasonable assumption since within-generation selection removes a fraction of
+phenotypic variation from the population, and the fraction of that phenotypic
+variation that is genetic is approximately the heritability $h^2$.
+
+Assuming the fitness
+variation $C_i^2$ in the segment is at an equilibrium between mutation and
+selection, $\Delta C_i^2 = 0$, and $\Delta C_i^2 = V_{m,i} - \kappa C_i^2$
+
+Note that
+the decay in fitness variation due to the substitution of weakly deleterious
+mutations happens on a slower timescale and is ignored here, but as we will
+impacts the overall level of $C_i^2$.
 
@@
 
--
+- we need to talk about Fisher's Fundamental Theorem and how this model
+connects fitness variation to absolute changes in fitness.
 
@@
 
-or for continuous $t$,
-
-\begin{align}
-  \label{eq:q_t}
-  Q_t &= 1+\frac{(1-\kappa) (1-r) \left(1-(1-\kappa)^t (1-r)^t\right)}{\kappa+(1-\kappa)r}
-\end{align}
-
+% or for continuous $t$,
+
+% \begin{align}
+%   \label{eq:q_t}
+%   Q_t &= 1+\frac{(1-\kappa) (1-r) \left(1-(1-\kappa)^t (1-r)^t\right)}{\kappa+(1-\kappa)r}
+% \end{align}
 which as $t \to \infty$, converges to,

### 2023-05-17 manuscript/manuscript.tex
more writing

@@
 a factor $Q_t^2$ due to the build up of autocorrelation to generation $t$. We
-let $Q^2 = Q_t^2$ as $t \to infty$. Intuitively, the product $V_A Q^2$
+let $Q^2 = Q_t^2$ as $t \to \infty$. Intuitively, the product $V_A Q^2$
 represents the expected total variance in reproductive success a neutral
@@
 
-(c.f. \cite{Robertson1961-ho,Santiago1995-hx}; see XXX for a proof). Note that
-because the variance under drift increases each generation due to the
-accumulation of drift perturbations and draft autocorrelations, the inflation
-term $Q^2$ has a temporal dimension. For notational efficiency, we let $Q^2$
-refer to the asymptotic $Q_t^2$ as $t \to \infty$ and $N_e$ is the
-\emph{asymptotic} weak draft effective population size. In general, however,
-statistics such as heterozygosity depend on the cumulative autocorrelation
-factor \emph{up to some time} $t$, $Q_t^2$ \parencite[p.
+(c.f. \cite{Robertson1961-ho,Santiago1995-hx}; see XXX for a proof). In
+general, however, statistics such as heterozygosity depend on the cumulative
+autocorrelation factor \emph{up to some time} $t$, $Q_t^2$ \parencite[p.
 2111]{Santiago1998-bs}. This reflects the fact that effective population size
@@
 $Q^2$. When fitness variation has a multiplicative polygenic basis, as it does
-for genome-wide negative selection processes, the asymptotic weak draft
-effective population size experienced by an arbitrary neutral site under the
-influence of all linked regions is,
+for genome-wide negative selection processes, the draft asymptotic effective
+population size experienced by an arbitrary neutral site under the influence of
+all linked regions is,
 
@@
 segment $i$ contributes additive fitness variance $V_{A,i}$ to the total
-additive genetic fitness variance. The degree to which segment $i$ inflates the
-total variance in reproductive success depends on the how quickly the
-autocorrelations in
-
-autocorrelation function in
-reproductive success. Assuming a constant selective environment, the
-autocorrelation in reproductive success of the neutral allele is given by a
-geometric autocorrelation function $C(t) = \[(1-r_i)(1-\kappa_i)\]^t$. The
-decay in autocorrelation is due to
+additive genetic fitness variance. Under draft, the autocorrelation function
+for a neutral allele associated with segment $i$ is the time-invariant function
+$C(t) = \[(1-r_i)(1-\kappa_i)\]^t$, where $r_i$ is the recombination fraction
+to the segment and $\kappa_i$ is the rate that the associated fitness decays
+due to selective dynamics. Then, the cumulative autocorrelation is,
 
@@
     Q_i &= 1 + \sum_{j=1}^\infty \left[(1-r_i)(1-\kappa_i)\right]^j \\
-        &= \frac{1}{\kappa_i + r_i(1-\kappa_i)} 
-\end{align}
-
-Segment $i$ inflates the total variance in
-reproductive success of the neutral allele by $Q_i^2$, according to two key
-parameters. First,
-
-according to the chance it stays
-associated each generation, given by the recombination fraction $r_i$, and the
-extent that the associated heritable fitness variation within the segment
-determines future reproductive success in later generations. Assuming a
-constant environment for selection, this
-
-, according to
-the recombination fraction $r_i$ to the segment and the extent to which the
-fitness variation in the segment perturbs the trajectory in later generations.
-
-the fitness
-variation associated with the neutral allele
-
-persistence of the
-associated fitness variance through the generations.
-
-Here, we consider each $i$ to refer
-to an $L_i$-basepair segment under negative selection (though this model
-accommodates other forms of polygenic selection, \cite[p.
-2109]{Santiago1998-bs}). The inflation factor $Q_i^2$ depends on the
-recombination fraction $r_i$ between the neutral site and selected segment, and
-the rate at which the fitness variance at segment $i$ associated with the
-
-neutral allele decays across generations, $1-\kappa_i$. This associated fitness
-variation decays at rate $1-\kappa_i$ due only to the dynamics of polygenic
-selection within the segment \parencite{Bulmer1980-zo,Turelli1990-kd}, assuming
-a constant environment. Then, the root of the asymptotic variance inflation
-factor $Q_i$ experienced by the neutral allele is,
-
-\begin{align}
-    Q_i &= 1 + \sum_{j=1}^\infty \left[(1-r_i)(1-\kappa_i)\right]^j \\
-        &= \frac{1}{\kappa_i + r_i(1-\kappa_i)} 
+        &= \frac{1}{\kappa_i + r_i(1-\kappa_i)}.
 \end{align}
@@
 reasonable assumption since within-generation selection removes a fraction of
-phenotypic variation from the population, and the fraction of that phenotypic
-variation that is genetic is approximately the heritability $h^2$.
-
-Assuming the fitness
-variation $C_i^2$ in the segment is at an equilibrium between mutation and
-selection, $\Delta C_i^2 = 0$, and $\Delta C_i^2 = V_{m,i} - \kappa C_i^2$
-
-Note that
-the decay in fitness variation due to the substitution of weakly deleterious
-mutations happens on a slower timescale and is ignored here, but as we will
-impacts the overall level of $C_i^2$.
+phenotypic variation from the population, and some fraction of that is additive
+genetic variation \parencite{Bulmer1971-ae,Keightley1988-eq}.
+
+The final pieces are the equilibrium additive fitness variance $V_A$ and the
+decay rate in associated fitness $1-\kappa$. At equilibrium,
 
@@
 
-% or for continuous $t$,
-
-% \begin{align}
-%   \label{eq:q_t}
-%   Q_t &= 1+\frac{(1-\kappa) (1-r) \left(1-(1-\kappa)^t (1-r)^t\right)}{\kappa+(1-\kappa)r}
-% \end{align}
+or for continuous $t$,
+
+\begin{align}
+  \label{eq:q_t}
+  Q_t &= 1+\frac{(1-\kappa) (1-r) \left(1-(1-\kappa)^t (1-r)^t\right)}{\kappa+(1-\kappa)r}
+\end{align}
 which as $t \to \infty$, converges to,

### 2023-05-17 manuscript/manuscript.tex
more theory

@@
 Robertson considered, this is because offspring from large families tend to
-beget many descendents themselves (and likewise with small families). However,
-the same process acts at the genomic level, as the perturbations to a neutral
+beget many descendents themselves (and likewise with small families). The same
+autocorrelation occurs at the genomic level, as the perturbations to a neutral
 allele's trajectory from its particular fitness background tend to occur in the
-same direction across generations. The additional heritable variation in
+same direction across generations as long as the linkage exists and the
+selective environment is constant. The additional heritable variation in
 offspring number is the additive fitness variation $V_A$, which is inflated by
@@
 mutation experiences over its lifetime in a system with draft. Incorporating
-the impact of time-invariant autocorrelation created by draft as rescaled rate
-of drift is analogous to using the Green--Kubo relation to find the transport
-coefficient \parencite{Green1954-kl,Kubo1957-va}.
+the impact of time-invariant autocorrelation created by draft as a rescaled
+rate of drift is analogous to using the Green--Kubo relation to find the
+transport coefficient in molecular dynamics with velocity autocorrelation
+\parencite{Green1954-kl,Kubo1957-va}.
 
@@
 pairwise coalescence rate per generation, which is not constant under draft.
+
+Equation \eqref{eq:main_Ne} is general, since different modes of selection and
+linkage can be accommodated by different expressions for the inflation factor
+$Q^2$. When fitness variation has a multiplicative polygenic basis, as it does
+for genome-wide negative selection processes, the draft asymptotic effective
+population size experienced by an arbitrary neutral site under the influence of
+all linked regions is,
+
+\begin{align}
+    \label{eq:polygenic_Ne}
+    N_e \approx N \exp\left(-\sum_{i=1}^n V_{A,i} \frac{Q_i^2}{2}\right)
+\end{align}
+
+where the factor of one-half comes from ignoring weak associations from
+unlinked regions and chromosomes (see XXX). In our genome-wide model, we
+consider the summation in Equation \eqref{eq:polygenic_Ne} over non-overlapping
+segments $i \in \{1, 2, \ldots, n\}$ each undergoing selection such that
+segment $i$ contributes additive fitness variance $V_{A,i}$ to the total
+additive genetic fitness variance. Under equilibrium levels of fitness
+variation, the autocorrelation function for a neutral allele associated with
+segment $i$ is the time-invariant function $C(t) = \[(1-r_i)(1-\kappa_i)\]^t$,
+where $r_i$ is the recombination fraction to the segment and $\kappa_i$ is the
+rate that the associated fitness variance decays due to selective dynamics.
+Then, the cumulative autocorrelation is,
+
+\begin{align}
+    Q_i &= 1 + \sum_{j=1}^\infty \left[(1-r_i)(1-\kappa_i)\right]^j \\
+        &= \frac{1}{\kappa_i + r_i(1-\kappa_i)}.
+\end{align}
+
+(see Appendix Equation \ref{eq:Qinf}). This general equation can accommodate
+models of polygenic selection as long as the equilibrium additive fitness
+variation $V_{A,i}$ can be specified and the change in variance due to
+selection can be approximated as a geometric decay, i.e. $\Delta V_{A,i} =
+-\kappa V_{A,i}$ \parencite{Bulmer1971-ae,Keightley1988-eq,Walsh2018-bt}. This
+is usually a reasonable assumption since within-generation selection removes a
+fraction of phenotypic variation from the population, and some fraction of that
+is additive genetic variation \parencite{Bulmer1971-ae,Keightley1988-eq}.
+
+The remaining pieces are expressions for the equilibrium additive fitness
+variance $V_A$ and the decay rate in associated fitness $1-\kappa$. At this
+point, we diverge from Santiago and Caballero (\citeyear{Santiago1998-bs},
+\citeyear{Santiago2016-mu}) to note that the additive fitness variation is the
+sum of additive \emph{genic} fitness variance $V_a$ (i.e. due to allele
+frequencies) and the total linkage disequilibria $\delta_{LD}$ between selected
+sites, $V_A = V_a + \delta_{LD}$. Considering only the genic variation, at
+equilibrium $\Delta V_{a,i} = 0$; thus, the loss in genic fitness each
+generation due to selective dynamics $-\kappa V_{A,i}$ must be equal to the
+increase in fitness variation to new mutations $V_m$ each generation and
+changes in the LD term \parencite{Bulmer1971-ae}. Note that the loss in genic
+fitness due to drift is not time-invariant due to the build up of
+autocorrelation, and considered elsewhere in the derivation (see Supplementary
+Materials Equation \ref{eq:vardecay}). Similarly, the loss in variation due to
+fixation of weakly deleterious occurs on much slower timescales and can be
+ignored here. Then, $\kappa_i = \nicefrac{V_{m,i}}{V_{a,i}}$ (see Supplementary
+Materials Equation \ref{eq:Z}).
+
+Under any selection model, the genic fitness variance created by a new mutation
+(at frequency $x=\nicefrac{1}{2N}$) is $2s^2 x(1-x) \approx \nicefrac{s^2}{N}$.
+For the entire population of $2N$ chromosomes, the fresh variance from mutation
+each generation in segment $i$ is $V_{m,i} \approx U_is^2$ where $U_i = 2\mu
+L_i$ is the diploid mutation rate per generation within the segment. Under
+mutation-selection balance assumed by BGS, an $L_i$-basepair segment has genic
+variance $V_{a,i}^{BGS} \approx U_i s$ (see Supplementary Materials Equation
+\ref{eq:va_bgs}) and thus $\kappa_i^{BGS} = s$. Combining this with $\kappa_i =
+\nicefrac{V_{m,i}}{V_{a,i}} = s$ in Equation \eqref{eq:simple_Ne}, we have
+
+\begin{align}
+    N_e = N \exp \left( - \sum_i^n \frac{\mu L_i}{s(1 + r_i(1-s)/s)^2} \right) 
+\end{align}
+
+which is exactly the genome-wide model of background selection used in previous
+studies \parencite{McVicker2009-ax,Elyashiv2016-vt,Murphy2022-sj} (note that
+\cite{Santiago1998-bs} have shown a similar correspondence between their work
+and BGS models with recombination from \cite{Nordborg1996-nq} and
+\cite{Hudson1995-xc}).
+
+When selection is weak, however, the additive fitness variance under the
+classic BGS model is no longer accurate. \parencite{Santiago2016-mu} suggested
+that this is due to the loss of fitness variation that occurs when a
+segregating site fixes and its heterozygosity becomes zero. If we let $R$
+represent twice the rate of the ratchet of deleterious mutations in fixations
+per generation, each fixation removes exactly the variation mutation puts in
+$V_m = Us$, thus under a general equilibrium negative selection process,
+
+\begin{align}
+  V_{a} = (U - R)s. 
+\end{align}
+
+Our results differ from \parencite{Santiago2016-mu}, since we find from
+simulations their expression models the additive \emph{genic} fitness
+variation, not additive genetic fitness variation. Intuitively, the equilibrium
+genic fitness variation is a balance between the rate of gain of genic fitness
+variation due to mutation and its loss due to fixation. Note that since
+mutation is the ultimate source of all mutations that fix or are lost, $U \ge
+R$.
+
 The benefit of using Robertson's forward-time model of draft is that the
@@
 its particular associated fitness background.
-
-Equation \eqref{eq:main_Ne} is general, since different modes of selection and
-linkage can be accommodated by different expressions for the inflation factor
-$Q^2$. When fitness variation has a multiplicative polygenic basis, as it does
-for genome-wide negative selection processes, the draft asymptotic effective
-population size experienced by an arbitrary neutral site under the influence of
-all linked regions is,
-
-\begin{align}
-    \label{eq:polygenic_Ne}
-    N_e \approx N \exp\left(-\sum_{i=1}^n V_{A,i} \frac{Q_i^2}{2}\right)
-\end{align}
-
-where the factor of one-half comes from ignoring weak associations from
-unlinked regions and chromosomes (see XXX). In our genome-wide model, we
-consider the summation in Equation \eqref{eq:polygenic_Ne} over non-overlapping
-segments $i \in \{1, 2, \ldots, n\}$ each undergoing selection such that
-segment $i$ contributes additive fitness variance $V_{A,i}$ to the total
-additive genetic fitness variance. Under draft, the autocorrelation function
-for a neutral allele associated with segment $i$ is the time-invariant function
-$C(t) = \[(1-r_i)(1-\kappa_i)\]^t$, where $r_i$ is the recombination fraction
-to the segment and $\kappa_i$ is the rate that the associated fitness decays
-due to selective dynamics. Then, the cumulative autocorrelation is,
-
-\begin{align}
-    Q_i &= 1 + \sum_{j=1}^\infty \left[(1-r_i)(1-\kappa_i)\right]^j \\
-        &= \frac{1}{\kappa_i + r_i(1-\kappa_i)}.
-\end{align}
-
-(see Appendix Equation \ref{eq:Qinf}). This general equation can accommodate
-models of polygenic selection as long as the equilibrium additive fitness
-variation $C_i^2$ can be specified and the change in variance due to selection
-can be approximated as a geometric decay, i.e. $\Delta C_i^2 = -\kappa C_i^2$
-\parencite{Bulmer1971-ae,Keightley1988-eq,Walsh2018-bt}. This is usually a
-reasonable assumption since within-generation selection removes a fraction of
-phenotypic variation from the population, and some fraction of that is additive
-genetic variation \parencite{Bulmer1971-ae,Keightley1988-eq}.
-
-The final pieces are the equilibrium additive fitness variance $V_A$ and the
-decay rate in associated fitness $1-\kappa$. At equilibrium,
 
@@
 \begin{align}
+    \label{eq:vardecay}
   \var(x_2 - p_0) &= \E({S_1}^2) + \E(S_1 S_2) + \E({S_2}^2) \\
@@
 \begin{align}
-  V_{A} = Us - 2Rs. 
-\end{align}
-
-where $U = 2u$ and $R$ is the diploid rate of the ratchet, which depends on
+  V_{A} = Us - Rs. 
+\end{align}
+
+where $U = 2u$ and $R=2r$ is the diploid rate of the ratchet, which depends on
 higher moments of the fitness distribution, but only differs from $r$ through a

### 2023-05-17 manuscript/manuscript.tex
draft up to theory sec

@@
 pairwise coalescence rate per generation, which is not constant under draft.
+The benefit of using Robertson's forward-time model of draft is that the
+inflation factor is invariant with respect to the particular fitness background
+the neutral allele becomes stochastically associated with. By contrast, the
+difficulty with modeling draft backwards in time is that the coalescence rates
+experienced by a lineage are not invariant to which lineage was sampled due to
+its particular associated fitness background.
 
@@
 variation, the autocorrelation function for a neutral allele associated with
-segment $i$ is the time-invariant function $C(t) = \[(1-r_i)(1-\kappa_i)\]^t$,
+segment $i$ is the time-invariant function $C(t) = [(1-r_i)(1-\kappa_i)]^t$,
 where $r_i$ is the recombination fraction to the segment and $\kappa_i$ is the
@@
 autocorrelation, and considered elsewhere in the derivation (see Supplementary
-Materials Equation \ref{eq:vardecay}). Similarly, the loss in variation due to
-fixation of weakly deleterious occurs on much slower timescales and can be
-ignored here. Then, $\kappa_i = \nicefrac{V_{m,i}}{V_{a,i}}$ (see Supplementary
-Materials Equation \ref{eq:Z}).
+Materials Equation \ref{eq:vardecay}). Then, $\kappa_i =
+\nicefrac{V_{m,i}}{V_{a,i}}$ (see Supplementary Materials Equation \ref{eq:Z}).
 
@@
 which is exactly the genome-wide model of background selection used in previous
-studies \parencite{McVicker2009-ax,Elyashiv2016-vt,Murphy2022-sj} (note that
-\cite{Santiago1998-bs} have shown a similar correspondence between their work
-and BGS models with recombination from \cite{Nordborg1996-nq} and
-\cite{Hudson1995-xc}).
+studies \parencite{McVicker2009-ax,Elyashiv2016-vt,Murphy2022-sj}. Thus, the
+classic background selection model is a special case of the more general theory
+of Santiago and Caballero (\citeyear{Santiago2016-mu}), which they have shown
+previously (\citeyear{Santiago1998-bs}).
 
@@
 per generation, each fixation removes exactly the variation mutation puts in
-$V_m = Us$, thus under a general equilibrium negative selection process,
-
-\begin{align}
+$V_m = Us$ (we omit the segment index for clarity). Thus, under a general
+equilibrium negative selection process,
+
+\begin{align}
+  \label{eq:Va}
   V_{a} = (U - R)s. 
@@
 
-Our results differ from \parencite{Santiago2016-mu}, since we find from
-simulations their expression models the additive \emph{genic} fitness
-variation, not additive genetic fitness variation. Intuitively, the equilibrium
-genic fitness variation is a balance between the rate of gain of genic fitness
-variation due to mutation and its loss due to fixation. Note that since
-mutation is the ultimate source of all mutations that fix or are lost, $U \ge
-R$.
-
-The benefit of using Robertson's forward-time model of draft is that the
-inflation factor is invariant with respect to the particular fitness background
-the neutral allele becomes stochastically associated with. By contrast, the
-difficulty with modeling draft backwards in time is that the coalescence rates
-experienced by a lineage are not invariant to which lineage was sampled due to
-its particular associated fitness background.
+Note that since mutation is the ultimate source of all mutations that fix or
+are lost, $U \ge R$. Intuitively, this equilibrium genic fitness variation is a
+balance between the rate of gain of genic fitness variation due to mutation and
+its loss due to fixation. Our results differ from \parencite{Santiago2016-mu},
+since we find from simulations their expression models the additive
+\emph{genic} fitness variation, not additive genetic fitness variation (see
+Section XXX), since it does not adequately capture linkage disequilibria
+dynamics. Santiago and Caballero demonstrate Equation \eqref{eq:Va} through
+Fisher's Fundamental Theorem of Natural Selection, but we find an alternative
+proof (\cite{Higgs1995-xc}; see Supplementary Materials Section
+\ref{supp:weak-strong}).
+
+However, closed-form expressions for the rate of the ratchet $R$ have generally
+been hard to find \parencite{Haigh1978-gt,Higgs1995-xc,Gessler1995-hz}. The key
+insight of \parencite{Santiago2016-mu} is that the ratchet rate under draft is
+given by the probability of fixation $p_f(N_e, s)$
+\parencite{Kimura1962-su,Malecot1948-zv} with the draft-rescaled effective
+population size, e.g. $R = N U p_f(N_e, s)$. Given this equation for the
+ratchet and Equation \eqref{eq:polygenic_Ne} for $N_e$ under draft, we have a
+system of non-linear equations that can be solved numerically for $N_e$ and
+$R$.
+
+\subsection*{Simulations of the Diploid Model}
 
@@
 
-\subsection*{Effective Population Size Under Weak and Strong BGS}
+\subsection*{Effective Population Size Under Weak and Strong BGS} \label{supp:weak-strong}
 

### 2023-05-18 manuscript/manuscript.tex
draft up to theory sec

@@
 
-\title{Towards a Complete Model of Mutation Selection Balance in Humans}
+\title{Towards a Complete Model of the Negative Selection Process in Humans}
 
@@
 \eqref{eq:simple_Ne} and assuming Wright--Fisher levels of non-heritable
-variation (i.e. $V = 2 + 4 Q^2 V_A$),
+variation (i.e. $V = 2 + 4 Q^2 V_A$), the draft-effective population size $N_d$
+is
 
@@
     \label{eq:main_Ne}
-    N_e = \frac{N}{Q^2 V_A + 1}
+    N_d = \frac{N}{Q^2 V_A + 1}
 \end{align}
@@
     \label{eq:polygenic_Ne}
-    N_e \approx N \exp\left(-\sum_{i=1}^n V_{A,i} \frac{Q_i^2}{2}\right)
+    N_d \approx N \exp\left(-\sum_{i=1}^n V_{A,i} \frac{Q_i^2}{2}\right)
 \end{align}
@@
 variance $V_{a,i}^{BGS} \approx U_i s$ (see Supplementary Materials Equation
-\ref{eq:va_bgs}) and thus $\kappa_i^{BGS} = s$. Combining this with $\kappa_i =
-\nicefrac{V_{m,i}}{V_{a,i}} = s$ in Equation \eqref{eq:simple_Ne}, we have
-
-\begin{align}
-    N_e = N \exp \left( - \sum_i^n \frac{\mu L_i}{s(1 + r_i(1-s)/s)^2} \right) 
-\end{align}
-
-which is exactly the genome-wide model of background selection used in previous
-studies \parencite{McVicker2009-ax,Elyashiv2016-vt,Murphy2022-sj}. Thus, the
-classic background selection model is a special case of the more general theory
-of Santiago and Caballero (\citeyear{Santiago2016-mu}), which they have shown
-previously (\citeyear{Santiago1998-bs}).
+\ref{eq:va_bgs}) and thus $\kappa_i^{BGS} = s$. Substituting $V_{a,i}^{BGS}$
+and $\kappa_i^{BGS}$ in Equation \eqref{eq:simple_Ne} and simplifying, we have
+
+\begin{align}
+    N_d = N \exp \left( - \sum_i^n \frac{\mu L_i}{s(1 + r_i(1-s)/s)^2} \right) 
+\end{align}
+
+which is identical to the genome-wide model of background selection used in
+previous studies \parencite{McVicker2009-ax,Elyashiv2016-vt,Murphy2022-sj}.
+Thus, the classic background selection model is a special case of the more
+general theory of Santiago and Caballero (\citeyear{Santiago2016-mu}), which
+they have shown previously (\citeyear{Santiago1998-bs}).
 
@@
 
-Note that since mutation is the ultimate source of all mutations that fix or
-are lost, $U \ge R$. Intuitively, this equilibrium genic fitness variation is a
-balance between the rate of gain of genic fitness variation due to mutation and
-its loss due to fixation. Our results differ from \parencite{Santiago2016-mu},
-since we find from simulations their expression models the additive
-\emph{genic} fitness variation, not additive genetic fitness variation (see
-Section XXX), since it does not adequately capture linkage disequilibria
-dynamics. Santiago and Caballero demonstrate Equation \eqref{eq:Va} through
-Fisher's Fundamental Theorem of Natural Selection, but we find an alternative
-proof (\cite{Higgs1995-xc}; see Supplementary Materials Section
-\ref{supp:weak-strong}).
+Note that since mutation is the ultimate source of all alleles that are fixed
+by the ratchet or are lost, $U \ge R$. Intuitively, this equilibrium genic
+fitness variation is a balance between the rate of gain of genic fitness
+variation due to mutation and its loss due to fixation. Our results differ from
+\parencite{Santiago2016-mu}, since we find from simulations their expression
+models the additive \emph{genic} fitness variation, not additive genetic
+fitness variation (see Section XXX), since it does not adequately capture
+linkage disequilibria dynamics. Santiago and Caballero demonstrate Equation
+\eqref{eq:Va} through Fisher's Fundamental Theorem of Natural Selection, but we
+find an alternative proof (\cite{Higgs1995-xc}; see Supplementary Materials
+Section \ref{supp:weak-strong}).
 
@@
 insight of \parencite{Santiago2016-mu} is that the ratchet rate under draft is
-given by the probability of fixation $p_f(N_e, s)$
+given by the probability of fixation $p_f(N_d, s)$
 \parencite{Kimura1962-su,Malecot1948-zv} with the draft-rescaled effective
-population size, e.g. $R = N U p_f(N_e, s)$. Given this equation for the
-ratchet and Equation \eqref{eq:polygenic_Ne} for $N_e$ under draft, we have a
-system of non-linear equations that can be solved numerically for $N_e$ and
-$R$.
-
-\subsection*{Simulations of the Diploid Model}
+population size, e.g. $R = N U p_f(N_d, s)$. Given this equation for the
+ratchet and Equation \eqref{eq:polygenic_Ne} for $N_d$ under draft, we have a
+system of two non-linear equations that can be solved numerically for $N_d$ and
+$R$ for each segment,
+
+\begin{align}
+  \label{eq:}
+  N_d &= N \exp \left( -V_a \frac{Q^2}{2} \right) & \text{\emph{draft-effective population equation}} \\
+  R &= \frac{4N_d U s}{\exp(4 N_d s)-1}  & \text{\emph{ratchet equation}}
+\end{align}
+
+where $V_a = (U-R)s$ as in Equation \eqref{eq:Va}.
+
+% where,
+
+% \begin{align}
+%   V_a &= (U - R(x)) s & \text{\emph{fitness variance equation}} \\
+%   Q^2 &= \frac{2}{(1-Z)(2-(2-M)Z)} & \text{\emph{linkage inflation factor}} \\
+%   Z &= 1 - \frac{Us}{U-{R(x)}} & \text{\emph{}} \text{\emph{variance decay rate}}
+% \end{align}
+
+\subsection*{Segment Simulations of the Diploid Model}
+
+Given that Santiago and Caballero's (\citeyear{Santiago2016-mu}) model is a
+haploid-model for a single region under equilibrium negative selection, we
+first sought to confirm it could accurately predict the reduction in realistic
+forward-time simulations of diploid populations with recombination. Using SLiM
+\parencite{Haller2019-vu,Haller2023-uk} we simulated a region of $10^5$
+basepairs under varying levels of mutation and selection. We find a close
+correspondence between the observed and predicted reduction in effective
+population size $B=\nicefrac{N_d}{N}$ over all selection and mutation
+parameters including weak selection (Figure 1A), in contrast to classic BGS
+theory. Furthermore, to investigate whether this accuracy was caused by the
+model capturing the equilibrium fitness variance and ratchet rate, we also
+measured these throughout the simulation. Again, we find their theory
+accurately predicts both the ratchet rate (Figure 1B) and the genic fitness
+variance (Figure 1C). Additionally, these figures provide further intuition for
+the underlying negative selection process. When selection is strong the ratchet
+rate is zero, as there is no probability that strongly deleterious can fix.
+However, around $2N_e s \approx 1$, the ratchet begins clicking as $p_F > 0$.
+Ultimately, the ratchet rate reaches the neutral mutation rate, as $V_a = 0$
+and $N_d = N$ (e.g. no draft). Likewise, the additive genic fitness variance
+matches the level predicted by BGS ($Us$), but begins to diverge when the
+ratchet begins clicking.
+
+\subsection*{Chromosome-wide Simulations and Models of Negative Selection}
+
+Given the accuracy of the SC16 model in predicting the reduction in effective
+population size and ratchet rate when equilibrium negative selection processes
+were acting in single segment, we extended their model so that it could be
+applied to whole-genome data.
 

### 2023-05-22 manuscript/manuscript.tex
ms changes

@@
 segregating site fixes and its heterozygosity becomes zero. If we let $R$
-represent twice the rate of the ratchet of deleterious mutations in fixations
-per generation, each fixation removes exactly the variation mutation puts in
-$V_m = Us$ (we omit the segment index for clarity). Thus, under a general
-equilibrium negative selection process,
+represent the rate of the ratchet of deleterious mutations in fixations per
+generation, each fixation removes the equivalent amount of equilibrium
+variation put in by mutation. Thus, the steady-state genic variance under
+mutation and negative selection is (omitting the segment index $i$ for
+clarity),
 
@@
   \label{eq:Va}
-  V_{a} = (U - R)s. 
-\end{align}
-
-Note that since mutation is the ultimate source of all alleles that are fixed
-by the ratchet or are lost, $U \ge R$. Intuitively, this equilibrium genic
-fitness variation is a balance between the rate of gain of genic fitness
-variation due to mutation and its loss due to fixation. Our results differ from
+  V_{a} = (U - 2 R)s. 
+\end{align}
+
+where the condition $V_a \ge 0$ is met when the probability of fixation is less
+than or equal to the neutral fixation probability of $\nicefrac{1}{2N}$, which
+is held when considering deleterious mutations. This equation connects the
+equilibrium additive genic fitness variance to the flux of new variation in due
+to new deleterious mutations and the flux out due to their substitution and the
+decline in mean population fitness. When $R=0$, the selection is so strong it
+cannot fix, and the equilibrium fitness variation is due entirely to young rare
+mutations before their extinction $V_a = V_a^{BGS} = Us$. Santiago and
+Caballero demonstrate Equation \eqref{eq:Va} through Fisher's Fundamental
+Theorem of Natural Selection, but we find an alternative proof
+(\cite{Higgs1995-xc}; see Supplementary Materials Section
+\ref{supp:weak-strong}). We find that the steady-state additive genic variance
+in Equation \eqref{eq:Va} also results from diffusion models with a flux of
+mutations into discrete sites (\cite{Kimura1969-jw}; see Supplementary
+Materials Section \ref{supp:kimura}). Our results differ from
 \parencite{Santiago2016-mu}, since we find from simulations their expression
 models the additive \emph{genic} fitness variation, not additive genetic
-fitness variation (see Section XXX), since it does not adequately capture
-linkage disequilibria dynamics. Santiago and Caballero demonstrate Equation
-\eqref{eq:Va} through Fisher's Fundamental Theorem of Natural Selection, but we
-find an alternative proof (\cite{Higgs1995-xc}; see Supplementary Materials
-Section \ref{supp:weak-strong}).
-
-However, closed-form expressions for the rate of the ratchet $R$ have generally
-been hard to find \parencite{Haigh1978-gt,Higgs1995-xc,Gessler1995-hz}. The key
-insight of \parencite{Santiago2016-mu} is that the ratchet rate under draft is
-given by the probability of fixation $p_f(N_d, s)$
+fitness variation (see Section XXX). This is because Equation \eqref{eq:Va}
+does not capture the expected build-up of negative linkage disequilibria when
+selection is only mildly deleterious.
+
+While using Equation \eqref{eq:Va} in Equation \eqref{eq:main_Ne} leads to a
+prediction for the draft-effective population size $N_d$, closed-form
+expressions for the rate of the ratchet $R$ have generally been hard to find
+\parencite{Haigh1978-gt,Higgs1995-xc,Gessler1995-hz}. The key insight of
+\parencite{Santiago2016-mu} is that the ratchet rate under draft is given by
+the probability of fixation $p_f(N_d, s)$
 \parencite{Kimura1962-su,Malecot1948-zv} with the draft-rescaled effective
@@
 accurately predicts both the ratchet rate (Figure 1B) and the genic fitness
-variance (Figure 1C). Additionally, these figures provide further intuition for
-the underlying negative selection process. When selection is strong the ratchet
-rate is zero, as there is no probability that strongly deleterious can fix.
-However, around $2N_e s \approx 1$, the ratchet begins clicking as $p_F > 0$.
-Ultimately, the ratchet rate reaches the neutral mutation rate, as $V_a = 0$
-and $N_d = N$ (e.g. no draft). Likewise, the additive genic fitness variance
-matches the level predicted by BGS ($Us$), but begins to diverge when the
-ratchet begins clicking.
+variance (Figure 1C).
+
+Additionally, these figures provide further intuition for the underlying
+negative selection process. When selection is strong the ratchet rate is zero,
+as there is no probability that strongly deleterious can fix. However, around
+$2N_e s \approx 1$, the ratchet begins clicking as $p_F > 0$. Ultimately, the
+ratchet rate reaches the neutral mutation rate, as $V_a = 0$ and $N_d = N$
+(e.g. no draft). Likewise, the additive genic fitness variance matches the
+level predicted by BGS ($Us$), but begins to diverge when the ratchet begins
+clicking. This occurs when the level of variation diverges from the
+deterministic mutation-selection equilibrium $V_a = Us$ due to the stochastic
+clicking of the ratchet at rate $R$.
 
@@
 were acting in single segment, we extended their model so that it could be
-applied to whole-genome data.
+applied to whole-genome data. Our method, implemented in the software
+\texttt{bgspy}, calculates the equilibrium additive genic fitness variance
+$V_a$ and ratchet rate $R$ for each specified segment in the genome for a range
+of mutation rates and selection coefficients and a fixed genome-wide
+drift-effective population size $N$ (see Methods XXX). Genomic regions
+putatively under negative selection (e.g. coding sequences or UTRs) and the
+recombination rate map are specified as arguments so that the predicted values
+accurately reflect local genomic details. Note that each equation is solved
+independently, since here we assume no selective interference between segments.
+
+To assess our genome-wide extension of the SC16 model, we simulated negative
+selection for fixed mutation rates and selection coefficient on human
+chromosome 10, using the HapMap recombination map (XXX) and in coding regions,
+PhastCons regions, and UTRs as an approximate set of putatively conserved
+segments. We then compared the reduction over an average of XXX simulation
+replicates with the predicted reduction using our method.
+
+We find that our theory closely matches simulations when selection is strong
+(Figure 1XXX and Figure 1XXX). Since our theory is identical to classic BGS in
+this domain, this provides the first genome-wide realistic forward-time
+simulation confirmation of this theory. However, we find slight discrepancies
+in regions with low recombination (Figure XXX). Second, we find our theory is
+vastly more accurate than classic BGS theory when selection is weak ($2N_e s
+\ll 1$; Figure 1XXX). However, when the strength of selection is at the
+drift-barrier of $2N_e s \approx 1$), we find our theory has about a 10\%
+relative error rate, which is slightly lower than classic BGS theory.
+
+We first hypothesized that this discrepancy could be because slight selective
+interference between segments
+
+could be due to the build up of negative
+linkage disequilibrium
 
@@
 \begin{align}
-  V_{A} = Us - Rs. 
-\end{align}
-
-where $U = 2u$ and $R=2r$ is the diploid rate of the ratchet, which depends on
+  \label{supp-eq:Va}
+  V_{A} = Us - 2Rs. 
+\end{align}
+
+where $U = 2u$ and $R=r$ is the diploid rate of the ratchet, which depends on
 higher moments of the fitness distribution, but only differs from $r$ through a
@@
 equilibrium variance $\widetilde{V} > 0$ for $s>0$.
+
+\subsection*{Steady-state Equilibrium Genic Variance} \label{supp:genic-va}
+
+We also find that Equation \eqref{eq:Va} (and Supplementary Materials Equation
+\ref{supp-eq:Va} above) can also be found through Kimura's diffusion models
+with a flux of mutations into infinite, discrete sites
+(\citeyear{Kimura1969-jw}). This model integrates the Kolmogorov backwards
+equation over time and forward variable to find the expected value of a
+functional of the initial frequency. Using the functional $I(p) = s^2 p(1-p)$
+for genic variance given mutations start at frequency $p$, one finds that the
+steady-state genic variance for $p=\nicefrac{1}{2N_e}$ is
+
+\begin{align}
+    V_a^{SS} &= 2 U N s \left(\frac{1}{2N_e} - p_f(N_e, s)\right) \\
+             &= \left(\frac{UN}{N_e} - 2 U Np_f(N_e, s))\right)s \nonumber \\
+             &\approx \left(U - 2 R\right)s \nonumber
+\end{align}
+
+which is approximately Equation \eqref{eq:Va}.
 

### 2023-05-24 manuscript/manuscript.tex
local fig AI stuff
new figures, captions

@@
 
-\begin{figure}[!htb]
+\begin{figure}[htbp]
   \centering
-  \includegraphics{figures/figure1.pdf}
-  \caption{}
+  \includegraphics[width=\textwidth]{figures/figure_1.pdf}
+  \caption{Theory compared to simulation results. (A) The predicted reduction
+      factor under classic B theory (dark gray line) and the diploid SC16 model
+      (colored lines corresponding to mutation rate) compared to average
+      reduction across XXX simulation replicates (points). (B) The predicted
+      ratchet rate under the SC16 model scaled by mutation rate (colored lines)
+      compared to the ratchet estimated from simulation (points). When $2Ns>1$,
+      the ratchet rate is near zero. (C) The genic variance from simulations
+      (points) against the predicted variance under the SC16 model (colored
+      lines). As the ratchet begins to click, the genic variance is decreased
+      from the level expected under strong BGS (dashed line). (D, bottom) The
+      mean squared error (MSE) between whole-chromosome simulations and
+      predicted classic B (dots), new B' (solid), and locally-rescaled B'
+      (dashed) for different mutation rates (colors). The dashed horizontal
+      line is the approximate theoretic minimum MSE  (D, top) The build up of
+      negative linkage disequilibria around $2Ns=1$ in whole-chromosome
+      simulations. (E) The average B across 100 chromosome 10 simulation
+  replicates (gray) against different predictions, for parameters that
+correspond to $2Ns < 1$, $2Ns = 1$, and $2Ns > 1$. }
   \label{fig:figure-1}

### 2023-05-25 manuscript/manuscript.tex
new ms secs

@@
 
-\section*{Results}
-
-Similar to previous work, our method characterizes the negative selection
-process through the signal left on pairwise diversity at linked sites. Although
-pairwise diversity is a less-informative statistic than the full site frequency
-spectrum, it is much easier to derive models of how selection alters expected
-pairwise coalescence times at linked sites than to describe the entire
-genealogy. Spatial patterns of pairwise diversity are highly variable across
-eukaryotic genomes, as linked selection leads diversity to become correlated
-with spatial heterogeneity in recombination rates and the density of conserved
-sites. Statistical methods then fit observed diversity to the expected level
-under a particular theoretic model of linked selection processes. Consequently,
-statistical inferences about selection may be incorrect when the underlying
-theory inaccurately predicts average pairwise coalescence times.
-
-\subsection*{Theory}
+\section*{Theory}
 
@@
 \begin{align}
-  \label{eq:}
+  \label{eq:main_eqns}
   N_d &= N \exp \left( -V_a \frac{Q^2}{2} \right) & \text{\emph{draft-effective population equation}} \\
@@
 
-\subsection*{Segment Simulations of the Diploid Model}
-
-Given that Santiago and Caballero's (\citeyear{Santiago2016-mu}) model is a
-haploid-model for a single region under equilibrium negative selection, we
-first sought to confirm it could accurately predict the reduction in realistic
-forward-time simulations of diploid populations with recombination. Using SLiM
-\parencite{Haller2019-vu,Haller2023-uk} we simulated a region of $10^5$
-basepairs under varying levels of mutation and selection. We find a close
-correspondence between the observed and predicted reduction in effective
-population size $B=\nicefrac{N_d}{N}$ over all selection and mutation
-parameters including weak selection (Figure 1A), in contrast to classic BGS
-theory. Furthermore, to investigate whether this accuracy was caused by the
-model capturing the equilibrium fitness variance and ratchet rate, we also
-measured these throughout the simulation. Again, we find their theory
-accurately predicts both the ratchet rate (Figure 1B) and the genic fitness
-variance (Figure 1C).
-
-Additionally, these figures provide further intuition for the underlying
-negative selection process. When selection is strong the ratchet rate is zero,
-as there is no probability that strongly deleterious can fix. However, around
-$2N_e s \approx 1$, the ratchet begins clicking as $p_F > 0$. Ultimately, the
-ratchet rate reaches the neutral mutation rate, as $V_a = 0$ and $N_d = N$
-(e.g. no draft). Likewise, the additive genic fitness variance matches the
-level predicted by BGS ($Us$), but begins to diverge when the ratchet begins
-clicking. This occurs when the level of variation diverges from the
-deterministic mutation-selection equilibrium $V_a = Us$ due to the stochastic
-clicking of the ratchet at rate $R$.
-
-\subsection*{Chromosome-wide Simulations and Models of Negative Selection}
-
-Given the accuracy of the SC16 model in predicting the reduction in effective
-population size and ratchet rate when equilibrium negative selection processes
-were acting in single segment, we extended their model so that it could be
-applied to whole-genome data. Our method, implemented in the software
-\texttt{bgspy}, calculates the equilibrium additive genic fitness variance
-$V_a$ and ratchet rate $R$ for each specified segment in the genome for a range
-of mutation rates and selection coefficients and a fixed genome-wide
-drift-effective population size $N$ (see Methods XXX). Genomic regions
-putatively under negative selection (e.g. coding sequences or UTRs) and the
-recombination rate map are specified as arguments so that the predicted values
-accurately reflect local genomic details. Note that each equation is solved
-independently, since here we assume no selective interference between segments.
-
-To assess our genome-wide extension of the SC16 model, we simulated negative
-selection for fixed mutation rates and selection coefficient on human
-chromosome 10, using the HapMap recombination map (XXX) and in coding regions,
-PhastCons regions, and UTRs as an approximate set of putatively conserved
-segments. We then compared the reduction over an average of XXX simulation
-replicates with the predicted reduction using our method.
-
-We find that our theory closely matches simulations when selection is strong
-(Figure 1XXX and Figure 1XXX). Since our theory is identical to classic BGS in
-this domain, this provides the first genome-wide realistic forward-time
-simulation confirmation of this theory. However, we find slight discrepancies
-in regions with low recombination (Figure XXX). Second, we find our theory is
-vastly more accurate than classic BGS theory when selection is weak ($2N_e s
-\ll 1$; Figure 1XXX). However, when the strength of selection is at the
-drift-barrier of $2N_e s \approx 1$), we find our theory has about a 10\%
-relative error rate, which is slightly lower than classic BGS theory.
-
-We first hypothesized that this discrepancy could be because slight selective
-interference between segments
-
-could be due to the build up of negative
-linkage disequilibrium
-
-\subsection*{DFE Estimates}
-% - Table s5 https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/genetics/206/1/10.1534_genetics.116.197145/8/files1.pdf?Expires=1686357613&Signature=cp9qoI2ChCb6GFZ74HI4eWcA~MQ4sHlWyeI3Zp49-5xEORzMfYlSRxzVvZlo8ZoGWDi0G3PnL00ZplM-~or4YIVm2uhkmqjf61AmC7RdZ1QTak4j0kFMQ0wAbM9pusiV3gIepTQHq2HKRKwgS~ypmv~0oqO5dDxaq9nhOWIBBmLAA49oHWuxUtnRe3sEYATwCwjIFEomhgJpg7vKGCF~87GdCV4JETW4PN2VvD4hsu2AVRd8yA~vUJvsXfhWwcElDeY4fsgpprl1lC3XGjJCnqdwbzRGPV5LNx5XEKLptcZtLeLcOmFJCvmVLe0bewJ7Vta0TSQM7T9AuklompjPPw__&Key-Pair-Id=APKAIE5G5CRDK6RD3PGA
+\section*{Results}
+
+Similar to previous work, our method characterizes the negative selection
+process through the signal left on pairwise diversity at linked sites. Although
+pairwise diversity is a less-informative statistic than the full site frequency
+spectrum, it is much easier to derive models of how selection alters expected
+pairwise coalescence times at linked sites than to describe the entire
+genealogy. Spatial patterns of pairwise diversity are highly variable across
+eukaryotic genomes, as linked selection leads diversity to become correlated
+with spatial heterogeneity in recombination rates and the density of conserved
+sites. Statistical methods then fit observed diversity to the expected level
+under a particular theoretic model of linked selection processes. Consequently,
+statistical inferences about selection may be incorrect when the underlying
+theory inaccurately predicts average pairwise coalescence times.
+
+\subsection*{Simulations of a Segment under Negative Selection}
 
@@
       (dashed) for different mutation rates (colors). The dashed horizontal
-      line is the approximate theoretic minimum MSE  (D, top) The build up of
+      line is the approximate theoretic minimum MSE. (D, top) The build up of
       negative linkage disequilibria around $2Ns=1$ in whole-chromosome
-      simulations. (E) The average B across 100 chromosome 10 simulation
+      simulations. (E) The average B map from 100 chromosome 10 simulation
   replicates (gray) against different predictions, for parameters that
-correspond to $2Ns < 1$, $2Ns = 1$, and $2Ns > 1$. }
+correspond to $2Ns < 1$, $2Ns = 1$, and $2Ns > 1$. The chromosome shows 
+the density of conserved sites and recombination map used in simulations. }
   \label{fig:figure-1}
 \end{figure}
+
+Given that Santiago and Caballero's (\citeyear{Santiago2016-mu}) model is a
+haploid-model for a single region under equilibrium negative selection, we
+first sought to confirm it could accurately predict the reduction in realistic
+forward-time simulations of diploid populations with recombination. Using SLiM
+\parencite{Haller2019-vu,Haller2023-uk} we simulated a region of $10^5$
+basepairs under varying levels of mutation and selection. We find a close
+correspondence between the observed and predicted reduction in effective
+population size $B=\nicefrac{N_d}{N}$ over all selection and mutation
+parameters including weak selection (Figure \ref{fig:figure-1}A), in contrast
+to classic BGS theory. Furthermore, to investigate whether this accuracy was
+caused by the model capturing the equilibrium fitness variance and ratchet
+rate, we also measured these throughout the simulation. Again, we find the
+diploid SC16 theory accurately predicts both the ratchet rate (Figure
+1\ref{fig:figure-1}B) and the genic fitness variance (Figure
+1\ref{fig:figure-1}C).
+
+Additionally, these figures provide further intuition for the underlying
+negative selection process. When selection is strong the ratchet rate is zero,
+as there is no probability that strongly deleterious can fix. However, around
+$2N_e s \approx 1$, the ratchet begins clicking as $p_F > 0$. Ultimately, the
+ratchet rate reaches the neutral mutation rate, as $V_a = 0$ and $N_d = N$
+(e.g. no draft). Likewise, the additive genic fitness variance matches the
+level predicted by BGS ($Us$), but begins to diverge when the ratchet begins
+clicking. This occurs when the level of variation diverges from the
+deterministic mutation-selection equilibrium $V_a = Us$ due to the stochastic
+clicking of the ratchet at rate $R$.
+
+\subsection*{Chromosome-wide Simulations and Models of Negative Selection}
+
+Given the accuracy of the SC16 model in predicting the reduction in effective
+population size and ratchet rate when equilibrium negative selection processes
+were acting in single segment, we extended their model so that it could be
+applied to whole-genome data. Our method, implemented in the software
+\texttt{bgspy}, calculates the equilibrium additive genic fitness variance
+$V_a$ and ratchet rate $R$ for each specified segment in the genome for a range
+of mutation rates and selection coefficients and a fixed genome-wide
+drift-effective population size $N$ (see Methods XXX). Genomic regions
+putatively under negative selection (e.g. coding sequences or UTRs) and the
+recombination rate map are specified as arguments so that the predicted values
+accurately reflect local genomic details. Note that each equation is solved
+independently, since here we assume no selective interference between segments.
+We refer to the prediction reductions along the genome from our approach as B'
+maps (analogous to B maps from \cite{McVicker2009-ax}).
+
+To assess our B' maps, we simulated negative selection for fixed mutation rates
+and selection coefficient on human chromosome 10 using realistic recombination
+and conserved feature maps. Averaging over $n=100$ simulation replicates, we
+estimated the simulation expected reduction map and compared this to the
+predicted B and B' maps for the same parameters. We find that our B' maps and
+the classic BGS theory B maps closely match simulations when selection is
+strong (top row of Figure \ref{fig:figure-1}E). This provides the first
+realistic forward-time simulation confirmation of classic background selection
+theory. However, we find slight discrepancies in regions with low recombination
+(Figure \ref{fig:figure-1}E). Second, we find our theory is vastly more
+accurate than the classic B maps when selection is very weak ($2N_e s \ll 1$;
+bottom row of Figure \ref{fig:figure-1}E), which is expected since BGS theory
+does not work in this domain. Across all mutation and selection parameters
+simulated, the relative error of the classic B maps is 14.6\% whereas the
+relative error in the new B' maps is 5\%. Additionally, we find that the mean
+squared error between simulations and B' maps is close to the theoretic lower
+bound set by the coalescence variance \parencite[see Methods
+XXX]{Tajima1983-gu}.
+
+Nearly all of the error between the new B' maps and chromosome-wide simulations
+occurs around the drift-barrier domain of $2Ns \approx 1$ (Figure
+\ref{fig:figure-1}D). The higher error in this domain occurs across all
+mutation rates. First, we hypothesized that the reason this occurred was the
+when we numerically solve Equations \eqref{eq:main_eqns} for each segment, this
+does does not consider the reduction due to selection at other linked segments.
+In particular, we use $N=1000$ corresponding to the number of diploids in the
+simulations rather than $B'(x) N$ at position $x$. To test this, we implemented
+a locally-rescaled version of the B' maps, which numerically solves Equations
+\eqref{eq:main_eqns} for each segment at position $x$ taking into account the
+reduction experienced due to selection at other segments by setting $B'(x)N$ as
+the population size (see Method XXX). We find the locally-rescaled B' maps
+reduce the relative error from 5\% to 0.4\% and mean squared error (Figure
+\ref{fig:figure-1}D, dashed colored lines), but does not entirely eliminate the
+error in the $2Ns \approx 1$ domain.
+
+Finally, we hypothesized that the remaining error is because of the build up of
+negative linkage disequilibria between selected sites due to Hill--Robertson
+interference \parencite{Hill1966-kd,McVean2000-bt,Comeron2007-wq}. To assess
+this possibility, we calculated the sum of all linkage disequilibria in our
+chromosome-wide simulations. We find negative linkage disequilibria build ups
+around $2Ns \approx 1$ (Figure \ref{fig:figure-1}D, top row) and is greater
+when mutation rates are higher (Supplementary Figure XXX). This is consistent
+our view that the equilibrium fitness variance under the SC16 model is the
+additive \emph{genic} fitness variance, which differs from the additive genetic
+variance by the sum of linkage disequilibria between selected sites,
+$\delta_{LD} = s^2 \sum_{i\nej} D_{i,k}$.
+
+\subsection*{Estimates of Negative Selection in Humans}
+
+\subsection*{DFE Estimates}
+% - Table s5 https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/genetics/206/1/10.1534_genetics.116.197145/8/files1.pdf?Expires=1686357613&Signature=cp9qoI2ChCb6GFZ74HI4eWcA~MQ4sHlWyeI3Zp49-5xEORzMfYlSRxzVvZlo8ZoGWDi0G3PnL00ZplM-~or4YIVm2uhkmqjf61AmC7RdZ1QTak4j0kFMQ0wAbM9pusiV3gIepTQHq2HKRKwgS~ypmv~0oqO5dDxaq9nhOWIBBmLAA49oHWuxUtnRe3sEYATwCwjIFEomhgJpg7vKGCF~87GdCV4JETW4PN2VvD4hsu2AVRd8yA~vUJvsXfhWwcElDeY4fsgpprl1lC3XGjJCnqdwbzRGPV5LNx5XEKLptcZtLeLcOmFJCvmVLe0bewJ7Vta0TSQM7T9AuklompjPPw__&Key-Pair-Id=APKAIE5G5CRDK6RD3PGA
 
@@
 \section*{Methods}
+
+\subsection*{Chromosome-wide Simulations}
+
+, using the HapMap recombination map (XXX) and in coding regions, PhastCons
+regions, and UTRs as an approximate set of putatively conserved segments
+
+\subsection*{Local-rescaled }
+We note that it is computationally intractable to solve this equation excluding
+the contribution of the focal segment
 

### 2023-06-07 manuscript/manuscript.tex
new supp, ms stuff

@@
 \begin{abstract} 
-
+% TODO: emphasize connection to nearly neutral theory
 \end{abstract}
@@
 \parencite{Haigh1978-gt,Higgs1995-xc,Gessler1995-hz}. The key insight of
-\parencite{Santiago2016-mu} is that the ratchet rate under draft is given by
-the probability of fixation $p_f(N_d, s)$
+\textcite{Santiago2016-mu} is that the ratchet rate under draft is determined
+by the probability of fixation $p_f(N_d, s)$
 \parencite{Kimura1962-su,Malecot1948-zv} with the draft-rescaled effective
@@
 
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=\textwidth]{figures/figure_1.pdf}
-  \caption{Theory compared to simulation results. (A) The predicted reduction
-      factor under classic B theory (dark gray line) and the diploid SC16 model
+\begin{figure}[htbp] \centering
+    \includegraphics[width=\textwidth]{figures/figure_1.pdf} \caption{Theory
+        compared to simulation results. (A) The predicted reduction factor
+        under classic B theory (dark gray line) and the diploid SC16 model
         (colored lines corresponding to mutation rate) compared to average
         reduction across XXX simulation replicates (points). (B) The predicted
-      ratchet rate under the SC16 model scaled by mutation rate (colored lines)
-      compared to the ratchet estimated from simulation (points). When $2Ns>1$,
-      the ratchet rate is near zero. (C) The genic variance from simulations
-      (points) against the predicted variance under the SC16 model (colored
-      lines). As the ratchet begins to click, the genic variance is decreased
-      from the level expected under strong BGS (dashed line). (D, bottom) The
-      mean squared error (MSE) between whole-chromosome simulations and
-      predicted classic B (dots), new B' (solid), and locally-rescaled B'
-      (dashed) for different mutation rates (colors). The dashed horizontal
-      line is the approximate theoretic minimum MSE. (D, top) The build up of
-      negative linkage disequilibria around $2Ns=1$ in whole-chromosome
-      simulations. (E) The average B map from 100 chromosome 10 simulation
-  replicates (gray) against different predictions, for parameters that
-correspond to $2Ns < 1$, $2Ns = 1$, and $2Ns > 1$. The chromosome shows 
-the density of conserved sites and recombination map used in simulations. }
+        ratchet rate under the SC16 model scaled by mutation rate (colored
+        lines) compared to the ratchet estimated from simulation (points). When
+        $2Ns>1$, the ratchet rate is near zero. (C) The genic variance from
+        simulations (points) against the predicted variance under the SC16
+        model (colored lines). As the ratchet begins to click, the genic
+        variance is decreased from the level expected under strong BGS (dashed
+        line). (D, bottom) The mean squared error (MSE) between
+        whole-chromosome simulations and predicted classic B (dots), new B'
+        (solid), and locally-rescaled B' (dashed) for different mutation rates
+        (colors). The dashed horizontal line is the approximate theoretic
+        minimum MSE. (D, top) The build up of negative linkage disequilibria
+        around $2Ns=1$ in whole-chromosome simulations in bottom panel. (E) The
+        average B map from 100 chromosome 10 simulation replicates (gray)
+    against different predictions, for parameters that correspond to $2Ns < 1$,
+$2Ns = 1$, and $2Ns > 1$. The chromosome shows the density of conserved sites
+and recombination map used in simulations. }
   \label{fig:figure-1}
@@
 chromosome-wide simulations. We find negative linkage disequilibria build ups
-around $2Ns \approx 1$ (Figure \ref{fig:figure-1}D, top row) and is greater
-when mutation rates are higher (Supplementary Figure XXX). This is consistent
-our view that the equilibrium fitness variance under the SC16 model is the
-additive \emph{genic} fitness variance, which differs from the additive genetic
-variance by the sum of linkage disequilibria between selected sites,
-$\delta_{LD} = s^2 \sum_{i\nej} D_{i,k}$.
+around $2Ns \approx 1$ (Figure \ref{fig:figure-1}D, top row) and is stronger
+when mutation rates are higher (Supplementary Figure XXX). As $Ns \to 0$, the
+variance in LD inflates as expected \parencite{Ohta1969-ae,Hill1968-ue}.
+Overall, the build up of negative LD is consistent our view that the
+equilibrium fitness variance modeled by the SC16 theory is the additive
+\emph{genic} fitness variance, which differs from the additive genetic variance
+by the sum of linkage disequilibria between selected sites (i.e. $\delta_{LD} =
+s^2 \sum_{i\ne j} D_{i,j}$ where $D_{i,j}$ is the LD between sites $i$ and
+$j$). However, according to theory, the reductions in diversity should be
+determined by levels of additive genetic fitness variance that include the
+contribution of LD (Supplementary Materials Section XXX).
 
 \subsection*{Estimates of Negative Selection in Humans}
+
+Next, we developed a composite-likelihood method to estimate pairwise diversity
+in the absence of linked selection $\pi_0$, the deleterious mutation rate
+($\mu_D$), and distribution of fitness effects ($\mathbf{W}$) under negative
+selection. Because the focus of our work is to fix deficiencies in the classic
+BGS model, our current implementation does not consider reductions in diversity
+due to recurrent hitchhiking or other positive selection processes. Since
+previous work has established that purifying selection is the dominant mode of
+linked selection in humans and hitchhiking plays a limited role
+\parencite{McVicker2009-ax,Murphy2022-sj}, humans are a suitable organism to
+test our method.
+
+Given that previous methods used classic BGS theory to fit genome-wide patterns
+of human diversity, we were curious whether their qualitative findings would
+change once weak selection was more accurately predicted. In particular, the
+U-shaped relationship between $B'$ and the strength of selection suggests that
+a given reduction factor could be explained by two different selection
+coefficients.
 
@@
 % - Table s5 https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/genetics/206/1/10.1534_genetics.116.197145/8/files1.pdf?Expires=1686357613&Signature=cp9qoI2ChCb6GFZ74HI4eWcA~MQ4sHlWyeI3Zp49-5xEORzMfYlSRxzVvZlo8ZoGWDi0G3PnL00ZplM-~or4YIVm2uhkmqjf61AmC7RdZ1QTak4j0kFMQ0wAbM9pusiV3gIepTQHq2HKRKwgS~ypmv~0oqO5dDxaq9nhOWIBBmLAA49oHWuxUtnRe3sEYATwCwjIFEomhgJpg7vKGCF~87GdCV4JETW4PN2VvD4hsu2AVRd8yA~vUJvsXfhWwcElDeY4fsgpprl1lC3XGjJCnqdwbzRGPV5LNx5XEKLptcZtLeLcOmFJCvmVLe0bewJ7Vta0TSQM7T9AuklompjPPw__&Key-Pair-Id=APKAIE5G5CRDK6RD3PGA
+
+Like other methods, our method divides the putatively conserved basepairs up
+
+Generally, when the putatively conserved region is larger, there is a tradeoff
+in two dimensions. First, more bases are conserved according to the same DFE,
+so the total mutational load increases. This causes the tradeoff between
+mutation rate estimate and CADD fraction we observe across both our B' model
+and the classic B when using sparse tracks (Figure 2XXX).
+
+We find that mutation rate estimates are fairly sensitive to model choice
+(Figure 2XXX).
+
+Since B' has a U-shaped relationship with the selection coefficient, we
+hypothesized that our model may reveal true identifiability issues in
+estimating the DFE that do not occur under the classic B model. We find that
+this is not the case for our models fit at the megabase-scale. The DFE
+estimates for our CADD6 and CADD8 models are consistent (Figure XXX), and both
+produce similar mutation rates.
+
+While we do not comprehensively fit all of our model to other spatial scales,
+we
+
+However, while a given reduction level may be explained by either a weak or
+strong selection coefficient, the spatial extent of these reductions caused by
+weak selection is much narrower than under strong selection.
+
+We see evidence this is the case in several of our models, where
+genome-wide diversity fits equally well between models with mutation rates
+
+Indeed, we see this by comparing parameter estimates of the CADD6
+sparse-track model using classic BGS B and our B' method.
+
+(which mimics the best-fitting model of
+\cite{Murphy2022-sj})
+
+(see also
+Appendix 1, Figure 26A of \cite{Murphy2022-sj})
+
+correlation between s and feature
+
+Overall, we find evidence that negative selection models are very sensitive to
+the input track. Because inference based on our B' method accommodates weaker
+selection, the primary models we fit here more robust to misspecification of
+the regions
 
@@
 regions, and UTRs as an approximate set of putatively conserved segments
+
+We calculate the mean squared error (MSE) and relative error between
+whole-chromosome simulations and the predicted B and B' maps by averaging
+branch diversity in 10 kilobase windows (tskit, XXX) and dividing by the
+neutral expectation of $4N$. The theoretic lower bound of the MSE is due the
+noise in estimating the expected reduction map from simulations. For $n=100$
+replicates, the sampling noise is minimal relative to the evolutionary noise in
+the coalescence process \parencite{Tajima1983-gu}. We use this as the theoretic
+lower bound assuming the maximum coalescence noise at $B=1$, $\var(\bar{B})
+\approx \nicefrac{2}{9n}$.
 

### 2023-06-08 manuscript/manuscript.tex
new figs

@@
 
+\subsection*{Annotation Model Comparison}
+
+Our method takes tracks of annotated features that are \emph{a priori} expected
+to have similar fitness effects, and estimates the distribution of fitness
+effects for each feature type. Additionally, our method estimates the
+per-basepair per-generation deleterious mutation rate ($\mu_\text{del}$) and
+the background level of pairwise diversity ($\pi_0$). Since the theory used by
+our method can in principle accommodate weak selection and neutrality. To test
+this, for each our annotation models, we fit (1) ``sparse" models that only
+cover the fraction of the genome that is the union of all tracks, and (2)
+``full`` models that are sparse models with the remaining complement of the
+genome labeled as ``other".
+
+Overall, we find that
+
+\begin{figure}[htbp] \centering
+    \includegraphics[width=\textwidth]{figures/figure_2.pdf} 
+    \caption{The $R^2$ estimates for a sparse (A) and full (B) models, for all 
+    populations (colors). Round points are our B' method; diamonds are the 
+    classic B. For our method, we also calculate leave-one-out out-sample $R^2$
+    (lighter filled points).}
+  \label{fig:figure-2}
+\end{figure}
+
 \subsection*{Estimates of Negative Selection in Humans}
+
+\begin{figure}[htbp] \centering
+    \includegraphics[width=\textwidth]{figures/figure_3.pdf} 
+    \caption{The distribution of fitness effects of new mutations estimates for
+        Yoruba individuals. (A) The DFEs using sparse (left column) and
+        full-coverage (right column) tracks, across different annotation models
+        (row). Color indicates the feature type. (B) The DFE of the
+        full-coverage CDS-priority model comparing the estimates across
+    populations.}
+  \label{fig:figure-3}
+\end{figure}
 
@@
 \subsection*{DFE Estimates}
+
+\begin{table}
+\centering
+\caption{$R^2$ and mutation rate estimates for all models}
+\begin{tabular}{lll|crr|cc}
+    \textbf{Model} & \textbf{Track type} & \textbf{Pop.} & \textbf{B' $R_\text{LOO}^2$} & \textbf{B' $R_\text{IS}^2$} & \textbf{B $R_\text{IS}^2$} & \textbf{B' $\hat{\mu} \times 10^{-8}$} & \textbf{B $\hat{\mu} \times 10^{-8}$} \\[0.5ex]
+\hline
+\hline
+    \textbf{phastcons$>$CDS$>$genes} &            sparse &          YRI &                        \textbf{68.45} &             68.16 &            65.23 &                                 1.8636 &                                0.2975 \\
+phastcons$>$CDS$>$genes &              full &          YRI &                        68.19 &             68.12 &            63.67 &                                 1.7082 &                                0.1666 \\
+               CADD 6\% &              full &          YRI &                        68.08 &             68.08 &            62.96 &                                 1.4612 &                                0.1721 \\
+               CADD 6\% &            sparse &          YRI &                        68.04 &             68.03 &            68.01 &                                 2.0912 &                                2.0716 \\
+               CADD 8\% &              full &          YRI &                        66.98 &             67.11 &            63.61 &                                 1.0126 &                                0.1713 \\
+               CADD 8\% &            sparse &          YRI &                        66.67 &             67.01 &            66.98 &                                 1.5688 &                                1.5473 \\
+CDS$>$genes$>$phastcons &              full &          YRI &                        64.90 &             65.51 &            61.23 &                                 4.1951 &                                0.1620 \\
+CDS$>$genes$>$phastcons &            sparse &          YRI &                        64.90 &             65.51 &            63.16 &                                 4.1795 &                                0.3070 \\
+\textbf{phastcons$>$CDS$>$genes} &            sparse &          CEU &                        \textbf{61.98} &             63.02 &            60.02 &                                 2.0194 &                                0.3036 \\
+phastcons$>$CDS$>$genes &              full &          CEU &                        61.78 &             63.01 &            58.57 &                                 2.3499 &                                0.1679 \\
+               CADD 6\% &              full &          CEU &                        61.74 &             62.94 &            57.42 &                                 1.5365 &                                0.1739 \\
+               CADD 6\% &            sparse &          CEU &                        61.66 &             62.86 &            62.84 &                                 2.1215 &                                2.1084 \\
+               CADD 8\% &              full &          CEU &                        60.80 &             62.12 &            57.95 &                                 1.1572 &                                0.1742 \\
+               CADD 8\% &            sparse &          CEU &                        60.59 &             62.06 &            62.04 &                                 1.6030 &                                1.5876 \\
+               CADD 6\% &              full &          CHB &                        59.62 &             61.31 &            55.31 &                                 1.5083 &                                0.1727 \\
+               CADD 6\% &            sparse &          CHB &                        59.54 &             61.21 &            61.20 &                                 2.1242 &                                2.1170 \\
+               \textbf{phastcons$>$CDS$>$genes} &            sparse &          CHB &                        \textbf{59.44} &             61.08 &            57.88 &                                 2.1982 &                                0.3060 \\
+phastcons$>$CDS$>$genes &              full &          CHB &                        59.29 &             61.08 &            56.71 &                                 2.5573 &                                0.1694 \\
+               CADD 8\% &              full &          CHB &                        58.88 &             60.63 &            55.86 &                                 1.1128 &                                0.1733 \\
+               CADD 8\% &            sparse &          CHB &                        58.65 &             60.57 &            60.55 &                                 1.6098 &                                1.6001 \\
+CDS$>$genes$>$phastcons &            sparse &          CEU &                        58.58 &             60.30 &            58.41 &                                 4.1041 &                                0.3110 \\
+CDS$>$genes$>$phastcons &              full &          CEU &                        58.58 &             60.30 &            56.75 &                                 4.1001 &                                0.1648 \\
+CDS$>$genes$>$phastcons &            sparse &          CHB &                        56.48 &             58.73 &            56.60 &                                 3.7195 &                                0.3136 \\
+CDS$>$genes$>$phastcons &              full &          CHB &                        56.48 &             58.73 &            54.95 &                                 3.7552 &                                0.1654 \\
+\hline
+\end{tabular}
+\end{table}
+
+Our best-fitting models have out-sample $R^2$ of around 68\% for Yoruba, 62\%
+European, and 60\% for Han Chinese ancestry individuals. However, $R^2$ has a
+theoretic upper bound due to coalescence noise around the predicted mean level
+$\widehat{\pi}(x)$.
+
 % - Table s5 https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/genetics/206/1/10.1534_genetics.116.197145/8/files1.pdf?Expires=1686357613&Signature=cp9qoI2ChCb6GFZ74HI4eWcA~MQ4sHlWyeI3Zp49-5xEORzMfYlSRxzVvZlo8ZoGWDi0G3PnL00ZplM-~or4YIVm2uhkmqjf61AmC7RdZ1QTak4j0kFMQ0wAbM9pusiV3gIepTQHq2HKRKwgS~ypmv~0oqO5dDxaq9nhOWIBBmLAA49oHWuxUtnRe3sEYATwCwjIFEomhgJpg7vKGCF~87GdCV4JETW4PN2VvD4hsu2AVRd8yA~vUJvsXfhWwcElDeY4fsgpprl1lC3XGjJCnqdwbzRGPV5LNx5XEKLptcZtLeLcOmFJCvmVLe0bewJ7Vta0TSQM7T9AuklompjPPw__&Key-Pair-Id=APKAIE5G5CRDK6RD3PGA

### 2023-06-08 manuscript/supplementary.tex
new supp, ms stuff
new figs

@@
 % - edge trunation
+
+\section*{Additional Results}
+
+\subsection*{Model Comparisons}
+
+% \begin{table}[ht]
+%     \caption{CADD 6\% Sparse Track B and B' Comparison}
+%     \label{supptab:cadd6}
+%     \centering
+% \begin{tabular}{lll}
+%                          & \strong{B}              & \strong{B'}                     \\
+%                          \hline
+%     \emph{negative log-likelihood} & 250,622,462,847.73962     & 250,622,458,048.0578      \\
+%     $R^2$                   & 53.89\%                 & 53.9\%                  \\
+%     \hline
+%     $\pi_0$                 & 0.0014272  & 0.0014297  \\
+%     $\mu_\text{del}$        & $1.88 \times 10^{-8}$ & $4.29 \times 10^{-8}$ \\
+%     \\
+%     \emph{sel. coef.} &\multicolumn{1}{c}{\emph{DFE estimates (percent)}} \\
+%     \hline
+%     $10^{-8}$                  & 0.0                    & 0.0                    \\
+%     $10^{-7}$                   & 0.0                    & 0.0                    \\
+%     $10^{-6}$                   & 0.0                    & 0.0                    \\
+%     $10^{-5}$                   & 0.3                    & 0.57                   \\
+%     $10^{-4}$                  & 10.27                  & 0.02                   \\
+%     $10^{-3}$                   & 25.22                  & 0.12                   \\
+%     $10^{-2}$                    & 1.91                   & 0.0                    \\
+%     $10^{-1}$                     & 62.3                   & 0.28                  \\
+%     \hline
+% \end{tabular}
+% \end{table}
+
+\section*{Approximate $R^2$ Limit}
+
+We derive a rough approximation of the theoretic upper bound of $R^2$ by
+noting that,
+
+\begin{align}
+    R^2 &= \frac{SS_\text{reg}}{SS_\text{reg} + SS_\text{res}} \\ 
+        &= \frac{\sum_i (\pi_i - \widehat{\pi}_i)^2}{\sum_i (\pi_i - \widehat{\pi}_i)^2 + SS_\text{res}}
+\end{align}
+
+where $\pi_i$ is the observed diversity in window $i$, and $\widehat{\pi}_i$ is
+the predicted diversity under a particular model.
+
+Then, $SS_\text{res}$ is the residual noise due to the randomness in mutation
+and average coalescence times in the window. If we make the assumptions that
+mutation rates are heterogeneous across the genome and the coalescence rate in
+a window is simply a background rate rescaled by the reduction factor $B_i$,
+then,
+
+\begin{align}
+    SS_\text{res} &= \frac{1}{n} \sum_{i=1}^n \var(w\pi_i) \\ 
+                  &= \frac{1}{n} \sum_{i=1}^n \var(w\pi_0 B_i) 
+\end{align}
+
+where $w$ is the width of the window such that the per-basepair $\pi$ is
+appropriately scaled. Then, we note that the variance in pairwise diversity is
+equivalent to the variance in the number of segregating sites for a sample of 
+two, $\var(\pi) = \var(S_2)$. From \textcite[eq. 7.20]{Wakeley2009-ua}
+
+\begin{align}
+    \var(S_2) = \theta + \theta^2 \frac{2}{\rho^2} \int_0^\rho (\rho - x) f_2(x) dx
+\end{align}
+where,
+\begin{align}
+    f_2(\rho) = \frac{\rho + 18}{\rho^2 + 13 \rho + 18}.
+\end{align}
+
+Since this is a rough approximation, we set $\theta = w \pi_0 B_i$ and fix
+$\rho = w \gamma \pi_0 B_i$ where $\gamma$ is the ratio of recombination to
+mutation rates, $\nicefrac{r}{\mu}$. In humans, $\gamma \approx 1$, but we
+wanted to check the sensitivity of this calculation to this.
+
+This back-of-the-envelope calculation finds that the upper bound of $R^2$ would
+be around 74\% for Yoruba, and 80-81\% for European and Han Chinese models
+(Figure \ref{suppfig:r2_upper} We note, however, that this rough approximation
+assumes constant demography, and that the out-of-Africa bottleneck causes a
+reduction in effective population size that would shrink the variance of
+exponential coalescence times, but a bottleneck could in fact inflate these
+variances. A more accurate approximate of the upper $R^2$ bound could rely on
+simulations with more realistic demography.
+
+\begin{figure}[htbp] \centering
+    \includegraphics[width=0.5\textwidth]{figures/supplementary/suppfigure_R2_upper.pdf}
+    \caption{A rough approximation to the theoretic $R^2$ upper bound assuming constant demography.}
+  \label{suppfig:r2_upper}
+\end{figure}
 

### 2023-06-09 bgspy/likelihood.py
new notebook stuff

@@ -668,13 +668,13 @@ class BGSLikelihood:
             df_merged = pd.merge(df_means_melted, df_errors_melted, on=['t', 'category'])
         return df_merged
 
-    def dfe_plot(self, add_legend=True, legend_kwargs={}, figax=None, ylabel='probability'):
+    def dfe_plot(self, add_legend=True, legend_kwargs={}, figax=None, ylabel='probability', barplot_kwargs={}):
         """
         Plot a boxplot of all features.
         """
         df = self.W_df()
         fig, ax = get_figax(figax)
-        sns.barplot(data=df, x='t', y='mean', hue='category', ax=ax)
+        sns.barplot(data=df, x='t', y='mean', hue='category', ax=ax, **barplot_kwargs)
         xticks = ax.get_xticks()
         xlabels = ax.get_xticklabels()
         ax.set_xticks(xticks, [f"$10^{{{int(np.log10(float(x.get_text())))}}}$" for x in xlabels])

### 2023-06-09 manuscript/manuscript.tex
MS updates, figs

--- 2023-06-07 manuscript/manuscript.tex
@@
 
+\subsection*{Annotation Model Comparison}
+
+Our method takes tracks of annotated features (an ``annotation model") that are
+\emph{a priori} expected to have similar fitness effects, and estimates the
+distribution of fitness effects for each feature type. We consider two classes
+of annotation models: (1) CADD score models, which assign a rank-order
+pathogenicity score to each basepair, and (2) and more interpretable gene
+feature-based models that includes protein coding regions, intron and UTRs, and
+PhastCons regions. We include PhastCons regions because there is abundant
+evidence of highly-conserved non-coding regions in humans that we would expect
+to be under negative selection and impact patterns of linked selection (XXX).
+Since our method fits a DFE for each feature type, PhastCons regions that
+overlap other gene annotations (e.g. coding sequences) must be assigned to
+either of the two feature types. Since this can impact estimates, we fit both
+alternatives: a PhastCons Priority model, where features that overlap PhastCons
+regions are classified as PhastCons, and a Feature Priority model, where gene
+structure features take priority and the PhastCons type catches all
+highly-conserved non-genic regions. In total, we fit four annotation models
+(CADD 6\%, CADD 8\%, PhastCons Priority, and Feature Priority) to high-coverage
+1000 Genome data for three populations: Yoruba (YRI), Han Chinese (CHB), and
+European (CEU) ancestries.
+
+Overall, we find the PhastCons Priority and CADD 6\% models fit about equally
+as well (Figure \ref{fig:figure-2}A), explaining $R^2=68.45$\% and
+$R^2=68.04$\% of the out-sample variance in Yoruba pairwise diversity at the
+megabase-scale, respectively. In populations impacted by the out-of-Africa
+bottleneck, the goodness-of-fit was lower (62.98\% and 59.98\% for CEU and CHB,
+respectively). The CADD 6\% model was previously found by
+\textcite{Murphy2022-sj} to be the best-fitting model using classic BGS. Our
+PhastCons Priority model fits marginally better than the CADD 6\% model
+consistently across all three populations (Supplementary Table XXX). This is
+likely because the PhastCons Priority model has more DFE parameters, and thus
+can fit finer-scale differences in the DFE across its four feature types. By
+contrast, the CADD 6\% track must fit patterns of diversity to conserved sites
+under a single DFE. The CADD 8\% model fit slightly worse than the CADD 6\%
+across all populations, and the Feature Priority model had the poorest-fit.
+
+Since our method is built upon theory that fixes the weak selection problem of
+classic BGS theory, it should in principle fit equally well when the annotation
+model covers the full genome, the bulk of which is expected to be neutrally
+evolution. To test this, for each annotation model (which is ``sparse") we fit
+a ``full" model that contains the remaining complement of the genome labeled as
+``other". Ideally, the goodness-of-fit of fit using our B' method should be
+invariant to whether an annotation model is sparse or full. Indeed, we find
+that
+
+NOTE: there looks like a bug in the table... repeated values.
+
+Each model was fit to the high-coverage 1000 human genome data for
+three populations, YRI, CEU, and CHB. Additionally, we also fit each of these
+models using our method, but using classic BGS theory rather than our B'
+theory. In total, we fit and compared 48 different models.
+
+Overall, we find that the goodness-of-fit, measured by both in- and out-sample
+$R^2$ are nearly identical across sparse and full fits using our B' method,
+regardless of the different annotation models (Figure \ref{fig:figure-2},
+Supplementary Materials Table XXX). By contrast, we find that when set our
+method to use classic BGS theory, the model fits are worse in full-coverage
+annotation models (Figure \ref{fig:figure-2}B). Additionally, when sparse
+annotation models contain genomic features that are likely under weak
+constraint (such as introns and UTRs), models fit worse under classic BGS
+theory than our B' method (Figure \ref{fig:figure-2}A). However, among the CADD
+annotation models, the goodness-of-fit is nearly identical between B' and
+classic BGS methods. This behavior is what we would expect given that the CADD
+annotation models only contain the most pathogenic sites, which are \emph{a
+prior} very likely under the strong selection domain under which B' and classic
+BGS theory agree.
+
+The out-sample $R^2$
+
+\begin{figure}[htbp] \centering
+    \includegraphics[width=\textwidth]{figures/figure_2.pdf} 
+    \caption{The $R^2$ estimates for a sparse (A) and full (B) models, for all 
+    populations (colors) fit at the megabase-scale. Round points are our B' 
+    method; diamonds are the classic B. For our method, we also calculate 
+    leave-one-out out-sample $R^2$ (lighter filled points).}
+  \label{fig:figure-2}
+\end{figure}
+
 \subsection*{Estimates of Negative Selection in Humans}
+
+\begin{figure}[htbp] \centering
+    \includegraphics[width=\textwidth]{figures/figure_3.pdf} 
+    \caption{The distribution of fitness effects of new mutations estimates for
+        Yoruba individuals. (A) The DFEs using sparse (left column) and
+        full-coverage (right column) tracks, across different annotation models
+        (row). Color indicates the feature type. (B) The DFE of the
+        full-coverage CDS-priority model comparing the estimates across
+    populations.}
+  \label{fig:figure-3}
+\end{figure}
 
@@
 \subsection*{DFE Estimates}
+
+Our best-fitting models have out-sample $R^2$ of around 68\% for Yoruba, 62\%
+European, and 60\% for Han Chinese ancestry individuals. However, $R^2$ has a
+theoretic upper bound due to coalescence noise around the predicted mean level
+$\widehat{\pi}(x)$.
+
 % - Table s5 https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/genetics/206/1/10.1534_genetics.116.197145/8/files1.pdf?Expires=1686357613&Signature=cp9qoI2ChCb6GFZ74HI4eWcA~MQ4sHlWyeI3Zp49-5xEORzMfYlSRxzVvZlo8ZoGWDi0G3PnL00ZplM-~or4YIVm2uhkmqjf61AmC7RdZ1QTak4j0kFMQ0wAbM9pusiV3gIepTQHq2HKRKwgS~ypmv~0oqO5dDxaq9nhOWIBBmLAA49oHWuxUtnRe3sEYATwCwjIFEomhgJpg7vKGCF~87GdCV4JETW4PN2VvD4hsu2AVRd8yA~vUJvsXfhWwcElDeY4fsgpprl1lC3XGjJCnqdwbzRGPV5LNx5XEKLptcZtLeLcOmFJCvmVLe0bewJ7Vta0TSQM7T9AuklompjPPw__&Key-Pair-Id=APKAIE5G5CRDK6RD3PGA
@@
 \section*{Methods}
+% (the method would support a new
+% ``interaction" type, but this greatly increases the number of parameters).
 

### 2023-06-09 manuscript/supplementary.tex
MS updates, figs

@@
 
+\section*{Table of $R^2$ Values}
+
+\begin{table}
+\centering
+\caption{$R^2$ and mutation rate estimates for all models}
+\begin{tabular}{lll|crr|cc}
+    \textbf{Model} & \textbf{Track type} & \textbf{Pop.} & \textbf{B' $R_\text{LOO}^2$} & \textbf{B' $R_\text{IS}^2$} & \textbf{B $R_\text{IS}^2$} & \textbf{B' $\hat{\mu} \times 10^{-8}$} & \textbf{B $\hat{\mu} \times 10^{-8}$} \\[0.5ex]
+\hline
+\hline
+    \textbf{phastcons$>$CDS$>$genes} &            sparse &          YRI &                        \textbf{68.45} &             68.16 &            65.23 &                                 1.8636 &                                0.2975 \\
+phastcons$>$CDS$>$genes &              full &          YRI &                        68.19 &             68.12 &            63.67 &                                 1.7082 &                                0.1666 \\
+               CADD 6\% &              full &          YRI &                        68.08 &             68.08 &            62.96 &                                 1.4612 &                                0.1721 \\
+               CADD 6\% &            sparse &          YRI &                        68.04 &             68.03 &            68.01 &                                 2.0912 &                                2.0716 \\
+               CADD 8\% &              full &          YRI &                        66.98 &             67.11 &            63.61 &                                 1.0126 &                                0.1713 \\
+               CADD 8\% &            sparse &          YRI &                        66.67 &             67.01 &            66.98 &                                 1.5688 &                                1.5473 \\
+CDS$>$genes$>$phastcons &              full &          YRI &                        64.90 &             65.51 &            61.23 &                                 4.1951 &                                0.1620 \\
+CDS$>$genes$>$phastcons &            sparse &          YRI &                        64.90 &             65.51 &            63.16 &                                 4.1795 &                                0.3070 \\
+\textbf{phastcons$>$CDS$>$genes} &            sparse &          CEU &                        \textbf{61.98} &             63.02 &            60.02 &                                 2.0194 &                                0.3036 \\
+phastcons$>$CDS$>$genes &              full &          CEU &                        61.78 &             63.01 &            58.57 &                                 2.3499 &                                0.1679 \\
+               CADD 6\% &              full &          CEU &                        61.74 &             62.94 &            57.42 &                                 1.5365 &                                0.1739 \\
+               CADD 6\% &            sparse &          CEU &                        61.66 &             62.86 &            62.84 &                                 2.1215 &                                2.1084 \\
+               CADD 8\% &              full &          CEU &                        60.80 &             62.12 &            57.95 &                                 1.1572 &                                0.1742 \\
+               CADD 8\% &            sparse &          CEU &                        60.59 &             62.06 &            62.04 &                                 1.6030 &                                1.5876 \\
+               CADD 6\% &              full &          CHB &                        59.62 &             61.31 &            55.31 &                                 1.5083 &                                0.1727 \\
+               CADD 6\% &            sparse &          CHB &                        59.54 &             61.21 &            61.20 &                                 2.1242 &                                2.1170 \\
+               \textbf{phastcons$>$CDS$>$genes} &            sparse &          CHB &                        \textbf{59.44} &             61.08 &            57.88 &                                 2.1982 &                                0.3060 \\
+phastcons$>$CDS$>$genes &              full &          CHB &                        59.29 &             61.08 &            56.71 &                                 2.5573 &                                0.1694 \\
+               CADD 8\% &              full &          CHB &                        58.88 &             60.63 &            55.86 &                                 1.1128 &                                0.1733 \\
+               CADD 8\% &            sparse &          CHB &                        58.65 &             60.57 &            60.55 &                                 1.6098 &                                1.6001 \\
+CDS$>$genes$>$phastcons &            sparse &          CEU &                        58.58 &             60.30 &            58.41 &                                 4.1041 &                                0.3110 \\
+CDS$>$genes$>$phastcons &              full &          CEU &                        58.58 &             60.30 &            56.75 &                                 4.1001 &                                0.1648 \\
+CDS$>$genes$>$phastcons &            sparse &          CHB &                        56.48 &             58.73 &            56.60 &                                 3.7195 &                                0.3136 \\
+CDS$>$genes$>$phastcons &              full &          CHB &                        56.48 &             58.73 &            54.95 &                                 3.7552 &                                0.1654 \\
+\hline
+\end{tabular}
+\end{table}
+
 \section*{Approximate $R^2$ Limit}
@@
 where $w$ is the width of the window such that the per-basepair $\pi$ is
-appropriately scaled. Then, we note that the variance in pairwise diversity is
-equivalent to the variance in the number of segregating sites for a sample of 
-two, $\var(\pi) = \var(S_2)$. From \textcite[eq. 7.20]{Wakeley2009-ua}
+appropriately scaled. We use the predicted $\hat{B}_i$ as the plugin estimator
+for $B_i$. Then, we note that the variance in pairwise diversity is equivalent
+to the variance in the number of segregating sites for a sample of two,
+$\var(\pi) = \var(S_2)$. From \textcite[eq. 7.20]{Wakeley2009-ua}
 
@@
 $\rho = w \gamma \pi_0 B_i$ where $\gamma$ is the ratio of recombination to
-mutation rates, $\nicefrac{r}{\mu}$. In humans, $\gamma \approx 1$, but we
-wanted to check the sensitivity of this calculation to this.
+mutation rates, $\nicefrac{r}{\mu}$. In humans, $\gamma \approx 1$, but explore
+different ratios to assess sensitivity of this calculation (Supplementary
+Figure \ref{suppfig:r2_upper}).
 
@@
 be around 74\% for Yoruba, and 80-81\% for European and Han Chinese models
-(Figure \ref{suppfig:r2_upper} We note, however, that this rough approximation
-assumes constant demography, and that the out-of-Africa bottleneck causes a
-reduction in effective population size that would shrink the variance of
-exponential coalescence times, but a bottleneck could in fact inflate these
-variances. A more accurate approximate of the upper $R^2$ bound could rely on
-simulations with more realistic demography.
+(Figure \ref{suppfig:r2_upper} This rough approximation assumes constant
+demography, which for bottlenecked out-of-Africa populations assumes that the
+variance in coalescence times is determined entirely by reducing the effective
+population size. However, the variance in coalescence times is \emph{higher} in
+bottlenecked populations (CEU and CHB) compared to Yoruba, which we confirm
+with simple simulations using the \texttt{OutOfAfrica\_3G09} model from
+\texttt{stdpopsim} \parencite{Gutenkunst2009-pg,Adrion2020-cf}. This higher
+than expected variance would inflate $SS_\text{res}$, which would decrease
+
+A more accurate approximate of the upper $R^2$ bound could rely on simulations
+with more realistic demography.
 

### 2023-06-13 manuscript/manuscript.tex
new ms stuff

@@
 
+\begin{figure}[htbp] \centering
+    \includegraphics[width=\textwidth]{figures/figure_2.pdf} 
+    \caption{The $R^2$ estimates for a sparse (A) and full (B) models, for all 
+    populations (colors) fit at the megabase-scale. Round points are our B' 
+    method and diamonds are the classic B. Lighter color round points are
+    the out-sample $R_\text{LOO}^2$ estimates for our B' method, and arrows show the decline
+    in goodness-of-fit due to in-sample overfitting (out-sample $R_\text{LOO}^2$ 
+    were not calculated for classic B values due to computational costs). The 
+    horizontal dashed lines are the $R_\text{drift}^2$ expected when the residual
+    variance is given by the theoretic variance in coalescence times due to drift alone.}
+  \label{fig:figure-2}
+\end{figure}
+
 Overall, we find the PhastCons Priority and CADD 6\% models fit about equally
@@
 megabase-scale, respectively. In populations impacted by the out-of-Africa
-bottleneck, the goodness-of-fit was lower (62.98\% and 59.98\% for CEU and CHB,
-respectively). The CADD 6\% model was previously found by
-\textcite{Murphy2022-sj} to be the best-fitting model using classic BGS. Our
-PhastCons Priority model fits marginally better than the CADD 6\% model
-consistently across all three populations (Supplementary Table XXX). This is
-likely because the PhastCons Priority model has more DFE parameters, and thus
-can fit finer-scale differences in the DFE across its four feature types. By
-contrast, the CADD 6\% track must fit patterns of diversity to conserved sites
-under a single DFE. The CADD 8\% model fit slightly worse than the CADD 6\%
-across all populations, and the Feature Priority model had the poorest-fit.
+bottleneck, the goodness-of-fit was lower across all models (e.g. 62.98\% and
+59.98\% for CEU and CHB respectively in the PhastCons Priority model). The CADD
+6\% model was previously found by \textcite{Murphy2022-sj} to be the
+best-fitting model using classic BGS. Our PhastCons Priority model fits
+marginally better than the CADD 6\% model consistently across all three
+populations (Supplementary Table XXX). This is likely because the PhastCons
+Priority model has more DFE parameters, and thus can fit finer-scale
+differences in the DFE across its four feature types. By contrast, the CADD 6\%
+track must fit patterns of diversity to conserved sites under a single DFE. The
+CADD 8\% model fit slightly worse than the CADD 6\% across all populations, and
+the Feature Priority model had the poorest-fit.
 
 Since our method is built upon theory that fixes the weak selection problem of
-classic BGS theory, it should in principle fit equally well when the annotation
-model covers the full genome, the bulk of which is expected to be neutrally
-evolution. To test this, for each annotation model (which is ``sparse") we fit
-a ``full" model that contains the remaining complement of the genome labeled as
-``other". Ideally, the goodness-of-fit of fit using our B' method should be
-invariant to whether an annotation model is sparse or full. Indeed, we find
-that
-
-NOTE: there looks like a bug in the table... repeated values.
-
-Each model was fit to the high-coverage 1000 human genome data for
-three populations, YRI, CEU, and CHB. Additionally, we also fit each of these
-models using our method, but using classic BGS theory rather than our B'
-theory. In total, we fit and compared 48 different models.
-
-Overall, we find that the goodness-of-fit, measured by both in- and out-sample
-$R^2$ are nearly identical across sparse and full fits using our B' method,
-regardless of the different annotation models (Figure \ref{fig:figure-2},
-Supplementary Materials Table XXX). By contrast, we find that when set our
-method to use classic BGS theory, the model fits are worse in full-coverage
-annotation models (Figure \ref{fig:figure-2}B). Additionally, when sparse
-annotation models contain genomic features that are likely under weak
-constraint (such as introns and UTRs), models fit worse under classic BGS
-theory than our B' method (Figure \ref{fig:figure-2}A). However, among the CADD
-annotation models, the goodness-of-fit is nearly identical between B' and
-classic BGS methods. This behavior is what we would expect given that the CADD
-annotation models only contain the most pathogenic sites, which are \emph{a
-prior} very likely under the strong selection domain under which B' and classic
-BGS theory agree.
-
-The out-sample $R^2$
-
-\begin{figure}[htbp] \centering
-    \includegraphics[width=\textwidth]{figures/figure_2.pdf} 
-    \caption{The $R^2$ estimates for a sparse (A) and full (B) models, for all 
-    populations (colors) fit at the megabase-scale. Round points are our B' 
-    method; diamonds are the classic B. For our method, we also calculate 
-    leave-one-out out-sample $R^2$ (lighter filled points).}
-  \label{fig:figure-2}
-\end{figure}
-
-\subsection*{Estimates of Negative Selection in Humans}
+classic BGS theory, it should in principle fit equally well when an annotation
+model includes regions that are likely neutrally evolution. To test this, for
+each annotation model (which are ``sparse") we fit a ``full" model that assigns
+the remaining complement of the genome as a feature called ``other". This other
+category is likely to predominantly neutrally evolving. Ideally, the
+goodness-of-fit of model using our B' method should be invariant to whether an
+annotation model is sparse or full. Indeed, we find this to be the case: both
+in-sample $R^2$ and out-sample $R_\text{LOO}^2$ values are nearly identical
+across full and sparse-track models (Figures \ref{fig:figure-2} A and B, round
+points).
+
+By contrast, full annotation models fit poorly when using classic BGS theory
+(Figure \ref{fig:figure-2}B, diamond-shaped points) and lead to unreasonable
+parameter estimates (discussed below). Additionally, when sparse annotation
+models contain genomic features that are likely under weak constraint (such as
+introns and UTRs), models fit worse under classic BGS theory than our B' method
+(Figure \ref{fig:figure-2}A). However, among the CADD annotation models, the
+goodness-of-fit is nearly identical between B' and classic BGS methods. This
+behavior is what we would expect given that the CADD annotation models contain
+only the most pathogenic sites, which are \emph{a prior} very likely under the
+strong selection domain under which B' and classic BGS theory agree.
+
+Along the chromosome, our approach models the observed megabase-scale diversity
+in window $i$ as the predicted level under our negative selection model,
+$B_i'$, plus some residual error $\varepsilon_i$, $\pi_i = B_i' +
+\varepsilon_i$. The out-sample $R_\text{LOO}^2$ estimates the ratio of variance
+explained by the negative selection model ($\var_i(\pi_i)$) to the total
+variance in diversity along the genome ($\var_i(\pi_i) = \var_i(B_i) +
+\var_i(\varepsilon_i) + 2\cov_i(B_i', \varepsilon_i)$, \cite{Murphy2022-sj}).
+Since our $R_\text{LOO}^2$ estimates suggest our models fit left-out
+chromosomes well even though our method assumes constant demography and
+homogeneous mutation rates along the genome, we wondered what $R^2$ we could
+expect under a simple theoretic model where $\cov_i(B_i', \varepsilon_i) = 0$
+and $\var_i(\varepsilon_i)$ was determined by the variance in coalescence times
+due to drift alone with an effective population size rescaled by $B_i'$. We
+calculate this ``drift $R^2$" by calculating the theoretic coalescence time
+variance in a megabase window with recombination and an effective population
+size of $B_i' N_e$ where we use our best-fitting model's as a plug-in estimator
+of $B_i'$. We note that this is a rough approximation, since this calculation
+assumes constant demography and that the variance in coalescence times under
+negative selection is likely lower than expected under drift alone. Still, we
+find that our out-sample $R_\text{LOO}^2$ for Yoruba individuals
+($R_\text{LOO}^2 \approx 68$\%) is close to the theoretic drift
+$R_\text{drift}^2 \approx 67$\%. For bottlenecked out-of-Africa populations,
+$R_\text{drift}^2 \approx 64$\% compared to the observed $R_\text{LOO}^2
+\approx 62$\% for CEU and $R_\text{LOO}^2 \approx 59$\% for CHB. We note that
+because bottlenecks increase the variance in coalescence times beyond the level
+implied by the effective population size, the disagreement between observed
+$R_\text{LOO}^2$ and the theoretic $R_\text{drift}^2$ is likely to shrink.
+Overall, this suggests that negative selection models explain the vast majority
+coalescence time variation at the megabase-scale that is capable of being
+explained (i.e. that is not coalescence noise).
+
+\subsection*{Estimates of Negative Selection Parameters in Humans}
 

### 2023-06-13 bgspy/bootstrap.py
fig cleanup

@@ -131,6 +131,7 @@ def load_from_bs_dir(bootstrap_dir):
     Because of the way parallelization was done, 
     this these assume there are B and B' results in each 
     .npz.
+    DEPRECATED
     """
     strap_files = os.listdir(bootstrap_dir)
     nlls_b, thetas_b = [], []
@@ -148,3 +149,16 @@ def load_from_bs_dir(bootstrap_dir):
     bp_boot_thetas = np.concatenate(thetas_bp, axis=1).T
 
     return BootstrapResults(b_boot_nlls, b_boot_thetas, bp_boot_nlls, bp_boot_thetas)
+
+
+def jackknife_stderr(thetas):
+    """
+    Jackknife standard errors, from http://harvard.edu/sites/reich.hms.harvard.edu/files/inline-files/lecture2.pdf
+    """
+    Tni = thetas
+    n = (~np.isnan(Tni)).sum(axis=0)
+    Tn = np.nanmean(Tni, axis=0)
+    Q = np.nansum((Tni - Tn[None, :])**2, axis=0)
+    sigma2_jack = (n-1)/n * Q
+    sigma_jack = np.sqrt(sigma2_jack)
+    return sigma_jack, n

### 2023-06-13 bgspy/likelihood.py

@@ -34,7 +34,7 @@ from bgspy.optim import nlopt_softmax_worker, nlopt_softmax_fixedmu_worker
 from bgspy.plots import model_diagnostic_plots, predict_chrom_plot
 from bgspy.plots import resid_fitted_plot, get_figax
 from bgspy.plots import chrom_resid_plot
-from bgspy.bootstrap import moving_block_bins
+from bgspy.bootstrap import moving_block_bins, jackknife_stderr
 
 
 # load the library (relative to this file in src/)
@@ -503,7 +503,7 @@ class BGSLikelihood:
             jackknife_results.append(obj)
         return jackknife_results, r2s
 
-    def load_jackknives(self, fit_dir, label=False):
+    def load_jackknives(self, fit_dir, label=False, trim=None):
         """
         Load the jackknife results by setting attributes.
         """
@@ -526,7 +526,7 @@ class BGSLikelihood:
         self.jack_thetas_ = thetas
         self.jack_indices = indices
         self.jack_r2s = r2s
-        self.jackknife_stderr()
+        self.jackknife_stderr(trim=trim)
 
     def load_loo(self, loo_dir):
         """
@@ -701,7 +701,7 @@ class BGSLikelihood:
             return R2(pred_pi, pi)
         return R2(pred_pi[_indices], pi[_indices])
 
-    def jackknife_stderr(self, use_loo_chrom=False):
+    def jackknife_stderr(self, use_loo_chrom=False, trim=None, iqr_factor=2):
         """
         Calculate the jackknife standard errors.
         """
@@ -716,6 +716,29 @@ class BGSLikelihood:
                    "load leave-one-out chromosome reults first")
             assert self.loo_thetas_ is not None, msg
             thetas = self.loo_thetas_
+
+        if trim:
+            if trim.upper() != "IQR":
+                if isinstance(trim, float):
+                    lower, upper = trim, 1-trim
+                else:
+                    lower, upper = trim
+                l, u = np.quantile(thetas, (lower, upper), axis=0)
+            else:
+                q1, q3 = np.percentile(thetas, (25, 75), axis=0)
+                iqr = q3 - q1
+                l, u = q1 - iqr_factor*iqr, q3 + iqr_factor*iqr
+            thetas = np.where((thetas < l) | (thetas > u), np.nan, thetas)
+
+        sigma_jack, n = jackknife_stderr(thetas)
+        self.sigma_trim_ = trim
[…]

### 2023-06-13 manuscript/manuscript.tex
ms r2 sec

@@
 
-Along the chromosome, our approach models the observed megabase-scale diversity
-in window $i$ as the predicted level under our negative selection model,
-$B_i'$, plus some residual error $\varepsilon_i$, $\pi_i = B_i' +
-\varepsilon_i$. The out-sample $R_\text{LOO}^2$ estimates the ratio of variance
-explained by the negative selection model ($\var_i(\pi_i)$) to the total
-variance in diversity along the genome ($\var_i(\pi_i) = \var_i(B_i) +
-\var_i(\varepsilon_i) + 2\cov_i(B_i', \varepsilon_i)$, \cite{Murphy2022-sj}).
 Since our $R_\text{LOO}^2$ estimates suggest our models fit left-out
@@
 assumes constant demography and that the variance in coalescence times under
-negative selection is likely lower than expected under drift alone. Still, we
-find that our out-sample $R_\text{LOO}^2$ for Yoruba individuals
+negative selection is likely lower than expected under drift alone.
+
+Still, we find that our out-sample $R_\text{LOO}^2$ for Yoruba individuals
 ($R_\text{LOO}^2 \approx 68$\%) is close to the theoretic drift

### 2023-06-13 manuscript/supplementary.tex
new ms stuff
ms r2 sec

@@
 \centering
-\caption{$R^2$ and mutation rate estimates for all models}
+\caption{$R^2$ and mutation rate estimates for all models. XXX note about repeats}
 \begin{tabular}{lll|crr|cc}
@@
 
-\section*{Approximate $R^2$ Limit}
+\section*{Derivation of Approximate Drift $R^2$}
 
-We derive a rough approximation of the theoretic upper bound of $R^2$ by
-noting that,
+Our approach models the observed megabase-scale diversity $y_i$ in window $i$
+as the predicted level under our negative selection model, $\pi_0 B_i$, plus
+some random residual error $\varepsilon_i$,
 
 \begin{align}
-    R^2 &= \frac{SS_\text{reg}}{SS_\text{reg} + SS_\text{res}} \\ 
-        &= \frac{\sum_i (\pi_i - \widehat{\pi}_i)^2}{\sum_i (\pi_i - \widehat{\pi}_i)^2 + SS_\text{res}}
+    y_i = \pi_0 B_i + \varepsilon_i
 \end{align}
 
-where $\pi_i$ is the observed diversity in window $i$, and $\widehat{\pi}_i$ is
-the predicted diversity under a particular model.
+where each $B_i = f(X_i | \Psi)$ for some data $X_i$ (i.e. the recombinaton map
+and annotation model) and parameters $\Psi$. To compare our models and assess
+model goodness-of-fit, we use the out-sample $R_\text{LOO}^2$ which is
+calculated by fitting the genome-wide model leaving one chromosome out, and
+calculating the residual variance between predictions and observed values for
+this out-sample chromosome.
 
-Then, $SS_\text{res}$ is the residual noise due to the randomness in mutation
-and average coalescence times in the window. If we make the assumptions that
-mutation rates are heterogeneous across the genome and the coalescence rate in
-a window is simply a background rate rescaled by the reduction factor $B_i$,
-then,
+In general, $R^2$ is calculated for a model fit across windows as,
 
 \begin{align}
-    SS_\text{res} &= \frac{1}{n} \sum_{i=1}^n \var(w\pi_i) \\ 
-                  &= \frac{1}{n} \sum_{i=1}^n \var(w\pi_0 B_i) 
+    R^2 = 1 - \frac{V_\text{res}}{V_\text{tot}}
 \end{align}
 
-where $w$ is the width of the window such that the per-basepair $\pi$ is
-appropriately scaled. We use the predicted $\hat{B}_i$ as the plugin estimator
-for $B_i$. Then, we note that the variance in pairwise diversity is equivalent
-to the variance in the number of segregating sites for a sample of two,
-$\var(\pi) = \var(S_2)$. From \textcite[eq. 7.20]{Wakeley2009-ua}
+where $V_\text{res}$ is the mean squared error across windows and
+$V_\text{tot}$ is the total variance in the predicator. Suppose we have
+estimates $\widehat{\pi}_0 \widehat{B}_i$ for $\pi_0 B_i$ from our model. Then,
+
+\begin{align}
+    V_\text{res} &= \frac{1}{n} \sum_{i=1}^n (y_i - \widehat{\pi}_0 \widehat{B}_i)^2  \\
+                 &= \underbrace{\frac{1}{n} \sum_{i=1}^n \varepsilon_i^2}_\text{irreducible error} + \underbrace{\left(\frac{1}{n} \sum_{i=1}^n (\pi_0 B_i - \widehat{\pi}_0 \widehat{B}_i)\right)^2}_\text{bias squared}.
+\end{align}
+
+The total variance in the predictor is,
+
+\begin{align}
+    V_\text{tot} = V_\text{res} + V_\text{model} + 2\widehat{\pi}_0\cov_i(\varepsilon_i, \widehat{B}_i)
+\end{align}
+where 
+\begin{align}
+    V_\text{model} = \frac{\widehat{\pi}_0^2}{n} \sum_{i=1}^n \left(\widehat{B}_i - \frac{1}{n}\sum_{i=1}^n\widehat{B}_i\right)^2.
+\end{align}
+
+Generally, fitting procedures act to minimize the squared deviation between
+predictors and observed values, which drives $\cov_i(\varepsilon_i,
+\widehat{B}_i)$ to zero XXX. \textcite{Murphy2022-sj} suggest  XXX
+
+Thus, our $R^2$ depends on (1) the bias of our predictors, (2) the covariance
+between residuals and predictors, and (3) the irreducible error due to variance
+in diversity within windows. However, it is of interest to approximate the
+irreducible error under theoretic models that assume all irreducible error is
+due entirely to the variance of coalescence times in a window, assuming that
+all of the effects of negative selection on the variance can be thought of as a
+local reduction in the effective population size to $B_i N_e$ and demographic
+effects amount to a simple rescaling of $N_e$. This also assumes mutation rates
+are constant across the genome, and do not contribute to the irreducible
+variance. We call this theoretic goodness-of-fit under drift
+$R_\text{drift}^2$, and it provides a rough approximation of the irreducible
+error in our model due to ``coalescence noise" around the modeled reductions
+due to negative selection.
+
+Then,
+
+\begin{align}
+    \E(\varepsilon_i^2) = \var(\pi_i)
+\end{align}
+
+since the mean residual is zero. The variance in pairwise diversity
+$\var(\pi_i)$ is equivalent to the variance in the number of segregating sites
+for a sample of two, $\var(\pi) = \var(S_2)$. The variance in coalescence times
+in a window of width $w$ basepairs with population-scaled recombination rate
+$\rho = 4 B_i N_e r w$ is,
 
@@
 \end{align}
-where,
+(\cite[eq. 7.20]{Wakeley2009-ua}) where,
 \begin{align}
@@
 $\rho = w \gamma \pi_0 B_i$ where $\gamma$ is the ratio of recombination to
-mutation rates, $\nicefrac{r}{\mu}$. In humans, $\gamma \approx 1$, but explore
-different ratios to assess sensitivity of this calculation (Supplementary
-Figure \ref{suppfig:r2_upper}).
+mutation rates, $\nicefrac{r}{\mu}$. In humans, $\gamma \approx 1$, but we
+explore different ratios to assess sensitivity of this calculation
+(Supplementary Figure \ref{suppfig:r2_upper}).
 

### 2023-06-13 manuscript/supplementary.tex
supp stuff
ms changes

@@
 This back-of-the-envelope calculation finds that the upper bound of $R^2$ would
-be around 74\% for Yoruba, and 80-81\% for European and Han Chinese models
-(Figure \ref{suppfig:r2_upper} This rough approximation assumes constant
-demography, which for bottlenecked out-of-Africa populations assumes that the
-variance in coalescence times is determined entirely by reducing the effective
-population size. However, the variance in coalescence times is \emph{higher} in
+be around 67\% for Yoruba, and 64\% for European and Han Chinese models (Figure
+\ref{suppfig:drift_r2} This rough approximation assumes constant demography,
+which for bottlenecked out-of-Africa populations assumes that the variance in
+coalescence times is determined entirely by reducing the effective population
+size. However, the variance in coalescence times is \emph{higher} in
 bottlenecked populations (CEU and CHB) compared to Yoruba, which we confirm
@@
 than expected variance would inflate $SS_\text{res}$, which would decrease
-
-A more accurate approximate of the upper $R^2$ bound could rely on simulations
-with more realistic demography.
+$R_\text{drift}^2$, bringing it closer to the observed $R_\text{LOO}^2$. A more
+accurate approximation of the $R_\text{drift}^2$ could be found with
+simulations with more realistic demography and varying coalescence rates along
+the genome.
 
 \begin{figure}[htbp] \centering
-    \includegraphics[width=0.5\textwidth]{figures/supplementary/suppfigure_R2_upper.pdf}
-    \caption{A rough approximation to the theoretic $R^2$ upper bound assuming constant demography.}
-  \label{suppfig:r2_upper}
+    \includegraphics[width=\textwidth]{figures/supplementary/suppfigure_drift_R2.pdf}
+    \caption{(A) the theoretic $R_\text{drift}^2$ using different plugin
+        estimators for $B_i$ under different models. This shows that among the
+        best-fitting models, the predicted $R_\text{drift}^2$ varies little;
+        most variation is among populations due to differing $\pi_0$. (B) The 
+        $R_\text{drift}^2$ across different populations, for varying ratio of recombination
+        to mutation rates (x-axis).}
+  \label{suppfig:drift_r2}
 \end{figure}

### 2023-06-17 bgspy/models.py
sub nb stuff

@@ -299,7 +299,6 @@ class BGSModel(object):
     def ratchet_df(self, fit, mu=None, ncores=None):
         """
         Output a combined ratchet, for all segments.
-        
         mu is optional mutation rate to predict under (otherwise 
         MLE will be used).
         """

### 2023-06-17 manuscript/manuscript.tex
ms changes

@@
 1000 Genome data for three populations: Yoruba (YRI), Han Chinese (CHB), and
-European (CEU) ancestries.
+European (CEU) ancestries. We assessed and compared models according to how
+well they predict the genome-wide patterns of diversity they were fit on. We
+use $R_\text{LOO}^2$, which is the proportion of the observed variance in
+megabase-scale diversity that is predicted by our model in out-samples (i.e.
+whole chromosomes left out during the fitting process).
 
@@
 expect under a simple theoretic model where $\cov_i(B_i', \varepsilon_i) = 0$
-and $\var_i(\varepsilon_i)$ was determined by the variance in coalescence times
-due to drift alone with an effective population size rescaled by $B_i'$. We
-calculate this ``drift $R^2$" by calculating the theoretic coalescence time
-variance in a megabase window with recombination and an effective population
-size of $B_i' N_e$ where we use our best-fitting model's as a plug-in estimator
-of $B_i'$. We note that this is a rough approximation, since this calculation
-assumes constant demography and that the variance in coalescence times under
-negative selection is likely lower than expected under drift alone.
+and $\var_i(\varepsilon_i)$ was determined by the variance in neutral
+coalescence times alone with an effective population size rescaled by $B_i'$.
+We calculate this ``coalescence $R^2$" by calculating the theoretic coalescence
+time variance in a megabase window with recombination and an effective
+population size of $B_i' N_e$ where we use our best-fitting model's as a
+plug-in estimator of $B_i'$. We note that this is a rough approximation, since
+this calculation assumes constant demography and that the variance in
+coalescence times under negative selection is likely lower than expected under
+drift alone.
 
 Still, we find that our out-sample $R_\text{LOO}^2$ for Yoruba individuals
-($R_\text{LOO}^2 \approx 68$\%) is close to the theoretic drift
-$R_\text{drift}^2 \approx 67$\%. For bottlenecked out-of-Africa populations,
-$R_\text{drift}^2 \approx 64$\% compared to the observed $R_\text{LOO}^2
-\approx 62$\% for CEU and $R_\text{LOO}^2 \approx 59$\% for CHB. We note that
-because bottlenecks increase the variance in coalescence times beyond the level
-implied by the effective population size, the disagreement between observed
-$R_\text{LOO}^2$ and the theoretic $R_\text{drift}^2$ is likely to shrink.
-Overall, this suggests that negative selection models explain the vast majority
-coalescence time variation at the megabase-scale that is capable of being
-explained (i.e. that is not coalescence noise).
+($R_\text{LOO}^2 \approx 68$\%) is close to the theoretic $R_\text{coal}^2
+\approx 67$\%. For bottlenecked out-of-Africa populations, $R_\text{coal}^2
+\approx 64$\% compared to the observed $R_\text{LOO}^2 \approx 62$\% for CEU
+and $R_\text{LOO}^2 \approx 59$\% for CHB. We note that because bottlenecks
+increase the variance in coalescence times beyond the level implied by the
+effective population size, the disagreement between observed $R_\text{LOO}^2$
+and the theoretic $R_\text{coal}^2$ is likely to shrink. Overall, this suggests
+that negative selection models explain the vast majority coalescence time
+variation at the megabase-scale that is capable of being explained (i.e. that
+is not coalescence noise).
 
@@
 
+\begin{figure}[htbp] \centering
+    \includegraphics{figures/figure_4_pop_mu_trimmed.pdf} 
+    \caption{}
+  \label{fig:figure-4}
+\end{figure}
+
 Given that previous methods used classic BGS theory to fit genome-wide patterns
@@
 
-\subsection*{DFE Estimates}
-
-Our best-fitting models have out-sample $R^2$ of around 68\% for Yoruba, 62\%
-European, and 60\% for Han Chinese ancestry individuals. However, $R^2$ has a
-theoretic upper bound due to coalescence noise around the predicted mean level
-$\widehat{\pi}(x)$.
+\subsection*{Predicted Diversity and $B'(x)$ maps}
+
+When we fit our negative selection model to genome-wide patterns of diversity,
+the parameter estimates generate three predicted quantities: (1) the $B(x)$
+map, (2) the substitution rates per segment, and (3) the fitness variance per
+segment. Here we focus on the predicted $B(x)$ map and implied levels of
+diversity at the megabase scale. We discuss substitution rates and fitness
+variance in the next section.
+
+We also obtain a prediction $\widehat{B}_i$ for the reduction factor $B_i$ in
+each window $i$, which can be multiplied by our estimate of $\widehat{\pi}_0$
+to obtain a prediction of diversity $\widehat{\pi}_i = \widehat{\pi}_0
+\widehat{B}_i$.
+
+\subsection*{Model Predictions and the Substitution Process}
+
+Our method accommodates weak selection by modeling the fitness variance in each
+segment under an equilibrium negative selection process. This fitness variance
+is found by simultaneously solving for the per-generation substitution rate $R$
+and the draft-effective population size $N_d$ for each segment. When our model
+is fit to genome-wide diversity data, the parameter estimates imply predictions
+of the deleterious substitution rate per basepair, per generation, for each
+feature. Our method only considers within-species pairwise diversity, so these
+predicted substitution rates constitute a test that our model is making
+reasonable out-sample predictions of the substitution rates across features.
+
+We focus specifically on the predictions from our feature-based annotation
+models, since we can readily estimate human substitution distances for each
+feature to compare to our predictions.
+
+The estimated substitution distance should in theory equal the product of the
+substitution rate per site, per year and the branch length in years.
+
+When comparing our predicted mutation rates to observed substitution distances,
+we must account for three major sources of uncertainty.
+
+First, the split time between humans and chimpanzees is unknown, with estimates
+ranging from 6Mya to 12Mya (XXX). Second, our predictions are in sites per
+generation, and the appropriate human generation time is unknown. Third,
+
+For a particular feature $f$, the estimated per-site substitution distance
+$S_f$ along a branch (which accounts for back mutations) should correspond to
+$S_f = T_g r_f$, where $T_g$ is the time in generations and $r_f$ is the
+substitution rate per basepair, per generation of that feature.
+
+Each feature is
+composed of basepairs with varying selection coefficients distributed according
+to the feature-specific DFE vector $\mathbf{w}_f$ and corresponding vector of
+substitution rates per selection coefficient, $\mathbf{r}_f$.
+
+Our estimate of a feature's DFE
+is $\widehat{\mathbf{w}}_f$.
+
+We calculate the
+
+\begin{align}
+    \label{eq:sub}
+    r_f = \mu \int_0^1 p_f(s) ds
+\end{align}
+
+Solving the XXX equations, we end up with predictions $\widehat{r}_f$ for the
+per-basepair, per-generation substitution rate for each feature. These are
+calibrated to the MLE mutation rate estimate $\widehat{\mu}_\text{del}$, which
+as mentioned, excludes the contribution of lethal mutations which act purely to
+reduce $N_e$ and thus only impact $\pi_0$. Using Equation \eqref{eq:sub} we can
+re-calibrate these estimates against other mutation rates.
+
+Because of it's greater interpretabilty, we initially focus on the substitution
+rate estimates from the Feature Priority model even though it had a poorer-fit
+and high mutation rate estimate.
+
+We can also re-calibrate this model to an alternate mutation rate.
 

### 2023-06-17 manuscript/supplementary.tex
ms changes

@@
 
+\section*{Ratchet Rate Predictions}
+
+\section*{Load Predictions}
+
+Each fixation of a mutant with effect $-s$ decreases the population fitness by
+$(1-2s)$ since we assume additive effects throughout. For each feature's DFE
+vector $\mathbf{w}$ and grid of selection coefficients $\mathbf{s}$, we have a
+corresponding vector $\widehat{\mathbf{r}}$ of predicted ratchet rates. Each
+generation $\widehat{\mathbf{r}}$ substitutions are fixed each generation.
+Since we assume multiplicative fitness, the log load for selection coefficient
+$i$ in the vector is,
+
+\begin{align}
+    l_i = \log(1-2 s_i) \widehat{r_i}.
+\end{align}
+
+The total log load is,
+
+\begin{align}
+    \log(L) = \sum_{i} l_i.
+\end{align}
+
 \section*{Likelihood}

### 2023-06-21 manuscript/manuscript.tex
ms changes

@@
 deleterious mutations from the population, which is essential to maintaining
-the integrity of information in the genome to maintain a lineage. Mutations
-under negative selection are kept at low frequencies (Haldane?), contributing
-to human traits (XXX) and diseases. As with most selection in the genome, this
-process perturbs the allele frequencies of neighboring linked variation. This
-reduces genetic variability around conserved segments and creates large-scale
-spatial patterns in genetic diversity. These patterns are mediated by the
-deleterious mutation rate, the strength of selection in conserved segments, and
-the spatial distribution of recombination rates and conserved segments. Since
-genome-wide recombination maps and putatively conserved segments are available
-for many species, researchers have combined these with population genetic
-theory to estimate the mutation rate and selection coefficients of deleterious
-mutations, as well as the overall reduction in genetic variability due to the
-negative selection process.
+the integrity of information in the genome needed to maintain a lineage.
+Mutations under negative selection are kept at low frequencies (Haldane?),
+contributing to human traits (XXX) and diseases. As with most selection in the
+genome, this process perturbs the allele frequencies of neighboring linked
+variation. This reduces genetic variability around conserved segments and
+creates large-scale spatial ``linked selection signal" in genetic diversity.
+Mutation and negative selection's contribution to this signal is mediated by
+the deleterious mutation rate, the strength of selection in conserved segments,
+and the spatial distribution of recombination rates and conserved segments.
+Since genome-wide recombination maps and putatively conserved segments are
+available for many species, researchers have combined these with population
+genetic theory to estimate the mutation rate and selection coefficients of
+deleterious mutations, as well as the overall reduction in genetic variability
+due to the negative selection process.
 
@@
 
-\subsection*{Predicted Diversity and $B'(x)$ maps}
+\subsection*{Predicted Diversity Substitution Rates}
+
+\begin{figure}[htbp] \centering
+    \includegraphics[width=\textwidth]{figures/figure_4.pdf} 
+    \caption{}
+  \label{fig:figure-4}
+\end{figure}
 
@@
 ranging from 6Mya to 12Mya (XXX). Second, our predictions are in sites per
-generation, and the appropriate human generation time is unknown. Third,
+generation, and the appropriate human generation time is unknown. Third, the
+appropriate mutation rate to use is debated, since there may have been a
+mutational slow down on the human lineage (XXX), which could explain why
+phylogenetic- and pedigree-based mutation rate estimates differ by roughly a
+factor of two (XXX). Finally, we are only considering modeling the deleterious
+substitution rate. In reality, some fraction of observed substitutions fixed
+because they were beneficial.
 

### 2023-06-22 manuscript/manuscript.tex
ms updates

@@
 variation. This reduces genetic variability around conserved segments and
-creates large-scale spatial ``linked selection signal" in genetic diversity.
-Mutation and negative selection's contribution to this signal is mediated by
-the deleterious mutation rate, the strength of selection in conserved segments,
-and the spatial distribution of recombination rates and conserved segments.
-Since genome-wide recombination maps and putatively conserved segments are
-available for many species, researchers have combined these with population
-genetic theory to estimate the mutation rate and selection coefficients of
-deleterious mutations, as well as the overall reduction in genetic variability
-due to the negative selection process.
-
-Classic Background Selection (BGS) theory is the predominant model for the
-reduction in linked neutral diversity used by statistical approaches to
-quantify the impacts of negative selection. However, this model makes some
-assumptions for mathematical tractability that could distort inferences about
-the negative selection process. First, since fixation probabilities ultimately
-depend on the product of the deleterious selection coefficient (s) and
-population size (N), the efficacy of negative selection should depend on
-demography. Unfortunately, accommodating demography into theoretic negative
-selection models remains an open, difficult problem (XXX). Second, building off
-classic models of mutation-selection balance (Crow 1970, Kimura and Maruyama
-196), the Background Selection (XXX) model assumes that new mutations are
-sufficiently deleterious that they are invariably driven to loss. Under this
-assumption, the effect of selection is well-approximated by simply rescaling
-the neutral coalescent by a reduction factor known as B (XXX). However, this is
-only appropriate in a limited domain of parameters relevant to the negative
-selection process (XXX). Because the BGS model cannot accommodate the
-possibility of fixation for mutations with $Ns \ll 1$, diversity is incorrectly
-predicted to fall to zero as the strength of selection diminishes. This weak
-selection problem is related to the wider problem of modeling the substitution
-rate of deleterious mutation (i.e. the “ratchet” rate). Third, the BGS model
-assumes segments under purifying selection are sufficiently far apart that
-there is no selective interference between them. This is a matter of degree,
-since under sufficiently high mutation rates or with little recombination,
-selective interference can occur with a segment.
+creates a large-scale spatial ``linked selection signal" in diversity alone the
+genome. Mutation and negative selection's contribution to this signal is
+mediated by the deleterious mutation rate, the strength of selection in
+conserved segments, and the spatial distribution of recombination rates and
+conserved segments. Since genome-wide recombination maps and putatively
+conserved segments are available for many species, researchers have combined
+these with population genetic theory to estimate the mutation rate and
+selection coefficients of deleterious mutations, as well as the overall
+reduction in genetic variability due to the negative selection process.
+
+Classic Background Selection (BGS) theory is the predominant theoretic model to
+predict the reduction in linked neutral diversity, and is used by statistical
+approaches that quantify the aspects of the negative selection. However, the
+BGS model makes some assumptions for mathematical tractability that could
+distort inferences about the negative selection process. First, since fixation
+probabilities ultimately depend on the product of the deleterious selection
+coefficient (s) and population size (N), the efficacy of negative selection
+should depend on demography. Unfortunately, accommodating demography into
+theoretic negative selection models remains an open, difficult problem (XXX).
+Second, building off classic models of mutation-selection balance (Crow 1970,
+Kimura and Maruyama 196), the Background Selection (XXX) model assumes that new
+mutations are sufficiently deleterious that they are invariably driven to loss.
+Under this assumption, the effect of selection is well-approximated by simply
+rescaling the neutral coalescent by a reduction factor known as B (XXX).
+However, this is only appropriate in a limited domain of parameters relevant to
+the negative selection process (XXX). Because the BGS model cannot accommodate
+the possibility of fixation for mutations with $Ns \ll 1$, diversity is
+incorrectly predicted to fall to zero as the strength of selection diminishes.
+This weak selection problem is related to the wider problem of modeling the
+substitution rate of deleterious mutation (i.e. the “ratchet” rate). Third, the
+BGS model assumes segments under purifying selection are sufficiently far apart
+that there is no selective interference between them. This is a matter of
+degree, since under sufficiently high mutation rates or with little
+recombination, selective interference can occur with a segment.
 
@@
 BGS model, simulations reveal another mysterious defect: a poor fit around
-$Ns=1$. We show that this can be partially corrected by incorporating the effects
-of selective interference on genic variation, but the remaining discrepancy is
-due to the build up of negative linkage disequilibrium that cannot be
-adequately captured by this model. We find accounting for genic selective
-interference between segments fits human population genomic data slightly
-better, suggesting levels of negative selection in humans are sufficient to
-perturb the selection at other sites.
+$Ns=1$. We show that this can be partially corrected by incorporating the
+effects of selective interference on genic variation, but the remaining
+discrepancy is due to the build up of negative linkage disequilibrium that
+cannot be adequately captured by this model. We find accounting for genic
+selective interference between segments fits human population genomic data
+slightly better, suggesting levels of negative selection in humans are
+sufficient to perturb the selection at other sites.
 
@@
 a factor $Q_t^2$ due to the build up of autocorrelation to generation $t$. We
-let $Q^2 = Q_t^2$ as $t \to \infty$. Intuitively, the product $V_A Q^2$
+let $Q^2 := Q_t^2$ as $t \to \infty$. Intuitively, the product $V_A Q^2$
 represents the expected total variance in reproductive success a neutral
-mutation experiences over its lifetime in a system with draft. Incorporating
-the impact of time-invariant autocorrelation created by draft as a rescaled
-rate of drift is analogous to using the Green--Kubo relation to find the
-transport coefficient in molecular dynamics with velocity autocorrelation
+mutation experiences over its lifetime in a system with weak selection draft.
+Incorporating the impact of time-invariant autocorrelation created by draft as
+a rescaled rate of drift is analogous to using the Green--Kubo relation to find
+the transport coefficient in molecular dynamics with velocity autocorrelation
 \parencite{Green1954-kl,Kubo1957-va}.
@@
 linkage can be accommodated by different expressions for the inflation factor
-$Q^2$. When fitness variation has a multiplicative polygenic basis, as it does
-for genome-wide negative selection processes, the draft asymptotic effective
-population size experienced by an arbitrary neutral site under the influence of
-all linked regions is,
+$Q^2$ \parencite{Santiago1995-hx,Santiago1998-bs}. When fitness variation has a
+multiplicative polygenic basis, as it does for genome-wide negative selection
+processes, the draft asymptotic effective population size experienced by an
+arbitrary neutral site under the influence of all linked regions is,
 
@@
 variation, the autocorrelation function for a neutral allele associated with
-segment $i$ is the time-invariant function $C(t) = [(1-r_i)(1-\kappa_i)]^t$,
-where $r_i$ is the recombination fraction to the segment and $\kappa_i$ is the
-rate that the associated fitness variance decays due to selective dynamics.
-Then, the cumulative autocorrelation is,
+segment $i$ is $C(t) = [(1-r_i)(1-\kappa_i)]^t$, where $r_i$ is the
+recombination fraction to the segment and $\kappa_i$ is the rate that the
+associated fitness variance decays due to selective dynamics. Then, the
+cumulative autocorrelation is,
 
@@
 sum of additive \emph{genic} fitness variance $V_a$ (i.e. due to allele
-frequencies) and the total linkage disequilibria $\delta_{LD}$ between selected
-sites, $V_A = V_a + \delta_{LD}$. Considering only the genic variation, at
-equilibrium $\Delta V_{a,i} = 0$; thus, the loss in genic fitness each
-generation due to selective dynamics $-\kappa V_{A,i}$ must be equal to the
-increase in fitness variation to new mutations $V_m$ each generation and
-changes in the LD term \parencite{Bulmer1971-ae}. Note that the loss in genic
-fitness due to drift is not time-invariant due to the build up of
+frequencies) and the fitness covariance due to linkage disequilibria between
+selected sites ($\delta_{LD}$), $V_A = V_a + \delta_{LD}$. Considering only the
+genic variation, at equilibrium $\Delta V_{a,i} = 0$; thus, the loss in genic
+fitness each generation due to selective dynamics $-\kappa V_{A,i}$ must be
+equal to the increase in fitness variation due to new mutations each generation
+($V_m$) and changes in the LD term \parencite{Bulmer1971-ae}. Note that the
+loss in genic fitness due to drift is not time-invariant due to the build up of
 autocorrelation, and considered elsewhere in the derivation (see Supplementary
-Materials Equation \ref{eq:vardecay}). Then, $\kappa_i =
+Materials Equation \ref{eq:vardecay}). Then, at equilibrium $\kappa_i =
 \nicefrac{V_{m,i}}{V_{a,i}}$ (see Supplementary Materials Equation \ref{eq:Z}).
@@
 When selection is weak, however, the additive fitness variance under the
-classic BGS model is no longer accurate. \parencite{Santiago2016-mu} suggested
+classic BGS model is no longer accurate. \textcite{Santiago2016-mu} suggested
 that this is due to the loss of fitness variation that occurs when a
@@
 in Equation \eqref{eq:Va} also results from diffusion models with a flux of
-mutations into discrete sites (\cite{Kimura1969-jw}; see Supplementary
-Materials Section \ref{supp:kimura}). Our results differ from
-\parencite{Santiago2016-mu}, since we find from simulations their expression
-models the additive \emph{genic} fitness variation, not additive genetic
-fitness variation (see Section XXX). This is because Equation \eqref{eq:Va}
-does not capture the expected build-up of negative linkage disequilibria when
-selection is only mildly deleterious.
+mutations into discrete sites (\cite{Kimura1969-jw}).
+
+%Our results differ from \parencite{Santiago2016-mu}, since we find from
+%simulations their expression models the additive \emph{genic} fitness
+%variation, not additive genetic fitness variation (see Section XXX). This is
+%because Equation \eqref{eq:Va} does not capture the expected build-up of
+%negative linkage disequilibria when selection is mildly deleterious.
 
@@
 \parencite{Kimura1962-su,Malecot1948-zv} with the draft-rescaled effective
-population size, e.g. $R = N U p_f(N_d, s)$. Given this equation for the
+population size, i.e. $R = N U p_f(N_d, s)$. Given this equation for the
 ratchet and Equation \eqref{eq:polygenic_Ne} for $N_d$ under draft, we have a
@@
 to classic BGS theory. Furthermore, to investigate whether this accuracy was
-caused by the model capturing the equilibrium fitness variance and ratchet
-rate, we also measured these throughout the simulation. Again, we find the
-diploid SC16 theory accurately predicts both the ratchet rate (Figure
+caused by the model accurately predicting the equilibrium fitness variance and
+ratchet rate, we also measured these throughout the simulation. Again, we find
+the diploid SC16 theory accurately predicts both the ratchet rate (Figure
 1\ref{fig:figure-1}B) and the genic fitness variance (Figure
@@
 
-Additionally, these figures provide further intuition for the underlying
+Additionally, these figures provide further intuition about the underlying
 negative selection process. When selection is strong the ratchet rate is zero,
 as there is no probability that strongly deleterious can fix. However, around
-$2N_e s \approx 1$, the ratchet begins clicking as $p_F > 0$. Ultimately, the
-ratchet rate reaches the neutral mutation rate, as $V_a = 0$ and $N_d = N$
-(e.g. no draft). Likewise, the additive genic fitness variance matches the
-level predicted by BGS ($Us$), but begins to diverge when the ratchet begins
-clicking. This occurs when the level of variation diverges from the
-deterministic mutation-selection equilibrium $V_a = Us$ due to the stochastic
-clicking of the ratchet at rate $R$.
+$2N_e s \approx 1$, the ratchet begins clicking as $p_F > 0$. When this occurs,
+fitness variation is reduced from the strong-selection equilibrium by the
+amount of population fitness lost per generation due to the ratchet.
+Ultimately, the ratchet rate reaches the neutral mutation rate, as $V_a = 0$
+and $N_d = N$ (e.g. no draft, pure drift). Likewise, the additive genic fitness
+variance matches the level predicted by BGS ($Us$), but begins to diverge when
+the ratchet begins clicking. This occurs when the level of variation diverges
+from the deterministic mutation-selection equilibrium $V_a = Us$ due to the
+stochastic clicking of the ratchet at rate $R$.
 
@@
 
-Given the accuracy of the SC16 model in predicting the reduction in effective
-population size and ratchet rate when equilibrium negative selection processes
-were acting in single segment, we extended their model so that it could be
-applied to whole-genome data. Our method, implemented in the software
-\texttt{bgspy}, calculates the equilibrium additive genic fitness variance
-$V_a$ and ratchet rate $R$ for each specified segment in the genome for a range
-of mutation rates and selection coefficients and a fixed genome-wide
-drift-effective population size $N$ (see Methods XXX). Genomic regions
-putatively under negative selection (e.g. coding sequences or UTRs) and the
-recombination rate map are specified as arguments so that the predicted values
-accurately reflect local genomic details. Note that each equation is solved
-independently, since here we assume no selective interference between segments.
-We refer to the prediction reductions along the genome from our approach as B'
-maps (analogous to B maps from \cite{McVicker2009-ax}).
-
-To assess our B' maps, we simulated negative selection for fixed mutation rates
-and selection coefficient on human chromosome 10 using realistic recombination
-and conserved feature maps. Averaging over $n=100$ simulation replicates, we
-estimated the simulation expected reduction map and compared this to the
-predicted B and B' maps for the same parameters. We find that our B' maps and
-the classic BGS theory B maps closely match simulations when selection is
-strong (top row of Figure \ref{fig:figure-1}E). This provides the first
-realistic forward-time simulation confirmation of classic background selection
-theory. However, we find slight discrepancies in regions with low recombination
-(Figure \ref{fig:figure-1}E). Second, we find our theory is vastly more
-accurate than the classic B maps when selection is very weak ($2N_e s \ll 1$;
-bottom row of Figure \ref{fig:figure-1}E), which is expected since BGS theory
-does not work in this domain. Across all mutation and selection parameters
-simulated, the relative error of the classic B maps is 14.6\% whereas the
-relative error in the new B' maps is 5\%. Additionally, we find that the mean
-squared error between simulations and B' maps is close to the theoretic lower
-bound set by the coalescence variance \parencite[see Methods
+Given the accuracy of the SC16 model in predicting the reduction factor $B$ and
+ratchet rate for a single segment under general equilibrium mutation negative
+selection processes, we extended their model so that it could be applied to
+whole-genome data. Our software method \texttt{bgspy} calculates the
+equilibrium additive genic fitness variance ($V_a$) and ratchet rate ($R$)
+across grids of mutation rates and selection coefficients for each
+pre-specified segment in the genome that may be under negative selection (e.g.
+coding sequences or UTRs). Then, our method computes a reduction map $B(x)$ for
+each position $x$ in the genome by combining the pre-computed $V_a$ for every
+segment and a user-supplied recombination map (see Methods XXX). To distinguish
+between McVicker's B maps based on classic BGS theory, we call our reduction
+map $B(x)$ a B' map.
+
+To assess our predicted B' reduction maps, we simulated negative selection for
+fixed mutation rates and selection coefficient on human chromosome 10 using
+realistic recombination and conserved feature maps. Averaging over $n=100$
+replicates, we estimated the simulation expected reduction map and compared
+this to the predicted B and B' maps for the same parameters. We find that our
+B' maps and the classic BGS theory B maps closely match simulations when
+selection is strong (top row of Figure \ref{fig:figure-1}E). This provides the
+first realistic forward-time chromosome-scale simulation confirmation of
+classic BGS theory. However, we find slight discrepancies in regions with low
+recombination (Figure \ref{fig:figure-1}E). Second, we find our theory is
+vastly more accurate than the classic B maps when selection is very weak ($2N_e
+s \ll 1$; bottom row of Figure \ref{fig:figure-1}E), which is expected since
+BGS theory does not work in this domain. Across all mutation and selection
+parameters simulated, the relative error of the classic B maps is 14.6\%
+whereas the relative error in the new B' maps is 5\%. Additionally, we find
+that the mean squared error between simulations and B' maps is close to the
+theoretic lower bound set by the coalescence variance \parencite[see Methods
 XXX]{Tajima1983-gu}.
@@
 \ref{fig:figure-1}D). The higher error in this domain occurs across all
-mutation rates. First, we hypothesized that the reason this occurred was the
-when we numerically solve Equations \eqref{eq:main_eqns} for each segment, this
-does does not consider the reduction due to selection at other linked segments.
-In particular, we use $N=1000$ corresponding to the number of diploids in the
-simulations rather than $B'(x) N$ at position $x$. To test this, we implemented
-a locally-rescaled version of the B' maps, which numerically solves Equations
-\eqref{eq:main_eqns} for each segment at position $x$ taking into account the
-reduction experienced due to selection at other segments by setting $B'(x)N$ as
-the population size (see Method XXX). We find the locally-rescaled B' maps
-reduce the relative error from 5\% to 0.4\% and mean squared error (Figure
-\ref{fig:figure-1}D, dashed colored lines), but does not entirely eliminate the
-error in the $2Ns \approx 1$ domain.
+mutation rates. We hypothesized that the higher error in this domain may be
+because when our method numerically solves Equations \eqref{eq:main_eqns} for
+each segment, it does not consider the experienced reduction due to selection
+at other linked segments. In particular, we use a fixed drift-effective
+$N=1000$ corresponding to the number of diploids in the simulations rather than
+$B(x) N$ at position $x$. To test this, we implemented a locally-rescaled
+version of the B' maps, which numerically solves Equations \eqref{eq:main_eqns}
+for each segment at position $x$ taking into account the reduction experienced
+due to selection at other segments by setting $B'(x)N$ as the population size
+(see Method XXX). We find the locally-rescaled B' maps reduce the relative
+error from 5\% to 0.4\% and mean squared error (Figure \ref{fig:figure-1}D,
+dashed colored lines), but does not entirely eliminate the error in the $2Ns
+\approx 1$ domain.
 
@@
 
-\subsection*{Annotation Model Comparison}
+Overall, while we find evidence that local-recalling can reduce error rate in
+the $2Ns \approx 1$ domain,
+
+it currently cannot be used during when fitting
+genome-wide models.
+
+\subsection*{Application to Humans and Model Comparison}
 
@@
 
+\subsection*{Segment Simulations}
+
+To assess the accuracy of the SC16 theory, we conducted forward simulations
+across grids of mutation rates ($\log_{10}(\mu) = \{10^{-9.5}, 10^{-9}, \ldots,
+10^{-8}\}$) and selection coefficients ($\log_{10}(\mu) = \{-1, 1.5, 2, \ldots,
+4.5\}$) using SLiM \parencite{Haller2023-uk,Haller2019-vu}. Each simulation
+evolved a 100kb region with deleterious mutations entering at rate $\mu$
+per-basepair, per generation, each with an additive selection coefficient of
+$-s$. Total fitness was multiplicative across sites. Each simulation recorded
+the number of fixations and the fitness variance per generation. After $10N$
+generations, the ARG of the tree was recorded and downstream Python scripts
+calculated diversity in the center of the region. We averaged the results of
+$10,000$ replicate simulations to compare against theory.
+
 \subsection*{Chromosome-wide Simulations}

### 2023-06-22 manuscript/supplementary.tex
ms updates

--- 2023-06-13 manuscript/supplementary.tex
@@
 \maketitle
+
+\tableofcontents
 
@@
 
+\section*{Predicted Rate of Fitness Change}
+
+Each fixation of a mutant with fitness effect $-s$ reduces the population
+fitness to $1-2s$ since all individuals are homozygotes for the deleterious
+mutation and we assume additive effects at a site. For each segment $l$, the
+maximum likelihood parameter estimates of our model imply a prediction of the
+deleterious substitution rate per segment $\widehat{R}_{l,i}$ for each
+selection coefficient $s_i$ in the $m_s$-element DFE grid (according to which
+feature class it belongs to).
+
+Since we assume multiplicative fitness effects, the log population fitness
+(considering only fixed sites) for segment $l$ in the next generation is,
+
+\begin{align}
+    \log(L_{l}) = \sum_{i=1}^{m_s} \log(1-2 s_i) \widehat{R}_{l,i}.
+\end{align}
+%
+Since we assume multiplicative fitness, the total genome-wide log population
+fitness in the next generation is
+
+\begin{align}
+    \log(L) = \sum_{l} \log{L_l}.
+\end{align}
+%
+This amounts to a predicted population fitness loss per generation of
+
+\begin{align}
+    \widehat{\Delta}_W &= 1 - L \\  \nonumber
+             &= 1 - \exp\left(\sum_{l} \sum_{i=1}^{m_s} \log(1-2s_i) \widehat{R}_{l,i} \right).
+\end{align}
+%
+We calculate this value for all segments within a feature class, and then sum
+over all features to arrive at our genome-wide estimate. We use a parametric
+bootstrap to carry forward the estimation uncertainty to the predicted fitness
+loss per generation. Briefly, we do this by sampling new estimates of the DFE
+per selection coefficient and feature from a normal distribution centered on
+the ML point estimate and with the jackknife standard deviation.
+
 \section*{Likelihood}
@@
 functions, so we opted for the simpler identity link function.
+
+\section{Simulations}
+
+\subsection{Segment Simulations}
+
+\section{Additional Theory}
+
+\subsection{supp:weak-strong}
 

### 2023-06-23 manuscript/manuscript.tex
ms updates

@@
 worthwhile to develop some intuition for why this is the case, how the classic
-background selection works, and how the model of \textcite{Santiago2016-mu}
-solves these problems and can be extended to genome-wide inference.
+background selection works, and how the quantitative genetic model of
+\textcite{Santiago2016-mu} solves these problems and can be extended to
+genome-wide inference.
 
@@
 heterozygotes and $2s$ in homozygotes. The classic background selection model
-considers when mutations are sufficiently deleterious that they are destined to
-extinction
+only consider mutations that are sufficiently deleterious that they are
+destined to extinction
 \parencite{Charlesworth1993-gb,Nordborg1996-nq,Hudson1995-pt,Hudson1994-oh}. In
@@
 deterministic mutation-selection model \parencite{Haldane1927-ga}. Since
-lineages carrying any strongly deleterious mutations cannot contribute to the
-genealogy, coalescence events only occur in the fraction
+lineages carrying strongly deleterious mutations cannot contribute to the
+genealogy in the long run, coalescence events only occur in the fraction
 $\exp(-\nicefrac{U}{s})$ of the population free of mutations. Second, the time
@@
 \parencite{Rouzine2008-qz,Good2013-lp,Gessler1995-hz} towards reduced
-population fitness (since beneficial mutations are ignored). Consequently,
-individuals in the least-loaded class are stochastically lost, fixing those
-mutations with a click of ``Muller's ratchet" \parencite{Muller1964-ki}.
-Determining both the rate of Muller's ratchet
+population fitness (since beneficial and back mutations are ignored).
+Consequently, individuals in the least-loaded class are stochastically lost,
+fixing those mutations with a click of ``Muller's ratchet"
+\parencite{Muller1964-ki}. Determining both the rate of Muller's ratchet
 \parencite{Haigh1978-gt,Gordo2002-dr,Gessler1995-hz} and the shape of
@@
 approximations ill-suited \parencite{Gillespie2000-mh,Der2011-it,Neher2013-dz}.
-However, when selection is weak, draft can be approximated as an increase in
+However, when selection is weak, draft is well-approximated as an increase in
 the rate of drift (or equivalently, a reduction in effective population size).
@@
 A key insight from Robertson (\citeyear{Robertson1961-ho}) is that in the
-long-run draft acts like an increase in the variance in offspring number.
+long-run weak draft acts like an increase in the variance in offspring number.
 \textcite{Wright1938-tv} showed that additional non-heritable variance in
@@
 \textcite{Santiago2016-mu} is that the ratchet rate under draft is determined
-by the probability of fixation $p_f(N_d, s)$
+by the probability of fixation $p_F(N_d, s)$
 \parencite{Kimura1962-su,Malecot1948-zv} with the draft-rescaled effective
-population size, i.e. $R = N U p_f(N_d, s)$. Given this equation for the
+population size, i.e. $R = N U p_F(N_d, s)$. Given this equation for the
 ratchet and Equation \eqref{eq:polygenic_Ne} for $N_d$ under draft, we have a
@@
 
-where $V_a = (U-R)s$ as in Equation \eqref{eq:Va}.
+where $V_a = (U-R)s$ as in Equation \eqref{eq:Va}. The predicted $V_a$, $R$,
+and $N_d$ values represent the equilibria expected under a
+mutation-selection-drift-draft process for neutral to strongly deleterious
+additive mutations.
+
+With the equilibrium $V_a$ calculated at each segment for a particular $\mu$
+and $s$, we can estimate the reduction factor $B(x) = \nicefrac{N_d}{N}$ at any
+genomic position $x$ (see Methods XXX). XXX
+
+Since it is too computationally-intensive to numerically solve this system of
+equations for all segments to estimate the reduction at every position
+
+for every focal position $x$
 
@@
 \parencite{Haller2019-vu,Haller2023-uk} we simulated a region of $10^5$
-basepairs under varying levels of mutation and selection. We find a close
-correspondence between the observed and predicted reduction in effective
-population size $B=\nicefrac{N_d}{N}$ over all selection and mutation
-parameters including weak selection (Figure \ref{fig:figure-1}A), in contrast
-to classic BGS theory. Furthermore, to investigate whether this accuracy was
-caused by the model accurately predicting the equilibrium fitness variance and
-ratchet rate, we also measured these throughout the simulation. Again, we find
-the diploid SC16 theory accurately predicts both the ratchet rate (Figure
-1\ref{fig:figure-1}B) and the genic fitness variance (Figure
-1\ref{fig:figure-1}C).
-
-Additionally, these figures provide further intuition about the underlying
-negative selection process. When selection is strong the ratchet rate is zero,
-as there is no probability that strongly deleterious can fix. However, around
-$2N_e s \approx 1$, the ratchet begins clicking as $p_F > 0$. When this occurs,
-fitness variation is reduced from the strong-selection equilibrium by the
-amount of population fitness lost per generation due to the ratchet.
-Ultimately, the ratchet rate reaches the neutral mutation rate, as $V_a = 0$
-and $N_d = N$ (e.g. no draft, pure drift). Likewise, the additive genic fitness
-variance matches the level predicted by BGS ($Us$), but begins to diverge when
-the ratchet begins clicking. This occurs when the level of variation diverges
-from the deterministic mutation-selection equilibrium $V_a = Us$ due to the
-stochastic clicking of the ratchet at rate $R$.
+basepairs under varying levels of mutation and selection (see Methods
+\ref{sec:methods-segsim}). We find a close correspondence between the observed
+and predicted reductions in effective population size $B=\nicefrac{N_d}{N}$
+over all selection and mutation parameters including weak selection (Figure
+\ref{fig:figure-1}A), in contrast to classic BGS theory. Furthermore, to
+investigate whether this accuracy was caused by the model correctly predicting
+the equilibrium fitness variance and ratchet rate, we also measured these
+throughout the simulation. Again, we find the diploid SC16 theory accurately
+predicts both the ratchet rate (Figure 1\ref{fig:figure-1}B) and the genic
+fitness variance (Figure 1\ref{fig:figure-1}C).
+
+Moreover, these simulations provide intuition about the underlying negative
+selection process. When mutations are strongly deleterious, there is no chance
+they can fix, and the ratchet rate is zero (Figure \ref{fig:figure-1}B for $2Ns
+> 1$). In this strong selection regime, the additive genic fitness variation
+closely matches the theoretic deterministic equilibrium of $V_a = Us$ (dashed
+gray line, Figure \ref{fig:figure-1}C) However, around $2N_e s \approx 1$, the
+ratchet begins clicking as $p_F > 0$. When this occurs, each click of the
+ratchet eliminates variation, and the equilibrium variation diverges from the
+deterministic mutation-selection equilibrium (Figure \ref{fig:figure-1}C).
 
@@
 Given the accuracy of the SC16 model in predicting the reduction factor $B$ and
-ratchet rate for a single segment under general equilibrium mutation negative
-selection processes, we extended their model so that it could be applied to
-whole-genome data. Our software method \texttt{bgspy} calculates the
-equilibrium additive genic fitness variance ($V_a$) and ratchet rate ($R$)
-across grids of mutation rates and selection coefficients for each
-pre-specified segment in the genome that may be under negative selection (e.g.
-coding sequences or UTRs). Then, our method computes a reduction map $B(x)$ for
-each position $x$ in the genome by combining the pre-computed $V_a$ for every
-segment and a user-supplied recombination map (see Methods XXX). To distinguish
+ratchet rate for a single segment under general mutation-selection processes,
+we extended their model so that it could be applied to whole-genome data. Our
+software method \texttt{bgspy} calculates the equilibrium additive genic
+fitness variance ($V_a$) and ratchet rate ($R$) across grids of mutation rates
+and selection coefficients for each pre-specified segment in the genome that
+may be under negative selection (e.g. coding sequences or UTRs). Then, our
+method computes a reduction map $B(x)$ for each position $x$ in the genome,
+which is determined by the fitness variation at each segment and their spacing
+along the user-supplied recombination map (see Methods XXX). To distinguish
 between McVicker's B maps based on classic BGS theory, we call our reduction
@@
 fixed mutation rates and selection coefficient on human chromosome 10 using
-realistic recombination and conserved feature maps. Averaging over $n=100$
-replicates, we estimated the simulation expected reduction map and compared
-this to the predicted B and B' maps for the same parameters. We find that our
-B' maps and the classic BGS theory B maps closely match simulations when
-selection is strong (top row of Figure \ref{fig:figure-1}E). This provides the
-first realistic forward-time chromosome-scale simulation confirmation of
+realistic recombination maps and conserved feature positions. Averaging over
+$100$ replicates, we estimated the simulation expected reduction map and
+compared this to the predicted B and B' maps for the same parameters. We find
+that our B' maps and the classic BGS theory B maps closely match simulations
+when selection is strong (top row of Figure \ref{fig:figure-1}E). This provides
+the first realistic forward-time chromosome-scale simulation confirmation of
 classic BGS theory. However, we find slight discrepancies in regions with low
@@
 because when our method numerically solves Equations \eqref{eq:main_eqns} for
-each segment, it does not consider the experienced reduction due to selection
+each segment, it does not consider the reduction experienced due to selection
 at other linked segments. In particular, we use a fixed drift-effective
@@
 
-Overall, while we find evidence that local-recalling can reduce error rate in
-the $2Ns \approx 1$ domain,
+While we find evidence that local-recalling reduces error in the $2Ns \approx
+1$ domain, we do not employ it in our genome-wide inference method for a few
+reasons. First, the reduction map
+
+numerically solving the system of equations
 
@@
     \label{eq:sub}
-    r_f = \mu \int_0^1 p_f(s) ds
+    r_f = \mu \int_0^1 p_F(s) ds
 \end{align}
@@
 
+\subsection*{Solving the B' Equations for each Segment}
+\label{sec:methods-bprime-eqns}
+
+In our implementation of the theory of \textcite{Santiago2016-mu}, we calculate
+the draft-effective population size $N_d = B N$ within a segment by solving a
+system of two non-linear equations,
+
+\begin{align}
+  \label{eq:}
+  {N}_{d} &= N \exp \left( -V_a \frac{Q^2}{2} \right) & \text{\emph{effective population equation}} \\
+  R &= \frac{4N_eU s}{\exp(4 N_e s)-1}  & \text{\emph{substitution equation}} 
+\end{align}
+
+where,
+
+\begin{align}
+  %V(x) &= U s - 2 {R}(x) s & \text{\emph{fitness variance equation}} \\
+  V_a &= (U  - 2 R) s & \text{\emph{fitness variance equation}} \\
+  Q^2 &= \frac{2}{(1-Z)(2-(2-M)Z)} & \text{\emph{linkage inflation factor}} \\
+  Z &= 1 - \frac{Us}{U-{R}} & \text{\emph{variance decay rate}}
+\end{align}
+
+and $U = \mu L$ and $M = r_\text{BP} L$ are the total mutation and
+recombination rates for a segment $L$ basepairs long.
+
+% Each mutation has a
+% selection coefficient of $s$, for which we assume additive effects. Then, the
+% local reduction factor is $b(x) = \nicefrac{N_e(x)}{N}$, and the per-basepair
+% per-generation substitution rate is $\lambda_D = \nicefrac{R(x)}{L}$. Note that
+% in the variance equation, the long-run variance is the balance between mutation
+% and twice the loss due to substitution, $(U-2R)s$, since for selective balance
+% $V(x) = 0$ for $s>0$, the per-basepair substitution rate must be half the mutation rate,
+% since each substitution removes twice the variance introduced by a mutation in
+% a diploid species.
+
+Our software \texttt{bgspy} solves these two equations numerically for each
+pre-specified segment across grids of mutation rates and selection
+coefficients. The recombination rate in a segment is determined by a
+user-supplied recombination map; segments that bridge a change in recombination
+rate are split, and solved independently with the appropriate rate. This
+pre-calculation step is done in parallel, and determines the equilibrium
+additive genic fitness variation for each segment under a
+mutation-selection-drift-draft process across the grids of parameters.
+
+\subsection*{Calculating the Reduction Maps}
+\label{sec:methods-maps}
+
+The pre-computed equilibria are used to calculate the $B(x)$ maps across
+parameter grids, which use in our comparisons to chromosome-scale forward
+simulations and our composite-likelihood method. The reduction at a focal site
+$x$ can be calculated as the product of all segments' contributing reduction
+factors accounting for the recombination distances between the segment and
+focal site. Additionally, to accommodate the estimation of separate DFEs for
+different types of features, each $B(x)$ maps is
+
+\begin{align}
+    \log(B(x)) = - \sum_{i=1} \sum_{j=1} 
+\end{align}
+
+where $b(x, v)$ is a modified version of $b(x)$ that accounts for the
+recombination between the segment $x$ and the neutral site $v$ (see XXX).
+
+In practice, we calculate B' using the variables determined by solving the B'
+equations for each mutation and selection coefficient in the grids. Since only
+the analog of $Q^2(x)$, $Q^2(x, v)$ changes, the pre-computed substitution rate
+(and thus local variance) can be re-used for each $v$. We calculate $B(v)$
+across a grid of focal points along the chromosome, spaced 10kb apart.
+
 \subsection*{Segment Simulations}
+\label{sec:methods-segsim}
 
@@
 the contribution of the focal segment
-
-\subsection*{Solving the B' Equations for each Segment}
-
-In our implementation of the theory of \textcite{Santiago2016-mu}, we calculate
-the local effective population size $N_e(x) = b(x) N$ within a segment $x$ by
-solving a system of two non-linear equations,
-
-\begin{align}
-  \label{eq:}
-  {N}_{e}(x) &= N \exp \left( -\frac{V(x)}{2} Q^2(x) \right) & \text{\emph{effective population equation}} \\
-  {R}(x) &= \frac{4N_e(x)U s}{\exp(4 N_e(x) s)-1}  & \text{\emph{substitution equation}} 
-\end{align}
-
-where,
-
-\begin{align}
-  %V(x) &= U s - 2 {R}(x) s & \text{\emph{fitness variance equation}} \\
-  V(x) &= (U  - 2 {R}(x)) s & \text{\emph{fitness variance equation}} \\
-  Q^2(x) &= \frac{2}{(1-Z)(2-(2-M)Z)} & \text{\emph{linkage inflation factor}} \\
-  Z &= 1 - \frac{Us}{U-{R(x)}} & \text{\emph{}} \text{\emph{variance decay rate}}
-\end{align}
-
-and $U = \mu L$ and $M = r_\text{BP} L$ are the total mutation and
-recombination rates for a segment $L$ basepairs long. Each mutation has a
-selection coefficient of $s$, for which we assume additive effects. Then, the
-local reduction factor is $b(x) = \nicefrac{N_e(x)}{N}$, and the per-basepair
-per-generation substitution rate is $\lambda_D = \nicefrac{R(x)}{L}$. Note that
-in the variance equation, the long-run variance is the balance between mutation
-and twice the loss due to substitution, $(U-2R)s$, since for selective balance
-$V(x) = 0$ for $s>0$, the per-basepair substitution rate must be half the mutation rate,
-since each substitution removes twice the variance introduced by a mutation in
-a diploid species.
-
-In our software \texttt{bgspy}, these two equations are solved for a grid of
-mutation rates and selection coefficients for each specified segment from a
-feature type (e.g. CDS, UTRs, etc.). This results in estimates of $N_e(x)$ and
-$R(x)$ per segment. In the first-round B' calculation, we use a single
-pre-specified drift-effective population size, $N$. The B' maps created here at
-this step reflect selection only in the segment, ignoring the impact of
-selection at other segments. By contrast, the locally-rescaled B' maps solve
-these equations, assuming the drift-effective population size is $B(x)N$. These
-reflect the segment's dynamics if they were effected by selection at other
-segments under selection in the genome. Consequently, we can compare and
-contrast models with local-recalling to test whether we achieve a better fit if
-we account selective interference from other segments.
-
-\subsection*{The B' Maps}
-
-The reduction at a focal site $x$ can be calculated as the product of all
-segments' contributing reduction factors,
-
-\begin{align}
-  B(v) = \prod_x b(x,v)
-\end{align}
-
-where $b(x, v)$ is a modified version of $b(x)$ that accounts for the
-recombination between the segment $x$ and the neutral site $v$ (see XXX).
-
-In practice, we calculate B' using the variables determined by solving the B'
-equations for each mutation and selection coefficient in the grids. Since only
-the analog of $Q^2(x)$, $Q^2(x, v)$ changes, the pre-computed substitution rate
-(and thus local variance) can be re-used for each $v$. We calculate $B(v)$
-across a grid of focal points along the chromosome, spaced 10kb apart.
 
@@
 TODO we need to give some intuition for what Vm and V are in humans, e.g. back of envelope type stuff
-
-\section*{Appendix}
-
-\subsection*{Overview of Quantitative Genetic Models of Effective Population Size}
-
-Here, we step through a quick derivation of \textcite{Santiago1995-hx}.
-Throughout, we assume random mating, hermaphroditic individuals, and a constant
-population size. The change in a neutral allele's frequency in one generation
-can be partitioned into the three sources of stochasticity: the random
-associations with fitness backgrounds (i.e. \emph{draft}), the non-heritable
-randomness in family size, and the Mendelian noise from heterozygotes
-segregating. If we let $x_{0,i} \in \{0, \nicefrac{1}{2}, 1\}$ be the frequency
-of neutral alleles individual $i$ in generation 0 carries, we can partition the
-random neutral allele frequency of the population ($x_t$ without the individual
-index) into the underlying stochastic causes,
-
-\begin{align}
-  x_1 = \frac{1}{2N} \sum_{i=1}^N \left( x_{0,i}k_{0,i} + \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \right)
-\end{align}
-
-where $k_{0,i}$ is the number of surviving gametes parent $i$ passes on, and
-$\delta_{0,i,j}$ is a random term that encapsulates the noise due to Mendelian
-segregation of heterozygotes. If parent $i$ is a homozygote ($x_{0,i} \in \{0,
-1\}$), then $\delta_{0,i,j} = 0$, whereas if $x_{0,i} = \nicefrac{1}{2}$ then
-$\delta_{0,i,j} = \pm \nicefrac{1}{2}$ with equal probability. This is because
-a heterozygous parent will transmit half a neutral allele in expectation, but
-each round of Mendelian segregation must pass on either 0 or 1 alleles, which
-the random $\pm \nicefrac{1}{2}$ term imposes. The factor of $\nicefrac{1}{2}$
-is due to the fact that we're summing over $N$ diploids, but considering the
-number of gametes they transmit. Since each diploid parent must have two
-offspring to maintain a constant population size, $\nicefrac{1}{N} \sum_i
-k_{0,i} = 2$.
-
-The frequency in the initial generation is $x_0 = \nicefrac{1}{N} \sum_{i=1}^N
-x_{0,i}$, though to indicate that we treat this as fixed, we use $p_0 = x_0$.
-Then, the allele frequency change is,
-
-\begin{align}
-  \Delta x_1 = x_1 - p_0 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (k_{0,i} - 1) + \frac{1}{2N} \sum_{i=1}^N \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \\
-  \Delta x_1 &= K_1 + H_1
-\end{align}
-
-where $K_1$ and $H_1$ are the random change in neutral allele frequency change
-due to offspring number (including heritable and non-heritable components), and
-Mendelian segregation.
-
-Now, let us look the variance of $\var(\Delta x_1)$ over evolutionary
-replicates. Since the Mendelian random process and the offspring random process
-are independent,
-
-\begin{align}
-  \var(\Delta x_1) &= \var(K_1) + \var(H_1).
-\end{align}
-
-Looking at each term,
-
-\begin{align}
-  \var(K_1) &\approx \frac{1}{4N^2} \sum_{i=1}^N \var\left(x_{0,i} (k_{0,i} - 1) \right)
-\end{align}
-
-where we ignore the covariance terms due to the sum, since these are on order
-$\nicefrac{1}{N^3}$. In the first generation from an arbitrary starting point,
-the neutral alleles assort independently into diploids with respect to their
-fitness, so
-
-\begin{align}
-  \var(K_1) &\approx \frac{1}{4N^2} \sum_{i=1}^N \var(x_{0,i}) \var(k_{0,i} - 1)  \\
-            &\approx \frac{1}{4N} \var(x_{0,i}) \var(k_{0,i}).
-\end{align}
-
-Assuming that no correlation between parental gametes (e.g. no inbreeding),
-$\var(x_{0,i})$ is the binomial variance in individual allele frequency, or
-$p_0(1-p_0)/2$, and $\var(k) := \var(k_{0,i})$ is the offspring variance of an
-individual given by the reproduction process. For example, if the reproduction
-process is a neutral multinomial Wright--Fisher, $\var(k) \approx 2$.
-
-\begin{align}
-  \var(K_1) &\approx \frac{p_0(1-p_0)}{2N} \frac{\var(k)}{4}.
-\end{align}
-
-The Mendelian noise variance term can be derived similarly. First, the sum over
-each parent's transmitted gametes can be simplified by noting that parents are
-exchangeable over evolutionary replicates with respect to their contribution to
-this term. In a constant population size, the double summation over parents and
-their offspring can be replaced by a summation over offspring, since both sum
-$N$ exchangeable terms. Then, note that $\var(\delta_{i,j}) = \nicefrac{1}{4}
-p_0(1-p_0)$, so
-
-\begin{align}
-  \var(H_1) &= \frac{p_0(1-p_0)}{2N} \frac{1}{2}.
-\end{align}
-
-Finally, we have
-
-\begin{align}
-  \var(\Delta x_1) &= \var(K_1) + \var(H_1) \\
-  &\approx \frac{p_0(1-p_0)}{2N} \frac{\var(k)}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2} \\
-                   &\approx \frac{p_0(1-p_0)}{2N}\left(\frac{\var(k)}{4} + \frac{1}{2}\right)
-\end{align}
-
-(c.f. \cite{Santiago1995-hx} equation 2 and \cite{Buffalo2019-qs} equation 30).
-Note that if the reproduction process is a neutral multinomial Wright--Fisher,
-$\var(k) \approx 2$, and this simplifies to the expected Wright--Fisher
-variance. If we were to \emph{define} a variance effective population size by
-setting this variance in neutral allele frequency change to the expected
-variance under a Wright--Fisher model ($V_\text{WF}$), we'd have
-
-\begin{align}
-  V_\text{WF} &:= \frac{p_0(1-p_0)}{2N_e} \\
-  V_\text{WF} &= \var(\Delta p_1) \\
-  N_e &= \frac{4N}{\var(k) + 2} \\
-\end{align}
-
-(c.f. \cite{Wright1938-tv}).
-
-Now, we look at this variance equation behaves with $t > 1$.
-
-\begin{align}
-  \V(x_1) = \E_1\left[(x_1 - x_0)^2\right]= \frac{x_0(1-x_0)}{2N},
-\end{align}
-
-and
-
-\begin{align}
-  \label{eq:var_wf}
-  \V(x_2) &= \E\left[(x_2 - x_0)^2\right] \\
-          &= \E_1\left[\E_2\left[((x_2-x_1) + (x_1 - x_0))^2 | x_1\right]\right] \\
-          &= \E_1\left[\E_2\left[(\Delta x_2 + \Delta x_1)^2 | x_1\right]\right] \\
-          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + 2\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
-          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] + \mathcal{C}_{1,2} \\
-          &= \frac{p_0(1-p_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{p_0(1-p_0)}{2N} + \mathcal{C}_{1,2}
-\end{align}
-
-Consequently, the variance of the neutral allele frequency changes each
-generation according to the probability of failing to coalesce each generation
-and the pairwise covariance terms $\mathcal{C}_{i,j}$ that build up due to associations
-with heritable fitness backgrounds. Note that the covariance term $\mathcal{C}_{1,2} =
-\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] = 0$ under neutral
-evolution, and this simplifies to the well-known equation for variance in a
-Wright--Fisher population.
-
-Now, we turn our attention to the case where there is heritable fitness
-variation, which leads to non-zero covariance terms $\mathcal{C}_{i,j}$. These
-terms emerge when $\var(k)$ has a heritable component of fitness that can be
-transmitted along with the neutral allele, thereby affecting the neutral
-allele's trajectory in later generations. If we partition the offspring number
-into heritable and non-heritable components, $k_i = 2f_i + \varepsilon_i$, then
-$\var(k_i) = 4 V_h + V_n$. When there is heritable fitness variation across
-individuals, $V_h > 0$. Because the population size is assumed to be constant,
-the mean heritable fitness across individuals is constrained to be $\E_i(f_i) =
-1$; this implies that the individual fitness values $f_i$ are relative
-fitnesses as is standard in population genetics, in this case to the population
-mean. Thus $V_h$ is the squared coefficient of heritable fitness variation
-(this is often denoted as $C^2 = \nicefrac{V_A}{\bar{w}^2}$, see
-\cite{Crow1958-pc,Charlesworth1987-ab,Houle1992-ur}).
-
-We can further decompose the $K_1$ term into $K_1 = S_1 + D_1$,
-
-\begin{align}
-  \var(\Delta x_1) &= \var(S_1) + \var(D_1) + \var(H_1) \\
-  &\approx \frac{p_0(1-p_0)}{2N} V_h + \frac{p_0(1-p_0)}{2N} \frac{V_n}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2}. \\
-\end{align}
-
-(c.f. \cite{Santiago1995-hx} equation 11).
-
-To see how the covariance terms accumulate, consider $\var(x_3 - p_0)$. Note
-that $\E(S_t) = \E(D_t) = \E(H_3) = 0$, since which neutral allele we track is
-arbitrary, so by symmetry the expected change is zero. Then,
-
-\begin{align}
-  \label{eq:var_x}
-  \var(x_3 - p_0) &= \E\left[ \left(S_1 + D_1 + H_1 + S_2 + D_2 + H_2 + S_3 + D_3 + H_3 \right)^2 \right] \\
-                  &= \E({S_1}^2) + \E({S_2}^2) + \E({S_3}^2)  \\
-                  &+ \E(S_1 S_2) + \E(S_1 S_3) + \E(S_2 S_3) \\
-                  &+ \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) + \E({D_3}^2) + \E({H_3}^2)
-\end{align}
-
-The covariance terms $\E(S_i S_j)$ for $j > i$ represent the expected neutral
-allele frequency change from a neutral allele becoming associated with a
-fitness background in generation $i$ and that fitness association persisting
-until generation $j$. However, the covariances terms $\E(S_j S_i)$ for $j > i$
-are zero since associations in the future cannot affect the past (see p. 1041
-of \cite{Buffalo2019-qs}). Let us consider $S_1$ and $S_2$. We have,
-
-\begin{align}
-  S_1 = \frac{1}{2N} \sum_{i=1}^N x_{0,i}(f_{0,i} - 1) = \cov(x_{0,i}, f_{0,i}) \\
-  S_2 = \frac{1}{2N} \sum_{i=1}^N x_{1,i}(f_{1,i} - 1) = \cov(x_{1,i}, f_{0,i})
-\end{align}
-
-which are chance covariances (across the population, \emph{not} evolutionary
-replicates) created by the random sorting of neutral alleles and fitness into
-individuals, \emph{each generation}. These covariances are equivalent to the
-Robertson-Price equation \parencite{Robertson1966-fs,Price1970-si}, which
-predict the change in neutral frequency due to heritable fitness (and likewise
-with the change due to non-heritable factors).
-
-Considering the underlying haplotypes that lead to $x_{0,i}$ and $f_{0,i}$ can
-give us an equation for the dynamics of the $\E(S_i S_j)$ (for $j > i$) terms
-over evolutionary replicates. Let us partition the allele frequency and fitness
-by into the average of the paternal contributions. The neutral allele frequency
-per gamete is either 0 or 1, and we assume the fitnesses across gametes are
-additive. Then, as long as there is random mating, we can simplify the
-associations in a diploid by looking at a single gamete,
-
-\begin{align}
-  S_1 = \cov(x_{0,i}, f_{0,i}) &= \cov\left(\frac{x_{0,i}' + x_{0,i}''}{2}, f_{0,i}' + f_{0,i}''\right)  \\
-                               &= \frac{1}{2}\left(\cov(x_{0,i}', f_{0,i}' + f_{0,i}'') + \cov(x_{0,i}'', f_{0,i}' + f_{0,i}'')\right)  \\
-                               &= \cov(x_{0,i}', f_{0,i}' + f_{0,i}'')  \\
-                               &= \cov(x_{0,i}', f_{0,i}') + \cov(x_{0,i}', f_{0,i}'')  \\
-                               &= S_1' + S_1''
-\end{align}
-
-which follows from symmetry, since it does not matter which gamete carrying the
-neutral allele we track (c.f. \cite{Santiago1998-bs} p. 2107). The primes now
-indicate whether the selected locus is on the same gamete as the neutral
-allele (single prime), or the homologous gamete (double prime).
-
-After meiosis, a fraction $1-r$ of these associations between the tracked
-neutral neutral allele $x_{0,1}'$ and fitness background $f_{0,i}'$ persist,
-and a fraction $r$ disassociate its currently coupled background and create an
-association with the homologous fitness background, $f_{0,i}''$. When linkage
-is tight, this latter term can be ignored as we will do here, though it does
-impact background levels of neutral diversity \parencite{Santiago1995-hx}.
-Additionally, the fitness effects of the selected background may change in
-later generations, in complex ways \parencite{Barton1986-yh,Turelli1990-kd}.
-\textcite{Santiago1995-hx} assume an equilibrium level of fitness variation
-$V_h$, with the fitness variance associated with a particular background
-decaying at a simple geometric decay at rate $1-\kappa$ each generation, which
-is sufficient for background selection. An approximate rate of the decay can be
-worked out for the particular selective system. In general, this can be a much
-more complicated function; \textcite{Buffalo2019-qs} show that even fluctuating
-selection can be accommodated.
-
-Now, the associations created in generation 1 and that persist in future
-generations as (ignoring $S_i''$ terms) follow the pattern
-
-\begin{align}
-  S_{2}' &= S_1 (1-\kappa)(1-r) \\
-  S_{3}' &= S_1 (1-\kappa)^2(1-r)^2 \\
-  S_{4}' &= S_1 (1-\kappa)^3(1-r)^3.
-\end{align}
-
-Then, the cumulative effect of the fitness associations created in generation
-$1$ can be written as,
-
-\begin{align}
-  \label{eq:qterm}
-  S_1 Q_t &= S_1 + S_2' + S_3' + \ldots + S_t' \\
-  Q_t &= 1 + \sum_{i=1}^t (1-r)^i(1-\kappa)^i
-\end{align}
-
-or for continuous $t$,
-
-\begin{align}
-  \label{eq:q_t}
-  Q_t &= 1+\frac{(1-\kappa) (1-r) \left(1-(1-\kappa)^t (1-r)^t\right)}{\kappa+(1-\kappa)r}
-\end{align}
-which as $t \to \infty$, converges to,
-
-\begin{align}
-  \label{eq:Qinf}
-  Q_\infty &= \frac{1}{\kappa + r(1-\kappa)}.
-\end{align}
-
-This shows the long reach of heritable fitness factors in altering allele
-frequency through the generations. In terms like $\E(S_1 S_2)$, a fraction
-$S_1'$ of the heritable allele frequency change $S_2$ were \emph{previously}
-built associations between the neutral allele and a fitness background from
-$S_1$. As long as the direction of selection is the same, the expected change
-of this fraction of associations formed in the first generation to later
-generations 2 and 3 would be,
-
-\begin{align}
-  \E(S_1 S_2) &= \E(S_1^2) (1-\kappa)(1-r) = \E(S_1^2) Q_2^2 \\
-  \E(S_1 S_3) &= \E(S_1^2) (1-\kappa)^2(1-r)^2 = \E(S_1^2) Q_3^2.
-\end{align}
-
-With these covariance terms, we now have a full model for $\var(p_t - p_0)$. We
-will look at the case for $t=2$. As shown in equation XXX, the second moment of
-every term is a function of the variance in neutral allele frequency in that
-generation, $x_t(1-x_t)$. This variance does not stay its initial value of
-$x_0(1-x_0)$; it decays each generation due to coalesce from the drift and
-selection processes modeled by this approach. The rate at which this decay
-happens is the effective population size each generation implied by this model. 
-We can write,
-
-\begin{align}
-    \label{eq:vardecay}
-  \var(x_2 - p_0) &= \E({S_1}^2) + \E(S_1 S_2) + \E({S_2}^2) \\
-                  & + \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) \\
-  \frac{\var(x_2 - p_0)}{p_0(1-p_0)} &= \frac{V_h}{2N_{e,1}} + \frac{V_h Q_2^2}{2N_{e,2}}\left(1-\frac{1}{2N_{e,1}}\right)  +  \frac{V_h}{2N_{e,2}}  \left(1-\frac{1}{2N_{e,1}}\right)  \\ 
-                                     &+ \frac{V_n}{8N_{e,1}} + \frac{1}{4N_{e,1}} + \frac{V_n}{8N_{e,2}}  \left(1 - \frac{1}{2N_{e,1}}\right) 
-+ \frac{1}{4N_{e,2}} \left(1 - \frac{1}{2N_{e,1}}\right).
-\end{align}
-
-In general, the evolution of the variance in allele frequency change in a
-system with heritable fitness due to a single locus $r$ recombination fraction
-away from the neutral site is given by,
-
-\begin{align}
-  \label{eqn:var_xt}
-  \frac{\var(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^\infty \left(\frac{V_h Q_i^2}{2 N_{e,i}} + \frac{V_n}{8N_{e,i}} + \frac{1}{4N_{e,i}} \right)\prod_{t=1}^{i-1} \left(1-\frac{1}{2 N_{e,t}}\right). \\
-\end{align}
-
-Since the probability of coalesce (or, equivalently, identity by descent) is
-proportional to the variance in allele frequency change
-\parencite{Barton2000-zg}, this encodes the pairwise coalescent rate of the
-population through time under selection through the $N_{e,t}$ terms. Others
-have found that the genealogies under purifying selection can be characterized
-by such a time-dependent effective population size
-\parencite{Nicolaisen2013-gv}. In the original derivation of
-\textcite{Santiago1995-hx}, they set each $N_{e,t}$ term to a single $N$ and
-solve this summation in the limit to infinity giving,
-
-\begin{align}
-  \label{eq:var_asymp}
-  \frac{\var(x_t - p_0)}{p_0(1-p_0)} &= \frac{4 V_h Q_\infty^2 +V_n+2}{4 (2 N-1)}.
-\end{align}
-
-To relate this to an asymptotic effective population size, note that in an
-ideal Wright--Fisher population,
-
-\begin{align}
-  \V(x_t) &= p_0(1-p_0) \left[1 - \left(1 - \frac{1}{2N_e}\right)^t \right]
-\end{align}
-
-such that if we look at the change in the variance in allele frequency change
-between adjacent generations $t$ and $t-1$, we get an implied $N_{e,t}$
-
-\begin{align}
-  \label{eqn:ne_var}
-  N_{e,t} = \frac{p_0(1-p_0) - \var(x_{t-1})}{2(\var(x_t) - \var(x_{t-1}))}
-\end{align}
-%
-(cf. p. 1018, \cite{Santiago1995-hx}).
-
-Solving this for equation \eqref{eq:var_asymp} gives the asymptotic effective
-population size,
-
-\begin{align}
-  %N_e &= \frac{2 (2 N-1)}{4 V_h Q_\infty^2 +V_n+2} \\
-  N_e &\approx \frac{4N}{4 V_h Q_\infty^2 +V_n+2}
-\end{align}
-
-(cf. equation 18, \cite{Santiago1995-hx}). Levels of pairwise diversity,
-however, depend on the full distribution of pairwise coalesce times, or the
-$N_{e,t}$. While this would require solving the recursion defined by equations
-\eqref{eqn:var_xt} and \eqref{eqn:ne_var}, \textcite{Santiago1998-bs} show that
-a reasonably good approximation follows from using a single effective
-population size for all timepoints, $N_{e,t} = N_e$, but use the time-indexed
-$Q_t$. This leads to,
-
-\begin{align}
-  \label{eq:ne_t}
-  N_{e,t} &\approx \frac{4N}{4 V_h Q_t^2 +V_n+2}
-\end{align}
-
-(c.f. equation 15 of \cite{Santiago1995-hx} and p. 1270 of \cite{Santiago2016-mu}).
-
-\subsection*{Effective Population Size Under Polygenic Selection}
-
-The results of the previous section assume only a single locus a recombination
-fraction $r$ apart from the neutral locus is determining the coalesce rate. The
-model of \textcite{Santiago1998-bs} extends this to the case where fitness is
-polygenic. As they show in their original paper, this system can accommodate a
-variety of different equilibrium selection systems as long as expressions for
-$V_h$ and $1-\kappa$ can be worked out. Other authors have extended similar models to
-more elaborate bbreeding structures \parencite{Wray1990-zf,Woolliams1993-qo}.
-
-First, we extend equation \eqref{eq:ne_t} to a polygenic system. Throughout, a
-multiplicative fitness model is assumed (i.e. independent effects, no
-epistasis), where the fitness of individual $i$ is the product of their fitness
-contributions from each locus $l$, $w_{i,l}$, giving $w_i = \prod_{l=1}^n
-w_{i,l}$. Then, assuming mean fitness is one and independence between sites,
-the total fitness variation is multiplicative across the locus-level fitnesses,
-
-\begin{align}
-  V_h %&= \E_i\left( w_i^2 \right) - 1 \\
-      %&= \prod_{l=1}^n (1 + \alpha_{l}^2 \V(g_l))  - 1  \\
-      &= \prod_{l=1}^n (1 + V_{h,l})  - 1
-\end{align}
-
-Similarly, the $Q_t^2 V_h$ term in the single-locus model is multiplicative
-under the polygenic model, as the change in site-specific fitness variances are
-assumed independent,
-
-\begin{align}
-  Q_t^2 V_h &= \prod_{l=1}^n \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)  - 1
-\end{align}
-
-where $Q_{t,l}$ is the cumulative impact of selection due to the selected site
-$l$ (c.f. equation 4 \cite{Santiago1998-bs}). The factor of $\nicefrac{1}{2}$
-is due to the fact that we are ignoring the chance that the fitness background
-on the homologous chromosome recombines onto the haplotype with the tracked
-neutral allele (i.e. due to the $Q_t''$ associations). In other words, only
-half the fitness variation in a diploid can stay associated with the neutral
-allele through the generations.
-
-To simplify the derivation, assume $V_n = 2$ as under a Wright--Fisher model.
-Then, the effective population size is,
-
-\begin{align}
-  N_{e,t} &\approx \frac{N}{V_h Q_t^2 + 1} \\
-          &\approx \frac{N}{\prod_{l=1}^n \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)} \\
-          &\approx N \exp\left(- \sum_{l=1}^n \frac{V_{h,l} Q_{t,l}^2}{2}\right)
-\end{align}
-
-(c.f. \cite{Santiago1998-bs} equation 4b). If we write each site $l$'s
-contribution to the variance as a deviation from the average contribution,
-$V_{h,l}^2 = \nicefrac{1}{n} V_h + \varepsilon_l$, then
-
-\begin{align}
-  N_{e,t} &\approx N \exp\left(- \frac{1}{2}\sum_{l=1}^n \left(\frac{1}{n}V_{h} + \varepsilon_l\right)  Q_{t,l}^2\right) \\
-          &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{t,l}^2+ \sum_{l=1}^n \varepsilon_l  Q_{t,l}^2\right)
-\end{align}
-
-Santiago and Caballero assume that over evolutionary replicates, as selected
-sites are randomly distributed, there is not systematic covariance between the
-fitness variation at a site and the decay of its association with the neutral
-site, e.g. $\E(\varepsilon_l Q_{t,l}^2) = 0$, and thus ignore the second sum
-term.
-
-\begin{align}
-      N_{e,t} &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{t,l}^2\right) \\
-\end{align}
-
-\subsection*{Continuous Approximation to the Segment Under Selection}
-
-Our model differs from that of \textcite{Santiago1998-bs} in that we do not
-integrate over the entire genome. In their 1998 model, Santiago and Caballero
-consider the impact of both sites under selection on the same chromosome, as
-well as sites on independently assorting chromosomes. We only consider the
-contribution of a single segment under purifying selection, and the reduction
-from additional segments accumulate multiplicatively (similar to our
-implementation of the classic BGS model, XXX). Like their model, we imagine the
-reduction experienced by a focal neutral site in the middle of the segment.
-The implementation of our method differs slightly on though, see XXX.
-
-Since $Q_{t,l}$ terms depend on $l$ only through the recombination fraction
-between the neutral allele and the selected site, $r(l)$. Assuming a constant
-per-basepair recombination rate of $r_\text{BP}$, $r(l) = r_\text{BP} l$ we can
-approximate the sum with an integral for the asymptotic $Q_\infty$. The
-reduction experienced in the middle of a chromosome is twice the reduction
-experienced from each half,
-
-\begin{align}
-      N_{e,\infty} &\approx N \exp\left(- \frac{V_{h}}{2L} \sum_{l=1}^L Q_{\infty,l}^2\right) \\
-              &\approx N \exp\left(- \frac{V_{h}}{M} \int_0^{\nicefrac{M}{2}} Q_{\infty}(r)^2 dr \right) \\
-\end{align}
-
-where $M = n r_\text{BP}$ is the total segment length in Morgans. Then, the
-total asymptotic effect of associations from all sites as $Q_\infty^2$.  Then,
-
-\begin{align}
-  Q_\infty^2 &= \frac{2}{M}\int_0^{\nicefrac{M}{2}} Q_{\infty}(r)^2 dr  \\
-  &=  \frac{2}{M} \int_0^{\nicefrac{M}{2}} \left(\frac{1}{1 - (1-r)Z}\right)^2 dr  \\
-  &= \frac{2}{(1-Z) (2-(2-M) Z)}
-\end{align}
-
-Santiago and Caballero further approximate this; we can derive their
-approximation by setting $Z = 1-\kappa$ and doing a first-order Taylor series
-expansion around $\kappa$ (since the loss in variance is presumed to be small).
-Then,
-
-\begin{align}
-  Q_\infty^2 = \frac{2}{M (1-Z)} + \frac{2 (M-2)}{M^2}+\frac{2 \left(M^2-4 M+4\right) (1-Z)}{M^3} + \mathcal{O}(\kappa^2)
-\end{align}
-
-Santiago and Caballero's approximation only keeps this first term, e.g. their
-$Q_\infty^2 \approx 2/(M(1-Z))$, which is only accurate in the domain $M >
-0.2$.
-
-\begin{align}
-  N_{e,\infty}^{\text(SC98)} \approx N \exp\left(-\frac{V_h}{(1-Z)M}\right) 
-\end{align}
-
-(c.f. \cite{Santiago1998-bs} equation 8). For small $M$, the small $\kappa$
-approximation strongly deviates from the full equation; thus, throughout, we
-use the unapproximated version,
-
-\begin{align}
-  \label{eq:Ne_inf}
-  N_{e,\infty} \approx N \exp\left(-\frac{V_h}{(1-Z) (2-(2-M) Z)} \right).
-\end{align}
-
-For the reasons described earlier, a single asymptotic $N_{e,\infty}$ cannot
-fully characterize pairwise coalescent rates in all cases, and thus neutral
-diversity levels depend on the full $N_{e,t}$, for which we integrate equation
-\eqref{eq:q_t} to get,
-
-\begin{align}
-  \label{eq:net_integral}
-  N_{e,t} \approx N \exp\left(-\frac{V_h}{M} \int_0^{\nicefrac{M}{2}} \left(1+\frac{(1-k) (1-r) \left(1-(1-k)^t (1-r)^t\right)}{k+(1-k)r} \right) dr \right).
-\end{align}
-
-\subsection*{The Variance Dynamics}
-
-The heritable variance in fitness $V_h$ associated with the tracked neutral
-allele does not remain constant through the generations, but decays due to
-selection and drift, and is maintained by the input of new mutations. For an
-arbitrary polygenic selective system, the dynamics are extraordinarily
-complicated due to the complexity of multilocus selection, and the dynamics of
-the trait response to selection depend on the variance, which in turn depends
-on higher moments of the fitness distribution. We discuss the simplification of
-Santiago and Caballero's model here; hereafter, we assume that $V_h = V_A$, the
-genetic variance in additive fitness put on a relative scale such that
-$\E_i(f_i) = 1$.
-
-Each generation, the genetic variance changes due to changes in the allele
-frequencies at selected sites and linkage disequilibria between these sites.
-These are typically expressed as recursions, but certain models permit these
-changes to be expressed as a proportional reduction (e.g. truncation selection
-and stabilizing selection models; \cite{Keightley1988-eq} p. 36, and
-\cite{Walsh2018-bt} p. 557). Santiago and Caballero's model follows this
-approach, assuming the variance decays due to selection and drift at a constant
-rate $Z := 1-\kappa$ through the generations, such that the variance in the next
-generation excluding mutation is $V_A' = (1-\kappa) V_A$ and $\Delta V_A =
--\kappa V_A$.
-
-With mutation, $\Delta V_A = V_m - \kappa V_A$. The model of Santiago and
-Caballero assumes that under the long-run equilibrium, $\Delta V_A = 0$, such
-that $\kappa V_A = V_m$, and thus the rate of loss is $\kappa =
-\nicefrac{V_m}{V_A}$. Then the variance for the dynamics simplify to,
-
-\begin{align}
-  \label{eq:Z}
-  V_A' &= \left(1 - \frac{V_m}{V_A}\right) V_A \\
-  Z &:= \frac{V_A'}{V_A} = 1 - \kappa = 1 - \frac{V_m}{V_A}
-\end{align}
-
-(c.f. \cite{Santiago1998-bs} equation 9).
-
-% TODO this is a good description that could make its way to maintext
-
-\subsection*{Effective Population Size Under Strong Background Selection}
-
-Under strong background selection, multiple segregating deleterious sites
-contribute to fitness variation. BGS models assume a multiplicative fitness
-model, such that the fitness of an individual is
-
-\begin{align}
-  w = (1-s)^{n}
-\end{align}
-
-where $n$ is the number of deleterious mutations this individual carries.
-
-Classic background selection theory assumes an infinite population size, such
-that mutation frequencies are at their mutation-selection equilibrium and the
-expected number of mutations across individuals is $\E(n) = \nicefrac{\mu}{s}L$
-where $L$ is the total number of basepairs that can be mutated (the mutational
-target size).  Under a multiplicative fitness function in an infinite
-population, selection cannot generate linkage disequilibria
-\parencite{Turelli1990-kd}. Thus the classic strong background selection model
-assumes additive genetic fitness is equal to the additive genic fitness (which
-excludes the contribution of covariance between selected sites),
-
-\begin{align}
-  \label{eq:va_bgs}
-  V_\text{BGS} &:= \V(w) = 2 s^2 \sum_{l=1}^L p_l(1-p_l)  \\
-            &= 2 L s^2 \frac{\mu}{s} \left( 1 - \frac{\mu}{s}\right) = U s + \mathcal{O}(\mu^2) \approx Us.
-\end{align}
-
-since under mutation-selection balance $p_l = \nicefrac{\mu}{s}$ and $U$ is
-defined $U := 2 L \mu$, the deleterious mutation rate per diploid genome, per
-generation. They then set $V_A = V_\text{BGS}$, and note that $V_m \approx
-Us^2$, since mutations have frequency of $\nicefrac{1}{2N}$ and each increase
-the variance by a factor of $s^2$. Note that because the classic BGS model
-assumes deterministic selection dynamics (i.e. an infinite population), the
-equilibrium $Z$ given by equation \eqref{eq:Z} excludes the reduction due to
-drift. The reduction factor under the strong BGS model is,
-
-\begin{align}
-  Z_\text{BGS} := 1 - \frac{Vm}{V_\text{BGS}} = 1 - \frac{Us^2}{Us} = 1 - s
-\end{align}
-
-(c.f. \cite{Santiago1998-bs} equation 10). An alternate derivation can be found
-by working out the deterministic reduction in variance from the single-locus
-dynamics.
-
-Now, we can use these values for $Z_\text{BGS}$, $V_m$, and $V$ in equation
-\eqref{eq:Ne_inf}. Through this alternate quantitative genetics derivation, we
-arrive at the same equation as classic BGS theory,
-
-\begin{align}
-  N_{e,\infty}^\text{BGS} = N \exp\left( - \frac{U}{2s + M - Ms} \right).
-\end{align}
-
-This matches equation 8 of \textcite{Hudson1995-xc} except for the $Ms$ term
-and equation 10 of \textcite{Nordborg1996-nq} except for the $2s$ term.
-
-\subsection*{Effective Population Size Under Weak and Strong BGS} \label{supp:weak-strong}
-
-With the possibility of weak selection, deleterious alleles can drift up in
-frequency and fix. Consequently, the fitness distribution is no longer
-stationary, as the set of haplotypes without any deleterious mutations can be
-lost due to drift. This changes the shape of the fitness distribution, and thus
-the variance parameter that our model relies upon
-\parencite{Gessler1995-hz,OFallon2010-my,Good2013-lp,Haigh1978-gt,Higgs1995-xc}.
-So far, finding an explicit equation for the variance has been difficult due to
-the ``moment-closure problem": the variance depends on higher moments such as
-the skew, which in turn depends on higher moments. While the weak selection
-regime is composed of an interaction of mutation, selection, drift, and linkage
-processes and the dynamics that differ qualitatively based on parameter
-combinations. In their approximation, \textcite{Santiago2016-mu} consider the
-modified additive genic variance due to the fixation of weakly deleterious
-mutations. This fixation process is related to the Muller's Ratchet, the rate
-$R$ of fixation of deleterious mutations per generation per region, since each
-fixation directly reduces the additive genic variance of the fitness
-distribution. However, in general, the rate of this ratchet is unknown since it
-depends on the fitness variance (and thus higher moments of the fitness
-distribution).
-
-% NOTE: The S&C math using Fisher's FToNS argues fitness decreases
-% at a rate of -s/T each generation. However their model is haploid;
-% in a diploid model, fixation makes everyone  a homozygote with fitness
-% effect -2s; thus the rate must be -2s/T which agrees with sim
-
-In their 2016 paper, Santiago and Caballero derive a general equation for the
-fitness variance that accounts for the reduction due to the fixation of weakly
-deleterious alleles (i.e. the rate of Muller's ratchet). There derivation for
-the variance uses a heuristic approach based on Fisher's Fundamental Theorem of
-Natural Selection (equation 2, \cite{Santiago2016-mu}, equation 15A
-\cite{Garcia-Dorado2007-jj}). A more formal derivation follows from
-\textcite{Higgs1995-xc} (though they assume a haploid asexual system, the end
-result is the same up to a factor of two). In their work, they derive equations
-for the change per generation for the moments and cross-moments of the fitness
-distribution under multiplicative selection in a haploid model. They find, the
-haploid rate of the ratchet is $r = u - s v_n$ (equation 13.3,
-\cite{Higgs1995-xc}), where we use lowercase to distinguish between haploid and
-diploid models. Here, $v_n$ is the variance in the \emph{number} of deleterious
-mutations; as long as selection is not too strong, the haploid additive
-variance is $v_{A} \approx s^2 v_n$. Substituting this for $v_n$ and
-rearranging, the haploid fitness variance is related to the $v_A = us - rs$,
-which is identical to Santiago and Caballero's (\citeyear{Santiago2016-mu})
-equation 2, noting that the waiting time between fixations (i.e. ratchet
-clicks) is $T = \nicefrac{1}{r}$. Intuitively, the fitness variance $Us$ is
-reduced by the rate at which deleterious mutations fix and remove fitness
-variation from the population.
-
-Now, note that under a diploid model, the fitness variance is twice that as the
-haploid model, so $V_A = 2 v_A$. Then, the fitness variance is,
-
-\begin{align}
-  \label{supp-eq:Va}
-  V_{A} = Us - 2Rs. 
-\end{align}
-
-where $U = 2u$ and $R=r$ is the diploid rate of the ratchet, which depends on
-higher moments of the fitness distribution, but only differs from $r$ through a
-rescaling of the population size. Note that if selected sites are sufficiently
-deleterious that they cannot fix (or the population is infinite), $R=0$ and the
-additive fitness variance is identical to that under strong background
-selection, $V_A = Us = V_\text{BGS}$ (equation \eqref{eq:va_bgs}). Then, the
-fixation probability of a new deleterious mutation with additive effects ($1-s$
-in the heterozygote, $1-2s$ in the homozygote) in a diploid population is,
-
-\begin{align}
-  p_\text{fix} = \frac{1-\exp(\nicefrac{2 N_e s}{N})}{1-\exp(4 N_e s)} \approx
-  \frac{2 N_e s}{N \left(\exp(4 N_e s)-1\right)}
-\end{align}
-
-(c.f. \cite{Durrett2008-ql}, equation 7.21; \cite{Kimura1957-rk}, equation
-5.6). Then, Santiago and Caballero argue that the rate of the ratchet is the
-inverse of the averaging waiting time until a fixation, $T \approx 1/(UN
-p_\text{fix})$ (note that since $U$ is the \emph{diploid} mutation rate, the
-total population mutation rate is $UN$ per generation), so
-
-\begin{align}
-  T \approx \frac{\exp(4 N_e s)-1}{2 U s N_e}
-\end{align}
-
-Since the ratchet rate depends on $N_e$, which in turn depends on the rate of
-the ratchet, we write the ratchet, or substitution, rate per basepair per
-generation as,
-
-\begin{align}
-  R(N_e) := \frac{2 U s N_e}{\exp(4 N_e s)-1}.
-\end{align}
-
-Under weak and strong background selection the variance input of mutations
-$V_m$ is the same, and using the $V_A$ derived above, the decay in variance is
-
-\begin{align}
-  Z_\text{WS} = 1 - \frac{V_m}{V_A} = 1 - \frac{Us^2}{(U-R(N_e))s} = 1 - \frac{Us}{U-R(N_e)}.
-\end{align}
-
-At equilibrium, the $N_e$ that determines fixation probability is the
-asymptotic $N_{e,\infty}$. Consequently, Santiago and Caballero argue that the
-asymptotic reduction in effective population size uses the $Q_\infty^2$ and is
-given by the following system of non-linear equations,
-
-\begin{align}
-  R(N_{e,\infty}) &= \frac{4 U s N_{e,\infty}}{\exp(4 N_{e,\infty} s)-1}\\
-  N_{e,\infty} &\approx N \exp\left(-\frac{V_A}{(1-Z) (2-(2-M) Z)} \right)
-\end{align}
-
-The solution to these equations is the asymptotic equilibrium $\widetilde{N}_e$
-and equilibrium ratchet rate $\widetilde{R}$. This sets the equilibrium
-variance, $\widetilde{V} = Us - 2\widetilde{R}s$, or $\widetilde{V} = (U -
-2\widetilde{R})s$. Note that because $p_\text{fix} \le \nicefrac{1}{2N}$ for
-deleterious to neutral variation, $2R \le U$, or the diploid mutation rate must
-be greater than twice the per-basepair per-generation substitution rate for
-equilibrium variance $\widetilde{V} > 0$ for $s>0$.
-
-\subsection*{Steady-state Equilibrium Genic Variance} \label{supp:genic-va}
-
-We also find that Equation \eqref{eq:Va} (and Supplementary Materials Equation
-\ref{supp-eq:Va} above) can also be found through Kimura's diffusion models
-with a flux of mutations into infinite, discrete sites
-(\citeyear{Kimura1969-jw}). This model integrates the Kolmogorov backwards
-equation over time and forward variable to find the expected value of a
-functional of the initial frequency. Using the functional $I(p) = s^2 p(1-p)$
-for genic variance given mutations start at frequency $p$, one finds that the
-steady-state genic variance for $p=\nicefrac{1}{2N_e}$ is
-
-\begin{align}
-    V_a^{SS} &= 2 U N s \left(\frac{1}{2N_e} - p_f(N_e, s)\right) \\
-             &= \left(\frac{UN}{N_e} - 2 U Np_f(N_e, s))\right)s \nonumber \\
-             &\approx \left(U - 2 R\right)s \nonumber
-\end{align}
-
-which is approximately Equation \eqref{eq:Va}.
-
-\subsection*{Misspecification of the Polygenic Selection Model}
-
-Thus far, we have developed the general model of equation \eqref{XXX} by
-assuming that the only source of $V_h$ is from the deleterious
-mutation-selection-draft-drift process. However, other polygenic evolutionary
-processes can also create, alter, and eliminate fitness variation. By assuming
-that the only fitness variation originates from deleterious
-mutation-selection-drift-draft processes, we introduce some model
-misspecification. There are two main types of misspecification under the
-selection model. First, non-polygenic positive selection processes, such as
-recurrent sweeps, may create small-scale distortions in local genealogies.
-These have been found to only impact diversity on small spatial scales
-\parencite{Murphy2022-sj} in humans, and explain little of the total variation
-in megabase-scale patterns of diversity. Second, there is model
-misspecification due to our model assuming that the only polygenic selection
-process happening is the flux of new deleterious mutations in, and their
-eventual loss or fixation. Other polygenic selection processes could be
-occurring, e.g. traits could be responding to directional selection. The
-Santiago and Caballero model our work extends is very general, and can account
-for multiple types of polygenic selection. The key input parameters are the
-equilibrium heritable fitness variation $V_h$, and the variation introduced by
-mutation each generation $V_m$.
-
-Let us assume for now that,
-
-\begin{align}
-  \label{eq:Ne_inf}
-  N_{e,\infty} \approx N \exp\left(-\frac{V_h + \varepsilon}{(1-Z) (2-(2-M) Z)} \right).
-\end{align}
-
-\subsection*{Heterozygosity and $B$ values Under Weak and Strong Selection}
-
-While the selection dynamics are set by the asymptotic $Q_\infty^2$ that
-determines the equilibrium $\widetilde{N}_e$ and $\widetilde{R}$,
-heterozygosity is determined by the sequence $\widetilde{N}_{e,t}$, where each
-term is given by \eqref{eq:net_integral}. We experimented with calculating this
-full sum in our methods, but it seemed to make little difference in our
-application. Additionally, it was too costly to calculate computationally
-for genome-wide calculations of the B' map.
-
-\subsection*{Numeric methods for calculating $B$}
-
-Depending on the number of features, this can be incredibly memory intensive.
-
-\subsection*{Optimization}
-
-Optimization in non-linear regression is notoriously difficult both in general
-(Bates) and in the specific problem of estimating the effects of genome-wide
-selection \parencite{Murphy2022-sj}. We experimented with a variety of global
-and local optimization approaches and parameterizations, but found that
 

### 2023-06-23 manuscript/supplementary.tex
ms updates

@@
 \tableofcontents
+
+\section{Theory}
+\subsection{Overview of Quantitative Genetic Models of Effective Population Size}
+
+Here, we step through a quick derivation of \textcite{Santiago1995-hx}.
+Throughout, we assume random mating, hermaphroditic individuals, and a constant
+population size. The change in a neutral allele's frequency in one generation
+can be partitioned into the three sources of stochasticity: the random
+associations with fitness backgrounds (i.e. \emph{draft}), the non-heritable
+randomness in family size, and the Mendelian noise from heterozygotes
+segregating. If we let $x_{0,i} \in \{0, \nicefrac{1}{2}, 1\}$ be the frequency
+of neutral alleles individual $i$ in generation 0 carries, we can partition the
+random neutral allele frequency of the population ($x_t$ without the individual
+index) into the underlying stochastic causes,
+
+\begin{align}
+  x_1 = \frac{1}{2N} \sum_{i=1}^N \left( x_{0,i}k_{0,i} + \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \right)
+\end{align}
+
+where $k_{0,i}$ is the number of surviving gametes parent $i$ passes on, and
+$\delta_{0,i,j}$ is a random term that encapsulates the noise due to Mendelian
+segregation of heterozygotes. If parent $i$ is a homozygote ($x_{0,i} \in \{0,
+1\}$), then $\delta_{0,i,j} = 0$, whereas if $x_{0,i} = \nicefrac{1}{2}$ then
+$\delta_{0,i,j} = \pm \nicefrac{1}{2}$ with equal probability. This is because
+a heterozygous parent will transmit half a neutral allele in expectation, but
+each round of Mendelian segregation must pass on either 0 or 1 alleles, which
+the random $\pm \nicefrac{1}{2}$ term imposes. The factor of $\nicefrac{1}{2}$
+is due to the fact that we're summing over $N$ diploids, but considering the
+number of gametes they transmit. Since each diploid parent must have two
+offspring to maintain a constant population size, $\nicefrac{1}{N} \sum_i
+k_{0,i} = 2$.
+
+The frequency in the initial generation is $x_0 = \nicefrac{1}{N} \sum_{i=1}^N
+x_{0,i}$, though to indicate that we treat this as fixed, we use $p_0 = x_0$.
+Then, the allele frequency change is,
+
+\begin{align}
+  \Delta x_1 = x_1 - p_0 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (k_{0,i} - 1) + \frac{1}{2N} \sum_{i=1}^N \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \\
+  \Delta x_1 &= K_1 + H_1
+\end{align}
+
+where $K_1$ and $H_1$ are the random change in neutral allele frequency change
+due to offspring number (including heritable and non-heritable components), and
+Mendelian segregation.
+
+Now, let us look the variance of $\var(\Delta x_1)$ over evolutionary
+replicates. Since the Mendelian random process and the offspring random process
+are independent,
+
+\begin{align}
+  \var(\Delta x_1) &= \var(K_1) + \var(H_1).
+\end{align}
+
+Looking at each term,
+
+\begin{align}
+  \var(K_1) &\approx \frac{1}{4N^2} \sum_{i=1}^N \var\left(x_{0,i} (k_{0,i} - 1) \right)
+\end{align}
+
+where we ignore the covariance terms due to the sum, since these are on order
+$\nicefrac{1}{N^3}$. In the first generation from an arbitrary starting point,
+the neutral alleles assort independently into diploids with respect to their
+fitness, so
+
+\begin{align}
+  \var(K_1) &\approx \frac{1}{4N^2} \sum_{i=1}^N \var(x_{0,i}) \var(k_{0,i} - 1)  \\
+            &\approx \frac{1}{4N} \var(x_{0,i}) \var(k_{0,i}).
+\end{align}
+
+Assuming that no correlation between parental gametes (e.g. no inbreeding),
+$\var(x_{0,i})$ is the binomial variance in individual allele frequency, or
+$p_0(1-p_0)/2$, and $\var(k) := \var(k_{0,i})$ is the offspring variance of an
+individual given by the reproduction process. For example, if the reproduction
+process is a neutral multinomial Wright--Fisher, $\var(k) \approx 2$.
+
+\begin{align}
+  \var(K_1) &\approx \frac{p_0(1-p_0)}{2N} \frac{\var(k)}{4}.
+\end{align}
+
+The Mendelian noise variance term can be derived similarly. First, the sum over
+each parent's transmitted gametes can be simplified by noting that parents are
+exchangeable over evolutionary replicates with respect to their contribution to
+this term. In a constant population size, the double summation over parents and
+their offspring can be replaced by a summation over offspring, since both sum
+$N$ exchangeable terms. Then, note that $\var(\delta_{i,j}) = \nicefrac{1}{4}
+p_0(1-p_0)$, so
+
+\begin{align}
+  \var(H_1) &= \frac{p_0(1-p_0)}{2N} \frac{1}{2}.
+\end{align}
+
+Finally, we have
+
+\begin{align}
+  \var(\Delta x_1) &= \var(K_1) + \var(H_1) \\
+  &\approx \frac{p_0(1-p_0)}{2N} \frac{\var(k)}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2} \\
+                   &\approx \frac{p_0(1-p_0)}{2N}\left(\frac{\var(k)}{4} + \frac{1}{2}\right)
+\end{align}
+
+(c.f. \cite{Santiago1995-hx} equation 2 and \cite{Buffalo2019-qs} equation 30).
+Note that if the reproduction process is a neutral multinomial Wright--Fisher,
+$\var(k) \approx 2$, and this simplifies to the expected Wright--Fisher
+variance. If we were to \emph{define} a variance effective population size by
+setting this variance in neutral allele frequency change to the expected
+variance under a Wright--Fisher model ($V_\text{WF}$), we'd have
+
+\begin{align}
+  V_\text{WF} &:= \frac{p_0(1-p_0)}{2N_e} \\
+  V_\text{WF} &= \var(\Delta p_1) \\
+  N_e &= \frac{4N}{\var(k) + 2} \\
+\end{align}
+
+(c.f. \cite{Wright1938-tv}).
+
+Now, we look at this variance equation behaves with $t > 1$.
+
+\begin{align}
+  \V(x_1) = \E_1\left[(x_1 - x_0)^2\right]= \frac{x_0(1-x_0)}{2N},
+\end{align}
+
+and
+
+\begin{align}
+  \label{eq:var_wf}
+  \V(x_2) &= \E\left[(x_2 - x_0)^2\right] \\
+          &= \E_1\left[\E_2\left[((x_2-x_1) + (x_1 - x_0))^2 | x_1\right]\right] \\
+          &= \E_1\left[\E_2\left[(\Delta x_2 + \Delta x_1)^2 | x_1\right]\right] \\
+          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + 2\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
+          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] + \mathcal{C}_{1,2} \\
+          &= \frac{p_0(1-p_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{p_0(1-p_0)}{2N} + \mathcal{C}_{1,2}
+\end{align}
+
+Consequently, the variance of the neutral allele frequency changes each
+generation according to the probability of failing to coalesce each generation
+and the pairwise covariance terms $\mathcal{C}_{i,j}$ that build up due to associations
+with heritable fitness backgrounds. Note that the covariance term $\mathcal{C}_{1,2} =
+\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] = 0$ under neutral
+evolution, and this simplifies to the well-known equation for variance in a
+Wright--Fisher population.
+
+Now, we turn our attention to the case where there is heritable fitness
+variation, which leads to non-zero covariance terms $\mathcal{C}_{i,j}$. These
+terms emerge when $\var(k)$ has a heritable component of fitness that can be
+transmitted along with the neutral allele, thereby affecting the neutral
+allele's trajectory in later generations. If we partition the offspring number
+into heritable and non-heritable components, $k_i = 2f_i + \varepsilon_i$, then
+$\var(k_i) = 4 V_h + V_n$. When there is heritable fitness variation across
+individuals, $V_h > 0$. Because the population size is assumed to be constant,
+the mean heritable fitness across individuals is constrained to be $\E_i(f_i) =
+1$; this implies that the individual fitness values $f_i$ are relative
+fitnesses as is standard in population genetics, in this case to the population
+mean. Thus $V_h$ is the squared coefficient of heritable fitness variation
+(this is often denoted as $C^2 = \nicefrac{V_A}{\bar{w}^2}$, see
+\cite{Crow1958-pc,Charlesworth1987-ab,Houle1992-ur}).
+
+We can further decompose the $K_1$ term into $K_1 = S_1 + D_1$,
+
+\begin{align}
+  \var(\Delta x_1) &= \var(S_1) + \var(D_1) + \var(H_1) \\
+  &\approx \frac{p_0(1-p_0)}{2N} V_h + \frac{p_0(1-p_0)}{2N} \frac{V_n}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2}. \\
+\end{align}
+
+(c.f. \cite{Santiago1995-hx} equation 11).
+
+To see how the covariance terms accumulate, consider $\var(x_3 - p_0)$. Note
+that $\E(S_t) = \E(D_t) = \E(H_3) = 0$, since which neutral allele we track is
+arbitrary, so by symmetry the expected change is zero. Then,
+
+\begin{align}
+  \label{eq:var_x}
+  \var(x_3 - p_0) &= \E\left[ \left(S_1 + D_1 + H_1 + S_2 + D_2 + H_2 + S_3 + D_3 + H_3 \right)^2 \right] \\
+                  &= \E({S_1}^2) + \E({S_2}^2) + \E({S_3}^2)  \\
+                  &+ \E(S_1 S_2) + \E(S_1 S_3) + \E(S_2 S_3) \\
+                  &+ \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) + \E({D_3}^2) + \E({H_3}^2)
+\end{align}
+
+The covariance terms $\E(S_i S_j)$ for $j > i$ represent the expected neutral
+allele frequency change from a neutral allele becoming associated with a
+fitness background in generation $i$ and that fitness association persisting
+until generation $j$. However, the covariances terms $\E(S_j S_i)$ for $j > i$
+are zero since associations in the future cannot affect the past (see p. 1041
+of \cite{Buffalo2019-qs}). Let us consider $S_1$ and $S_2$. We have,
+
+\begin{align}
+  S_1 = \frac{1}{2N} \sum_{i=1}^N x_{0,i}(f_{0,i} - 1) = \cov(x_{0,i}, f_{0,i}) \\
+  S_2 = \frac{1}{2N} \sum_{i=1}^N x_{1,i}(f_{1,i} - 1) = \cov(x_{1,i}, f_{0,i})
+\end{align}
+
+which are chance covariances (across the population, \emph{not} evolutionary
+replicates) created by the random sorting of neutral alleles and fitness into
+individuals, \emph{each generation}. These covariances are equivalent to the
+Robertson-Price equation \parencite{Robertson1966-fs,Price1970-si}, which
+predict the change in neutral frequency due to heritable fitness (and likewise
+with the change due to non-heritable factors).
+
+Considering the underlying haplotypes that lead to $x_{0,i}$ and $f_{0,i}$ can
+give us an equation for the dynamics of the $\E(S_i S_j)$ (for $j > i$) terms
+over evolutionary replicates. Let us partition the allele frequency and fitness
+by into the average of the paternal contributions. The neutral allele frequency
+per gamete is either 0 or 1, and we assume the fitnesses across gametes are
+additive. Then, as long as there is random mating, we can simplify the
+associations in a diploid by looking at a single gamete,
+
+\begin{align}
+  S_1 = \cov(x_{0,i}, f_{0,i}) &= \cov\left(\frac{x_{0,i}' + x_{0,i}''}{2}, f_{0,i}' + f_{0,i}''\right)  \\
+                               &= \frac{1}{2}\left(\cov(x_{0,i}', f_{0,i}' + f_{0,i}'') + \cov(x_{0,i}'', f_{0,i}' + f_{0,i}'')\right)  \\
+                               &= \cov(x_{0,i}', f_{0,i}' + f_{0,i}'')  \\
+                               &= \cov(x_{0,i}', f_{0,i}') + \cov(x_{0,i}', f_{0,i}'')  \\
+                               &= S_1' + S_1''
+\end{align}
+
+which follows from symmetry, since it does not matter which gamete carrying the
+neutral allele we track (c.f. \cite{Santiago1998-bs} p. 2107). The primes now
+indicate whether the selected locus is on the same gamete as the neutral
+allele (single prime), or the homologous gamete (double prime).
+
+After meiosis, a fraction $1-r$ of these associations between the tracked
+neutral neutral allele $x_{0,1}'$ and fitness background $f_{0,i}'$ persist,
+and a fraction $r$ disassociate its currently coupled background and create an
+association with the homologous fitness background, $f_{0,i}''$. When linkage
+is tight, this latter term can be ignored as we will do here, though it does
+impact background levels of neutral diversity \parencite{Santiago1995-hx}.
+Additionally, the fitness effects of the selected background may change in
+later generations, in complex ways \parencite{Barton1986-yh,Turelli1990-kd}.
+\textcite{Santiago1995-hx} assume an equilibrium level of fitness variation
+$V_h$, with the fitness variance associated with a particular background
+decaying at a simple geometric decay at rate $1-\kappa$ each generation, which
+is sufficient for background selection. An approximate rate of the decay can be
+worked out for the particular selective system. In general, this can be a much
+more complicated function; \textcite{Buffalo2019-qs} show that even fluctuating
+selection can be accommodated.
+
+Now, the associations created in generation 1 and that persist in future
+generations as (ignoring $S_i''$ terms) follow the pattern
+
+\begin{align}
+  S_{2}' &= S_1 (1-\kappa)(1-r) \\
+  S_{3}' &= S_1 (1-\kappa)^2(1-r)^2 \\
+  S_{4}' &= S_1 (1-\kappa)^3(1-r)^3.
+\end{align}
+
+Then, the cumulative effect of the fitness associations created in generation
+$1$ can be written as,
+
+\begin{align}
+  \label{eq:qterm}
+  S_1 Q_t &= S_1 + S_2' + S_3' + \ldots + S_t' \\
+  Q_t &= 1 + \sum_{i=1}^t (1-r)^i(1-\kappa)^i
+\end{align}
+
+or for continuous $t$,
+
+\begin{align}
+  \label{eq:q_t}
+  Q_t &= 1+\frac{(1-\kappa) (1-r) \left(1-(1-\kappa)^t (1-r)^t\right)}{\kappa+(1-\kappa)r}
+\end{align}
+which as $t \to \infty$, converges to,
+
+\begin{align}
+  \label{eq:Qinf}
+  Q_\infty &= \frac{1}{\kappa + r(1-\kappa)}.
+\end{align}
+
+This shows the long reach of heritable fitness factors in altering allele
+frequency through the generations. In terms like $\E(S_1 S_2)$, a fraction
+$S_1'$ of the heritable allele frequency change $S_2$ were \emph{previously}
+built associations between the neutral allele and a fitness background from
+$S_1$. As long as the direction of selection is the same, the expected change
+of this fraction of associations formed in the first generation to later
+generations 2 and 3 would be,
+
+\begin{align}
+  \E(S_1 S_2) &= \E(S_1^2) (1-\kappa)(1-r) = \E(S_1^2) Q_2^2 \\
+  \E(S_1 S_3) &= \E(S_1^2) (1-\kappa)^2(1-r)^2 = \E(S_1^2) Q_3^2.
+\end{align}
+
+With these covariance terms, we now have a full model for $\var(p_t - p_0)$. We
+will look at the case for $t=2$. As shown in equation XXX, the second moment of
+every term is a function of the variance in neutral allele frequency in that
+generation, $x_t(1-x_t)$. This variance does not stay its initial value of
+$x_0(1-x_0)$; it decays each generation due to coalesce from the drift and
+selection processes modeled by this approach. The rate at which this decay
+happens is the effective population size each generation implied by this model. 
+We can write,
+
+\begin{align}
+    \label{eq:vardecay}
+  \var(x_2 - p_0) &= \E({S_1}^2) + \E(S_1 S_2) + \E({S_2}^2) \\
+                  & + \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) \\
+  \frac{\var(x_2 - p_0)}{p_0(1-p_0)} &= \frac{V_h}{2N_{e,1}} + \frac{V_h Q_2^2}{2N_{e,2}}\left(1-\frac{1}{2N_{e,1}}\right)  +  \frac{V_h}{2N_{e,2}}  \left(1-\frac{1}{2N_{e,1}}\right)  \\ 
+                                     &+ \frac{V_n}{8N_{e,1}} + \frac{1}{4N_{e,1}} + \frac{V_n}{8N_{e,2}}  \left(1 - \frac{1}{2N_{e,1}}\right) 
++ \frac{1}{4N_{e,2}} \left(1 - \frac{1}{2N_{e,1}}\right).
+\end{align}
+
+In general, the evolution of the variance in allele frequency change in a
+system with heritable fitness due to a single locus $r$ recombination fraction
+away from the neutral site is given by,
+
+\begin{align}
+  \label{eqn:var_xt}
+  \frac{\var(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^\infty \left(\frac{V_h Q_i^2}{2 N_{e,i}} + \frac{V_n}{8N_{e,i}} + \frac{1}{4N_{e,i}} \right)\prod_{t=1}^{i-1} \left(1-\frac{1}{2 N_{e,t}}\right). \\
+\end{align}
+
+Since the probability of coalesce (or, equivalently, identity by descent) is
+proportional to the variance in allele frequency change
+\parencite{Barton2000-zg}, this encodes the pairwise coalescent rate of the
+population through time under selection through the $N_{e,t}$ terms. Others
+have found that the genealogies under purifying selection can be characterized
+by such a time-dependent effective population size
+\parencite{Nicolaisen2013-gv}. In the original derivation of
+\textcite{Santiago1995-hx}, they set each $N_{e,t}$ term to a single $N$ and
+solve this summation in the limit to infinity giving,
+
+\begin{align}
+  \label{eq:var_asymp}
+  \frac{\var(x_t - p_0)}{p_0(1-p_0)} &= \frac{4 V_h Q_\infty^2 +V_n+2}{4 (2 N-1)}.
+\end{align}
+
+To relate this to an asymptotic effective population size, note that in an
+ideal Wright--Fisher population,
+
+\begin{align}
+  \V(x_t) &= p_0(1-p_0) \left[1 - \left(1 - \frac{1}{2N_e}\right)^t \right]
+\end{align}
+
+such that if we look at the change in the variance in allele frequency change
+between adjacent generations $t$ and $t-1$, we get an implied $N_{e,t}$
+
+\begin{align}
+  \label{eqn:ne_var}
+  N_{e,t} = \frac{p_0(1-p_0) - \var(x_{t-1})}{2(\var(x_t) - \var(x_{t-1}))}
+\end{align}
+%
+(cf. p. 1018, \cite{Santiago1995-hx}).
+
+Solving this for equation \eqref{eq:var_asymp} gives the asymptotic effective
+population size,
+
+\begin{align}
+  %N_e &= \frac{2 (2 N-1)}{4 V_h Q_\infty^2 +V_n+2} \\
+  N_e &\approx \frac{4N}{4 V_h Q_\infty^2 +V_n+2}
+\end{align}
+
+(cf. equation 18, \cite{Santiago1995-hx}). Levels of pairwise diversity,
+however, depend on the full distribution of pairwise coalesce times, or the
+$N_{e,t}$. While this would require solving the recursion defined by equations
+\eqref{eqn:var_xt} and \eqref{eqn:ne_var}, \textcite{Santiago1998-bs} show that
+a reasonably good approximation follows from using a single effective
+population size for all timepoints, $N_{e,t} = N_e$, but use the time-indexed
+$Q_t$. This leads to,
+
+\begin{align}
+  \label{eq:ne_t}
+  N_{e,t} &\approx \frac{4N}{4 V_h Q_t^2 +V_n+2}
+\end{align}
+
+(c.f. equation 15 of \cite{Santiago1995-hx} and p. 1270 of \cite{Santiago2016-mu}).
+
+\subsection{Effective Population Size Under Polygenic Selection}
+
+The results of the previous section assume only a single locus a recombination
+fraction $r$ apart from the neutral locus is determining the coalesce rate. The
+model of \textcite{Santiago1998-bs} extends this to the case where fitness is
+polygenic. As they show in their original paper, this system can accommodate a
+variety of different equilibrium selection systems as long as expressions for
+$V_h$ and $1-\kappa$ can be worked out. Other authors have extended similar models to
+more elaborate bbreeding structures \parencite{Wray1990-zf,Woolliams1993-qo}.
+
+First, we extend equation \eqref{eq:ne_t} to a polygenic system. Throughout, a
+multiplicative fitness model is assumed (i.e. independent effects, no
+epistasis), where the fitness of individual $i$ is the product of their fitness
+contributions from each locus $l$, $w_{i,l}$, giving $w_i = \prod_{l=1}^n
+w_{i,l}$. Then, assuming mean fitness is one and independence between sites,
+the total fitness variation is multiplicative across the locus-level fitnesses,
+
+\begin{align}
+  V_h %&= \E_i\left( w_i^2 \right) - 1 \\
+      %&= \prod_{l=1}^n (1 + \alpha_{l}^2 \V(g_l))  - 1  \\
+      &= \prod_{l=1}^n (1 + V_{h,l})  - 1
+\end{align}
+
+Similarly, the $Q_t^2 V_h$ term in the single-locus model is multiplicative
+under the polygenic model, as the change in site-specific fitness variances are
+assumed independent,
+
+\begin{align}
+  Q_t^2 V_h &= \prod_{l=1}^n \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)  - 1
+\end{align}
+
+where $Q_{t,l}$ is the cumulative impact of selection due to the selected site
+$l$ (c.f. equation 4 \cite{Santiago1998-bs}). The factor of $\nicefrac{1}{2}$
+is due to the fact that we are ignoring the chance that the fitness background
+on the homologous chromosome recombines onto the haplotype with the tracked
+neutral allele (i.e. due to the $Q_t''$ associations). In other words, only
+half the fitness variation in a diploid can stay associated with the neutral
+allele through the generations.
+
+To simplify the derivation, assume $V_n = 2$ as under a Wright--Fisher model.
+Then, the effective population size is,
+
+\begin{align}
+  N_{e,t} &\approx \frac{N}{V_h Q_t^2 + 1} \\
+          &\approx \frac{N}{\prod_{l=1}^n \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)} \\
+          &\approx N \exp\left(- \sum_{l=1}^n \frac{V_{h,l} Q_{t,l}^2}{2}\right)
+\end{align}
+
+(c.f. \cite{Santiago1998-bs} equation 4b). If we write each site $l$'s
+contribution to the variance as a deviation from the average contribution,
+$V_{h,l}^2 = \nicefrac{1}{n} V_h + \varepsilon_l$, then
+
+\begin{align}
+  N_{e,t} &\approx N \exp\left(- \frac{1}{2}\sum_{l=1}^n \left(\frac{1}{n}V_{h} + \varepsilon_l\right)  Q_{t,l}^2\right) \\
+          &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{t,l}^2+ \sum_{l=1}^n \varepsilon_l  Q_{t,l}^2\right)
+\end{align}
+
+Santiago and Caballero assume that over evolutionary replicates, as selected
+sites are randomly distributed, there is not systematic covariance between the
+fitness variation at a site and the decay of its association with the neutral
+site, e.g. $\E(\varepsilon_l Q_{t,l}^2) = 0$, and thus ignore the second sum
+term.
+
+\begin{align}
+      N_{e,t} &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{t,l}^2\right) \\
+\end{align}
+
+\subsection{Continuous Approximation to the Segment Under Selection}
+
+Our model differs from that of \textcite{Santiago1998-bs} in that we do not
+integrate over the entire genome. In their 1998 model, Santiago and Caballero
+consider the impact of both sites under selection on the same chromosome, as
+well as sites on independently assorting chromosomes. We only consider the
+contribution of a single segment under purifying selection, and the reduction
+from additional segments accumulate multiplicatively (similar to our
+implementation of the classic BGS model, XXX). Like their model, we imagine the
+reduction experienced by a focal neutral site in the middle of the segment.
+The implementation of our method differs slightly on though, see XXX.
+
+Since $Q_{t,l}$ terms depend on $l$ only through the recombination fraction
+between the neutral allele and the selected site, $r(l)$. Assuming a constant
+per-basepair recombination rate of $r_\text{BP}$, $r(l) = r_\text{BP} l$ we can
+approximate the sum with an integral for the asymptotic $Q_\infty$. The
+reduction experienced in the middle of a chromosome is twice the reduction
+experienced from each half,
+
+\begin{align}
+      N_{e,\infty} &\approx N \exp\left(- \frac{V_{h}}{2L} \sum_{l=1}^L Q_{\infty,l}^2\right) \\
+              &\approx N \exp\left(- \frac{V_{h}}{M} \int_0^{\nicefrac{M}{2}} Q_{\infty}(r)^2 dr \right) \\
+\end{align}
+
+where $M = n r_\text{BP}$ is the total segment length in Morgans. Then, the
+total asymptotic effect of associations from all sites as $Q_\infty^2$.  Then,
+
+\begin{align}
+  Q_\infty^2 &= \frac{2}{M}\int_0^{\nicefrac{M}{2}} Q_{\infty}(r)^2 dr  \\
+  &=  \frac{2}{M} \int_0^{\nicefrac{M}{2}} \left(\frac{1}{1 - (1-r)Z}\right)^2 dr  \\
+  &= \frac{2}{(1-Z) (2-(2-M) Z)}
+\end{align}
+
+Santiago and Caballero further approximate this; we can derive their
+approximation by setting $Z = 1-\kappa$ and doing a first-order Taylor series
+expansion around $\kappa$ (since the loss in variance is presumed to be small).
+Then,
+
+\begin{align}
+  Q_\infty^2 = \frac{2}{M (1-Z)} + \frac{2 (M-2)}{M^2}+\frac{2 \left(M^2-4 M+4\right) (1-Z)}{M^3} + \mathcal{O}(\kappa^2)
+\end{align}
+
+Santiago and Caballero's approximation only keeps this first term, e.g. their
+$Q_\infty^2 \approx 2/(M(1-Z))$, which is only accurate in the domain $M >
+0.2$.
+
+\begin{align}
+  N_{e,\infty}^{\text(SC98)} \approx N \exp\left(-\frac{V_h}{(1-Z)M}\right) 
+\end{align}
+
+(c.f. \cite{Santiago1998-bs} equation 8). For small $M$, the small $\kappa$
+approximation strongly deviates from the full equation; thus, throughout, we
+use the unapproximated version,
+
+\begin{align}
+  \label{eq:Ne_inf}
+  N_{e,\infty} \approx N \exp\left(-\frac{V_h}{(1-Z) (2-(2-M) Z)} \right).
+\end{align}
+
+For the reasons described earlier, a single asymptotic $N_{e,\infty}$ cannot
+fully characterize pairwise coalescent rates in all cases, and thus neutral
+diversity levels depend on the full $N_{e,t}$, for which we integrate equation
+\eqref{eq:q_t} to get,
+
+\begin{align}
+  \label{eq:net_integral}
+  N_{e,t} \approx N \exp\left(-\frac{V_h}{M} \int_0^{\nicefrac{M}{2}} \left(1+\frac{(1-k) (1-r) \left(1-(1-k)^t (1-r)^t\right)}{k+(1-k)r} \right) dr \right).
+\end{align}
+
+\subsection{The Variance Dynamics}
+
+The heritable variance in fitness $V_h$ associated with the tracked neutral
+allele does not remain constant through the generations, but decays due to
+selection and drift, and is maintained by the input of new mutations. For an
+arbitrary polygenic selective system, the dynamics are extraordinarily
+complicated due to the complexity of multilocus selection, and the dynamics of
+the trait response to selection depend on the variance, which in turn depends
+on higher moments of the fitness distribution. We discuss the simplification of
+Santiago and Caballero's model here; hereafter, we assume that $V_h = V_A$, the
+genetic variance in additive fitness put on a relative scale such that
+$\E_i(f_i) = 1$.
+
+Each generation, the genetic variance changes due to changes in the allele
+frequencies at selected sites and linkage disequilibria between these sites.
+These are typically expressed as recursions, but certain models permit these
+changes to be expressed as a proportional reduction (e.g. truncation selection
+and stabilizing selection models; \cite{Keightley1988-eq} p. 36, and
+\cite{Walsh2018-bt} p. 557). Santiago and Caballero's model follows this
+approach, assuming the variance decays due to selection and drift at a constant
+rate $Z := 1-\kappa$ through the generations, such that the variance in the next
+generation excluding mutation is $V_A' = (1-\kappa) V_A$ and $\Delta V_A =
+-\kappa V_A$.
+
+With mutation, $\Delta V_A = V_m - \kappa V_A$. The model of Santiago and
+Caballero assumes that under the long-run equilibrium, $\Delta V_A = 0$, such
+that $\kappa V_A = V_m$, and thus the rate of loss is $\kappa =
+\nicefrac{V_m}{V_A}$. Then the variance for the dynamics simplify to,
+
+\begin{align}
+  \label{eq:Z}
+  V_A' &= \left(1 - \frac{V_m}{V_A}\right) V_A \\
+  Z &:= \frac{V_A'}{V_A} = 1 - \kappa = 1 - \frac{V_m}{V_A}
+\end{align}
+
+(c.f. \cite{Santiago1998-bs} equation 9).
+
+% TODO this is a good description that could make its way to maintext
+
+\subsection{Effective Population Size Under Strong Background Selection}
+
+Under strong background selection, multiple segregating deleterious sites
+contribute to fitness variation. BGS models assume a multiplicative fitness
+model, such that the fitness of an individual is
+
+\begin{align}
+  w = (1-s)^{n}
+\end{align}
+
+where $n$ is the number of deleterious mutations this individual carries.
+
+Classic background selection theory assumes an infinite population size, such
+that mutation frequencies are at their mutation-selection equilibrium and the
+expected number of mutations across individuals is $\E(n) = \nicefrac{\mu}{s}L$
+where $L$ is the total number of basepairs that can be mutated (the mutational
+target size).  Under a multiplicative fitness function in an infinite
+population, selection cannot generate linkage disequilibria
+\parencite{Turelli1990-kd}. Thus the classic strong background selection model
+assumes additive genetic fitness is equal to the additive genic fitness (which
+excludes the contribution of covariance between selected sites),
+
+\begin{align}
+  \label{eq:va_bgs}
+  V_\text{BGS} &:= \V(w) = 2 s^2 \sum_{l=1}^L p_l(1-p_l)  \\
+            &= 2 L s^2 \frac{\mu}{s} \left( 1 - \frac{\mu}{s}\right) = U s + \mathcal{O}(\mu^2) \approx Us.
+\end{align}
+
+since under mutation-selection balance $p_l = \nicefrac{\mu}{s}$ and $U$ is
+defined $U := 2 L \mu$, the deleterious mutation rate per diploid genome, per
+generation. They then set $V_A = V_\text{BGS}$, and note that $V_m \approx
+Us^2$, since mutations have frequency of $\nicefrac{1}{2N}$ and each increase
+the variance by a factor of $s^2$. Note that because the classic BGS model
+assumes deterministic selection dynamics (i.e. an infinite population), the
+equilibrium $Z$ given by equation \eqref{eq:Z} excludes the reduction due to
+drift. The reduction factor under the strong BGS model is,
+
+\begin{align}
+  Z_\text{BGS} := 1 - \frac{Vm}{V_\text{BGS}} = 1 - \frac{Us^2}{Us} = 1 - s
+\end{align}
+
+(c.f. \cite{Santiago1998-bs} equation 10). An alternate derivation can be found
+by working out the deterministic reduction in variance from the single-locus
+dynamics.
+
+Now, we can use these values for $Z_\text{BGS}$, $V_m$, and $V$ in equation
+\eqref{eq:Ne_inf}. Through this alternate quantitative genetics derivation, we
+arrive at the same equation as classic BGS theory,
+
+\begin{align}
+  N_{e,\infty}^\text{BGS} = N \exp\left( - \frac{U}{2s + M - Ms} \right).
+\end{align}
+
+This matches equation 8 of \textcite{Hudson1995-xc} except for the $Ms$ term
+and equation 10 of \textcite{Nordborg1996-nq} except for the $2s$ term.
+
+\subsection{Effective Population Size Under Weak and Strong BGS} \label{supp:weak-strong}
+
+With the possibility of weak selection, deleterious alleles can drift up in
+frequency and fix. Consequently, the fitness distribution is no longer
+stationary, as the set of haplotypes without any deleterious mutations can be
+lost due to drift. This changes the shape of the fitness distribution, and thus
+the variance parameter that our model relies upon
+\parencite{Gessler1995-hz,OFallon2010-my,Good2013-lp,Haigh1978-gt,Higgs1995-xc}.
+So far, finding an explicit equation for the variance has been difficult due to
+the ``moment-closure problem": the variance depends on higher moments such as
+the skew, which in turn depends on higher moments. While the weak selection
+regime is composed of an interaction of mutation, selection, drift, and linkage
+processes and the dynamics that differ qualitatively based on parameter
+combinations. In their approximation, \textcite{Santiago2016-mu} consider the
+modified additive genic variance due to the fixation of weakly deleterious
+mutations. This fixation process is related to the Muller's Ratchet, the rate
+$R$ of fixation of deleterious mutations per generation per region, since each
+fixation directly reduces the additive genic variance of the fitness
+distribution. However, in general, the rate of this ratchet is unknown since it
+depends on the fitness variance (and thus higher moments of the fitness
+distribution).
+
+% NOTE: The S&C math using Fisher's FToNS argues fitness decreases
+% at a rate of -s/T each generation. However their model is haploid;
+% in a diploid model, fixation makes everyone  a homozygote with fitness
+% effect -2s; thus the rate must be -2s/T which agrees with sim
+
+In their 2016 paper, Santiago and Caballero derive a general equation for the
+fitness variance that accounts for the reduction due to the fixation of weakly
+deleterious alleles (i.e. the rate of Muller's ratchet). There derivation for
+the variance uses a heuristic approach based on Fisher's Fundamental Theorem of
+Natural Selection (equation 2, \cite{Santiago2016-mu}, equation 15A
+\cite{Garcia-Dorado2007-jj}). A more formal derivation follows from
+\textcite{Higgs1995-xc} (though they assume a haploid asexual system, the end
+result is the same up to a factor of two). In their work, they derive equations
+for the change per generation for the moments and cross-moments of the fitness
+distribution under multiplicative selection in a haploid model. They find, the
+haploid rate of the ratchet is $r = u - s v_n$ (equation 13.3,
+\cite{Higgs1995-xc}), where we use lowercase to distinguish between haploid and
+diploid models. Here, $v_n$ is the variance in the \emph{number} of deleterious
+mutations; as long as selection is not too strong, the haploid additive
+variance is $v_{A} \approx s^2 v_n$. Substituting this for $v_n$ and
+rearranging, the haploid fitness variance is related to the $v_A = us - rs$,
+which is identical to Santiago and Caballero's (\citeyear{Santiago2016-mu})
+equation 2, noting that the waiting time between fixations (i.e. ratchet
+clicks) is $T = \nicefrac{1}{r}$. Intuitively, the fitness variance $Us$ is
+reduced by the rate at which deleterious mutations fix and remove fitness
+variation from the population.
+
+Now, note that under a diploid model, the fitness variance is twice that as the
+haploid model, so $V_A = 2 v_A$. Then, the fitness variance is,
+
+\begin{align}
+  \label{supp-eq:Va}
+  V_{A} = Us - 2Rs. 
+\end{align}
+
+where $U = 2u$ and $R=r$ is the diploid rate of the ratchet, which depends on
+higher moments of the fitness distribution, but only differs from $r$ through a
+rescaling of the population size. Note that if selected sites are sufficiently
+deleterious that they cannot fix (or the population is infinite), $R=0$ and the
+additive fitness variance is identical to that under strong background
+selection, $V_A = Us = V_\text{BGS}$ (equation \eqref{eq:va_bgs}). Then, the
+fixation probability of a new deleterious mutation with additive effects ($1-s$
+in the heterozygote, $1-2s$ in the homozygote) in a diploid population is,
+
+\begin{align}
+  p_\text{fix} = \frac{1-\exp(\nicefrac{2 N_e s}{N})}{1-\exp(4 N_e s)} \approx
+  \frac{2 N_e s}{N \left(\exp(4 N_e s)-1\right)}
+\end{align}
+
+(c.f. \cite{Durrett2008-ql}, equation 7.21; \cite{Kimura1957-rk}, equation
+5.6). Then, Santiago and Caballero argue that the rate of the ratchet is the
+inverse of the averaging waiting time until a fixation, $T \approx 1/(UN
+p_\text{fix})$ (note that since $U$ is the \emph{diploid} mutation rate, the
+total population mutation rate is $UN$ per generation), so
+
+\begin{align}
+  T \approx \frac{\exp(4 N_e s)-1}{2 U s N_e}
+\end{align}
+
+Since the ratchet rate depends on $N_e$, which in turn depends on the rate of
+the ratchet, we write the ratchet, or substitution, rate per basepair per
+generation as,
+
+\begin{align}
+  R(N_e) := \frac{2 U s N_e}{\exp(4 N_e s)-1}.
+\end{align}
+
+Under weak and strong background selection the variance input of mutations
+$V_m$ is the same, and using the $V_A$ derived above, the decay in variance is
+
+\begin{align}
+  Z_\text{WS} = 1 - \frac{V_m}{V_A} = 1 - \frac{Us^2}{(U-R(N_e))s} = 1 - \frac{Us}{U-R(N_e)}.
+\end{align}
+
+At equilibrium, the $N_e$ that determines fixation probability is the
+asymptotic $N_{e,\infty}$. Consequently, Santiago and Caballero argue that the
+asymptotic reduction in effective population size uses the $Q_\infty^2$ and is
+given by the following system of non-linear equations,
+
+\begin{align}
+  R(N_{e,\infty}) &= \frac{4 U s N_{e,\infty}}{\exp(4 N_{e,\infty} s)-1}\\
+  N_{e,\infty} &\approx N \exp\left(-\frac{V_A}{(1-Z) (2-(2-M) Z)} \right)
+\end{align}
+
+The solution to these equations is the asymptotic equilibrium $\widetilde{N}_e$
+and equilibrium ratchet rate $\widetilde{R}$. This sets the equilibrium
+variance, $\widetilde{V} = Us - 2\widetilde{R}s$, or $\widetilde{V} = (U -
+2\widetilde{R})s$. Note that because $p_\text{fix} \le \nicefrac{1}{2N}$ for
+deleterious to neutral variation, $2R \le U$, or the diploid mutation rate must
+be greater than twice the per-basepair per-generation substitution rate for
+equilibrium variance $\widetilde{V} > 0$ for $s>0$.
+
+\subsection{Steady-state Equilibrium Genic Variance} \label{supp:genic-va}
+
+We also find that Equation \eqref{eq:Va} (and Supplementary Materials Equation
+\ref{supp-eq:Va} above) can also be found through Kimura's diffusion models
+with a flux of mutations into infinite, discrete sites
+(\citeyear{Kimura1969-jw}). This model integrates the Kolmogorov backwards
+equation over time and forward variable to find the expected value of a
+functional of the initial frequency. Using the functional $I(p) = s^2 p(1-p)$
+for genic variance given mutations start at frequency $p$, one finds that the
+steady-state genic variance for $p=\nicefrac{1}{2N_e}$ is
+
+\begin{align}
+    V_a^{SS} &= 2 U N s \left(\frac{1}{2N_e} - p_F(N_e, s)\right) \\
+             &= \left(\frac{UN}{N_e} - 2 U Np_F(N_e, s))\right)s \nonumber \\
+             &\approx \left(U - 2 R\right)s \nonumber
+\end{align}
+
+which is approximately Equation \eqref{eq:Va}.
+
+\subsection{Misspecification of the Polygenic Selection Model}
+
+Thus far, we have developed the general model of equation \eqref{XXX} by
+assuming that the only source of $V_h$ is from the deleterious
+mutation-selection-draft-drift process. However, other polygenic evolutionary
+processes can also create, alter, and eliminate fitness variation. By assuming
+that the only fitness variation originates from deleterious
+mutation-selection-drift-draft processes, we introduce some model
+misspecification. There are two main types of misspecification under the
+selection model. First, non-polygenic positive selection processes, such as
+recurrent sweeps, may create small-scale distortions in local genealogies.
+These have been found to only impact diversity on small spatial scales
+\parencite{Murphy2022-sj} in humans, and explain little of the total variation
+in megabase-scale patterns of diversity. Second, there is model
+misspecification due to our model assuming that the only polygenic selection
+process happening is the flux of new deleterious mutations in, and their
+eventual loss or fixation. Other polygenic selection processes could be
+occurring, e.g. traits could be responding to directional selection. The
+Santiago and Caballero model our work extends is very general, and can account
+for multiple types of polygenic selection. The key input parameters are the
+equilibrium heritable fitness variation $V_h$, and the variation introduced by
+mutation each generation $V_m$.
+
+Let us assume for now that,
+
+\begin{align}
+  \label{eq:Ne_inf}
+  N_{e,\infty} \approx N \exp\left(-\frac{V_h + \varepsilon}{(1-Z) (2-(2-M) Z)} \right).
+\end{align}
+
+\subsection{Heterozygosity and $B$ values Under Weak and Strong Selection}
+
+While the selection dynamics are set by the asymptotic $Q_\infty^2$ that
+determines the equilibrium $\widetilde{N}_e$ and $\widetilde{R}$,
+heterozygosity is determined by the sequence $\widetilde{N}_{e,t}$, where each
+term is given by \eqref{eq:net_integral}. We experimented with calculating this
+full sum in our methods, but it seemed to make little difference in our
+application. Additionally, it was too costly to calculate computationally
+for genome-wide calculations of the B' map.
+
+\subsection{Numeric methods for calculating $B$}
+
+Depending on the number of features, this can be incredibly memory intensive.
+
+\subsection{Optimization}
+
+Optimization in non-linear regression is notoriously difficult both in general
+(Bates) and in the specific problem of estimating the effects of genome-wide
+selection \parencite{Murphy2022-sj}. We experimented with a variety of global
+and local optimization approaches and parameterizations, but found that
 
@@
 
-\section*{Additional Results}
-
-\subsection*{Model Comparisons}
+\section{Additional Results}
+
+\subsection{Model Comparisons}
 
@@
 
-\section*{Table of $R^2$ Values}
+\section{Table of $R^2$ Values}
 
@@
 
-\section*{Derivation of Approximate Drift $R^2$}
+\section{Derivation of Approximate Drift $R^2$}
 
@@
 
-\section*{Predicted Rate of Fitness Change}
+\section{Predicted Rate of Fitness Change}
 
@@
 
-\section*{Likelihood}
+\section{Likelihood}
 
@@
 \subsection{Reduction Under the Modified Santiago and Caballero Model}
-
-When weak deleterious mutations can fix, causing substitutions, the classic
-background selection models break down. We find that the generalized background
-selection model of \textcite{Santiago2016-mu} works well (see Section XXX for
-our derivation). Unlike the classic BGS models which only produce B maps, that
-only estimate the reduction in effective population size, The generalized BGS
-model of \textcite{Santiago2016-mu} also predicts the deleterious substitution
-rate per generation per segment (the ``ratchet" rate), which we call the R map.
-We describe how these are calculated substitution rates per segment, for each
-mutation rate and selection coefficient combination in the grids. We describe
-how these are calculated here.
-
-First, the reduction factor experienced by a focal site within the segment $g$
-(and thus is similar to what selected alleles in $g$ experience) under
-generalized background selection is calculated as
-
-\begin{align}
-  \widetilde{b}'(\mu, s, L, r_\text{BP}) = \log\left(\frac{\widetilde{N}_{e,\infty}}{N} \right)
-\end{align}
-
-where $\widetilde{N}_{e,\infty}$ is the asymptotic effective population size,
-which is the solution to the following system of two equations (see XXX for
-derivation),
-
-\begin{align}
-    {N}_{e,\infty} &= N \exp \left( -\frac{V}{2} Q_\infty^2 \right) \label{eqn:Ne_VQ2} \\
-  {R} &= \frac{4 U s N_{e,\infty}}{\exp(4 N_{e,\infty} s)-1} \nonumber \\
-\end{align}
-
-where
-
-\begin{align}
-  V &= U s - 2 {R} s \\
-  Q_\infty^2 &= \frac{2}{(1-Z)(2-(2-M)Z)} \\
-  Z &= 1 - \frac{Us}{U-{R}}
-\end{align}
-
-and $U = 2\mu L$ and $M = r_\text{BP} L$ as above. In this section, we assume
-$\mu$ is the mutation rate to variants with heterozygous selection coefficient
-$s$ to simplify notation; later, $\mu$ refers to the genome-wide average
-mutation rate and the mutation rate is scaled by the DFE. Solving this system
-of two equations for each segment also gives the total substitution or
-``ratchet" rate per generation per segment ($\widetilde{R}$). The per-basepair
-substitution rate of a segment with mutation rate $\mu$ and selection
-coefficient $s$ is $\lambda_d(\mu, s) = \nicefrac{\widetilde{R}}{L}$
 

### 2023-06-23 manuscript/supplementary.tex
ms updates

@@
 due to offspring number (including heritable and non-heritable components), and
-Mendelian segregation.
+Mendelian segregation in generation 1.
 
 Now, let us look the variance of $\var(\Delta x_1)$ over evolutionary
-replicates. Since the Mendelian random process and the offspring random process
+replicates. Since the Mendelian segregation and the offspring random process
 are independent,
@@
 \end{align}
-
+%
 Looking at each term,
@@
 \end{align}
-
+%
 where we ignore the covariance terms due to the sum, since these are on order
@@
 the neutral alleles assort independently into diploids with respect to their
-fitness, so
+fitness, so we can simply the variance as
 
@@
 
-Assuming that no correlation between parental gametes (e.g. no inbreeding),
+Assuming no correlation between parental gametes (e.g. no inbreeding),
 $\var(x_{0,i})$ is the binomial variance in individual allele frequency, or
@@
 individual given by the reproduction process. For example, if the reproduction
-process is a neutral multinomial Wright--Fisher, $\var(k) \approx 2$.
+process is a neutral multinomial Wright--Fisher, $\var(k) \approx 2$. Then, the
+variance in allele frequency change due to non-heritable offspring number
+variation is,
 
@@
 
-\subsection{Misspecification of the Polygenic Selection Model}
-
-Thus far, we have developed the general model of equation \eqref{XXX} by
-assuming that the only source of $V_h$ is from the deleterious
-mutation-selection-draft-drift process. However, other polygenic evolutionary
-processes can also create, alter, and eliminate fitness variation. By assuming
-that the only fitness variation originates from deleterious
-mutation-selection-drift-draft processes, we introduce some model
-misspecification. There are two main types of misspecification under the
-selection model. First, non-polygenic positive selection processes, such as
-recurrent sweeps, may create small-scale distortions in local genealogies.
-These have been found to only impact diversity on small spatial scales
-\parencite{Murphy2022-sj} in humans, and explain little of the total variation
-in megabase-scale patterns of diversity. Second, there is model
-misspecification due to our model assuming that the only polygenic selection
-process happening is the flux of new deleterious mutations in, and their
-eventual loss or fixation. Other polygenic selection processes could be
-occurring, e.g. traits could be responding to directional selection. The
-Santiago and Caballero model our work extends is very general, and can account
-for multiple types of polygenic selection. The key input parameters are the
-equilibrium heritable fitness variation $V_h$, and the variation introduced by
-mutation each generation $V_m$.
-
-Let us assume for now that,
-
-\begin{align}
-  \label{eq:Ne_inf}
-  N_{e,\infty} \approx N \exp\left(-\frac{V_h + \varepsilon}{(1-Z) (2-(2-M) Z)} \right).
-\end{align}
+% \subsection{Misspecification of the Polygenic Selection Model}
+
+% Thus far, we have developed the general model of equation \eqref{XXX} by
+% assuming that the only source of $V_h$ is from the deleterious
+% mutation-selection-draft-drift process. However, other polygenic evolutionary
+% processes can also create, alter, and eliminate fitness variation. By assuming
+% that the only fitness variation originates from deleterious
+% mutation-selection-drift-draft processes, we introduce some model
+% misspecification. There are two main types of misspecification under the
+% selection model. First, non-polygenic positive selection processes, such as
+% recurrent sweeps, may create small-scale distortions in local genealogies.
+% These have been found to only impact diversity on small spatial scales
+% \parencite{Murphy2022-sj} in humans, and explain little of the total variation
+% in megabase-scale patterns of diversity. Second, there is model
+% misspecification due to our model assuming that the only polygenic selection
+% process happening is the flux of new deleterious mutations in, and their
+% eventual loss or fixation. Other polygenic selection processes could be
+% occurring, e.g. traits could be responding to directional selection. The
+% Santiago and Caballero model our work extends is very general, and can account
+% for multiple types of polygenic selection. The key input parameters are the
+% equilibrium heritable fitness variation $V_h$, and the variation introduced by
+% mutation each generation $V_m$.
+
+% Let us assume for now that,
+
+% \begin{align}
+%   \label{eq:Ne_inf}
+%   N_{e,\infty} \approx N \exp\left(-\frac{V_h + \varepsilon}{(1-Z) (2-(2-M) Z)} \right).
+% \end{align}
 

### 2023-06-24 manuscript/manuscript.tex
ms updates

@@
 of all adaptations, but the vast majority of mutations either do not affect
-fitness or are deleterious. Negative selection works to eliminate such
-deleterious mutations from the population, which is essential to maintaining
-the integrity of information in the genome needed to maintain a lineage.
-Mutations under negative selection are kept at low frequencies (Haldane?),
-contributing to human traits (XXX) and diseases. As with most selection in the
-genome, this process perturbs the allele frequencies of neighboring linked
+fitness or are deleterious. Negative selection (i.e. purifying selection) works
+to eliminate such deleterious mutations from the population, which is essential
+to maintaining the integrity of information in the genome needed to maintain a
+lineage. Mutations under negative selection are kept at low frequencies
+(Haldane?), contributing to human traits (XXX) and diseases. Moreover, the
+negative selection process determines sequence conservation across phylogenetic
+timescales, which is evident in genome-wide cross-species comparisons
+\parencite{Siepel2005-wh}. As with most selection in the genome, the negative
+selection process perturbs the allele frequencies of neighboring linked
 variation. This reduces genetic variability around conserved segments and
@@
 PhastCons regions. We include PhastCons regions because there is abundant
-evidence of highly-conserved non-coding regions in humans that we would expect
-to be under negative selection and impact patterns of linked selection (XXX).
-Since our method fits a DFE for each feature type, PhastCons regions that
-overlap other gene annotations (e.g. coding sequences) must be assigned to
-either of the two feature types. Since this can impact estimates, we fit both
-alternatives: a PhastCons Priority model, where features that overlap PhastCons
-regions are classified as PhastCons, and a Feature Priority model, where gene
-structure features take priority and the PhastCons type catches all
-highly-conserved non-genic regions. In total, we fit four annotation models
-(CADD 6\%, CADD 8\%, PhastCons Priority, and Feature Priority) to high-coverage
-1000 Genome data for three populations: Yoruba (YRI), Han Chinese (CHB), and
-European (CEU) ancestries. We assessed and compared models according to how
-well they predict the genome-wide patterns of diversity they were fit on. We
-use $R_\text{LOO}^2$, which is the proportion of the observed variance in
-megabase-scale diversity that is predicted by our model in out-samples (i.e.
-whole chromosomes left out during the fitting process).
+evidence of highly-conserved non-coding regions in humans under negative
+selection
+\parencite{Meader2010-hm,Harmston2013-tt,Katzman2007-gq,Siepel2005-wh}, but
+would not be included in gene-based annotation. Since our method fits a DFE for
+each feature type, PhastCons regions that overlap other gene annotations (e.g.
+coding sequences) must be assigned to either of the two feature types. Since
+this can impact estimates, we fit both alternatives: a PhastCons Priority
+model, where genic features that overlap PhastCons regions are classified as
+PhastCons, and a Feature Priority model, where gene structure features take
+priority and the PhastCons type catches all highly-conserved non-genic regions.
+In total, we fit four annotation models (CADD 6\%, CADD 8\%, PhastCons
+Priority, and Feature Priority) to high-coverage 1000 Genome data for three
+populations: Yoruba (YRI), Han Chinese (CHB), and European (CEU) ancestries. We
+assess and compare our models according to how well they predict patterns of
+diversity on whole chromosomes left-out during the model fitting process. We
+use the metric $R_\text{LOO}^2$, which is the proportion of the observed
+variance in genomic diversity at the megabase scale predicted by our model. We
+experimented with a few smaller spatial scales, but our results were consistent
+with previous results suggesting the human linked selection signal fits best at
+the megabase scale \parencite{Murphy2022-sj}.
 
@@
 
-\begin{figure}[htbp] \centering
-    \includegraphics{figures/figure_4_pop_mu_trimmed.pdf} 
+\begin{figure}[htbp] 
+    \centering
+    \includegraphics[width=\textwidth]{figures/figure_4.pdf} 
     \caption{}
@@
 \begin{figure}[htbp] \centering
-    \includegraphics[width=\textwidth]{figures/figure_4.pdf} 
+    \includegraphics[width=\textwidth]{figures/figure_5.pdf} 
     \caption{}

### 2023-06-26 manuscript/supplementary.tex
ms updates

@@
 
-\subsection{YRI Samples}
+\subsection{}
 
@@
 
-\subsection{Site-based Filtering of Counts}
+\subsection{Non-accessible and Masked Regions}
+\label{supp:accessible}
 
@@
 \begin{enumerate}
+
   \item \textbf{Non-accessible regions}: masks out centromeres (\texttt{acen} entries in
@@
       determined by Ensembl (XXX), which uses Repeat Masker (XXX) and Dust
-      (XXX).
-
-  \item \textbf{Non-``putatively" neutral regions}: Additionally, for fitting our
-    likelihood and estimating observed pairwise diversity, we only consider .
-    This masks out phastCons regions (from
-    \path{phastConsElements100way.bed.gz}) and Ensembl gene regions (from
-    annotation file \path{Homo_sapiens.GRCh38.107.chr.gff3.gz}).  While introns
-    are possibly under some weak selection, they collectively make up nearly
-    40\% of the human genome and are included so genome-wide diversity can be
-    estimated more precisely (possibly at the expense of some bias).
+      (XXX). Additionally, we used the 1000 Genomes strict filter (XXX).
+
 \end{enumerate}
 
-These files are all produced by the Snakemake file \path{data/annotation/Snakefile}.
-Note that the levels of accessible putatively neutral bases
+\subsection{Putatively Neutral Tracks}
 
@@
   \centering
-  \includegraphics{figures/supplementary/chr1_diversity_filtering.pdf}
+  \includegraphics[width=\textwidth]{figures/supplementary/neutral_tracks.pdf}
 
   \caption{Estimates of chromosome 1 YRI diversity in non-overlapping megabase
-    windows, under different filtering criteria. The filtering criteria are,
-    (1) ``neutral, accessible" which includes only putatively neutral sites, and
-    ignores regions masked as inaccessible, (2) ``accessible" which is only ignores
-  sites masked as inaccessible, and (3) ``no filtering" which uses all available
-data. Note that filtering changes the absolute level of diversity, but has more
-minor effects on regional patterns of diversity at the megabase scale.}
-
-  \label{suppfig:chr1-diversity}
+      windows, under different neutral tracks (colors) and the filtering
+      criteria described in Supplementary Section \ref{supp:accessible}. To the
+      right of the chromosome subfigure are the genome-wide boxplots of
+      megabase-scale diversity. }
+
+  \label{suppfig:neutral-tracks}
 \end{figure}
+
+Linked selection theory typically assumes that focal sites are not under
+purifying selection themselves. Consequently, linked selection statistical
+methods often only consider the fraction of the genome not \emph{a priori}
+under selective constraint \parencite[e.g. Appendix Section
+3.1]{Murphy2022-sj}. Our methods follow suit, and first create a track
+consisting of the union (i.e. \texttt{bedtools merge}) of: exons and coding
+sequences from Ensembl Version 107
+\parencite[\path{Homo_sapiens.GRCh38.107.chr.gff3.gz}]{Cunningham2022-vk} and
+PhastCons regions
+\parencite[\path{phastConsElements100way.bed.gz}]{Siepel2005-wh}. Then, we
+remove 200 basepairs of ``slop" around each of these regions (i.e.
+\texttt{bedtools slop}). We show in Supplementary Figure
+\ref{suppfig:neutral-tracks} that average diversity (we only consider YRI
+samples for this analysis) is fairly insensitive to the slop width, but
+increasing slop widths around putatively conserved segments does lead to higher
+diversity estimates. This is expected, since increasing the slop width
+considers a smaller subset of sites, further away from putatively conserved
+regions. For our analyses we choose the 200bp slop version, which covers
+63.66\% of the autosomal reference sequence.
+
+Additionally, we compared our neutral track (and the slop width alternatives)
+against the bottom 70\% of CADD site scores (this works out to be 64.70\% of
+the autosomal reference sequence). The correlation between these two tracks is
+extraordinarily high (Pearson $r = 0.98$, p-value $=0$), but the average
+pairwise diversity is 13\% higher (two-tailed t-test p-value $1.29 \times
+10^{-71}$) in the CADD 70\% neutral track. However, we did not use the CADD
+70\%  neutral track in our analysis of human population genomic data. Our
+concern was that this neutral track would be too close to the complement of the
+CADD 6\% track, which we use as one of our annotation models.
+
+These files are all produced by the Snakemake file
+\path{data/fit_annotation/Snakefile} in this study's GitHub repository
+(\url{http://github.com/vsbuffalo/bprime}).
 
@@
 \centering
+\caption{XXX}
 \begin{tabular}{lrrr}

### 2023-06-27 manuscript/supplementary.tex
ms updates

@@
 \RequirePackage{bbm}
+\RequirePackage{xr}
 %\RequirePackage[osf]{mathpazo}
@@
 
+% for cross references
+\externaldocument{manuscript}
+
 % line numbers:
@@
 \newcommand{\V}{\text{V}}
-\DeclareMathOperator{\var}{var}
+\DeclareMathOperator{\var}{V}
+\newcommand{\V}{\text{V}}
 \DeclareMathOperator{\cov}{cov}
@@
 
-\section{Theory}
-\subsection{Overview of Quantitative Genetic Models of Effective Population Size}
+\section{Quantitative Genetic Linked Selection Model Theory}
 
@@
 The frequency in the initial generation is $x_0 = \nicefrac{1}{N} \sum_{i=1}^N
-x_{0,i}$, though to indicate that we treat this as fixed, we use $p_0 = x_0$.
-Then, the allele frequency change is,
-
-\begin{align}
-  \Delta x_1 = x_1 - p_0 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (k_{0,i} - 1) + \frac{1}{2N} \sum_{i=1}^N \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \\
+x_{0,i}$, though to indicate that we treat this as fixed rather than random, we
+use $p_0 := x_0$. Then, the allele frequency change is,
+
+\begin{align}
+  \Delta x_1 = x_1 - p_0 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (k_{0,i} - 1) + \frac{1}{2N} \sum_{i=1}^N \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \nonumber \\
   \Delta x_1 &= K_1 + H_1
@@
 
-Now, let us look the variance of $\var(\Delta x_1)$ over evolutionary
+Now, let us look the variance of $\V(\Delta x_1)$ over evolutionary
 replicates. Since the Mendelian segregation and the offspring random process
@@
 \begin{align}
-  \var(\Delta x_1) &= \var(K_1) + \var(H_1).
+  \V(\Delta x_1) &= \V(K_1) + \V(H_1).
 \end{align}
@@
 \begin{align}
-  \var(K_1) &\approx \frac{1}{4N^2} \sum_{i=1}^N \var\left(x_{0,i} (k_{0,i} - 1) \right)
+  \V(K_1) &\approx \frac{1}{4N^2} \sum_{i=1}^N \V\left(x_{0,i} (k_{0,i} - 1) \right)
 \end{align}
@@
 \begin{align}
-  \var(K_1) &\approx \frac{1}{4N^2} \sum_{i=1}^N \var(x_{0,i}) \var(k_{0,i} - 1)  \\
-            &\approx \frac{1}{4N} \var(x_{0,i}) \var(k_{0,i}).
+  \V(K_1) &\approx \frac{1}{4N^2} \sum_{i=1}^N \V(x_{0,i}) \V(k_{0,i} - 1) \nonumber  \\
+            &\approx \frac{1}{4N} \V(x_{0,i}) \V(k_{0,i}).
 \end{align}
@@
 Assuming no correlation between parental gametes (e.g. no inbreeding),
-$\var(x_{0,i})$ is the binomial variance in individual allele frequency, or
-$p_0(1-p_0)/2$, and $\var(k) := \var(k_{0,i})$ is the offspring variance of an
+$\V(x_{0,i})$ is the binomial variance in individual allele frequency, or
+$p_0(1-p_0)/2$, and $\V(k) := \V(k_{0,i})$ is the offspring variance of an
 individual given by the reproduction process. For example, if the reproduction
-process is a neutral multinomial Wright--Fisher, $\var(k) \approx 2$. Then, the
+process is a neutral multinomial Wright--Fisher, $\V(k) \approx 2$. Then, the
 variance in allele frequency change due to non-heritable offspring number
@@
 \begin{align}
-  \var(K_1) &\approx \frac{p_0(1-p_0)}{2N} \frac{\var(k)}{4}.
+  \V(K_1) &\approx \frac{p_0(1-p_0)}{2N} \frac{\V(k)}{4}.
 \end{align}
@@
 their offspring can be replaced by a summation over offspring, since both sum
-$N$ exchangeable terms. Then, note that $\var(\delta_{i,j}) = \nicefrac{1}{4}
+$N$ exchangeable terms. Then, note that $\V(\delta_{i,j}) = \nicefrac{1}{4}
 p_0(1-p_0)$, so
@@
 \begin{align}
-  \var(H_1) &= \frac{p_0(1-p_0)}{2N} \frac{1}{2}.
+  \V(H_1) &= \frac{p_0(1-p_0)}{2N} \frac{1}{2}.
 \end{align}
@@
 \begin{align}
-  \var(\Delta x_1) &= \var(K_1) + \var(H_1) \\
-  &\approx \frac{p_0(1-p_0)}{2N} \frac{\var(k)}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2} \\
-                   &\approx \frac{p_0(1-p_0)}{2N}\left(\frac{\var(k)}{4} + \frac{1}{2}\right)
-\end{align}
-
+  \V(\Delta x_1) &= \V(K_1) + \V(H_1) \\
+  &\approx \frac{p_0(1-p_0)}{2N} \frac{\V(k)}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2} \nonumber \\
+                   &\approx \frac{p_0(1-p_0)}{2N}\left(\frac{\V(k)}{4} + \frac{1}{2}\right)
+\end{align}
+%
 (c.f. \cite{Santiago1995-hx} equation 2 and \cite{Buffalo2019-qs} equation 30).
 Note that if the reproduction process is a neutral multinomial Wright--Fisher,
-$\var(k) \approx 2$, and this simplifies to the expected Wright--Fisher
-variance. If we were to \emph{define} a variance effective population size by
-setting this variance in neutral allele frequency change to the expected
-variance under a Wright--Fisher model ($V_\text{WF}$), we'd have
-
-\begin{align}
-  V_\text{WF} &:= \frac{p_0(1-p_0)}{2N_e} \\
-  V_\text{WF} &= \var(\Delta p_1) \\
-  N_e &= \frac{4N}{\var(k) + 2} \\
-\end{align}
-
-(c.f. \cite{Wright1938-tv}).
-
-Now, we look at this variance equation behaves with $t > 1$.
-
-\begin{align}
-  \V(x_1) = \E_1\left[(x_1 - x_0)^2\right]= \frac{x_0(1-x_0)}{2N},
-\end{align}
-
-and
+$\V(k) \approx 2$, and this simplifies to the expected Wright--Fisher
+variance. If we were to define a variance \emph{drift-effective} population
+size $N_e$ by setting this variance in neutral allele frequency change to the
+expected variance under a Wright--Fisher model ($V_\text{WF}$), we'd have
+
+\begin{align}
+  V_\text{WF} &:= \frac{p_0(1-p_0)}{2N_e} \nonumber \\
+  V_\text{WF} &= \V(\Delta p_1) \nonumber \\
+  N_e &= \frac{4N}{\V(k) + 2}
+\end{align}
+%
+(c.f. \cite{Wright1938-tv}). Now, we look at the variance in allele frequency
+change across two generations for a neutral allele. The change in neutral
+allele frequency, regardless of the selective system, is directionless in
+expectation. This is because whichever allele we track is arbitrary by the
+definition of neutrality, which imposes a symmetry, $\E(\Delta x) = 0$. Then, the
+variance of change in the first generation is,
+
+\begin{align}
+  \V(\Delta x_1) = \E_1\left[(x_1 - x_0)^2\right]= \frac{x_0(1-x_0)}{2N},
+\end{align}
+%
+and across two generations,
 
@@
   \label{eq:var_wf}
-  \V(x_2) &= \E\left[(x_2 - x_0)^2\right] \\
-          &= \E_1\left[\E_2\left[((x_2-x_1) + (x_1 - x_0))^2 | x_1\right]\right] \\
-          &= \E_1\left[\E_2\left[(\Delta x_2 + \Delta x_1)^2 | x_1\right]\right] \\
-          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + 2\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
-          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] + \mathcal{C}_{1,2} \\
-          &= \frac{p_0(1-p_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{p_0(1-p_0)}{2N} + \mathcal{C}_{1,2}
+  \V(x_2 - x_0) &= \E\left[(x_2 - x_0)^2\right] \nonumber \\
+          &= \E_1\left[\E_2\left[((x_2-x_1) + (x_1 - x_0))^2 | x_1\right]\right] \nonumber \\
+          &= \E_1\left[\E_2\left[(\Delta x_2 + \Delta x_1)^2 | x_1\right]\right] \nonumber \\
+          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] + 2\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] \nonumber \\
+          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] + 2\mathcal{C}_{1,2} \nonumber \\
+          &= \frac{p_0(1-p_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{p_0(1-p_0)}{2N} + 2\mathcal{C}_{1,2}
 \end{align}
@@
 Consequently, the variance of the neutral allele frequency changes each
-generation according to the probability of failing to coalesce each generation
-and the pairwise covariance terms $\mathcal{C}_{i,j}$ that build up due to associations
-with heritable fitness backgrounds. Note that the covariance term $\mathcal{C}_{1,2} =
-\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] = 0$ under neutral
-evolution, and this simplifies to the well-known equation for variance in a
-Wright--Fisher population.
+generation according to the probability of failing to coalescence each
+generation and the pairwise covariance terms $\mathcal{C}_{i,j}$ that build up
+due to associations with heritable fitness backgrounds. Note that the
+covariance term $\mathcal{C}_{1,2} = \E_1\left[\E_2[\Delta x_2 \Delta x_1 |
+x_1]\right] = 0$ under neutral evolution, and this simplifies to the well-known
+equation for variance in a Wright--Fisher population.
 
@@
 variation, which leads to non-zero covariance terms $\mathcal{C}_{i,j}$. These
-terms emerge when $\var(k)$ has a heritable component of fitness that can be
+terms emerge when $\V(k)$ has a heritable component of fitness that can be
 transmitted along with the neutral allele, thereby affecting the neutral
 allele's trajectory in later generations. If we partition the offspring number
-into heritable and non-heritable components, $k_i = 2f_i + \varepsilon_i$, then
-$\var(k_i) = 4 V_h + V_n$. When there is heritable fitness variation across
+into heritable and non-heritable components, $k_i = 2f_i + \Vepsilon_i$, then
+$\V(k_i) = 4 V_h + V_n$. When there is heritable fitness variation across
 individuals, $V_h > 0$. Because the population size is assumed to be constant,
@@
 \begin{align}
-  \var(\Delta x_1) &= \var(S_1) + \var(D_1) + \var(H_1) \\
-  &\approx \frac{p_0(1-p_0)}{2N} V_h + \frac{p_0(1-p_0)}{2N} \frac{V_n}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2}. \\
-\end{align}
-
+  \V(\Delta x_1) &= \V(S_1) + \V(D_1) + \V(H_1) \nonumber \\
+                   &\approx \frac{p_0(1-p_0)}{2N} V_h + \frac{p_0(1-p_0)}{2N} \frac{V_n}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2} \label{supp-eq:varterms} 
+\end{align}
+%
 (c.f. \cite{Santiago1995-hx} equation 11).
 
-To see how the covariance terms accumulate, consider $\var(x_3 - p_0)$. Note
+To see how the covariance terms accumulate, consider $\V(x_3 - p_0)$. Note
 that $\E(S_t) = \E(D_t) = \E(H_3) = 0$, since which neutral allele we track is
@@
   \label{eq:var_x}
-  \var(x_3 - p_0) &= \E\left[ \left(S_1 + D_1 + H_1 + S_2 + D_2 + H_2 + S_3 + D_3 + H_3 \right)^2 \right] \\
-                  &= \E({S_1}^2) + \E({S_2}^2) + \E({S_3}^2)  \\
-                  &+ \E(S_1 S_2) + \E(S_1 S_3) + \E(S_2 S_3) \\
-                  &+ \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) + \E({D_3}^2) + \E({H_3}^2)
-\end{align}
-
+  \V(x_3 - p_0) &= \E\left[ \left(S_1 + D_1 + H_1 + S_2 + D_2 + H_2 + S_3 + D_3 + H_3 \right)^2 \right] \nonumber \\
+                  &= \E({S_1}^2) + \E({S_2}^2) + \E({S_3}^2) \nonumber \\
+                  &+ \E(S_1 S_2) + \E(S_1 S_3) + \E(S_2 S_3)\nonumber \\
+                  &+ \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) + \E({D_3}^2) + \E({H_3}^2).
+\end{align}
+%
 The covariance terms $\E(S_i S_j)$ for $j > i$ represent the expected neutral
@@
 \begin{align}
-  S_1 = \frac{1}{2N} \sum_{i=1}^N x_{0,i}(f_{0,i} - 1) = \cov(x_{0,i}, f_{0,i}) \\
+  S_1 = \frac{1}{2N} \sum_{i=1}^N x_{0,i}(f_{0,i} - 1) = \cov(x_{0,i}, f_{0,i}) \nonumber \\
   S_2 = \frac{1}{2N} \sum_{i=1}^N x_{1,i}(f_{1,i} - 1) = \cov(x_{1,i}, f_{0,i})
 \end{align}
-
+%
 which are chance covariances (across the population, \emph{not} evolutionary
@@
 \begin{align}
-  S_1 = \cov(x_{0,i}, f_{0,i}) &= \cov\left(\frac{x_{0,i}' + x_{0,i}''}{2}, f_{0,i}' + f_{0,i}''\right)  \\
-                               &= \frac{1}{2}\left(\cov(x_{0,i}', f_{0,i}' + f_{0,i}'') + \cov(x_{0,i}'', f_{0,i}' + f_{0,i}'')\right)  \\
-                               &= \cov(x_{0,i}', f_{0,i}' + f_{0,i}'')  \\
-                               &= \cov(x_{0,i}', f_{0,i}') + \cov(x_{0,i}', f_{0,i}'')  \\
+  S_1 = \cov(x_{0,i}, f_{0,i}) &= \cov\left(\frac{x_{0,i}' + x_{0,i}''}{2}, f_{0,i}' + f_{0,i}''\right) \nonumber \\
+                               &= \frac{1}{2}\left(\cov(x_{0,i}', f_{0,i}' + f_{0,i}'') + \cov(x_{0,i}'', f_{0,i}' + f_{0,i}'')\right) \nonumber  \\
+                               &= \cov(x_{0,i}', f_{0,i}' + f_{0,i}'') \nonumber \\
+                               &= \cov(x_{0,i}', f_{0,i}') + \cov(x_{0,i}', f_{0,i}'') \nonumber \\
                                &= S_1' + S_1''
 \end{align}
-
-which follows from symmetry, since it does not matter which gamete carrying the
-neutral allele we track (c.f. \cite{Santiago1998-bs} p. 2107). The primes now
-indicate whether the selected locus is on the same gamete as the neutral
-allele (single prime), or the homologous gamete (double prime).
+%
+which follows from symmetry again, since it does not matter which gamete
+carrying the neutral allele we track (c.f. \cite{Santiago1998-bs} p. 2107). The
+primes now indicate whether the selected locus is on the same gamete as the
+neutral allele (single prime), or the homologous gamete (double prime).
 
@@
 \begin{align}
-  S_{2}' &= S_1 (1-\kappa)(1-r) \\
-  S_{3}' &= S_1 (1-\kappa)^2(1-r)^2 \\
+  S_{2}' &= S_1 (1-\kappa)(1-r) \nonumber \\
+  S_{3}' &= S_1 (1-\kappa)^2(1-r)^2 \nonumber \\
   S_{4}' &= S_1 (1-\kappa)^3(1-r)^3.
 \end{align}
-
+%
 Then, the cumulative effect of the fitness associations created in generation
@@
   \label{eq:qterm}
-  S_1 Q_t &= S_1 + S_2' + S_3' + \ldots + S_t' \\
+  S_1 Q_t &= S_1 + S_2' + S_3' + \ldots + S_t' \nonumber \\
   Q_t &= 1 + \sum_{i=1}^t (1-r)^i(1-\kappa)^i
 \end{align}
-
+%
 or for continuous $t$, 
-
+%
 \begin{align}
@@
 \begin{align}
-  \E(S_1 S_2) &= \E(S_1^2) (1-\kappa)(1-r) = \E(S_1^2) Q_2^2 \\
+  \E(S_1 S_2) &= \E(S_1^2) (1-\kappa)(1-r) = \E(S_1^2) Q_2^2 \nonumber \\
   \E(S_1 S_3) &= \E(S_1^2) (1-\kappa)^2(1-r)^2 = \E(S_1^2) Q_3^2.
 \end{align}
-
-With these covariance terms, we now have a full model for $\var(p_t - p_0)$. We
-will look at the case for $t=2$. As shown in equation XXX, the second moment of
-every term is a function of the variance in neutral allele frequency in that
-generation, $x_t(1-x_t)$. This variance does not stay its initial value of
-$x_0(1-x_0)$; it decays each generation due to coalesce from the drift and
-selection processes modeled by this approach. The rate at which this decay
-happens is the effective population size each generation implied by this model. 
-We can write,
+%
+The $Q_t^2$ terms represent the inflation of variance due to autocovariance in
+frequency change across generations due to weak draft.
+
+With these covariance terms, we now have a full model for $\V(p_t - p_0)$. We
+will look at the case for $t=2$. As shown in equation \eqref{supp-eq:varterms},
+the second moment of every term is a function of the variance in neutral allele
+frequency in that generation, $x_t(1-x_t)$. This variance does not stay its
+initial value of $x_0(1-x_0)$; it decays each generation due to coalescence
+from the drift and selection processes modeled by this approach. The rate at
+which this decay happens is the effective population size each generation
+implied by this model. We can write,
 
@@
     \label{eq:vardecay}
-  \var(x_2 - p_0) &= \E({S_1}^2) + \E(S_1 S_2) + \E({S_2}^2) \\
-                  & + \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) \\
-  \frac{\var(x_2 - p_0)}{p_0(1-p_0)} &= \frac{V_h}{2N_{e,1}} + \frac{V_h Q_2^2}{2N_{e,2}}\left(1-\frac{1}{2N_{e,1}}\right)  +  \frac{V_h}{2N_{e,2}}  \left(1-\frac{1}{2N_{e,1}}\right)  \\ 
+  \V(x_2 - p_0) &= \E({S_1}^2) + \E(S_1 S_2) + \E({S_2}^2) \nonumber \\
+                  & + \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) \nonumber \\
+  \frac{\V(x_2 - p_0)}{p_0(1-p_0)} &= \frac{V_h}{2N_{e,1}} + \frac{V_h Q_2^2}{2N_{e,2}}\left(1-\frac{1}{2N_{e,1}}\right)  +  \frac{V_h}{2N_{e,2}}  \left(1-\frac{1}{2N_{e,1}}\right)  \nonumber \\ 
                                      &+ \frac{V_n}{8N_{e,1}} + \frac{1}{4N_{e,1}} + \frac{V_n}{8N_{e,2}}  \left(1 - \frac{1}{2N_{e,1}}\right) 
@@
 \end{align}
-
-In general, the evolution of the variance in allele frequency change in a
+%
+In general, the variance in allele frequency change after $t$ generations in a
 system with heritable fitness due to a single locus $r$ recombination fraction
@@
   \label{eqn:var_xt}
-  \frac{\var(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^\infty \left(\frac{V_h Q_i^2}{2 N_{e,i}} + \frac{V_n}{8N_{e,i}} + \frac{1}{4N_{e,i}} \right)\prod_{t=1}^{i-1} \left(1-\frac{1}{2 N_{e,t}}\right). \\
-\end{align}
-
-Since the probability of coalesce (or, equivalently, identity by descent) is
-proportional to the variance in allele frequency change
-\parencite{Barton2000-zg}, this encodes the pairwise coalescent rate of the
-population through time under selection through the $N_{e,t}$ terms. Others
-have found that the genealogies under purifying selection can be characterized
-by such a time-dependent effective population size
-\parencite{Nicolaisen2013-gv}. In the original derivation of
-\textcite{Santiago1995-hx}, they set each $N_{e,t}$ term to a single $N$ and
-solve this summation in the limit to infinity giving,
+  \frac{\V(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^t \left(\frac{V_h Q_i^2}{2 N_{e,i}} + \frac{V_n}{8N_{e,i}} + \frac{1}{4N_{e,i}} \right)\prod_{j=1}^{i-1} \left(1-\frac{1}{2 N_{e,j}}\right).
+\end{align}
+
+We will see later that with the correct form for $Q_t^2$, this equation holds
+for polygenic selection (see Supplementary Section \ref{supp:ne-polygenic}).
+Additionally, we note that since the probability of coalescence (or,
+equivalently, identity by descent) is proportional to the variance in allele
+frequency change \parencite{Barton2000-zg}, this equation for the variance
+encodes the pairwise coalescent rate of the population through time under weak
+draft in the $N_{e,t}$ terms. Others have found that the genealogies under
+purifying selection can be characterized by such a time-dependent effective
+population size \parencite{Nicolaisen2013-gv}.
+
+Now, we consider how to turn this expression for the variance in neutral allele
+frequency at time $t$ under weak draft into a \emph{draft-effective} population
+size. In an ideal Wright--Fisher population, the variance after $t$ generations
+is
+
+\begin{align}
+    \label{supp-eq:drift-vt}
+    \V(x_t) &= p_0(1-p_0) \left[1 - \left(1 - \frac{1}{2N_e}\right)^t \right].
+\end{align}
+
+To find an analogous expression for the asymptotic \emph{draft-effective}
+population size $N_d$, we let $N_e = N_d$ in the above equation, and consider
+the change in variance between generations $t$ and $t-1$, $\V(x_t) -
+\V(x_{t-1})$. Using Supplementary Equation \eqref{supp-eq:drift-vt}, we have,
+
+\begin{align}
+    \V(x_t) - \V(x_{t-1}) = p_0(1-p_0) \left[ \left(1 - \frac{1}{2N_e}\right)^{t-1} - \left(1 - \frac{1}{2N_e}\right)^t  \right]
+\end{align}
+
+\textcite{Santiago1995-hx} assume the change in population size per generation
+is small enough that the coalescence term each generation is constant, i.e. $(1
+- \nicefrac{1}{2N_{e,t}}) \approx (1 - \nicefrac{1}{2N_e})$ for each $t$.
+
+\begin{align}
+    \frac{\V(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^t \left(\frac{V_h Q_i^2}{2 N_e} + \frac{V_n}{8N_e} + \frac{1}{4N_e} \right)\left(1-\frac{1}{2 N_{e}}\right)^{i-1}.
+\end{align}
+
+One
+
+Then, imagine the process burns in for $t$ generations and asking what the
+equivalent
+
+\begin{align}
+    \V(x_t) &= \V(x_{t-1})
+
+  p_0(1-p_0) \left[1 - \left(1 - \frac{1}{2N_d}\right)^t \right].
+\end{align}
+
+\begin{align}
+  \label{eqn:ne_var}
+  N_{d,t} = \frac{p_0(1-p_0) - \V(x_{t-1})}{2(\V(x_t) - \V(x_{t-1}))}
+\end{align}
+%
+(cf. p. 1018, \cite{Santiago1995-hx}).
+
+XXX here
+
+\textcite{Santiago1995-hx} assume the change in population size per generation
+is small enough that the coalescence term each generation is constant, i.e. $(1
+- \nicefrac{1}{2N_{e,t}}) \approx (1 - \nicefrac{1}{2N_e})$ for each $t$.
+
+\begin{align}
+    \frac{\V(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^t \left(\frac{V_h Q_i^2}{2 N_{e,i}} + \frac{V_n}{8N_{e,i}} + \frac{1}{4N_{e,i}} \right)\left(1-\frac{1}{2 N_{e}}\right)^{i-1}.
+\end{align}
 
@@
   \label{eq:var_asymp}
-  \frac{\var(x_t - p_0)}{p_0(1-p_0)} &= \frac{4 V_h Q_\infty^2 +V_n+2}{4 (2 N-1)}.
-\end{align}
-
-To relate this to an asymptotic effective population size, note that in an
-ideal Wright--Fisher population,
-
-\begin{align}
-  \V(x_t) &= p_0(1-p_0) \left[1 - \left(1 - \frac{1}{2N_e}\right)^t \right]
-\end{align}
-
-such that if we look at the change in the variance in allele frequency change
-between adjacent generations $t$ and $t-1$, we get an implied $N_{e,t}$
-
-\begin{align}
-  \label{eqn:ne_var}
-  N_{e,t} = \frac{p_0(1-p_0) - \var(x_{t-1})}{2(\var(x_t) - \var(x_{t-1}))}
-\end{align}
-%
-(cf. p. 1018, \cite{Santiago1995-hx}).
-
+  \frac{\V(x_t - p_0)}{p_0(1-p_0)} &= \frac{4 V_h Q_\infty^2 +V_n+2}{4 (2 N-1)}.
+\end{align}
+
+Note here, to differentiate from the drift-effective population size $N_e$ due
+under non-selective causes, we use $N_d$ as the \emph{draft-effective}
+population size. Then, if we look at the change in the variance in allele
+frequency change between adjacent generations $t$ and $t-1$, we get an implied
+$N_{d,t}$
+
+---
 Solving this for equation \eqref{eq:var_asymp} gives the asymptotic effective
@@
   %N_e &= \frac{2 (2 N-1)}{4 V_h Q_\infty^2 +V_n+2} \\
-  N_e &\approx \frac{4N}{4 V_h Q_\infty^2 +V_n+2}
-\end{align}
-
-(cf. equation 18, \cite{Santiago1995-hx}). Levels of pairwise diversity,
-however, depend on the full distribution of pairwise coalesce times, or the
-$N_{e,t}$. While this would require solving the recursion defined by equations
-\eqref{eqn:var_xt} and \eqref{eqn:ne_var}, \textcite{Santiago1998-bs} show that
-a reasonably good approximation follows from using a single effective
-population size for all timepoints, $N_{e,t} = N_e$, but use the time-indexed
-$Q_t$. This leads to,
+  N_d &\approx \frac{4N}{4 V_h Q_\infty^2 +V_n+2}
+\end{align}
+%
+(cf. equation 18, \cite{Santiago1995-hx}). Note if $V_h = 0$, this equation
+reduces to Wright's (\citeyear{Wright1938-tv}) equation.
+
+Levels of pairwise diversity, however, depend on the full distribution of
+pairwise coalescence times, or the $N_{d,t}$. While this would require solving the
+recursion defined by equations \eqref{eqn:var_xt} and \eqref{eqn:ne_var},
+\textcite{Santiago1998-bs} show that a reasonably good approximation follows
+from using a single effective population size for all timepoints, $N_{d,t} =
+N_d$, but use the time-indexed $Q_t$. This leads to,
 
@@
   \label{eq:ne_t}
-  N_{e,t} &\approx \frac{4N}{4 V_h Q_t^2 +V_n+2}
-\end{align}
-
+  N_{d,t} &\approx \frac{4N}{4 V_h Q_t^2 +V_n+2}
+\end{align}
+%
 (c.f. equation 15 of \cite{Santiago1995-hx} and p. 1270 of \cite{Santiago2016-mu}).
@@
 \subsection{Effective Population Size Under Polygenic Selection}
+\label{supp:ne-polygenic}
 
 The results of the previous section assume only a single locus a recombination
-fraction $r$ apart from the neutral locus is determining the coalesce rate. The
+fraction $r$ apart from the neutral locus is determining the coalescence rate. The
 model of \textcite{Santiago1998-bs} extends this to the case where fitness is
@@
 \end{align}
-
+%
 where $Q_{t,l}$ is the cumulative impact of selection due to the selected site
@@
 \begin{align}
-  N_{e,t} &\approx \frac{N}{V_h Q_t^2 + 1} \\
-          &\approx \frac{N}{\prod_{l=1}^n \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)} \\
+  N_{d,t} &\approx \frac{N}{V_h Q_t^2 + 1} \nonumber \\
+          &\approx \frac{N}{\prod_{l=1}^n \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)} \nonumber \\
           &\approx N \exp\left(- \sum_{l=1}^n \frac{V_{h,l} Q_{t,l}^2}{2}\right)
 \end{align}
-
+%
 (c.f. \cite{Santiago1998-bs} equation 4b). If we write each site $l$'s
 contribution to the variance as a deviation from the average contribution,
-$V_{h,l}^2 = \nicefrac{1}{n} V_h + \varepsilon_l$, then
-
-\begin{align}
-  N_{e,t} &\approx N \exp\left(- \frac{1}{2}\sum_{l=1}^n \left(\frac{1}{n}V_{h} + \varepsilon_l\right)  Q_{t,l}^2\right) \\
-          &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{t,l}^2+ \sum_{l=1}^n \varepsilon_l  Q_{t,l}^2\right)
-\end{align}
-
+$V_{h,l}^2 = \nicefrac{1}{n} V_h + \Vepsilon_l$, then
+
+\begin{align}
+  N_{d,t} &\approx N \exp\left(- \frac{1}{2}\sum_{l=1}^n \left(\frac{1}{n}V_{h} + \Vepsilon_l\right)  Q_{t,l}^2\right) \nonumber \\
+          &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{t,l}^2+ \sum_{l=1}^n \Vepsilon_l  Q_{t,l}^2\right)
+\end{align}
+%
 Santiago and Caballero assume that over evolutionary replicates, as selected
@@
 fitness variation at a site and the decay of its association with the neutral
-site, e.g. $\E(\varepsilon_l Q_{t,l}^2) = 0$, and thus ignore the second sum
-term.
-
-\begin{align}
-      N_{e,t} &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{t,l}^2\right) \\
+site, e.g. $\E(\Vepsilon_l Q_{t,l}^2) = 0$, and thus ignore the second sum
+term, leading to
+
+\begin{align}
+      N_{d,t} &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{t,l}^2\right).
 \end{align}
@@
 \begin{align}
-      N_{e,\infty} &\approx N \exp\left(- \frac{V_{h}}{2L} \sum_{l=1}^L Q_{\infty,l}^2\right) \\
+      N_{d,\infty} &\approx N \exp\left(- \frac{V_{h}}{2L} \sum_{l=1}^L Q_{\infty,l}^2\right) \\
               &\approx N \exp\left(- \frac{V_{h}}{M} \int_0^{\nicefrac{M}{2}} Q_{\infty}(r)^2 dr \right) \\
@@
 \begin{align}
-  N_{e,\infty}^{\text(SC98)} \approx N \exp\left(-\frac{V_h}{(1-Z)M}\right) 
+  N_{d,\infty}^{\text(SC98)} \approx N \exp\left(-\frac{V_h}{(1-Z)M}\right) 
 \end{align}
@@
   \label{eq:Ne_inf}
-  N_{e,\infty} \approx N \exp\left(-\frac{V_h}{(1-Z) (2-(2-M) Z)} \right).
-\end{align}
-
-For the reasons described earlier, a single asymptotic $N_{e,\infty}$ cannot
+  N_{d,\infty} \approx N \exp\left(-\frac{V_h}{(1-Z) (2-(2-M) Z)} \right).
+\end{align}
+
+For the reasons described earlier, a single asymptotic $N_{d,\infty}$ cannot
 fully characterize pairwise coalescent rates in all cases, and thus neutral
-diversity levels depend on the full $N_{e,t}$, for which we integrate equation
+diversity levels depend on the full $N_{d,t}$, for which we integrate equation
 \eqref{eq:q_t} to get,
@@
   \label{eq:net_integral}
-  N_{e,t} \approx N \exp\left(-\frac{V_h}{M} \int_0^{\nicefrac{M}{2}} \left(1+\frac{(1-k) (1-r) \left(1-(1-k)^t (1-r)^t\right)}{k+(1-k)r} \right) dr \right).
+  N_{d,t} \approx N \exp\left(-\frac{V_h}{M} \int_0^{\nicefrac{M}{2}} \left(1+\frac{(1-k) (1-r) \left(1-(1-k)^t (1-r)^t\right)}{k+(1-k)r} \right) dr \right).
 \end{align}
@@
 \begin{align}
-  N_{e,\infty}^\text{BGS} = N \exp\left( - \frac{U}{2s + M - Ms} \right).
+  N_{d,\infty}^\text{BGS} = N \exp\left( - \frac{U}{2s + M - Ms} \right).
 \end{align}
@@
 
-\subsection{Effective Population Size Under Weak and Strong BGS} \label{supp:weak-strong}
+\subsection{Effective Population Size Under Weak and Strong BGS}
+\label{supp:weak-strong}
 
@@
 At equilibrium, the $N_e$ that determines fixation probability is the
-asymptotic $N_{e,\infty}$. Consequently, Santiago and Caballero argue that the
+asymptotic $N_{d,\infty}$. Consequently, Santiago and Caballero argue that the
 asymptotic reduction in effective population size uses the $Q_\infty^2$ and is
@@
 \begin{align}
-  R(N_{e,\infty}) &= \frac{4 U s N_{e,\infty}}{\exp(4 N_{e,\infty} s)-1}\\
-  N_{e,\infty} &\approx N \exp\left(-\frac{V_A}{(1-Z) (2-(2-M) Z)} \right)
+  R(N_{d,\infty}) &= \frac{4 U s N_{d,\infty}}{\exp(4 N_{d,\infty} s)-1}\\
+  N_{d,\infty} &\approx N \exp\left(-\frac{V_A}{(1-Z) (2-(2-M) Z)} \right)
 \end{align}
@@
 
-We also find that Equation \eqref{eq:Va} (and Supplementary Materials Equation
-\ref{supp-eq:Va} above) can also be found through Kimura's diffusion models
-with a flux of mutations into infinite, discrete sites
-(\citeyear{Kimura1969-jw}). This model integrates the Kolmogorov backwards
-equation over time and forward variable to find the expected value of a
-functional of the initial frequency. Using the functional $I(p) = s^2 p(1-p)$
-for genic variance given mutations start at frequency $p$, one finds that the
-steady-state genic variance for $p=\nicefrac{1}{2N_e}$ is
+We also find that Equation \eqref{eq:Va} and Supplementary Materials Equation
+\ref{supp-eq:Va} above can also be found through Kimura's diffusion models with
+a flux of mutations into infinite, discrete sites (\citeyear{Kimura1969-jw}).
+This model integrates the Kolmogorov backwards equation over time and forward
+variable to find the expected value of a functional of the initial frequency.
+Using the functional $I(p) = s^2 p(1-p)$ for genic variance given mutations
+start at frequency $p$, one finds that the steady-state genic variance for
+$p=\nicefrac{1}{2N_e}$ is
 
@@
 \end{align}
-
-which is approximately Equation \eqref{eq:Va}.
+%
+which is approximately Equation \eqref{eq:Va} in the main paper.
 
@@
 %   \label{eq:Ne_inf}
-%   N_{e,\infty} \approx N \exp\left(-\frac{V_h + \varepsilon}{(1-Z) (2-(2-M) Z)} \right).
+%   N_{e,\infty} \approx N \exp\left(-\frac{V_h + \Vepsilon}{(1-Z) (2-(2-M) Z)} \right).
 % \end{align}
 
-\subsection{Heterozygosity and $B$ values Under Weak and Strong Selection}
-
-While the selection dynamics are set by the asymptotic $Q_\infty^2$ that
-determines the equilibrium $\widetilde{N}_e$ and $\widetilde{R}$,
-heterozygosity is determined by the sequence $\widetilde{N}_{e,t}$, where each
-term is given by \eqref{eq:net_integral}. We experimented with calculating this
-full sum in our methods, but it seemed to make little difference in our
-application. Additionally, it was too costly to calculate computationally
-for genome-wide calculations of the B' map.
+\section{Statistical Methods}
+
+\subsection{Parametric Forms of Negative Linked Selection Models}
+
+Both classic BGS and the B' method we build off the SC16 model have the same
+parametric form, since both (1) model the impact of linked selection as a
+rescaling of effective population size, and (2) both assume multiplicative
+fitness across segments. Thus, the total reduction experienced by a neutral
+site at position $x$ due to segments across the genome is the product of the
+individual reductions is
+
+\begin{align}
+  \label{supp-eq:b}
+  B(x, \Psi) &= \prod_g^S \exp \left( \int_0^1 b(\mu(s, k(g)), s, L_g, r_g, d(x, g)) ds \right) \\
+             &= \exp \left( \sum_g^S \int_0^1 b(\mu(s, k(g)), s, L_g, r_g, d(x, g)) ds \right)
+\end{align}
+
+where, $b(\mu(s, k(g)), s, L_g, r_g, d(x,g))$ is the log reduction due to
+segment $g$, and there are $S$ total segments. Here, $\mu(s, k(g))$ is the
+per-basepair per-genome rate that mutations with selection coefficient $s$
+enter segments with feature class $k(g)$ (e.g. it incorporates the distribution
+of fitness effects and mutation rate), $L_g$ and $r_g$ are the length in
+basepairs and recombination rate per basepair of segment $g$, and the
+recombination distance between focal site $x$ and segment $g$ is $d(x, g)$. The
+functional form of $b$ varies depending on whether the classic BGS model is
+used (Supplementary Section \ref{supp:reduction-classic}), or our form of
+Santiago and Caballero's (\citeyear{Santiago2016-mu}) equations (see
+Supplementary Section \ref{supp:reduction-sc16}).
+
+Under both background selection models used in our likelihood, the diversity in
+window $v$ is $\bar{\pi}(v, \Psi) = \bar{B}(v, \Psi) \pi_0$. Here, $\bar{B}(v |
+\Psi)$ is the predicted reduction in diversity due to BGS in window $v$, given
+background selection parameters $\Psi$. In practice, we pre-calculate $B(x |
+\Psi)$ at fixed sites $x$ across the genome, and take the average of the fixed
+sites $B$ values within window $v$ for the average $\bar{B}(v | \Psi)$.
+
+\subsubsection{Reductions Under the Classic BGS Model}
+\label{supp:reduction-classic}
+
+The classic BGS model, which assumes deleterious mutations cannot fix,
+expresses the reduction for a focal site in the middle of a segment
+\parencite{Hudson1995-xc,Hudson1994-oh,Nordborg1996-nq}; here we consider a
+focal site directly to an arbitrary side of a segment under a deleterious
+mutations, and extend the model to handle a focal segment arbitrarily far from
+the segment. To simplify notation, we will only consider a single feature class
+in this and the next section. Using our notation from equation
+\eqref{supp-eq:b}, the log reduction under the classic
+mutation-selection-balance BGS model is,
+
+\begin{align}
+  b(\mu, s, L, r, 0) &=  - \int_0^L \frac{2\mu(l)}{s (1 + (1-s) r(l)/s)^2} dl 
+\end{align}
+
+(c.f. \cite{Hudson1995-xc} equation 5), where $\mu(l)$ is the \emph{haploid}
+per-basepair mutation rate at $l$ and $r(l)$ is the recombination fraction
+between the focal site (at position 0) and basepair $l$. Assuming constant
+per-basepair recombination and mutation rates, $r(l) = r_\text{BP}l$ and
+$\mu(l) = \mu$, and
+
+\begin{align}
+  b(\mu, s, L, r_\text{BP}, 0) &= - \frac{2\mu L}{s + (1-s) r_\text{BP} L} 
+\end{align}
+
+(c.f. \cite{Hudson1995-xc} equation 8, where the factor of two difference is
+due to the position of the focal site). Often per-region rates are defined $U =
+2 \mu L$ and $M = r_\text{BP} L$ are the segment-wide mutation diploid mutation
+rates (mutations per diploid segment per generation) and recombination length
+(Morgans).
+
+For computational efficiency, we adjust this model by setting $r(l) = d +
+r_\text{BP} l$, where $d$ is the recombination fraction between the focal site
+$x$ and the start of segment $g$. This allows us to pre-compute the local
+effects of the segment, and substitute in $b$ as the focal site changes. This
+integral substitution has a closed form,
+
+\begin{align}
+  b(\mu, s, L, r_\text{BP}, d) &= \frac{-2\mu L}{(d(1-s) + s)(s + (d + r_\text{BP}L)(1-s))}
+\end{align}
+
+and terms can be collected in powers of $d$ and pre-computed for all segments
+for the grid of $\mu$ and $s$. Our implementation pre-computes these segment
+components, and computes the final $b$ value for an $x$ and $g$ by calculating
+the distance $d(x,g) = |M(x) - M(g)|$, where $M(x)$ is the cumulative
+recombination map length at position $x$ in Morgans. Since $r(l)$ could vary by
+position, segments with differing recombination rates are split into segments
+of the same annotation class by their recombination rate. This has no effect on
+the calculation due to the multiplicative fitness, but ensures more accurate
+estimates of the reduction factor $B$.
+
+% Each B map is determined by the annotation track of putatively conserved
+% segments, which includes (1) the segment positions and lengths, (2) the type of
+% feature of each segment (e.g.  exons, UTRs, phastcons conserved elements,
+% etc.), as well as (3) the recombination map, and (3) the deleterious selection
+% coefficient $s(i)$ and mutation rate at which these deleterious enter the
+% population for feature type $i$.
+
+\subsubsection{Reduction Under the Modified Santiago and Caballero Model}
+\label{supp:reduction-sc16}
+
+Under Santiago and Caballero's model, the correct effective population size to
+gauge reduction in heterozygosity or coalescent times is given by their
+equation 4 \parencite{Santiago2016-mu}. This equation considers how the
+inflation factor $Q_t^2$ that sets the draft-effective population size varies
+over $t$ (c.f. Supplementary equation \ref{eq:net_integral}). This is because
+selective processes can lead to non-constant pairwise coalescent rates, which
+impact pairwise coalescent rates and thus the reduction factor. We experimented
+with calculating this full sum and integral approximations to it in our
+methods, but it seemed to make little difference in our application.
+Additionally, it was too costly to calculate computationally for genome-wide
+calculations of the B' map. Instead, we solve for the asymptotic
+$\widetilde{N}_{e,\infty}$ for each segment, and use that to calculate the log
+reduction factor under the SC16 model, $b'$.
+
+Note that the asymptotic model above determines the reduction in effective
+population size experienced by a focal site in the middle of a $L$-basepair
+segment, but B maps require the reduction $b'$ experienced by a focal site $x$
+at an arbitrary recombination distance apart. For each segment, we pre-compute
+the solution to the system of two equations, giving us $V$ and $Q_\infty^2$,
+and then use these values in the equation \eqref{eqn:Ne_VQ2},
+
+\begin{align} 
+    b'(\mu, s, L, r, d) = -\frac{V}{2} \left(\frac{1}{(1-(1-d)Z}\right)^2 
+\end{align}
+
+where $Z = 1 - Us / (U-\widetilde{R})$ and $V = Us - 2\widetilde{R}s$. TODO
+check d.
+
+\subsection{Discretization of Parameters and Annotation Feature Classes}
+
+In practice, to fit such a model, we must discretize both $\mu(s, k(g))$ and
+$s$ in Equation \eqref{supp-eq:b}. We define $\mu_i$ to be the rate of
+mutations entering the population per basepair, per generation with selection
+coefficient $s_i$ (i.e. the product of the mutation rate and the distribution
+of deleterious fitness effects). Then, the reduction at position $x$ is,
+
+\begin{align}
+  \label{supp-eq:b}
+  B(x, \Psi) = \exp \left( \sum_g^S \sum_{i} b\left( \mu_{i, \;k(g)}, s_i, L_g, r_g, d(x, g) \right)\right).
+\end{align}
+
+%TODO double check
+
+Because it is computationally-intensive to numerically solve the system of two
+non-linear equations to calculate $b$ at each segment, these are pre-computed
+for an $m_s$-element grid of selection coefficients and an $m_\mu$-element grid
+of mutation rates--DFE products. The number of annotation features $K$ is
+variable and set by the annotation data specified. There are two ways to
+parameterize the mutation rates and DFE for all annotation classes in
+background selection models. First, for the \emph{free-mutation}
+parameterization, each annotation feature class has a free mutation rate
+parameter for each of these selection coefficients (as in equation
+\eqref{supp-eq:b}).
+
+In the free-mutation parameterization, there are $1 + K \times m_s$ for $\pi_0$
+and the free mutation rate parameters, where $K$ is the number of annotation
+features. The MLE optimization for this approach is unconstrained, though
+$\pi_0$ is bounded to be within $10^{-5} \le \pi_0 \le 10^{-1}$ and mutation
+rates are bounded within $10^{-11} \le \mu \le 10^{-7}$. The BGS model requires
+a mutation rate for each selection coefficient in the grid, for each annotation
+class (here, as an example, CDS, UTRs, and phastcons), which are stored in the
+$m_s \times K$ matrix $\mathbf{M}_\text{F}$,
+
+\begin{align}
+  \mathbf{M}_\text{F} = \begin{bmatrix}
+    \mu_{10^{-6},\text{CDS}} & \mu_{10^{-6},\text{UTR}} & \mu_{10^{-6},\text{PC}} \\
+    \mu_{10^{-5},\text{CDS}} & \mu_{10^{-5},\text{UTR}} & \mu_{10^{-5},\text{PC}} \\
+    \vdots & \vdots & \vdots \\
+    \mu_{10^{-1},\text{CDS}} & \mu_{10^{-1},\text{UTR}} & \mu_{10^{-1},\text{PC}} \\
+  \end{bmatrix}
+\end{align}
+
+That is, the distribution of fitness effect (DFE) is implied by
+the total mutation rate across selection coefficients for an annotation class.
+We can normalize by the total mutation rate estimate to get the estimated DFE,
+giving us the DFE weight for annotation class $k$ and selection coefficient
+with index $i$,
+
+\begin{align}
+  \widehat{w}_{i,k} = \frac{\widehat{\mu}_{i,k}}{\sum_i \widehat{\mu}_{i,k}}.
+\end{align}
+
+The second parametrization is the \emph{simplex} parametrization, which has a
+single mutation rate across all features, so it is a DFE weight matrix times
+the mutation rate $\mu$,
+
+\begin{align}
+  \mathbf{M}_\text{S} = \mu \begin{bmatrix}
+    w_{10^{-6},\text{CDS}} & w_{10^{-6},\text{UTR}} & w_{10^{-6},\text{PC}} \\
+    w_{10^{-5},\text{CDS}} & w_{10^{-5},\text{UTR}} & w_{10^{-5},\text{PC}} \\
+    \vdots & \vdots & \vdots \\
+    w_{10^{-1},\text{CDS}} & w_{10^{-1},\text{UTR}} & w_{10^{-1},\text{PC}} \\
+  \end{bmatrix}
+\end{align}
+
+which has $2 + K \times (m_s-1)$ free parameters, since each column must sum to
+one. This imposes the constraint that $\sum_i w_{i,k} = 1$, requiring
+constrained MLE optimization.
+
+\subsection{Data Summary Matrices}
+
+Our underlying data for all likelihood and pairwise diversity estimates is the
+allele count matrix $\mathbf{C}$ with dimensions $L \times 2$, where $L$ is the
+chromosome length. This is transformed to a pairwise summary matrix with
+identical dimensions, $\mathbf{Y}$. The first column of $\mathbf{Y}$ is the
+number of pairwise comparisons between chromosomes that are identical, and the
+second column is the number that are different. Both of these columns are
+combinatoric summaries of the raw allele counts matrix needed for the binomial
+likelihood and pairwise diversity estimates. Let $[c_1, c_2]$ be a row of
+$\mathbf{C}$ for basepair $l$ (the $l$ index is omitted for clarity), and $n =
+c_1 + c_2$. Then, the (1) total number of pairwise combinations of chromosomes
+$n_T$, (2) the number of pairwise with identical alleles $n_S$, and (3) the
+number of pairwise combinations with differing alleles $n_D$ are respectively,
+
+\begin{align*}
+  n_T &= \frac{n(n-1)}{2} \\
+  n_S &= {c_1 \choose 2} + {c_2 \choose 2} \\
+  n_D &= n_T - n_S
+\end{align*}
+%
+which would be stored in row $\mathbf{Y}_l = [n_S, n_D]$. Note that the
+per-site $\mathbf{Y}$ handles non-polymorphic sites and missing data.
+Non-polymorphic sites have $n_S = {n \choose 2}$ and $n_D = 0$, and missing
+data has $n_S = n_D = 0$.
+
+\subsection{Pairwise Diversity Estimates}
+\label{suppsec:diversity}
+
+The pairwise diversity at site $l$ across the $n$ sampled chromosomes can be
+calculated from row $l$ of the $\mathbf{Y}$ matrix as follows,
+
+\begin{align}
+  \pi_l = \frac{n_D}{n_T}
+\end{align}
+
+which is identical to the more common expression of this estimator,
+
+\begin{align}
+  \pi_l = \frac{2}{n(n-1)} \sum_{i < j}^n k_{i,j}
+\end{align}
+
+where $k_{i,j}$ is 1 if the alleles at this site differ at site $l$, and 0
+otherwise.
+
+There are three ways to aggregate $\pi_l$ across all sites. The first is,
+
+\begin{align}
+  \label{eq:}
+  \pi^{(1)} = \frac{1}{L} \sum_{i=1}^L \frac{n_{D,i}}{n_{T,i}}
+\end{align}
+
+which if the number of samples across loci is constant, simplifies to an
+unweighted average across sites. Second, one can take a weighted average, with
+weights determined by the total number of samples present at a site,
+
+\begin{align}
+  \label{eq:}
+  \pi^{(2)} = \frac{1}{\sum_{i=1}^L n_i} \sum_{i=1}^L n_i \frac{n_{D,i}}{n_{T,i}}
+\end{align}
+
+Third, one can weight by the number of pairwise comparisons at a site,
+$n_{T,i}$, rather than total number of samples, $n_i$, which leads to a ratio
+of sums,
+
+\begin{align}
+  \label{eq:}
+  \pi^{(3)} = \frac{\sum_{i=1}^L n_{D,i}}{\sum_{i=1}^L n_{T,i}}.
+\end{align}
+
+We predominantly use the estimator $\pi^{(3)}$, as it corresponds to how we
+summarize the matrix $\mathbf{Y}$ across windows for our likelihood. All
+methods have mean squared errors and biases very close to one another (TODO).
+
+Note, however, that estimates of pairwise diversity often condition on the
+accessible bases, and thus treat this as fixed. However, the number of
+accessible bases varies across the chromosome; this can lead to a source of
+apparent bias during block-bootstrap estimates of uncertainty.  In this case,
+pairwise diversity is a ratio estimator, and is thus biased, since by Jensen's
+inequality $\E(\nicefrac{y}{x}) \ge \nicefrac{\E(y)}{\E(x)}$ for random
+variables $x$ and $y$.
+
+\subsection{Window-based Summaries and filtering}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics{figures/supplementary/diversity_trimming_dist.pdf}
+
+  \caption{The distribution of diversity across the genome at the megabase
+  scale, with outliers trimmed. The blue line is the normal CDF, fit with MLE
+parameters XXX. The inset figure is the untrimmed data, with trimmed points
+shown in red.}
+
+  \label{suppfig:trimming}
+\end{figure}
+
+The likelihood method is fit to non-overlapping binned summaries of the allele
+counts matrix.
+
+Based on exploratory data analyses, some bins were outliers and
+excluded (XXX).
+
+\begin{enumerate}
+  \item \textbf{Fraction of inaccessible sites per window}.  \path{mask_inaccessible_bins_frac}
+  % \item \textbf{Edge truncation}
+
+  \item \textbf{Outliers}: Based on exploratory analysis, there were some
+    regions with very high diversity. The 0.05\% right tail was excluded.
+
+\end{enumerate}
+
+% supp data
+% - npy files
+% - edge trunation
+
+\subsection{Annotation Models}
+
+CADD 6\%: 6.49\% of autosomal sequence.
+CADD 8\%: 8.56\% of autosomal sequence.
+
+\subsection{Ratchet Rate Prediction}
+
+After rescaling each segment $N_e$ by the local predicted reduction
+$\hat{B}(x)$, the ratchet rates and $B'$ values are re-calculated.
+
+The rescaled $B'$ calculation outputs a $m_\mu \times m_s \times S$
+multidimensional array $\mathbf{R}$ of ratchet rates per segment. These are
+rescaled by the segment lengths $\mathbf{l}$, giving the per-basepair mutation
+rate array. The per-basepair ratchet rate given maximum likelihood estimates of
+$\hat{\mu}$ and $\widehat{\mathbf{W}}$ is used to predicted the ratchet rate
+for each segment $g$. The DFE estimate for feature $k$ is a column vector of
+$\widehat{\mathbf{W}}$, i.e. $\widehat{\mathbf{w}}_k$
+
+\begin{align}
+  \lambda_D(g) = 
+\end{align}
+
+The predicted ratchet rate for segment $g$ $\lambda_d(g)$ is based on these
+values.
+
+\subsection{Numeric Optimization}
+
+mu bounds
+
+bounds and time, bounds and convergence
+
+Same optima, but took longer
 
@@
 regions. For our analyses we choose the 200bp slop version, which covers
-63.66\% of the autosomal reference sequence.
+63.66\% of the autosomal reference sequence and consists of 1.26 million
+contiguous genomic regions.
 
@@
 pairwise diversity is 13\% higher (two-tailed t-test p-value $1.29 \times
-10^{-71}$) in the CADD 70\% neutral track. However, we did not use the CADD
-70\%  neutral track in our analysis of human population genomic data. Our
-concern was that this neutral track would be too close to the complement of the
-CADD 6\% track, which we use as one of our annotation models.
-
-These files are all produced by the Snakemake file
+10^{-71}$) in the CADD bottom 70\% neutral track. However, we did not use the
+CADD bottom 70\%  neutral track in our analysis of human population genomic
+data. Our concern was that because CADD models are trained by differentiating
+high-frequency derived SNPs in humans from simulated variants at every
+basepair, there could be circularity in basing our neutral regions on
+predictions based on modeling frequency data. Moreover, this neutral track
+would be close to a complement of the CADD 6\% track, which we use as one of
+our annotation models, and might lead to issues. Additionally, the CADD bottom
+70\% neutral track is much less sparse than the track we used in our analysis,
+with a total of $\approx 20$ times more contiguous regions than the neutral
+track used in analysis (25.38 million total). Many of these ranges are at very
+small spatial scales (average: $\approx 7$ basepairs).
+
+Additionally, we ran one test model with the CADD bottom 70\% as the neutral
+track as a test. Using the CADD bottom 70\% neutral track with the CADD 6\%
+annotation model, the in-sample was $R^2=63.99$\%; by comparison, the in-sample
+$R^2=68.03$\% using the constructed neutral track with 200bp slop. Overall,
+future work calibrating the neutral track is needed.
+
+These neutral tracks are all produced by the Snakemake file
 \path{data/fit_annotation/Snakefile} in this study's GitHub repository
@@
 \centering
-\caption{XXX}
-\begin{tabular}{lrrr}
+\caption{Percent of basepairs remaining that are accessible and 
+         putatively neutral.}
+\begin{tabular}{|l|rrr|}
 \hline
-chrom &  accessible &  neutral &  both \\
+Chromosome & Accessible & Neutral & Intersection \\
 \hline
- chr1 &        42.4 &     64.4 &  22.9 \\
- chr2 &        46.8 &     58.7 &  23.7 \\
- chr3 &        44.1 &     62.9 &  24.8 \\
- chr4 &        44.0 &     59.3 &  21.5 \\
- chr5 &        44.1 &     58.0 &  21.4 \\
- chr6 &        45.0 &     58.0 &  22.1 \\
- chr7 &        44.4 &     65.8 &  26.4 \\
- chr8 &        43.8 &     61.2 &  23.1 \\
- chr9 &        39.2 &     64.5 &  20.3 \\
-chr10 &        44.6 &     62.8 &  25.3 \\
-chr11 &        42.7 &     63.5 &  23.4 \\
-chr12 &        41.7 &     62.7 &  22.7 \\
-chr13 &        39.7 &     62.0 &  18.3 \\
-chr14 &        37.6 &     65.4 &  19.0 \\
-chr15 &        37.0 &     69.4 &  20.3 \\
-chr16 &        39.4 &     63.8 &  20.9 \\
-chr17 &        40.4 &     64.0 &  22.3 \\
-chr18 &        40.9 &     59.8 &  19.8 \\
-chr19 &        30.9 &     69.8 &  18.0 \\
-chr20 &        36.9 &     61.6 &  19.8 \\
-chr21 &        31.6 &     68.6 &  18.2 \\
-chr22 &        26.9 &     75.0 &  16.4 \\
+ chr1 &        39.7 &     61.2 &  23.4 \\
+ chr2 &        45.3 &     67.3 &  27.5 \\
+ chr3 &        43.2 &     67.3 &  25.7 \\
+ chr4 &        42.7 &     70.5 &  27.4 \\
+ chr5 &        42.4 &     67.5 &  26.0 \\
+ chr6 &        43.8 &     67.9 &  26.7 \\
+ chr7 &        41.9 &     67.8 &  26.2 \\
+ chr8 &        42.3 &     70.9 &  27.5 \\
+ chr9 &        36.2 &     58.1 &  22.0 \\
+chr10 &        42.9 &     68.2 &  26.8 \\
+chr11 &        41.3 &     67.0 &  25.0 \\
+chr12 &        40.6 &     67.2 &  24.1 \\
+chr13 &        38.8 &     59.8 &  25.2 \\
+chr14 &        36.5 &     54.9 &  21.5 \\
+chr15 &        34.4 &     53.7 &  20.6 \\
+chr16 &        36.0 &     62.5 &  23.1 \\
+chr17 &        37.2 &     63.3 &  21.6 \\
+chr18 &        40.3 &     65.7 &  26.3 \\
+chr19 &        28.1 &     66.1 &  17.0 \\
+chr20 &        36.1 &     65.1 &  22.4 \\
+chr21 &        29.8 &     55.9 &  20.9 \\
+chr22 &        25.5 &     50.9 &  17.1 \\
+\hline
 \end{tabular}
 \end{table}
-
-\subsection{Data Summary Matrices}
-
-Our underlying data for all likelihood and pairwise diversity estimates is the
-allele count matrix $\mathbf{C}$ with dimensions $L \times 2$, where $L$ is the
-chromosome length. This is transformed to a pairwise summary matrix with
-identical dimensions, $\mathbf{Y}$. The first column of $\mathbf{Y}$ is the
-number of pairwise comparisons between chromosomes that are identical, and the
-second column is the number that are different. Both of these columns are
-combinatoric summaries of the raw allele counts matrix needed for the binomial
-likelihood and pairwise diversity estimates. Let $[c_1, c_2]$ be a row of
-$\mathbf{C}$ for basepair $l$ (the $l$ index is omitted for clarity), and $n =
-c_1 + c_2$. Then, the (1) total number of pairwise combinations of chromosomes
-$n_T$, (2) the number of pairwise with identical alleles $n_S$, and (3) the
-number of pairwise combinations with differing alleles $n_D$ are respectively,
-
-\begin{align*}
-  n_T &= \frac{n(n-1)}{2} \\
-  n_S &= {c_1 \choose 2} + {c_2 \choose 2} \\
-  n_D &= n_T - n_S
-\end{align*}
-%
-which would be stored in row $\mathbf{Y}_l = [n_S, n_D]$. Note that the
-per-site $\mathbf{Y}$ handles non-polymorphic sites and missing data.
-Non-polymorphic sites have $n_S = {n \choose 2}$ and $n_D = 0$, and missing
-data has $n_S = n_D = 0$.
-
-\subsection{Pairwise Diversity Estimates}
-\label{suppsec:diversity}
-
-The pairwise diversity at site $l$ across the $n$ sampled chromosomes can be
-calculated from row $l$ of the $\mathbf{Y}$ matrix as follows,
-
-\begin{align}
-  \pi_l = \frac{n_D}{n_T}
-\end{align}
-
-which is identical to the more common expression of this estimator,
-
-\begin{align}
-  \pi_l = \frac{2}{n(n-1)} \sum_{i < j}^n k_{i,j}
-\end{align}
-
-where $k_{i,j}$ is 1 if the alleles at this site differ at site $l$, and 0
-otherwise.
-
-There are three ways to aggregate $\pi_l$ across all sites. The first is,
-
-\begin{align}
-  \label{eq:}
-  \pi^{(1)} = \frac{1}{L} \sum_{i=1}^L \frac{n_{D,i}}{n_{T,i}}
-\end{align}
-
-which if the number of samples across loci is constant, simplifies to an
-unweighted average across sites. Second, one can take a weighted average, with
-weights determined by the total number of samples present at a site,
-
-\begin{align}
-  \label{eq:}
-  \pi^{(2)} = \frac{1}{\sum_{i=1}^L n_i} \sum_{i=1}^L n_i \frac{n_{D,i}}{n_{T,i}}
-\end{align}
-
-Third, one can weight by the number of pairwise comparisons at a site,
-$n_{T,i}$, rather than total number of samples, $n_i$, which leads to a ratio
-of sums,
-
-\begin{align}
-  \label{eq:}
-  \pi^{(3)} = \frac{\sum_{i=1}^L n_{D,i}}{\sum_{i=1}^L n_{T,i}}.
-\end{align}
-
-We predominantly use the estimator $\pi^{(3)}$, as it corresponds to how we
-summarize the matrix $\mathbf{Y}$ across windows for our likelihood. All
-methods have mean squared errors and biases very close to one another (TODO).
-
-Note, however, that estimates of pairwise diversity often condition on the
-accessible bases, and thus treat this as fixed. However, the number of
-accessible bases varies across the chromosome; this can lead to a source of
-apparent bias during block-bootstrap estimates of uncertainty.  In this case,
-pairwise diversity is a ratio estimator, and is thus biased, since by Jensen's
-inequality $\E(\nicefrac{y}{x}) \ge \nicefrac{\E(y)}{\E(x)}$ for random
-variables $x$ and $y$.
-
-\subsection{Window-based Summaries and filtering}
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics{figures/supplementary/diversity_trimming_dist.pdf}
-
-  \caption{The distribution of diversity across the genome at the megabase
-  scale, with outliers trimmed. The blue line is the normal CDF, fit with MLE
-parameters XXX. The inset figure is the untrimmed data, with trimmed points
-shown in red.}
-
-  \label{suppfig:trimming}
-\end{figure}
-
-The likelihood method is fit to non-overlapping binned summaries of the allele
-counts matrix.
-
-Based on exploratory data analyses, some bins were outliers and
-excluded (XXX).
-
-\begin{enumerate}
-  \item \textbf{Fraction of inaccessible sites per window}.  \path{mask_inaccessible_bins_frac}
-  % \item \textbf{Edge truncation}
-
-  \item \textbf{Outliers}: Based on exploratory analysis, there were some
-    regions with very high diversity. The 0.05\% right tail was excluded.
-
-\end{enumerate}
-
-% supp data
-% - npy files
-% - edge trunation
-
-\section{Additional Results}
 
@@
 
-\section{Table of $R^2$ Values}
+\subsection{Table of $R^2$ Values}
 
@@
 
-\section{Derivation of Approximate Drift $R^2$}
+\subsection{Derivation of Approximate Drift $R^2$}
 
@@
 as the predicted level under our negative selection model, $\pi_0 B_i$, plus
-some random residual error $\varepsilon_i$,
-
-\begin{align}
-    y_i = \pi_0 B_i + \varepsilon_i
+some random residual error $\Vepsilon_i$,
+
+\begin{align}
+    y_i = \pi_0 B_i + \Vepsilon_i
 \end{align}
@@
     V_\text{res} &= \frac{1}{n} \sum_{i=1}^n (y_i - \widehat{\pi}_0 \widehat{B}_i)^2  \\
-                 &= \underbrace{\frac{1}{n} \sum_{i=1}^n \varepsilon_i^2}_\text{irreducible error} + \underbrace{\left(\frac{1}{n} \sum_{i=1}^n (\pi_0 B_i - \widehat{\pi}_0 \widehat{B}_i)\right)^2}_\text{bias squared}.
+                 &= \underbrace{\frac{1}{n} \sum_{i=1}^n \Vepsilon_i^2}_\text{irreducible error} + \underbrace{\left(\frac{1}{n} \sum_{i=1}^n (\pi_0 B_i - \widehat{\pi}_0 \widehat{B}_i)\right)^2}_\text{bias squared}.
 \end{align}
@@
 \begin{align}
-    V_\text{tot} = V_\text{res} + V_\text{model} + 2\widehat{\pi}_0\cov_i(\varepsilon_i, \widehat{B}_i)
+    V_\text{tot} = V_\text{res} + V_\text{model} + 2\widehat{\pi}_0\cov_i(\Vepsilon_i, \widehat{B}_i)
 \end{align}
@@
 Generally, fitting procedures act to minimize the squared deviation between
-predictors and observed values, which drives $\cov_i(\varepsilon_i,
+predictors and observed values, which drives $\cov_i(\Vepsilon_i,
 \widehat{B}_i)$ to zero XXX. \textcite{Murphy2022-sj} suggest  XXX
@@
 \begin{align}
-    \E(\varepsilon_i^2) = \var(\pi_i)
+    \E(\Vepsilon_i^2) = \V(\pi_i)
 \end{align}
@@
 since the mean residual is zero. The variance in pairwise diversity
-$\var(\pi_i)$ is equivalent to the variance in the number of segregating sites
-for a sample of two, $\var(\pi) = \var(S_2)$. The variance in coalescence times
+$\V(\pi_i)$ is equivalent to the variance in the number of segregating sites
+for a sample of two, $\V(\pi) = \V(S_2)$. The variance in coalescence times
 in a window of width $w$ basepairs with population-scaled recombination rate
@@
 \begin{align}
-    \var(S_2) = \theta + \theta^2 \frac{2}{\rho^2} \int_0^\rho (\rho - x) f_2(x) dx
+    \V(S_2) = \theta + \theta^2 \frac{2}{\rho^2} \int_0^\rho (\rho - x) f_2(x) dx
 \end{align}
@@
 
-\section{Predicted Rate of Fitness Change}
+\subsection{Predicted Rate of Fitness Change}
 
@@
 
-\section{Likelihood}
+\subsection{Likelihood}
 
@@
 
-\section{Additional Theory}
-
-\subsection{supp:weak-strong}
-
-\section{Background Selection Models}
-
-The BGS models considered here both assume multiplicative fitness effects
-across sites. Thus, the total reduction due to segments across the genome is
-the product of the individual reductions,
-
-\begin{align}
-  \label{suppeqn:b}
-  B(x, \Psi) &= \prod_g^S \exp \left( \int_0^1 b(\mu(s, k(g)), s, L_g, r_g, d(x, g)) ds \right) \\
-             &= \exp \left( \sum_g^S \int_0^1 b(\mu(s, k(g)), s, L_g, r_g, d(x, g)) ds \right)
-\end{align}
-
-where, $b(\mu(s, k(g)), s, L_g, r_g, d(x,g))$ is the log reduction due to
-segment $g$, and there are $S$ total segments. Here, $\mu(s, k(g))$ is the
-per-basepair per-genome rate that mutations with selection coefficient $s$
-enter segments with feature class $k(g)$, $L_g$ and $r_g$ are the length in
-basepairs and recombination rate per basepair of segment $g$, and the
-recombination distance between focal site $x$ and segment $g$ is $d(x, g)$. The
-functional form of $b$ varies depending on whether the classic BGS model is
-used, or our form of Santiago and Caballero's (\citeyear{Santiago2016-mu})
-equations (see Section XXX).
-
-Under both background selection models used in our likelihood, the diversity in
-window $b$ is $\bar{\pi}(b, \Psi) = \bar{B}(b, \Psi) \pi_0$. Here, $\bar{B}(b |
-\Psi)$ is the predicted reduction in diversity due to BGS in window $b$, given
-background selection parameters $\Psi$. In practice, we pre-calculate $B(x |
-\Psi)$ at fixed sites $x$ across the genome, and take the average of the fixed
-sites $B$ values within widow $b$ for the average $\bar{B}(b | \Psi)$.
-
-\subsection{Reductions Under the Classic BGS Model}
-
-The classic BGS model, which assumes deleterious mutations cannot fix,
-expresses the reduction for a focal site in the middle of a segment
-\parencite{Hudson1995-xc,Hudson1994-oh,Nordborg1996-nq}; here we consider a
-focal site directly to an arbitrary side of a segment under a deleterious
-mutations, and extend the model to handle a focal segment arbitrarily far from
-the segment. To simplify notation, we will only consider a single feature class
-in this and the next section. Using our notation from equation
-\eqref{suppeqn:b}, the classic mutation-selection-balance BGS model is,
-
-\begin{align}
-  b(\mu, s, L, r, 0) &=  - \int_0^L \frac{2\mu(l)}{s (1 + (1-s) r(l)/s)^2} dl 
-\end{align}
-
-(c.f. \cite{Hudson1995-xc} equation 5), where $\mu(l)$ is the \emph{haploid}
-per-basepair mutation rate at $l$ and $r(l)$ is the recombination fraction
-between the focal site (at position 0) and basepair $l$. Assuming constant
-per-basepair recombination and mutation rates, $r(l) = r_\text{BP}l$ and
-$\mu(l) = \mu$, and
-
-\begin{align}
-  b(\mu, s, L, r_\text{BP}, 0) &= - \frac{2\mu L}{s + (1-s) r_\text{BP} L} 
-\end{align}
-
-(c.f. \cite{Hudson1995-xc} equation 8, where the factor of two difference is
-due to the position of the focal site). Often per-region rates are defined $U =
-2 \mu L$ and $M = r_\text{BP} L$ are the segment-wide mutation diploid mutation
-rates (mutations per diploid segment per generation) and recombination length
-(Morgans).
-
-For computational efficiency, we adjust this model by setting $r(l) = d +
-r_\text{BP} l$, where $d$ is the recombination fraction between the focal site
-$x$ and the start of segment $g$. This allows us to pre-compute the local
-effects of the segment, and substitute in $b$ as the focal site changes. This
-integral substitution has a closed form,
-
-\begin{align}
-  b(\mu, s, L, r_\text{BP}, d) &= \frac{-2\mu L}{(d(1-s) + s)(s + (d + r_\text{BP}L)(1-s))}
-\end{align}
-
-and terms can be collected in powers of $d$ and pre-computed for all segments
-for the grid of $\mu$ and $s$. Our implementation pre-computes these segment
-components, and computes the final $b$ value for an $x$ and $g$ by calculating
-the distance $d(x,g) = |M(x) - M(g)|$, where $M(x)$ is the cumulative
-recombination map length at position $x$ in Morgans. Since $r(l)$ could vary by
-position, segments with differing recombination rates are split into segments
-of the same annotation class by their recombination rate. This has no effect on
-the calculation due to the multiplicative fitness, but ensures more accurate
-estimates of the reduction factor $B$.
-
-% Each B map is determined by the annotation track of putatively conserved
-% segments, which includes (1) the segment positions and lengths, (2) the type of
-% feature of each segment (e.g.  exons, UTRs, phastcons conserved elements,
-% etc.), as well as (3) the recombination map, and (3) the deleterious selection
-% coefficient $s(i)$ and mutation rate at which these deleterious enter the
-% population for feature type $i$.
-
-\subsection{Reduction Under the Modified Santiago and Caballero Model}
-
-Under Santiago and Caballero's model, the correct effective population size to
-gauge reduction in heterozygosity or coalescent times is given by equation XXX.
-This is because selective processes can lead to non-constant pairwise
-coalescent rates, which impact pairwise coalescent rates and thus the reduction
-factor. We experimented with the corrections described in Section XXX, and
-found (1) they had little effect on the accuracy of the B maps, and (2) they
-were incredibly costly to calculate, even with the approximations described in
-equations XXX. Overall, we solve for the asymptotic $\widetilde{N}_{e,\infty}$
-for each segment, and use that to calculate $b'$.
-
-Note that the asymptotic model above determines the reduction in effective
-population size experienced by a focal site in the middle of a $L$-basepair
-segment, but B maps require the reduction $b'$ experienced by a focal site $x$
-at an arbitrary recombination distance apart. For each segment, we pre-compute
-the solution to the system of two equations, giving us $V$ and $Q_\infty^2$,
-and then use these values in the equation \eqref{eqn:Ne_VQ2},
-
-\begin{align} 
-    b'(\mu, s, L, r, d) = -\frac{V}{2} \left(\frac{1}{(1-(1-d)Z}\right)^2 
-\end{align}
-
-where $Z = 1 - Us / (U-\widetilde{R})$ and $V = Us - 2\widetilde{R}s$. TODO
-check d.
-
-\subsection{Discretization of Parameters and Annotation Feature Classes}
-
-In practice, to fit such a model, we must discretize both $\mu(s, k(g))$ and
-$s$ in Equation \eqref{suppeqn:b}. We define $\mu_i$ to be the rate of
-mutations entering the population per basepair, per generation with selection
-coefficient $s_i$ (i.e. the product of the mutation rate and the distribution
-of deleterious fitness effects). Then, the reduction at position $x$ is,
-
-\begin{align}
-  \label{suppeqn:b}
-  B(x, \Psi) = \exp \left( \sum_g^S \sum_{i} b\left( \mu_{i, \;k(g)}, s_i, L_g, r_g, d(x, g) \right)\right).
-\end{align}
-
-%TODO double check
-
-Because it is computationally-intensive to numerically solve the system of two
-non-linear equations to calculate $b$ at each segment, these are pre-computed
-for an $m_s$-element grid of selection coefficients and an $m_\mu$-element grid
-of mutation rates--DFE products. The number of annotation features $K$ is
-variable and set by the annotation data specified. There are two ways to
-parameterize the mutation rates and DFE for all annotation classes in
-background selection models. First, for the \emph{free-mutation}
-parameterization, each annotation feature class has a free mutation rate
-parameter for each of these selection coefficients (as in equation
-\eqref{suppeqn:b}).
-
-In the free-mutation parameterization, there are $1 + K \times m_s$ for $\pi_0$
-and the free mutation rate parameters, where $K$ is the number of annotation
-features. The MLE optimization for this approach is unconstrained, though
-$\pi_0$ is bounded to be within $10^{-5} \le \pi_0 \le 10^{-1}$ and mutation
-rates are bounded within $10^{-11} \le \mu \le 10^{-7}$. The BGS model requires
-a mutation rate for each selection coefficient in the grid, for each annotation
-class (here, as an example, CDS, UTRs, and phastcons), which are stored in the
-$m_s \times K$ matrix $\mathbf{M}_\text{F}$,
-
-\begin{align}
-  \mathbf{M}_\text{F} = \begin{bmatrix}
-    \mu_{10^{-6},\text{CDS}} & \mu_{10^{-6},\text{UTR}} & \mu_{10^{-6},\text{PC}} \\
-    \mu_{10^{-5},\text{CDS}} & \mu_{10^{-5},\text{UTR}} & \mu_{10^{-5},\text{PC}} \\
-    \vdots & \vdots & \vdots \\
-    \mu_{10^{-1},\text{CDS}} & \mu_{10^{-1},\text{UTR}} & \mu_{10^{-1},\text{PC}} \\
-  \end{bmatrix}
-\end{align}
-
-That is, the distribution of fitness effect (DFE) is implied by
-the total mutation rate across selection coefficients for an annotation class.
-We can normalize by the total mutation rate estimate to get the estimated DFE,
-giving us the DFE weight for annotation class $k$ and selection coefficient
-with index $i$,
-
-\begin{align}
-  \widehat{w}_{i,k} = \frac{\widehat{\mu}_{i,k}}{\sum_i \widehat{\mu}_{i,k}}.
-\end{align}
-
-The second parametrization is the \emph{simplex} parametrization, which has a
-single mutation rate across all features, so it is a DFE weight matrix times
-the mutation rate $\mu$,
-
-\begin{align}
-  \mathbf{M}_\text{S} = \mu \begin{bmatrix}
-    w_{10^{-6},\text{CDS}} & w_{10^{-6},\text{UTR}} & w_{10^{-6},\text{PC}} \\
-    w_{10^{-5},\text{CDS}} & w_{10^{-5},\text{UTR}} & w_{10^{-5},\text{PC}} \\
-    \vdots & \vdots & \vdots \\
-    w_{10^{-1},\text{CDS}} & w_{10^{-1},\text{UTR}} & w_{10^{-1},\text{PC}} \\
-  \end{bmatrix}
-\end{align}
-
-which has $2 + K \times (m_s-1)$ free parameters, since each column must sum to
-one. This imposes the constraint that $\sum_i w_{i,k} = 1$, requiring
-constrained MLE optimization.
-
-\section{Ratchet Rate Prediction}
-
-After rescaling each segment $N_e$ by the local predicted reduction
-$\hat{B}(x)$, the ratchet rates and $B'$ values are re-calculated.
-
-The rescaled $B'$ calculation outputs a $m_\mu \times m_s \times S$
-multidimensional array $\mathbf{R}$ of ratchet rates per segment. These are
-rescaled by the segment lengths $\mathbf{l}$, giving the per-basepair mutation
-rate array. The per-basepair ratchet rate given maximum likelihood estimates of
-$\hat{\mu}$ and $\widehat{\mathbf{W}}$ is used to predicted the ratchet rate
-for each segment $g$. The DFE estimate for feature $k$ is a column vector of
-$\widehat{\mathbf{W}}$, i.e. $\widehat{\mathbf{w}}_k$
-
-\begin{align}
-  \lambda_D(g) = 
-\end{align}
-
-The predicted ratchet rate for segment $g$ $\lambda_d(g)$ is based on these
-values.
-
-\subsection{Numeric Optimization}
-
-mu bounds
-
-bounds and time, bounds and convergence
-
-Same optima, but took longer
+\section{Additional Results and Figures}
+
+\subsection{Chromosome Fit Figures}
+\clearpage
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr2.pdf}
+  \label{suppfig:fit-chr1}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr3.pdf}
+  \label{suppfig:fit-chr3}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr4.pdf}
+  \label{suppfig:fit-chr4}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr5.pdf}
+  \label{suppfig:fit-chr5}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr6.pdf}
+  \label{suppfig:fit-chr6}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr7.pdf}
+  \label{suppfig:fit-chr7}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr8.pdf}
+  \label{suppfig:fit-chr8}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr9.pdf}
+  \label{suppfig:fit-chr9}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr10.pdf}
+  \label{suppfig:fit-chr10}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr11.pdf}
+  \label{suppfig:fit-chr11}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr12.pdf}
+  \label{suppfig:fit-chr12}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr13.pdf}
+  \label{suppfig:fit-chr13}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr14.pdf}
+  \label{suppfig:fit-chr14}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr15.pdf}
+  \label{suppfig:fit-chr15}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr16.pdf}
+  \label{suppfig:fit-chr16}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr17.pdf}
+  \label{suppfig:fit-chr17}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr18.pdf}
+  \label{suppfig:fit-chr18}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr19.pdf}
+  \label{suppfig:fit-chr19}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr20.pdf}
+  \label{suppfig:fit-chr20}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr21.pdf}
+  \label{suppfig:fit-chr21}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr22.pdf}
+  \label{suppfig:fit-chr22}
+\end{figure}
+
+\clearpage
 

### 2023-06-28 manuscript/supplementary.tex
supp cleanup

@@
 
+\subsection{Sources of Neutral Allele Frequency Change}
+
 Here, we step through a quick derivation of \textcite{Santiago1995-hx}.
@@
 \end{align}
-
+%
 where $k_{0,i}$ is the number of surviving gametes parent $i$ passes on, and
@@
 \end{align}
-
+%
 where $K_1$ and $H_1$ are the random change in neutral allele frequency change
@@
 Mendelian segregation in generation 1.
+
+\subsection{The Variance in Neutral Allele Frequency Change}
 
@@
 \end{align}
-
+%
 Assuming no correlation between parental gametes (e.g. no inbreeding),
@@
 \end{align}
-
+%
 Finally, we have
@@
   V_\text{WF} &= \V(\Delta p_1) \nonumber \\
-  N_e &= \frac{4N}{\V(k) + 2}
-\end{align}
-%
-(c.f. \cite{Wright1938-tv}). Now, we look at the variance in allele frequency
-change across two generations for a neutral allele. The change in neutral
-allele frequency, regardless of the selective system, is directionless in
-expectation. This is because whichever allele we track is arbitrary by the
-definition of neutrality, which imposes a symmetry, $\E(\Delta x) = 0$. Then, the
-variance of change in the first generation is,
+  N_e &= \frac{4N}{\V(k) + 2} \label{supp-eq:var-ne}
+\end{align}
+%
+(c.f. \cite{Wright1938-tv}).
+
+Now, we look at the variance in allele frequency change across two generations
+for a neutral allele. The change in neutral allele frequency, regardless of the
+selective system, is directionless in expectation. This is because whichever
+allele we track is arbitrary by the definition of neutrality, which imposes a
+symmetry, $\E(\Delta x) = 0$. Then, the variance of change in the first
+generation is,
 
@@
 covariance term $\mathcal{C}_{1,2} = \E_1\left[\E_2[\Delta x_2 \Delta x_1 |
-x_1]\right] = 0$ under neutral evolution, and this simplifies to the well-known
-equation for variance in a Wright--Fisher population.
-
-Now, we turn our attention to the case where there is heritable fitness
+x_1]\right] = 0$ when there is no heritable fitness variation, and this
+simplifies to the well-known equation for variance in a Wright--Fisher
+population.
+
+\subsection{Heritable Fitness and the Accumulation of Autocovariance}
+
+Next, we turn our attention to the case where there is heritable fitness
 variation, which leads to non-zero covariance terms $\mathcal{C}_{i,j}$. These
@@
 allele's trajectory in later generations. If we partition the offspring number
-into heritable and non-heritable components, $k_i = 2f_i + \Vepsilon_i$, then
-$\V(k_i) = 4 V_h + V_n$. When there is heritable fitness variation across
-individuals, $V_h > 0$. Because the population size is assumed to be constant,
+into heritable and non-heritable components, $k_i = 2f_i + \varepsilon_i$, then
+$\V(k_i) = 4 V_A + V_n$. When there is heritable fitness variation across
+individuals, $V_A > 0$. Because the population size is assumed to be constant,
 the mean heritable fitness across individuals is constrained to be $\E_i(f_i) =
@@
 fitnesses as is standard in population genetics, in this case to the population
-mean. Thus $V_h$ is the squared coefficient of heritable fitness variation
+mean. Thus $V_A$ is the squared coefficient of heritable fitness variation
 (this is often denoted as $C^2 = \nicefrac{V_A}{\bar{w}^2}$, see
@@
   \V(\Delta x_1) &= \V(S_1) + \V(D_1) + \V(H_1) \nonumber \\
-                   &\approx \frac{p_0(1-p_0)}{2N} V_h + \frac{p_0(1-p_0)}{2N} \frac{V_n}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2} \label{supp-eq:varterms} 
+                   &\approx \frac{p_0(1-p_0)}{2N} V_A + \frac{p_0(1-p_0)}{2N} \frac{V_n}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2} \label{supp-eq:varterms} 
 \end{align}
@@
 \textcite{Santiago1995-hx} assume an equilibrium level of fitness variation
-$V_h$, with the fitness variance associated with a particular background
+$V_A$, with the fitness variance associated with a particular background
 decaying at a simple geometric decay at rate $1-\kappa$ each generation, which
@@
 With these covariance terms, we now have a full model for $\V(p_t - p_0)$. We
-will look at the case for $t=2$. As shown in equation \eqref{supp-eq:varterms},
+will look at the case for $t=2$. As shown in Supplementary Equation \eqref{supp-eq:varterms},
 the second moment of every term is a function of the variance in neutral allele
@@
                   & + \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) \nonumber \\
-  \frac{\V(x_2 - p_0)}{p_0(1-p_0)} &= \frac{V_h}{2N_{e,1}} + \frac{V_h Q_2^2}{2N_{e,2}}\left(1-\frac{1}{2N_{e,1}}\right)  +  \frac{V_h}{2N_{e,2}}  \left(1-\frac{1}{2N_{e,1}}\right)  \nonumber \\ 
-                                     &+ \frac{V_n}{8N_{e,1}} + \frac{1}{4N_{e,1}} + \frac{V_n}{8N_{e,2}}  \left(1 - \frac{1}{2N_{e,1}}\right) 
-+ \frac{1}{4N_{e,2}} \left(1 - \frac{1}{2N_{e,1}}\right).
+  \frac{\V(x_2 - p_0)}{p_0(1-p_0)} &= \frac{V_A}{2N} + \frac{V_A Q_2^2}{2N}\left(1-\frac{1}{2N}\right)  +  \frac{V_A}{2N_}  \left(1-\frac{1}{2N}\right)  \nonumber \\ 
+                                     &+ \frac{V_n}{8N} + \frac{1}{4N} + \frac{V_n}{8N}  \left(1 - \frac{1}{2N}\right) 
++ \frac{1}{4N} \left(1 - \frac{1}{2N}\right).
 \end{align}
@@
 \begin{align}
-  \label{eqn:var_xt}
-  \frac{\V(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^t \left(\frac{V_h Q_i^2}{2 N_{e,i}} + \frac{V_n}{8N_{e,i}} + \frac{1}{4N_{e,i}} \right)\prod_{j=1}^{i-1} \left(1-\frac{1}{2 N_{e,j}}\right).
-\end{align}
-
-We will see later that with the correct form for $Q_t^2$, this equation holds
-for polygenic selection (see Supplementary Section \ref{supp:ne-polygenic}).
-Additionally, we note that since the probability of coalescence (or,
-equivalently, identity by descent) is proportional to the variance in allele
-frequency change \parencite{Barton2000-zg}, this equation for the variance
-encodes the pairwise coalescent rate of the population through time under weak
-draft in the $N_{e,t}$ terms. Others have found that the genealogies under
-purifying selection can be characterized by such a time-dependent effective
-population size \parencite{Nicolaisen2013-gv}.
+  \label{supp-eq:var-xt}
+  \frac{\V(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^t \frac{1}{2 N} \left(V_A Q_i^2 + \frac{V_n}{4} + \frac{1}{2} \right) \left(1-\frac{1}{2 N}\right)^{i-1}.
+\end{align}
+%
+We will see later that with the appropriate form for $Q_t^2$, this equation
+holds for polygenic selection (see Supplementary Section
+\ref{supp:ne-polygenic}). Additionally, we note that since the probability of
+coalescence (or, equivalently, identity by descent) is proportional to the
+variance in allele frequency change \parencite{Barton2000-zg}, this equation
+for the variance encodes the pairwise coalescent rate of the population through
+time under weak draft. Others have found that the genealogies under purifying
+selection can be characterized by such a time-dependent effective population
+size \parencite{Nicolaisen2013-gv}.
+
+\subsection{The Draft-Effective Population Size}
 
 Now, we consider how to turn this expression for the variance in neutral allele
-frequency at time $t$ under weak draft into a \emph{draft-effective} population
-size. In an ideal Wright--Fisher population, the variance after $t$ generations
-is
-
-\begin{align}
-    \label{supp-eq:drift-vt}
-    \V(x_t) &= p_0(1-p_0) \left[1 - \left(1 - \frac{1}{2N_e}\right)^t \right].
-\end{align}
-
-To find an analogous expression for the asymptotic \emph{draft-effective}
-population size $N_d$, we let $N_e = N_d$ in the above equation, and consider
-the change in variance between generations $t$ and $t-1$, $\V(x_t) -
-\V(x_{t-1})$. Using Supplementary Equation \eqref{supp-eq:drift-vt}, we have,
-
-\begin{align}
-    \V(x_t) - \V(x_{t-1}) = p_0(1-p_0) \left[ \left(1 - \frac{1}{2N_e}\right)^{t-1} - \left(1 - \frac{1}{2N_e}\right)^t  \right]
-\end{align}
-
-\textcite{Santiago1995-hx} assume the change in population size per generation
-is small enough that the coalescence term each generation is constant, i.e. $(1
-- \nicefrac{1}{2N_{e,t}}) \approx (1 - \nicefrac{1}{2N_e})$ for each $t$.
-
-\begin{align}
-    \frac{\V(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^t \left(\frac{V_h Q_i^2}{2 N_e} + \frac{V_n}{8N_e} + \frac{1}{4N_e} \right)\left(1-\frac{1}{2 N_{e}}\right)^{i-1}.
-\end{align}
-
-One
-
-Then, imagine the process burns in for $t$ generations and asking what the
-equivalent
-
-\begin{align}
-    \V(x_t) &= \V(x_{t-1})
-
-  p_0(1-p_0) \left[1 - \left(1 - \frac{1}{2N_d}\right)^t \right].
-\end{align}
-
-\begin{align}
-  \label{eqn:ne_var}
-  N_{d,t} = \frac{p_0(1-p_0) - \V(x_{t-1})}{2(\V(x_t) - \V(x_{t-1}))}
-\end{align}
-%
-(cf. p. 1018, \cite{Santiago1995-hx}).
-
-XXX here
-
-\textcite{Santiago1995-hx} assume the change in population size per generation
-is small enough that the coalescence term each generation is constant, i.e. $(1
-- \nicefrac{1}{2N_{e,t}}) \approx (1 - \nicefrac{1}{2N_e})$ for each $t$.
-
-\begin{align}
-    \frac{\V(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^t \left(\frac{V_h Q_i^2}{2 N_{e,i}} + \frac{V_n}{8N_{e,i}} + \frac{1}{4N_{e,i}} \right)\left(1-\frac{1}{2 N_{e}}\right)^{i-1}.
-\end{align}
-
-\begin{align}
-  \label{eq:var_asymp}
-  \frac{\V(x_t - p_0)}{p_0(1-p_0)} &= \frac{4 V_h Q_\infty^2 +V_n+2}{4 (2 N-1)}.
-\end{align}
-
-Note here, to differentiate from the drift-effective population size $N_e$ due
-under non-selective causes, we use $N_d$ as the \emph{draft-effective}
-population size. Then, if we look at the change in the variance in allele
-frequency change between adjacent generations $t$ and $t-1$, we get an implied
-$N_{d,t}$
-
----
-Solving this for equation \eqref{eq:var_asymp} gives the asymptotic effective
-population size,
-
-\begin{align}
-  %N_e &= \frac{2 (2 N-1)}{4 V_h Q_\infty^2 +V_n+2} \\
-  N_d &\approx \frac{4N}{4 V_h Q_\infty^2 +V_n+2}
-\end{align}
-%
-(cf. equation 18, \cite{Santiago1995-hx}). Note if $V_h = 0$, this equation
+frequency at time $t$ under weak draft into an expression for the
+\emph{draft-effective} population size analogous to Supplementary Equation
+\eqref{supp-eq:var-ne}. In an ideal Wright--Fisher population, the
+instantaneous variance is $\V(\Delta x_t) =
+\nicefrac{p_{t-1}(1-p_{t-1})}{2N_{d,t-1}}$. Santiago and Caballero consider the
+allele frequency change at some $t$ to define their effective population size
+at time $t$ (which we here are calling draft-effective population size,
+$N_{d,t}$),
+
+\begin{align}
+    N_{d,t} := \frac{p_{t-1} (1-p_{t-1})}{2\V(x_t) - \V(x_{t-1})}.
+\end{align}
+%
+The time-dependency is due to $Q_i^2$, which reflects the fact that
+autocovariances build up, but at a decreasing rate. Note that levels of
+pairwise diversity depend on the full distribution of pairwise coalescence
+times, or the $N_{d,t}$ (e.g. see equation 4 of \cite{Santiago2016-mu} and
+equation 16 of \cite{Santiago1998-bs}).
+
+Using Supplementary Equation \eqref{supp-eq:var-xt} we have,
+
+\begin{align}
+    \label{supp-eq:var-ndt}
+    N_{d,t} &\approx \frac{4N}{4 V_A Q_t^2 +V_n+2}
+\end{align}
+%
+(cf. equation 18, \cite{Santiago1995-hx}). Note if $V_A = 0$, this equation
 reduces to Wright's (\citeyear{Wright1938-tv}) equation.
 
-Levels of pairwise diversity, however, depend on the full distribution of
-pairwise coalescence times, or the $N_{d,t}$. While this would require solving the
-recursion defined by equations \eqref{eqn:var_xt} and \eqref{eqn:ne_var},
-\textcite{Santiago1998-bs} show that a reasonably good approximation follows
-from using a single effective population size for all timepoints, $N_{d,t} =
-N_d$, but use the time-indexed $Q_t$. This leads to,
-
-\begin{align}
-  \label{eq:ne_t}
-  N_{d,t} &\approx \frac{4N}{4 V_h Q_t^2 +V_n+2}
-\end{align}
-%
-(c.f. equation 15 of \cite{Santiago1995-hx} and p. 1270 of \cite{Santiago2016-mu}).
-
-\subsection{Effective Population Size Under Polygenic Selection}
+In many cases, we care about the asymptotic $N_d$,
+
+\begin{align}
+    N_d := \lim_{t \to \infty} N_{d,t} \approx \frac{4N}{4 V_A Q_\infty^2 +V_n+2}.
+\end{align}
+
+We now turn to different expressions for $Q_t^2$ and $V_A$.
+
+\subsection{The Draft-Effective Population Size Under Polygenic Selection}
 \label{supp:ne-polygenic}
@@
 variety of different equilibrium selection systems as long as expressions for
-$V_h$ and $1-\kappa$ can be worked out. Other authors have extended similar models to
+$V_A$ and $1-\kappa$ can be worked out. Other authors have extended similar models to
 more elaborate bbreeding structures \parencite{Wray1990-zf,Woolliams1993-qo}.
 
-First, we extend equation \eqref{eq:ne_t} to a polygenic system. Throughout, a
-multiplicative fitness model is assumed (i.e. independent effects, no
-epistasis), where the fitness of individual $i$ is the product of their fitness
-contributions from each locus $l$, $w_{i,l}$, giving $w_i = \prod_{l=1}^n
-w_{i,l}$. Then, assuming mean fitness is one and independence between sites,
-the total fitness variation is multiplicative across the locus-level fitnesses,
-
-\begin{align}
-  V_h %&= \E_i\left( w_i^2 \right) - 1 \\
+First, we extend Supplementary Equation \eqref{supp-eq:var-ndt} to a polygenic
+system. Throughout, we assume a multiplicative fitness model (i.e. independent
+effects, no epistasis), where the fitness of individual $i$ is the product of
+their fitness contributions from each locus $l$, $w_{i,l}$, giving $w_i =
+\prod_{l=1}^n w_{i,l}$. Then, assuming mean fitness is one and independence
+between sites, the total fitness variation is multiplicative across the
+locus-level fitnesses,
+
+\begin{align}
+  V_A %&= \E_i\left( w_i^2 \right) - 1 \\
       %&= \prod_{l=1}^n (1 + \alpha_{l}^2 \V(g_l))  - 1  \\
@@
 
-Similarly, the $Q_t^2 V_h$ term in the single-locus model is multiplicative
+Similarly, the $Q_t^2 V_A$ term in the single-locus model is multiplicative
 under the polygenic model, as the change in site-specific fitness variances are
@@
 \begin{align}
-  Q_t^2 V_h &= \prod_{l=1}^n \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)  - 1
+  Q_t^2 V_A &= \prod_{l=1}^n \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)  - 1
 \end{align}
@@
 \begin{align}
-  N_{d,t} &\approx \frac{N}{V_h Q_t^2 + 1} \nonumber \\
+  N_{d,t} &\approx \frac{N}{V_A Q_t^2 + 1} \nonumber \\
           &\approx \frac{N}{\prod_{l=1}^n \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)} \nonumber \\
@@
 contribution to the variance as a deviation from the average contribution,
-$V_{h,l}^2 = \nicefrac{1}{n} V_h + \Vepsilon_l$, then
-
-\begin{align}
-  N_{d,t} &\approx N \exp\left(- \frac{1}{2}\sum_{l=1}^n \left(\frac{1}{n}V_{h} + \Vepsilon_l\right)  Q_{t,l}^2\right) \nonumber \\
-          &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{t,l}^2+ \sum_{l=1}^n \Vepsilon_l  Q_{t,l}^2\right)
+$V_{h,l}^2 = \nicefrac{1}{n} V_A + \varepsilon_l$, then
+
+\begin{align}
+  N_{d,t} &\approx N \exp\left(- \frac{1}{2}\sum_{l=1}^n \left(\frac{1}{n}V_{h} + \varepsilon_l\right)  Q_{t,l}^2\right) \nonumber \\
+          &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{t,l}^2+ \sum_{l=1}^n \varepsilon_l  Q_{t,l}^2\right)
 \end{align}
@@
 fitness variation at a site and the decay of its association with the neutral
-site, e.g. $\E(\Vepsilon_l Q_{t,l}^2) = 0$, and thus ignore the second sum
+site, e.g. $\E(\varepsilon_l Q_{t,l}^2) = 0$, and thus ignore the second sum
 term, leading to
@@
 contribution of a single segment under purifying selection, and the reduction
-from additional segments accumulate multiplicatively (similar to our
-implementation of the classic BGS model, XXX). Like their model, we imagine the
-reduction experienced by a focal neutral site in the middle of the segment.
-The implementation of our method differs slightly on though, see XXX.
-
-Since $Q_{t,l}$ terms depend on $l$ only through the recombination fraction
-between the neutral allele and the selected site, $r(l)$. Assuming a constant
+from additional segments accumulate multiplicatively. Like their model, we
+imagine the reduction experienced by a focal neutral site in the middle of the
+segment. The implementation of our method differs slightly on though, see XXX.
+
+The $Q_{t,l}$ terms depend on $l$ only through the recombination fraction
+$r(l)$ between the neutral allele and the selected site. Assuming a constant
 per-basepair recombination rate of $r_\text{BP}$, $r(l) = r_\text{BP} l$ we can
@@
 \begin{align}
-      N_{d,\infty} &\approx N \exp\left(- \frac{V_{h}}{2L} \sum_{l=1}^L Q_{\infty,l}^2\right) \\
-              &\approx N \exp\left(- \frac{V_{h}}{M} \int_0^{\nicefrac{M}{2}} Q_{\infty}(r)^2 dr \right) \\
-\end{align}
-
+      N_{d,\infty} &\approx N \exp\left(- \frac{V_{h}}{2L} \sum_{l=1}^L Q_{\infty,l}^2\right) \nonumber \\
+              &\approx N \exp\left(- \frac{V_{h}}{M} \int_0^{\nicefrac{M}{2}} Q_{\infty}(r)^2 dr \right) 
+\end{align}
+%
 where $M = n r_\text{BP}$ is the total segment length in Morgans. Then, the
@@
 \begin{align}
-  Q_\infty^2 &= \frac{2}{M}\int_0^{\nicefrac{M}{2}} Q_{\infty}(r)^2 dr  \\
-  &=  \frac{2}{M} \int_0^{\nicefrac{M}{2}} \left(\frac{1}{1 - (1-r)Z}\right)^2 dr  \\
+  Q_\infty^2 &= \frac{2}{M}\int_0^{\nicefrac{M}{2}} Q_{\infty}(r)^2 dr  \nonumber \\
+  &=  \frac{2}{M} \int_0^{\nicefrac{M}{2}} \left(\frac{1}{1 - (1-r)Z}\right)^2 dr. \nonumber  \\
   &= \frac{2}{(1-Z) (2-(2-M) Z)}
 \end{align}
-
+%
 Santiago and Caballero further approximate this; we can derive their
@@
 \end{align}
-
+%
 Santiago and Caballero's approximation only keeps this first term, e.g. their
@@
 \begin{align}
-  N_{d,\infty}^{\text(SC98)} \approx N \exp\left(-\frac{V_h}{(1-Z)M}\right) 
-\end{align}
-
+  N_{d,\infty}^{\text(SC98)} \approx N \exp\left(-\frac{V_A}{(1-Z)M}\right) 
+\end{align}
+%
 (c.f. \cite{Santiago1998-bs} equation 8). For small $M$, the small $\kappa$
@@
   \label{eq:Ne_inf}
-  N_{d,\infty} \approx N \exp\left(-\frac{V_h}{(1-Z) (2-(2-M) Z)} \right).
-\end{align}
-
+  N_{d,\infty} \approx N \exp\left(-\frac{V_A}{(1-Z) (2-(2-M) Z)} \right).
+\end{align}
+%
 For the reasons described earlier, a single asymptotic $N_{d,\infty}$ cannot
@@
   \label{eq:net_integral}
-  N_{d,t} \approx N \exp\left(-\frac{V_h}{M} \int_0^{\nicefrac{M}{2}} \left(1+\frac{(1-k) (1-r) \left(1-(1-k)^t (1-r)^t\right)}{k+(1-k)r} \right) dr \right).
+  N_{d,t} \approx N \exp\left(-\frac{V_A}{M} \int_0^{\nicefrac{M}{2}} \left(1+\frac{(1-k) (1-r) \left(1-(1-k)^t (1-r)^t\right)}{k+(1-k)r} \right) dr \right).
 \end{align}
@@
 
-The heritable variance in fitness $V_h$ associated with the tracked neutral
+The heritable variance in fitness $V_A$ associated with the tracked neutral
 allele does not remain constant through the generations, but decays due to
@@
 the trait response to selection depend on the variance, which in turn depends
-on higher moments of the fitness distribution. We discuss the simplification of
-Santiago and Caballero's model here; hereafter, we assume that $V_h = V_A$, the
-genetic variance in additive fitness put on a relative scale such that
-$\E_i(f_i) = 1$.
+on higher moments of the fitness distribution (the ``moment-closure problem").
+We discuss the simplification of Santiago and Caballero's model here.
 
@@
   \label{eq:Z}
-  V_A' &= \left(1 - \frac{V_m}{V_A}\right) V_A \\
+  V_A' &= \left(1 - \frac{V_m}{V_A}\right) V_A \nonumber \\
   Z &:= \frac{V_A'}{V_A} = 1 - \kappa = 1 - \frac{V_m}{V_A} 
 \end{align}
-
+%
 (c.f. \cite{Santiago1998-bs} equation 9).
@@
 
-\subsection{Effective Population Size Under Strong Background Selection}
+\subsection{The Draft-Effective Population Size Under Strong Background Selection}
 
@@
 \end{align}
-
+%
 where $n$ is the number of deleterious mutations this individual carries.
@@
   \label{eq:va_bgs}
-  V_\text{BGS} &:= \V(w) = 2 s^2 \sum_{l=1}^L p_l(1-p_l)  \\
+  V_\text{BGS} &:= \V(w) = 2 s^2 \sum_{l=1}^L p_l(1-p_l) \nonumber  \\
             &= 2 L s^2 \frac{\mu}{s} \left( 1 - \frac{\mu}{s}\right) = U s + \mathcal{O}(\mu^2) \approx Us.
 \end{align}
-
+%
 since under mutation-selection balance $p_l = \nicefrac{\mu}{s}$ and $U$ is
@@
 \end{align}
-
+%
 (c.f. \cite{Santiago1998-bs} equation 10). An alternate derivation can be found
@@
 \end{align}
-
+%
 This matches equation 8 of \textcite{Hudson1995-xc} except for the $Ms$ term
@@
 
-\subsection{Effective Population Size Under Weak and Strong BGS}
+\subsection{The Draft-Effective Population Size Under Weak and Strong BGS}
 \label{supp:weak-strong}
@@
 So far, finding an explicit equation for the variance has been difficult due to
-the ``moment-closure problem": the variance depends on higher moments such as
-the skew, which in turn depends on higher moments. While the weak selection
-regime is composed of an interaction of mutation, selection, drift, and linkage
-processes and the dynamics that differ qualitatively based on parameter
-combinations. In their approximation, \textcite{Santiago2016-mu} consider the
-modified additive genic variance due to the fixation of weakly deleterious
-mutations. This fixation process is related to the Muller's Ratchet, the rate
-$R$ of fixation of deleterious mutations per generation per region, since each
-fixation directly reduces the additive genic variance of the fitness
-distribution. However, in general, the rate of this ratchet is unknown since it
-depends on the fitness variance (and thus higher moments of the fitness
-distribution).
+the ``moment-closure problem" previously mentioned. In their approximation,
+\textcite{Santiago2016-mu} consider a modified additive genic variance that is
+below $V_\text{BGS}$ due to the loss of variation from each fixation of a
+weakly deleterious mutation. This fixation process is related to the Muller's
+Ratchet, the rate $R$ of fixation of deleterious mutations per generation per
+region, since each fixation directly reduces the additive genic variance of the
+fitness distribution. However, in general, the rate of this ratchet is unknown
+since it depends on the fitness variance (and thus higher moments of the
+fitness distribution).
 
@@
 
-In their 2016 paper, Santiago and Caballero derive a general equation for the
-fitness variance that accounts for the reduction due to the fixation of weakly
-deleterious alleles (i.e. the rate of Muller's ratchet). There derivation for
-the variance uses a heuristic approach based on Fisher's Fundamental Theorem of
-Natural Selection (equation 2, \cite{Santiago2016-mu}, equation 15A
-\cite{Garcia-Dorado2007-jj}). A more formal derivation follows from
-\textcite{Higgs1995-xc} (though they assume a haploid asexual system, the end
-result is the same up to a factor of two). In their work, they derive equations
-for the change per generation for the moments and cross-moments of the fitness
-distribution under multiplicative selection in a haploid model. They find, the
-haploid rate of the ratchet is $r = u - s v_n$ (equation 13.3,
-\cite{Higgs1995-xc}), where we use lowercase to distinguish between haploid and
-diploid models. Here, $v_n$ is the variance in the \emph{number} of deleterious
-mutations; as long as selection is not too strong, the haploid additive
-variance is $v_{A} \approx s^2 v_n$. Substituting this for $v_n$ and
-rearranging, the haploid fitness variance is related to the $v_A = us - rs$,
-which is identical to Santiago and Caballero's (\citeyear{Santiago2016-mu})
-equation 2, noting that the waiting time between fixations (i.e. ratchet
-clicks) is $T = \nicefrac{1}{r}$. Intuitively, the fitness variance $Us$ is
-reduced by the rate at which deleterious mutations fix and remove fitness
-variation from the population.
-
-Now, note that under a diploid model, the fitness variance is twice that as the
+In their 2016 paper, Santiago and Caballero derive use an approach based on
+Fisher's Fundamental Theorem of Natural Selection (equation 2,
+\cite{Santiago2016-mu}, equation 15A \cite{Garcia-Dorado2007-jj}). A more
+formal derivation follows from \textcite{Higgs1995-xc} (though they assume a
+haploid asexual system, the end result is the same up to a factor of two). In
+their work, they derive equations for the change per generation for the moments
+and cross-moments of the fitness distribution under multiplicative selection in
+a haploid model. They find, the haploid rate of the ratchet is $r = u - s v_n$
+(equation 13.3, \cite{Higgs1995-xc}), where we use lowercase to distinguish
+between haploid and diploid models. Here, $v_n$ is the variance in the
+\emph{number} of deleterious mutations; as long as selection is not too strong,
+the haploid additive variance is $v_{A} \approx s^2 v_n$. Substituting this for
+$v_n$ and rearranging, the haploid fitness variance is related to the $v_A = us
+- rs$, which is identical to Santiago and Caballero's
+(\citeyear{Santiago2016-mu}) equation 2, noting that the waiting time between
+fixations (i.e. ratchet clicks) is $T = \nicefrac{1}{r}$. Intuitively, the
+fitness variance $V_\text{BGS} = Us$ is reduced by the rate at which
+deleterious mutations fix and remove fitness variation from the population.
+
+Now, note that under a diploid model, the fitness variance is twice that of the
 haploid model, so $V_A = 2 v_A$. Then, the fitness variance is,
@@
 \end{align}
-
+%
 where $U = 2u$ and $R=r$ is the diploid rate of the ratchet, which depends on
 higher moments of the fitness distribution, but only differs from $r$ through a
-rescaling of the population size. Note that if selected sites are sufficiently
-deleterious that they cannot fix (or the population is infinite), $R=0$ and the
-additive fitness variance is identical to that under strong background
-selection, $V_A = Us = V_\text{BGS}$ (equation \eqref{eq:va_bgs}). Then, the
-fixation probability of a new deleterious mutation with additive effects ($1-s$
-in the heterozygote, $1-2s$ in the homozygote) in a diploid population is,
-
-\begin{align}
-  p_\text{fix} = \frac{1-\exp(\nicefrac{2 N_e s}{N})}{1-\exp(4 N_e s)} \approx
+rescaling of the population size. We also find that Equation \eqref{eq:Va} and
+Supplementary Materials Equation \ref{supp-eq:Va} above can also be found
+through Kimura's diffusion models with a flux of mutations into infinite,
+discrete sites (\citeyear{Kimura1969-jw}). Note that if selected sites are
+sufficiently deleterious that they cannot fix (or the population is infinite),
+$R=0$ and the additive fitness variance is identical to that under strong
+background selection, $V_A = Us = V_\text{BGS}$ (equation \ref{eq:va_bgs}).
+We note that while we use $V_A$ here following \textcite{Santiago2016-mu}, 
+this is only the \emph{additive genic} component of additive fitness variation,
+as our simulations confirm in the main text.
+
+Next, we turn to the fixation probability of a new deleterious mutation with
+additive effects ($1-s$ in the heterozygote, $1-2s$ in the homozygote). In a
+diploid population with drift-effective population size $N_e$ this is,
+
+\begin{align}
+  p_\text{fix} := \frac{1-\exp(\nicefrac{2 N_e s}{N})}{1-\exp(4 N_e s)} \approx
   \frac{2 N_e s}{N \left(\exp(4 N_e s)-1\right)}
 \end{align}
-
+%
 (c.f. \cite{Durrett2008-ql}, equation 7.21; \cite{Kimura1957-rk}, equation
@@
 \begin{align}
-  T \approx \frac{\exp(4 N_e s)-1}{2 U s N_e}
-\end{align}
-
+  T \approx \frac{\exp(4 N_e s)-1}{2 U s N_e}.
+\end{align}
+%
 Since the ratchet rate depends on $N_e$, which in turn depends on the rate of
-the ratchet, we write the ratchet, or substitution, rate per basepair per
-generation as,
+the ratchet, we write the ratchet, or substitution, rate per $L$-basepair
+region per generation as,
 
@@
 \end{align}
-
+%
 Under weak and strong background selection the variance input of mutations
@@
 \end{align}
-
+%
 At equilibrium, the $N_e$ that determines fixation probability is the
 asymptotic $N_{d,\infty}$. Consequently, Santiago and Caballero argue that the
-asymptotic reduction in effective population size uses the $Q_\infty^2$ and is
+asymptotic reduction in draft-effective population size uses the $Q_\infty^2$ and is
 given by the following system of non-linear equations,
@@
 \end{align}
-
-The solution to these equations is the asymptotic equilibrium $\widetilde{N}_e$
+%
+The solution to these equations is the asymptotic equilibrium $\widetilde{N}_d$
 and equilibrium ratchet rate $\widetilde{R}$. This sets the equilibrium
 variance, $\widetilde{V} = Us - 2\widetilde{R}s$, or $\widetilde{V} = (U -
-2\widetilde{R})s$. Note that because $p_\text{fix} \le \nicefrac{1}{2N}$ for
-deleterious to neutral variation, $2R \le U$, or the diploid mutation rate must
-be greater than twice the per-basepair per-generation substitution rate for
+2\widetilde{R})s$. Note that because $p_\text{fix} < \nicefrac{1}{2N}$ for
+deleterious variation, $2R < U$, or the diploid mutation rate must be greater
+than or equal to twice the per-basepair per-generation substitution rate for
 equilibrium variance $\widetilde{V} > 0$ for $s>0$.
-
-\subsection{Steady-state Equilibrium Genic Variance} \label{supp:genic-va}
-
-We also find that Equation \eqref{eq:Va} and Supplementary Materials Equation
-\ref{supp-eq:Va} above can also be found through Kimura's diffusion models with
-a flux of mutations into infinite, discrete sites (\citeyear{Kimura1969-jw}).
-This model integrates the Kolmogorov backwards equation over time and forward
-variable to find the expected value of a functional of the initial frequency.
-Using the functional $I(p) = s^2 p(1-p)$ for genic variance given mutations
-start at frequency $p$, one finds that the steady-state genic variance for
-$p=\nicefrac{1}{2N_e}$ is
-
-\begin{align}
-    V_a^{SS} &= 2 U N s \left(\frac{1}{2N_e} - p_F(N_e, s)\right) \\
-             &= \left(\frac{UN}{N_e} - 2 U Np_F(N_e, s))\right)s \nonumber \\
-             &\approx \left(U - 2 R\right)s \nonumber
-\end{align}
-%
-which is approximately Equation \eqref{eq:Va} in the main paper.
-
-% \subsection{Misspecification of the Polygenic Selection Model}
-
-% Thus far, we have developed the general model of equation \eqref{XXX} by
-% assuming that the only source of $V_h$ is from the deleterious
-% mutation-selection-draft-drift process. However, other polygenic evolutionary
-% processes can also create, alter, and eliminate fitness variation. By assuming
-% that the only fitness variation originates from deleterious
-% mutation-selection-drift-draft processes, we introduce some model
-% misspecification. There are two main types of misspecification under the
-% selection model. First, non-polygenic positive selection processes, such as
-% recurrent sweeps, may create small-scale distortions in local genealogies.
-% These have been found to only impact diversity on small spatial scales
-% \parencite{Murphy2022-sj} in humans, and explain little of the total variation
-% in megabase-scale patterns of diversity. Second, there is model
-% misspecification due to our model assuming that the only polygenic selection
-% process happening is the flux of new deleterious mutations in, and their
-% eventual loss or fixation. Other polygenic selection processes could be
-% occurring, e.g. traits could be responding to directional selection. The
-% Santiago and Caballero model our work extends is very general, and can account
-% for multiple types of polygenic selection. The key input parameters are the
-% equilibrium heritable fitness variation $V_h$, and the variation introduced by
-% mutation each generation $V_m$.
-
-% Let us assume for now that,
-
-% \begin{align}
-%   \label{eq:Ne_inf}
-%   N_{e,\infty} \approx N \exp\left(-\frac{V_h + \Vepsilon}{(1-Z) (2-(2-M) Z)} \right).
-% \end{align}
 
@@
 as the predicted level under our negative selection model, $\pi_0 B_i$, plus
-some random residual error $\Vepsilon_i$,
-
-\begin{align}
-    y_i = \pi_0 B_i + \Vepsilon_i
+some random residual error $\varepsilon_i$,
+
+\begin{align}
+    y_i = \pi_0 B_i + \varepsilon_i
 \end{align}
@@
     V_\text{res} &= \frac{1}{n} \sum_{i=1}^n (y_i - \widehat{\pi}_0 \widehat{B}_i)^2  \\
-                 &= \underbrace{\frac{1}{n} \sum_{i=1}^n \Vepsilon_i^2}_\text{irreducible error} + \underbrace{\left(\frac{1}{n} \sum_{i=1}^n (\pi_0 B_i - \widehat{\pi}_0 \widehat{B}_i)\right)^2}_\text{bias squared}.
+                 &= \underbrace{\frac{1}{n} \sum_{i=1}^n \varepsilon_i^2}_\text{irreducible error} + \underbrace{\left(\frac{1}{n} \sum_{i=1}^n (\pi_0 B_i - \widehat{\pi}_0 \widehat{B}_i)\right)^2}_\text{bias squared}.
 \end{align}
@@
 \begin{align}
-    V_\text{tot} = V_\text{res} + V_\text{model} + 2\widehat{\pi}_0\cov_i(\Vepsilon_i, \widehat{B}_i)
+    V_\text{tot} = V_\text{res} + V_\text{model} + 2\widehat{\pi}_0\cov_i(\varepsilon_i, \widehat{B}_i)
 \end{align}
@@
 Generally, fitting procedures act to minimize the squared deviation between
-predictors and observed values, which drives $\cov_i(\Vepsilon_i,
+predictors and observed values, which drives $\cov_i(\varepsilon_i,
 \widehat{B}_i)$ to zero XXX. \textcite{Murphy2022-sj} suggest  XXX
@@
 \begin{align}
-    \E(\Vepsilon_i^2) = \V(\pi_i)
+    \E(\varepsilon_i^2) = \V(\pi_i)
 \end{align}

### 2023-06-28 manuscript/supplementary.tex
supp cleanup

@@
 \newcommand{\E}{\mathbb{E}}
-\newcommand{\V}{\text{V}}
 \DeclareMathOperator{\var}{V}
@@
 
-\subsection{Sources of Neutral Allele Frequency Change}
+\subsection{Stochastic Sources of Neutral Allele Frequency Change}
 
@@
                   & + \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) \nonumber \\
-  \frac{\V(x_2 - p_0)}{p_0(1-p_0)} &= \frac{V_A}{2N} + \frac{V_A Q_2^2}{2N}\left(1-\frac{1}{2N}\right)  +  \frac{V_A}{2N_}  \left(1-\frac{1}{2N}\right)  \nonumber \\ 
+  \frac{\V(x_2 - p_0)}{p_0(1-p_0)} &= \frac{V_A}{2N} + \frac{V_A Q_2^2}{2N}\left(1-\frac{1}{2N}\right)  +  \frac{V_A}{2N}  \left(1-\frac{1}{2N}\right)  \nonumber \\ 
                                      &+ \frac{V_n}{8N} + \frac{1}{4N} + \frac{V_n}{8N}  \left(1 - \frac{1}{2N}\right) 
@@
 frequency at time $t$ under weak draft into an expression for the
-\emph{draft-effective} population size analogous to Supplementary Equation
-\eqref{supp-eq:var-ne}. In an ideal Wright--Fisher population, the
-instantaneous variance is $\V(\Delta x_t) =
+\emph{draft-effective} population size analogous to drift-effective population
+size given in Supplementary Equation \eqref{supp-eq:var-ne}. In an ideal
+Wright--Fisher population, the instantaneous variance is $\V(\Delta x_t) =
 \nicefrac{p_{t-1}(1-p_{t-1})}{2N_{d,t-1}}$. Santiago and Caballero consider the
@@
 \end{align}
-
-We now turn to different expressions for $Q_t^2$ and $V_A$.
+%
+Now, we turn to different expressions for $Q_t^2$ and $V_A$ under different
+types of polygenic selection.
 
@@
 imagine the reduction experienced by a focal neutral site in the middle of the
-segment. The implementation of our method differs slightly on though, see XXX.
+segment.
 
@@
 
-\subsection{The Draft-Effective Population Size Under Weak and Strong BGS}
+\subsection{The Draft-Effective Population Size Under Weak and Strong Negative Selection}
 \label{supp:weak-strong}
@@
 rescaling of the population size. We also find that Equation \eqref{eq:Va} and
-Supplementary Materials Equation \ref{supp-eq:Va} above can also be found
+Supplementary Equation \ref{supp-eq:Va} above can also be found
 through Kimura's diffusion models with a flux of mutations into infinite,
@@
 \begin{align}
+    \label{supp-eq:sc16-sys}
   R(N_{d,\infty}) &= \frac{4 U s N_{d,\infty}}{\exp(4 N_{d,\infty} s)-1}\\
@@
 
+The previous section outlined the basis of quantitative genetic linked
+selection models that predict the draft-effective population size under
+polygenic selection and linkage. Our work here only considers the contribution
+of strongly and weakly deleterious mutations to fitness variation. In this
+section, we outline more details about the statistical models used to construct
+the genome-wide B and B' maps that are used in model fitting.
+
 \subsection{Parametric Forms of Negative Linked Selection Models}
 
-Both classic BGS and the B' method we build off the SC16 model have the same
-parametric form, since both (1) model the impact of linked selection as a
-rescaling of effective population size, and (2) both assume multiplicative
-fitness across segments. Thus, the total reduction experienced by a neutral
-site at position $x$ due to segments across the genome is the product of the
-individual reductions is
+Both the classic BGS and the SC16 model we extend have the same parametric
+form, since both (1) model the impact of linked selection as a rescaling of
+effective population size, and (2) both assume multiplicative fitness across
+segments. Thus, the total reduction experienced by a neutral site at position
+$x$ due to segments across the genome is the product of the individual
+reductions,
 
@@
   \label{supp-eq:b}
-  B(x, \Psi) &= \prod_g^S \exp \left( \int_0^1 b(\mu(s, k(g)), s, L_g, r_g, d(x, g)) ds \right) \\
+  B(x, \Psi) &= \prod_g^S \exp \left( \int_0^1 b(\mu(s, k(g)), s, L_g, r_g, d(x, g)) ds \right) \nonumber  \\
              &= \exp \left( \sum_g^S \int_0^1 b(\mu(s, k(g)), s, L_g, r_g, d(x, g)) ds \right)
 \end{align}
-
+%
 where, $b(\mu(s, k(g)), s, L_g, r_g, d(x,g))$ is the log reduction due to
@@
 
-Under both background selection models used in our likelihood, the diversity in
-window $v$ is $\bar{\pi}(v, \Psi) = \bar{B}(v, \Psi) \pi_0$. Here, $\bar{B}(v |
-\Psi)$ is the predicted reduction in diversity due to BGS in window $v$, given
-background selection parameters $\Psi$. In practice, we pre-calculate $B(x |
-\Psi)$ at fixed sites $x$ across the genome, and take the average of the fixed
-sites $B$ values within window $v$ for the average $\bar{B}(v | \Psi)$.
-
-\subsubsection{Reductions Under the Classic BGS Model}
+Under both negative selection models implemented in our software, the diversity
+in window $v$ is $\bar{\pi}(v, \Psi) = \bar{B}(v, \Psi) \pi_0$. Here,
+$\bar{B}(v | \Psi)$ is the predicted reduction in diversity due to BGS in
+window $v$, given background selection parameters $\Psi$. In practice, we
+pre-calculate $B(x | \Psi)$ at fixed sites $x$ across the genome, and take the
+average of the fixed sites $B$ values within window $v$ for the average
+$\bar{B}(v | \Psi)$.
+
+\subsection{Discretization of Parameters and Annotation Feature Classes}
+
+In practice, to fit negative selection models, we must discretize both $\mu(s,
+k(g))$ and $s$ in Supplementary Equation \eqref{supp-eq:b}. We define
+$\mu_{i,k(g)}$ to be the rate of mutations entering the population per basepair
+per generation with selection coefficient $s_i$ for the feature class $k(g)$ of
+segment $g$. This $\mu_{i, k(g)}$ term reflects the product of the mutation
+rate and the distribution of deleterious fitness effects for selection
+coefficient $s_i$ and feature class $k(g)$. Then, the reduction at position $x$
+is,
+
+\begin{align}
+  \label{supp-eq:b-disc}
+  B(x, \Psi) = \exp \left( \sum_g^S \sum_{i}^{m_s} b(\mu_{i, \;k(g)}, s_i, L_g, r_g, d(x, g))\right).
+\end{align}
+%
+Because it is computationally-intensive to numerically solve the system of two
+non-linear equations to calculate $b$ at each segment, these are pre-computed
+for an $m_s$-element grid of selection coefficients and an $m_\mu$-element grid
+of mutation rates--DFE products. Likewise, when we calculate $b$ under classic
+BGS models, we pre-calculate segment components across these same grids, and
+combine them for each $x$.
+
+The number of annotation features $K$ is variable and set by the annotation
+data specified. There are two ways to parameterize the mutation rates and DFE
+for all annotation classes in background selection models. First, for the
+\emph{free-mutation} parameterization, each annotation feature class has a free
+mutation rate parameter for each of these selection coefficients (as in
+Supplementary Equation \ref{supp-eq:b}). We experimented with this approach,
+but ultimately found its results less intuitive and it has more degrees of
+freedom.
+
+% In the free-mutation parameterization, there are $1 + K \times m_s$ for $\pi_0$
+% and the free mutation rate parameters, where $K$ is the number of annotation
+% features. The MLE optimization for this approach is unconstrained, though
+% $\pi_0$ is bounded to be within $10^{-5} \le \pi_0 \le 10^{-1}$ and mutation
+% rates are bounded within $10^{-11} \le \mu \le 10^{-7}$. The BGS model requires
+% a mutation rate for each selection coefficient in the grid, for each annotation
+% class (here, as an example, CDS, UTRs, and phastcons), which are stored in the
+% $m_s \times K$ matrix $\mathbf{M}_\text{F}$,
+
+% \begin{align}
+%   \mathbf{M}_\text{F} = \begin{bmatrix}
+%     \mu_{10^{-6},\text{CDS}} & \mu_{10^{-6},\text{UTR}} & \mu_{10^{-6},\text{PC}} \\
+%     \mu_{10^{-5},\text{CDS}} & \mu_{10^{-5},\text{UTR}} & \mu_{10^{-5},\text{PC}} \\
+%     \vdots & \vdots & \vdots \\
+%     \mu_{10^{-1},\text{CDS}} & \mu_{10^{-1},\text{UTR}} & \mu_{10^{-1},\text{PC}} \\
+%   \end{bmatrix}
+% \end{align}
+
+% That is, the distribution of fitness effect (DFE) is implied by
+% the total mutation rate across selection coefficients for an annotation class.
+% We can normalize by the total mutation rate estimate to get the estimated DFE,
+% giving us the DFE weight for annotation class $k$ and selection coefficient
+% with index $i$,
+
+% \begin{align}
+%   \widehat{w}_{i,k} = \frac{\widehat{\mu}_{i,k}}{\sum_i \widehat{\mu}_{i,k}}.
+% \end{align}
+
+Instead, we opted for a \emph{simplex} parameterization, which has a single
+shared mutation rate across all features, so the mutation parameter $\mu_{i,
+\;k(g)}$ above is given by the product of the mutation rate and the DFE weight
+matrix. For example, for a $\log_{10}$ selection grid and three features, CDS,
+UTRs, and PhastCons (PC), we'd have a DFE matrix like,
+
+\begin{align}
+  \mathbf{M}_\text{S} = \mu \begin{bmatrix}
+    w_{10^{-6},\text{CDS}} & w_{10^{-6},\text{UTR}} & w_{10^{-6},\text{PC}} \\
+    w_{10^{-5},\text{CDS}} & w_{10^{-5},\text{UTR}} & w_{10^{-5},\text{PC}} \\
+    \vdots & \vdots & \vdots \\
+    w_{10^{-1},\text{CDS}} & w_{10^{-1},\text{UTR}} & w_{10^{-1},\text{PC}} \\
+  \end{bmatrix}
+\end{align}
+%
+which has $2 + K \times (m_s-1)$ free parameters, since each column must sum to
+one. This imposes the constraint that $\sum_i w_{i,k} = 1$, requiring
+constrained MLE optimization. We implement this using softmax, which describe 
+further in Supplementary Section \ref{supp:optim}.
+
+\subsection{Reductions Under the Classic BGS Model}
 \label{supp:reduction-classic}
@@
 the segment. To simplify notation, we will only consider a single feature class
-in this and the next section. Using our notation from equation
-\eqref{supp-eq:b}, the log reduction under the classic
+in this and the next section. The log reduction under the classic
 mutation-selection-balance BGS model is,
@@
 \begin{align}
-  b(\mu, s, L, r, 0) &=  - \int_0^L \frac{2\mu(l)}{s (1 + (1-s) r(l)/s)^2} dl 
-\end{align}
-
-(c.f. \cite{Hudson1995-xc} equation 5), where $\mu(l)$ is the \emph{haploid}
-per-basepair mutation rate at $l$ and $r(l)$ is the recombination fraction
-between the focal site (at position 0) and basepair $l$. Assuming constant
-per-basepair recombination and mutation rates, $r(l) = r_\text{BP}l$ and
-$\mu(l) = \mu$, and
-
-\begin{align}
-  b(\mu, s, L, r_\text{BP}, 0) &= - \frac{2\mu L}{s + (1-s) r_\text{BP} L} 
-\end{align}
-
-(c.f. \cite{Hudson1995-xc} equation 8, where the factor of two difference is
-due to the position of the focal site). Often per-region rates are defined $U =
-2 \mu L$ and $M = r_\text{BP} L$ are the segment-wide mutation diploid mutation
-rates (mutations per diploid segment per generation) and recombination length
-(Morgans).
+    \label{supp-eq:classic-b-int}
+  b(\mu, s, L, r, 0) &=  - \int_0^L \frac{\mu(l)}{s (1 + (1-s) r(l)/s)^2} dl 
+\end{align}
+% 
+(c.f. \cite{McVicker2009-ax} p. 11, \cite{Hudson1995-xc} equation 5,), where
+$\mu(l)$ is the \emph{haploid} per-basepair mutation rate at $l$ and $r(l)$ is
+the recombination fraction between the focal site (at position 0) and basepair
+$l$. Assuming constant per-basepair recombination and mutation rates, $r(l) =
+r_\text{BP}l$ and $\mu(l) = \mu$,
+
+\begin{align}
+  b(\mu, s, L, r_\text{BP}, 0) &= - \frac{\mu L}{s + (1-s) r_\text{BP} L} 
+\end{align}
+%
+(c.f. \cite{Hudson1995-xc} equation 8; their model considers a neutral site in
+the middle of the segment and drops small terms). Often per-region rates are
+defined $U = 2 \mu L$ and $M = r_\text{BP} L$ are the segment-wide mutation
+diploid mutation rates (mutations per diploid segment per generation) and
+recombination length (Morgans).
 
 For computational efficiency, we adjust this model by setting $r(l) = d +
-r_\text{BP} l$, where $d$ is the recombination fraction between the focal site
-$x$ and the start of segment $g$. This allows us to pre-compute the local
-effects of the segment, and substitute in $b$ as the focal site changes. This
-integral substitution has a closed form,
-
-\begin{align}
-  b(\mu, s, L, r_\text{BP}, d) &= \frac{-2\mu L}{(d(1-s) + s)(s + (d + r_\text{BP}L)(1-s))}
-\end{align}
-
-and terms can be collected in powers of $d$ and pre-computed for all segments
-for the grid of $\mu$ and $s$. Our implementation pre-computes these segment
-components, and computes the final $b$ value for an $x$ and $g$ by calculating
-the distance $d(x,g) = |M(x) - M(g)|$, where $M(x)$ is the cumulative
-recombination map length at position $x$ in Morgans. Since $r(l)$ could vary by
-position, segments with differing recombination rates are split into segments
-of the same annotation class by their recombination rate. This has no effect on
-the calculation due to the multiplicative fitness, but ensures more accurate
-estimates of the reduction factor $B$.
+r_\text{BP} l$ in Supplementary Equation \eqref{supp-eq:classic-b-int} and
+integrating, where $d$ is the recombination fraction between the focal site $x$
+and the start of segment $g$. This allows us to pre-compute the local effects
+of the segment, and substitute in $b$ as the focal site changes. This integral
+substitution has a closed form,
+
+\begin{align}
+  b(\mu, s, L, r_\text{BP}, d) &= - \frac{\mu L}{(d(1-s) + s)(s + (d + r_\text{BP}L)(1-s))}.
+\end{align}
+%
+Then, terms can be collected in powers of $d$ and pre-computed for all segments for
+the grid of $\mu$ and $s$. Our implementation of classic BGS in \texttt{bgspy}
+pre-computes these segment components, and then calculates the final $b$ value
+for an $x$ and $g$ by calculating the distance $d(x,g) = |M(x) - M(g)|$, where
+$M(x)$ is the cumulative recombination map length at position $x$ in Morgans.
+Since $r(l)$ could vary by position, segments with differing recombination
+rates are split in two of the same annotation class by their recombination
+rate. This has no effect on the calculation due to the multiplicative fitness,
+but ensures more accurate estimates of the reduction factor $B$.
 
@@
 
-\subsubsection{Reduction Under the Modified Santiago and Caballero Model}
+\subsection{Reduction Under the Modified Santiago and Caballero Model}
 \label{supp:reduction-sc16}
@@
 inflation factor $Q_t^2$ that sets the draft-effective population size varies
-over $t$ (c.f. Supplementary equation \ref{eq:net_integral}). This is because
-selective processes can lead to non-constant pairwise coalescent rates, which
-impact pairwise coalescent rates and thus the reduction factor. We experimented
-with calculating this full sum and integral approximations to it in our
-methods, but it seemed to make little difference in our application.
-Additionally, it was too costly to calculate computationally for genome-wide
-calculations of the B' map. Instead, we solve for the asymptotic
-$\widetilde{N}_{e,\infty}$ for each segment, and use that to calculate the log
-reduction factor under the SC16 model, $b'$.
+over $t$. This is because selective processes can lead to non-constant pairwise
+coalescent rates, which impact pairwise coalescent rates and thus the reduction
+factor. We experimented with calculating this full sum and integral
+approximations to it in our methods, but it seemed to make little difference in
+our application. Additionally, it was too costly to calculate computationally
+for genome-wide calculations of the B' map. Instead, we solve for the
+asymptotic $\widetilde{N}_{e,\infty}$ for each segment, and use that to
+calculate the log reduction factor under the SC16 model, $b'$.
 
@@
 at an arbitrary recombination distance apart. For each segment, we pre-compute
-the solution to the system of two equations, giving us $V$ and $Q_\infty^2$,
-and then use these values in the equation \eqref{eqn:Ne_VQ2},
-
-\begin{align} 
-    b'(\mu, s, L, r, d) = -\frac{V}{2} \left(\frac{1}{(1-(1-d)Z}\right)^2 
-\end{align}
-
-where $Z = 1 - Us / (U-\widetilde{R})$ and $V = Us - 2\widetilde{R}s$. TODO
-check d.
-
-\subsection{Discretization of Parameters and Annotation Feature Classes}
-
-In practice, to fit such a model, we must discretize both $\mu(s, k(g))$ and
-$s$ in Equation \eqref{supp-eq:b}. We define $\mu_i$ to be the rate of
-mutations entering the population per basepair, per generation with selection
-coefficient $s_i$ (i.e. the product of the mutation rate and the distribution
-of deleterious fitness effects). Then, the reduction at position $x$ is,
-
-\begin{align}
-  \label{supp-eq:b}
-  B(x, \Psi) = \exp \left( \sum_g^S \sum_{i} b\left( \mu_{i, \;k(g)}, s_i, L_g, r_g, d(x, g) \right)\right).
-\end{align}
-
-%TODO double check
-
-Because it is computationally-intensive to numerically solve the system of two
-non-linear equations to calculate $b$ at each segment, these are pre-computed
-for an $m_s$-element grid of selection coefficients and an $m_\mu$-element grid
-of mutation rates--DFE products. The number of annotation features $K$ is
-variable and set by the annotation data specified. There are two ways to
-parameterize the mutation rates and DFE for all annotation classes in
-background selection models. First, for the \emph{free-mutation}
-parameterization, each annotation feature class has a free mutation rate
-parameter for each of these selection coefficients (as in equation
-\eqref{supp-eq:b}).
-
-In the free-mutation parameterization, there are $1 + K \times m_s$ for $\pi_0$
-and the free mutation rate parameters, where $K$ is the number of annotation
-features. The MLE optimization for this approach is unconstrained, though
-$\pi_0$ is bounded to be within $10^{-5} \le \pi_0 \le 10^{-1}$ and mutation
-rates are bounded within $10^{-11} \le \mu \le 10^{-7}$. The BGS model requires
-a mutation rate for each selection coefficient in the grid, for each annotation
-class (here, as an example, CDS, UTRs, and phastcons), which are stored in the
-$m_s \times K$ matrix $\mathbf{M}_\text{F}$,
-
-\begin{align}
-  \mathbf{M}_\text{F} = \begin{bmatrix}
-    \mu_{10^{-6},\text{CDS}} & \mu_{10^{-6},\text{UTR}} & \mu_{10^{-6},\text{PC}} \\
-    \mu_{10^{-5},\text{CDS}} & \mu_{10^{-5},\text{UTR}} & \mu_{10^{-5},\text{PC}} \\
-    \vdots & \vdots & \vdots \\
-    \mu_{10^{-1},\text{CDS}} & \mu_{10^{-1},\text{UTR}} & \mu_{10^{-1},\text{PC}} \\
-  \end{bmatrix}
-\end{align}
-
-That is, the distribution of fitness effect (DFE) is implied by
-the total mutation rate across selection coefficients for an annotation class.
-We can normalize by the total mutation rate estimate to get the estimated DFE,
-giving us the DFE weight for annotation class $k$ and selection coefficient
-with index $i$,
-
-\begin{align}
-  \widehat{w}_{i,k} = \frac{\widehat{\mu}_{i,k}}{\sum_i \widehat{\mu}_{i,k}}.
-\end{align}
-
-The second parametrization is the \emph{simplex} parametrization, which has a
-single mutation rate across all features, so it is a DFE weight matrix times
-the mutation rate $\mu$,
-
-\begin{align}
-  \mathbf{M}_\text{S} = \mu \begin{bmatrix}
-    w_{10^{-6},\text{CDS}} & w_{10^{-6},\text{UTR}} & w_{10^{-6},\text{PC}} \\
-    w_{10^{-5},\text{CDS}} & w_{10^{-5},\text{UTR}} & w_{10^{-5},\text{PC}} \\
-    \vdots & \vdots & \vdots \\
-    w_{10^{-1},\text{CDS}} & w_{10^{-1},\text{UTR}} & w_{10^{-1},\text{PC}} \\
-  \end{bmatrix}
-\end{align}
-
-which has $2 + K \times (m_s-1)$ free parameters, since each column must sum to
-one. This imposes the constraint that $\sum_i w_{i,k} = 1$, requiring
-constrained MLE optimization.
+the solution to the system of two equations in Supplementary Equation
+\eqref{supp-eq:sc16-sys} for the mutation and selection coefficient grids,
+giving us $\widetilde{V}$ and $\widetilde{Q}_\infty^2$ for each set of
+parameters. Then,
+
+\begin{align} 
+    b'(\mu, s, L, r, d) = -\frac{\widetilde{V}}{2} \left(\frac{1}{(1-(1-d)Z}\right)^2 
+\end{align}
+%
+where $Z = 1 - Us / (U-\widetilde{R})$ and $\widetilde{V} = Us - 2\widetilde{R}s$. As
+before, $d$ is the recombination fraction between the segment and the focal
+neutral site.
 
@@
 \subsection{Numeric Optimization}
+\label{supp:optim}
 

### 2023-06-29 manuscript/manuscript.tex
ms updates
methods

@@
 \RequirePackage{bbm}
+
+\RequirePackage{cleveref}
+\RequirePackage{xr}
+\externaldocument{supplementary}
+
 %\RequirePackage[osf]{mathpazo}
@@
 
-\title{Towards a Complete Model of the Negative Selection Process in Humans}
+%\title{Towards a Complete Model of the Negative Selection Process in Humans}
+\title{An Evolutionary Quantitative Genetic Model of the Negative Selection Process in Humans}
 
@@
 population size $N_e$ (e.g. due to only non-selective process). The insight of
-nearly neutral theory (XXX) was that the qualitative behavior of selected sites
-is determined by the compound parameter $2N_e s$. When $2N_e s \gg 1$, the
-stochastic component of allele frequency change is minuscule relative to
-selection and can be ignored; this is why the factor that rescales the
-drift-effective population size under the BGS model does not depend on $N_e$.
-However, when $2N_e s \le 1$, the stochastic perturbations are on a scale equal
-to or greater than the changes due to selection. In this regime, weakly
-deleterious alleles can drift up to intermediate frequencies before their
-eventual loss or fixation. When this occurs, classic BGS theory breaks down for
-several reasons. First, the ``delay phase" is no longer negligible as lineages
-carrying weakly deleterious mutations can persist on coalescent timescales.
-This distorts genealogies away those expected under neutrality
+nearly neutral theory \parencite{Ohta1971-gq,Ohta1992-yi} is determined by the
+compound parameter $2N_e s$. When $2N_e s \gg 1$, the stochastic component of
+allele frequency change is minuscule relative to selection and can be ignored;
+this is why the factor that rescales the drift-effective population size under
+the BGS model does not depend on $N_e$. However, when $2N_e s \le 1$, the
+stochastic perturbations are on a scale equal to or greater than the changes
+due to selection. In this regime, weakly deleterious alleles can drift up to
+intermediate frequencies before their eventual loss or fixation. When this
+occurs, classic BGS theory breaks down for several reasons. First, the ``delay
+phase" is no longer negligible as lineages carrying weakly deleterious
+mutations can persist on coalescent timescales. This distorts genealogies away
+those expected under neutrality
 \parencite{Przeworski1999-mb,OFallon2010-my,Higgs1995-xc}. Second, the
@@
 \end{align}
+%
+where $V$ is the extra variance in offspring number.
 
@@
 same direction across generations as long as the linkage exists and the
-selective environment is constant. The additional heritable variation in
-offspring number is the additive fitness variation $V_A$, which is inflated by
-a factor $Q_t^2$ due to the build up of autocorrelation to generation $t$. We
-let $Q^2 := Q_t^2$ as $t \to \infty$. Intuitively, the product $V_A Q^2$
-represents the expected total variance in reproductive success a neutral
-mutation experiences over its lifetime in a system with weak selection draft.
-Incorporating the impact of time-invariant autocorrelation created by draft as
-a rescaled rate of drift is analogous to using the Green--Kubo relation to find
-the transport coefficient in molecular dynamics with velocity autocorrelation
-\parencite{Green1954-kl,Kubo1957-va}.
+selective environment is constant. Thus, the total additional heritable
+variation in offspring number is the product of the equilibrium additive
+fitness variation $V_A$ and an inflation factor $Q_t^2$ due to the build up of
+autocorrelation to generation $t$. We let $Q^2 := Q_t^2$ as $t \to \infty$.
+Intuitively, the product $V_A Q^2$ represents the expected total variance in
+reproductive success a neutral mutation experiences over its lifetime in a
+system with weak selection draft. Incorporating the impact of time-invariant
+autocorrelation created by draft as a rescaled rate of drift is analogous to
+using the Green--Kubo relation to find the transport coefficient in molecular
+dynamics with velocity autocorrelation \parencite{Green1954-kl,Kubo1957-va}.
 
@@
 
-(c.f. \cite{Robertson1961-ho,Santiago1995-hx}; see XXX for a proof). In
-general, however, statistics such as heterozygosity depend on the cumulative
-autocorrelation factor \emph{up to some time} $t$, $Q_t^2$ \parencite[p.
-2111]{Santiago1998-bs}. This reflects the fact that effective population size
-experienced by a mutation newly associated with a fitness background is larger
-than the asymptotic effective population size, since not much autocorrelation
-has accumulated in the stochastic perturbations caused by drift yet. From a
-backwards-time perspective, this reflects the fact that pairwise diversity
-(which is approximately heterozygosity when $2N\mu \ll 1$) depends on the
-pairwise coalescence rate per generation, which is not constant under draft.
-The benefit of using Robertson's forward-time model of draft is that the
-inflation factor is invariant with respect to the particular fitness background
-the neutral allele becomes stochastically associated with. By contrast, the
-difficulty with modeling draft backwards in time is that the coalescence rates
-experienced by a lineage are not invariant to which lineage was sampled due to
-its particular associated fitness background.
+(c.f. \cite{Robertson1961-ho,Santiago1995-hx}; see Supplementary Section
+\ref{suppsec:theory} for a proof). In general, however, statistics such as
+heterozygosity depend on the cumulative autocorrelation factor up to some time
+$t$, $Q_t^2$ \parencite[p. 2111]{Santiago1998-bs}. This reflects the fact that
+effective population size experienced by a mutation newly associated with a
+fitness background is larger than the asymptotic effective population size,
+since not much autocorrelation has accumulated in the stochastic perturbations
+caused by drift yet. From a backwards-time perspective, this reflects the fact
+that pairwise diversity (which is approximately heterozygosity when $2N\mu \ll
+1$) depends on the pairwise coalescence rate per generation, which is not
+constant under draft. The benefit of using Robertson's forward-time model of
+draft is that the inflation factor is invariant with respect to the particular
+fitness background the neutral allele becomes stochastically associated with.
+By contrast, the difficulty with modeling draft backwards in time is that the
+coalescence rates experienced by a lineage are not invariant to which lineage
+was sampled due to its particular associated fitness background.
 
@@
 $Q^2$ \parencite{Santiago1995-hx,Santiago1998-bs}. When fitness variation has a
-multiplicative polygenic basis, as it does for genome-wide negative selection
-processes, the draft asymptotic effective population size experienced by an
-arbitrary neutral site under the influence of all linked regions is,
+multiplicative polygenic basis, as is often assumed for genome-wide negative
+selection processes, the draft asymptotic effective population size experienced
+by an arbitrary neutral site under the influence of all linked regions is,
 
@@
 where the factor of one-half comes from ignoring weak associations from
-unlinked regions and chromosomes (see XXX). In our genome-wide model, we
-consider the summation in Equation \eqref{eq:polygenic_Ne} over non-overlapping
-segments $i \in \{1, 2, \ldots, n\}$ each undergoing selection such that
-segment $i$ contributes additive fitness variance $V_{A,i}$ to the total
-additive genetic fitness variance. Under equilibrium levels of fitness
-variation, the autocorrelation function for a neutral allele associated with
-segment $i$ is $C(t) = [(1-r_i)(1-\kappa_i)]^t$, where $r_i$ is the
-recombination fraction to the segment and $\kappa_i$ is the rate that the
-associated fitness variance decays due to selective dynamics. Then, the
-cumulative autocorrelation is,
+unlinked regions and chromosomes (see Supplementary Section
+\ref{suppsec:heritable-fitness}). In our genome-wide model, we consider the
+summation in Equation \eqref{eq:polygenic_Ne} over non-overlapping segments $i
+\in \{1, 2, \ldots, S\}$ each undergoing selection such that segment $i$
+contributes additive fitness variance $V_{A,i}$ to the total additive genetic
+fitness variance. Under equilibrium levels of fitness variation, the
+autocorrelation function for a neutral allele associated with segment $i$ is
+$C(t) = [(1-r_i)(1-\kappa_i)]^t$, where $r_i$ is the recombination fraction to
+the segment and $\kappa_i$ is the rate that the associated fitness variance
+decays due to selective dynamics. Then, the cumulative autocorrelation is,
 
 \begin{align}
-    Q_i &= 1 + \sum_{j=1}^\infty \left[(1-r_i)(1-\kappa_i)\right]^j \\
+    Q_i &= 1 + \sum_{j=1}^\infty \left[(1-r_i)(1-\kappa_i)\right]^j \nonumber \\
         &= \frac{1}{\kappa_i + r_i(1-\kappa_i)}.
@@
 general theory of Santiago and Caballero (\citeyear{Santiago2016-mu}), which
-they have shown previously (\citeyear{Santiago1998-bs}).
+they had shown previously (\citeyear{Santiago1998-bs}).
 
@@
 segregating site fixes and its heterozygosity becomes zero. If we let $R$
-represent the rate of the ratchet of deleterious mutations in fixations per
-generation, each fixation removes the equivalent amount of equilibrium
-variation put in by mutation. Thus, the steady-state genic variance under
-mutation and negative selection is (omitting the segment index $i$ for
-clarity),
+represent the fixation rate (i.e. ratchet rate) in the region per generation,
+each fixation removes the equivalent amount of equilibrium variation put in by
+mutation. Thus, the steady-state genic variance under mutation and negative
+selection is (omitting the segment index $i$ for clarity),
 
@@
 to new deleterious mutations and the flux out due to their substitution and the
-decline in mean population fitness. When $R=0$, the selection is so strong it
+decline in mean population fitness. When $R=0$, selection is so strong it
 cannot fix, and the equilibrium fitness variation is due entirely to young rare
-mutations before their extinction $V_a = V_a^{BGS} = Us$. Santiago and
-Caballero demonstrate Equation \eqref{eq:Va} through Fisher's Fundamental
-Theorem of Natural Selection, but we find an alternative proof
-(\cite{Higgs1995-xc}; see Supplementary Materials Section
-\ref{supp:weak-strong}). We find that the steady-state additive genic variance
-in Equation \eqref{eq:Va} also results from diffusion models with a flux of
-mutations into discrete sites (\cite{Kimura1969-jw}).
+mutations before their extinction $V_a = V_a^{BGS} \approx Us$. Santiago and
+Caballero derive Equation \eqref{eq:Va} through Fisher's Fundamental Theorem of
+Natural Selection, but we find an alternative proof (\cite{Higgs1995-xc}; see
+Supplementary Materials Section \ref{supp:weak-strong}). We also find that the
+steady-state additive genic variance in Equation \eqref{eq:Va} results from
+diffusion models with a flux of mutations into discrete sites
+(\cite{Kimura1969-jw}).
 
@@
 by the probability of fixation $p_F(N_d, s)$
-\parencite{Kimura1962-su,Malecot1948-zv} with the draft-rescaled effective
+\parencite{Kimura1962-su,Malecot1948-zv} using the draft-rescaled effective
 population size, i.e. $R = N U p_F(N_d, s)$. Given this equation for the
@@
 mutation-selection-drift-draft process for neutral to strongly deleterious
-additive mutations.
+additive mutations. The collection of these $V_a$ and $R$ equilibria can then
+be used to calculate the reduction at an arbitrary position in the genome (see
+Methods
 
@@
 and $s$, we can estimate the reduction factor $B(x) = \nicefrac{N_d}{N}$ at any
-genomic position $x$ (see Methods XXX). XXX
-
-Since it is too computationally-intensive to numerically solve this system of
-equations for all segments to estimate the reduction at every position
-
-for every focal position $x$
-
-% where,
-
-% \begin{align}
-%   V_a &= (U - R(x)) s & \text{\emph{fitness variance equation}} \\
-%   Q^2 &= \frac{2}{(1-Z)(2-(2-M)Z)} & \text{\emph{linkage inflation factor}} \\
-%   Z &= 1 - \frac{Us}{U-{R(x)}} & \text{\emph{}} \text{\emph{variance decay rate}}
-% \end{align}
+genomic position $x$ (see Methods XXX).
 
@@
 
+Measuring selection in the Genome Versus Finding Selection in the Genome
+
 \section*{Methods}
@@
 
-In our implementation of the theory of \textcite{Santiago2016-mu}, we calculate
-the draft-effective population size $N_d = B N$ within a segment by solving a
-system of two non-linear equations,
+Our software \texttt{bgspy} first calculates the equilibrium additive genic
+fitness variation $\widetilde{V}_a$ and ratchet rate $\widetilde{R}$ for each
+user-specified segments in the genome. These equilibria are calculated across
+grids of mutation rate weighted by the DFE $m_i$ and selection coefficient
+$s_j$, by numerically solving the following system of equations,
 
@@
   \label{eq:}
-  {N}_{d} &= N \exp \left( -V_a \frac{Q^2}{2} \right) & \text{\emph{effective population equation}} \\
-  R &= \frac{4N_eU s}{\exp(4 N_e s)-1}  & \text{\emph{substitution equation}} 
+  {N}_{d} &= N \exp \left( -V_a \frac{Q^2(m_i, s_j)}{2} \right) & \text{\emph{effective population equation}} \\
+  R &= \frac{4N_dU s_j}{\exp(4 N_d s_j)-1}  & \text{\emph{substitution equation}} 
 \end{align}
@@
   %V(x) &= U s - 2 {R}(x) s & \text{\emph{fitness variance equation}} \\
-  V_a &= (U  - 2 R) s & \text{\emph{fitness variance equation}} \\
-  Q^2 &= \frac{2}{(1-Z)(2-(2-M)Z)} & \text{\emph{linkage inflation factor}} \\
-  Z &= 1 - \frac{Us}{U-{R}} & \text{\emph{variance decay rate}}
+  V_a &= (U - 2 R) s_j & \text{\emph{fitness variance equation}} \\
+  Q^2(m_i, s_j) &= \frac{2}{(1-Z)(2-(2-M)Z)} & \text{\emph{linkage inflation factor}} \\
+  Z &= 1 - \frac{Us_j}{U-{M}} & \text{\emph{variance decay rate}}
 \end{align}
-
-and $U = \mu L$ and $M = r_\text{BP} L$ are the total mutation and
-recombination rates for a segment $L$ basepairs long.
-
-% Each mutation has a
-% selection coefficient of $s$, for which we assume additive effects. Then, the
-% local reduction factor is $b(x) = \nicefrac{N_e(x)}{N}$, and the per-basepair
-% per-generation substitution rate is $\lambda_D = \nicefrac{R(x)}{L}$. Note that
-% in the variance equation, the long-run variance is the balance between mutation
-% and twice the loss due to substitution, $(U-2R)s$, since for selective balance
-% $V(x) = 0$ for $s>0$, the per-basepair substitution rate must be half the mutation rate,
-% since each substitution removes twice the variance introduced by a mutation in
-% a diploid species.
-
-Our software \texttt{bgspy} solves these two equations numerically for each
-pre-specified segment across grids of mutation rates and selection
-coefficients. The recombination rate in a segment is determined by a
-user-supplied recombination map; segments that bridge a change in recombination
-rate are split, and solved independently with the appropriate rate. This
-pre-calculation step is done in parallel, and determines the equilibrium
-additive genic fitness variation for each segment under a
-mutation-selection-drift-draft process across the grids of parameters.
+%
+and $U = m_i L$ and $M = r_\text{BP} L$ are the total mutation and
+recombination rates in the segment. A detailed derivation of these equations
+can be found in Supplementary Section \ref{suppsec:theory}. The recombination
+rate in a segment is determined by a user-supplied recombination map.
 
@@
 
-The pre-computed equilibria are used to calculate the $B(x)$ maps across
-parameter grids, which use in our comparisons to chromosome-scale forward
-simulations and our composite-likelihood method. The reduction at a focal site
-$x$ can be calculated as the product of all segments' contributing reduction
-factors accounting for the recombination distances between the segment and
-focal site. Additionally, to accommodate the estimation of separate DFEs for
-different types of features, each $B(x)$ maps is
+We use these pre-computed equilibria to calculate the reduction map $B(x; m_i,
+s_j)$ at positions $x$ across the mutation grids described above. Since we
+assume multiplicative fitness, the reduction is the product of each segment's
+contribution accounting for the recombination is,
 
 \begin{align}
-    \log(B(x)) = - \sum_{i=1} \sum_{j=1} 
+    B(x; m_i, s_j) = \exp\left(- \frac{1}{2}\sum_{g \in G} \sum_{i=1}^{n_s} V_g(m_i, s_j) Q_g^2(m_i, s_j, r_{x, g})\right)
 \end{align}
 
-where $b(x, v)$ is a modified version of $b(x)$ that accounts for the
-recombination between the segment $x$ and the neutral site $v$ (see XXX).
-
-In practice, we calculate B' using the variables determined by solving the B'
-equations for each mutation and selection coefficient in the grids. Since only
-the analog of $Q^2(x)$, $Q^2(x, v)$ changes, the pre-computed substitution rate
-(and thus local variance) can be re-used for each $v$. We calculate $B(v)$
-across a grid of focal points along the chromosome, spaced 10kb apart.
+where $r_{x, g}$ is the recombination fraction between the focal site and
+segment $g$. A separate reduction map is calculated for all features $G$ within
+a specific feature type. Our B' maps are calculated for $\log_{10}$-spaced
+grids over $10^{-1} \le s \le 10^{-8}$ and $10^{-11} \le m \le 10^{-7}$, in
+10kb increments across the genome.
+
+\subsection*{Composite Likelihood}
+\label{sec:methods-likelihood}
+
+\begin{align}
+    \log\left(B(x; W, s_j) \right) = - \frac{1}{2} \sum_{g \in G} \sum_{i=1}^{n_s} V_g(\mu W_{j, k(g)}, s_j) Q_g^2(m_i, s_j, r_{x, g})
+\end{align}
 
@@
 \approx \nicefrac{2}{9n}$.
+
+\subsection*{Human Population Genomic Diversity Estimates}
+
+We estimate pairwise diversity for Yoruba (YRI), European (CEU), and Han
+Chinese (CHB) ancestry individuals in the high-coverage 1000 Human Genomes
+dataset (XXX). We
 

### 2023-06-29 manuscript/supplementary.tex
methods

@@
 \section{Quantitative Genetic Linked Selection Model Theory}
+\label{suppsec:theory}
 
@@
 \subsection{Heritable Fitness and the Accumulation of Autocovariance}
+\label{suppsec:heritable-fitness}
 
@@
   \label{supp-eq:b}
-  B(x, \Psi) &= \prod_g^S \exp \left( \int_0^1 b(\mu(s, k(g)), s, L_g, r_g, d(x, g)) ds \right) \nonumber  \\
-             &= \exp \left( \sum_g^S \int_0^1 b(\mu(s, k(g)), s, L_g, r_g, d(x, g)) ds \right)
+  B(x, \Psi) &= \prod_{g \in G} \exp \left( \int_0^1 b(\mu(s, k(g)), s, L_g, r_g, d(x, g)) ds \right) \nonumber  \\
+             &= \exp \left( \sum_{g \in G} \int_0^1 b(\mu(s, k(g)), s, L_g, r_g, d(x, g)) ds \right)
 \end{align}
@@
 non-linear equations to calculate $b$ at each segment, these are pre-computed
-for an $m_s$-element grid of selection coefficients and an $m_\mu$-element grid
+for an $n_s$-element grid of selection coefficients and an $n_\mu$-element grid
 of mutation rates--DFE products. Likewise, when we calculate $b$ under classic
@@
 %
-which has $2 + K \times (m_s-1)$ free parameters, since each column must sum to
+which has $2 + K \times (n_s-1)$ free parameters, since each column must sum to
 one. This imposes the constraint that $\sum_i w_{i,k} = 1$, requiring
@@
 \eqref{supp-eq:sc16-sys} for the mutation and selection coefficient grids,
-giving us $\widetilde{V}$ and $\widetilde{Q}_\infty^2$ for each set of
+giving us $\widetilde{V}_a$ and $\widetilde{Q}_\infty^2$ for each set of
 parameters. Then,
@@
 \begin{align} 
-    b'(\mu, s, L, r, d) = -\frac{\widetilde{V}}{2} \left(\frac{1}{(1-(1-d)Z}\right)^2 
-\end{align}
-%
-where $Z = 1 - Us / (U-\widetilde{R})$ and $\widetilde{V} = Us - 2\widetilde{R}s$. As
+    b'(\mu, s, L, r, d) &= -\frac{\widetilde{V}_a(\mu, s, r)}{2} Q^2(\mu, s, r) \\
+                        Q^2(\mu, s, r)&= \left(\frac{1}{(1-(1-d)\widetilde{Z}}\right)^2 
+\end{align}
+%
+where $\widetilde{Z} = 1 - Us / (U-\widetilde{R})$ and $\widetilde{V}_a = Us - 2\widetilde{R}s$. As
 before, $d$ is the recombination fraction between the segment and the focal
@@
 \end{align}
-
+%
 which is identical to the more common expression of this estimator,
@@
 \end{align}
-
+%
 where $k_{i,j}$ is 1 if the alleles at this site differ at site $l$, and 0
@@
 \end{align}
-
+%
 which if the number of samples across loci is constant, simplifies to an
@@
 \end{align}
-
+%
 Third, one can weight by the number of pairwise comparisons at a site,
@@
 summarize the matrix $\mathbf{Y}$ across windows for our likelihood. All
-methods have mean squared errors and biases very close to one another (TODO).
+methods have mean squared errors and biases very close to one another.
 
@@
 
-\subsection{Window-based Summaries and filtering}
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics{figures/supplementary/diversity_trimming_dist.pdf}
-
-  \caption{The distribution of diversity across the genome at the megabase
-  scale, with outliers trimmed. The blue line is the normal CDF, fit with MLE
-parameters XXX. The inset figure is the untrimmed data, with trimmed points
-shown in red.}
-
-  \label{suppfig:trimming}
-\end{figure}
-
-The likelihood method is fit to non-overlapping binned summaries of the allele
-counts matrix.
-
-Based on exploratory data analyses, some bins were outliers and
-excluded (XXX).
-
-\begin{enumerate}
-  \item \textbf{Fraction of inaccessible sites per window}.  \path{mask_inaccessible_bins_frac}
-  % \item \textbf{Edge truncation}
-
-  \item \textbf{Outliers}: Based on exploratory analysis, there were some
-    regions with very high diversity. The 0.05\% right tail was excluded.
-
-\end{enumerate}
-
-% supp data
-% - npy files
-% - edge trunation
-
-\subsection{Annotation Models}
-
-CADD 6\%: 6.49\% of autosomal sequence.
-CADD 8\%: 8.56\% of autosomal sequence.
-
 \subsection{Ratchet Rate Prediction}
@@
 
-The rescaled $B'$ calculation outputs a $m_\mu \times m_s \times S$
+The rescaled $B'$ calculation outputs a $n_\mu \times n_s \times S$
 multidimensional array $\mathbf{R}$ of ratchet rates per segment. These are
@@
 \end{table}
+
+\subsection{Window-based Summaries and filtering}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics{figures/supplementary/diversity_trimming_dist.pdf}
+
+  \caption{The distribution of diversity across the genome at the megabase
+  scale, with outliers trimmed. The blue line is the normal CDF, fit with MLE
+parameters XXX. The inset figure is the untrimmed data, with trimmed points
+shown in red.}
+  \label{suppfig:trimming}
+\end{figure}
+
+The likelihood method is fit to non-overlapping binned summaries of the allele
+counts matrix.
+
+Based on exploratory data analyses, some bins were outliers and
+excluded (XXX).
+
+\begin{enumerate}
+  \item \textbf{Fraction of inaccessible sites per window}.  \path{mask_inaccessible_bins_frac}
+  % \item \textbf{Edge truncation}
+
+  \item \textbf{Outliers}: Based on exploratory analysis, there were some
+    regions with very high diversity. The 0.05\% right tail was excluded.
+
+\end{enumerate}
+
+% supp data
+% - npy files
+% - edge trunation
+
+\subsection{Annotation Models}
+
+CADD 6\%: 6.49\% of autosomal sequence.
+CADD 8\%: 8.56\% of autosomal sequence.
 

### 2023-06-30 manuscript/manuscript.tex
methods

@@
 \begin{align}
+    \label{eq:Q}
     Q_i &= 1 + \sum_{j=1}^\infty \left[(1-r_i)(1-\kappa_i)\right]^j \nonumber \\
@@
 
-where $V_a = (U-R)s$ as in Equation \eqref{eq:Va}. The predicted $V_a$, $R$,
-and $N_d$ values represent the equilibria expected under a
-mutation-selection-drift-draft process for neutral to strongly deleterious
-additive mutations. The collection of these $V_a$ and $R$ equilibria can then
-be used to calculate the reduction at an arbitrary position in the genome (see
-Methods
+where $V_a = (U-R)s$ as in Equation \eqref{eq:Va}. We denote the solutions to
+these equation, which represent equilibria under mutation-selection-drift-draft
+process, as $\widetilde{N}_d$ and $\widetilde{R}$. These equilibria also imply
+an equilibrium level of additive fitness variation $\widetilde{V}_a$ in the
+segment, are used to calculate the reduction at an arbitrary position in the
+genome (see Methods
 
@@
 
-We use these pre-computed equilibria to calculate the reduction map $B(x; m_i,
-s_j)$ at positions $x$ across the mutation grids described above. Since we
-assume multiplicative fitness, the reduction is the product of each segment's
-contribution accounting for the recombination is,
+Our method uses the pre-computed equilibria $\widetilde{V}_{a,g}$ for each
+segment $g$ to calculate the reduction map $B(x; m_i, s_j)$ at positions $x$
+across the parameter grids described above. Since we assume multiplicative
+fitness, the reduction is the product of each segment's contribution accounting
+for the recombination is,
 
 \begin{align}
-    B(x; m_i, s_j) = \exp\left(- \frac{1}{2}\sum_{g \in G} \sum_{i=1}^{n_s} V_g(m_i, s_j) Q_g^2(m_i, s_j, r_{x, g})\right)
+    B(x; m_i, s_j) = \exp\left(- \frac{1}{2}\sum_{g \in G} \sum_{i=1}^{n_s} \widetilde{V}_{a,g}(m_i, s_j) Q_g^2(m_i, s_j, r_{x, g})\right)
 \end{align}
-
+%
 where $r_{x, g}$ is the recombination fraction between the focal site and
-segment $g$. A separate reduction map is calculated for all features $G$ within
-a specific feature type. Our B' maps are calculated for $\log_{10}$-spaced
-grids over $10^{-1} \le s \le 10^{-8}$ and $10^{-11} \le m \le 10^{-7}$, in
-10kb increments across the genome.
-
-\subsection*{Composite Likelihood}
+segment $g$. Here, $Q^2(m_i, s_j, r_{x,g})$ is given by Equation \eqref{eq:Q}
+squared. A separate reduction map is calculated for all features $G$ within a
+specific feature type. We calculate B' calculate for $\log_{10}$-spaced grids
+over $10^{-1} \le s \le 10^{-8}$ and $10^{-11} \le m \le 10^{-7}$, in 10kb
+increments across the genome.
+
+\subsection*{Composite Likelihood and Optimization}
 \label{sec:methods-likelihood}
 
+Following previous approaches
+\parencite{McVicker2009-ax,Elyashiv2016-vt,Murphy2022-sj}, we use a composite
+likelihood approach to fit our negative selection model. Per-basepair allele
+count data (described below) is summarized into the number of same and
+different pairwise differences per window. All of our primary models were fit
+with megabase windows, since previous work has found the strongest selection
+signal at this scale (we confirm this with one XXX fit at the $10^5$bp scale).
+
+We use a binomial likelihood with probability $\bar{\pi}(b; \Psi) = \pi_0
+\bar{B}(b; \mu, \mathbf{W})$, where bars indicate averages over some bin width.
+With multiple features classes and a map $k(g)$ between a segment $g$ to its
+class, the reduction map is,
+
 \begin{align}
-    \log\left(B(x; W, s_j) \right) = - \frac{1}{2} \sum_{g \in G} \sum_{i=1}^{n_s} V_g(\mu W_{j, k(g)}, s_j) Q_g^2(m_i, s_j, r_{x, g})
+    \log\left(B(x; W, s_j) \right) = - \frac{1}{2} \sum_{g \in G} \sum_{i=1}^{n_s} V_g(\mu W_{j, k(g)}, s_j) Q_g^2(m_i, s_j, r_{x, g}).
 \end{align}
+
+Here, $\mu$ is the mutation rate and $\mathbf{W}$ is the $n_s \times K$ DFE
+matrix over the selection grid and features. Since each column of the DFE
+$\mathbf{W}$ must sum to one, numeric optimization was done in unconstrained
+$n_s-1$ dimensional softmax space. Our binomial likelihood modeled the number
+of different pairwise comparisons observed per window given the predicted
+pairwise diversity and total number of comparisons implied per window (see
+Supplementary Materials Section \ref{supp:data-summary}).
+
+We tested multiple different optimization routines, finding that BOBYQA
+outperformed alternatives \parencite{Powell_undated-ec}. We inspected and
+confirmed convergence with diagnostic plots finding stable optima across 10,000
+random starts (see Supplementary Materials Section XXX).
 

### 2023-06-30 manuscript/supplementary.tex
methods

@@
 \begin{align}
-  \mathbf{M}_\text{S} = \mu \begin{bmatrix}
+    \mathbf{M} = \mu \mathbf{W} &= \begin{bmatrix} 
     w_{10^{-6},\text{CDS}} & w_{10^{-6},\text{UTR}} & w_{10^{-6},\text{PC}} \\
@@
 \subsection{Data Summary Matrices}
+\label{supp:data-summary}
 
@@
 \path{data/fit_annotation/Snakefile} in this study's GitHub repository
-(\url{http://github.com/vsbuffalo/bprime}).
+(\url{http://github.com/vsbuffalo/bprime}). We summarize the percentage of
+basepairs after masking per chromosome in Supplementary Table
+\ref{supp-table:masks}.
 
 \begin{table}
-\centering
-\caption{Percent of basepairs remaining that are accessible and 
-         putatively neutral.}
-\begin{tabular}{|l|rrr|}
+    \label{supp-table:masks}
+\centering
+\caption{Accessible and putatively neutral sequence mask statistics, for the criteria used in 
+         the main model fits.}
+\begin{tabular}{|c|cccc|}
 \hline
-Chromosome & Accessible & Neutral & Intersection \\
+Chromosome & Accessible (\%) & Neutral (\%) & Intersection (\%) & Total Basepairs ($\times 10^6$) \\
 \hline 
- chr1 &        39.7 &     61.2 &  23.4 \\
- chr2 &        45.3 &     67.3 &  27.5 \\
- chr3 &        43.2 &     67.3 &  25.7 \\
- chr4 &        42.7 &     70.5 &  27.4 \\
- chr5 &        42.4 &     67.5 &  26.0 \\
- chr6 &        43.8 &     67.9 &  26.7 \\
- chr7 &        41.9 &     67.8 &  26.2 \\
- chr8 &        42.3 &     70.9 &  27.5 \\
- chr9 &        36.2 &     58.1 &  22.0 \\
-chr10 &        42.9 &     68.2 &  26.8 \\
-chr11 &        41.3 &     67.0 &  25.0 \\
-chr12 &        40.6 &     67.2 &  24.1 \\
-chr13 &        38.8 &     59.8 &  25.2 \\
-chr14 &        36.5 &     54.9 &  21.5 \\
-chr15 &        34.4 &     53.7 &  20.6 \\
-chr16 &        36.0 &     62.5 &  23.1 \\
-chr17 &        37.2 &     63.3 &  21.6 \\
-chr18 &        40.3 &     65.7 &  26.3 \\
-chr19 &        28.1 &     66.1 &  17.0 \\
-chr20 &        36.1 &     65.1 &  22.4 \\
-chr21 &        29.8 &     55.9 &  20.9 \\
-chr22 &        25.5 &     50.9 &  17.1 \\
+1 &        39.7 &     62.0 &  15.7 &     39.09 \\
+2 &        45.3 &     60.1 &  18.9 &     45.77 \\
+3 &        43.2 &     59.6 &  16.9 &     33.51 \\
+4 &        42.7 &     65.9 &  20.2 &     38.42 \\
+5 &        42.4 &     62.2 &  18.1 &     32.86 \\
+6 &        43.8 &     61.4 &  18.4 &     31.43 \\
+7 &        41.9 &     64.0 &  18.6 &     29.64 \\
+8 &        42.3 &     66.1 &  20.2 &     29.32 \\
+9 &        36.2 &     65.4 &  14.6 &     20.21 \\
+10 &        42.9 &     62.2 &  18.8 &     25.15 \\
+11 &        41.3 &     61.5 &  17.0 &     22.96 \\
+12 &        40.6 &     60.4 &  15.9 &     21.19 \\
+13 &        38.8 &     70.7 &  18.8 &     21.50 \\
+14 &        36.5 &     64.6 &  13.8 &     14.77 \\
+15 &        34.4 &     66.3 &  13.5 &     13.77 \\
+16 &        36.0 &     66.0 &  15.5 &     14.00 \\
+17 &        37.2 &     58.8 &  14.2 &     11.82 \\
+18 &        40.3 &     66.1 &  19.6 &     15.75 \\
+19 &        28.1 &     67.5 &  11.5 &      6.74 \\
+20 &        36.1 &     62.5 &  15.4 &      9.92 \\
+21 &        29.8 &     76.9 &  17.3 &      8.08 \\
+22 &        25.5 &     75.5 &  12.6 &      6.40 \\
+\hline
+weighted average & 40.1 & 63.6 & 17.1 & - \\
 \hline
@@
 
-$\Phi$ is the set of BGS parameters (i.e. the DFE for each feature type and
-mutation rate), and $\pi_0 = 4 N_e \mu$ is determined by the genome-wide
+$\Phi$ is the set of BGS parameters (i.e. the DFE matrix $\mathbf{W}$ and
+mutation rate $\mu$), and $\pi_0 = 4 N_e \mu$ is determined by the genome-wide
 drift-effective population size $N_e$, set by only reproductive and demographic
@@
 
+\subsection{Optimization Diagnostics}
+
+\begin{figure}[!htb]
+  \label{suppfig:diag-plot}
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/figure_feature_priority_yri_full_diag.pdf}
+  \caption{
+      An optimization diagnostic plot showing the }
+\end{figure}
+
 \subsection{Chromosome Fit Figures}

### 2023-06-30 manuscript/manuscript.tex
supp, methods

@@
 confirmed convergence with diagnostic plots finding stable optima across 10,000
-random starts (see Supplementary Materials Section XXX).
+random starts (see Supplementary Materials Section \ref{suppsec:optima}).
 

### 2023-06-30 manuscript/supplementary.tex
supp, methods

@@
 \subsection{Optimization Diagnostics}
-
-\begin{figure}[!htb]
+\label{suppsec:optim}
+
+We only show one optimization diagnostic plot here (Supplementary Figure
+\ref{suppfig:diag-plot}), but the can be easily calculated from the model
+pickle files in the XXX repository (XXX). In one case, our optimization
+diagnostics indicated that the best optima across 10,000 random starts was one
+of few of the top 100 hits that were hitting either an alternate likelihood
+mode, or the optimization routine had numeric issues. In this one case, we
+removed the top hit and used the second, which was the consensus among the
+other top 100 optima.
+
+\begin{figure}[htbp]
   \label{suppfig:diag-plot}
@@
   \includegraphics[width=\textwidth]{figures/supplementary/figure_feature_priority_yri_full_diag.pdf}
-  \caption{
-      An optimization diagnostic plot showing the }
-\end{figure}
+
+  \caption{ An optimization diagnostic plot showing the 100 top optima. Heatmap
+  colors indicate probability mass for the DFE parameters (top four rows).
+Estimates of $\pi_0$ and $\mu$ are shown in the bottom two rows. This is the
+optimization plot for YRI Feature Priority model.}
+
+\end{figure}
+\clearpage
 
 \subsection{Chromosome Fit Figures}
-\clearpage
 
@@
   \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr6.pdf}
+  \caption{Note: we mask the MHC region during model fit.}
   \label{suppfig:fit-chr6}

### 2023-06-30 manuscript/supplementary.tex
supp, methods

@@
   \item \textbf{Non-accessible regions}: masks out centromeres (\texttt{acen} entries in
-    \texttt{cytoBand.txt}), with 5Mbp padding on either side. The file of
+    \texttt{cytoBand.txt}), with 5 Mbp padding on either side. The file of
     passing masks is \path{data/annotation/no_centro.bed}.
@@
 
+\subsection{Out-Sample Error Estimation and $R_\text{LOO}^2$}
+
+We fit each of our B' model fits 22 times, excluding a whole chromosome each
+time to calculate out-sample prediction error. Each fit used 10,000 random
+starting positions as the main fits. The out-sample prediction error for all
+bins in chromosome $C$ left out is,
+
+\begin{align}
+    SSE(C) = \sum_{b \in C} (\widehat{\pi}_{b,Y\not \in C}  - \pi_b)^2
+\end{align}
+%
+
+where $\widehat{\pi}_{b,Y\not \in C}$ denotes the predictions for window $b$
+fit on data excluding chromosome $C$. The average out-sample error $SSE$ across
+chromosomes is then the average of $SSE(C)$ across chromosomes, weighted by the
+number of bins per chromosome. We then calculate the out-sample coefficient of
+determination $R_\text{LOO}^2$
+
+\begin{align}
+    R_\text{LOO}^2 = \frac{SSE}{}
+\end{align}
+
+XXX
+
+\subsection{Jackknife Procedures}
+
+\begin{figure}[htbp]
+  \label{suppfig:jackknife-iqr}
+  \centering
+  \includegraphics[width=0.7\textwidth]{figures/supplementary/iqr_jackknife_trim.pdf}
+
+  \caption{The distribution of block jackknife samples for the Feature Priority
+  model fit on YRI data. The ranges at the top of the figure indicate the
+untrimmed and the 2 IQR-trimmed jackknife intervals. for XXX fix fig }
+
+\end{figure}
+
+We use a block jackknife procedure to calculate the uncertainty of our
+estimates. This approach is unfortunately computationally intensive, since a
+10Mbp block width requires $\approx 280$ blocks to cover 2.8Gbp of autosomal
+sequence and each fit requires thousands of random starts to reliably find
+consistent optima. We use a parallelization strategy where a set number of
+evenly-spaced positions along the genome that are chosen to approximately cover
+all 2.8Gbps of autosomal sequence. Each of these positions is sent out across a
+computing cluster, which excludes the 10Mbp block there and fits on the
+remaining data. Then, using these jackknife samples we calculated the jackknife
+standard error for each of parameters \parencite{Wasserman2004-ip}. In some
+cases, our jackknife standard error estimates were very large. Upon inspection
+of optimization diagnostic plots (see Supplementary Materials Section
+\ref{suppfig:diag-plot}) and histograms of jackknife fits (Supplementary Figure
+\ref{suppfig:jackknife-iqr}), these were due to rare aberrant optima. We
+removed these using a trimmed-interquartile range, which removes points that
+are 2 interquartile ranges below the first quartile or above the third quartile
+(1.5 is standard, but we chose to be more conservative to not underestimate
+uncertainty).
+
 \section{Simulations}
@@
 
-\subsection{Optimization Diagnostics}
+\subsection{Optimization and Diagnostics}
 \label{suppsec:optim}
@@
 optimization plot for YRI Feature Priority model.}
-
-\end{figure}
+\end{figure}
+
 \clearpage

### 2023-07-01 manuscript/manuscript.tex
supp, methods
more methods

@@
 (c.f. \cite{Robertson1961-ho,Santiago1995-hx}; see Supplementary Section
-\ref{suppsec:theory} for a proof). In general, however, statistics such as
+\ref{supp:theory} for a proof). In general, however, statistics such as
 heterozygosity depend on the cumulative autocorrelation factor up to some time
@@
 unlinked regions and chromosomes (see Supplementary Section
-\ref{suppsec:heritable-fitness}). In our genome-wide model, we consider the
+\ref{supp:heritable-fitness}). In our genome-wide model, we consider the
 summation in Equation \eqref{eq:polygenic_Ne} over non-overlapping segments $i
@@
 Similar to previous work, our method characterizes the negative selection
-process through the signal left on pairwise diversity at linked sites. Although
-pairwise diversity is a less-informative statistic than the full site frequency
-spectrum, it is much easier to derive models of how selection alters expected
-pairwise coalescence times at linked sites than to describe the entire
-genealogy. Spatial patterns of pairwise diversity are highly variable across
-eukaryotic genomes, as linked selection leads diversity to become correlated
-with spatial heterogeneity in recombination rates and the density of conserved
-sites. Statistical methods then fit observed diversity to the expected level
-under a particular theoretic model of linked selection processes. Consequently,
-statistical inferences about selection may be incorrect when the underlying
-theory inaccurately predicts average pairwise coalescence times.
+process through the signal left on nucleotide diversity at linked sites.
+Although nucleotide diversity is a less-informative statistic than the full
+site frequency spectrum, it is much easier to derive models of how selection
+alters expected pairwise coalescence times at linked sites than to describe the
+entire genealogy. Spatial patterns of pairwise diversity are highly variable
+across eukaryotic genomes, as linked selection leads diversity to become
+correlated with spatial heterogeneity in recombination rates and the density of
+conserved sites. Statistical methods then fit observed diversity to the
+expected level under a particular theoretic model of linked selection
+processes. Consequently, statistical inferences about selection may be
+incorrect when the underlying theory inaccurately predicts average pairwise
+coalescence times.
 
@@
 basepairs under varying levels of mutation and selection (see Methods
-\ref{sec:methods-segsim}). We find a close correspondence between the observed
+\ref{sec:methods-sim}). We find a close correspondence between the observed
 and predicted reductions in effective population size $B=\nicefrac{N_d}{N}$
@@
 Priority, and Feature Priority) to high-coverage 1000 Genome data for three
-populations: Yoruba (YRI), Han Chinese (CHB), and European (CEU) ancestries. We
-assess and compare our models according to how well they predict patterns of
-diversity on whole chromosomes left-out during the model fitting process. We
-use the metric $R_\text{LOO}^2$, which is the proportion of the observed
-variance in genomic diversity at the megabase scale predicted by our model. We
-experimented with a few smaller spatial scales, but our results were consistent
-with previous results suggesting the human linked selection signal fits best at
-the megabase scale \parencite{Murphy2022-sj}.
+populations: Yoruba (YRI), Han Chinese (CHB), and European (CEU) reference
+samples. We assess and compare our models according to how well they predict
+patterns of diversity on whole chromosomes left-out during the model fitting
+process. We use the metric $R_\text{LOO}^2$, which is the proportion of the
+observed variance in genomic diversity at the megabase scale predicted by our
+model. We experimented with a few smaller spatial scales, but our results were
+consistent with previous results suggesting the human linked selection signal
+fits best at the megabase scale \parencite{Murphy2022-sj}.
 
@@
 
-Next, we developed a composite-likelihood method to estimate pairwise diversity
-in the absence of linked selection $\pi_0$, the deleterious mutation rate
-($\mu_D$), and distribution of fitness effects ($\mathbf{W}$) under negative
-selection. Because the focus of our work is to fix deficiencies in the classic
-BGS model, our current implementation does not consider reductions in diversity
-due to recurrent hitchhiking or other positive selection processes. Since
-previous work has established that purifying selection is the dominant mode of
-linked selection in humans and hitchhiking plays a limited role
+Next, we developed a composite-likelihood method to estimate expected diversity
+without linked selection $\pi_0$, the deleterious mutation rate ($\mu_D$), and
+distribution of fitness effects ($\mathbf{W}$) under negative selection.
+Because the focus of our work is to fix deficiencies in the classic BGS model,
+our current implementation does not consider reductions in diversity due to
+recurrent hitchhiking or other positive selection processes. Since previous
+work has established that purifying selection is the dominant mode of linked
+selection in humans and hitchhiking plays a limited role
 \parencite{McVicker2009-ax,Murphy2022-sj}, humans are a suitable organism to
@@
 \section*{Discussion}
+
+Regularization / overfitting
 
@@
 \bar{B}(b; \mu, \mathbf{W})$, where bars indicate averages over some bin width.
-With multiple features classes and a map $k(g)$ between a segment $g$ to its
+With multiple features classes and a map $k(g)$ between segments $g$ to their
 class, the reduction map is,
@@
 
-Here, $\mu$ is the mutation rate and $\mathbf{W}$ is the $n_s \times K$ DFE
-matrix over the selection grid and features. Since each column of the DFE
-$\mathbf{W}$ must sum to one, numeric optimization was done in unconstrained
-$n_s-1$ dimensional softmax space. Our binomial likelihood modeled the number
-of different pairwise comparisons observed per window given the predicted
-pairwise diversity and total number of comparisons implied per window (see
-Supplementary Materials Section \ref{supp:data-summary}).
+Here, $\mu$ is the mutation rate and $W_{j,k(g)}$ is a DFE entry in the
+$\mathbf{W}$ is the $n_s \times K$ DFE matrix over the selection grid and
+features. Since each column of the DFE $\mathbf{W}$ must sum to one, numeric
+optimization was done in unconstrained $n_s-1$ dimensional softmax space. Our
+binomial likelihood modeled the number of different pairwise comparisons
+observed per window given the predicted pairwise diversity and total number of
+comparisons implied per window (see Supplementary Materials Sections
+\ref{supp:data-summary} and \ref{supp:likelihood}).
 
@@
 confirmed convergence with diagnostic plots finding stable optima across 10,000
-random starts (see Supplementary Materials Section \ref{suppsec:optima}).
-
-\subsection*{Segment Simulations}
-\label{sec:methods-segsim}
-
-To assess the accuracy of the SC16 theory, we conducted forward simulations
-across grids of mutation rates ($\log_{10}(\mu) = \{10^{-9.5}, 10^{-9}, \ldots,
-10^{-8}\}$) and selection coefficients ($\log_{10}(\mu) = \{-1, 1.5, 2, \ldots,
-4.5\}$) using SLiM \parencite{Haller2023-uk,Haller2019-vu}. Each simulation
-evolved a 100kb region with deleterious mutations entering at rate $\mu$
-per-basepair, per generation, each with an additive selection coefficient of
-$-s$. Total fitness was multiplicative across sites. Each simulation recorded
-the number of fixations and the fitness variance per generation. After $10N$
-generations, the ARG of the tree was recorded and downstream Python scripts
-calculated diversity in the center of the region. We averaged the results of
-$10,000$ replicate simulations to compare against theory.
-
-\subsection*{Chromosome-wide Simulations}
-
-, using the HapMap recombination map (XXX) and in coding regions, PhastCons
-regions, and UTRs as an approximate set of putatively conserved segments
-
-We calculate the mean squared error (MSE) and relative error between
-whole-chromosome simulations and the predicted B and B' maps by averaging
-branch diversity in 10 kilobase windows (tskit, XXX) and dividing by the
-neutral expectation of $4N$. The theoretic lower bound of the MSE is due the
-noise in estimating the expected reduction map from simulations. For $n=100$
-replicates, the sampling noise is minimal relative to the evolutionary noise in
-the coalescence process \parencite{Tajima1983-gu}. We use this as the theoretic
-lower bound assuming the maximum coalescence noise at $B=1$, $\var(\bar{B})
-\approx \nicefrac{2}{9n}$.
-
-\subsection*{Human Population Genomic Diversity Estimates}
-
-We estimate pairwise diversity for Yoruba (YRI), European (CEU), and Han
-Chinese (CHB) ancestry individuals in the high-coverage 1000 Human Genomes
-dataset (XXX). We
-
-\subsection*{Local-rescaled }
-We note that it is computationally intractable to solve this equation excluding
-the contribution of the focal segment
-
-\subsection*{Assessing the B' Maps}
-
-We first assessed the B' maps against the observed average reduction across
-$XXX$ forward-simulations, $\bar{B}_\text{sim}(x)$. The mean-squared error has
-two sources, (1) bias from a breakdown in our theory $B(x)$ for the average
-reduction, and (2) variance from the simulations. We show that for most of the
-parameter space, the measured mean-squared error between simulation averages
-and our B' maps is at the theoretic minimum. This minimum variance is entirely
-from the coalescent noise when estimating $\bar{B}\text{sim}(x)$, which
-
-\subsection*{Maximum Likelihood Estimation}
-
-\subsection*{Local Rescaling and Substitution Rate Prediction}
+random starts (see Supplementary Materials Section \ref{suppsec:optima}). We
+assessed model fit using out-sample predictive error, calculated by leaving out
+a whole chromosome during fitting and predicting its diversity. To calculate
+uncertainty, we used a block jackknife approach in 10 Mbp windows
+(Supplementary Section \ref{supp:jackknife}).
+
+\subsection*{Human Population Genomic Data}
+\label{sec:methods-data}
+
+Our analyses was conducted on the Yoruba (YRI), European (CEU), and Han Chinese
+(CHB) reference samples individuals from the high-coverage 1000 Human Genomes
+data aligned to GRCh38/hg38 \parencite{Byrska-Bishop2022-tn}. Since nucleotide
+diversity is a ratio estimator, it can be biased when subtly different
+filtering criteria are applied to variant and invariant sites. To prevent this,
+we conducted our analyses on Genomic VCF (gVCF) files that contain genotype
+calls for both variant and invariant sites \parencite{Illumina_Inc2020-dk}.
+Then, we apply the same genotype filtering criteria to all called sites
+(Supplementary Materials Section \ref{supp:filter}). We also applied sequence
+accessibility masks that containing only non-repeat, non-centromeric sequence
+that passed the 1000 Genomes strict filter
+(\cite{1000_Genomes_Project_Consortium2015-mi}; Supplementary Material Section
+\ref{supp:accessible}). Since our theory only considers the indirect effects of
+linked selection on a site, we additionally masked sites that are likely under
+direct selective constraint (see Supplementary Materials Section
+\ref{supp:neutral}). Finally, for every basepair passing these filtering and
+masking criteria, we counted the number of reference and alternative allele
+counts (excluding all multiallelic, indel, and CNV variants).
+
+For all of our main analyses, we used the recombination map from
+\textcite{Halldorsson2019-ey} estimated from a trio-based design to avoid
+circularity that could occur by using LD-based maps. We use Ensembl gene
+annotation \parencite{Cunningham2022-vk}, a special CADD Score dataset with
+McVicker B scores removed (to avoid circularity;
+\cite{Kircher2014-bv,Rentzsch2019-lr}), and PhastCons regions
+\parencite{Siepel2005-wh}.
+
+\subsection*{Forward Simulations}
+\label{sec:methods-sim}
+
+We use two types of forward simulations in this paper; both use SLiM version
+4.0.1 \parencite{Haller2023-uk,Haller2019-vu}. First, we simulated a 100 kbp
+segment in a population of $N=1000$ diploids across grids of mutation and
+selection parameters to assess the accuracy of the SC16 theory. Each simulation
+recorded the number of fixations and the fitness variance per generation, and
+at the end $10N$ generations, the ancestral recombination graph was recorded so
+downstream Python scripts could calculate diversity. Each of these segment
+simulations was replicated $10,000$ times and averaged to compare against
+theoretic expectations.
+
+Second, we conducted forward simulations of negative selection on whole human
+chromosomes. These simulations used the HapMap recombination map
+\parencite{International_HapMap_Consortium2007-nu} and treated coding,
+PhastCons, and UTR regions as an approximate set of conserved regions. We used
+these whole-chromosome simulations for three purposes. First, we simulated
+negative selection on chromosome 10 to confirm our method to estimate classic B
+and new B' maps matched the average empirical reduction map across a grid of
+mutation and selection parameters. Second, we evaluated our composite
+likelihood method by simulating negative selection on the first five human
+chromosomes, across grids of mutation and selection parameters. Multiple
+independently-evolved chromosomes were combined into a synthetic genome in
+order to have enough data for our procedure. Third, we conducted additional
+simulations like these, but varying the dominance coefficient and expanding the
+population by a factor of 1.004 each generation, starting $9.3$ generations in
+to mimic the human expansion out-of-Africa (XXX). More details about these
+simulations can be found in Supplementary Section \ref{supp:simulation}.
 

### 2023-07-01 manuscript/supplementary.tex
more methods

@@
 \section{Quantitative Genetic Linked Selection Model Theory}
-\label{suppsec:theory}
+\label{supp:theory}
 
@@
 \subsection{Heritable Fitness and the Accumulation of Autocovariance}
-\label{suppsec:heritable-fitness}
+\label{supp:heritable-fitness}
 
@@
 \subsection{Pairwise Diversity Estimates}
-\label{suppsec:diversity}
+\label{supp:diversity}
 
@@
 
-\subsection{}
-
 In order to try to ensure accurate estimation of pairwise diversity, which is a
@@
 per-basepair genotype calls (gVCF) produced by Illumina's DRAGEN pipeline
-\parencite{Illumina_Inc2020-dk}. The original samples were from 178 Yoruban
-individuals sequenced to 30x by the New York Genome Center
-\parencite{Byrska-Bishop2022-tn}. This allows for filtering to be applied to
-the entire genome at once, rather than just variants, so the denominator does
-not need to be estimated separately.
+\parencite{Illumina_Inc2020-dk}. The original samples were sequenced to 30x by
+the New York Genome Center \parencite{Byrska-Bishop2022-tn}. This allows for
+filtering to be applied to the entire genome at once, rather than just
+variants, so the denominator does not need to be estimated separately.
 
@@
 
+XXX
+
 \subsection{Filtering gVCFs}
+\label{supp:filter}
 
@@
       GRCh38 reference genome are also masked. Soft-masked regions were
-      determined by Ensembl (XXX), which uses Repeat Masker (XXX) and Dust
-      (XXX). Additionally, we used the 1000 Genomes strict filter (XXX).
+      determined by Ensembl \parencite{Cunningham2022-vk}, which uses Repeat
+      Masker \parencite{Smit2015-nm} and Dust \parencite{Morgulis2006-uo}. We
+      also only used sites that pass the 1000 Genomes strict filter
+      \parencite{1000_Genomes_Project_Consortium2015-mi}.
 
@@
 \subsection{Putatively Neutral Tracks}
+\label{supp:neutral}
 
@@
 where $V_\text{res}$ is the mean squared error across windows and
-$V_\text{tot}$ is the total variance in the predicator. Suppose we have
+$V_\text{tot}$ is the total variance in the observed values. Suppose we have
 estimates $\widehat{\pi}_0 \widehat{B}_i$ for $\pi_0 B_i$ from our model. Then,
@@
 
-The total variance in the predictor is,
+The total variance among the observed diversity values is,
 
@@
 \subsection{Likelihood}
+\label{supp:likelihood}
 
@@
 
-As in Section \ref{suppsec:diversity}, the number of pairwise differences and
+As in Section \ref{supp:diversity}, the number of pairwise differences and
 the total number of pairwise comparison at a site are sufficient statistics for
@@
 
-\subsection{Out-Sample Error Estimation and $R_\text{LOO}^2$}
-
-We fit each of our B' model fits 22 times, excluding a whole chromosome each
-time to calculate out-sample prediction error. Each fit used 10,000 random
-starting positions as the main fits. The out-sample prediction error for all
-bins in chromosome $C$ left out is,
-
-\begin{align}
-    SSE(C) = \sum_{b \in C} (\widehat{\pi}_{b,Y\not \in C}  - \pi_b)^2
-\end{align}
-%
-
-where $\widehat{\pi}_{b,Y\not \in C}$ denotes the predictions for window $b$
-fit on data excluding chromosome $C$. The average out-sample error $SSE$ across
-chromosomes is then the average of $SSE(C)$ across chromosomes, weighted by the
+\subsection{Out-Sample Error Estimation and $\widehat{R}_\text{LOCO}^2$}
+
+To estimate out-sample prediction error, we fit each model leaving one
+chromosome out (LOCO) and calculating the out-sample prediction error on
+these excluded chromosomes. Each fit used 10,000 random starting positions as
+the main fits. The out-sample prediction error (i.e. sum of residuals
+squared, $\widehat{\mathrm{SSR}}_\text{LOCO}$) for all bins in chromosome $C$
+left out is,
+
+\begin{align}
+    \widehat{\mathrm{SSR}}_\text{LOCO}(C) = \sum_{b \in C} (\widehat{\pi}_{b, (C)}  - \pi_b)^2
+\end{align}
+%
+where $\widehat{\pi}_{b,(C)}$ denotes the predictions for window $b$ fit on
+data excluding chromosome $C$. The average out-sample error
+$\widehat{\mathrm{SSR}}_\text{LOCO}$ across chromosomes is then the average of
+$\widehat{\mathrm{SSR}}_\text{LOCO}(C)$ across chromosomes, weighted by the
 number of bins per chromosome. We then calculate the out-sample coefficient of
-determination $R_\text{LOO}^2$
-
-\begin{align}
-    R_\text{LOO}^2 = \frac{SSE}{}
-\end{align}
-
-XXX
+determination $\widehat{R}_\text{LOCO}^2$
+
+\begin{align}
+    \widehat{R}_\text{LOCO}^2 = 1-\frac{\widehat{\mathrm{SSR}}_\text{LOCO}}{\widehat{\mathrm{SST}}_\text{LOCO}}
+\end{align}
+%
+where,
+
+\begin{align}
+    \widehat{\mathrm{SST}}_\text{LOCO}(C) = \sum_{b \in C} (\widehat{\pi}_{b, (C)}  - \bar{\pi})^2
+\end{align}
+%
+and $\bar{\pi}$ is the average diversity across all windows.
 
 \subsection{Jackknife Procedures}
+\label{supp:jackknife}
 
@@
 
+\subsection{Chromosome 10 Simulations}
+
+XXX
+, using the HapMap recombination map (XXX) and in coding regions, PhastCons
+regions, and UTRs as an approximate set of putatively conserved segments
+
+We calculate the mean squared error (MSE) and relative error between
+whole-chromosome simulations and the predicted B and B' maps by averaging
+branch diversity in 10 kilobase windows (tskit, XXX) and dividing by the
+neutral expectation of $4N$. The theoretic lower bound of the MSE is due the
+noise in estimating the expected reduction map from simulations. For $n=100$
+replicates, the sampling noise is minimal relative to the evolutionary noise in
+the coalescence process \parencite{Tajima1983-gu}. We use this as the theoretic
+lower bound assuming the maximum coalescence noise at $B=1$, $\var(\bar{B})
+\approx \nicefrac{2}{9n}$.
+
 \section{Additional Results and Figures}
@@
 \subsection{Optimization and Diagnostics}
-\label{suppsec:optim}
+\label{supp:optim}
 

### 2023-07-01 manuscript/manuscript.tex
more methods

@@
 we extended their model so that it could be applied to whole-genome data. Our
-software method \texttt{bgspy} calculates the equilibrium additive genic
-fitness variance ($V_a$) and ratchet rate ($R$) across grids of mutation rates
-and selection coefficients for each pre-specified segment in the genome that
-may be under negative selection (e.g. coding sequences or UTRs). Then, our
-method computes a reduction map $B(x)$ for each position $x$ in the genome,
-which is determined by the fitness variation at each segment and their spacing
-along the user-supplied recombination map (see Methods XXX). To distinguish
+software method \texttt{bgspy} numerically solves Equations XXX to compute the
+equilibrium additive genic fitness variance ($V_a$) and ratchet rate ($R$)
+across grids of mutation rates and selection coefficients. This is done for
+each pre-specified segment in the genome that may be under negative selection
+(e.g. coding sequences or UTRs). Then, these equilibrium can be used to
+calculate a reduction map $B(x)$ across genomic positions $x$. To distinguish
 between McVicker's B maps based on classic BGS theory, we call our reduction
@@
 
+Ipsita Agarwal article
+
+what time scale is the average mutation being lost at?
+
 Regularization / overfitting

### 2023-07-02 manuscript/manuscript.tex
supp sim info

@@
 recombination rates in the segment. A detailed derivation of these equations
-can be found in Supplementary Section \ref{suppsec:theory}. The recombination
+can be found in Supplementary Section \ref{supp:theory}. The recombination
 rate in a segment is determined by a user-supplied recombination map.
@@
 
-We use a binomial likelihood with probability $\bar{\pi}(b; \Psi) = \pi_0
-\bar{B}(b; \mu, \mathbf{W})$, where bars indicate averages over some bin width.
-With multiple features classes and a map $k(g)$ between segments $g$ to their
-class, the reduction map is,
+Our binomial likelihood models the number of different pairwise comparisons
+observed per window given the total number of pairwise comparisons. The
+probability is given by $\bar{\pi}(b; \Psi) = \pi_0 \bar{B}(b; \mu,
+\mathbf{W})$, where bars indicate averages over some bin width, and the
+reduction is
 
 \begin{align}
-    \log\left(B(x; W, s_j) \right) = - \frac{1}{2} \sum_{g \in G} \sum_{i=1}^{n_s} V_g(\mu W_{j, k(g)}, s_j) Q_g^2(m_i, s_j, r_{x, g}).
+    \log\left(B(x; \mu, \mathbf{W}) \right) = - \frac{1}{2} \sum_{g \in G} \sum_{j=1}^{n_s} V_g(\mu W_{j, k(g)}, s_j) Q_g^2(m_i, s_j, r_{x, g}).
 \end{align}
-
+%
 Here, $\mu$ is the mutation rate and $W_{j,k(g)}$ is a DFE entry in the
 $\mathbf{W}$ is the $n_s \times K$ DFE matrix over the selection grid and
-features. Since each column of the DFE $\mathbf{W}$ must sum to one, numeric
-optimization was done in unconstrained $n_s-1$ dimensional softmax space. Our
-binomial likelihood modeled the number of different pairwise comparisons
-observed per window given the predicted pairwise diversity and total number of
-comparisons implied per window (see Supplementary Materials Sections
-\ref{supp:data-summary} and \ref{supp:likelihood}).
-
-We tested multiple different optimization routines, finding that BOBYQA
-outperformed alternatives \parencite{Powell_undated-ec}. We inspected and
-confirmed convergence with diagnostic plots finding stable optima across 10,000
-random starts (see Supplementary Materials Section \ref{suppsec:optima}). We
-assessed model fit using out-sample predictive error, calculated by leaving out
-a whole chromosome during fitting and predicting its diversity. To calculate
-uncertainty, we used a block jackknife approach in 10 Mbp windows
-(Supplementary Section \ref{supp:jackknife}).
+features. See Supplementary Materials Sections \ref{supp:data-summary} and
+\ref{supp:likelihood} for more details.
+
+Our method uses two tricks to improve optimization over the mutation and DFE
+parameters. First, we find $B(x; \mu, \mathbf{W})$ is exponential over columns
+of $\mu \mathbf{W}$, which allows for optimization over this smooth function
+rather than the grid. Second, we use softmax to convert constrained
+optimization over the DFE columns (which must sum to one) to unconstrained. We
+tested multiple different optimization routines, finding that BOBYQA
+outperformed alternatives \parencite{Powell2009-jm,Johnson2007-tl}. We
+inspected and confirmed convergence with diagnostic plots finding stable optima
+across 10,000 random starts (see Supplementary Materials Section
+\ref{supp:optima}). We assessed model fit using out-sample predictive error,
+calculated by leaving out a whole chromosome during fitting and predicting its
+diversity. To calculate uncertainty, we used a block jackknife approach in 10
+Mbp windows (Supplementary Section \ref{supp:jackknife}).
 
@@
 
-We use two types of forward simulations in this paper; both use SLiM version
-4.0.1 \parencite{Haller2023-uk,Haller2019-vu}. First, we simulated a 100 kbp
-segment in a population of $N=1000$ diploids across grids of mutation and
-selection parameters to assess the accuracy of the SC16 theory. Each simulation
-recorded the number of fixations and the fitness variance per generation, and
-at the end $10N$ generations, the ancestral recombination graph was recorded so
-downstream Python scripts could calculate diversity. Each of these segment
-simulations was replicated $10,000$ times and averaged to compare against
-theoretic expectations.
-
-Second, we conducted forward simulations of negative selection on whole human
-chromosomes. These simulations used the HapMap recombination map
-\parencite{International_HapMap_Consortium2007-nu} and treated coding,
-PhastCons, and UTR regions as an approximate set of conserved regions. We used
-these whole-chromosome simulations for three purposes. First, we simulated
-negative selection on chromosome 10 to confirm our method to estimate classic B
-and new B' maps matched the average empirical reduction map across a grid of
-mutation and selection parameters. Second, we evaluated our composite
-likelihood method by simulating negative selection on the first five human
-chromosomes, across grids of mutation and selection parameters. Multiple
-independently-evolved chromosomes were combined into a synthetic genome in
-order to have enough data for our procedure. Third, we conducted additional
-simulations like these, but varying the dominance coefficient and expanding the
-population by a factor of 1.004 each generation, starting $9.3$ generations in
-to mimic the human expansion out-of-Africa (XXX). More details about these
-simulations can be found in Supplementary Section \ref{supp:simulation}.
+We conducted forward simulations of negative selection on whole human
+chromosomes to validate our method at two stages. First, we simulated negative
+selection on chromosome 10 using a realistic recombination map and putatively
+conserved features to confirm that our classic B and new B' maps matched the
+average simulation reduction map across mutation and selection parameters.
+Second, we evaluated our composite likelihood method by simulating negative
+selection on the first five human chromosomes, across grids of fixed mutation
+and selection parameters. We then combined these into a synthetic genome, and
+overlaid mutations on the ARG. Then, we ran our likelihood methods on the
+resulting allele count data to assess model accuracy. We ran additional
+synthetic genome simulations like these to evaluate the impact of two model
+violations: recessivity of deleterious mutations and expanding populations. For
+the latter, after $9.3$ generations, we grew the population by factor of 1.004
+each generation to mimic the human expansion out-of-Africa
+\parencite{Gutenkunst2009-pg}. We did not simulate population bottlenecks since
+our analyses showed little difference between bottlenecked out-of-Africa
+samples (CEU and CHB) and YRI samples. More details about these and the segment
+simulations shown in Figure \ref{fig:figure-1}A-C can be found Supplementary
+Section \ref{supp:simulation}.
 

### 2023-07-02 manuscript/supplementary.tex
supp sim info

@@
 
-\subsection{Derivation of Approximate Drift $R^2$}
+\subsection{Derivation of Approximate Coalescence $R^2$}
+\label{supp:r2-coal}
 
@@
 \section{Simulations}
+\label{supp:simulation}
+
+All of our forward simulations used SLiM version 4.0.1
+\parencite{Haller2019-vu,Haller2023-uk} and all Eidos scripts are available in
+the \path{slim_sims/} directory in the GitHub repository
+(\url{https://github.com/vsbuffalo/bprime}). These Eidos scripts take command
+line arguments for the parameters. Our simulation pipelines use YAML
+configuration files that specify the parameter grids and other settings, which
+are read and dispatched across a cluster using Snakemake
+\parencite{Koster2012-iv}.
 
@@
 
-\subsection{Chromosome 10 Simulations}
-
-XXX
-, using the HapMap recombination map (XXX) and in coding regions, PhastCons
-regions, and UTRs as an approximate set of putatively conserved segments
-
-We calculate the mean squared error (MSE) and relative error between
-whole-chromosome simulations and the predicted B and B' maps by averaging
-branch diversity in 10 kilobase windows (tskit, XXX) and dividing by the
-neutral expectation of $4N$. The theoretic lower bound of the MSE is due the
-noise in estimating the expected reduction map from simulations. For $n=100$
-replicates, the sampling noise is minimal relative to the evolutionary noise in
-the coalescence process \parencite{Tajima1983-gu}. We use this as the theoretic
-lower bound assuming the maximum coalescence noise at $B=1$, $\var(\bar{B})
-\approx \nicefrac{2}{9n}$.
+To confirm the SC16 theory, we simulated a 100 kbp segment in a population of
+$N=1000$ diploids across grids of mutation and selection parameters. The Eidos
+script for these simulations can be found at
+\path{slim_sims/region/region.slim} and the analysis notebook at
+\path{notebooks/region_simulations.ipynb} in the GitHub repository
+(\url{https://github.com/vsbuffalo/bprime}). We simulated over grids of
+mutation rates ($\mu \in \{10^{-9.5}, 10^{-9}, 10^{-8.5}, 10^{-8}\}$) and
+selection coefficients ($s \in \{10^{-1}, 10^{-1.5}, 10^{-2}, \ldots,
+10^{-4.5}\}$). Selection was additive within a site and multiplicative across
+sites, and recombination was fixed at $10^{-8}$ per basepair.
+
+We replicated $10,000$ simulations for each parameter combination to average
+and compare against theory. Each simulation recorded the number of fixations
+and the fitness variance per generation. At the end $10N$ generations, the
+ancestral recombination graph was recorded. Then, downstream Python scripts
+calculate branch-length diversity using \texttt{tskit}
+\parencite{Kelleher2018-qu}, and we estimate the reduction factor in a window
+centered in the region with $B = \nicefrac{\pi}{4N}$ (since branch length
+diversity implies $\mu=1$).
+
+\subsection{Whole-chromosome Simulations}
+
+Our whole-chromosome simulations use realistic human recombination maps and
+putatively conserved segments. These annotation differ slightly from our
+primary model fits, since these simulations were written before the method was
+finalized. We use the HapMap recombination map
+\parencite{International_HapMap_Consortium2007-nu}, PhastCons regions, and UTRs
+as an approximate set of putatively conserved segments.
+
+Our first set of simulations were used to validate that our methods were
+producing accurate classic B and B' maps. We replicated $100$ whole-chromosome
+simulations over grids of mutation rates ($\mu \in \{10^{-10}, 10^{-9.5},
+\ldots, 10^{-8}, 10^{-7.5}\}$) and selection coefficients ($s \in \{10^{-1},
+10^{-1.5}, 10^{-2}, \ldots, 10^{-6}\}$). For each parameter combination, we
+calculate mean branch diversity in 10 kbp windows across the 100 replicates to
+create the average simulation reduction map.
+
+We then compare this average simulation map to our method's theoretic B' maps
+(see Figure \ref{fig:figure-1} D and E). We calculate the MSE per 10 kbp window
+to estimate the total rate across the chromosome. We can assess the accuracy
+relative to a theoretic lower bound of the MSE if the noise were due to
+coalescence noise. We approximate this by noting that the MSE can be decomposed
+into a squared bias term and a variance term. If our methods are unbiased in
+estimating the true reduction at a position, the MSE is just the stochastic
+variance due to the $n=100$ replicates we simulate and the coalescence noise in
+each one. \textcite{Tajima1983-gu} showed that variance in diversity due to
+coalescence noise is much larger than sampling noise, and is given by
+$\var(\pi) \approx \nicefrac{2\theta^2}{9}$ in an infinite sample. One can
+derive an analogous equation for the variance of average simulation $\bar{B}$
+across $n$ replicate simulations as, $\var(\bar{B}) \approx
+\nicefrac{2B^2}{9n}$. We note that this assumes no recombination but this
+approximation is fine over short scales (such as the 10 kbp scale used here).
 

### 2023-07-02 bgspy/nonparametric.py
figs checked in

@@ -1,6 +1,7 @@
 import numpy as np
 from sklearn.utils.validation import check_X_y
 import warnings
+from scipy import stats
 from sklearn.model_selection import KFold, LeaveOneOut
 from sklearn.model_selection import cross_validate
 from sklearn.base import BaseEstimator
@@ -142,3 +143,25 @@ def lowess_bootstrap(x, y, frac, nboot=200, alpha=0.05):
     lb = np.percentile(res, 100 * alpha/2, axis=0)
     ub = np.percentile(res, 100 * (1-alpha/2), axis=0)
     return np.array(res), lb, ub
+
+
+def linregress_bootstrap(x, y, nboot=200, alpha=0.05):
+    n = len(x)
+    res = []
+    for i in range(nboot):
+        idx = np.random.choice(range(n), size=n, replace=True)
+        x_rs = x[idx]
+        y_rs = y[idx]
+        xs = np.sort(x_rs)
+        slope, yint, *lrres = stats.linregress(x_rs, y_rs)
+        pred = xs * slope + yint 
+        res.append(pred)
+    res = np.array(res)
+    lb = np.percentile(res, 100 * alpha/2, axis=0)
+    ub = np.percentile(res, 100 * (1-alpha/2), axis=0)
+    lb = np.percentile(res, 100 * alpha/2, axis=0)
+    ub = np.percentile(res, 100 * (1-alpha/2), axis=0)
+    return np.array(res), lb, ub
+
+
+

### 2023-07-02 bgspy/plots.py

@@ -313,7 +313,7 @@ def predicted_observed(fit, n_bins=40, smooth_col = 'orange',
 def binned_means_plot(df, min_n=None, gen=None,
                       stat='mean', s=5,
                       c='0.22', linewidth=0.5,
-                      thin=None,
+                      thin=None, zorder=None,
                       label=None, figax=None):
     fig, ax = get_figax(figax)
 
@@ -328,7 +328,7 @@ def binned_means_plot(df, min_n=None, gen=None,
         points = ax.scatter(x, mean, c=c, s=s, alpha=1, zorder=10, label=label)
         ax.errorbar(x, mean, 2*sd/np.sqrt(n), fmt='none',
                     c=points.get_edgecolor(),
-                    elinewidth=linewidth)
+                    elinewidth=linewidth, zorder=zorder)
     else:
         x, mean = x, mean
         se = (2*sd/np.sqrt(n))
@@ -337,7 +337,7 @@ def binned_means_plot(df, min_n=None, gen=None,
         points = ax.scatter(x, mean, c=c, s=s, alpha=1,
                             zorder=10, label=label)
         ax.errorbar(x, mean, se,
-                    fmt='none', c=points.get_edgecolor(), elinewidth=linewidth)
+                    fmt='none', c=points.get_edgecolor(), elinewidth=linewidth, zorder=zorder)
     return fig, ax
 
 

### 2023-07-02 manuscript/manuscript.tex
minor ms changes

@@
 %\title{Towards a Complete Model of the Negative Selection Process in Humans}
-\title{An Evolutionary Quantitative Genetic Model of the Negative Selection Process in Humans}
+%\title{An Evolutionary Quantitative Genetic Model of the Negative Selection Process in Humans}
+\title{A Quantitative Genetic Linked Selection Model fit to the Negative Selection Process in Humans}
 
@@
 \parencite{Siepel2005-wh}. As with most selection in the genome, the negative
-selection process perturbs the allele frequencies of neighboring linked
-variation. This reduces genetic variability around conserved segments and
-creates a large-scale spatial ``linked selection signal" in diversity alone the
-genome. Mutation and negative selection's contribution to this signal is
-mediated by the deleterious mutation rate, the strength of selection in
-conserved segments, and the spatial distribution of recombination rates and
-conserved segments. Since genome-wide recombination maps and putatively
-conserved segments are available for many species, researchers have combined
-these with population genetic theory to estimate the mutation rate and
-selection coefficients of deleterious mutations, as well as the overall
-reduction in genetic variability due to the negative selection process.
+selection process stochastically perturbs the allele frequencies of neighboring
+linked variation. This reduces genetic variability to create a large-scale
+spatial linked selection ``reduction signal" in genome-wide diversity. Mutation
+and negative selection's contribution to this signal is mediated by the
+deleterious mutation rate, the strength of selection in conserved segments, and
+the spatial distribution of recombination rates and conserved segments. Since
+genome-wide recombination maps and putatively conserved segments are available
+for many species, researchers have combined these with population genetic
+theory to estimate the mutation rate and selection coefficients of deleterious
+mutations, as well as the overall reduction in genetic variability due to the
+negative selection process.
 
@@
 In this work, we extend another class of linked selection models that derive
-from quantitative genetics to address limitations of the classic BGS model.
-These models consider how polygenic fitness variation increases the rate of
-stochastic allele frequency change as alleles become randomly linked to fitness
-backgrounds. While these models can theoretically accommodate polygenic fitness
-variation from any source as long as its rate of change is not too rapid, we
-focus specifically on a deleterious-mutations-only model from Santiago and
-Caballero (2016). This model is identical to the classic BGS model when
-selection against deleterious mutations is strong, but it also correctly
-predicts the reduction in diversity and substitutions rates of weakly
-deleterious mutations by accurately modeling levels of genic fitness variation.
-We extend Santiago and Caballero’s (hereafter the SC16) model so that it can be
-used to model genome-wide patterns of diversity under negative selection
-processes, and develop Python software to use this theoretic approach to fit
-genome-wide models of negative selection. Using forward simulations, we show
-this model leads to more accurate DFE estimates under weak selection. While
-this new model accommodates the possibility that a given reduction B could be
-due to weaker selection, in applying our composite-likelihood method to human
-population genomic data, we confirm earlier findings that strong negative
-selection at a small fraction of sites fits human data across populations best.
-Finally, we show that while this model solves some limitations of the classic
-BGS model, simulations reveal another mysterious defect: a poor fit around
-$Ns=1$. We show that this can be partially corrected by incorporating the
-effects of selective interference on genic variation, but the remaining
-discrepancy is due to the build up of negative linkage disequilibrium that
-cannot be adequately captured by this model. We find accounting for genic
-selective interference between segments fits human population genomic data
-slightly better, suggesting levels of negative selection in humans are
-sufficient to perturb the selection at other sites.
+from quantitative genetics to address limitations of classic background
+selection theory. These models suggest that the reduction in pairwise
+coalescence times at a position in the genome is determined by the
+recombination map and distribution of polygenic additive fitness variance along
+the chromosome \parencite{Santiago1998-bs}. This reduction is due to cumulative
+effect of stochastic perturbations from selection at other sites. While these
+models can theoretically accommodate polygenic fitness variation from any
+source as long as its rate of change is not too rapid, we focus specifically on
+a deleterious-mutations-only model from Santiago and Caballero (2016). This
+model is identical to the classic BGS model when selection against deleterious
+mutations is strong, but it also correctly predicts the reduction in diversity
+and substitutions rates of weakly deleterious mutations by accurately modeling
+levels of genic fitness variation. We extend Santiago and Caballero’s
+(hereafter the SC16) model so that it can be used to model genome-wide patterns
+of diversity under negative selection processes, and develop Python software to
+use this theoretic approach to fit genome-wide models of negative selection.
+Using forward simulations, we show this model leads to more accurate DFE
+estimates under weak selection. While this new model accommodates the
+possibility that a given reduction B could be due to weaker selection, in
+applying our composite-likelihood method to human population genomic data, we
+confirm earlier findings that strong negative selection at a small fraction of
+sites fits human data across populations best. Finally, we show that while this
+model solves some limitations of the classic BGS model, simulations reveal
+another mysterious defect: a poor fit around $Ns=1$. We show that this can be
+partially corrected by incorporating the effects of selective interference on
+genic variation, but the remaining discrepancy is due to the build up of
+negative linkage disequilibrium that cannot be adequately captured by this
+model. We find accounting for genic selective interference between segments
+fits human population genomic data slightly better, suggesting levels of
+negative selection in humans are sufficient to perturb the selection at other
+sites.
 
@@
 dashed colored lines), but does not entirely eliminate the error in the $2Ns
-\approx 1$ domain.
+\approx 1$ domain. Overall, while we find evidence that local-rescaling
+slightly reduces error in the $2Ns \approx 1$ domain, we do not use it during
+in our genome-wide inference method. This is because to avoid circularity and
+possible overfitting, the method would need to solve the equilibria equations
+for all segments simultaneously, but this is currently computational
+unfeasible.
 
@@
 around $2Ns \approx 1$ (Figure \ref{fig:figure-1}D, top row) and is stronger
-when mutation rates are higher (Supplementary Figure XXX). As $Ns \to 0$, the
-variance in LD inflates as expected \parencite{Ohta1969-ae,Hill1968-ue}.
-Overall, the build up of negative LD is consistent our view that the
-equilibrium fitness variance modeled by the SC16 theory is the additive
-\emph{genic} fitness variance, which differs from the additive genetic variance
-by the sum of linkage disequilibria between selected sites (i.e. $\delta_{LD} =
-s^2 \sum_{i\ne j} D_{i,j}$ where $D_{i,j}$ is the LD between sites $i$ and
-$j$). However, according to theory, the reductions in diversity should be
-determined by levels of additive genetic fitness variance that include the
-contribution of LD (Supplementary Materials Section XXX).
-
-While we find evidence that local-recalling reduces error in the $2Ns \approx
-1$ domain, we do not employ it in our genome-wide inference method for a few
-reasons. First, the reduction map
-
-numerically solving the system of equations
-
-it currently cannot be used during when fitting
-genome-wide models.
+when mutation rates are higher. As $Ns \to 0$, the variance in LD inflates as
+expected \parencite{Ohta1969-ae,Hill1968-ue}. Overall, the build up of negative
+LD is consistent our view that the equilibrium fitness variance modeled by the
+SC16 theory is the additive \emph{genic} fitness variance, which differs from
+the additive genetic variance by the sum of linkage disequilibria between
+selected sites (i.e. $\delta_{LD} = s^2 \sum_{i\ne j} D_{i,j}$ where $D_{i,j}$
+is the LD between sites $i$ and $j$). However, according to theory, the
+reductions in diversity should be determined by levels of additive genetic
+fitness variance that include the contribution of LD (Supplementary Materials
+Section \ref{supp:theory}).
 
@@
 
-Our method takes tracks of annotated features (an ``annotation model") that are
-\emph{a priori} expected to have similar fitness effects, and estimates the
-distribution of fitness effects for each feature type. We consider two classes
-of annotation models: (1) CADD score models, which assign a rank-order
-pathogenicity score to each basepair, and (2) and more interpretable gene
-feature-based models that includes protein coding regions, intron and UTRs, and
-PhastCons regions. We include PhastCons regions because there is abundant
-evidence of highly-conserved non-coding regions in humans under negative
-selection
+Our composite likelihood method takes tracks of annotated features (an
+``annotation model") that are \emph{a priori} expected to have similar fitness
+effects, and estimates the overall mutation rate and distribution of fitness
+effects for each feature type. We consider two classes of annotation models:
+(1) CADD score models, which assign a rank-order pathogenicity score to each
+basepair, and (2) and more interpretable gene feature-based models that
+includes protein coding regions, intron and UTRs, and PhastCons regions. We
+include PhastCons regions because there is abundant evidence of
+highly-conserved non-coding regions in humans under negative selection
 \parencite{Meader2010-hm,Harmston2013-tt,Katzman2007-gq,Siepel2005-wh}, but
@@
 patterns of diversity on whole chromosomes left-out during the model fitting
-process. We use the metric $R_\text{LOO}^2$, which is the proportion of the
-observed variance in genomic diversity at the megabase scale predicted by our
-model. We experimented with a few smaller spatial scales, but our results were
-consistent with previous results suggesting the human linked selection signal
-fits best at the megabase scale \parencite{Murphy2022-sj}.
+process (e.g. leave-one-chromosome-out, LOCO). We use the metric
+$R_\text{LOCO}^2$, which is the proportion of the observed variance in genomic
+diversity at the megabase scale predicted by our model. We experimented with a
+few smaller spatial scales, but our results were consistent with previous
+results suggesting the human linked selection signal fits best at the megabase
+scale \parencite{Murphy2022-sj}.
 
@@
     method and diamonds are the classic B. Lighter color round points are
-    the out-sample $R_\text{LOO}^2$ estimates for our B' method, and arrows show the decline
-    in goodness-of-fit due to in-sample overfitting (out-sample $R_\text{LOO}^2$ 
+    the out-sample $R_\text{LOCO}^2$ estimates for our B' method, and arrows show the decline
+    in goodness-of-fit due to in-sample overfitting (out-sample $R_\text{LOCO}^2$ 
     were not calculated for classic B values due to computational costs). The 
@@
 marginally better than the CADD 6\% model consistently across all three
-populations (Supplementary Table XXX). This is likely because the PhastCons
-Priority model has more DFE parameters, and thus can fit finer-scale
+populations (Supplementary Table \ref{supp:tbl-r2}). This is likely because the
+PhastCons Priority model has more DFE parameters, and thus can fit finer-scale
 differences in the DFE across its four feature types. By contrast, the CADD 6\%
@@
 classic BGS theory, it should in principle fit equally well when an annotation
-model includes regions that are likely neutrally evolution. To test this, for
-each annotation model (which are ``sparse") we fit a ``full" model that assigns
-the remaining complement of the genome as a feature called ``other". This other
-category is likely to predominantly neutrally evolving. Ideally, the
+model includes regions that are under no selective constrain and neutrally
+evolving. To test this, for each annotation model (which are ``sparse") we fit
+a corresponding ``full" model that assigns the remaining complement of the
+genome as a feature called ``other". This category contains most of the genome,
+and is likely to be predominantly neutrally evolving. Ideally, the
 goodness-of-fit of model using our B' method should be invariant to whether an
 annotation model is sparse or full. Indeed, we find this to be the case: both
-in-sample $R^2$ and out-sample $R_\text{LOO}^2$ values are nearly identical
-across full and sparse-track models (Figures \ref{fig:figure-2} A and B, round
-points).
+in-sample $R^2$ and out-sample $R_\text{LOCO}^2$ values are nearly identical
+across full and sparse-track models (Figures \ref{fig:figure-2}A and
+\ref{fig:figure-2}B, round points).
 
@@
 behavior is what we would expect given that the CADD annotation models contain
-only the most pathogenic sites, which are \emph{a prior} very likely under the
+only the most pathogenic sites, which are \emph{a priori} very likely under the
 strong selection domain under which B' and classic BGS theory agree.
 
-Since our $R_\text{LOO}^2$ estimates suggest our models fit left-out
+Since our $R_\text{LOCO}^2$ estimates suggest our models fit left-out
 chromosomes well even though our method assumes constant demography and
@@
 
-Still, we find that our out-sample $R_\text{LOO}^2$ for Yoruba individuals
-($R_\text{LOO}^2 \approx 68$\%) is close to the theoretic $R_\text{coal}^2
+Still, we find that our out-sample $R_\text{LOCO}^2$ for the Yoruba samples
+($R_\text{LOCO}^2 \approx 68$\%) is close to the theoretic $R_\text{coal}^2
 \approx 67$\%. For bottlenecked out-of-Africa populations, $R_\text{coal}^2
-\approx 64$\% compared to the observed $R_\text{LOO}^2 \approx 62$\% for CEU
-and $R_\text{LOO}^2 \approx 59$\% for CHB. We note that because bottlenecks
+\approx 64$\% compared to the observed $R_\text{LOCO}^2 \approx 62$\% for CEU
+and $R_\text{LOCO}^2 \approx 59$\% for CHB. We note that because bottlenecks
 increase the variance in coalescence times beyond the level implied by the
-effective population size, the disagreement between observed $R_\text{LOO}^2$
+effective population size, the disagreement between observed $R_\text{LOCO}^2$
 and the theoretic $R_\text{coal}^2$ is likely to shrink. Overall, this suggests
@@
 
+Linked selection theory suggests the reduction signal is proportional to the
+spatial distributions of recombination and fitness variance along the
+chromosome.
+
 Next, we developed a composite-likelihood method to estimate expected diversity
@@
 \subsection*{Model Predictions and the Substitution Process}
+
+Each of our maximum likelihood estimates $\widehat{\Psi}$ imply other
+corresponding
 
@@
 observed per window given the total number of pairwise comparisons. The
-probability is given by $\bar{\pi}(b; \Psi) = \pi_0 \bar{B}(b; \mu,
-\mathbf{W})$, where bars indicate averages over some bin width, and the
-reduction is
+binomial probability for window $b$ is $\bar{\pi}(b; \Psi) = \pi_0 \bar{B}(b;
+\mu, \mathbf{W})$, where bars indicate averages over some bin width. The free
+parameters $\Psi = \{\pi_0, \mu, \mathbf{W}\}$ are the expected diversity in
+the absence of selection ($\pi_0$), the mutation rate ($\mu$), and the
+distribution of fitness effects for the discretized selection grid and $K$
+features ($\mathbf{W}$). The reduction at position $x$ is then,
 
@@
 %
-Here, $\mu$ is the mutation rate and $W_{j,k(g)}$ is a DFE entry in the
-$\mathbf{W}$ is the $n_s \times K$ DFE matrix over the selection grid and
-features. See Supplementary Materials Sections \ref{supp:data-summary} and
+See Supplementary Materials Sections \ref{supp:data-summary} and
 \ref{supp:likelihood} for more details.
@@
 
-\subsection*{Substitution Rate Predictions}
-
-Our MLE fit $\hat{\Phi}$ also implies a set of corresponding substitution rate
-predictions for each segment under mutation-selection-draft process. There are
-two sources of variation in these predictions in this model. First, there is
-the variation across the pre-specified functional categories, e.g. CDS, UTRs,
-and PhastCons. Second, there is region-specific variation
+\subsection*{Substitution Rate and Total Genic Fitness Variance Predictions}
+
+a set of corresponding substitution rate predictions for each segment under
+mutation-selection-draft process. There are two sources of variation in these
+predictions in this model. First, there is the variation across the
+pre-specified functional categories, e.g. CDS, UTRs, and PhastCons. Second,
+there is region-specific variation
+
+The SC16 theory models linked selection in quantitative genetics terms and
+relies on simultaneously predicting the deleterious substitution rate.
 

### 2023-07-02 manuscript/supplementary.tex
minor ms changes

@@
   \caption{The distribution of diversity across the genome at the megabase
-  scale, with outliers trimmed. The blue line is the normal CDF, fit with MLE
-parameters XXX. The inset figure is the untrimmed data, with trimmed points
-shown in red.}
+  scale, with outliers trimmed. The blue line is the best-fitting normal CDF.
+  We note that this implies the distribution of megabase-scale diversity 
+  is slightly fat-tailed compared to a normal distribution.
+  The inset figure is the untrimmed data, with trimmed points shown in red. }
   \label{suppfig:trimming}
@@
 
-The likelihood method is fit to non-overlapping binned summaries of the allele
-counts matrix.
-
-Based on exploratory data analyses, some bins were outliers and
-excluded (XXX).
-
-\begin{enumerate}
-  \item \textbf{Fraction of inaccessible sites per window}.  \path{mask_inaccessible_bins_frac}
-  % \item \textbf{Edge truncation}
-
-  \item \textbf{Outliers}: Based on exploratory analysis, there were some
-    regions with very high diversity. The 0.05\% right tail was excluded.
-
-\end{enumerate}
-
-% supp data
-% - npy files
-% - edge trunation
+Based on exploratory data analyses, we noticed some rare windows that were
+outliers. Given that maximum likelihood estimation can be easily biased by
+outliers, we took a conservative approach and removed the top 0.5\% of
+high-diversity windows. The results of this can be seen in Supplementary
+Figure \ref{suppfig:trimming}. We also remove 1cM from each end of the
+recombination map. These window-based filters are in addition to the
+accessible and neutral sequence masked described in Supplementary Sections
+\ref{supp:filter} and \ref{supp:accessible}.
 
@@
 \begin{table}
-\centering
-\caption{$R^2$ and mutation rate estimates for all models. XXX note about repeats}
+    \label{supp:tbl-r2}
+\centering
+\caption{$R^2$ and mutation rate estimates for all models. Note that the repeated values in the last 
+rows are correct and due to rounding; see \path{notebooks/main_fits.ipnb} in the GitHub repository for 
+more information.}
 \begin{tabular}{lll|crr|cc}
@@
 
-\subsection{Derivation of Approximate Coalescence $R^2$}
+\subsection{Derivation of Approximate Coalescence $R_\text{coal}^2$}
 \label{supp:r2-coal}

### 2023-07-03 manuscript/manuscript.tex
minor ms changes

@@
 
+\begin{figure}[htbp] \centering
+    \includegraphics[width=\textwidth]{figures/figure_6.pdf}
+
+    \caption{Evidence that negative selection signal remains. (A) Residuals
+        from the YRI CADD 6\% model plotted against the average LoF selection
+        coefficient across genes in a megabase window (estimated from
+        \cite{Agarwal2023-un}). Our model is systematically over predicting
+        diversity in regions estimated to be sensitive to LoF variants. The
+        ordinary least squares (OLS) estimate suggests that 2.1\% of the
+        variance in residuals can be explained by reductions caused by the
+        local sensitivity to LoF mutations. (B) Residuals from the YRI CADD 6\%
+        model plotted against the fraction of CADD 6\% basepairs in the
+        megabase window. Here, only 0.8\% of the variance is explained by
+        remaining spatial variance in CADD 6\% regions (blue). Overlaid is the CADD
+        2\% OLS curve (orange). }
+
+  \label{fig:figure-6}
+\end{figure}
+
 Each of our maximum likelihood estimates $\widehat{\Psi}$ imply other

### 2023-07-03 manuscript/supplementary.tex
minor ms changes
minor ms changesul

@@
 
+\subsection{Additional Residual Diagnostics}
+
+\begin{figure}[htbp]
+  \label{suppfig:resid-b}
+  \centering
+  \includegraphics[width=0.6\textwidth]{figures/supplementary/b_resid.pdf}
+  \caption{Residuals from the CADD 6\% model plotted against the predicted $B(x)$
+           value.}
+\end{figure}
+
+\begin{figure}[htbp]
+  \label{suppfig:resid-pc-fit}
+  \centering
+  \includegraphics[width=0.6\textwidth]{figures/supplementary/phastcons_resid_lof.pdf}
+  \caption{Residuals from the PhastCons Priority model plotted against the
+  average LoF selection coefficient across genes in a megabase window
+(estimated from \cite{Agarwal2023-un}). These results show that the remaining likely 
+selection signal is not dependent on the model used (PhastCons Priority versus CADD 6\%).}
+\end{figure}
+
 \clearpage
@@
   \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr1.pdf}
+  \label{suppfig:fit-chr1}
+\end{figure}
+
+\begin{figure}[!htb]
+  \centering
   \includegraphics[width=\textwidth]{figures/supplementary/pred_plot_chr2.pdf}
-  \label{suppfig:fit-chr1}
+  \label{suppfig:fit-chr2}
 \end{figure}

### 2023-07-03 bgspy/likelihood.py
new figs

@@ -413,6 +413,8 @@ class BGSLikelihood:
         kwargs = {**defaults, **kwargs}
         mids = self.bins.midpoints()
         idx = self.bins.chrom_indices(chrom)
+        if not len(idx):
+            return np.nan
         y = self.predict(B=True)
         func = interpolate.interp1d(mids[chrom], y[idx],
                                     fill_value=(y[0], y[-1]),

### 2023-07-03 manuscript/supplementary.tex
minor ms changes

@@
   model fit on YRI data. The ranges at the top of the figure indicate the
-untrimmed and the 2 IQR-trimmed jackknife intervals. for XXX fix fig }
+untrimmed and the 2 IQR-trimmed jackknife intervals. Trimming off the numeric
+optimization outliers leads to more realistic estimates of uncertainty.}
 
@@
 
+\subsection{Additional DFE Figures}
+
+\begin{figure}[htbp]
+  \label{suppfig:dfe-ceu}
+  \centering
+  \includegraphics[width=0.6\textwidth]{figures/supplementary/figure_ceu_dfe.pdf}
+  \caption{The DFE across models for the CEU reference sample (analogous to the
+  DFE for YRI samples in Figure \ref{fig:figure-3}.}
+\end{figure}
+
+\begin{figure}[htbp]
+    \label{suppfig:dfe-chb}
+  \centering
+  \includegraphics[width=0.6\textwidth]{figures/supplementary/figure_chb_dfe.pdf}
+  \caption{The DFE across models for the CHB reference sample (analogous to the
+  DFE for YRI samples in Figure \ref{fig:figure-3}.}
+\end{figure}
+
 \clearpage

### 2023-07-03 bgspy/command_line.py
talapas model fits

@@ -36,13 +36,18 @@ LogLiks = namedtuple('LogLiks', ('pi0', 'pi0_ll', 'w', 't', 'll'))
 # sorry non-human researchers, I like other organisms too, just for
 # my sims :)
 # midpooint between 1e-7 and 1e-8 on a log10 scale
-MIN_W = np.sqrt(1e-8 * 1e-7)
+#MIN_W = np.sqrt(1e-8 * 1e-7)
 #HUMAN_W = (-10, np.log10(MIN_W))
 HUMAN_W = (-11, -7)
 HUMAN_T = (-8, -1)
+#HUMAN_T = (-8, -2)
 
 HUMAN_W_STR = f"{HUMAN_W[0]}:{HUMAN_W[1]}:6"
-HUMAN_T_STR = f"{HUMAN_T[0]}:{HUMAN_T[1]}:8"
+# use in all main fits:
+#HUMAN_T_STR = f"{HUMAN_T[0]}:{HUMAN_T[1]}:8"
+
+# new test:
+HUMAN_T_STR = f"{HUMAN_T[0]}:{HUMAN_T[1]}:7"
 
 
 def sterialize_mu(mu):
@@ -52,15 +57,19 @@ def sterialize_mu(mu):
         raise ValueError("mu must be 'free'/None, or a float")
     return mu
 
-
-def grid_maker(nw, nt, w_range=HUMAN_W, t_range=HUMAN_T):
-    return Grid(w=np.logspace(*w_range, nw), t=np.logspace(*t_range, nt))
-
-
 def grid_maker_from_str(x):
-    nw, nt = tuple(map(int, x.split('x')))
-    return grid_maker(nw, nt)
-
+    """
+    Pre-set grids, for humans. 
+    """
+    if x == '6x8':
+        # the default fits, from log10(s) -1 to -8 
+        return Grid(w=np.logspace(*HUMAN_W, 6), t=np.logspace(*HUMAN_T, 8)) 
+    elif x == '6x7':
+        # the grid from log10(s) -2 to -8
+        return Grid(w=np.logspace(*HUMAN_W, 6), t=np.logspace(*(-8, -2), 7)) 
+    else:
+        raise ValueError("x must be '6x8' (log10(s) from -1 to -8) or '6x7' (log10(s) from -2 to -8)")
+        
 
 def make_bgs_model(seqlens, annot, recmap, conv_factor, w, t, g=None,
                    chroms=None, name=None, genome_only=False,

### 2023-07-03 bgspy/likelihood.py

@@ -662,6 +662,8 @@ class BGSLikelihood:
         draw = np.random.normal(self.theta_, self.sigma_)
         # truncate the bounds of W
         W = draw[2:].reshape((self.nt, self.nf))
+        # small number; if zero, the mutation rate = 0 and downstream stuff
+        # fails
         W[W < 0] = 1e-100
         W[W > 1] = 1
         W = W / W.sum(axis=0)

### 2023-07-03 bgspy/pipeline.py

@@ -17,12 +17,26 @@ from bgspy.likelihood import SimplexModel
 
 AVOID_CHRS = set(('M', 'chrM', 'chrX', 'chrY', 'Y', 'X'))
 
+def calc_pred_stats(df):
+    """
+    Calc the stats from a predictions dataframe (of all segments)
+    """
+    V = df['V'].values.sum()
+    Vm = df['Vm'].values.sum()
+    load = df['load'].values.sum()
+    ave_r = np.average(df['r'].values, weights=df['seglen'])
+    # feature-level average sub rates
+    dfg = df.groupby('feature').mean().reset_index()
+    r = dict(zip(dfg['feature'], dfg['r']))
+    return dict(V=V, Vm=Vm, load=load, ave_r=ave_r, r=r)
+
 
 def load_model(base_dir='./', jackwidth=20_000_000):
     initial = {}
     rescaled = {}
     initial_predict = {}
     mu_predicts = defaultdict(dict)
+    mu_predicts_boot = defaultdict(dict)
     rescaled_predict = {}
     pop_dirs = [d for d in os.listdir(base_dir) if isdir(join(base_dir, d)) and "pop_" in d]
     for pop_dir in pop_dirs:
@@ -39,6 +53,8 @@ def load_model(base_dir='./', jackwidth=20_000_000):
                 dir_paths = [fit_dir_path, rescale_dir_path]
                 runs = 'initial', 'rescaled'
                 for i, dir_path in enumerate(dir_paths):
+                    if not os.path.exists(dir_path):
+                        continue
                     # load the main results if mle.pkl exists
                     mle_file_path = join(dir_path, 'mle.pkl')
                     pred_file_path = join(dir_path, 'predicted_subrates.tsv')
@@ -71,8 +87,18 @@ def load_model(base_dir='./', jackwidth=20_000_000):
                         for pred in other_preds:
                            mu = float(pred.replace('predicted_subrates_', '').replace('.tsv', ''))
                            mu_predicts[(pop, window, type_)][mu] = pd.read_csv(join(dir_path, pred), sep='\t')
+                           # check if bootstraps exist
+                           boot_dir = join(dir_path, f"predict_boots__{mu}")
+                           if os.path.exists(boot_dir):
+                               boots = os.listdir(boot_dir)
+                               if len(boots):
+                                   mu_predicts_boot[(pop, window, type_)][mu] = []
+                                   for boot in boots:
[…]

### 2023-07-03 manuscript/supplementary.tex
minor ms changes

@@
 \begin{tabular}{lll|crr|cc}
-    \textbf{Model} & \textbf{Track type} & \textbf{Pop.} & \textbf{B' $R_\text{LOO}^2$} & \textbf{B' $R_\text{IS}^2$} & \textbf{B $R_\text{IS}^2$} & \textbf{B' $\hat{\mu} \times 10^{-8}$} & \textbf{B $\hat{\mu} \times 10^{-8}$} \\[0.5ex]
+    \textbf{Model} & \textbf{Track type} & \textbf{Pop.} & \textbf{B' $R_\text{LOCO}^2$} & \textbf{B' $R_\text{IS}^2$} & \textbf{B $R_\text{IS}^2$} & \textbf{B' $\hat{\mu} \times 10^{-8}$} & \textbf{B $\hat{\mu} \times 10^{-8}$} \\[0.5ex]
 \hline
@@
 and annotation model) and parameters $\Psi$. To compare our models and assess
-model goodness-of-fit, we use the out-sample $R_\text{LOO}^2$ which is
+model goodness-of-fit, we use the out-sample $R_\text{LOCO}^2$ which is
 calculated by fitting the genome-wide model leaving one chromosome out, and
@@
 than expected variance would inflate $SS_\text{res}$, which would decrease
-$R_\text{drift}^2$, bringing it closer to the observed $R_\text{LOO}^2$. A more
+$R_\text{drift}^2$, bringing it closer to the observed $R_\text{LOCO}^2$. A more
 accurate approximation of the $R_\text{drift}^2$ could be found with
@@
 
+\begin{figure}[htbp]
+    \label{suppfig:dfe-pop-smallgrid}
+  \centering
+  \includegraphics[width=0.6\textwidth]{figures/supplementary/phastcons_priority_pop_dfe_smallselgrid.pdf}
+  \caption{The DFE for the PhastCons Priority model, fit with a smaller DFE grid from 
+  $\mu \in \{10^{-8}, 10^{-7}, \ldots, 10^{-2}\}$.}
+\end{figure}
+
 \clearpage

### 2023-07-03 manuscript/supplementary.tex
minor ms changes

@@
 
-\subsection{Additional Residual Diagnostics}
+\subsection{Demographic Expansion and Recessive Simulation Tests}
+
+\begin{figure}[htbp]
+    \label{suppfig:sim-exp}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/supplementary/subfigure_sim_mle_expansion.pdf}
+  \caption{}
+\end{figure}
+
+\begin{figure}[htbp]
+  \label{suppfig:sim-rec}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/supplementary/subfigure_sim_mle_recessive.pdf}
+  \caption{}
+\end{figure}
+
+\subsection{Residual Covariate Diagnostics}
 
@@
   \centering
-  \includegraphics[width=0.6\textwidth]{figures/supplementary/figure_ceu_dfe.pdf}
+  \includegraphics[width=0.9\textwidth]{figures/supplementary/figure_ceu_dfe.pdf}
   \caption{The DFE across models for the CEU reference sample (analogous to the
@@
   \centering
-  \includegraphics[width=0.6\textwidth]{figures/supplementary/figure_chb_dfe.pdf}
+  \includegraphics[width=0.9\textwidth]{figures/supplementary/figure_chb_dfe.pdf}
   \caption{The DFE across models for the CHB reference sample (analogous to the

### 2023-07-03 manuscript/manuscript.tex
minor ms changes
rework secs

@@
 possibility that a given reduction B could be due to weaker selection, in
-applying our composite-likelihood method to human population genomic data, we
+applying our composite likelihood method to human population genomic data, we
 confirm earlier findings that strong negative selection at a small fraction of
@@
 $R^2=68.04$\% of the out-sample variance in Yoruba pairwise diversity at the
-megabase-scale, respectively. In populations impacted by the out-of-Africa
-bottleneck, the goodness-of-fit was lower across all models (e.g. 62.98\% and
-59.98\% for CEU and CHB respectively in the PhastCons Priority model). The CADD
-6\% model was previously found by \textcite{Murphy2022-sj} to be the
-best-fitting model using classic BGS. Our PhastCons Priority model fits
-marginally better than the CADD 6\% model consistently across all three
+megabase-scale, respectively. Generally, our models predict out-sample
+diversity levels slightly better than previous methods; the best-fitting CADD
+6\% model from \textcite{Murphy2022-sj} explained 60\% of diversity in left-out
+2 Mbp windows across YRI samples.
+
+In populations impacted by the out-of-Africa bottleneck, the goodness-of-fit
+was lower across all models (e.g. 62.98\% and 59.98\% for CEU and CHB
+respectively in the PhastCons Priority model). Our PhastCons Priority model
+fits marginally better than the CADD 6\% model consistently across all three
 populations (Supplementary Table \ref{supp:tbl-r2}). This is likely because the
 PhastCons Priority model has more DFE parameters, and thus can fit finer-scale
-differences in the DFE across its four feature types. By contrast, the CADD 6\%
-track must fit patterns of diversity to conserved sites under a single DFE. The
-CADD 8\% model fit slightly worse than the CADD 6\% across all populations, and
-the Feature Priority model had the poorest-fit.
+differences with the four feature types. By contrast, the CADD 6\% track must
+fit patterns of diversity to conserved sites under a single DFE. The CADD 8\%
+model fit slightly worse than the CADD 6\% across all populations, and the
+Feature Priority model had the poorest-fit.
 
@@
 
-Since our $R_\text{LOCO}^2$ estimates suggest our models fit left-out
-chromosomes well even though our method assumes constant demography and
-homogeneous mutation rates along the genome, we wondered what $R^2$ we could
-expect under a simple theoretic model where $\cov_i(B_i', \varepsilon_i) = 0$
-and $\var_i(\varepsilon_i)$ was determined by the variance in neutral
-coalescence times alone with an effective population size rescaled by $B_i'$.
-We calculate this ``coalescence $R^2$" by calculating the theoretic coalescence
-time variance in a megabase window with recombination and an effective
-population size of $B_i' N_e$ where we use our best-fitting model's as a
-plug-in estimator of $B_i'$. We note that this is a rough approximation, since
-this calculation assumes constant demography and that the variance in
-coalescence times under negative selection is likely lower than expected under
-drift alone.
-
-Still, we find that our out-sample $R_\text{LOCO}^2$ for the Yoruba samples
-($R_\text{LOCO}^2 \approx 68$\%) is close to the theoretic $R_\text{coal}^2
-\approx 67$\%. For bottlenecked out-of-Africa populations, $R_\text{coal}^2
-\approx 64$\% compared to the observed $R_\text{LOCO}^2 \approx 62$\% for CEU
-and $R_\text{LOCO}^2 \approx 59$\% for CHB. We note that because bottlenecks
-increase the variance in coalescence times beyond the level implied by the
-effective population size, the disagreement between observed $R_\text{LOCO}^2$
-and the theoretic $R_\text{coal}^2$ is likely to shrink. Overall, this suggests
+Overall, our $R_\text{LOCO}^2$ estimates suggest our negative selection model
+explains up to 68\% of out-sample variance in diversity of the megabase scale,
+even though our method assumes constant demography and homogeneous mutation
+rates along the genome. A worthwhile question is: how much variation
+\emph{could} we expect to fit at this scale? Given that selection alters
+genealogies in ways beyond just decreasing mean pairwise coalescence time and
+populations have non-constant demography, an exact analytic answer is
+intractable. However, we can get an approximate idea if we assume that the
+residual variance $\sum_b (\pi_b - \widehat{\pi}_b)^2$ is determined entirely
+by the expected neutral coalescence noise around the expected coalescence time
+$2B(b)N$ (see Method XXX). Thus, we define $R_\text{coal}^2$ as the theoretic
+proportion of variance explainable under this model to ballpark the variance
+capable of being explained.
+
+We find that our out-sample $R_\text{LOCO}^2$ for the Yoruba samples
+($R_\text{LOCO}^2 \approx 68$\%) is slightly above the theoretic
+$R_\text{coal}^2 \approx 67$\%. This suggests our model is in the vicinity of
+fitting all possible signal possible under these simplifying assumption. It
+should be noted that selection is expected to \emph{decrease} the variance in
+coalescence times beyond simply rescaling effective population size implies;
+better future models for $R_\text{coal}^2$ could show there is more linked
+selection signal to explain.
+
+For bottlenecked out-of-Africa populations, we find a larger discrepancy
+between $R_\text{coal}^2$ and observed out-sample $R_\text{LOCO}^2$. The
+theoretic $R_\text{coal}^2 \approx 64$\% for both populations, compared to the
+observed $R_\text{LOCO}^2 \approx 62$\% for CEU and $R_\text{LOCO}^2 \approx
+59$\% for CHB. Given that bottlenecks would act to increase the residual
+variance in coalescence times beyond the level implied by the effective
+population size, the gap would likely shrink under more realistic models or
+simulation-based approximations for $R_\text{coal}^2$. Overall, this suggests
 that negative selection models explain the vast majority coalescence time
@@
 
-Linked selection theory suggests the reduction signal is proportional to the
+Our composite likelihood method has three sets of parameters: expected
+diversity in the absence of linked selection diversity $\pi_0$, the mutation
+rate $\mu$, and the distribution of fitness effects $\mathbf{W}$ across the
+selection grid for each of the $K$ feature types. Given that the relationship
+between $\pi$ and the strength selection is U-shaped, we wondered whether the
+observed genome-wide patterns of diversity could fit under weaker selection
+than found in previous methods based on classic BGS theory.
+
+However, across all of our annotation models, new deleterious mutations in
+putatively conserved features are consistently estimated to have strongly
+deleterious effects (Figure \ref{fig:figure-3}A). This includes features such
+as coding sequences, PhastCons regions, or the top 6\% of basepairs with the
+most pathogenic CADD scores. Thus, even with a model that can better
+distinguish weak versus strong selection, our DFE estimates suggest
+
+qualitatively
+agree with recent work \parencite{Murphy2022-sj}.
+
+Moreover, the estimated mutation rates are close
+to pedigree-based estimates for most of these annotation models (XXX).
+
+Second, we find evidence that our method will indeed estimate weak selection
+across features
+
+Given that the B' method more
+accurately predicts diversity across weaker selection coefficients, we wondered
+whether our inferred DFEs could
+
+Linked selection theory suggests the reduction signal is determined by the
 spatial distributions of recombination and fitness variance along the
-chromosome.
-
-Next, we developed a composite-likelihood method to estimate expected diversity
+chromosome. When fitness variation is determined mostly by negative selection,
+the
+
+Where does the information for the DFE come from?
+
+All of our models show a general trade off in fitting the linked selection
+signal.
+
+Next, we developed a composite likelihood method to estimate expected diversity
 without linked selection $\pi_0$, the deleterious mutation rate ($\mu_D$), and

### 2023-07-04 manuscript/manuscript.tex
ms changes

@@
 between $\pi$ and the strength selection is U-shaped, we wondered whether the
-observed genome-wide patterns of diversity could fit under weaker selection
-than found in previous methods based on classic BGS theory.
+observed genome-wide patterns of diversity could be explained by weaker
+selection than found in previous methods based on classic BGS theory.
 
@@
 
-\subsection*{Predicted Diversity Substitution Rates}
+\subsection*{Predicted Diversity, Substitution Rates, and Fitness Variance}
 
@@
     \caption{}
-  \label{fig:figure-4}
+  \label{fig:figure-5}
 \end{figure}

### 2023-07-04 manuscript/supplementary.tex
ms changes

@@
 \centering
-\caption{$R^2$ and mutation rate estimates for all models. Note that the repeated values in the last 
-rows are correct and due to rounding; see \path{notebooks/main_fits.ipnb} in the GitHub repository for 
-more information.}
+\caption{$R^2$ and mutation rate estimates for all primary models (top table)
+    and models fit on a smaller DFE grid (excluding the possibility of strong
+    selection $s=10^{-1}$). Note that the repeated values in the last rows are
+    correct and due to rounding; see \path{notebooks/main_fits.ipynb} in the
+    GitHub repository for more information.}
 \begin{tabular}{lll|crr|cc}
@@
 \hline
+\hline
+\emph{smaller DFE grid} & & & & & & \\
+\hline
+\textbf{phastcons$>$CDS$>$genes}  &              full &          YRI &               \textbf{67.27} &             67.36 &             &                                 2.0333 &                                \\
+               CADD 6\%  &            sparse &          YRI &                        66.66 &             66.81 &             &                                 1.6565 &                                \\
+               CADD 6\%  &              full &          YRI &                        66.50 &             66.68 &             &                                 1.5556 &                                \\
+CDS$>$genes$>$phastcons  &              full &          YRI &                        63.61 &             64.28 &             &                                 3.3493 &                                \\
+\textbf{phastcons$>$CDS$>$genes}  &              full &          CEU &               \textbf{61.04} &             62.25 &             &                                 1.9128 &                                \\
+               CADD 6\%  &            sparse &          CEU &                        60.37 &             61.80 &             &                                 1.6797 &                                \\
+               CADD 6\%  &              full &          CEU &                        60.24 &             61.77 &             &                                 1.6347 &                                \\
+               \textbf{phastcons$>$CDS$>$genes}  &     full &          CHB &                        \textbf{58.76} &             60.45 &             &                                 1.8691 &                                \\
+               CADD 6\%  &            sparse &          CHB &                        58.67 &             60.49 &             &                                 1.7454 &                                \\
+               CADD 6\%  &              full &          CHB &                        58.48 &             60.40 &             &                                 1.5981 &                                \\
+CDS$>$genes$>$phastcons  &              full &          CEU &                        57.24 &             59.11 &             &                                 3.0461 &                                \\
+CDS$>$genes$>$phastcons  &              full &          CHB &                        55.63 &             57.88 &             &                                 3.1723 &                                \\
 \end{tabular}
@@
 (Supplementary Figure \ref{suppfig:r2_upper}).
+
+\begin{figure}[htbp] \centering
+    \includegraphics[width=\textwidth]{figures/supplementary/suppfigure_drift_R2.pdf}
+    \caption{(A) the theoretic $R_\text{drift}^2$ using different plugin
+        estimators for $B_i$ under different models. This shows that among the
+        best-fitting models, the predicted $R_\text{drift}^2$ varies little;
+        most variation is among populations due to differing $\pi_0$. (B) The 
+        $R_\text{drift}^2$ across different populations, for varying ratio of recombination
+        to mutation rates (x-axis).}
+  \label{suppfig:drift_r2}
+\end{figure}
 
@@
 \begin{figure}[htbp] \centering
-    \includegraphics[width=\textwidth]{figures/supplementary/suppfigure_drift_R2.pdf}
-    \caption{(A) the theoretic $R_\text{drift}^2$ using different plugin
-        estimators for $B_i$ under different models. This shows that among the
-        best-fitting models, the predicted $R_\text{drift}^2$ varies little;
-        most variation is among populations due to differing $\pi_0$. (B) The 
-        $R_\text{drift}^2$ across different populations, for varying ratio of recombination
-        to mutation rates (x-axis).}
-  \label{suppfig:drift_r2}
+    \includegraphics[width=0.5\textwidth]{figures/supplementary/subfigure_R2_scale.pdf}
+    \caption{XXX}
+  \label{suppfig:coal_r2_scale}
 \end{figure}

### 2023-07-05 manuscript/manuscript.tex
ms changes

@@
 However, across all of our annotation models, new deleterious mutations in
-putatively conserved features are consistently estimated to have strongly
-deleterious effects (Figure \ref{fig:figure-3}A). This includes features such
-as coding sequences, PhastCons regions, or the top 6\% of basepairs with the
-most pathogenic CADD scores. Thus, even with a model that can better
-distinguish weak versus strong selection, our DFE estimates suggest
+likely conserved features are consistently estimated to have strongly
+deleterious effects (Figure \ref{fig:figure-3}A) consistent with previous work
+\parencite{McVicker2009-ax,Murphy2022-sj}. This includes features such as
+coding sequences, PhastCons regions, or the top 6\% of basepairs with the most
+pathogenic CADD scores. However, we find that there is considerable uncertainty
+in how strong selection is against the most deleterious mutations. For example,
+in the CADD 6\% full-track model, new mutations are estimated to have a
+selection coefficients of $s = 10^{-1}$ with 42\% chance, $s=10^{-2}$ with
+7.8\% chance, and $s=10^{3}$ with 41\% chance. These estimates suggest a
+bimodal DFE (Figure \ref{fig:figure-3}A, second row, left column), and are
+mirrored across the other CADD models.
+
+Given that previous methods have typically not estimated the DFE up to
+selection coefficients of $s=10^{-1}$, we ran additional models using a DFE
+constrained to $s=10^{-2}$. Each of these constrained-DFE models across all
+populations fit the data less well than models run with the grid up to
+$s=10^{-1}$, explaining 1\% less of the leave-one-chromosome-out variation on
+average. As expected, the constrained-DFE models put more mass on $s=10^{-2}$
+(see Supplementary Table \ref{supp:tbl-r2}).
+
+Our method's estimates for the DFE rely on a
+pre-specified logarithmic grid; we use a fairly fine-scale grid from $10^{-8}$
+to $10^{-1}$.
+
+Across the CADD models, we note a discontinuity in DFE estimates,
+where the largest mass is placed on the strongest selection coefficient ($s =
+10^{-1}$), very little on the next coefficient ($s=10^{-2}$), and mo($\approx
+41\%-67\%$)
+
+These DFE estimates also suggest that our method
+can accurately estimate when regions are under weak or no constraint.
+
+Across our full-coverage models, the estimated DFE for the ``other" feature
+containing the rest of the genome is estimated to nearly 100\% neutral
+($s=10^{-8}$). In some cases (e.g. the CADD 6\% and 8\% models) the ``other"
+estimates a small fraction ($\approx 3\%-5\%$) of mutations in the ``other"
+region have a strongly deleterious impact ($s=10^{-1}$). This is likely due to
+non-identifiability between $\pi_0$ and a small chance of deleterious mutations
+occurring on the unannotated remainder of the genome.
+
+Our DFE estimates
+
+Thus, even though our model can better distinguish
+weak versus strong selection, our DFE estimates suggest
 

### 2023-07-05 manuscript/supplementary.tex
ms changes

@@
 \RequirePackage{wasysym}
+\RequirePackage{listings}
 \RequirePackage{authblk}
@@
 \hline
-\emph{smaller DFE grid} & & & & & & \\
+\emph{constrained DFE grid} & & & & & & \\
 \hline
@@
 \begin{figure}[htbp]
-    \label{suppfig:dfe-pop-smallgrid}
+    \label{suppfig:dfe-pop-smallgrid-feature}
+  \centering
+  \includegraphics[width=0.6\textwidth]{figures/supplementary/feature_priority_pop_dfe_smallselgrid.pdf}
+  \caption{The DFE for the Feature Priority model, fit with a smaller DFE grid from 
+  $\mu \in \{10^{-8}, 10^{-7}, \ldots, 10^{-2}\}$.}
+\end{figure}
+
+\begin{figure}[htbp]
+    \label{suppfig:dfe-pop-smallgrid-phastcons}
   \centering
@@
 
-%Our $B$ is,
-%
-%\begin{align}
-%  B = \frac{1}{2N} \sum_{t=0}^\infty \prod_{i=0}^t \left(1 - \frac{1}{2N_e(i)}\right).
-%\end{align}
-%
-%This sum takes many terms to converge, which makes calculating the $B$ between
-%each focal neutral site and segment intractably slow (even when the equilibrium
-%$\widetilde{V}$ for the segment pre-calculated). We approximate this in the
-%following way for faster numerical calculations. Let $g(t) = \prod_{i=0}^t (1 -
-%\nicefrac{1}{2N_e(i)})$. Then, taking the first-order series expansion around
-%the logarithm of the terms and approximating the sum with an integral,
-%
-%\begin{align}
-%  g(t) &\approx \exp\left(- \frac{1}{2N} \int_0^t \exp\left(\frac{V}{2} Q(x)^2 \right) dx \right) 
-%       %&\approx \exp\left(- \int_0^t \frac{\exp\left(\frac{V}{2} Q(i)^2\right)}{2N}  dx \right)
-%\end{align}
-%
-%$B$ is the case of $\lim b \to \infty$ for $B(b)$. Approximating the sum in
-%$B(b)$ with an integral,
-%
-%\begin{align}
-%  B(b) &= \int_0^b \exp\left(- \frac{1}{2N} \int_0^t \exp\left(\frac{V}{2} Q(x)^2 \right) dx \right) dt \\
-%       %&\approx \exp\left(- \int_0^t \frac{\exp\left(\frac{V}{2} Q(i)^2\right)}{2N}  dx \right)
-%\end{align}
-%
-%Now, let $z = \log(t)$, then
-%
-%\begin{align}
-%  B(w) &= \int_0^w \exp\left(- \frac{1}{2N} \int_0^{\log(z)} \exp\left(\frac{V}{2} Q(x)^2 \right) dx \right) \exp(t) dz \\
-%       %&\approx \exp\left(- \int_0^t \frac{\exp\left(\frac{V}{2} Q(i)^2\right)}{2N}  dx \right)
-%\end{align}
-%
-%
+\section{All Model Fits}
+\label{supp:fits}
+
+%\printbibliography
+%\end{document}
+
+\subsection{Phastcons Priority, Full Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 280
+standard error method: moving block jackknife
+negative log-likelihood: 197825873893.4953
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.0011279 (0.00113, 0.00113)
+pi  = 0.000801495
+mut. rate = 2.35e-08  (1.98e-08, 2.72e-08)
+Ne = 28,197 (if mu=1e-8), Ne = 14,098 (if mu=2e-8)
+R2 = 63.0106% (in-sample)  61.7759% (out-sample)
+W = 
+        CDS                      gene                     other                    phastcons
+------  -----------------------  -----------------------  -----------------------  -----------------------
+1e-08   0.000 (-0.0410, 0.0410)  0.901 (0.6521, 1.1503)   0.984 (0.9739, 0.9938)   0.034 (-2.8217, 2.8898)
+1e-07   0.000 (-0.1314, 0.1314)  0.026 (-0.2223, 0.2745)  0.000 (-0.0096, 0.0096)  0.376 (-2.4935, 3.2456)
+1e-06   0.000 (-0.0088, 0.0088)  0.000 (-0.0018, 0.0027)  0.000 (-0.0001, 0.0001)  0.001 (-0.0308, 0.0318)
+1e-05   0.000 (-0.0002, 0.0002)  0.019 (0.0133, 0.0249)   0.000 (-0.0000, 0.0000)  0.000 (-0.0003, 0.0003)
+0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
+0.001   0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.472 (0.4050, 0.5388)
+0.01    0.000 (-0.0001, 0.0001)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.046 (0.0351, 0.0571)
+0.1     1.000 (0.8620, 1.1380)   0.053 (0.0454, 0.0608)   0.016 (0.0138, 0.0185)   0.071 (0.0571, 0.0858)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 280
+standard error method: moving block jackknife
+negative log-likelihood: 61821783504.67356
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00106467 (0.00106, 0.00107)
+pi  = 0.000752601
+mut. rate = 2.557e-08  (-8.06e-10, 5.2e-08)
+Ne = 26,616 (if mu=1e-8), Ne = 13,308 (if mu=2e-8)
+R2 = 61.0847% (in-sample)  59.2923% (out-sample)
+W = 
+        CDS                      gene                     other                    phastcons
+------  -----------------------  -----------------------  -----------------------  -----------------------
+1e-08   0.000 (-0.5839, 0.5839)  0.931 (0.6381, 1.2238)   0.985 (0.9661, 1.0044)   0.478 (-2.5565, 3.5128)
+1e-07   0.000 (-0.2506, 0.2506)  0.005 (-0.2858, 0.2949)  0.000 (-0.0165, 0.0165)  0.002 (-3.0192, 3.0225)
+1e-06   0.000 (-0.0119, 0.0119)  0.000 (-0.0043, 0.0045)  0.000 (-0.0002, 0.0002)  0.000 (-0.0172, 0.0177)
+1e-05   0.000 (-0.0001, 0.0001)  0.013 (-0.0013, 0.0268)  0.000 (-0.0000, 0.0000)  0.000 (-0.0002, 0.0002)
+0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
+0.001   0.000 (-0.0000, 0.0000)  0.003 (0.0005, 0.0049)   0.000 (-0.0000, 0.0000)  0.430 (0.1533, 0.7071)
+0.01    0.000 (-0.0001, 0.0001)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.090 (0.0316, 0.1477)
+0.1     1.000 (0.3574, 1.6426)   0.049 (0.0174, 0.0804)   0.015 (0.0053, 0.0243)   0.000 (-0.0005, 0.0005)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 280
+standard error method: moving block jackknife
+negative log-likelihood: 250656989872.19766
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00147277 (0.00147, 0.00147)
+pi  = 0.00106447
+mut. rate = 1.708e-08  (-4.08e-08, 7.49e-08)
+Ne = 36,819 (if mu=1e-8), Ne = 18,409 (if mu=2e-8)
+R2 = 68.1177% (in-sample)  68.1947% (out-sample)
+W = 
+        CDS                      gene                     other                    phastcons
+------  -----------------------  -----------------------  -----------------------  -----------------------
+1e-08   1.000 (-3.4930, 5.4930)  0.902 (0.6716, 1.1325)   0.979 (0.9416, 1.0156)   0.000 (-1.2400, 1.2407)
+1e-07   0.000 (-4.4936, 4.4936)  0.000 (-0.1820, 0.1820)  0.000 (-0.0194, 0.0194)  0.004 (-0.5459, 0.5537)
+1e-06   0.000 (-0.0107, 0.0107)  0.000 (-0.0024, 0.0024)  0.000 (-0.0003, 0.0003)  0.001 (-0.0353, 0.0365)
+1e-05   0.000 (-0.0001, 0.0001)  0.042 (-0.0177, 0.1024)  0.000 (-0.0000, 0.0000)  0.000 (-0.0005, 0.0006)
+0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0001, 0.0001)
+0.001   0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.523 (-0.2146, 1.2611)
+0.01    0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.212 (-0.0866, 0.5104)
+0.1     0.000 (-0.0006, 0.0006)  0.056 (-0.0228, 0.1339)  0.021 (-0.0088, 0.0516)  0.260 (-0.1070, 0.6270)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{Phastcons Priority, Sparse Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 197826612684.87143
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00112308
+pi  = 0.000801495
+mut. rate = 2.019e-08 
+Ne = 28,076 (if mu=1e-8), Ne = 14,038 (if mu=2e-8)
+R2 = 63.0152% (in-sample)  61.9834% (out-sample)
+W = 
+          CDS    gene    phastcons
+------  -----  ------  -----------
+1e-08   0       0.913        0.227
+1e-07   0       0            0.004
+1e-06   0.004   0            0
+1e-05   0       0.019        0
+0.0001  0       0            0
+0.001   0       0            0.546
+0.01    0       0            0.061
+0.1     0.996   0.069        0.161
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 61822016128.79776
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00106069
+pi  = 0.000752601
+mut. rate = 5.887e-08 
+Ne = 26,517 (if mu=1e-8), Ne = 13,258 (if mu=2e-8)
+R2 = 61.0804% (in-sample)  59.4419% (out-sample)
+W = 
+          CDS    gene    phastcons
+------  -----  ------  -----------
+1e-08   0.617   0.968        0.744
+1e-07   0.01    0            0.02
+1e-06   0       0            0
+1e-05   0       0.005        0
+0.0001  0       0            0
+0.001   0       0.001        0.185
+0.01    0       0            0.051
+0.1     0.373   0.026        0
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 250657908564.79294
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00146801
+pi  = 0.00106447
+mut. rate = 1.864e-08 
+Ne = 36,700 (if mu=1e-8), Ne = 18,350 (if mu=2e-8)
+R2 = 68.1582% (in-sample)  68.4462% (out-sample)
+W = 
+          CDS    gene    phastcons
+------  -----  ------  -----------
+1e-08       1   0.907        0
+1e-07       0   0.001        0
+1e-06       0   0            0
+1e-05       0   0.035        0
+0.0001      0   0            0
+0.001       0   0            0.477
+0.01        0   0            0.199
+0.1         0   0.057        0.323
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{Feature Priority, Full Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 457
+standard error method: moving block jackknife
+negative log-likelihood: 197848627167.32043
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00115698 (0.00116, 0.00116)
+pi  = 0.000801495
+mut. rate = 4.1e-08  (2.2e-08, 6e-08)
+Ne = 28,924 (if mu=1e-8), Ne = 14,462 (if mu=2e-8)
+R2 = 60.3026% (in-sample)  58.5808% (out-sample)
+W = 
+        CDS                      gene                     other                    phastcons
+------  -----------------------  -----------------------  -----------------------  -----------------------
+1e-08   0.107 (-1.0244, 1.2375)  0.946 (0.7974, 1.0947)   1.000 (0.9991, 1.0009)   0.000 (-0.2751, 0.2757)
+1e-07   0.000 (-1.0319, 1.0319)  0.001 (-0.1448, 0.1458)  0.000 (-0.0009, 0.0009)  0.000 (-0.2482, 0.2482)
+1e-06   0.000 (-0.0087, 0.0087)  0.000 (-0.0013, 0.0021)  0.000 (-0.0000, 0.0000)  0.000 (-0.0150, 0.0150)
+1e-05   0.000 (-0.0001, 0.0001)  0.023 (0.0088, 0.0368)   0.000 (-0.0000, 0.0000)  0.000 (-0.0003, 0.0003)
+0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
+0.001   0.144 (0.0831, 0.2059)   0.014 (0.0080, 0.0196)   0.000 (-0.0000, 0.0000)  0.229 (0.1327, 0.3256)
+0.01    0.000 (-0.0012, 0.0012)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.086 (0.0462, 0.1253)
+0.1     0.749 (0.4358, 1.0622)   0.016 (0.0095, 0.0233)   0.000 (-0.0000, 0.0000)  0.685 (0.3987, 0.9709)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 454
+standard error method: moving block jackknife
+negative log-likelihood: 61827828856.16418
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00108991 (0.00109, 0.00109)
+pi  = 0.000752601
+mut. rate = 3.755e-08  (1.55e-08, 5.96e-08)
+Ne = 27,247 (if mu=1e-8), Ne = 13,623 (if mu=2e-8)
+R2 = 58.7328% (in-sample)  56.4805% (out-sample)
+W = 
+        CDS                      gene                     other                    phastcons
+------  -----------------------  -----------------------  -----------------------  -----------------------
+1e-08   0.088 (-0.9233, 1.0989)  0.940 (0.7665, 1.1139)   1.000 (0.9988, 1.0012)   0.006 (-0.3608, 0.3730)
+1e-07   0.010 (-0.8568, 0.8763)  0.000 (-0.1685, 0.1691)  0.000 (-0.0011, 0.0011)  0.003 (-0.2610, 0.2674)
+1e-06   0.000 (-0.0075, 0.0078)  0.000 (-0.0021, 0.0021)  0.000 (-0.0000, 0.0000)  0.000 (-0.0145, 0.0151)
+1e-05   0.000 (-0.0001, 0.0001)  0.023 (0.0037, 0.0422)   0.000 (-0.0000, 0.0000)  0.000 (-0.0003, 0.0003)
+0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
+0.001   0.156 (0.0732, 0.2384)   0.017 (0.0081, 0.0260)   0.000 (-0.0000, 0.0000)  0.244 (0.1185, 0.3696)
+0.01    0.000 (-0.0139, 0.0140)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.254 (0.1215, 0.3871)
+0.1     0.746 (0.3590, 1.1338)   0.020 (0.0094, 0.0298)   0.000 (-0.0000, 0.0000)  0.492 (0.2392, 0.7448)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 280
+standard error method: moving block jackknife
+negative log-likelihood: 250679665159.79324
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00151818 (0.00152, 0.00152)
+pi  = 0.00106447
+mut. rate = 4.195e-08  (2.76e-08, 5.63e-08)
+Ne = 37,954 (if mu=1e-8), Ne = 18,977 (if mu=2e-8)
+R2 = 65.5097% (in-sample)  64.9049% (out-sample)
+W = 
+        CDS                      gene                     other                    phastcons
+------  -----------------------  -----------------------  -----------------------  -----------------------
+1e-08   0.364 (-0.4599, 1.1874)  0.931 (0.8360, 1.0260)   1.000 (0.9993, 1.0007)   0.003 (-0.1973, 0.2040)
+1e-07   0.001 (-0.7930, 0.7942)  0.000 (-0.0901, 0.0901)  0.000 (-0.0007, 0.0007)  0.000 (-0.1669, 0.1675)
+1e-06   0.000 (-0.0038, 0.0040)  0.000 (-0.0009, 0.0009)  0.000 (-0.0000, 0.0000)  0.000 (-0.0106, 0.0109)
+1e-05   0.000 (-0.0001, 0.0001)  0.039 (0.0240, 0.0532)   0.000 (-0.0000, 0.0000)  0.000 (-0.0002, 0.0002)
+0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
+0.001   0.091 (0.0617, 0.1212)   0.011 (0.0077, 0.0150)   0.000 (-0.0000, 0.0000)  0.151 (0.1030, 0.1986)
+0.01    0.000 (-0.0001, 0.0001)  0.000 (-0.0001, 0.0001)  0.000 (-0.0000, 0.0000)  0.328 (0.2249, 0.4316)
+0.1     0.544 (0.3725, 0.7157)   0.019 (0.0130, 0.0250)   0.000 (-0.0000, 0.0000)  0.517 (0.3548, 0.6796)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{Feature Priority, Sparse Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 27
+standard error method: moving block jackknife
+negative log-likelihood: 197848626646.91736
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00115698 (0.00116, 0.00116)
+pi  = 0.000801495
+mut. rate = 4.104e-08  (3.87e-08, 4.34e-08)
+Ne = 28,924 (if mu=1e-8), Ne = 14,462 (if mu=2e-8)
+R2 = 60.3027% (in-sample)  58.5823% (out-sample)
+W = 
+        CDS                      gene                     phastcons
+------  -----------------------  -----------------------  -----------------------
+1e-08   0.107 (0.0490, 0.1649)   0.947 (0.9402, 0.9539)   0.001 (-0.0478, 0.0491)
+1e-07   0.000 (-0.0273, 0.0279)  0.000 (-0.0052, 0.0052)  0.000 (-0.0073, 0.0075)
+1e-06   0.000 (-0.0003, 0.0006)  0.000 (-0.0001, 0.0001)  0.001 (-0.0008, 0.0019)
+1e-05   0.000 (-0.0000, 0.0000)  0.023 (0.0187, 0.0268)   0.000 (-0.0000, 0.0000)
+0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
+0.001   0.144 (0.1365, 0.1522)   0.014 (0.0130, 0.0146)   0.229 (0.2150, 0.2428)
+0.01    0.000 (-0.0001, 0.0001)  0.000 (-0.0000, 0.0000)  0.086 (0.0733, 0.0981)
+0.1     0.748 (0.7078, 0.7887)   0.016 (0.0155, 0.0173)   0.684 (0.6463, 0.7219)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 26
+standard error method: moving block jackknife
+negative log-likelihood: 61827828699.43107
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00108991 (0.00109, 0.00109)
+pi  = 0.000752601
+mut. rate = 3.719e-08  (3.2e-08, 4.24e-08)
+Ne = 27,247 (if mu=1e-8), Ne = 13,623 (if mu=2e-8)
+R2 = 58.7329% (in-sample)  56.4805% (out-sample)
+W = 
+        CDS                      gene                     phastcons
+------  -----------------------  -----------------------  -----------------------
+1e-08   0.089 (-0.0565, 0.2340)  0.938 (0.9196, 0.9573)   0.000 (-0.1223, 0.1223)
+1e-07   0.000 (-0.0861, 0.0861)  0.001 (-0.0154, 0.0180)  0.000 (-0.0145, 0.0146)
+1e-06   0.000 (-0.0005, 0.0010)  0.000 (-0.0001, 0.0002)  0.000 (-0.0009, 0.0009)
+1e-05   0.000 (-0.0000, 0.0000)  0.023 (0.0203, 0.0260)   0.000 (-0.0000, 0.0000)
+0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
+0.001   0.157 (0.1380, 0.1766)   0.017 (0.0151, 0.0194)   0.246 (0.2161, 0.2767)
+0.01    0.000 (-0.0004, 0.0006)  0.000 (-0.0000, 0.0000)  0.257 (0.2252, 0.2883)
+0.1     0.754 (0.6611, 0.8461)   0.020 (0.0173, 0.0222)   0.497 (0.4358, 0.5578)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 17
+standard error method: moving block jackknife
+negative log-likelihood: 250679664591.4161
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00151817 (0.00152, 0.00152)
+pi  = 0.00106447
+mut. rate = 4.179e-08  (3.84e-08, 4.52e-08)
+Ne = 37,954 (if mu=1e-8), Ne = 18,977 (if mu=2e-8)
+R2 = 65.5099% (in-sample)  64.9049% (out-sample)
+W = 
+        CDS                      gene                     phastcons
+------  -----------------------  -----------------------  -----------------------
+1e-08   0.360 (0.3063, 0.4146)   0.931 (0.9239, 0.9375)   0.000 (-0.0647, 0.0647)
+1e-07   0.002 (-0.0234, 0.0265)  0.000 (-0.0051, 0.0052)  0.000 (-0.0097, 0.0097)
+1e-06   0.000 (-0.0003, 0.0003)  0.000 (-0.0000, 0.0001)  0.000 (-0.0009, 0.0009)
+1e-05   0.000 (-0.0000, 0.0000)  0.039 (0.0359, 0.0417)   0.000 (-0.0000, 0.0000)
+0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
+0.001   0.092 (0.0849, 0.0987)   0.011 (0.0105, 0.0122)   0.151 (0.1401, 0.1627)
+0.01    0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.329 (0.3048, 0.3540)
+0.1     0.546 (0.5054, 0.5870)   0.019 (0.0176, 0.0205)   0.519 (0.4804, 0.5580)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{CADD 6\%, Full Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 280
+standard error method: moving block jackknife
+negative log-likelihood: 197826953189.02365
+number of successful starts: 20000 (100.0% total)
+pi_0 = 0.00109999 (0.00110, 0.00110)
+pi  = 0.000801495
+mut. rate = 1.537e-08  (8e-09, 2.27e-08)
+Ne = 27,499 (if mu=1e-8), Ne = 13,749 (if mu=2e-8)
+R2 = 62.9394% (in-sample)  61.7433% (out-sample)
+W = 
+        cadd6                    other
+------  -----------------------  -----------------------
+1e-08   0.000 (-0.2792, 0.2795)  0.965 (0.9473, 0.9832)
+1e-07   0.000 (-0.2355, 0.2355)  0.000 (-0.0109, 0.0109)
+1e-06   0.000 (-0.0371, 0.0371)  0.000 (-0.0001, 0.0001)
+1e-05   0.000 (-0.0011, 0.0013)  0.000 (-0.0000, 0.0000)
+0.0001  0.026 (0.0129, 0.0393)   0.000 (-0.0000, 0.0000)
+0.001   0.555 (0.3257, 0.7834)   0.000 (-0.0000, 0.0000)
+0.01    0.000 (-0.0001, 0.0001)  0.000 (-0.0000, 0.0000)
+0.1     0.419 (0.2460, 0.5921)   0.035 (0.0204, 0.0491)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 280
+standard error method: moving block jackknife
+negative log-likelihood: 61821323176.26884
+number of successful starts: 20000 (100.0% total)
+pi_0 = 0.00103624 (0.00104, 0.00104)
+pi  = 0.000752601
+mut. rate = 1.508e-08  (-2.43e-09, 3.26e-08)
+Ne = 25,905 (if mu=1e-8), Ne = 12,952 (if mu=2e-8)
+R2 = 61.3115% (in-sample)  59.6228% (out-sample)
+W = 
+        cadd6                    other
+------  -----------------------  -----------------------
+1e-08   0.000 (-0.6278, 0.6278)  0.961 (0.9187, 1.0036)
+1e-07   0.007 (-0.4928, 0.5076)  0.000 (-0.0227, 0.0227)
+1e-06   0.009 (-0.0393, 0.0570)  0.000 (-0.0002, 0.0002)
+1e-05   0.000 (-0.0014, 0.0015)  0.000 (-0.0000, 0.0000)
+0.0001  0.004 (-0.0082, 0.0171)  0.000 (-0.0000, 0.0000)
+0.001   0.670 (0.0898, 1.2509)   0.000 (-0.0000, 0.0000)
+0.01    0.000 (-0.0001, 0.0001)  0.000 (-0.0000, 0.0000)
+0.1     0.309 (0.0412, 0.5766)   0.039 (0.0051, 0.0726)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 280
+standard error method: moving block jackknife
+negative log-likelihood: 250659830283.90668
+number of successful starts: 20000 (100.0% total)
+pi_0 = 0.00145893 (0.00146, 0.00146)
+pi  = 0.00106447
+mut. rate = 1.461e-08  (8.61e-09, 2.06e-08)
+Ne = 36,473 (if mu=1e-8), Ne = 18,236 (if mu=2e-8)
+R2 = 68.0839% (in-sample)  68.0751% (out-sample)
+W = 
+        cadd6                    other
+------  -----------------------  -----------------------
+1e-08   0.000 (-0.2318, 0.2327)  0.958 (0.9399, 0.9768)
+1e-07   0.023 (-0.2328, 0.2797)  0.000 (-0.0083, 0.0089)
+1e-06   0.001 (-0.0153, 0.0176)  0.000 (-0.0001, 0.0001)
+1e-05   0.000 (-0.0003, 0.0004)  0.000 (-0.0000, 0.0000)
+0.0001  0.065 (0.0378, 0.0921)   0.000 (-0.0000, 0.0000)
+0.001   0.407 (0.2491, 0.5649)   0.000 (-0.0000, 0.0000)
+0.01    0.078 (0.0439, 0.1126)   0.000 (-0.0000, 0.0000)
+0.1     0.425 (0.2604, 0.5891)   0.041 (0.0253, 0.0573)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{CADD 6\%, Sparse Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 197829116669.21103
+number of successful starts: 20000 (100.0% total)
+pi_0 = 0.00107623
+pi  = 0.000801495
+mut. rate = 2.121e-08 
+Ne = 26,905 (if mu=1e-8), Ne = 13,452 (if mu=2e-8)
+R2 = 62.8555% (in-sample)  61.6574% (out-sample)
+W = 
+          cadd6
+------  -------
+1e-08     0
+1e-07     0
+1e-06     0
+1e-05     0
+0.0001    0.014
+0.001     0.408
+0.01      0
+0.1       0.578
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 61822132892.58568
+number of successful starts: 20000 (100.0% total)
+pi_0 = 0.00101201
+pi  = 0.000752601
+mut. rate = 2.124e-08 
+Ne = 25,300 (if mu=1e-8), Ne = 12,650 (if mu=2e-8)
+R2 = 61.2064% (in-sample)  59.5382% (out-sample)
+W = 
+          cadd6
+------  -------
+1e-08     0
+1e-07     0
+1e-06     0
+1e-05     0
+0.0001    0
+0.001     0.477
+0.01      0
+0.1       0.523
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 250663447300.65836
+number of successful starts: 20000 (100.0% total)
+pi_0 = 0.00142415
+pi  = 0.00106447
+mut. rate = 2.091e-08 
+Ne = 35,603 (if mu=1e-8), Ne = 17,801 (if mu=2e-8)
+R2 = 68.028% (in-sample)  68.0414% (out-sample)
+W = 
+          cadd6
+------  -------
+1e-08     0
+1e-07     0
+1e-06     0
+1e-05     0
+0.0001    0.038
+0.001     0.302
+0.01      0.033
+0.1       0.627
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{CADD 6\%, Full Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 140
+standard error method: moving block jackknife
+negative log-likelihood: 197834651568.3971
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00109606 (0.00110, 0.00110)
+pi  = 0.000801495
+mut. rate = 1.157e-08  (9.01e-09, 1.41e-08)
+Ne = 27,401 (if mu=1e-8), Ne = 13,700 (if mu=2e-8)
+R2 = 62.1245% (in-sample)  60.8009% (out-sample)
+W = 
+        cadd8                    other
+------  -----------------------  -----------------------
+1e-08   0.000 (-0.0835, 0.0835)  0.952 (0.9415, 0.9624)
+1e-07   0.004 (-0.1451, 0.1523)  0.000 (-0.0043, 0.0043)
+1e-06   0.000 (-0.0145, 0.0150)  0.000 (-0.0001, 0.0001)
+1e-05   0.000 (-0.0005, 0.0005)  0.000 (-0.0000, 0.0000)
+0.0001  0.052 (0.0393, 0.0656)   0.000 (-0.0000, 0.0000)
+0.001   0.537 (0.4316, 0.6417)   0.000 (-0.0000, 0.0000)
+0.01    0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
+0.1     0.407 (0.3276, 0.4867)   0.048 (0.0386, 0.0575)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 140
+standard error method: moving block jackknife
+negative log-likelihood: 61823378586.32404
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00103245 (0.00103, 0.00103)
+pi  = 0.000752601
+mut. rate = 1.113e-08  (4.22e-09, 1.8e-08)
+Ne = 25,811 (if mu=1e-8), Ne = 12,905 (if mu=2e-8)
+R2 = 60.6265% (in-sample)  58.8768% (out-sample)
+W = 
+        cadd8                    other
+------  -----------------------  -----------------------
+1e-08   0.000 (-0.3716, 0.3716)  0.944 (0.9141, 0.9736)
+1e-07   0.000 (-0.2895, 0.2903)  0.001 (-0.0075, 0.0094)
+1e-06   0.000 (-0.0376, 0.0378)  0.000 (-0.0001, 0.0001)
+1e-05   0.000 (-0.0005, 0.0005)  0.000 (-0.0000, 0.0000)
+0.0001  0.026 (0.0069, 0.0447)   0.000 (-0.0000, 0.0000)
+0.001   0.690 (0.3451, 1.0351)   0.000 (-0.0000, 0.0000)
+0.01    0.000 (-0.0001, 0.0001)  0.000 (-0.0000, 0.0000)
+0.1     0.284 (0.1418, 0.4254)   0.055 (0.0275, 0.0829)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 140
+standard error method: moving block jackknife
+negative log-likelihood: 250669695762.61414
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00145867 (0.00146, 0.00146)
+pi  = 0.00106447
+mut. rate = 1.013e-08  (5.65e-09, 1.46e-08)
+Ne = 36,466 (if mu=1e-8), Ne = 18,233 (if mu=2e-8)
+R2 = 67.1076% (in-sample)  66.9835% (out-sample)
+W = 
+        cadd8                    other
+------  -----------------------  -----------------------
+1e-08   0.000 (-0.3206, 0.3212)  0.932 (0.9054, 0.9583)
+1e-07   0.002 (-0.1725, 0.1762)  0.000 (-0.0062, 0.0062)
+1e-06   0.000 (-0.0130, 0.0130)  0.000 (-0.0001, 0.0001)
+1e-05   0.000 (-0.0004, 0.0004)  0.000 (-0.0000, 0.0000)
+0.0001  0.103 (0.0627, 0.1425)   0.000 (-0.0000, 0.0000)
+0.001   0.438 (0.2744, 0.6008)   0.000 (-0.0000, 0.0000)
+0.01    0.000 (-0.0003, 0.0003)  0.000 (-0.0000, 0.0000)
+0.1     0.458 (0.2870, 0.6282)   0.068 (0.0427, 0.0936)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{CADD 6\%, Sparse Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 140
+standard error method: moving block jackknife
+negative log-likelihood: 197836763801.90414
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00107057 (0.00107, 0.00107)
+pi  = 0.000801495
+mut. rate = 1.603e-08  (1.57e-08, 1.63e-08)
+Ne = 26,764 (if mu=1e-8), Ne = 13,382 (if mu=2e-8)
+R2 = 62.061% (in-sample)  60.5937% (out-sample)
+W = 
+        cadd8
+------  -----------------------
+1e-08   0.000 (-0.0171, 0.0171)
+1e-07   0.000 (-0.0007, 0.0007)
+1e-06   0.000 (-0.0000, 0.0000)
+1e-05   0.000 (-0.0000, 0.0000)
+0.0001  0.032 (0.0268, 0.0372)
+0.001   0.400 (0.3894, 0.4110)
+0.01    0.000 (-0.0000, 0.0000)
+0.1     0.568 (0.5575, 0.5781)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 140
+standard error method: moving block jackknife
+negative log-likelihood: 61824178668.93152
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00100597 (0.00101, 0.00101)
+pi  = 0.000752601
+mut. rate = 1.61e-08  (-4.15e-08, 7.37e-08)
+Ne = 25,149 (if mu=1e-8), Ne = 12,574 (if mu=2e-8)
+R2 = 60.5692% (in-sample)  58.6497% (out-sample)
+W = 
+        cadd8
+------  -----------------------
+1e-08   0.000 (-0.7779, 0.7779)
+1e-07   0.000 (-0.0032, 0.0033)
+1e-06   0.000 (-0.0008, 0.0008)
+1e-05   0.000 (-0.0000, 0.0000)
+0.0001  0.011 (-0.0013, 0.0236)
+0.001   0.492 (0.1090, 0.8748)
+0.01    0.000 (-0.0000, 0.0000)
+0.1     0.497 (0.1101, 0.8838)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 140
+standard error method: moving block jackknife
+negative log-likelihood: 250673972070.93787
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00141668 (0.00142, 0.00142)
+pi  = 0.00106447
+mut. rate = 1.569e-08  (1.49e-08, 1.64e-08)
+Ne = 35,416 (if mu=1e-8), Ne = 17,708 (if mu=2e-8)
+R2 = 67.0084% (in-sample)  66.6718% (out-sample)
+W = 
+        cadd8
+------  -----------------------
+1e-08   0.000 (-0.0445, 0.0445)
+1e-07   0.000 (-0.0136, 0.0136)
+1e-06   0.000 (-0.0002, 0.0002)
+1e-05   0.000 (-0.0000, 0.0000)
+0.0001  0.059 (0.0526, 0.0655)
+0.001   0.298 (0.2808, 0.3142)
+0.01    0.000 (-0.0000, 0.0000)
+0.1     0.643 (0.6132, 0.6737)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{Phastcons Priority (small DFE grid), Full Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 197835157685.628
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00105552
+pi  = 0.000801495
+mut. rate = 1.913e-08 
+Ne = 26,387 (if mu=1e-8), Ne = 13,193 (if mu=2e-8)
+R2 = 62.2548% (in-sample)  61.0392% (out-sample)
+W = 
+          CDS    gene    other    phastcons
+------  -----  ------  -------  -----------
+1e-08       1   0.947        1        0.001
+1e-07       0   0            0        0
+1e-06       0   0            0        0
+1e-05       0   0.03         0        0
+0.0001      0   0            0        0
+0.001       0   0            0        0.413
+0.01        0   0.023        0        0.586
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 61824235925.00332
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00100129
+pi  = 0.000752601
+mut. rate = 1.869e-08 
+Ne = 25,032 (if mu=1e-8), Ne = 12,516 (if mu=2e-8)
+R2 = 60.4486% (in-sample)  58.7636% (out-sample)
+W = 
+          CDS    gene    other    phastcons
+------  -----  ------  -------  -----------
+1e-08       1   0.926        1        0
+1e-07       0   0.013        0        0
+1e-06       0   0            0        0
+1e-05       0   0.031        0        0
+0.0001      0   0            0        0
+0.001       0   0            0        0.457
+0.01        0   0.03         0        0.543
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 250667315910.91177
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00139482
+pi  = 0.00106447
+mut. rate = 2.033e-08 
+Ne = 34,870 (if mu=1e-8), Ne = 17,435 (if mu=2e-8)
+R2 = 67.3564% (in-sample)  67.274% (out-sample)
+W = 
+          CDS    gene    other    phastcons
+------  -----  ------  -------  -----------
+1e-08       1   0.956        1        0.005
+1e-07       0   0            0        0.007
+1e-06       0   0            0        0.002
+1e-05       0   0.028        0        0
+0.0001      0   0            0        0.027
+0.001       0   0            0        0.224
+0.01        0   0.016        0        0.735
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{Feature Priority (small DFE grid), Full Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 197860660949.1676
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00109138
+pi  = 0.000801495
+mut. rate = 3.046e-08 
+Ne = 27,284 (if mu=1e-8), Ne = 13,642 (if mu=2e-8)
+R2 = 59.1052% (in-sample)  57.2393% (out-sample)
+W = 
+          CDS    gene    other    phastcons
+------  -----  ------  -------  -----------
+1e-08   0.412   0.887        1        0
+1e-07   0       0            0        0
+1e-06   0       0            0        0
+1e-05   0       0.074        0        0
+0.0001  0       0            0        0
+0.001   0.066   0.008        0        0.136
+0.01    0.521   0.031        0        0.863
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 61830715800.047676
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00103388
+pi  = 0.000752601
+mut. rate = 3.172e-08 
+Ne = 25,847 (if mu=1e-8), Ne = 12,923 (if mu=2e-8)
+R2 = 57.8792% (in-sample)  55.631% (out-sample)
+W = 
+          CDS    gene    other    phastcons
+------  -----  ------  -------  -----------
+1e-08   0.432   0.898        1        0
+1e-07   0       0            0        0
+1e-06   0       0            0        0
+1e-05   0       0.064        0        0
+0.0001  0       0            0        0
+0.001   0.052   0.012        0        0.166
+0.01    0.516   0.026        0        0.834
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 250692005501.01343
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00144148
+pi  = 0.00106447
+mut. rate = 3.349e-08 
+Ne = 36,037 (if mu=1e-8), Ne = 18,018 (if mu=2e-8)
+R2 = 64.2844% (in-sample)  63.6077% (out-sample)
+W = 
+          CDS    gene    other    phastcons
+------  -----  ------  -------  -----------
+1e-08   0.72    0.875        1         0
+1e-07   0.033   0.003        0         0
+1e-06   0       0            0         0
+1e-05   0       0.085        0         0
+0.0001  0       0            0         0
+0.001   0.1     0.003        0         0.07
+0.01    0.148   0.033        0         0.93
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{CADD 6\% (small DFE grid), Full Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 197839911915.8117
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00102951
+pi  = 0.000801495
+mut. rate = 1.635e-08 
+Ne = 25,737 (if mu=1e-8), Ne = 12,868 (if mu=2e-8)
+R2 = 61.767% (in-sample)  60.2433% (out-sample)
+W = 
+          cadd6    other
+------  -------  -------
+1e-08     0.002    0.997
+1e-07     0.006    0
+1e-06     0        0
+1e-05     0        0
+0.0001    0.049    0
+0.001     0.325    0
+0.01      0.617    0.003
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 61824516495.02472
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.000977579
+pi  = 0.000752601
+mut. rate = 1.598e-08 
+Ne = 24,439 (if mu=1e-8), Ne = 12,219 (if mu=2e-8)
+R2 = 60.4032% (in-sample)  58.4847% (out-sample)
+W = 
+          cadd6    other
+------  -------  -------
+1e-08     0        0.992
+1e-07     0.001    0.001
+1e-06     0        0
+1e-05     0        0
+0.0001    0.03     0
+0.001     0.44     0
+0.01      0.529    0.008
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 250675195472.4508
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00136706
+pi  = 0.00106447
+mut. rate = 1.556e-08 
+Ne = 34,176 (if mu=1e-8), Ne = 17,088 (if mu=2e-8)
+R2 = 66.6755% (in-sample)  66.5047% (out-sample)
+W = 
+          cadd6    other
+------  -------  -------
+1e-08     0        0.994
+1e-07     0.006    0
+1e-06     0        0
+1e-05     0        0
+0.0001    0.099    0
+0.001     0.115    0
+0.01      0.78     0.006
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
 \printbibliography
-
 \end{document}

### 2023-07-05 bgspy/likelihood.py
nb updates and figs

@@ -1151,7 +1151,7 @@ class SimplexModel(BGSLikelihood):
         i = 1 + int(self._fixed_mu is None)
         return self.theta_[i:].reshape(self.nt, self.nf)
 
-    def summary(self, index=None):
+    def summary(self, no_unicode=False, index=None):
         if self.theta_ is None:
             return "\n[not fit yet]\n"  # not fit yet
         if index is None:
@@ -1198,16 +1198,25 @@ class SimplexModel(BGSLikelihood):
         nstarts = self.optim.thetas.shape[0]
         frac = np.round(self.optim.frac_success, 2)*100
         base_rows += f"number of successful starts: {nstarts} ({frac}% total)\n"
-        base_rows += f"π0 = {pi0:0.6g}"
+        if not no_unicode:
+            base_rows += f"π0 = {pi0:0.6g}"
+        else:
+            base_rows += f"pi_0 = {pi0:0.6g}"
         if cis is not None:
             l, u = cis[0][0], cis[1][0]
             base_rows += f" ({l:.5f}, {u:.5f})\n"
         else:
             base_rows += "\n"
         pi = pi_from_pairwise_summaries(self.Y.sum(axis=0))
-        base_rows += f"π  = {pi:0.6g}\n"
+        if not no_unicode:
+            base_rows += f"π  = {pi:0.6g}\n"
+        else:
+            base_rows += f"pi  = {pi:0.6g}\n"
         fixed_str = "(FIXED)" if self._fixed_mu is not None else ""
-        base_rows += f"μ_del  = {mu:0.4g} {fixed_str}"
+        if not no_unicode:
+            base_rows += f"μ_del  = {mu:0.4g} {fixed_str}"
+        else:
+            base_rows += f"mut. rate = {mu:0.4g} {fixed_str}"
         if cis:
             l, u = cis[0][1], cis[1][1]
             base_rows += f" ({l:.3g}, {u:.3g})\n"
@@ -1215,9 +1224,14 @@ class SimplexModel(BGSLikelihood):
             base_rows += "\n"
 
         #base_rows += f"Ne (del) = {int(Ne):,} (implied from π0 and μ)\n"
-        base_rows += f"Ne = {int(pi0 / (4 * 1e-8)):,} (if μ=1e-8), "
[…]

### 2023-07-05 manuscript/manuscript.tex
ms changes

--- 2023-07-04 manuscript/manuscript.tex
@@
 
-\subsection*{Estimates of Negative Selection Parameters in Humans}
+\subsection*{Estimates of the Distribution of Fitness Effects in Humans}
 
@@
 However, across all of our annotation models, new deleterious mutations in
-putatively conserved features are consistently estimated to have strongly
-deleterious effects (Figure \ref{fig:figure-3}A). This includes features such
-as coding sequences, PhastCons regions, or the top 6\% of basepairs with the
-most pathogenic CADD scores. Thus, even with a model that can better
-distinguish weak versus strong selection, our DFE estimates suggest
-
-qualitatively
-agree with recent work \parencite{Murphy2022-sj}.
-
-Moreover, the estimated mutation rates are close
-to pedigree-based estimates for most of these annotation models (XXX).
-
-Second, we find evidence that our method will indeed estimate weak selection
-across features
-
-Given that the B' method more
-accurately predicts diversity across weaker selection coefficients, we wondered
-whether our inferred DFEs could
-
-Linked selection theory suggests the reduction signal is determined by the
-spatial distributions of recombination and fitness variance along the
-chromosome. When fitness variation is determined mostly by negative selection,
-the
-
-Where does the information for the DFE come from?
-
-All of our models show a general trade off in fitting the linked selection
-signal.
-
-Next, we developed a composite likelihood method to estimate expected diversity
-without linked selection $\pi_0$, the deleterious mutation rate ($\mu_D$), and
-distribution of fitness effects ($\mathbf{W}$) under negative selection.
-Because the focus of our work is to fix deficiencies in the classic BGS model,
-our current implementation does not consider reductions in diversity due to
-recurrent hitchhiking or other positive selection processes. Since previous
-work has established that purifying selection is the dominant mode of linked
-selection in humans and hitchhiking plays a limited role
-\parencite{McVicker2009-ax,Murphy2022-sj}, humans are a suitable organism to
-test our method.
+likely conserved features are consistently estimated to have strongly
+deleterious effects (Figure \ref{fig:figure-3}A) consistent with previous work
+\parencite{McVicker2009-ax,Murphy2022-sj}. This includes features such as
+coding sequences, PhastCons regions, or the top 6\% of basepairs with the most
+pathogenic CADD scores. However, we find that there is considerable uncertainty
+in how strong selection is against the most deleterious mutations. For example,
+in the CADD 6\% full-track model, new mutations are estimated to have a
+selection coefficients of $s = 10^{-1}$ with 42\% chance, $s=10^{-2}$ with
+7.8\% chance, and $s=10^{-3}$ with 41\% chance. These estimates suggest a
+bimodal DFE (Figure \ref{fig:figure-3}A, second row, left column), and are
+mirrored across the other CADD models.
+
+Given that previous methods have typically not estimated the DFE up to
+selection coefficients of $s=10^{-1}$, we ran additional models using a DFE
+constrained to $s=10^{-2}$. Each of these constrained-DFE models across all
+populations fit the data less well than models run with the grid up to
+$s=10^{-1}$, explaining 1\% less of the leave-one-chromosome-out variation on
+average. As expected, the constrained-DFE models put more mass on $s=10^{-2}$
+(see Supplementary Table \ref{supp:tbl-r2}). Overall, this common behavior
+across models seems likely to reflect a non-identifiability issue between
+selective effects of $s=10^{-1}$ and $s=10^{-2}$, which we refer to as the
+\emph{strong-selection non-identifiability} issue.
+
+We propose that one mechanism for this non-identifiability is that mutations
+extremely harmful to fitness are difficult to distinguish from a smaller
+effective population. Intuitively, as $s$ increases, the average persistence
+time (in generations) of the mutation in the population goes below one
+generation, i.e. death of the carrier of the \emph{de novo} mutation. In this
+hypothetical limit, each occurrence of a very harmful mutation is like an
+instantaneous death and thus acts like a genome-wide reduced effective
+population size. In our model, this genome-wide effective population size is
+parameterized by $\pi_0$; thus, one line of evidence for this effect is a
+correlation between estimates of $\pi_0$ and average selection $\bar{s}$.
+Looking at the CADD 6\% models fit across the default and constrained DFE
+grids, we find this is the case. Across every population, average selection was
+about 7 times larger using the default grid, and $\pi_0$ was $5.6\%$ larger.
+However, there was no similar consistent change in mutation rate estimates
+among populations. Overall, this suggests that the linked selection signal
+alone carries little information that can differentiate very strong selection
+from a slightly smaller drift-effective population size.
+
+Despite the strong-selection non-identifiability issue, we still were curious
+how stable these estimates are across samples from reference populations. In
+particular, we focused on the Feature Priority model, which fits the data less
+well but is more interpretable since its annotation is based on gene features.
+Overall, estimates of the distribution of selection coefficients are stable
+across populations and had narrow uncertainty as estimated by a block jackknife
+procedure (Figure \ref{figure-3}B). Coding sequences show the strong-selection
+non-identifiability issue, and an uncertain signal of a neutral mode in the
+DFE. This would be expected given the very different levels of selective
+constraint synonymous and non-synonymous sites in coding sequences are under.
+The PhastCons class of this Feature Priority model, which contains non-coding
+highly conserved elements (XXX), is under strong selective constraint as
+expected.
+
+\subsection*{Estimates of the Deleterious Mutation Rate in Humans}
 
@@
 \end{figure}
-
-Given that previous methods used classic BGS theory to fit genome-wide patterns
-of human diversity, we were curious whether their qualitative findings would
-change once weak selection was more accurately predicted. In particular, the
-U-shaped relationship between $B'$ and the strength of selection suggests that
-a given reduction factor could be explained by two different selection
-coefficients.
 
@@
 \section*{Discussion}
+
+- the $pi_0$ problem
 

### 2023-07-05 manuscript/manuscript.tex
ms changes

@@
 about 7 times larger using the default grid, and $\pi_0$ was $5.6\%$ larger.
-However, there was no similar consistent change in mutation rate estimates
-among populations. Overall, this suggests that the linked selection signal
-alone carries little information that can differentiate very strong selection
-from a slightly smaller drift-effective population size.
+There was no similar consistent change in mutation rate estimates among
+populations. In the CADD 6\% model, genome-wide average reduction factor
+$\bar{B}$ was $\approx 6.1\%$ lower in the default versus constrained grid.
+Overall, this suggests that the linked selection signal alone carries little
+information that can differentiate very strong selection from a slightly
+smaller drift-effective population size.
 
@@
 constraint synonymous and non-synonymous sites in coding sequences are under.
-The PhastCons class of this Feature Priority model, which contains non-coding
-highly conserved elements (XXX), is under strong selective constraint as
-expected.
+The PhastCons class of this Feature Priority model, which contains highly
+conserved non-coding elements (XXX), is under strong selective constraint as
+expected. Interestingly, in our PhastCons Priority annotation model (Figure
+\ref{fig:figure-3}A bottom row), the DFE estimate for coding sequences
+excluding PhastCons regions is neutral. This is expected; the selection signal
+in coding regions is absorbed by the PhastCons feature.
 
@@
     \includegraphics[width=\textwidth]{figures/figure_4.pdf} 
-    \caption{}
+    \caption{(A) }
   \label{fig:figure-4}
 \end{figure}
+
+Early genome-wide models of background selection closely fit megabase-scale
+patterns in sequence divergence among five primates
+\parencite{McVicker2009-ax}. However, the mutation rate estimates from this
+model were unreasonably high compared to pedigree-based estimates (XXX). This
+mutation rate discordance has led to the hypotheses that positive selection
+could explain some fraction of the genome-wide linked selection signal (e.g.
+\cite{Enard2014-kz}); however, evidence suggests that the role of classic
+sweeps in humans is minuscule
+\parencite{Pickrell2009-tt,Hernandez2011-gs,Murphy2022-sj}. Furthermore, Murphy
+et al. find the classic BGS model fit using the CADD 6\% track leads to
+mutation rate estimates that are consistent with pedigree-based rates. Given
+that our method is based on a different theoretic approach, we curious whether
+it would lead to estimates consistent with pedigree-based mutation rates.
+
+Overall, across our CADD-based models and all populations, our mutation rate
+estimates are roughly concordant with pedigree-based estimates (Figure
+\ref{fig:figure-4}A). Our full-track CADD 6\% model estimates a mutation rate
+of $\mu = 1.46 \times 10^{-8}$ for YRI, $1.53 \times 10^{-8}$ for CEU, and $1.5
+\times 10^{-8}$ for CHB reference samples (Supplementary Table
+\ref{supp:tbl-r2}). Mutation rate estimates for our feature-based annotation
+models are unrealistically higher. We believe this could be caused by
+per-basepair heterogeneity in selective effects within each of these features.
+For example, mutations in coding sequences are known to have highly
+heterogeneous effects depending on whether the result is synonymous,
+non-synonymous, or mis-sense. If some fraction of the impact of negative
+selection is due to strong selection at a few sites, sparser annotation models
+like these cannot isolate these effects like CADD-based models. Instead, XXXHERE
+
+reflects a general problem
+with these more interpretable, but sparser feature-based tracks: heterogeneity
+in selective effects
+
+However, we noticed that estimates of mutation rate is very sensitive to the
+choice of annotation model. Between CADD 6\% and 6\% models, the
+
+\subsection*{Tests of Parameter Inference using Forward Simulations}
+
+Given
 
@@
 \widehat{B}_i$.
-
-\subsection*{Model Predictions and the Substitution Process}
 

### 2023-07-06 manuscript/manuscript.tex
ms changes

@@
 
-Overall, across our CADD-based models and all populations, our mutation rate
-estimates are roughly concordant with pedigree-based estimates (Figure
-\ref{fig:figure-4}A). Our full-track CADD 6\% model estimates a mutation rate
-of $\mu = 1.46 \times 10^{-8}$ for YRI, $1.53 \times 10^{-8}$ for CEU, and $1.5
-\times 10^{-8}$ for CHB reference samples (Supplementary Table
-\ref{supp:tbl-r2}). Mutation rate estimates for our feature-based annotation
-models are unrealistically higher. We believe this could be caused by
-per-basepair heterogeneity in selective effects within each of these features.
-For example, mutations in coding sequences are known to have highly
-heterogeneous effects depending on whether the result is synonymous,
-non-synonymous, or mis-sense. If some fraction of the impact of negative
-selection is due to strong selection at a few sites, sparser annotation models
-like these cannot isolate these effects like CADD-based models. Instead, XXXHERE
-
-reflects a general problem
-with these more interpretable, but sparser feature-based tracks: heterogeneity
-in selective effects
-
-However, we noticed that estimates of mutation rate is very sensitive to the
-choice of annotation model. Between CADD 6\% and 6\% models, the
+Across all populations, our mutation rate estimates from CADD-based models are
+roughly concordant with pedigree-based estimates (Figure \ref{fig:figure-4}A).
+Our full-track CADD 6\% model estimates the mutation rate as $\widehat{\mu} =
+1.46 \times 10^{-8}$ for YRI, $1.53 \times 10^{-8}$ for CEU, and $1.5 \times
+10^{-8}$ for CHB reference samples (Supplementary Table \ref{supp:tbl-r2}). As
+expected, the sparse-track CADD model mutation rate estimates are identical
+between the B' and classic BGS methods (Figure \ref{fig:figure-4}A top row).
+
+However, mutation rate estimates for feature-based annotation models show
+several issues. First, mutation rate estimates under using our implementation
+of classic BGS theory are an order of magnitude below the expected range
+(Figure \ref{fig:figure-4}A top row). We observe similar behavior when we use
+the classic BGS model to fit full-coverage annotation models (Figure
+\ref{fig:figure-4}A bottom row). This behavior is consistent with the fact that
+classic BGS theory cannot fit weak selection, and to accommodate features under
+weak constrain (e.g. introns, UTRs, and the ``other" feature) it must severally
+underestimate the mutation rate.
+
+Second, we noticed that across all populations and sparse/full tracks, the CADD
+6\% model consistently led to slightly higher mutation rates than the CADD 8\%
+model (Figure \ref{fig:figure-4}A bottom row; Supplementary Table
+\ref{supp:tbl-r2}). This pattern was observed in Murphy et al. too (2023;
+Appendix 1, Figure 16). This behavior is consistent with a non-identifiability
+issue between a slightly higher per-basepair mutation rate and annotation
+tracks that contain more conserved sequence. This is expected from theory,
+since both classic BGS and SC16 models only depend on mutation rate through the
+compound parameter $\mu L$ where $L$ is the length of the conserved segment.
+Even though our method is much more robust to including non-conserved regions
+like introns, we still observe this non-identifiability issue.
+
+Finally, we note that mutation rate estimates in the Feature Priority model are
+unreasonably high ($\widehat{\mu} \approx 4.2 \times 10^{-8}$), reminiscent of
+the high mutation rate estimates found under McVicker et al.'s
+(\citeyear{McVicker2009-ax}) model. While both our and Murphy et al.'s CADD and
+PhastCons-based models alleviate this issue, it is worth discussing why this
+could occur. We gain some insight from comparing the estimated mutation rates
+between our Feature and PhastCons Priority annotation models, which both
+contain the exact same number of basepairs, but these are assigned to different
+feature classes based on the priority of overlapping features. That one of
+these models is our best-fitting model, and the other our worst indicates a
+large role in per-basepair heterogeneity in selective effects in determining
+the genome-wide linked selection signal. CADD-based models fit well in part due
+to their fine-scale resolution of selective effects across the genome.
 
@@
 
-Given
+Given these the sensitivity of mutation rate estimates and the strong-selection
+non-identifiability issue, we next sought to assess the estimation accuracy of
+our composite likelihood method using forward simulations. In order to have
+enough megabase-scale data to apply our method, we independently evolved the
+first five human chromosomes. Due to the immense computational expense of
+conducting forward simulations at this scale over numerous parameter
+combinations, we used fixed mutation and selection parameters in putatively
+conserved regions (see Methods XXX). We note three important findings from
+these simulations.
+
+First, the linked selection signal, as measured by the $R^2$ between predicted
+and simulated megabase-scale diversity, increases with the intensity of
+selection against new deleterious mutations and mutation rate (Figure
+\ref{fig:figure-4}B, top row). This is consistent across both classic BGS and
+our B' method; the only exception is the classic BGS model run with a
+full-coverage track for the reasons described above. However, if new mutations
+are highly deleterious ($s=0.05$), $R^2$ is reduced. This is consistent with
+very strong selection having less localized effects and acting more like a
+genome-wide faster rate of drift.
+
+Second, both classic BGS theory and our B' method accurately infer the average
+selection coefficient under strong selection (Figure \ref{fig:figure-4}B,
+middle row). However, when selection was weak, the classic BGS model
+erroneously estimated strong selection and a very low mutation rate. By
+contrast, our B' method estimated selection coefficients much closer to their
+true value. A minor discrepancy occurs around $2Ns = 1$, for reasons discussed
+in Section XXX. Because we only simulated fixed selection coefficients and five
+chromosomes, we only assessed the accuracy of average selection coefficients
+and not the DFE probabilities.
+
+Third, with both our B' and the classic BGS models, we find mutation rate
+estimates show slight biases in the strong-selection regime (Figure
+\ref{fig:figure-3}B, bottom row). Still, our B' method leads to more accurate
+mutation rate estimates than classic BGS theory across a variety of selection
+coefficients. When BGS is not acting, either due to weak selection or a low
+rate of deleterious mutations (Figure \ref{fig:figure-4}B right column), all
+estimates are worse. However, we note that this is an unlikely region of
+parameter space and this issue is readily diagnosable from the poor model fit 
+as evident by low $R^2$ values.
 

### 2023-07-06 manuscript/manuscript.tex
ms changes

@@
 
-\subsection*{Predicted Diversity, Substitution Rates, and Fitness Variance}
+Finally, we also used simulations to test the sensitivity of our method to
+violations of two assumptions: constant population size and recessivity of
+selective effects. Since we find little difference in parameter estimates among
+YRI and bottlenecked out-of-Africa (CEU and CHB) samples, we tested whether a
+population expansion would bias our method. We expansion rates like those
+estimated for Europeans by \textcite{Gutenkunst2009-pg} had little effect on
+parameter estimates (Supplementary Section \ref{supp:sim-assum}). By contrast,
+we find our estimates are downwardly biased when mutational effects are weak
+and recessive, but are accurate under recessive strong selection. This occurs
+since the fixation probabilities of recessive mutations are in reality much
+lower than they are under our model assuming additivity.
+
+\subsection*{Predicted Diversity and Residual Signal}
 
@@
 
-When we fit our negative selection model to genome-wide patterns of diversity,
-the parameter estimates generate three predicted quantities: (1) the $B(x)$
-map, (2) the substitution rates per segment, and (3) the fitness variance per
-segment. Here we focus on the predicted $B(x)$ map and implied levels of
-diversity at the megabase scale. We discuss substitution rates and fitness
-variance in the next section.
-
-We also obtain a prediction $\widehat{B}_i$ for the reduction factor $B_i$ in
-each window $i$, which can be multiplied by our estimate of $\widehat{\pi}_0$
-to obtain a prediction of diversity $\widehat{\pi}_i = \widehat{\pi}_0
-\widehat{B}_i$.
+While we have previously shown that out-sample predicted diversity matches
+observed diversity (Section XXX), we also investigated whether there were
+regions of the genome that were particularly poor-fitting. Comparing predicted
+against observed diversity along chromosomes (see Supplementary Materials
+Section \ref{supp:chrom-fit}) we overall find close correspondence consistent
+with the high $R_\text{LOCO}^2$. However, we find a few large (tens of
+megabases) regions with systematically poorer fit (Supplementary Materials
+Figure \ref{supp:spatial-residuals}). In Figure \ref{fig:figure-5}A we see one
+such region from XXX to XXX. Interestingly, predicted diversity closely follows
+the peaks and troughs of this region. We note that a smaller segment within
+this region has been found by a genome-wide scan for associative overdominance
+\parencite{Gilbert2020-aw}.  For model XXX
+
+Another model diagnostic is to inspect whether observations are systematically
+different from predictions. We confirm the finding of \textcite{Murphy2022-sj}
+that regions predicted to experience little reduction in diversity due to
+background selection have higher diversity than expected (Figure
+\ref{fig:figure-5}B, orange line). Their work suggested that this could reflect
+ancient introgression between archaic humans and ancestors of contemporary
+humans. Finally, the variance around observed and predicted diversity levels
+falls very close to what we would expect under the theoretic
+coalescence-noise-only expectation ($R_\text{coal}^2$).
+
+\subsection*{Predicted Substitution Rates, and Fitness Variance}
 
@@
 
+Since our B' method is based upon quantitative genetic models of linked
+selection that explicitly incorporates the substitution process, fitting our
+model to genomic data leads to predictions that can be explicitly tested. In
+particular, there are two key predicted quantities: (1) the substitution rates
+per segment, and (3) the fitness variance per segment.
+
 Each of our maximum likelihood estimates $\widehat{\Psi}$ imply other
@@
 the regions
-
-\subsection*{Evolutionary Models}
-
-We model the diversity at punitively neutral sites. These are sites in regions
-we have \emph{a prior} reason to believe have a DFE nearly entirely
-concentrated in the neutral range. While this may not be the case in reality,
-it is necessary given the computational demands of modeling whole-genome
-selection.
-
-CDS+phastcons+UTR model
-
- - UTR as neutral control
-
- - CDS takes priority over phastcons, phastcons is fallback. Otherwise, the
-   effect of CDS would be biased by the fraction removed and classified as
-   phastcons. This is the right comparison for the substitution model.
-
-\subsection*{Predicted Substitution Rates}
-
-- mention neg. auto correlation?
-
-- Gene level analysis
-
-To evaluate the predicted substitution rate of our CDS+phastcons+UTR model,
-
-Additionally, our GBGS model makes predictions about the rate of substitutions
-in different segments under purifying selection. There are two stages of the
-prediction process. First, when the initial B' maps are calculated, the
-predicted substitution rate is calculated as a byproduct of
 
@@
 
-% As genomes are transmitted through the generations, lifetime events selectively
-% alter what is passed to the next generation. While selection directly acts on
-% the genetic variants that cause heritable fitness differences among
-% individuals, neighboring regions experience an indirect contraction in genetic
-% variability. This byproduct is known as \emph{linked selection}, and it occurs
-% when selection changes a site's frequency on timescale faster than
-% recombination can decouple that site from its genetic background. Because
-% linked selection modifies the genetic variability in a regionally-correlated
-% way, it creates observable patterns in genomic data that are used to understand
-% the type of selection processes acting on genomes and estimate key evolutionary
-% quantities.
-
-% , with characteristic length
-% $\nicefrac{s}{r}$ basepairs, where $s$ is the heterozygous selection
-% coefficient and $r$ is the per-basepair recombination rate. Sweeps were found
-% in humans, \emph{Drosophila}, and other species (XXX), though it was unclear
-% whether hitchhiking alone could explain the observed genome-wide correlation
-% between recombination rate and genetic diversity in some species (XXX).
-
-% Much work has been done on developing mathematical theory for two linked
-% selection processes: hitchhiking and background selection. These are
-% distinguished by the evolutionary processes that generate fitness variation and
-% the resulting selection dynamics. Under the classic hitchhiking model, a new
-% beneficial mutation sweeps fixation, leading to a dip in genetic diversity on a
-% small spatial scale. Since the sweeps leave relatively narrow footprint, they
-% influence genome-wide diversity only when beneficial mutations are recurring
-% and recombination is relatively limited, as in \emph{Drosophila}. These models
-% assume sweeps are sufficiently rare as to not occur simultaneously, which is
-% consistent with the slow estimated rate of strongly beneficial mutations (XXX).
-% Overall, the classic hitchhiking model is simple because the underlying
-% selection process does not generate polygenic fitness variation, even though
-% this is type of variation plays a dominant role in adaptation (XXX).
-
-% By contrast, background selection is a consequence of the fitness variation
-% created by a steady influx of new deleterious mutations. The classic background
-% selection model (XXX) considers how the balance between mutation introducing
-% deleterious variants into the population and selection removing them, reduces
-% nearby genetic variability. A critical assumption of this model is that such an
-% equilibrium between mutation and selection exists, which is the case only when
-% these mutations have a large enough fitness cost to ensure that they cannot fix
-% in the population. Under this process, the number of deleterious mutations per
-% gamete reaches a Poisson stationary distribution with rate $\nicefrac{U}{s}$,
-% where $U = \mu_D L$ is the deleterious mutation rate ($\mu_D$ per basepair per
-% generation) in a region of $L$ basepairs for mutations with a heterozygous
-% selection coefficient of $s$ (XXX). The fraction of gametes without any
-% mutations is then $f_0 = \exp(-\nicefrac{\mu_D}{s})$. Without recombination,
-% any gamete with deleterious mutations is destined to bring its linked neutral
-% variation with it to extinction, effectively reducing diversity levels to those
-% expected in a neutrally-evolving population of size $f_0 N$.
-
-% Remarkably, predictions from simple background selection models with
-% recombination closely match large-scale patterns of human diversity and
-% divergence (XXX). Since background approximately rescales a neutral coalescent
-% process by $f_0$, the predicted diversity is reduced from neutral levels by a
-% factor $B = \nicefrac{N_e}{N}$, which is the ratio of effective population size
-% under background selection to the population size. In background selection
-% models with recombination, $B$ is determined by the number of conserved
-% basepairs and recombination in a region. Consequently the predicted reduction
-% along the genome (known as $B$ maps) mirror the regional density of conserved
-% elements under purifying selection and recombination rates. These $B$ maps have
-% been used in studies of XXX, are inputs into machine learning models that
-% predict functional constraint, and XXX. Furthermore, since new deleterious
-% mutations are perpetually flowing into populations $B$ maps may provide a
-% better null model for evolution than the standard neutral model.
-
-% While background selection models do predict diversity levels  XXX
-
-% % outline?
-% - Disentangling evolutionary processes from the patterns of genetic variation
-% across the genome. The structure of what can be under selection shapes these
-% patterns when selection acts; when drift acts, these patterns should not exist.
-
-% - For species with small effective population sizes (and perhaps large map
-% lengths), BGS seems to be a dominant selective force in shaping patterns of
-% diversity (e.g. at the megabase scale in humans).
-
-% - However, current approaches to fitting BGS models fail under the weak
-% selection domain. This weak selection domain is important (discuss
-% HRI/ratchet/etc). Importantly, the characteristic scale of the impact of linked
-% selection is $\nicefrac{s}{r}$, meaning weak selection acts on smaller scales
-% than strong selection. Our ability to see selection at small scales may be
-% inhibited by the fact that our current BGS theory only works for strong
-% selection.
-
-% - We need to say briefly that we're using BGS to mean linked selection due to
-% purifying selection, and thus the strong and weak regimes (classically it has
-% just meant the strong regime).
-
-% \section*{Conclusion}
-
-% - care should be taken applying these methods to \emph{Drosophila}, since the
-% $\delta_{LD}$ term could be much larger.
-
-% \section*{Results}
-
-% \subsection*{Theory}
-
-% - overview of why BGS models fail?
-
-% From an evolutionary perspective, linked selection is important because by
-% modifying neighboring levels of genetic diversity, it is also directly altering
-% the selection dynamics at these neighboring sites.
-
-% The other use of these models is to understand how linked selection processes
-% differentially alter the genetic variability, or equivalently, the genealogical
-% trees, along the genome.
-
-% XXX how many sites are under selection at any moment? What's fitness variation?
-
-% Early on, some pointed out that evidence of strong selective sweeps in the
-% human genome was not the full story
-
-% Generally, in organisms with large amounts of recombination, linked selection
-% has
-
-% Weak selection processes
-
-% distorts
-% genome-wide
-
-% diversity.
-
-% The linked selection literature has primarily focused on developing theory for,
-% and quantifying the effects of two linked selection processes: hitchhiking and
-% background selection. Genome-wide models of the hitchhiking process consider
-% how beneficial mutations that sporadically enter the population
-
-% - QGLS polygenic linked selection
-
-% - H and W
-
-% - inflation factor
-
-% \subsection*{Statistical Methods}
-
-TODO we need to make appoint about how Ben G's paper found the V(n)->VA connectoin -- this is their patching model
-
-TODO we need to give some intuition for what Vm and V are in humans, e.g. back of envelope type stuff
-
-% The $N_{e,t}$ terms are analogous to an instantaneous coalesce rate at
-% generation $t$, which can be expressed in terms of the variance in allele
-% frequency change by noting that in an ideal Wright--Fisher population,
-
-% \begin{align}
-%   \V(x_t) &= p_0(1-p_0) \left[1 - \left(1 - \frac{1}{2N}\right)^t \right]
-% \end{align}
-
-% This is an asymptotic $N_e$, but this does not accurately describe the levels
-% of diversity (\cite{Santiago1998-bs} equation 16, \cite{Santiago2016-mu}
-% equation 4).
-
-% \begin{align}
-%   \V(x_t) &= \V(x_1) \left(1 + \sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i} \right) + \sum_{i \ne j} C_{i,j}
-% \end{align}
-
-% In general, the variance in allele frequency after $t$ generations
-% of neutral evolution is,
-
-%\begin{align}
-%  \label{eq:var_freq}
-%  %\V(x_t) &= \frac{x_0(1-x_0)}{2N}\left(\sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i-1} \right)
-%  \V(x_t) &= \V(x_1) \left(1 + \sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i} \right)
-%\end{align}
-
-% Each term is time-indexed because it is conditioned on the starting neutral
-% frequency in each individual, $x_{0,1}$. This equation is very general, and
-% allows for any reproduction model with finite offspring variance and can
-% account for the impact of linked selection under a variety of selection
-% processes.
-
-% Next, we can partition the gamete number into heritable ($f_i$) and
-% non-heritable ($\varepsilon_i$) fitness components of each individual $i$, $k_i
-% = f_i + \varepsilon_i$. We define $f_i$ such that $\nicefrac{1}{N} \sum_i f_i =
-% 2$ in a constant population, and $\varepsilon_i$ is noise term centered around
-% zero. Then, the neutral allele frequency change is can be written as the sum of
-% three independent stochastic terms,
-
-% \begin{align}
-%   \Delta p_1 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (f_i - 1) +
-%                \frac{1}{2N} \sum_{i=1}^N x_{0,i} \varepsilon_i
-%   + \frac{1}{N} \sum_{i=1}^N \sum_{j=1}^{k_i} \delta_{0,i,j} \\
-%   % \Delta p_1 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (f_i - 1) +
-%                % \frac{1}{2N} \sum_{i=1}^N x_{0,i} \varepsilon_i
-%   % + \frac{1}{N} \sum_{i=1}^N \sum_{j=1}^{k_i} \delta_{0,i,j} \\
-%   \Delta p_1 &= S_1 + D_1 + H_1
-% \end{align}
-
-% (c.f. \cite{Santiago1995-hx} equation 2 and \cite{Buffalo2019-qs} equation 30).
-
-% Each term is time-indexed because it is conditioned on the starting neutral
-% frequency in each individual, $x_{0,1}$. This equation is very general, and
-% allows for any reproduction model with finite offspring variance and can
-% account for the impact of linked selection under a variety of selection
-% processes.
-
-% Now, let us look the variance of $\var(\Delta p_1)$ over evolutionary
-% replicates.
-
-% \begin{align}
-%   \var(\Delta p_1) &= \var(S_1) + \var(D_1) + \var(H_1)
-% \end{align}
-
-% To introduce how this equation can be used to quantify the variance in allele
-% frequency change through time, $\var(p_t - p_0$ for $t > 0$, we first look at
-% evolution in a neutral Wright--Fisher model. Neutrality ensures that $S_t = 0$
-% for all $t$.
-
-%%% OLD
-
-%There has been considerable past work to find both the rate of the ratchet
-%(XXX), and summaries of the genealogies created under weak purifying selection
-%(XXX). Our approximation combines a few theoretic avenues from past work. Our
-%primary goal is to find a theoretic scaling function $B$ that rescales the neutral
-%pairwise coalescent times the appropriate amount to
-
-%Strictly speaking, there is no way to rescale the coalescent under general
-%background selection process, as genealogies deviate from the Kingman
-%coalescent slightly under strong selection
-%\parencite{Walczak2012-fi,Cvijovic2018-vd} and greatly under weak selection
-%\parencite{OFallon2010-my}
-
-%First, we approximate the coalescent effective population size, which
-%determines expected pairwise diversity, from the variance effective population
-%size. Then, we use a general quantitative genetic model of linked selection to
-%find the variance effective population size under background selection. Since
-%the impact of selection under this model is parameterized by the additive
-%genetic fitness variance, we use an approximation for this based on the
-%additive \emph{genic} variance, which depends on the rate of the ratchet.
-%Previous work has been unable to find a simple closed-form solution to the rate
-%of the ratchet because it depends on moments of the fitness distribution (e.g.
-%fitness variation), which in turn depend on even higher-moments of the fitness
-%distribution (e.g. the skew of the fitness distribution). While the ratchet
-%should scale with the probability of fixation, the probability of fixation
-%(Kimura, Malecot, XXX) is not accurate when multiple deleterious mutations are
-%segregating simultaneously and in close linkage with one another. A key insight
-%of \textcite{Santiago2016-mu} was that the classic equation for the probability
-%of fixation is fairly robust as long as the correct effective population size
-%is used. Their approach is to solve a system of two non-linear equations: one
-%for the variance effective population size (which depends on the rate of the
-%ratchet), and another for the time to fixation (which depends on the effective
-%population size).
-
-%Below, we step through the theory we use in our statistical methods. We
-%re-derive central parts to illustrate where approximations have been made,
-%which allows us to understand how these approximations may impact our
-%statistical methods and inspect model fit in regions where these approximations
-%are less likely to hold.
-
-%\subsection*{Pairwise Diversity under Background Selection}
-
-%The central quantity of interest for our statistical inference methods is
-%average pairwise diversity along the chromosome. While other summaries of the
-%data, such as the allele frequency spectrum, may be more informative about the
-%evolutionary processes acting in a region, we lack population genetic theory
-%for these quantities under most selective regimes and they require more degrees
-%of freedom to characterize. By contrast, neutral diversity under an
-%infinite-sites model is determined only by the per-basepair mutation rate $\mu$
-%and the expected pairwise coalescent time $\E[T_2]$,
-
-%\begin{align}
-%  \E(\pi) = 2 \mu \E(T_2)
-%\end{align}
-
-%Let us assume that there is a constant \emph{drift-effective} population size
-%$N_d$ which determines the rate of coalescence due only to neutral demographic
-%processes. Additionally, we assume some background selection process that in
-%expectation scales the effective population size by $b(t)$ at time $t$ (though
-%this could in theory be due to any process). Then, the expected pairwise
-%coalescent time is
-
-%\begin{align}
-%  \label{eq:T_2}
-%  \E(T_2) = \sum_{i=1}^\infty \frac{i}{2 b(i) N_d}\prod_{t=1}^{i-1} \left(1-\frac{1}{2 b(t) N_d}\right).
-%\end{align}
-
-%The time-dependency of $b(t)$ is required because the background selection
-%process will usually alter the rate of coalesce through time. For example,
-%under weak and intermediate selection, coalesce rates speed up as lineages move
-%in the past under intermediate to weak selection, similar to population growth
-%\parencite{Walczak2012-fi,Seger2010-rs,Good2014-yz}.
-
-%% Intuitively, this is
-%% because lineages are unlikely to coalesce with numerous less fit lineages they
-%% encounter as they go back in time, until ultimately coalescing in one of the
-%% least-loaded fitness classes.
-
-%This formulation is general and classic background selection theory (in the
-%no-ratchet domain) is a special case of it. Under classic BGS and load theory,
-%two randomly sampled lineages will each carry a number of deleterious mutations
-%sampled from a Poisson stationary distribution. Because any lineage carrying $j
-%> 0$ strongly deleterious mutations is destined to extinction, these lineages
-%can only coalesce in a common ancestor in the least-loaded class with zero
-%deleterious mutations. Under the Poisson stationary distribution, the size of
-%this class is $f_0 N$, where $f_0 = e^{-\nicefrac{U}{s}}$, so the coalescent
-%rate is scaled by $\nicefrac{1}{f_0}$. Once the lineages reach this fitness
-%class, there are no fitness differences among individuals, so they coalesce
-%following the Kingman process. A lineage moves backwards in time $t_d$
-%generations through the ``delay phase" to the least-loaded class. Classic BGS
-%theory ignores this delay phase, since $t_d \sim \log(j+1)/s << f_0 N$
-%(Durrett, p. 213).
-
-%%Here, we model the background selection process forward in time. We approximate
-%%the expected pairwise diversity with the probability of heterozygosity,
-%%$\E(\pi) \approx \mathcal{H}$, which holds under weak mutation and random
-%%mating, such that
-
-%%\begin{align}
-%%  %\mathcal{H} =  2 \mu \sum_{i=1}^\infty \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right)
-%%  \mathcal{H} &= 1 - \sum_{i=1}^\infty (1-\mu)^{2i} \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right) \\
-%%              &\approx 2 \mu \sum_{i=1}^\infty \prod_{t=1}^i \left(1-\frac{1}{2 b(t) N_d} \right)
-%%\end{align}
-
-%%in the limit where $\mu \to 0$. This is equation 4 of
-%%\textcite{Santiago2016-mu}; they define the sum as the heterozygosity effective
-%%population size.
-
-%\subsection*{Forward in Time Models of Linked Selection}
-
-%Next, show how quantitative genetic models of linked selection
-%\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu}
-%find expressions for $b(t)$ with by modeling the selective process forward in
-%time. Their approach is to consider how the variance in neutral allele
-%frequency change is impacted by a linked selection process.
-
-%\subsubsection*{Variance in Neutral Allele Frequency in a Neutral System}
-
-%For comparison, we start with the recursion for the variance in allele
-%frequency change in a neutral Wright--Fisher model. Under binomial sampling of
-%alleles, the random frequency $x_t$ is $2N x_t \sim \mathrm{Binom}(x_{t-1},
-%2N)$. If we look forward in time on generation and treat $x_0$ as fixed,
-
-%\begin{align}
-%  \V(x_1) = \E_1\left[(x_1 - x_0)^2\right]= \frac{x_0(1-x_0)}{2N},
-%\end{align}
-
-%and
-
-%\begin{align}
-%  \V(x_2) &= \E\left[(x_2 - x_0)^2\right] \\
-%          &= \E_1\left[\E_2\left[((x_2-x_1) + (x_1 - x_0))^2 | x_1\right]\right] \\
-%          &= \E_1\left[\E_2\left[(\Delta x_2 + \Delta x_1)^2 | x_1\right]\right] \\
-%          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + 2\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
-%          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
-%          &= \frac{x_0(1-x_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{x_0(1-x_0)}{2N}
-%\end{align}
-
-%Note that the cross-moment $\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right]
-%= 0$ under neutral evolution. In general, the variance in allele frequency
-%after $t$ generations of neutral evolution is,
-
-%\begin{align}
-%  \label{eq:var_freq}
-%  %\V(x_t) &= \frac{x_0(1-x_0)}{2N}\left(\sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i-1} \right)
-%  \V(x_t) &= \V(x_1) \left(1 + \sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i} \right)
-%\end{align}
-
-%This geometric series converges to the familiar expression,
-
-%\begin{align}
-%  \V(x_t) &= x_0(1-x_0)\left(1 - \left(1 - \frac{1}{2N}\right)^t \right).
-%\end{align}
-
-%which in the limit $t \to \infty$, converges to the Bernoulli variance
-%$x_0(1-x_0)$ as the frequency either goes to 1 with probability $x_0$ or 0 with
-%probability $1-x_0$.
-
-%% Now, if the effective population size were to change each generation, equation
-%% \eqref{eq:var_freq} would be,
-
-%% \begin{align}
-%%   \V(x_t) &= \V(x_0) \left(1 + \sum_{i=1}^t \prod_{j=1}^{i} \left(1 - \frac{1}{2N(j)}\right) \right)
-%% \end{align}
-
-%% Note that this too converges to $x_0(1-x_0)$ as $t \to \infty$.
-
-%\subsubsection*{Variance in Neutral Allele Frequency with Linked Selection}
-
-%Note that equation XXX is a special case, when the evolutionary process has
-%independent allele frequency changes through the generations and thus the
-%cross-moment is zero. Selection generates correlations between allele frequency
-%changes at nearby neutral sites. These correlations created by linked selection
-%inflate the variance in neutral allele frequency $\V(x_t)$ beyond what it would
-%be in a system without linked selection operating. The magnitude of this
-%inflation term is set by the selection process and its dynamics, linkage
-%disequilibrium, and the recombination rate. Even k
-
-%The theory of this began with \textcite{Robertson1961-ho},
-%and was thoroughly worked out by Santiago and Caballero
-%(\citeyear{Santiago1995-hx}, \citeyear{Santiago1998-bs}; see also
-%\cite{Barton2000-zg} and \cite{Buffalo2019-qs}). Our BGS statistical methods
-%employ the extension of this model for weak BGS \parencite{Santiago2016-mu}.
-%For completeness, we step through a quick derivation of the linked selection
-%inflation factor models here, based on
-%\textcite{Santiago1995-hx,Santiago1998-bs}.
-
-%Quantitative genetic models of linked selection
-
-%%%%%%
-
-%One way we can model these dependencies is to partition $\Delta_i$ into drift
-%and selection components. Over evolutionary replicates, the drift components
-%are independent and the selection components have some degree of
-%autocorrelation. Let us partition $\Delta x_i = \Delta_S x_i + \Delta_D x_i$,
-
-%\begin{align}
-%  \V(x_2) &= \V\left(\Delta x_2 + \Delta x_1\right) \\
-%          &= \V(\Delta_D x_2) + \V(\Delta_D x_1) + 2 \cov(\Delta_S x_2, \Delta_S x_1) \\
-%          &= \frac{x_0(1-x_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{x_0(1-x_0)}{2N} +  2 \cov(\Delta_S x_2, \Delta_S x_1) 
-%\end{align}
-
-%since $\Delta_D x_i \perp \Delta_D x_j$ for $i \ne j$. The covariances are
-%determined by the selective process dynamics, linkage disequilibrium, and the
-%recombination rate. These covariances are throughly worked out in
-%\textcite{Santiago1995-hx} and \textcite{Buffalo2019-qs}, so we give a short
-%derivation here. Throughout, we assume random mating, hermaphroditic
-%individuals, and a constant population size. We will use angular brackets to
-%indicate an average over all diploid individuals in the population, of some
-%value with an individual-level index. For example, if we denote the number of
-%allele an individual carries in generation $t$ as $z_{t,i} \in \{0, 1, 2\}$,
-%the population frequency at time $x_t$ is
-
-%\begin{align}
-%  x_t = \langle z_t \rangle = \frac{1}{2N} \sum_{i=1}^N z_{t,i}
-%\end{align}
-
-%Across evolutionary replicates, each parent $i$ passes one of their two neutral
-%alleles to a random number $k_i$ offspring. The \emph{heritable fitness} of an
-%individual $f_i$ is defined as the expected number of offspring passed down
-%across evolutionary replicates, such that $k_i = f_i + \varepsilon_i$.  Some
-%selection process determines the distribution of $f_i$, but the this occurs
-%independently of the process that sorts neutral alleles into diploid
-%individuals. We are unsure of the neutral allele count in an individual in the
-%initial generation, $z_{t,0}$, so we condition on it and take expectations
-%later. Then, the frequency in the initial and first generations are,
-
-%\begin{align}
-%  x_0 = \langle z_0 \rangle = \frac{1}{2N} \sum_{i=1}^N z_{0,i} \\
-%  x_1 = \langle z_1 \rangle = \frac{1}{2N} \sum_{i=1}^N \left( k_i \frac{z_{0,i}}{2} + \sum_{j=1}^{k_i} \delta_{i,j} \right)
-%\end{align}
-
-%The first term is the average number of neutral alleles each parent passes to
-%its $k_i$ offspring. In reality, alleles are discrete, so each heterozygous
-%parent also passes a random $\pm \nicefrac{1}{2}$ to each offspring, which
-%leads to the second term. The variable $\delta_{i,j}$ is 0 if the parent is a
-%homozygote, but if the parent is a heterozygote, is $\pm \nicefrac{1}{2}$ with
-%equal probability over evolutionary replicates.
-
-%Since individuals are exchangeable with their contribution to the
-%heterozygosity term and population size is constant,
-
-%\begin{align}
-%  \sum_{i=1}^N \sum_{j=1}^{k_i} \delta_{i,j} = \sum_{i=1}^N \delta_{i} 
-%\end{align}
-
-%We can further partition the offspring number into heritable and non-heritable
-%parts,
-
-%\begin{align}
-%  x_1 &= \frac{1}{2N} \sum_{i=1}^N \left( (f_i + \varepsilon_i) \frac{z_{0,i}}{2} + \delta_{i} \right) \\
-%      &= \frac{1}{4N} \sum_{i=1}^N f_i z_{0,i} + \frac{1}{4N} \sum_{i=1}^N \varepsilon_i z_{0,i} + \frac{1}{2N} \sum_{i=1}^N \delta_{i} \\
-%      &=  \frac{1}{4} \langle f \; z_{0} \rangle + \frac{1}{4} \langle \varepsilon \; z_{0} \rangle + \frac{1}{2}\langle \delta \rangle \\
-%\end{align}
-
-%% TODO factor 1/2 error
-
-%\begin{align}
-%  x_1 - x_0 &= \frac{1}{4} \langle f \; z_{0} \rangle + \frac{1}{4} \langle \varepsilon \; z_{0} \rangle + \frac{1}{2}\langle \delta \rangle - \langle z_0 \rangle \\
-%            &= \frac{1}{4} \langle (f-1) \; z_{0} \rangle + \frac{1}{4} \langle \varepsilon \; z_{0} \rangle + \frac{1}{2}\langle \delta \rangle 
-%\end{align}
-
-%Now, we look at the variance of this over evolutionary replicates. By symmetry,
-%the expected change in a neutral allele is $\E(\Delta x_t) = 0$ since we track
-%an arbitrary neutral allele. Then the variance $\V(\Delta x_1) = \E( {\Delta
-%x_1}^2)$,
-
-%\begin{align}
-%  \V(\Delta x_1) &= \frac{1}{4} \langle (f-1) \; z_{0} \rangle + \frac{1}{4} \langle \varepsilon \; z_{0} \rangle + \frac{1}{2}\langle \delta \rangle 
-%\end{align}
-
-%OLD
-%Let us imagine there is some stochastic process $b(t)$ that rescales the coalescent
-%rate,
-
-%\begin{align}
-%  \V(x_t) &= \frac{x_0(1-x_0)}{2N(0)}\left(\sum_{i=1}^t \prod_{j=1}^{i-1} \left(1 - \frac{1}{2b(j)N}\right) \right)
-%\end{align}
-
-%\begin{align}
-%  \var(p_3 - p_0) = \E & \left( \bigg( \underbrace{S_1(1 + (1-r) G + (1-r)^2 G^2)}_{\text{creation and persistence of generation 1 associations} \;\; := \; S_1 Q_3} \right. +  D_1 + H_1+ \nonumber \\
-%                       & \underbrace{S_2(1 + (1-r) G)}_{\text{creation and persistence of generation 2 associations} \;\; := \; S_2 Q_2}+ D_2 + H_2 + \nonumber \\
-%                       &  \underbrace{S_3}_{\text{creation of generation 3 associations} \;\; := \; S_3 Q_1  }\left. + D_3 + H_3 \bigg)^2 \right). &
-%    \label{eq:scr-var2}
-%\end{align}
-%%
-
-%Note the similarity between the equations.
-
-%\begin{align}
-%  V &\propto \frac{1}{2b(0)N_d}\left(\sum_{i=1}^\infty \prod_{t=1}^{i-1} \left(1 - \frac{1}{2b(t)N_d}\right) \right) \\
-%  W &\propto \sum_{i=1}^\infty \frac{i}{2 b(i) N_d} \prod_{t=1}^{i-1} \left(1-\frac{1}{2 b(t) N_d}\right) \\
-%\end{align}
-
-%Let us define the function $a(t)$ to be the product, which is the cumulative
-%failure to coalesce up to time $t$.
-
-%\begin{align}
-%  a(t) = \prod_{i=1}^{t-1} \left( 1-\frac{1}{2 b(i) N_d} \right)
-%\end{align}
-
-%Comparing this to equation \eqref{eq:T_2} up to constant factors,
-
-%\begin{align}
-%  V &\propto \sum_{i=1}^\infty a(t) \\
-%  W &\propto \sum_{i=1}^\infty i \; a(t) \\
-%\end{align}
-
-%end{align}
-
-%XXX we swap indices
-
-%Now, if we assume that the final coalesce in the waiting time process were to
-%have a fixed rate of $\nicefrac{1}{2b(0)N_d}$, this
-
-%we see that both the waiting time process and the variance process
-
-%Intuitively, these define a stochastic selection process in two ways through
-%the function $b(t)$ that modulates the rate of pairwise coalesce.
-
-%where if $N_e(t) = N_e$ and $t=1$, this is the familiar Wright--Fisher variance
-%$\V(p_1) = p_0(1-p_0) / 2N_e$. When selection is not acting, allele frequency
-%changes per generation are independent, and neutral diversity is set by this
-%single $N_e$.
-
-%XXX
-
-%Since we track an arbitrary neutral allele's frequency $p_i$, by symmetry,
-%linked selection has no impact on the direction of change, $\E(\Delta p_i) =
-%0$. We will work through a short proof of $\V(p_t)$ by induction following the
-%approach of \textcite{Santiago1995-hx}. Throughout, we will use angular
-%brackets to indicate an average over all diploid individuals in the population,
-%of some value with an individual-level index. For example, if we denote the
-%number of allele an individual carries in generation $t$ as $x_{t,i} \in \{0,
-%1, 2\}$, the population frequency at time $p_t$ is
-
-%\begin{align}
-%  p_t = \langle x_t \rangle = \frac{1}{N} \sum_{i=1}^N x_{t,i}
-%\end{align}
-
-%\begin{align}
-%  \Delta_H p &= \langle x (f-1) \rangle \\
-%  \Delta_N p &= \langle x d \rangle \\
-%  \Delta_M p &= \langle \Delta \rangle \\
-%\end{align}
-
-%Assuming uninherited, random environments, only the $\Delta_{_H} p$ terms
-%create autocorrelation in $p_t$. The autocorrelation is created because during
-%the random pairing of neutral alleles to fitness backgrounds, associations are
-%created between an individual $i$'s allele frequency and their fitness $f_i$.
-%This is because both fitness and the neutral alleles reside in the same
-%individual in the present generation.
-
-%We can imagine now how these associations change over evolutionary replicates
-%in the next generation. In a fraction $r$ of the cases, the neutral allele 
-%recombines off (or segregates away, if they are on different chromosomes) the
-%fitness background. In a faction of $1-r$ XXX
-
-%\begin{align}
-%  \cov(\Delta_H p_t, \Delta_H p_s) &= \E(\langle x_t (f_t-1) \rangle , \langle x_s (f_s-1) \rangle)
-%\end{align}
-
-%Note that this covariance is only
-
-%In the first generation, the frequency change
-%can be partitioned into three components. Conditioning on the parent's
-%heritable fitnesses, we can imagine constructing an infinite number of
-%evolutionary replicates. Across all replicates, a random neutral allele's is
-%perturbed
-
-%\begin{align}
-%  p_1 - p_0 = \sum_{i=1}^{N} 
-%\end{align}
-
-%Since the variance in neutral allele frequency change is also determined by the
-%pairwise coalescence rate, this provides us with a way to model $b(t)$ under
-%linked selection.
-
-%These models derive an expression for the variance-effective population size of
-%a focal neutral site, $N_v(t)$, based on how linked fitness variation perturbs
-%the frequency trajectory at this site.  The variance-effective population size
-%for a constant $N_e$ is defined by the variance in allele frequency change at
-%the neutral site, over evolutionary replicates,
-
-%where if $t=1$, this is the familiar Wright--Fisher variance $\V(p_1) =
-%p_0(1-p_0) / 2N_e$. When selection is not acting, allele frequency changes per
-%generation are independent, and neutral diversity is set by this
-%single $N_e$.
-
-%However when selection is acting at linked sites, neutral diversity is
-%determined by the long-run $\V(p_t)$ as $t \to \infty$. Intuitively, this is
-%because the rate of drift is magnified by autocorrelation in allele frequency
-%change (XXX), and diversity is determined by the cumulative impact of these
-%autocorrelations in the long-run. This autocorrelation effectively reduces the
-%coalescence rate in a time-dependent manner, analogous to equation
-%\eqref{eq:T_2}. Thus, a single constant $N_e$ is insufficient to describe
-%levels of neutral variation, since it is insufficient
-
-%The variance in allele frequency change after $t$ generations is
-
-%\begin{align}
-%  \V(p_t) = p_0(1-p_0)\left(1 - \prod_{i=1}^t \left(1-\frac{1}{2N_e(t)}\right)\right)
-%\end{align}
-
-%This is the basis of the quantitative genetic models of linked selection: a
-%forward in time model for $\V(p_t)$ implies a pairwise coalescence rate
-%function $b(t)$, which determines statistics based on pairwise coalescence such
-%as $\E(\pi)$.
-
-%Next, we consider how fitness variation impacts $\V(p_t)$.
-
-%Throughout, we will take the variance $\V(p_t)$ to be the variance over
-%evolutionary replicates. Since the variance is for a random neutral allele's
-%frequency,
-
-%a reasonable approximation to how background selection process
-%rescales the coalescent rate at this site is given by $b(t) =
-%\nicefrac{N_v(t)}{N_d}$.
-
-%\begin{align}
-%  N_v = \lim_{t \to \infty} \frac{p_0 (1 - p_0) - \var(p_{t-1})}{2\var(p_t) - \var(p_{t-1})}
-%\end{align}
-
-%\subsection*{Linked Selection Models for Variance-Effective Population Size}
-
-%We wish to derive an expression for the reduction factor $B$. Ideally, this
-%reduction factor would rescale the coalescent-effective population size, but
-%this would require knowing the coalescent rate function $\lambda(t)$ over the
-%evolutionary history $0 < t < \infty$. Instead, following previous work
-%\parencite{Santiago1995-hx,Santiago1998-bs}, we derive the rescaling function
-%$B$ relative to the variance-effective population size.
-
-%Following past quantitative genetic models of linked selection
-%\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu},
-%we consider the long-run variance effective population size,
-
-%\begin{align}
-%  N_v := \lim_{t \to \infty} \frac{p_0(1-p_0)}{2\var\left(p_t - p_0\right)}
-%\end{align}
-
-%% ADD TO INTRO TODO
-%Comment about characteristic scale $\nicefrac{\mu}{s}$. Is the fact that we see
-%BGS on megabase scales because this is the only scale we can detect strong
-%selection under?
-
-%approximately $\pi \approx 4N_e \mu$ (for low mutation rate). However, this
-%assumes a constant Kingman coalescent with coalescent effective population size
-%$N_e$.
-
-%\subsection*{}
-
-%We model the reduction in pairwise coalescence times along the genome due to
-%background selection by extending the work of \textcite{Santiago2016-mu}
-%(hereafter, the SC16 model). The SC16 model derives from a class of linked
-%selection \parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs} that
-%predict the reduction in pairwise coalescence time at a focal neutral site
-%using macro-state parameters such as fitness variation, rather than micro-state
-%parameters like the frequency trajectories of selected alleles (e.g.
-%\textcite{Maynard_Smith1974-zr}). We will briefly introduce the SC16 model's
-%approach and how we validated this model could be extended to whole-genome
-%inference.
-
-%Classic background selection theory models how coalescence times are reduced by
-%deleterious mutation-selection balance at neighboring sites. Building upon
-%older models of genetic load \parencite{Crow1970-wj,Kimura1966-bk}, these
-%models consider the equilibrium distribution of the \emph{number} of
-%deleterious mutations segregating on haplotypes
-%\parencite{Charlesworth1993-gb,Nordborg1996-nq}. Given an infinitely large
-%population, or sufficiently strong selection, the deleterious mutations in
-%these models cannot fix; consequently, any lineage containing
-
-%While the effect of selection on linked selection distorts neighboring
-%genealogies (XXX), skewing the site frequency spectrum (XXX) and altering
-%linkage disequilibria along the chromosome, we focus on the impact of selection
-%on pairwise coalescence rates. This simple summary is amenable to mathematical
-%and statistical inference, and is robust to difficulties in identifying rare
-%variants. The effect of background selection on pairwise coalescence rates is
-%expressed as the reduction ratio $B = \nicefrac{N_e}{N}$ where $N$ is the
-%drift-effective population size (i.e. including demographic processes) and
-%$N_e$ is the effective population size under background selection.
-
-%Under quantitative genetic models of linked selection, the reduction in
-%effective population size is determined by the product of fitness variation,
-%$V_f$, and an inflation factor $Q$. It is helpful to understand this by comparing
-%it to a model of $N_e$ without \emph{heritable} fitness variation but with
-%\emph{non-heritable} fitness variation. In this case,
-
-%\begin{align}
-%  N_e = \frac{4N}{2 + V_k},
-%\end{align}
-
-%where $V_k$ is the variance in offspring number across individuals. Under a
-%constant population size Wright--Fisher model of reproduction, $V_k \approx 2$,
-%and this simplifies to $N_e = N$. However, as first noted by
-%\textcite{Robertson1961-ho}, linked selection inflates the variance in
-%offspring number, as a neutral allele that becomes associated with a
-%high-fitness or low-fitness genetic background has an $1-r$ chance of remaining
-%associated with that background in the next generation, where $r$ is the chance
-%of recombination (or independent assortment, $r=\nicefrac{1}{2}$).
-
-%This persistence of this association inflates the effective variance in fitness
-%by a factory $Q^2$,
-
-%\begin{align}
-%  N_e = \frac{4N}{2 + V_k + 4 Q^2 V_f},
-%\end{align}
-
-%\begin{align}
-%  N_e = \frac{4N}{2 + V_k + 4 Q(r)^2 V_f},
-%\end{align}
-
-%\begin{align}
-%  N_e = N \exp\left(-Q^2 V_f\right)
-%\end{align}
-
-%In a region of $L$ basepairs under purifying selection, there is a net flux of
-%deleterious mutations in at rate $U = 2 \mu L$ per generation (where the factor
-%of two is because in diploids, $2L$ is the total number of basepairs per genome).
-
-%Typical models of background selection consider the equilibrium distribution of
-%the \emph{count} of deleterious mutations, rather than then level of fitness
-%variation implied by this count. If we assume fixed additive effects across
-%alleles and multiplicative fitness effects across loci, each copy of a
-%deleterious mutation an individual carries reduces their fitness by a factor of
-%$1-s$. The total reduction for an individual is
-
-%\begin{align}
-%  x = (1-s)^n
-%\end{align}
-
-%where $n$ is the number of deleterious alleles an individual carries. Under
-%mutation-selection balance, the population evolves to either a stationary
-%distribution under the strong BGS domain, or a traveling distribution in weak
-%selection case. Under some assumptions, equilibrium models of the strong domain
-%show that the distribution $f(n) \sim \text{Pois}(\nicefrac{U}{s})$
-%\parencite{Kimura1966-bk,Crow1970-wj} (CHECK) in the asexual case, and the
-%fixation rate of deleterious mutations per generation, or ratchet rate, is $R =
-%0$ since the probability of fixation, $p_F$ is zero. By contrast, as the
-%selection coefficient grows weaker, $s \to 0$, the probability of fixation $p_F
-%\to \nicefrac{1}{2N}$, and the ratchet approaches the neutral rate of $R =
-%\mu$. The weak selection domain is of importance because both because it
-%implies a certain likely irreversible rate of fitness reduction, and because it
-%could possibly confound population genetic inference which assumes dynamics
-%only under the strong model.
-
-%As $s \to 0$, the fitness distribution, put on a scale reflecting the fitness
-%relative at one moment in time, travels leftwards and changes shape. It has
-%been difficult to find dynamic equations for the rate of the ratchet and
-%fitness distribution, since lower moments of the distribution, such as the
-%variation, depend on higher moments, such as the skew
-%\parencite{Good2013-lp,Haigh1978-gt,Higgs1995-xc}. This is known as the ``moment
-%closure problem" and it occurs regardless of whether one models the count
-%number of deleterious mutations (XXX) or the fitness distribution (XXX).
-
-%We can approximate the mapping between the distribution of deleterious
-%mutations in the population, and the fitness distribution. It is worth noting
-%that the count of deleterious mutations in an individual $i$ can be written
-%$n_i = \sum_l^{L} g_l$ where each $g_l$ is an indicator variable for whether
-%the individual carries a mutation at position $l$. Then, $\V_i(n) = \sum_l
-%\V(g_l) + \sum_{i,j}\cov(g_i, g_j)$.
-
-%the
-%number of deleterious mutations in gamete $i$; thus $V(n) = \sum_i^{2N} V(x_i)
-%+ \sum_{i \ne j} \cov(x_i, x_j)$ where the last term is the correlation among
-%gamete mutation counts, which is an LD quantity. Then, assuming $s$ is small,
-%we can write express this in terms of the additive genetic variance,
-
-%\begin{align}
-%  V(f) &\approx s^2 \V_a(n) + \Delta_\text{LD}
-%\end{align}
-
-%The first term is the additive genic variance, which is
-
-%Under this model, the rate of the ratchet $R = U  - s V(n)$, which implies,
-
-%\begin{align}
-%  R = U - \nicefrac{V(f)}{s} \\
-%  V(f) = s(U - R)
-%\end{align}
-
-%Throughout, it's worthwhile to keep into mind many population-level
-%macro-states are simply distributional summaries of the current population. For
-%example, the variance in deleterious mutations in the population is simply,
-%$\V(n) = \nicefrac{1}{2N} \sum_i (x_i - \mu_x)^2$
-
-%- We need to mention the two heterozygosities (S\&C '16 says they diverge under
-%$N_e s < 1$ -- could explain anything we see?)
-
-%- A point about reconciling the two micro and macro-state theories of
-%population and quantitative genetics.
-
-% \subsection*{The Rate of the Ratchet}
-
-% \section*{Methods}
-
-% \subsection*{Data Preprocessing}
-
-% \subsection*{Human}
-
-% Site-level neutral and accessibility masks
-
-% Bin-level masks based on fraction of accessible bases in a window (7\%).
-
-% Bin-level tail filtering.
-
 \printbibliography
@@
 \end{document}
-
-% A central goal of evolutionary genetics is to quantify the relative roles
-% different evolutionary processes have on genetic variation. Certain
-% evolutionary processes, such as demographic bottlenecks or expansions, act
-% uniformly across the genome as a whole. Other processes, such as selection, are
-% hypothesized to act differently across the genome depending on the density of
-% genomic features, and local heterogeneity in recombination and mutation rate.
-% With the emergence of population genomic data, a variety of statistical models
-% have teased apart the contribution of these various processes. These
-% statistical models differ in how they reduce large genomic data to summary
-% statistics, as well as in their statistical methodology. Some models reduce
-% genomic data to the histograms of the frequency spectrum (XXX), while others
-% model scalar summaries of the frequency spectrum along the genome.
-
-% One class of models, parametric statistical models, leverage population genetic
-% theory to inform expectations of the mean pairwise diversity along the
-% chromosome. Estimation under these models often uses a composite likelihood
-% approach, where the joint distribution is assumed independent across positions.
-% While composite likelihoods are simpler to formulate and lead to consistent
-% parameter estimation, ignoring dependencies between neighboring sites has two
-% primary disadvantages. First, unlike proper likelihoods, the composite
-% likelihood's curvature is uninformative about parameter estimate uncertainty.
-% Previous work has circumvented this shortcoming through (1) resampling-based
-% uncertainty estimates, or (2) adjusting the composite likelihood's curvature to
-% account for these dependencies. Second, similar to inference under phylogenetic
-% comparative models, the dependency structure tells us how to properly weight
-% information across sites. For example, imagine a conserved segment 10 kilobases
-% long right on the border of two genomic bins, which due to linked selection,
-% bot experience a reduction in diversity. When estimating parameters, the
-% independence assumption of composite likelihoods would weight these observed
-% reductions equal to the amount of a reduction observed on an entirely different
-% independent chromosome.
-
-% patterns of genetic variability along the genome are treated as
-% coming from an indeprocess $m(z)$.
-
-% Here, we focus on solving an issue processes: background
-
-% BGS as evolution around the static features of the genome. Due to the slow rate
-% of genomic structural evolution, much of the genome can be treated as static
-% throughout the genome. These regions are under differing levels of selective
-% constraint, which induces background selection at neighboring regions.
-
-% Signal processing
-
-% Genomes have inherent structure that impacts the linked selection map. There is
-% tremendous heterogeneity across species in the organization of likely selected
-% regions on the genetic and physical maps.
-
-% Within an organism, there are two key determinants of B maps: recombination and
-% gene density.
-
-% Initially, the relationship between recombination and diversity was regressed
-% with no regard to the spatial organization of the genome.
-
-% Certain regions are pinchpoints of selection: some combination of low
-% recombination and high gene density lead to very strong reductions in genetic
-% diversity. In these regions, coalescent dynamics change considerably; where as
-% under strong background selection, deleterious alleles would not fix, in these
-% regions the can fix. Additionally, the selective interference impacts positive
-% selection.
-
-% Understanding the extent to which linked selection influences other
-% evolutionary processes is a central question in evolutionary genetics. Two
-% central questions have emerged. First, while genetic drift is thought to be the
-% main source of random fluctuations in neutral allele frequencies, linked
-% selection can have a similar impact on neutral sites, increasing the variance
-% in allele frequency change as if the population size were effectively smaller.
-% Although linked selection causes significant reductions in genetic diversity in
-% some species, it is unlikely to explain the narrow range of observed diversity
-% across species \parencite{Buffalo2021-uf}. Second, by perturbing the
-% frequencies of neighboring alleles, selection can interfere with their
-% selection dynamics, known as Hill-Robertson interference (HRI, XXX). When
-% recombination is limited, simultaneous selection across the genome reduces the
-% odds beneficial mutations are fixed (HRI), effectively like selection was
-% occurring in an smaller population size. The Hill-Robertson effect is thought
-% to be weak in organisms in organisms with large recombination maps, perhaps
-% because of selection for increased recombination.
-
-% In organisms with limited recombination such as
-% \emph{Drosophila}, estimates suggest background selection plays a marginally
-% larger role than hard sweeps in reducing diversity \parencite{Elyashiv2016-vt}.
-
-% closely predict large-scale patterns of
-% diversity across the genome
-
-% Since far more new mutations are deleterious than beneficial in functional
-% regions, it would seem like background selection would have a dominant role in
-% shaping genetic diversity. However, deleterious mutations segregate on a
-% timescale inversely proportional to their selection cost (XXX). Strongly
-% deleterious mutations eliminate neighboring genetic variability on the few
-% chromosomes they may recombine onto before the eventual extinction, and
-% consequently have very weak effects on larger spatial scales. Moderately weakly
-% selected mutations can persist for much longer periods, allowing for more
-% recombination before their eventual loss and produce reductions in diversity on
-% shorter spatial scales.  The reduction in diversity under background selection
-% is determined approximately by $\nicefrac{U}{R}$, where $U = \mu_D L$ is the
-% number of deleterious mutations (at rate $\mu_D$ per basepair, per generation)
-% in a conserved region of $L$ basepairs per generation, and $R = r L$, which is
-% the total recombination per generation in Morgans. Since the deleterious
-% mutation rate is relatively constant across conserved regions, the reduction
-% due to background selection closely matches the composition of conserved
-% regions and recombination along the genome. Remarkably, simple mathematical
-% models of background selection closely predicts levels of human-chimp
-% divergence at the megabase scale (XXX) from genome-wide patterns of
-% recombination and regions conserved between species (McVicker).
-
-% This leads to some of the unidentifiability of hitchhiking versus background
-% selection.
-
-% and the reduction is determined
-% by
-
-% While the rate
-% new beneficial mutations enter the population is low (XXX),
-
-% The spatial scale of background selection is
-
-% kkkkkk
-
-% not all
-% adaptation occurs from such new beneficial mutations sweeping to fixation (XXX)
-
-% Other processes also act
-% on beneficial fitness variation to engender linked selection, such as polygenic
-% selection after an optimum shift \parencite{Buffalo2019-qs,Buffalo2020-my} or
-% soft sweeps from standing variation (XXX).
-
-% In turn, statistical methods have use this theory for a variety of wide-ranging
-% empirical problems.  Early on, researchers
-
-% While it is difficult to differentiate the selective processes
-% leading to linked selection, studies indicate background selection
-
-% hitchhiking events where a low-frequency
-
-% studies
-% in \emph{Drosophila} and humans indicate background selection has a large role
-% in shaping patterns of diversity, especially at large spatial scales (i.e.
-% megabases).
-
-% Another central empirical goal is to estimate the relative strengths of the
-% these two selection processes on levels of genetic diversity (Eyalshiv). Since
-% the effects of linked selection are determined in part by mutation rate and
-% distribution of fitness effects (DFE), these statistical methods also estimate
-% these important evolutionary parameters from genomic data. These estimates
-% should, in theory, align with others using different approaches, such as
-% trio-based estimates of the mutation rate and estimates of the DFE from
-% site-frequency spectra.
-
-% Our focus here is to begin to fix some deficiencies in the background selection
-% model. In particular,

### 2023-07-06 manuscript/supplementary.tex
ms changes

@@
 \subsection{Demographic Expansion and Recessive Simulation Tests}
+\label{supp:sim-assum}
 
@@
   \includegraphics[width=0.8\textwidth]{figures/supplementary/subfigure_sim_mle_expansion.pdf}
-  \caption{}
+
+  \caption{Simulations to diagnose the effect of population exponential
+      expansion on our composite likelihood estimates. The simulated population
+      expansion occurs at a rate of 0.4\% a generation, starting $9.3N$ (out of
+      $10N$) generations in. These parameters were meant to approximate the
+  European expansion parameters estimated by \textcite{Gutenkunst2009-pg}.
+Overall, we see little impact on estimated parameters.}
+
 \end{figure}
@@
   \includegraphics[width=0.8\textwidth]{figures/supplementary/subfigure_sim_mle_recessive.pdf}
-  \caption{}
+
+  \caption{Simulations to diagnose the effect of non-additive mutations. These
+      simulations use a dominance coefficient of $h=0.3$, i.e. the selective
+      effects are $1$ ancestral state, $1+hs$ for heterozygotes, and $1+s$ for
+      homozygotes. We find that slightly recessive mutations have little effect
+      on the estimated selection coefficient when selection is strong,
+      consistent with the fact that only the selective effect in heterozygotes
+      matters when mutations are kept at low frequency by strong selection.
+      However, our B' method underestimates weak selection coefficients when
+      mutation effects are recessive. This is because recessive mutations are
+      removed far more efficiently once they reach appreciable frequencies and
+      are found in heterozygotes. Since our theory assumes additive effects, it
+      accommodates by estimating weaker selection coefficients.}
+
 \end{figure}
@@
 
+\subsection{Spatial Residuals}
+\label{supp:spatial-residuals}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/supplementary/spatial_residual.pdf}
+  \label{suppfig:spatial-residuals}
+\end{figure}
+
 \clearpage
@@
 \subsection{Chromosome Fit Figures}
+\label{supp:chrom-fit}
 

### 2023-07-06 manuscript/manuscript.tex
ms changes

@@
 
-\section*{Results}
+\section*{Results: Confirmation of Theory}
 
@@
 
-\subsection*{Application to Humans and Model Comparison}
+\section*{Results: Application to Human Genomic Data}
+
+\subsection*{Annotation Model Comparison}
 
@@
 effects for each feature type. We consider two classes of annotation models:
-(1) CADD score models, which assign a rank-order pathogenicity score to each
-basepair, and (2) and more interpretable gene feature-based models that
-includes protein coding regions, intron and UTRs, and PhastCons regions. We
-include PhastCons regions because there is abundant evidence of
-highly-conserved non-coding regions in humans under negative selection
-\parencite{Meader2010-hm,Harmston2013-tt,Katzman2007-gq,Siepel2005-wh}, but
-would not be included in gene-based annotation. Since our method fits a DFE for
-each feature type, PhastCons regions that overlap other gene annotations (e.g.
-coding sequences) must be assigned to either of the two feature types. Since
-this can impact estimates, we fit both alternatives: a PhastCons Priority
-model, where genic features that overlap PhastCons regions are classified as
-PhastCons, and a Feature Priority model, where gene structure features take
-priority and the PhastCons type catches all highly-conserved non-genic regions.
+(1) CADD-based models, which consider the top $x\%$ of basepairs according to
+the CADD pathogenicity score assigned to each basepair, and (2) and more
+interpretable gene feature-based models that includes protein coding regions,
+intron and UTRs, and PhastCons regions. We include PhastCons regions because
+there is strong evidence of highly-conserved non-coding regions
+\parencite{Meader2010-hm,Harmston2013-tt,Katzman2007-gq,Siepel2005-wh}, that
+would be missed by gene feature only annotation. These two classes of
+annotation models have a trade-off between fine-scaled specificity to which
+basepairs are likely to be under negative selection, and interpretability of
+the DFE estimates for each feature.
+
+Our method estimates the distribution of fitness effects (DFE) for each feature
+class. While CADD-based models only have a single conserved feature class (e.g.
+CADD 6\%), feature-based models can have multiple feature classes under varying
+levels of selective constrain. However, overlapping features (e.g. a basepair
+that is annotated as both PhastCons and coding sequence) must be assigned to
+one category or the other. Since this assignment impact DFE estimates, we fit
+both of the two alternatives. First, a \emph{PhastCons Priority} model, where
+genic features that overlap PhastCons regions are classified as PhastCons, and
+all remaining coding basepairs are labeled as CDS. Second, a \emph{Feature
+Priority} model, where all coding basepairs are assigned to CDS, and the
+PhastCons class catches the remaining highly-conserved non-genic regions.
+
 In total, we fit four annotation models (CADD 6\%, CADD 8\%, PhastCons
@@
 diversity at the megabase scale predicted by our model. We experimented with a
-few smaller spatial scales, but our results were consistent with previous
-results suggesting the human linked selection signal fits best at the megabase
-scale \parencite{Murphy2022-sj}.
+few smaller spatial scales (e.g. 100 kbp), but our results were consistent with
+previous results suggesting the human linked selection signal fits best at the
+megabase scale \parencite{Murphy2022-sj}.
 
@@
 Overall, we find the PhastCons Priority and CADD 6\% models fit about equally
-as well (Figure \ref{fig:figure-2}A), explaining $R^2=68.45$\% and
-$R^2=68.04$\% of the out-sample variance in Yoruba pairwise diversity at the
-megabase-scale, respectively. Generally, our models predict out-sample
-diversity levels slightly better than previous methods; the best-fitting CADD
-6\% model from \textcite{Murphy2022-sj} explained 60\% of diversity in left-out
-2 Mbp windows across YRI samples.
+as well (Figure \ref{fig:figure-2}A), consistent with recent work using classic
+BGS theory \parencite{Murphy2022-sj}. However, we find that our models predict
+out-sample diversity levels slightly better than previous methods. For these
+two models, we find that our B' method predicts $R_\text{LOCO}^2=68.45$\% and
+$R_\text{LOCO}^2=68.04$\% of the out-sample variance in Yoruba pairwise
+diversity at the megabase-scale, respectively. By contrast, the best-fitting
+CADD 6\% model from \textcite{Murphy2022-sj} explained 60\% of diversity in
+left-out 2 Mbp windows across YRI samples. We note that this might be explained
+by other differences in data process, reference genome, etc.
 
@@
 PhastCons Priority model has more DFE parameters, and thus can fit finer-scale
-differences with the four feature types. By contrast, the CADD 6\% track must
+differences among the four feature types. By contrast, the CADD 6\% track must
 fit patterns of diversity to conserved sites under a single DFE. The CADD 8\%
@@
 classic BGS theory, it should in principle fit equally well when an annotation
-model includes regions that are under no selective constrain and neutrally
-evolving. To test this, for each annotation model (which are ``sparse") we fit
-a corresponding ``full" model that assigns the remaining complement of the
-genome as a feature called ``other". This category contains most of the genome,
-and is likely to be predominantly neutrally evolving. Ideally, the
-goodness-of-fit of model using our B' method should be invariant to whether an
-annotation model is sparse or full. Indeed, we find this to be the case: both
-in-sample $R^2$ and out-sample $R_\text{LOCO}^2$ values are nearly identical
-across full and sparse-track models (Figures \ref{fig:figure-2}A and
+model includes regions that are under no selective constraint and thus
+neutrally evolving. To test this, for each of our annotation models (which are
+``sparse") we fit a corresponding ``full" model that assigns the remaining
+complement of the genome to a feature called ``other" that contains most of the
+genome. Ideally, how a model fits using our B' method should be invariant to
+whether an annotation model is sparse or full, since our method in principle
+can accommodate weak selection and neutrality. Indeed, we find this to be the
+case: both in-sample $R^2$ and out-sample $R_\text{LOCO}^2$ values are nearly
+identical across full and sparse-track models (Figures \ref{fig:figure-2}A and
 \ref{fig:figure-2}B, round points).
 
-By contrast, full annotation models fit poorly when using classic BGS theory
-(Figure \ref{fig:figure-2}B, diamond-shaped points) and lead to unreasonable
-parameter estimates (discussed below). Additionally, when sparse annotation
+By contrast, full annotation models fit poorly under classic BGS theory (Figure
+\ref{fig:figure-2}B, diamond-shaped points) and lead to unreasonable parameter
+estimates (discussed in Section XXX). Additionally, when sparse annotation
 models contain genomic features that are likely under weak constraint (such as
@@
 goodness-of-fit is nearly identical between B' and classic BGS methods. This
-behavior is what we would expect given that the CADD annotation models contain
-only the most pathogenic sites, which are \emph{a priori} very likely under the
-strong selection domain under which B' and classic BGS theory agree.
+behavior is what we would expect given that the CADD models contain only the
+most pathogenic sites, which are \emph{a priori} very likely under the strong
+selection domain under which B' and classic BGS theory agree.
 
@@
 by the expected neutral coalescence noise around the expected coalescence time
-$2B(b)N$ (see Method XXX). Thus, we define $R_\text{coal}^2$ as the theoretic
-proportion of variance explainable under this model to ballpark the variance
-capable of being explained.
+$2B(b)N$. This can be be found analytically, plugging-in our predictions for
+$B(b)$. This allows us to calculate $R_\text{coal}^2$ to ballpark the theoretic
+variance that is capable of being explained, assuming this coalescence-only
+noise process (see Method XXX). We note that selection is expected to
+\emph{decrease} the variance in coalescence times beyond a rescaled effective
+population size implies; our $R_\text{coal}^2$ would be an underestimate under
+selection.
 
@@
 ($R_\text{LOCO}^2 \approx 68$\%) is slightly above the theoretic
-$R_\text{coal}^2 \approx 67$\%. This suggests our model is in the vicinity of
-fitting all possible signal possible under these simplifying assumption. It
-should be noted that selection is expected to \emph{decrease} the variance in
-coalescence times beyond simply rescaling effective population size implies;
-better future models for $R_\text{coal}^2$ could show there is more linked
-selection signal to explain.
-
-For bottlenecked out-of-Africa populations, we find a larger discrepancy
-between $R_\text{coal}^2$ and observed out-sample $R_\text{LOCO}^2$. The
-theoretic $R_\text{coal}^2 \approx 64$\% for both populations, compared to the
-observed $R_\text{LOCO}^2 \approx 62$\% for CEU and $R_\text{LOCO}^2 \approx
-59$\% for CHB. Given that bottlenecks would act to increase the residual
-variance in coalescence times beyond the level implied by the effective
-population size, the gap would likely shrink under more realistic models or
-simulation-based approximations for $R_\text{coal}^2$. Overall, this suggests
-that negative selection models explain the vast majority coalescence time
-variation at the megabase-scale that is capable of being explained (i.e. that
-is not coalescence noise).
-
-\subsection*{Estimates of the Distribution of Fitness Effects in Humans}
+$R_\text{coal}^2 \approx 67$\%. This suggests our model is close to the
+vicinity of fitting all the signal possible, under the coalescence-only noise
+assumption. By contrast, for bottlenecked out-of-Africa populations, we find a
+larger discrepancy between $R_\text{coal}^2$ and observed out-sample
+$R_\text{LOCO}^2$. The theoretic $R_\text{coal}^2 \approx 64$\% for both
+populations, compared to the observed $R_\text{LOCO}^2 \approx 62$\% for CEU
+and $R_\text{LOCO}^2 \approx 59$\% for CHB. Given that bottlenecks would act to
+increase the residual variance in coalescence times beyond the level implied by
+the effective population size, the gap would likely shrink under more realistic
+models or simulation-based approximations for $R_\text{coal}^2$. Overall, this
+suggests that negative selection models explain the vast majority coalescence
+time variation at the megabase-scale that is capable of being explained (i.e.
+that is not coalescence noise).
+
+\subsection*{Estimated Distribution of Fitness Effects}
 
@@
 selection grid for each of the $K$ feature types. Given that the relationship
-between $\pi$ and the strength selection is U-shaped, we wondered whether the
-observed genome-wide patterns of diversity could be explained by weaker
-selection than found in previous methods based on classic BGS theory.
+between $\pi$ and the strength selection is U-shaped, our model in theory could
+fit under different combinations of parameters.
 
 However, across all of our annotation models, new deleterious mutations in
-likely conserved features are consistently estimated to have strongly
-deleterious effects (Figure \ref{fig:figure-3}A) consistent with previous work
+conserved feature classes were consistently estimated to have strongly
+deleterious effects (Figure \ref{fig:figure-3}A), consistent with previous work
 \parencite{McVicker2009-ax,Murphy2022-sj}. This includes features such as
@@
 selection coefficients of $s = 10^{-1}$ with 42\% chance, $s=10^{-2}$ with
-7.8\% chance, and $s=10^{-3}$ with 41\% chance. These estimates suggest a
-bimodal DFE (Figure \ref{fig:figure-3}A, second row, left column), and are
-mirrored across the other CADD models.
+7.8\% chance, and $s=10^{-3}$ with 41\% chance. This creates a bimodal DFE
+(Figure \ref{fig:figure-3}A, second row, left column), which is mirrored across
+the other CADD models. Moreover, these DFE estimates suggest a very high
+average selection coefficient across CADD 6\% regions of $\wisehat{\bar{s}}
+\approx 0.038$.
 
 Given that previous methods have typically not estimated the DFE up to
-selection coefficients of $s=10^{-1}$, we ran additional models using a DFE
-constrained to $s=10^{-2}$. Each of these constrained-DFE models across all
-populations fit the data less well than models run with the grid up to
-$s=10^{-1}$, explaining 1\% less of the leave-one-chromosome-out variation on
-average. As expected, the constrained-DFE models put more mass on $s=10^{-2}$
-(see Supplementary Table \ref{supp:tbl-r2}). Overall, this common behavior
-across models seems likely to reflect a non-identifiability issue between
-selective effects of $s=10^{-1}$ and $s=10^{-2}$, which we refer to as the
+selection coefficients of $s=10^{-1}$, we ran additional ``constrained grid"
+models using a DFE only up to $s=10^{-2}$. Across all populations, these
+constrained grid models fit the data less well than the default grid. On
+average, the constrained grid fits had $R_\text{LOCO}^2$ one percentage point
+lower (see Supplementary Table \ref{supp:tbl-r2}). As expected, the constrained
+grid models distributed mass onto the next strongest selection coefficient,
+$s=10^{-2}$. While the DFE could be truly bimodal for highly-conserved regions
+with a very high average selection coefficient, a likely alternative is that
+these results reflect a non-identifiability issue between selective effects of
+$s=10^{-1}$ and $s=10^{-2}$. We will refer to this as the
 \emph{strong-selection non-identifiability} issue.
@@
 generation, i.e. death of the carrier of the \emph{de novo} mutation. In this
-hypothetical limit, each occurrence of a very harmful mutation is like an
-instantaneous death and thus acts like a genome-wide reduced effective
-population size. In our model, this genome-wide effective population size is
-parameterized by $\pi_0$; thus, one line of evidence for this effect is a
-correlation between estimates of $\pi_0$ and average selection $\bar{s}$.
-Looking at the CADD 6\% models fit across the default and constrained DFE
-grids, we find this is the case. Across every population, average selection was
-about 7 times larger using the default grid, and $\pi_0$ was $5.6\%$ larger.
-There was no similar consistent change in mutation rate estimates among
-populations. In the CADD 6\% model, genome-wide average reduction factor
-$\bar{B}$ was $\approx 6.1\%$ lower in the default versus constrained grid.
-Overall, this suggests that the linked selection signal alone carries little
-information that can differentiate very strong selection from a slightly
-smaller drift-effective population size.
-
-Despite the strong-selection non-identifiability issue, we still were curious
-how stable these estimates are across samples from reference populations. In
-particular, we focused on the Feature Priority model, which fits the data less
-well but is more interpretable since its annotation is based on gene features.
-Overall, estimates of the distribution of selection coefficients are stable
-across populations and had narrow uncertainty as estimated by a block jackknife
-procedure (Figure \ref{figure-3}B). Coding sequences show the strong-selection
-non-identifiability issue, and an uncertain signal of a neutral mode in the
-DFE. This would be expected given the very different levels of selective
-constraint synonymous and non-synonymous sites in coding sequences are under.
-The PhastCons class of this Feature Priority model, which contains highly
-conserved non-coding elements (XXX), is under strong selective constraint as
-expected. Interestingly, in our PhastCons Priority annotation model (Figure
-\ref{fig:figure-3}A bottom row), the DFE estimate for coding sequences
-excluding PhastCons regions is neutral. This is expected; the selection signal
-in coding regions is absorbed by the PhastCons feature.
+hypothetical limit, each occurrence of a very harmful mutation nearly acts like
+an instantaneous death that in effect is just a smaller population size. One
+way to test this hypothesis is to look to see if there is a systematic positive
+relationship across models between average selection coefficient and $\pi_0$,
+which is includes the drift-effective population size $N_e$.
+
+We find this is the case for all of our CADD 6\% models. Across every
+population, average selection was about $7.1$ times larger using the default
+grid, and $\pi_0$ was $5.6\%$ larger. There was no similar consistent change in
+mutation rate estimates among populations. In the CADD 6\% model, genome-wide
+average reduction factor $\bar{B}$ was $\approx 6.1\%$ lower in the default
+versus constrained grid. Overall, this suggests that the linked selection
+signal alone cannot differentiate very strong selection from a slightly smaller
+drift-effective population size.
+
+Despite the strong-selection non-identifiability issue, we were curious how
+stable these estimates are across different reference populations. Overall, the
+DFE estimates were consistent over all populations and models (Supplementary
+Materials Section \ref{supp:fits}). Here, we highlight the Feature Priority
+model since though it is poor-fitting, the DFE estimates are more readily
+interpretable since it is based on gene features. Figure \ref{fig:figure-3}B
+shows consistent DFE estimates across populations, with narrow uncertainty as
+estimated by a block jackknife procedure. Coding sequences have a bimodal DFE
+consistent with the strong-selection non-identifiability issue. Additionally,
+coding sequences have a very uncertain neutral mode in the DFE. This is
+expected given the different level of selective constraint between synonymous
+and non-synonymous sites.
+
+Additionally, the PhastCons class of this Feature Priority model, which
+contains highly conserved non-coding elements, is under strong selective
+constraint as expected. Finally, we highlight one result from our PhastCons
+Priority annotation model (Figure \ref{fig:figure-3}A bottom row): the DFE
+estimate for coding sequences excluding PhastCons regions is estimated as
+neutral. This too is expected; the selection signal in coding regions is
+absorbed by the PhastCons feature, leaving only conditionally neutral sites.
 
@@
 
-Early genome-wide models of background selection closely fit megabase-scale
-patterns in sequence divergence among five primates
-\parencite{McVicker2009-ax}. However, the mutation rate estimates from this
-model were unreasonably high compared to pedigree-based estimates (XXX). This
-mutation rate discordance has led to the hypotheses that positive selection
-could explain some fraction of the genome-wide linked selection signal (e.g.
-\cite{Enard2014-kz}); however, evidence suggests that the role of classic
-sweeps in humans is minuscule
-\parencite{Pickrell2009-tt,Hernandez2011-gs,Murphy2022-sj}. Furthermore, Murphy
-et al. find the classic BGS model fit using the CADD 6\% track leads to
-mutation rate estimates that are consistent with pedigree-based rates. Given
-that our method is based on a different theoretic approach, we curious whether
-it would lead to estimates consistent with pedigree-based mutation rates.
-
-Across all populations, our mutation rate estimates from CADD-based models are
-roughly concordant with pedigree-based estimates (Figure \ref{fig:figure-4}A).
+Early genome-wide classic BGS models fit well, but lead to unusually high
+estimates of the mutation rate \parencite{McVicker2009-ax}. This lead to the
+hypothesis that these models could be absorbing the signal of positive
+selection \parencite{Enard2014-kz}, though other work has found a limited role
+for hitchhiking \parencite{Pickrell2009-tt,Hernandez2011-gs,Murphy2022}. Still,
+a central benchmark of fitting genome-wide models is that mutation rate
+estimates are in accordance with pedigree-based estimates (XXX).
+
+We find across all populations, our mutation rate estimates from CADD-based
+models are roughly concordant with pedigree-based estimates (Figure
+\ref{fig:figure-4}A), consistent with recent work \parencite{Murphy2022-sj}.
 Our full-track CADD 6\% model estimates the mutation rate as $\widehat{\mu} =
-1.46 \times 10^{-8}$ for YRI, $1.53 \times 10^{-8}$ for CEU, and $1.5 \times
+1.46 \times 10^{-8}$ for YRI, $1.53 \times 10^{-8}$ for CEU, and $1.51 \times
 10^{-8}$ for CHB reference samples (Supplementary Table \ref{supp:tbl-r2}). As
@@
 However, mutation rate estimates for feature-based annotation models show
-several issues. First, mutation rate estimates under using our implementation
-of classic BGS theory are an order of magnitude below the expected range
-(Figure \ref{fig:figure-4}A top row). We observe similar behavior when we use
-the classic BGS model to fit full-coverage annotation models (Figure
-\ref{fig:figure-4}A bottom row). This behavior is consistent with the fact that
-classic BGS theory cannot fit weak selection, and to accommodate features under
-weak constrain (e.g. introns, UTRs, and the ``other" feature) it must severally
-underestimate the mutation rate.
+several issues. First, mutation rate estimates under our implementation of
+classic BGS theory are an order of magnitude below the expected range (Figure
+\ref{fig:figure-4}A top row). We observe similar behavior when we use the
+classic BGS model to fit full-coverage annotation models (Figure
+\ref{fig:figure-4}A bottom row). This behavior is consistent with classic BGS
+theory being unable to fit the DFE to features under weak constraint (e.g.
+introns, UTRs, and the ``other" feature), and consequently severally
+underestimating the mutation rate.
 
@@
 since both classic BGS and SC16 models only depend on mutation rate through the
-compound parameter $\mu L$ where $L$ is the length of the conserved segment.
+compound parameter $\mu L$, where $L$ is the length of the conserved segment.
 Even though our method is much more robust to including non-conserved regions

### 2023-07-06 manuscript/manuscript.tex
more ms stuff

@@
 the other CADD models. Moreover, these DFE estimates suggest a very high
-average selection coefficient across CADD 6\% regions of $\wisehat{\bar{s}}
-\approx 0.038$.
+average selection coefficient across CADD 6\% regions of $\widehat{\bar{s}}
+\approx 0.043$ for YRI.
 
@@
 models using a DFE only up to $s=10^{-2}$. Across all populations, these
-constrained grid models fit the data less well than the default grid. On
-average, the constrained grid fits had $R_\text{LOCO}^2$ one percentage point
-lower (see Supplementary Table \ref{supp:tbl-r2}). As expected, the constrained
-grid models distributed mass onto the next strongest selection coefficient,
-$s=10^{-2}$. While the DFE could be truly bimodal for highly-conserved regions
-with a very high average selection coefficient, a likely alternative is that
-these results reflect a non-identifiability issue between selective effects of
+constrained grid models fit the data less well than the default grid by about
+one percentage point less in $R_\text{LOCO}^2$ (see Supplementary Table
+\ref{supp:tbl-r2}). As expected, the constrained grid models distributed mass
+onto the next strongest selection coefficient, $s=10^{-2}$, and the average
+selection was reduced about six-fold to $\widehat{\bar{s}} \approx 0.0079$ for
+YRI. While the DFE may truly be bimodal for highly-conserved regions and have a
+very high average selection coefficient, a likely alternative is that these
+results reflect a non-identifiability issue between selective effects of
 $s=10^{-1}$ and $s=10^{-2}$. We will refer to this as the
@@
 feature classes based on the priority of overlapping features. That one of
-these models is our best-fitting model, and the other our worst indicates a
-large role in per-basepair heterogeneity in selective effects in determining
-the genome-wide linked selection signal. CADD-based models fit well in part due
-to their fine-scale resolution of selective effects across the genome.
+these models is our best-fitting model, and the other our worst indicates the
+sensitivity these models have when features classes have heterogeneous DFEs.
+CADD-based models fit better in part due to their fine-scale resolution of
+selective effects across the genome. While ideally we would fit a CADD model
+with different features corresponding to the different percentiles of
+pathogenicity, these features are on the basepair scale and thus too
+memory-intensive for our method to currently accommodate.
 
@@
 Given these the sensitivity of mutation rate estimates and the strong-selection
-non-identifiability issue, we next sought to assess the estimation accuracy of
-our composite likelihood method using forward simulations. In order to have
-enough megabase-scale data to apply our method, we independently evolved the
-first five human chromosomes. Due to the immense computational expense of
-conducting forward simulations at this scale over numerous parameter
-combinations, we used fixed mutation and selection parameters in putatively
-conserved regions (see Methods XXX). We note three important findings from
-these simulations.
+non-identifiability issue, we next sought to assess the estimation overall
+accuracy of our composite likelihood method using forward simulations. In order
+to have enough megabase-scale data to apply our method, we independently
+evolved the first five human chromosomes and combined them into a ``synthetic"
+genome. Due to the immense computational expense of conducting forward
+simulations at this scale over numerous parameter combinations, we used fixed
+mutation and selection parameters in putatively conserved regions (see Methods
+XXX). We note three important findings from these simulations.
 
@@
 selection against new deleterious mutations and mutation rate (Figure
-\ref{fig:figure-4}B, top row). This is consistent across both classic BGS and
-our B' method; the only exception is the classic BGS model run with a
-full-coverage track for the reasons described above. However, if new mutations
-are highly deleterious ($s=0.05$), $R^2$ is reduced. This is consistent with
-very strong selection having less localized effects and acting more like a
-genome-wide faster rate of drift.
+\ref{fig:figure-4}B, top row). This shows the ``strength" of the linked
+selection signal varies as expected with these key parameters. However, if new
+mutations are highly deleterious ($s=0.05$), $R^2$ is reduced. This is
+consistent with very strong selection having less localized effects and acting
+more like a genome-wide faster rate of drift.
 
@@
 middle row). However, when selection was weak, the classic BGS model
-erroneously estimated strong selection and a very low mutation rate. By
+erroneously estimates strong selection and a very low mutation rate. By
 contrast, our B' method estimated selection coefficients much closer to their
@@
 mutation rate estimates than classic BGS theory across a variety of selection
-coefficients. When BGS is not acting, either due to weak selection or a low
+coefficients. These biases in mutation rate estimates suggest that benchmarking
+genome-wide models via their concordance with pedigree-based rate estimates may
+not be suitable. When BGS is not acting, either due to weak selection or a low
 rate of deleterious mutations (Figure \ref{fig:figure-4}B right column), all
-estimates are worse. However, we note that this is an unlikely region of
-parameter space and this issue is readily diagnosable from the poor model fit 
-as evident by low $R^2$ values.
+estimates are worse. This makes intuitive sense; the overall linked selection
+signal is weaker relative to drift-based noise. However, we note that this is
+an unlikely region of parameter space and this issue is readily diagnosable
+from the low $R^2$ values.
 
@@
 YRI and bottlenecked out-of-Africa (CEU and CHB) samples, we tested whether a
-population expansion would bias our method. We expansion rates like those
-estimated for Europeans by \textcite{Gutenkunst2009-pg} had little effect on
-parameter estimates (Supplementary Section \ref{supp:sim-assum}). By contrast,
-we find our estimates are downwardly biased when mutational effects are weak
-and recessive, but are accurate under recessive strong selection. This occurs
-since the fixation probabilities of recessive mutations are in reality much
-lower than they are under our model assuming additivity.
+population expansion would bias our method. We use an expansion rate and timing
+based on those estimated for Europeans by \textcite{Gutenkunst2009-pg}, finding
+they had little effect on parameter estimates (Supplementary Section
+\ref{supp:sim-assum}). By contrast, we find our estimates are downwardly biased
+when mutational effects are weak and recessive, but are accurate under
+recessive strong selection. This occurs because the fixation probabilities of
+recessive mutations are in reality much lower than under our model that assumes
+additivity.
 
@@
     \includegraphics[width=\textwidth]{figures/figure_5.pdf} 
-    \caption{}
+    \caption{(A) CADD 6\% full-track }
   \label{fig:figure-5}
@@
 this region has been found by a genome-wide scan for associative overdominance
-\parencite{Gilbert2020-aw}.  For model XXX
+\parencite{Gilbert2020-aw}.
 

### 2023-07-06 manuscript/supplementary.tex
ms changes
sup tbl

@@
 
+\subsection{Strong Selection Non-Identifiability}
+\label{supp:}
+
+\begin{figure}[htbp]
+    \label{suppfig:sel-ident}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/supplementary/ave_sel_vs_pi0.pdf}
+  \caption{XXX}
+\end{figure}
+
 \subsection{Demographic Expansion and Recessive Simulation Tests}
@@
 
+\subsection{Average Selection Coefficients}
+
+There are the average selection coefficients ($\bar{s}$) calculated from the
+estimated DFEs, across both default and constrained grids.
+
+\begin{tabular}{|l|llrr|}
+\hline
+Annotation Model &  Population &    Feature &   ${\bar{s}}$ (default grid) & ${\bar{s}}$ (constrained grid) \\
+\hline
+ PhastCons Priority &  CEU &        CDS &  0.100000 &  0.000000 \\
+ PhastCons Priority &  CEU &       gene &  0.005308 &  0.000226 \\
+ PhastCons Priority &  CEU &  PhastCons &  0.008076 &  0.006275 \\
+ PhastCons Priority &  CHB &        CDS &  0.100000 &  0.000000 \\
+ PhastCons Priority &  CHB &       gene &  0.004896 &  0.000296 \\
+ PhastCons Priority &  CHB &  PhastCons &  0.001327 &  0.005886 \\
+ PhastCons Priority &  YRI &        CDS &  0.000000 &  0.000000 \\
+ PhastCons Priority &  YRI &       gene &  0.005556 &  0.000160 \\
+ PhastCons Priority &  YRI &  PhastCons &  0.028642 &  0.007575 \\
+   Feature Priority &  CEU &        CDS &  0.075041 &  0.005281 \\
+   Feature Priority &  CEU &       gene &  0.001654 &  0.000315 \\
+   Feature Priority &  CEU &  PhastCons &  0.069563 &  0.008767 \\
+   Feature Priority &  CHB &        CDS &  0.074797 &  0.005211 \\
+   Feature Priority &  CHB &       gene &  0.001974 &  0.000271 \\
+   Feature Priority &  CHB &  PhastCons &  0.051988 &  0.008507 \\
+   Feature Priority &  YRI &        CDS &  0.054503 &  0.001576 \\
+   Feature Priority &  YRI &       gene &  0.001912 &  0.000336 \\
+   Feature Priority &  YRI &  PhastCons &  0.055153 &  0.009369 \\
+            CADD 6\% &  CEU &      CADD 6\% &  0.042458 &  0.006500 \\
+            CADD 6\% &  CHB &      CADD 6\% &  0.031565 &  0.005733 \\
+            CADD 6\% &  YRI &      CADD 6\% &  0.043671 &  0.007926 \\
+            CADD 8\% &  CEU &      CADD 8\% &  0.041254 &       - \\
+            CADD 8\% &  CHB &      CADD 8\% &  0.029052 &       - \\
+            CADD 8\% &  YRI &      CADD 8\% &  0.046212 &       - \\
+\hline
+\end{tabular}
+
 \subsection{Residual Covariate Diagnostics}

### 2023-07-07 manuscript/manuscript.tex
paper changes

@@
                 hyperref=true]{biblatex}
-\usepackage{color}
-\usepackage{nicefrac}
+\RequirePackage{color}
+\RequirePackage{nicefrac}
+
+\addbibresource{biblio.bib}
 
@@
 
-\addbibresource{biblio.bib}
-
 %\title{Towards a Complete Model of the Negative Selection Process in Humans}
 %\title{An Evolutionary Quantitative Genetic Model of the Negative Selection Process in Humans}
-\title{A Quantitative Genetic Linked Selection Model fit to the Negative Selection Process in Humans}
+\title{A Quantitative Genetic Linked Selection Model of the Negative Selection Process in Humans}
 
@@
 % TODO: emphasize connection to nearly neutral theory
+
+    % uniting quant gen/fitness variation, linked selection, and substitution
+    % The interesting thing about connecting all these models is that we might
+    % have hope of calibrating these models to the true process.
 \end{abstract}
@@
 the other CADD models. Moreover, these DFE estimates suggest a very high
-average selection coefficient across CADD 6\% regions of $\widehat{\bar{s}}
+average selection coefficient across CADD 6\% regions of ${\bar{s}}
 \approx 0.043$ for YRI.
@@
 models using a DFE only up to $s=10^{-2}$. Across all populations, these
-constrained grid models fit the data less well than the default grid by about
-one percentage point less in $R_\text{LOCO}^2$ (see Supplementary Table
-\ref{supp:tbl-r2}). As expected, the constrained grid models distributed mass
-onto the next strongest selection coefficient, $s=10^{-2}$, and the average
-selection was reduced about six-fold to $\widehat{\bar{s}} \approx 0.0079$ for
-YRI. While the DFE may truly be bimodal for highly-conserved regions and have a
-very high average selection coefficient, a likely alternative is that these
-results reflect a non-identifiability issue between selective effects of
-$s=10^{-1}$ and $s=10^{-2}$. We will refer to this as the
-\emph{strong-selection non-identifiability} issue.
+constrained grid models fit the data slightly less well than the default grid
+by about one percentage point less in $R_\text{LOCO}^2$ (see Supplementary
+Table \ref{supp:tbl-r2}). As expected, the constrained grid models distributed
+mass onto the next strongest selection coefficient, $s=10^{-2}$, and the
+average selection was reduced about six-fold to ${\bar{s}} \approx 0.0079$ for
+YRI. Our constrained-grid DFE estimates for the Feature Priority model are
+qualitatively consistent with the U-shaped DFEs found for amino acids through
+Poisson Random Field method \parencite{Boyko2008-tj}. While the DFE may truly
+be bimodal for highly-conserved regions and have a very high average selection
+coefficient, a likely alternative is that these results reflect a
+non-identifiability issue between selective effects of $s=10^{-1}$ and
+$s=10^{-2}$. We will refer to this as the \emph{strong-selection
+non-identifiability} issue.
 
@@
     \includegraphics[width=\textwidth]{figures/figure_5.pdf}
-    \caption{(A) CADD 6\% full-track }
+
+    \caption{(A) Observed and predicted diversity under the CADD 6\% full-track
+        model. Once scaled by average diversity, populations differ little in
+        their predicted diversity levels (colors). Additionally, we show
+        summaries of CADD density and recombination rate along the chromosome
+        below. (B) Predicted $B$ and observed $\pi$ for each window. The
+        residual variance is in the vicinity of the levels expected from the
+        $R_\text{LOCO}^2$ calculation. (C) Evidence that negative selection
+    signal remains in CADD 6\% residuals (YRI shown). Residuals are plotted
+against the average LoF selection coefficient across genes in a megabase window
+(estimated from \cite{Agarwal2023-un}).}
+
   \label{fig:figure-5}
@@
 against observed diversity along chromosomes (see Supplementary Materials
-Section \ref{supp:chrom-fit}) we overall find close correspondence consistent
-with the high $R_\text{LOCO}^2$. However, we find a few large (tens of
-megabases) regions with systematically poorer fit (Supplementary Materials
-Figure \ref{supp:spatial-residuals}). In Figure \ref{fig:figure-5}A we see one
-such region from XXX to XXX. Interestingly, predicted diversity closely follows
-the peaks and troughs of this region. We note that a smaller segment within
-this region has been found by a genome-wide scan for associative overdominance
-\parencite{Gilbert2020-aw}.
+Section \ref{supp:chrom-fit}) we find a close correspondence consistent with
+the high $R_\text{LOCO}^2$. However, we find a few large (tens of megabases)
+regions with systematically poorer fit (Supplementary Materials Figure
+\ref{supp:spatial-residuals}). In Figure \ref{fig:figure-5}A we see one such
+region on the short arm of chromosome 2, from 30 Mbp to 60Mbp. Interestingly,
+predicted diversity closely follows the peaks and troughs of this region. We
+note that a small region within this stretch had been found by a genome-wide
+scan for associative overdominance \parencite{Gilbert2020-aw}.
 
@@
 
-\subsection*{Predicted Substitution Rates, and Fitness Variance}
+Earlier, we found that the goodness-of-fit and DFE estimates of our
+feature-based annotation models are sensitive to how basepairs are grouped into
+feature classes. Our models are in essence fitting a single distribution across
+sites within a feature class, yet there may be local heterogeneity among sites
+that is poorly captured by this approach. We investigated this possibility by
+looking for lingering selection signal in our model residuals. First, we
+inspected whether there was a relationship between the fraction of CADD 2\% and
+6\% basepairs and the residual across megabase windows (Supplementary Figure
+\label{suppfig:resid-cadd}), finding a negative significant relationship in
+both cases. Moreover, our model over-predicted diversity in windows containing
+more CADD 2\% basepairs moreso than CADD 6\%, consistent with heterogeneity in
+site pathogenicity being poorly fit by our model. However, the total residual
+variance explained is $R^2 = 0.4\%$ and $R^2 = 0.8\%$ for the CADD 2\% and 6\%
+tracks respectively, suggesting only modest amount of selection signal remains.
+
+We further tested for remaining selection signal in our CADD 6\% model
+residuals by using gene-specific estimates of the fitness cost of
+loss-of-function (LoF) mutations from \textcite{Agarwal2023-un}. These
+estimates are based on an Approximate Bayesian Computation approach that
+estimates the posterior distribution over LoF fitness costs from the observed
+dearth of LoF mutations per gene, and thus is an independent methodological
+approach than ours. We averaged the estimated LoF fitness costs across genes
+for each of our megabase windows, and plotted our residuals against these
+average LoF fitness costs. Contrary to the weak CADD residual signal described
+above, we find evidence of a fairly strong relationship between our residuals
+and average LoF fitness cost (Figure \ref{fig:figure-5}C; $R^2 = 2.1\%$,
+p-value $1.27 \times 10^{-10}$). In other words, roughly 2\% of the variance in
+these residuals is explained by the average fitness costs of LoF mutations in
+the window. Consequently, our model over-predicts diversity by about
+$\nicefrac{\sigma}{2}$ or more in windows harboring the top 1.7\% most
+LoF-intolerant genes.
+
+\subsection*{Predicted Substitution Rate and Additive Fitness Variance}
 
@@
 
-    \caption{Evidence that negative selection signal remains. (A) Residuals
-        from the YRI CADD 6\% model plotted against the average LoF selection
-        coefficient across genes in a megabase window (estimated from
-        \cite{Agarwal2023-un}). Our model is systematically over predicting
-        diversity in regions estimated to be sensitive to LoF variants. The
-        ordinary least squares (OLS) estimate suggests that 2.1\% of the
-        variance in residuals can be explained by reductions caused by the
-        local sensitivity to LoF mutations. (B) Residuals from the YRI CADD 6\%
-        model plotted against the fraction of CADD 6\% basepairs in the
-        megabase window. Here, only 0.8\% of the variance is explained by
-        remaining spatial variance in CADD 6\% regions (blue). Overlaid is the CADD
-        2\% OLS curve (orange). }
+    \caption{}
 
@@
 selection that explicitly incorporates the substitution process, fitting our
-model to genomic data leads to predictions that can be explicitly tested. In
-particular, there are two key predicted quantities: (1) the substitution rates
-per segment, and (3) the fitness variance per segment.
-
-Each of our maximum likelihood estimates $\widehat{\Psi}$ imply other
-corresponding
-
-Our method accommodates weak selection by modeling the fitness variance in each
-segment under an equilibrium negative selection process. This fitness variance
-is found by simultaneously solving for the per-generation substitution rate $R$
-and the draft-effective population size $N_d$ for each segment. When our model
-is fit to genome-wide diversity data, the parameter estimates imply predictions
-of the deleterious substitution rate per basepair, per generation, for each
-feature. Our method only considers within-species pairwise diversity, so these
-predicted substitution rates constitute a test that our model is making
-reasonable out-sample predictions of the substitution rates across features.
-
-We focus specifically on the predictions from our feature-based annotation
-models, since we can readily estimate human substitution distances for each
-feature to compare to our predictions.
-
-The estimated substitution distance should in theory equal the product of the
-substitution rate per site, per year and the branch length in years.
-
-When comparing our predicted mutation rates to observed substitution distances,
-we must account for three major sources of uncertainty.
-
-First, the split time between humans and chimpanzees is unknown, with estimates
-ranging from 6Mya to 12Mya (XXX). Second, our predictions are in sites per
-generation, and the appropriate human generation time is unknown. Third, the
-appropriate mutation rate to use is debated, since there may have been a
-mutational slow down on the human lineage (XXX), which could explain why
-phylogenetic- and pedigree-based mutation rate estimates differ by roughly a
-factor of two (XXX). Finally, we are only considering modeling the deleterious
-substitution rate. In reality, some fraction of observed substitutions fixed
-because they were beneficial.
-
-For a particular feature $f$, the estimated per-site substitution distance
-$S_f$ along a branch (which accounts for back mutations) should correspond to
-$S_f = T_g r_f$, where $T_g$ is the time in generations and $r_f$ is the
-substitution rate per basepair, per generation of that feature.
-
-Each feature is
-composed of basepairs with varying selection coefficients distributed according
-to the feature-specific DFE vector $\mathbf{w}_f$ and corresponding vector of
-substitution rates per selection coefficient, $\mathbf{r}_f$.
-
-Our estimate of a feature's DFE
-is $\widehat{\mathbf{w}}_f$.
-
-We calculate the
-
-\begin{align}
-    \label{eq:sub}
-    r_f = \mu \int_0^1 p_F(s) ds
-\end{align}
-
-Solving the XXX equations, we end up with predictions $\widehat{r}_f$ for the
-per-basepair, per-generation substitution rate for each feature. These are
-calibrated to the MLE mutation rate estimate $\widehat{\mu}_\text{del}$, which
-as mentioned, excludes the contribution of lethal mutations which act purely to
-reduce $N_e$ and thus only impact $\pi_0$. Using Equation \eqref{eq:sub} we can
-re-calibrate these estimates against other mutation rates.
-
-Because of it's greater interpretabilty, we initially focus on the substitution
-rate estimates from the Feature Priority model even though it had a poorer-fit
-and high mutation rate estimate.
-
-We can also re-calibrate this model to an alternate mutation rate.
-
-% - Table s5 https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/genetics/206/1/10.1534_genetics.116.197145/8/files1.pdf?Expires=1686357613&Signature=cp9qoI2ChCb6GFZ74HI4eWcA~MQ4sHlWyeI3Zp49-5xEORzMfYlSRxzVvZlo8ZoGWDi0G3PnL00ZplM-~or4YIVm2uhkmqjf61AmC7RdZ1QTak4j0kFMQ0wAbM9pusiV3gIepTQHq2HKRKwgS~ypmv~0oqO5dDxaq9nhOWIBBmLAA49oHWuxUtnRe3sEYATwCwjIFEomhgJpg7vKGCF~87GdCV4JETW4PN2VvD4hsu2AVRd8yA~vUJvsXfhWwcElDeY4fsgpprl1lC3XGjJCnqdwbzRGPV5LNx5XEKLptcZtLeLcOmFJCvmVLe0bewJ7Vta0TSQM7T9AuklompjPPw__&Key-Pair-Id=APKAIE5G5CRDK6RD3PGA
-
-Like other methods, our method divides the putatively conserved basepairs up
-
-Generally, when the putatively conserved region is larger, there is a tradeoff
-in two dimensions. First, more bases are conserved according to the same DFE,
-so the total mutational load increases. This causes the tradeoff between
-mutation rate estimate and CADD fraction we observe across both our B' model
-and the classic B when using sparse tracks (Figure 2XXX).
-
-We find that mutation rate estimates are fairly sensitive to model choice
-(Figure 2XXX).
-
-Since B' has a U-shaped relationship with the selection coefficient, we
-hypothesized that our model may reveal true identifiability issues in
-estimating the DFE that do not occur under the classic B model. We find that
-this is not the case for our models fit at the megabase-scale. The DFE
-estimates for our CADD6 and CADD8 models are consistent (Figure XXX), and both
-produce similar mutation rates.
-
-While we do not comprehensively fit all of our model to other spatial scales,
-we
-
-However, while a given reduction level may be explained by either a weak or
-strong selection coefficient, the spatial extent of these reductions caused by
-weak selection is much narrower than under strong selection.
-
-We see evidence this is the case in several of our models, where
-genome-wide diversity fits equally well between models with mutation rates
-
-Indeed, we see this by comparing parameter estimates of the CADD6
-sparse-track model using classic BGS B and our B' method.
-
-(which mimics the best-fitting model of
-\cite{Murphy2022-sj})
-
-(see also
-Appendix 1, Figure 26A of \cite{Murphy2022-sj})
-
-correlation between s and feature
-
-Overall, we find evidence that negative selection models are very sensitive to
-the input track. Because inference based on our B' method accommodates weaker
-selection, the primary models we fit here more robust to misspecification of
-the regions
+model to genomic data leads to predictions: (1) the substitution rates per
+segment, (2) the fitness variance per segment, and (3) the total number of
+strongly deleterious mutations per diploid genome per generation. Given
+predicted substitution rates for each feature can be compared to estimated
+sequence divergence between species, this offers a qualitative check our model
+is producing sensible results.
+
+To compare our models' predicted substitution rates against observed
+divergence, we estimated sequence divergence on the human lineage using a
+multiple alignment of five primates for each feature in our feature-based
+models (Methods XXX). Since we have observed that mutation rate estimates from
+our method can be unreliable (and simulation results confirm this), we
+predicted substitution rates under a fixed mutation rate of $\mu = 1.5 \times
+10^{-8}$, rather than the maximum likelihood estimates. Unfortunately, a
+careful comparison between our predictions and observed divergence rates is
+hindered by considerable uncertainty in the rate of the molecular clock,
+generation times, and the human-chimpanzee divergence time. We assume a
+generation time of 29 years, and calculate the sequence divergence implied by
+our predicted substitution rates over a range of divergence times, from 6 Mya
+to 12 Mya \parencite{Moorjani2016-tb,Nachman2000-te,Yi2002-pw,Steiper2006-xx}.
+
+Our predictions are qualitatively consistent with the observed divergence along
+the human lineage (Figure \ref{fig:figure-6}). As expected, the predicted
+substitution rate in features unlikely to be under selective constraint is very
+close to the mutation rate. Throughout, we report our substitution rates as a
+percent relative to the total mutation rate, $\mu$ (here fixed to $1.5 \times
+10^{-8}$). In our PhastCons Priority model, introns and UTRs are estimated to
+have have substitution rates that are $93.6\%$ of the mutation rate, and the
+``other" feature is estimated at $97.8\%$. When we use the constrained grid,
+these change to $97.85\%$ and $99.98\%$ respectively. By contrast, our
+predicted substitution rates for PhastCons regions are much lower: $0.48\%$ for
+the default grid, and $1.65\%$ for the constrained grid. This is likely due to
+the constrained grid placing more of the DFE mass on weaker selection
+coefficients, which in turn have a higher chance of drifting up to fixation.
+
+Finally, we entertain two more speculative predictions. First, the equilibrium
+additive genic fitness variance. Our PhastCons Priority model DFE estimates and
+negative selection model imply a $\widehat{V}_a = 0.45$ for the default grid,
+and $\widehat{V}_a = 0.040$ for the constrained grid. This is a considerable
+level of uncertainty---over an order of magnitude---due to a difference in the
+selection grid. There are no other known direct estimates in humans of this
+quantity for comparison. When $2N \gg 1$, theory suggests $V_a \approx
+U_D\bar{s}$ where $U_D$ is the diploid rate of strongly deleterious mutations.
+Thus, the uncertainty is a byproduct of the different average estimates of
+$\bar{s}$ across the alternate grids. Since both this work and
+\textcite{Murphy2022-sj} find the single-feature CADD 6\% track fits well, we
+can use this to gain intuition through some rough calculations. In this case,
+$U_D \approx 6\% \times 2 \times 3 \times 10^{9} \mu \approx 3.6 \times 10^{8}
+\mu$. For $\mu \approx 1.5 \times 10^{-8}$, $U_D \approx 5.2$. Under the
+constrained grid, ${\bar{s}} \approx 0.0079$ and thus $V_a \approx 0.043$; with
+the default grid $\bar{s} \approx 0.043$, and $V_a \approx 0.23$. Often fitness
+reductions of $\approx 1\%$ are assumed for the most highly-conserved regions
+(XXX, Keightly, Lynch); for comparison, this leads to $V_a \approx 0.054$.
+
+XXX HERE  USE CADD PREDICTIONS
+
+By theory, this relative fitness (scaled by the mean squared) is
+bounded from above by the Index of Selection, defined as the theoretic relative
+fitness variance if fitness were perfectly heritable \parencite{Crow1958-pc}.
+Based on rough calculation by Crow, this could range between 0.982-3.286 in
+humans. However, this is based on modern-day mortality figures that may not
+extrapolate to the timescales that determine genome-wide patterns of diversity.
+
+Second, we can calculate the reduction in relative fitness per generation due
+to deleterious substitutions. Since \textcite{Muller1950-ta}, there has been
+considerable interest in whether mildly deleterious mutations, perhaps due to
+the relaxation of selection pressures, could accumulate and fix at a faster
+pace in humans.
+
+Here, the selection grid choice has little impact
+on the estimates: we calculate that across the genome, the loss in relative
+fitness per generation is 0.00160\% for the default grid, and 0.00119\% for the
+constrained grid.
 
 \section*{Discussion}
+
+- parametric DFEs are trash
 

### 2023-07-07 manuscript/supplementary.tex
paper changes

@@
 
-The full list of samples is available in TSV format in the GitHub repository
-(\path{data/h1kg/yri_samples.tsv}).
-
-XXX
-
 \subsection{Filtering gVCFs}
@@
 
-gVCFs were filtered using a custom Python tool, \texttt{gvcf2counts.py} (in
+The gVCFs were filtered using a custom Python tool, \texttt{gvcf2counts.py} (in
 \path{tools/gvcf2counts.py}), which reads the gVCFs, filters them according to
@@
 
-Generally, fitting procedures act to minimize the squared deviation between
-predictors and observed values, which drives $\cov_i(\varepsilon_i,
-\widehat{B}_i)$ to zero XXX. \textcite{Murphy2022-sj} suggest  XXX
-
 Thus, our $R^2$ depends on (1) the bias of our predictors, (2) the covariance
@@
 
-\begin{figure}[htbp] \centering
-    \includegraphics[width=0.5\textwidth]{figures/supplementary/subfigure_R2_scale.pdf}
-    \caption{XXX}
-  \label{suppfig:coal_r2_scale}
-\end{figure}
+% \begin{figure}[htbp] \centering
+%     \includegraphics[width=0.5\textwidth]{figures/supplementary/subfigure_R2_scale.pdf}
+
+%     \caption{How the window size effects our theoretic $R_\text{coal}^2$ versus
+%         our CADD 6\% models classic BGS (B) and new B' methods at two scales (1
+%         Mbp and 100 kbp). Additionally, we have overlaid the CADD 6\%
+%         out-sample $R_{OS}^2$ estimates across many window sizes from
+%     \textcite{Murphy2022-sj}. We note that the theoretic $R_\text{coal}^2$
+% quickly declines to zero since the coalescence and mutational noise quickly
+% begins the linked selection signal.}
+
+%   \label{suppfig:coal_r2_scale}
+% \end{figure}
 
@@
 \begin{figure}[htbp]
+  \label{suppfig:resid-cadd}
+  \centering
+  \includegraphics[width=0.6\textwidth]{figures/supplementary/cadd_residual.pdf}
+  \caption{Residuals from the CADD 6\% sparse track plotted against the fraction of
+  basepairs in a window annotated by a CADD region.}
+\end{figure}
+
+\begin{figure}[htbp]
   \label{suppfig:resid-b}

### 2023-07-10 manuscript/manuscript.tex
fig checking

@@
 \begin{abstract}
+
+    Previous work has shown that negative selection against the steady influx
+    of deleterious mutations is the predominant process shaping genome-wide
+    patterns of genetic diversity in humans. This process leaves a ``linked
+    selection signal"
+
+    The classic population genetic
+    model of this process, background selection, considers
+
     % TODO: emphasize connection to nearly neutral theory
@@
 
-\section*{Results: Confirmation of Theory}
+\section*{Results: Validation of Theory}
 
@@
 ``other" feature is estimated at $97.8\%$.  When we use the constrained grid,
-these change to $97.85\%$ and $99.98\%$ respectively. By contrast, our
-predicted substitution rates for PhastCons regions are much lower: $0.48\%$ for
-the default grid, and $1.65\%$ for the constrained grid. This is likely due to
-the constrained grid placing more of the DFE mass on weaker selection
-coefficients, which in turn have a higher chance of drifting up to fixation.
+these change to $97.85\%$ and $99.98\%$ respectively. In our PhyloFit
+estimates, the divergence rate in introns is $92.44\%$, 5' UTRs $75.5\%$, and
+3' UTRs $72.65\%$. By contrast, our predicted substitution rates for PhastCons
+regions are much lower: $0.48\%$ for the default grid, and $1.65\%$ for the
+constrained grid. For comparison, the PhyloFit estimates for PhastCons regions
+is much higher, at $51.23\%$. Since this would include beneficial substitutions
+that would not be predicted by our model, this suggests a very high proportion
+of beneficial substitutions XXX check
+
+% This is likely due to the constrained grid placing more of the DFE mass on
+% weaker selection coefficients, which in turn have a higher chance of drifting
+% up to fixation.
 
@@
 \section*{Discussion}
+
+the PhastCons Priority Model is very close to CADD 6\%, and interpretable,
+similiar mutation rates. This
+
+weak selection remains elusive
+
+UTR rates
 

### 2023-07-10 manuscript/manuscript.tex
abstract draft

@@
 
-    Previous work has shown that negative selection against the steady influx
-    of deleterious mutations is the predominant process shaping genome-wide
-    patterns of genetic diversity in humans. This process leaves a ``linked
-    selection signal"
-
-    The classic population genetic
-    model of this process, background selection, considers
-
-    % TODO: emphasize connection to nearly neutral theory
-
-    % uniting quant gen/fitness variation, linked selection, and substitution
-    % The interesting thing about connecting all these models is that we might
-    % have hope of calibrating these models to the true process.
+    Across the human genome, there are large-scale fluctuations in genetic
+    diversity caused by the indirect effects of selection. This can be thought
+    of as a ``linked selection signal" that is shaped by selection acting on
+    the placement of functional regions and recombination rates along the
+    genome. Previous work has shown that negative selection against the steady
+    influx of new deleterious mutations into conserved regions is the
+    predominant mode of selection generating this signal in humans. However,
+    the theoretic model that underpins these results, classic Background
+    Selection theory, is only applicable when new mutations are so deleterious
+    that they cannot drift up in frequency and ultimately fix in the
+    population. Here, we develop an alternate theoretic basis for genome-wide
+    linked selection based on a quantitative genetics view, which considers the
+    distribution of additive fitness variance along the genome. Using a recent
+    model that jointly predicts the equilibrium fitness variance and
+    substitution rate due to deleterious mutations, we fit this model to
+    samples from three human populations to estimate the distribution of
+    selection coefficients and mutation rate. While our model can accommodate
+    weaker selection, we find consistent evidence across all populations of
+    very strong selection against deleterious mutations. We also find that our
+    model's predicted substitution rates are qualitatively consistent with
+    observed divergences in features under weak selective constraint like
+    introns and UTRs, but we observe vastly more substitutions in regions under
+    strong selective constrain, consistent with positive selection. While our
+    model explains up to 67\% of the out-sample variance in diversity at the
+    megabase scale, we find evidence of residual strong selection against
+    loss-of-function variants, indicating the dominant role negative selection
+    plays in the human genome. Altogether, our genome-wide model of the linked
+    selection is a step towards uniting evolutionary quantitative genetics with
+    the molecular evolutionary processes.
+
 \end{abstract}

### 2023-07-11 manuscript/manuscript.tex
working in andys edits

--- 2023-07-10 manuscript/manuscript.tex
@@
 \begin{abstract}
-
-    Previous work has shown that negative selection against the steady influx
-    of deleterious mutations is the predominant process shaping genome-wide
-    patterns of genetic diversity in humans. This process leaves a ``linked
-    selection signal"
-
-    The classic population genetic
-    model of this process, background selection, considers
-
-    % TODO: emphasize connection to nearly neutral theory
-
-    % uniting quant gen/fitness variation, linked selection, and substitution
-    % The interesting thing about connecting all these models is that we might
-    % have hope of calibrating these models to the true process.
+    Across the human genome, there are large-scale fluctuations in genetic
+    diversity caused by the indirect effects of selection. This can be thought
+    of as a ``linked selection signal" that reflects selection's effects
+    varying according to the placement of functional regions and recombination
+    rates along the genome. Previous work has shown that negative selection
+    against the steady influx of new deleterious mutations into conserved
+    regions is the predominant mode of selection in humans. However, the
+    theoretic model that underpins these results, classic Background Selection
+    theory, is only applicable when new mutations are so deleterious that they
+    cannot drift up in frequency and ultimately fix in the population. Here, we
+    develop an alternate theoretic basis for genome-wide linked selection based
+    on a quantitative genetics view, which considers the distribution of
+    additive fitness variance along the genome. Using a recent model that
+    jointly predicts the equilibrium fitness variance and substitution rates
+    due to deleterious mutations, we estimate the distribution of selection
+    coefficients and mutation rate across three human populations. While our
+    model can accommodate weaker selection, we find consistent evidence across
+    all populations of very strong selection against deleterious mutations. We
+    also find that our model's predicted substitution rates are qualitatively
+    consistent with observed divergences in features under weak selective
+    constraint like introns and UTRs, but less so in regions under strong
+    selective constraint. While our model explains up to 67\% of the out-sample
+    variance in diversity at the megabase scale, we find evidence of residual
+    strong selection against loss-of-function variants unfit by our model,
+    indicating the dominant role negative selection plays in the human genome.
+    Altogether, our genome-wide model of the linked selection is a step towards
+    uniting evolutionary quantitative genetics with substitution processes.
 \end{abstract}
@@
 of all adaptations, but the vast majority of mutations either do not affect
-fitness or are deleterious. Negative selection (i.e. purifying selection) works
-to eliminate such deleterious mutations from the population, which is essential
-to maintaining the integrity of information in the genome needed to maintain a
-lineage. Mutations under negative selection are kept at low frequencies
-(Haldane?), contributing to human traits (XXX) and diseases. Moreover, the
-negative selection process determines sequence conservation across phylogenetic
-timescales, which is evident in genome-wide cross-species comparisons
-\parencite{Siepel2005-wh}. As with most selection in the genome, the negative
-selection process stochastically perturbs the allele frequencies of neighboring
-linked variation. This reduces genetic variability to create a large-scale
-spatial linked selection ``reduction signal" in genome-wide diversity. Mutation
-and negative selection's contribution to this signal is mediated by the
-deleterious mutation rate, the strength of selection in conserved segments, and
-the spatial distribution of recombination rates and conserved segments. Since
-genome-wide recombination maps and putatively conserved segments are available
-for many species, researchers have combined these with population genetic
-theory to estimate the mutation rate and selection coefficients of deleterious
-mutations, as well as the overall reduction in genetic variability due to the
-negative selection process.
+fitness or are deleterious. The best predictor of whether a mutation will be
+harmful to fitness is if it lands in a functional region, which are often
+conserved over deep phylogenetic timescales \parencite{Siepel2005-wh}. Since
+such high conservation reflects the product of millions or even billions of
+years of evolutionary optimization, it follows that deleterious variants should
+be the overwhelming majority of segregating variation in functional regions,
+and be responsible for a significant portion of phenotypic variation and
+disease \parencite{Zeng2018-ci} (XXX).
+
+Both positive and negative selection stochastically perturb the allele
+frequencies of neighboring linked variation in a phenomenon known as linked
+selection \parencite{Nordborg1996-nq,Maynard_Smith1974-zr,Barton1998-aj}. While
+linked selection processes like sweeps can leave signals on small spatial
+scales (XXX), the precise arrangement of functional regions and recombination
+rates across the genome leads to a large-scale linked selection signal.
+
+Mutation and negative selection's contribution to this signal is mediated by
+the deleterious mutation rate, the strength of selection in conserved segments,
+and the spatial distribution of recombination rates and conserved segments.
+Since genome-wide recombination maps and putatively conserved segments are
+available for many species, researchers have combined these with population
+genetic theory to estimate the mutation rate and selection coefficients of
+deleterious mutations, as well as the overall reduction in genetic variability
+due to the negative selection process.
+
+HERE
+
+As deleterious variation is clustered in functional portions of the genome,
+we expect linked selection to reduce levels of diversity surrounding evolutionarily
+conserved segments (e.g. genes, regulatory elements, etc.) 
+and to shape large-scale spatial patterns in genetic diversity. 
+This landscape of diversity is the product of a complex interplay between the
+deleterious mutation rate, the local recombination rate, the strength of selection, and
+the spatial distribution of conserved segments along a chromosome. 
+Since genome-wide recombination maps and putatively conserved segments are available
+for many species, there has been consistent effort to fit models of
+linked selection to patterns of diversity to estimate population genetic
+parameters such as the strength of selection and the deleterious mutation rate 
+\parencite{Hudson1995-pt,McVicker2009-ax},
+the overall reduction in genetic variability due to the
+selection on deleterious variants \parencite{Elyashiv2016-vt,Murphy2022-sj},
+and in some cases the rate of beneficial mutations \parencite{corbett2015natural,Elyashiv2016-vt}.
 

### 2023-07-11 manuscript/manuscript.tex
ms changes

@@
 \begin{abstract}
+
     Across the human genome, there are large-scale fluctuations in genetic
@@
     develop an alternate theoretic basis for genome-wide linked selection based
-    on a quantitative genetics view, which considers the distribution of
-    additive fitness variance along the genome. Using a recent model that
-    jointly predicts the equilibrium fitness variance and substitution rates
-    due to deleterious mutations, we estimate the distribution of selection
-    coefficients and mutation rate across three human populations. While our
-    model can accommodate weaker selection, we find consistent evidence across
-    all populations of very strong selection against deleterious mutations. We
-    also find that our model's predicted substitution rates are qualitatively
-    consistent with observed divergences in features under weak selective
-    constraint like introns and UTRs, but less so in regions under strong
-    selective constraint. While our model explains up to 67\% of the out-sample
-    variance in diversity at the megabase scale, we find evidence of residual
-    strong selection against loss-of-function variants unfit by our model,
-    indicating the dominant role negative selection plays in the human genome.
-    Altogether, our genome-wide model of the linked selection is a step towards
-    uniting evolutionary quantitative genetics with substitution processes.
+    on a quantitative genetics view, which considers how diversity is reduced
+    according to the arrangement of additive fitness variance along the genome.
+    Using a recent model that jointly predicts the equilibrium fitness variance
+    and substitution rates due to deleterious mutations, we estimate the
+    distribution of selection coefficients and mutation rate across three human
+    populations. While our model can accommodate weaker selection, we find
+    consistent evidence across all populations of very strong selection against
+    deleterious mutations. We also find that our model's predicted substitution
+    rates are qualitatively consistent with observed divergences in features
+    under weak selective constraint like introns and UTRs, but less so in
+    regions under strong selective constraint. While our model explains up to
+    67\% of the out-sample variance in diversity at the megabase scale, we find
+    evidence of residual strong selection against loss-of-function variants
+    unfit by our model, indicating the dominant role negative selection plays
+    in the human genome. Altogether, our genome-wide model of the linked
+    selection is a step towards uniting evolutionary quantitative genetics with
+    substitution processes.
+
 \end{abstract}
@@
 of all adaptations, but the vast majority of mutations either do not affect
-fitness or are deleterious. The best predictor of whether a mutation will be
-harmful to fitness is if it lands in a functional region, which are often
-conserved over deep phylogenetic timescales \parencite{Siepel2005-wh}. Since
-such high conservation reflects the product of millions or even billions of
-years of evolutionary optimization, it follows that deleterious variants should
-be the overwhelming majority of segregating variation in functional regions,
-and be responsible for a significant portion of phenotypic variation and
-disease \parencite{Zeng2018-ci} (XXX).
-
-Both positive and negative selection stochastically perturb the allele
-frequencies of neighboring linked variation in a phenomenon known as linked
-selection \parencite{Nordborg1996-nq,Maynard_Smith1974-zr,Barton1998-aj}. While
-linked selection processes like sweeps can leave signals on small spatial
-scales (XXX), the precise arrangement of functional regions and recombination
-rates across the genome leads to a large-scale linked selection signal.
-
-Mutation and negative selection's contribution to this signal is mediated by
-the deleterious mutation rate, the strength of selection in conserved segments,
-and the spatial distribution of recombination rates and conserved segments.
-Since genome-wide recombination maps and putatively conserved segments are
-available for many species, researchers have combined these with population
-genetic theory to estimate the mutation rate and selection coefficients of
-deleterious mutations, as well as the overall reduction in genetic variability
-due to the negative selection process.
-
-HERE
-
-As deleterious variation is clustered in functional portions of the genome,
-we expect linked selection to reduce levels of diversity surrounding evolutionarily
-conserved segments (e.g. genes, regulatory elements, etc.) 
-and to shape large-scale spatial patterns in genetic diversity. 
-This landscape of diversity is the product of a complex interplay between the
-deleterious mutation rate, the local recombination rate, the strength of selection, and
-the spatial distribution of conserved segments along a chromosome. 
-Since genome-wide recombination maps and putatively conserved segments are available
-for many species, there has been consistent effort to fit models of
-linked selection to patterns of diversity to estimate population genetic
+fitness or are deleterious. Negative selection (i.e. purifying selection) works
+to eliminate these deleterious mutations from the population, thus we expect
+them to appear at low frequencies within populations
+\parencite{Haldane1927-ga}, and less likely to fix between lineages. The best
+predictor of whether a new mutation will reduce fitness is if it occurs in a
+functional region of the genome, which are often conserved over deep
+phylogenetic timescales \parencite{Siepel2005-wh,Margulies2003-sn}. Since these
+conserved regions reflect the product of millions of years of evolutionary
+optimization, it follows that deleterious variants should be the overwhelming
+majority of segregating variation in these regions. Furthermore, rare variation
+in these regions is responsible for a significant portion of phenotypic
+variation and disease in humans
+\parencite{Zeng2018-ci,Lek2016-cb,Karczewski2020-ky,Tennessen2012-ge}.
+
+Selection on both beneficial and deleterious variants perturbs the allele
+frequencies of neighboring linked sites, a phenomenon known as linked selection
+\parencite{Nordborg1996-nq,Maynard_Smith1974-zr,Barton1998-aj,Charlesworth1993-gb}.
+Since deleterious variation is clustered in functional portions of the genome,
+we expect linked selection to reduce levels of diversity around conserved
+segments (e.g. genes, regulatory elements, etc.). The genomic arrangement of
+these evolutionarily conserved regions and recombination rates create a
+large-scale spatial signal of the linked selection process in genetic diversity
+along the genome. Since genome-wide recombination maps and putatively conserved
+segments are available for many species, there has been consistent effort to
+fit models to this linked selection signal to estimate population genetic
 parameters such as the strength of selection and the deleterious mutation rate
-\parencite{Hudson1995-pt,McVicker2009-ax},
-the overall reduction in genetic variability due to the
-selection on deleterious variants \parencite{Elyashiv2016-vt,Murphy2022-sj},
-and in some cases the rate of beneficial mutations \parencite{corbett2015natural,Elyashiv2016-vt}.
-
-Classic Background Selection (BGS) theory is the predominant theoretic model to
-predict the reduction in linked neutral diversity, and is used by statistical
-approaches that quantify the aspects of the negative selection. However, the
-BGS model makes some assumptions for mathematical tractability that could
-distort inferences about the negative selection process. First, since fixation
-probabilities ultimately depend on the product of the deleterious selection
-coefficient (s) and population size (N), the efficacy of negative selection
-should depend on demography. Unfortunately, accommodating demography into
-theoretic negative selection models remains an open, difficult problem (XXX).
-Second, building off classic models of mutation-selection balance (Crow 1970,
-Kimura and Maruyama 196), the Background Selection (XXX) model assumes that new
-mutations are sufficiently deleterious that they are invariably driven to loss.
-Under this assumption, the effect of selection is well-approximated by simply
-rescaling the neutral coalescent by a reduction factor known as B (XXX).
-However, this is only appropriate in a limited domain of parameters relevant to
-the negative selection process (XXX). Because the BGS model cannot accommodate
-the possibility of fixation for mutations with $Ns \ll 1$, diversity is
-incorrectly predicted to fall to zero as the strength of selection diminishes.
-This weak selection problem is related to the wider problem of modeling the
-substitution rate of deleterious mutation (i.e. the “ratchet” rate). Third, the
-BGS model assumes segments under purifying selection are sufficiently far apart
-that there is no selective interference between them. This is a matter of
-degree, since under sufficiently high mutation rates or with little
-recombination, selective interference can occur with a segment.
+\parencite{Hudson1995-pt,McVicker2009-ax}, as well as distinguish and quantify
+the roles of positive and negative selection and estimate the rate of
+beneficial mutations \parencite{Elyashiv2016-vt,Murphy2022-sj}. In humans,
+previous work has shown that negative selection plays the dominant role in
+shaping megabase-scale patterns of diversity, with positive selection playing a
+nearly negligible role \parencite{Murphy2022-sj}.
+
+Prior work to model the reduction in linked diversity due to deleterious
+mutations has largely relied on the classic Background Selection (BGS) model
+\parencite{Charlesworth1993-gb,Nordborg1996-nq,Hudson1995-pt,Hudson1995-xc}.
+While this model has been successful in explaining many patterns of diversity,
+some its simplifying assumptions may distort inferences about the negative
+selection process. First, since fixation probabilities ultimately depend on the
+product of the deleterious selection coefficient (s) and population size (N),
+the efficacy of negative selection should depend on demography. Unfortunately,
+accommodating demography into theoretic negative selection models remains an
+open, difficult problem \parencite{Zeng2013-ep,Johri2020-oj}. Second, building
+off classic models of mutation-selection balance
+\parencite{Crow1970-wm,Kimura1966-bk}, the BGS model assumes that new mutations
+are sufficiently deleterious that they are invariably driven to loss. Under
+this assumption, the effect of selection is well-approximated by simply
+rescaling the neutral coalescent by a reduction factor known as $B =
+\nicefrac{N}{N_e}$ \parencite{Charlesworth2013-kl}. However, this simple
+rescaling approach is not appropriate across parts of parameter space that are
+relevant to natural populations \parencite{McVean2000-bt,Good2014-yz}. In
+particular, the BGS model cannot accommodate the possibility of fixation for
+mutations with $Ns \ll 1$, diversity is incorrectly predicted to fall to zero
+as the strength of selection diminishes. This weak selection problem is related
+to the wider problem of modeling the substitution rate of deleterious mutation
+(i.e. the “ratchet” rate) and incorporating the impact of perturbations from
+selective interference.
 
 In this work, we extend another class of linked selection models that derive
-from quantitative genetics to address limitations of classic background
-selection theory. These models suggest that the reduction in pairwise
-coalescence times at a position in the genome is determined by the
-recombination map and distribution of polygenic additive fitness variance along
-the chromosome \parencite{Santiago1998-bs}. This reduction is due to cumulative
-effect of stochastic perturbations from selection at other sites. While these
-models can theoretically accommodate polygenic fitness variation from any
-source as long as its rate of change is not too rapid, we focus specifically on
-a deleterious-mutations-only model from Santiago and Caballero (2016). This
-model is identical to the classic BGS model when selection against deleterious
-mutations is strong, but it also correctly predicts the reduction in diversity
-and substitutions rates of weakly deleterious mutations by accurately modeling
-levels of genic fitness variation. We extend Santiago and Caballero’s
-(hereafter the SC16) model so that it can be used to model genome-wide patterns
-of diversity under negative selection processes, and develop Python software to
-use this theoretic approach to fit genome-wide models of negative selection.
-Using forward simulations, we show this model leads to more accurate DFE
-estimates under weak selection. While this new model accommodates the
-possibility that a given reduction B could be due to weaker selection, in
-applying our composite likelihood method to human population genomic data, we
-confirm earlier findings that strong negative selection at a small fraction of
-sites fits human data across populations best. Finally, we show that while this
-model solves some limitations of the classic BGS model, simulations reveal
-another mysterious defect: a poor fit around $Ns=1$. We show that this can be
-partially corrected by incorporating the effects of selective interference on
-genic variation, but the remaining discrepancy is due to the build up of
-negative linkage disequilibrium that cannot be adequately captured by this
-model. We find accounting for genic selective interference between segments
-fits human population genomic data slightly better, suggesting levels of
-negative selection in humans are sufficient to perturb the selection at other
-sites.
+from quantitative genetics to address limitations of the classic BGS model.
+These models express linked selection in terms of how polygenic fitness
+variance creates weak genetic draft, and which increases pairwise coalescence
+rates. While these models can theoretically accommodate additive fitness
+variance from any source as long as its rate of change is not too rapid, we
+focus specifically on a deleterious-mutations-only model of fitness variance
+from \textcite{Santiago2016-mu}. This model is identical to BGS when selection
+against deleterious mutations is strong, but it also correctly predicts the
+reduction in diversity when selection is weak by jointly predicting the
+deleterious substitution rate. We extend the Santiago and Caballero (hereafter
+the SC16) model of the negative selection process so that it can be fit to
+patterns of genome-wide diversity, according to the spatial distribution of
+genomic features that could harbor deleterious fitness variation. Using forward
+simulations, we show this model leads to more accurate estimates of the
+distribution of fitness effects under weak selection. Finally, we apply our
+composite-likelihood method to human population genomic data and provide new
+estimates of the negative selection process in humans.
 
@@
 
-Our work is motivated by a well-known limitation of classic background
-selection theory: it is only accurate for strongly deleterious mutations
-\parencite{Charlesworth1993-gb,McVean2000-bt,Good2013-lp,Gordo2002-dr}. Given
-statistical methods that fit linked selection processes to genome-wide
-diversity data rely on the classic BGS model to estimate the distribution of
-fitness effects and rate of deleterious mutations, it is important to
-characterize and remedy these shortcomings. However, extending background
-selection theory to model weak negative selection is challenging, and it is
-worthwhile to develop some intuition for why this is the case, how the classic
-background selection works, and how the quantitative genetic model of
-\textcite{Santiago2016-mu} solves these problems and can be extended to
-genome-wide inference.
+Similar to previous work, we aim to characterize the negative selection process
+through the signal left on nucleotide diversity at linked sites. Although
+nucleotide diversity is a less-informative statistic than the full site
+frequency spectrum, it is much easier to derive theoretic models of how
+selection alters expected pairwise coalescence times at linked sites than to
+describe the entire genealogy. Still, inferences drawn from the linked
+selection signal about selective processes are only as accurate as the
+underlying theoretic model. Our work is motivated by a well-known limitation of
+classic background selection theory: it only accurately predicts levels of
+linked diversity when new mutations are strongly deleterious
+\parencite{Charlesworth1993-gb,McVean2000-bt,Good2013-lp,Gordo2002-dr}.
+However, developing linked selection theory that accommodates weak selection
+has been challenging problem \parencite{Good2014-yz,Haigh1978-gt,Higgs1995-xc},
+and it is worthwhile to develop in intuition for how the classic BGS model
+works, and how quantitative genetic models of linked selection can alleviate
+these problems and be extended to genome-wide inference.
 
@@
 heterozygotes and $2s$ in homozygotes. The classic background selection model
-only consider mutations that are sufficiently deleterious that they are
+only considers mutations that are sufficiently deleterious that they are
 destined to extinction
@@
 
-When mutations are only weakly deleterious, their allele frequencies are
-strongly influenced by stochastic perturbations and their frequency
-trajectories are no longer well-approximated by deterministic models. When
-genetic drift is the only source of randomness in allele frequency change, the
-relevant scale of stochastic perturbations is determined by the drift-effective
-population size $N_e$ (e.g. due to only non-selective process). The insight of
-nearly neutral theory \parencite{Ohta1971-gq,Ohta1992-yi} is determined by the
-compound parameter $2N_e s$. When $2N_e s \gg 1$, the stochastic component of
-allele frequency change is minuscule relative to selection and can be ignored;
-this is why the factor that rescales the drift-effective population size under
-the BGS model does not depend on $N_e$. However, when $2N_e s \le 1$, the
+By contrast, the frequency trajectories of weakly deleterious mutations are
+strongly impacted by stochastic perturbations, so deterministic models no
+longer provide a good approximation. When genetic drift is the only source of
+these randomness in allele frequency change, the relevant scale of stochastic
+perturbations is determined by the drift-effective population size $N_e$
+\parencite{Ohta1971-gq,Ohta1992-yi}. Under weak selection ($2N_e s \le 1$), the
 stochastic perturbations are on a scale equal to or greater than the changes
-due to selection. In this regime, weakly deleterious alleles can drift up to
-intermediate frequencies before their eventual loss or fixation. When this
-occurs, classic BGS theory breaks down for several reasons. First, the ``delay
-phase" is no longer negligible as lineages carrying weakly deleterious
-mutations can persist on coalescent timescales. This distorts genealogies away
-those expected under neutrality
-\parencite{Przeworski1999-mb,OFallon2010-my,Higgs1995-xc}. Second, the
-distribution of the number of deleterious mutations (and its corresponding
+due to selection. Then, weakly deleterious alleles can by chance drift up to
+intermediate frequencies before their eventual loss or fixation. Classic BGS
+breaks down in this domain, since the ``delay phase" is no longer negligible
+and genealogies are no longer well-approximated by a neutral coalescence
+process \parencite{Przeworski1999-mb,OFallon2010-my,Higgs1995-xc}. Moreover,
+the distribution of the number of deleterious mutations (and its corresponding
 fitness distribution) are no longer stationary and become a traveling wave
 \parencite{Rouzine2008-qz,Good2013-lp,Gessler1995-hz} towards reduced
-population fitness (since beneficial and back mutations are ignored).
-Consequently, individuals in the least-loaded class are stochastically lost,
-fixing those mutations with a click of ``Muller's ratchet"
-\parencite{Muller1964-ki}. Determining both the rate of Muller's ratchet
+population fitness (since beneficial and back mutations are ignored). When this
+occurs, individuals in the least-loaded class are stochastically lost, fixing
+those mutations with a click of ``Muller's ratchet" \parencite{Muller1964-ki}.
+Determining both the rate of Muller's ratchet
 \parencite{Haigh1978-gt,Gordo2002-dr,Gessler1995-hz} and the shape of
@@
 
-To further complicate matters, both strong and weakly deleterious mutations
-create \emph{heritable} fitness variation. In turn, the presence of heritable
-fitness variation generates another source of randomness in allele frequency
-change known as genetic \emph{draft} \parencite{Neher2013-dz}. An allele
-experiences genetic draft when it becomes randomly associated with a fitness
-background, which stochastically perturbs its trajectory across generations.
-Like genetic drift, perturbations due to draft are directionless but overall
-increase the variance in allele frequency change. Unlike drift, stochastic
-perturbations caused by draft are autocorrelated across generations for as long
-an allele does not recombine or segregate off its fitness background and the
-selective environment does not change
-\parencite{Robertson1961-ho,Santiago1995-hx,Buffalo2019-qs}. When selection is
-strong, draft can generate multiple-merger coalescences that create
-discontinuities in allele frequency trajectories, making diffusion
+To further complicate matters, when selection is weak, the stochastic
+perturbations from selection at neighboring sites generates another source of
+randomness in allele frequency change known as genetic \emph{draft}
+\parencite{Neher2013-dz}. An allele experiences genetic draft when it becomes
+randomly associated with a fitness background, which stochastically perturbs
+its trajectory across generations. Like genetic drift, perturbations due to
+draft are directionless but overall increase the variance in allele frequency
+change. Unlike drift, stochastic perturbations caused by draft are
+autocorrelated across generations for as long an allele does not recombine or
+segregate off its fitness background and the selective environment does not
+change \parencite{Robertson1961-ho,Santiago1995-hx,Buffalo2019-qs}. When
+selection is strong, draft can generate multiple-merger coalescences that
+create discontinuities in allele frequency trajectories, making diffusion
 approximations ill-suited \parencite{Gillespie2000-mh,Der2011-it,Neher2013-dz}.
@@
 
-A key insight from Robertson (\citeyear{Robertson1961-ho}) is that in the
-long-run weak draft acts like an increase in the variance in offspring number.
-\textcite{Wright1938-tv} showed that additional non-heritable variance in
-offspring number (e.g. due to mating strategy) rescales the population size
-according to,
-
-\begin{align}
-    \label{eq:simple_Ne}
-    N_e = \frac{2N}{V + 2}
-\end{align}
-%
-where $V$ is the extra variance in offspring number.
-
-However, unlike non-heritable variation in offspring number, the heritable
-variation is autocorrelated across generations. At the individual level
-Robertson considered, this is because offspring from large families tend to
-beget many descendents themselves (and likewise with small families). The same
-autocorrelation occurs at the genomic level, as the perturbations to a neutral
-allele's trajectory from its particular fitness background tend to occur in the
-same direction across generations as long as the linkage exists and the
-selective environment is constant. Thus, the total additional heritable
-variation in offspring number is the product of the equilibrium additive
-fitness variation $V_A$ and an inflation factor $Q_t^2$ due to the build up of
-autocorrelation to generation $t$. We let $Q^2 := Q_t^2$ as $t \to \infty$.
-Intuitively, the product $V_A Q^2$ represents the expected total variance in
-reproductive success a neutral mutation experiences over its lifetime in a
-system with weak selection draft. Incorporating the impact of time-invariant
-autocorrelation created by draft as a rescaled rate of drift is analogous to
-using the Green--Kubo relation to find the transport coefficient in molecular
-dynamics with velocity autocorrelation \parencite{Green1954-kl,Kubo1957-va}.
-
-Including the total asymptotic variance in reproductive success in Equation
-\eqref{eq:simple_Ne} and assuming Wright--Fisher levels of non-heritable
-variation (i.e. $V = 2 + 4 Q^2 V_A$), the draft-effective population size $N_d$
-is
+Quantitative genetics models of linked selection formulate the impact of weak
+draft in terms of additive genetic fitness variance, which can more readily
+accommodate weak selection. This idea derives from a key insight from Robertson
+(\citeyear{Robertson1961-ho}): in the long-run, the effect of weak draft on
+neutral allele frequencies acts like an increase in the variance in offspring
+number. However, unlike non-heritable variation in offspring number, the
+heritable variation that generates draft is autocorrelated across generations.
+At the individual level Robertson considered, this is because offspring from
+large families tend to beget many descendents themselves (and likewise with
+small families). The same autocorrelation occurs at the genomic level, as the
+perturbations to a neutral allele's trajectory from its particular fitness
+background tend to occur in the same direction across generations as long as
+the linkage exists and the selective environment is constant. Thus, the total
+additional heritable variation in offspring number is the product of the
+equilibrium additive fitness variation $V_A$ in a generation and an inflation
+factor $Q_t^2$ due to the build up of autocorrelation to generation $t$. We let
+$Q^2 := Q_t^2$ as $t \to \infty$. Intuitively, the product $V_A Q^2$ represents
+the expected total variance in reproductive success a neutral mutation
+experiences over its lifetime in a system with weak selection draft.
+
+We can include this total additional variance into expressions for effective
+population size as we would larger variance in family sizes due to
+non-heritable causes \parencite{Wright1938-tv}. We define the draft-effective
+population size $N_d$ as
 
@@
 \end{align}
-
+%
 (c.f. \cite{Robertson1961-ho,Santiago1995-hx}; see Supplementary Section
-\ref{supp:theory} for a proof). In general, however, statistics such as
-heterozygosity depend on the cumulative autocorrelation factor up to some time
-$t$, $Q_t^2$ \parencite[p. 2111]{Santiago1998-bs}. This reflects the fact that
-effective population size experienced by a mutation newly associated with a
-fitness background is larger than the asymptotic effective population size,
-since not much autocorrelation has accumulated in the stochastic perturbations
-caused by drift yet. From a backwards-time perspective, this reflects the fact
-that pairwise diversity (which is approximately heterozygosity when $2N\mu \ll
-1$) depends on the pairwise coalescence rate per generation, which is not
-constant under draft. The benefit of using Robertson's forward-time model of
-draft is that the inflation factor is invariant with respect to the particular
-fitness background the neutral allele becomes stochastically associated with.
-By contrast, the difficulty with modeling draft backwards in time is that the
-coalescence rates experienced by a lineage are not invariant to which lineage
-was sampled due to its particular associated fitness background.
+\ref{supp:theory} for a proof). The benefit of modeling draft with Robertson's
+forward-time model is that the inflation factor is invariant with respect to
+the particular fitness background the neutral allele becomes stochastically
+associated with. By contrast, the difficulty with modeling draft backwards in
+time is that the coalescence rates experienced by a lineage are not invariant
+to which lineage was sampled due to its particular associated fitness
+background.
 
@@
 by the probability of fixation $p_F(N_d, s)$
-\parencite{Kimura1962-su,Malecot1948-zv} using the draft-rescaled effective
+\parencite{Kimura1962-su,Malecot1952-qh} using the draft-rescaled effective
 population size, i.e. $R = N U p_F(N_d, s)$. Given this equation for the

### 2023-07-11 manuscript/supplementary.tex
ms changes

@@
 The $Q_t^2$ terms represent the inflation of variance due to autocovariance in
-frequency change across generations due to weak draft.
+frequency change across generations due to weak draft. Incorporating the impact
+of time-invariant autocorrelation created by draft as a rescaled rate of drift
+is analogous to using the Green--Kubo relation to find the transport
+coefficient in molecular dynamics with velocity autocorrelation
+\parencite{Green1954-kl,Kubo1957-va}.
 
@@
 inflation factor $Q_t^2$ that sets the draft-effective population size varies
-over $t$. This is because selective processes can lead to non-constant pairwise
-coalescent rates, which impact pairwise coalescent rates and thus the reduction
-factor. We experimented with calculating this full sum and integral
-approximations to it in our methods, but it seemed to make little difference in
-our application. Additionally, it was too costly to calculate computationally
-for genome-wide calculations of the B' map. Instead, we solve for the
-asymptotic $\widetilde{N}_{e,\infty}$ for each segment, and use that to
-calculate the log reduction factor under the SC16 model, $b'$.
+over $t$. This is because statistics such as heterozygosity depend on the
+cumulative autocorrelation factor up to some time $t$, $Q_t^2$ \parencite[p.
+2111]{Santiago1998-bs}. This reflects the fact that effective population size
+experienced by a mutation newly associated with a fitness background is larger
+than the asymptotic effective population size, since not much autocorrelation
+has accumulated in the stochastic perturbations caused by drift yet. From a
+backwards-time perspective, this reflects the fact that pairwise diversity
+(which is approximately heterozygosity when $2N\mu \ll 1$) depends on the
+pairwise coalescence rate per generation, which is not constant under draft.
+
+We experimented with calculating this full sum and integral approximations to
+it in our methods, but it seemed to make little difference in our application.
+Additionally, it was too costly to calculate computationally for genome-wide
+calculations of the B' map. Instead, we solve for the asymptotic
+$\widetilde{N}_{e,\infty}$ for each segment, and use that to calculate the log
+reduction factor under the SC16 model, $b'$.
 

### 2023-07-12 manuscript/manuscript.tex
ms changes

@@
 nucleotide diversity is a less-informative statistic than the full site
-frequency spectrum, it is much easier to derive theoretic models of how
-selection alters expected pairwise coalescence times at linked sites than to
-describe the entire genealogy. Still, inferences drawn from the linked
-selection signal about selective processes are only as accurate as the
-underlying theoretic model. Our work is motivated by a well-known limitation of
-classic background selection theory: it only accurately predicts levels of
-linked diversity when new mutations are strongly deleterious
+frequency spectrum, it is much easier to derive theoretic models of how linked
+selection alters expected pairwise coalescence times than to describe the
+entire genealogy. Still, inferences drawn from the linked selection signal
+about selective processes are only as accurate as the underlying theoretic
+model. Our work is motivated by a well-known limitation of classic background
+selection theory: it only accurately predicts levels of linked diversity when
+new mutations are strongly deleterious
 \parencite{Charlesworth1993-gb,McVean2000-bt,Good2013-lp,Gordo2002-dr}.
-However, developing linked selection theory that accommodates weak selection
-has been challenging problem \parencite{Good2014-yz,Haigh1978-gt,Higgs1995-xc},
-and it is worthwhile to develop in intuition for how the classic BGS model
-works, and how quantitative genetic models of linked selection can alleviate
-these problems and be extended to genome-wide inference.
-
-Both the classic background selection and SC16 models imagine a negative
-selection process where a continual influx of deleterious mutations flow into
-the population at a rate of $\mu$ per basepair per generation in a conserved
-region of $L$ basepairs, such that the region-wide per generation mutation rate
-is $U = \mu L$. Each mutation imposes an additive selective cost of $s$ in
-heterozygotes and $2s$ in homozygotes. The classic background selection model
-only considers mutations that are sufficiently deleterious that they are
-destined to extinction
-\parencite{Charlesworth1993-gb,Nordborg1996-nq,Hudson1995-pt,Hudson1994-oh}. In
-this case, the genealogy is well-approximated by a neutral coalescence process
-with a rescaled effective population size of $N_0 = N \exp(-\nicefrac{U}{s})$.
-This approximation works reasonably well for two reasons (though see
-\cite{Cvijovic2018-vd,Walczak2012-fi,Nicolaisen2012-vs}). First, when selection
-is strong, the number of deleterious mutations per haplotype has a stationary
-Poisson distribution with an average equal to the equilibrium under the
-deterministic mutation-selection model \parencite{Haldane1927-ga}. Since
-lineages carrying strongly deleterious mutations cannot contribute to the
-genealogy in the long run, coalescence events only occur in the fraction
-$\exp(-\nicefrac{U}{s})$ of the population free of mutations. Second, the time
-it takes for a present-day lineage to trace their ancestry back to a
-mutation-free ancestor (the ``delay phase") is negligible compared to the
-coalescence timescale of $N_0$ among these ancestors (the ``coalescence
-phase"), and can be ignored (\cite{Durrett2008-ql}, p. 213;
-\cite{Good2014-yz}).
-
-By contrast, the frequency trajectories of weakly deleterious mutations are
-strongly impacted by stochastic perturbations, so deterministic models no
-longer provide a good approximation. When genetic drift is the only source of
-these randomness in allele frequency change, the relevant scale of stochastic
-perturbations is determined by the drift-effective population size $N_e$
-\parencite{Ohta1971-gq,Ohta1992-yi}. Under weak selection ($2N_e s \le 1$), the
-stochastic perturbations are on a scale equal to or greater than the changes
-due to selection. Then, weakly deleterious alleles can by chance drift up to
-intermediate frequencies before their eventual loss or fixation. Classic BGS
-breaks down in this domain, since the ``delay phase" is no longer negligible
-and genealogies are no longer well-approximated by a neutral coalescence
-process \parencite{Przeworski1999-mb,OFallon2010-my,Higgs1995-xc}. Moreover,
-the distribution of the number of deleterious mutations (and its corresponding
-fitness distribution) are no longer stationary and become a traveling wave
-\parencite{Rouzine2008-qz,Good2013-lp,Gessler1995-hz} towards reduced
-population fitness (since beneficial and back mutations are ignored). When this
-occurs, individuals in the least-loaded class are stochastically lost, fixing
-those mutations with a click of ``Muller's ratchet" \parencite{Muller1964-ki}.
-Determining both the rate of Muller's ratchet
-\parencite{Haigh1978-gt,Gordo2002-dr,Gessler1995-hz} and the shape of
-genealogies under weak selection have been stubborn open problems in
-evolutionary genetics.
-
-To further complicate matters, when selection is weak, the stochastic
-perturbations from selection at neighboring sites generates another source of
-randomness in allele frequency change known as genetic \emph{draft}
-\parencite{Neher2013-dz}. An allele experiences genetic draft when it becomes
-randomly associated with a fitness background, which stochastically perturbs
-its trajectory across generations. Like genetic drift, perturbations due to
-draft are directionless but overall increase the variance in allele frequency
-change. Unlike drift, stochastic perturbations caused by draft are
-autocorrelated across generations for as long an allele does not recombine or
-segregate off its fitness background and the selective environment does not
-change \parencite{Robertson1961-ho,Santiago1995-hx,Buffalo2019-qs}. When
-selection is strong, draft can generate multiple-merger coalescences that
-create discontinuities in allele frequency trajectories, making diffusion
-approximations ill-suited \parencite{Gillespie2000-mh,Der2011-it,Neher2013-dz}.
-However, when selection is weak, draft is well-approximated as an increase in
-the rate of drift (or equivalently, a reduction in effective population size).
-
-Quantitative genetics models of linked selection formulate the impact of weak
-draft in terms of additive genetic fitness variance, which can more readily
-accommodate weak selection. This idea derives from a key insight from Robertson
-(\citeyear{Robertson1961-ho}): in the long-run, the effect of weak draft on
-neutral allele frequencies acts like an increase in the variance in offspring
-number. However, unlike non-heritable variation in offspring number, the
-heritable variation that generates draft is autocorrelated across generations.
-At the individual level Robertson considered, this is because offspring from
-large families tend to beget many descendents themselves (and likewise with
-small families). The same autocorrelation occurs at the genomic level, as the
-perturbations to a neutral allele's trajectory from its particular fitness
-background tend to occur in the same direction across generations as long as
-the linkage exists and the selective environment is constant. Thus, the total
-additional heritable variation in offspring number is the product of the
-equilibrium additive fitness variation $V_A$ in a generation and an inflation
-factor $Q_t^2$ due to the build up of autocorrelation to generation $t$. We let
-$Q^2 := Q_t^2$ as $t \to \infty$. Intuitively, the product $V_A Q^2$ represents
-the expected total variance in reproductive success a neutral mutation
-experiences over its lifetime in a system with weak selection draft.
+Unfortunately, modifying background selection theory so that it accommodates
+weak selection has been challenging problem
+\parencite{Good2014-yz,Haigh1978-gt,Higgs1995-xc}. Here, we develop intuition
+for why the classic BGS model works well for strong selection but not weak, and
+how formulating selection in terms of the additive fitness variance it creates
+can alleviate these problems.
+
+Throughout, we imagine a negative selection process where a continual influx of
+deleterious mutations occur at a rate of $\mu$ per basepair per generation in a
+conserved region of $L$ basepairs, such that the region-wide per generation
+mutation rate is $U = \mu L$. Each mutation imposes a selective cost of $s$ in
+heterozygotes and $2s$ in homozygotes, and fitness effects are multiplicative
+across sites. When selection against a new mutation is strong, the stochastic
+perturbations it experiences due to random genetic drift are weak relative to
+the frequency changes caused by selection. In this case, every deleterious
+mutation is destined to eventual extinction, and only the fraction of lineages
+without any mutations contribute to the long-run ancestry of future
+populations. The classic background selection approximates this process by
+rescaling the neutral coalescent with an effective population size of $N_0 = N
+\exp(-\nicefrac{U}{s})$
+\parencite{Charlesworth1993-gb,Nordborg1996-nq,Hudson1995-pt,Hudson1994-oh}.
+This approximation works well for two reasons (though see
+\cite{Cvijovic2018-vd,Walczak2012-fi,Nicolaisen2012-vs}). First, he number of
+mutations per haplotype has has a stationary Poisson distribution with an
+average equal to the deterministic mutation-selection equilibrium
+\parencite{Haldane1927-ga}. Second, the time it takes for a present-day lineage
+to trace their ancestry back to a mutation-free ancestor (the ``delay phase")
+is negligible compared to the coalescence timescale of $N_0$ among
+mutation-free ancestors (\cite{Durrett2008-ql}, p. 213; \cite{Good2014-yz}).
+
+By contrast, when selection is weak, the frequency trajectories of both neutral
+and selected sites are strongly impacted by stochastic perturbations from two
+sources. First, genetic drift is due to Mendelian segregation and non-heritable
+fitness differences; the magnitude of its perturbations are set by the
+\emph{drift-effective} population size, $N_e$. Second, genetic draft occurs
+when an allele becomes randomly associated with a fitness background that
+perturbs its trajectory across generations \parencite{Neher2013-dz}. Like
+genetic drift, the perturbations due to draft are directionless but overall
+increase the variance in allele frequency change. Unlike drift, the impact of
+the stochastic perturbations due to draft are autocorrelated and build up over
+generations until the fitness background recombines off
+\parencite{Robertson1961-ho,Santiago1995-hx,Buffalo2019-qs}. While draft can
+generate multiple-merger coalescences that create discontinuities in allele
+frequency trajectories, making diffusion approximations ill-suited
+\parencite{Gillespie2000-mh,Der2011-it,Neher2013-dz}. However, when selection
+is weak, draft is well-approximated as an increase in the rate of drift (or
+equivalently, a reduction in effective population size). The stochastic
+perturbations between selecting sites creates Hill--Robertson interference
+\parencite{Hill1966-kd}.
+
+The stochastic perturbations from genetic drift and draft impact both
+genealogies and selection dynamics in complex ways, which has made finding
+analytic theory predicting their joint effects difficult. When drift acts
+alone, the relevant scale of perturbations is determined by the
+\emph{drift-effective population size} $N_e$
+\parencite{Ohta1971-gq,Ohta1992-yi}. When the random perturbations due to drift
+are stronger than the changes due to selection ($2N_es \le 1$), weakly
+deleterious mutations can drift up to intermediate frequencies before their
+eventual loss or fixation. The assumptions of classic BGS suddenly breakdown in
+two ways. First, the ``delay phase" is no longer negligible and genealogies are
+no longer well-approximated by a rescaled neutral coalescence process
+\parencite{Przeworski1999-mb,OFallon2010-my,Higgs1995-xc}. Second, the
+distribution of the number of deleterious mutations (and its corresponding
+fitness distribution) is no longer a stationary Poisson distribution, instead
+becoming a traveling wave \parencite{Rouzine2008-qz,Good2013-lp,Gessler1995-hz}
+towards reduced population fitness (since we ignore beneficial mutations). When
+this occurs, individuals in the least-loaded class are stochastically lost,
+fixing those mutations with a click of ``Muller's ratchet"
+\parencite{Muller1964-ki,Charlesworth1997-qn}. Unfortunately, determining the
+rate of Muller's ratchet is another difficult, open problem
+\parencite{Haigh1978-gt,Gordo2002-dr,Gessler1995-hz} related to Hill--Robertson
+interference \parencite{Felsenstein1974-xm}.
+
+A recent model from \textcite{Santiago2016-mu} addresses these limitations by
+formulating the negative selection process in terms of the equilibrium fitness
+variance $V_A$ under mutation, drift, and draft.
+
+Since their earlier work has
+related the additive fitness variance in a region to the draft-effective
+population experienced by a neutral allele arbitrarily far away
+\parencite{Santiago1998-bs}, this can be
+
+HERE
+
+that derives the effective population
+size experienced by a neutral allele when there is heritable fitness in the
+population. This theory is in essence a model for the magnitude of stochastic
+perturbations (what we think of as a \emph{draft-effective} population size)
+when heritable fitness variation generates weak genetic draft. As mentioned,
+perturbations due to draft are autocorrelated through time; at the individual
+level Robertson considered, this is because offspring from large families tend
+to beget many descendents themselves (and likewise with small families).
+However, the same autocorrelation occurs at the genomic level with linkage
+\parencite{Santiago1998-bs}, as the perturbations to a neutral allele's
+trajectory from its particular fitness background tend to occur in the same
+direction across generations until recombination.
+
+The benefit of this quantitative genetics formulation of linked selection is
+two-fold. First, as we will see, Santiago and Caballero's
+(\citeyear{Santiago2016-mu}) model for the additive fitness variance under both
+strong and weak negative selection is quite accurate. Second, these models of
+linked selection can 
+They do this through
+
+Thus, the total additional
+heritable variation in offspring number is the product of the equilibrium
+additive fitness variation $V_A$ in a generation and an inflation factor
+$Q_t^2$ due to the build up of autocorrelation to generation $t$. We let $Q^2
+:= Q_t^2$ as $t \to \infty$. Intuitively, the product $V_A Q^2$ represents the
+expected total variance in reproductive success a neutral mutation experiences
+over its lifetime in a system with weak selection draft.
 
@@
 
+draft limits discussion - we were heavily constrained by simulations / regions
+etc
+
 the PhastCons Priority Model is very close to CADD 6\%, and interpretable,

### 2023-07-12 manuscript/manuscript.tex
theory sec changes

@@
 This approximation works well for two reasons (though see
-\cite{Cvijovic2018-vd,Walczak2012-fi,Nicolaisen2012-vs}). First, he number of
+\cite{Cvijovic2018-vd,Walczak2012-fi,Nicolaisen2012-vs}). First, the number of
 mutations per haplotype has has a stationary Poisson distribution with an
@@
 generations until the fitness background recombines off
-\parencite{Robertson1961-ho,Santiago1995-hx,Buffalo2019-qs}. While draft can
-generate multiple-merger coalescences that create discontinuities in allele
-frequency trajectories, making diffusion approximations ill-suited
+\parencite{Robertson1961-ho,Santiago1995-hx,Buffalo2019-qs}. The stochastic
+perturbations due to selection also impact selected sites, and alter their
+dynamics, creating Hill--Robertson interference \parencite{Hill1966-kd}. Strong
+draft can generate multiple-merger coalescences that create discontinuities in
+allele frequency trajectories, making diffusion approximations ill-suited
 \parencite{Gillespie2000-mh,Der2011-it,Neher2013-dz}. However, when selection
 is weak, draft is well-approximated as an increase in the rate of drift (or
-equivalently, a reduction in effective population size). The stochastic
-perturbations between selecting sites creates Hill--Robertson interference
-\parencite{Hill1966-kd}.
+equivalently, a reduction in effective population size).
 
@@
 
-A recent model from \textcite{Santiago2016-mu} addresses these limitations by
-formulating the negative selection process in terms of the equilibrium fitness
-variance $V_A$ under mutation, drift, and draft.
-
-Since their earlier work has
-related the additive fitness variance in a region to the draft-effective
-population experienced by a neutral allele arbitrarily far away
-\parencite{Santiago1998-bs}, this can be
-
-HERE
-
-that derives the effective population
-size experienced by a neutral allele when there is heritable fitness in the
-population. This theory is in essence a model for the magnitude of stochastic
-perturbations (what we think of as a \emph{draft-effective} population size)
-when heritable fitness variation generates weak genetic draft. As mentioned,
-perturbations due to draft are autocorrelated through time; at the individual
-level Robertson considered, this is because offspring from large families tend
-to beget many descendents themselves (and likewise with small families).
-However, the same autocorrelation occurs at the genomic level with linkage
-\parencite{Santiago1998-bs}, as the perturbations to a neutral allele's
-trajectory from its particular fitness background tend to occur in the same
-direction across generations until recombination.
-
-The benefit of this quantitative genetics formulation of linked selection is
-two-fold. First, as we will see, Santiago and Caballero's
-(\citeyear{Santiago2016-mu}) model for the additive fitness variance under both
-strong and weak negative selection is quite accurate. Second, these models of
-linked selection can 
-They do this through
-
-Thus, the total additional
-heritable variation in offspring number is the product of the equilibrium
-additive fitness variation $V_A$ in a generation and an inflation factor
-$Q_t^2$ due to the build up of autocorrelation to generation $t$. We let $Q^2
-:= Q_t^2$ as $t \to \infty$. Intuitively, the product $V_A Q^2$ represents the
+Our work extends theory from \textcite{Santiago2016-mu}, which suggests both
+strong and weak negative selection can be be accommodated with a quantitative
+genetic model of linked selection. These linked selection models stem from
+\textcite{Robertson1961-ho}, which in essence describes how heritable polygenic
+fitness variation generates weak genetic draft. At the individual level
+Robertson considered, selection generates autocorrelation as offspring from
+large families tend to beget many descendents themselves (and likewise with
+small families) when fitness is heritable. This same autocorrelation occurs at
+the genomic level with linkage \parencite{Santiago1998-bs,Barton2000-zg}, as
+the perturbations to a neutral allele's trajectory from its particular fitness
+background tend to occur in the same direction across generations until the
+background recombines off. These models quantify the total impact of the
+autocorrelation generated by weak draft in terms of what we think of as a
+\emph{draft-effective} population size $N_d$. The key insight is that in the
+long-run, steady, weak draft created by the presence of additive genetic
+fitness variance $V_A > 0$ acts like an extra variance in offspring number
+would under pure drift \parencite{Wright1938-tv}. However, due to
+autocorrelation, the cumulative effect of this fitness variance is inflated by
+by a factor of $Q^2$. Intuitively, the product $V_A Q^2$ represents the
 expected total variance in reproductive success a neutral mutation experiences
@@
 
-We can include this total additional variance into expressions for effective
-population size as we would larger variance in family sizes due to
-non-heritable causes \parencite{Wright1938-tv}. We define the draft-effective
-population size $N_d$ as
+We define the draft-effective population size $N_d$ by including the total
+additional variance created by heritable fitness in Wright's
+(\citeyear{Wright1938-tv}) equation,
 

### 2023-07-12 bgspy/likelihood.py
nb changes, new fits

@@ -753,7 +753,7 @@ class BGSLikelihood:
     def ave_B(self):
         return self.predict(B=True).mean()
 
-    def jackknife_stderr(self, use_loo_chrom=False, trim=None, iqr_factor=2):
+    def jackknife_stderr(self, use_loo_chrom=False, trim=None, iqr_factor=3):
         """
         Calculate the jackknife standard errors.
         """

### 2023-07-12 manuscript/manuscript.tex
fig shuffle

@@
 %
-(c.f. \cite{Robertson1961-ho,Santiago1995-hx}; see Supplementary Section
+(c.f. \cite{Robertson1961-ho,Santiago1995-hx}; see Supplementary Materials Section
 \ref{supp:theory} for a proof). The benefit of modeling draft with Robertson's
@@
     \label{eq:polygenic_Ne}
-    N_d \approx N \exp\left(-\sum_{i=1}^n V_{A,i} \frac{Q_i^2}{2}\right)
+    N_d \approx N \exp\left(-\sum_{i=1}^S V_{A,i} \frac{Q_i^2}{2}\right)
 \end{align}
@@
 where the factor of one-half comes from ignoring weak associations from
-unlinked regions and chromosomes (see Supplementary Section
+unlinked regions and chromosomes (see Supplementary Materials Section
 \ref{supp:heritable-fitness}). In our genome-wide model, we consider the
@@
 \begin{align}
-    N_d = N \exp \left( - \sum_i^n \frac{\mu L_i}{s(1 + r_i(1-s)/s)^2} \right) 
+    N_d = N \exp \left( - \sum_i^S \frac{\mu L_i}{s(1 + r_i(1-s)/s)^2} \right) 
 \end{align}
@@
   \label{eq:main_eqns}
-  N_d &= N \exp \left( -V_a \frac{Q^2}{2} \right) & \text{\emph{draft-effective population equation}} \\
+  N_d &= N \exp \left( -(U-R)s \frac{Q^2}{2} \right) & \text{\emph{draft-effective population equation}} \\
   R &= \frac{4N_d U s}{\exp(4 N_d s)-1}  & \text{\emph{ratchet equation}}
 \end{align}
-
-where $V_a = (U-R)s$ as in Equation \eqref{eq:Va}. We denote the solutions to
-these equation, which represent equilibria under mutation-selection-drift-draft
-process, as $\widetilde{N}_d$ and $\widetilde{R}$. These equilibria also imply
-an equilibrium level of additive fitness variation $\widetilde{V}_a$ in the
-segment, are used to calculate the reduction at an arbitrary position in the
-genome (see Methods
-
-With the equilibrium $V_a$ calculated at each segment for a particular $\mu$
-and $s$, we can estimate the reduction factor $B(x) = \nicefrac{N_d}{N}$ at any
-genomic position $x$ (see Methods XXX).
-
-\section*{Results: Validation of Theory}
-
-Similar to previous work, our method characterizes the negative selection
-process through the signal left on nucleotide diversity at linked sites.
-Although nucleotide diversity is a less-informative statistic than the full
-site frequency spectrum, it is much easier to derive models of how selection
-alters expected pairwise coalescence times at linked sites than to describe the
-entire genealogy. Spatial patterns of pairwise diversity are highly variable
-across eukaryotic genomes, as linked selection leads diversity to become
-correlated with spatial heterogeneity in recombination rates and the density of
-conserved sites. Statistical methods then fit observed diversity to the
-expected level under a particular theoretic model of linked selection
-processes. Consequently, statistical inferences about selection may be
-incorrect when the underlying theory inaccurately predicts average pairwise
-coalescence times.
+%
+
+We denote the solutions to these equation, which represent equilibria under
+mutation-selection-drift-draft process, as $\widetilde{N}_d$ and
+$\widetilde{R}$. These equilibria also imply an equilibrium level of additive
+fitness variation $\widetilde{V}_a$ in the segment, which are used to calculate
+the reduction factor $B(x) = \nicefrac{N_d}{N}$ at any genomic position $x$
+(see Methods \ref{sec:methods-maps}).
+
+\section*{Results: Simulation Validation of Theory and Methods}
+
+Given that modeling the interplay of mutation, drift, and draft under both weak
+and strong selection has proven to be difficult problem, we first sought to
+verify the SC16 theory and our genome-wide implementation of it with a three
+levels of simulations.
 
 \subsection*{Simulations of a Segment under Negative Selection}
+\label{sec:segment-sims}
 
@@
 
-Given that Santiago and Caballero's (\citeyear{Santiago2016-mu}) model is a
-haploid-model for a single region under equilibrium negative selection, we
-first sought to confirm it could accurately predict the reduction in realistic
-forward-time simulations of diploid populations with recombination. Using SLiM
-\parencite{Haller2019-vu,Haller2023-uk} we simulated a region of $10^5$
-basepairs under varying levels of mutation and selection (see Methods
-\ref{sec:methods-sim}). We find a close correspondence between the observed
-and predicted reductions in effective population size $B=\nicefrac{N_d}{N}$
-over all selection and mutation parameters including weak selection (Figure
-\ref{fig:figure-1}A), in contrast to classic BGS theory. Furthermore, to
-investigate whether this accuracy was caused by the model correctly predicting
-the equilibrium fitness variance and ratchet rate, we also measured these
-throughout the simulation. Again, we find the diploid SC16 theory accurately
-predicts both the ratchet rate (Figure 1\ref{fig:figure-1}B) and the genic
-fitness variance (Figure 1\ref{fig:figure-1}C).
+Our first set of forward simulations were to ensure that the SC16 model
+adequately captures selective dynamics in a single $10^5$ basepair region
+undergoing negative selection, across a variety of mutation rates and selection
+coefficients (see Methods \ref{sec:methods-sim}). We find a close
+correspondence between the observed and predicted reductions in effective
+population size $B=\nicefrac{N_d}{N}$ over all selection and mutation
+parameters including weak selection (Figure \ref{fig:figure-1}A), in contrast
+to classic BGS theory. Furthermore, to investigate whether this accuracy was
+caused by the model correctly predicting the equilibrium fitness variance and
+ratchet rate, we also measured these throughout the simulation. Again, we find
+the diploid SC16 theory accurately predicts both the ratchet rate (Figure
+1\ref{fig:figure-1}B) and the genic fitness variance (Figure
+1\ref{fig:figure-1}C).
 
@@
 \subsection*{Chromosome-wide Simulations and Models of Negative Selection}
+\label{sec:chrom-sims}
 
@@
 ratchet rate for a single segment under general mutation-selection processes,
-we extended their model so that it could be applied to whole-genome data. Our
-software method \texttt{bgspy} numerically solves Equations XXX to compute the
-equilibrium additive genic fitness variance ($V_a$) and ratchet rate ($R$)
-across grids of mutation rates and selection coefficients. This is done for
-each pre-specified segment in the genome that may be under negative selection
-(e.g. coding sequences or UTRs). Then, these equilibrium can be used to
-calculate a reduction map $B(x)$ across genomic positions $x$. To distinguish
-between McVicker's B maps based on classic BGS theory, we call our reduction
-map $B(x)$ a B' map.
-
-To assess our predicted B' reduction maps, we simulated negative selection for
-fixed mutation rates and selection coefficient on human chromosome 10 using
-realistic recombination maps and conserved feature positions. Averaging over
-$100$ replicates, we estimated the simulation expected reduction map and
-compared this to the predicted B and B' maps for the same parameters. We find
-that our B' maps and the classic BGS theory B maps closely match simulations
-when selection is strong (top row of Figure \ref{fig:figure-1}E). This provides
-the first realistic forward-time chromosome-scale simulation confirmation of
-classic BGS theory. However, we find slight discrepancies in regions with low
-recombination (Figure \ref{fig:figure-1}E). Second, we find our theory is
-vastly more accurate than the classic B maps when selection is very weak ($2N_e
-s \ll 1$; bottom row of Figure \ref{fig:figure-1}E), which is expected since
-BGS theory does not work in this domain. Across all mutation and selection
-parameters simulated, the relative error of the classic B maps is 14.6\%
-whereas the relative error in the new B' maps is 5\%. Additionally, we find
-that the mean squared error between simulations and B' maps is close to the
-theoretic lower bound set by the coalescence variance \parencite[see Methods
-XXX]{Tajima1983-gu}.
-
-Nearly all of the error between the new B' maps and chromosome-wide simulations
-occurs around the drift-barrier domain of $2Ns \approx 1$ (Figure
-\ref{fig:figure-1}D). The higher error in this domain occurs across all
-mutation rates. We hypothesized that the higher error in this domain may be
-because when our method numerically solves Equations \eqref{eq:main_eqns} for
-each segment, it does not consider the reduction experienced due to selection
-at other linked segments. In particular, we use a fixed drift-effective
-$N=1000$ corresponding to the number of diploids in the simulations rather than
-$B(x) N$ at position $x$. To test this, we implemented a locally-rescaled
-version of the B' maps, which numerically solves Equations \eqref{eq:main_eqns}
-for each segment at position $x$ taking into account the reduction experienced
-due to selection at other segments by setting $B'(x)N$ as the population size
-(see Method XXX). We find the locally-rescaled B' maps reduce the relative
-error from 5\% to 0.4\% and mean squared error (Figure \ref{fig:figure-1}D,
-dashed colored lines), but does not entirely eliminate the error in the $2Ns
-\approx 1$ domain. Overall, while we find evidence that local-rescaling
-slightly reduces error in the $2Ns \approx 1$ domain, we do not use it during
-in our genome-wide inference method. This is because to avoid circularity and
-possible overfitting, the method would need to solve the equilibria equations
-for all segments simultaneously, but this is currently computational
-unfeasible.
+we next extended their model so that it could be applied to whole-genome data.
+Our software method \texttt{bgspy} numerically solves Equations
+\eqref{eq:main_eqns} to compute the equilibrium additive genic fitness variance
+($V_a$) and ratchet rate ($R$) across grids of mutation rates and selection
+coefficients. This is done for each pre-specified segment in the genome that
+may be under negative selection (e.g. coding sequences or UTRs). Then, these
+equilibrium can be used to calculate a reduction map $B(x)$ across genomic
+positions $x$. To distinguish between McVicker's B maps based on classic BGS
+theory, we call our reduction map $B(x)$ a B' map.
+
+We next validated our predicted B' reduction maps with realistic
+chromosome-scale forward simulations of negative selection in conserved
+regions. Averaging over $100$ replicates, we estimated the simulation empirical
+reduction map and compared this to the predicted B and B' maps for the same
+parameters. We find that our B' maps and the classic BGS theory B maps closely
+match simulations when selection is strong (top row of Figure
+\ref{fig:figure-1}E). This provides the first realistic forward-time
+chromosome-scale simulation confirmation of classic BGS theory. However, we
+find slight discrepancies in regions with low recombination (Figure
+\ref{fig:figure-1}E). Second, we find our theory is vastly more accurate than
+the classic B maps when selection is very weak ($2N_e s \ll 1$; bottom row of
+Figure \ref{fig:figure-1}E), which is expected since BGS theory does not work
+in this domain. Across all mutation and selection parameters simulated, the
+relative error of the classic B maps is 14.6\% whereas the relative error in
+the new B' maps is 5\%. Additionally, we find that the mean squared error
+between simulations and B' maps is close to the theoretic lower bound set by
+the coalescence variance \parencite[see Methods XXX]{Tajima1983-gu}.
+
+The discrepancy between our theoretic B' maps and chromosome-wide simulations
+that leads to the 5\% error occurs entirely around the drift-barrier domain of
+$2Ns \approx 1$ (Figure \ref{fig:figure-1}D). The higher error in this domain
+occurs across all mutation rates. We hypothesized this error may be due to
+selective interference between segments that is not taken into account when we
+numerically solve Equations \eqref{eq:main_eqns} for each segment
+independently. In particular, we use a fixed drift-effective $N=1000$
+corresponding to the number of diploids in the simulations rather than $B(x) N$
+at position $x$. To test this, we implemented a locally-rescaled version of the
+B' maps, which numerically solves Equations \eqref{eq:main_eqns} for each
+segment at position $x$ taking into account the reduction experienced due to
+selection at other segments by setting $B'(x)N$ as the population size. We find
+the locally-rescaled B' maps reduce the relative error from 5\% to 0.4\% and
+mean squared error (Figure \ref{fig:figure-1}D, dashed colored lines), but does
+not entirely eliminate the error in the $2Ns \approx 1$ domain. Overall, while
+we find evidence that local-rescaling slightly reduces error in the $2Ns
+\approx 1$ domain, we do not use it during in our genome-wide inference method.
+This is because to avoid circularity and possible overfitting, our method would
+need to solve the equilibria equations for all segments simultaneously;
+unfortunately, this is currently computational unfeasible.
 
@@
 
+\subsection*{Validation of Composite-likelihood Method using Forward
+Simulations}
+
+\begin{figure}[htbp] \centering
+    \includegraphics[width=\textwidth]{figures/figure_2.pdf} 
+    \caption{Comparison of parameter estimates using classic BGS theory (green
+        lines) with our new B' method (blue lines) across both full and sparse
+        track types (dark versus light hue), and different mutation rates
+        (columns). In the first row, $R^2$ between predictions and observations
+        increases with selection intensity. Both classic BGS and B' methods fit
+        the data equally well with sparse tracks, though the estimated
+        selection coefficients are inaccurate when selection is weak (second
+        row). Mutation rate estimates (third row) are more accurately estimated
+        by the B' method than classic BGS, but overall show slight biases.
+        Overall, classic BGS methods fail as expected when full-coverage tracks
+    are used, since it cannot accommodate weak selection and neutrality. }
+  \label{fig:figure-2}
+\end{figure}
+
+We next developed a composite-likelihood method that estimates the distribution
+of selection coefficients for each feature type, the mutation rate, and the
+diversity in the absence of linked selection ($\pi_0$) by fitting the theoretic
+reduction map to genome-wide diversity (see Methods
+\ref{sec:methods-likelihood}). We validated that our method can accurately
+estimate the parameters of negative selection by simulation a ``synthetic
+genome" of the first five human chromosomes (see Methods
+\ref{sec:methods-sim}). We note three findings from these simulations.
+
+First, we find the $R^2$ between predicted and simulated megabase-scale
+diversity serves as a measure of the strength of the linked selection signal in
+genome-wide data. The $R^2$ increases with the intensity of selection against
+new deleterious mutations and mutation rate (Figure \ref{fig:figure-2}, top
+row). Under just drift or weak negative selection, the variance in diversity is
+driven by unstructured coalescence noise along the genome and the predicted
+reduction map $B(x)$ does not fit the data well. However, we note that under
+very strong selection ($s=0.05$), $R^2$ is reduced. This is like due to very
+strong selection having less localized effects and impacting overall
+genome-wide diversity \parencite{Santiago1995-hx,Robertson1961-ho}.
+
+Second, both our implementation of classic BGS theory and our B' method
+accurately infer the average selection coefficient under strong selection
+(Figure \ref{fig:figure-2}, middle row). However, when selection was weak, the
+classic BGS model erroneously estimates strong selection and a very low
+mutation rate. By contrast, our B' method estimated selection coefficients much
+closer to their true value. A minor discrepancy occurs around $2Ns = 1$, likely
+due to the sensitivity of mutations in this region to selective interference
+(discussed in Section \ref{sec:chrom-sims}). Because we only simulated fixed
+selection coefficients and five chromosomes, we only assessed the accuracy of
+average selection coefficients and not the DFE probabilities.
+
+Third, we find slight biases in mutation rated estimated by both our B' and the
+classic BGS methods (Figure \ref{fig:figure-2}, bottom row). However, mutation
+rate estimates based on our B' method are more accurate than classic BGS theory
+across a range of selection coefficients. Overall, the bias in estimated
+mutation rates suggests that benchmarking genome-wide negative selection models
+based on their agreement with pedigree-based rate estimates may not be
+appropriate. When BGS is not functioning, either due to weak selection or a low
+rate of deleterious mutations (Figure \ref{fig:figure-2}, right column), all
+estimates deteriorate. This is understandable, as the overall signal from
+linked selection weakens relative to drift-based noise. We should note, though,
+that this is an unlikely region of parameter space and this issue can be
+readily diagnosed from the low $R^2$ values.
+
+Finally, we tested the sensitivity of our method to violations of two
+assumptions---constant population size and recessivity of selective
+effects---through simulations. Using an expansion rate and timing as estimated
+for Europeans by \textcite{Gutenkunst2009-pg}, we found that a population
+expansion had minimal impact on parameter estimates (Supplementary Materials
+Section \ref{supp:sim-assum}). We did not test the influence of population
+bottlenecks, since parameter estimates of out-of-Africa bottlenecked
+populations (CHB and CEU) did not differ much from YRI estimates (see below).
+However, these simulations did find that selection coefficient estimates were
+downwardly biased when deleterious mutations were recessive and weak.
+Recessivity did not bias selection coefficient estimates when selection was
+strong; this is expected since strongly deleterious mutations are predominantly
+are sufficiently rare that they are not often found in a homozygote. Overall,
+the sensitivity to recessivity is likely due to the ratchet rate being
+overestimated when selection is weak and recessive mutations drift up to higher
+frequency and experience stronger than expected selection in homozygotes.
+
 \section*{Results: Application to Human Genomic Data}
@@
 effects for each feature type. We consider two classes of annotation models:
-(1) CADD-based models, which consider the top $x\%$ of basepairs according to
-the CADD pathogenicity score assigned to each basepair, and (2) and more
-interpretable gene feature-based models that includes protein coding regions,
-intron and UTRs, and PhastCons regions. We include PhastCons regions because
-there is strong evidence of highly-conserved non-coding regions
+(1) CADD-based models, which consider the top $x\%$ most pathogenic basepairs
+according to the CADD score, and (2) and more interpretable gene feature-based
+models that includes protein coding regions, intron and UTRs, and PhastCons
+regions. We include PhastCons regions because there is strong evidence of
+highly-conserved non-coding regions
 \parencite{Meader2010-hm,Harmston2013-tt,Katzman2007-gq,Siepel2005-wh}, that
@@
 that is annotated as both PhastCons and coding sequence) must be assigned to
-one category or the other. Since this assignment impact DFE estimates, we fit
-both of the two alternatives. First, a \emph{PhastCons Priority} model, where
-genic features that overlap PhastCons regions are classified as PhastCons, and
-all remaining coding basepairs are labeled as CDS. Second, a \emph{Feature
-Priority} model, where all coding basepairs are assigned to CDS, and the
-PhastCons class catches the remaining highly-conserved non-genic regions.
+one category or the other. Since this assignment impacts DFE estimates, we fit
+both of the two alternative models. First, a \emph{PhastCons Priority} model,
+where genic features that overlap PhastCons regions are classified as
+PhastCons, and all remaining coding basepairs are labeled as CDS. Second, a
+\emph{Feature Priority} model, where all coding basepairs are assigned to CDS,
+and the PhastCons class catches the remaining highly-conserved non-genic
+regions.
 
@@
 \begin{figure}[htbp] \centering
-    \includegraphics[width=\textwidth]{figures/figure_2.pdf} 
+    \includegraphics[width=\textwidth]{figures/figure_3.pdf}
+
     \caption{The $R^2$ estimates for a sparse (A) and full (B) models, for all
         populations (colors) fit at the megabase-scale. Round points are our B'
-    method and diamonds are the classic B. Lighter color round points are
-    the out-sample $R_\text{LOCO}^2$ estimates for our B' method, and arrows show the decline
-    in goodness-of-fit due to in-sample overfitting (out-sample $R_\text{LOCO}^2$ 
-    were not calculated for classic B values due to computational costs). The 
-    horizontal dashed lines are the $R_\text{drift}^2$ expected when the residual
-    variance is given by the theoretic variance in coalescence times due to drift alone.}
+        method and diamonds are the classic BGS (we exclude classic BGS in the
+        full track subfigure, since these all fit very poorly). Lighter color
+        round points are the out-sample $R_\text{LOCO}^2$ estimates for our B'
+        method, and arrows show the decline in goodness-of-fit due to in-sample
+        overfitting (out-sample $R_\text{LOCO}^2$ were not calculated for
+        classic B values due to computational costs). The horizontal dashed
+        lines are the $R_\text{drift}^2$ expected when the residual variance is
+    given by the theoretic variance in coalescence times due to drift alone.}
+
   \label{fig:figure-2}
@@
 out-sample diversity levels slightly better than previous methods. For these
-two models, we find that our B' method predicts $R_\text{LOCO}^2=68.45$\% and
-$R_\text{LOCO}^2=68.04$\% of the out-sample variance in Yoruba pairwise
+two models, we find that our B' method predicts $R_\text{LOCO}^2=67.3$\% and
+$R_\text{LOCO}^2=68.7$\% of the out-sample variance in Yoruba pairwise
 diversity at the megabase-scale, respectively. By contrast, the best-fitting
 CADD 6\% model from \textcite{Murphy2022-sj} explained 60\% of diversity in
-left-out 2 Mbp windows across YRI samples. We note that this might be explained
-by other differences in data process, reference genome, etc.
+left-out 2 Mbp windows across YRI samples. We note that this difference could
+be explained by other differences in data processing, optimization, etc.
 
 In populations impacted by the out-of-Africa bottleneck, the goodness-of-fit
-was lower across all models (e.g. 62.98\% and 59.98\% for CEU and CHB
-respectively in the PhastCons Priority model). Our PhastCons Priority model
-fits marginally better than the CADD 6\% model consistently across all three
+was lower across all models (e.g. 60.4\% and 58.7\% for CEU and CHB
+respectively in the CADD 6\% model).
+
+Our PhastCons Priority model fits
+marginally better than the CADD 6\% model consistently across all three
 populations (Supplementary Table \ref{supp:tbl-r2}). This is likely because the
@@
 \begin{figure}[htbp] \centering
-    \includegraphics[width=\textwidth]{figures/figure_3.pdf} 
+    \includegraphics[width=\textwidth]{figures/figure_4.pdf} 
     \caption{The distribution of fitness effects of new mutations estimates for
@@
 diversity in the absence of linked selection diversity $\pi_0$, the mutation
-rate $\mu$, and the distribution of fitness effects $\mathbf{W}$ across the
-selection grid for each of the $K$ feature types. Given that the relationship
-between $\pi$ and the strength selection is U-shaped, our model in theory could
-fit under different combinations of parameters.
+rate $\mu$, and the matrix of distribution of fitness effects $\mathbf{W}$
+across the selection grid for each of the $K$ feature types. Given that the
+relationship between $\pi$ and the strength selection is U-shaped, we wondered
+whether the linked selection signal in humans could fit under a weak selection
+regime.
 
@@
 coding sequences, PhastCons regions, or the top 6\% of basepairs with the most
-pathogenic CADD scores. However, we find that there is considerable uncertainty
+pathogenic CADD scores.
+
+However, we find that there is considerable uncertainty
 in how strong selection is against the most deleterious mutations. For example,
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/figure_4.pdf} 
+    \includegraphics[width=\textwidth]{figures/figure_5.pdf} 
     \caption{(A) }
@@
 
-\subsection*{Tests of Parameter Inference using Forward Simulations}
-
-Given these the sensitivity of mutation rate estimates and the strong-selection
-non-identifiability issue, we next sought to assess the estimation overall
-accuracy of our composite likelihood method using forward simulations. In order
-to have enough megabase-scale data to apply our method, we independently
-evolved the first five human chromosomes and combined them into a ``synthetic"
-genome. Due to the immense computational expense of conducting forward
-simulations at this scale over numerous parameter combinations, we used fixed
-mutation and selection parameters in putatively conserved regions (see Methods
-XXX). We note three important findings from these simulations.
-
-First, the linked selection signal, as measured by the $R^2$ between predicted
-and simulated megabase-scale diversity, increases with the intensity of
-selection against new deleterious mutations and mutation rate (Figure
-\ref{fig:figure-4}B, top row). This shows the ``strength" of the linked
-selection signal varies as expected with these key parameters. However, if new
-mutations are highly deleterious ($s=0.05$), $R^2$ is reduced. This is
-consistent with very strong selection having less localized effects and acting
-more like a genome-wide faster rate of drift.
-
-Second, both classic BGS theory and our B' method accurately infer the average
-selection coefficient under strong selection (Figure \ref{fig:figure-4}B,
-middle row). However, when selection was weak, the classic BGS model
-erroneously estimates strong selection and a very low mutation rate. By
-contrast, our B' method estimated selection coefficients much closer to their
-true value. A minor discrepancy occurs around $2Ns = 1$, for reasons discussed
-in Section XXX. Because we only simulated fixed selection coefficients and five
-chromosomes, we only assessed the accuracy of average selection coefficients
-and not the DFE probabilities.
-
-Third, with both our B' and the classic BGS models, we find mutation rate
-estimates show slight biases in the strong-selection regime (Figure
-\ref{fig:figure-3}B, bottom row). Still, our B' method leads to more accurate
-mutation rate estimates than classic BGS theory across a variety of selection
-coefficients. These biases in mutation rate estimates suggest that benchmarking
-genome-wide models via their concordance with pedigree-based rate estimates may
-not be suitable. When BGS is not acting, either due to weak selection or a low
-rate of deleterious mutations (Figure \ref{fig:figure-4}B right column), all
-estimates are worse. This makes intuitive sense; the overall linked selection
-signal is weaker relative to drift-based noise. However, we note that this is
-an unlikely region of parameter space and this issue is readily diagnosable
-from the low $R^2$ values.
-
-Finally, we also used simulations to test the sensitivity of our method to
-violations of two assumptions: constant population size and recessivity of
-selective effects. Since we find little difference in parameter estimates among
-YRI and bottlenecked out-of-Africa (CEU and CHB) samples, we tested whether a
-population expansion would bias our method. We use an expansion rate and timing
-based on those estimated for Europeans by \textcite{Gutenkunst2009-pg}, finding
-they had little effect on parameter estimates (Supplementary Section
-\ref{supp:sim-assum}). By contrast, we find our estimates are downwardly biased
-when mutational effects are weak and recessive, but are accurate under
-recessive strong selection. This occurs because the fixation probabilities of
-recessive mutations are in reality much lower than under our model that assumes
-additivity.
-
 \subsection*{Predicted Diversity and Residual Signal}
@@
 \begin{figure}[htbp] \centering
-    \includegraphics[width=\textwidth]{figures/figure_5.pdf}
+    \includegraphics[width=\textwidth]{figures/figure_6.pdf}
 
@@
 \begin{figure}[htbp] \centering
-    \includegraphics[width=\textwidth]{figures/figure_6.pdf}
+    \includegraphics[width=\textwidth]{figures/figure_7.pdf}
 
@@
 recombination rates in the segment. A detailed derivation of these equations
-can be found in Supplementary Section \ref{supp:theory}. The recombination
+can be found in Supplementary Materials Section \ref{supp:theory}. The recombination
 rate in a segment is determined by a user-supplied recombination map.
@@
 diversity. To calculate uncertainty, we used a block jackknife approach in 10
-Mbp windows (Supplementary Section \ref{supp:jackknife}).
+Mbp windows (Supplementary Materials Section \ref{supp:jackknife}).
 

### 2023-07-12 manuscript/supplementary.tex
fig shuffle

@@
 \centering
-\caption{$R^2$ and mutation rate estimates for all primary models (top table)
-    and models fit on a smaller DFE grid (excluding the possibility of strong
-    selection $s=10^{-1}$). Note that the repeated values in the last rows are
+
+\caption{$R^2$ and mutation rate estimates for all primary models fit on the
+    default DFE grid (up to selection $s=10^{-1}$).}
+
+\begin{tabular}{lll|crr|cc}
+    \textbf{Model} & \textbf{Track type} & \textbf{Pop.} & \textbf{B' $R_\text{LOCO}^2$} & \textbf{B' $R_\text{IS}^2$} & \textbf{B $R_\text{IS}^2$} & \textbf{B' $\hat{\mu} \times 10^{-8}$} & \textbf{B $\hat{\mu} \times 10^{-8}$} \\[0.5ex]
+\hline
+\hline
+PhastCons Priority &            sparse &          YRI &                        67.29 &             67.36 &            62.82 &                                 2.0039 &                                0.2070 \\
+PhastCons Priority &              full &          YRI &                        67.27 &             67.36 &            58.26 &                                 2.0037 &                                0.1037 \\
+          CADD 6\% &            sparse &          YRI &                        66.66 &             66.81 &            66.79 &                                 1.6565 &                                1.6328 \\
+          CADD 6\% &              full &          YRI &                        66.50 &             66.68 &            58.76 &                                 1.5556 &                                0.1074 \\
+          CADD 8\% &            sparse &          YRI &                        65.30 &             65.58 &            65.56 &                                 1.2229 &                                1.1997 \\
+          CADD 8\% &              full &          YRI &                        65.13 &             65.41 &            59.16 &                                 1.1260 &                                0.1067 \\
+  Feature Priority &            sparse &          YRI &                        63.61 &             64.28 &            60.43 &                                 3.3528 &                                0.2259 \\
+  Feature Priority &              full &          YRI &                        63.06 &             64.28 &            56.08 &                                 3.3490 &                                0.1000 \\
+PhastCons Priority &            sparse &          CEU &                        61.04 &             62.25 &            58.37 &                                 2.3133 &                                0.2160 \\
+PhastCons Priority &              full &          CEU &                        61.04 &             62.25 &            53.95 &                                 1.9117 &                                0.1044 \\
+          CADD 6\% &            sparse &          CEU &                        60.37 &             61.80 &            61.79 &                                 1.6797 &                                1.6670 \\
+          CADD 6\% &              full &          CEU &                        60.24 &             61.77 &            53.99 &                                 1.6347 &                                0.1096 \\
+          CADD 8\% &            sparse &          CEU &                        59.38 &             60.92 &            60.91 &                                 1.2547 &                                1.2418 \\
+          CADD 8\% &              full &          CEU &                        59.24 &             60.88 &            54.31 &                                 1.2078 &                                0.1090 \\
+PhastCons Priority &            sparse &          CHB &                        58.84 &             60.45 &            56.55 &                                 1.9202 &                                0.2240 \\
+PhastCons Priority &              full &          CHB &                        58.76 &             60.45 &            52.93 &                                 1.8689 &                                0.1101 \\
+          CADD 6\% &            sparse &          CHB &                        58.67 &             60.49 &            60.48 &                                 1.7454 &                                1.7334 \\
+          CADD 6\% &              full &          CHB &                        58.48 &             60.40 &            52.76 &                                 1.5981 &                                0.1166 \\
+          CADD 8\% &            sparse &          CHB &                        57.91 &             59.78 &            59.76 &                                 1.3093 &                                1.3022 \\
+          CADD 8\% &              full &          CHB &                        57.74 &             59.69 &            53.14 &                                 1.2038 &                                0.1161 \\
+  Feature Priority &            sparse &          CEU &                        57.24 &             59.11 &            56.35 &                                 3.0442 &                                0.2313 \\
+  Feature Priority &              full &          CEU &                        57.14 &             59.11 &            52.21 &                                 3.0441 &                                0.1004 \\
+  Feature Priority &            sparse &          CHB &                        56.00 &             57.88 &            54.76 &                                 3.1722 &                                0.2362 \\
+  Feature Priority &              full &          CHB &                        55.63 &             57.88 &            51.39 &
+\end{tabular}
+\end{table}
+
+\begin{table}
+    \label{supp:tbl-r2-strongsel}
+\centering
+\caption{$R^2$ and mutation rate estimates for all primary models fit 
+    with the strong selection grid (up to $s=10^{-1}$). Note that the repeated values in the last rows are
     correct and due to rounding; see \path{notebooks/main_fits.ipynb} in the
@@
 \hline
-    \textbf{phastcons$>$CDS$>$genes} &            sparse &          YRI &                        \textbf{68.45} &             68.16 &            65.23 &                                 1.8636 &                                0.2975 \\
-phastcons$>$CDS$>$genes &              full &          YRI &                        68.19 &             68.12 &            63.67 &                                 1.7082 &                                0.1666 \\
+PhastCons Priority &            sparse &          YRI &                        68.45 &             68.16 &            65.23 &                                 1.8636 &                                0.2975 \\
+PhastCons Priority &              full &          YRI &                        68.19 &             68.12 &            63.67 &                                 1.7082 &                                0.1666 \\
           CADD 6\% &              full &          YRI &                        68.08 &             68.08 &            62.96 &                                 1.4612 &                                0.1721 \\
@@
           CADD 8\% &            sparse &          YRI &                        66.67 &             67.01 &            66.98 &                                 1.5688 &                                1.5473 \\
-CDS$>$genes$>$phastcons &              full &          YRI &                        64.90 &             65.51 &            61.23 &                                 4.1951 &                                0.1620 \\
-CDS$>$genes$>$phastcons &            sparse &          YRI &                        64.90 &             65.51 &            63.16 &                                 4.1795 &                                0.3070 \\
-\textbf{phastcons$>$CDS$>$genes} &            sparse &          CEU &                        \textbf{61.98} &             63.02 &            60.02 &                                 2.0194 &                                0.3036 \\
-phastcons$>$CDS$>$genes &              full &          CEU &                        61.78 &             63.01 &            58.57 &                                 2.3499 &                                0.1679 \\
+  Feature Priority &              full &          YRI &                        64.90 &             65.51 &            61.23 &                                 4.1951 &                                0.1620 \\
+  Feature Priority &            sparse &          YRI &                        64.90 &             65.51 &            63.16 &                                 4.1795 &                                0.3070 \\
+PhastCons Priority &            sparse &          CEU &                        61.98 &             63.02 &            60.02 &                                 2.0194 &                                0.3036 \\
+PhastCons Priority &              full &          CEU &                        61.78 &             63.01 &            58.57 &                                 2.3499 &                                0.1679 \\
           CADD 6\% &              full &          CEU &                        61.74 &             62.94 &            57.42 &                                 1.5365 &                                0.1739 \\
@@
           CADD 6\% &            sparse &          CHB &                        59.54 &             61.21 &            61.20 &                                 2.1242 &                                2.1170 \\
-               \textbf{phastcons$>$CDS$>$genes} &            sparse &          CHB &                        \textbf{59.44} &             61.08 &            57.88 &                                 2.1982 &                                0.3060 \\
-phastcons$>$CDS$>$genes &              full &          CHB &                        59.29 &             61.08 &            56.71 &                                 2.5573 &                                0.1694 \\
+PhastCons Priority &            sparse &          CHB &                        59.44 &             61.08 &            57.88 &                                 5.8874 &                                0.3060 \\
+PhastCons Priority &              full &          CHB &                        59.29 &             61.08 &            56.71 &                                 2.5573 &                                0.1694 \\
           CADD 8\% &              full &          CHB &                        58.88 &             60.63 &            55.86 &                                 1.1128 &                                0.1733 \\
           CADD 8\% &            sparse &          CHB &                        58.65 &             60.57 &            60.55 &                                 1.6098 &                                1.6001 \\
-CDS$>$genes$>$phastcons &            sparse &          CEU &                        58.58 &             60.30 &            58.41 &                                 4.1041 &                                0.3110 \\
-CDS$>$genes$>$phastcons &              full &          CEU &                        58.58 &             60.30 &            56.75 &                                 4.1001 &                                0.1648 \\
-CDS$>$genes$>$phastcons &            sparse &          CHB &                        56.48 &             58.73 &            56.60 &                                 3.7195 &                                0.3136 \\
-CDS$>$genes$>$phastcons &              full &          CHB &                        56.48 &             58.73 &            54.95 &                                 3.7552 &                                0.1654 \\
+  Feature Priority &            sparse &          CEU &                        58.58 &             60.30 &            58.41 &                                 4.1041 &                                0.3110 \\
+  Feature Priority &              full &          CEU &                        58.58 &             60.30 &            56.75 &                                 4.1001 &                                0.1648 \\
+  Feature Priority &            sparse &          CHB &                        56.48 &             58.73 &            56.60 &                                 3.7195 &                                0.3136 \\
+  Feature Priority &              full &          CHB &                        56.48 &             58.73 &            54.95 &                                 3.7552 &                                0.1654 \\
 \hline
-\hline
-\emph{constrained DFE grid} & & & & & & \\
-\hline
-\textbf{phastcons$>$CDS$>$genes}  &              full &          YRI &               \textbf{67.27} &             67.36 &             &                                 2.0333 &                                \\
-               CADD 6\%  &            sparse &          YRI &                        66.66 &             66.81 &             &                                 1.6565 &                                \\
-               CADD 6\%  &              full &          YRI &                        66.50 &             66.68 &             &                                 1.5556 &                                \\
-CDS$>$genes$>$phastcons  &              full &          YRI &                        63.61 &             64.28 &             &                                 3.3493 &                                \\
-\textbf{phastcons$>$CDS$>$genes}  &              full &          CEU &               \textbf{61.04} &             62.25 &             &                                 1.9128 &                                \\
-               CADD 6\%  &            sparse &          CEU &                        60.37 &             61.80 &             &                                 1.6797 &                                \\
-               CADD 6\%  &              full &          CEU &                        60.24 &             61.77 &             &                                 1.6347 &                                \\
-               \textbf{phastcons$>$CDS$>$genes}  &     full &          CHB &                        \textbf{58.76} &             60.45 &             &                                 1.8691 &                                \\
-               CADD 6\%  &            sparse &          CHB &                        58.67 &             60.49 &             &                                 1.7454 &                                \\
-               CADD 6\%  &              full &          CHB &                        58.48 &             60.40 &             &                                 1.5981 &                                \\
-CDS$>$genes$>$phastcons  &              full &          CEU &                        57.24 &             59.11 &             &                                 3.0461 &                                \\
-CDS$>$genes$>$phastcons  &              full &          CHB &                        55.63 &             57.88 &             &                                 3.1723 &                                \\
 \end{tabular}

### 2023-07-12 manuscript/manuscript.tex
ms updates

@@
 \begin{figure}[htbp] \centering
-    \includegraphics[width=\textwidth]{figures/figure_2.pdf} 
+    \includegraphics[width=0.8\textwidth]{figures/figure_2.pdf} 
     \caption{Comparison of parameter estimates using classic BGS theory (green
@@
 two models, we find that our B' method predicts $R_\text{LOCO}^2=67.3$\% and
-$R_\text{LOCO}^2=68.7$\% of the out-sample variance in Yoruba pairwise
+$R_\text{LOCO}^2=66.7$\% of the out-sample variance in Yoruba pairwise
 diversity at the megabase-scale, respectively. By contrast, the best-fitting
@@
 left-out 2 Mbp windows across YRI samples. We note that this difference could
-be explained by other differences in data processing, optimization, etc.
-
-In populations impacted by the out-of-Africa bottleneck, the goodness-of-fit
-was lower across all models (e.g. 60.4\% and 58.7\% for CEU and CHB
-respectively in the CADD 6\% model).
-
-Our PhastCons Priority model fits
-marginally better than the CADD 6\% model consistently across all three
-populations (Supplementary Table \ref{supp:tbl-r2}). This is likely because the
-PhastCons Priority model has more DFE parameters, and thus can fit finer-scale
-differences among the four feature types. By contrast, the CADD 6\% track must
-fit patterns of diversity to conserved sites under a single DFE. The CADD 8\%
-model fit slightly worse than the CADD 6\% across all populations, and the
-Feature Priority model had the poorest-fit.
+be explained by other differences in data processing, optimization, etc. For
+populations impacted by the out-of-Africa bottleneck, the goodness-of-fit was
+lower across all models (e.g. 61.0\% and 58.8\% for CEU and CHB respectively in
+the PhastCons Priority model).
+
+Our PhastCons Priority model fits marginally better than the CADD 6\% model
+consistently across all three populations (Supplementary Table
+\ref{supp:tbl-r2}). This is likely because the PhastCons Priority model has
+more DFE parameters, and thus can fit finer-scale differences among the four
+feature types. By contrast, the CADD 6\% track must fit patterns of diversity
+to conserved sites under a single DFE. The CADD 8\% model fit slightly worse
+than the CADD 6\% across all populations, and the Feature Priority model had
+the poorest-fit.
 
@@
 ``sparse") we fit a corresponding ``full" model that assigns the remaining
-complement of the genome to a feature called ``other" that contains most of the
-genome. Ideally, how a model fits using our B' method should be invariant to
-whether an annotation model is sparse or full, since our method in principle
-can accommodate weak selection and neutrality. Indeed, we find this to be the
-case: both in-sample $R^2$ and out-sample $R_\text{LOCO}^2$ values are nearly
-identical across full and sparse-track models (Figures \ref{fig:figure-2}A and
-\ref{fig:figure-2}B, round points).
-
-By contrast, full annotation models fit poorly under classic BGS theory (Figure
-\ref{fig:figure-2}B, diamond-shaped points) and lead to unreasonable parameter
-estimates (discussed in Section XXX). Additionally, when sparse annotation
-models contain genomic features that are likely under weak constraint (such as
-introns and UTRs), models fit worse under classic BGS theory than our B' method
-(Figure \ref{fig:figure-2}A). However, among the CADD annotation models, the
-goodness-of-fit is nearly identical between B' and classic BGS methods. This
-behavior is what we would expect given that the CADD models contain only the
-most pathogenic sites, which are \emph{a priori} very likely under the strong
-selection domain under which B' and classic BGS theory agree.
+portion of the genome to a feature called ``other". Ideally, how a model fits
+using our B' method should be invariant to whether an annotation model is
+sparse or full, since our method in principle can accommodate weak selection
+and neutrality. Indeed, we find this to be the case: both in-sample $R^2$ and
+out-sample $R_\text{LOCO}^2$ values are nearly identical across full and
+sparse-track models (Figures \ref{fig:figure-2}A and \ref{fig:figure-2}B, round
+points).
+
+By contrast, full annotation models fit poorly under classic BGS theory, and
+lead to unreasonable parameter estimates (discussed in Section XXX).
+Additionally, when sparse annotation models contain genomic features that are
+likely under weak constraint (such as introns and UTRs), models fit worse under
+classic BGS theory than our B' method (Figure \ref{fig:figure-2}A). However,
+among the CADD annotation models, the goodness-of-fit is nearly identical
+between B' and classic BGS methods. This behavior is what we would expect given
+that the CADD models contain only the most pathogenic sites, which are \emph{a
+priori} very likely under the strong selection domain under which B' and
+classic BGS theory agree.
 
 Overall, our $R_\text{LOCO}^2$ estimates suggest our negative selection model
-explains up to 68\% of out-sample variance in diversity of the megabase scale,
+explains up to 67\% of out-sample variance in diversity of the megabase scale,
 even though our method assumes constant demography and homogeneous mutation
@@
 We find that our out-sample $R_\text{LOCO}^2$ for the Yoruba samples
-($R_\text{LOCO}^2 \approx 68$\%) is slightly above the theoretic
-$R_\text{coal}^2 \approx 67$\%. This suggests our model is close to the
-vicinity of fitting all the signal possible, under the coalescence-only noise
-assumption. By contrast, for bottlenecked out-of-Africa populations, we find a
-larger discrepancy between $R_\text{coal}^2$ and observed out-sample
-$R_\text{LOCO}^2$. The theoretic $R_\text{coal}^2 \approx 64$\% for both
-populations, compared to the observed $R_\text{LOCO}^2 \approx 62$\% for CEU
-and $R_\text{LOCO}^2 \approx 59$\% for CHB. Given that bottlenecks would act to
-increase the residual variance in coalescence times beyond the level implied by
-the effective population size, the gap would likely shrink under more realistic
-models or simulation-based approximations for $R_\text{coal}^2$. Overall, this
-suggests that negative selection models explain the vast majority coalescence
-time variation at the megabase-scale that is capable of being explained (i.e.
-that is not coalescence noise).
+($R_\text{LOCO}^2 \approx 67$\%) is slightly above the theoretic
+$R_\text{coal}^2 \approx 66.6$\%. This suggests our model is in the vicinity of
+fitting all the signal possible, under the coalescence-only noise assumption.
+By contrast, for bottlenecked out-of-Africa populations, we find a larger
+discrepancy between $R_\text{coal}^2$ and observed out-sample
+$R_\text{LOCO}^2$. The theoretic $R_\text{coal}^2 \approx 63\%-64$\% for both
+populations, compared to the observed $R_\text{LOCO}^2 \approx 61$\% for CEU
+and $R_\text{LOCO}^2 \approx 59$\% for CHB under the PhastCons Priority model.
+Given that bottlenecks would act to increase the residual variance in
+coalescence times beyond the level implied by the effective population size,
+this gap would likely shrink under more realistic models or simulation-based
+approximations for $R_\text{coal}^2$. Overall, this suggests that negative
+selection models explain the vast majority coalescence time variation at the
+megabase-scale that is capable of being explained (i.e. that is not coalescence
+noise).
 
@@
 
-Our composite likelihood method has three sets of parameters: expected
+Our composite likelihood method has three sets of parameters: the expected
 diversity in the absence of linked selection diversity $\pi_0$, the mutation
@@
 relationship between $\pi$ and the strength selection is U-shaped, we wondered
-whether the linked selection signal in humans could fit under a weak selection
-regime.
+whether our new B' model accommodating weak selection could fit the linked
+selection signal under a different combination of weak and strong selection
+parameters than found by previous work.
 
 However, across all of our annotation models, new deleterious mutations in
-conserved feature classes were consistently estimated to have strongly
-deleterious effects (Figure \ref{fig:figure-3}A), consistent with previous work
-\parencite{McVicker2009-ax,Murphy2022-sj}. This includes features such as
-coding sequences, PhastCons regions, or the top 6\% of basepairs with the most
-pathogenic CADD scores.
-
-However, we find that there is considerable uncertainty
-in how strong selection is against the most deleterious mutations. For example,
-in the CADD 6\% full-track model, new mutations are estimated to have a
-selection coefficients of $s = 10^{-1}$ with 42\% chance, $s=10^{-2}$ with
-7.8\% chance, and $s=10^{-3}$ with 41\% chance. This creates a bimodal DFE
-(Figure \ref{fig:figure-3}A, second row, left column), which is mirrored across
-the other CADD models. Moreover, these DFE estimates suggest a very high
-average selection coefficient across CADD 6\% regions of ${\bar{s}}
-\approx 0.043$ for YRI.
+conserved feature classes (CADD tracks, CDS, and PhastCons regions) were
+consistently estimated to have strongly deleterious effects (Figure
+\ref{fig:figure-4}A), consistent with previous work
+\parencite{McVicker2009-ax,Murphy2022-sj}. The DFE estimates for CADD and
+PhastCons regions consistently places $\ge 75\%$ of mass on the largest
+selection coefficient, $s=10^{-2}$. The CADD 6\% DFE estimates imply an average
+selection coefficient of $\bar{s} = 0.65\%$ for CEU, $\bar{s} = 0.57\%$ for
+CHB, and $\bar{s} = 0.79\%$ for YRI. Similarly, the PhastCons Priority model
+implies average selection coefficient estimates of $\bar{s} = 0.63\%$, $\bar{s}
+= 0.58\%$, and $\bar{s}= 0.77\%$ for CEU, CHB, and YRI respectively for
+PhastCons regions.
+
+Following previous work, our method used a grid of selection coefficients up to
+$s=10^{-2}$. However, we also experimented with a strong selection grid that
+includes $s=10^{-1}$. We find that models fit with the strong selection grid
+had $R_\text{LOCO}^2$ that were about one percentage point higher, which is
+suggestive of stronger selection than has previously been estimated using
+constrained grids (see Supplementary Materials Table
+\ref{supp:tbl-r2-strongsel}). For this strong selection grid, we estimate
+average selection coefficients for the CADD 6\% model of $\bar{s} = 4.3\%$ for
+CEU, $3.2\%$ for CHB, and $4.4\%$ for YRI. Thus, across different selection
+coefficient grids, there is considerable uncertainty in the average strength of
+selection against deleterious mutations in conserved regions.
+
+However, across the strong selection grid runs,
+
+However, we find that there is considerable uncertainty in how strong selection
+is against the most deleterious mutations. For example, in the CADD 6\%
+full-track model, new mutations are estimated to have a selection coefficients
+of $s = 10^{-1}$ with 42\% chance, $s=10^{-2}$ with 7.8\% chance, and
+$s=10^{-3}$ with 41\% chance. This creates a bimodal DFE (Figure
+\ref{fig:figure-3}A, second row, left column), which is mirrored across the
+other CADD models. Moreover, these DFE estimates suggest a very high average
+selection coefficient across CADD 6\% regions of ${\bar{s}} \approx 0.043$ for
+YRI.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/figure_5.pdf} 
-    \caption{(A) }
+    \includegraphics[width=0.5\textwidth]{figures/figure_5.pdf}
+
+    \caption{Mutation rate estimates across sparse track (top row) and
+        full-coverage track (bottom row) models, for the new B' (circles) and
+        classic BGS (diamonds) methods. Estimates of the mutation rate are
+        consistent between classic BGS and B' methods for sparse tracks CADD
+        models (overlapping diamonds and circles, top row). Overall, mutation
+        rate estimates are sensitive to the underlying annotation model.}
   \label{fig:figure-4}

### 2023-07-13 manuscript/manuscript.tex
ms updates

@@
     \includegraphics[width=\textwidth]{figures/figure_4.pdf}
+
     \caption{The distribution of fitness effects of new mutations estimates for
-        Yoruba individuals. (A) The DFEs using sparse (left column) and
+        YRI reference samples. (A) The DFEs using sparse (left column) and
         full-coverage (right column) tracks, across different annotation models
         (row). Color indicates the feature type. (B) The DFE of the
-        full-coverage CDS-priority model comparing the estimates across
-    populations.}
+    full-coverage Feature Priority model comparing the estimates across
+populations. Although this model fit the data less well than alternatives, its
+results are more interpretable.}
+
   \label{fig:figure-3}
@@
 = 0.58\%$, and $\bar{s}= 0.77\%$ for CEU, CHB, and YRI respectively for
-PhastCons regions.
+PhastCons regions. Our DFE estimates for the Feature Priority model are
+qualitatively consistent with the U-shaped DFEs found for amino acids through
+Poisson Random Field method \parencite{Boyko2008-tj}.
 
@@
 
-However, across the strong selection grid runs,
-
-However, we find that there is considerable uncertainty in how strong selection
-is against the most deleterious mutations. For example, in the CADD 6\%
-full-track model, new mutations are estimated to have a selection coefficients
-of $s = 10^{-1}$ with 42\% chance, $s=10^{-2}$ with 7.8\% chance, and
-$s=10^{-3}$ with 41\% chance. This creates a bimodal DFE (Figure
-\ref{fig:figure-3}A, second row, left column), which is mirrored across the
-other CADD models. Moreover, these DFE estimates suggest a very high average
-selection coefficient across CADD 6\% regions of ${\bar{s}} \approx 0.043$ for
-YRI.
-
-Given that previous methods have typically not estimated the DFE up to
-selection coefficients of $s=10^{-1}$, we ran additional ``constrained grid"
-models using a DFE only up to $s=10^{-2}$. Across all populations, these
-constrained grid models fit the data slightly less well than the default grid
-by about one percentage point less in $R_\text{LOCO}^2$ (see Supplementary
-Table \ref{supp:tbl-r2}). As expected, the constrained grid models distributed
-mass onto the next strongest selection coefficient, $s=10^{-2}$, and the
-average selection was reduced about six-fold to ${\bar{s}} \approx 0.0079$ for
-YRI. Our constrained-grid DFE estimates for the Feature Priority model are
-qualitatively consistent with the U-shaped DFEs found for amino acids through
-Poisson Random Field method \parencite{Boyko2008-tj}. While the DFE may truly
-be bimodal for highly-conserved regions and have a very high average selection
-coefficient, a likely alternative is that these results reflect a
-non-identifiability issue between selective effects of $s=10^{-1}$ and
-$s=10^{-2}$. We will refer to this as the \emph{strong-selection
-non-identifiability} issue.
-
-We propose that one mechanism for this non-identifiability is that mutations
-extremely harmful to fitness are difficult to distinguish from a smaller
-effective population. Intuitively, as $s$ increases, the average persistence
-time (in generations) of the mutation in the population goes below one
-generation, i.e. death of the carrier of the \emph{de novo} mutation. In this
-hypothetical limit, each occurrence of a very harmful mutation nearly acts like
-an instantaneous death that in effect is just a smaller population size. One
-way to test this hypothesis is to look to see if there is a systematic positive
+However, across the strong selection grid runs, we find indications of model
+non-identifiability. First, estimates of the DFE with the strong grid are
+strangely bimodal (Supplementary Figure \ref{suppfig:strong-sel-dfe}). For
+example, under the CADD 6\% strong selection grid model, new mutations are
+estimated to have a selection coefficients of $s = 10^{-1}$ with 43\% chance,
+$s=10^{-2}$ with 7.8\% chance, and $s=10^{-3}$ with 41\% chance. We propose
+that one mechanism for this non-identifiability is that mutations extremely
+harmful to fitness are difficult to distinguish from a smaller effective
+population. Intuitively, as $s$ increases, the average persistence time (in
+generations) of the mutation in the population goes below one generation, i.e.
+death of the carrier of the \emph{de novo} mutation. In this hypothetical
+limit, each occurrence of a very harmful mutation nearly acts like an
+instantaneous death that in effect is just a smaller population size. One way
+to test this hypothesis is to look to see if there is a systematic positive
 relationship across models between average selection coefficient and $\pi_0$,
@@
 population, average selection was about $7.1$ times larger using the default
-grid, and $\pi_0$ was $5.6\%$ larger. There was no similar consistent change in
-mutation rate estimates among populations. In the CADD 6\% model, genome-wide
-average reduction factor $\bar{B}$ was $\approx 6.1\%$ lower in the default
-versus constrained grid. Overall, this suggests that the linked selection
-signal alone cannot differentiate very strong selection from a slightly smaller
+grid, and $\pi_0$ was $5.6\%$ larger (see Supplementary Materials Figure
+\ref{suppfig:sel-ident}). There was no similar consistent change in mutation
+rate estimates among populations. In the CADD 6\% model, genome-wide average
+reduction factor $\bar{B}$ was $\approx 6.1\%$ lower in the default versus
+constrained grid. Overall, this suggests that the linked selection signal alone
+cannot differentiate very strong selection from a slightly smaller
 drift-effective population size.
 
-Despite the strong-selection non-identifiability issue, we were curious how
-stable these estimates are across different reference populations. Overall, the
-DFE estimates were consistent over all populations and models (Supplementary
-Materials Section \ref{supp:fits}). Here, we highlight the Feature Priority
-model since though it is poor-fitting, the DFE estimates are more readily
-interpretable since it is based on gene features. Figure \ref{fig:figure-3}B
-shows consistent DFE estimates across populations, with narrow uncertainty as
-estimated by a block jackknife procedure. Coding sequences have a bimodal DFE
-consistent with the strong-selection non-identifiability issue. Additionally,
-coding sequences have a very uncertain neutral mode in the DFE. This is
-expected given the different level of selective constraint between synonymous
-and non-synonymous sites.
-
-Additionally, the PhastCons class of this Feature Priority model, which
-contains highly conserved non-coding elements, is under strong selective
-constraint as expected. Finally, we highlight one result from our PhastCons
+Given that it is debated how strongly demography impacts the deleterious
+mutation load
+\parencite{Torres2018-ni,Torres2020-hc,Lohmueller2008-qi,Simons2014-sv,Simons2016-cs},
+we were curious how consistent our DFE estimates are across populations.
+Overall, we find DFE estimates are relatively stable across populations and
+models (Supplementary Materials Section \ref{supp:fits}). Only in our Feature
+Priority model (Figure \ref{fig:figure-4}B, top row) do we see a slightly
+different DFE estimate for coding sequences between YRI and CEU/CHB
+populations, but this could be due to the poorer fit this model has to data.
+
+Although the Feature Priority model fits the data less well than alternative
+models, its features are more interpretable. We find that our B' method
+estimates a bimodal DFE for coding sequences for the Feature Priority model,
+with a large mass placed on $10^{-3} \le s \le 10^{-2}$ and another on the
+neutral class $s=10^{-8}$. This is expected, given that the synonymous and
+non-synonymous sites that constitute coding sequences are under vastly
+different levels of constraint. Moreover, features expected to be only weakly
+constrained such as introns and UTRs have the bulk of DFE mass on the neutral
+class, with a small but significant amount of mass ($\approx 3\%$) placed on
+$s=10^{-2}$. As expected, the DFE for PhastCons regions suggests it is under
+strong selective constraint; however we note that block jackknife based
+uncertainty estimates suggest the model is uncertain whether there is some mass
+on the neutral class. Finally, we highlight one result from our PhastCons
 Priority annotation model (Figure \ref{fig:figure-3}A bottom row): the DFE

### 2023-07-13 manuscript/supplementary.tex
ms updates

@@
 
-\subsection{Strong Selection Non-Identifiability}
-\label{supp:}
+\subsection{Strong Selection Grids}
+\label{supp:strong-sel-grid}
+
+\begin{figure}[htbp]
+    \label{suppfig:strong-sel-dfe}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/supplementary/strongsel_grid_dfe.pdf}
+  \caption{The DFE estimates for the strong selection grid (up to $s=10^{-1}$).}
+\end{figure}
 
@@
   \includegraphics[width=0.8\textwidth]{figures/supplementary/ave_sel_vs_pi0.pdf}
-  \caption{XXX}
+
+  \caption{The maximum likelihood estimates of $\pi_0$ and average selection
+  coefficient implied by the estimated DFE for the CADD 6\% models. Diamonds
+indicate estimates under the strong selection grid (up to $s=10^{-1}$) and
+circles indicate estimates under the default grid (up to $s=10^{-2}$).}
+
 \end{figure}
@@
 
-\subsection{Phastcons Priority, Full Track}
-\subsubsection*{CEU}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-number jackknife samples: 280
-standard error method: moving block jackknife
-negative log-likelihood: 197825873893.4953
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.0011279 (0.00113, 0.00113)
-pi  = 0.000801495
-mut. rate = 2.35e-08  (1.98e-08, 2.72e-08)
-Ne = 28,197 (if mu=1e-8), Ne = 14,098 (if mu=2e-8)
-R2 = 63.0106% (in-sample)  61.7759% (out-sample)
-W = 
-        CDS                      gene                     other                    phastcons
-------  -----------------------  -----------------------  -----------------------  -----------------------
-1e-08   0.000 (-0.0410, 0.0410)  0.901 (0.6521, 1.1503)   0.984 (0.9739, 0.9938)   0.034 (-2.8217, 2.8898)
-1e-07   0.000 (-0.1314, 0.1314)  0.026 (-0.2223, 0.2745)  0.000 (-0.0096, 0.0096)  0.376 (-2.4935, 3.2456)
-1e-06   0.000 (-0.0088, 0.0088)  0.000 (-0.0018, 0.0027)  0.000 (-0.0001, 0.0001)  0.001 (-0.0308, 0.0318)
-1e-05   0.000 (-0.0002, 0.0002)  0.019 (0.0133, 0.0249)   0.000 (-0.0000, 0.0000)  0.000 (-0.0003, 0.0003)
-0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
-0.001   0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.472 (0.4050, 0.5388)
-0.01    0.000 (-0.0001, 0.0001)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.046 (0.0351, 0.0571)
-0.1     1.000 (0.8620, 1.1380)   0.053 (0.0454, 0.0608)   0.016 (0.0138, 0.0185)   0.071 (0.0571, 0.0858)
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{CHB}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-number jackknife samples: 280
-standard error method: moving block jackknife
-negative log-likelihood: 61821783504.67356
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00106467 (0.00106, 0.00107)
-pi  = 0.000752601
-mut. rate = 2.557e-08  (-8.06e-10, 5.2e-08)
-Ne = 26,616 (if mu=1e-8), Ne = 13,308 (if mu=2e-8)
-R2 = 61.0847% (in-sample)  59.2923% (out-sample)
-W = 
-        CDS                      gene                     other                    phastcons
-------  -----------------------  -----------------------  -----------------------  -----------------------
-1e-08   0.000 (-0.5839, 0.5839)  0.931 (0.6381, 1.2238)   0.985 (0.9661, 1.0044)   0.478 (-2.5565, 3.5128)
-1e-07   0.000 (-0.2506, 0.2506)  0.005 (-0.2858, 0.2949)  0.000 (-0.0165, 0.0165)  0.002 (-3.0192, 3.0225)
-1e-06   0.000 (-0.0119, 0.0119)  0.000 (-0.0043, 0.0045)  0.000 (-0.0002, 0.0002)  0.000 (-0.0172, 0.0177)
-1e-05   0.000 (-0.0001, 0.0001)  0.013 (-0.0013, 0.0268)  0.000 (-0.0000, 0.0000)  0.000 (-0.0002, 0.0002)
-0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
-0.001   0.000 (-0.0000, 0.0000)  0.003 (0.0005, 0.0049)   0.000 (-0.0000, 0.0000)  0.430 (0.1533, 0.7071)
-0.01    0.000 (-0.0001, 0.0001)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.090 (0.0316, 0.1477)
-0.1     1.000 (0.3574, 1.6426)   0.049 (0.0174, 0.0804)   0.015 (0.0053, 0.0243)   0.000 (-0.0005, 0.0005)
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{YRI}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-number jackknife samples: 280
-standard error method: moving block jackknife
-negative log-likelihood: 250656989872.19766
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00147277 (0.00147, 0.00147)
-pi  = 0.00106447
-mut. rate = 1.708e-08  (-4.08e-08, 7.49e-08)
-Ne = 36,819 (if mu=1e-8), Ne = 18,409 (if mu=2e-8)
-R2 = 68.1177% (in-sample)  68.1947% (out-sample)
-W = 
-        CDS                      gene                     other                    phastcons
-------  -----------------------  -----------------------  -----------------------  -----------------------
-1e-08   1.000 (-3.4930, 5.4930)  0.902 (0.6716, 1.1325)   0.979 (0.9416, 1.0156)   0.000 (-1.2400, 1.2407)
-1e-07   0.000 (-4.4936, 4.4936)  0.000 (-0.1820, 0.1820)  0.000 (-0.0194, 0.0194)  0.004 (-0.5459, 0.5537)
-1e-06   0.000 (-0.0107, 0.0107)  0.000 (-0.0024, 0.0024)  0.000 (-0.0003, 0.0003)  0.001 (-0.0353, 0.0365)
-1e-05   0.000 (-0.0001, 0.0001)  0.042 (-0.0177, 0.1024)  0.000 (-0.0000, 0.0000)  0.000 (-0.0005, 0.0006)
-0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0001, 0.0001)
-0.001   0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.523 (-0.2146, 1.2611)
-0.01    0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.212 (-0.0866, 0.5104)
-0.1     0.000 (-0.0006, 0.0006)  0.056 (-0.0228, 0.1339)  0.021 (-0.0088, 0.0516)  0.260 (-0.1070, 0.6270)
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsection{Phastcons Priority, Sparse Track}
-\subsubsection*{CEU}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-standard error method: None
-negative log-likelihood: 197826612684.87143
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00112308
-pi  = 0.000801495
-mut. rate = 2.019e-08 
-Ne = 28,076 (if mu=1e-8), Ne = 14,038 (if mu=2e-8)
-R2 = 63.0152% (in-sample)  61.9834% (out-sample)
-W = 
-          CDS    gene    phastcons
-------  -----  ------  -----------
-1e-08   0       0.913        0.227
-1e-07   0       0            0.004
-1e-06   0.004   0            0
-1e-05   0       0.019        0
-0.0001  0       0            0
-0.001   0       0            0.546
-0.01    0       0            0.061
-0.1     0.996   0.069        0.161
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{CHB}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-standard error method: None
-negative log-likelihood: 61822016128.79776
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00106069
-pi  = 0.000752601
-mut. rate = 5.887e-08 
-Ne = 26,517 (if mu=1e-8), Ne = 13,258 (if mu=2e-8)
-R2 = 61.0804% (in-sample)  59.4419% (out-sample)
-W = 
-          CDS    gene    phastcons
-------  -----  ------  -----------
-1e-08   0.617   0.968        0.744
-1e-07   0.01    0            0.02
-1e-06   0       0            0
-1e-05   0       0.005        0
-0.0001  0       0            0
-0.001   0       0.001        0.185
-0.01    0       0            0.051
-0.1     0.373   0.026        0
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{YRI}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-standard error method: None
-negative log-likelihood: 250657908564.79294
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00146801
-pi  = 0.00106447
-mut. rate = 1.864e-08 
-Ne = 36,700 (if mu=1e-8), Ne = 18,350 (if mu=2e-8)
-R2 = 68.1582% (in-sample)  68.4462% (out-sample)
-W = 
-          CDS    gene    phastcons
-------  -----  ------  -----------
-1e-08       1   0.907        0
-1e-07       0   0.001        0
-1e-06       0   0            0
-1e-05       0   0.035        0
-0.0001      0   0            0
-0.001       0   0            0.477
-0.01        0   0            0.199
-0.1         0   0.057        0.323
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsection{Feature Priority, Full Track}
-\subsubsection*{CEU}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-number jackknife samples: 457
-standard error method: moving block jackknife
-negative log-likelihood: 197848627167.32043
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00115698 (0.00116, 0.00116)
-pi  = 0.000801495
-mut. rate = 4.1e-08  (2.2e-08, 6e-08)
-Ne = 28,924 (if mu=1e-8), Ne = 14,462 (if mu=2e-8)
-R2 = 60.3026% (in-sample)  58.5808% (out-sample)
-W = 
-        CDS                      gene                     other                    phastcons
-------  -----------------------  -----------------------  -----------------------  -----------------------
-1e-08   0.107 (-1.0244, 1.2375)  0.946 (0.7974, 1.0947)   1.000 (0.9991, 1.0009)   0.000 (-0.2751, 0.2757)
-1e-07   0.000 (-1.0319, 1.0319)  0.001 (-0.1448, 0.1458)  0.000 (-0.0009, 0.0009)  0.000 (-0.2482, 0.2482)
-1e-06   0.000 (-0.0087, 0.0087)  0.000 (-0.0013, 0.0021)  0.000 (-0.0000, 0.0000)  0.000 (-0.0150, 0.0150)
-1e-05   0.000 (-0.0001, 0.0001)  0.023 (0.0088, 0.0368)   0.000 (-0.0000, 0.0000)  0.000 (-0.0003, 0.0003)
-0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
-0.001   0.144 (0.0831, 0.2059)   0.014 (0.0080, 0.0196)   0.000 (-0.0000, 0.0000)  0.229 (0.1327, 0.3256)
-0.01    0.000 (-0.0012, 0.0012)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.086 (0.0462, 0.1253)
-0.1     0.749 (0.4358, 1.0622)   0.016 (0.0095, 0.0233)   0.000 (-0.0000, 0.0000)  0.685 (0.3987, 0.9709)
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{CHB}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-number jackknife samples: 454
-standard error method: moving block jackknife
-negative log-likelihood: 61827828856.16418
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00108991 (0.00109, 0.00109)
-pi  = 0.000752601
-mut. rate = 3.755e-08  (1.55e-08, 5.96e-08)
-Ne = 27,247 (if mu=1e-8), Ne = 13,623 (if mu=2e-8)
-R2 = 58.7328% (in-sample)  56.4805% (out-sample)
-W = 
-        CDS                      gene                     other                    phastcons
-------  -----------------------  -----------------------  -----------------------  -----------------------
-1e-08   0.088 (-0.9233, 1.0989)  0.940 (0.7665, 1.1139)   1.000 (0.9988, 1.0012)   0.006 (-0.3608, 0.3730)
-1e-07   0.010 (-0.8568, 0.8763)  0.000 (-0.1685, 0.1691)  0.000 (-0.0011, 0.0011)  0.003 (-0.2610, 0.2674)
-1e-06   0.000 (-0.0075, 0.0078)  0.000 (-0.0021, 0.0021)  0.000 (-0.0000, 0.0000)  0.000 (-0.0145, 0.0151)
-1e-05   0.000 (-0.0001, 0.0001)  0.023 (0.0037, 0.0422)   0.000 (-0.0000, 0.0000)  0.000 (-0.0003, 0.0003)
-0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
-0.001   0.156 (0.0732, 0.2384)   0.017 (0.0081, 0.0260)   0.000 (-0.0000, 0.0000)  0.244 (0.1185, 0.3696)
-0.01    0.000 (-0.0139, 0.0140)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.254 (0.1215, 0.3871)
-0.1     0.746 (0.3590, 1.1338)   0.020 (0.0094, 0.0298)   0.000 (-0.0000, 0.0000)  0.492 (0.2392, 0.7448)
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{YRI}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-number jackknife samples: 280
-standard error method: moving block jackknife
-negative log-likelihood: 250679665159.79324
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00151818 (0.00152, 0.00152)
-pi  = 0.00106447
-mut. rate = 4.195e-08  (2.76e-08, 5.63e-08)
-Ne = 37,954 (if mu=1e-8), Ne = 18,977 (if mu=2e-8)
-R2 = 65.5097% (in-sample)  64.9049% (out-sample)
-W = 
-        CDS                      gene                     other                    phastcons
-------  -----------------------  -----------------------  -----------------------  -----------------------
-1e-08   0.364 (-0.4599, 1.1874)  0.931 (0.8360, 1.0260)   1.000 (0.9993, 1.0007)   0.003 (-0.1973, 0.2040)
-1e-07   0.001 (-0.7930, 0.7942)  0.000 (-0.0901, 0.0901)  0.000 (-0.0007, 0.0007)  0.000 (-0.1669, 0.1675)
-1e-06   0.000 (-0.0038, 0.0040)  0.000 (-0.0009, 0.0009)  0.000 (-0.0000, 0.0000)  0.000 (-0.0106, 0.0109)
-1e-05   0.000 (-0.0001, 0.0001)  0.039 (0.0240, 0.0532)   0.000 (-0.0000, 0.0000)  0.000 (-0.0002, 0.0002)
-0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
-0.001   0.091 (0.0617, 0.1212)   0.011 (0.0077, 0.0150)   0.000 (-0.0000, 0.0000)  0.151 (0.1030, 0.1986)
-0.01    0.000 (-0.0001, 0.0001)  0.000 (-0.0001, 0.0001)  0.000 (-0.0000, 0.0000)  0.328 (0.2249, 0.4316)
-0.1     0.544 (0.3725, 0.7157)   0.019 (0.0130, 0.0250)   0.000 (-0.0000, 0.0000)  0.517 (0.3548, 0.6796)
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsection{Feature Priority, Sparse Track}
-\subsubsection*{CEU}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-number jackknife samples: 27
-standard error method: moving block jackknife
-negative log-likelihood: 197848626646.91736
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00115698 (0.00116, 0.00116)
-pi  = 0.000801495
-mut. rate = 4.104e-08  (3.87e-08, 4.34e-08)
-Ne = 28,924 (if mu=1e-8), Ne = 14,462 (if mu=2e-8)
-R2 = 60.3027% (in-sample)  58.5823% (out-sample)
-W = 
-        CDS                      gene                     phastcons
-------  -----------------------  -----------------------  -----------------------
-1e-08   0.107 (0.0490, 0.1649)   0.947 (0.9402, 0.9539)   0.001 (-0.0478, 0.0491)
-1e-07   0.000 (-0.0273, 0.0279)  0.000 (-0.0052, 0.0052)  0.000 (-0.0073, 0.0075)
-1e-06   0.000 (-0.0003, 0.0006)  0.000 (-0.0001, 0.0001)  0.001 (-0.0008, 0.0019)
-1e-05   0.000 (-0.0000, 0.0000)  0.023 (0.0187, 0.0268)   0.000 (-0.0000, 0.0000)
-0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
-0.001   0.144 (0.1365, 0.1522)   0.014 (0.0130, 0.0146)   0.229 (0.2150, 0.2428)
-0.01    0.000 (-0.0001, 0.0001)  0.000 (-0.0000, 0.0000)  0.086 (0.0733, 0.0981)
-0.1     0.748 (0.7078, 0.7887)   0.016 (0.0155, 0.0173)   0.684 (0.6463, 0.7219)
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{CHB}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-number jackknife samples: 26
-standard error method: moving block jackknife
-negative log-likelihood: 61827828699.43107
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00108991 (0.00109, 0.00109)
-pi  = 0.000752601
-mut. rate = 3.719e-08  (3.2e-08, 4.24e-08)
-Ne = 27,247 (if mu=1e-8), Ne = 13,623 (if mu=2e-8)
-R2 = 58.7329% (in-sample)  56.4805% (out-sample)
-W = 
-        CDS                      gene                     phastcons
-------  -----------------------  -----------------------  -----------------------
-1e-08   0.089 (-0.0565, 0.2340)  0.938 (0.9196, 0.9573)   0.000 (-0.1223, 0.1223)
-1e-07   0.000 (-0.0861, 0.0861)  0.001 (-0.0154, 0.0180)  0.000 (-0.0145, 0.0146)
-1e-06   0.000 (-0.0005, 0.0010)  0.000 (-0.0001, 0.0002)  0.000 (-0.0009, 0.0009)
-1e-05   0.000 (-0.0000, 0.0000)  0.023 (0.0203, 0.0260)   0.000 (-0.0000, 0.0000)
-0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
-0.001   0.157 (0.1380, 0.1766)   0.017 (0.0151, 0.0194)   0.246 (0.2161, 0.2767)
-0.01    0.000 (-0.0004, 0.0006)  0.000 (-0.0000, 0.0000)  0.257 (0.2252, 0.2883)
-0.1     0.754 (0.6611, 0.8461)   0.020 (0.0173, 0.0222)   0.497 (0.4358, 0.5578)
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{YRI}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-number jackknife samples: 17
-standard error method: moving block jackknife
-negative log-likelihood: 250679664591.4161
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00151817 (0.00152, 0.00152)
-pi  = 0.00106447
-mut. rate = 4.179e-08  (3.84e-08, 4.52e-08)
-Ne = 37,954 (if mu=1e-8), Ne = 18,977 (if mu=2e-8)
-R2 = 65.5099% (in-sample)  64.9049% (out-sample)
-W = 
-        CDS                      gene                     phastcons
-------  -----------------------  -----------------------  -----------------------
-1e-08   0.360 (0.3063, 0.4146)   0.931 (0.9239, 0.9375)   0.000 (-0.0647, 0.0647)
-1e-07   0.002 (-0.0234, 0.0265)  0.000 (-0.0051, 0.0052)  0.000 (-0.0097, 0.0097)
-1e-06   0.000 (-0.0003, 0.0003)  0.000 (-0.0000, 0.0001)  0.000 (-0.0009, 0.0009)
-1e-05   0.000 (-0.0000, 0.0000)  0.039 (0.0359, 0.0417)   0.000 (-0.0000, 0.0000)
-0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
-0.001   0.092 (0.0849, 0.0987)   0.011 (0.0105, 0.0122)   0.151 (0.1401, 0.1627)
-0.01    0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.329 (0.3048, 0.3540)
-0.1     0.546 (0.5054, 0.5870)   0.019 (0.0176, 0.0205)   0.519 (0.4804, 0.5580)
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsection{CADD 6\%, Full Track}
-\subsubsection*{CEU}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-number jackknife samples: 280
-standard error method: moving block jackknife
-negative log-likelihood: 197826953189.02365
-number of successful starts: 20000 (100.0% total)
-pi_0 = 0.00109999 (0.00110, 0.00110)
-pi  = 0.000801495
-mut. rate = 1.537e-08  (8e-09, 2.27e-08)
-Ne = 27,499 (if mu=1e-8), Ne = 13,749 (if mu=2e-8)
-R2 = 62.9394% (in-sample)  61.7433% (out-sample)
-W = 
-        cadd6                    other
-------  -----------------------  -----------------------
-1e-08   0.000 (-0.2792, 0.2795)  0.965 (0.9473, 0.9832)
-1e-07   0.000 (-0.2355, 0.2355)  0.000 (-0.0109, 0.0109)
-1e-06   0.000 (-0.0371, 0.0371)  0.000 (-0.0001, 0.0001)
-1e-05   0.000 (-0.0011, 0.0013)  0.000 (-0.0000, 0.0000)
-0.0001  0.026 (0.0129, 0.0393)   0.000 (-0.0000, 0.0000)
-0.001   0.555 (0.3257, 0.7834)   0.000 (-0.0000, 0.0000)
-0.01    0.000 (-0.0001, 0.0001)  0.000 (-0.0000, 0.0000)
-0.1     0.419 (0.2460, 0.5921)   0.035 (0.0204, 0.0491)
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{CHB}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-number jackknife samples: 280
-standard error method: moving block jackknife
-negative log-likelihood: 61821323176.26884
-number of successful starts: 20000 (100.0% total)
-pi_0 = 0.00103624 (0.00104, 0.00104)
-pi  = 0.000752601
-mut. rate = 1.508e-08  (-2.43e-09, 3.26e-08)
-Ne = 25,905 (if mu=1e-8), Ne = 12,952 (if mu=2e-8)
-R2 = 61.3115% (in-sample)  59.6228% (out-sample)
-W = 
-        cadd6                    other
-------  -----------------------  -----------------------
-1e-08   0.000 (-0.6278, 0.6278)  0.961 (0.9187, 1.0036)
-1e-07   0.007 (-0.4928, 0.5076)  0.000 (-0.0227, 0.0227)
-1e-06   0.009 (-0.0393, 0.0570)  0.000 (-0.0002, 0.0002)
-1e-05   0.000 (-0.0014, 0.0015)  0.000 (-0.0000, 0.0000)
-0.0001  0.004 (-0.0082, 0.0171)  0.000 (-0.0000, 0.0000)
-0.001   0.670 (0.0898, 1.2509)   0.000 (-0.0000, 0.0000)
-0.01    0.000 (-0.0001, 0.0001)  0.000 (-0.0000, 0.0000)
-0.1     0.309 (0.0412, 0.5766)   0.039 (0.0051, 0.0726)
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{YRI}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-number jackknife samples: 280
-standard error method: moving block jackknife
-negative log-likelihood: 250659830283.90668
-number of successful starts: 20000 (100.0% total)
-pi_0 = 0.00145893 (0.00146, 0.00146)
-pi  = 0.00106447
-mut. rate = 1.461e-08  (8.61e-09, 2.06e-08)
-Ne = 36,473 (if mu=1e-8), Ne = 18,236 (if mu=2e-8)
-R2 = 68.0839% (in-sample)  68.0751% (out-sample)
-W = 
-        cadd6                    other
-------  -----------------------  -----------------------
-1e-08   0.000 (-0.2318, 0.2327)  0.958 (0.9399, 0.9768)
-1e-07   0.023 (-0.2328, 0.2797)  0.000 (-0.0083, 0.0089)
-1e-06   0.001 (-0.0153, 0.0176)  0.000 (-0.0001, 0.0001)
-1e-05   0.000 (-0.0003, 0.0004)  0.000 (-0.0000, 0.0000)
-0.0001  0.065 (0.0378, 0.0921)   0.000 (-0.0000, 0.0000)
-0.001   0.407 (0.2491, 0.5649)   0.000 (-0.0000, 0.0000)
-0.01    0.078 (0.0439, 0.1126)   0.000 (-0.0000, 0.0000)
-0.1     0.425 (0.2604, 0.5891)   0.041 (0.0253, 0.0573)
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsection{CADD 6\%, Sparse Track}
-\subsubsection*{CEU}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-standard error method: None
-negative log-likelihood: 197829116669.21103
-number of successful starts: 20000 (100.0% total)
-pi_0 = 0.00107623
-pi  = 0.000801495
-mut. rate = 2.121e-08 
-Ne = 26,905 (if mu=1e-8), Ne = 13,452 (if mu=2e-8)
-R2 = 62.8555% (in-sample)  61.6574% (out-sample)
-W = 
-          cadd6
-------  -------
-1e-08     0
-1e-07     0
-1e-06     0
-1e-05     0
-0.0001    0.014
-0.001     0.408
-0.01      0
-0.1       0.578
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{CHB}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-standard error method: None
-negative log-likelihood: 61822132892.58568
-number of successful starts: 20000 (100.0% total)
-pi_0 = 0.00101201
-pi  = 0.000752601
-mut. rate = 2.124e-08 
-Ne = 25,300 (if mu=1e-8), Ne = 12,650 (if mu=2e-8)
-R2 = 61.2064% (in-sample)  59.5382% (out-sample)
-W = 
-          cadd6
-------  -------
-1e-08     0
-1e-07     0
-1e-06     0
-1e-05     0
-0.0001    0
-0.001     0.477
-0.01      0
-0.1       0.523
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{YRI}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-standard error method: None
-negative log-likelihood: 250663447300.65836
-number of successful starts: 20000 (100.0% total)
-pi_0 = 0.00142415
-pi  = 0.00106447
-mut. rate = 2.091e-08 
-Ne = 35,603 (if mu=1e-8), Ne = 17,801 (if mu=2e-8)
-R2 = 68.028% (in-sample)  68.0414% (out-sample)
-W = 
-          cadd6
-------  -------
-1e-08     0
-1e-07     0
-1e-06     0
-1e-05     0
-0.0001    0.038
-0.001     0.302
-0.01      0.033
-0.1       0.627
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsection{CADD 6\%, Full Track}
-\subsubsection*{CEU}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-number jackknife samples: 140
-standard error method: moving block jackknife
-negative log-likelihood: 197834651568.3971
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00109606 (0.00110, 0.00110)
-pi  = 0.000801495
-mut. rate = 1.157e-08  (9.01e-09, 1.41e-08)
-Ne = 27,401 (if mu=1e-8), Ne = 13,700 (if mu=2e-8)
-R2 = 62.1245% (in-sample)  60.8009% (out-sample)
-W = 
-        cadd8                    other
-------  -----------------------  -----------------------
-1e-08   0.000 (-0.0835, 0.0835)  0.952 (0.9415, 0.9624)
-1e-07   0.004 (-0.1451, 0.1523)  0.000 (-0.0043, 0.0043)
-1e-06   0.000 (-0.0145, 0.0150)  0.000 (-0.0001, 0.0001)
-1e-05   0.000 (-0.0005, 0.0005)  0.000 (-0.0000, 0.0000)
-0.0001  0.052 (0.0393, 0.0656)   0.000 (-0.0000, 0.0000)
-0.001   0.537 (0.4316, 0.6417)   0.000 (-0.0000, 0.0000)
-0.01    0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
-0.1     0.407 (0.3276, 0.4867)   0.048 (0.0386, 0.0575)
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{CHB}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-number jackknife samples: 140
-standard error method: moving block jackknife
-negative log-likelihood: 61823378586.32404
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00103245 (0.00103, 0.00103)
-pi  = 0.000752601
-mut. rate = 1.113e-08  (4.22e-09, 1.8e-08)
-Ne = 25,811 (if mu=1e-8), Ne = 12,905 (if mu=2e-8)
-R2 = 60.6265% (in-sample)  58.8768% (out-sample)
-W = 
-        cadd8                    other
-------  -----------------------  -----------------------
-1e-08   0.000 (-0.3716, 0.3716)  0.944 (0.9141, 0.9736)
-1e-07   0.000 (-0.2895, 0.2903)  0.001 (-0.0075, 0.0094)
-1e-06   0.000 (-0.0376, 0.0378)  0.000 (-0.0001, 0.0001)
-1e-05   0.000 (-0.0005, 0.0005)  0.000 (-0.0000, 0.0000)
-0.0001  0.026 (0.0069, 0.0447)   0.000 (-0.0000, 0.0000)
-0.001   0.690 (0.3451, 1.0351)   0.000 (-0.0000, 0.0000)
-0.01    0.000 (-0.0001, 0.0001)  0.000 (-0.0000, 0.0000)
-0.1     0.284 (0.1418, 0.4254)   0.055 (0.0275, 0.0829)
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{YRI}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-number jackknife samples: 140
-standard error method: moving block jackknife
-negative log-likelihood: 250669695762.61414
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00145867 (0.00146, 0.00146)
-pi  = 0.00106447
-mut. rate = 1.013e-08  (5.65e-09, 1.46e-08)
-Ne = 36,466 (if mu=1e-8), Ne = 18,233 (if mu=2e-8)
-R2 = 67.1076% (in-sample)  66.9835% (out-sample)
-W = 
-        cadd8                    other
-------  -----------------------  -----------------------
-1e-08   0.000 (-0.3206, 0.3212)  0.932 (0.9054, 0.9583)
-1e-07   0.002 (-0.1725, 0.1762)  0.000 (-0.0062, 0.0062)
-1e-06   0.000 (-0.0130, 0.0130)  0.000 (-0.0001, 0.0001)
-1e-05   0.000 (-0.0004, 0.0004)  0.000 (-0.0000, 0.0000)
-0.0001  0.103 (0.0627, 0.1425)   0.000 (-0.0000, 0.0000)
-0.001   0.438 (0.2744, 0.6008)   0.000 (-0.0000, 0.0000)
-0.01    0.000 (-0.0003, 0.0003)  0.000 (-0.0000, 0.0000)
-0.1     0.458 (0.2870, 0.6282)   0.068 (0.0427, 0.0936)
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsection{CADD 6\%, Sparse Track}
-\subsubsection*{CEU}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-number jackknife samples: 140
-standard error method: moving block jackknife
-negative log-likelihood: 197836763801.90414
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00107057 (0.00107, 0.00107)
-pi  = 0.000801495
-mut. rate = 1.603e-08  (1.57e-08, 1.63e-08)
-Ne = 26,764 (if mu=1e-8), Ne = 13,382 (if mu=2e-8)
-R2 = 62.061% (in-sample)  60.5937% (out-sample)
-W = 
-        cadd8
-------  -----------------------
-1e-08   0.000 (-0.0171, 0.0171)
-1e-07   0.000 (-0.0007, 0.0007)
-1e-06   0.000 (-0.0000, 0.0000)
-1e-05   0.000 (-0.0000, 0.0000)
-0.0001  0.032 (0.0268, 0.0372)
-0.001   0.400 (0.3894, 0.4110)
-0.01    0.000 (-0.0000, 0.0000)
-0.1     0.568 (0.5575, 0.5781)
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{CHB}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-number jackknife samples: 140
-standard error method: moving block jackknife
-negative log-likelihood: 61824178668.93152
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00100597 (0.00101, 0.00101)
-pi  = 0.000752601
-mut. rate = 1.61e-08  (-4.15e-08, 7.37e-08)
-Ne = 25,149 (if mu=1e-8), Ne = 12,574 (if mu=2e-8)
-R2 = 60.5692% (in-sample)  58.6497% (out-sample)
-W = 
-        cadd8
-------  -----------------------
-1e-08   0.000 (-0.7779, 0.7779)
-1e-07   0.000 (-0.0032, 0.0033)
-1e-06   0.000 (-0.0008, 0.0008)
-1e-05   0.000 (-0.0000, 0.0000)
-0.0001  0.011 (-0.0013, 0.0236)
-0.001   0.492 (0.1090, 0.8748)
-0.01    0.000 (-0.0000, 0.0000)
-0.1     0.497 (0.1101, 0.8838)
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{YRI}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-number jackknife samples: 140
-standard error method: moving block jackknife
-negative log-likelihood: 250673972070.93787
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00141668 (0.00142, 0.00142)
-pi  = 0.00106447
-mut. rate = 1.569e-08  (1.49e-08, 1.64e-08)
-Ne = 35,416 (if mu=1e-8), Ne = 17,708 (if mu=2e-8)
-R2 = 67.0084% (in-sample)  66.6718% (out-sample)
-W = 
-        cadd8
-------  -----------------------
-1e-08   0.000 (-0.0445, 0.0445)
-1e-07   0.000 (-0.0136, 0.0136)
-1e-06   0.000 (-0.0002, 0.0002)
-1e-05   0.000 (-0.0000, 0.0000)
-0.0001  0.059 (0.0526, 0.0655)
-0.001   0.298 (0.2808, 0.3142)
-0.01    0.000 (-0.0000, 0.0000)
-0.1     0.643 (0.6132, 0.6737)
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsection{Phastcons Priority (small DFE grid), Full Track}
-\subsubsection*{CEU}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-standard error method: None
-negative log-likelihood: 197835157685.628
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00105552
-pi  = 0.000801495
-mut. rate = 1.913e-08 
-Ne = 26,387 (if mu=1e-8), Ne = 13,193 (if mu=2e-8)
-R2 = 62.2548% (in-sample)  61.0392% (out-sample)
-W = 
-          CDS    gene    other    phastcons
-------  -----  ------  -------  -----------
-1e-08       1   0.947        1        0.001
-1e-07       0   0            0        0
-1e-06       0   0            0        0
-1e-05       0   0.03         0        0
-0.0001      0   0            0        0
-0.001       0   0            0        0.413
-0.01        0   0.023        0        0.586
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{CHB}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-standard error method: None
-negative log-likelihood: 61824235925.00332
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00100129
-pi  = 0.000752601
-mut. rate = 1.869e-08 
-Ne = 25,032 (if mu=1e-8), Ne = 12,516 (if mu=2e-8)
-R2 = 60.4486% (in-sample)  58.7636% (out-sample)
-W = 
-          CDS    gene    other    phastcons
-------  -----  ------  -------  -----------
-1e-08       1   0.926        1        0
-1e-07       0   0.013        0        0
-1e-06       0   0            0        0
-1e-05       0   0.031        0        0
-0.0001      0   0            0        0
-0.001       0   0            0        0.457
-0.01        0   0.03         0        0.543
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{YRI}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-standard error method: None
-negative log-likelihood: 250667315910.91177
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00139482
-pi  = 0.00106447
-mut. rate = 2.033e-08 
-Ne = 34,870 (if mu=1e-8), Ne = 17,435 (if mu=2e-8)
-R2 = 67.3564% (in-sample)  67.274% (out-sample)
-W = 
-          CDS    gene    other    phastcons
-------  -----  ------  -------  -----------
-1e-08       1   0.956        1        0.005
-1e-07       0   0            0        0.007
-1e-06       0   0            0        0.002
-1e-05       0   0.028        0        0
-0.0001      0   0            0        0.027
-0.001       0   0            0        0.224
-0.01        0   0.016        0        0.735
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsection{Feature Priority (small DFE grid), Full Track}
-\subsubsection*{CEU}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-standard error method: None
-negative log-likelihood: 197860660949.1676
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00109138
-pi  = 0.000801495
-mut. rate = 3.046e-08 
-Ne = 27,284 (if mu=1e-8), Ne = 13,642 (if mu=2e-8)
-R2 = 59.1052% (in-sample)  57.2393% (out-sample)
-W = 
-          CDS    gene    other    phastcons
-------  -----  ------  -------  -----------
-1e-08   0.412   0.887        1        0
-1e-07   0       0            0        0
-1e-06   0       0            0        0
-1e-05   0       0.074        0        0
-0.0001  0       0            0        0
-0.001   0.066   0.008        0        0.136
-0.01    0.521   0.031        0        0.863
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{CHB}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-standard error method: None
-negative log-likelihood: 61830715800.047676
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00103388
-pi  = 0.000752601
-mut. rate = 3.172e-08 
-Ne = 25,847 (if mu=1e-8), Ne = 12,923 (if mu=2e-8)
-R2 = 57.8792% (in-sample)  55.631% (out-sample)
-W = 
-          CDS    gene    other    phastcons
-------  -----  ------  -------  -----------
-1e-08   0.432   0.898        1        0
-1e-07   0       0            0        0
-1e-06   0       0            0        0
-1e-05   0       0.064        0        0
-0.0001  0       0            0        0
-0.001   0.052   0.012        0        0.166
-0.01    0.516   0.026        0        0.834
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{YRI}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-standard error method: None
-negative log-likelihood: 250692005501.01343
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00144148
-pi  = 0.00106447
-mut. rate = 3.349e-08 
-Ne = 36,037 (if mu=1e-8), Ne = 18,018 (if mu=2e-8)
-R2 = 64.2844% (in-sample)  63.6077% (out-sample)
-W = 
-          CDS    gene    other    phastcons
-------  -----  ------  -------  -----------
-1e-08   0.72    0.875        1         0
-1e-07   0.033   0.003        0         0
-1e-06   0       0            0         0
-1e-05   0       0.085        0         0
-0.0001  0       0            0         0
-0.001   0.1     0.003        0         0.07
-0.01    0.148   0.033        0         0.93
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsection{CADD 6\% (small DFE grid), Full Track}
-\subsubsection*{CEU}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-standard error method: None
-negative log-likelihood: 197839911915.8117
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00102951
-pi  = 0.000801495
-mut. rate = 1.635e-08 
-Ne = 25,737 (if mu=1e-8), Ne = 12,868 (if mu=2e-8)
-R2 = 61.767% (in-sample)  60.2433% (out-sample)
-W = 
-          cadd6    other
-------  -------  -------
-1e-08     0.002    0.997
-1e-07     0.006    0
-1e-06     0        0
-1e-05     0        0
-0.0001    0.049    0
-0.001     0.325    0
-0.01      0.617    0.003
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{CHB}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-standard error method: None
-negative log-likelihood: 61824516495.02472
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.000977579
-pi  = 0.000752601
-mut. rate = 1.598e-08 
-Ne = 24,439 (if mu=1e-8), Ne = 12,219 (if mu=2e-8)
-R2 = 60.4032% (in-sample)  58.4847% (out-sample)
-W = 
-          cadd6    other
-------  -------  -------
-1e-08     0        0.992
-1e-07     0.001    0.001
-1e-06     0        0
-1e-05     0        0
-0.0001    0.03     0
-0.001     0.44     0
-0.01      0.529    0.008
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
-\subsubsection*{YRI}
-\begin{minipage}{\linewidth}\begin{footnotesize}
-\begin{verbatim}
-
-ML estimates: (whole genome)
-standard error method: None
-negative log-likelihood: 250675195472.4508
-number of successful starts: 10000 (100.0% total)
-pi_0 = 0.00136706
-pi  = 0.00106447
-mut. rate = 1.556e-08 
-Ne = 34,176 (if mu=1e-8), Ne = 17,088 (if mu=2e-8)
-R2 = 66.6755% (in-sample)  66.5047% (out-sample)
-W = 
-          cadd6    other
-------  -------  -------
-1e-08     0        0.994
-1e-07     0.006    0
-1e-06     0        0
-1e-05     0        0
-0.0001    0.099    0
-0.001     0.115    0
-0.01      0.78     0.006
-\end{verbatim}
-\end{footnotesize}\end{minipage}
-
 \printbibliography

### 2023-07-13 bgspy/likelihood.py
new supp fig

@@ -753,7 +753,7 @@ class BGSLikelihood:
     def ave_B(self):
         return self.predict(B=True).mean()
 
-    def jackknife_stderr(self, use_loo_chrom=False, trim=None, iqr_factor=3):
+    def jackknife_stderr(self, use_loo_chrom=False, trim=None, iqr_factor=5):
         """
         Calculate the jackknife standard errors.
         """

### 2023-07-13 manuscript/manuscript.tex
ms updates

@@
 would under pure drift \parencite{Wright1938-tv}. However, due to
-autocorrelation, the cumulative effect of this fitness variance is inflated by
+autocorrelation, the cumulative effect of this fitness variance is inflated 
 by a factor of $Q^2$. Intuitively, the product $V_A Q^2$ represents the
@@
 
-We define the draft-effective population size $N_d$ by including the total
+We can define the draft-effective population size $N_d$ by including the total
 additional variance created by heritable fitness in Wright's
-(\citeyear{Wright1938-tv}) equation,
+(\citeyear{Wright1938-tv}) equation for effective population size,
 
@@
 Equation \eqref{eq:main_Ne} is general, since different modes of selection and
-linkage can be accommodated by different expressions for the inflation factor
-$Q^2$ \parencite{Santiago1995-hx,Santiago1998-bs}. When fitness variation has a
-multiplicative polygenic basis, as is often assumed for genome-wide negative
-selection processes, the draft asymptotic effective population size experienced
-by an arbitrary neutral site under the influence of all linked regions is,
+linkage can be accommodated by different expressions for $V_A$ and the
+inflation factor $Q^2$ \parencite{Santiago1995-hx,Santiago1998-bs}. When
+fitness variation has a multiplicative polygenic basis, as is often assumed for
+genome-wide negative selection processes, the draft asymptotic effective
+population size experienced by an arbitrary neutral site under the influence of
+all linked regions is,
 
@@
 \end{align}
-
+%
 where the factor of one-half comes from ignoring weak associations from
@@
 
-Under any selection model, the genic fitness variance created by a new mutation
-(at frequency $x=\nicefrac{1}{2N}$) is $2s^2 x(1-x) \approx \nicefrac{s^2}{N}$.
-For the entire population of $2N$ chromosomes, the fresh variance from mutation
-each generation in segment $i$ is $V_{m,i} \approx U_is^2$ where $U_i = 2\mu
-L_i$ is the diploid mutation rate per generation within the segment. Under
-mutation-selection balance assumed by BGS, an $L_i$-basepair segment has genic
-variance $V_{a,i}^{BGS} \approx U_i s$ (see Supplementary Materials Equation
-\ref{eq:va_bgs}) and thus $\kappa_i^{BGS} = s$. Substituting $V_{a,i}^{BGS}$
-and $\kappa_i^{BGS}$ in Equation \eqref{eq:simple_Ne} and simplifying, we have
+Under any selection model, the additive genic fitness variance created by a new
+mutation (at frequency $x=\nicefrac{1}{2N}$) is $2s^2 x(1-x) \approx
+\nicefrac{s^2}{N}$. For the entire population of $2N$ chromosomes, the fresh
+variance from mutation each generation in segment $i$ is $V_{m,i} \approx
+U_is^2$ where $U_i = 2\mu L_i$ is the diploid mutation rate per generation
+within the segment. Under mutation-selection balance assumed by BGS, an
+$L_i$-basepair segment has genic variance $V_{a,i}^{BGS} \approx U_i s$ (see
+Supplementary Materials Equation \ref{supp-eq:va_bgs}) and thus $\kappa_i^{BGS}
+= s$. Substituting $V_{a,i}^{BGS}$ and $\kappa_i^{BGS}$ in Equation
+\eqref{eq:polygenic_Ne} and simplifying, we have
 
@@
 and strong selection has proven to be difficult problem, we first sought to
-verify the SC16 theory and our genome-wide implementation of it with a three
-levels of simulations.
+verify the SC16 theory and our genome-wide extension of this theory with three
+levels of simulations: forward simulations of negative selection in a region,
+chromosome-scale forward simulations of negative selection, and simulations of
+a ``synthetic genome" to test our composite-likelihood method based on this
+theory.
 
@@
         (colored lines corresponding to mutation rate) compared to average
-        reduction across XXX simulation replicates (points). (B) The predicted
+        reduction across 10,000 simulation replicates (points). (B) The predicted
         ratchet rate under the SC16 model scaled by mutation rate (colored
@@
 \subsection*{Annotation Model Comparison}
+\label{sec:annotation}
 
@@
 variance that is capable of being explained, assuming this coalescence-only
-noise process (see Method XXX). We note that selection is expected to
-\emph{decrease} the variance in coalescence times beyond a rescaled effective
-population size implies; our $R_\text{coal}^2$ would be an underestimate under
-selection.
+noise process (see Supplementary Materials \ref{supp:r2-coal}). We note that
+selection is expected to \emph{decrease} the variance in coalescence times
+beyond a rescaled effective population size implies; our $R_\text{coal}^2$
+would be an underestimate under selection.
 
@@
 
-Early genome-wide classic BGS models fit well, but lead to unusually high
-estimates of the mutation rate \parencite{McVicker2009-ax}. This lead to the
-hypothesis that these models could be absorbing the signal of positive
-selection \parencite{Enard2014-kz}, though other work has found a limited role
-for hitchhiking \parencite{Pickrell2009-tt,Hernandez2011-gs,Murphy2022}. Still,
-a central benchmark of fitting genome-wide models is that mutation rate
-estimates are in accordance with pedigree-based estimates (XXX).
+Early genome-wide implementations of the classic BGS model fit well, but lead
+to unusually high estimates of the mutation rate \parencite{McVicker2009-ax}.
+This lead to the hypothesis that these models could be absorbing the signal of
+positive selection \parencite{Enard2014-kz}, though other work has found a
+limited role for hitchhiking
+\parencite{Pickrell2009-tt,Hernandez2011-gs,Murphy2022}. While our simulation
+results suggest estimates of the mutation rate from linked selection models are
+biased, we still our estimates were in rough accordance with pedigree-based
+estimates \parencite{Kong2013-fc,Tian2019-so}.
 
 We find across all populations, our mutation rate estimates from CADD-based
-models are roughly concordant with pedigree-based estimates (Figure
+models are roughly consistent with pedigree-based estimates (Figure
 \ref{fig:figure-4}A), consistent with recent work \parencite{Murphy2022-sj}.
 Our full-track CADD 6\% model estimates the mutation rate as $\widehat{\mu} =
-1.46 \times 10^{-8}$ for YRI, $1.53 \times 10^{-8}$ for CEU, and $1.51 \times
+1.56 \times 10^{-8}$ for YRI, $1.63 \times 10^{-8}$ for CEU, and $1.60 \times
 10^{-8}$ for CHB reference samples (Supplementary Table \ref{supp:tbl-r2}). As
@@
 
-However, mutation rate estimates for feature-based annotation models show
-several issues. First, mutation rate estimates under our implementation of
-classic BGS theory are an order of magnitude below the expected range (Figure
-\ref{fig:figure-4}A top row). We observe similar behavior when we use the
-classic BGS model to fit full-coverage annotation models (Figure
-\ref{fig:figure-4}A bottom row). This behavior is consistent with classic BGS
-theory being unable to fit the DFE to features under weak constraint (e.g.
-introns, UTRs, and the ``other" feature), and consequently severally
-underestimating the mutation rate.
+However, mutation rate estimates for feature-based annotation models are less
+concordant with pedigree-based estimates. First, mutation rate estimates under
+our implementation of classic BGS theory are an order of magnitude below the
+expected range (Figure \ref{fig:figure-4}A top row). We observe similar
+behavior when we use the classic BGS model to fit full-coverage annotation
+models (Figure \ref{fig:figure-4}A bottom row). This behavior is consistent
+with classic BGS theory being unable to fit the DFE to features under weak
+constraint (e.g. introns, UTRs, and the ``other" feature), and most compensate 
+by estimating an unusually low mutation rate.
 
@@
 \ref{supp:tbl-r2}). This pattern was observed in Murphy et al. too (2023;
-Appendix 1, Figure 16). This behavior is consistent with a non-identifiability
-issue between a slightly higher per-basepair mutation rate and annotation
-tracks that contain more conserved sequence. This is expected from theory,
-since both classic BGS and SC16 models only depend on mutation rate through the
-compound parameter $\mu L$, where $L$ is the length of the conserved segment.
-Even though our method is much more robust to including non-conserved regions
-like introns, we still observe this non-identifiability issue.
+Appendix 1, Figure 16). This behavior suggests a non-identifiability issue
+between a slightly higher per-basepair mutation rate and annotation tracks that
+contain more conserved sequence. This is expected from theory, since both
+classic BGS and SC16 models only depend on mutation rate through the compound
+parameter $\mu L$, where $L$ is the length of the conserved segment. Even
+though our method is much more robust to including non-conserved regions like
+introns, we still observe this non-identifiability issue.
 
 Finally, we note that mutation rate estimates in the Feature Priority model are
-unreasonably high ($\widehat{\mu} \approx 4.2 \times 10^{-8}$), reminiscent of
+unreasonably high ($\widehat{\mu} \approx 3 \times 10^{-8}$), reminiscent of
 the high mutation rate estimates found under McVicker et al.'s
@@
 against the average LoF selection coefficient across genes in a megabase window
-(estimated from \cite{Agarwal2023-un}).}
+(estimated by \cite{Agarwal2023-un}).}
 
@@
 While we have previously shown that out-sample predicted diversity matches
-observed diversity (Section XXX), we also investigated whether there were
-regions of the genome that were particularly poor-fitting. Comparing predicted
-against observed diversity along chromosomes (see Supplementary Materials
-Section \ref{supp:chrom-fit}) we find a close correspondence consistent with
-the high $R_\text{LOCO}^2$. However, we find a few large (tens of megabases)
-regions with systematically poorer fit (Supplementary Materials Figure
-\ref{supp:spatial-residuals}). In Figure \ref{fig:figure-5}A we see one such
-region on the short arm of chromosome 2, from 30 Mbp to 60Mbp. Interestingly,
-predicted diversity closely follows the peaks and troughs of this region. We
-note that a small region within this stretch had been found by a genome-wide
-scan for associative overdominance \parencite{Gilbert2020-aw}.
+observed diversity (Section \ref{sec:annotation}), we also investigated whether
+there were regions of the genome that were particularly poor-fitting. Comparing
+predicted against observed diversity along chromosomes (see Supplementary
+Materials Section \ref{supp:chrom-fit}) we find a close correspondence
+consistent with the high $R_\text{LOCO}^2$. However, we find a few large (tens
+of megabases) regions with systematically poorer fit (Supplementary Materials
+Figure \ref{supp:spatial-residuals}). In Figure \ref{fig:figure-5}A we see one
+such region on the short arm of chromosome 2, from 30 Mbp to 60Mbp.
+Interestingly, predicted diversity closely follows the peaks and troughs of
+this region. We note that a small region within this stretch had been found by
+a genome-wide scan for associative overdominance \parencite{Gilbert2020-aw}.
 
@@
 that regions predicted to experience little reduction in diversity due to
-background selection have higher diversity than expected (Figure
-\ref{fig:figure-5}B, orange line). Their work suggested that this could reflect
-ancient introgression between archaic humans and ancestors of contemporary
-humans. Finally, the variance around observed and predicted diversity levels
-falls very close to what we would expect under the theoretic
-coalescence-noise-only expectation ($R_\text{coal}^2$).
+background selection (i.e. $B\approx 1$) have higher diversity than predicted
+(Figure \ref{fig:figure-5}B, orange line). They suggested that this could
+reflect ancient introgression between archaic humans and ancestors of
+contemporary humans. Despite this aberration, the variance around observed and
+predicted diversity levels falls very close to what we would expect under the
+theoretic coalescence-noise-only expectation ($R_\text{coal}^2$).
 
@@
 site pathogenicity being poorly fit by our model. However, the total residual
-variance explained is $R^2 = 0.4\%$ and $R^2 = 0.8\%$ for the CADD 2\% and 6\%
+variance explained is $R^2 = 0.3\%$ and $R^2 = 0.9\%$ for the CADD 2\% and 6\%
 tracks respectively, suggesting only modest amount of selection signal remains.
+
+Since our method does not include the possible effects of linked positive
+selection, we might expect windows containing hard or soft sweeps would have
+systematically lower diversity levels than predicted. Using the locations of
+soft and hard sweeps detected using a machine learning approach
+\parencite{Schrider2017-yx,Kern2018-jd}, we tested whether the residuals of the
+CADD 6\% model containing sweeps were systematically different than those not
+containing sweeps. We find no significant difference between the magnitude of
+residuals of windows containing sweeps versus those that do not (Supplementary
+Materials Figure \ref{suppfig:resid-sweeps}; Kolmogorov--Smirnov p-value =
+0.71).
 
@@
 selection that explicitly incorporates the substitution process, fitting our
-model to genomic data leads to predictions: (1) the substitution rates per
-segment, (2) the fitness variance per segment, and (3) the total number of
-strongly deleterious mutations per diploid genome per generation. Given
-predicted substitution rates for each feature can be compared to estimated
-sequence divergence between species, this offers a qualitative check our model
-is producing sensible results.
+model to genomic diversity data leads to additional predictions. In particular,
+our model predicts (1) the substitution rates per segment, (2) the fitness
+variance per segment, and (3) the total number of strongly deleterious
+mutations per diploid genome per generation. Given predicted substitution rates
+for each feature can be compared to estimated sequence divergence between
+species, this offers a qualitative check our model is producing sensible
+results.
 
@@
 multiple alignment of five primates for each feature in our feature-based
-models (Methods XXX). Since we have observed that mutation rate estimates from
-our method can be unreliable (and simulation results confirm this), we
-predicted substitution rates under a fixed mutation rate of $\mu = 1.5 \times
-10^{-8}$, rather than the maximum likelihood estimates. Unfortunately, a
-careful comparison between our predictions and observed divergence rates is
+models (Methods XXX). Since our simulations show that mutation rate estimates
+can be biased, we predicted substitution rates under a fixed mutation rate of
+$\mu = 1.5 \times 10^{-8}$. Fixing the mutation rate also allows us to more
+easily compare the predictions across our feature-based models. Unfortunately,
+a careful comparison between our predictions and observed divergence rates is
 hindered by considerable uncertainty in the rate of the molecular clock,
@@
 Our predictions are qualitatively consistent with the observed divergence along
-the human lineage (Figure \ref{fig:figure-6}). As expected, the predicted
+the human lineage (Figure \ref{fig:figure-7}). As expected, the predicted
 substitution rate in features unlikely to be under selective constraint is very

### 2023-07-13 manuscript/supplementary.tex
ms updates

@@
 \begin{align}
-  \label{eq:va_bgs}
+  \label{supp-eq:va_bgs}
   V_\text{BGS} &:= \V(w) = 2 s^2 \sum_{l=1}^L p_l(1-p_l) \nonumber  \\
@@
 \caption{$R^2$ and mutation rate estimates for all primary models fit on the
-    default DFE grid (up to selection $s=10^{-1}$).}
+default DFE grid (up to selection $s=10^{-1}$). $R_\text{IS}^2$ is the
+in-sample $R^2$, and $R_\text{LOCO}^2$ is the leave-one-chromosome-out $R^2$.}
 
@@
   Feature Priority &            sparse &          CHB &                        56.00 &             57.88 &            54.76 &                                 3.1722 &                                0.2362 \\
-  Feature Priority &              full &          CHB &                        55.63 &             57.88 &            51.39 &
+  Feature Priority &              full &          CHB &                        55.63 &             57.88 &            51.39 &                                 3.1722 &                                0.1056 \\
 \end{tabular}
@@
 \centering
-\caption{$R^2$ and mutation rate estimates for all primary models fit 
-    with the strong selection grid (up to $s=10^{-1}$). Note that the repeated values in the last rows are
-    correct and due to rounding; see \path{notebooks/main_fits.ipynb} in the
-    GitHub repository for more information.}
+
+\caption{$R^2$ and mutation rate estimates for all primary models fit with the
+    strong selection grid (up to $s=10^{-1}$). Note that the repeated values in
+    the last rows are correct and due to rounding; see
+    \path{notebooks/main_fits.ipynb} in the GitHub repository for more
+information. $R_\text{IS}^2$ is the in-sample $R^2$, and $R_\text{LOCO}^2$ is
+the leave-one-chromosome-out $R^2$. }
+
 \begin{tabular}{lll|crr|cc}
@@
 PhastCons Priority & CEU &      gene &         0.005308 &       0.000226 \\
- PhastCons Priority &  CEU &  PhastCons &  0.008076 &  0.006275 \\
+PhastCons Priority & CEU & phastcons &         0.008076 &       0.006278 \\
 PhastCons Priority & CHB &       CDS &         0.100000 &       0.000000 \\
 PhastCons Priority & CHB &      gene &         0.004896 &       0.000296 \\
- PhastCons Priority &  CHB &  PhastCons &  0.001327 &  0.005886 \\
+PhastCons Priority & CHB & phastcons &         0.001327 &       0.005887 \\
 PhastCons Priority & YRI &       CDS &         0.000000 &       0.000000 \\
- PhastCons Priority &  YRI &       gene &  0.005556 &  0.000160 \\
- PhastCons Priority &  YRI &  PhastCons &  0.028642 &  0.007575 \\
-   Feature Priority &  CEU &        CDS &  0.075041 &  0.005281 \\
+PhastCons Priority & YRI &      gene &         0.005556 &       0.000163 \\
+PhastCons Priority & YRI & phastcons &         0.028642 &       0.007687 \\
+  Feature Priority & CEU &       CDS &         0.075041 &       0.005284 \\
   Feature Priority & CEU &      gene &         0.001654 &       0.000315 \\
-   Feature Priority &  CEU &  PhastCons &  0.069563 &  0.008767 \\
+  Feature Priority & CEU & phastcons &         0.069563 &       0.008772 \\
   Feature Priority & CHB &       CDS &         0.074797 &       0.005211 \\
   Feature Priority & CHB &      gene &         0.001974 &       0.000271 \\
-   Feature Priority &  CHB &  PhastCons &  0.051988 &  0.008507 \\
+  Feature Priority & CHB & phastcons &         0.051988 &       0.008507 \\
   Feature Priority & YRI &       CDS &         0.054503 &       0.001576 \\
   Feature Priority & YRI &      gene &         0.001912 &       0.000336 \\
-   Feature Priority &  YRI &  PhastCons &  0.055153 &  0.009369 \\
-            CADD 6\% &  CEU &      CADD 6\% &  0.042458 &  0.006500 \\
-            CADD 6\% &  CHB &      CADD 6\% &  0.031565 &  0.005733 \\
-            CADD 6\% &  YRI &      CADD 6\% &  0.043671 &  0.007926 \\
-            CADD 8\% &  CEU &      CADD 8\% &  0.041254 &       - \\
-            CADD 8\% &  CHB &      CADD 8\% &  0.029052 &       - \\
-            CADD 8\% &  YRI &      CADD 8\% &  0.046212 &       - \\
+  Feature Priority & YRI & phastcons &         0.055153 &       0.009370 \\
+           CADD 6\% & CEU &     cadd6 &         0.042458 &       0.006500 \\
+           CADD 6\% & CHB &     cadd6 &         0.031565 &       0.005733 \\
+           CADD 6\% & YRI &     cadd6 &         0.043671 &       0.007926 \\
+           CADD 8\% & CEU &     cadd8 &         0.041254 &       0.006017 \\
+           CADD 8\% & CHB &     cadd8 &         0.029052 &       0.005068 \\
+           CADD 8\% & YRI &     cadd8 &         0.046212 &       0.007543 \\
 \hline
@@
   basepairs in a window annotated by a CADD region.}
+\end{figure}
+
+\begin{figure}[htbp]
+  \label{suppfig:resid-sweeps}
+  \centering
+  \includegraphics[width=0.6\textwidth]{figures/supplementary/sweep_residuals.pdf}
+
+  \caption{The distributions of residuals in windows containing hard or soft
+  sweeps (blue) and not containing sweeps (orange) found by
+\parencite{Schrider2017-yx}.}
+
 \end{figure}

### 2023-07-18 manuscript/manuscript.tex
ms updates

@@
     in the human genome. Altogether, our genome-wide model of the linked
-    selection is a step towards uniting evolutionary quantitative genetics with
-    substitution processes.
+    selection is a step towards uniting population and quantitative genetic
+    models of selection with the substitution processes.
 
@@
 fitness or are deleterious. Negative selection (i.e. purifying selection) works
-to eliminate these deleterious mutations from the population, thus we expect
+to eliminate these deleterious mutations from the population; thus we expect
 them to appear at low frequencies within populations
-\parencite{Haldane1927-ga}, and less likely to fix between lineages. The best
-predictor of whether a new mutation will reduce fitness is if it occurs in a
-functional region of the genome, which are often conserved over deep
+\parencite{Haldane1927-ga}, and be less likely to fix between lineages. The
+best predictor of whether a new mutation will reduce fitness is if it occurs in
+a functional region of the genome, which are often conserved over deep
 phylogenetic timescales \parencite{Siepel2005-wh,Margulies2003-sn}. Since these
@@
 previous work has shown that negative selection plays the dominant role in
-shaping megabase-scale patterns of diversity, with positive selection playing a
-nearly negligible role \parencite{Murphy2022-sj}.
+shaping megabase-scale patterns of diversity, with positive selection having a
+nearly negligible impact \parencite{Murphy2022-sj}.
 
@@
 selection process. First, since fixation probabilities ultimately depend on the
-product of the deleterious selection coefficient (s) and population size (N),
-the efficacy of negative selection should depend on demography. Unfortunately,
-accommodating demography into theoretic negative selection models remains an
-open, difficult problem \parencite{Zeng2013-ep,Johri2020-oj}. Second, building
-off classic models of mutation-selection balance
+product of the deleterious selection coefficient ($s$) and population size
+($N$), the efficacy of negative selection should depend on demography.
+Unfortunately, accommodating demography into theoretic negative selection
+models remains an open, difficult problem \parencite{Zeng2013-ep,Johri2020-oj}.
+Second, building off classic models of mutation-selection balance
 \parencite{Crow1970-wm,Kimura1966-bk}, the BGS model assumes that new mutations
@@
 particular, the BGS model cannot accommodate the possibility of fixation for
-mutations with $Ns \ll 1$, diversity is incorrectly predicted to fall to zero
-as the strength of selection diminishes. This weak selection problem is related
-to the wider problem of modeling the substitution rate of deleterious mutation
-(i.e. the “ratchet” rate) and incorporating the impact of perturbations from
-selective interference.
-
-In this work, we extend another class of linked selection models that derive
-from quantitative genetics to address limitations of the classic BGS model.
-These models express linked selection in terms of how polygenic fitness
-variance creates weak genetic draft, and which increases pairwise coalescence
-rates. While these models can theoretically accommodate additive fitness
-variance from any source as long as its rate of change is not too rapid, we
-focus specifically on a deleterious-mutations-only model of fitness variance
-from \textcite{Santiago2016-mu}. This model is identical to BGS when selection
+mutations with $2Ns \ll 1$, leading diversity to be incorrectly predicted to
+fall to zero as the strength of selection diminishes. This weak selection
+problem is related to the wider problem of modeling the substitution rate of
+deleterious mutation (i.e. the ``ratchet" rate) and incorporating the impact of
+interference from other neighboring selected sites.
+
+In this work, we use another class of linked selection models that derive from
+quantitative genetics to address limitations of the classic BGS model. These
+models express linked selection in terms of how polygenic fitness variance
+creates weak genetic draft, and which increases pairwise coalescence rates.
+While these models can theoretically accommodate additive fitness variance from
+any source as long as its rate of change is not too rapid, we focus
+specifically on a deleterious-mutations-only model of fitness variance from
+\textcite{Santiago2016-mu}. This model is identical to BGS when selection
 against deleterious mutations is strong, but it also correctly predicts the
@@
 for why the classic BGS model works well for strong selection but not weak, and
-how formulating selection in terms of the additive fitness variance it creates
-can alleviate these problems.
+how formulating the equilibrium between new deleterious mutations and negative
+selection in terms of the additive fitness variance can alleviate these
+problems.
 
@@
 conserved region of $L$ basepairs, such that the region-wide per generation
-mutation rate is $U = \mu L$. Each mutation imposes a selective cost of $s$ in
-heterozygotes and $2s$ in homozygotes, and fitness effects are multiplicative
-across sites. When selection against a new mutation is strong, the stochastic
-perturbations it experiences due to random genetic drift are weak relative to
-the frequency changes caused by selection. In this case, every deleterious
-mutation is destined to eventual extinction, and only the fraction of lineages
-without any mutations contribute to the long-run ancestry of future
+diploid mutation rate is $U = 2 \mu L$. Each mutation imposes a selective cost
+of $s$ in heterozygotes and $2s$ in homozygotes, and fitness effects are
+multiplicative across sites. When selection against a new mutation is strong,
+the stochastic perturbations it experiences due to random genetic drift are
+weak relative to the frequency changes caused by selection. In this case, every
+deleterious mutation is destined to eventual extinction, and only the fraction
+of lineages without any mutations contribute to the long-run ancestry of future
 populations. The classic background selection approximates this process by
@@
 and selected sites are strongly impacted by stochastic perturbations from two
-sources. First, genetic drift is due to Mendelian segregation and non-heritable
-fitness differences; the magnitude of its perturbations are set by the
-\emph{drift-effective} population size, $N_e$. Second, genetic draft occurs
-when an allele becomes randomly associated with a fitness background that
-perturbs its trajectory across generations \parencite{Neher2013-dz}. Like
-genetic drift, the perturbations due to draft are directionless but overall
-increase the variance in allele frequency change. Unlike drift, the impact of
-the stochastic perturbations due to draft are autocorrelated and build up over
-generations until the fitness background recombines off
+sources. First, genetic drift, which is due to Mendelian segregation and
+non-heritable fitness differences; the magnitude of its perturbations are set
+by the \emph{drift-effective} population size, $N_e$. Second, genetic draft
+occurs when an allele becomes randomly associated with a fitness background
+that perturbs its trajectory across generations
+\parencite{Neher2013-dz,Gillespie2000-mh,Gillespie2001-mv}. Like genetic drift,
+the perturbations due to draft are directionless but overall increase the
+variance in allele frequency change. Unlike drift, the impact of the stochastic
+perturbations due to draft are autocorrelated and build up over generations
+until the fitness background recombines off
 \parencite{Robertson1961-ho,Santiago1995-hx,Buffalo2019-qs}. The stochastic
@@
 would under pure drift \parencite{Wright1938-tv}. However, due to
-autocorrelation, the cumulative effect of this fitness variance is inflated 
-by a factor of $Q^2$. Intuitively, the product $V_A Q^2$ represents the
-expected total variance in reproductive success a neutral mutation experiences
-over its lifetime in a system with weak selection draft.
-
-We can define the draft-effective population size $N_d$ by including the total
+autocorrelation, the cumulative effect of this fitness variance is inflated by
+a factor of $Q^2$. Intuitively, the product $V_A Q^2$ represents the expected
+total variance in reproductive success a neutral mutation experiences over its
+lifetime in a system with weak selection draft.
+
+We define the draft-effective population size $N_d$ by including the total
 additional variance created by heritable fitness in Wright's
@@
 
-We next validated our predicted B' reduction maps with realistic
-chromosome-scale forward simulations of negative selection in conserved
-regions. Averaging over $100$ replicates, we estimated the simulation empirical
-reduction map and compared this to the predicted B and B' maps for the same
-parameters. We find that our B' maps and the classic BGS theory B maps closely
-match simulations when selection is strong (top row of Figure
-\ref{fig:figure-1}E). This provides the first realistic forward-time
-chromosome-scale simulation confirmation of classic BGS theory. However, we
-find slight discrepancies in regions with low recombination (Figure
-\ref{fig:figure-1}E). Second, we find our theory is vastly more accurate than
-the classic B maps when selection is very weak ($2N_e s \ll 1$; bottom row of
-Figure \ref{fig:figure-1}E), which is expected since BGS theory does not work
-in this domain. Across all mutation and selection parameters simulated, the
-relative error of the classic B maps is 14.6\% whereas the relative error in
-the new B' maps is 5\%. Additionally, we find that the mean squared error
-between simulations and B' maps is close to the theoretic lower bound set by
-the coalescence variance \parencite[see Methods XXX]{Tajima1983-gu}.
+We validated our predicted B' reduction maps with realistic chromosome-scale
+forward simulations of negative selection in conserved regions. Averaging over
+$100$ replicates, we estimated the simulation empirical reduction map and
+compared this to the predicted B and B' maps for the same parameters. We find
+that our B' maps and the classic BGS theory B maps closely match simulations
+when selection is strong (top row of Figure \ref{fig:figure-1}E). This provides
+the first realistic forward-time chromosome-scale simulation confirmation of
+classic BGS theory. However, we find slight discrepancies in regions with low
+recombination (Figure \ref{fig:figure-1}E). Second, we find our theory is
+vastly more accurate than the classic B maps when selection is very weak ($2N_e
+s \ll 1$; bottom row of Figure \ref{fig:figure-1}E), which is expected since
+BGS theory does not work in this domain. Across all mutation and selection
+parameters simulated, the relative error of the classic B maps is 14.6\%
+whereas the relative error in the new B' maps is 5\%. Additionally, we find
+that the mean squared error between simulations and B' maps is close to the
+theoretic lower bound set by the coalescence variance
+\parencite{Tajima1983-gu}.
 
@@
 
-Our predictions are qualitatively consistent with the observed divergence along
-the human lineage (Figure \ref{fig:figure-7}). As expected, the predicted
-substitution rate in features unlikely to be under selective constraint is very
-close to the mutation rate. Throughout, we report our substitution rates as a
-percent relative to the total mutation rate, $\mu$ (here fixed to $1.5 \times
+We find that our predicted substitution rates are qualitatively consistent with
+the observed divergence along the human lineage for all features except
+PhastCons (Figure \ref{fig:figure-7}).
+
+As expected, the predicted substitution
+rate in features unlikely to be under selective constraint is very close to the
+mutation rate. Throughout, we report our substitution rates as a percent
+relative to the total mutation rate, $\mu$ (here fixed to $1.5 \times
 10^{-8}$). In our PhastCons Priority model, introns and UTRs are estimated to

### 2023-07-19 manuscript/manuscript.tex
ms updates

@@
 corresponding to the number of diploids in the simulations rather than $B(x) N$
-at position $x$. To test this, we implemented a locally-rescaled version of the
-B' maps, which numerically solves Equations \eqref{eq:main_eqns} for each
+at position $x$. To test this, we implemented a ``locally-rescaled" version of
+the B' maps, which numerically solves Equations \eqref{eq:main_eqns} for each
 segment at position $x$ taking into account the reduction experienced due to
@@
 implies average selection coefficient estimates of $\bar{s} = 0.63\%$, $\bar{s}
-= 0.58\%$, and $\bar{s}= 0.77\%$ for CEU, CHB, and YRI respectively for
+= 0.59\%$, and $\bar{s}= 0.77\%$ for CEU, CHB, and YRI respectively for
 PhastCons regions. Our DFE estimates for the Feature Priority model are
@@
 \ref{supp:tbl-r2-strongsel}). For this strong selection grid, we estimate
-average selection coefficients for the CADD 6\% model of $\bar{s} = 4.3\%$ for
-CEU, $3.2\%$ for CHB, and $4.4\%$ for YRI. Thus, across different selection
-coefficient grids, there is considerable uncertainty in the average strength of
-selection against deleterious mutations in conserved regions.
+average selection coefficients for the CADD 6\% model of $\bar{s} = 4.2\%$ for
+CEU, $3.2\%$ for CHB, and $4.4\%$ for YRI. Thus, there is quite a degree of
+sensitivity to DFEs and average selection estimates based on the selection
+coefficient grid.
 
@@
 estimated to have a selection coefficients of $s = 10^{-1}$ with 43\% chance,
-$s=10^{-2}$ with 7.8\% chance, and $s=10^{-3}$ with 41\% chance. We propose
-that one mechanism for this non-identifiability is that mutations extremely
-harmful to fitness are difficult to distinguish from a smaller effective
-population. Intuitively, as $s$ increases, the average persistence time (in
-generations) of the mutation in the population goes below one generation, i.e.
-death of the carrier of the \emph{de novo} mutation. In this hypothetical
-limit, each occurrence of a very harmful mutation nearly acts like an
-instantaneous death that in effect is just a smaller population size. One way
-to test this hypothesis is to look to see if there is a systematic positive
-relationship across models between average selection coefficient and $\pi_0$,
-which is includes the drift-effective population size $N_e$.
-
-We find this is the case for all of our CADD 6\% models. Across every
-population, average selection was about $7.1$ times larger using the default
-grid, and $\pi_0$ was $5.6\%$ larger (see Supplementary Materials Figure
-\ref{suppfig:sel-ident}). There was no similar consistent change in mutation
-rate estimates among populations. In the CADD 6\% model, genome-wide average
-reduction factor $\bar{B}$ was $\approx 6.1\%$ lower in the default versus
-constrained grid. Overall, this suggests that the linked selection signal alone
-cannot differentiate very strong selection from a slightly smaller
+$s=10^{-2}$ with 7.8\% chance, and $s=10^{-3}$ with 41\% chance in the YRI
+samples. We propose that one mechanism for this non-identifiability is that
+very deleterious mutations lead to larger whole-genome reductions, which are
+difficult to distinguish from a smaller drift effective population size (i.e.
+the $\pi_0$ parameter). One way to test this hypothesis is to look to see if
+there is a systematic positive relationship across models between average
+selection coefficient and $\pi_0$, which is includes the drift-effective
+population size $N_e$.
+
+We find this is the case for all of our CADD 6\% models. Across all
+populations, average selection was about $7.1$ times larger using the strong
+selection grid, and $\pi_0$ was $5.6\%$ higher (see Supplementary Materials
+Figure \ref{suppfig:sel-ident}). There was no similar consistent change in
+mutation rate estimates among populations. In the CADD 6\% model, genome-wide
+average reduction factor $\bar{B}$ was $\approx 6.1\%$ lower in the default
+versus constrained grid. Overall, this suggests that the linked selection
+signal alone cannot differentiate very strong selection from a slightly smaller
 drift-effective population size.
@@
 Although the Feature Priority model fits the data less well than alternative
-models, its features are more interpretable. We find that our B' method
-estimates a bimodal DFE for coding sequences for the Feature Priority model,
-with a large mass placed on $10^{-3} \le s \le 10^{-2}$ and another on the
-neutral class $s=10^{-8}$. This is expected, given that the synonymous and
+models, its DFE estimates are are more interpretable. We find that our B'
+method estimates a bimodal DFE for coding sequences for the Feature Priority
+model, with a large mass placed on $10^{-3} \le s \le 10^{-2}$ and another on
+the neutral class $s=10^{-8}$. This is expected, given that the synonymous and
 non-synonymous sites that constitute coding sequences are under vastly
@@
 class, with a small but significant amount of mass ($\approx 3\%$) placed on
-$s=10^{-2}$. As expected, the DFE for PhastCons regions suggests it is under
-strong selective constraint; however we note that block jackknife based
-uncertainty estimates suggest the model is uncertain whether there is some mass
-on the neutral class. Finally, we highlight one result from our PhastCons
-Priority annotation model (Figure \ref{fig:figure-3}A bottom row): the DFE
-estimate for coding sequences excluding PhastCons regions is estimated as
-neutral. This too is expected; the selection signal in coding regions is
-absorbed by the PhastCons feature, leaving only conditionally neutral sites.
+$s=10^{-2}$. As expected, the DFE for PhastCons regions (which in this model
+correspond to highly-conserved non-coding elements) suggests it is under strong
+selective constraint; however we note that block jackknife based uncertainty
+estimates suggest the model is uncertain whether there is some mass on the
+neutral class. Finally, we highlight one result from our PhastCons Priority
+annotation model (Figure \ref{fig:figure-3}A bottom row): the DFE estimate for
+coding sequences excluding PhastCons regions is estimated as neutral. This too
+is expected; the selection signal in coding regions is absorbed by the
+PhastCons feature, leaving only conditionally neutral sites.
 
@@
 
-Early genome-wide implementations of the classic BGS model fit well, but lead
-to unusually high estimates of the mutation rate \parencite{McVicker2009-ax}.
-This lead to the hypothesis that these models could be absorbing the signal of
-positive selection \parencite{Enard2014-kz}, though other work has found a
-limited role for hitchhiking
+Prior work on genome-wide inference using the classic BGS model fit the linked
+selection signal well, but led to unusually high estimates of the mutation rate
+\parencite{McVicker2009-ax}. This lead to the hypothesis that these models
+could be absorbing the signal of positive selection \parencite{Enard2014-kz},
+though other work has found a limited role for hitchhiking
 \parencite{Pickrell2009-tt,Hernandez2011-gs,Murphy2022}. While our simulation
 results suggest estimates of the mutation rate from linked selection models are
-biased, we still our estimates were in rough accordance with pedigree-based
-estimates \parencite{Kong2013-fc,Tian2019-so}.
+biased, we still check for rough agreement with pedigree-based estimates
+\parencite{Kong2013-fc,Tian2019-so}.
 
@@
 Our full-track CADD 6\% model estimates the mutation rate as $\widehat{\mu} =
-1.56 \times 10^{-8}$ for YRI, $1.63 \times 10^{-8}$ for CEU, and $1.60 \times
+1.56 \times 10^{-8}$ for YRI, $1.64 \times 10^{-8}$ for CEU, and $1.60 \times
 10^{-8}$ for CHB reference samples (Supplementary Table \ref{supp:tbl-r2}). As
-expected, the sparse-track CADD model mutation rate estimates are identical
-between the B' and classic BGS methods (Figure \ref{fig:figure-4}A top row).
-
-However, mutation rate estimates for feature-based annotation models are less
-concordant with pedigree-based estimates. First, mutation rate estimates under
-our implementation of classic BGS theory are an order of magnitude below the
-expected range (Figure \ref{fig:figure-4}A top row). We observe similar
-behavior when we use the classic BGS model to fit full-coverage annotation
-models (Figure \ref{fig:figure-4}A bottom row). This behavior is consistent
-with classic BGS theory being unable to fit the DFE to features under weak
-constraint (e.g. introns, UTRs, and the ``other" feature), and most compensate 
-by estimating an unusually low mutation rate.
+expected, the sparse-track CADD model mutation rate estimates are nearly
+identical between the B' and classic BGS methods (Figure \ref{fig:figure-4}A
+top row).
+
+However, mutation rate estimates for feature-based annotation models are
+implausibly higher than pedigree-based estimates. First, mutation rate
+estimates under our implementation of classic BGS theory are an order of
+magnitude below the expected range (Figure \ref{fig:figure-4}A top row). We
+observe similar behavior when we use the classic BGS model to fit full-coverage
+annotation models (Figure \ref{fig:figure-4}A bottom row). This behavior is
+consistent with classic BGS theory being unable to fit the DFE to features
+under weak constraint (e.g. introns, UTRs, and the ``other" feature), and most
+compensate by estimating an unusually low mutation rate.
 
@@
 model (Figure \ref{fig:figure-4}A bottom row; Supplementary Table
-\ref{supp:tbl-r2}). This pattern was observed in Murphy et al. too (2023;
+\ref{supp:tbl-r2}). This same pattern was observed in Murphy et al. too (2023;
 Appendix 1, Figure 16). This behavior suggests a non-identifiability issue
@@
 parameter $\mu L$, where $L$ is the length of the conserved segment. Even
-though our method is much more robust to including non-conserved regions like
-introns, we still observe this non-identifiability issue.
+though our method is much more robust to inclusion of non-conserved regions
+like introns, we still observe this non-identifiability issue.
 
@@
 between our Feature and PhastCons Priority annotation models, which both
-contain the exact same number of basepairs, but these are assigned to different
-feature classes based on the priority of overlapping features. That one of
-these models is our best-fitting model, and the other our worst indicates the
-sensitivity these models have when features classes have heterogeneous DFEs.
-CADD-based models fit better in part due to their fine-scale resolution of
-selective effects across the genome. While ideally we would fit a CADD model
-with different features corresponding to the different percentiles of
-pathogenicity, these features are on the basepair scale and thus too
-memory-intensive for our method to currently accommodate.
+contain the exact same number of feature basepairs, but these are assigned to
+different feature classes based on the priority of overlapping features. That
+one of these models is our best-fitting model, and the other our worst
+indicates the sensitivity these models have when features classes have
+heterogeneous DFEs. CADD-based models fit better in part due to their
+fine-scale resolution of selective effects across the genome. While ideally we
+would fit a CADD model with different features corresponding to the different
+percentiles of pathogenicity, these features are on the basepair scale and thus
+too memory-intensive for our method to currently accommodate.
 
@@
 multiple alignment of five primates for each feature in our feature-based
-models (Methods XXX). Since our simulations show that mutation rate estimates
-can be biased, we predicted substitution rates under a fixed mutation rate of
-$\mu = 1.5 \times 10^{-8}$. Fixing the mutation rate also allows us to more
-easily compare the predictions across our feature-based models. Unfortunately,
-a careful comparison between our predictions and observed divergence rates is
-hindered by considerable uncertainty in the rate of the molecular clock,
-generation times, and the human-chimpanzee divergence time. We assume a
-generation time of 29 years, and calculate the sequence divergence implied by
-our predicted substitution rates over a range of divergence times, from 6 Mya
-to 12 Mya \parencite{Moorjani2016-tb,Nachman2000-te,Yi2002-pw,Steiper2006-xx}.
+models (Methods \ref{sec:methods-sub}). We compared these to the predicted
+substitution rates per feature, averaging over all segments in the genome.
+Since our simulations show that mutation rate estimates can be biased, we
+predicted substitution rates under a fixed mutation rate of $\mu = 1.5 \times
+10^{-8}$. Fixing the mutation rate also allows us to more easily compare the
+predictions across our feature-based models. Unfortunately, a careful
+comparison between our predictions and observed divergence rates is hindered by
+considerable uncertainty in the rate of the molecular clock, generation times,
+and the human-chimpanzee divergence time. We assume a generation time of 29
+years, and calculate the sequence divergence implied by our predicted
+substitution rates over a range of divergence times, from 6 Mya to 12 Mya
+\parencite{Moorjani2016-tb,Nachman2000-te,Yi2002-pw,Steiper2006-xx}.
 
@@
 the observed divergence along the human lineage for all features except
-PhastCons (Figure \ref{fig:figure-7}).
-
-As expected, the predicted substitution
-rate in features unlikely to be under selective constraint is very close to the
-mutation rate. Throughout, we report our substitution rates as a percent
-relative to the total mutation rate, $\mu$ (here fixed to $1.5 \times
-10^{-8}$). In our PhastCons Priority model, introns and UTRs are estimated to
-have have substitution rates that are $93.6\%$ of the mutation rate, and the
-``other" feature is estimated at $97.8\%$.  When we use the constrained grid,
-these change to $97.85\%$ and $99.98\%$ respectively. In our PhyloFit
-estimates, the divergence rate in introns is $92.44\%$, 5' UTRs $75.5\%$, and
-3' UTRs $72.65\%$. By contrast, our predicted substitution rates for PhastCons
-regions are much lower: $0.48\%$ for the default grid, and $1.65\%$ for the
-constrained grid. For comparison, the PhyloFit estimates for PhastCons regions
-is much higher, at $51.23\%$. Since this would include beneficial substitutions
-that would not be predicted by our model, this suggests a very high proportion
-of beneficial substitutions XXX check
-
-% This is likely due to the constrained grid placing more of the DFE mass on
-% weaker selection coefficients, which in turn have a higher chance of drifting
-% up to fixation.
+PhastCons (Figure \ref{fig:figure-7}). As expected, the predicted substitution
+rates in features unlikely to be under selective constraint (introns and UTRs,
+and the ``other" feature) are very close to the mutation rate. Throughout, we
+report our substitution rates as a percent relative to the total mutation rate,
+$\mu$ (here fixed to $1.5 \times 10^{-8}$). In our Feature Priority model,
+coding sequences are predicted to have a substitution rate of 41.20\% of the
+mutation rate, introns and UTRs 94.71\%, coding sequences 0\%, and the ``other"
+feature 99.98\%. For comparison, the substitution rates along the human lineage
+(as a proportion to the substitution rate in putatively neutral regions) are
+74.15\% in UTRs, 92.44\% in introns, 50.96\% in coding sequences, and 49.56\%
+in PhastCons regions. The large discrepancy between predicted and observed
+PhastCons substitution rates is driven by our DFE estimates suggesting that the
+bulk of mass is on selection coefficients greater than $10^{-3}$, which have no
+chance of fixation in a population of $N_e \approx 10,000$. We note that our
+DFE estimates are qualitatively similar to those inferred using the classic BGS
+model; thus this DFE-substitution rate contradiction is not unique to our
+estimates, and could indicate a potential model misspecfication problem.
+
+XXX HERE
 
@@
 
-\subsection*{Substitution Rate and Total Genic Fitness Variance Predictions}
-
-a set of corresponding substitution rate predictions for each segment under
-mutation-selection-draft process. There are two sources of variation in these
-predictions in this model. First, there is the variation across the
-pre-specified functional categories, e.g. CDS, UTRs, and PhastCons. Second,
-there is region-specific variation
-
-The SC16 theory models linked selection in quantitative genetics terms and
-relies on simultaneously predicting the deleterious substitution rate.
+\subsection*{Substitution Rate Estimation and Prediction}
+\label{sec:methods-sub}
+
+We estimated the divergence along the human branch using PhyloFit
+\parencite{Siepel2004-ql} run on the Multiz alignments
+\parencite{Blanchette2004-bb} of humans and four other primates (\emph{Pongo
+abelii}, \emph{Pan troglodytes}, \emph{Pan paniscus}, \emph{Gorilla gorilla}).
+PhyloFit was run using the HKY85 substitution model per-feature; estimates from
+alternate substitution models yielded equivalent results. Further details about
+this process can be be found in the GitHub repository
+(\url{https://github.com/vsbuffalo/bprime}).
 

### 2023-07-19 manuscript/supplementary.tex
ms updates

@@
 
+\begin{figure}[htbp]
+    \label{suppfig:strong-sel-dfe-pop-phastcons}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/supplementary/feature_priority_pop_dfe_strongselgrid.pdf}
+
+  \caption{The DFE estimates for the strong selection grid (up to $s=10^{-1}$)
+  under the Feature Priority model, across populations.}
+\end{figure}
+
+\begin{figure}[htbp]
+    \label{suppfig:strong-sel-dfe-pop-feature}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/supplementary/phastcons_priority_pop_dfe_strongselgrid.pdf}
+
+  \caption{The DFE estimates for the strong selection grid (up to $s=10^{-1}$)
+  under the PhastCons Priority model, across populations.}
+\end{figure}
+
 \subsection{Demographic Expansion and Recessive Simulation Tests}
@@
 
-\begin{figure}[htbp]
-    \label{suppfig:dfe-pop-smallgrid-feature}
-  \centering
-  \includegraphics[width=0.6\textwidth]{figures/supplementary/feature_priority_pop_dfe_smallselgrid.pdf}
-  \caption{The DFE for the Feature Priority model, fit with a smaller DFE grid from 
-  $\mu \in \{10^{-8}, 10^{-7}, \ldots, 10^{-2}\}$.}
-\end{figure}
-
-\begin{figure}[htbp]
-    \label{suppfig:dfe-pop-smallgrid-phastcons}
-  \centering
-  \includegraphics[width=0.6\textwidth]{figures/supplementary/phastcons_priority_pop_dfe_smallselgrid.pdf}
-  \caption{The DFE for the PhastCons Priority model, fit with a smaller DFE grid from 
-  $\mu \in \{10^{-8}, 10^{-7}, \ldots, 10^{-2}\}$.}
-\end{figure}
-
 \subsection{Spatial Residuals}
@@
 
+\subsection{Phastcons Priority (strong sel. grid), Full Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: moving block jackknife
+negative log-likelihood: 197825873893.4953
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.0011279
+pi  = 0.000801495
+mut. rate = 2.35e-08 
+Ne = 28,197 (if mu=1e-8), Ne = 14,098 (if mu=2e-8)
+R2 = 63.0106% (in-sample)  61.7759% (out-sample)
+W = 
+          CDS    gene    other    phastcons
+------  -----  ------  -------  -----------
+1e-08       0   0.901    0.984        0.034
+1e-07       0   0.026    0            0.376
+1e-06       0   0        0            0.001
+1e-05       0   0.019    0            0
+0.0001      0   0        0            0
+0.001       0   0        0            0.472
+0.01        0   0        0            0.046
+0.1         1   0.053    0.016        0.071
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+ML estimates: (whole genome)
+standard error method: moving block jackknife
+negative log-likelihood: 61821783504.67356
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00106467
+pi  = 0.000752601
+mut. rate = 2.557e-08 
+Ne = 26,616 (if mu=1e-8), Ne = 13,308 (if mu=2e-8)
+R2 = 61.0847% (in-sample)  59.2923% (out-sample)
+W = 
+          CDS    gene    other    phastcons
+------  -----  ------  -------  -----------
+1e-08       0   0.931    0.985        0.478
+1e-07       0   0.005    0            0.002
+1e-06       0   0        0            0
+1e-05       0   0.013    0            0
+0.0001      0   0        0            0
+0.001       0   0.003    0            0.43
+0.01        0   0        0            0.09
+0.1         1   0.049    0.015        0
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: moving block jackknife
+negative log-likelihood: 250656989872.19766
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00147277
+pi  = 0.00106447
+mut. rate = 1.708e-08 
+Ne = 36,819 (if mu=1e-8), Ne = 18,409 (if mu=2e-8)
+R2 = 68.1177% (in-sample)  68.1947% (out-sample)
+W = 
+          CDS    gene    other    phastcons
+------  -----  ------  -------  -----------
+1e-08       1   0.902    0.979        0
+1e-07       0   0        0            0.004
+1e-06       0   0        0            0.001
+1e-05       0   0.042    0            0
+0.0001      0   0        0            0
+0.001       0   0        0            0.523
+0.01        0   0        0            0.212
+0.1         0   0.056    0.021        0.26
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{Phastcons Priority (strong sel. grid), Sparse Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 197826612684.87143
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00112308
+pi  = 0.000801495
+mut. rate = 2.019e-08 
+Ne = 28,076 (if mu=1e-8), Ne = 14,038 (if mu=2e-8)
+R2 = 63.0152% (in-sample)  61.9834% (out-sample)
+W = 
+          CDS    gene    phastcons
+------  -----  ------  -----------
+1e-08   0       0.913        0.227
+1e-07   0       0            0.004
+1e-06   0.004   0            0
+1e-05   0       0.019        0
+0.0001  0       0            0
+0.001   0       0            0.546
+0.01    0       0            0.061
+0.1     0.996   0.069        0.161
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 61822016128.79776
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00106069
+pi  = 0.000752601
+mut. rate = 5.887e-08 
+Ne = 26,517 (if mu=1e-8), Ne = 13,258 (if mu=2e-8)
+R2 = 61.0804% (in-sample)  59.4419% (out-sample)
+W = 
+          CDS    gene    phastcons
+------  -----  ------  -----------
+1e-08   0.617   0.968        0.744
+1e-07   0.01    0            0.02
+1e-06   0       0            0
+1e-05   0       0.005        0
+0.0001  0       0            0
+0.001   0       0.001        0.185
+0.01    0       0            0.051
+0.1     0.373   0.026        0
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 250657908564.79294
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00146801
+pi  = 0.00106447
+mut. rate = 1.864e-08 
+Ne = 36,700 (if mu=1e-8), Ne = 18,350 (if mu=2e-8)
+R2 = 68.1582% (in-sample)  68.4462% (out-sample)
+W = 
+          CDS    gene    phastcons
+------  -----  ------  -----------
+1e-08       1   0.907        0
+1e-07       0   0.001        0
+1e-06       0   0            0
+1e-05       0   0.035        0
+0.0001      0   0            0
+0.001       0   0            0.477
+0.01        0   0            0.199
+0.1         0   0.057        0.323
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{Feature Priority (strong sel. grid), Full Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: moving block jackknife
+negative log-likelihood: 197848627167.32043
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00115698
+pi  = 0.000801495
+mut. rate = 4.1e-08 
+Ne = 28,924 (if mu=1e-8), Ne = 14,462 (if mu=2e-8)
+R2 = 60.3026% (in-sample)  58.5808% (out-sample)
+W = 
+          CDS    gene    other    phastcons
+------  -----  ------  -------  -----------
+1e-08   0.107   0.946        1        0
+1e-07   0       0.001        0        0
+1e-06   0       0            0        0
+1e-05   0       0.023        0        0
+0.0001  0       0            0        0
+0.001   0.144   0.014        0        0.229
+0.01    0       0            0        0.086
+0.1     0.749   0.016        0        0.685
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: moving block jackknife
+negative log-likelihood: 61827828856.16418
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00108991
+pi  = 0.000752601
+mut. rate = 3.755e-08 
+Ne = 27,247 (if mu=1e-8), Ne = 13,623 (if mu=2e-8)
+R2 = 58.7328% (in-sample)  56.4805% (out-sample)
+W = 
+          CDS    gene    other    phastcons
+------  -----  ------  -------  -----------
+1e-08   0.088   0.94         1        0.006
+1e-07   0.01    0            0        0.003
+1e-06   0       0            0        0
+1e-05   0       0.023        0        0
+0.0001  0       0            0        0
+0.001   0.156   0.017        0        0.244
+0.01    0       0            0        0.254
+0.1     0.746   0.02         0        0.492
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: moving block jackknife
+negative log-likelihood: 250679665159.79324
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00151818
+pi  = 0.00106447
+mut. rate = 4.195e-08 
+Ne = 37,954 (if mu=1e-8), Ne = 18,977 (if mu=2e-8)
+R2 = 65.5097% (in-sample)  64.9049% (out-sample)
+W = 
+          CDS    gene    other    phastcons
+------  -----  ------  -------  -----------
+1e-08   0.364   0.931        1        0.003
+1e-07   0.001   0            0        0
+1e-06   0       0            0        0
+1e-05   0       0.039        0        0
+0.0001  0       0            0        0
+0.001   0.091   0.011        0        0.151
+0.01    0       0            0        0.328
+0.1     0.544   0.019        0        0.517
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{Feature Priority (strong sel. grid), Sparse Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: moving block jackknife
+negative log-likelihood: 197848626646.91736
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00115698
+pi  = 0.000801495
+mut. rate = 4.104e-08 
+Ne = 28,924 (if mu=1e-8), Ne = 14,462 (if mu=2e-8)
+R2 = 60.3027% (in-sample)  58.5823% (out-sample)
+W = 
+          CDS    gene    phastcons
+------  -----  ------  -----------
+1e-08   0.107   0.947        0.001
+1e-07   0       0            0
+1e-06   0       0            0.001
+1e-05   0       0.023        0
+0.0001  0       0            0
+0.001   0.144   0.014        0.229
+0.01    0       0            0.086
+0.1     0.748   0.016        0.684
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: moving block jackknife
+negative log-likelihood: 61827828699.43107
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00108991
+pi  = 0.000752601
+mut. rate = 3.719e-08 
+Ne = 27,247 (if mu=1e-8), Ne = 13,623 (if mu=2e-8)
+R2 = 58.7329% (in-sample)  56.4805% (out-sample)
+W = 
+          CDS    gene    phastcons
+------  -----  ------  -----------
+1e-08   0.089   0.938        0
+1e-07   0       0.001        0
+1e-06   0       0            0
+1e-05   0       0.023        0
+0.0001  0       0            0
+0.001   0.157   0.017        0.246
+0.01    0       0            0.257
+0.1     0.754   0.02         0.497
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: moving block jackknife
+negative log-likelihood: 250679664591.4161
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00151817
+pi  = 0.00106447
+mut. rate = 4.179e-08 
+Ne = 37,954 (if mu=1e-8), Ne = 18,977 (if mu=2e-8)
+R2 = 65.5099% (in-sample)  64.9049% (out-sample)
+W = 
+          CDS    gene    phastcons
+------  -----  ------  -----------
+1e-08   0.36    0.931        0
+1e-07   0.002   0            0
+1e-06   0       0            0
+1e-05   0       0.039        0
+0.0001  0       0            0
+0.001   0.092   0.011        0.151
+0.01    0       0            0.329
+0.1     0.546   0.019        0.519
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{CADD 6\% (strong selection DFE grid), Full Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: moving block jackknife
+negative log-likelihood: 197826953189.02365
+number of successful starts: 20000 (100.0% total)
+pi_0 = 0.00109999
+pi  = 0.000801495
+mut. rate = 1.537e-08 
+Ne = 27,499 (if mu=1e-8), Ne = 13,749 (if mu=2e-8)
+R2 = 62.9394% (in-sample)  61.7433% (out-sample)
+W = 
+          cadd6    other
+------  -------  -------
+1e-08     0        0.965
+1e-07     0        0
+1e-06     0        0
+1e-05     0        0
+0.0001    0.026    0
+0.001     0.555    0
+0.01      0        0
+0.1       0.419    0.035
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: moving block jackknife
+negative log-likelihood: 61821323176.26884
+number of successful starts: 20000 (100.0% total)
+pi_0 = 0.00103624
+pi  = 0.000752601
+mut. rate = 1.508e-08 
+Ne = 25,905 (if mu=1e-8), Ne = 12,952 (if mu=2e-8)
+R2 = 61.3115% (in-sample)  59.6228% (out-sample)
+W = 
+          cadd6    other
+------  -------  -------
+1e-08     0        0.961
+1e-07     0.007    0
+1e-06     0.009    0
+1e-05     0        0
+0.0001    0.004    0
+0.001     0.67     0
+0.01      0        0
+0.1       0.309    0.039
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: moving block jackknife
+negative log-likelihood: 250659830283.90668
+number of successful starts: 20000 (100.0% total)
+pi_0 = 0.00145893
+pi  = 0.00106447
+mut. rate = 1.461e-08 
+Ne = 36,473 (if mu=1e-8), Ne = 18,236 (if mu=2e-8)
+R2 = 68.0839% (in-sample)  68.0751% (out-sample)
+W = 
+          cadd6    other
+------  -------  -------
+1e-08     0        0.958
+1e-07     0.023    0
+1e-06     0.001    0
+1e-05     0        0
+0.0001    0.065    0
+0.001     0.407    0
+0.01      0.078    0
+0.1       0.425    0.041
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{CADD 8\% (strong selection DFE grid), Full Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: moving block jackknife
+negative log-likelihood: 197834651568.3971
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00109606
+pi  = 0.000801495
+mut. rate = 1.157e-08 
+Ne = 27,401 (if mu=1e-8), Ne = 13,700 (if mu=2e-8)
+R2 = 62.1245% (in-sample)  60.8009% (out-sample)
+W = 
+          cadd8    other
+------  -------  -------
+1e-08     0        0.952
+1e-07     0.004    0
+1e-06     0        0
+1e-05     0        0
+0.0001    0.052    0
+0.001     0.537    0
+0.01      0        0
+0.1       0.407    0.048
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: moving block jackknife
+negative log-likelihood: 61823378586.32404
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00103245
+pi  = 0.000752601
+mut. rate = 1.113e-08 
+Ne = 25,811 (if mu=1e-8), Ne = 12,905 (if mu=2e-8)
+R2 = 60.6265% (in-sample)  58.8768% (out-sample)
+W = 
+          cadd8    other
+------  -------  -------
+1e-08     0        0.944
+1e-07     0        0.001
+1e-06     0        0
+1e-05     0        0
+0.0001    0.026    0
+0.001     0.69     0
+0.01      0        0
+0.1       0.284    0.055
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: moving block jackknife
+negative log-likelihood: 250669695762.61414
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00145867
+pi  = 0.00106447
+mut. rate = 1.013e-08 
+Ne = 36,466 (if mu=1e-8), Ne = 18,233 (if mu=2e-8)
+R2 = 67.1076% (in-sample)  66.9835% (out-sample)
+W = 
+          cadd8    other
+------  -------  -------
+1e-08     0        0.932
+1e-07     0.002    0
+1e-06     0        0
+1e-05     0        0
+0.0001    0.103    0
+0.001     0.438    0
+0.01      0        0
+0.1       0.458    0.068
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{Phastcons Priority, Full Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 264
+standard error method: moving block jackknife
+negative log-likelihood: 197835157725.57724
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00105552 (0.00106, 0.00106)
+pi  = 0.000801495
+mut. rate = 1.912e-08  (1.32e-08, 2.5e-08)
+Ne = 26,387 (if mu=1e-8), Ne = 13,193 (if mu=2e-8)
+R2 = 62.2547% (in-sample)  61.0392% (out-sample)
+W = 
+        CDS                      gene                     other                    phastcons
+------  -----------------------  -----------------------  -----------------------  -----------------------
+1e-08   1.000 (0.9744, 1.0256)   0.947 (0.8398, 1.0542)   1.000 (0.9959, 1.0041)   0.000 (-0.2261, 0.2262)
+1e-07   0.000 (-0.0250, 0.0250)  0.000 (-0.1053, 0.1053)  0.000 (-0.0041, 0.0041)  0.000 (-0.1200, 0.1200)
+1e-06   0.000 (-0.0021, 0.0021)  0.000 (-0.0010, 0.0010)  0.000 (-0.0001, 0.0001)  0.000 (-0.0280, 0.0280)
+1e-05   0.000 (-0.0000, 0.0000)  0.030 (0.0209, 0.0400)   0.000 (-0.0000, 0.0000)  0.000 (-0.0004, 0.0004)
+0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0001)
+0.001   0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.413 (0.2960, 0.5308)
+0.01    0.000 (-0.0001, 0.0001)  0.023 (0.0160, 0.0292)   0.000 (-0.0000, 0.0000)  0.586 (0.4205, 0.7525)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 266
+standard error method: moving block jackknife
+negative log-likelihood: 61824235931.79921
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00100129 (0.00100, 0.00100)
+pi  = 0.000752601
+mut. rate = 1.869e-08  (1.16e-08, 2.57e-08)
+Ne = 25,032 (if mu=1e-8), Ne = 12,516 (if mu=2e-8)
+R2 = 60.4486% (in-sample)  58.7636% (out-sample)
+W = 
+        CDS                      gene                     other                    phastcons
+------  -----------------------  -----------------------  -----------------------  -----------------------
+1e-08   1.000 (0.8992, 1.1008)   0.940 (0.7946, 1.0845)   1.000 (0.9996, 1.0004)   0.000 (-0.2201, 0.2202)
+1e-07   0.000 (-0.1006, 0.1006)  0.000 (-0.1420, 0.1420)  0.000 (-0.0003, 0.0003)  0.000 (-0.2016, 0.2017)
+1e-06   0.000 (-0.0021, 0.0021)  0.000 (-0.0025, 0.0025)  0.000 (-0.0001, 0.0001)  0.000 (-0.0118, 0.0119)
+1e-05   0.000 (-0.0000, 0.0000)  0.031 (0.0161, 0.0457)   0.000 (-0.0000, 0.0000)  0.000 (-0.0002, 0.0002)
+0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
+0.001   0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.457 (0.2998, 0.6141)
+0.01    0.000 (-0.0000, 0.0000)  0.030 (0.0193, 0.0399)   0.000 (-0.0000, 0.0000)  0.543 (0.3558, 0.7302)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 264
+standard error method: moving block jackknife
+negative log-likelihood: 250667315874.309
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00139482 (0.00139, 0.00140)
+pi  = 0.00106447
+mut. rate = 2.004e-08  (1.53e-08, 2.48e-08)
+Ne = 34,870 (if mu=1e-8), Ne = 17,435 (if mu=2e-8)
+R2 = 67.3564% (in-sample)  67.274% (out-sample)
+W = 
+        CDS                      gene                     other                    phastcons
+------  -----------------------  -----------------------  -----------------------  -----------------------
+1e-08   1.000 (0.9646, 1.0354)   0.955 (0.8905, 1.0192)   1.000 (0.9972, 1.0028)   0.000 (-0.1515, 0.1516)
+1e-07   0.000 (-0.0352, 0.0352)  0.000 (-0.0636, 0.0636)  0.000 (-0.0028, 0.0028)  0.000 (-0.1435, 0.1435)
+1e-06   0.000 (-0.0008, 0.0008)  0.000 (-0.0009, 0.0011)  0.000 (-0.0000, 0.0000)  0.000 (-0.0101, 0.0101)
+1e-05   0.000 (-0.0000, 0.0000)  0.029 (0.0192, 0.0384)   0.000 (-0.0000, 0.0000)  0.000 (-0.0002, 0.0002)
+0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.027 (0.0108, 0.0435)
+0.001   0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.227 (0.1661, 0.2882)
+0.01    0.000 (-0.0000, 0.0000)  0.016 (0.0124, 0.0201)   0.000 (-0.0000, 0.0000)  0.746 (0.5776, 0.9138)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{Phastcons Priority, Sparse Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 197835157553.76733
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00105552
+pi  = 0.000801495
+mut. rate = 2.313e-08 
+Ne = 26,387 (if mu=1e-8), Ne = 13,193 (if mu=2e-8)
+R2 = 62.2548% (in-sample)  61.0392% (out-sample)
+W = 
+          CDS    gene    phastcons
+------  -----  ------  -----------
+1e-08       1   0.954        0.174
+1e-07       0   0.002        0
+1e-06       0   0            0
+1e-05       0   0.025        0
+0.0001      0   0            0
+0.001       0   0            0.342
+0.01        0   0.019        0.485
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 61824235897.32505
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00100129
+pi  = 0.000752601
+mut. rate = 1.92e-08 
+Ne = 25,032 (if mu=1e-8), Ne = 12,516 (if mu=2e-8)
+R2 = 60.4487% (in-sample)  58.8423% (out-sample)
+W = 
+          CDS    gene    phastcons
+------  -----  ------  -----------
+1e-08       1   0.939        0.003
+1e-07       0   0.002        0.022
+1e-06       0   0            0.002
+1e-05       0   0.03         0
+0.0001      0   0            0
+0.001       0   0            0.445
+0.01        0   0.029        0.528
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 250667315734.13507
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00139482
+pi  = 0.00106447
+mut. rate = 2.004e-08 
+Ne = 34,870 (if mu=1e-8), Ne = 17,435 (if mu=2e-8)
+R2 = 67.3565% (in-sample)  67.2897% (out-sample)
+W = 
+          CDS    gene    phastcons
+------  -----  ------  -----------
+1e-08       1   0.954        0
+1e-07       0   0.001        0
+1e-06       0   0            0
+1e-05       0   0.029        0
+0.0001      0   0            0.027
+0.001       0   0            0.227
+0.01        0   0.016        0.746
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{Feature Priority, Full Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 116
+standard error method: moving block jackknife
+negative log-likelihood: 197860660948.8551
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00109138 (0.00109, 0.00109)
+pi  = 0.000801495
+mut. rate = 3.044e-08  (2.79e-08, 3.3e-08)
+Ne = 27,284 (if mu=1e-8), Ne = 13,642 (if mu=2e-8)
+R2 = 59.1052% (in-sample)  57.1447% (out-sample)
+W = 
+        CDS                      gene                     other                    phastcons
+------  -----------------------  -----------------------  -----------------------  -----------------------
+1e-08   0.412 (0.2565, 0.5672)   0.884 (0.8329, 0.9356)   1.000 (0.9996, 1.0003)   0.000 (-0.0666, 0.0666)
+1e-07   0.000 (-0.1509, 0.1511)  0.003 (-0.0480, 0.0534)  0.000 (-0.0003, 0.0004)  0.000 (-0.0248, 0.0248)
+1e-06   0.000 (-0.0021, 0.0025)  0.000 (-0.0005, 0.0005)  0.000 (-0.0000, 0.0000)  0.000 (-0.0054, 0.0054)
+1e-05   0.000 (-0.0000, 0.0000)  0.074 (0.0608, 0.0875)   0.000 (-0.0000, 0.0000)  0.000 (-0.0001, 0.0001)
+0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
+0.001   0.066 (0.0340, 0.0982)   0.008 (0.0064, 0.0102)   0.000 (-0.0000, 0.0000)  0.136 (0.1225, 0.1503)
+0.01    0.522 (0.4497, 0.5939)   0.031 (0.0276, 0.0336)   0.000 (-0.0000, 0.0000)  0.864 (0.7932, 0.9340)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 256
+standard error method: moving block jackknife
+negative log-likelihood: 61830715796.965416
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00103388 (0.00103, 0.00104)
+pi  = 0.000752601
+mut. rate = 3.172e-08  (2.73e-08, 3.62e-08)
+Ne = 25,847 (if mu=1e-8), Ne = 12,923 (if mu=2e-8)
+R2 = 57.8792% (in-sample)  55.631% (out-sample)
+W = 
+        CDS                      gene                     other                    phastcons
+------  -----------------------  -----------------------  -----------------------  -----------------------
+1e-08   0.430 (0.1137, 0.7471)   0.894 (0.8175, 0.9702)   1.000 (0.9993, 1.0007)   0.000 (-0.1107, 0.1107)
+1e-07   0.001 (-0.2978, 0.3001)  0.004 (-0.0686, 0.0772)  0.000 (-0.0007, 0.0007)  0.000 (-0.0393, 0.0393)
+1e-06   0.000 (-0.0035, 0.0042)  0.000 (-0.0006, 0.0006)  0.000 (-0.0000, 0.0000)  0.000 (-0.0053, 0.0053)
+1e-05   0.000 (-0.0000, 0.0000)  0.064 (0.0437, 0.0848)   0.000 (-0.0000, 0.0000)  0.000 (-0.0001, 0.0001)
+0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
+0.001   0.052 (0.0045, 0.1000)   0.012 (0.0086, 0.0148)   0.000 (-0.0000, 0.0000)  0.166 (0.1391, 0.1926)
+0.01    0.516 (0.4059, 0.6258)   0.026 (0.0216, 0.0302)   0.000 (-0.0000, 0.0000)  0.834 (0.7222, 0.9461)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 241
+standard error method: moving block jackknife
+negative log-likelihood: 250692005490.22156
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00144148 (0.00144, 0.00144)
+pi  = 0.00106447
+mut. rate = 3.349e-08  (2.85e-08, 3.85e-08)
+Ne = 36,037 (if mu=1e-8), Ne = 18,018 (if mu=2e-8)
+R2 = 64.2844% (in-sample)  63.058% (out-sample)
+W = 
+        CDS                      gene                     other                    phastcons
+------  -----------------------  -----------------------  -----------------------  -----------------------
+1e-08   0.752 (0.4939, 1.0110)   0.878 (0.8209, 0.9353)   1.000 (0.9998, 1.0002)   0.000 (-0.1066, 0.1066)
+1e-07   0.000 (-0.2548, 0.2552)  0.000 (-0.0533, 0.0533)  0.000 (-0.0002, 0.0002)  0.000 (-0.0669, 0.0669)
+1e-06   0.000 (-0.0028, 0.0029)  0.000 (-0.0005, 0.0005)  0.000 (-0.0000, 0.0000)  0.000 (-0.0087, 0.0087)
+1e-05   0.000 (-0.0000, 0.0000)  0.085 (0.0721, 0.0986)   0.000 (-0.0000, 0.0000)  0.000 (-0.0002, 0.0002)
+0.0001  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)  0.000 (-0.0000, 0.0000)
+0.001   0.100 (0.0858, 0.1136)   0.003 (0.0026, 0.0040)   0.000 (-0.0000, 0.0000)  0.070 (0.0577, 0.0823)
+0.01    0.148 (0.1266, 0.1688)   0.033 (0.0285, 0.0379)   0.000 (-0.0000, 0.0000)  0.930 (0.8006, 1.0594)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{Feature Priority, Sparse Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 197860660635.007
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00109138
+pi  = 0.000801495
+mut. rate = 3.044e-08 
+Ne = 27,284 (if mu=1e-8), Ne = 13,642 (if mu=2e-8)
+R2 = 59.1053% (in-sample)  57.2393% (out-sample)
+W = 
+          CDS    gene    phastcons
+------  -----  ------  -----------
+1e-08   0.412   0.887        0
+1e-07   0       0            0
+1e-06   0.001   0            0
+1e-05   0       0.074        0
+0.0001  0       0            0
+0.001   0.066   0.008        0.136
+0.01    0.522   0.031        0.864
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 61830715695.97452
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00103388
+pi  = 0.000752601
+mut. rate = 3.172e-08 
+Ne = 25,847 (if mu=1e-8), Ne = 12,923 (if mu=2e-8)
+R2 = 57.8793% (in-sample)  55.9962% (out-sample)
+W = 
+          CDS    gene    phastcons
+------  -----  ------  -----------
+1e-08   0.418   0.898        0
+1e-07   0.014   0            0
+1e-06   0       0            0
+1e-05   0       0.064        0
+0.0001  0       0            0
+0.001   0.052   0.012        0.166
+0.01    0.516   0.026        0.834
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 250692005023.91275
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00144148
+pi  = 0.00106447
+mut. rate = 3.353e-08 
+Ne = 36,037 (if mu=1e-8), Ne = 18,018 (if mu=2e-8)
+R2 = 64.2844% (in-sample)  63.6078% (out-sample)
+W = 
+          CDS    gene    phastcons
+------  -----  ------  -----------
+1e-08   0.747   0.875        0
+1e-07   0.006   0.004        0.001
+1e-06   0       0            0
+1e-05   0       0.085        0
+0.0001  0       0            0
+0.001   0.1     0.003        0.07
+0.01    0.148   0.033        0.929
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{CADD 6\%, Full Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 280
+standard error method: moving block jackknife
+negative log-likelihood: 197839911915.8117
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00102951 (0.00103, 0.00103)
+pi  = 0.000801495
+mut. rate = 1.635e-08  (1.2e-08, 2.07e-08)
+Ne = 25,737 (if mu=1e-8), Ne = 12,868 (if mu=2e-8)
+R2 = 61.767% (in-sample)  60.2433% (out-sample)
+W = 
+        cadd6                    other
+------  -----------------------  -----------------------
+1e-08   0.002 (-0.1758, 0.1800)  0.997 (0.9918, 1.0019)
+1e-07   0.006 (-0.1224, 0.1353)  0.000 (-0.0049, 0.0049)
+1e-06   0.000 (-0.0079, 0.0084)  0.000 (-0.0001, 0.0001)
+1e-05   0.000 (-0.0002, 0.0002)  0.000 (-0.0000, 0.0000)
+0.0001  0.049 (0.0332, 0.0649)   0.000 (-0.0000, 0.0000)
+0.001   0.325 (0.2429, 0.4072)   0.000 (-0.0000, 0.0000)
+0.01    0.617 (0.4630, 0.7711)   0.003 (0.0022, 0.0042)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 279
+standard error method: moving block jackknife
+negative log-likelihood: 61824516495.02472
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.000977579 (0.00098, 0.00098)
+pi  = 0.000752601
+mut. rate = 1.598e-08  (1.15e-08, 2.05e-08)
+Ne = 24,439 (if mu=1e-8), Ne = 12,219 (if mu=2e-8)
+R2 = 60.4032% (in-sample)  58.4847% (out-sample)
+W = 
+        cadd6                    other
+------  -----------------------  -----------------------
+1e-08   0.000 (-0.1922, 0.1922)  0.992 (0.9829, 1.0002)
+1e-07   0.001 (-0.1431, 0.1441)  0.001 (-0.0076, 0.0088)
+1e-06   0.000 (-0.0068, 0.0068)  0.000 (-0.0001, 0.0001)
+1e-05   0.000 (-0.0002, 0.0002)  0.000 (-0.0000, 0.0000)
+0.0001  0.030 (0.0129, 0.0476)   0.000 (-0.0000, 0.0000)
+0.001   0.440 (0.3238, 0.5568)   0.000 (-0.0000, 0.0000)
+0.01    0.529 (0.3902, 0.6676)   0.008 (0.0056, 0.0101)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 280
+standard error method: moving block jackknife
+negative log-likelihood: 250675195472.4508
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00136706 (0.00137, 0.00137)
+pi  = 0.00106447
+mut. rate = 1.556e-08  (1.25e-08, 1.86e-08)
+Ne = 34,176 (if mu=1e-8), Ne = 17,088 (if mu=2e-8)
+R2 = 66.6755% (in-sample)  66.5047% (out-sample)
+W = 
+        cadd6                    other
+------  -----------------------  -----------------------
+1e-08   0.000 (-0.1074, 0.1079)  0.994 (0.9892, 0.9985)
+1e-07   0.006 (-0.1023, 0.1133)  0.000 (-0.0045, 0.0045)
+1e-06   0.000 (-0.0038, 0.0038)  0.000 (-0.0001, 0.0001)
+1e-05   0.000 (-0.0001, 0.0001)  0.000 (-0.0000, 0.0000)
+0.0001  0.099 (0.0783, 0.1204)   0.000 (-0.0000, 0.0000)
+0.001   0.115 (0.0848, 0.1448)   0.000 (-0.0000, 0.0000)
+0.01    0.780 (0.6381, 0.9221)   0.006 (0.0049, 0.0074)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{CADD 6\%, Sparse Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 197839997860.2112
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00102636
+pi  = 0.000801495
+mut. rate = 1.68e-08 
+Ne = 25,659 (if mu=1e-8), Ne = 12,829 (if mu=2e-8)
+R2 = 61.8005% (in-sample)  60.3663% (out-sample)
+W = 
+          cadd6
+------  -------
+1e-08     0
+1e-07     0
+1e-06     0
+1e-05     0
+0.0001    0.046
+0.001     0.317
+0.01      0.638
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 61824669694.580444
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.000970441
+pi  = 0.000752601
+mut. rate = 1.745e-08 
+Ne = 24,261 (if mu=1e-8), Ne = 12,130 (if mu=2e-8)
+R2 = 60.4898% (in-sample)  58.67% (out-sample)
+W = 
+          cadd6
+------  -------
+1e-08     0.003
+1e-07     0
+1e-06     0
+1e-05     0
+0.0001    0.024
+0.001     0.403
+0.01      0.57
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 250675581171.07547
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00135936
+pi  = 0.00106447
+mut. rate = 1.657e-08 
+Ne = 33,983 (if mu=1e-8), Ne = 16,991 (if mu=2e-8)
+R2 = 66.8084% (in-sample)  66.6575% (out-sample)
+W = 
+          cadd6
+------  -------
+1e-08     0
+1e-07     0
+1e-06     0
+1e-05     0
+0.0001    0.09
+0.001     0.109
+0.01      0.801
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{CADD 8\%, Full Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 241
+standard error method: moving block jackknife
+negative log-likelihood: 197849369068.89285
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00102211 (0.00102, 0.00102)
+pi  = 0.000801495
+mut. rate = 1.208e-08  (8.74e-09, 1.54e-08)
+Ne = 25,552 (if mu=1e-8), Ne = 12,776 (if mu=2e-8)
+R2 = 60.8766% (in-sample)  59.2433% (out-sample)
+W = 
+        cadd8                    other
+------  -----------------------  -----------------------
+1e-08   0.001 (-0.1544, 0.1557)  0.995 (0.9888, 1.0016)
+1e-07   0.002 (-0.1762, 0.1804)  0.000 (-0.0061, 0.0061)
+1e-06   0.000 (-0.0102, 0.0108)  0.000 (-0.0001, 0.0001)
+1e-05   0.000 (-0.0002, 0.0003)  0.000 (-0.0000, 0.0000)
+0.0001  0.068 (0.0475, 0.0880)   0.000 (-0.0000, 0.0000)
+0.001   0.365 (0.2695, 0.4596)   0.000 (-0.0000, 0.0000)
+0.01    0.565 (0.4218, 0.7074)   0.005 (0.0034, 0.0062)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 214
+standard error method: moving block jackknife
+negative log-likelihood: 61827005013.645355
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00097036 (0.00097, 0.00097)
+pi  = 0.000752601
+mut. rate = 1.204e-08  (9.22e-09, 1.49e-08)
+Ne = 24,259 (if mu=1e-8), Ne = 12,129 (if mu=2e-8)
+R2 = 59.6862% (in-sample)  57.7389% (out-sample)
+W = 
+        cadd8                    other
+------  -----------------------  -----------------------
+1e-08   0.000 (-0.1398, 0.1399)  0.990 (0.9834, 0.9965)
+1e-07   0.000 (-0.1330, 0.1331)  0.000 (-0.0061, 0.0061)
+1e-06   0.000 (-0.0085, 0.0085)  0.000 (-0.0001, 0.0001)
+1e-05   0.000 (-0.0002, 0.0002)  0.000 (-0.0000, 0.0000)
+0.0001  0.040 (0.0316, 0.0492)   0.000 (-0.0000, 0.0000)
+0.001   0.504 (0.3939, 0.6132)   0.000 (-0.0000, 0.0000)
+0.01    0.456 (0.3568, 0.5552)   0.010 (0.0079, 0.0122)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+number jackknife samples: 263
+standard error method: moving block jackknife
+negative log-likelihood: 250687699969.473
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00135876 (0.00136, 0.00136)
+pi  = 0.00106447
+mut. rate = 1.126e-08  (7.74e-09, 1.48e-08)
+Ne = 33,968 (if mu=1e-8), Ne = 16,984 (if mu=2e-8)
+R2 = 65.4089% (in-sample)  65.1251% (out-sample)
+W = 
+        cadd8                    other
+------  -----------------------  -----------------------
+1e-08   0.013 (-0.2171, 0.2439)  0.989 (0.9817, 0.9956)
+1e-07   0.000 (-0.1466, 0.1466)  0.000 (-0.0059, 0.0059)
+1e-06   0.000 (-0.0058, 0.0058)  0.000 (-0.0001, 0.0001)
+1e-05   0.000 (-0.0002, 0.0002)  0.000 (-0.0000, 0.0000)
+0.0001  0.136 (0.0944, 0.1781)   0.000 (-0.0000, 0.0000)
+0.001   0.108 (0.0673, 0.1492)   0.000 (-0.0000, 0.0000)
+0.01    0.742 (0.5313, 0.9529)   0.011 (0.0079, 0.0148)
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsection{CADD 8\%, Sparse Track}
+\subsubsection*{CEU}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 197849464223.25174
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00101859
+pi  = 0.000801495
+mut. rate = 1.255e-08 
+Ne = 25,464 (if mu=1e-8), Ne = 12,732 (if mu=2e-8)
+R2 = 60.9207% (in-sample)  59.3772% (out-sample)
+W = 
+          cadd8
+------  -------
+1e-08     0
+1e-07     0
+1e-06     0
+1e-05     0
+0.0001    0.063
+0.001     0.354
+0.01      0.583
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CHB}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 61827132377.38223
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.000963444
+pi  = 0.000752601
+mut. rate = 1.309e-08 
+Ne = 24,086 (if mu=1e-8), Ne = 12,043 (if mu=2e-8)
+R2 = 59.7755% (in-sample)  57.9086% (out-sample)
+W = 
+          cadd8
+------  -------
+1e-08     0
+1e-07     0
+1e-06     0
+1e-05     0
+0.0001    0.032
+0.001     0.469
+0.01      0.499
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 250688310460.67032
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00134845
+pi  = 0.00106447
+mut. rate = 1.223e-08 
+Ne = 33,711 (if mu=1e-8), Ne = 16,855 (if mu=2e-8)
+R2 = 65.5761% (in-sample)  65.3002% (out-sample)
+W = 
+          cadd8
+------  -------
+1e-08     0
+1e-07     0
+1e-06     0
+1e-05     0
+0.0001    0.12
+0.001     0.106
+0.01      0.774
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
 \printbibliography

### 2023-07-20 bgspy/plots.py
rescale stuff

@@ -210,7 +210,7 @@ def savitzky_golay(y, window_size, order, deriv=0, rate=1):
 
 
 def predict_chrom_plot(model, chrom, ratio=True, center_scale=False,
-                       label='prediction', lw=2,
+                       label='prediction', data_label='data', lw=2,
                        pred_col='cornflowerblue', pi_col='0.22',
                        alpha_predict=1, alpha_pi=1,
                        xlab='position',
@@ -248,7 +248,7 @@ def predict_chrom_plot(model, chrom, ratio=True, center_scale=False,
     if center_scale:
         assert not ratio, msg
         pi = center_and_scale(pi)
-    ax.plot(midpoints, pi, label='data', linewidth=lw, c=pi_col,
+    ax.plot(midpoints, pi, label=data_label, linewidth=lw, c=pi_col,
             alpha=alpha_pi)
     if add_r2:
         ax.set_title(f"$R^2 = {np.round(model.R2(), 2)}$")

### 2023-07-20 manuscript/manuscript.tex
ms updates

@@
 
-XXX HERE
-
 Finally, we entertain two more speculative predictions. First, the equilibrium
@@
 
+\section*{Locally-rescaled Fits}
+
+Given that our simulations indicated that $B(x)$ is more accurately predicted
+when we use local-rescaling, we modified our composite-likelihood method to be
+run a second time. Intuitively, this is based on the notion that if a neutral
+allele experiences an effective population size of $B(x)N$, so too should a
+selected allele, and this should be considered in how the SC16 equations are
+solved. Immediately, there is an important caveats to note about this analysis.
+Local-rescaling is done using the first round of maximum likelihood estimates,
+and then re-solving the system of equations with a population size of $B(x)N$.
+This could introduce the possibility of some circularity, since the $B(x)$ at
+this position includes the contribution of the focal segment that is being
+rescaled. However, the qualitative results do not change based on whether the
+$B(x)$ is at averaged at the 1 kbp level or at the megabase scale; for the
+latter, a single segment makes little contribution to the average.
+
+With this caveat noted, there are three important but tentative results to draw
+from this analysis. First, the estimated mutation rate is implausibly higher,
+reaching the upper limit of our optimization boundary of $\mu = 8 \times
+10^{-8}$ across all populations in or PhastCons Priority model (Supplementary
+Material XXX). Second, all of our leave-one-chromosome-out $R_\text{LOCO}^2$
+are about one percentage point higher than the unrescaled model. This improved
+fit could be an artifact of the circularity, so it is unimportant. Third, the
+DFE estimates for PhastCons regions is now U-shaped, with $75-77\%$ of mass
+being placed on a weakly deleterious class, $s=10{^-5}$. Interestingly, this is
+the first of all of our models where such an appreciable mass has been placed
+on a midpoint in our grid; in all other cases, non-strongly estimates were
+neutral ($s=10^{-8}$). Finally, predicted pairwise diversity is nearly
+identical to our original, non-locally-rescaled fit (see Supplementary
+Materials XXX). This
+
+Overall, this leads to an open end in our work. If selected sites along the
+genome were to experience a draft effective population size that is closer to
+our locally-rescaled version, then these quantitative genetic models of linked
+selection
+
+the fit to the linked selection signal is very close under
+different parameter combinations (Supplementary Materials XXX). The fits with
+local-rescaling
+
+Intuitively, this is based on the notion that if a neutral allele experiences
+an effective population size of $B(x)N$, so too should a selected allele. We
+find in simulations this reduces error around $2 N_e s \approx 1$.
+Unfortunately, employing this in our composite-likelihood method would create
+some circularity, since the local $B(x)$ experienced at a conserved segment
+includes the reduction due to the segment itself. There are not computationally
+feasible ways to avoid this issue in the current implementation.
+
 \section*{Discussion}
+
+Linked selection processes alter genetic diversity on multiple scales, creating
+patterns based on their strength relative to the stochasticity of coalescence
+times and mutations. Sweeps typically leave signals on shorter genomic scales
+(i.e. kilobases; \cite{Akey2004-zg}) which combined with the paucity of
+adaptive substitutions in humans suggests positive selection is generating the
+apparent large, megabase-scale linked selection signal
+\parencite{Hernandez2011-gs,Murphy2022-sj}. By contrast, there is reasonable
+evidence that negative selection against new mutations occurring in
+highly-conserved sequences is generating large-scale chromosomal patterns of
+diversity.
+
+Population genetic theory has been productively applied to fit this linked
+selection signal, producing plausible estimates of the strength of selection at
+these conserved regions and mutations rates consistent with more precise
+pedigree-based estimates. However, these models rely on the canonical
+background selection model, which due to its assumption that mutations are
+sufficiently deleterious they cannot fix, has constrained statistical methods
+to only fit background selection models in regions \emph{a priori} under
+constraint. In reality, the relationship between diversity levels and the
+strength of selection at deleterious mutations is U-shaped, which implies there
+could be more uncertainty that previously appreciated in the distribution of
+weak and strongly deleterious mutations.
+
+We sought to address this uncertainty in this work, by leveraging a different
+class of linked selection models based on modeling the equilibrium fitness
+variance \parencite{Santiago2016-mu}. This model is more general than classic
+background selection theory; the background selection model is analytically a
+special case when the deleterious substitution rate is zero. Moreover, the SC16
+model explains equilibrium fitness variance across all selection coefficients
+by jointly predicting another central quantity in evolutionary genomics: the
+deleterious substitution rate in conserved features. This offers a further way
+to confirm or falsify that our understanding of the negative selection process
+is accordance with other observed quantities.
+
+However, while our improved negative selection model leads to nearly identical
+parameter estimates as previous working employing classic BGS theory, we find a
+fundamental contradiction in the implied rate of deleterious substitution and
+observed divergence in conserved regions. Unfortunately, this suggests a
+potentially serious model misspecfication or non-identifiability issue. Since
+the rate of substitutions is determined by the fraction of DFE mass below $2N_e
+s \le 1$, the consistency between our DFE estimates and those from
+\textcite{Murphy2022-sj} suggests this discrepancy is shared across models of
+negative selection.
+
+One clue towards solving this contradiction is given by our simulation work.
+While our linked selection model accurately predicts the substitution rates,
+fitness variance, and diversity for $2N_e s > 1$ and $2N_e s \ll 1$, there's a
+nominal error around the drift boundary of $2N_e s \approx 1$. This region of
+parameter space is at a delicate balance between stochasticity dominating
+allele frequency dynamics (for $2N_e s \ll 1$) and having little influence
+($2N_e s > 1$). Our implementation of the SC16 model in essence only considers
+the reduction due to selection \emph{within} the segment and not neighboring
+regions. Thus, like classic BGS methods, our model does not consider selective
+interference due to neighboring regions under selective constraint. Solving the
+SC16 equations using a locally-rescaled population size of $B(x) N$ rather than
+$N$ are in effect approximating the reduction in effective population due to
+selective interference. Intuitively, this is based on the notion that if a
+neutral allele experiences an effective population size of $B(x)N$, so too
+should a selected allele. We find in simulations this reduces error around $2
+N_e s \approx 1$. Unfortunately, employing this in our composite-likelihood
+method would create some circularity, since the local $B(x)$ experienced at a
+conserved segment includes the reduction due to the segment itself. There are
+not computationally feasible ways to avoid this issue in the current
+implementation.
+
+Out of curiosity, we experimented with fitting human diversity data, using the
+predicted $\widehat{B}(x)$ for local-rescaling, and rerunning our
+composite-likelihood framework (see Supplementary Mathematical XXX). While the
+circularity issue means these results should be interpreted with caution, we
+generally find these models fit nearly as well
+
+it creates a bit of circularity.
 

### 2023-07-20 manuscript/manuscript.tex
ms updates

@@
 identical to our original, non-locally-rescaled fit (see Supplementary
-Materials XXX). This
+Materials Figure \ref{suppfig:rescale}). However, this is not the case when the
+locally-rescaled maximum likelihood estimates are used to predict on the
+non-locally-rescaled B' maps.
 

### 2023-07-20 manuscript/supplementary.tex
ms updates

@@
 
+\subsection{Locally-Rescaled Fit Results}
+
+\begin{figure}[htbp]
+    \label{suppfig:rescale}
+  \centering
+  \includegraphics[]{figures/supplementary/rescale_pi_pred_plot.pdf}
+
+  \caption{Chromosome 2 predictions on the YRI samples, with locally-rescaled
+      model fits. Observed data is the dark gray line, and the normal MLE for
+      the PhastCons Priority model is the blue line. The locally-rescaled
+      predictions are the green line. The dashed red line are the prediction
+      using the standard B' map (without local-rescaled) and the maximum likelihood
+      estimates from the locally-rescaled fits. The large discrepancy in this shows
+      that estimates are highly dependent on the B' map.}
+\end{figure}
+
+\begin{figure}[htbp]
+    \label{suppfig:rescale-b}
+  \centering
+  \includegraphics[]{figures/supplementary/rescale_Bs.pdf}
+
+  \caption{The standard (dark gray) and rescaled (blue) B' maps for different
+  $\mu = 1.6 \times 10^{-8}$ and three different selection coefficients. For
+$s=10^{-5}$ ($2N_e s = 0.2$), locally-rescaling alters the predicted reduction
+so that it is essentially insignificant ($B \approx 1$). }
+
+\end{figure}
+
 \subsection{Strong Selection Grids}

### 2023-07-21 manuscript/supplementary.tex
ms updates

@@
   \caption{The standard (dark gray) and rescaled (blue) B' maps for different
-  $\mu = 1.6 \times 10^{-8}$ and three different selection coefficients. For
-$s=10^{-5}$ ($2N_e s = 0.2$), locally-rescaling alters the predicted reduction
-so that it is essentially insignificant ($B \approx 1$). }
+      $\mu = 1.6 \times 10^{-8}$ and three different selection coefficients.
+      For $s=10^{-5}$ ($2N_e s = 0.2$), locally-rescaling alters the predicted
+      reduction so that it is essentially insignificant ($B \approx 1$). For
+  mid-strength selection $s=10^{-4}$ ($2N_e s = 2$), there is only a very
+slight difference between standard and locally-rescaled B' maps. Finally, for
+strong selection $s=10^{-3}$ ($2N_e s = 20$), locally-rescaling does not change
+the B' maps, as expected. }
+\end{figure}
+
+\begin{figure}[htbp]
+    \label{suppfig:rescale-dfe}
+  \centering
+  \includegraphics[]{figures/supplementary/rescale_DFE.pdf}
+
+  \caption{The DFE for YRI samples with locally-rescaling. The maximum
+  likelihood mutation rate estimate for this is $\widehat{\mu} = 8 \times
+10^{-8}$, which is the upper boundary of the range used during optimization.}
 

### 2023-07-21 manuscript/manuscript.tex
ms updates

--- 2023-07-20 manuscript/manuscript.tex
@@
 \begin{figure}[htbp] \centering
-    \includegraphics[width=\textwidth]{figures/figure_7.pdf}
+    \includegraphics[width=\textwidth]{figures/figure_7_mod.pdf}
 
@@
 
-Finally, we entertain two more speculative predictions. First, the equilibrium
-additive genic fitness variance. Our PhastCons Priority model DFE estimates and
-negative selection model imply a $\widehat{V}_a = 0.45$ for the default grid,
-and $\widehat{V}_a = 0.040$ for the constrained grid. This is a considerable
-level of uncertainty---over an order of magnitude---due to a difference in the
-selection grid. There are no other known direct estimates in humans of this
-quantity for comparison. When $2N \gg 1$, theory suggests $V_a \approx
-U_D\bar{s}$ where $U_D$ is the diploid rate of strongly deleterious mutations.
-Thus, the uncertainty is a byproduct of the different average estimates of
-$\bar{s}$ across the alternate grids. Since both this work and
-\textcite{Murphy2022-sj} find the single-feature CADD 6\% track fits well, we
-can use this to gain intuition through some rough calculations. In this case,
-$U_D \approx 6\% \times 2 \times 3 \times 10^{9} \mu \approx 3.6 \times 10^{8}
-\mu$. For $\mu \approx 1.5 \times 10^{-8}$, $U_D \approx 5.2$. Under the
-constrained grid, ${\bar{s}} \approx 0.0079$ and thus $V_a \approx 0.043$; with
-the default grid $\bar{s} \approx 0.043$, and $V_a \approx 0.23$. Often fitness
-reductions of $\approx 1\%$ are assumed for the most highly-conserved regions
-(XXX, Keightly, Lynch); for comparison, this leads to $V_a \approx 0.054$.
-
-XXX HERE  USE CADD PREDICTIONS
-
-By theory, this relative fitness (scaled by the mean squared) is
-bounded from above by the Index of Selection, defined as the theoretic relative
-fitness variance if fitness were perfectly heritable \parencite{Crow1958-pc}.
-Based on rough calculation by Crow, this could range between 0.982-3.286 in
-humans. However, this is based on modern-day mortality figures that may not
-extrapolate to the timescales that determine genome-wide patterns of diversity.
-
-Second, we can calculate the reduction in relative fitness per generation due
-to deleterious substitutions. Since \textcite{Muller1950-ta}, there has been
-considerable interest in whether mildly deleterious mutations, perhaps due to
-the relaxation of selection pressures, could accumulate and fix at a faster
-pace in humans.
-
-Here, the selection grid choice has little impact
-on the estimates: we calculate that across the genome, the loss in relative
-fitness per generation is 0.00160\% for the default grid, and 0.00119\% for the
-constrained grid.
-
 \section*{Locally-rescaled Fits}
 
-Given that our simulations indicated that $B(x)$ is more accurately predicted
-when we use local-rescaling, we modified our composite-likelihood method to be
-run a second time. Intuitively, this is based on the notion that if a neutral
-allele experiences an effective population size of $B(x)N$, so too should a
-selected allele, and this should be considered in how the SC16 equations are
-solved. Immediately, there is an important caveats to note about this analysis.
-Local-rescaling is done using the first round of maximum likelihood estimates,
-and then re-solving the system of equations with a population size of $B(x)N$.
-This could introduce the possibility of some circularity, since the $B(x)$ at
-this position includes the contribution of the focal segment that is being
-rescaled. However, the qualitative results do not change based on whether the
-$B(x)$ is at averaged at the 1 kbp level or at the megabase scale; for the
-latter, a single segment makes little contribution to the average.
-
-With this caveat noted, there are three important but tentative results to draw
+Given the observed DFE-substitution rate contradiction and that our simulations
+indicated that $B(x)$ is more accurately predicted when we use local-rescaling,
+we modified our composite-likelihood method to so that it can be run a second
+time, on B' maps locally-rescaled by the predicted $\widehat{B}(x)$ from the
+initial fit. Intuitively, this is based on the notion that if a neutral allele
+experiences an effective population size of $B(x)N$, so too should a selected
+allele, and this should be considered in how the SC16 equations are solved.
+There is an important caveat about this analysis. Since local-rescaling is done
+using the first round of maximum likelihood estimates, there is a possibility
+of some circularity, since the $B(x)$ at this position includes the
+contribution of the focal segment that is being rescaled. However, the results
+do not change based on whether the $B(x)$ is at averaged at the 1 kbp level or
+at the megabase scale; for the latter, a single segment makes little
+contribution to the average.
+
+With this caveat noted, there are five important but tentative results to draw
 from this analysis. First, the estimated mutation rate is implausibly higher,
@@
 10^{-8}$ across all populations in or PhastCons Priority model (Supplementary
-Material XXX). Second, all of our leave-one-chromosome-out $R_\text{LOCO}^2$
-are about one percentage point higher than the unrescaled model. This improved
-fit could be an artifact of the circularity, so it is unimportant. Third, the
-DFE estimates for PhastCons regions is now U-shaped, with $75-77\%$ of mass
-being placed on a weakly deleterious class, $s=10{^-5}$. Interestingly, this is
-the first of all of our models where such an appreciable mass has been placed
-on a midpoint in our grid; in all other cases, non-strongly estimates were
-neutral ($s=10^{-8}$). Finally, predicted pairwise diversity is nearly
-identical to our original, non-locally-rescaled fit (see Supplementary
-Materials XXX). This
-
-Overall, this leads to an open end in our work. If selected sites along the
-genome were to experience a draft effective population size that is closer to
-our locally-rescaled version, then these quantitative genetic models of linked
-selection
-
-the fit to the linked selection signal is very close under
-different parameter combinations (Supplementary Materials XXX). The fits with
-local-rescaling
-
-Intuitively, this is based on the notion that if a neutral allele experiences
-an effective population size of $B(x)N$, so too should a selected allele. We
-find in simulations this reduces error around $2 N_e s \approx 1$.
-Unfortunately, employing this in our composite-likelihood method would create
-some circularity, since the local $B(x)$ experienced at a conserved segment
-includes the reduction due to the segment itself. There are not computationally
-feasible ways to avoid this issue in the current implementation.
+Material XXX). When we re-ran optimization with an upper limit of $\mu = 1
+\times 10^{-7}$, the mutation rate was again estimated to this (implausibly
+large) upper limit, with no considerable changes to model fit or DFE estimates.
+Second, all of our leave-one-chromosome-out $R_\text{LOCO}^2$ are about one
+percentage point higher than the unrescaled model. This improved fit could be
+an artifact of the circularity, so it is unimportant. Third, the DFE estimates
+for PhastCons regions is now U-shaped (Supplementary Materials Figure
+\ref{suppfig:rescale-dfe}), with $75-77\%$ of mass being placed on a weakly
+deleterious class, $s=10{^-5}$. Interestingly, this is the first of all of our
+models where such an appreciable mass has been placed on a midpoint in our
+grid; in all other cases, non-strongly estimates were neutral ($s=10^{-8}$).
+Fourth, predicted pairwise diversity is nearly identical to our original,
+non-locally-rescaled fit (see Supplementary Materials Figure
+\ref{suppfig:rescale}). Finally, since local-rescaling increases the DFE mass
+over $s < 10^{-3}$, mutations in PhastCons regions now have the possibility of
+fixation. We find that local-rescaling the PhastCons Priority, leads the
+predicted substitution rates in PhastCons regions to be much closer to observed
+levels (blue line, Figure \ref{fig:figure-6}).
+
+Finally, as a robustness check, we investigated the extent to which
+local-rescaling differed from the B' maps across selection parameters. We find
+minor differences between the locally-rescaled and standard B' maps in the $0.2
+\le 2N_e s \le 2$ range; the maps were identical under strong selection ($2N_e
+s= 20$) as expected (Supplementary Materials Figure \ref{suppfig:rescale-b}).
+Overall, this suggests that models that fit the linked selection are worryingly
+sensitive the theoretic B' values in the $2Ns \le 1$ domain. Local-rescaling
+turns out to only slightly alter the B' maps, yet significantly modifies the
+DFE estimates. This brings the deleterious substitution rate in agreement with
+observations (since this is predicted with a fixed $\mu = 1.5 \times 10^{-8}$;
+however, the maximum likelihood estimate of mutation is implausibly high.
 

### 2023-07-21 manuscript/manuscript.tex
ms updates

@@
 
-    \caption{}
+    \caption{Predicted versus observed divergence along the human lineage.
+        Black points are the PhyloFit divergence rate estimates per features
+        (on x-axis). Line ranges are the implied divergences across a range of
+        human-chimpanzee divergence times of 6-12 Mya (using a generate time of
+        29 years). We show the predicted divergences for our Feature
+        (turquoise) and PhastCons priority (green) annotation models.
+        Additionally, we show the predicted PhastCons region divergences when
+    local-rescaling is applied (we omit other locally-rescaled predictions
+since these to not differ substantially).}
 
@@
 the observed divergence along the human lineage for all features except
-PhastCons (Figure \ref{fig:figure-7}). As expected, the predicted substitution
+PhastCons (Figure \ref{fig:figure-6}). As expected, the predicted substitution
 rates in features unlikely to be under selective constraint (introns and UTRs,
@@
 (i.e. kilobases; \cite{Akey2004-zg}) which combined with the paucity of
-adaptive substitutions in humans suggests positive selection is generating the
-apparent large, megabase-scale linked selection signal
+adaptive substitutions in humans suggests positive selection is not generating
+the apparent large, megabase-scale linked selection signal
 \parencite{Hernandez2011-gs,Murphy2022-sj}. By contrast, there is reasonable
@@
 strength of selection at deleterious mutations is U-shaped, which implies there
-could be more uncertainty that previously appreciated in the distribution of
+could be more uncertainty than previously appreciated in the distribution of
 weak and strongly deleterious mutations.
@@
 class of linked selection models based on modeling the equilibrium fitness
-variance \parencite{Santiago2016-mu}. This model is more general than classic
-background selection theory; the background selection model is analytically a
-special case when the deleterious substitution rate is zero. Moreover, the SC16
-model explains equilibrium fitness variance across all selection coefficients
-by jointly predicting another central quantity in evolutionary genomics: the
-deleterious substitution rate in conserved features. This offers a further way
-to confirm or falsify that our understanding of the negative selection process
-is accordance with other observed quantities.
-
-However, while our improved negative selection model leads to nearly identical
-parameter estimates as previous working employing classic BGS theory, we find a
-fundamental contradiction in the implied rate of deleterious substitution and
-observed divergence in conserved regions. Unfortunately, this suggests a
-potentially serious model misspecfication or non-identifiability issue. Since
-the rate of substitutions is determined by the fraction of DFE mass below $2N_e
-s \le 1$, the consistency between our DFE estimates and those from
-\textcite{Murphy2022-sj} suggests this discrepancy is shared across models of
-negative selection.
-
-One clue towards solving this contradiction is given by our simulation work.
-While our linked selection model accurately predicts the substitution rates,
-fitness variance, and diversity for $2N_e s > 1$ and $2N_e s \ll 1$, there's a
-nominal error around the drift boundary of $2N_e s \approx 1$. This region of
-parameter space is at a delicate balance between stochasticity dominating
-allele frequency dynamics (for $2N_e s \ll 1$) and having little influence
-($2N_e s > 1$). Our implementation of the SC16 model in essence only considers
-the reduction due to selection \emph{within} the segment and not neighboring
-regions. Thus, like classic BGS methods, our model does not consider selective
-interference due to neighboring regions under selective constraint. Solving the
-SC16 equations using a locally-rescaled population size of $B(x) N$ rather than
-$N$ are in effect approximating the reduction in effective population due to
-selective interference. Intuitively, this is based on the notion that if a
-neutral allele experiences an effective population size of $B(x)N$, so too
-should a selected allele. We find in simulations this reduces error around $2
-N_e s \approx 1$. Unfortunately, employing this in our composite-likelihood
-method would create some circularity, since the local $B(x)$ experienced at a
-conserved segment includes the reduction due to the segment itself. There are
-not computationally feasible ways to avoid this issue in the current
-implementation.
-
-Out of curiosity, we experimented with fitting human diversity data, using the
-predicted $\widehat{B}(x)$ for local-rescaling, and rerunning our
-composite-likelihood framework (see Supplementary Mathematical XXX). While the
-circularity issue means these results should be interpreted with caution, we
-generally find these models fit nearly as well
-
-it creates a bit of circularity.
-
-draft limits discussion - we were heavily constrained by simulations / regions
-etc
-
-the PhastCons Priority Model is very close to CADD 6\%, and interpretable,
-similiar mutation rates. This
-
-weak selection remains elusive
-
-UTR rates
-
-- parametric DFEs are trash
-
-- the $pi_0$ problem
-
-Ipsita Agarwal article
-
-what time scale is the average mutation being lost at?
-
-Regularization / overfitting
-
-It is worth considering why this model performs well for models of negative
-selection. The fitness variance equation, $V(x) = (U-2R)s$ tells us that under
-models, the long-run variance is set by the balance of mutation and fixation.
-Whether a particular basepair's behavior follows this equation depends upon
-whether the long-run average is close to ``typical" for that basepair at a
-given time. Good things fix so quickly, that we're left with a constant
-churning of new variation
-
-- need for coalescent simulations with reduced $N_e$ along the genome.
-
-- we need to talk about Fisher's Fundamental Theorem and how this model
-connects fitness variation to absolute changes in fitness.
-
-Measuring selection in the Genome Versus Finding Selection in the Genome
+variance \parencite{Santiago2016-mu}. At its core, the theory underlying this
+model states that the linked selection signal is a function of how the fitness
+variance is distributed along the recombination map of the genome
+\parencite{Santiago1998-bs}. We fit a specific process model for this fitness
+variance that supposes all fitness variance is the result of selection against
+new deleterious mutations \parencite{Santiago2016-mu}. Unlike classic
+background selection theory, the SC16 model explains equilibrium fitness
+variance across all selection coefficients by jointly predicting another
+central quantity in evolutionary genomics: the deleterious substitution rate in
+conserved features. This model is more general than classic background
+selection theory; background selection theory is analytically a special case
+when the deleterious substitution rate is zero.
+
+Generally, our method lead to better fits to data that those based on classic
+BGS, as measured by predicted out-sample diversity. Furthermore, unlike
+BGS-based methods, our model is capable of fitting weak selection; we find
+evidence of this since we can robustly estimate the DFE for weakly constrained
+features like introns and UTRs. Yet even with the possibility of fitting weak
+selection, our main model fits are in agreement with the prior results from
+BGS-based methods. Both mutation rate estimates and DFE estimates are
+consistent with prior work. This, at first glance, suggests further
+confirmation that negative selection against new deleterious mutations is the
+dominant mode of linked selection in the genome.
+
+Since the SC16 theory calculates fitness variance by jointly predicting the
+deleterious substitution rate, this provides another route to verify our main
+findings. We find that predicted substitution rates for highly-conserved
+PhastCons features are in strong disagreement with observed substitution rates
+along the human lineage: the DFE-substitution rate contradiction. This is a
+consequence of our DFE estimate for PhastCons regions containing little mass
+over weakly deleterious and neutral selection coefficients --- a characteristic
+of DFE estimates from other work too \parencite{Murphy2022-sj}. One clue to
+this discrepancy is that our simulation results indicate an appreciable level
+of error between observed reductions and the B' maps in the weak selection
+domain of $2N_e s \approx 1$.
+
+We hypothesized that this could be because in this domain, a segment under
+selective constraint experiences a local draft-effective population size of
+$B(x)N$, and not just $N$. This local draft-effective population size is
+induced by selection at other segments that is not being taken into account by
+classic BGS theory or our standard model. When we experimented by fitting our
+model and then using the predicted reduction map to local-rescale $N_e$ to the
+draft-effective population size $N_d = \widehat{B}(x)$, we find the
+disagreement between predicted and observed substitution rates disappears. This
+is expected, since locally-rescaled fits result in DFE estimates to have an
+appreciable mass on weakly deleterious selection coefficients. While this
+brings substitution rates into agreement, we encounter the same problem found
+by \textcite{McVicker2009-ax}: the estimated mutation rate is too high. We
+emphasize that although there are several caveats with the local-rescaling
+approach, in practice it is altering the B' maps only slightly in the domain of
+$2N_e s \approx 1$. Even if the local-rescaling approximation to selective
+interference is incorrect, these results suggest parameter estimates are
+extremely sensitivity to how accurate the B' maps are in this domain.
+
+Unfortunately, this work opens up more questions about the nature of selection
+in humans that it answers. The implausibly high estimate of mutation rates
+could suggest that negative selection is not the only source of fitness
+variance generating the genome-wide linked selection signal in humans. Positive
+selection on polygenic variation along the genome could be another source of
+fitness variance contributing to the weak selection signal. Alternatively,
+negative selection could be the main source of fitness variance, but the
+complexity of selective interference may be poorly approximated by the
+local-rescaling approach. Overall, understanding how the sources of human
+fitness variance shape the striking patterns of diversity along the human
+genome still remains an open riddle.
 
 \section*{Methods}
-% (the method would support a new
-% ``interaction" type, but this greatly increases the number of parameters).
 
@@
 \cite{Kircher2014-bv,Rentzsch2019-lr}), and PhastCons regions
-\parencite{Siepel2005-wh}.
+\parencite{Siepel2005-wh}. We did not account for mutation rate heterogeneity
+along the genome, since this would require using divergence-based estimates of
+local mutation rates that would introduce circularity when we predict
+divergence rates under our model.
 

### 2023-07-21 manuscript/manuscript.tex
ms updates before rework

@@
     diversity caused by the indirect effects of selection. This can be thought
-    of as a ``linked selection signal" that reflects selection's effects
+    of as a ``linked selection signal" that reflects the effects of selection
     varying according to the placement of functional regions and recombination
@@
     cannot drift up in frequency and ultimately fix in the population. Here, we
-    develop an alternate theoretic basis for genome-wide linked selection based
-    on a quantitative genetics view, which considers how diversity is reduced
-    according to the arrangement of additive fitness variance along the genome.
-    Using a recent model that jointly predicts the equilibrium fitness variance
-    and substitution rates due to deleterious mutations, we estimate the
-    distribution of selection coefficients and mutation rate across three human
-    populations. While our model can accommodate weaker selection, we find
-    consistent evidence across all populations of very strong selection against
-    deleterious mutations. We also find that our model's predicted substitution
-    rates are qualitatively consistent with observed divergences in features
-    under weak selective constraint like introns and UTRs, but less so in
-    regions under strong selective constraint. While our model explains up to
-    67\% of the out-sample variance in diversity at the megabase scale, we find
-    evidence of residual strong selection against loss-of-function variants
-    unfit by our model, indicating the dominant role negative selection plays
-    in the human genome. Altogether, our genome-wide model of the linked
-    selection is a step towards uniting population and quantitative genetic
-    models of selection with the substitution processes.
-
+    develop and apply a statistical method to fit the linked selection signal
+    based on an alternate theoretic approach. This theory is based on a
+    quantitative genetics view of the linked selection process, which models
+    the cumulative effects of weak draft created according the arrangement of
+    polygenic additive fitness variance along the genome. Using a recent model
+    that jointly predicts the equilibrium fitness variance and substitution
+    rates due to deleterious mutations, we estimate the distribution of fitness
+    effects (DFE) and mutation rate across three human populations. While our
+    model can accommodate weaker selection, we initially find evidence across
+    three human populations of very strong selection against deleterious
+    mutations consistent with previous work. However, the corollary predictions
+    of the substitution rates from our model suggest an unreasonably low rate
+    of divergence in highly-conserved region which disagree with observations.
+    We hypothesize that this is driven by selected sites within conserved
+    segments experiencing a reduced effective population size due to selection
+    across other segments. Extending our method to approximate these effects
+    through a locally-reduced effective population size brings predicted
+    substitution rates into agreement with observed divergence rates and
+    revises DFE estimates to include a sizable portion of weakly deleterious
+    mutations. However, this also leads to implausibly large mutation rate
+    estimates. Overall, while our model of the genomic linked selection signal
+    brings us a step towards uniting population and quantitative genetic
+    selection models with the substitution process, our work suggests
+    considerable uncertainty remains about the processes generating fitness
+    variance in humans.
 \end{abstract}
@@
 
-Similar to previous work, we aim to characterize the negative selection process
+Similar to previous work, we aim to characterize the selection processes
 through the signal left on nucleotide diversity at linked sites. Although
@@
 problems.
+
+The theoretic basis of our method models the effects of weak draft on diversity
+in terms of polygenic additive fitness variance. This fitness variance could
+occur due to any weakly selected
 
@@
 the reduction factor $B(x) = \nicefrac{N_d}{N}$ at any genomic position $x$
-(see Methods \ref{sec:methods-maps}).
+(see Methods Section \nameref{sec:methods-maps}).
 
@@
 
-  \label{fig:figure-2}
+  \label{fig:figure-3}
 \end{figure}
@@
 
-  \label{fig:figure-3}
+  \label{fig:figure-4}
 \end{figure}
@@
         rate estimates are sensitive to the underlying annotation model.}
-  \label{fig:figure-4}
+  \label{fig:figure-5}
 \end{figure}
@@
 though other work has found a limited role for hitchhiking
-\parencite{Pickrell2009-tt,Hernandez2011-gs,Murphy2022}. While our simulation
+\parencite{Pickrell2009-tt,Hernandez2011-gs,Murphy2022-sj}. While our simulation
 results suggest estimates of the mutation rate from linked selection models are
@@
 
-  \label{fig:figure-5}
+  \label{fig:figure-6}
 \end{figure}
@@
 6\% basepairs and the residual across megabase windows (Supplementary Figure
-\label{suppfig:resid-cadd}), finding a negative significant relationship in
+\ref{suppfig:resid-cadd}), finding a negative significant relationship in
 both cases. Moreover, our model over-predicted diversity in windows containing
@@
 
-  \label{fig:figure-6}
+  \label{fig:figure-7}
 \end{figure}
@@
 \begin{align}
-  \label{eq:}
+  \label{eq:method-2eqns}
   {N}_{d} &= N \exp \left( -V_a \frac{Q^2(m_i, s_j)}{2} \right) & \text{\emph{effective population equation}} \\

### 2023-07-21 manuscript/supplementary.tex
ms updates before rework

@@
 \begin{align}
-  \label{eq:}
   \pi^{(1)} = \frac{1}{L} \sum_{i=1}^L \frac{n_{D,i}}{n_{T,i}}
@@
 \begin{align}
-  \label{eq:}
   \pi^{(2)} = \frac{1}{\sum_{i=1}^L n_i} \sum_{i=1}^L n_i \frac{n_{D,i}}{n_{T,i}}
@@
 \begin{align}
-  \label{eq:}
   \pi^{(3)} = \frac{\sum_{i=1}^L n_{D,i}}{\sum_{i=1}^L n_{T,i}}.
@@
 values.
-
-\subsection{Numeric Optimization}
-\label{supp:optim}
-
-mu bounds
-
-bounds and time, bounds and convergence
-
-Same optima, but took longer
 
@@
 
+As described in the Methods section we use the BOBYQA algorithm implemented by
+the NLopt library for numeric optimization to find maximum likelihood estimates
+\parencite{Powell2009-jm,Johnson2007-tl}. The DFE is a simplex, and thus
+requires constrained optimization. We compared constrained optimization
+algorithms to an unconstrained softmax approach, finding the latter vastly
+outperformed the former. Bounds for $\pi_0$ and $\mu$ can be set manually in
+\texttt{bgspy}; our main models used $10^{-8} \le \mu \le 5 \times 10^{-8}$ and
+$0.0005 \le \pi_0 \le 0.008$. All main model fits were not at bounds. For
+locally-rescaled fits, we found our $\widehat{\mu}$ estimate was continually
+the upper bound; for these fits we increased this first to $\mu = 8 \times
+10^{-8}$ and then again to $\mu = \times 10^{-7}$, find it still always reached
+the upper bound. Since $10^{-7}$ is an implausibly high mutation rate, we
+stopped there.
+
 We only show one optimization diagnostic plot here (Supplementary Figure

### 2023-07-21 manuscript/manuscript.tex
ms updates from the frankfurt airport

@@
 deleterious mutation (i.e. the ``ratchet" rate) and incorporating the impact of
-interference from other neighboring selected sites.
+interference from other neighboring selected sites. Finally, the classic BGS
+model assumes that selection occurring within other segments does not interfere
+with selection occurring in other regions, i.e. no selective or
+``Hill--Robertson" interference
+\parencite{Hill1966-kd,McVean2000-bt,Felsenstein1974-xm}.
 
@@
 
-Similar to previous work, we aim to characterize the selection processes
-through the signal left on nucleotide diversity at linked sites. Although
-nucleotide diversity is a less-informative statistic than the full site
+Similar to previous work, we aim to characterize selection processes through
+the signal left on nucleotide diversity at linked neutral sites. Although
+nucleotide diversity is a less informative statistic than the full site
 frequency spectrum, it is much easier to derive theoretic models of how linked
@@
 about selective processes are only as accurate as the underlying theoretic
-model. Our work is motivated by a well-known limitation of classic background
-selection theory: it only accurately predicts levels of linked diversity when
-new mutations are strongly deleterious
-\parencite{Charlesworth1993-gb,McVean2000-bt,Good2013-lp,Gordo2002-dr}.
-Unfortunately, modifying background selection theory so that it accommodates
-weak selection has been challenging problem
-\parencite{Good2014-yz,Haigh1978-gt,Higgs1995-xc}. Here, we develop intuition
-for why the classic BGS model works well for strong selection but not weak, and
-how formulating the equilibrium between new deleterious mutations and negative
-selection in terms of the additive fitness variance can alleviate these
-problems.
-
-The theoretic basis of our method models the effects of weak draft on diversity
-in terms of polygenic additive fitness variance. This fitness variance could
-occur due to any weakly selected
-
-Throughout, we imagine a negative selection process where a continual influx of
-deleterious mutations occur at a rate of $\mu$ per basepair per generation in a
-conserved region of $L$ basepairs, such that the region-wide per generation
-diploid mutation rate is $U = 2 \mu L$. Each mutation imposes a selective cost
-of $s$ in heterozygotes and $2s$ in homozygotes, and fitness effects are
-multiplicative across sites. When selection against a new mutation is strong,
-the stochastic perturbations it experiences due to random genetic drift are
-weak relative to the frequency changes caused by selection. In this case, every
-deleterious mutation is destined to eventual extinction, and only the fraction
-of lineages without any mutations contribute to the long-run ancestry of future
-populations. The classic background selection approximates this process by
-rescaling the neutral coalescent with an effective population size of $N_0 = N
-\exp(-\nicefrac{U}{s})$
-\parencite{Charlesworth1993-gb,Nordborg1996-nq,Hudson1995-pt,Hudson1994-oh}.
-This approximation works well for two reasons (though see
-\cite{Cvijovic2018-vd,Walczak2012-fi,Nicolaisen2012-vs}). First, the number of
-mutations per haplotype has has a stationary Poisson distribution with an
-average equal to the deterministic mutation-selection equilibrium
-\parencite{Haldane1927-ga}. Second, the time it takes for a present-day lineage
-to trace their ancestry back to a mutation-free ancestor (the ``delay phase")
-is negligible compared to the coalescence timescale of $N_0$ among
-mutation-free ancestors (\cite{Durrett2008-ql}, p. 213; \cite{Good2014-yz}).
-
-By contrast, when selection is weak, the frequency trajectories of both neutral
-and selected sites are strongly impacted by stochastic perturbations from two
-sources. First, genetic drift, which is due to Mendelian segregation and
-non-heritable fitness differences; the magnitude of its perturbations are set
-by the \emph{drift-effective} population size, $N_e$. Second, genetic draft
-occurs when an allele becomes randomly associated with a fitness background
-that perturbs its trajectory across generations
-\parencite{Neher2013-dz,Gillespie2000-mh,Gillespie2001-mv}. Like genetic drift,
-the perturbations due to draft are directionless but overall increase the
-variance in allele frequency change. Unlike drift, the impact of the stochastic
-perturbations due to draft are autocorrelated and build up over generations
-until the fitness background recombines off
-\parencite{Robertson1961-ho,Santiago1995-hx,Buffalo2019-qs}. The stochastic
-perturbations due to selection also impact selected sites, and alter their
-dynamics, creating Hill--Robertson interference \parencite{Hill1966-kd}. Strong
-draft can generate multiple-merger coalescences that create discontinuities in
-allele frequency trajectories, making diffusion approximations ill-suited
-\parencite{Gillespie2000-mh,Der2011-it,Neher2013-dz}. However, when selection
-is weak, draft is well-approximated as an increase in the rate of drift (or
-equivalently, a reduction in effective population size).
-
-The stochastic perturbations from genetic drift and draft impact both
-genealogies and selection dynamics in complex ways, which has made finding
-analytic theory predicting their joint effects difficult. When drift acts
-alone, the relevant scale of perturbations is determined by the
-\emph{drift-effective population size} $N_e$
-\parencite{Ohta1971-gq,Ohta1992-yi}. When the random perturbations due to drift
-are stronger than the changes due to selection ($2N_es \le 1$), weakly
-deleterious mutations can drift up to intermediate frequencies before their
-eventual loss or fixation. The assumptions of classic BGS suddenly breakdown in
-two ways. First, the ``delay phase" is no longer negligible and genealogies are
-no longer well-approximated by a rescaled neutral coalescence process
-\parencite{Przeworski1999-mb,OFallon2010-my,Higgs1995-xc}. Second, the
-distribution of the number of deleterious mutations (and its corresponding
-fitness distribution) is no longer a stationary Poisson distribution, instead
-becoming a traveling wave \parencite{Rouzine2008-qz,Good2013-lp,Gessler1995-hz}
-towards reduced population fitness (since we ignore beneficial mutations). When
-this occurs, individuals in the least-loaded class are stochastically lost,
-fixing those mutations with a click of ``Muller's ratchet"
-\parencite{Muller1964-ki,Charlesworth1997-qn}. Unfortunately, determining the
-rate of Muller's ratchet is another difficult, open problem
-\parencite{Haigh1978-gt,Gordo2002-dr,Gessler1995-hz} related to Hill--Robertson
-interference \parencite{Felsenstein1974-xm}.
-
-Our work extends theory from \textcite{Santiago2016-mu}, which suggests both
-strong and weak negative selection can be be accommodated with a quantitative
-genetic model of linked selection. These linked selection models stem from
+model. Since we base our method on a fundamentally different theoretic approach
+to modeling linked selection, we briefly introduce the theory was extend to
+apply across whole-genomes.
+
+Our work extends quantitative genetic models of linked selection
+\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu},
+which model the effects linked selection in terms of polygenic additive fitness
+variance ($V_A$). These linked selection models stem from
 \textcite{Robertson1961-ho}, which in essence describes how heritable polygenic
@@
 
-We define the draft-effective population size $N_d$ by including the total
-additional variance created by heritable fitness in Wright's
-(\citeyear{Wright1938-tv}) equation for effective population size,
+Following \textcite{Robertson1961-ho} and \textcite{Santiago1995-hx}, we define
+the draft-effective population size $N_d$ by including the total additional
+variance created by heritable fitness into Wright's (\citeyear{Wright1938-tv})
+equation for effective population size,
 
@@
 population size experienced by an arbitrary neutral site under the influence of
-all linked regions is,
+all $S$ linked regions is,
 
@@
 
-The remaining pieces are expressions for the equilibrium additive fitness
-variance $V_A$ and the decay rate in associated fitness $1-\kappa$. At this
-point, we diverge from Santiago and Caballero (\citeyear{Santiago1998-bs},
-\citeyear{Santiago2016-mu}) to note that the additive fitness variation is the
-sum of additive \emph{genic} fitness variance $V_a$ (i.e. due to allele
-frequencies) and the fitness covariance due to linkage disequilibria between
-selected sites ($\delta_{LD}$), $V_A = V_a + \delta_{LD}$. Considering only the
-genic variation, at equilibrium $\Delta V_{a,i} = 0$; thus, the loss in genic
-fitness each generation due to selective dynamics $-\kappa V_{A,i}$ must be
-equal to the increase in fitness variation due to new mutations each generation
-($V_m$) and changes in the LD term \parencite{Bulmer1971-ae}. Note that the
-loss in genic fitness due to drift is not time-invariant due to the build up of
-autocorrelation, and considered elsewhere in the derivation (see Supplementary
-Materials Equation \ref{eq:vardecay}). Then, at equilibrium $\kappa_i =
-\nicefrac{V_{m,i}}{V_{a,i}}$ (see Supplementary Materials Equation \ref{eq:Z}).
-
-Under any selection model, the additive genic fitness variance created by a new
-mutation (at frequency $x=\nicefrac{1}{2N}$) is $2s^2 x(1-x) \approx
+The remaining required expressions are the equilibrium additive fitness
+variance $V_A$ and the decay rate in associated fitness $1-\kappa$. Fitness
+variance could arise from positively or negatively selected alleles, but given
+prior work has found negative selection against new deleterious mutations play
+a dominant role in shaping genome-wide patterns of diversity
+\parencite{McVicker2009-ax,Murphy2022-sj}, we focus on models of fitness
+variance created by the continual influx of deleterious mutations into a
+population. We imagine a mutation-selection process that creates fitness
+variance through a continual influx of deleterious mutations occur at a rate of
+$\mu$ per basepair per generation in a conserved region of $L$ basepairs, such
+that the region-wide per generation diploid mutation rate is $U = 2 \mu L$.
+Each mutation imposes a selective cost of $s$ in heterozygotes and $2s$ in
+homozygotes, and fitness effects are multiplicative across sites.
+
+Under this selection model, the additive genic fitness variance created by a
+new mutation (at frequency $x=\nicefrac{1}{2N}$) is $2s^2 x(1-x) \approx
 \nicefrac{s^2}{N}$. For the entire population of $2N$ chromosomes, the fresh
@@
 U_is^2$ where $U_i = 2\mu L_i$ is the diploid mutation rate per generation
-within the segment. Under mutation-selection balance assumed by BGS, an
-$L_i$-basepair segment has genic variance $V_{a,i}^{BGS} \approx U_i s$ (see
-Supplementary Materials Equation \ref{supp-eq:va_bgs}) and thus $\kappa_i^{BGS}
-= s$. Substituting $V_{a,i}^{BGS}$ and $\kappa_i^{BGS}$ in Equation
-\eqref{eq:polygenic_Ne} and simplifying, we have
+within the segment. Under the mutation-selection balance assumed by classic BGS
+theory, an $L_i$-basepair segment has equilibrium genic variance $V_{a,i}^{BGS}
+\approx U_i s$ (see Supplementary Materials Equation \ref{supp-eq:va_bgs}) and
+thus $\kappa_i^{BGS} = s$. Substituting $V_{a,i}^{BGS}$ and $\kappa_i^{BGS}$ in
+Equation \eqref{eq:polygenic_Ne} and simplifying, we have
 
@@
 
-When selection is weak, however, the additive fitness variance under the
-classic BGS model is no longer accurate. \textcite{Santiago2016-mu} suggested
-that this is due to the loss of fitness variation that occurs when a
-segregating site fixes and its heterozygosity becomes zero. If we let $R$
-represent the fixation rate (i.e. ratchet rate) in the region per generation,
-each fixation removes the equivalent amount of equilibrium variation put in by
-mutation. Thus, the steady-state genic variance under mutation and negative
-selection is (omitting the segment index $i$ for clarity),
+However, when new mutations are only weakly deleterious, they can drift up in
+frequency before their eventual loss or fixation. At this point, the number of
+deleterious mutations is no longer well-approximated by the deterministic
+mutation-selection-balance theory; their dynamics are strongly influenced by
+stochastic perturbations due to both drift and draft. In this weak selection
+regime, classic BGS no longer accurately predicts levels of linked diversity
+\parencite{Charlesworth1993-gb,McVean2000-bt,Good2013-lp,Gordo2002-dr}.
+Moreover, selection against weakly deleterious mutations alters the
+genealogies, such that they are no longer well-approximated by a rescaled
+neutral coalescent as under the classic BGS model
+\parencite{Przeworski1999-mb,OFallon2010-my,Higgs1995-xc}. To further
+complicate matters, the distribution of the number of deleterious mutations
+(and its corresponding fitness distribution) is no longer a stationary Poisson
+distribution, instead becoming a traveling wave
+\parencite{Rouzine2008-qz,Good2013-lp,Gessler1995-hz} towards reduced
+population fitness (since we ignore beneficial mutations). When this occurs,
+individuals in the least-loaded class are stochastically lost, fixing those
+mutations with a click of ``Muller's ratchet"
+\parencite{Muller1964-ki,Charlesworth1997-qn}. Unfortunately, determining the
+rate of Muller's ratchet is another difficult, open problem
+\parencite{Haigh1978-gt,Gordo2002-dr,Gessler1995-hz,Higgs1995-xc} related to
+Hill--Robertson interference \parencite{Felsenstein1974-xm}.
+
+Under qualitative genetic models of linked selection, these complications can
+can be avoided by finding an expression for the equilibrium fitness variance
+that is correct for both strong and weak deleterious mutations.
+
+\textcite{Santiago2016-mu} suggested that this is due to the loss of fitness
+variation that occurs when a segregating site fixes and its heterozygosity
+becomes zero. If we let $R$ represent the fixation rate (i.e. ratchet rate) in
+the region per generation, each fixation removes the equivalent amount of
+equilibrium variation put in by mutation. Thus, the steady-state genic variance
+under mutation and negative selection is (omitting the segment index $i$ for
+clarity),
 

### 2023-07-21 manuscript/supplementary.tex
ms updates from the frankfurt airport

@@
 
-  \caption{The standard (dark gray) and rescaled (blue) B' maps for different
+  \caption{The standard (dark gray) and locally-rescaled (blue) B' maps for different
       $\mu = 1.6 \times 10^{-8}$ and three different selection coefficients.

### 2023-07-23 manuscript/manuscript.tex
ms updates from italy

@@
     We hypothesize that this is driven by selected sites within conserved
-    segments experiencing a reduced effective population size due to selection
-    across other segments. Extending our method to approximate these effects
-    through a locally-reduced effective population size brings predicted
-    substitution rates into agreement with observed divergence rates and
-    revises DFE estimates to include a sizable portion of weakly deleterious
-    mutations. However, this also leads to implausibly large mutation rate
-    estimates. Overall, while our model of the genomic linked selection signal
-    brings us a step towards uniting population and quantitative genetic
-    selection models with the substitution process, our work suggests
-    considerable uncertainty remains about the processes generating fitness
-    variance in humans.
+    segments experiencing a further reduced effective population size due to
+    selection across other segments. Extending our method to approximate these
+    effects through a locally-reduced effective population size brings
+    predicted substitution rates into agreement with observed divergence rates
+    and revises DFE estimates to include a sizable portion of weakly
+    deleterious mutations. However, this also leads to implausibly large
+    mutation rate estimates. Overall, while our model of the genomic linked
+    selection signal brings us a step towards uniting population and
+    quantitative genetic selection models with the substitution process, our
+    work suggests considerable uncertainty remains about the processes
+generating fitness variance in humans.
+
 \end{abstract}
@@
 rescaling the neutral coalescent by a reduction factor known as $B =
-\nicefrac{N}{N_e}$ \parencite{Charlesworth2013-kl}. However, this simple
+\nicefrac{N_e}{N}$ \parencite{Charlesworth2013-kl}. However, this simple
 rescaling approach is not appropriate across parts of parameter space that are
@@
 simulations, we show this model leads to more accurate estimates of the
-distribution of fitness effects under weak selection. Finally, we apply our
+distribution of fitness effects under weak selection. We apply our
 composite-likelihood method to human population genomic data and provide new
-estimates of the negative selection process in humans.
+estimates of the negative selection process in humans. However, we find that
+there may be considerable uncertainty around these estimates if selective
+interference is occurring across segments under purifying selection.
 
@@
 
-The remaining required expressions are the equilibrium additive fitness
+The remaining required expressions are for the equilibrium additive fitness
 variance $V_A$ and the decay rate in associated fitness $1-\kappa$. Fitness
@@
 genealogies, such that they are no longer well-approximated by a rescaled
-neutral coalescent as under the classic BGS model
+neutral coalescent as assumed under the classic BGS model
 \parencite{Przeworski1999-mb,OFallon2010-my,Higgs1995-xc}. To further
@@
 
-Under qualitative genetic models of linked selection, these complications can
-can be avoided by finding an expression for the equilibrium fitness variance
-that is correct for both strong and weak deleterious mutations.
-
-\textcite{Santiago2016-mu} suggested that this is due to the loss of fitness
-variation that occurs when a segregating site fixes and its heterozygosity
-becomes zero. If we let $R$ represent the fixation rate (i.e. ratchet rate) in
-the region per generation, each fixation removes the equivalent amount of
-equilibrium variation put in by mutation. Thus, the steady-state genic variance
-under mutation and negative selection is (omitting the segment index $i$ for
-clarity),
+Under qualitative genetic models of linked selection, these complications are
+avoided by finding a general expression for the equilibrium fitness variance
+that is applicable for both strong and weakly deleterious mutations.
+\textcite{Santiago2016-mu} suggest that a general equation for equilibrium
+fitness variance must account for the chance fixations of deleterious mutations
+that can occur. Each fixation (or click of Muller's ``ratchet") decreases the
+fitness variance, as a previously segregating polymorphism fixes and its
+heterozygosity becomes zero. If we let $R$ represent the fixation rate (i.e.
+ratchet rate) in the region per generation, each fixation removes the
+equivalent amount of equilibrium variation put in by mutation. Thus, the
+steady-state genic variance under mutation and negative selection is (omitting
+the segment index $i$ for clarity),
 
@@
 coefficients. This is done for each pre-specified segment in the genome that
-may be under negative selection (e.g. coding sequences or UTRs). Then, these
-equilibrium can be used to calculate a reduction map $B(x)$ across genomic
-positions $x$. To distinguish between McVicker's B maps based on classic BGS
-theory, we call our reduction map $B(x)$ a B' map.
+may be under negative selection (e.g. coding sequences or UTRs). We call the
+set of predicted reductions across these grids the B' maps (to distinguish them
+from McVicker's B maps); these can be used then to find the equilibrium
+reduction factor $B(x)$ for any genomic position $x$.
 
@@
 independently. In particular, we use a fixed drift-effective $N=1000$
-corresponding to the number of diploids in the simulations rather than $B(x) N$
-at position $x$. To test this, we implemented a ``locally-rescaled" version of
-the B' maps, which numerically solves Equations \eqref{eq:main_eqns} for each
-segment at position $x$ taking into account the reduction experienced due to
-selection at other segments by setting $B'(x)N$ as the population size. We find
-the locally-rescaled B' maps reduce the relative error from 5\% to 0.4\% and
-mean squared error (Figure \ref{fig:figure-1}D, dashed colored lines), but does
-not entirely eliminate the error in the $2Ns \approx 1$ domain. Overall, while
-we find evidence that local-rescaling slightly reduces error in the $2Ns
-\approx 1$ domain, we do not use it during in our genome-wide inference method.
-This is because to avoid circularity and possible overfitting, our method would
-need to solve the equilibria equations for all segments simultaneously;
-unfortunately, this is currently computational unfeasible.
+corresponding to the number of diploids in the simulations rather than $B(x)
+N$, which is an approximation to the draft-effective population size
+experienced by a selected segment at position $x$. To test this, we implemented
+a ``locally-rescaled" version of the B' maps, which numerically solves
+Equations \eqref{eq:main_eqns} for each segment at position $x$ taking into
+account the reduction experienced due to selection at other segments by setting
+$B(x)N$ as the population size. We find the locally-rescaled B' maps reduce the
+relative error from 5\% to 0.4\% and mean squared error (Figure
+\ref{fig:figure-1}D, dashed colored lines), but does not entirely eliminate the
+error in the $2Ns \approx 1$ domain. Overall, while we find evidence that
+local-rescaling slightly reduces error in the $2Ns \approx 1$ domain, we do not
+use it during in our genome-wide inference method. This is because to avoid
+circularity and possible overfitting, our method would need to solve the
+equilibria equations for all segments simultaneously; unfortunately, this is
+currently computational unfeasible.
 
@@
 simulations shown in Figure \ref{fig:figure-1}A-C can be found Supplementary
-Section \ref{supp:simulation}.
+Materials Section \ref{supp:simulation}.
 

### 2023-07-23 manuscript/manuscript.tex
ms updates from italy

@@
 
-(see Appendix Equation \ref{eq:Qinf}). This general equation can accommodate
+(see Supplementary Materials Equation \ref{supp-eq:Qinf}). This general equation can accommodate
 models of polygenic selection as long as the equilibrium additive fitness
@@
 Our PhastCons Priority model fits marginally better than the CADD 6\% model
-consistently across all three populations (Supplementary Table
+consistently across all three populations (Supplementary Materials Table
 \ref{supp:tbl-r2}). This is likely because the PhastCons Priority model has
@@
 1.56 \times 10^{-8}$ for YRI, $1.64 \times 10^{-8}$ for CEU, and $1.60 \times
-10^{-8}$ for CHB reference samples (Supplementary Table \ref{supp:tbl-r2}). As
-expected, the sparse-track CADD model mutation rate estimates are nearly
-identical between the B' and classic BGS methods (Figure \ref{fig:figure-4}A
-top row).
+10^{-8}$ for CHB reference samples (Supplementary Materials Table
+\ref{supp:tbl-r2}). As expected, the sparse-track CADD model mutation rate
+estimates are nearly identical between the B' and classic BGS methods (Figure
+\ref{fig:figure-4}A top row).
 
@@
 6\% model consistently led to slightly higher mutation rates than the CADD 8\%
-model (Figure \ref{fig:figure-4}A bottom row; Supplementary Table
+model (Figure \ref{fig:figure-4}A bottom row; Supplementary Materials Table
 \ref{supp:tbl-r2}). This same pattern was observed in Murphy et al. too (2023;

### 2023-07-23 manuscript/supplementary.tex
ms updates from italy

@@
 \begin{align}
-  \label{eq:Qinf}
+  \label{supp-eq:Qinf}
   Q_\infty &= \frac{1}{\kappa + r(1-\kappa)}.
@@
 \subsection{Table of $R^2$ Values}
-
+\label{supp:tbl-r2}
 \begin{table}
-    \label{supp:tbl-r2}
 \centering
@@
 
+\subsection{Table of $R^2$ Values (Strong Selection Grid)}
+\label{supp:tbl-r2-strongsel}
 \begin{table}
-    \label{supp:tbl-r2-strongsel}
 \centering
@@
 
-\begin{figure}[htbp]
-  \label{suppfig:jackknife-iqr}
-  \centering
-  \includegraphics[width=0.7\textwidth]{figures/supplementary/iqr_jackknife_trim.pdf}
-
-  \caption{The distribution of block jackknife samples for the Feature Priority
-  model fit on YRI data. The ranges at the top of the figure indicate the
-untrimmed and the 2 IQR-trimmed jackknife intervals. Trimming off the numeric
-optimization outliers leads to more realistic estimates of uncertainty.}
-
-\end{figure}
-
 We use a block jackknife procedure to calculate the uncertainty of our
@@
 remaining data. Then, using these jackknife samples we calculated the jackknife
-standard error for each of parameters \parencite{Wasserman2004-ip}. In some
-cases, our jackknife standard error estimates were very large. Upon inspection
-of optimization diagnostic plots (see Supplementary Materials Section
-\ref{suppfig:diag-plot}) and histograms of jackknife fits (Supplementary Figure
-\ref{suppfig:jackknife-iqr}), these were due to rare aberrant optima. We
-removed these using a trimmed-interquartile range, which removes points that
-are 2 interquartile ranges below the first quartile or above the third quartile
-(1.5 is standard, but we chose to be more conservative to not underestimate
-uncertainty).
+standard error for each of parameters \parencite{Wasserman2004-ip}.
+
+\begin{figure}[htbp]
+  \label{suppfig:diag-plot-mode}
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/phastcons_diag_plot.pdf}
+
+  \caption{An optimization diagnostic plot showing an example of the bimodal likelihood
+  optima, or a numeric optimization issue.}
+\end{figure}
+
+In one case, for our PhastCons Priority model mutation rate estimtes, our
+jackknife standard error estimates were very large. Upon inspection of
+optimization diagnostic plots (see Supplementary Materials Section
+\ref{suppfig:diag-plot-mode}) and histograms of jackknife estimates, these
+looked like they were due to rare aberrant optima. These were either caused by
+a numeric optimization issue or another likelihood mode. Given that this looked
+to be artificially inflating the uncertainty compared to all other models, we
+filtered out these jackknife results before computing the standard error.
 
@@
 \begin{figure}[htbp]
-    \label{suppfig:rescale}
   \centering
@@
       that estimates are highly dependent on the B' map.}
+    \label{suppfig:rescale}
 \end{figure}
@@
 \begin{figure}[htbp]
-    \label{suppfig:rescale-b}
   \centering
@@
 the B' maps, as expected. }
+    \label{suppfig:rescale-b}
 \end{figure}
@@
 \begin{figure}[htbp]
-    \label{suppfig:rescale-dfe}
   \centering
@@
 
+    \label{suppfig:rescale-dfe}
 \end{figure}
@@
 \begin{figure}[htbp]
-    \label{suppfig:strong-sel-dfe}
   \centering
@@
   \caption{The DFE estimates for the strong selection grid (up to $s=10^{-1}$).}
+    \label{suppfig:strong-sel-dfe}
 \end{figure}
@@
 \begin{figure}[htbp]
-    \label{suppfig:sel-ident}
   \centering
@@
 
+    \label{suppfig:sel-ident}
 \end{figure}
@@
 \begin{figure}[htbp]
-    \label{suppfig:strong-sel-dfe-pop-phastcons}
   \centering
@@
   under the Feature Priority model, across populations.}
+
+    \label{suppfig:strong-sel-dfe-pop-phastcons}
 \end{figure}
@@
 \begin{figure}[htbp]
-    \label{suppfig:strong-sel-dfe-pop-feature}
   \centering
@@
   under the PhastCons Priority model, across populations.}
+    \label{suppfig:strong-sel-dfe-pop-feature}
 \end{figure}
@@
 \begin{figure}[htbp]
-    \label{suppfig:sim-exp}
   \centering
@@
 
+    \label{suppfig:sim-exp}
 \end{figure}
@@
 \begin{figure}[htbp]
-  \label{suppfig:sim-rec}
   \centering
@@
 
+  \label{suppfig:sim-rec}
 \end{figure}
@@
 \begin{figure}[htbp]
-  \label{suppfig:resid-cadd}
   \centering
@@
   basepairs in a window annotated by a CADD region.}
+  \label{suppfig:resid-cadd}
 \end{figure}
@@
 \begin{figure}[htbp]
-  \label{suppfig:resid-sweeps}
   \centering
@@
 
+  \label{suppfig:resid-sweeps}
 \end{figure}
@@
 \begin{figure}[htbp]
-  \label{suppfig:resid-b}
   \centering
@@
            value.}
+  \label{suppfig:resid-b}
 \end{figure}
@@
 \begin{figure}[htbp]
-  \label{suppfig:resid-pc-fit}
   \centering
@@
 selection signal is not dependent on the model used (PhastCons Priority versus CADD 6\%).}
+  \label{suppfig:resid-pc-fit}
 \end{figure}
@@
 \begin{figure}[htbp]
-  \label{suppfig:dfe-ceu}
   \centering
@@
   DFE for YRI samples in Figure \ref{fig:figure-3}.}
+  \label{suppfig:dfe-ceu}
 \end{figure}
@@
 \begin{figure}[htbp]
-    \label{suppfig:dfe-chb}
   \centering
@@
   DFE for YRI samples in Figure \ref{fig:figure-3}.}
+    \label{suppfig:dfe-chb}
 \end{figure}

### 2023-07-23 manuscript/manuscript.tex
ms updates from italy

@@
 allele, and this should be considered in how the SC16 equations are solved.
-There is an important caveat about this analysis. Since local-rescaling is done
-using the first round of maximum likelihood estimates, there is a possibility
-of some circularity, since the $B(x)$ at this position includes the
-contribution of the focal segment that is being rescaled. However, the results
-do not change based on whether the $B(x)$ is at averaged at the 1 kbp level or
-at the megabase scale; for the latter, a single segment makes little
-contribution to the average.
-
-With this caveat noted, there are five important but tentative results to draw
-from this analysis. First, the estimated mutation rate is implausibly higher,
-reaching the upper limit of our optimization boundary of $\mu = 8 \times
-10^{-8}$ across all populations in or PhastCons Priority model (Supplementary
-Material XXX). When we re-ran optimization with an upper limit of $\mu = 1
-\times 10^{-7}$, the mutation rate was again estimated to this (implausibly
-large) upper limit, with no considerable changes to model fit or DFE estimates.
-Second, all of our leave-one-chromosome-out $R_\text{LOCO}^2$ are about one
-percentage point higher than the unrescaled model. This improved fit could be
-an artifact of the circularity, so it is unimportant. Third, the DFE estimates
-for PhastCons regions is now U-shaped (Supplementary Materials Figure
+There is an important caveat about this analysis.
+
+There are five important but tentative results to draw from this analysis.
+First, the estimated mutation rate is implausibly higher, reaching the upper
+limit of our optimization boundary of $\mu = 8 \times 10^{-8}$ across all
+populations in or PhastCons Priority model (Supplementary Material XXX). When
+we re-ran optimization with an upper limit of $\mu = 1 \times 10^{-7}$, the
+mutation rate was again estimated to this (implausibly large) upper limit, with
+no considerable changes to model fit or DFE estimates. Second, all of our
+leave-one-chromosome-out $R_\text{LOCO}^2$ are about one percentage point
+higher than the unrescaled model. This improved fit could be an artifact of the
+circularity, so it is unimportant. Third, the DFE estimates for PhastCons
+regions is now U-shaped (Supplementary Materials Figure
 \ref{suppfig:rescale-dfe}), with $75-77\%$ of mass being placed on a weakly
@@
 
-Finally, as a robustness check, we investigated the extent to which
-local-rescaling differed from the B' maps across selection parameters. We find
-minor differences between the locally-rescaled and standard B' maps in the $0.2
-\le 2N_e s \le 2$ range; the maps were identical under strong selection ($2N_e
-s= 20$) as expected (Supplementary Materials Figure \ref{suppfig:rescale-b}).
+Finally, we one important caveat about this analysis. Since local-rescaling is
+done using the first round of maximum likelihood estimates, there is a
+possibility of some circularity, since the $B(x)$ at this position includes the
+contribution of the focal segment that is being rescaled. However, two
+observations indicate this would matter little. First, the results do not
+change based on whether the $B(x)$ is at averaged at the 1 kbp level or at the
+megabase scale; for the latter, a single segment makes little contribution to
+the average. Second, we investigated the extent to which local-rescaling
+differed from the B' maps across selection parameters. We find minor
+differences between the locally-rescaled and standard B' maps in the
+nearly-neutral domain ($0.2 \le 2N_e s \le 2$); the maps were identical under
+strong selection ($2N_e s= 20$) as expected (Supplementary Materials Figure
+\ref{suppfig:rescale-b}). Moreover, the correlations between the standard and
+local-rescaled B' maps across the genome are high (100\% for $2N_e s = 20$,
+96.5\% for $2N_e s = 2$, and 60.21\% for $2N_e s = 0.2$). The overall realized
+effect of local-rescaling is to just alter how deep the ``U" is in the
+relationship between the reduction factor $B$ and the selection coefficient
+(Supplementary Materials Figure \ref{suppfig:rescale-effects}).
+
 Overall, this suggests that models that fit the linked selection are worryingly
@@
 appreciable mass on weakly deleterious selection coefficients. While this
-brings substitution rates into agreement, we encounter the same problem found
+brings substitution rates into agreement, we encounter a similar problem found
 by \textcite{McVicker2009-ax}: the estimated mutation rate is too high. We
@@
 approach, in practice it is altering the B' maps only slightly in the domain of
-$2N_e s \approx 1$. Even if the local-rescaling approximation to selective
-interference is incorrect, these results suggest parameter estimates are
-extremely sensitivity to how accurate the B' maps are in this domain.
+$2N_e s \approx 1$ (deepening the bottom of the ``U"). Even if the local-rescaling
+approximation to selective interference is incorrect, these results suggest
+parameter estimates are extremely sensitivity to how accurate the B' maps are
+in this domain.
 
@@
 complexity of selective interference may be poorly approximated by the
-local-rescaling approach. Overall, understanding how the sources of human
-fitness variance shape the striking patterns of diversity along the human
-genome still remains an open riddle.
+local-rescaling approach. Another possibility is that our assumption throughout
+of additive effects may lead to biases given deleterious mutations likely
+recessive effects. Overall, understanding how the sources of human fitness
+variance shape the striking patterns of diversity along the human genome still
+remains an open riddle.
 

### 2023-07-24 manuscript/manuscript.tex
ms updates from italy

@@
     diversity caused by the indirect effects of selection. This can be thought
-    of as a ``linked selection signal" that reflects the effects of selection
+    of as a ``linked selection signal" that reflects the impact of selection
     varying according to the placement of functional regions and recombination
@@
     theory, is only applicable when new mutations are so deleterious that they
-    cannot drift up in frequency and ultimately fix in the population. Here, we
-    develop and apply a statistical method to fit the linked selection signal
-    based on an alternate theoretic approach. This theory is based on a
-    quantitative genetics view of the linked selection process, which models
-    the cumulative effects of weak draft created according the arrangement of
-    polygenic additive fitness variance along the genome. Using a recent model
-    that jointly predicts the equilibrium fitness variance and substitution
-    rates due to deleterious mutations, we estimate the distribution of fitness
-    effects (DFE) and mutation rate across three human populations. While our
-    model can accommodate weaker selection, we initially find evidence across
-    three human populations of very strong selection against deleterious
-    mutations consistent with previous work. However, the corollary predictions
-    of the substitution rates from our model suggest an unreasonably low rate
-    of divergence in highly-conserved region which disagree with observations.
-    We hypothesize that this is driven by selected sites within conserved
-    segments experiencing a further reduced effective population size due to
-    selection across other segments. Extending our method to approximate these
-    effects through a locally-reduced effective population size brings
-    predicted substitution rates into agreement with observed divergence rates
-    and revises DFE estimates to include a sizable portion of weakly
-    deleterious mutations. However, this also leads to implausibly large
-    mutation rate estimates. Overall, while our model of the genomic linked
-    selection signal brings us a step towards uniting population and
-    quantitative genetic selection models with the substitution process, our
-    work suggests considerable uncertainty remains about the processes
-generating fitness variance in humans.
+    cannot fix in the population. Here, we develop a statistical method based
+    on a quantitative genetics view of the linked selection, which models the
+    effects of weak draft created according to how polygenic additive fitness
+    variance is distributed along the genome. We use a recent model that
+    jointly predicts the equilibrium fitness variance and substitution rates
+    due to both strong and weakly deleterious mutations, we estimate the
+    distribution of fitness effects (DFE) and mutation rate across three human
+    populations. While our model can accommodate weaker selection, we initially
+    find evidence across three human populations of very strong selection
+    against deleterious mutations consistent with previous work. However, the
+    corollary predicted substitution rates for conserved regions are
+    unreasonably low, and in disagreement with observed rates. We hypothesize
+    this could be due to selected sites experiencing a further diminished
+    population size due to selective interference. When we account for this in
+    our method, we find evidence of weakly deleterious mutations in conserved
+    regions which brings the predicted substitution rate into agreement with
+    observations. However, these models lead to implausibly large mutation rate
+    estimates. Overall, while our model of the genomic linked selection signal
+    brings us a step towards uniting population and quantitative genetic
+    selection models with the substitution process, our work suggests
+    considerable uncertainty remains about the processes generating fitness
+    variance in humans.
 
@@
 
-Similar to previous work, we aim to characterize selection processes through
-the signal left on nucleotide diversity at linked neutral sites. Although
-nucleotide diversity is a less informative statistic than the full site
-frequency spectrum, it is much easier to derive theoretic models of how linked
-selection alters expected pairwise coalescence times than to describe the
-entire genealogy. Still, inferences drawn from the linked selection signal
-about selective processes are only as accurate as the underlying theoretic
-model. Since we base our method on a fundamentally different theoretic approach
-to modeling linked selection, we briefly introduce the theory was extend to
-apply across whole-genomes.
+% Similar to previous work, we aim to characterize selection processes through
+% the signal left on nucleotide diversity at linked neutral sites. Although
+% nucleotide diversity is a less informative statistic than the full site
+% frequency spectrum, it is much easier to derive theoretic models of how linked
+% selection alters expected pairwise coalescence times than to describe the
+% entire genealogy. Still, inferences drawn from the linked selection signal
+% about selective processes are only as accurate as the underlying theoretic
+% model. Since we base our method on a fundamentally different theoretic approach
+% to modeling linked selection, we briefly introduce the theory was extend to
+% apply across whole-genomes.
 
@@
 \textcite{Robertson1961-ho}, which in essence describes how heritable polygenic
-fitness variation generates weak genetic draft. At the individual level
-Robertson considered, selection generates autocorrelation as offspring from
-large families tend to beget many descendents themselves (and likewise with
-small families) when fitness is heritable. This same autocorrelation occurs at
-the genomic level with linkage \parencite{Santiago1998-bs,Barton2000-zg}, as
-the perturbations to a neutral allele's trajectory from its particular fitness
+fitness variation causes weak genetic draft. At the individual level Robertson
+considered, selection generates autocorrelation as offspring from large
+families tend to beget many descendents themselves (and likewise with small
+families) when fitness is heritable. This same autocorrelation occurs at the
+genomic level with linkage \parencite{Santiago1998-bs,Barton2000-zg}, as the
+perturbations to a neutral allele's trajectory from its particular fitness
 background tend to occur in the same direction across generations until the
@@
 contributes additive fitness variance $V_{A,i}$ to the total additive genetic
-fitness variance. Under equilibrium levels of fitness variation, the
-autocorrelation function for a neutral allele associated with segment $i$ is
-$C(t) = [(1-r_i)(1-\kappa_i)]^t$, where $r_i$ is the recombination fraction to
-the segment and $\kappa_i$ is the rate that the associated fitness variance
-decays due to selective dynamics. Then, the cumulative autocorrelation is,
+fitness variance. The autocorrelation function for a neutral allele associated
+with segment $i$ is $C(t) = [(1-r_i)(1-\kappa_i)]^t$, where $r_i$ is the
+recombination fraction to the segment and $\kappa_i$ is the rate that the
+associated fitness variance decays due to selective dynamics. Then, the
+cumulative autocorrelation over the lifespan of the allele is,
 
@@
 \end{align}
-
+%
 (see Supplementary Materials Equation \ref{supp-eq:Qinf}). This general equation can accommodate
@@
 \end{align}
-
+%
 which is identical to the genome-wide model of background selection used in
@@
 frequency before their eventual loss or fixation. At this point, the number of
-deleterious mutations is no longer well-approximated by the deterministic
-mutation-selection-balance theory; their dynamics are strongly influenced by
-stochastic perturbations due to both drift and draft. In this weak selection
-regime, classic BGS no longer accurately predicts levels of linked diversity
+deleterious mutations per haplotype is no longer well-approximated by the
+deterministic mutation-selection-balance theory; their dynamics are strongly
+influenced by stochastic perturbations due to both drift and draft. In this
+weak selection regime, classic BGS no longer accurately predicts levels of
+linked diversity
 \parencite{Charlesworth1993-gb,McVean2000-bt,Good2013-lp,Gordo2002-dr}.
@@
 
-Under qualitative genetic models of linked selection, these complications are
-avoided by finding a general expression for the equilibrium fitness variance
-that is applicable for both strong and weakly deleterious mutations.
-\textcite{Santiago2016-mu} suggest that a general equation for equilibrium
-fitness variance must account for the chance fixations of deleterious mutations
-that can occur. Each fixation (or click of Muller's ``ratchet") decreases the
-fitness variance, as a previously segregating polymorphism fixes and its
-heterozygosity becomes zero. If we let $R$ represent the fixation rate (i.e.
-ratchet rate) in the region per generation, each fixation removes the
-equivalent amount of equilibrium variation put in by mutation. Thus, the
-steady-state genic variance under mutation and negative selection is (omitting
-the segment index $i$ for clarity),
+Under qualitative genetic models of linked selection, these complications can
+be avoided by finding a general expression for the equilibrium $V_A$ applicable
+for both strong and weakly deleterious mutations. \textcite{Santiago2016-mu}
+suggest that equilibrium fitness variance is lower than predicted by
+$V_{a}^{BGS}$ once weakly deleterious mutations begin to have an appreciable
+rate of fixation, $R > 0$, since each substitution (or click of Muller's
+``ratchet") removes a segregating site and its heterozygosity becomes zero.
+Thus, the steady-state genic variance under mutation and negative selection is
+(omitting the segment index $i$ for clarity),
 
@@
 is held when considering deleterious mutations. This equation connects the
-equilibrium additive genic fitness variance to the flux of new variation in due
-to new deleterious mutations and the flux out due to their substitution and the
-decline in mean population fitness. When $R=0$, selection is so strong it
+equilibrium additive genic fitness variance to the flux of new variation in
+from new deleterious mutations, and the flux out due to their substitution (and
+the decline in mean population fitness). When $R=0$, selection is so strong it
 cannot fix, and the equilibrium fitness variation is due entirely to young rare
@@
 \eqref{eq:main_eqns} to compute the equilibrium additive genic fitness variance
-($V_a$) and ratchet rate ($R$) across grids of mutation rates and selection
-coefficients. This is done for each pre-specified segment in the genome that
-may be under negative selection (e.g. coding sequences or UTRs). We call the
-set of predicted reductions across these grids the B' maps (to distinguish them
-from McVicker's B maps); these can be used then to find the equilibrium
-reduction factor $B(x)$ for any genomic position $x$.
+($\widetilde{V}_a$) and ratchet rate ($\widetilde{R}$) across grids of mutation
+rates and selection coefficients. This is done for each pre-specified segment
+in the genome that may be under negative selection (e.g. coding sequences or
+UTRs). We call the set of predicted reductions across these grids the B' maps
+(to distinguish them from McVicker's B maps); these can be used then to find
+the equilibrium reduction factor $B(x)$ for any genomic position $x$.
 
 We validated our predicted B' reduction maps with realistic chromosome-scale
-forward simulations of negative selection in conserved regions. Averaging over
-$100$ replicates, we estimated the simulation empirical reduction map and
-compared this to the predicted B and B' maps for the same parameters. We find
-that our B' maps and the classic BGS theory B maps closely match simulations
-when selection is strong (top row of Figure \ref{fig:figure-1}E). This provides
-the first realistic forward-time chromosome-scale simulation confirmation of
-classic BGS theory. However, we find slight discrepancies in regions with low
-recombination (Figure \ref{fig:figure-1}E). Second, we find our theory is
-vastly more accurate than the classic B maps when selection is very weak ($2N_e
-s \ll 1$; bottom row of Figure \ref{fig:figure-1}E), which is expected since
-BGS theory does not work in this domain. Across all mutation and selection
-parameters simulated, the relative error of the classic B maps is 14.6\%
-whereas the relative error in the new B' maps is 5\%. Additionally, we find
-that the mean squared error between simulations and B' maps is close to the
-theoretic lower bound set by the coalescence variance
+forward simulations of negative selection in conserved regions. We find that
+our B' maps and the classic BGS theory B maps closely match simulations when
+selection is strong (top row of Figure \ref{fig:figure-1}E), apart from slight
+discrepancies in low recombination regions (Figure \ref{fig:figure-1}E).
+Second, we find our theory is vastly more accurate than the classic BGS when
+selection is very weak ($2N_e s \ll 1$; bottom row of Figure
+\ref{fig:figure-1}E). Across all mutation and selection parameters simulated,
+the relative error of the classic B maps is 14.6\% whereas the relative error
+in the new B' maps is 5\%. Nearly all of this error is in the drift-barrier
+domain ($2N_e s \approx 1$ domain); for strong and weak selection, the mean
+squared error between simulations and B' maps is close to the theoretic lower
+bound of $\V(B) \approx \nicefrac{2}{9n}$ set by the coalescence variance
 \parencite{Tajima1983-gu}.
 
-The discrepancy between our theoretic B' maps and chromosome-wide simulations
-that leads to the 5\% error occurs entirely around the drift-barrier domain of
-$2Ns \approx 1$ (Figure \ref{fig:figure-1}D). The higher error in this domain
-occurs across all mutation rates. We hypothesized this error may be due to
-selective interference between segments that is not taken into account when we
-numerically solve Equations \eqref{eq:main_eqns} for each segment
-independently. In particular, we use a fixed drift-effective $N=1000$
-corresponding to the number of diploids in the simulations rather than $B(x)
-N$, which is an approximation to the draft-effective population size
-experienced by a selected segment at position $x$. To test this, we implemented
-a ``locally-rescaled" version of the B' maps, which numerically solves
-Equations \eqref{eq:main_eqns} for each segment at position $x$ taking into
-account the reduction experienced due to selection at other segments by setting
-$B(x)N$ as the population size. We find the locally-rescaled B' maps reduce the
-relative error from 5\% to 0.4\% and mean squared error (Figure
-\ref{fig:figure-1}D, dashed colored lines), but does not entirely eliminate the
-error in the $2Ns \approx 1$ domain. Overall, while we find evidence that
-local-rescaling slightly reduces error in the $2Ns \approx 1$ domain, we do not
-use it during in our genome-wide inference method. This is because to avoid
-circularity and possible overfitting, our method would need to solve the
-equilibria equations for all segments simultaneously; unfortunately, this is
-currently computational unfeasible.
-
-Finally, we hypothesized that the remaining error is because of the build up of
-negative linkage disequilibria between selected sites due to Hill--Robertson
-interference \parencite{Hill1966-kd,McVean2000-bt,Comeron2007-wq}. To assess
-this possibility, we calculated the sum of all linkage disequilibria in our
-chromosome-wide simulations. We find negative linkage disequilibria build ups
-around $2Ns \approx 1$ (Figure \ref{fig:figure-1}D, top row) and is stronger
-when mutation rates are higher. As $Ns \to 0$, the variance in LD inflates as
-expected \parencite{Ohta1969-ae,Hill1968-ue}. Overall, the build up of negative
-LD is consistent our view that the equilibrium fitness variance modeled by the
-SC16 theory is the additive \emph{genic} fitness variance, which differs from
-the additive genetic variance by the sum of linkage disequilibria between
-selected sites (i.e. $\delta_{LD} = s^2 \sum_{i\ne j} D_{i,j}$ where $D_{i,j}$
-is the LD between sites $i$ and $j$). However, according to theory, the
-reductions in diversity should be determined by levels of additive genetic
+We hypothesized this error in the drift-barrier domain may be due to selective
+interference between segments that is not taken into account when we
+numerically solve Equations \eqref{eq:main_eqns} independently for each
+segment. In particular, we use a fixed drift-effective $N=1,000$ corresponding
+to the number of diploids in the simulations rather than $B(x) N$, which is an
+approximation to the draft-effective population size experienced by a selected
+segment at position $x$. To test this, we implemented a ``locally-rescaled"
+version of the B' maps, which uses $B(x)N$ as the population size. We find the
+locally-rescaled B' maps reduce the relative error from 5\% to 0.4\% and mean
+squared error (Figure \ref{fig:figure-1}D, dashed colored lines), but does not
+entirely eliminate the error in the $2Ns \approx 1$ domain.
+
+Finally, we hypothesized that the remaining error could be due to the build up
+of negative linkage disequilibria between selected sites due to Hill--Robertson
+interference \parencite{Hill1966-kd,McVean2000-bt,Comeron2007-wq}. By
+calculating the sum of all linkage disequilibria pairs in our chromosome-wide
+simulations, we find such a build up of negative LD around $2Ns \approx 1$
+(Figure \ref{fig:figure-1}D, top row). As $Ns \to 0$, the variance in LD
+inflates as expected \parencite{Ohta1969-ae,Hill1968-ue}. Overall, the build up
+of negative LD is consistent our view that the equilibrium fitness variance
+modeled by the SC16 theory is the additive \emph{genic} fitness variance, which
+differs from the additive genetic variance by the sum of linkage disequilibria
+between selected sites (i.e. $\delta_{LD} = s^2 \sum_{i\ne j} D_{i,j}$ where
+$D_{i,j}$ is the LD between sites $i$ and $j$). However, according to theory,
+the reductions in diversity should be determined by levels of additive genetic
 fitness variance that include the contribution of LD (Supplementary Materials
@@
     \includegraphics[width=0.8\textwidth]{figures/figure_2.pdf}
+
     \caption{Comparison of parameter estimates using classic BGS theory (green
@@
         by the B' method than classic BGS, but overall show slight biases.
-        Overall, classic BGS methods fail as expected when full-coverage tracks
+        Overall, classic BGS methods break down as expected when full-coverage tracks
     are used, since it cannot accommodate weak selection and neutrality. }
+
   \label{fig:figure-2}
@@
 driven by unstructured coalescence noise along the genome and the predicted
-reduction map $B(x)$ does not fit the data well. However, we note that under
-very strong selection ($s=0.05$), $R^2$ is reduced. This is like due to very
-strong selection having less localized effects and impacting overall
-genome-wide diversity \parencite{Santiago1995-hx,Robertson1961-ho}.
+reduction map $B(x)$ does not fit the data well. Under very strong selection
+($s=0.05$), $R^2$ is reduced; this is likely due to very strong selection
+having less localized effects and impacting overall genome-wide diversity
+\parencite{Santiago1995-hx,Robertson1961-ho}.
 
@@
 due to the sensitivity of mutations in this region to selective interference
-(discussed in Section \ref{sec:chrom-sims}). Because we only simulated fixed
-selection coefficients and five chromosomes, we only assessed the accuracy of
-average selection coefficients and not the DFE probabilities.
+(these do not use local-rescaling). Because we only simulated fixed selection
+coefficients and five chromosomes, we only assessed the accuracy of average
+selection coefficients and not the DFE probabilities.
 
@@
 based on their agreement with pedigree-based rate estimates may not be
-appropriate. When BGS is not functioning, either due to weak selection or a low
+appropriate. When BGS is not occurring, either due to weak selection or a low
 rate of deleterious mutations (Figure \ref{fig:figure-2}, right column), all
@@
 that this is an unlikely region of parameter space and this issue can be
-readily diagnosed from the low $R^2$ values.
-
-Finally, we tested the sensitivity of our method to violations of two
-assumptions---constant population size and recessivity of selective
-effects---through simulations. Using an expansion rate and timing as estimated
-for Europeans by \textcite{Gutenkunst2009-pg}, we found that a population
-expansion had minimal impact on parameter estimates (Supplementary Materials
-Section \ref{supp:sim-assum}). We did not test the influence of population
-bottlenecks, since parameter estimates of out-of-Africa bottlenecked
-populations (CHB and CEU) did not differ much from YRI estimates (see below).
-However, these simulations did find that selection coefficient estimates were
-downwardly biased when deleterious mutations were recessive and weak.
-Recessivity did not bias selection coefficient estimates when selection was
-strong; this is expected since strongly deleterious mutations are predominantly
-are sufficiently rare that they are not often found in a homozygote. Overall,
-the sensitivity to recessivity is likely due to the ratchet rate being
-overestimated when selection is weak and recessive mutations drift up to higher
-frequency and experience stronger than expected selection in homozygotes.
+readily diagnosed from the low $R^2$ values. Finally, we find in additional
+tests that our estimates are robust to demographic expansions but sensitive
+when mutuations have recessive effects, since our model assumes additive
+effects (see Supplementary Materials Section \ref{supp:sim-assum}). We did not
+test the influence of population bottlenecks, since parameter estimates of
+out-of-Africa bottlenecked populations (CHB and CEU) did not differ much from
+YRI estimates (see below).
 
@@
 
-Our method estimates the distribution of fitness effects (DFE) for each feature
+Our method estimates a distribution of fitness effects (DFE) for each feature
 class. While CADD-based models only have a single conserved feature class (e.g.
@@
 
-Our PhastCons Priority model fits marginally better than the CADD 6\% model
-consistently across all three populations (Supplementary Materials Table
-\ref{supp:tbl-r2}). This is likely because the PhastCons Priority model has
-more DFE parameters, and thus can fit finer-scale differences among the four
-feature types. By contrast, the CADD 6\% track must fit patterns of diversity
-to conserved sites under a single DFE. The CADD 8\% model fit slightly worse
-than the CADD 6\% across all populations, and the Feature Priority model had
-the poorest-fit.
+% Our PhastCons Priority model fits marginally better than the CADD 6\% model
+% consistently across all three populations (Supplementary Materials Table
+% \ref{supp:tbl-r2}). This is likely because the PhastCons Priority model has
+% more DFE parameters, and thus can fit finer-scale differences among the four
+% feature types. By contrast, the CADD 6\% track must fit patterns of diversity
+% to conserved sites under a single DFE. The CADD 8\% model fit slightly worse
+% than the CADD 6\% across all populations, and the Feature Priority model had
+% the poorest-fit.
 
@@
 By contrast, full annotation models fit poorly under classic BGS theory, and
-lead to unreasonable parameter estimates (discussed in Section XXX).
-Additionally, when sparse annotation models contain genomic features that are
-likely under weak constraint (such as introns and UTRs), models fit worse under
-classic BGS theory than our B' method (Figure \ref{fig:figure-2}A). However,
-among the CADD annotation models, the goodness-of-fit is nearly identical
-between B' and classic BGS methods. This behavior is what we would expect given
-that the CADD models contain only the most pathogenic sites, which are \emph{a
-priori} very likely under the strong selection domain under which B' and
-classic BGS theory agree.
+lead to unreasonable parameter estimates. Additionally, when sparse annotation
+models contain genomic features that are likely under weak constraint (such as
+introns and UTRs), models fit worse under classic BGS theory than our B' method
+(Figure \ref{fig:figure-2}A). However, among the CADD annotation models, the
+goodness-of-fit is nearly identical between B' and classic BGS methods. This
+behavior is what we would expect given that the CADD models contain only the
+most pathogenic sites, which are \emph{a priori} very likely under the strong
+selection domain under which B' and classic BGS theory agree.
 
@@
 = 0.59\%$, and $\bar{s}= 0.77\%$ for CEU, CHB, and YRI respectively for
-PhastCons regions. Our DFE estimates for the Feature Priority model are
-qualitatively consistent with the U-shaped DFEs found for amino acids through
-Poisson Random Field method \parencite{Boyko2008-tj}.
+PhastCons regions. Our DFE estimates for CDS under the Feature Priority model
+are qualitatively consistent with the U-shaped DFEs found for amino acids
+through Poisson Random Field method \parencite{Boyko2008-tj}.
 
@@
 
-While we have previously shown that out-sample predicted diversity matches
-observed diversity (Section \ref{sec:annotation}), we also investigated whether
-there were regions of the genome that were particularly poor-fitting. Comparing
-predicted against observed diversity along chromosomes (see Supplementary
-Materials Section \ref{supp:chrom-fit}) we find a close correspondence
-consistent with the high $R_\text{LOCO}^2$. However, we find a few large (tens
-of megabases) regions with systematically poorer fit (Supplementary Materials
-Figure \ref{supp:spatial-residuals}). In Figure \ref{fig:figure-5}A we see one
-such region on the short arm of chromosome 2, from 30 Mbp to 60Mbp.
-Interestingly, predicted diversity closely follows the peaks and troughs of
-this region. We note that a small region within this stretch had been found by
-a genome-wide scan for associative overdominance \parencite{Gilbert2020-aw}.
+Comparing predicted against observed diversity along chromosomes (see
+Supplementary Materials Section \ref{supp:chrom-fit}) we find a close
+correspondence consistent with the high $R_\text{LOCO}^2$. However, we find a
+few large (tens of megabases) regions with systematically poorer fit
+(Supplementary Materials Figure \ref{supp:spatial-residuals}). In Figure
+\ref{fig:figure-5}A we see one such region on the short arm of chromosome 2,
+from 30 Mbp to 60Mbp. Interestingly, predicted diversity closely follows the
+peaks and troughs of this region. We note that a small region within this
+stretch had been found by a genome-wide scan for associative overdominance
+\parencite{Gilbert2020-aw}.
 
@@
 
-Since our B' method is based upon quantitative genetic models of linked
-selection that explicitly incorporates the substitution process, fitting our
-model to genomic diversity data leads to additional predictions. In particular,
-our model predicts (1) the substitution rates per segment, (2) the fitness
-variance per segment, and (3) the total number of strongly deleterious
-mutations per diploid genome per generation. Given predicted substitution rates
-for each feature can be compared to estimated sequence divergence between
-species, this offers a qualitative check our model is producing sensible
-results.
-
-To compare our models' predicted substitution rates against observed
-divergence, we estimated sequence divergence on the human lineage using a
-multiple alignment of five primates for each feature in our feature-based
-models (Methods \ref{sec:methods-sub}). We compared these to the predicted
+Since our B' method relies on theory that jointly predicts the deleterious
+substitution rate, we can also check whether our model produces substitution
+rate predictions in agreement with observed divergence levels across features.
+We estimated sequence divergence on the human lineage using a multiple
+alignment of five primates for each feature in our feature-based models
+(Methods Section \nameref{sec:methods-sub}). We compared these to the predicted
 substitution rates per feature, averaging over all segments in the genome.
@@
 allele, and this should be considered in how the SC16 equations are solved.
-There is an important caveat about this analysis.
 
@@
 limit of our optimization boundary of $\mu = 8 \times 10^{-8}$ across all
-populations in or PhastCons Priority model (Supplementary Material XXX). When
-we re-ran optimization with an upper limit of $\mu = 1 \times 10^{-7}$, the
-mutation rate was again estimated to this (implausibly large) upper limit, with
-no considerable changes to model fit or DFE estimates. Second, all of our
-leave-one-chromosome-out $R_\text{LOCO}^2$ are about one percentage point
-higher than the unrescaled model. This improved fit could be an artifact of the
-circularity, so it is unimportant. Third, the DFE estimates for PhastCons
-regions is now U-shaped (Supplementary Materials Figure
+populations in or PhastCons Priority model (Supplementary Material
+\ref{supp:locally-rescaled-fits}). When we re-ran optimization with an upper
+limit of $\mu = 1 \times 10^{-7}$, the mutation rate was again estimated to
+this (implausibly large) upper limit, with no considerable changes to model fit
+or DFE estimates. Second, all of our leave-one-chromosome-out $R_\text{LOCO}^2$
+are about one percentage point higher than the unrescaled model. Third, the DFE
+estimates for PhastCons regions is now U-shaped (Supplementary Materials Figure
 \ref{suppfig:rescale-dfe}), with $75-77\%$ of mass being placed on a weakly
@@
 possibility of some circularity, since the $B(x)$ at this position includes the
-contribution of the focal segment that is being rescaled. However, two
-observations indicate this would matter little. First, the results do not
-change based on whether the $B(x)$ is at averaged at the 1 kbp level or at the
-megabase scale; for the latter, a single segment makes little contribution to
-the average. Second, we investigated the extent to which local-rescaling
-differed from the B' maps across selection parameters. We find minor
-differences between the locally-rescaled and standard B' maps in the
-nearly-neutral domain ($0.2 \le 2N_e s \le 2$); the maps were identical under
-strong selection ($2N_e s= 20$) as expected (Supplementary Materials Figure
-\ref{suppfig:rescale-b}). Moreover, the correlations between the standard and
-local-rescaled B' maps across the genome are high (100\% for $2N_e s = 20$,
-96.5\% for $2N_e s = 2$, and 60.21\% for $2N_e s = 0.2$). The overall realized
-effect of local-rescaling is to just alter how deep the ``U" is in the
-relationship between the reduction factor $B$ and the selection coefficient
-(Supplementary Materials Figure \ref{suppfig:rescale-effects}).
+contribution of the focal segment that is being rescaled. Ideally, one would
+exclude this segment's contribution to $B(x)$; however, this is computationally
+unfeasible. However, two observations indicate our findings here are robust
+despite this limitation. First, the results do not change based on whether the
+$B(x)$ is averaged at the 1 kbp level or at the megabase scale; for the latter,
+a single segment makes little contribution to the average. Second, we
+investigated the extent to which local-rescaling differed from the B' maps
+across selection parameters. We find minor differences between the
+locally-rescaled and standard B' maps in the nearly-neutral domain ($0.2 \le
+2N_e s \le 2$); the maps were identical under strong selection ($2N_e s= 20$)
+as expected (Supplementary Materials Figure \ref{suppfig:rescale-b}). Moreover,
+the correlations between the standard and local-rescaled B' maps across the
+genome are high (100\% for $2N_e s = 20$, 96.5\% for $2N_e s = 2$, and 60.21\%
+for $2N_e s = 0.2$). The overall realized effect of local-rescaling is to just
+alter how deep the ``U" is in the relationship between the reduction factor $B$
+and the selection coefficient (Supplementary Materials Figure
+\ref{suppfig:rescale-effects}).
 
 Overall, this suggests that models that fit the linked selection are worryingly
-sensitive the theoretic B' values in the $2Ns \le 1$ domain. Local-rescaling
-turns out to only slightly alter the B' maps, yet significantly modifies the
-DFE estimates. This brings the deleterious substitution rate in agreement with
-observations (since this is predicted with a fixed $\mu = 1.5 \times 10^{-8}$;
-however, the maximum likelihood estimate of mutation is implausibly high.
+sensitive the theoretic B' values in the $2Ns \le 1$ domain. The fact that
+predicted diversity differs little between standard and locally-rescaled B'
+methods indicates there may not be enough information in the linked selection
+signal alone to differentiate when interference is occurring or the causes of
+fitness variance along the genome. Moreover, local-rescaling turns out to only
+slightly alter the B' maps, yet significantly modifies the DFE estimates. This
+brings the deleterious substitution rate in agreement with observations (since
+this is predicted with a fixed $\mu = 1.5 \times 10^{-8}$; however, the maximum
+likelihood estimate of mutation is implausibly high.
 

### 2023-07-24 manuscript/supplementary.tex
ms updates from italy

@@
 
+\clearpage
 \subsection{Locally-Rescaled Fit Results}
@@
 
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[]{figures/supplementary/b-rescaling-effect.pdf}
+
+  \caption{Genome-wide average $B(x)$ values across selection coefficients for
+  $\mu = 1.58 \times 10^{-8}$, for both the standard (blue) and
+locally-rescaled B' (orange) maps. This indicates that locally-rescaling the B'
+maps only in practice changes how deep the ``U" is.}
+
+    \label{suppfig:rescale-effects}
+\end{figure}
+
+\clearpage
+\subsubsection{Locally-Rescaled Model Fit Results}
+\label{supp:locally-rescaled-fits}
+
+Below are the model fit results for locally-rescaled model runs, with the
+upper mutation rate boundary set to $\mu = 8 \times 10^{-8}$.
+
+\subsubsection*{CHB}
+
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+SimplexModel (interpolated w): 6 x 7 x 3
+    w grid: [1.000e-11 6.310e-11 3.981e-10 2.512e-09 1.585e-08 1.000e-07] (before interpolation)
+    t grid: [1.e-08 1.e-07 1.e-06 1.e-05 1.e-04 1.e-03 1.e-02]
+    window size: 1.0 Mbp
+
+  ML estimates: (whole genome)
+  standard error method: None
+  negative log-likelihood: 61820142829.01206
+  number of successful starts: 10000 (100.0% total)
+  pi_0 = 0.00106289
+  pi  = 0.000752601
+  mut. rate  = 8e-08 
+  Ne = 26,572 (if μ=1e-8), Ne = 13,286 (if μ=2e-8)
+  R² = 60.8648% (in-sample)  59.1654% (out-sample)
+  W = 
+            CDS    gene    phastcons
+  ------  -----  ------  -----------
+  1e-08       1   0.991        0
+  1e-07       0   0            0
+  1e-06       0   0            0
+  1e-05       0   0            0.767
+  0.0001      0   0            0.03
+  0.001       0   0            0.069
+  0.01        0   0.008        0.134
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CEU}
+
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+SimplexModel (interpolated w): 6 x 7 x 3
+    w grid: [1.000e-11 6.310e-11 3.981e-10 2.512e-09 1.585e-08 1.000e-07] (before interpolation)
+    t grid: [1.e-08 1.e-07 1.e-06 1.e-05 1.e-04 1.e-03 1.e-02]
+    window size: 1.0 Mbp
+
+  ML estimates: (whole genome)
+  standard error method: None
+  negative log-likelihood: 250647896892.40094
+  number of successful starts: 10000 (100.0% total)
+  pi_0 = 0.001479
+  pi  = 0.00106447
+  mut. rate  = 8e-08 
+  Ne = 36,974 (if μ=1e-8), Ne = 18,487 (if μ=2e-8)
+  R² = 68.7046% (in-sample)  67.8854% (out-sample)
+  W = 
+            CDS    gene    phastcons
+  ------  -----  ------  -----------
+  1e-08       1   0.994        0
+  1e-07       0   0            0
+  1e-06       0   0            0
+  1e-05       0   0            0.751
+  0.0001      0   0            0.039
+  0.001       0   0            0.028
+  0.01        0   0.006        0.183
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+SimplexModel (interpolated w): 6 x 7 x 3
+    w grid: [1.000e-11 6.310e-11 3.981e-10 2.512e-09 1.585e-08 1.000e-07] (before interpolation)
+    t grid: [1.e-08 1.e-07 1.e-06 1.e-05 1.e-04 1.e-03 1.e-02]
+    window size: 1.0 Mbp
+
+  ML estimates: (whole genome)
+  standard error method: None
+  negative log-likelihood: 197824525643.2897
+  number of successful starts: 10000 (100.0% total)
+  pi_0 = 0.00111021
+  pi  = 0.000801495
+  mut. rate  = 8e-08 
+  Ne = 27,755 (if μ=1e-8), Ne = 13,877 (if μ=2e-8)
+  R² = 62.5591% (in-sample)  61.367% (out-sample)
+  W = 
+            CDS    gene    phastcons
+  ------  -----  ------  -----------
+  1e-08       1   0.989        0
+  1e-07       0   0.003        0
+  1e-06       0   0            0
+  1e-05       0   0            0.774
+  0.0001      0   0            0.025
+  0.001       0   0            0.067
+  0.01        0   0.008        0.134
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\clearpage
 \subsection{Strong Selection Grids}

### 2023-07-25 manuscript/manuscript.tex
ms updates from italy, this time with covid

@@
 %\title{An Evolutionary Quantitative Genetic Model of the Negative Selection Process in Humans}
-\title{A Quantitative Genetic Linked Selection Model of the Negative Selection Process in Humans}
+\title{A Quantitative Genetic Linked Selection Model of Linked Selection in Humans}
 
@@
 
-\subsection*{Predicted Substitution Rate and Additive Fitness Variance}
+\subsection*{Predicted Substitution Rates}
 
@@
         Additionally, we show the predicted PhastCons region divergences when
-    local-rescaling is applied (we omit other locally-rescaled predictions
+    local-rescaling is applied (blue; we omit other locally-rescaled predictions
 since these to not differ substantially).}
@@
 There are five important but tentative results to draw from this analysis.
-First, the estimated mutation rate is implausibly higher, reaching the upper
-limit of our optimization boundary of $\mu = 8 \times 10^{-8}$ across all
-populations in or PhastCons Priority model (Supplementary Material
-\ref{supp:locally-rescaled-fits}). When we re-ran optimization with an upper
-limit of $\mu = 1 \times 10^{-7}$, the mutation rate was again estimated to
-this (implausibly large) upper limit, with no considerable changes to model fit
-or DFE estimates. Second, all of our leave-one-chromosome-out $R_\text{LOCO}^2$
-are about one percentage point higher than the unrescaled model. Third, the DFE
-estimates for PhastCons regions is now U-shaped (Supplementary Materials Figure
-\ref{suppfig:rescale-dfe}), with $75-77\%$ of mass being placed on a weakly
+First, estimated mutation rates are in general higher. Under the CADD 6\%
+model, they are $\widehat{\mu} = 6-7.3 \times 10^{-8}$ across populations; for
+the PhastCons Priority model, they reach the upper limit of our optimization
+boundary of $\mu = 8 \times 10^{-8}$ (Supplementary Material
+\ref{supp:locally-rescaled-fits}). Second, all of our leave-one-chromosome-out
+$R_\text{LOCO}^2$ are about one percentage point higher than the unrescaled
+model. Third, the DFE estimates for both CADD 6\% and PhastCons regions in the
+PhastCons Priority model is now U-shaped (Supplementary Materials Figure
+\ref{suppfig:rescale-dfe}), with $70-77\%$ of mass being placed on a weakly
 deleterious class, $s=10{^-5}$. Interestingly, this is the first of all of our
 models where such an appreciable mass has been placed on a midpoint in our
-grid; in all other cases, non-strongly estimates were neutral ($s=10^{-8}$).
-Fourth, predicted pairwise diversity is nearly identical to our original,
-non-locally-rescaled fit (see Supplementary Materials Figure
+selection coefficient grid; in all other cases, non-strongly estimates were
+neutral ($s=10^{-8}$). Fourth, predicted pairwise diversity is nearly identical
+to our original, non-locally-rescaled fit (see Supplementary Materials Figure
 \ref{suppfig:rescale}). Finally, since local-rescaling increases the DFE mass
@@
 
-Finally, we one important caveat about this analysis. Since local-rescaling is
-done using the first round of maximum likelihood estimates, there is a
+Finally, we note an important caveat about this analysis. Since local-rescaling
+is done using the first round of maximum likelihood estimates, there is a
 possibility of some circularity, since the $B(x)$ at this position includes the
@@
 exclude this segment's contribution to $B(x)$; however, this is computationally
-unfeasible. However, two observations indicate our findings here are robust
-despite this limitation. First, the results do not change based on whether the
-$B(x)$ is averaged at the 1 kbp level or at the megabase scale; for the latter,
-a single segment makes little contribution to the average. Second, we
-investigated the extent to which local-rescaling differed from the B' maps
-across selection parameters. We find minor differences between the
-locally-rescaled and standard B' maps in the nearly-neutral domain ($0.2 \le
-2N_e s \le 2$); the maps were identical under strong selection ($2N_e s= 20$)
-as expected (Supplementary Materials Figure \ref{suppfig:rescale-b}). Moreover,
-the correlations between the standard and local-rescaled B' maps across the
-genome are high (100\% for $2N_e s = 20$, 96.5\% for $2N_e s = 2$, and 60.21\%
-for $2N_e s = 0.2$). The overall realized effect of local-rescaling is to just
+unfeasible. However, two observations indicate our findings here are relatively
+robust despite this limitation. First, the results do not change based on
+whether the $B(x)$ is averaged at the 1 kbp level or at the megabase scale; for
+the latter, a single segment makes little contribution to the average. Second,
+we investigated the extent to which local-rescaling modified the B' maps across
+selection parameters. We find minor differences between the locally-rescaled
+and standard B' maps in the nearly-neutral domain ($0.2 \le 2N_e s \le 2$); the
+maps were identical under strong selection ($2N_e s= 20$) as expected
+(Supplementary Materials Figure \ref{suppfig:rescale-b}). Moreover, the
+correlations between the standard and local-rescaled B' maps across the genome
+are high (100\% for $2N_e s = 20$, 96.5\% for $2N_e s = 2$, and 60.21\% for
+$2N_e s = 0.2$). The overall realized effect of local-rescaling is to just
 alter how deep the ``U" is in the relationship between the reduction factor $B$
@@
 Overall, this suggests that models that fit the linked selection are worryingly
-sensitive the theoretic B' values in the $2Ns \le 1$ domain. The fact that
+sensitive to the theoretic B' values in the $2Ns \le 1$ domain. The fact that
 predicted diversity differs little between standard and locally-rescaled B'
@@
 this is predicted with a fixed $\mu = 1.5 \times 10^{-8}$; however, the maximum
-likelihood estimate of mutation is implausibly high.
+likelihood estimate of mutation is implausibly high. This suggests either the
+local-rescaling approximation to interference is not suitable (though
+chromosome-wide simulations show it results in lower error), or there is
+another selection process contributing to total fitness variance.
 
@@
 Since the SC16 theory calculates fitness variance by jointly predicting the
-deleterious substitution rate, this provides another route to verify our main
-findings. We find that predicted substitution rates for highly-conserved
+deleterious substitution rate, this provides another avenue to verify or refute
+our models. We find that predicted substitution rates for highly-conserved
 PhastCons features are in strong disagreement with observed substitution rates
@@
 approach, in practice it is altering the B' maps only slightly in the domain of
-$2N_e s \approx 1$ (deepening the bottom of the ``U"). Even if the local-rescaling
-approximation to selective interference is incorrect, these results suggest
-parameter estimates are extremely sensitivity to how accurate the B' maps are
-in this domain.
+$2N_e s \approx 1$ (deepening the bottom of the ``U"). Even if the
+local-rescaling approximation to selective interference is incorrect (though
+simulations indicate it does reduce error), these results suggest parameter
+estimates are extremely sensitivity to how accurate the B' maps are in this
+domain. Moreover, we find that predicted diversity differs little across DFE
+and mutation rate estimates, suggesting there may be limited information in the
+pairwise diversity linked selection signal to differentiate models. While
+nucleotide diversity is a less informative statistic than the full site
+frequency spectrum, it is much easier to derive theoretic models of how linked
+selection alters expected pairwise coalescence times than to describe the
+entire genealogy. While one could imagine using more informative SFS-based
+models to infer selection parameters, accounting for selective interference in
+these models would be difficult.
 
 Unfortunately, this work opens up more questions about the nature of selection
-in humans that it answers. The implausibly high estimate of mutation rates
-could suggest that negative selection is not the only source of fitness
-variance generating the genome-wide linked selection signal in humans. Positive
-selection on polygenic variation along the genome could be another source of
-fitness variance contributing to the weak selection signal. Alternatively,
-negative selection could be the main source of fitness variance, but the
-complexity of selective interference may be poorly approximated by the
-local-rescaling approach. Another possibility is that our assumption throughout
-of additive effects may lead to biases given deleterious mutations likely
-recessive effects. Overall, understanding how the sources of human fitness
-variance shape the striking patterns of diversity along the human genome still
-remains an open riddle.
+in humans that it answers. The high estimate of mutation rates could suggest
+that negative selection is not the only source of fitness variance generating
+the genome-wide linked selection signal in humans. Positive selection on
+polygenic variation along the genome could be another source of fitness
+variance contributing to the weak selection signal. Alternatively, negative
+selection could be the main source of fitness variance, but the complexity of
+selective interference may be poorly approximated by the local-rescaling
+approach. Another possibility is that our assumption throughout of additive
+effects may lead to biases given deleterious mutations likely recessive
+effects. As we find evidence of strong selection against loss-of-function in
+our model residuals, it is also possible that the bulk of fitness variance
+\emph{is} due to purifying selection, but our model is unable to account for
+local strong effects since the DFE is estimated per feature. Overall,
+understanding how the sources of human fitness variance shape the striking
+patterns of diversity along the human genome still remains an open riddle.
+
+\subsection*{Acknowledgments}
+
+We would like to thank Nate Pope and Rasmus Nielsen for helpful advice, and
+Martin Kircher for providing special CADD tracks. This research was supported
+by NIH awards R35GM148253 and R01HG010774 to ADK
 
@@
 where,
-
+%
 \begin{align}
@@
 with megabase windows, since previous work has found the strongest selection
-signal at this scale (we confirm this with one XXX fit at the $10^5$bp scale).
+signal at this scale (we confirm this with one CADD 6\% fit at the $100$ kbp
+scale).
 
@@
 across 10,000 random starts (see Supplementary Materials Section
-\ref{supp:optima}). We assessed model fit using out-sample predictive error,
+\ref{supp:optim}). We assessed model fit using out-sample predictive error,
 calculated by leaving out a whole chromosome during fitting and predicting its
@@
 
+\subsection*{Data Availability}
+
+All code from \texttt{bgspy} and our Jupyter Lab \parencite{Kluyver_undated-ol}
+notebooks for analysis are available on GitHub
+(\url{https://github.com/vsbuffalo/bprime}). The main model fits are available
+as Python Pickle objects on FigShare (XXX).
+
 \printbibliography

### 2023-07-25 manuscript/supplementary.tex
ms updates from italy, this time with covid

@@
 sequences from Ensembl Version 107
-\parencite[\path{Homo_sapiens.GRCh38.107.chr.gff3.gz}]{Cunningham2022-vk} and
-PhastCons regions
-\parencite[\path{phastConsElements100way.bed.gz}]{Siepel2005-wh}. Then, we
-remove 200 basepairs of ``slop" around each of these regions (i.e.
-\texttt{bedtools slop}). We show in Supplementary Figure
-\ref{suppfig:neutral-tracks} that average diversity (we only consider YRI
-samples for this analysis) is fairly insensitive to the slop width, but
-increasing slop widths around putatively conserved segments does lead to higher
-diversity estimates. This is expected, since increasing the slop width
-considers a smaller subset of sites, further away from putatively conserved
-regions. For our analyses we choose the 200bp slop version, which covers
-63.66\% of the autosomal reference sequence and consists of 1.26 million
-contiguous genomic regions.
+\path{Homo_sapiens.GRCh38.107.chr.gff3.gz}]{Cunningham2022-vk} and PhastCons
+regions \path{phastConsElements100way.bed.gz}]{Siepel2005-wh}. Then, we remove
+200 basepairs of ``slop" around each of these regions (i.e. \texttt{bedtools
+slop}). We show in Supplementary Figure \ref{suppfig:neutral-tracks} that
+average diversity (we only consider YRI samples for this analysis) is fairly
+insensitive to the slop width, but increasing slop widths around putatively
+conserved segments does lead to higher diversity estimates. This is expected,
+since increasing the slop width considers a smaller subset of sites, further
+away from putatively conserved regions. For our analyses we choose the 200bp
+slop version, which covers 63.66\% of the autosomal reference sequence and
+consists of 1.26 million contiguous genomic regions.
 
@@
 \subsection{Locally-Rescaled Fit Results}
+\label{supp:locally-rescaled-fits}
 
@@
 \clearpage
-\subsubsection{Locally-Rescaled Model Fit Results}
-\label{supp:locally-rescaled-fits}
+
+\subsubsection{PhastCons Priority Locally-Rescaled Model Fit Results}
 
@@
   0.01        0   0.008        0.134
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection{CADD 6\% Locally-Rescaled Model Fit Results}
+
+\subsubsection*{CHB}
+
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+SimplexModel (interpolated w): 6 x 7 x 1
+  w grid: [1.000e-11 6.310e-11 3.981e-10 2.512e-09 1.585e-08 1.000e-07] (before interpolation)
+  t grid: [1.e-08 1.e-07 1.e-06 1.e-05 1.e-04 1.e-03 1.e-02]
+  window size: 1.0 Mbp
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 61820130669.36475
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00102026
+pi  = 0.000752601
+mut. rate  = 7.291e-08
+Ne = 25,506 (if μ=1e-8), Ne = 12,753 (if μ=2e-8)
+R² = 60.6057% (in-sample)
+W =
+          cadd6
+------  -------
+1e-08     0.009
+1e-07     0
+1e-06     0
+1e-05     0.749
+0.0001    0.014
+0.001     0.067
+0.01      0.161
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{CEU}
+
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+SimplexModel (interpolated w): 6 x 7 x 1
+  w grid: [1.000e-11 6.310e-11 3.981e-10 2.512e-09 1.585e-08 1.000e-07] (before interpolation)
+  t grid: [1.e-08 1.e-07 1.e-06 1.e-05 1.e-04 1.e-03 1.e-02]
+  window size: 1.0 Mbp
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 197825902519.15137
+number of successful starts: 10000 (100.0% total)
+pi_0 = 0.00107426
+pi0  = 0.000801495
+mut. rate  = 5.991e-08
+Ne = 26,856 (if μ=1e-8), Ne = 13,428 (if μ=2e-8)
+R² = 62.0668% (in-sample)
+W =
+          cadd6
+------  -------
+1e-08     0.001
+1e-07     0
+1e-06     0
+1e-05     0.71
+0.0001    0.027
+0.001     0.065
+0.01      0.197
+\end{verbatim}
+\end{footnotesize}\end{minipage}
+
+\subsubsection*{YRI}
+
+\begin{minipage}{\linewidth}\begin{footnotesize}
+\begin{verbatim}
+SimplexModel (interpolated w): 6 x 7 x 1
+  w grid: [1.000e-11 6.310e-11 3.981e-10 2.512e-09 1.585e-08 1.000e-07] (before interpolation)
+  t grid: [1.e-08 1.e-07 1.e-06 1.e-05 1.e-04 1.e-03 1.e-02]
+  window size: 1.0 Mbp
+
+ML estimates: (whole genome)
+standard error method: None
+negative log-likelihood: 250648005868.94275
+number of successful starts: 10000 (100.0% total)
+pi0 = 0.0014366
+pi  = 0.00106447
+mut. rate  = 6.024e-08
+Ne = 35,914 (if μ=1e-8), Ne = 17,957 (if μ=2e-8)
+R² = 68.6747% (in-sample)
+W =
+          cadd6
+------  -------
+1e-08     0
+1e-07     0
+1e-06     0
+1e-05     0.707
+0.0001    0.048
+0.001     0.012
+0.01      0.233
 \end{verbatim}

### 2023-07-25 manuscript/manuscript.tex
data dryad

@@
 
-We would like to thank Nate Pope and Rasmus Nielsen for helpful advice, and
-Martin Kircher for providing special CADD tracks. This research was supported
-by NIH awards R35GM148253 and R01HG010774 to ADK
+We would like to thank Graham McVicker, Priya Moorjani, Rasmus Nielsen, Nate
+Pope, Guy Sella, and Peter Sudmant for helpful advice, and Martin Kircher for
+providing modified CADD tracks. This research was supported by NIH awards
+R35GM148253 and R01HG010774 to ADK
 
@@
 (\url{https://github.com/vsbuffalo/bprime}). The main model fits are available
-as Python Pickle objects on FigShare (XXX).
+as Python Pickle objects on Data Dryad
+(\url{https://doi.org/10.5061/dryad.qnk98sfnv}).
 

### 2023-09-08 manuscript/manuscript.tex
new updated ms from overleaf

--- 2023-07-25 manuscript/manuscript.tex
@@
 \RequirePackage{xr}
-\externaldocument{supplementary}
+
+\newcommand*{\addFileDependency}[1]{
+\typeout{(#1)}
+
+\IfFileExists{#1}{}{\typeout{No file #1.}}
+}\makeatother
+
+\newcommand*{\myexternaldocument}[1]{%
+\externaldocument{#1}%
+\addFileDependency{#1.tex}%
+\addFileDependency{#1.aux}%
+}
+
+\myexternaldocument{supplementary}
 
@@
 % line numbers:
-\RequirePackage{lineno}
-%\modulolinenumbers[5]
-\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{black}}
+\RequirePackage[left, modulo]{lineno}
+\modulolinenumbers[2]
+
+% spacing
+\RequirePackage{setspace}
+\onehalfspacing
 
@@
 %\title{An Evolutionary Quantitative Genetic Model of the Negative Selection Process in Humans}
-\title{A Quantitative Genetic Linked Selection Model of Linked Selection in Humans}
-
-\author{Vince Buffalo and Andrew Kern}
+\title{A Quantitative Genetic Model of Background Selection in Humans}
+
+\author[1,2,$\dag$]{Vince Buffalo}
+\author[2]{Andrew D. Kern}
+\affil[1]{\small{Department of Integrative Biology, University of California, Berkeley}}
+\affil[2]{\small{Institute of Ecology and Evolution and Department of Biology, University of Oregon }}
+\affil[*]{Corresponding author email: \href{vsb@berkeley.edu}{vsb@berkeley.edu}}
 
@@
 
+\linenumbers
 \begin{abstract}
-
-    Across the human genome, there are large-scale fluctuations in genetic
-    diversity caused by the indirect effects of selection. This can be thought
-    of as a ``linked selection signal" that reflects the impact of selection
-    varying according to the placement of functional regions and recombination
-    rates along the genome. Previous work has shown that negative selection
-    against the steady influx of new deleterious mutations into conserved
-    regions is the predominant mode of selection in humans. However, the
-    theoretic model that underpins these results, classic Background Selection
-    theory, is only applicable when new mutations are so deleterious that they
-    cannot fix in the population. Here, we develop a statistical method based
-    on a quantitative genetics view of the linked selection, which models the
-    effects of weak draft created according to how polygenic additive fitness
-    variance is distributed along the genome. We use a recent model that
-    jointly predicts the equilibrium fitness variance and substitution rates
+    Across the human genome, there are large-scale fluctuations in genetic diversity caused by the indirect effects of selection. 
+    This ``linked selection signal" reflects the impact of selection according to the physical placement of functional regions and recombination
+    rates along chromosomes. Previous work has shown that purifying selection
+    acting against the steady influx of new deleterious mutations at functional portions
+    of the genome shapes patterns of genomic variation. To date, statistical
+    efforts to estimate purifying selection parameters from linked selection models have relied on
+    classic Background Selection theory, which is only applicable when new mutations are so deleterious that they cannot fix in the population. 
+    Here, we develop a statistical method based
+    on a quantitative genetics view of linked selection, that models how
+    polygenic additive fitness variance distributed along the genome increases 
+    the rate of stochastic allele frequency change. By 
+    jointly predicting the equilibrium fitness variance and substitution rate
     due to both strong and weakly deleterious mutations, we estimate the
-    distribution of fitness effects (DFE) and mutation rate across three human
-    populations. While our model can accommodate weaker selection, we initially
-    find evidence across three human populations of very strong selection
-    against deleterious mutations consistent with previous work. However, the
-    corollary predicted substitution rates for conserved regions are
-    unreasonably low, and in disagreement with observed rates. We hypothesize
-    this could be due to selected sites experiencing a further diminished
-    population size due to selective interference. When we account for this in
-    our method, we find evidence of weakly deleterious mutations in conserved
-    regions which brings the predicted substitution rate into agreement with
-    observations. However, these models lead to implausibly large mutation rate
-    estimates. Overall, while our model of the genomic linked selection signal
-    brings us a step towards uniting population and quantitative genetic
-    selection models with the substitution process, our work suggests
-    considerable uncertainty remains about the processes generating fitness
-    variance in humans.
-
+    distribution of fitness effects (DFE) and mutation rate across three geographically distinct human samples.
+While our model can accommodate weaker selection, we find evidence of strong selection operating similarly across all human samples.
+While our quantitative genetic model of linked selection fits better than previous models, substitution rates of the most constrained sites are poorly predicted.
+We find evidence that a model incorporating selective interference better predicts observed divergence, but overall our results ultimately suggest considerable uncertainty remains about the processes generating fitness variation in humans.
 \end{abstract}
@@
 
-The continual influx of new mutations into populations is the ultimate source
-of all adaptations, but the vast majority of mutations either do not affect
-fitness or are deleterious. Negative selection (i.e. purifying selection) works
-to eliminate these deleterious mutations from the population; thus we expect
-them to appear at low frequencies within populations
-\parencite{Haldane1927-ga}, and be less likely to fix between lineages. The
-best predictor of whether a new mutation will reduce fitness is if it occurs in
-a functional region of the genome, which are often conserved over deep
-phylogenetic timescales \parencite{Siepel2005-wh,Margulies2003-sn}. Since these
-conserved regions reflect the product of millions of years of evolutionary
-optimization, it follows that deleterious variants should be the overwhelming
-majority of segregating variation in these regions. Furthermore, rare variation
-in these regions is responsible for a significant portion of phenotypic
-variation and disease in humans
-\parencite{Zeng2018-ci,Lek2016-cb,Karczewski2020-ky,Tennessen2012-ge}.
-
-Selection on both beneficial and deleterious variants perturbs the allele
-frequencies of neighboring linked sites, a phenomenon known as linked selection
-\parencite{Nordborg1996-nq,Maynard_Smith1974-zr,Barton1998-aj,Charlesworth1993-gb}.
+The continual influx of new mutations into populations is the ultimate source of all adaptations, but the vast majority of mutations either do not affect fitness or are deleterious.
+Natural selection works to eliminate these deleterious mutations from the population, thus we expect them to appear at low frequencies within populations \parencite{Haldane1927-ga}, and be less likely to fix between lineages.
+Conserved genomic regions reflect the product of hundreds of millions of years of evolutionary optimization; thus the overwhelming majority of segregating variation in these regions will have deleterious fitness effects.
+Consequently, a good predictor of whether a new mutation will reduce fitness is if it occurs in a region of the genome that has been conserved over phylogenetic timescales \parencite{Siepel2005-wh,Margulies2003-sn}.
+Moreover, segregating rare variation in these regions is responsible for a significant proportion of the genetic contribution to phenotypic variation and disease in humans\parencite{Zeng2018-ci,Lek2016-cb,Karczewski2020-ky,Tennessen2012-ge}.
+
+Selection on both beneficial and deleterious variants perturbs the allele frequencies of neighboring linked sites, a phenomenon known as linked selection
+\parencite{Nordborg1996-nq,Maynard_Smith1974-zr,Barton1998-aj,Charlesworth1993-gb,Kaplan1989-sc}.
 Since deleterious variation is clustered in functional portions of the genome,
-we expect linked selection to reduce levels of diversity around conserved
+we expect linked selection to reduce levels of diversity around evolutionarily constrained
 segments (e.g. genes, regulatory elements, etc.). The genomic arrangement of
-these evolutionarily conserved regions and recombination rates create a
-large-scale spatial signal of the linked selection process in genetic diversity
-along the genome. Since genome-wide recombination maps and putatively conserved
-segments are available for many species, there has been consistent effort to
-fit models to this linked selection signal to estimate population genetic
+these evolutionarily conserved regions coupled with heterogeneous recombination rates create a
+large-scale spatial signal of linked selection of genetic diversity along chromosomes.
+Since genome-wide recombination maps and functional annotations are available for many species, there has been consistent effort to fit models of linked selection to patterns of diversity.
+This general approach provides estimates of population genetic
 parameters such as the strength of selection and the deleterious mutation rate
-\parencite{Hudson1995-pt,McVicker2009-ax}, as well as distinguish and quantify
-the roles of positive and negative selection and estimate the rate of
+\parencite{Hudson1995-pt,McVicker2009-ax}, and potentially distinguishes the roles of positive and negative selection and estimate the rate of
 beneficial mutations \parencite{Elyashiv2016-vt,Murphy2022-sj}. In humans,
@@
 \parencite{Charlesworth1993-gb,Nordborg1996-nq,Hudson1995-pt,Hudson1995-xc}.
-While this model has been successful in explaining many patterns of diversity,
-some its simplifying assumptions may distort inferences about the negative
-selection process. First, since fixation probabilities ultimately depend on the
+While the BGS model has been successful in fitting many patterns of diversity,
+some of its simplifying assumptions may distort inferences about the selective process. First, since fixation probabilities ultimately depend on the
 product of the deleterious selection coefficient ($s$) and population size
-($N$), the efficacy of negative selection should depend on demography.
-Unfortunately, accommodating demography into theoretic negative selection
-models remains an open, difficult problem \parencite{Zeng2013-ep,Johri2020-oj}.
-Second, building off classic models of mutation-selection balance
-\parencite{Crow1970-wm,Kimura1966-bk}, the BGS model assumes that new mutations
-are sufficiently deleterious that they are invariably driven to loss. Under
+($N$), the efficacy of selection depends on past population sizes.
+Unfortunately, accommodating such demography into purifying selection models remains an open, difficult problem \parencite{Zeng2013-ep,Johri2020-oj}.
+Second as the BGS model builds off classic models of mutation-selection balance
+\parencite{Crow1970-wm,Kimura1966-bk}, it assumes that new mutations are sufficiently deleterious that they are invariably driven to loss. Under
 this assumption, the effect of selection is well-approximated by simply
@@
 relevant to natural populations \parencite{McVean2000-bt,Good2014-yz}. In
-particular, the BGS model cannot accommodate the possibility of fixation for
-mutations with $2Ns \ll 1$, leading diversity to be incorrectly predicted to
-fall to zero as the strength of selection diminishes. This weak selection
-problem is related to the wider problem of modeling the substitution rate of
-deleterious mutation (i.e. the ``ratchet" rate) and incorporating the impact of
-interference from other neighboring selected sites. Finally, the classic BGS
-model assumes that selection occurring within other segments does not interfere
-with selection occurring in other regions, i.e. no selective or
+particular, the BGS model cannot accommodate the possibility of weakly deleterious mutations (those with fitness effects $2Ns \le 1$) reaching fixation, which leads to incorrect
+predictions of diversity levels as the strength of selection diminishes. 
+Finally, the classic BGS
+model assumes that the selective dynamics occurring within a segment are not impacted by selection occurring outside the segment, i.e. no selective or
 ``Hill--Robertson" interference
@@
 In this work, we use another class of linked selection models that derive from
-quantitative genetics to address limitations of the classic BGS model. These
-models express linked selection in terms of how polygenic fitness variance
-creates weak genetic draft, and which increases pairwise coalescence rates.
-While these models can theoretically accommodate additive fitness variance from
-any source as long as its rate of change is not too rapid, we focus
+quantitative genetics to address limitations of the classic BGS model \parencite{Santiago1998-bs, Santiago2016-mu,Santiago1995-hx,Robertson1961-ho}.
+These models consider how polygenic fitness variance spread along the genome increases the variance of stochastic allele frequency change, as alleles become randomly linked to fitness backgrounds over time and their frequency trajectories are perturbed by selection at other sites.
+While these models can theoretically accommodate additive fitness variance from any source as long as its rate of change is not too rapid, we focus
 specifically on a deleterious-mutations-only model of fitness variance from
@@
 deleterious substitution rate. We extend the Santiago and Caballero (hereafter
-the SC16) model of the negative selection process so that it can be fit to
-patterns of genome-wide diversity, according to the spatial distribution of
-genomic features that could harbor deleterious fitness variation. Using forward
-simulations, we show this model leads to more accurate estimates of the
-distribution of fitness effects under weak selection. We apply our
+the SC16) model of the negative selection process so that it can be fit using a composite likelihood approach to patterns of genome-wide diversity, according to the spatial distribution of genomic features that could harbor deleterious fitness variation.
+Using forward simulations, we show this model leads to more accurate estimates of the distribution of fitness effects (DFE) under weak selection.
+We apply our
 composite-likelihood method to human population genomic data and provide new
-estimates of the negative selection process in humans. However, we find that
-there may be considerable uncertainty around these estimates if selective
-interference is occurring across segments under purifying selection.
+parameter estimates of the genome-wide impact of purifying selection in humans. 
+We show that our new method is better able to predict the patterns of diversity along human chromosomes than previous models. 
+However, our model leads to predictions of the deleterious substitution rate that disagree with observed levels of divergence.
+We discuss the potential causes and implications of such discrepancies and what it might mean for future efforts to fit linked selection models to genomic patterns of variation.
 
 \section*{Theory}
-
-% Similar to previous work, we aim to characterize selection processes through
-% the signal left on nucleotide diversity at linked neutral sites. Although
-% nucleotide diversity is a less informative statistic than the full site
-% frequency spectrum, it is much easier to derive theoretic models of how linked
-% selection alters expected pairwise coalescence times than to describe the
-% entire genealogy. Still, inferences drawn from the linked selection signal
-% about selective processes are only as accurate as the underlying theoretic
-% model. Since we base our method on a fundamentally different theoretic approach
-% to modeling linked selection, we briefly introduce the theory was extend to
-% apply across whole-genomes.
 
@@
 \parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu},
-which model the effects linked selection in terms of polygenic additive fitness
-variance ($V_A$). These linked selection models stem from
-\textcite{Robertson1961-ho}, which in essence describes how heritable polygenic
-fitness variation causes weak genetic draft. At the individual level Robertson
-considered, selection generates autocorrelation as offspring from large
-families tend to beget many descendents themselves (and likewise with small
-families) when fitness is heritable. This same autocorrelation occurs at the
-genomic level with linkage \parencite{Santiago1998-bs,Barton2000-zg}, as the
+which model the effects of linked selection in terms of polygenic additive fitness
+variance ($V_A$). Here we review the relevant theory before introducing our genome-wide extension. These linked selection models stem from
+\textcite{Robertson1961-ho}, which in essence describes how polygenic additive
+fitness variation increases the pace of stochastic allele frequency change, thus reducing effective population size. At the individual level Robertson considered, 
+selection generates an autocorrelation in fecundity as offspring from large
+families tend to beget many descendants themselves (and likewise with small
+families) when fitness is heritable. This same across-generation autocorrelation occurs at the
+genomic level due to linkage \parencite{Santiago1998-bs,Barton2000-zg}, as the
 perturbations to a neutral allele's trajectory from its particular fitness
 background tend to occur in the same direction across generations until the
-background recombines off. These models quantify the total impact of the
-autocorrelation generated by weak draft in terms of what we think of as a
-\emph{draft-effective} population size $N_d$. The key insight is that in the
-long-run, steady, weak draft created by the presence of additive genetic
-fitness variance $V_A > 0$ acts like an extra variance in offspring number
-would under pure drift \parencite{Wright1938-tv}. However, due to
-autocorrelation, the cumulative effect of this fitness variance is inflated by
-a factor of $Q^2$. Intuitively, the product $V_A Q^2$ represents the expected
-total variance in reproductive success a neutral mutation experiences over its
-lifetime in a system with weak selection draft.
+background recombines off. Quantitative genetic models such as \textcite{Santiago1998-bs}
+quantify the total impact of the autocorrelation generated by selection in terms of what we think of as a
+\emph{fitness-effective} population size $N_f$ (to differentiate it from the \emph{drift}-effective population size, which is the size of the ideal population when there is no fitness variation).
+
+The key insight is that in the long run, the steady presence of additive genetic fitness variance ($V_A > 0$) contributes an extra source of variance in offspring number beyond the variance expected under pure drift \parencite{Wright1938-tv}.
+However, because heritable fitness variation generates across-generation autocorrelation, the cumulative effect of this fitness variance on the variance in allele frequency change is inflated by a factor of $Q^2$.
+Intuitively, the product $V_A Q^2$ represents the expected
+total variance in reproductive success a neutral mutation experiences over its lifetime in a system with weak selection at linked sites.
 
 Following \textcite{Robertson1961-ho} and \textcite{Santiago1995-hx}, we define
-the draft-effective population size $N_d$ by including the total additional
+the fitness-effective population size $N_f$ by including the total additional
 variance created by heritable fitness into Wright's (\citeyear{Wright1938-tv})
@@
     \label{eq:main_Ne}
-    N_d = \frac{N}{Q^2 V_A + 1}
+    N_f = \frac{N}{Q^2 V_A + 1}
 \end{align}
@@
 (c.f. \cite{Robertson1961-ho,Santiago1995-hx}; see Supplementary Materials Section
-\ref{supp:theory} for a proof). The benefit of modeling draft with Robertson's
-forward-time model is that the inflation factor is invariant with respect to
-the particular fitness background the neutral allele becomes stochastically
-associated with. By contrast, the difficulty with modeling draft backwards in
-time is that the coalescence rates experienced by a lineage are not invariant
-to which lineage was sampled due to its particular associated fitness
-background.
+\ref{supp:theory} for a proof). The benefit of modeling linked selection with Robertson's
+forward-time model is that the inflation factor is invariant with respect to the particular fitness background the neutral allele becomes stochastically associated with.
+By contrast,  modeling diversity levels under linked selection backwards in time requires tracking the particular associated fitness backgrounds, as coalescence rates experienced by a lineage are not invariant with respect to their fitness background.
 
@@
 fitness variation has a multiplicative polygenic basis, as is often assumed for
-genome-wide negative selection processes, the draft asymptotic effective
+genome-wide selection processes, the fitness-effective
 population size experienced by an arbitrary neutral site under the influence of
@@
     \label{eq:polygenic_Ne}
-    N_d \approx N \exp\left(-\sum_{i=1}^S V_{A,i} \frac{Q_i^2}{2}\right)
+    N_f \approx N \exp\left(-\sum_{i=1}^S V_{A,i} \frac{Q_i^2}{2}\right)
 \end{align}
 %
-where the factor of one-half comes from ignoring weak associations from
-unlinked regions and chromosomes (see Supplementary Materials Section
+where the factor of one-half comes from ignoring the short-lived 
+associations with the homologous chromosome, 
+which have a weak effect on the focal allele (see Supplementary Materials Section
 \ref{supp:heritable-fitness}). In our genome-wide model, we consider the
@@
 contributes additive fitness variance $V_{A,i}$ to the total additive genetic
-fitness variance. The autocorrelation function for a neutral allele associated
-with segment $i$ is $C(t) = [(1-r_i)(1-\kappa_i)]^t$, where $r_i$ is the
-recombination fraction to the segment and $\kappa_i$ is the rate that the
+fitness variance. The impact this segment's fitness variance $V_{A,i}$ has on 
+the fitness-effective population size is mediated by the autocorrelation term 
+$Q_i$, which is a decaying function of the recombination rate between the 
+segment and focal neutral allele. Specifically, the autocorrelation function
+for a neutral allele associated with segment $i$ is $C(t) = [(1-r_i)(1-\kappa_i)]^t$, 
+where $r_i$ is the recombination fraction to the segment and $\kappa_i$ is the rate that the
 associated fitness variance decays due to selective dynamics. Then, the
@@
 %
-(see Supplementary Materials Equation \ref{supp-eq:Qinf}). This general equation can accommodate
-models of polygenic selection as long as the equilibrium additive fitness
-variation $V_{A,i}$ can be specified and the change in variance due to
-selection can be approximated as a geometric decay, i.e. $\Delta V_{A,i} =
--\kappa V_{A,i}$ \parencite{Bulmer1971-ae,Keightley1988-eq,Walsh2018-bt}. This
-is usually a reasonable assumption since within-generation selection removes a
-fraction of phenotypic variation from the population, and some fraction of that
-is additive genetic variation \parencite{Bulmer1971-ae,Keightley1988-eq}.
+(see Supplementary Materials Equation \ref{supp-eq:Qinf}).
+This general equation can accommodate models of polygenic selection as long as the equilibrium additive fitness variation $V_{A,i}$ can be specified and the change in variance due to selection can be approximated as a geometric decay, i.e.
+$\Delta V_{A,i} = -\kappa V_{A,i}$ \parencite{Bulmer1971-ae,Keightley1988-eq,Walsh2018-bt, Santiago1998-bs}.
+This is usually a reasonable assumption since within-generation selection removes a fraction of phenotypic variation from the population, and some fraction of that is additive genetic variation \parencite{Bulmer1971-ae,Keightley1988-eq}.
 
@@
 variance $V_A$ and the decay rate in associated fitness $1-\kappa$. Fitness
-variance could arise from positively or negatively selected alleles, but given
-prior work has found negative selection against new deleterious mutations play
-a dominant role in shaping genome-wide patterns of diversity
-\parencite{McVicker2009-ax,Murphy2022-sj}, we focus on models of fitness
-variance created by the continual influx of deleterious mutations into a
-population. We imagine a mutation-selection process that creates fitness
-variance through a continual influx of deleterious mutations occur at a rate of
+variance could arise from beneficial or deleterious alleles, but given
+prior work has found selection against new deleterious mutations in conserved regions plays a dominant role in shaping genome-wide patterns of diversity \parencite{McVicker2009-ax,Murphy2022-sj}, we focus specifically on purifying selection.
+We imagine a mutation-selection process that creates fitness variation as deleterious mutations enter a population at rate
 $\mu$ per basepair per generation in a conserved region of $L$ basepairs, such
@@
 new mutation (at frequency $x=\nicefrac{1}{2N}$) is $2s^2 x(1-x) \approx
-\nicefrac{s^2}{N}$. For the entire population of $2N$ chromosomes, the fresh
-variance from mutation each generation in segment $i$ is $V_{m,i} \approx
+\nicefrac{s^2}{N}$. For the entire population of $2N$ chromosomes, the mutational variance input each generation in segment $i$ is $V_{M,i} \approx
 U_is^2$ where $U_i = 2\mu L_i$ is the diploid mutation rate per generation
@@
 \begin{align}
-    N_d = N \exp \left( - \sum_i^S \frac{\mu L_i}{s(1 + r_i(1-s)/s)^2} \right) 
+    N_f = N \exp \left( - \sum_i^S \frac{\mu L_i}{s(1 + r_i(1-s)/s)^2} \right) 
 \end{align}
@@
 deleterious mutations per haplotype is no longer well-approximated by the
-deterministic mutation-selection-balance theory; their dynamics are strongly
-influenced by stochastic perturbations due to both drift and draft. In this
-weak selection regime, classic BGS no longer accurately predicts levels of
-linked diversity
+deterministic mutation-selection-balance theory, and their dynamics are strongly
+influenced by stochastic perturbations due to both drift and linked selection. 
+In this weak selection regime, classic BGS theory no longer accurately predicts levels of linked diversity
 \parencite{Charlesworth1993-gb,McVean2000-bt,Good2013-lp,Gordo2002-dr}.
-Moreover, selection against weakly deleterious mutations alters the
-genealogies, such that they are no longer well-approximated by a rescaled
+Moreover, selection against weakly deleterious mutations alters the topology of genealogies, such that they are no longer well-approximated by a rescaled
 neutral coalescent as assumed under the classic BGS model
@@
 complicate matters, the distribution of the number of deleterious mutations
-(and its corresponding fitness distribution) is no longer a stationary Poisson
-distribution, instead becoming a traveling wave
-\parencite{Rouzine2008-qz,Good2013-lp,Gessler1995-hz} towards reduced
-population fitness (since we ignore beneficial mutations). When this occurs,
-individuals in the least-loaded class are stochastically lost, fixing those
-mutations with a click of ``Muller's ratchet"
-\parencite{Muller1964-ki,Charlesworth1997-qn}. Unfortunately, determining the
-rate of Muller's ratchet is another difficult, open problem
+(and its corresponding fitness distribution) is no longer a stationary Poisson distribution, instead becoming a traveling wave
+\parencite{Rouzine2008-qz,Good2013-lp,Gessler1995-hz} towards increased numbers
+of deleterious alleles per chromosome and reduced mean population fitness.
+In asexual populations, substitutions correspond to a click of ``Muller's ratchet", which is 
+the stochastic loss of the least-loaded class \parencite{Muller1964-ki,Charlesworth1997-qn}.
+Unfortunately, determining the rate of Muller's ratchet is another difficult problem
 \parencite{Haigh1978-gt,Gordo2002-dr,Gessler1995-hz,Higgs1995-xc} related to
@@
 
-Under qualitative genetic models of linked selection, these complications can
+Under quantitative genetic models of linked selection, these complications can
 be avoided by finding a general expression for the equilibrium $V_A$ applicable
-for both strong and weakly deleterious mutations. \textcite{Santiago2016-mu}
+to both strong and weakly deleterious mutations while concurrently modeling the rate
+of fixation of deleterious alleles in the region, $R$. \textcite{Santiago2016-mu}
 suggest that equilibrium fitness variance is lower than predicted by
 $V_{a}^{BGS}$ once weakly deleterious mutations begin to have an appreciable
-rate of fixation, $R > 0$, since each substitution (or click of Muller's
-``ratchet") removes a segregating site and its heterozygosity becomes zero.
-Thus, the steady-state genic variance under mutation and negative selection is
+rate of fixation $R > 0$ per generation in the region. The substitution rate $R$
+decreases fitness variance since each substitution removes a segregating site 
+and thus its contribution to fitness variance. Thus the steady-state additive 
+genetic variance of fitness under mutation and negative selection is
 (omitting the segment index $i$ for clarity),
@@
 where the condition $V_a \ge 0$ is met when the probability of fixation is less
-than or equal to the neutral fixation probability of $\nicefrac{1}{2N}$, which
-is held when considering deleterious mutations. This equation connects the
-equilibrium additive genic fitness variance to the flux of new variation in
-from new deleterious mutations, and the flux out due to their substitution (and
-the decline in mean population fitness). When $R=0$, selection is so strong it
-cannot fix, and the equilibrium fitness variation is due entirely to young rare
-mutations before their extinction $V_a = V_a^{BGS} \approx Us$. Santiago and
+than or equal to the neutral fixation probability of $\nicefrac{1}{2N}$, as is true for all deleterious mutations. This equation describes the
+equilibrium additive genetic fitness variance as the balance of the flux of new variation in to the population from deleterious mutations, and the removal of variation due to their substitution (and the decline in mean population fitness).
+When $R=0$, selection is so strong deleterious alleles cannot fix, and the equilibrium fitness variation is due entirely to young rare mutations before their extinction $V_a = V_a^{BGS} \approx Us$.
+Santiago and
 Caballero derive Equation \eqref{eq:Va} through Fisher's Fundamental Theorem of
@@
 
-%Our results differ from \parencite{Santiago2016-mu}, since we find from
-%simulations their expression models the additive \emph{genic} fitness
-%variation, not additive genetic fitness variation (see Section XXX). This is
-%because Equation \eqref{eq:Va} does not capture the expected build-up of
-%negative linkage disequilibria when selection is mildly deleterious.
-
 While using Equation \eqref{eq:Va} in Equation \eqref{eq:main_Ne} leads to a
-prediction for the draft-effective population size $N_d$, closed-form
-expressions for the rate of the ratchet $R$ have generally been hard to find
-\parencite{Haigh1978-gt,Higgs1995-xc,Gessler1995-hz}. The key insight of
-\textcite{Santiago2016-mu} is that the ratchet rate under draft is determined
-by the probability of fixation $p_F(N_d, s)$
-\parencite{Kimura1962-su,Malecot1952-qh} using the draft-rescaled effective
-population size, i.e. $R = N U p_F(N_d, s)$. Given this equation for the
-ratchet and Equation \eqref{eq:polygenic_Ne} for $N_d$ under draft, we have a
-system of two non-linear equations that can be solved numerically for $N_d$ and
+prediction for the fitness-effective population size $N_f$, closed-form
+expressions for the deleterious substitution rate $R$ have generally been hard to find
+\parencite{Haigh1978-gt,Higgs1995-xc,Gessler1995-hz}. A key insight of
+\textcite{Santiago2016-mu} is that the deleterious substitution rate with linked selection is determined by the probability of fixation $p_F(N_f, s)$
+\parencite{Kimura1962-su,Malecot1952-qh} using the rescaled fitness-effective
+population size, i.e. $R = N U p_F(N_f, s)$. Given this equation for the
+substitution rate and Equation \eqref{eq:polygenic_Ne} for $N_f$ under linked selection, we have a
+system of two non-linear equations that can be solved numerically for $N_f$ and
 $R$ for each segment,
@@
   \label{eq:main_eqns}
-  N_d &= N \exp \left( -(U-R)s \frac{Q^2}{2} \right) & \text{\emph{draft-effective population equation}} \\
-  R &= \frac{4N_d U s}{\exp(4 N_d s)-1}  & \text{\emph{ratchet equation}}
+  N_f &= N \exp \left( -(U-R)s \frac{Q^2}{2} \right) & \text{\emph{fitness-effective population equation}} \\
+  R &= \frac{4N_f U s}{\exp(4 N_f s)-1}  & \text{\emph{substitution rate equation}}
 \end{align}
@@
 
-We denote the solutions to these equation, which represent equilibria under
-mutation-selection-drift-draft process, as $\widetilde{N}_d$ and
+We denote the solutions to these equations, which represent equilibria under
+mutation-selection-drift process, as $\widetilde{N}_f$ and
 $\widetilde{R}$. These equilibria also imply an equilibrium level of additive
 fitness variation $\widetilde{V}_a$ in the segment, which are used to calculate
-the reduction factor $B(x) = \nicefrac{N_d}{N}$ at any genomic position $x$
+the reduction factor $B(x) = \nicefrac{N_f}{N}$ at any genomic position $x$
 (see Methods Section \nameref{sec:methods-maps}).
 
-\section*{Results: Simulation Validation of Theory and Methods}
-
-Given that modeling the interplay of mutation, drift, and draft under both weak
-and strong selection has proven to be difficult problem, we first sought to
-verify the SC16 theory and our genome-wide extension of this theory with three
-levels of simulations: forward simulations of negative selection in a region,
-chromosome-scale forward simulations of negative selection, and simulations of
-a ``synthetic genome" to test our composite-likelihood method based on this
+\section*{Results}
+
+We provide two main classes of results. First, we show simulation results which demonstrate
+the accuracy of the SC16 model over the BGS approximation across the parameter space,
+as well as validations of the composite likelihood strategy we use to fit the SC16 model.
+Second, we provide fits of our method to human genome data, where we show comparison of models fit using different annotations, the estimated DFEs, and predictions of the deleterious substitution
+rate.
+
+\subsection*{Simulation Validation of Theory and Methods}
+
+Given that modeling the interplay of mutation, drift, and linked selection under both weak and strong selection has proven to be a difficult problem, we first sought to verify the SC16 theory and our genome-wide extension with three
+levels of simulations: forward simulations of purifying selection in a region,
+chromosome-scale forward simulations of purifying selection, and simulations of
+a ``synthetic genome" (i.e. by combining independently simulated chromosomes) to test our composite-likelihood method based on this
 theory.
 
-\subsection*{Simulations of a Segment under Negative Selection}
+\subsubsection*{Simulations of a Segment under Purifying Selection}
 \label{sec:segment-sims}
@@
 \begin{figure}[htbp] \centering
-    \includegraphics[width=\textwidth]{figures/figure_1.pdf} \caption{Theory
-        compared to simulation results. (A) The predicted reduction factor
+    \includegraphics[width=\textwidth]{figures/figure_1.pdf} \caption{Santiago and Caballero (2016) theory models the weak selection regime better than classic BGS theory.
+(A) The predicted reduction factor
         under classic B theory (dark gray line) and the diploid SC16 model
         (colored lines corresponding to mutation rate) compared to average
-        reduction across 10,000 simulation replicates (points). (B) The predicted
-        ratchet rate under the SC16 model scaled by mutation rate (colored
-        lines) compared to the ratchet estimated from simulation (points). When
-        $2Ns>1$, the ratchet rate is near zero. (C) The genic variance from
+        reduction across 10,000 simulation replicates (points). The inset figure is zoomed out to show extent of disagreement under classic BGS. (B) The predicted
+        deleterious substitution rate under the SC16 model, scaled by mutation rate (colored
+        lines) compared to the substitution rate estimated from simulation (points). When
+        $2Ns>1$, the substitution rate is near zero. (C) The genic variance from
         simulations (points) against the predicted variance under the SC16
-        model (colored lines). As the ratchet begins to click, the genic
+        model (colored lines). As substitutions begin to occur, the genic
         variance is decreased from the level expected under strong BGS (dashed
@@
         whole-chromosome simulations and predicted classic B (dots), new B'
-        (solid), and locally-rescaled B' (dashed) for different mutation rates
-        (colors). The dashed horizontal line is the approximate theoretic
-        minimum MSE. (D, top) The build up of negative linkage disequilibria
-        around $2Ns=1$ in whole-chromosome simulations in bottom panel. (E) The
+        (solid), and locally rescaled B' (dashed) for different mutation rates
+        (colors). Locally rescaled B' (yellow lines) are omitted for clarity in the top and bottom 
+        rows, since they are identical to B'; Local rescaling only impacts B'
+        in the $2Ns \approx = 1$ domain. The dashed horizontal line is the approximate theoretic
+        minimum MSE. (D, top) The build-up of negative linkage disequilibria
+        around $2Ns=1$ in whole-chromosome simulations shown in the bottom panel. (E) The
         average B map from 100 chromosome 10 simulation replicates (gray)
@@
 
-Our first set of forward simulations were to ensure that the SC16 model
-adequately captures selective dynamics in a single $10^5$ basepair region
-undergoing negative selection, across a variety of mutation rates and selection
-coefficients (see Methods \ref{sec:methods-sim}). We find a close
+Our first set of forward simulations was to ensure that the SC16 model
+adequately captures selective dynamics in a single $100$ Mbp basepair region
+under selection, across a variety of mutation rates and selection
+coefficients (see Methods \nameref{sec:methods-sim}). We find a close
 correspondence between the observed and predicted reductions in effective
-population size $B=\nicefrac{N_d}{N}$ over all selection and mutation
+population size $B=\nicefrac{N_f}{N}$ over all selection and mutation
 parameters including weak selection (Figure \ref{fig:figure-1}A), in contrast
@@
 caused by the model correctly predicting the equilibrium fitness variance and
-ratchet rate, we also measured these throughout the simulation. Again, we find
-the diploid SC16 theory accurately predicts both the ratchet rate (Figure
+substitution rate, we also measured these throughout the simulation. Again, we find
+diploid SC16 theory accurately predicts both the deleterious substitution rate (Figure
 1\ref{fig:figure-1}B) and the genic fitness variance (Figure
@@
 
-Moreover, these simulations provide intuition about the underlying negative
+Moreover, these simulations provide intuition about the underlying 
 selection process. When mutations are strongly deleterious, there is no chance
-they can fix, and the ratchet rate is zero (Figure \ref{fig:figure-1}B for $2Ns
+they can fix, and the substitution rate is zero (Figure \ref{fig:figure-1}B for $2Ns
 > 1$). In this strong selection regime, the additive genic fitness variation
@@
 gray line, Figure \ref{fig:figure-1}C) However, around $2N_e s \approx 1$, the
-ratchet begins clicking as $p_F > 0$. When this occurs, each click of the
-ratchet eliminates variation, and the equilibrium variation diverges from the
+substitutions begin to occur as $p_F$ moves away from zero. When this occurs, each fixation eliminates variation, and the equilibrium variation diverges from the
 deterministic mutation-selection equilibrium (Figure \ref{fig:figure-1}C).
 
-\subsection*{Chromosome-wide Simulations and Models of Negative Selection}
+\subsubsection*{Chromosome-wide Simulations and Models of Negative Selection}
 \label{sec:chrom-sims}
@@
 Given the accuracy of the SC16 model in predicting the reduction factor $B$ and
-ratchet rate for a single segment under general mutation-selection processes,
-we next extended their model so that it could be applied to whole-genome data.
-Our software method \texttt{bgspy} numerically solves Equations
+the deleterious substitution rate for a single segment under general mutation-selection processes, we next extended their model so that it could be fit to patterns of windowed genome-wide
+diversity through a composite likelihood approach. Our software method 
+\texttt{bgspy} numerically solves Equations
 \eqref{eq:main_eqns} to compute the equilibrium additive genic fitness variance
-($\widetilde{V}_a$) and ratchet rate ($\widetilde{R}$) across grids of mutation
+($\widetilde{V}_a$) and the deleterious substitution rate ($\widetilde{R}$) across grids of mutation
 rates and selection coefficients. This is done for each pre-specified segment
-in the genome that may be under negative selection (e.g. coding sequences or
-UTRs). We call the set of predicted reductions across these grids the B' maps
-(to distinguish them from McVicker's B maps); these can be used then to find
+in the genome (potentially tens of millions of small regions) that may be under purifying selection (e.g.
+coding sequences or UTRs).
+We call the set of theoretic predicted reductions across these grids the B' maps
+(to distinguish them from McVicker's B maps \parencite{McVicker2009-ax});
+these can be used to find
 the equilibrium reduction factor $B(x)$ for any genomic position $x$.
@@
 We validated our predicted B' reduction maps with realistic chromosome-scale
-forward simulations of negative selection in conserved regions. We find that
+forward simulations of purifying selection using putatively conserved regions and recombination maps for the human genome. We find that
 our B' maps and the classic BGS theory B maps closely match simulations when
@@
 the relative error of the classic B maps is 14.6\% whereas the relative error
-in the new B' maps is 5\%. Nearly all of this error is in the drift-barrier
+in the new B' maps is 5\%. Nearly all of this error is in the nearly neutral
 domain ($2N_e s \approx 1$ domain); for strong and weak selection, the mean
 squared error between simulations and B' maps is close to the theoretic lower
-bound of $\V(B) \approx \nicefrac{2}{9n}$ set by the coalescence variance
+bound of the mean squared error, $\approx \nicefrac{2}{9n}$ set by the coalescence variance for a 10 kb region
 \parencite{Tajima1983-gu}.
 
-We hypothesized this error in the drift-barrier domain may be due to selective
+We hypothesized this error in the nearly neutral domain may be due to selective
 interference between segments that is not taken into account when we
 numerically solve Equations \eqref{eq:main_eqns} independently for each
-segment. In particular, we use a fixed drift-effective $N=1,000$ corresponding
-to the number of diploids in the simulations rather than $B(x) N$, which is an
-approximation to the draft-effective population size experienced by a selected
-segment at position $x$. To test this, we implemented a ``locally-rescaled"
-version of the B' maps, which uses $B(x)N$ as the population size. We find the
-locally-rescaled B' maps reduce the relative error from 5\% to 0.4\% and mean
-squared error (Figure \ref{fig:figure-1}D, dashed colored lines), but does not
-entirely eliminate the error in the $2Ns \approx 1$ domain.
-
-Finally, we hypothesized that the remaining error could be due to the build up
+segment. In particular, we use a fixed drift-effective population size, $N=1,000$, corresponding 
+to the number of diploids in the simulations rather than $B(x) N$, which is an approximation to 
+the fitness-effective population size experienced by a selected segment at position $x$. To test 
+this, we implemented a ``locally rescaled" version of the B' maps, which uses $B(x)N$ as the population size.
+We find the locally rescaled B' maps reduce the relative error from 5\% to 0.4\% and mean squared error (Figure \ref{fig:figure-1}D, dashed colored lines), but does not entirely eliminate the error in the $2Ns \approx 1$ domain.
+
+We hypothesized that the remaining error could be due to the build-up
 of negative linkage disequilibria between selected sites due to Hill--Robertson
 interference \parencite{Hill1966-kd,McVean2000-bt,Comeron2007-wq}. By
-calculating the sum of all linkage disequilibria pairs in our chromosome-wide
-simulations, we find such a build up of negative LD around $2Ns \approx 1$
+calculating the sum of all linkage disequilibria among all pairs of selected sites in our chromosome-wide simulations, we find such a build-up of negative LD around $2Ns \approx 1$
 (Figure \ref{fig:figure-1}D, top row). As $Ns \to 0$, the variance in LD
-inflates as expected \parencite{Ohta1969-ae,Hill1968-ue}. Overall, the build up
+inflates as expected \parencite{Ohta1969-ae,Hill1968-ue}. Overall, the build-up
 of negative LD is consistent our view that the equilibrium fitness variance
-modeled by the SC16 theory is the additive \emph{genic} fitness variance, which
-differs from the additive genetic variance by the sum of linkage disequilibria
+modeled by the SC16 theory is predominantly the additive \emph{genic} fitness variance, which differs from the additive genetic variance by the sum of linkage disequilibria
 between selected sites (i.e. $\delta_{LD} = s^2 \sum_{i\ne j} D_{i,j}$ where
@@
 fitness variance that include the contribution of LD (Supplementary Materials
-Section \ref{supp:theory}).
-
-\subsection*{Validation of Composite-likelihood Method using Forward
+Section \ref{supp:theory}); thus the fitted SC16 model has predictable bias in the nearly neutral selection regime.
+
+\subsubsection*{Validation of Composite-likelihood Method using Forward
 Simulations}
@@
         Overall, classic BGS methods break down as expected when full-coverage tracks
-    are used, since it cannot accommodate weak selection and neutrality. }
+    are used, since it cannot accommodate weak selection and neutrality in putatively
+    conserved regions. See methods for details on sparse versus full tracks.
+    }
 
@@
 
-We next developed a composite-likelihood method that estimates the distribution
+Our composite-likelihood method estimates the distribution
 of selection coefficients for each feature type, the mutation rate, and the
 diversity in the absence of linked selection ($\pi_0$) by fitting the theoretic
-reduction map to genome-wide diversity (see Methods
-\ref{sec:methods-likelihood}). We validated that our method can accurately
-estimate the parameters of negative selection by simulation a ``synthetic
+reduction map to windowed genome-wide diversity (see Methods
+\nameref{sec:methods-likelihood}). We validated that our method can accurately
+estimate the selective parameters by simulation a ``synthetic
 genome" of the first five human chromosomes (see Methods
-\ref{sec:methods-sim}). We note three findings from these simulations.
-
-First, we find the $R^2$ between predicted and simulated megabase-scale
-diversity serves as a measure of the strength of the linked selection signal in
-genome-wide data. The $R^2$ increases with the intensity of selection against
-new deleterious mutations and mutation rate (Figure \ref{fig:figure-2}, top
-row). Under just drift or weak negative selection, the variance in diversity is
-driven by unstructured coalescence noise along the genome and the predicted
-reduction map $B(x)$ does not fit the data well. Under very strong selection
-($s=0.05$), $R^2$ is reduced; this is likely due to very strong selection
-having less localized effects and impacting overall genome-wide diversity
-\parencite{Santiago1995-hx,Robertson1961-ho}.
-
-Second, both our implementation of classic BGS theory and our B' method
+\nameref{sec:methods-sim}). We note three findings from these simulations.
+
+First, both our implementation of classic BGS theory and our B' method
 accurately infer the average selection coefficient under strong selection
 (Figure \ref{fig:figure-2}, middle row). However, when selection was weak, the
-classic BGS model erroneously estimates strong selection and a very low
+classic BGS model erroneously estimated strong selection and a very low
 mutation rate. By contrast, our B' method estimated selection coefficients much
-closer to their true value. A minor discrepancy occurs around $2Ns = 1$, likely
+more accurately. A minor discrepancy occurs around $2Ns = 1$, likely
 due to the sensitivity of mutations in this region to selective interference
-(these do not use local-rescaling). Because we only simulated fixed selection
-coefficients and five chromosomes, we only assessed the accuracy of average
-selection coefficients and not the DFE probabilities.
-
-Third, we find slight biases in mutation rated estimated by both our B' and the
+(these results do not use local rescaling). 
+To ease computational costs, we only simulated fixed selection coefficients and five chromosomes, 
+and we only assessed the accuracy of average selection coefficients rather than the full estimated DFE.
+
+Second, we find slight biases in mutation rate estimates from both our B' and the
 classic BGS methods (Figure \ref{fig:figure-2}, bottom row). However, mutation
@@
 readily diagnosed from the low $R^2$ values. Finally, we find in additional
-tests that our estimates are robust to demographic expansions but sensitive
-when mutuations have recessive effects, since our model assumes additive
+tests that our estimates are robust to demographic expansions but inaccurate
+when mutations have recessive effects, since our model assumes additive
 effects (see Supplementary Materials Section \ref{supp:sim-assum}). We did not
@@
 
-\section*{Results: Application to Human Genomic Data}
-
-\subsection*{Annotation Model Comparison}
+Third, we find the coefficient of determination, $R^2$, between predicted and simulated megabase-scale diversity serves as a measure of the strength of the linked selection signal in genome-wide data.
+$R^2$ increases with the intensity of selection against new deleterious mutations and mutation rate (Figure \ref{fig:figure-2}, top
+row). Under just drift or weak purifying selection, the variance in diversity is
+driven by unstructured coalescence noise along the genome and the predicted
+reduction map, $B(x)$, and does not fit the data well. Under very strong selection
+($s=0.05$), $R^2$ is reduced; this is likely due to very strong selection
+having less localized effects and impacting overall genome-wide diversity
+\parencite{Santiago1995-hx,Robertson1961-ho}.
+
+\subsection*{Application to Human Genomic Data}
+
+\subsubsection*{Annotation Model Comparison}
 \label{sec:annotation}
@@
 (1) CADD-based models, which consider the top $x\%$ most pathogenic basepairs
-according to the CADD score, and (2) and more interpretable gene feature-based
-models that includes protein coding regions, intron and UTRs, and PhastCons
-regions. We include PhastCons regions because there is strong evidence of
-highly-conserved non-coding regions
+according to the CADD score, and (2) and more interpretable, feature-based
+models that includes protein coding regions, introns and UTRs, and PhastCons
+regions. We include PhastCons regions because they include 
+highly-conserved, non-coding regions known to harbor important functions
 \parencite{Meader2010-hm,Harmston2013-tt,Katzman2007-gq,Siepel2005-wh}, that
@@
 Priority, and Feature Priority) to high-coverage 1000 Genome data for three
-populations: Yoruba (YRI), Han Chinese (CHB), and European (CEU) reference
-samples. We assess and compare our models according to how well they predict
+reference samples: Yoruba (YRI), Han Chinese (CHB), and European (CEU). We assess and compare our models according to how well they predict
 patterns of diversity on whole chromosomes left-out during the model fitting
@@
 $R_\text{LOCO}^2$, which is the proportion of the observed variance in genomic
-diversity at the megabase scale predicted by our model. We experimented with a
-few smaller spatial scales (e.g. 100 kbp), but our results were consistent with
-previous results suggesting the human linked selection signal fits best at the
-megabase scale \parencite{Murphy2022-sj}.
+diversity at the megabase scale predicted by our model on held out data. 
+We experimented with a few smaller spatial scales (e.g. 100 kbp), but our results were consistent 
+with previous results suggesting the human linked selection signal due to purifying selection
+fits best at the megabase scale \parencite{Murphy2022-sj}.
 
@@
 
-    \caption{The $R^2$ estimates for a sparse (A) and full (B) models, for all
-        populations (colors) fit at the megabase-scale. Round points are our B'
+    \caption{The $R^2$ estimates for sparse (A) and full (B) models, for all
+        samples (colors) fit at the megabase-scale. Round points are our B'
         method and diamonds are the classic BGS (we exclude classic BGS in the
@@
 
-Overall, we find the PhastCons Priority and CADD 6\% models fit about equally
-as well (Figure \ref{fig:figure-2}A), consistent with recent work using classic
+Overall, we find the PhastCons Priority and CADD 6\% models fit equally
+well (Figure \ref{fig:figure-2}A), consistent with recent work using classic
 BGS theory \parencite{Murphy2022-sj}. However, we find that our models predict
-out-sample diversity levels slightly better than previous methods. For these
+out-of-sample diversity levels slightly better than previous methods. For these
 two models, we find that our B' method predicts $R_\text{LOCO}^2=67.3$\% and
-$R_\text{LOCO}^2=66.7$\% of the out-sample variance in Yoruba pairwise
+$R_\text{LOCO}^2=66.7$\% of the out-of-sample variance in Yoruba pairwise
 diversity at the megabase-scale, respectively. By contrast, the best-fitting
@@
 be explained by other differences in data processing, optimization, etc. For
-populations impacted by the out-of-Africa bottleneck, the goodness-of-fit was
+lineages impacted by the out-of-Africa bottleneck, the goodness-of-fit was
 lower across all models (e.g. 61.0\% and 58.8\% for CEU and CHB respectively in
@@
 
-% Our PhastCons Priority model fits marginally better than the CADD 6\% model
-% consistently across all three populations (Supplementary Materials Table
-% \ref{supp:tbl-r2}). This is likely because the PhastCons Priority model has
-% more DFE parameters, and thus can fit finer-scale differences among the four
-% feature types. By contrast, the CADD 6\% track must fit patterns of diversity
-% to conserved sites under a single DFE. The CADD 8\% model fit slightly worse
-% than the CADD 6\% across all populations, and the Feature Priority model had
-% the poorest-fit.
-
 Since our method is built upon theory that fixes the weak selection problem of
 classic BGS theory, it should in principle fit equally well when an annotation
-model includes regions that are under no selective constraint and thus
-neutrally evolving. To test this, for each of our annotation models (which are
+model includes regions that are under no or little selective constraint and thus
+(nearly) neutrally evolving. To test this, for each of our annotation models (which are
 ``sparse") we fit a corresponding ``full" model that assigns the remaining
@@
 sparse or full, since our method in principle can accommodate weak selection
-and neutrality. Indeed, we find this to be the case: both in-sample $R^2$ and
-out-sample $R_\text{LOCO}^2$ values are nearly identical across full and
+and neutrality. Indeed, we find that both in-sample $R^2$ and
+out-of-sample $R_\text{LOCO}^2$ values are nearly identical across full and
 sparse-track models (Figures \ref{fig:figure-2}A and \ref{fig:figure-2}B, round
-points).
+points), which demonstrates that our method is able to deal with weak selection
+and that there is little more predictive power to gain from including sites considered ``other'' as annotations.
 
@@
 
-Overall, our $R_\text{LOCO}^2$ estimates suggest our negative selection model
+Overall, our $R_\text{LOCO}^2$ estimates suggest our model
 explains up to 67\% of out-sample variance in diversity of the megabase scale,
@@
 $B(b)$. This allows us to calculate $R_\text{coal}^2$ to ballpark the theoretic
-variance that is capable of being explained, assuming this coalescence-only
-noise process (see Supplementary Materials \ref{supp:r2-coal}). We note that
+variance that is capable of being explained, assuming this coalescent
+noise process alone (see Supplementary Materials \ref{supp:r2-coal}). We note that
 selection is expected to \emph{decrease} the variance in coalescence times
-beyond a rescaled effective population size implies; our $R_\text{coal}^2$
-would be an underestimate under selection.
+beyond a rescaled effective population size implies, thus our $R_\text{coal}^2$
+would be an underestimate under models with selection.
 
@@
 fitting all the signal possible, under the coalescence-only noise assumption.
-By contrast, for bottlenecked out-of-Africa populations, we find a larger
+By contrast, for bottlenecked out-of-Africa samples, we find a larger
 discrepancy between $R_\text{coal}^2$ and observed out-sample
 $R_\text{LOCO}^2$. The theoretic $R_\text{coal}^2 \approx 63\%-64$\% for both
-populations, compared to the observed $R_\text{LOCO}^2 \approx 61$\% for CEU
+samples, compared to the observed $R_\text{LOCO}^2 \approx 61$\% for CEU
 and $R_\text{LOCO}^2 \approx 59$\% for CHB under the PhastCons Priority model.
@@
 this gap would likely shrink under more realistic models or simulation-based
-approximations for $R_\text{coal}^2$. Overall, this suggests that negative
-selection models explain the vast majority coalescence time variation at the
-megabase-scale that is capable of being explained (i.e. that is not coalescence
+approximations for $R_\text{coal}^2$. Overall, this suggests that purifying
+selection models fit the vast majority coalescence time variation at the
+megabase-scale that is capable of being explained (i.e. that is not coalescent
 noise).
 
-\subsection*{Estimated Distribution of Fitness Effects}
+\subsubsection*{Estimated Distribution of Fitness Effects}
 
@@
     full-coverage Feature Priority model comparing the estimates across
-populations. Although this model fit the data less well than alternatives, its
+reference population samples. Although this model fit the data less well than alternatives, its
 results are more interpretable.}
@@
 across the selection grid for each of the $K$ feature types. Given that the
-relationship between $\pi$ and the strength selection is U-shaped, we wondered
-whether our new B' model accommodating weak selection could fit the linked
+relationship between $\pi$ and the strength selection is U-shaped (i.e., see \ref{fig:figure-1}A), we wondered
+whether our new B' model accommodating weak selection would fit the linked
 selection signal under a different combination of weak and strong selection
-parameters than found by previous work.
-
+parameters than observed previously.
 However, across all of our annotation models, new deleterious mutations in
@@
 PhastCons regions consistently places $\ge 75\%$ of mass on the largest
-selection coefficient, $s=10^{-2}$. The CADD 6\% DFE estimates imply an average
-selection coefficient of $\bar{s} = 0.65\%$ for CEU, $\bar{s} = 0.57\%$ for
-CHB, and $\bar{s} = 0.79\%$ for YRI. Similarly, the PhastCons Priority model
-implies average selection coefficient estimates of $\bar{s} = 0.63\%$, $\bar{s}
-= 0.59\%$, and $\bar{s}= 0.77\%$ for CEU, CHB, and YRI respectively for
+selection coefficient we used, $s=10^{-2}$. The CADD 6\% DFE estimates imply an average
+selection coefficient of $\bar{s} = 0.0065$ for CEU, $\bar{s} = 0.0057$ for
+CHB, and $\bar{s} = 0.0079$ for YRI. Similarly, the PhastCons Priority model
+implies average selection coefficient estimates of $\bar{s} = 0.0063$, $\bar{s}
+= 0.0059$, and $\bar{s}= 0.0077$ for CEU, CHB, and YRI respectively for
 PhastCons regions. Our DFE estimates for CDS under the Feature Priority model
@@
 includes $s=10^{-1}$. We find that models fit with the strong selection grid 
-had $R_\text{LOCO}^2$ that were about one percentage point higher, which is
-suggestive of stronger selection than has previously been estimated using
+have predictive accuracy, as measured with $R_\text{LOCO}^2$, that were about one percentage point higher. This is suggestive of stronger selection than has previously been estimated using
 constrained grids (see Supplementary Materials Table
 \ref{supp:tbl-r2-strongsel}). For this strong selection grid, we estimate
-average selection coefficients for the CADD 6\% model of $\bar{s} = 4.2\%$ for
-CEU, $3.2\%$ for CHB, and $4.4\%$ for YRI. Thus, there is quite a degree of
-sensitivity to DFEs and average selection estimates based on the selection
-coefficient grid.
-
-However, across the strong selection grid runs, we find indications of model
-non-identifiability. First, estimates of the DFE with the strong grid are
-strangely bimodal (Supplementary Figure \ref{suppfig:strong-sel-dfe}). For
+average selection coefficients for the CADD 6\% model of $\bar{s} = 0.042$ for
+CEU, $0.032$ for CHB, and $0.044$ for YRI. Thus, the predefined selection coefficient grid affects ultimate estimates of the DFE and average selection coefficient.
+
+However, we find indications of model non-identifiability across the strong selection grid runs. First, estimates of the DFE with the strong grid are
+bimodal (Supplementary Figure \ref{suppfig:strong-sel-dfe}). For
 example, under the CADD 6\% strong selection grid model, new mutations are
@@
 samples. We propose that one mechanism for this non-identifiability is that
-very deleterious mutations lead to larger whole-genome reductions, which are
+very deleterious mutations lead to larger whole-genome reductions in diversity, which are
 difficult to distinguish from a smaller drift effective population size (i.e.
@@
 selection coefficient and $\pi_0$, which is includes the drift-effective
-population size $N_e$.
-
-We find this is the case for all of our CADD 6\% models. Across all
-populations, average selection was about $7.1$ times larger using the strong
+population size $N_e$. We find this is the case for all of our CADD 6\% models. Across all reference samples, average selection was about $7.1$ times larger using the strong
 selection grid, and $\pi_0$ was $5.6\%$ higher (see Supplementary Materials
 Figure \ref{suppfig:sel-ident}). There was no similar consistent change in
-mutation rate estimates among populations. In the CADD 6\% model, genome-wide
+mutation rate estimates among reference samples. In the CADD 6\% model, genome-wide
 average reduction factor $\bar{B}$ was $\approx 6.1\%$ lower in the default
@@
 \parencite{Torres2018-ni,Torres2020-hc,Lohmueller2008-qi,Simons2014-sv,Simons2016-cs},
-we were curious how consistent our DFE estimates are across populations.
-Overall, we find DFE estimates are relatively stable across populations and
-models (Supplementary Materials Section \ref{supp:fits}). Only in our Feature
+we were curious how consistent our DFE estimates are across samples from different reference populations.
+Overall, we find DFE estimates are relatively stable across samples from different reference populations and annotation models (Supplementary Materials Section \ref{supp:fits}).
+Only in our Feature
 Priority model (Figure \ref{fig:figure-4}B, top row) do we see a slightly
 different DFE estimate for coding sequences between YRI and CEU/CHB
-populations, but this could be due to the poorer fit this model has to data.
+samples, but this could be due to the poorer fit this model has to data.
 
 Although the Feature Priority model fits the data less well than alternative
-models, its DFE estimates are are more interpretable. We find that our B'
+models, its DFE estimates are more interpretable. We find that our B'
 method estimates a bimodal DFE for coding sequences for the Feature Priority
@@
 non-synonymous sites that constitute coding sequences are under vastly
-different levels of constraint. Moreover, features expected to be only weakly
+different levels of constraint and are lumped together in our annotation class.
+Moreover, features expected to be only weakly
 constrained such as introns and UTRs have the bulk of DFE mass on the neutral
@@
 correspond to highly-conserved non-coding elements) suggests it is under strong
-selective constraint; however we note that block jackknife based uncertainty
+selective constraint; however, we note that block jackknife-based uncertainty
 estimates suggest the model is uncertain whether there is some mass on the
@@
 
-\subsection*{Estimates of the Deleterious Mutation Rate in Humans}
+\subsubsection*{Estimates of the Deleterious Mutation Rate Are Sensitive to Model Choice}
 
@@
 
-    \caption{Mutation rate estimates across sparse track (top row) and
-        full-coverage track (bottom row) models, for the new B' (circles) and
+    \caption{Mutation rate estimates across the sparse (top row) and
+        full-coverage tracks (bottom row) models, for the new B' (circles) and
         classic BGS (diamonds) methods. Estimates of the mutation rate are
@@
 
-Prior work on genome-wide inference using the classic BGS model fit the linked
-selection signal well, but led to unusually high estimates of the mutation rate
-\parencite{McVicker2009-ax}. This lead to the hypothesis that these models
+Prior work on genome-wide inference using the classic BGS model fit the patterns of diversity well,
+but led to unusually high estimates of the mutation rate
+\parencite{McVicker2009-ax}. This led to the hypothesis that these models
 could be absorbing the signal of positive selection \parencite{Enard2014-kz},
-though other work has found a limited role for hitchhiking
-\parencite{Pickrell2009-tt,Hernandez2011-gs,Murphy2022-sj}. While our simulation
+though other work has found a limited role for hitchhiking at amino acid substitutions \parencite{Pickrell2009-tt,Hernandez2011-gs,Murphy2022-sj}. While our simulation
 results suggest estimates of the mutation rate from linked selection models are
@@
 \parencite{Kong2013-fc,Tian2019-so}. 
-
 We find across all populations, our mutation rate estimates from CADD-based
@@
 
-However, mutation rate estimates for feature-based annotation models are
-implausibly higher than pedigree-based estimates. First, mutation rate
-estimates under our implementation of classic BGS theory are an order of
+However, mutation rate estimates for feature-based annotation models
+do not agree with pedigree-based estimates. First, mutation rate
+estimates under from classic BGS theory are an order of
 magnitude below the expected range (Figure \ref{fig:figure-4}A top row). We
@@
 consistent with classic BGS theory being unable to fit the DFE to features
-under weak constraint (e.g. introns, UTRs, and the ``other" feature), and most
-compensate by estimating an unusually low mutation rate.
-
-Second, we noticed that across all populations and sparse/full tracks, the CADD
-6\% model consistently led to slightly higher mutation rates than the CADD 8\%
+under weak constraint (e.g. introns, UTRs, and the ``other" feature), and thus must compensate by estimating too low a mutation rate.
+
+Second, we noticed that across all populations and sparse and full tracks, the CADD 6\% model consistently led to slightly higher mutation rates than the CADD 8\%
 model (Figure \ref{fig:figure-4}A bottom row; Supplementary Materials Table
@@
 parameter $\mu L$, where $L$ is the length of the conserved segment. Even
-though our method is much more robust to inclusion of non-conserved regions
+though our method is much more robust to the inclusion of non-conserved regions
 like introns, we still observe this non-identifiability issue.
 
-Finally, we note that mutation rate estimates in the Feature Priority model are
-unreasonably high ($\widehat{\mu} \approx 3 \times 10^{-8}$), reminiscent of
-the high mutation rate estimates found under McVicker et al.'s
+Finally, we note that mutation rate estimates from the Feature Priority model are themselves too high ($\widehat{\mu} \approx 3 \times 10^{-8}$), reminiscent of the high mutation rate estimates found under McVicker et al.'s
 (\citeyear{McVicker2009-ax}) model. While both our and Murphy et al.'s CADD and
-PhastCons-based models alleviate this issue, it is worth discussing why this
-could occur. We gain some insight from comparing the estimated mutation rates
-between our Feature and PhastCons Priority annotation models, which both
-contain the exact same number of feature basepairs, but these are assigned to
-different feature classes based on the priority of overlapping features. That
-one of these models is our best-fitting model, and the other our worst
-indicates the sensitivity these models have when features classes have
-heterogeneous DFEs. CADD-based models fit better in part due to their
+PhastCons-based models alleviate this issue, it is worth considering why this
+could occur.
+We can potentially gain some insight from comparing the estimated mutation rates from our Feature and PhastCons Priority annotation models, which each contain the exact same number of feature basepairs, but whose composition varies based on the priority of overlapping features.
+That
+one of these models is our best-fitting model and the other our worst
+indicates that model fits are sensitive to feature classes which themselves have heterogeneous DFEs. CADD-based models fit better in part due to their
 fine-scale resolution of selective effects across the genome. While ideally we
@@
 
-\subsection*{Predicted Diversity and Residual Signal}
+\subsubsection*{Despite Close Fit, Residual Purifying Selection Signal Remains}
 
@@
 
-    \caption{(A) Observed and predicted diversity under the CADD 6\% full-track
-        model. Once scaled by average diversity, populations differ little in
-        their predicted diversity levels (colors). Additionally, we show
+    \caption{(A) Observed and predicted diversity of the B' model fit with the
+       CADD 6\% full-track annotation.
+Once scaled by average diversity, predicted diversity for populations (colored lines) differs little across populations, and closely matches observed diversity within each population (light gray lines).
+Additionally, we show
         summaries of CADD density and recombination rate along the chromosome
-        below. (B) Predicted $B$ and observed $\pi$ for each window. The
-        residual variance is in the vicinity of the levels expected from the
-        $R_\text{LOCO}^2$ calculation. (C) Evidence that negative selection
-    signal remains in CADD 6\% residuals (YRI shown). Residuals are plotted
-against the average LoF selection coefficient across genes in a megabase window
-(estimated by \cite{Agarwal2023-un}).}
+        below.
+(B) Predicted $B$ and observed $\pi$ for each window.
+The red dashed line indicates the observed 2 standard deviation ellipsoid, which has nearly the same width as the expected by $R_\text{coal}^2$, indicating the residual variance is close to theoretic expectations.
+The yellow points are binned means, and the yellow line is the lowess curve through predicted and observed values.
+(C) CADD 6\% residuals (YRI shown) plotted against the average LoF selection coefficient across genes in megabase windows (estimated by \cite{Agarwal2023-un}).}
 
@@
 
-Comparing predicted against observed diversity along chromosomes (see
-Supplementary Materials Section \ref{supp:chrom-fit}) we find a close
-correspondence consistent with the high $R_\text{LOCO}^2$. However, we find a
-few large (tens of megabases) regions with systematically poorer fit
+Comparing predicted against observed diversity along chromosomes, we find a close
+correspondence consistent with the high $R_\text{LOCO}^2$ (Figure \ref{fig:figure-5}A).
+Once scaled by the genome-wide average, predicted and observed diversity levels across the genome differ little across samples from reference populations.
+Given that the CEU and CHB samples are from bottlenecked out-of-Africa populations and their mutation rate and DFE estimates are similar, this is an empirical demonstration that our model is fairly robust to violations of the constant population size assumption of the theory.
+
+However, we note a few large (tens of megabases) regions with systematically poorer fit
 (Supplementary Materials Figure \ref{supp:spatial-residuals}). In Figure
 \ref{fig:figure-5}A we see one such region on the short arm of chromosome 2,
-from 30 Mbp to 60Mbp. Interestingly, predicted diversity closely follows the
-peaks and troughs of this region. We note that a small region within this
-stretch had been found by a genome-wide scan for associative overdominance
-\parencite{Gilbert2020-aw}.
-
-Another model diagnostic is to inspect whether observations are systematically
-different from predictions. We confirm the finding of \textcite{Murphy2022-sj}
-that regions predicted to experience little reduction in diversity due to
-background selection (i.e. $B\approx 1$) have higher diversity than predicted
-(Figure \ref{fig:figure-5}B, orange line). They suggested that this could
-reflect ancient introgression between archaic humans and ancestors of
-contemporary humans. Despite this aberration, the variance around observed and
-predicted diversity levels falls very close to what we would expect under the
-theoretic coalescence-noise-only expectation ($R_\text{coal}^2$).
-
-Earlier, we found that the goodness-of-fit and DFE estimates of our
-feature-based annotation models are sensitive to how basepairs are grouped into
-feature classes. Our models are in essence fitting a single distribution across
-sites within a feature class, yet there may be local heterogeneity among sites
-that is poorly captured by this approach. We investigated this possibility by
-looking for lingering selection signal in our model residuals. First, we
-inspected whether there was a relationship between the fraction of CADD 2\% and
-6\% basepairs and the residual across megabase windows (Supplementary Figure
+from 30 Mbp to 60 Mbp. Interestingly, predicted diversity closely follows the
+peaks and troughs of this region, however, predicted diversity is lower than observed.
+We note that a small region within this stretch had been found by a genome-wide scan for associative overdominance \parencite{Gilbert2020-aw}.
+We further investigate this by inspecting whether observations are systematically different from predictions.
+We confirm a finding of \textcite{Murphy2022-sj} that regions predicted to experience little reduction in diversity due to background selection (i.e.
+$B\approx 1$) have higher diversity than predicted (Figure \ref{fig:figure-5}B, orange line).
+\textcite{Murphy2022-sj} suggested that this could reflect ancient introgression between archaic humans and ancestors of contemporary humans.
+Despite the prediction error in this region, the variance around observed and predicted diversity levels falls very close to what we would expect under the theoretic coalescent-noise-only expectation ($R_\text{coal}^2$).
+
+As DFE heterogeneity within a class of sites may be poorly fit by our model,
+we looked for unaccounted selection in our model residuals. First, we
+inspected whether there was a relationship between the fraction of CADD 2\% and 6\% basepairs and the residual across megabase windows (Supplementary Figure
 \ref{suppfig:resid-cadd}), finding a negative significant relationship in
-both cases. Moreover, our model over-predicted diversity in windows containing
-more CADD 2\% basepairs moreso than CADD 6\%, consistent with heterogeneity in
+both cases. CADD 2\% was used in this case to search for a residual signal from highly-constrained regions. Moreover, our model over-predicted diversity in windows containing
+more CADD 2\% basepairs than CADD 6\%, consistent with heterogeneity in
 site pathogenicity being poorly fit by our model. However, the total residual
 variance explained is $R^2 = 0.3\%$ and $R^2 = 0.9\%$ for the CADD 2\% and 6\%
-tracks respectively, suggesting only modest amount of selection signal remains.
+tracks respectively, suggesting only a modest amount of selection signal remains within the CADD annotations.
 
@@
 soft and hard sweeps detected using a machine learning approach
-\parencite{Schrider2017-yx,Kern2018-jd}, we tested whether the residuals of the
+\parencite{Schrider2016-xn, Schrider2017-yx}, we tested whether the residuals of the
 CADD 6\% model containing sweeps were systematically different than those not
@@
 Materials Figure \ref{suppfig:resid-sweeps}; Kolmogorov--Smirnov p-value =
-0.71).
+0.71). The same was true if we looked at hard or soft sweeps individually as a class.
 
@@
 estimates the posterior distribution over LoF fitness costs from the observed
-dearth of LoF mutations per gene, and thus is an independent methodological
-approach than ours. We averaged the estimated LoF fitness costs across genes
+dearth of LoF mutations per gene, and thus is an independent
+approach to assess the strength of purifying selection. We averaged the estimated LoF fitness costs across genes
 for each of our megabase windows, and plotted our residuals against these
@@
 
-\subsection*{Predicted Substitution Rates}
+\subsubsection*{Predicted Substitution Rates Indicate Potential Model Misspecification}
 
 \begin{figure}[htbp] \centering
-    \includegraphics[width=\textwidth]{figures/figure_7_mod.pdf}
-
-    \caption{Predicted versus observed divergence along the human lineage.
-        Black points are the PhyloFit divergence rate estimates per features
+    \includegraphics[width=\textwidth]{figures/figure_7.pdf}
+
+    \caption{The divergence implied from predicted substitution rates under the B' model versus observed divergence along the human lineage.
+        Black points are the PhyloFit divergence rate estimates per feature
         (on x-axis). Line ranges are the implied divergences across a range of
-        human-chimpanzee divergence times of 6-12 Mya (using a generate time of
-        29 years). We show the predicted divergences for our Feature
+        human-chimpanzee divergence times of 6-12 Mya (using a generation time of
+        28 years). We show the predicted divergences for our Feature
         (turquoise) and PhastCons priority (green) annotation models.
         Additionally, we show the predicted PhastCons region divergences when
-    local-rescaling is applied (blue; we omit other locally-rescaled predictions
+    local rescaling is applied (blue; we omit other locally rescaled predictions
 since these to not differ substantially).}
@@
 
-Since our B' method relies on theory that jointly predicts the deleterious
-substitution rate, we can also check whether our model produces substitution
-rate predictions in agreement with observed divergence levels across features.
+Since our B' method also predicts deleterious substitution rates ($\widehat{R}$) for each feature class, it allows for another check of model sufficiency by comparing the predicted substitution rates to observed levels of divergence.
 We estimated sequence divergence on the human lineage using a multiple
 alignment of five primates for each feature in our feature-based models
-(Methods Section \nameref{sec:methods-sub}). We compared these to the predicted
-substitution rates per feature, averaging over all segments in the genome.
-Since our simulations show that mutation rate estimates can be biased, we
-predicted substitution rates under a fixed mutation rate of $\mu = 1.5 \times
+(Methods Section \nameref{sec:methods-sub}).
+We compared these to the predicted substitution rates per feature, averaging over all segments in the genome.
+Since our simulations show that mutation rate estimates can be biased, we predicted substitution rates under a fixed mutation rate of $\mu = 1.5 \times
 10^{-8}$. Fixing the mutation rate also allows us to more easily compare the
 predictions across our feature-based models. Unfortunately, a careful
-comparison between our predictions and observed divergence rates is hindered by
-considerable uncertainty in the rate of the molecular clock, generation times,
-and the human-chimpanzee divergence time. We assume a generation time of 29
-years, and calculate the sequence divergence implied by our predicted
+comparison between our predictions and observed divergence rates is hindered by considerable uncertainty in the rate of the molecular clock, generation times, and the human-chimpanzee divergence time.
+We assume a generation time of 28 years \parencite{Fenner2005-yi}, and calculate the sequence divergence implied by our predicted
 substitution rates over a range of divergence times, from 6 Mya to 12 Mya
@@
 
-We find that our predicted substitution rates are qualitatively consistent with
-the observed divergence along the human lineage for all features except
-PhastCons (Figure \ref{fig:figure-6}). As expected, the predicted substitution
-rates in features unlikely to be under selective constraint (introns and UTRs,
-and the ``other" feature) are very close to the mutation rate. Throughout, we
-report our substitution rates as a percent relative to the total mutation rate,
-$\mu$ (here fixed to $1.5 \times 10^{-8}$). In our Feature Priority model,
-coding sequences are predicted to have a substitution rate of 41.20\% of the
-mutation rate, introns and UTRs 94.71\%, coding sequences 0\%, and the ``other"
-feature 99.98\%. For comparison, the substitution rates along the human lineage
+We find that predicted substitution rates are qualitatively consistent with the observed divergence along the human lineage for all features except
+the PhastCons regions (Figure \ref{fig:figure-6}).
+As expected, the predicted substitution rates in features under reduced selective constraint (introns and UTRs, and the ``other" feature) are very close to the mutation rate.
+Throughout, we report our substitution rates as a percent relative to the total mutation rate, $\mu$ (here fixed to $1.5 \times 10^{-8}$).
+In our Feature Priority model, coding sequences are predicted to have a substitution rate of 41.20\% of the mutation rate, introns and UTRs 94.71\%, PhastCons regions 0\%, and the ``other" feature 99.98\%.
+For comparison, the substitution rates along the human lineage
 (as a proportion to the substitution rate in putatively neutral regions) are
@@
 in PhastCons regions. The large discrepancy between predicted and observed
-PhastCons substitution rates is driven by our DFE estimates suggesting that the
-bulk of mass is on selection coefficients greater than $10^{-3}$, which have no
-chance of fixation in a population of $N_e \approx 10,000$. We note that our
-DFE estimates are qualitatively similar to those inferred using the classic BGS
-model; thus this DFE-substitution rate contradiction is not unique to our
-estimates, and could indicate a potential model misspecfication problem.
-
-\section*{Locally-rescaled Fits}
-
-Given the observed DFE-substitution rate contradiction and that our simulations
-indicated that $B(x)$ is more accurately predicted when we use local-rescaling,
-we modified our composite-likelihood method to so that it can be run a second
-time, on B' maps locally-rescaled by the predicted $\widehat{B}(x)$ from the
-initial fit. Intuitively, this is based on the notion that if a neutral allele
-experiences an effective population size of $B(x)N$, so too should a selected
-allele, and this should be considered in how the SC16 equations are solved.
+PhastCons substitution rates is driven by our DFE estimates suggesting that the bulk of mass is on selection coefficients greater than $10^{-3}$, which have no chance of fixation in a population of $N_e \approx 10,000$.
+We note that our DFE estimates are qualitatively similar to those inferred using the classic BGS model, so the disagreement between observed divergence and predicted substitution rates could indicate a potential model misspecification problem.
+
+\subsubsection*{Possible Signal of Selective Interference}
+
+Given the prediction error for substitution rates in highly-conserved regions and that simulation indicates that $B(x)$ is more accurately predicted when we use local rescaling, we modified our composite-likelihood method so that it can be run a second time, on B' maps locally rescaled by the predicted $\widehat{B}(x)$ from the initial fit.
+Intuitively, this is based on the notion that if a neutral allele experiences a fitness-effective population size of $B(x)N$, so too should a selected allele, and this should be considered in how the SC16 equations are solved.
 
@@
 selection coefficient grid; in all other cases, non-strongly estimates were
-neutral ($s=10^{-8}$). Fourth, predicted pairwise diversity is nearly identical
-to our original, non-locally-rescaled fit (see Supplementary Materials Figure
-\ref{suppfig:rescale}). Finally, since local-rescaling increases the DFE mass
-over $s < 10^{-3}$, mutations in PhastCons regions now have the possibility of
-fixation. We find that local-rescaling the PhastCons Priority, leads the
-predicted substitution rates in PhastCons regions to be much closer to observed
-levels (blue line, Figure \ref{fig:figure-6}).
-
-Finally, we note an important caveat about this analysis. Since local-rescaling
-is done using the first round of maximum likelihood estimates, there is a
-possibility of some circularity, since the $B(x)$ at this position includes the
-contribution of the focal segment that is being rescaled. Ideally, one would
-exclude this segment's contribution to $B(x)$; however, this is computationally
-unfeasible. However, two observations indicate our findings here are relatively
+neutral ($s=10^{-8}$).
+Fourth, predicted pairwise diversity is nearly identical to our original, non-rescaled fit (see Supplementary Materials Figure \ref{suppfig:rescale}).
+Finally, since local rescaling increases the DFE mass over $s < 10^{-3}$, mutations in PhastCons regions now have the possibility of fixation.
+We find that local rescaling the PhastCons Priority, leads the predicted substitution rates in PhastCons regions to be much closer to observed levels (blue line, Figure \ref{fig:figure-6}).
+
+Finally, we note an important caveat about this analysis.
+Since local rescaling is done using the first round of maximum likelihood estimates, there is some possibility of statistical ``double dipping'', since the $B(x)$ at this position includes the contribution of the focal segment that is being rescaled.
+Ideally, one would exclude this segment's contribution to $B(x)$; however, this is computationally unfeasible.
+However, two observations indicate our findings here are relatively
 robust despite this limitation. First, the results do not change based on
-whether the $B(x)$ is averaged at the 1 kbp level or at the megabase scale; for
-the latter, a single segment makes little contribution to the average. Second,
-we investigated the extent to which local-rescaling modified the B' maps across
-selection parameters. We find minor differences between the locally-rescaled
-and standard B' maps in the nearly-neutral domain ($0.2 \le 2N_e s \le 2$); the
-maps were identical under strong selection ($2N_e s= 20$) as expected
-(Supplementary Materials Figure \ref{suppfig:rescale-b}). Moreover, the
-correlations between the standard and local-rescaled B' maps across the genome
-are high (100\% for $2N_e s = 20$, 96.5\% for $2N_e s = 2$, and 60.21\% for
-$2N_e s = 0.2$). The overall realized effect of local-rescaling is to just
-alter how deep the ``U" is in the relationship between the reduction factor $B$
-and the selection coefficient (Supplementary Materials Figure
+whether the $B(x)$ is averaged at the 1 kbp level or at the megabase scale; for the latter, a single segment makes little contribution to the average.
+Second, we investigated the extent to which local rescaling modified the B' maps across selection parameters.
+We find minor differences between the locally rescaled and standard B' maps for fixed selection coefficients (i.e.
+before model fitting) in the nearly-neutral domain ($0.2 \le 2N_e s \le 2$).
+Additionally, the locally rescaled and standard maps are identical under strong selection ($2N_e s= 20$) as expected (Supplementary Materials Figure \ref{suppfig:rescale-b}).
+Moreover, the correlations between the standard and locally rescaled B' maps across the genome are high (100\% for $2N_e s = 20$, 96.5\% for $2N_e s = 2$, and 60.21\% for $2N_e s = 0.2$).
+The overall realized effect of local rescaling is to just alter how deep the ``U" is in the relationship between the reduction factor $B$ and the selection coefficient (Supplementary Materials Figure
 \ref{suppfig:rescale-effects}).
 
-Overall, this suggests that models that fit the linked selection are worryingly
-sensitive to the theoretic B' values in the $2Ns \le 1$ domain. The fact that
-predicted diversity differs little between standard and locally-rescaled B'
-methods indicates there may not be enough information in the linked selection
-signal alone to differentiate when interference is occurring or the causes of
-fitness variance along the genome. Moreover, local-rescaling turns out to only
-slightly alter the B' maps, yet significantly modifies the DFE estimates. This
-brings the deleterious substitution rate in agreement with observations (since
-this is predicted with a fixed $\mu = 1.5 \times 10^{-8}$; however, the maximum
-likelihood estimate of mutation is implausibly high. This suggests either the
-local-rescaling approximation to interference is not suitable (though
-chromosome-wide simulations show it results in lower error), or there is
-another selection process contributing to total fitness variance.
+Overall, this suggests that models of the signal of linked selection are worryingly sensitive to the theoretic B' values in the $2Ns \le 1$ domain.
+The fact that predicted diversity differs little between standard and locally rescaled B' methods indicates there may not be enough information in pairwise diversity alone to differentiate when interference is occurring or the causes of fitness variance along the genome.
+Moreover, local rescaling turns out to only slightly alter the B' maps, yet significantly modifies the DFE estimates.
+This brings the deleterious substitution rate in agreement with observations (since this is predicted with a fixed $\mu = 1.5 \times 10^{-8}$; however, the maximum likelihood estimate of mutation is implausibly high.
+This suggests either the local rescaling approximation to interference is not suitable (though our chromosome-wide simulations show locally rescaled B' maps are close to the reductions observed from simulations), or that the deleterious mutations-only model does not adequately describe the processes generating fitness variance.
 
@@
 
-Linked selection processes alter genetic diversity on multiple scales, creating
-patterns based on their strength relative to the stochasticity of coalescence
-times and mutations. Sweeps typically leave signals on shorter genomic scales
-(i.e. kilobases; \cite{Akey2004-zg}) which combined with the paucity of
-adaptive substitutions in humans suggests positive selection is not generating
-the apparent large, megabase-scale linked selection signal
-\parencite{Hernandez2011-gs,Murphy2022-sj}. By contrast, there is reasonable
-evidence that negative selection against new mutations occurring in
-highly-conserved sequences is generating large-scale chromosomal patterns of
-diversity.
-
-Population genetic theory has been productively applied to fit this linked
-selection signal, producing plausible estimates of the strength of selection at
-these conserved regions and mutations rates consistent with more precise
-pedigree-based estimates. However, these models rely on the canonical
-background selection model, which due to its assumption that mutations are
-sufficiently deleterious they cannot fix, has constrained statistical methods
-to only fit background selection models in regions \emph{a priori} under
-constraint. In reality, the relationship between diversity levels and the
-strength of selection at deleterious mutations is U-shaped, which implies there
-could be more uncertainty than previously appreciated in the distribution of
-weak and strongly deleterious mutations.
-
-We sought to address this uncertainty in this work, by leveraging a different
-class of linked selection models based on modeling the equilibrium fitness
-variance \parencite{Santiago2016-mu}. At its core, the theory underlying this
-model states that the linked selection signal is a function of how the fitness
-variance is distributed along the recombination map of the genome
-\parencite{Santiago1998-bs}. We fit a specific process model for this fitness
-variance that supposes all fitness variance is the result of selection against
-new deleterious mutations \parencite{Santiago2016-mu}. Unlike classic
-background selection theory, the SC16 model explains equilibrium fitness
-variance across all selection coefficients by jointly predicting another
-central quantity in evolutionary genomics: the deleterious substitution rate in
-conserved features. This model is more general than classic background
-selection theory; background selection theory is analytically a special case
-when the deleterious substitution rate is zero.
-
-Generally, our method lead to better fits to data that those based on classic
-BGS, as measured by predicted out-sample diversity. Furthermore, unlike
-BGS-based methods, our model is capable of fitting weak selection; we find
-evidence of this since we can robustly estimate the DFE for weakly constrained
-features like introns and UTRs. Yet even with the possibility of fitting weak
-selection, our main model fits are in agreement with the prior results from
-BGS-based methods. Both mutation rate estimates and DFE estimates are
-consistent with prior work. This, at first glance, suggests further
-confirmation that negative selection against new deleterious mutations is the
-dominant mode of linked selection in the genome.
-
-Since the SC16 theory calculates fitness variance by jointly predicting the
-deleterious substitution rate, this provides another avenue to verify or refute
-our models. We find that predicted substitution rates for highly-conserved
-PhastCons features are in strong disagreement with observed substitution rates
-along the human lineage: the DFE-substitution rate contradiction. This is a
-consequence of our DFE estimate for PhastCons regions containing little mass
-over weakly deleterious and neutral selection coefficients --- a characteristic
-of DFE estimates from other work too \parencite{Murphy2022-sj}. One clue to
-this discrepancy is that our simulation results indicate an appreciable level
-of error between observed reductions and the B' maps in the weak selection
-domain of $2N_e s \approx 1$.
-
-We hypothesized that this could be because in this domain, a segment under
-selective constraint experiences a local draft-effective population size of
-$B(x)N$, and not just $N$. This local draft-effective population size is
-induced by selection at other segments that is not being taken into account by
-classic BGS theory or our standard model. When we experimented by fitting our
-model and then using the predicted reduction map to local-rescale $N_e$ to the
-draft-effective population size $N_d = \widehat{B}(x)$, we find the
-disagreement between predicted and observed substitution rates disappears. This
-is expected, since locally-rescaled fits result in DFE estimates to have an
-appreciable mass on weakly deleterious selection coefficients. While this
-brings substitution rates into agreement, we encounter a similar problem found
-by \textcite{McVicker2009-ax}: the estimated mutation rate is too high. We
-emphasize that although there are several caveats with the local-rescaling
-approach, in practice it is altering the B' maps only slightly in the domain of
-$2N_e s \approx 1$ (deepening the bottom of the ``U"). Even if the
-local-rescaling approximation to selective interference is incorrect (though
-simulations indicate it does reduce error), these results suggest parameter
-estimates are extremely sensitivity to how accurate the B' maps are in this
-domain. Moreover, we find that predicted diversity differs little across DFE
-and mutation rate estimates, suggesting there may be limited information in the
-pairwise diversity linked selection signal to differentiate models. While
-nucleotide diversity is a less informative statistic than the full site
-frequency spectrum, it is much easier to derive theoretic models of how linked
-selection alters expected pairwise coalescence times than to describe the
-entire genealogy. While one could imagine using more informative SFS-based
-models to infer selection parameters, accounting for selective interference in
-these models would be difficult.
-
-Unfortunately, this work opens up more questions about the nature of selection
-in humans that it answers. The high estimate of mutation rates could suggest
-that negative selection is not the only source of fitness variance generating
-the genome-wide linked selection signal in humans. Positive selection on
-polygenic variation along the genome could be another source of fitness
-variance contributing to the weak selection signal. Alternatively, negative
-selection could be the main source of fitness variance, but the complexity of
-selective interference may be poorly approximated by the local-rescaling
-approach. Another possibility is that our assumption throughout of additive
-effects may lead to biases given deleterious mutations likely recessive
-effects. As we find evidence of strong selection against loss-of-function in
-our model residuals, it is also possible that the bulk of fitness variance
-\emph{is} due to purifying selection, but our model is unable to account for
-local strong effects since the DFE is estimated per feature. Overall,
-understanding how the sources of human fitness variance shape the striking
-patterns of diversity along the human genome still remains an open riddle.
-
-\subsection*{Acknowledgments}
-
-We would like to thank Nate Pope and Rasmus Nielsen for helpful advice, and
-Martin Kircher for providing special CADD tracks. This research was supported
-by NIH awards R35GM148253 and R01HG010774 to ADK
+New mutations at functionally important regions of the genome are a major source of fitness variation in natural populations, as the vast majority of such mutations are deleterious.
+Purifying selection, working to remove these deleterious variants, perturbs genealogies at linked sites, creating large-scale patterns in genomic diversity.
+While this has been recognized for decades (e.g., \textcite{Nordborg1996-nq,Hudson1995-xc}), the availability of genomic data allows for methods to estimate the degree to which purifying selection shapes genomic variation and at what scale.
+
+Accordingly, there have been a number of recent efforts to fit parametric models of linked selection to polymorphism and divergence data in Drosophila \parencite{Elyashiv2016-vt} and humans \parencite{McVicker2009-ax, Murphy2022-sj}.
+These efforts have yielded reasonable estimates of the strength of selection on new mutations as well as provided mutation rate estimates that largely agree with pedigree-based estimates.
+However, previous methods have relied on the canonical background selection model, which assumes that mutations are sufficiently deleterious such that they cannot fix.
+Consequently, statistical methods using the BGS model should only be expected to fit well when some regions are \emph{a priori} under strong selective constraint.
+In reality, the relationship between neutral diversity levels and the strength of selection from purifying selection in linked regions is U-shaped, which implies there could be more uncertainty than previously appreciated in the distribution of weak and strongly deleterious mutations.
+
+In this work, we developed and fit a different class of linked selection models based on the equilibrium fitness variance \parencite{Santiago1998-bs,Santiago2016-mu}.
+Fundamentally, we model the reduction in diversity as a
+function of how additive fitness variance is distributed along the recombination map of the genome \parencite{Santiago1998-bs}. We fit a specific model for this fitness variance that supposes all 
+variation is the result of selection against new additive deleterious mutations \parencite{Santiago2016-mu}. 
+Unlike classic background selection theory \parencite{Charlesworth1993-gb,Nordborg1996-nq,Hudson1995-pt,Hudson1995-xc}, 
+the SC16 model explains equilibrium fitness variance across all selection coefficients by jointly predicting another central quantity in evolutionary genomics: the 
+substitution rate of deleterious alleles.
+
+Our method has at least four improvements over previous whole-genome linked selection methods based on the BGS model.
+First, our model leads to better fits to data than those based on classic BGS, as measured by predicted out-of-sample diversity.
+Second, unlike BGS-based methods, our model is capable of fitting weak selection.
+When regions under weak or little selective constraint are included in methods using classic BGS, parameter estimates can become severely biased.
+By contrast, we have demonstrated via simulation that our method can estimate the strength of selection even for weakly constrained features (e.g.
+introns and UTRs), as well as remaining unannotated regions of the genome.
+Third, fitting our model produces a simultaneous prediction of substitution rates, which can be compared to observed divergence rates.
+Finally, the effect of selective interference can be approximated by locally rescaling the B' maps, which our forward simulations show reduce prediction error of genome-wide diversity levels.
+
+Even though our model is able to fit weak selection, our initial estimates of mutation rate and DFEs were consistent with prior work \parencite{Murphy2022-sj}.
+This, at first glance, suggests further confirmation that strong purifying selection is the dominant mode of linked selection in the genome.
+However, we find that predicted substitution rates for highly-conserved PhastCons features disagree with observed rates of divergence along the human lineage.
+This disparity between observed divergence and predicted substitution rates is likely a consequence of our DFE estimate for PhastCons regions containing little mass over weakly deleterious and neutral selection coefficients that would have some possibility of fixation --- a characteristic of DFE estimates from other work too \parencite{Murphy2022-sj}.
+
+Our simulation results reveal another possible source of disagreement: in the weak selection domain of $2N_e s \approx 1$ there is an appreciable level of disagreement between theory and simulation.
+We hypothesize that this could be because as $2N_e s$ approaches $1$, a segment under
+selective constraint experiences a local fitness-effective population size of $B(x)N$, and not just $N$.
+This local fitness-effective population size is induced by selection at other segments that is not being taken into account by classic BGS theory or our standard SC16 model.
+When we experimented by fitting our model and then using the predicted reduction map to locally rescale $N_e$ to the fitness-effective population size $N_f = \widehat{B}(x)$, we found the disagreement between predicted and observed substitution rates disappears.
+This is expected since locally rescaled DFE estimates have an appreciable mass on weakly deleterious selection coefficients, contrary to the standard fits.
+
+This pattern is consistent with a scenario where the same selection processes that reduce diversity over long stretches along the chromosome also decrease the efficacy of purifying selection.
+This idea has been proposed before in an extension to the McDonald--Kreitman test that accounts for how
+background selection can bias estimates of the proportion of adaptive substitutions 
+\parencite{Uricchio2019-wp}.
+While our simulation results indicate local rescaling reduces error in the weak selection domain, it is worth noting some caveats about this approach.
+First, local rescaling is only an approximation to selective interference; as our simulations show, this approximation reduces error in the predicted reduction $B(x)$, but this may not fully account for how negative linkage disequilibrium builds up and reduces fitness variance.
+A benefit of the SC16 approach is that the equations
+can be solved with a locally rescaled effective population that approximates this process.
+Second, there is the possibility of circularity, since a preliminary fit must be made to estimate $B(x)$, which is then
+used to re-solve the SC16 equations with a local fitness-effective population size.
+
+Despite these caveats, the local rescaling approach suggests selective interference could alter inferences about the DFE and bring predicted substitution rates into agreement with observed divergence rates.
+However, these results also demonstrate that parameter estimates are extremely sensitive to how accurate the theoretic B' maps
+are in this domain.
+Moreover, we find that predicted diversity differs little across DFE and mutation rate estimates, suggesting there may be limited information in pairwise diversity to differentiate between models, thus inclusion
+of allele or haplotype frequency information might be informative in the future.
+While local rescaling brings substitution rates into agreement, it also re-introduces a similar problem found by \textcite{McVicker2009-ax}: the
+estimated mutation rate is too high. While our mutation rate includes point mutations as well as all other forms of 
+deleterious variation (e.g. insertions/deletions, copy number variants, etc.), our estimations suggest exceedingly high 
+rates of deleterious variation per generation.
+
+Examination of the local residuals of our predicted diversity levels demonstrated no systematic effect of previously identified hard and soft selective sweeps \parencite{Schrider2017-yx}. 
+This result echoes what has been observed in previous efforts to look at genome-wide patterns of 
+linked selection \parencite{Hernandez2011-gs, Murphy2022-sj}, and suggests that the scale of perturbation due to selection sweeps is more restricted (e.g.
+at the kilobase scale, \cite{Akey2004-zg}) than the scale at which we are modeling variation.
+Taken together this suggests that selective sweeps are not likely responsible for shaping the
+majority of variance in large-scale patterns of chromosomal variation in humans.
+
+Given that our model is essentially parameterized by levels of fitness 
+variance along the genome, the high estimates of mutation rate could suggest that purifying selection is not the only source of fitness variance generating the genome-wide linked selection signal in humans.
+Selection on polygenic traits could be another source of fitness variance.
+Alternatively, purifying selection could be the main source of fitness variance, but the complexity of selective interference may be poorly approximated by the local rescaling approach.
+Another possibility is that our assumption throughout of additive effects may lead to biases given deleterious mutations likely recessive effects.
+Future theoretic work, as well as realistic across-species forward simulations of multiple modes of linked selection (e.g.
+\cite{Rodrigues2023-rj}) are needed to fully disentangle these processes.
+As we find evidence of strong selection against loss-of-function in our model residuals, it is also possible that the bulk of fitness variance \emph{is} due to purifying selection, but our model is unable to account for strong heterogeneity in the DFE per annotation class.
+
+Moving forward it remains a central goal to understand how the sources of fitness variation shape the striking patterns of diversity along the human genome.
+Our work embeds this question in the quantitative genetic framework that is more accurate and flexible than proceeding models, but there is much work yet to do to incorporate important population genetic features such as dominance effects and selective interference.
+Overall, the complex interplay of mutation, selection, drift, and interference may confound our understanding of selection in the human genome for some time.
+
+\section*{Acknowledgments}
+
+We would like to thank Doc Edge, Ben Good, Taylor Kessinger, Graham McVicker, Priya Moorjani, Rasmus Nielsen, Guy Sella, Joshua Schraiber, and Peter Sudmant for helpful discussions, and Martin Kircher for providing modified CADD tracks.
+We thank Graham Coop, Matt Hahn, Nate Pope, and Enrique Santiago for comments on the manuscript.
+This research was supported by NIH awards R35GM148253 and R01HG010774 to ADK
 
@@
 Our software \texttt{bgspy} first calculates the equilibrium additive genic
-fitness variation $\widetilde{V}_a$ and ratchet rate $\widetilde{R}$ for each
+fitness variation $\widetilde{V}_a$ and deleterious substitution rate $\widetilde{R}$ for each
 user-specified segments in the genome. These equilibria are calculated across
@@
   \label{eq:method-2eqns}
-  {N}_{d} &= N \exp \left( -V_a \frac{Q^2(m_i, s_j)}{2} \right) & \text{\emph{effective population equation}} \\
-  R &= \frac{4N_dU s_j}{\exp(4 N_d s_j)-1}  & \text{\emph{substitution equation}} 
+  {N}_f &= N \exp \left( -V_a \frac{Q^2(m_i, s_j)}{2} \right) & \text{\emph{effective population equation}} \\
+  R &= \frac{4N_fU s_j}{\exp(4 N_f s_j)-1}  & \text{\emph{substitution equation}} 
 \end{align}
@@
 Our analyses was conducted on the Yoruba (YRI), European (CEU), and Han Chinese
-(CHB) reference samples individuals from the high-coverage 1000 Human Genomes
+(CHB) reference sample individuals from the high-coverage 1000 Human Genomes
 data aligned to GRCh38/hg38 \parencite{Byrska-Bishop2022-tn}. Since nucleotide
@@
 
-\subsection*{Substitution Rate Estimation and Prediction}
+\subsection*{Substitution Rate Prediction and Divergence Estimates}
 \label{sec:methods-sub}
 
+Substitution rates were predicted by resolving Equations \eqref{eq:main_eqns} for the given estimated product between mutation rate and DFE weight, $w_{i,j} = \mu W_{i,j}$.
 We estimated the divergence along the human branch using PhyloFit
-\parencite{Siepel2004-ql} run on the Multiz alignments
-\parencite{Blanchette2004-bb} of humans and four other primates (\emph{Pongo
+\parencite{Siepel2004-ql} run on a subset of the UCSC 17-way Multiz alignments
+\parencite{Blanchette2004-bb} consisting of humans and four other primates (\emph{Pongo
 abelii}, \emph{Pan troglodytes}, \emph{Pan paniscus}, \emph{Gorilla gorilla}).
@@
 (\url{https://github.com/vsbuffalo/bprime}). 
-
 \subsection*{Data Availability}
@@
 (\url{https://github.com/vsbuffalo/bprime}). The main model fits are available
-as Python Pickle objects on FigShare (XXX).
+as Python Pickle objects on Data Dryad repository (\url{https://doi.org/10.5061/dryad.qnk98sfnv}).
 

### 2023-09-08 manuscript/supplementary.tex
data dryad
new updated ms from overleaf

@@
 \RequirePackage{xr}
+
+\newcommand*{\addFileDependency}[1]{
+\typeout{(#1)}
+
+\IfFileExists{#1}{}{\typeout{No file #1.}}
+}\makeatother
+
+\newcommand*{\myexternaldocument}[1]{%
+\externaldocument{#1}%
+\addFileDependency{#1.tex}%
+\addFileDependency{#1.aux}%
+}
+
+\myexternaldocument{manuscript}
+
 %\RequirePackage[osf]{mathpazo}
@@
 % for cross references
-\externaldocument{manuscript}
+%\externaldocument{manuscript}
 
 % line numbers:
-\RequirePackage{lineno}
-%\modulolinenumbers[5]
-\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{black}}
+\RequirePackage[left, modulo]{lineno}
+\modulolinenumbers[2]
+\linenumbers
+
+% spacing
+\RequirePackage{setspace}
+\onehalfspacing
 
@@
 can be partitioned into the three sources of stochasticity: the random
-associations with fitness backgrounds (i.e. \emph{draft}), the non-heritable
+associations with fitness backgrounds, the non-heritable
 randomness in family size, and the Mendelian noise from heterozygotes
@@
 The $Q_t^2$ terms represent the inflation of variance due to autocovariance in
-frequency change across generations due to weak draft. Incorporating the impact
-of time-invariant autocorrelation created by draft as a rescaled rate of drift
+frequency change across generations due to linked polygenic selection. Incorporating the impact
+of time-invariant autocorrelation created by linked selection as a rescaled rate of drift
 is analogous to using the Green--Kubo relation to find the transport
@@
 for the variance encodes the pairwise coalescent rate of the population through
-time under weak draft. Others have found that the genealogies under purifying
+time under weak polygenic linked selection. Others have found that the genealogies under purifying
 selection can be characterized by such a time-dependent effective population
@@
 
-\subsection{The Draft-Effective Population Size}
+\subsection{The Fitness-Effective Population Size}
 
 Now, we consider how to turn this expression for the variance in neutral allele
-frequency at time $t$ under weak draft into an expression for the
-\emph{draft-effective} population size analogous to drift-effective population
+frequency at time $t$ under weak linked selection into an expression for the
+\emph{fitness-effective} population size analogous to drift-effective population
 size given in Supplementary Equation \eqref{supp-eq:var-ne}. In an ideal
@@
 allele frequency change at some $t$ to define their effective population size
-at time $t$ (which we here are calling draft-effective population size,
+at time $t$ (which we here are calling fitness-effective population size,
 $N_{d,t}$),
@@
 
-In many cases, we care about the asymptotic $N_d$,
-
-\begin{align}
-    N_d := \lim_{t \to \infty} N_{d,t} \approx \frac{4N}{4 V_A Q_\infty^2 +V_n+2}.
+In many cases, we care about the asymptotic $N_f$,
+
+\begin{align}
+    N_f := \lim_{t \to \infty} N_{d,t} \approx \frac{4N}{4 V_A Q_\infty^2 +V_n+2}.
 \end{align}
@@
 
-\subsection{The Draft-Effective Population Size Under Polygenic Selection}
+\subsection{The Fitness-Effective Population Size Under Polygenic Selection}
 \label{supp:ne-polygenic}
@@
 $V_A$ and $1-\kappa$ can be worked out. Other authors have extended similar models to
-more elaborate bbreeding structures \parencite{Wray1990-zf,Woolliams1993-qo}.
+more elaborate breeding structures \parencite{Wray1990-zf,Woolliams1993-qo}.
 
@@
 
-\subsection{The Draft-Effective Population Size Under Strong Background Selection}
+\subsection{The Fitness-Effective Population Size Under Strong Background Selection}
 
@@
 
-\subsection{The Draft-Effective Population Size Under Weak and Strong Negative Selection}
+\subsection{The Fitness-Effective Population Size Under Weak and Strong Negative Selection}
 \label{supp:weak-strong}
@@
 weakly deleterious mutation. This fixation process is related to the Muller's
-Ratchet, the rate $R$ of fixation of deleterious mutations per generation per
+Ratchet in assexual populations, the rate $R$ of fixation of deleterious mutations per generation per
 region, since each fixation directly reduces the additive genic variance of the
@@
 and cross-moments of the fitness distribution under multiplicative selection in
-a haploid model. They find, the haploid rate of the ratchet is $r = u - s v_n$
+a haploid model. They find the haploid rate of the ratchet is $r = u - s v_n$
 (equation 13.3, \cite{Higgs1995-xc}), where we use lowercase to distinguish
@@
 $R=0$ and the additive fitness variance is identical to that under strong
-background selection, $V_A = Us = V_\text{BGS}$ (equation \ref{eq:va_bgs}).
+background selection, $V_A = Us = V_\text{BGS}$ (equation \ref{supp-eq:va_bgs}).
 We note that while we use $V_A$ here following \textcite{Santiago2016-mu}, 
@@
 asymptotic $N_{d,\infty}$. Consequently, Santiago and Caballero argue that the
-asymptotic reduction in draft-effective population size uses the $Q_\infty^2$ and is
+asymptotic reduction in fitness-effective population size uses the $Q_\infty^2$ and is
 given by the following system of non-linear equations,
@@
 The previous section outlined the basis of quantitative genetic linked
-selection models that predict the draft-effective population size under
+selection models that predict the fitness-effective population size under
 polygenic selection and linkage. Our work here only considers the contribution
-of strongly and weakly deleterious mutations to fitness variation. In this
+of deleterious mutations to fitness variation. In this
 section, we outline more details about the statistical models used to construct
@@
 
-\subsection{Parametric Forms of Negative Linked Selection Models}
+\subsection{Models of Linked Selection}
 
@@
 
-In practice, to fit negative selection models, we must discretize both $\mu(s,
+In practice to fit negative selection models we must discretize both $\mu(s,
 k(g))$ and $s$ in Supplementary Equation \eqref{supp-eq:b}. We define
@@
 Because it is computationally-intensive to numerically solve the system of two
-non-linear equations to calculate $b$ at each segment, these are pre-computed
+non-linear equations (Equation \ref{eq:main_eqns} in the main paper) to calculate $b$ at each segment, these are pre-computed
 for an $n_s$-element grid of selection coefficients and an $n_\mu$-element grid
@@
 equation 4 \parencite{Santiago2016-mu}. This equation considers how the
-inflation factor $Q_t^2$ that sets the draft-effective population size varies
+inflation factor $Q_t^2$ that sets the fitness-effective population size varies
 over $t$. This is because statistics such as heterozygosity depend on the
@@
 (which is approximately heterozygosity when $2N\mu \ll 1$) depends on the
-pairwise coalescence rate per generation, which is not constant under draft.
+pairwise coalescence rate per generation, which is not constant under linked selection.
 
@@
 
-\begin{align*}
-  n_T &= \frac{n(n-1)}{2} \\
-  n_S &= {c_1 \choose 2} + {c_2 \choose 2} \\
+\begin{align}
+  n_T &= \frac{n(n-1)}{2} \nonumber \\
+  n_S &= \binom{c_1}{2} + \binom{c_2}{2} \nonumber \\
   n_D &= n_T - n_S
-\end{align*}
+\end{align}
 %
@@
 per-site $\mathbf{Y}$ handles non-polymorphic sites and missing data.
-Non-polymorphic sites have $n_S = {n \choose 2}$ and $n_D = 0$, and missing
+Non-polymorphic sites have $n_S = \binom{n}{2}$ and $n_D = 0$, and missing
 data has $n_S = n_D = 0$.
@@
 
-\subsection{Ratchet Rate Prediction}
-
-After rescaling each segment $N_e$ by the local predicted reduction
-$\hat{B}(x)$, the ratchet rates and $B'$ values are re-calculated.
-
-The rescaled $B'$ calculation outputs a $n_\mu \times n_s \times S$
-multidimensional array $\mathbf{R}$ of ratchet rates per segment. These are
-rescaled by the segment lengths $\mathbf{l}$, giving the per-basepair mutation
-rate array. The per-basepair ratchet rate given maximum likelihood estimates of
-$\hat{\mu}$ and $\widehat{\mathbf{W}}$ is used to predicted the ratchet rate
-for each segment $g$. The DFE estimate for feature $k$ is a column vector of
-$\widehat{\mathbf{W}}$, i.e. $\widehat{\mathbf{w}}_k$
-
-\begin{align}
-  \lambda_D(g) = 
-\end{align}
-
-The predicted ratchet rate for segment $g$ $\lambda_d(g)$ is based on these
-values.
-
-\subsection{Numeric methods for calculating $B$}
-
-Depending on the number of features, this can be incredibly memory intensive.
-
-\subsection{Optimization}
-
-Optimization in non-linear regression is notoriously difficult both in general
-(Bates) and in the specific problem of estimating the effects of genome-wide
-selection \parencite{Murphy2022-sj}. We experimented with a variety of global
-and local optimization approaches and parameterizations, but found that
-
-\section{Human Genomic Data}
+\section{Analysis Details}
+
+\subsection{Human Genomic Data}
 
@@
 ratio estimator that is sensitivity to its denominator, we use the complete
-per-basepair genotype calls (gVCF) produced by Illumina's DRAGEN pipeline
-\parencite{Illumina_Inc2020-dk}. The original samples were sequenced to 30x by
+per-basepair genotype calls (gVCF) produced by Illumina's DRAGEN pipeline for the high-coverage 1000 Genome samples
+\parencite{Illumina_Inc2020-dk,Byrska-Bishop2022-tn}. The original samples were sequenced to 30x by
 the New York Genome Center \parencite{Byrska-Bishop2022-tn}. This allows for
@@
 
-The gVCFs were filtered using a custom Python tool, \texttt{gvcf2counts.py} (in
+The gVCFs were filtered using a custom Python tool, \texttt{gvcf2counts.py} (available in the repository associated with the manuscript at
 \path{tools/gvcf2counts.py}), which reads the gVCFs, filters them according to
@@
 regions \path{phastConsElements100way.bed.gz}]{Siepel2005-wh}. Then, we remove
-200 basepairs of ``slop" around each of these regions (i.e. \texttt{bedtools
+200 basepairs of buffer around each of these regions (i.e. \texttt{bedtools
 slop}). We show in Supplementary Figure \ref{suppfig:neutral-tracks} that
 average diversity (we only consider YRI samples for this analysis) is fairly
-insensitive to the slop width, but increasing slop widths around putatively
+insensitive to the buffer width, but increasing slop widths around putatively
 conserved segments does lead to higher diversity estimates. This is expected,
@@
 \begin{table}
-    \label{supp-table:masks}
 \centering
@@
 \end{tabular}
+\label{supp-table:masks}
 \end{table}
@@
 \ref{supp:filter} and \ref{supp:accessible}.
-
-\subsection{Annotation Models}
-
-CADD 6\%: 6.49\% of autosomal sequence.
-CADD 8\%: 8.56\% of autosomal sequence.
 
@@
 
-\subsection{Table of $R^2$ Values}
+\subsection{Table of \texorpdfstring{$R^2$}{R2} Values}
 \label{supp:tbl-r2}
@@
 
-\subsection{Table of $R^2$ Values (Strong Selection Grid)}
+\subsection{Table of \texorpdfstring{$R^2$}{R2} Values (Strong Selection Grid)}
 \label{supp:tbl-r2-strongsel}
@@
     the last rows are correct and due to rounding; see
-    \path{notebooks/main_fits.ipynb} in the GitHub repository for more
-information. $R_\text{IS}^2$ is the in-sample $R^2$, and $R_\text{LOCO}^2$ is
-the leave-one-chromosome-out $R^2$. }
+    the main analysis notebook in the GitHub repository for more information. $R_\mathrm{IS}^2$ is the in-sample $R^2$, and $R_\mathrm{LOCO}^2$ is the leave-one-chromosome-out $R^2$. }
 
@@
 
-\subsection{Derivation of Approximate Coalescence $R_\text{coal}^2$}
+\subsection{Derivation of Approximate Coalescence \texorpdfstring{$R_\text{coal}^2$}{R2}}
 \label{supp:r2-coal}
@@
 explore different ratios to assess sensitivity of this calculation
-(Supplementary Figure \ref{suppfig:r2_upper}).
+(Supplementary Figure \ref{suppfig:drift_r2}B).
 
@@
 nucleotide differences in genomic windows to the expected pairwise diversity
-under some evolutionary model. We only consider BGS models here, so the mean
-function for position $x$ is the product of the proportion by which BGS reduces
+under some evolutionary model. We only consider  models of purifying selection here, so the mean
+function for position $x$ is the product of the proportion by which selection reduces
 diversity at position $x$, $B(x)$, and genome-wide neutral diversity $\pi_0$,
@@
 
-$\Phi$ is the set of BGS parameters (i.e. the DFE matrix $\mathbf{W}$ and
+$\Phi$ is the set of parameters (i.e. the DFE matrix $\mathbf{W}$ and
 mutation rate $\mu$), and $\pi_0 = 4 N_e \mu$ is determined by the genome-wide
@@
 
+\subsection{Optimization and Diagnostics}
+\label{supp:optim}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/figure_feature_priority_yri_full_diag.pdf}
+
+  \caption{ An optimization diagnostic plot showing the 100 top optima. Heatmap
+  colors indicate probability mass for the DFE parameters (top four rows).
+  Estimates of $\pi_0$ and $\mu$ are shown in the bottom two rows. This is the
+optimization plot for YRI Feature Priority model.}
+  \label{suppfig:diag-plot}
+\end{figure}
+
+As described in the Methods section we use the BOBYQA algorithm implemented by
+the NLopt library for numeric optimization to find maximum likelihood estimates
+\parencite{Powell2009-jm,Johnson2007-tl}. The DFE is a simplex, and thus
+requires constrained optimization. We compared constrained optimization
+algorithms to an unconstrained softmax approach, finding the latter vastly
+outperformed the former. Bounds for $\pi_0$ and $\mu$ can be set manually in
+\texttt{bgspy}; our main models used $10^{-8} \le \mu \le 5 \times 10^{-8}$ and
+$0.0005 \le \pi_0 \le 0.008$. All main model fits were not at bounds. For
+locally rescaled fits, we found our $\widehat{\mu}$ estimate was continually
+the upper bound for the PhastCons Priority model; for these fits we increased this first to $\mu = 8 \times
+10^{-8}$ and then again to $\mu = \times 10^{-7}$, find it still always reached the upper bound. This did not occur for CADD 6\% locally rescaled models.
+
+We only show one optimization diagnostic plot here (Supplementary Figure
+\ref{suppfig:diag-plot}), but the can be easily calculated from the model
+pickle files in the Data Dryad repository (\url{https://doi.org/10.5061/dryad.qnk98sfnv}). In one case, our optimization
+diagnostics indicated that the best optima across 10,000 random starts was one
+of few of the top 100 hits that were hitting either an alternate likelihood
+mode, or the optimization routine had numeric issues. In this one case, we
+removed the top hit and used the second, which was the consensus among the
+other top 100 optima.
+
 \subsection{The scale of processes}
@@
 
-\subsection{Out-Sample Error Estimation and $\widehat{R}_\text{LOCO}^2$}
+\subsection{Out-Sample Error Estimation and \texorpdfstring{$\widehat{R}_\text{LOCO}^2$}{R2}}
 
 To estimate out-sample prediction error, we fit each model leaving one
-chromosome out (LOCO) and calculating the out-sample prediction error on
+chromosome out (LOCO) and calculating the out-of-sample prediction error on
 these excluded chromosomes. Each fit used 10,000 random starting positions as
@@
 \begin{figure}[htbp]
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/phastcons_diag_plot.pdf}
+
+  \caption{An optimization diagnostic plot showing an example of the bimodal likelihood optima, or a numeric optimization issue.
+The first four rows are the DFE estimates, where color indicates the parameter
+  value (brighter colors closer to 1, darker colors closer to 0). 
+  The $\pi_0$ and $\mu$ estimates are shown in the bottom two 
+  rows. Note that different parameter combinations co-occur, 
+  indicating in the top 100 optima, there is an apparent other
+  mode that varies with mutation rate.}
     \label{suppfig:diag-plot-mode}
-  \centering
-  \includegraphics[width=\textwidth]{figures/supplementary/phastcons_diag_plot.pdf}
-
-  \caption{An optimization diagnostic plot showing an example of the bimodal likelihood
-  optima, or a numeric optimization issue.}
 \end{figure}
@@
 looked like they were due to rare aberrant optima. These were either caused by
-a numeric optimization issue or another likelihood mode. Given that this looked
+a numeric optimization issue or the presence of multiple local optima. Given that this looked
 to be artificially inflating the uncertainty compared to all other models, we
@@
 finalized. We use the HapMap recombination map
-\parencite{International_HapMap_Consortium2007-nu}, PhastCons regions, and UTRs
+\parencite{International_HapMap_Consortium2007-nu}, PhastCons regions \parencite{Siepel2005-wh}, and UTRs
 as an approximate set of putatively conserved segments.
@@
 coalescence noise is much larger than sampling noise, and is given by
-$\var(\pi) \approx \nicefrac{2\theta^2}{9}$ in an infinite sample. One can
+$\var(\pi) \approx \nicefrac{2\theta^2}{9}$ in an infinite sample at a non-recombining locus. One can
 derive an analogous equation for the variance of average simulation $\bar{B}$
@@
 
-\subsection{Optimization and Diagnostics}
-\label{supp:optim}
-
-As described in the Methods section we use the BOBYQA algorithm implemented by
-the NLopt library for numeric optimization to find maximum likelihood estimates
-\parencite{Powell2009-jm,Johnson2007-tl}. The DFE is a simplex, and thus
-requires constrained optimization. We compared constrained optimization
-algorithms to an unconstrained softmax approach, finding the latter vastly
-outperformed the former. Bounds for $\pi_0$ and $\mu$ can be set manually in
-\texttt{bgspy}; our main models used $10^{-8} \le \mu \le 5 \times 10^{-8}$ and
-$0.0005 \le \pi_0 \le 0.008$. All main model fits were not at bounds. For
-locally-rescaled fits, we found our $\widehat{\mu}$ estimate was continually
-the upper bound; for these fits we increased this first to $\mu = 8 \times
-10^{-8}$ and then again to $\mu = \times 10^{-7}$, find it still always reached
-the upper bound. Since $10^{-7}$ is an implausibly high mutation rate, we
-stopped there.
-
-We only show one optimization diagnostic plot here (Supplementary Figure
-\ref{suppfig:diag-plot}), but the can be easily calculated from the model
-pickle files in the XXX repository (XXX). In one case, our optimization
-diagnostics indicated that the best optima across 10,000 random starts was one
-of few of the top 100 hits that were hitting either an alternate likelihood
-mode, or the optimization routine had numeric issues. In this one case, we
-removed the top hit and used the second, which was the consensus among the
-other top 100 optima.
+\clearpage
+\subsection{Locally Rescaled Fit Results}
+\label{supp:locally-rescaled-fits}
 
 \begin{figure}[htbp]
-  \label{suppfig:diag-plot}
-  \centering
-  \includegraphics[width=\textwidth]{figures/supplementary/figure_feature_priority_yri_full_diag.pdf}
-
-  \caption{ An optimization diagnostic plot showing the 100 top optima. Heatmap
-  colors indicate probability mass for the DFE parameters (top four rows).
-  Estimates of $\pi_0$ and $\mu$ are shown in the bottom two rows. This is the
-optimization plot for YRI Feature Priority model.}
-\end{figure}
-
-\clearpage
-\subsection{Locally-Rescaled Fit Results}
-\label{supp:locally-rescaled-fits}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[]{figures/supplementary/rescale_pi_pred_plot.pdf}
-
-  \caption{Chromosome 2 predictions on the YRI samples, with locally-rescaled
-      model fits. Observed data is the dark gray line, and the normal MLE for
-      the PhastCons Priority model is the blue line. The locally-rescaled
+  \centering
+  \includegraphics[width=\textwidth]{figures/supplementary/rescale_pi_pred_plot.pdf}
+
+  \caption{Chromosome 2 predictions on the YRI samples, with locally rescaled
+      model fits. The observed data is the dark gray line, and the normal MLE for
+      the PhastCons Priority model is the blue line. The locally rescaled
       predictions are the green line. The dashed red line are the prediction
-      using the standard B' map (without local-rescaled) and the maximum likelihood
-      estimates from the locally-rescaled fits. The large discrepancy in this shows
+      using the standard B' map (without local rescaling) and the maximum likelihood
+      estimates from the locally rescaled fits. The large discrepancy in this shows
       that estimates are highly dependent on the B' map.}
@@
 
-  \caption{The standard (dark gray) and locally-rescaled (blue) B' maps for different
+  \caption{The standard (dark gray) and locally rescaled (blue) B' maps for different
       $\mu = 1.6 \times 10^{-8}$ and three different selection coefficients.
-      For $s=10^{-5}$ ($2N_e s = 0.2$), locally-rescaling alters the predicted
+      For $s=10^{-5}$ ($2N_e s = 0.2$), locally rescaling alters the predicted
       reduction so that it is essentially insignificant ($B \approx 1$). For
   mid-strength selection $s=10^{-4}$ ($2N_e s = 2$), there is only a very
-slight difference between standard and locally-rescaled B' maps. Finally, for
-strong selection $s=10^{-3}$ ($2N_e s = 20$), locally-rescaling does not change
+slight difference between standard and locally rescaled B' maps. Finally, for
+strong selection $s=10^{-3}$ ($2N_e s = 20$), local rescaling does not change
 the B' maps, as expected. }
@@
 
-  \caption{The DFE for YRI samples with locally-rescaling. The maximum
+  \caption{The DFE for YRI samples with local rescaling. The maximum
   likelihood mutation rate estimate for this is $\widehat{\mu} = 8 \times
@@
   $\mu = 1.58 \times 10^{-8}$, for both the standard (blue) and
-locally-rescaled B' (orange) maps. This indicates that locally-rescaling the B'
+locally rescaled B' (orange) maps. This indicates that locally rescaling the B'
 maps only in practice changes how deep the ``U" is.}
@@
 
-\subsubsection{PhastCons Priority Locally-Rescaled Model Fit Results}
-
-Below are the model fit results for locally-rescaled model runs, with the
-upper mutation rate boundary set to $\mu = 8 \times 10^{-8}$.
+\subsubsection{PhastCons Priority Locally Rescaled Model Fit Results}
+
+Below are the model fit results for locally rescaled model runs, with the upper mutation rate boundary set to $\mu = 8 \times 10^{-8}$. Each model fit is the direct output from a
+\texttt{bgspy}.
+The \texttt{w grid} and \texttt{t grid} are the grids the B and B' maps are built on; note that the mutation grid \texttt{w} is an exponential that is inferred later, so optimization is done continuously over this.
+Each model output shows the maximum likelihood estimates, with uncertainty where available.
 
@@
   mut. rate  = 8e-08 
-  Ne = 26,572 (if μ=1e-8), Ne = 13,286 (if μ=2e-8)
+  Ne = 26,572 (if mu=1e-8), Ne = 13,286 (if mu=2e-8)
   R² = 60.8648% (in-sample)  59.1654% (out-sample)
@@
   mut. rate  = 8e-08 
-  Ne = 36,974 (if μ=1e-8), Ne = 18,487 (if μ=2e-8)
+  Ne = 36,974 (if mu=1e-8), Ne = 18,487 (if mu=2e-8)
   R² = 68.7046% (in-sample)  67.8854% (out-sample)
@@
   mut. rate  = 8e-08 
-  Ne = 27,755 (if μ=1e-8), Ne = 13,877 (if μ=2e-8)
+  Ne = 27,755 (if mu=1e-8), Ne = 13,877 (if mu=2e-8)
   R² = 62.5591% (in-sample)  61.367% (out-sample)
@@
 
-\subsubsection{CADD 6\% Locally-Rescaled Model Fit Results}
+\subsubsection{CADD 6\% Locally Rescaled Model Fit Results}
 
@@
 mut. rate  = 7.291e-08
-Ne = 25,506 (if μ=1e-8), Ne = 12,753 (if μ=2e-8)
+Ne = 25,506 (if mu=1e-8), Ne = 12,753 (if mu=2e-8)
 R² = 60.6057% (in-sample)
@@
 mut. rate  = 5.991e-08
-Ne = 26,856 (if μ=1e-8), Ne = 13,428 (if μ=2e-8)
+Ne = 26,856 (if mu=1e-8), Ne = 13,428 (if mu=2e-8)
 R² = 62.0668% (in-sample)
@@
 mut. rate  = 6.024e-08
-Ne = 35,914 (if μ=1e-8), Ne = 17,957 (if μ=2e-8)
+Ne = 35,914 (if mu=1e-8), Ne = 17,957 (if mu=2e-8)
 R² = 68.6747% (in-sample)
@@
   \centering
-  \includegraphics[width=0.8\textwidth]{figures/supplementary/feature_priority_pop_dfe_strongselgrid.pdf}
+  \includegraphics[width=0.7\textwidth]{figures/supplementary/feature_priority_pop_dfe_strongselgrid.pdf}
 
@@
   \centering
-  \includegraphics[width=0.8\textwidth]{figures/supplementary/phastcons_priority_pop_dfe_strongselgrid.pdf}
+  \includegraphics[width=0.7\textwidth]{figures/supplementary/phastcons_priority_pop_dfe_strongselgrid.pdf}
 
@@
 
+\clearpage
 \subsection{Demographic Expansion and Recessive Simulation Tests}
@@
   \includegraphics[width=0.8\textwidth]{figures/supplementary/spatial_residual.pdf}
+  \caption{Blue lines represent the residual between predicted an observed diversity levels along chromosomes. The orange line is a lowess curve, with bandwidth chosen by cross-validation.}
   \label{suppfig:spatial-residuals}
